### 2003-11-24 Documentation/Manual.tex
*** empty log message ***

\documentclass[a4paper,12pt,twoside]{book}

\sloppy
% \frenchspacing
\setlength{\unitlength}{1cm}

\usepackage{cite}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage{a4wide}
\usepackage{fancyheadings}
\usepackage[footnotesize,bf]{caption}
\usepackage{ae,aecompl} % only for Type 1 fonts

\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\ra}{$\rightarrow$}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}
\newcommand{\nn}{\nonumber}
\newcommand{\ldm}{$\Delta m_{31}^2$}
\newcommand{\sdm}{$\Delta m_{21}^2$}
\newcommand{\deltacp}{$\delta_{\mathrm{CP}}$}
\newcommand{\stheta}{$\sin^2 2 \theta_{13}$}

\newcommand{\ie}{{\it i.e.}}
\newcommand{\Ie}{{\it I.e.}}
\newcommand{\eg}{{\it e.g.}}
\newcommand{\Eg}{{\it E.g.}}
\newcommand{\cf}{{\it cf.}}
\newcommand{\etc}{{\it etc.}}
\newcommand{\eq}{Eq.}
\newcommand{\eqs}{Eqs.}
\newcommand{\Def}{Definition}
\newcommand{\fig}{Fig.}
\newcommand{\Fig}{Fig.}
\newcommand{\figs}{Figs.}
\newcommand{\Figs}{Figs.}
\newcommand{\Ref}{Ref.}
\newcommand{\Refs}{Refs.}
\newcommand{\Sec}{Sec.}
\newcommand{\Secs}{Secs.}
\newcommand{\App}{the Appendix}
\newcommand{\Apps}{Appendices}
\newcommand{\Tab}{Table}
\newcommand{\Tabs}{Tables}

\newcommand{\JHFSK}{{\sc JHF-SK}}
\newcommand{\NUMI }{{\sc NuMI}}
\newcommand{\ReactorI}{{\sc Reactor-I}}
\newcommand{\ReactorII}{{\sc Reactor-II}}
\newcommand{\JHFHK}{{\sc JHF-HK}}
\newcommand{\NuFactI}{{\sc NuFact-I}}
\newcommand{\NuFactII}{{\sc NuFact-II}}

\newcommand{\GLOBES}{{\sf GLoBES}}
\newcommand{\EDM}{{\sf EDM}}

\newcommand{\equ}[1]{\eq~(\ref{equ:#1})}
\newcommand{\figu}[1]{\fig~\ref{fig:#1}}

%-- page parameters -------------------------------------------------

\pagestyle{fancyplain}

% Olddefs:

%\addtolength{\textwidth}{-1truecm}
%\addtolength{\oddsidemargin}{1.0truecm}
%\addtolength{\evensidemargin}{-0.3truecm}

% Owndefs:

%\addtolength{\textwidth}{0.0truecm}
%\addtolength{\textheight}{1.8truecm}
%\addtolength{\topmargin}{-0.6truecm}
%\addtolength{\oddsidemargin}{0.65truecm}
%\addtolength{\evensidemargin}{-1.0truecm}

% Headdefs:

\advance \headheight by 3.0truept       % for 12pt mandatory...
\lhead[\fancyplain{}{\thepage}]{\fancyplain{}{\rightmark}}
\rhead[\fancyplain{}{\leftmark}]{\fancyplain{\thepage}{\thepage}}
\cfoot{}

\renewcommand{\chaptermark}[1]{
% chapter title im Seitenkopf
\markboth{\uppercase{\chaptername}\ \thechapter.\ \ #1}
          {\uppercase{\chaptername}\ \thechapter.\ \ #1}}
% section title im Seitenkopf
\renewcommand{\sectionmark}[1]{\markright{\thesection\ \ #1}}

\begin{document}

\pagenumbering{Roman}
\thispagestyle{empty}
{
\setlength{\parindent}{0cm}

%\begin{titlepage}

\font\fa=cmssbx14 scaled 1440
\font\fb=cmbx10 scaled 1200
\font\fc=cmb10 scaled 1200
\font\fd=cmr10 scaled 1000

{\setlength{\baselineskip}{1.2cm}}

\input{tumlogo}

\begin{minipage}{7cm}
\begin{center}
{\bf Technische Universit{\"a}t M{\"u}nchen} \\
Physik-Department \\
Institut f{\"u}r Theoretische Physik, T30d \\
\end{center}
\end{minipage}

\vspace{-2cm}

\hfill \oTUM{3.5cm}

\vspace{3cm}

\begin{center}
{ \Large \bf
\GLOBES\ 1.X \\
Global Long Baseline Experiment Simulator \\ }
\vspace*{0.5cm}
{\large \bf User's and reference manual }
\end{center}

\vspace{1cm}

\begin{center}
{\large Patrick Huber, Manfred Lindner, Walter Winter}
\end{center}

\vspace{1cm}

\begin{center}
Here comes our logo!
\end{center}

\vspace{1cm}

\begin{center}
Version from \today
\end{center}

%\end{titlepage}

%\title{}
%\date{}
%\author{}
%\maketitle

}

\clearpage
\thispagestyle{empty}

\cleardoublepage
\setcounter{page}{1}

\chapter*{What is \GLOBES\ good for?}

Here comes a maximum of one page what \GLOBES\ is good for and what main features it has.
\cleardoublepage
\tableofcontents

\cleardoublepage
\setcounter{page}{1}
\pagenumbering{arabic}

\chapter*{How to use this manual}
\addcontentsline{toc}{chapter}{How to use this manual}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PART I: User's manual
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\part{User's manual}

\chapter*{Getting started}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PART II: Experiment definition module
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\part{The Experiment Definition Module (\EDM )}

\chapter{Basics: Source, Oscillation, Detection}

Here comes how a experiment is defined (what is a channel, a rule, how are they treated, ...). Splitting in source, oscillation, and detector, etc.

\chapter{Getting started: Editing existing experiments}

Here comes how to load and modify existing experiments;
also: how does the main screen work and the most important functions at exampled

\chapter{Features of the \EDM }

\section{Source properties}

\section{Oscillation properties}

\section{Detector properties}

\chapter{Creating new experiments}

Most advanced step: How to create a new experiment from scratch.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PART III: Reference manual
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\part{Reference manual}

Here comes a list of all functions and their function in alphabetical order.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{appendix}

\chapter{\GLOBES\ installation}

\chapter{Statistics and \GLOBES }

Here comes a statistics summary (similar to Superbeams appendix):
\bi
\item
 What $\chi^2$ is used
\item
 How channels/rules are treated
\item
 How does the flux-folding work
\item
 ...
\ei

\end{appendix}
\end{document}

### 2003-11-24 Documentation/Manual.tex
*** empty log message ***

@@
 
-\part{User's manual}
-
-\chapter*{Getting started}
+\input{part1}
 
@@
 
-\part{The Experiment Definition Module (\EDM )}
-
-\chapter{Basics: Source, Oscillation, Detection}
-
-Here comes how a experiment is defined (what is a channel, a rule, how are they treated, ...). Splitting in source, oscillation, and detector, etc.
-
-\chapter{Getting started: Editing existing experiments}
-
-Here comes how to load and modify existing experiments;
-also: how does the main screen work and the most important functions at exampled
-
-\chapter{Features of the \EDM }
-
-\section{Source properties}
-
-\section{Oscillation properties}
-
-\section{Detector properties}
-
-\chapter{Creating new experiments}
-
-Most advanced step: How to create a new experiment from scratch.
+\input{part2}
 
@@
 
-\part{Reference manual}
-
-Here comes a list of all functions and their function in alphabetical order.
+\input{part3}
 
@@
 
-\begin{appendix}
+\input{appendix}
 
-\chapter{\GLOBES\ installation}
-
-\chapter{Statistics and \GLOBES }
-
-Here comes a statistics summary (similar to Superbeams appendix):
-\bi
-\item
- What $\chi^2$ is used
-\item
- How channels/rules are treated
-\item
- How does the flux-folding work
-\item
- ...
-\ei
-
-\end{appendix}
 \end{document}

### 2003-11-24 Documentation/appendix.tex
*** empty log message ***

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{appendix}

\chapter{\GLOBES\ installation}

\chapter{Statistics and \GLOBES }

Here comes a statistics summary (similar to Superbeams appendix):
\bi
\item
 What $\chi^2$ is used
\item
 How channels/rules are treated
\item
 How does the flux-folding work
\item
 ...
\ei

\chapter{Licence agreement}

\chapter*{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}

\end{appendix}

### 2003-11-24 Documentation/part1.tex
*** empty log message ***

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PART I: User's manual
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\part{User's manual}
\chapter{Getting started: A short \GLOBES\ tour}

Also: which units, which order of parameters (table for different functions!) etc. (somewhere)

\chapter{Setting up \GLOBES }

\bi
\item
 How to install and start \GLOBES
\item
 How to set the vacuum parameters 
\item
 Which units
\item
 How to load one or more experiments and re-obtain formation about them
\item
Table: pre-defined experiment types
\item
 How to clear experiments
\ei

\chapter[Calculating $\chi^2$ with systematics only]{Calculating $\boldsymbol{\chi^2}$ with systematics only}

\bi
\item
 More technical: what does it do; how to imagine; what functions are there; examples
 Single/multiple experiments and the connection (sum of individual $\chi$'s )
\item
 Advanced topic: Changing the systematics treatment (SetErrorDim)
\ei

\chapter[Calculating $\chi^2$-projections: how to include correlations]{Calculating $\boldsymbol{\chi^2}$-projections: how to include correlations}

\section[Projection onto $\theta_{13}$ or $\delta$]{Projection onto $\boldsymbol{\theta_{13}}$ or $\boldsymbol{\delta}$}

\bi
\item
 Single and multiple experiments
\item
 Figure: how does that work: local routine
\item
 Treatment of external input
\ei

\section[Projection onto any set of $N \ge 1$ parameters]{Projection onto any set of $\boldsymbol{N \ge 1}$ parameters}

General form and especially: DeltaTheta-Minimizer; mention calculation time!

\chapter{Finding degenerate solutions}

\bi
\item
 Single- and multiple experiment support
\item
 What does that do?
\item
 Example: how to treat sign-degeneracy in th13-limit
\ei

\chapter{Obtaining low-level information}

\bi
\item
 Obtaining rate vectors
\item
 Obtaining fluxes/cross sections etc.
\item
 ``Check''-functions
\ei

\chapter{Changing experiment parameters at the running time}

\bi
\item
 Baseline, Target mass
\item
 Matter density profile
\item
 Threshold function, efficiencies etc.
\ei

### 2003-11-24 Documentation/part2.tex
*** empty log message ***

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PART II: Experiment definition module
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\part{The Experiment Definition Module (\EDM )}

\chapter{Basics: Source, Oscillation, Detection}

Here comes how a experiment is defined (what is a channel, a rule, how are they treated, ...). Splitting in source, oscillation, and detector, etc.

\chapter{Getting started: Editing existing experiments}

Here comes how to load and modify existing experiments;
also: how does the main screen work and the most important functions at exampled

\chapter{Features of the \EDM }

\section{Source properties}

\section{Oscillation properties}

\section{Detector properties}

\chapter{Creating new experiments}

Most advanced step: How to create a new experiment from scratch.

### 2003-11-24 Documentation/part3.tex
*** empty log message ***

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PART III: Reference manual
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\part{Reference manual}

Here comes a list of all functions and their function in alphabetical order.

### 2003-11-27 Documentation/part1.tex
*** empty log message ***

@@
 
+\chapter{Setting up \GLOBES }
+\textsc{} 
 Also: which units, which order of parameters (table for different functions!) etc. (somewhere)
 
-\chapter{Setting up \GLOBES }
+\begin{table}[t]
+\begin{center}
+\begin{tabular}{lll}
+\hline
+Quantities & Examples & Units \\
+\hline
+Angles & $\theta_{13}$, $\theta_{12}$, $\theta_{23}$ & Radians  \\
+Mass squared differences & $\sdm$, $\ldm$ & $\mathrm{eV}^2$ \\
+Matter densitities & $\rho_i$ & $\mathrm{g}/\mathrm{cm}^3$ \\
+Baseline lengths & $L_i$ & $\mathrm{km}$ \\
+Energies & $E_\nu$ & $\mathrm{GeV}$ \\  
+Fiducial masses & $m_{\mathrm{Det}}$ & $\mathrm{kt}$ \\
+Time intervals & $t_{\mathrm{run}}$ & $\mathrm{yr}$ \\
+Integrated luminosities & $m_{\mathrm{Det}} \, t_{\mathrm{run}}$ & $\mathrm{kt \cdot yr}$ \\
+Cross sections & $\sigma_{\mathrm{CC}}$ &  ???? \\
+\hline
+\end{tabular}
+\mycaption{\label{tab:units} Quantities used in \GLOBES , examples of these quantities, and their standard units in the application software.}
+\end{center}
+\end{table}
+
+\begin{table}[t]
+\begin{center}
+\begin{tabular}{llp{7cm}c}
+\hline
+Experiment & Filename & Short description & Refs. \\
+\hline 
+\multicolumn{3}{l}{\underline{Conventional beams:}} \\
+??? & & \\[0.1cm]
+
+\multicolumn{3}{l}{\underline{First-generation superbeams:}} \\
+\JHFSK\ ($\nu$) & {\tt JHFSK.exp} & JHF (J-PARC) to Super-Kamiokande, neutrino running &  \cite{Huber:2002mx,Huber:2002rs} \\
+\JHFSK\ ($\bar\nu$)& {\tt JHFSKanti.exp} & JHF (J-PARC) to Super-Kamiokande, antineutrino running &  \cite{Huber:2002rs} \\
+\NUMI\  ($\nu$), OA $9 \, \mathrm{km}$ & {\tt NUMI9.exp} & NuMI with off-axis angle of $9 \, \mathrm{km}$ for $L=712 \, \mathrm{km}$, neutrino running & \cite{Huber:2002rs} \\
+\NUMI\  ($\bar{\nu}$), OA $9 \, \mathrm{km}$ & {\tt NUMI9anti.exp} & NuMI with off-axis angle of $9 \, \mathrm{km}$ for $L=712 \, \mathrm{km}$, antineutrino running & \cite{Huber:2002rs} \\
+\NUMI\  ($\nu$), OA $12 \, \mathrm{km}$ & {\tt NUMI12.exp} & NuMI with off-axis angle of $12 \, \mathrm{km}$ for $L=712 \, \mathrm{km}$, neutrino running & \cite{Huber:2002rs} \\
+\NUMI\  ($\bar{\nu}$), OA $12 \, \mathrm{km}$ & {\tt NUMI12anti.exp} & NuMI with off-axis angle of $12 \, \mathrm{km}$ for $L=712 \, \mathrm{km}$, antineutrino running & \cite{Huber:2002rs} \\
+\SPL\  ($\nu$) & {\tt SPL.exp} & SPL (CERN), neutrino running &  ??? \\
+\SPL\  ($\bar\nu$) & {\tt SPLanti.exp} & SPL (CERN), antineutrino running & ??? \\[0.1cm]
+
+\multicolumn{3}{l}{\underline{Superbeam upgrades:}} \\
+\JHFHK\ ($\nu$) & {\tt JHFHK.exp} & JHF (J-PARC) to Hyper-Kamiokande superbeam upgrade, neutrino running &  \cite{Huber:2002mx,Huber:2002rs} \\
+\JHFHK\ ($\bar\nu$)& {\tt JHFHKanti.exp} & JHF (J-PARC) to Hyper-Kamiokande superbeam upgrade, antineutrino running &  \cite{Huber:2002mx,Huber:2002rs} \\[0.1cm]
+
+\multicolumn{3}{l}{\underline{Neutrino factories:}} \\
+\NuFactI\ & {\tt NuFact.exp} & Initial stage neutrino factory, symmetric operation in both polarities & \cite{Huber:2002mx} \\
+\NuFactII\  & {\tt NuFact2.exp} & Advanced stage neutrino factory, symmetric operation in both polarities & \cite{Huber:2002mx,Huber:2003ak} \\[0.1cm]
+
+\multicolumn{3}{l}{\underline{Reactor experiments:}} \\
+\ReactorI\ & {\tt Reactor.exp} & Small reactor experiment with identical near and far detectors & \cite{Huber:2003pm} \\
+\ReactorII\ & {\tt Reactor2.exp} & Large reactor experiment with identical near and far detectors & \cite{Huber:2003pm} \\[0.1cm]
+
+\multicolumn{3}{l}{\underline{$\beta$-Beams:}} \\
+\Beta\ ($\nu$) & {\tt BETA.exp} & $\beta$-Beam, neutrino running & ??? \\
+\Beta\ ($\bar\nu$) & {\tt BETAanti.exp} & $\beta$-Beam, antineutrino running & ??? \\
+\hline
+\end{tabular}
+\end{center}
+\mycaption{\label{tab:experiments} Different pre-defined experiments, their filenames (to be used in {\tt LoadExperiment}), their short description, and the references in which they are defined.
+Note that all experiments use in their standard configurations one year of running time.
+Details about the experiment parameters can be obtained with {\tt InfoExperiment}(Experiment number) after they have been loaded.}
+\end{table}
 

### 2003-11-28 Documentation/part1.tex
*** empty log message ***

@@
 \chapter{Getting started: A short \GLOBES\ tour}
+
+\begin{table}[t]
+\begin{center}
+\begin{tabular}{p{1.8cm}p{4.5cm}p{8.6cm}}
+\hline
+Function & Purpose & Parameters \ra\ Result \\
+\hline
+{\tt Chi} & $\chi^2$ with systematics only \newline (all initialized exps.) & ($\{ \theta_{12}, \theta_{13}, \theta_{23}, \deltacp , \sdm , \ldm, \frac{\rho_1}{\bar{\rho}_1},...
+, \frac{\rho_n}{\bar{\rho}_n} \}$)  \newline \ra\  $\chi^2$ \\[0.1cm]
+{\tt SingleChi} & $\chi^2$ with systematics only \newline (only one experiment) & $(\{ \theta_{12}, \theta_{13}, \theta_{23}, \deltacp , \sdm , \ldm, \frac{\rho_{N_{\mathrm{exp}}}}{\bar{\rho}_{N_{\mathrm{exp}}}} \}, \, N_{\mathrm{exp}} )$   \newline \ra\ $\chi^2$ \\[0.1cm]
+{\tt ChiTheta} & $\chi^2$ with systematics and correlations: Projection onto $\theta_{13}$-axis (all exps.) &  ($ \theta_{13}, \, \{ \theta_{12}, \theta_{23}, \deltacp , \sdm , \ldm, \frac{\rho_1}{\bar{\rho}_1}, ...
+, \frac{\rho_n}{\bar{\rho}_n} \}$) \newline \ra\  $\{ \chi^2, \theta_{12}, \theta_{23}, \deltacp , \sdm , \ldm, \frac{\rho_1}{\bar{\rho}_1}, ...
+, \frac{\rho_n}{\bar{\rho}_n} , N_{\mathrm{Iter}} \}$ \\[0.1cm]
+{\tt Single-} \newline {\tt ChiTheta} & $\chi^2$ with systematics and correlations: Projection onto $\theta_{13}$-axis (one exp.) &  ($ \theta_{13}, \, \{ \theta_{12}, \theta_{23}, \deltacp , \sdm , \ldm,  \frac{\rho_{N_{\mathrm{exp}}}}{\bar{\rho}_{N_{\mathrm{exp}}}} \}, \, N_{\mathrm{exp}}$) \newline \ra\  $\{ \chi^2, \theta_{12}, \theta_{23}, \deltacp , \sdm , \ldm, \frac{\rho_{N_{\mathrm{exp}}}}{\bar{\rho}_{N_{\mathrm{exp}}}} , N_{\mathrm{Iter}} \}$ \\[0.1cm]
+{\tt ChiDelta} & $\chi^2$ with systematics and correlations: Projection onto $\deltacp$-axis (all exps.) &  ($\deltacp, \, \{ \theta_{12}, \theta_{13}, \theta_{23},  \sdm , \ldm, \frac{\rho_1}{\bar{\rho}_1}, ...
+, \frac{\rho_n}{\bar{\rho}_n} \}$) \newline \ra\  $\{ \chi^2, \theta_{12}, \theta_{13}, \theta_{23}, \sdm , \ldm, \frac{\rho_1}{\bar{\rho}_1}, ...
+, \frac{\rho_n}{\bar{\rho}_n} , N_{\mathrm{Iter}} \}$ \\[0.1cm]
+{\tt Single-} \newline {\tt ChiDelta} & $\chi^2$ with systematics and correlations: Projection onto $\deltacp$-axis (one exp.) &  ($ \deltacp, \, \{ \theta_{12}, \theta_{13}, \theta_{23}, \sdm , \ldm,  \frac{\rho_{N_{\mathrm{exp}}}}{\bar{\rho}_{N_{\mathrm{exp}}}} \}, \, N_{\mathrm{exp}}$) \newline \ra\  $\{ \chi^2, \theta_{12}, \theta_{13}, \theta_{23},  \sdm , \ldm, \frac{\rho_{N_{\mathrm{exp}}}}{\bar{\rho}_{N_{\mathrm{exp}}}} , N_{\mathrm{Iter}} \}$ \\[0.1cm]
+{\tt ChiTheta-} {\tt Delta} & $\chi^2$ with systematics and correlations: Projection onto $\deltacp$-$\theta_{13}$-plane (all exps.) &  ($\theta_{13}, \, \deltacp, \, \{ \theta_{12}, \theta_{23},  \sdm , \ldm, \frac{\rho_1}{\bar{\rho}_1}, ...
+, \frac{\rho_n}{\bar{\rho}_n} \}$) \newline \ra\  $\{ \chi^2, \theta_{12}, \theta_{23}, \sdm , \ldm, \frac{\rho_1}{\bar{\rho}_1}, ...
+, \frac{\rho_n}{\bar{\rho}_n} , N_{\mathrm{Iter}} \}$ \\[0.1cm]
+{\tt SingleChi-} {\tt Theta- } {\tt Delta} & $\chi^2$ with systematics and correlations: Projection onto $\deltacp$-$\theta_{13}$-plane (one exp.) &  ($ \theta_{13}, \, \deltacp, \, \{ \theta_{12}, \theta_{23}, \sdm , \ldm,  \frac{\rho_{N_{\mathrm{exp}}}}{\bar{\rho}_{N_{\mathrm{exp}}}} \}, \, N_{\mathrm{exp}}$) \newline \ra\  $\{ \chi^2, \theta_{12},  \theta_{23},  \sdm , \ldm, \frac{\rho_{N_{\mathrm{exp}}}}{\bar{\rho}_{N_{\mathrm{exp}}}} , N_{\mathrm{Iter}} \}$ \\[0.1cm]
+{\tt ChiNP} & $\chi^2$ with systematics and correlations: Projection onto $N$-parameter hyper-plane (one exp.) & ??? to be defined \\[0.1cm]
+{\tt Single-} {\tt ChiNP} & $\chi^2$ with systematics and correlations: Projection onto $N$-parameter hyper-plane (one exp.) & ??? to be defined \\[0.1cm]
+{\tt ChiAll} & Lokal minimum of $\chi^2$ with respect to all parameters (all exps.) &
+($ \{ \theta_{13}, \theta_{12}, \theta_{23}, \deltacp , \sdm , \ldm, \frac{\rho_1}{\bar{\rho}_1}, ...
+, \frac{\rho_n}{\bar{\rho}_n} \}$) \newline \ra\  $\{ \chi^2, \theta_{13}, \theta_{12}, \theta_{23},$ \newline \hspace*{1.4cm} $ \deltacp , \sdm , \ldm, \frac{\rho_1}{\bar{\rho}_1}, ...
+, \frac{\rho_n}{\bar{\rho}_n} , N_{\mathrm{Iter}} \}$
+\\[0.1cm]
+{\tt Single-} {\tt ChiAll} & Lokal minimum of $\chi^2$ with respect to all parameters (one exp.) &  ($ \{ \theta_{13}, \theta_{12}, \theta_{23}, \deltacp , \sdm , \ldm,  \frac{\rho_{N_{\mathrm{exp}}}}{\bar{\rho}_{N_{\mathrm{exp}}}} \}, \, N_{\mathrm{exp}}$) \newline \ra\  $\{ \chi^2, \theta_{13}, \theta_{12}, \theta_{23}, \deltacp , \sdm , \ldm, \frac{\rho_{N_{\mathrm{exp}}}}{\bar{\rho}_{N_{\mathrm{exp}}}} , N_{\mathrm{Iter}} \}$ \\[0.1cm]
+\hline
+\end{tabular}
+\end{center}
+\caption{\label{tab:stdfunctions} The \GLOBES\ standard function to obtain a $\chi^2$-value for all or one of the initialized experiments.
+The curly brackets refer to the parameters to be transferred in form of a list.
+Note that all functions but {\tt Chi} and {\tt SingleChi} are using minimizers which have to be initialiued with {\tt SetInputErrors} and {\tt SetStartingValues} first.}
+\end{table}
 
@@
 \hline
-Angles & $\theta_{13}$, $\theta_{12}$, $\theta_{23}$ & Radians  \\
+Angles & $\theta_{13}$, $\theta_{12}$, $\theta_{23}$, $\deltacp$ & Radians  \\
 Mass squared differences & $\sdm$, $\ldm$ & $\mathrm{eV}^2$ \\

### 2003-12-01 Documentation/part1.tex
*** empty log message ***

@@
 \part{User's manual}
-\chapter{Getting started: A short \GLOBES\ tour}
+\chapter{A short \GLOBES\ tour}
 
@@
 
-\chapter{Setting up \GLOBES }
-\textsc{} 
-Also: which units, which order of parameters (table for different functions!) etc. (somewhere)
+\chapter{Getting started with \GLOBES }
+
+In this first chapter of the user's manual, we assume that the \GLOBES\ software is readily installed on your computer system.
+We demonstrate how to load pre-defined experiments and re-obtain information about them.
+However, we only secondarily discuss the usage of \GLOBES\ in your specific programming language, such as C, Mathematica, or others.
+Thus, you should be familiar of how to load \GLOBES\ on your computer system before reading this chapter.
+An example of how to use \GLOBES\ with C can be found on page~\pageref{ex:c}.
+
+\example{Using \GLOBES\ with C}{\label{ex:c}
+
+Here comes some complete C-code with a very simple example of how to use \GLOBES .
+
+}
 
@@
 Time intervals & $t_{\mathrm{run}}$ & $\mathrm{yr}$ \\
-Integrated luminosities & $m_{\mathrm{Det}} \, t_{\mathrm{run}}$ & $\mathrm{kt \cdot yr}$ \\
+Source powers & $P_{\mathrm{Source}}$ & ??? \\
+% Integrated luminosities & $m_{\mathrm{Det}} \, t_{\mathrm{run}}$ & $\mathrm{kt \cdot yr}$ \\
 Cross sections & $\sigma_{\mathrm{CC}}$ &  ???? \\
@@
 \end{table}
+
+Throughout the programming interface of \GLOBES , the software needs to transfer parameters to and from the software core.
+Unless the Experiment Definition Module \EDM , the programming interface only uses one set of units for each type of quantity in order to avoid confusion about the definition of individual parameters.
+\tabl{units} summarizes the units of the most important quantities used in \GLOBES .
 
@@
 
-\bi
-\item
- How to install and start \GLOBES
-\item
- How to set the vacuum parameters 
-\item
- Which units
-\item
- How to load one or more experiments and re-obtain formation about them
-\item
-Table: pre-defined experiment types
-\item
- How to clear experiments
-\ei
+In principle, \GLOBES\ can handle any number of different long-baseline experiments simultaneously.
+This means that their $\chi^2$-values are added {\em after} the minimization over the independent systematics parameters and {\em before} any minimization over the oscillation parameters.
+Though the simplest case of only one experiment may be most often used, more experiments are useful in many cases.
+For example, running a superbeam some years in the neutrino mode and some years in the antineutrino mode can be, to a first approximation, simulated by the combination of two such experiments.\footnote{Note that in this case the systematics parameters are minimized over independently, which means that this approach does not allow correlations among the systematics parameters.
+Therefore, the neutrino factory with the symmetric operation of both polarities is encapsulated into a single experiment.} Another example is the test of synergetic effects among different experiment types.
+Thus, \GLOBES\ has an internal (initially empty) list of currently initialized experiments.
+To add a pre-defined experiment to this list, one can use the function {\tt LoadExperiment}:
+\begin{function}
+{\tt LoadExperiment}$($``filename'' $)$ adds a single experiment to the list of currently loaded experiments.
+All currently loaded experiments are evaluated simultaneously (if not explicitely stated otherwise), \ie , their $\chi^2$-values are added.
+\end{function}
+A list of pre-defined experiment types, their filenames, their short descriptions, and the references of their definitions can be found in \tabl{experiments}.
+To remove all experiments from the evaluation list, one uses {\tt ClearExp}:
+\begin{function}
+{\tt ClearExp}$()$ removes all experiments from the evaluation list.   
+\end{function}
+Both functions do not return anything.
+Thus, one can either add an experiment to the internal evaluation list, or remove all experiments from this list.
+After adding an experiment, it gets an internal experiment number $N_{\mathrm{exp}}$ assigned in the order of the addition, which is starting from zero and running to the number of experiments minus one.
+Therefore, one will be able to access the individual experiment by its number later.
+
+Since the pre-defined experiments in \tabl{experiments} are given for one year running time, specific target masses, and specific source powers, it is useful to change these parameters of the individual experiments:
+\begin{function}
+{\tt SetRunningTime}$(N_{\mathrm{exp}},t_{\mathrm{run}})$ sets the running time of experiment number $N_{\mathrm{exp}}$ to $t_{\mathrm{run}}$ years.
+\end{function}
+ \begin{function}
+{\tt SetTargetMass}$(N_{\mathrm{exp}},m_{\mathrm{Det}})$ sets the fiducial mass of experiment number $N_{\mathrm{exp}}$ to $m_{\mathrm{Det}}$ kilotons.
+\end{function}
+\begin{function}
+{\tt SetSourcePower}$(N_{\mathrm{exp}},P_{\mathrm{Source}})$ sets the source power of experiment number $N_{\mathrm{exp}}$ to $P_{\mathrm{source}}$.
+The definition of the source power depends on the experiment type: ...
+(MISSING).
+\end{function}
+Thus, these functions also demonstrate how to use the assigned experiment number.
+
+A useful function to re-obtain the information about the initialized experiments is the function {\tt InfoExperiment}:
+\begin{function}
+{\tt InfoExperiment}$()$ prints a list of the initialized functions with their experiment numbers and their most important parameters to the standard output.
+\end{function} 
+Especially, after changing individual parameters, such as baseline or target mass, this information can be useful to check the changes.
+Another useful function is {\tt ShowChannels}, which prints the initialized oscillation channels for a specified experiment:
+\begin{function}
+{\tt ShowChannels}$(N_{\mathrm{exp}})$ prints the information about the oscillation channels of the experiment with the number $N_{\mathrm{exp}}$ to the standard output.
+\end{function}
+
+Compared to an existing experiment, which uses real data, a future experiment uses simulated data.
+Thus, the {\em true parameter values} and their results in form of the reference rate vectors are simulated.
+After setting the true parameter values, the {\em fit parameter values} can be varied in order to obtain information on the measurement performance for the given set of true parameter values.
+Therefore, it is often useful to show the results of a future measurement as function of the true parameter values for which the reference rate vectors are computed -- at least within the currently allowed ranges.
+The true parameter values for the vacuum neutrino oscillation parameters have to be set by the functions {\tt SetVacuumParameters} and {\tt SetRates} {\em before} any evaluation function is used and {\em after} the experiments have been initialized and the experiment parameters have been adjusted which could change the rates (such as baseline or target mass).
+Any matter effects are then included automatically depending on the experiment definitions.
+\begin{function}
+{\tt SetVacuumParameters}$(\{\theta_{12}, \theta_{13}, \theta_{23}, \deltacp , \sdm , \ldm \})$ sets the neutrino oscillation parameters to be used to compute the reference rate vector in vacuum.
+\end{function}
+\begin{function}
+{\tt SetRates}$( )$ computes the reference rate vector for the neutrino oscillation parameters set with {\tt SetVacuumParameters}. 
+\end{function}
+Finally, an initialization sequence for \GLOBES\ could look like this:
+\begin{quote}
+{\tt ClearExp}$()$; \\
+{\tt InitExperiment}$($``{\tt JHFHK.exp}''$)$; \\
+{\tt SetRunningTime}$(0,2.0)$; \\
+{\tt InitExperiment}$($``{\tt JHFHKanti.exp}''$)$; \\
+{\tt SetRunningTime}$(1,6.0)$;\\
+{\tt MInfoExperiment}$()$; \\
+{\tt SetVacuumParameters}$(\{ 0.55, \, 0.16, \, 3.14/4, \, 3.14/2, \, 7e-5 , \, 2e-3 \} )$; \\
+{\tt SetRates();} 
+\end{quote}
+This piece of code initializes the JHF (J-PARC) to Hyper-Kamiokande superbeam upgrade with two years of neutrino running and six years of antineutrino running, \ie, an overall running time of eight years.
+The final configuration is then printed to the standard output and the reference rate vector is set to the chosen parameter values.
 
@@
 \ei
+
+\example{Correlation between $\stheta$ and $\deltacp$}{Text}
 
@@
 \item
- Baseline, Target mass
+ Baseline, Target mass, Source power, running time
 \item

### 2003-12-02 Documentation/part1.tex
*** empty log message ***

@@
 \chapter{A short \GLOBES\ tour}
+
+To be written later: Short introduction such as ``With the following lines we obtain ...''
+without detailed description of parameters. Should contain all most important functions of \GLOBES\ --
+overview of main functions in \tabl{stdfunctions}.
+Maybe: in form of long example which involves everything (for example: our famous bar plot, which involves
+systematics, correlations, degeneracies and thus the full set of \GLOBES\ functions without being too
+complicated in the application software part).
 
@@
 \item
- More technical: what does it do; how to imagine; what functions are there; examples
+ More technical: what does it do (minimizer); how to imagine (simple example); 
+what functions are there; examples
+\item
  Single/multiple experiments and the connection (sum of individual $\chi$'s )
 \item
- Advanced topic: Changing the systematics treatment (SetErrorDim)
+ Advanced topic: Changing the systematics treatment (SetSystematics)
 \ei

### 2003-12-02 Documentation/part1.tex
*** empty log message ***

@@
 \hline
-{\tt Chi} & $\chi^2$ with systematics only \newline (all initialized exps.) & ($\{ \theta_{12}, \theta_{13}, \theta_{23}, \deltacp , \sdm , \ldm, \frac{\rho_1}{\bar{\rho}_1},...
-, \frac{\rho_n}{\bar{\rho}_n} \}$)  \newline \ra\  $\chi^2$ \\[0.1cm]
-{\tt SingleChi} & $\chi^2$ with systematics only \newline (only one experiment) & $(\{ \theta_{12}, \theta_{13}, \theta_{23}, \deltacp , \sdm , \ldm, \frac{\rho_{N_{\mathrm{exp}}}}{\bar{\rho}_{N_{\mathrm{exp}}}} \}, \, N_{\mathrm{exp}} )$   \newline \ra\ $\chi^2$ \\[0.1cm]
-{\tt ChiTheta} & $\chi^2$ with systematics and correlations: Projection onto $\theta_{13}$-axis (all exps.) &  ($ \theta_{13}, \, \{ \theta_{12}, \theta_{23}, \deltacp , \sdm , \ldm, \frac{\rho_1}{\bar{\rho}_1}, ...
-, \frac{\rho_n}{\bar{\rho}_n} \}$) \newline \ra\  $\{ \chi^2, \theta_{12}, \theta_{23}, \deltacp , \sdm , \ldm, \frac{\rho_1}{\bar{\rho}_1}, ...
-, \frac{\rho_n}{\bar{\rho}_n} , N_{\mathrm{Iter}} \}$ \\[0.1cm]
-{\tt Single-} \newline {\tt ChiTheta} & $\chi^2$ with systematics and correlations: Projection onto $\theta_{13}$-axis (one exp.) &  ($ \theta_{13}, \, \{ \theta_{12}, \theta_{23}, \deltacp , \sdm , \ldm,  \frac{\rho_{N_{\mathrm{exp}}}}{\bar{\rho}_{N_{\mathrm{exp}}}} \}, \, N_{\mathrm{exp}}$) \newline \ra\  $\{ \chi^2, \theta_{12}, \theta_{23}, \deltacp , \sdm , \ldm, \frac{\rho_{N_{\mathrm{exp}}}}{\bar{\rho}_{N_{\mathrm{exp}}}} , N_{\mathrm{Iter}} \}$ \\[0.1cm]
-{\tt ChiDelta} & $\chi^2$ with systematics and correlations: Projection onto $\deltacp$-axis (all exps.) &  ($\deltacp, \, \{ \theta_{12}, \theta_{13}, \theta_{23},  \sdm , \ldm, \frac{\rho_1}{\bar{\rho}_1}, ...
-, \frac{\rho_n}{\bar{\rho}_n} \}$) \newline \ra\  $\{ \chi^2, \theta_{12}, \theta_{13}, \theta_{23}, \sdm , \ldm, \frac{\rho_1}{\bar{\rho}_1}, ...
-, \frac{\rho_n}{\bar{\rho}_n} , N_{\mathrm{Iter}} \}$ \\[0.1cm]
-{\tt Single-} \newline {\tt ChiDelta} & $\chi^2$ with systematics and correlations: Projection onto $\deltacp$-axis (one exp.) &  ($ \deltacp, \, \{ \theta_{12}, \theta_{13}, \theta_{23}, \sdm , \ldm,  \frac{\rho_{N_{\mathrm{exp}}}}{\bar{\rho}_{N_{\mathrm{exp}}}} \}, \, N_{\mathrm{exp}}$) \newline \ra\  $\{ \chi^2, \theta_{12}, \theta_{13}, \theta_{23},  \sdm , \ldm, \frac{\rho_{N_{\mathrm{exp}}}}{\bar{\rho}_{N_{\mathrm{exp}}}} , N_{\mathrm{Iter}} \}$ \\[0.1cm]
-{\tt ChiTheta-} {\tt Delta} & $\chi^2$ with systematics and correlations: Projection onto $\deltacp$-$\theta_{13}$-plane (all exps.) &  ($\theta_{13}, \, \deltacp, \, \{ \theta_{12}, \theta_{23},  \sdm , \ldm, \frac{\rho_1}{\bar{\rho}_1}, ...
-, \frac{\rho_n}{\bar{\rho}_n} \}$) \newline \ra\  $\{ \chi^2, \theta_{12}, \theta_{23}, \sdm , \ldm, \frac{\rho_1}{\bar{\rho}_1}, ...
-, \frac{\rho_n}{\bar{\rho}_n} , N_{\mathrm{Iter}} \}$ \\[0.1cm]
-{\tt SingleChi-} {\tt Theta- } {\tt Delta} & $\chi^2$ with systematics and correlations: Projection onto $\deltacp$-$\theta_{13}$-plane (one exp.) &  ($ \theta_{13}, \, \deltacp, \, \{ \theta_{12}, \theta_{23}, \sdm , \ldm,  \frac{\rho_{N_{\mathrm{exp}}}}{\bar{\rho}_{N_{\mathrm{exp}}}} \}, \, N_{\mathrm{exp}}$) \newline \ra\  $\{ \chi^2, \theta_{12},  \theta_{23},  \sdm , \ldm, \frac{\rho_{N_{\mathrm{exp}}}}{\bar{\rho}_{N_{\mathrm{exp}}}} , N_{\mathrm{Iter}} \}$ \\[0.1cm]
+{\tt Chi} & $\chi^2$ with systematics only \newline (all initialized exps.) & ($\{ \theta_{12}, \theta_{13}, \theta_{23}, \deltacp , \sdm , \ldm, \hat{\rho_1},...
+, \hat{\rho_n} \}$)  \newline \ra\  $\chi^2$ \\[0.1cm]
+{\tt SingleChi} & $\chi^2$ with systematics only \newline (only one experiment) & $(\{ \theta_{12}, \theta_{13}, \theta_{23}, \deltacp , \sdm , \ldm, \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}} )$   \newline \ra\ $\chi^2$ \\[0.1cm]
+{\tt ChiTheta} & $\chi^2$ with systematics and correlations: Projection onto $\theta_{13}$-axis (all exps.) &  ($ \theta_{13}, \, \{ \theta_{12}, \theta_{23}, \deltacp , \sdm , \ldm, \hat{\rho_1}, ...
+, \hat{\rho_n} \}$) \newline \ra\  $\{ \chi^2, \theta_{12}, \theta_{23}, \deltacp , \sdm , \ldm, \hat{\rho_1}, ...
+, \hat{\rho_n} , N_{\mathrm{Iter}} \}$ \\[0.1cm]
+{\tt Single-} \newline {\tt ChiTheta} & $\chi^2$ with systematics and correlations: Projection onto $\theta_{13}$-axis (one exp.) &  ($ \theta_{13}, \, \{ \theta_{12}, \theta_{23}, \deltacp , \sdm , \ldm,  \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}}$) \newline \ra\  $\{ \chi^2, \theta_{12}, \theta_{23}, \deltacp , \sdm , \ldm, \hat{\rho}_{N_{\mathrm{exp}}} , N_{\mathrm{Iter}} \}$ \\[0.1cm]
+{\tt ChiDelta} & $\chi^2$ with systematics and correlations: Projection onto $\deltacp$-axis (all exps.) &  ($\deltacp, \, \{ \theta_{12}, \theta_{13}, \theta_{23},  \sdm , \ldm, \hat{\rho_1}, ...
+, \hat{\rho_n} \}$) \newline \ra\  $\{ \chi^2, \theta_{12}, \theta_{13}, \theta_{23}, \sdm , \ldm, \hat{\rho_1}, ...
+, \hat{\rho_n} , N_{\mathrm{Iter}} \}$ \\[0.1cm]
+{\tt Single-} \newline {\tt ChiDelta} & $\chi^2$ with systematics and correlations: Projection onto $\deltacp$-axis (one exp.) &  ($ \deltacp, \, \{ \theta_{12}, \theta_{13}, \theta_{23}, \sdm , \ldm,  \hat{\rho}_{N_{\mathrm{exp}}}  \}, \, N_{\mathrm{exp}}$) \newline \ra\  $\{ \chi^2, \theta_{12}, \theta_{13}, \theta_{23},  \sdm , \ldm, \hat{\rho}_{N_{\mathrm{exp}}} , N_{\mathrm{Iter}} \}$ \\[0.1cm]
+{\tt ChiTheta-} {\tt Delta} & $\chi^2$ with systematics and correlations: Projection onto $\deltacp$-$\theta_{13}$-plane (all exps.) &  ($\theta_{13}, \, \deltacp, \, \{ \theta_{12}, \theta_{23},  \sdm , \ldm,  \hat{\rho_1}, ...
+, \hat{\rho_n} \}$) \newline \ra\  $\{ \chi^2, \theta_{12}, \theta_{23}, \sdm , \ldm,  \hat{\rho_1}, ...
+, \hat{\rho_n} , N_{\mathrm{Iter}} \}$ \\[0.1cm]
+{\tt SingleChi-} {\tt Theta- } {\tt Delta} & $\chi^2$ with systematics and correlations: Projection onto $\deltacp$-$\theta_{13}$-plane (one exp.) &  ($ \theta_{13}, \, \deltacp, \, \{ \theta_{12}, \theta_{23}, \sdm , \ldm,  \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}}$) \newline \ra\  $\{ \chi^2, \theta_{12},  \theta_{23},  \sdm , \ldm, \hat{\rho}_{N_{\mathrm{exp}}} , N_{\mathrm{Iter}} \}$ \\[0.1cm]
 {\tt ChiNP} & $\chi^2$ with systematics and correlations: Projection onto $N$-parameter hyper-plane (one exp.) & ??? to be defined \\[0.1cm]
@@
 {\tt ChiAll} & Lokal minimum of $\chi^2$ with respect to all parameters (all exps.) &
-($ \{ \theta_{13}, \theta_{12}, \theta_{23}, \deltacp , \sdm , \ldm, \frac{\rho_1}{\bar{\rho}_1}, ...
-, \frac{\rho_n}{\bar{\rho}_n} \}$) \newline \ra\  $\{ \chi^2, \theta_{13}, \theta_{12}, \theta_{23},$ \newline \hspace*{1.4cm} $ \deltacp , \sdm , \ldm, \frac{\rho_1}{\bar{\rho}_1}, ...
-, \frac{\rho_n}{\bar{\rho}_n} , N_{\mathrm{Iter}} \}$
+($ \{ \theta_{13}, \theta_{12}, \theta_{23}, \deltacp , \sdm , \ldm,  \hat{\rho_1}, ...
+, \hat{\rho_n} \}$) \newline \ra\  $\{ \chi^2, \theta_{13}, \theta_{12}, \theta_{23},$ \newline \hspace*{1.4cm} $ \deltacp , \sdm , \ldm,  \hat{\rho_1}, ...
+, \hat{\rho_n} , N_{\mathrm{Iter}} \}$
 \\[0.1cm]
-{\tt Single-} {\tt ChiAll} & Lokal minimum of $\chi^2$ with respect to all parameters (one exp.) &  ($ \{ \theta_{13}, \theta_{12}, \theta_{23}, \deltacp , \sdm , \ldm,  \frac{\rho_{N_{\mathrm{exp}}}}{\bar{\rho}_{N_{\mathrm{exp}}}} \}, \, N_{\mathrm{exp}}$) \newline \ra\  $\{ \chi^2, \theta_{13}, \theta_{12}, \theta_{23}, \deltacp , \sdm , \ldm, \frac{\rho_{N_{\mathrm{exp}}}}{\bar{\rho}_{N_{\mathrm{exp}}}} , N_{\mathrm{Iter}} \}$ \\[0.1cm]
+{\tt Single-} {\tt ChiAll} & Lokal minimum of $\chi^2$ with respect to all parameters (one exp.) &  ($ \{ \theta_{13}, \theta_{12}, \theta_{23}, \deltacp , \sdm , \ldm,  \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}}$) \newline \ra\  $\{ \chi^2, \theta_{13}, \theta_{12}, \theta_{23}, \deltacp , \sdm , \ldm, \hat{\rho}_{N_{\mathrm{exp}}} , N_{\mathrm{Iter}} \}$ \\[0.1cm]
 \hline
@@
 The curly brackets refer to the parameters to be transferred in form of a list.
-Note that all functions but {\tt Chi} and {\tt SingleChi} are using minimizers which have to be initialiued with {\tt SetInputErrors} and {\tt SetStartingValues} first.}
+The parameter $\hat{\rho}_i \equiv \rho_i/\bar{\rho}_i$ refers to the matter density scaling factor, which is $1.0$ for leaving the matter density profile unchanged.
+Note that all functions but {\tt Chi} and {\tt SingleChi} are using minimizers which have to be initialized with {\tt SetInputErrors} and {\tt SetStartingValues} first.}
 \end{table}
@@
 \begin{quote}
-{\tt ClearExp}$()$; \\
-{\tt InitExperiment}$($``{\tt JHFHK.exp}''$)$; \\
-{\tt SetRunningTime}$(0,2.0)$; \\
-{\tt InitExperiment}$($``{\tt JHFHKanti.exp}''$)$; \\
-{\tt SetRunningTime}$(1,6.0)$;\\
-{\tt MInfoExperiment}$()$; \\
-{\tt SetVacuumParameters}$(\{ 0.55, \, 0.16, \, 3.14/4, \, 3.14/2, \, 7e-5 , \, 2e-3 \} )$; \\
-{\tt SetRates();} 
+{\tt
+ClearExp(); \\
+InitExperiment("JHFHK.exp"); \\
+SetRunningTime(0, 2.0); \\
+InitExperiment("JHFHKanti.exp"); \\
+SetRunningTime(1, 6.0);\\
+MInfoExperiment(); \\
+SetVacuumParameters(\{0.55, 0.16, 3.14/4, 3.14/2, 7e-5, 2e-3\}); \\
+SetRates();
+} 
 \end{quote}
@@
 
-\bi
-\item
- More technical: what does it do (minimizer); how to imagine (simple example); 
-what functions are there; examples
-\item
- Single/multiple experiments and the connection (sum of individual $\chi$'s )
-\item
- Advanced topic: Changing the systematics treatment (SetSystematics)
-\ei
+Calculating a $\chi^2$-value with or without systematics, but no correlations and degeneracies, is the simplest and fastest possibility to obtain high-level information on an experiment.
+I general, \GLOBES\ uses the six independent oscillation parameters $\theta_{12}$, $\theta_{13}$, $\theta_{23}$, $\deltacp$, $\sdm$, $\ldm$, as well as the matter density of each experiment.
+Thus, there are six plus the number of experiments parameters determining the rate vectors.
+Using the matter densities in addition to the oscillation parameters will allow the simulation of matter density uncertainties: In this approach, the matter densities can be treated as parameters to be measured by the experiments within certain limits, where the limits are given by the observed precision of the matter density profile.
+Defining a independent matter density for each experiment is necessary if the baselines are completely uncorrelated.
+For correlated or even identical baselines, one may rather want to encapsulate the experiments into a single experiment definition.
 
-\example{Correlation between $\stheta$ and $\deltacp$}{Text}
+\example{Correlation between $\stheta$ and $\deltacp$}{
+\label{ex:corrth13dcp}
+
+A typical application for {\tt Chi} and {\tt SingleChi} is the visualization of two-parameter correlations.
+For example, to show the correlation between $\stheta$ and $\deltacp$ at a large neutrino factory with four years of running time for each polarity, one can use the following code:
+\begin{quote}
+{\tt
+/\% Experiment initialization: \%/ \\
+ClearExp(); \\
+ InitExperiment("{\tt NuFact2.exp}"); \\
+ SetRunningTime(0,4.0); \\
+ MInfoExperiment(); \\
+\\
+/\% Set reference rate vector: \%/ \\
+th12=arcsin(sqrt(0.8))/2; sdm=7e-5; \\
+th23=3.14/4; ldm=2e-3; \\
+th13=arcsin(sqrt(0.001))/2; dcp=3.14/2; \\
+\mbox{SetVacuumParameters(\{th12, \, th13, \, th23, \, dcp, \, sdm, \, ldm\}); }\\
+ SetRates(); \\
+\\
+/\% Compute chi-square matrix: \%/ \\
+for (float x=-2; x<-4; x=x+2/50) \\
+\hspace*{0.5cm} for (float y=0; y<200; y=y+200/50) \{ \\
+\hspace*{1cm} theth13 = arcsin(sqrt(10$\hat{\, \, \,}$x))/2; \\
+\hspace*{1cm} res[x,y]=Chi(\{th12, theth13,  th23,  y, sdm, ldm, 1.0\}) \};
+}
+\end{quote}
+
+The result {\tt res[x,y]} can then be plotted as a contour plot (2 d.o.f.):
+\begin{center}
+\includegraphics[width=8cm]{correx}
+\end{center}
+
+}
+
+Keeping all oscillation parameters and matter densities fixed, one can use the following functions to obtain the total $\chi^2$ of all specified oscillation channels including systematics:
+\begin{function} 
+{\tt Chi}$(\{ \theta_{12}, \theta_{13}, \theta_{23}, \deltacp , \sdm , \ldm, \hat{\rho_1}, \hdots , \hat{\rho_n} \})$  returns the total added $\chi^2$ of all loaded experiments.
+\end{function}
+\begin{function}
+ {\tt SingleChi}$(\{ \theta_{12}, \theta_{13}, \theta_{23}, \deltacp , \sdm , \ldm, \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}} )$  returns the total $\chi^2$ of the experiment number $N_{\mathrm{exp}}$.
+\end{function}
+Note that the result of {\tt Chi} corresponds to the sum of all of the {\tt SingleChi}'s of the loaded experiment.
+This equality will not hold for the minimizors in the next sections anymore.
+In both functions, one has to give the matter density scaling factor $\hat{\rho_i}$ for each of the used experiments.
+The effect of this scaling factor depends on the type of the matter density profile, which is given in the  experiment definition.
+For a constant matter density, it is simply the ratio of the matter density and the average matter density specified in the experiment definition, \ie , $\hat{\rho_i} \equiv \rho_i/\bar{\rho}_i$.
+For a matter density profile, it is an overall scaling factor: The matter density in each layer is multiplied by this factor.
+In most cases one wants to take a scaling factor of $1.0$ here, which simply means taking the matter density profile as it is given in the experiment definition.
+Moreover, note that its effect is in general small for short baselines.
+An example of how to use  {\tt Chi} (or {\tt SingleChi}) can be found on page~\pageref{ex:corrth13dcp}.
+
+The treatment of systematics is done by the usage of auxiliary systematics parameters, which are taken completely uncorrelated among different oscillation channels, and treated with simple Gaussian statistics.
+One such example is the signal normalization error, \ie, an error to the overall normalization of the signal.
+For illustration, we assume that the signal event rate in the $i$th bin $s_i^0$ of one oscillation channel is altered by the overall normalization auxiliary parameter of this channel, \ie ,
+\be
+ s_i = s_i(n_s) = s_i^0 \cdot (1 + n_s),
+\ee
+where $n_s$ is the signal normalization parameter.
+The total number of events in the $i$th bin $x_i$ also includes the background event rates $b_i$, \ie, $x_i = s_i + b_i$, which may have their own systematics parameters.
+In order to implement an overall signal normalization error $\sigma_{n_s}$,  the $\chi^2$, which includes all event rates $x_i$ of all bins, is minimized over the auxiliary parameter $n_s$:
+\be
+ \hat{\chi^2} = \underset{n_s}{\mathrm{min}} \left(  \chi^2(n_s, \hdots) + \frac{(n_s)^2}{\sigma_{n_s}^2} \right).
+\ee 
+This minimization is done independently for all auxiliary parameters of this oscillation channel.
+The total $\chi^2$ for the considered experiment is finally obtained by repreating this procedure for all oscillation channels and adding their $\chi^2$-values.
+In general, the situation is more complicated because of the usage of many systematical errors.
+More details about systematics parameters and the definition of signal, background, and oscillation channels can be found in the \EDM\ part of this book, too.
+
+The systematics minimization of an experiment can be easily switched on and off, \ie, one can also compute the $\chi^2$ without even taking into account systematics:
+\begin{function}
+{\tt SetSystematics}$(N_{mathrm{exp}}, \, sys)$ switches the systematics minimization for experiment $N_{\mathrm{exp}}$ on (sys=1) or off (sys=0).
+\end{function}
+This function can be especially useful for the test of the impact of systematics.
+??? Wie einzelne Systematik-Parameter ein und ausschalten? Sonst individuell nicht testbar! ??? Verweis auf appendix?
 

### 2003-12-03 Documentation/part1.tex
*** empty log message ***

@@
 /\% Compute chi-square matrix: \%/ \\
-for (float x=-2; x<-4; x=x+2/50) \\
+for (float x=-4; x<-2; x=x+2/50) \\
 \hspace*{0.5cm} for (float y=0; y<200; y=y+200/50) \{ \\
 \hspace*{1cm} theth13 = arcsin(sqrt(10$\hat{\, \, \,}$x))/2; \\
-\hspace*{1cm} res[x,y]=Chi(\{th12, theth13,  th23,  y, sdm, ldm, 1.0\}) \};
+\mbox{\hspace*{1cm} Print(\{x,y,Chi(\{th12, theth13,  th23,  y, sdm, ldm, 1.0\}) \}) \};}
 }
@@
 
-The result {\tt res[x,y]} can then be plotted as a contour plot (2 d.o.f.):
+The resulting matrix can then be plotted as a contour plot (2 d.o.f.):
 \begin{center}
-\includegraphics[width=8cm]{correx}
+\colorbox{white}{\includegraphics[width=8cm]{correx}}
 \end{center}
@@
 
-\chapter[Calculating $\chi^2$-projections: how to include correlations]{Calculating $\boldsymbol{\chi^2}$-projections: how to include correlations}
+\chapter[Calculating $\chi^2$-projections: how one can include correlations]{Calculating $\boldsymbol{\chi^2}$-projections: how one can include correlations}
 
-\section[Projection onto $\theta_{13}$ or $\delta$]{Projection onto $\boldsymbol{\theta_{13}}$ or $\boldsymbol{\delta}$}
+This chapter deals with the rather complicated issue of $n$-parameter correlations.
+Since it has before this software  not been possible to include the full $n$-parameter correlations in the high-dimensional parameter space with reasonable effort, it is the core part of this software  -- as well as its strength.
+Of course, calculating $\chi^2$-projections is somewhat more complicated than using systematics only.
+Therefore, we use a simple step by step introduction to the problem.
+
+\section{Introduction}
+
+In principle, the precision of an individual parameter measurement including correlations can be obtained as the projection of the $n$-dimensional fit manifold onto the respective axis.
+Similarly, one can project the fit manifold onto a plane, such as the $\stheta$-$\deltacp$-plane, if one wants to explicitely show this correlation with all the other parameter correlations included.
+In practice, this projection is very difficult: a grid-based method would need $(N_{\mathrm{grid}})^n$ function calls of {\tt Chi} or {\tt SingleChi} to calculate the precision including the full $n$-parameter correlation, where $N_{\mathrm{grid}}$ is the number of points in each direction of the lattice.
+For example, taking only $N_{\mathrm{grid}}=20$ and $n=7$ (six oscillation parameters and matter density) would mean more than one billion function calls of {\tt Chi} or {\tt SingleChi}.
+One can easily image that this is too much for any sophisticated application.
+
+The solution to this problem is using a local $n$-dimensional minimizer instead of a grid-based method.
+It turns out that such a minimizer can include a full $6$-parameter correlation with of the order of $1\, 000$ function calls of {\tt Chi} or {\tt ChiNew}.
+It is a standard method which can be found in every good book for standard numerical calculation routines.
+Thus, for each point on the projection axis/plane, one can obtain a result within about $10$ to $30$ seconds on a modern computer, which means that the complete measurement precision for one fixed true parameter set can be obtained in as much as $10$ to $15$ minutes.
+One can easily imagine that such a minimizer makes more sophisticated applications possible with the help of overnight calculations, such as showing the dependencies on the true parameter values.
+
+This approach also has a major disadvantage: One can not simply program a robust grid-based code and let it run, since using a local minimizer always means that one may end up in an unwanted local minimum and not in the investigated one.
+Thus, one has to use some (analytical or numerical) knowledge on the topology of the fit manifold and start the local minimizer close enough to the investigated solution.
+Fortunately, this can be done quite straightfoward in most cases, since the structure of the neutrino oscillation formulas does not cause very complicated topologies of the fit manifolds.
+Especially, the are plenty of analytical discussions of this issue, which means that one can implicitely use this knowledge to obtain better predictions for the measurement performances.
+Note that, since one can easily find the global fit minimum at the best-fit values, any solution found with the local minimizer makes the measurement performance worse.
+Thus, one can only run the danger to obtain a too optimistic solution if one does not find the other local minima below the chosen confidence level.
+
+In many cases, the fit manifold is restricted by the knowledge from earlier experiments.
+For example, the knowledge on the solar parameters will in most cases be supplied by the solar neutrino experiments.
+If the external precision of a parameter is at the time of the measurement better than the one of the experiment itself, one has to impose some external knowledge on this parameter.
+This external knowledge may reduce the $n$-dimensional fit manifold in the respective direction.
+In the most extreme case, keeping all parameters but the measured one fixed in the analysis means that all parameters are determined externally with infinitively high precisions.
+Thus, using the projection method on the axis/plane of interest is a reasonable approach.
+The inclusion of external input in \GLOBES\ is done by the use of Gaussian {\em priors}: We assume that an external measurement has determined the measured parameter to be at the central value (called {\em starting value}) with a $1 \sigma$ Gaussian error (called {\em input error}).
+The explicit definition of these priors will be shown in the next section.
+
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=16cm]{projex}
+\end{center}
+\mycaption{\label{fig:projex} Left plot: The correlation between $\stheta$ and $\deltacp$ as calculated in the example on page~\pageref{ex:corrth13dcp}, but for 1 d.o.f.
+only.
+Right plot: The $\chi^2$-value of the projection onto the $\stheta$-axis as function of $\stheta$.
+The projection onto the  $\stheta$-axis is obtained by finding the minimum $\chi^2$-value for each fixed value of $\stheta$ in the left-hand plot, \ie, along the gray vertical lines.
+The thick gray curve marks the position of these minima in the left-hand plot.
+The arrows mark the obtained fit ranges for $\stheta$ at the $3 \sigma$ confidence level (1 d.o.f.), \ie , the precision of $\stheta$.}
+\end{figure}
+
+\example{Projection of two- and $n$-dimensional manifold onto $\stheta$-axis}{
+\label{ex:corrproj}
+
+This example demonstrates how to use  {\tt ChiTheta} to project the fit manifold onto the $\stheta$-axis, \ie, how one can include correlations.
+We compute two sets of data: one for keeping all parameters but $\deltacp$ fixed (two-parameter correlations), and one for keeping all parameters free (multi-parameter correlation), but some external input for the solar parameters and $\rho$.
+Note that fixing all parameters but $\deltacp$ corresponds to imposing external knowledge on these parameters.
+
+\begin{quote}
+{\tt {\footnotesize
+/\% Experiment initialization, set reference rate vector: \%/ \\
+\mbox{ClearExp(); InitExperiment("{\tt NuFact2.exp}"); SetRunningTime(0,4.0); }\\
+th12=arcsin(sqrt(0.8))/2; sdm=7e-5; th23=3.14/4; ldm=2e-3; \\
+th13=arcsin(sqrt(0.001))/2; dcp=3.14/2; \\
+\mbox{SetVacuumParameters(\{th12, th13, th23, dcp, sdm, ldm\}); SetRates();}\\
+\\
+/\%  Compute chi-square list with all but dcp fixed: \%/ \\
+SetStartingValues(\{th12, th13, th23, dcp, sdm, ldm, 1.0\}); \\
+\mbox{SetInputErrors(\{th12*0.001, 10, th23*0.001, 10, sdm*0.001, ldm*0.001, 0.001\});} \\
+for (float x=-4; x<-2; x=x+2/20) \\
+\mbox{\hspace*{0.5cm} Print(\{x,ChiDelta(x,\{th12, th13, th23, dcp, sdm, ldm, 1.0\}) \}); }\\
+\\
+/\% Compute chi-square list with all parameters free: \%/ \\
+\mbox{/\% (but: 10\% prec. for solar params, 5\% for matter density) \%/} \\
+SetStartingValues(\{th12, th13, th23, dcp, sdm, ldm, 1.0\}); \\
+SetInputErrors(\{th12*0.1, 10, 10, 10, sdm*0.1, ldm, 0.05\}); \\
+for (float x=-4; x<-2; x=x+2/20) \\
+\mbox{\hspace*{0.5cm} Print(\{x,ChiDelta(x,\{th12, th13, th23, dcp, sdm, ldm, 1.0\}) \});}
+}}
+\end{quote}
+The two lists of data then represent the $\stheta$ precisions with two-parameter correlations (gray-shaded) and multi-parameter correlations (arrows):
+\begin{center}
+\colorbox{white}{\includegraphics[width=6cm]{projallex}}
+
+\vspace*{0.1cm}
+
+\footnotesize{(Same parameters as on page~\pageref{ex:corrth13dcp} and in \figu{projex}, but 1 d.o.f.)}
+\end{center}
+}
+
+\section{The treatment of external input}
 
@@
 \item
- Single and multiple experiments
+ Treatment of external input: short introduction to priors
 \item
- Figure: how does that work: local routine
+ The effect of the starting values and input errors
 \item
- Treatment of external input
+ Functions SetStartingValues, SetInputErrors to be called first
+\item
+ What for negative hierarchy? Defintion of $-\ldm - \sdm$?
 \ei
 
-\section[Projection onto any set of $N \ge 1$ parameters]{Projection onto any set of $\boldsymbol{N \ge 1}$ parameters}
+\section[Projection onto the $\theta_{13}$-axis or $\deltacp$-axis]{Projection onto the $\boldsymbol{\theta_{13}}$- or $\boldsymbol{\deltacp}$-axis}
+
+\bi
+\item
+ Call SetStartingValues and InputErrors first \ra\ three sets of parameters 
+\item
+ Single and multiple experiments: ChiDelta and ChiTheta + single funtions \ra\ return values;
+mention that $\chi^2$ does not add anymore; general parameter transfer concept
+\item
+ The effect of the starting point of the minimizer
+\item
+ What these functions do \ra\ figure
+\item
+ Refer to example: code
+\item
+ When does it not work: jumps \ra\ modify starting point accordingly
+\ei
+
+\section[Projection onto any hyperplane]{Projection onto any  hyperplane}
 
@@
  Single- and multiple experiment support
+\item 
+Example: Finding degenerate solutions with multiple experiments!?
 \item
@@
 
-\chapter{Changing experiment parameters at the running time}
+\chapter{Changing experiment parameters at running time}
 

### 2003-12-04 Documentation/part1.tex
*** empty log message ***

@@
 
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=16cm]{projex}
-\end{center}
-\mycaption{\label{fig:projex} Left plot: The correlation between $\stheta$ and $\deltacp$ as calculated in the example on page~\pageref{ex:corrth13dcp}, but for 1 d.o.f.
-only.
-Right plot: The $\chi^2$-value of the projection onto the $\stheta$-axis as function of $\stheta$.
-The projection onto the  $\stheta$-axis is obtained by finding the minimum $\chi^2$-value for each fixed value of $\stheta$ in the left-hand plot, \ie, along the gray vertical lines.
-The thick gray curve marks the position of these minima in the left-hand plot.
-The arrows mark the obtained fit ranges for $\stheta$ at the $3 \sigma$ confidence level (1 d.o.f.), \ie , the precision of $\stheta$.}
-\end{figure}
+\section{The treatment of external input}
+
+It is one of the strengths of the \GLOBES\ software to use external input to reduce the fit manifold with the knowledge from external (eariler) measurements.
+The treatment of external input is done by the addition of Gaussian {\em priors} to the final\footnote{After the systematics minimization and after all $\chi^2$'s are added for all channels} $\chi^2$-function  and the minimization over the respective parameters.
+For example, for the matter density, one obtains as the minimized $\chi^2_F$ after minimzation over $\rho$
+\be
+ \chi^2_F = \underset{\rho}{\mathrm{min}} \left( \chi^2(\rho) + \frac{(\rho - \rho^0)^2}{\sigma_\rho^2} \right).
+\label{equ:priors}
+\ee
+In practice, this minimzation is done simultaneously over all priors and free oscillation parameters, but this example serves as a very simple illustration.
+In \equ{priors}, $\rho^0$ is the {\em starting value} of the prior, and $\sigma_\rho$ the $1 \sigma$ absolute {\em input error}.
+Thus, it is assumed that an external measurement has determined the matter density with a precision (input error) $\sigma_\rho$ at the central value $\rho^0$.
+Usually, the starting value corresponds to the best-fit value and the input error to the $1 \sigma$ half width of the external measurement.
+For the matter density, $\rho^0$ can be set to the average matter density $\bar{\rho}$ and $\sigma_\rho$ to the matter density uncertainty.
+Since \GLOBES\ may also use a matter density profile, it actually uses a scaling factor $\hat\rho$ instead of the average matter density (see last section), \ie, $\rho^0=1.0$ is the central value in either case.
+Thus, $\sigma_\rho$ directly correponds to the uncertainty of this scaling factor.
+For instance, using the PREM (``Preliminary Reference Earth Model'') profile, $\rho^0 = 1.0$ and $\sigma_\rho = 0.05$ is a conservative estimate of the PREM profile uncertainty.
+
+In principle, one can set the priors for the matter density and all oscillation parameters.
+For example, if the disappearance channels of the experiment determine the leading oscillation parameters, once can set very large values for the input errors ($\sigma$'s).
+If, however, earlier external measurements provide better information, one can set their absolute precisions with the input errors.
+The starting values are usually set to the best-fit values.
+In some cases, it may be necessary to adjust them, such as for $\ldm$ and the negative mass hierarchy to $\rho^0_{\ldm} = - |\ldm|$ if some external precision on $| \ldm |$ is imposed and one is investigating the negative-sign solution.
+In other cases, minor modifications of the starting values can cause a faster convergence of the algorithm.
+In either case, two function have to be called {\em before the usage of any minimizer}:
+\begin{function}
+{\tt SetStartingValues}$(\{\rho^0_{\theta_{12}}, \, \rho^0_{\theta_{13}}, \, \rho^0_{\theta_{23}}, \, \rho^0_{\deltacp}, \, \rho^0_{\sdm}, \, \rho^0_{\ldm}, \, \rho^0_{\hat\rho} \})$ sets the starting values for all of the following minimizer calls.
+\end{function}
+\begin{function}
+{\tt SetInputErrors}$(\{\sigma_{\theta_{12}}, \, \sigma_{\theta_{13}}, \, \sigma_{\theta_{23}}, \, \sigma_{\deltacp}, \, \sigma_{\sdm}, \, \sigma_{\ldm}, \, \sigma_{\hat\rho} \})$ sets the input errors for all of the following minimizer calls.
+\end{function}
+Both functions may take one matter density parameter or as many as there are experiments.
+Since the assumptions about the matter density profile are usually similar for the loaded experiments, using only one matter density parameter is a useful tool to apply the same starting value  and input error to all loaded experiments.
+Eventually, a typical initialization of the external input may look like this:
+\begin{quote}
+{\tt
+SetStartingValues(\{th12, th13, th23, dcp, sdm, ldm, 1.0\}); \\
+SetInputErrors(\{th12*0.1, 10, 10, 10, sdm*0.1, ldm, 0.05\}); 
+}
+\end{quote}
+In this example, the starting values are set to the best-fit values.
+The input errors for $\theta_{13}$, $\theta_{23}$, $\deltacp$, and $\ldm$ are kept large, since the experiment measures these parameters itself.
+However, some precision on $\ldm$ is imposed to avoid ending up in the negative-sign solution.
+The solar parameters $\theta_{12}$ and $\sdm$ are assumed to be known with $10\%$ precision each.\footnote{In fact, accelerator-based long-baseline experiments are only sensitive to the product $\sin 2 \theta_{12} \cdot \sdm$, which means that these errors effectively add up to an error of this product.} Finally, the matter density uncertainties are assumed to have an amplitude of $5\% \,  \bar\rho$.
 
@@
 
-\section{The treatment of external input}
+\section[Projection onto the $\stheta$-axis or $\deltacp$-axis]{Projection onto the $\boldsymbol{\stheta}$- or $\boldsymbol{\deltacp}$-axis}
 
-\bi
-\item
- Treatment of external input: short introduction to priors
-\item
- The effect of the starting values and input errors
-\item
- Functions SetStartingValues, SetInputErrors to be called first
-\item
- What for negative hierarchy? Defintion of $-\ldm - \sdm$?
-\ei
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=16cm]{projex}
+\end{center}
+\mycaption{\label{fig:projex} Left plot: The correlation between $\stheta$ and $\deltacp$ as calculated in the example on page~\pageref{ex:corrth13dcp}, but for 1 d.o.f.
+only.
+Right plot: The $\chi^2$-value of the projection onto the $\stheta$-axis as function of $\stheta$.
+The projection onto the  $\stheta$-axis is obtained by finding the minimum $\chi^2$-value for each fixed value of $\stheta$ in the left-hand plot, \ie, along the gray vertical lines.
+The thick gray curve marks the position of these minima in the left-hand plot.
+The arrows mark the obtained fit ranges for $\stheta$ at the $3 \sigma$ confidence level (1 d.o.f.), \ie , the precision of $\stheta$.}
+\end{figure}
 
-\section[Projection onto the $\theta_{13}$-axis or $\deltacp$-axis]{Projection onto the $\boldsymbol{\theta_{13}}$- or $\boldsymbol{\deltacp}$-axis}
+The projection onto the $\stheta$- (or $\deltacp$-) axis is performed by fixing $\stheta$ (or $\deltacp$) and minimizing the $\chi^2$-function over all free fit parameters and the matter densities.
+We illustrate this method at the example of the projection of the two-dimensional manifold in the $\stheta$-$\deltacp$-plane onto the $\stheta$-axis in \figu{projex}.
+In this figure, the left-hand plot shows the correlation in the $\stheta$-$\deltacp$-plane computed with {\tt Chi} or {\tt SingleChi}.
+The right-hand plot shows the projection of this two-dimensional manifold onto the $\stheta$ axis by minimizing $\chi^2$ over $\deltacp$.
+In this simple example, the minimization is done along the vertical gray lines in the left hand plot.
+The obtained minima are located on the thick gray curve, which means the the right-hand plot represents the $\chi^2$-value along this curve.
+In fact, one can easily see that one obtains the correct projected $3 \sigma$ errors in this example (\cf, arrows).
+This figure illustrates the projection of a two-parameter correlation.
+In general, the full $n$-parameter correlation is treated similarly by the simultaneous (local) minimization over all free fit parameters.
+The example on page~\pageref{ex:corrproj} demonstrates how one can obtain \figu{projex} (right) with keeping all parameters but $\deltacp$ fixed, as well as how one can inlcude the full $n$-parameter correlation with external input.
+It also demonstrates how these two compare to each other.
 
-\bi
-\item
- Call SetStartingValues and InputErrors first \ra\ three sets of parameters 
-\item
- Single and multiple experiments: ChiDelta and ChiTheta + single funtions \ra\ return values;
-mention that $\chi^2$ does not add anymore; general parameter transfer concept
-\item
- The effect of the starting point of the minimizer
-\item
- What these functions do \ra\ figure
-\item
- Refer to example: code
-\item
- When does it not work: jumps \ra\ modify starting point accordingly
-\ei
+One mentioned disadvantage of the local minimizer is the possibility to end up in a local minimum.
+Usually it works very well to start the minimizer close to the best-fit values.
+In the example in \figu{projex}, the minimizer is always started at $\deltacp=90^\circ$ and finds its way to the appropriate minimum.
+In some cases, however, the topology may be more difficult and the true minimum is not correctly found.
+A strong indication for such a situation are discontinuities in the projected $\chi^2$-function, where the minimizer jumos from one minimum to the other.
+In such a case, the starting point of the minimizer has to be adjusted to help it find the true minimum.
+In other cases, the convergence of the minimizer may be infinitively small.
+Then it usually helps to start the minimizer somewhat off the best-fit values.
+The following functions are the simplest minimizers provided by \GLOBES :
+\begin{function}
+{\tt ChiTheta}$( \theta_{13}, \, \{ \theta_{12}, \, \theta_{23}, \, \deltacp , \, \sdm , \, \ldm, \, \hat{\rho_1}, \,  \hdots ,  \,\hat{\rho_n} \})$ returns the projection onto the $\theta_{13}$-axis for all experiments.
+The parameters are the fixed value of $\theta_{13}$ and the starting point of the minimizer including all starting values for the matter density scaling factors (without $\theta_{13}$, which is fixed).
+The return value is a list $\{ \chi^2, \, \theta_{12}, \, \theta_{23}, \, \deltacp , \, \sdm , \, \ldm, \, \hat{\rho_1}, \, \hdots , \, \hat{\rho_n} , \, N_{\mathrm{Iter}} \}$ with the minimum $\chi^2$ found, the coordinates of the local minimum (without $\theta_{13}$), and the number of iterations $N_{\mathrm{Iter}}$ used by the minimizer (number of function calls of {\tt Chi}).
+\end{function}
+\begin{function}
+{\tt SingleChiTheta}$( \theta_{13}, \, \{ \theta_{12}, \, \theta_{23}, \, \deltacp , \, \sdm , \, \ldm,  \, \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}})$ works similar to {\tt ChiTheta}, but only for one of the loaded experiments $N_{\mathrm{exp}}$.
+It returns the list $\{ \chi^2, \, \theta_{12}, \,  \theta_{23}, \, \deltacp , \, \sdm , \, \ldm, \, \hat{\rho}_{N_{\mathrm{exp}}} , \, N_{\mathrm{Iter}} \}$.
+\end{function}
+\begin{function}
+{\tt ChiDelta}$( \deltacp, \, \{ \theta_{12}, \, \theta_{13}, \, \theta_{23}, \, \sdm , \, \ldm, \, \hat{\rho_1}, \,  \hdots ,  \,\hat{\rho_n} \})$ returns the projection onto the $\deltacp$-axis for all experiments.
+The parameters are the fixed value of $\deltacp$ and the starting point of the minimizer including all starting values for the matter density scaling factors (without $\deltacp$, which is fixed).
+The return value is a list $\{ \chi^2, \, \theta_{12}, \, \theta_{13}, \, \theta_{23}, \, \sdm , \, \ldm, \, \hat{\rho_1}, \, \hdots , \, \hat{\rho_n} , \, N_{\mathrm{Iter}} \}$ with the minimum $\chi^2$ found, the coordinates of the local minimum (without $\deltacp$), and the number of iterations $N_{\mathrm{Iter}}$ used by the minimizer (number of function calls of {\tt Chi}).
+\end{function}
+\begin{function}
+{\tt SingleChiDelta}$( \deltacp, \, \{ \theta_{12}, \, \theta_{13}, \, \theta_{23}, \, \sdm , \, \ldm, \,  \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}})$ works similar to {\tt ChiDelta}, but only for one of the loaded experiments $N_{\mathrm{exp}}$.
+It returns the list $\{ \chi^2, \, \theta_{12}, \,  \theta_{13}, \,  \theta_{23}, \, \sdm , \, \ldm, \, \hat{\rho}_{N_{\mathrm{exp}}} , \, N_{\mathrm{Iter}} \}$.
+\end{function}
+All of these function have the same parameter structure: The fixed parameter is not transferred in the list, but as a the first separate parameter.
+The number of matter density scaling factors corresponds to the number of experiments used, since each experiment may face other matter density conditions.
+Note that before any of these function calls, {\tt SetStartingValues} and {\tt SetInputErrors} have to be used at least once.
+In addition, note that the resulting $\chi^2$ of {\tt ChiTheta} (or {\tt ChiDelta}) is not the sum of the $\chi^2$-values over all {\tt Single}-functions of all experiments anymore.
+This has two reasons: First, the topology of the fit manifold is altered by the addition of $\chi^2$-values of different experiments.
+Thus, after the minimization, the position of the minimum can be different to the ones of the individual experiments.
+Second, the priors for the external knowledge on the parameters are only added once -- independent of the number of experiments.
+A simple application of {\tt ChiTheta} can be found in the example on page~\pageref{ex:corrproj}.
+
+The return values of these functions do not only contain the minimum $\chi^2$-value found, but also the position of the minimum.
+This information can often be valuable, since one can often quickly locate irregularities: If one of the parameters is far off the best-fit value, the minimzer may have ended up in a local unwanted minimum.
+For instance, a negative value of $\ldm$ (with a positive best-fit value) indicates that the minimizer ended up in the opposite-sign solution.
+Another possibility is that the experiment may not be able to determine the measured parameter at all without external knowledge about the parameter far off its best-fit value.
+In both cases, an externally imposed precision on the run-off parameter can help to solve the problem.
 
@@
 
-General form and especially: DeltaTheta-Minimizer; mention calculation time!
+In general, one can show the measurement result in any $k$-dimensional hyperplane, where $k$ is smaller than the dimension of the parameter space $n$, and thus the dimension of the fit manifold.
+In this case, $k$ parameters are fixed and $n-k$ parameters are minimized over.
+One such example is the projection of the fit manifold onto the $\stheta$-$\deltacp$-plane, \ie, $k=2$ here.
+This projection can be performed with the following functions:
+\begin{function}
+{\tt ChiThetaDelta}$(\theta_{13}, \, \deltacp, \, \{ \theta_{12}, \, \theta_{23},  \, \sdm , \, \ldm, \,   \hat{\rho_1}, ...
+, \hat{\rho_n} \})$ returns the list  $\{ \chi^2, \, \theta_{12}, \, \theta_{23}, \, \sdm , \, \ldm,  \, \hat{\rho_1}, \, \hdots , \, \hat{\rho_n} , \, N_{\mathrm{Iter}} \}$ for the projection into the $\theta_{13}$-$\deltacp$-plane for all experiments.
+\end{function}
+\begin{function}
+{\tt SingleChiThetaDelta}$( \theta_{13}, \, \deltacp, \, \{ \theta_{12}, \, \theta_{23}, \, \sdm , \, \ldm, \,   \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}})$ returns the list $\{ \chi^2, \, \theta_{12}, \,   \theta_{23},  \, \sdm , \, \ldm, \, \hat{\rho}_{N_{\mathrm{exp}}} ,  \, N_{\mathrm{Iter}} \}$ for the projection into the $\theta_{13}$-$\deltacp$-plane for one experiment $N_{\mathrm{exp}}$.
+\end{function}
+These functions work analogously to the ones in the last section.
+They can, for example, be used to obtain a figure similar to \figu{projex}, left.
+However, the result would not be a two-dimensional cut through the fit-manifold, but a two-dimensional projection.
+Though the running time for one call of these functions is somewhat shorter than the one for the $\stheta$- or $\deltacp$-projections, one has to compute a two-dimensional array for such a figure (instead of a one-dimensional list).
+Therefore, the overall computational effort is much higher.
+
+In principle, one can also use three- or more-dimensional projections.
+In addition, one may want to use any other set of parameters for single- or two-parameter projections.
+The functions {\tt ChiNP} and {\tt SingleChiNP} are designed for this purpose:
+
+FUNCTIONS TO BE DEFINED. (WORK IN PROGRESS).
 

### 2003-12-04 Documentation/part1.tex
*** empty log message ***

@@
 
+\example{Finding the $\mathrm{sgn}(\ldm)$-degeneracy}{
+\label{ex:sgndeg}
+
+In many cases, one can find the $\mathrm{sgn}(\ldm)$-degeneracy with {\tt ChiAll}: One lets the local minimizer start at the position of the suspected solution at the opposite sign of $\ldm$.
+The following example corresponds to finding the degenerate solution for the example on page~\pageref{ex:corrth13dcp}.
+\begin{quote}
+{\tt
+{\footnotesize
+/\% Experiment initialization: \%/ \\
+ClearExp(); InitExperiment("{\tt NuFact2.exp}"); SetRunningTime(0,4.0); \\
+\\
+/\% Set reference rate vector: \%/ \\
+th12=arcsin(sqrt(0.8))/2; sdm=7e-5; th23=3.14/4; ldm=2e-3; \\
+th13=arcsin(sqrt(0.001))/2; dcp=3.14/2; \\
+\mbox{SetVacuumParameters(\{th12, \, th13, \, th23, \, dcp, \, sdm, \, ldm\});  SetRates();} \\
+\\
+/\% Find negative-sign solution: \%/ \\
+SetStartingValues(\{th12, \, th13, \, th23, \, dcp, \, sdm, \, -ldm, \, 1.0\}); \\
+SetInputErrors(\{th12*0.1, \, 10, \, 10, \, 10, \, sdm*0.1, \, ldm/3, \, 0.05\}); \\
+d = ChiAll(\{th12, \, th13, \, th23, \, dcp, \, sdm, \, ldm, \, 1.0\}); \\
+\\
+/\% Compute chi-square matrix: \%/ \\
+for (float x=-4; x<-2; x=x+2/50) \\
+\hspace*{0.5cm} for (float y=0; y<200; y=y+200/50) \{ \\
+\hspace*{1cm} theth13 = arcsin(sqrt(10$\hat{\, \, \,}$x))/2; \\
+\mbox{\hspace*{1cm} Print(\{x,y,Chi(\{d[2], theth13,  d[4],  y, d[6], d[7], d[8]\}) \}) \};}
+}
+}
+\end{quote}
+
+The resulting matrix can then be plotted as a contour plot in addition to the original solution (2 d.o.f., gray contours):
+\begin{center}
+\colorbox{white}{\includegraphics[width=7.5cm]{correntex}}
+\end{center}
+
+}
 \bi

### 2003-12-05 Documentation/part1.tex
*** empty log message ***

@@
 \end{center}
-\caption{\label{tab:stdfunctions} The \GLOBES\ standard function to obtain a $\chi^2$-value for all or one of the initialized experiments.
+\caption{\label{tab:stdfunctions} \index{Standard functions (table)} The \GLOBES\ standard function to obtain a $\chi^2$-value for all or one of the initialized experiments.
 The curly brackets refer to the parameters to be transferred in form of a list.
@@
 \example{Using \GLOBES\ with C}{\label{ex:c}
+\index{C-Code}
 
@@
 \end{tabular}
-\mycaption{\label{tab:units} Quantities used in \GLOBES , examples of these quantities, and their standard units in the application software.}
+\mycaption{\label{tab:units} \index{Units in \GLOBES } Quantities used in \GLOBES , examples of these quantities, and their standard units in the application software.}
 \end{center}
@@
 \end{center}
-\mycaption{\label{tab:experiments} Different pre-defined experiments, their filenames (to be used in {\tt LoadExperiment}), their short description, and the references in which they are defined.
+\mycaption{\label{tab:experiments} \index{Experiment data files (table)} Different pre-defined experiments, their filenames (to be used in {\tt LoadExperiment}), their short description, and the references in which they are defined.
 Note that all experiments use in their standard configurations one year of running time.
@@
 \begin{function}
+\index{{\tt LoadExperiment}}
+\index{Experiment initialization}
 {\tt LoadExperiment}$($``filename'' $)$ adds a single experiment to the list of currently loaded experiments.
@@
 \begin{function}
+\index{{\tt ClearExp}}
 {\tt ClearExp}$()$ removes all experiments from the evaluation list.   
@@
 \begin{function}
+\index{{\tt SetRunningTime}}
 {\tt SetRunningTime}$(N_{\mathrm{exp}},t_{\mathrm{run}})$ sets the running time of experiment number $N_{\mathrm{exp}}$ to $t_{\mathrm{run}}$ years.
@@
  \begin{function}
+\index{{\tt SetTargetMass}}
 {\tt SetTargetMass}$(N_{\mathrm{exp}},m_{\mathrm{Det}})$ sets the fiducial mass of experiment number $N_{\mathrm{exp}}$ to $m_{\mathrm{Det}}$ kilotons.
@@
 \begin{function}
+\index{{\tt SetSourcePower}}
 {\tt SetSourcePower}$(N_{\mathrm{exp}},P_{\mathrm{Source}})$ sets the source power of experiment number $N_{\mathrm{exp}}$ to $P_{\mathrm{source}}$.
@@
 \begin{function}
+\index{{\tt InfoExperiment}}
 {\tt InfoExperiment}$()$ prints a list of the initialized functions with their experiment numbers and their most important parameters to the standard output.
@@
 \begin{function}
+\index{{\tt ShowCannels}}
 {\tt ShowChannels}$(N_{\mathrm{exp}})$ prints the information about the oscillation channels of the experiment with the number $N_{\mathrm{exp}}$ to the standard output.
@@
 \begin{function}
+\index{{\tt SetVacuumParameters}}
 {\tt SetVacuumParameters}$(\{\theta_{12}, \theta_{13}, \theta_{23}, \deltacp , \sdm , \ldm \})$ sets the neutrino oscillation parameters to be used to compute the reference rate vector in vacuum.
@@
 \begin{function}
+\index{{\tt SetRates}}
 {\tt SetRates}$( )$ computes the reference rate vector for the neutrino oscillation parameters set with {\tt SetVacuumParameters}. 
@@
 
+\index{Systematics}
 Calculating a $\chi^2$-value with or without systematics, but no correlations and degeneracies, is the simplest and fastest possibility to obtain high-level information on an experiment.
@@
 
+\index{Two-parameter correlation}
 A typical application for {\tt Chi} and {\tt SingleChi} is the visualization of two-parameter correlations.
@@
 \begin{function} 
+\index{{\tt Chi}}
 {\tt Chi}$(\{ \theta_{12}, \theta_{13}, \theta_{23}, \deltacp , \sdm , \ldm, \hat{\rho_1}, \hdots , \hat{\rho_n} \})$  returns the total added $\chi^2$ of all loaded experiments.
@@
 \begin{function}
+\index{{\tt SingleChi}}
  {\tt SingleChi}$(\{ \theta_{12}, \theta_{13}, \theta_{23}, \deltacp , \sdm , \ldm, \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}} )$  returns the total $\chi^2$ of the experiment number $N_{\mathrm{exp}}$.
@@
 The effect of this scaling factor depends on the type of the matter density profile, which is given in the  experiment definition.
-For a constant matter density, it is simply the ratio of the matter density and the average matter density specified in the experiment definition, \ie , $\hat{\rho_i} \equiv \rho_i/\bar{\rho}_i$.
-For a matter density profile, it is an overall scaling factor: The matter density in each layer is multiplied by this factor.
+For a constant matter density, it is simply the ratio of the matter density and the average matter density specified in the experiment definition, \ie , $\hat{\rho_i} \equiv \rho_i/\bar{\rho}_i$.\index{Matter scaling factor} For a matter density profile, it is an overall scaling factor: The matter density in each layer is multiplied by this factor.
 In most cases one wants to take a scaling factor of $1.0$ here, which simply means taking the matter density profile as it is given in the experiment definition.
@@
 
+\index{Systematics}
 The treatment of systematics is done by the usage of auxiliary systematics parameters, which are taken completely uncorrelated among different oscillation channels, and treated with simple Gaussian statistics.
 One such example is the signal normalization error, \ie, an error to the overall normalization of the signal.
-For illustration, we assume that the signal event rate in the $i$th bin $s_i^0$ of one oscillation channel is altered by the overall normalization auxiliary parameter of this channel, \ie ,
+For illustration, we assume that the signal event rate in the $i$th bin $s_i^0$ of one oscillation channel is altered by the overall normalization auxiliary parameter\index{Auxiliary parameter} of this channel, \ie ,
 \be
@@
 \begin{function}
+\index{{\tt SetSystematics}}
 {\tt SetSystematics}$(N_{mathrm{exp}}, \, sys)$ switches the systematics minimization for experiment $N_{\mathrm{exp}}$ on (sys=1) or off (sys=0).
@@
 
+\index{Multi-parameter correlation}
 This chapter deals with the rather complicated issue of $n$-parameter correlations.
@@
 
+\index{Projection of manifold}
 In principle, the precision of an individual parameter measurement including correlations can be obtained as the projection of the $n$-dimensional fit manifold onto the respective axis.
@@
 
+\index{Minimizer}
 The solution to this problem is using a local $n$-dimensional minimizer instead of a grid-based method.
@@
 
+\index{Priors} \index{Input errors} \index{Starting values}
 In many cases, the fit manifold is restricted by the knowledge from earlier experiments.
@@
 
+\index{External input}
 It is one of the strengths of the \GLOBES\ software to use external input to reduce the fit manifold with the knowledge from external (eariler) measurements.
@@
 For instance, using the PREM (``Preliminary Reference Earth Model'') profile, $\rho^0 = 1.0$ and $\sigma_\rho = 0.05$ is a conservative estimate of the PREM profile uncertainty.
+\index{PREM profile}
 
@@
 \begin{function}
+\index{{\tt SetStartingValues}}
 {\tt SetStartingValues}$(\{\rho^0_{\theta_{12}}, \, \rho^0_{\theta_{13}}, \, \rho^0_{\theta_{23}}, \, \rho^0_{\deltacp}, \, \rho^0_{\sdm}, \, \rho^0_{\ldm}, \, \rho^0_{\hat\rho} \})$ sets the starting values for all of the following minimizer calls.
@@
 \begin{function}
+\index{{\tt SetInputErrors}}
 {\tt SetInputErrors}$(\{\sigma_{\theta_{12}}, \, \sigma_{\theta_{13}}, \, \sigma_{\theta_{23}}, \, \sigma_{\deltacp}, \, \sigma_{\sdm}, \, \sigma_{\ldm}, \, \sigma_{\hat\rho} \})$ sets the input errors for all of the following minimizer calls.
@@
 \label{ex:corrproj}
+\index{Projection of manifold}
+\index{Two-parameter correlation}
+\index{Multi-parameter correlation}
 
@@
 \section[Projection onto the $\stheta$-axis or $\deltacp$-axis]{Projection onto the $\boldsymbol{\stheta}$- or $\boldsymbol{\deltacp}$-axis}
+\index{Projection onto axis}
 
@@
 In some cases, however, the topology may be more difficult and the true minimum is not correctly found.
-A strong indication for such a situation are discontinuities in the projected $\chi^2$-function, where the minimizer jumos from one minimum to the other.
+A strong indication for such a situation are discontinuities in the projected $\chi^2$-function, where the minimizer jumps from one minimum to the other.
 In such a case, the starting point of the minimizer has to be adjusted to help it find the true minimum.
@@
 \begin{function}
+\index{{\tt ChiTheta}}
 {\tt ChiTheta}$( \theta_{13}, \, \{ \theta_{12}, \, \theta_{23}, \, \deltacp , \, \sdm , \, \ldm, \, \hat{\rho_1}, \,  \hdots ,  \,\hat{\rho_n} \})$ returns the projection onto the $\theta_{13}$-axis for all experiments.
@@
 \begin{function}
+\index{{\tt SingleChiTheta}}
 {\tt SingleChiTheta}$( \theta_{13}, \, \{ \theta_{12}, \, \theta_{23}, \, \deltacp , \, \sdm , \, \ldm,  \, \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}})$ works similar to {\tt ChiTheta}, but only for one of the loaded experiments $N_{\mathrm{exp}}$.
@@
 \begin{function}
+\index{{\tt ChiDelta}}
 {\tt ChiDelta}$( \deltacp, \, \{ \theta_{12}, \, \theta_{13}, \, \theta_{23}, \, \sdm , \, \ldm, \, \hat{\rho_1}, \,  \hdots ,  \,\hat{\rho_n} \})$ returns the projection onto the $\deltacp$-axis for all experiments.
@@
 \begin{function}
+\index{{\tt SingleChiDelta}}
 {\tt SingleChiDelta}$( \deltacp, \, \{ \theta_{12}, \, \theta_{13}, \, \theta_{23}, \, \sdm , \, \ldm, \,  \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}})$ works similar to {\tt ChiDelta}, but only for one of the loaded experiments $N_{\mathrm{exp}}$.
@@
 \section[Projection onto any hyperplane]{Projection onto any  hyperplane}
+\index{Projection onto hyperplane}
 
@@
 \begin{function}
+\index{{\tt ChiThetaDelta}}
 {\tt ChiThetaDelta}$(\theta_{13}, \, \deltacp, \, \{ \theta_{12}, \, \theta_{23},  \, \sdm , \, \ldm, \,   \hat{\rho_1}, ...
@@
 \begin{function}
+\index{{\tt SingleChiThetaDelta}}
 {\tt SingleChiThetaDelta}$( \theta_{13}, \, \deltacp, \, \{ \theta_{12}, \, \theta_{23}, \, \sdm , \, \ldm, \,   \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}})$ returns the list $\{ \chi^2, \, \theta_{12}, \,   \theta_{23},  \, \sdm , \, \ldm, \, \hat{\rho}_{N_{\mathrm{exp}}} ,  \, N_{\mathrm{Iter}} \}$ for the projection into the $\theta_{13}$-$\deltacp$-plane for one experiment $N_{\mathrm{exp}}$.
@@
 \chapter{Finding degenerate solutions}
+\index{Degenerate solutions}
+
+In the last chapter, we introduced the projection of any set of $k$ parameter onto any $n-k$ dimensional hyperplane, which was done by the minimization over the $k$ free fit parameters.
+Similarly, one can minimize over {\em all} $n$ parameters to find the local minimum close to any starting point.
+This approach is very useful for the exact numerical location of a degenerate if its approximate position is known.
+For the determination of the approximate positon, one can use analytical approaches, such as in \Refs~TO BE ENTERED, or an educated guess.
+Though the usage of the all-parameter minimizers is quite simple, one should keep in mind that they are local minimizers.
+Therefore, one may need a very sophisticated application software to actually find all degenerate solutions.
 
@@
 }
-\bi
-\item
- Single- and multiple experiment support
-\item 
-Example: Finding degenerate solutions with multiple experiments!?
-\item
- What does that do?
-\item
- Example: how to treat sign-degeneracy in th13-limit
-\ei
+
+\index{All-parameter minimization}
+The functions to perform the all-parameter minimization are {\tt ChiAll} and {\tt SingleChiAll}:
+\begin{function}
+\index{{\tt ChiAll}}
+{\tt ChiAll}$(\{ \theta_{13}, \, \theta_{12}, \, \theta_{23}, \, \deltacp , \, \sdm , \, \ldm,  \, \hat{\rho_1}, \, \hdots , \, \hat{\rho_n} \})$ finds the local minimum close to the given starting point (list) for all initialized experiments.
+It returns the list  $\{ \chi^2, \, \theta_{13}, \, \theta_{12}, \, \theta_{23}, \, \deltacp , \, \sdm , \,  \ldm,  \, \hat{\rho_1}, \, \hdots , \hat{\rho_n} , \, N_{\mathrm{Iter}} \}$ with the minimum $\chi^2$-value, the actual position of the local minimum, and the number of iterations $N_{\mathrm{Iter}}$ used.
+\end{function}
+\begin{function}
+\index{{\tt SingleChiAll}} 
+{\tt SingleChiAll}$( \{ \theta_{13}, \, \theta_{12}, \, \theta_{23}, \, \deltacp , \, \sdm , \, \ldm, \,   \hat{\rho}_{N_{\mathrm{exp}}} \}, \, \, N_{\mathrm{exp}})$ finds the local minimum close to the given starting point (list) for the experiment $N_{\mathrm{exp}}$ only.
+It returns the list
+ $\{ \chi^2, \, \theta_{13}, \, \theta_{12}, \, \theta_{23}, \, \deltacp , \, \sdm , \, \ldm, \, \hat{\rho}_{N_{\mathrm{exp}}} ,  \, N_{\mathrm{Iter}} \}$ with the minimum $\chi^2$-value, the actual position of the local minimum, and the number of iterations $N_{\mathrm{Iter}}$ used.
+\end{function}
+%
+Both functions take the suspected position of the local minimum and return its true position.
+The application of  {\tt SingleChiAll} can be especially usefull for many different experiments evaluated simultaneously: It turns out to be useful to find the positions of the degeneracies for the individual experiments first, and then test all of these with the combination of all experiments in order not to miss a degenerate solution.
+The example on page~\pageref{ex:sgndeg} illustrates how to locate the $\mathrm{sgn}(\ldm)$-degeneracy and show the corresponding degenerate solution in the $\stheta$-$\deltacp$-plane together with the original solution.
+In this case, the position of the degeneracy can be easily guessed to be at the best-fit parameter values but the $\ldm$ inverted.
+The minimizer then runs off the plane of the best-fit parameters into the local minimum.
+It is very important to take into account the position of the degeneracy off this plane, since the actual $\chi^2$ in the minimum is certainly lower than on the plane of the best-fit parameter values.
+Thus, the degeneracy may not even appear at the chosen confidence level on the plane, but it does appear at the true minimum.
+The two cuts\footnote{The discussed figure on page~\pageref{ex:sgndeg} is produced by  {\tt Chi} and thus only represents a cut through the fit manifold.
+For the projection including correlations, one may rather want to use {\tt ChiThetaDelta}.} through the fit manifold shown in the figure on page~\pageref{ex:sgndeg} therefore do not have the same oscillation parameter values (except from the ones shown in the figure).
+
+In practice, a number of tricks can be useful for the treatment of degenerate solutions:
+\begin{description}
+\item[Minimum $\boldsymbol{\chi^2}$ larger than threshold.] If a located degeneracy has a minimum $\chi^2$ larger than the corresponding confidence level threshold for the discussed quantity of interest, the degeneracy can be immediately ignored.
+This saves a lot of computation time.
+\item[Locating degeneracies with more complicated topologies.] For more complicated topologies, such as for neutrino factories, it is often useful to use multi-step location procedures or analytical knowledge.
+For example, for a numerical procedure, one may first of all switch off the systematics and keep $\stheta$ or $\deltacp$ fixed, \ie, use {\tt SingleChiTheta} or {\tt SingleChiDelta}, where $\stheta$ or $\deltacp$ is fixed to the best-fit value.
+The result can then be used as a starting point for {\tt SingleChiAll} with the systematics switched on again.
+In addition to switching off the systematics, it can be useful to reduce the input errors during the first steps in order to make the minimizer not to run away too much from the true solution.
+\item[Finding degeneracies with multiple experiments.] For multiple experiments, the degeneracies should be located for each of the individual experiments first.
+Then, all of the found degeneracies below the threshold can be tested for the combination of experiments.
+\end{description}
+Finally, note that any degenerate solution below the confidence level threshold which can not be located makes the result appear better than it actually is.
+Thus, one should be careful with the determination of the degenerate solutions.
 

### 2004-04-21 Documentation/part2.tex
part of the edm documentation

@@
 
-Here comes how a experiment is defined (what is a channel, a rule, how are they treated, ...). Splitting in source, oscillation, and detector, etc.
+Here comes how a experiment is defined (what is a channel, a rule, how are 
+they treated, ...). Splitting in source, oscillation, and detector, etc.
 
@@
 Here comes how to load and modify existing experiments;
-also: how does the main screen work and the most important functions at exampled
+also: how does the main screen work and the most important functions at 
+exampled
 
@@
 
+\chapter{General concept}
 \section{Source properties}
+
+\GLOBES\ handles only point sources, thus it is not possible
+to study effects from the finit size of the neutrino production
+region like in the Sun. Furthermore only one source for each experiment
+is possible, this make it impossible\footnote{at least rather tricky} 
+to simulate multiple source experiments like KamLAND. All neutrino
+sources useable within \GLOBES\ therefore can be specified by the
+flux spectra for each flavour and CP sign and the total luminosity
+of the source. There are two principal ways to provide a flux:
+use a built-in source or a file provided by the user.
+
+The built-in source is muon decay, since the neutrino
+spectra and flavour composition are knon analytically in that case. For 
+this sources the energy of the parent particle has to be specified
+and the number of useful decays per year.
+
+In the latter case the fluxes are read from a file with seven rows and
+501 lines.
+\begin{quotation}
+$ E\quad
+\Phi_{\nu_e}\quad
+\Phi_{\nu_\mu}\quad
+\Phi_{\nu_\tau}\quad
+\Phi_{\bar\nu_e}\quad
+\Phi_{\bar\nu_\mu}\quad
+\Phi_{\bar\nu_\tau}$
+\end{quotation}
+where the energy values are evenly spaced. 
+For accessing fluxes at 
+arbitrary energies linear interpolation is used. Once the energy leaves the
+range of values given in the file $0.0$ is returned.
+
+Since many sources have two modes of operation, it is possible to define
+a pair of fluxes by using the suffixes ``plus'' and ``minus'' in the file 
+names. The individual members of the pair can be accessed by the polarity
+variable. Also the built-in fluxes provide this choice.
+
+For sources based on $\pi$-decay and nuclear reactors the total power
+involved defines how neutrino are produced therefore this variable is
+used to define the luminosity.
+
+Furthemore it is useful to define the operation time of an experiment
+in years.
 
@@
 
-\section{Detector properties}
+\subsection{Baseline \& matter profile}
+The purpose of this module is to specify the path the neutrinos
+are traveling. The central piece of information is the total length
+of the path - the baseline\index{Baseline}.
+\begin{equation}
+L\,\left[\mathrm{km}\right]
+\end{equation}
+
+Furthermore there is
+the matter profile encountered along the baseline which can be set in
+several ways. The simplest way is to use a constant density corresponding
+to the average density along the baseline calculated using the 
+PREM\cite{Stacey} onion shell model of the 
+Earth\index{PREM}\index{Earth matterdensity}.
+
+Another possibility is to use several layers of constant density each in 
+order to improve the approximation to the actual Earth matter profile. In this
+case the user specifies the total baseline and the number of layers. The
+time consumed in computing the oscillation probabilities is directly 
+proportional to number of layers.
+
+The third possibility to specify the matter profile is to provide a list
+of thicknesses and densities for the layers.
+
+For any type of profile a central value $a$ and a corresponding error 
+$\delta a$ have to be specified\index{Density error}\index{Density center}.
+\begin{equation}
+\label{eq:density_error}
+\rho(x)=a\cdot\rho_0(x)\,\left[\frac{\mathrm{g}}{\mathrm{cm}^2}\right]\,
+\end{equation}
+where $\rho_0(x)$ is the matter profile as chosen by the user and $\rho(x)$ is
+the one used in the calculation of the oscillation probabilities. The default
+value for $a$ is $1$ and a typical value for $\delta a = 0.05$.
+
+\section{Detection properties}
+
+The most obvious property of the neutrino detection is the total
+active mass of the target.
+
+\subsection{Cross section}
+\label{sec:cross_section}
+
+Cross sections\index{Cross sections} are used as part of the 
+channel definition 
+(see section~\ref{sec:channel}) and have to be provided by the user in form
+of a file. Cross section are in general given as differential cross divided
+by the energy
+\begin{equation}
+\hat\sigma(E)=\sigma(E)/E\,\left[ 10^{-38}\,
+\frac{\mathrm{cm}^2}{\mathrm{GeV}^2} \right]
+\end{equation}
+The cross section files have seven rows and $1001$ lines.
+\begin{quotation}
+$log_{10} E\quad
+\hat\sigma_{\nu_e}\quad
+\hat\sigma_{\nu_\mu}\quad
+\hat\sigma_{\nu_\tau}\quad
+\hat\sigma_{\bar\nu_e}\quad
+\hat\sigma_{\bar\nu_\mu}\quad
+\hat\sigma_{\bar\nu_\tau}$
+\end{quotation}
+where the logarithm of the energy values is evenly spaced. 
+For accessing cross section at 
+arbitrary energies linear interpolation is used. Once the energy leaves the
+range of values given in the file $0.0$ is returned.
+
+\subsection{Channels}
+\label{sec:channel}
+
+Channels\index{Channel} represent an intermediate level in bewteen 
+the pure oscillation 
+physics given by the oscillation probablity $P_{\alpha\beta}$ and the
+detected signal and background. A channel basicaly decsribes the path from one
+initial state in the source to one interaction type (IT) in  the detector.  
+Therefore a channel consists of the description of the inital state given 
+by the polarity 
+of the source, \eg\ muons or anti-muons in a neutrino factory, by the CP 
+eigenvalue of the intial state, \ie\ neutrino or anti-neutrino and by the 
+intial flavour. A channel furthermore consists of the interaction type in the
+detector given by the final flavour of the neutrino, the interaction cross
+section and the energy resolution function.
+
+For the calculation of event rates, the first step is to compute the number of
+events for each IT in the fiducial mass of the detector for each 
+neutrino flavor and energy bin. The second step is to include the detector 
+effects coming from
+the insufficient knowledge used in the event reconstruction. We combine these
+two steps in order to obtain the differential event rate spectrum for each
+flavor and IT as seen by the detector, which we call a ``channel''. 
+The channels
+for all ITs then have to be combined in a way which takes into account
+that the ITs or flavors can, either because of
+physical reasons (\eg, the flavor--blindness of NC interactions) or 
+because of detector effects (\eg, charge misidentification), not be 
+ separated cleanly.
+
+The master formula for the differential event rate for each channel 
+is given by
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{eqnarray}
+\label{eq:master_event}
+\frac{dn_{\beta}^{\text{IT}}}{dE'}=&&N\,\int \int dE\,d\hat{E}\quad
+\underbrace{\Phi_{\alpha} (E)}_{\mathrm{Production}} \times \nonumber\\
+&&\underbrace{\frac{1}{L^2} P_{(\alpha\rightarrow\beta)}(E,L,\rho;\theta_{23},
+\theta_{12},\theta_{13},
+\Delta m^2_{31},\Delta m^2_{21},\deltacp)}_{\mathrm{Propagation}}
+\times \nonumber \\ &&\underbrace{\sigma^{\text{IT}}_f(E)
+k_f^{\text{IT}}(E-\hat{E})}_{\mathrm{Interaction}} \times \nonumber \\
+&&\underbrace{ T_f(\hat{E}) V_f(\hat{E}-E')}_{\mathrm{Detection}}\,,
+\end{eqnarray}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+where $\alpha$ is the initial flavor of the neutrino and 
+$\beta$ is the final flavour, 
+$E$ is the incident neutrino energy, $\Phi_{\alpha} (E)$ is the flux of the 
+initial flavor at the
+source, $L$ is the baseline length, $N$ is a normalization factor, and 
+$\rho$ is the matter density. The interaction term is composed of 
+two factors, which are the total cross section 
+$\sigma^{\text{IT}}_\beta(E)$ for the flavor $f$ and
+the interaction type IT, and the energy distribution of the 
+secondary particle $k_\beta^{\text{IT}}(E-\hat{E})$ with 
+$\hat{E}$ the energy of the secondary particle. The detector properties are 
+modeled by the threshold function $T_\beta(\hat{E})$, coming from the the 
+limited resolution or the cuts in the analysis, and the energy resolution 
+function $V_\beta(\hat{E}-E')$ of the secondary particle. Thus, $E'$ is the 
+{\em reconstructed} neutrino energy.
+
+Since it is rather cumbersome to numerically solve this double integral,
+we split the two integration. We first evaluate the integral over
+$\hat{E}$, where the only terms containing $\hat{E}$ are
+$k_\beta^{\text{IT}}(E-\hat{E})$,  $ T_\beta(\hat{E})$, and 
+$ V_\beta(\hat{E}-E')$, and define:
+\begin{eqnarray}
+\label{eq:e_res} 
+R_\beta^{\text{IT}}(E,E')\,\epsilon_\beta^{\text{IT}}(E')
+ \equiv
+\int d\hat{E} \quad T_\beta(\hat{E})\,k_\beta^{\text{IT}}(E-\hat{E})
+\,V_\beta(\hat{E}-E')\,, 
+\end{eqnarray}
+Thus $R_\beta^{\text{IT}}(E,E')$ described the energy response of 
+the detector, \ie\ a neutrino with a (true) energy of $E$ has a probablity
+of $R_\beta^{\text{IT}}(E,E') dE'$ to have a reconstucted energy 
+in bewteen $E'$ and $E'+dE'$. The function $R(E,E')$ is also refered to
+as energy resolution function or smearing matrix in its actual representation
+in the program\index{Energy resolution}\index{Smear matrix}. The detailed
+defintion and initialization of the energy resolution function is described
+in section~\ref{sec:energy}.
+
+\subsection{Rules}
+\label{sec:rules}
+
+The set of rules\index{Rule} for an experiment is the final 
+link between the event rate
+computation and the statistical analysis. The information given by the rules
+specifies how the $\chi^2$ is computed based on the raw event rates as 
+described by the channels including possible systematical errors. 
+Therefore a rule has two parts: the first part describes how signal and 
+background events are composed out of the channels and the second part
+specifies which systematical errors are considered and what the the size
+of the relevant errors is.
+
+Based on the definition of channels it is possible to assemble the 
+signal and background event numbers for each bin.  
+The number of events in one channel and  in one bin\index{Bin} $i$ is given by
+\begin{equation}
+n_i^c=\int_{E_i-\Delta E/2}^{E_i+\Delta E/2} dE' \quad
+\frac{dn_{\beta}^{\text{IT}}}{dE'} (E')
+\end{equation}
+where $c$ is the channel index.
+Note that the events are binned according to their \emph{reconstructed} energy.
+The signal $s_i$ in one bin for a rule now can be composed out of 
+several channels by
+\begin{equation}
+s_i=\epsilon_{c_1}\cdot n_i^{c_1}\,+\,\epsilon_{c_2}\cdot n_i^{c_2}\,+\,\ldots
+\end{equation}
+where the $\epsilon$'s are weight factors determined by the properties
+of the detector.
+Simlarly the background in one bin $i$ can be specified by
+\begin{equation}
+b_i=\epsilon_{c_1}\cdot n_i^{c_1}\,+\,\epsilon_{c_2}\cdot n_i^{c_2}\,+\,\ldots
+\end{equation}
+
+The set of all $s_i$'s and $b_i$'s now can be composed in a form that takes
+some types of systematical error into account. The two most important and
+most easily parameterized systematics are a normalization and an energy
+calibration error. These error are considered seperately for the signal events
+and the background events. The implementation of the normalization error
+is straight forward
+\begin{equation}
+s_i(a):=a\cdot s_i
+\end{equation} 
+with an anloguous definition for the background events.
+
+For the parameterization of an energy calibration error two possibilities
+are implemented. The first one being somewhat simpler and second one
+being more accurate. The first option is
+\begin{equation}
+s_i(a,b):=s_i(a)+b\cdot s_i\, E_i/(E_\mathrm{max}-E_\mathrm{min})
+\end{equation}
+and it is refered to as tilt since it describes a linear distortion, a tilt
+of the event rate spectrum. The second option is closer to an actual energy
+calibration error and is an approximation to the exact form
+\begin{equation}
+s_i(b) = \int_{(1+b)(E_i-\Delta E/2)}^{(1+b)(E_i+\Delta E/2)} dE' \quad
+\frac{dn_{\beta}^{\text{IT}}}{dE'} ( E')
+\end{equation}
+or something like this, to be continued ...
+
+Thus the total event rate $x_i$ in a bin $i$ is given by
+\begin{equation}
+x_i(a,b,c,d)=s_i(a,b)+b_i(c,d)
+\end{equation}
+and thus a function of four parameters. The central values for all of
+the four parameters are always needed. They are called signal normalization
+($a$), signal tilt/calibration ($b$), background  normalization ($c$) and
+background tilt/calibration ($d$). The default values are
+\begin{equation}
+a=1\,,\quad b=0\,,\quad c={\tt NaN}\,,\quad d=0\,,
+\end{equation}
+thus for the background normalization $c$ a value has to be specified in 
+\emph{all} cases.
+
+The four parameters $a,b,c,d$ have been introduced in order to describe
+systematical uncertainties, they are so called nuisance parameters.
+For the analysis of the systematical errors the so called 
+ pull method is 
+used~\cite{Fogli:2002pt}\footnote{In fact the pull method
+was employed already in~\cite{SBvNF} before \cite{Fogli:2002pt} appeared.}. 
+In the  pull method $k$ systematical errors are included by introducing 
+$k$ additional variables $\zeta_k$, which will  be called 
+nuisance parameters in the following. 
+The nuisance parameters describe the dependence of the event rates on the 
+various sources of systematical errors, \eg\ an error on the total 
+normalization is included by multiplying the expected number of events in 
+each bin by a factor $(1+\zeta_1)$. The variation of $\zeta_1$ in the fit 
+is constrained by adding a penalty $p_1$ to the $\chi^2$-function. In case 
+of a  Gau\ss ian distributed systematical error this penalty is 
+given by
+\begin{equation}
+\label{eq:penalty}
+p_i=\frac{(\zeta_i-\zeta_i^0)^2}{\sigma_{\zeta_i}^2}\,,
+\end{equation}
+where $\zeta_i^0$ denotes the mean and $\sigma_{\zeta_i}$ the standard 
+deviation of the corresponding nuisance parameter, \ie\ the amount of 
+systematical uncertainty. 
+The resulting $\chi^2$ is then minimized with respect to all nuisance 
+parameters $\zeta_i$ and this yields $\chi^2_\mathrm{pull}$
+\begin{equation}
+\chi^2_\mathrm{pull}(\boldsymbol{\lambda}):=\min_{\{\zeta_i\} } \,\, \left( 
+\chi^2(\boldsymbol{\lambda},
+\zeta_1, \ldots, \zeta_k)+ \sum_{j=1}^{k} p_j(\zeta_j)\right)\,,
+\end{equation}
+where $\boldsymbol{\lambda}$ denotes the oscillation parameters 
+including the matter density
+$\rho$. One advantage of the pull method is that whenever the number $N$ of 
+data points is much larger than $k$, it is numerically easier to compute 
+$\chi^2_\mathrm{pull}$ than to invert the $N\times N$ covariance matrix. For
+the experiments considered here $N$ is typically $20$ and $k\sim 4$, thus
+the pull method is numerically much faster. Moreover it is more flexible and 
+allows the inclusion of systematical errors also for a 
+Poissonian $\chi^2$-function.
+In~\cite{Fogli:2002pt} it was shown that the pull method and the covariance
+based approach are equivalent for a Gau\ss ian and linear model. In general
+there is a separate $(\chi^2_\mathrm{pull})^\alpha$ for each rule $\alpha$, 
+\ie\ pair of signal and background spectra, with a separate set of 
+nuisance parameters $\zeta_i^\alpha$. In this case $\chi^2_\mathrm{pull}$
+is the sum of all individual  $(\chi^2_\mathrm{pull})^\alpha$.
+In this way the dependence on the $k$ nuisance parameters has been eliminated 
+from $\chi^2_\mathrm{pull}$.
+
+The user has the possibility to choose the set $\{\zeta_i\}$ of nuisance 
+parameters which are marginalized, this achieved with the error dimension 
+variable\index{Error dimension}, the different possibilities are shown in
+table~\ref{tab:error_dim}. If a parameter is desginated with $+$ it will
+be marginalized and therefore the corresponding error needs to have a non-zero
+value. In the cases labeled (number 4 and 8) ``total rates'' 
+the summation over the bins is performed \emph{before} computing 
+the $\chi^2$. The case labeled (number 7) ``spectrum only'' leaves the 
+normalization free ($\sigma_a=\sigma_c=\infty$) and therefore counts only the 
+spectral information in the date, therefore
+the settings for the normalization error will be ignored. It is furthermore
+possible to extend the set of error dimensions and thus the set of possible 
+systematical errors to be studied, but this requires proper C programing
+and recompiling \GLOBES. The interface for user defined error dimensions is
+described in detail in section~\ref{sec:ud_error_dim}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{center}
+\begin{table}[hbt!]
+\begin{center}
+\begin{tabular}[h]{|c|cccc|c|c|}
+\hline
+Error dimension&$a$&$b$&$c$&$d$&Tilt/Calibration&Remarks\\
+\hline
+\hline
+0&+&+&+&+&T&\\
+2&-&-&-&-&-&\\
+4&+&+&+&+&T&total rates\\
+7&o&-&o&-&-&spectrum only\\
+8&-&-&-&-&-&total rates\\
+9&+&+&+&+&C&\\
+\hline
+\end{tabular}
+\caption[Table of error dimensions]{\label{tab:error_dim}
+Values of the error dimension variable and their meaning.
+ }
+\end{center} 
+\end{table} 
+\end{center}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In each rule it is possible to constrain the energy range from which the
+data are taken for the analysis\index{Energy window}. This achieved by 
+setting the energy window
+variable. The purpose is mainly to allow a fine tuning of the description
+
+of an energy calibration error. The default setting is that the energy window
+coincides with the range defined by the minimal and maximal reconstructed 
+energy. For the error dimension using the calibration method (C) it is 
+advisable to reduce the energy window by two or three bins at both the lower
+and upper end. When studying the impact of the energy calibration
+error in detail and it turns out not to be small, make sure that your result
+does not change by reducing the energy window by 1 or 2 bins.
+
+\subsection{Energy resolution function}
+\label{sec:energy}
+
+The origin and meaning of the energy resolution function 
+$R^c(E,E')$ was already discussed in 
+section~\ref{sec:channel} and  a mathematical definition
+was given in equation~\ref{eq:e_res}. Here we deal with
+how to tell \GLOBES\ which energy resolution function it should
+use. This also naturally involves the issue how the actual convolution
+of the differntial event rate spectrum with $R_\beta^{\text{IT}}(E,E')$
+is performed. First we will discuss the mathematical structure of
+the problem and then the various algorithms and possibilities which
+\GLOBES\ offers are introduced. In order to obtain the number of events 
+$n_i^c$ from one channel $c$ in one bin $i$ one has to solve this integral
+%
+\begin{eqnarray}
+\label{eq:events_bin}
+n_i^c=N/L^2\,\int_{E_i-\Delta E/2}^{E_i+\Delta E/2} dE' 
+\quad \int dE \,\, \Phi^c(E)\,
+P^c(E)\,
+\sigma^c(E)\,
+R^c(E,E')\,
+\epsilon^c(E')\,.
+\end{eqnarray} 
+%
+The integration with respect to the reconstructed energy $E'$ can be
+performed independently of the oscillation parameters and we define
+the bin kernel for the $i$th bin $K_i^c$
+\begin{equation}
+\label{eq:kernel}
+K_i^c(E):=\int_{E_i-\Delta E/2}^{E_i+\Delta E/2} dE' 
+\quad R^c(E,E')\,
+\epsilon^c(E')\,.
+\end{equation}
+With this definition equation~\ref{eq:events_bin} can be rewritten as
+\begin{equation}
+\label{eq:simple_int}
+n_i^c=N/L^2 \int dE\quad  \underbrace{\Phi^c(E)\,
+P^c(E)\,
+\sigma^c(E)\,
+K_i^c(E)\,}_{f(E)}. 
+\end{equation}
+
+There is no basic reason why one could not evaluate this integral directly
+by the usual numerical methods. However it turns out that this may be rather
+slow therefore we introduce two different approximation schemes which have
+a varying degree of acuracy and speed. The simplest, fastest and
+least acurate is algorithm one. The key idea is to choose a set of assumptions
+which allow to evaluate the integrand of equation~\ref{eq:simple_int}
+at fixed values of the energy $E$ and to use the same number of points $N$ in
+the calculation as there are bins. Thus equation~\ref{eq:simple_int} is reduced
+to
+\begin{equation}
+\label{eq:algo_one}
+n_i^c=N/L^2 \Delta E \sum_{j=1}^N \quad  \Phi^c(E_j)\,
+P^c(E_j)\,
+\sigma^c(E_j)\,
+K_i^c(E_j)\,.
+\end{equation}
+The advantage of using equation~\ref{eq:algo_one} is that all factors
+independent of the oscillation paramters have to be evaluated only at 
+values of $E$ which are known in advance, \ie\ they can be put 
+into a look-up table. Another advantage is that also the probability
+has to be evaluated only at previously known values of the energy, which
+makes it possible to compute the transition amplitudes for all channels
+in one step. The key assumption is that all involved factors are piece-wise
+constant, \ie\ they do not change within one bin. This assumption seems to be
+very restrictive however whenever one is analyzing simulated 
+data\footnote{simulated with the same algorithm} the errors
+made by violating this assumption nearly cancel between the simulated 
+data and the fitted data. This algorithm is the fastest which \GLOBES\ offers
+and is suitable whenever there are no fast oscillations. Otherwise severe
+aliasing effects may occur which usually lead to a failure of one of the 
+minimizing routines. The computation of the bin kernel $K_i^c$ is performed
+by \GLOBES\, to this end it requires the specification of the number of bins
+and the minimum and maximum energy. Furthemore following parameterization
+for $R^c(E,E')$ is used
+\begin{equation}
+R^c(E,E')=\frac{1}{\sigma(E)\,\sqrt{2\pi}}\,e^{-\frac{(E-E')^2}{2\sigma^2(E)}}
+\end{equation} 
+where the energy resolution $\sigma(E)$ is defined by
+\begin{equation}
+\label{eq:sigma_e}
+\sigma(E)=\alpha\cdot E + \beta \cdot \sqrt{E} +\gamma\,.
+\end{equation}
+The parameters $\alpha, \beta$ and $\gamma$ are provided by the user.
+Furthemore the user has the possibility to provide $\epsilon(E')$ by specifying
+the value of  $\epsilon(E')$ at each $E_j$ since this algroithm assumes
+everything else to be constant within one bin, the  $\epsilon(E')$ are
+also taken as constant within one bin. In the actual implementation the sum
+in equation one only extends over those $E_j$ where $K(E_j)$ above a certain
+treshold, by default $10^{-7}$.
+
+The second available algorithm is more complicated, a little slower but
+much more accurate and much less susceptible to aliasing effects. The starting
+point again is equation~\ref{eq:simple_int}. The idea now is to
+ have a good approximation to the integral but still keeping the fixed sampling
+point in $E$. To this end we rewrite the integrand of 
+equation~\ref{eq:simple_int} using the sampling theorem~\cite{NRC,Rybicki}
+\begin{equation}
+f(E)=\sum_{k=-\infty}^{\infty}\,f(E_k) 
+\times \mathrm{sinc}\left(\frac{\pi}{h}(E-E_k)\right)+e(E)
+\end{equation}
+with $\mathrm{sinc}(x):=\sin(x)/x$ and $E_k=E^0+h\cdot k$. 
+This the so called sampling 
+representation of the integrand $f(E)$ and $e(E)$ is the error term.
 
@@
 
-Most advanced step: How to create a new experiment from scratch.
+Most advanced step: How to create a new experiment from scratch .
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: Manual.tex
+%%% End:

### 2004-04-23 Documentation/fdl.tex
two files added for the GPL and FDL

\chapter{GNU Free Documentation License}
%\label{label_fdl}

 \begin{center}

       Version 1.2, November 2002

 Copyright \copyright 2000,2001,2002  Free Software Foundation, Inc.

 \bigskip

     59 Temple Place, Suite 330, Boston, MA  02111-1307  USA

 \bigskip

 Everyone is permitted to copy and distribute verbatim copies
 of this license document, but changing it is not allowed.
\end{center}

\begin{center}
{\bf\large Preamble}
\end{center}

The purpose of this License is to make a manual, textbook, or other
functional and useful document "free" in the sense of freedom: to
assure everyone the effective freedom to copy and redistribute it,
with or without modifying it, either commercially or noncommercially.
Secondarily, this License preserves for the author and publisher a way
to get credit for their work, while not being considered responsible
for modifications made by others.

This License is a kind of "copyleft", which means that derivative
works of the document must themselves be free in the same sense.  It
complements the GNU General Public License, which is a copyleft
license designed for free software.

We have designed this License in order to use it for manuals for free
software, because free software needs free documentation: a free
program should come with manuals providing the same freedoms that the
software does.  But this License is not limited to software manuals;
it can be used for any textual work, regardless of subject matter or
whether it is published as a printed book.  We recommend this License
principally for works whose purpose is instruction or reference.

\begin{center}
{\Large\bf 1. APPLICABILITY AND DEFINITIONS}
\addcontentsline{toc}{section}{1. APPLICABILITY AND DEFINITIONS}
\end{center}

This License applies to any manual or other work, in any medium, that
contains a notice placed by the copyright holder saying it can be
distributed under the terms of this License.  Such a notice grants a
world-wide, royalty-free license, unlimited in duration, to use that
work under the conditions stated herein.  The \textbf{"Document"}, below,
refers to any such manual or work.  Any member of the public is a
licensee, and is addressed as \textbf{"you"}.  You accept the license if you
copy, modify or distribute the work in a way requiring permission
under copyright law.

A \textbf{"Modified Version"} of the Document means any work containing the
Document or a portion of it, either copied verbatim, or with
modifications and/or translated into another language.

A \textbf{"Secondary Section"} is a named appendix or a front-matter section of
the Document that deals exclusively with the relationship of the
publishers or authors of the Document to the Document's overall subject
(or to related matters) and contains nothing that could fall directly
within that overall subject.  (Thus, if the Document is in part a
textbook of mathematics, a Secondary Section may not explain any
mathematics.)  The relationship could be a matter of historical
connection with the subject or with related matters, or of legal,
commercial, philosophical, ethical or political position regarding
them.

The \textbf{"Invariant Sections"} are certain Secondary Sections whose titles
are designated, as being those of Invariant Sections, in the notice
that says that the Document is released under this License.  If a
section does not fit the above definition of Secondary then it is not
allowed to be designated as Invariant.  The Document may contain zero
Invariant Sections.  If the Document does not identify any Invariant
Sections then there are none.

The \textbf{"Cover Texts"} are certain short passages of text that are listed,
as Front-Cover Texts or Back-Cover Texts, in the notice that says that
the Document is released under this License.  A Front-Cover Text may
be at most 5 words, and a Back-Cover Text may be at most 25 words.

A \textbf{"Transparent"} copy of the Document means a machine-readable copy,
represented in a format whose specification is available to the
general public, that is suitable for revising the document
straightforwardly with generic text editors or (for images composed of
pixels) generic paint programs or (for drawings) some widely available
drawing editor, and that is suitable for input to text formatters or
for automatic translation to a variety of formats suitable for input
to text formatters.  A copy made in an otherwise Transparent file
format whose markup, or absence of markup, has been arranged to thwart
or discourage subsequent modification by readers is not Transparent.
An image format is not Transparent if used for any substantial amount
of text.  A copy that is not "Transparent" is called \textbf{"Opaque"}.

Examples of suitable formats for Transparent copies include plain
ASCII without markup, Texinfo input format, LaTeX input format, SGML
or XML using a publicly available DTD, and standard-conforming simple
HTML, PostScript or PDF designed for human modification.  Examples of
transparent image formats include PNG, XCF and JPG.  Opaque formats
include proprietary formats that can be read and edited only by
proprietary word processors, SGML or XML for which the DTD and/or
processing tools are not generally available, and the
machine-generated HTML, PostScript or PDF produced by some word
processors for output purposes only.

The \textbf{"Title Page"} means, for a printed book, the title page itself,
plus such following pages as are needed to hold, legibly, the material
this License requires to appear in the title page.  For works in
formats which do not have any title page as such, "Title Page" means
the text near the most prominent appearance of the work's title,
preceding the beginning of the body of the text.

A section \textbf{"Entitled XYZ"} means a named subunit of the Document whose
title either is precisely XYZ or contains XYZ in parentheses following
text that translates XYZ in another language.  (Here XYZ stands for a
specific section name mentioned below, such as \textbf{"Acknowledgements"},
\textbf{"Dedications"}, \textbf{"Endorsements"}, or \textbf{"History"}.)  
To \textbf{"Preserve the Title"}
of such a section when you modify the Document means that it remains a
section "Entitled XYZ" according to this definition.

The Document may include Warranty Disclaimers next to the notice which
states that this License applies to the Document.  These Warranty
Disclaimers are considered to be included by reference in this
License, but only as regards disclaiming warranties: any other
implication that these Warranty Disclaimers may have is void and has
no effect on the meaning of this License.

\begin{center}
{\Large\bf 2. VERBATIM COPYING}
\addcontentsline{toc}{section}{2. VERBATIM COPYING}
\end{center}

You may copy and distribute the Document in any medium, either
commercially or noncommercially, provided that this License, the
copyright notices, and the license notice saying this License applies
to the Document are reproduced in all copies, and that you add no other
conditions whatsoever to those of this License.  You may not use
technical measures to obstruct or control the reading or further
copying of the copies you make or distribute.  However, you may accept
compensation in exchange for copies.  If you distribute a large enough
number of copies you must also follow the conditions in section 3.

You may also lend copies, under the same conditions stated above, and
you may publicly display copies.

\begin{center}
{\Large\bf 3. COPYING IN QUANTITY}
\addcontentsline{toc}{section}{3. COPYING IN QUANTITY}
\end{center}

If you publish printed copies (or copies in media that commonly have
printed covers) of the Document, numbering more than 100, and the
Document's license notice requires Cover Texts, you must enclose the
copies in covers that carry, clearly and legibly, all these Cover
Texts: Front-Cover Texts on the front cover, and Back-Cover Texts on
the back cover.  Both covers must also clearly and legibly identify
you as the publisher of these copies.  The front cover must present
the full title with all words of the title equally prominent and
visible.  You may add other material on the covers in addition.
Copying with changes limited to the covers, as long as they preserve
the title of the Document and satisfy these conditions, can be treated
as verbatim copying in other respects.

If the required texts for either cover are too voluminous to fit
legibly, you should put the first ones listed (as many as fit
reasonably) on the actual cover, and continue the rest onto adjacent
pages.

If you publish or distribute Opaque copies of the Document numbering
more than 100, you must either include a machine-readable Transparent
copy along with each Opaque copy, or state in or with each Opaque copy
a computer-network location from which the general network-using
public has access to download using public-standard network protocols
a complete Transparent copy of the Document, free of added material.
If you use the latter option, you must take reasonably prudent steps,
when you begin distribution of Opaque copies in quantity, to ensure
that this Transparent copy will remain thus accessible at the stated
location until at least one year after the last time you distribute an
Opaque copy (directly or through your agents or retailers) of that
edition to the public.

It is requested, but not required, that you contact the authors of the
Document well before redistributing any large number of copies, to give
them a chance to provide you with an updated version of the Document.

\begin{center}
{\Large\bf 4. MODIFICATIONS}
\addcontentsline{toc}{section}{4. MODIFICATIONS}
\end{center}

You may copy and distribute a Modified Version of the Document under
the conditions of sections 2 and 3 above, provided that you release
the Modified Version under precisely this License, with the Modified
Version filling the role of the Document, thus licensing distribution
and modification of the Modified Version to whoever possesses a copy
of it.  In addition, you must do these things in the Modified Version:

\begin{itemize}
\item[A.] 
   Use in the Title Page (and on the covers, if any) a title distinct
   from that of the Document, and from those of previous versions
   (which should, if there were any, be listed in the History section
   of the Document).  You may use the same title as a previous version
   if the original publisher of that version gives permission.

\item[B.]
   List on the Title Page, as authors, one or more persons or entities
   responsible for authorship of the modifications in the Modified
   Version, together with at least five of the principal authors of the
   Document (all of its principal authors, if it has fewer than five),
   unless they release you from this requirement.

\item[C.]
   State on the Title page the name of the publisher of the
   Modified Version, as the publisher.

\item[D.]
   Preserve all the copyright notices of the Document.

\item[E.]
   Add an appropriate copyright notice for your modifications
   adjacent to the other copyright notices.

\item[F.]
   Include, immediately after the copyright notices, a license notice
   giving the public permission to use the Modified Version under the
   terms of this License, in the form shown in the Addendum below.

\item[G.]
   Preserve in that license notice the full lists of Invariant Sections
   and required Cover Texts given in the Document's license notice.

\item[H.]
   Include an unaltered copy of this License.

\item[I.]
   Preserve the section Entitled "History", Preserve its Title, and add
   to it an item stating at least the title, year, new authors, and
   publisher of the Modified Version as given on the Title Page.  If
   there is no section Entitled "History" in the Document, create one
   stating the title, year, authors, and publisher of the Document as
   given on its Title Page, then add an item describing the Modified
   Version as stated in the previous sentence.

\item[J.]
   Preserve the network location, if any, given in the Document for
   public access to a Transparent copy of the Document, and likewise
   the network locations given in the Document for previous versions
   it was based on.  These may be placed in the "History" section.
   You may omit a network location for a work that was published at
   least four years before the Document itself, or if the original
   publisher of the version it refers to gives permission.

\item[K.]
   For any section Entitled "Acknowledgements" or "Dedications",
   Preserve the Title of the section, and preserve in the section all
   the substance and tone of each of the contributor acknowledgements
   and/or dedications given therein.

\item[L.]
   Preserve all the Invariant Sections of the Document,
   unaltered in their text and in their titles.  Section numbers
   or the equivalent are not considered part of the section titles.

\item[M.]
   Delete any section Entitled "Endorsements".  Such a section
   may not be included in the Modified Version.

\item[N.]
   Do not retitle any existing section to be Entitled "Endorsements"
   or to conflict in title with any Invariant Section.

\item[O.]
   Preserve any Warranty Disclaimers.
\end{itemize}

If the Modified Version includes new front-matter sections or
appendices that qualify as Secondary Sections and contain no material
copied from the Document, you may at your option designate some or all
of these sections as invariant.  To do this, add their titles to the
list of Invariant Sections in the Modified Version's license notice.
These titles must be distinct from any other section titles.

You may add a section Entitled "Endorsements", provided it contains
nothing but endorsements of your Modified Version by various
parties--for example, statements of peer review or that the text has
been approved by an organization as the authoritative definition of a
standard.

You may add a passage of up to five words as a Front-Cover Text, and a
passage of up to 25 words as a Back-Cover Text, to the end of the list
of Cover Texts in the Modified Version.  Only one passage of
Front-Cover Text and one of Back-Cover Text may be added by (or
through arrangements made by) any one entity.  If the Document already
includes a cover text for the same cover, previously added by you or
by arrangement made by the same entity you are acting on behalf of,
you may not add another; but you may replace the old one, on explicit
permission from the previous publisher that added the old one.

The author(s) and publisher(s) of the Document do not by this License
give permission to use their names for publicity for or to assert or
imply endorsement of any Modified Version.

\begin{center}
{\Large\bf 5. COMBINING DOCUMENTS}
\addcontentsline{toc}{section}{5. COMBINING DOCUMENTS}
\end{center}

You may combine the Document with other documents released under this
License, under the terms defined in section 4 above for modified
versions, provided that you include in the combination all of the
Invariant Sections of all of the original documents, unmodified, and
list them all as Invariant Sections of your combined work in its
license notice, and that you preserve all their Warranty Disclaimers.

The combined work need only contain one copy of this License, and
multiple identical Invariant Sections may be replaced with a single
copy.  If there are multiple Invariant Sections with the same name but
different contents, make the title of each such section unique by
adding at the end of it, in parentheses, the name of the original
author or publisher of that section if known, or else a unique number.
Make the same adjustment to the section titles in the list of
Invariant Sections in the license notice of the combined work.

In the combination, you must combine any sections Entitled "History"
in the various original documents, forming one section Entitled
"History"; likewise combine any sections Entitled "Acknowledgements",
and any sections Entitled "Dedications".  You must delete all sections
Entitled "Endorsements".

\begin{center}
{\Large\bf 6. COLLECTIONS OF DOCUMENTS}
\addcontentsline{toc}{section}{6. COLLECTIONS OF DOCUMENTS}
\end{center}

You may make a collection consisting of the Document and other documents
released under this License, and replace the individual copies of this
License in the various documents with a single copy that is included in
the collection, provided that you follow the rules of this License for
verbatim copying of each of the documents in all other respects.

You may extract a single document from such a collection, and distribute
it individually under this License, provided you insert a copy of this
License into the extracted document, and follow this License in all
other respects regarding verbatim copying of that document.

\begin{center}
{\Large\bf 7. AGGREGATION WITH INDEPENDENT WORKS}
\addcontentsline{toc}{section}{7. AGGREGATION WITH INDEPENDENT WORKS}
\end{center}

A compilation of the Document or its derivatives with other separate
and independent documents or works, in or on a volume of a storage or
distribution medium, is called an "aggregate" if the copyright
resulting from the compilation is not used to limit the legal rights
of the compilation's users beyond what the individual works permit.
When the Document is included in an aggregate, this License does not
apply to the other works in the aggregate which are not themselves
derivative works of the Document.

If the Cover Text requirement of section 3 is applicable to these
copies of the Document, then if the Document is less than one half of
the entire aggregate, the Document's Cover Texts may be placed on
covers that bracket the Document within the aggregate, or the
electronic equivalent of covers if the Document is in electronic form.
Otherwise they must appear on printed covers that bracket the whole
aggregate.

\begin{center}
{\Large\bf 8. TRANSLATION}
\addcontentsline{toc}{section}{8. TRANSLATION}
\end{center}

Translation is considered a kind of modification, so you may
distribute translations of the Document under the terms of section 4.
Replacing Invariant Sections with translations requires special
permission from their copyright holders, but you may include
translations of some or all Invariant Sections in addition to the
original versions of these Invariant Sections.  You may include a
translation of this License, and all the license notices in the
Document, and any Warranty Disclaimers, provided that you also include
the original English version of this License and the original versions
of those notices and disclaimers.  In case of a disagreement between
the translation and the original version of this License or a notice
or disclaimer, the original version will prevail.

If a section in the Document is Entitled "Acknowledgements",
"Dedications", or "History", the requirement (section 4) to Preserve
its Title (section 1) will typically require changing the actual
title.

\begin{center}
{\Large\bf 9. TERMINATION}
\addcontentsline{toc}{section}{9. TERMINATION}
\end{center}

You may not copy, modify, sublicense, or distribute the Document except
as expressly provided for under this License.  Any other attempt to
copy, modify, sublicense or distribute the Document is void, and will
automatically terminate your rights under this License.  However,
parties who have received copies, or rights, from you under this
License will not have their licenses terminated so long as such
parties remain in full compliance.

\begin{center}
{\Large\bf 10. FUTURE REVISIONS OF THIS LICENSE}
\addcontentsline{toc}{section}{10. FUTURE REVISIONS OF THIS LICENSE}
\end{center}

The Free Software Foundation may publish new, revised versions
of the GNU Free Documentation License from time to time.  Such new
versions will be similar in spirit to the present version, but may
differ in detail to address new problems or concerns.  See
http://www.gnu.org/copyleft/.

Each version of the License is given a distinguishing version number.
If the Document specifies that a particular numbered version of this
License "or any later version" applies to it, you have the option of
following the terms and conditions either of that specified version or
of any later version that has been published (not as a draft) by the
Free Software Foundation.  If the Document does not specify a version
number of this License, you may choose any version ever published (not
as a draft) by the Free Software Foundation.

\begin{center}
{\Large\bf ADDENDUM: How to use this License for your documents}
\addcontentsline{toc}{section}{ADDENDUM: How to use this License for your documents}
\end{center}

To use this License in a document you have written, include a copy of
the License in the document and put the following copyright and
license notices just after the title page:

\bigskip
\begin{quote}
    Copyright \copyright  YEAR  YOUR NAME.
    Permission is granted to copy, distribute and/or modify this document
    under the terms of the GNU Free Documentation License, Version 1.2
    or any later version published by the Free Software Foundation;
    with no Invariant Sections, no Front-Cover Texts, and no Back-Cover Texts.
    A copy of the license is included in the section entitled "GNU
    Free Documentation License".
\end{quote}
\bigskip

If you have Invariant Sections, Front-Cover Texts and Back-Cover Texts,
replace the "with...Texts." line with this:

\bigskip
\begin{quote}
    with the Invariant Sections being LIST THEIR TITLES, with the
    Front-Cover Texts being LIST, and with the Back-Cover Texts being LIST.
\end{quote}
\bigskip

If you have Invariant Sections without Cover Texts, or some other
combination of the three, merge those two alternatives to suit the
situation.

If your document contains nontrivial examples of program code, we
recommend releasing these examples in parallel under your choice of
free software license, such as the GNU General Public License,
to permit their use in free software.

### 2004-04-23 Documentation/part2.tex
two files added for the GPL and FDL

@@
 \chapter{General concept}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Source properties}
+\label{sec:source}
 
 \GLOBES\ handles only point sources, thus it is not possible
-to study effects from the finit size of the neutrino production
+to study effects from the finite size of the neutrino production
 region like in the Sun. Furthermore only one source for each experiment
@@
 to simulate multiple source experiments like KamLAND. All neutrino
-sources useable within \GLOBES\ therefore can be specified by the
+sources usable within \GLOBES\ therefore can be specified by the
 flux spectra for each flavour and CP sign and the total luminosity
@@
 The built-in source is muon decay, since the neutrino
-spectra and flavour composition are knon analytically in that case. For 
+spectra and flavour composition are known analytically in that case. For 
 this sources the energy of the parent particle has to be specified
@@
 arbitrary energies linear interpolation is used. Once the energy leaves the
-range of values given in the file $0.0$ is returned.
+range of values given in the file $0.0$ is returned. The units used for the
+fluxes are not specified and right now each flux has its own fudge factor
+which is hard encoded in the file {\tt nufact.c}. The problem is that
+in computing the normalization one has to know things like number of target
+particles per unit mass asf. which is not always straitforward. Thus a fudge
+factor will be usefule in any case. Right now however this is not the only 
+fudge factor since for some experiments the definition of their target mass
+involves another fudge factor.
 
 Since many sources have two modes of operation, it is possible to define
-a pair of fluxes by using the suffixes ``plus'' and ``minus'' in the file 
-names. The individual members of the pair can be accessed by the polarity
+a pair of fluxes by using the polarity flag when loading the files. Usually
+files connected in this way use the suffixes ``plus'' and ``minus'' in the 
+file names. The individual members of the pair can be accessed by the polarity
 variable. Also the built-in fluxes provide this choice.
@@
 For sources based on $\pi$-decay and nuclear reactors the total power
-involved defines how neutrino are produced therefore this variable is
-used to define the luminosity.
-
-Furthemore it is useful to define the operation time of an experiment
-in years.
-
+involved defines how many neutrinos are produced therefore this variable is
+used to define the luminosity. Furthermore it is useful to define 
+the operation time of an experiment in years. Note however that the 
+number of events used in analysis only depends on the product of
+\begin{equation}
+\mathrm{mass}\,\left[10^6\,\mathrm{kg}\right]\times \mathrm{time}
+\,\left[\mathrm{years}\right]\times\left\{ \begin{array}{c}
+\mathrm{source~power}\,\left[\mathrm{MWy}\right]\\
+\mathrm{useful~decays}\,\left[\mathrm{year}^{-1}\right]
+\end{array}\right.\,.
+\end{equation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Oscillation properties}
@@
 are traveling. The central piece of information is the total length
-of the path - the baseline\index{Baseline}.
+of the path - the baseline\index{Baseline}
 \begin{equation}
@@
 PREM\cite{Stacey} onion shell model of the 
-Earth\index{PREM}\index{Earth matterdensity}.
+Earth\index{PREM}\index{Earth matter density}.
 
@@
 time consumed in computing the oscillation probabilities is directly 
-proportional to number of layers.
+proportional to the number of layers.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Detection properties}
@@
 The most obvious property of the neutrino detection is the total
-active mass of the target.
+active mass of the target as already defined in section~\ref{sec:source}.
 
@@
 (see section~\ref{sec:channel}) and have to be provided by the user in form
-of a file. Cross section are in general given as differential cross divided
-by the energy
+of a file. Cross section are in general given as differential cross section
+divided by the energy
 \begin{equation}
@@
 
-Channels\index{Channel} represent an intermediate level in bewteen 
+Channels\index{Channel} represent an intermediate level in between 
 the pure oscillation 
-physics given by the oscillation probablity $P_{\alpha\beta}$ and the
-detected signal and background. A channel basicaly decsribes the path from one
+physics given by the oscillation probability $P_{\alpha\beta}$ and the
+detected signal and background. A channel basically describes the path from one
 initial state in the source to one interaction type (IT) in  the detector.  
-Therefore a channel consists of the description of the inital state given 
+Therefore a channel consists of the description of the initial state given 
 by the polarity 
 of the source, \eg\ muons or anti-muons in a neutrino factory, by the CP 
-eigenvalue of the intial state, \ie\ neutrino or anti-neutrino and by the 
-intial flavour. A channel furthermore consists of the interaction type in the
+eigenvalue of the initial state, \ie\ neutrino or anti-neutrino and by the 
+initial flavour. A channel furthermore consists of the interaction type in the
 detector given by the final flavour of the neutrino, the interaction cross
@@
 flavor and IT as seen by the detector, which we call a ``channel''. 
-The channels
-for all ITs then have to be combined in a way which takes into account
-that the ITs or flavors can, either because of
-physical reasons (\eg, the flavor--blindness of NC interactions) or 
-because of detector effects (\eg, charge misidentification), not be 
- separated cleanly.
-
 The master formula for the differential event rate for each channel 
@@
 Since it is rather cumbersome to numerically solve this double integral,
-we split the two integration. We first evaluate the integral over
+we split the two integrations. We first evaluate the integral over
 $\hat{E}$, where the only terms containing $\hat{E}$ are
@@
 \end{eqnarray}
-Thus $R_\beta^{\text{IT}}(E,E')$ described the energy response of 
-the detector, \ie\ a neutrino with a (true) energy of $E$ has a probablity
-of $R_\beta^{\text{IT}}(E,E') dE'$ to have a reconstucted energy 
-in bewteen $E'$ and $E'+dE'$. The function $R(E,E')$ is also refered to
+
+!!!!!!!!!!!!
+Do we need $\epsilon(E')$ ???
+!!!!!!!!!!!!
+
+Thus $R_\beta^{\text{IT}}(E,E')$ describes the energy response of 
+the detector, \ie\ a neutrino with a (true) energy of $E$ has a probability
+of $R_\beta^{\text{IT}}(E,E') dE'$ to have a reconstructed energy 
+in between $E'$ and $E'+dE'$. The function $R(E,E')$ is also refered to
 as energy resolution function or smearing matrix in its actual representation
 in the program\index{Energy resolution}\index{Smear matrix}. The detailed
-defintion and initialization of the energy resolution function is described
+definition and initialization of the energy resolution function is described
 in section~\ref{sec:energy}.
@@
 background events are composed out of the channels and the second part
-specifies which systematical errors are considered and what the the size
+specifies which systematical errors are considered and what the size
 of the relevant errors is.
@@
 of the detector.
-Simlarly the background in one bin $i$ can be specified by
+Similarly the background in one bin $i$ can be specified by
 \begin{equation}
@@
 \end{equation}
+Quite often it is not sufficient to have constant efficiencies and therefore
+it is possible to define efficiencies for each bin and channel $\alpha_i^c$
+and thus the event rates in one bin become
+\begin{equation}
+s_i=\epsilon_{c_1}\alpha_i^{c_1}\cdot n_i^{c_1}\,+
+\,\epsilon_{c_2}\alpha_i^{c_2}\cdot n_i^{c_2}\,+\,\ldots\,,
+\end{equation}
+and analogous for $b_i$.
 
@@
 most easily parameterized systematics are a normalization and an energy
-calibration error. These error are considered seperately for the signal events
+calibration error. These error are considered separately for the signal events
 and the background events. The implementation of the normalization error
@@
 \end{equation} 
-with an anloguous definition for the background events.
+with an analogous definition for the background events.
 
 For the parameterization of an energy calibration error two possibilities
-are implemented. The first one being somewhat simpler and second one
-being more accurate. The first option is
+are implemented. The first one being somewhat simpler, whereas the second one
+is more accurate but requires a careful choice of parameters. The first 
+option is
 \begin{equation}
@@
 of the event rate spectrum. The second option is closer to an actual energy
-calibration error and is an approximation to the exact form
-\begin{equation}
-s_i(b) = \int_{(1+b)(E_i-\Delta E/2)}^{(1+b)(E_i+\Delta E/2)} dE' \quad
-\frac{dn_{\beta}^{\text{IT}}}{dE'} ( E')
-\end{equation}
-or something like this, to be continued ...
+calibration error, which basicaly amounts to replacing the reconstructed 
+energy $E'$ by $(1+b)\cdot E'$ and we use following approximation
+\begin{eqnarray}
+s_i(b)&=& (1+b)\cdot(s_{k+1}-s_k)\cdot\delta+s_i\,,\\
+\delta&=&b\cdot(i+\Delta E+ 1/2)+i\,,\nonumber\\
+k&=&\mod(\delta,1)\,.\nonumber
+\end{eqnarray}
+Special care is required at the boundaries when $k<1$ or $k+1>N_\mathrm{bins}$,
+since there $s_k$ or $s_{k+1}$ may not have been calculated. One possibility
+to deal with those cases is to assume that for those values of $k$ $s_k$ is
+zero. This is the default. In cases where $s$ at the boundary is however still
+sizeable this may lead to untolerable errors and subsequently to a wrong
+estimate of the impact of a calibration error. Therefore in those cases it
+is advisable to truncate the analysis range by a few bins at the boundary
+and therefore ensure in this way that only those $s_i$ are used whose index
+$k$ is within the range $1,\ldots, N_\mathrm{bins}-1$.
+
+In order to perform this fine tuning, it is possible to constrain 
+the energy range from which the data are taken for the 
+analysis\index{Energy window}. This is achieved by setting the energy window
+variable. The purpose is mainly to allow a fine tuning of the description
+of an energy calibration error. The default setting is that the energy window
+coincides with the range defined by the minimal and maximal reconstructed 
+energy. For the error dimension using the calibration method (C) it is 
+advisable to reduce the energy window by two or three bins at both the lower
+and upper end. As a rule of
+thumb consider the three standard deviations error on the calibration and
+evaluate the corresponding shift in $E'$ at the lower and upper energy limit 
+and reduce the analysis range by this amount.
 
@@
 used~\cite{Fogli:2002pt}\footnote{In fact the pull method
-was employed already in~\cite{SBvNF} before \cite{Fogli:2002pt} appeared.}. 
+was employed already in~\cite{Huber:2002mx} before 
+\cite{Fogli:2002pt} appeared.}. 
 In the  pull method $k$ systematical errors are included by introducing 
@@
 The user has the possibility to choose the set $\{\zeta_i\}$ of nuisance 
-parameters which are marginalized, this achieved with the error dimension 
-variable\index{Error dimension}, the different possibilities are shown in
-table~\ref{tab:error_dim}. If a parameter is desginated with $+$ it will
+parameters which are marginalized. This achieved with the error dimension 
+variable\index{Error dimension}, and the different possibilities are shown in
+table~\ref{tab:error_dim}. If a parameter is designated with $+$, it will
 be marginalized and therefore the corresponding error needs to have a non-zero
@@
 normalization free ($\sigma_a=\sigma_c=\infty$) and therefore counts only the 
-spectral information in the date, therefore
+spectral information in the data. As a consequence
 the settings for the normalization error will be ignored. It is furthermore
 possible to extend the set of error dimensions and thus the set of possible 
-systematical errors to be studied, but this requires proper C programing
-and recompiling \GLOBES. The interface for user defined error dimensions is
-described in detail in section~\ref{sec:ud_error_dim}.
+systematical errors to be studied, but this requires proper C programing, 
+compiling and linking with \GLOBES. The interface for user defined error 
+dimensions is described in detail in section~\ref{sec:ud_error_dim}.
 
@@
 
-In each rule it is possible to constrain the energy range from which the
-data are taken for the analysis\index{Energy window}. This achieved by 
-setting the energy window
-variable. The purpose is mainly to allow a fine tuning of the description
-
-of an energy calibration error. The default setting is that the energy window
-coincides with the range defined by the minimal and maximal reconstructed 
-energy. For the error dimension using the calibration method (C) it is 
-advisable to reduce the energy window by two or three bins at both the lower
-and upper end. When studying the impact of the energy calibration
-error in detail and it turns out not to be small, make sure that your result
-does not change by reducing the energy window by 1 or 2 bins.
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Energy resolution function}
@@
 use. This also naturally involves the issue how the actual convolution
-of the differntial event rate spectrum with $R_\beta^{\text{IT}}(E,E')$
+of the differential event rate spectrum with $R_\beta^{\text{IT}}(E,E')$
 is performed. First we will discuss the mathematical structure of
 the problem and then the various algorithms and possibilities which
-\GLOBES\ offers are introduced. In order to obtain the number of events 
+\GLOBES\ offers are introduced.
+
+Before going into the details we define that the range from $E_\mathrm{min}$
+to $E_\mathrm{max}$ is diveded in a certain number of bins $N_\mathrm{bins}$.
+These bins form a partition, which not necessarily is equidistant. The 
+partition is any case completely defined by specifying $E_\mathrm{min}$
+and $E_\mathrm{max}$ and the size of each bin  $\Delta E_i$. It is also
+straightforward to compute the midpoint of each bin $E_i$.
+In order to obtain the number of events 
 $n_i^c$ from one channel $c$ in one bin $i$ one has to solve this integral
@@
 \label{eq:events_bin}
-n_i^c=N/L^2\,\int_{E_i-\Delta E/2}^{E_i+\Delta E/2} dE' 
+n_i^c=N/L^2\,\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
 \quad \int dE \,\, \Phi^c(E)\,
@@
 \label{eq:kernel}
-K_i^c(E):=\int_{E_i-\Delta E/2}^{E_i+\Delta E/2} dE' 
+K_i^c(E):=\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
 \quad R^c(E,E')\,
@@
 by the usual numerical methods. However it turns out that this may be rather
-slow therefore we introduce two different approximation schemes which have
-a varying degree of acuracy and speed. The simplest, fastest and
-least acurate is algorithm one. The key idea is to choose a set of assumptions
+slow, therefore we introduce two different approximation schemes which have
+a varying degree of accuracy and speed.
+
+The simplest, fastest and
+least accurate is algorithm A. The key idea is to choose a set of assumptions
 which allow to evaluate the integrand of equation~\ref{eq:simple_int}
@@
 \label{eq:algo_one}
-n_i^c=N/L^2 \Delta E \sum_{j=1}^N \quad  \Phi^c(E_j)\,
+n_i^c=N/L^2 \Delta E_i \sum_{j=1}^N \quad  \Phi^c(E_j)\,
 P^c(E_j)\,
@@
 The advantage of using equation~\ref{eq:algo_one} is that all factors
-independent of the oscillation paramters have to be evaluated only at 
+independent of the oscillation parameters have to be evaluated only at 
 values of $E$ which are known in advance, \ie\ they can be put 
@@
 constant, \ie\ they do not change within one bin. This assumption seems to be
-very restrictive however whenever one is analyzing simulated 
+very restrictive, note however whenever one is analyzing simulated 
 data\footnote{simulated with the same algorithm} the errors
@@
 minimizing routines. The computation of the bin kernel $K_i^c$ is performed
-by \GLOBES\, to this end it requires the specification of the number of bins
-and the minimum and maximum energy. Furthemore following parameterization
+by \GLOBES. To this end it requires the specification of the number of bins
+ $N_\mathrm{bins}$
+and the minimum $E_\mathrm{min}$ and maximum energy $E_\mathrm{max}$. 
+Furthermore following parameterization
 for $R^c(E,E')$ is used
@@
 The parameters $\alpha, \beta$ and $\gamma$ are provided by the user.
-Furthemore the user has the possibility to provide $\epsilon(E')$ by specifying
-the value of  $\epsilon(E')$ at each $E_j$ since this algroithm assumes
-everything else to be constant within one bin, the  $\epsilon(E')$ are
-also taken as constant within one bin. In the actual implementation the sum
-in equation one only extends over those $E_j$ where $K(E_j)$ above a certain
-treshold, by default $10^{-7}$.
+Furthermore the user has the possibility to provide 
+$\epsilon(E')$ by specifying
+the value of  $\epsilon(E')$ at each $E_j$. Thus the  $\epsilon(E')$ are
+also taken as constant within one bin\footnote{since algorithm A assumes
+everything else to be constant within one bin}. In the actual implementation 
+the sum in equation one only extends over those $E_j$ where $K(E_j)$ 
+is above a certain threshold, by default $10^{-7}$. This threshold is 
+defined at compile time.
 
 The second available algorithm is more complicated, a little slower but
-much more accurate and much less susceptible to aliasing effects. The starting
+much more accurate and much less susceptible to aliasing effects
+\index{Aliasing}. The starting
 point again is equation~\ref{eq:simple_int}. The idea now is to
@@
 \begin{equation}
+\label{eq:sampling}
 f(E)=\sum_{k=-\infty}^{\infty}\,f(E_k) 
@@
 This the so called sampling 
-representation of the integrand $f(E)$ and $e(E)$ is the error term.
+representation of the integrand $f(E)$ and $e(E)$ is an error term.
+The error term vanishes if the Fourier transform $\hat f(\omega)$ of $f(E)$
+is zero outside the interval $[-\pi/h,\pi/h]$. Thus one suspects that
+the error term is small once the power beyond $\pi/h$ is low. Furthermore
+the sum in equation~\ref{eq:sampling} can be truncated to a few terms if
+$f(E)$ has a compact support, \ie\ it is different from zero only 
+on a finite interval. Thus the sampling representation is most useful
+for functions which have both a compact support in real \emph{and} Fourier
+space. This of course is impossible, however the function which comes
+most closely to this, is well known -- it is a Gau\ss ian. Our $f(E)$ is not
+exactly a Gau\ss ian  but the convolution of a top hat and a Gau\ss ian, thus
+its real and Fourier space components still behave rather Gau\ss ian 
+especially for large values of $E$ and $\omega$ respectively. The advantage of
+using the sampling representation is that the integration with respect to $E$
+is easily performed analytically
+  \begin{eqnarray}
+\label{eq:int_sampling}
+\int dE \quad f(E)&\simeq& \sum_{k=k_{l}}^{k_{u}}\,f(E_k) 
+\times \mathrm{Si}\left(\frac{\pi}{h}(E-E_k)\right)\,,\\
+\mathrm{Si}(x)&:=&\int_{0}^x dx'\quad \sin(x')/x'\,,\nonumber 
+\end{eqnarray}
+where $k_l$ and $k_u$ are suitable chosen bounds for the summation.
+
+In order to have a good compromise between speed an accuracy the choice
+of $h$, $k_l$ and $k_u$ is crucial. The choice of $h$ is governed by the
+need to sample $f(E)$ at a sufficient rate in order not to produce aliasing.
+For this purpose it useful to consider some properties of the power spectrum
+of $f(E)$. Provided the cross section and flux is smooth enough  the only
+high power components in $\hat f(\omega)$ are due to a fast oscillating
+$P(E)$. We can estimate the largest frequency $\omega_m$ quite easily, it is
+given by the largest $\Delta m^2_m$ and the lowest energy $E_l$
+\begin{equation}
+\omega_m \simeq \frac{\Delta m^2_m \cdot L}{E_l^2}
+\end{equation}
+On the other hand the finite energy resolution of any real detector anyhow
+will act as a low-pass filter eliminating any power beyond $\sim 1/\sigma(E)$.
+For now we assume that there are no frequencies which are not resolved by
+the detector thus the sampling frequency is determined by the energy 
+resolution of the detector. In that case the correct sampling 
+width $h$ is approximately
+\begin{equation}
+h\simeq 1/\sigma(E_l)\,.
+\end{equation}
+
+The range for the summation $k_l^i$ and $k_u^i$ is given by the range of 
+integration $a_i$ and $b_i$ in equation~\ref{eq:simple_int}, the integration 
+limits there are $0$ to $\infty$. For practical purposes however the range
+can be reduced to $a_i$ and $b_i$. They are  determined by requiring
+that the area under the bin kernel $K_i^c$ has reached a specific 
+value close to one,
+the so called confidence level $c$ (new name would be better). $1-c$ is a 
+small number and specifies the fraction of  events in bin which lie outside
+the summation range, usually $1-c=10^{-3}$ is sufficient. Furthemore the
+value of the bin kernel at the lower and upper boundary should
+be the same
+\begin{eqnarray}
+\int_{a_i}^{b_i} dE\quad K_i^c(E) &=& c\,,\\
+K_i^c(a_i)&=&K_i^c(b_i)\,.
+\end{eqnarray} 
+Knowing $a_i$ and $b_i$ it is easy to obtain $k_l$ and $k_u$ by requiring
+that the summation extends one bin beyond  each $a_i$ and $b_i$. Furthermore
+it is possible to specify an offset which extends the index range on each side
+by its value.
+
+Since the $a_i$'s and $b_i$'s are not known in advance it could be possible
+that $a_0$ and/or $b_{N_\mathrm{bins}}$ lie outside the computational domain,
+\eg\ no cross section or flux is defined at those energies. To remidy this 
+problem the computational domain can be specified. No calculation in \GLOBES\
+will leave this domain.
+
+In order to ensure that fast oscillating probablities do not lead to aliasing
+it is desirable to impose a low-pass filter already during the calculation
+of the probablities itself. This is possible and implemented has a highly
+experimental feature called ``filter''\index{Filter}. 
+The calculation of oscillation
+probabilities basically is a computation of phase differences. Restricting
+the maximum admissable size of those phase difference effectively filters
+the high frequency component of the oscillation probability. This idea is
+implemented according to
+\begin{eqnarray}
+\label{eq:filter_a}
+P_{\alpha\beta}(E)&=&\sum_{ij}
+U_{\alpha j} U^*_{\beta j} U^*_{\alpha i} U_{\beta i} 
+e^{-i\Phi_{ij}}\times 
+e^{ -\Phi_{ij}^2/\sigma_f(E)^2 }\,,
+\end{eqnarray}
+where $\Phi_{ij}:=\Delta m_{ij}^2 L/2E$ is the usual phase difference and
+the last term is a Gau\ss ian filter with width $\sigma_f(E)$. Choosing
+$\sigma_f(E):=\sigma_f^0 \cdot E$ ensures that this filter behaves 
+approximately like an energy resolution function with constant width 
+$\sigma_e=\sqrt{2}/\sigma_f^0$, \ie\
+\begin{equation}
+\label{eq:filter_b}
+\int d\tilde E\quad P(\tilde E) \frac{1}{\sigma_e\,\sqrt{2\pi}}\,
+e^{-\frac{(E-\tilde E)^2}{2\sigma^2_e}}\,.
+\end{equation}
+The equivalence of equation~\ref{eq:filter_a} and equation~\ref{eq:filter_b}
+is not obvious and connected to the properties of $P_{\alpha\beta}$, 
+see~\cite{Kiers:1996zj,Giunti:2003ax}. This feature works \emph{only} 
+for vacuum and constant denisties and controled
+by the filer state variable and $\sigma_f^0$ is set by the filter value 
+variable. Assuming that $P_{\alpha\beta}$ still is the source for the highest
+frequencies one can splitt the energy resolution function $\sigma(E)^c$ in
+two parts by
+\begin{equation}
+\sigma_c^2(E)=\underbrace{\sigma_c^2(E)-\sigma_c^2(E_\mathrm{min})}_
+{\tilde\sigma^2_c(E)}+\underbrace{\sigma_c^2(E_\mathrm{min})}_{\sigma_e^2}\,,
+\end{equation}
+where $\tilde\sigma_c(E)$ is used instead of $\sigma_c(E)$ in computing the
+smearing data and $\sigma_e$ is used to determine $\sigma_f^0$.
+
+Finally the user has the possibility to directly specify the smearing data.
+The format requires to provide $k_l^i$ and $k_u^i$ for each bin, the number of
+bins, the number of sampling points, the lower and upper limit of the sampling
+range. Furthemore the numbers $K_i(E_j)^c$ have to provide for all values
+in between $k_l^i$ and $k_u^i$. The convolution is then performed according to
+equation~\ref{eq:algo_one}.
 

### 2004-04-28 Documentation/part2.tex
Documentation part2.tex

@@
 
-\part{The Experiment Definition Module (\EDM )}
+\part{The Abstract Experiment Definition Language -- AEDL}
 
-\chapter{Basics: Source, Oscillation, Detection}
+\chapter{General Concept}
 
-Here comes how a experiment is defined (what is a channel, a rule, how are 
-they treated, ...). Splitting in source, oscillation, and detector, etc.
+The aim of AEDL is to describe a large variety of complex experiments
+by a limited number of parameters and to allow a representation of
+very different setups within one data structure, thus enabling a
+universal rate and $\chi^2$ computation. Usually in dealing with
+experiment simulations a new piece of code is written and compiled
+for each different experiments. In many cases even changes of some
+parameter like the number of bins requires the recompilation of
+large portions of the source code. This technique may however soon
+reach its limits in cases where the simulated experiments are rather complex
+and/or more than one kind of experiment is going to examined. Furthermore
+is is very difficult to assure the correctness of the results obtained
+in this way since every time a new piece of code is added to handle a new
+experiment new errors will be introduced.
 
-\chapter{Getting started: Editing existing experiments}
+Thus a general and flexible experiment description language is needed. 
+A neutrino experiment can be split into three parts: source, oscillation
+and detection. The sources within \GLOBES\ are assumed to be stationary
+point sources and each experiment has only one source. Specifically this
+does not allow to study cases where the time dependence of the neutrino flux
+is of physical significance like \eg\ in supernov\ae. It however is possible
+to study beams which have a bunch structure, since the time dependence of the
+neutrino source is physically only in as far important as it is used to
+suppress backgrounds. Also geometrical effects of a source distribution like
+\eg\ in the Sun can not described in this way. For this kind of sources
+the oscillation part of an 
+experiment is simply
+given by the baseline and the matter density profile.
 
-Here comes how to load and modify existing experiments;
-also: how does the main screen work and the most important functions at 
-exampled
+While
+it is comparatively simple to define a general neutrino source 
+and the oscillation physics, the general properties of a detector are more
+complicated. The basic assumption in building an abstract detector description
+is \emph{linearity}, \ie\ two neutrino events do not interfere with each other.
+Furthermore it is assumed that all information on the oscillation physics 
+is given by the \emph{reconstructed} flavour and energy of a 
+neutrino event. The term ``reconstructed'' implies that there is not a single
+well defined value for each event and variable but a distribution of possible
+values. This illustrated in figure~\ref{fig:distro} for the energy variable.
+The same is of course possible with the flavour variable. In that case however
+only discrete values are possible.
+%
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[width=0.6\textwidth]{mapping}
+\end{center}
+\caption{\label{fig:distro} A detector maps the true values into
+a distribution of reconstructed values of a parameter. Here the energy is 
+shown.}
+\end{figure}
+%
 
-\chapter{Features of the \EDM }
+This picture also can be formulated in a more mathematical language. Let $x$
+be the true variable and $x'$ the reconstructed variable then $f(x)$ is the
+distribution of true values and $p(x')$ is the distribution of reconstructed
+values. Then the detector function  $D(x,x')$ which describes 
+the mapping performed by the detector is defined by
+\begin{eqnarray}
+\label{eq:mapping}
+p(x')&=&\int dx\, f(x)\cdot D(x,x')\,.
+\end{eqnarray}
+Obviously equation~\ref{eq:mapping} can only describe the detector correctly
+when the condition of linearity is fulfilled. Within this model a detector
+is completely specified by a set of $D(E,E')$ for the energy variable $E$
+and a set $D(F,F')$ for the flavour variable $F$, where usually $D(E,E')$ also
+depends on the true flavour $F$ as well as $D(F,F')$ depends on the true energy
+$E$. Those sets of mapping functions usually will be obtained by a 
+full detector simulation and can be obtained by using as input 
+distribution $f(x)$ a delta distribution $\delta(x-x_0)$.
 
-\chapter{General concept}
+In order to allow a systematic experiment definition including various
+sources of systematical errors we use several layers. The first layer is
+the so called ``channel'', it is the link between the oscillation physics
+and the detector. A channel specifies how a certain neutrino type in the 
+source is mapped into a reconstructed neutrino type, \eg\ a muon neutrino
+oscillates into an electron neutrino and subsequently interacts via
+quasi-elastic charged current scattering. The measured energy and direction of
+the electron in the detector finally allow to reconstruct the neutrino energy.
+The channels are the building blocks for so called ``rules'' which
+contain information like the efficiencies and systematical errors. An example
+could look like this: the efficiency for quasi-elastics electron neutrino
+events is $0.4$. There is a fraction of $0.01$ of all neutral current events
+which are mis-identified as quasi-elastic electron neutrino events. 
+The neutral current fraction is only known within $10\%$ and there is
+an energy scale uncertainty of $100\,\mathrm{MeV}$. Thus a rule connects the
+event rates to the calculation of a $\chi^2$ which properly includes 
+systematical errors. This $\chi^2$ is then the starting point for the
+oscillation physics analysis.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\chapter{Getting started}
+
+As a first example we discuss how to define a neutrino factory experiment using
+AEDL. The first line of every experiment definition file is
+\begin{quote}
+{\tt !\%GLoBES}
+\end{quote}
+
+We tell \GLOBES\ that we want to use the built-in neutrino factory
+flux, by setting the beam type variable to $1$. The target mass 
+is $50\,\mathrm{kt}$ and a total of $ 4.264\cdot10^{21}$
+muons is stored. The data is divided into 20 energy bins ranging from
+$4\,\mathrm{GeV}$ to $50\,\mathrm{GeV}$. 
+\begin{quote}
+{\tt /* beam */}\\
+{\tt \$beam\_type = 1}\\
+{\tt \$target\_mass = 50}\\
+{\tt \$stored\_muons = 4.264e+21}\\
+{\tt \$numofbins = 20}\\
+{\tt \$emin = 4.0}\\
+{\tt \$emax = 50.0}
+\end{quote}
+
+Next a file has to be specified which contains the cross section we want to 
+use.
+\begin{quote}
+{\tt /* cross section */}\\
+{\tt cross(\#CC)<}\\
+{\tt \hspace*{3ex}@cross\_file = XCC.dat}\\
+{\tt >}
+\end{quote}
+The command {\tt cross} tells the parser that a cross section environment
+begins and has the name {\tt \#CC}. These names are used to refer later on
+to this specific environment and each name begins with a leading {\tt \#}.
+Whenever we need to specify a cross section from here on we can use {\tt \#CC}
+to refer to the cross section as given in the file {\tt XCC.dat}.
+
+Of course also the baseline and matter profile have to be specified, which
+ looks like this
+\begin{quote}
+{\tt /* baseline */}\\
+{\tt \$baseline = 3000.0}\\
+{\tt \$densitytab = \{3.5\}}\\
+{\tt \$lengthtab = \{3000.0\}}\\
+{\tt \$density\_error = 0.05}
+\end{quote}
+The curly brackets used for the definition of {\tt \$densitytab} and
+{\tt \$lengthtab} delimit as simple list of numbers -- here the lists
+have length $1$. {\tt \$density\_error} is the relative $1\,\sigma$ uncertainty
+on the density.
+
+The energy resolution function can easily be specified with
+\begin{quote}
+{\tt /* energy resolution */}\\
+{\tt energy(\#MINOS)<}\\
+{\tt \hspace*{3ex}@algorithm = 1}\\
+{\tt \hspace*{3ex}@bins = 20}\\
+{\tt \hspace*{3ex}@emin = 4.0}\\
+{\tt \hspace*{3ex}@emax = 50.0}\\
+{\tt \hspace*{3ex}@energy\_resolution = \{0.15,0.0,0.0\}}\\
+{\tt >}
+\end{quote}
+The {\tt energy} command opens the corresponding energy environment and
+the energy resolution function defined here is refered to by the name 
+{\tt \#MINOS}. Out of several possibilities it employs algorithm $1$,
+the simplest and fastest one. The actual energy resolution is specified
+by the energy resolution variable which is a list of three elements. Each 
+element is one parameter of a general resolution function as defined in 
+equation ~\ref{eq:sigma_e}.
+
+Now we define the one appearance and the corresponding disappearance 
+channel of a neutrino factory: 
+$\nu_e\rightarrow\nu_\mu$  and $\bar\nu_\mu\rightarrow\bar\nu_\mu$ 
+($\mu^+$ stored).
+\begin{quote}
+{\tt /* channels */}\\
+{\tt channel(\#appearance)<}\\
+{\tt \hspace*{3ex}@channel = +: +: electron: muon: \#CC: \#MINOS}\\
+{\tt >}\\
+{\tt channel(\#disappearance)<}\\
+{\tt \hspace*{3ex}@channel = +: -: muon: muon: \#CC: \#MINOS}\\
+{\tt >}
+\end{quote}
+The first $\pm$ selects the polarity, \ie\ wether $\mu^+$ or $\mu^-$ are 
+stored. The second $\pm$ determines whether neutrinos or anti-neutrinos are
+taken from the flux table. The third position defines the initial flavour
+and the forth position the final flavour, followed by the name of the cross
+section and energy resolution function.
+
+The last step is to assemble the channels into a rule.
+\begin{quote}
+{\tt /* rules */}\\
+{\tt rule(\#rule1)<}\\
+{\tt \hspace*{3ex}@signal = 0.45 @ \#appearance}\\
+{\tt \hspace*{3ex}@signalerror = 0.001 : 0.0001}\\
+{\tt \hspace*{3ex}@background = 1.0e-05 @ \#disappearance}\\
+{\tt \hspace*{3ex}@backgroundcenter = 1 : 0.0}\\
+{\tt \hspace*{3ex}@backgrounderror = 0.05 : 0.0001}\\
+{\tt \hspace*{3ex}@errordim = 0}\\
+{\tt \hspace*{3ex}@energy\_window = 4.0 : 50.0}\\
+{\tt >}
+\end{quote}
+The signal variable says that the signal in our experiment is the 
+channel named {\tt \#appearance} with an constant efficiency of $0.45$.
+The signal error variable has two components: the first one is the
+normalization error of the signal, here it is $0.1\%$. The second component
+signifies the energy calibration error of the signal, whose definition 
+can be found in section~\ref{sec:energy}. The background variable
+specifies the composition of the beam background and this case is
+given by the channel named {\tt \#disappearance}, \ie\ this are
+the muons whose charge has been mis-identified. They enter the calculation
+with an efficiency of $1\cdot 10^{-5}$. The background center variable allows
+to rescale the total background contribution which is especially useful if
+there are many background components. 
+The background error variable is defined in the same way as the signal error
+variable, \ie\ we have a $5\%$ background uncertainty.
+The error dimension variable selects how
+the systematical errors are treated according to table~\ref{tab:error_dim}.
+
+The above defined experiment represents a first version of a neutrino factory
+experiment. It still lacks the correct energy dependence of the efficiencies
+and of course the channels and rules for the operation with $\mu^-$ stored
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\chapter{Features of the  AEDL}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 in computing the normalization one has to know things like number of target
-particles per unit mass asf. which is not always straitforward. Thus a fudge
-factor will be usefule in any case. Right now however this is not the only 
+particles per unit mass asf. which is not always straightforward. Thus a fudge
+factor will be useful in any case. Right now however this is not the only 
 fudge factor since for some experiments the definition of their target mass
@@
 of the event rate spectrum. The second option is closer to an actual energy
-calibration error, which basicaly amounts to replacing the reconstructed 
+calibration error, which basically amounts to replacing the reconstructed 
 energy $E'$ by $(1+b)\cdot E'$ and we use following approximation
@@
 zero. This is the default. In cases where $s$ at the boundary is however still
-sizeable this may lead to untolerable errors and subsequently to a wrong
+sizeable this may lead to intolerable errors and subsequently to a wrong
 estimate of the impact of a calibration error. Therefore in those cases it
@@
 Before going into the details we define that the range from $E_\mathrm{min}$
-to $E_\mathrm{max}$ is diveded in a certain number of bins $N_\mathrm{bins}$.
+to $E_\mathrm{max}$ is divided in a certain number of bins $N_\mathrm{bins}$.
 These bins form a partition, which not necessarily is equidistant. The 
@@
 small number and specifies the fraction of  events in bin which lie outside
-the summation range, usually $1-c=10^{-3}$ is sufficient. Furthemore the
+the summation range, usually $1-c=10^{-3}$ is sufficient. Furthermore the
 value of the bin kernel at the lower and upper boundary should
@@
 that $a_0$ and/or $b_{N_\mathrm{bins}}$ lie outside the computational domain,
-\eg\ no cross section or flux is defined at those energies. To remidy this 
+\eg\ no cross section or flux is defined at those energies. To remedy this 
 problem the computational domain can be specified. No calculation in \GLOBES\
@@
 
-In order to ensure that fast oscillating probablities do not lead to aliasing
+In order to ensure that fast oscillating probabilities do not lead to aliasing
 it is desirable to impose a low-pass filter already during the calculation
-of the probablities itself. This is possible and implemented has a highly
+of the probabilities itself. This is possible and implemented has a highly
 experimental feature called ``filter''\index{Filter}. 
@@
 probabilities basically is a computation of phase differences. Restricting
-the maximum admissable size of those phase difference effectively filters
+the maximum admissible size of those phase difference effectively filters
 the high frequency component of the oscillation probability. This idea is
@@
 see~\cite{Kiers:1996zj,Giunti:2003ax}. This feature works \emph{only} 
-for vacuum and constant denisties and controled
+for vacuum and constant densities and controlled
 by the filer state variable and $\sigma_f^0$ is set by the filter value 
 variable. Assuming that $P_{\alpha\beta}$ still is the source for the highest
-frequencies one can splitt the energy resolution function $\sigma(E)^c$ in
+frequencies one can split the energy resolution function $\sigma(E)^c$ in
 two parts by
@@
 bins, the number of sampling points, the lower and upper limit of the sampling
-range. Furthemore the numbers $K_i(E_j)^c$ have to provide for all values
+range. Furthermore the numbers $K_i(E_j)^c$ have to provide for all values
 in between $k_l^i$ and $k_u^i$. The convolution is then performed according to
 equation~\ref{eq:algo_one}.
-
-\chapter{Creating new experiments}
-
-Most advanced step: How to create a new experiment from scratch .
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{User-defined systematics}
+\label{sec:ud_error_dim}.
 

### 2004-05-04 Documentation/part2.tex
missing files and plots added

@@
 {\tt \$target\_mass = 50}\\
-{\tt \$stored\_muons = 4.264e+21}\\
+{\tt \$time = 8}\\
+{\tt \$stored\_muons = 5.33e+20}\\
 {\tt \$numofbins = 20}\\
@@
 of the source. There are two principal ways to provide a flux:
-use a built-in source or a file provided by the user.
+use a built-in source or a file provided by the user. Built-in sources
+as well as user defined fluxes are referred to internally by the
+ {\tt \$beam\_type} variable. Certain values of  {\tt \$beam\_type} are
+pre-defined and used for the built-in fluxes, \eg\ {\tt \$beam\_type = 1}
+selects the neutrino factory flux.
 
@@
 this sources the energy of the parent particle has to be specified
-and the number of useful decays per year.
+and the number of useful decays per year. Thus the sequence for setting
+up a neutrino factory is
+\begin{quote}
+{\tt \$beam\_type = 1}\\
+{\tt \$parent\_energy = 50.0}\\
+{\tt \$stored\_muons = 5.33e+20}\\
+\end{quote}
+The values of {\tt \$parent\_energy} and {\tt emax} are always the same.
 
-In the latter case the fluxes are read from a file with seven rows and
+User defined fluxes are dealt with a little different at initialization
+and the value of {\tt \$beam\_type} is automatically assigned. This is
+necessary to avoid any problems with multiple experiments, since the fluxes
+are read only once. Reading a flux file looks like
+\begin{quote}
+{\tt flux(\#dummy)<}\\
+{\tt \hspace*{3ex}@flux\_file\_p = user\_file\_1.dat}\\
+{\tt \hspace*{3ex}@flux\_file\_m = user\_file\_2.dat}\\
+{\tt >}\\
+{\tt \$target\_power = 4.0 }
+\end{quote}
+
+Since many sources have two modes of operation, it is possible to define
+a pair of fluxes by using the polarity flag when loading the files. Usually
+files connected in this way use the suffixes ``plus'' and ``minus'' in the 
+file names. The individual members of the pair can be accessed by the polarity
+variable. Also the built-in fluxes provide this choice.
+The flux read by {\tt @flux\_file\_p} can be accesed by selecting positive
+polarity, whereas the flux read by {\tt @flux\_file\_m} is selected by a 
+negative value of the polarity.
+
+For sources based on $\pi$-decay and nuclear reactors the total power
+involved defines how many neutrinos are produced therefore this variable is
+used to define the luminosity. Furthermore it is useful to define 
+the operation time of an experiment in years. Note however that the 
+number of events used in analysis only depends on the product of
+\begin{equation}
+\mathrm{mass}\,\left[10^6\,\mathrm{kg}\right]\times \mathrm{time}
+\,\left[\mathrm{years}\right]\times\left\{ \begin{array}{c}
+\mathrm{source~power}\,\left[\mathrm{MWy}\right]\\
+\mathrm{useful~decays}\,\left[\mathrm{year}^{-1}\right]
+\end{array}\right.\,.
+\end{equation}
+
+The format of the flux files is seven rows and
 501 lines.
@@
 involves another fudge factor. 
-
-Since many sources have two modes of operation, it is possible to define
-a pair of fluxes by using the polarity flag when loading the files. Usually
-files connected in this way use the suffixes ``plus'' and ``minus'' in the 
-file names. The individual members of the pair can be accessed by the polarity
-variable. Also the built-in fluxes provide this choice.
-
-For sources based on $\pi$-decay and nuclear reactors the total power
-involved defines how many neutrinos are produced therefore this variable is
-used to define the luminosity. Furthermore it is useful to define 
-the operation time of an experiment in years. Note however that the 
-number of events used in analysis only depends on the product of
-\begin{equation}
-\mathrm{mass}\,\left[10^6\,\mathrm{kg}\right]\times \mathrm{time}
-\,\left[\mathrm{years}\right]\times\left\{ \begin{array}{c}
-\mathrm{source~power}\,\left[\mathrm{MWy}\right]\\
-\mathrm{useful~decays}\,\left[\mathrm{year}^{-1}\right]
-\end{array}\right.\,.
-\end{equation}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{equation} 
+The baseline is set by 
+\begin{quote}
+{\tt \$baseline = 3000.0 }
+\end{quote}
 
@@
 Earth\index{PREM}\index{Earth matter density}.
+\begin{quote}
+TBD
+\end{quote}
 
@@
 proportional to the number of layers.
+\begin{quote}
+TBD
+\end{quote}
 
@@
 of thicknesses and densities for the layers. 
+\begin{quote}
+{\tt \$densitytab=\{ 2.8, 3.5\}}\\
+{\tt \$lengthtab=\{ 1000.0, 2000.0\}}\\
+\end{quote}
+It is important that both list are of equal length and that the sum
+of the thicknesses given in  {\tt \$lengthtab} add up to the value of
+{\tt \$baseline}.
 
@@
 The most obvious property of the neutrino detection is the total
-active mass of the target as already defined in section~\ref{sec:source}.
+active mass of the target as already defined in section~\ref{sec:source} and
+its value is defined by 
+\begin{quote}
+{\tt \$target\_mass = 50.0 }
+\end{quote}
+Furthermore the running time has to be specified by
+\begin{quote}
+{\tt \$time = 8.0 }
+\end{quote}
 
@@
 range of values given in the file $0.0$ is returned.
+
+A cross section file is read by 
+\begin{quote}
+{\tt flux(\#name)<}\\
+{\tt \hspace*{3ex}@cross\_file =user\_file\_1.dat}\\
+{\tt >}
+\end{quote}  
+and can later be refered to by {\tt \#name}.
 
@@
 
-!!!!!!!!!!!!
-Do we need $\epsilon(E')$ ???
-!!!!!!!!!!!!
-
 Thus $R_\beta^{\text{IT}}(E,E')$ describes the energy response of 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Rules}
@@
 %
+
+Instead of using directly equation~\ref{eq:e_res} we will use a different
+definition concerning $\epsilon(E')$. In general  $\epsilon(E')$ has to be
+determined by means of a Monte Carlo simulation of the experiment. This
+usually involves anyhow a binning of the simulated events in the reconstructed
+energy $E'$, therefore one more easily can obtain  $\hat\epsilon_i$ defined
+by
+\begin{equation}
+\label{eq:post_smearing}
+\hat\epsilon_i \cdot \int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
+\quad R^c(E,E')\,.
+\end{equation}
+These  $\hat\epsilon_i$ are refered to as so called 
+``post-smearing'' efficiencies.
+
 The integration with respect to the reconstructed energy $E'$ can be
@@
 K_i^c(E):=\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
-\quad R^c(E,E')\,
-\epsilon^c(E')\,.
+\quad R^c(E,E')\,.
 \end{equation}

### 2004-05-06 Documentation/part2.tex
latest version of part2

@@
 We tell \GLOBES\ that we want to use the built-in neutrino factory
-flux, by setting the beam type variable to $1$. The target mass 
+flux originating from stored $\mu^+$'s. This is achieved by setting 
+the builtin variable to $1$.  Next the energy of the muons is specified
+to be $50\,\mathrm{GeV}$ by the parent energy variable. It is assumed
+that there will be $5.33\cdot 10^{20}$ useful muon decays per year
+and that this luminosity is available for $8$ years.
+\begin{quote}
+{\tt /* beam */}\\
+{\tt flux(\#mu\_plus)<\\
+\tb  @builtin = 1\\
+\tb  @parent\_energy = 50.0\\
+\tb  @stored\_muons = 5.33e+20\\
+\tb  @time = 8.0\\
+>}\\
+\end{quote}
+This specific flux can be later on refered to as {\tt \#mu\_plus}.
+
+\begin{quote}
+{\tt \$target\_mass = 50}\\
+{\tt \$numofbins = 20}\\
+{\tt \$emin = 4.0}\\
+{\tt \$emax = 50.0}
+\end{quote}
+The target mass 
 is $50\,\mathrm{kt}$ and a total of $ 4.264\cdot10^{21}$
@@
 $4\,\mathrm{GeV}$ to $50\,\mathrm{GeV}$.
-\begin{quote}
-{\tt /* beam */}\\
-{\tt \$beam\_type = 1}\\
-{\tt \$target\_mass = 50}\\
-{\tt \$time = 8}\\
-{\tt \$stored\_muons = 5.33e+20}\\
-{\tt \$numofbins = 20}\\
-{\tt \$emin = 4.0}\\
-{\tt \$emax = 50.0}
-\end{quote}
 
@@
 {\tt cross(\#CC)<}\\
-{\tt \hspace*{3ex}@cross\_file = XCC.dat}\\
+{\tt \tb @cross\_file = XCC.dat}\\
 {\tt >}
@@
 {\tt energy(\#MINOS)<}\\
-{\tt \hspace*{3ex}@algorithm = 1}\\
-{\tt \hspace*{3ex}@bins = 20}\\
-{\tt \hspace*{3ex}@emin = 4.0}\\
-{\tt \hspace*{3ex}@emax = 50.0}\\
-{\tt \hspace*{3ex}@energy\_resolution = \{0.15,0.0,0.0\}}\\
+{\tt \tb @type = 1}\\
+{\tt \tb @sigma\_e = \{0.15,0.0,0.0\}}\\
 {\tt >}
@@
 {\tt channel(\#appearance)<}\\
-{\tt \hspace*{3ex}@channel = +: +: electron: muon: \#CC: \#MINOS}\\
+{\tt \tb @channel = \#mu\_plus: +: electron: muon: \#CC: \#MINOS}\\
 {\tt >}\\
 {\tt channel(\#disappearance)<}\\
-{\tt \hspace*{3ex}@channel = +: -: muon: muon: \#CC: \#MINOS}\\
+{\tt \tb @channel = \#mu\_plus: -: muon: muon: \#CC: \#MINOS}\\
 {\tt >}
 \end{quote}
-The first $\pm$ selects the polarity, \ie\ wether $\mu^+$ or $\mu^-$ are 
-stored. The second $\pm$ determines whether neutrinos or anti-neutrinos are
+The first element is the name of the flux as defined previously, here we use
+the flux named {\tt \#mu\_plus}. The second element $\pm$ determines whether 
+neutrinos or anti-neutrinos are
 taken from the flux table. The third position defines the initial flavour
@@
 {\tt rule(\#rule1)<}\\
-{\tt \hspace*{3ex}@signal = 0.45 @ \#appearance}\\
-{\tt \hspace*{3ex}@signalerror = 0.001 : 0.0001}\\
-{\tt \hspace*{3ex}@background = 1.0e-05 @ \#disappearance}\\
-{\tt \hspace*{3ex}@backgroundcenter = 1 : 0.0}\\
-{\tt \hspace*{3ex}@backgrounderror = 0.05 : 0.0001}\\
-{\tt \hspace*{3ex}@errordim = 0}\\
-{\tt \hspace*{3ex}@energy\_window = 4.0 : 50.0}\\
+{\tt \tb @signal = 0.45 @ \#appearance}\\
+{\tt \tb @signalerror = 0.001 : 0.0001}\\
+{\tt \tb @background = 1.0e-05 @ \#disappearance}\\
+{\tt \tb @backgroundcenter = 1 : 0.0}\\
+{\tt \tb @backgrounderror = 0.05 : 0.0001}\\
+{\tt \tb @errordim = 0}\\
+{\tt \tb @energy\_window = 4.0 : 50.0}\\
 {\tt >}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Syntax and types of  AEDL}
+\label{sec:syntax}
+
+ The first eight characters have to be {\tt \%!GLoBES} 
+ this serves to protect from parsing MBytes of chunk
+ and cluttering the screen with error messages.
+
+Comments
+ are entered like this one -- everything between the 
+ delimiters is ignored. 
+\begin{quote}
+{\tt /* This starts a comment\\
+ and here the comment ends */
+}
+\end{quote}
+thus the comments in AEDL work like in C
+
+There are so called  named quantities, they all start with {\tt \$} 
+and are predefined. They are checked for being within range, \eg\ {\tt 
+\$bins} can be only in between 0 and 150. If you give a {\tt float} where
+ an {\tt int} is expected the {\tt float} gets converted to an {\tt int} 
+in the usual way in C.  We have named scalar quantities
+\begin{quote}
+{\tt
+\$bins = 10\\
+\$baseline = 1200.0
+}
+\end{quote}
+as well as named simple lists
+\begin{quote}
+{\tt
+\$densitytab=\{1.0,2.2343,3.3432\}
+}
+\end{quote}
+
+Since there are some very special data to parse, some special
+provisions have been made so called named environments. This is illustrated 
+with the channel definition part
+\begin{quote}
+{\tt channel(\#ch1)<\\
+\tb  $\ldots$\\
+>
+}
+\end{quote}
+
+The first part tells the type of environment, here it is {\tt channel}. 
+There are the following types 
+\begin{quote}
+{\tt flux\\
+cross\\
+channel\\
+energy\\
+rule
+}
+\end{quote}
+Besides the type there is a user-defined name beginning with {\tt \#}, in the
+above example {\tt \#ch1}, which can be used later on to refer to the 
+channel defined in between {\tt <$\ldots$>}. Those names are so called
+automatic variables. They start all with {\tt \#} and do not behave as 
+usual variables. New names can only be
+created within the above construction. Only names which have been created
+can be refered to. Using the same name for different types of environments 
+does not work.  Each type has its own counter, \eg\ for channel the counter is 
+{\tt numofchannels}. The counter keeps track of how many different names 
+there are for one type, \ie\ they count the number of channels, 
+the number of rules, the number of energy resolution functions asf. 
+The value of those automatic variables is the number which can be used 
+to refer to them in your C code. You can create a variable for later use by
+\begin{quote}
+{\tt    channel(\#ch2)<>}
+\end{quote}
+Now you can refer to {\tt \#ch2} without any trouble and you now could define
+a rule using {\tt \#ch2}. Later on of course you still have to specify the
+actual channel as described in section~\ref{sec:channel}.
+
+Variables beginning with {\tt @} are named and can be used only within the 
+appropriate type of environment, since we have to keep track where 
+they belong to. Moreover many of them have a special syntax, like 
+{\tt @channel}.
+
+You can have several experiments in one file, just separate the different
+ experiments by 
+\begin{quote}
+{\tt    \#NEXT\#}
+\end{quote}
+This command resets the counters for channels, rules and energy resolution
+functions. All variables have their scope only within a region delimited
+by either {\tt \%!GLoBES, \#NEXT\#} or {\tt EOF}. The only exception are 
+the counters defined with {\tt cross} and {\tt flux}. They are valid for 
+the complete session, since you never must overload the flux and 
+cross section information. This ensures  a consistent treatment of various 
+experiments in one file - there is no  shadowing of any variables.
+
+Using multiple files to keep things clean and clear is supported by 
+ an {\tt include} command. Includes can be nested up 
+ to {\tt MAX\_INCLUSION\_DEPTH}, which is set to 10. Error reporting works 
+ also for nested includes. The included file is not required to begin 
+ with {\tt \%!GLoBES} to facilitate cut and paste
+\begin{quote}
+{\tt include ./file\_1}
+\end{quote}
+
+Thus it is possible and advised to use constructs as the following
+\begin{quote}
+{\tt    include NuFact.gls\\
+        \#NEXT\#\\
+        include JHFHK.gls
+}
+\end{quote}
+to define a combined analysis of the experiments defined in the files 
+{\tt NuFact.gls} and {\tt JHFHK.gls}. Even if you use the named 
+variable {\tt \#CC} in both experiments, but the cross section data used 
+is different (\eg\ because of different target nuclei), the correct 
+cross section data is used for each of the experiments, since {\tt \#CC} 
+in {\tt NuFact.gls} has the value $0$ and
+in {\tt JHFHK.gls} it has the value $1$, thus no shadowing occurs.
+
+Furthermore you can store numbers in variables
+\begin{quote}
+{\tt
+Pi = 3.141
+}
+\end{quote}
+And additionally we have some simple calculator functionality like
+\begin{quote}
+{\tt
+Pi+1\\
+\verb+Pi^2+\\
+sin(Pi/2)\\
+}
+\end{quote}
+The following mathematical functions from {\tt <math.h>} are available 
+{\tt sin}, {\tt cos}, {\tt atan}, {\tt ln}, {\tt exp}, {\tt sqrt}. 
+Those calculations work everywhere, where
+a simple number could stand. However threading over lists does not
+ work, \eg\ {\tt sin(\{1,2,3\})} will not work.
+
+One nasty thing is that a line feed character \verb+\n+ is necessary at
+ the end of input, alternatively you can put a comment at the end.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Source properties}
@@
 of the source. There are two principal ways to provide a flux:
-use a built-in source or a file provided by the user. Built-in sources
-as well as user defined fluxes are referred to internally by the
- {\tt \$beam\_type} variable. Certain values of  {\tt \$beam\_type} are
-pre-defined and used for the built-in fluxes, \eg\ {\tt \$beam\_type = 1}
-selects the neutrino factory flux.
-
-The built-in source is muon decay, since the neutrino
-spectra and flavour composition are known analytically in that case. For 
-this sources the energy of the parent particle has to be specified
+use a built-in source or a file provided by the user. In both cases
+a flux is defined by a named environment.
+\begin{quote}
+  {\tt flux(\#name)<\\
+\tb $\ldots$\\
+>}
+\end{quote}
+
+In case a built-in source is used the user has to specify which
+built-in spectrum is used. Currently two built-in fluxes are available:
+$\mu^+$-decay ({\tt @builtin = 1}) and $\mu^-$-decay ({\tt @builtin = 2}).
+In any case  the energy of the parent particle has to be specified
 and the number of useful decays per year. Thus the sequence for setting
 up a neutrino factory is
-\begin{quote}
-{\tt \$beam\_type = 1}\\
-{\tt \$parent\_energy = 50.0}\\
-{\tt \$stored\_muons = 5.33e+20}\\
-\end{quote}
-The values of {\tt \$parent\_energy} and {\tt emax} are always the same.
-
-User defined fluxes are dealt with a little different at initialization
-and the value of {\tt \$beam\_type} is automatically assigned. This is
-necessary to avoid any problems with multiple experiments, since the fluxes
-are read only once. Reading a flux file looks like
-\begin{quote}
-{\tt flux(\#dummy)<}\\
-{\tt \hspace*{3ex}@flux\_file\_p = user\_file\_1.dat}\\
-{\tt \hspace*{3ex}@flux\_file\_m = user\_file\_2.dat}\\
-{\tt >}\\
-{\tt \$target\_power = 4.0 }
-\end{quote}
-
-Since many sources have two modes of operation, it is possible to define
-a pair of fluxes by using the polarity flag when loading the files. Usually
-files connected in this way use the suffixes ``plus'' and ``minus'' in the 
-file names. The individual members of the pair can be accessed by the polarity
-variable. Also the built-in fluxes provide this choice.
-The flux read by {\tt @flux\_file\_p} can be accesed by selecting positive
-polarity, whereas the flux read by {\tt @flux\_file\_m} is selected by a 
-negative value of the polarity.
+{\tt flux(\#mu\_plus)<\\
+\tb  @builtin = 1\\
+\tb  @parent\_energy = 50.0\\
+\tb  @stored\_muons = 5.33e+20\\
+\tb  @time = 8.0\\
+>}\\
+
+Alternatively user defined fluxes may used like this
+\begin{quote}
+{\tt flux(\#user)<}\\
+{\tt \tb @flux\_file = user\_file\_1.dat\\
+\tb @time = 2.0\\
+\tb @power = 4.0\\
+\tb @norm = 1e+8}\\
+{\tt >}
+\end{quote}
+
+The norm variable allows to set the conversion factor from the fluxes in
+the file to the units in \GLOBES. The power variable allows to
+specify a target or reactor power. The units used for power are determined
+by the units in the flux file and by the value of the norm variable. It is
+recommended to use $\mathrm{MW}$ in case a proton based beam is used and
+$\mathrm{GW}_\mathrm{thermal}$ for the reactor experiments.
 
@@
 range of values given in the file $0.0$ is returned. The units used for the
-fluxes are not specified and right now each flux has its own fudge factor
-which is hard encoded in the file {\tt nufact.c}. The problem is that
+fluxes are not specified and right now each flux has its own normalization 
+factor as specified by {\tt norm}. The problem is that
 in computing the normalization one has to know things like number of target
-particles per unit mass asf. which is not always straightforward. Thus a fudge
-factor will be useful in any case. Right now however this is not the only 
-fudge factor since for some experiments the definition of their target mass
-involves another fudge factor. 
+particles per unit mass asf. which is not always straightforward. The fluxes
+will be rescaled by $1/L^2$, thus the normalization must contain a factor
+$L_0^2$, when $L_0$ is the distance from the source for which the flux is
+given.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 {\tt \$target\_mass = 50.0 }
-\end{quote}
-Furthermore the running time has to be specified by
-\begin{quote}
-{\tt \$time = 8.0 }
 \end{quote}
@@
 {\tt flux(\#name)<}\\
-{\tt \hspace*{3ex}@cross\_file =user\_file\_1.dat}\\
+{\tt \tb @cross\_file =user\_file\_1.dat}\\
 {\tt >}
@@
 \,V_\beta(\hat{E}-E')\,, 
-\end{eqnarray}
+e\end{eqnarray}
 
@@
 in section~\ref{sec:energy}.
+
+The number of events in one channel and  in one bin\index{Bin} $i$ is given by
+\begin{equation}
+\label{eq:channel}
+n_i^c=\int_{E_i-\Delta E/2}^{E_i+\Delta E/2} dE' \quad
+\frac{dn_{\beta}^{\text{IT}}}{dE'} (E')
+\end{equation}
+where $c$ is the channel index.
+Note that the events are binned according to their \emph{reconstructed} energy.
+In order to perform the computation \GLOBES\ needs to know the flux,
+the CP-sign of the initial state, the initial flavour, the final flavour,
+the cross section and the energy resolution function. Employing the notation
+of section~\ref{sec:energy} the full formula is
+
+Thus the actual definition of a channel looks like
+\begin{quote}
+{\tt channel(\#channel\_1)<\\
+\tb @channel = \#flux : $+$: muon: muon: \#cross: \#energy\\
+>}
+\end{quote}
+Furthermore is is possible to define energy dependent efficiencies and
+backgrounds (non-beam). These additional components can be either added
+before the convolution in with $R(E,E')$ is done, \ie\ the integration with
+respect to $E$ in equation~\ref{eq:e_res}. This is called 
+{\tt @pre\_smearing\_$\ldots$}. The pre-smearing efficiencies and backgrounds
+have to have as many elements as there are sampling points 
+(see section~\ref{sec:energy}). In same cases the number of sampling points
+is determined automatically by \GLOBES and the number of elements in the 
+pre-smearing vectors has to be adjusted in the second run.
+Of course, it is possible after all
+integrations been performed to have similar factors  and this is called  
+{\tt @post\_smearing\_$\ldots$}. All post-smearing factors have to have as 
+many elements as there are bins. Efficiencies are multiplicative and their
+default value is $1$ whereas backgrounds are additive and thus their default
+value is $0$. Thus a more elaborate channel looks like
+\begin{quote}
+{\tt channel(\#channel\_1)<\\
+\tb @channel = \#flux : $+$: muon: muon: \#cross: \#energy\\
+\tb @pre\_smearing\_background = \{1,2,3,4,5,6,7,8,9,10\}\\
+\tb @post\_smearing\_efficiencies = \{0.1,0.2,0.3,0.4,0.5\}\\
+>}
+\end{quote}
+This experiment obviously uses $10$ sampling points and $5$ bins.
 
@@
 signal and background event numbers for each bin.  
-The number of events in one channel and  in one bin\index{Bin} $i$ is given by
-\begin{equation}
-n_i^c=\int_{E_i-\Delta E/2}^{E_i+\Delta E/2} dE' \quad
-\frac{dn_{\beta}^{\text{IT}}}{dE'} (E')
-\end{equation}
-where $c$ is the channel index.
-Note that the events are binned according to their \emph{reconstructed} energy.
 The signal $s_i$ in one bin for a rule now can be composed out of 
@@
 \end{equation}
+This basic building blocks of a rule are specified by
+\begin{quote}
+{\tt \tb @signal = 0.5 @ \#channel\_1\\
+\tb @background = 0.001 @ \#channel\_2 :  0.005 @ \#channel\_3
+}
+\end{quote}
+
 Quite often it is not sufficient to have constant efficiencies and therefore
@@
 s_i(b)&=& (1+b)\cdot(s_{k+1}-s_k)\cdot\delta+s_i\,,\\
-\delta&=&b\cdot(i+\Delta E+ 1/2)+i\,,\nonumber\\
-k&=&\mod(\delta,1)\,.\nonumber
+\delta&=&b\cdot(i+t_0+ 1/2)+i\,,\nonumber\\
+k&=&\mod(\delta,1)\,,\nonumber\\
+t_0&=&E_\mathrm{min}/\Delta E_0\,.\nonumber
 \end{eqnarray}
@@
 analysis\index{Energy window}. This is achieved by setting the energy window
-variable. The purpose is mainly to allow a fine tuning of the description
+variable. 
+\begin{quote}
+{\tt 
+\tb @energy\_window = 4.0 : 50.0 
+}
+\end{quote}
+The purpose is mainly to allow a fine tuning of the description
 of an energy calibration error. The default setting is that the energy window
@@
 thus for the background normalization $c$ a value has to be specified in 
-\emph{all} cases.
+\emph{all} cases. The central values for the normalization and the 
+corresponding central value of tilt/calibration are regarded
+as a pair. The errors are treated in the same way.
+\begin{quote}
+{\tt
+\tb @signalerror =       0.001  :       0.01\\
+\tb @backgroundcenter =  0.1 :       0.0\\
+\tb @backgrounderror =   0.001 :       0.01
+}
+\end{quote}
+There is no {\tt @signalcenter} since by default the central value for the
+signal normalization is $1$ and the central value for the tilt/calibration 
+is $0$.
 
@@
 
+The error dimensions can be chosen for each rule by
+\begin{quote}
+{\tt 
+\tb @errordim = 1
+}
+\end{quote}
+
+Next we can assemble the various pieces into one rule
+\begin{quote}
+{\tt rule(\#rule\_1)<\\
+\tb @signal = 0.5 @ \#channel\_1\\
+\tb @background = 0.001 @ \#channel\_2 :  0.005 @ \#channel\_3\\
+\tb @signalerror =       0.001  :       0.01\\
+\tb @backgroundcenter =  0.1 :       0.0\\
+\tb @backgrounderror =   0.001 :       0.01\\
+\tb @errordim = 1\\
+\tb @energy\_window = 4.0 : 50.0\\ 
+>}
+\end{quote}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 and $E_\mathrm{max}$ and the size of each bin  $\Delta E_i$. It is also
-straightforward to compute the midpoint of each bin $E_i$.
+straightforward to compute the midpoint of each bin $E_i$. This information
+has to be provided by the user. Equidistant bins are obtained by
+\begin{quote}
+{\tt
+\$emin = 4.0\\
+\$emax = 50.0\\
+\$bins = 20
+}
+\end{quote}
+Another possibility is to specify arbitrary bin sizes and to leave out
+the number of bins itself, by
+\begin{quote}
+{\tt
+\$emin = 4.0\\
+\$emax = 50.0\\
+\$binsize = \{  15.0 , 5.0 , 20.0, 6.0 \} 
+}
+\end{quote}
+
 In order to obtain the number of events 
@@
 These  $\hat\epsilon_i$ are refered to as so called 
-``post-smearing'' efficiencies.
-
+``post-smearing'' efficiencies and set within the corresponding {\tt channel}
+environment.
 The integration with respect to the reconstructed energy $E'$ can be
@@
 slow, therefore we introduce two different approximation schemes which have
-a varying degree of accuracy and speed.
-
-The simplest, fastest and
+a varying degree of accuracy and speed. In any case each scheme to evaluate this
+integral involves a the evaluation of the integrand at fixed sampling points.
+These sampling points can be either defined by the user or be computed by
+\GLOBES, the choice between these possibilities also depends on the algorithm
+which is used for the actual calculation, as explained in the following.
+If the user chooses to set the sampling points the procedure is analogous
+to the one setting the bins. For equidistant sampling points use
+\begin{quote}
+{\tt
+\$simbins = 20\\
+\$simtresh =          4.0\\
+\$simbeam =         50.0
+}
+\end{quote}
+The advantage in setting the sampling points in this way is that in case many
+different energy resolution functions are used the user has the control
+how the different sampling requirements are fulfilled. In case a user-defined
+smear matrix is used this data is mandatory.
+
+All energy resolution function are defined within an {\tt energy} environment
+and can be refered to by {\tt \#name}.
+\begin{quote}
+  {\tt energy(\#name)<\\
+\tb $\ldots$\\
+>}
+\end{quote}
+
+First we discuss the builtin algorithms. The simplest, fastest and
 least accurate is algorithm A. The key idea is to choose a set of assumptions
@@
 aliasing effects may occur which usually lead to a failure of one of the 
-minimizing routines. The computation of the bin kernel $K_i^c$ is performed
-by \GLOBES. To this end it requires the specification of the number of bins
+minimizing routines. This algorithm is selected by
+\begin{quote}
+{\tt \tb @type = 1}
+\end{quote}
+The computation of the bin kernel $K_i^c$ is performed
+by \GLOBES. To this end it requires  the number of bins
  $N_\mathrm{bins}$
@@
 \end{equation} 
-where the energy resolution $\sigma(E)$ is defined by
+There are several energy resolution functions available and by default
+{\tt \#standard} is used. They can be chosen
+by
+\begin{quote}
+{\tt \tb @sigma\_function = \#standard}
+\end{quote}
+where the energy resolution function named {\tt \#standard} is defined by
 \begin{equation}
@@
 The parameters $\alpha, \beta$ and $\gamma$ are provided by the user.
-Furthermore the user has the possibility to provide 
-$\epsilon(E')$ by specifying
-the value of  $\epsilon(E')$ at each $E_j$. Thus the  $\epsilon(E')$ are
-also taken as constant within one bin\footnote{since algorithm A assumes
-everything else to be constant within one bin}. In the actual implementation 
+\begin{quote}
+{\tt \tb @sigma\_e = \{ 0.15, 0.0, 0.0\}}
+\end{quote}
+The other possible choice for {\tt @sigma\_function} is {\tt \#inverse\_beta}
+which has only one parameter $\alpha$ and is defined by
+\begin{equation}
+\sigma(E)= \left\{\begin{array}{cl}
+ \alpha \cdot \sqrt{1000}^{-1}\,\sqrt{x-8\cdot10^{-4}}\,,&\mathrm{for}\,\, 
+x>1.8\cdot10^{-3}\\
+5\cdot10^{-5} \,,&\mathrm{for}\,\, x \leq 1.8\cdot10^{-3}
+\end{array} \right.
+\end{equation}
+
+In the actual implementation 
 the sum in equation one only extends over those $E_j$ where $K(E_j)$ 
 is above a certain threshold, by default $10^{-7}$. This threshold is 
-defined at compile time.
+defined at compile time. Thus a complete energy resolution definition of this 
+type looks like
+\begin{quote}
+{\tt energy(\#name)<\\
+\tb @type = 1\\
+\tb @sigma\_function = \#standard\\
+\tb @sigma\_e = \{ 0.15 ,0.0 ,0.0\}\\
+>
+}
+\end{quote}
 
@@
 in between $k_l^i$ and $k_u^i$. The convolution is then performed according to
-equation~\ref{eq:algo_one}.
+equation~\ref{eq:algo_one}. Thus a complete example for this case would be
+\begin{quote}
+{\tt
+\$simbins = 20\\
+\$simtresh =          4.0\\
+\$simbeam =         50.0\\
+}
+
+{\tt 
+energy(\#name)<\\
+\tb @energy =   \{0,2, 0.8634265, 0.0682827,     4e-06\}:\\
+\tb\tb \{0,4, 0.1507103, 0.6965592, 0.1507103,   0.00101,     1e-07\}:\\
+\tb\tb $\ldots$\\
+>
+}
+\end{quote}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2004-05-10 Documentation/part2.tex
*** empty log message ***

@@
 
-\chapter{General Concept}
-
-The aim of AEDL is to describe a large variety of complex experiments
-by a limited number of parameters and to allow a representation of
-very different setups within one data structure, thus enabling a
-universal rate and $\chi^2$ computation. Usually in dealing with
-experiment simulations a new piece of code is written and compiled
-for each different experiments. In many cases even changes of some
-parameter like the number of bins requires the recompilation of
-large portions of the source code. This technique may however soon
-reach its limits in cases where the simulated experiments are rather complex
-and/or more than one kind of experiment is going to examined. Furthermore
-is is very difficult to assure the correctness of the results obtained
-in this way since every time a new piece of code is added to handle a new
-experiment new errors will be introduced.
-
-Thus a general and flexible experiment description language is needed. 
-A neutrino experiment can be split into three parts: source, oscillation
-and detection. The sources within \GLOBES\ are assumed to be stationary
-point sources and each experiment has only one source. Specifically this
-does not allow to study cases where the time dependence of the neutrino flux
-is of physical significance like \eg\ in supernov\ae. It however is possible
-to study beams which have a bunch structure, since the time dependence of the
-neutrino source is physically only in as far important as it is used to
-suppress backgrounds. Also geometrical effects of a source distribution like
-\eg\ in the Sun can not described in this way. For this kind of sources
-the oscillation part of an 
-experiment is simply
-given by the baseline and the matter density profile.
-
-While
-it is comparatively simple to define a general neutrino source 
-and the oscillation physics, the general properties of a detector are more
-complicated. The basic assumption in building an abstract detector description
-is \emph{linearity}, \ie\ two neutrino events do not interfere with each other.
+\chapter{Getting started}
+
+Here the general concept of the AEDL is described and illustrated by an example. In addition, a short introduction to the syntax of the AEDL is given.
+
+\section{General concept of the experiment simulation}
+
+The goal of AEDL is to describe a large number of complex and very 
+different experiments by a limited number of parameters. It allows a
+representation of very different setups within one data structure, and thus implements universal rate and $\chi^2$ computation methods.
+For experiment simulations, a new piece of code is usually written and compiled
+for each different experiment. In many cases, even parameter changes, such as
+the number of bins, requires the recompilation of the source code. 
+However, such a technique soon reaches its limits when the simulated experiments are rather complex, or more than one type of experiment is studied simultaneously.
+Furthermore, it is very difficult to verify the correctness of the obtained results, since every time a new piece of code is added to
+deal with a new experiment type, new errors will be introduced.
+
+Thus, a general and flexible experiment description language is needed.  
+The description of a neutrino experiment can be split into three parts: Source, oscillation, and detection. The neutrino sources within \GLOBES\ 
+are assumed to be stationary point sources, where each experiment has only 
+one source. This restricts the classes of neutrino sources which can be studied with \GLOBES :
+\begin{itemize}
+\item
+ Experiments using many point-like sources can only be approximated. One example are reactor experiments using many distant reactor blocks.
+\item
+ Geometrical effects of a source distribution, such as in the sun or the atmosphere, can not be described.
+\item
+ Sources with a physically significant time dependencies  can not be studied, such as  supernov\ae. It is, however, possible
+to study beams with bunch structure, since the time dependence of the
+neutrino source is physically only important to suppress backgrounds. 
+\end{itemize}
+
+The description of the neutrino oscillation physics is, at least numerically, relatively simple.
+We use the {\em evolution operator method} to compute the neutrino oscillation probabilities and divide the matter density profile into layers with constant matter densities.
+For each of these layers, the Hamiltonian in matter is diagonalized in order to propagate the neutrino transition amplitudes.
+Finally, the transition probability is obtained by the absolute square of the total neutrino transition amplitudes.
+Depending on the precision of the studied experiment, this approach turns out to be precise enough in Earth matter even for a small number matter density steps.
+Since we allow an uncertainty of the matter density profile, it is, in fact, in most cases sufficient to consider only one density step with the average matter density together with a matter density uncertainty~\cite{Ohlsson:2003ip}.
+Note that this approach may not be applicable to quickly varying extraterrestrial matter density profiles.
+
+While it is comparatively simple to define a general neutrino source 
+and to compute the oscillation physics, the general properties of a detector simulation are much more complicated.
+The basic assumption in building an abstract detector description is \emph{linearity}, \ie , that two neutrino events do not interfere with each other.
 Furthermore it is assumed that all information on the oscillation physics
-is given by the \emph{reconstructed} flavour and energy of a 
-neutrino event. The term ``reconstructed'' implies that there is not a single
-well defined value for each event and variable but a distribution of possible
-values. This illustrated in figure~\ref{fig:distro} for the energy variable.
-The same is of course possible with the flavour variable. In that case however
-only discrete values are possible.
+is given by the \emph{reconstructed} flavor and energy of a 
+neutrino event.
+The term ``reconstructed'' implies that the well-defined energy of the incident neutrino, which can not be directly observed, translates via secondary particles and the detection properties into a distribution of possible energy values.
+This process is illustrated in \figu{distro} for the energy variable.
 %
-\begin{figure}[h!]
+\begin{figure}[ht]
 \begin{center}
@@
 \end{center}
-\caption{\label{fig:distro} A detector maps the true values into
-a distribution of reconstructed values of a parameter. Here the energy is 
-shown.}
+\caption{\label{fig:distro} A detector maps a true parameter value onto
+a distribution of reconstructed parameter values. This is illustrated here for there energy.}
 \end{figure}
 % 
-
-This picture also can be formulated in a more mathematical language. Let $x$
-be the true variable and $x'$ the reconstructed variable then $f(x)$ is the
-distribution of true values and $p(x')$ is the distribution of reconstructed
-values. Then the detector function  $D(x,x')$ which describes 
-the mapping performed by the detector is defined by
+The same, in principle, applies to the nature of the neutrino flavor.
+However, in this case only discrete values are applicable.
+Note that the reconstructed neutrino energy and the neutrino flavor are the only observables in \GLOBES .
+
+This picture can also be formulated in a more mathematical way.
+Let us define $x$ as the true parameter value and $x'$ as the reconstructed parameter value.
+Similarly, $f(x)$ is the distribution of true parameters values and $p(x')$ is the distribution of reconstructed parameter values.
+Then the detector function  $D(x,x')$, which describes the mapping performed by the detector, is given by
 \begin{eqnarray}
-\label{eq:mapping}
+\label{equ:mapping}
 p(x')&=&\int dx\, f(x)\cdot D(x,x')\,.
 \end{eqnarray}
-Obviously equation~\ref{eq:mapping} can only describe the detector correctly
-when the condition of linearity is fulfilled. Within this model a detector
-is completely specified by a set of $D(E,E')$ for the energy variable $E$
-and a set $D(F,F')$ for the flavour variable $F$, where usually $D(E,E')$ also
-depends on the true flavour $F$ as well as $D(F,F')$ depends on the true energy
-$E$. Those sets of mapping functions usually will be obtained by a 
+Obviously \equ{mapping} only describes the detector properly
+if the linearity condition is fulfilled. Within this model, a detector
+is completely specified by a set of $D(E,E')$ for the energy variable $E$,
+and a set $D(F,F')$ for the flavor variable $F$.
+In general, $D(E,E')$ also depends on the true flavor $F$, as well as $D(F,F')$ depends on the true energy $E$.
+These sets of mapping functions usually are obtained from a
 full detector simulation and can be obtained by using as input 
@@
 
-In order to allow a systematic experiment definition including various
-sources of systematical errors we use several layers. The first layer is
-the so called ``channel'', it is the link between the oscillation physics
-and the detector. A channel specifies how a certain neutrino type in the 
-source is mapped into a reconstructed neutrino type, \eg\ a muon neutrino
-oscillates into an electron neutrino and subsequently interacts via
-quasi-elastic charged current scattering. The measured energy and direction of
-the electron in the detector finally allow to reconstruct the neutrino energy.
-The channels are the building blocks for so called ``rules'' which
-contain information like the efficiencies and systematical errors. An example
-could look like this: the efficiency for quasi-elastics electron neutrino
-events is $0.4$. There is a fraction of $0.01$ of all neutral current events
-which are mis-identified as quasi-elastic electron neutrino events. 
-The neutral current fraction is only known within $10\%$ and there is
-an energy scale uncertainty of $100\,\mathrm{MeV}$. Thus a rule connects the
-event rates to the calculation of a $\chi^2$ which properly includes 
-systematical errors. This $\chi^2$ is then the starting point for the
-oscillation physics analysis.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Getting started}
-
-As a first example we discuss how to define a neutrino factory experiment using
-AEDL. The first line of every experiment definition file is
+In order to implement a experiment definition including various
+sources of systematical errors, we use several abstraction levels. 
+The first level is the so-called ``channel'', which is the link between 
+the oscillation physics and the detection properties for a specfific oscillation pattern.
+A channel specifies the mapping of a specific neutrino flavor produced by the source onto a reconstructed neutrino flavor.
+For example, a muon neutrino oscillates into an electron neutrino and subsequently interacts via quasi-elastic charged current scattering. The measured energy and direction of
+the secondary electron in the detector then allows to reconstruct the neutrino energy.
+The connection from the source flux of the muon neutrino, via the  probability to appear as a electron neutrino, to its detection properties (such as cross sections and energy smearing) is encapsulated into the channel.
+
+The channels are the building blocks for the so-called ``rules''.
+In general, a rule consists of one or more ``signal'' and ``background'' oscillation channels, which are normalized with efficiencies.
+The event numbers from these channels are added {\em before} the $\chi^2$-value is calculated.
+In addition, each rule implements an {\em independent} systematics, such as signal and background normalization errors.
+Eventually, each rule gives a $\chi^2$-value, and the total $\chi^2$ is obtained by adding the $\chi^2$'s of all rules.
+ An example for a rule could look like this: We want to detect electron neutrino appearance (``signal''), where the overall efficiency for quasi-elastics electron neutrino events is $0.4$.
+There is a fraction of $0.01$ of all neutral current events which are mis-identified as quasi-elastic electron neutrino events (``background'').
+The neutral current fraction is only known within $10\%$ (``background uncertainty'') and there is
+an energy scale uncertainty of $100\,\mathrm{MeV}$ (``energy calibration error'').
+All this systematics is independent of the other rules.  Thus, a rule connects the event rates to the calculation of a $\chi^2$ which properly includes 
+systematical errors. The resulting $\chi^2$ is then the starting point for the
+oscillation physics analysis. Note again that
+\begin{itemize}
+\item
+ Within each rule the event numbers are added.
+\item
+ Within each rule the systematics is treated independently from the other rules.
+\item
+ For each rule the $\chi^2$ is computed; the $\chi^2$'s from all rules are added.
+\end{itemize}
+
+Of course, an abstract experiment definition language can not simulate all possible types of experiments.
+As we have seen, there are several assumptions for source and detector.
+However, it turns out that \GLOBES\ can be applied to a large number of experiment types, such as conventional beams, superbeams, neutrino factories, $\beta$-Beams, and reactor experiments.
+
+\section{A simple example for the AEDL}
+
+Experiments are in \GLOBES\ defined by the Abstract Experiment Definition Language (AEDL).
+The experiment definition is written into a text file using the AEDL syntax.
+Currently, a number of pre-defined experiment definition files are provided with \GLOBES , which have to be modified manually in order to define new experiments.
+An IDE (Integrated Developement Environment) for AEDL files is currently in preparation.
+The application software then uses this text file to initialize the experiment, where other secondary files might read for source fluxes, cross sections \etc .
+In this section, we show the definition of a very simple neutrino factory in AEDL, where we do not go into details.
+In the next chapter, we will discuss each of the individual steps in detail.
+
+The first line of every experiment definition file has to be
 \begin{quote}
@@
 \end{quote}
-
-We tell \GLOBES\ that we want to use the built-in neutrino factory
-flux originating from stored $\mu^+$'s. This is achieved by setting 
-the builtin variable to $1$.  Next the energy of the muons is specified
-to be $50\,\mathrm{GeV}$ by the parent energy variable. It is assumed
+in order not to confuse it with some other file format.
+%
+First, we instruct \GLOBES\ to use the built-in source flux for a neutrino factory
+originating from stored $\mu^+$'s. Therefore, we set the {\tt @builtin} variable to $1$. Next, we specify the muon energy to be $50\,\mathrm{GeV}$ by the {\tt @parent\_energy} variable. We assume 
 that there will be $5.33\cdot 10^{20}$ useful muon decays per year
-and that this luminosity is available for $8$ years.
+and that this luminosity is available for $8$ years, \ie , a total number
+of $ 4.264\cdot10^{21}$ muons is stored:
 \begin{quote}
@@
 \end{quote}
-This specific flux can be later on refered to as {\tt \#mu\_plus}.
-
+Note that we tell \GLOBES\ that we want to refer to this neutrino source later as as {\tt \#mu\_plus}. 
+%
+Let us now define a very simple detector with a target mass 
+of $50\,\mathrm{kt}$ and $20$ energy bins between
+$4\,\mathrm{GeV}$ and $50\,\mathrm{GeV}$: 
 \begin{quote}
@@
 \end{quote}
-The target mass 
-is $50\,\mathrm{kt}$ and a total of $ 4.264\cdot10^{21}$
-muons is stored. The data is divided into 20 energy bins ranging from
-$4\,\mathrm{GeV}$ to $50\,\mathrm{GeV}$.
-
-Next a file has to be specified which contains the cross section we want to 
-use.
+Then we specify the file which contains the cross sections we want to 
+use:
 \begin{quote}
@@
 The command {\tt cross} tells the parser that a cross section environment
-begins and has the name {\tt \#CC}. These names are used to refer later on
-to this specific environment and each name begins with a leading {\tt \#}.
-Whenever we need to specify a cross section from here on we can use {\tt \#CC}
-to refer to the cross section as given in the file {\tt XCC.dat}.
-
-Of course also the baseline and matter profile have to be specified, which
- looks like this
+begins. It has the name {\tt \#CC}, which can later be used to refer 
+to this specific environment, and thus to the file {\tt XCC.dat}. Note that each name begins with a leading {\tt \#}.
+%
+Of course, the baseline and matter profile have to be specified, too:
 \begin{quote}
@@
 The curly brackets used for the definition of {\tt \$densitytab} and
-{\tt \$lengthtab} delimit as simple list of numbers -- here the lists
-have length $1$. {\tt \$density\_error} is the relative $1\,\sigma$ uncertainty
-on the density.
-
-The energy resolution function can easily be specified with
+{\tt \$lengthtab} refer to a list of numbers. Here, the lists contain only
+one element, because we only use one density layer: We initialize a baseline length of $3000 \, \mathrm{km}$ with a constant matter density of $3.5 \, \mathrm{g/cm^3}$ and a matter density uncertainty of $5\%$ ($1 \sigma$ relative error).
+Note that the elements in {\tt \$lengthtab} have to add up to the baseline length.
+%
+As another ingredient, we have to define the energy resolution function:
 \begin{quote}
@@
 \end{quote}
-The {\tt energy} command opens the corresponding energy environment and
-the energy resolution function defined here is refered to by the name 
-{\tt \#MINOS}. Out of several possibilities it employs algorithm $1$,
+The {\tt energy} command opens the energy environmen, which has the name 
+{\tt \#MINOS} here. Out of several possibilities, it uses algorithm one,
 the simplest and fastest one. The actual energy resolution is specified
-by the energy resolution variable which is a list of three elements. Each 
-element is one parameter of a general resolution function as defined in 
-equation ~\ref{eq:sigma_e}.
-
-Now we define the one appearance and the corresponding disappearance 
-channel of a neutrino factory: 
+by the energy resolution variable, which is a list of three elements. Each 
+element is one parameter of the general resolution function as defined in 
+\eq~\ref{eq:sigma_e}.
+%
+Now we have all pieces to be able to define the appearance and the corresponding disappearance channel of a neutrino factory: 
 $\nu_e\rightarrow\nu_\mu$  and $\bar\nu_\mu\rightarrow\bar\nu_\mu$ 
@@
 \end{quote}
-The first element is the name of the flux as defined previously, here we use
-the flux named {\tt \#mu\_plus}. The second element $\pm$ determines whether 
-neutrinos or anti-neutrinos are
-taken from the flux table. The third position defines the initial flavour
-and the forth position the final flavour, followed by the name of the cross
-section and energy resolution function.
-
-The last step is to assemble the channels into a rule.
+The first element is the name of the flux, which we have already defined above as defined above. The second element $\pm$ determines whether 
+neutrinos or anti-neutrinos are taken from the flux table (two different polarities allowed). The third position defines the initial flavor,
+and the forth position the final flavor, followed by the name of the cross
+section and energy resolution function as defined before.
+%
+The last step is to encapsulate the channels into a rule:
 \begin{quote}
@@
 \end{quote}
-The signal variable says that the signal in our experiment is the 
-channel named {\tt \#appearance} with an constant efficiency of $0.45$.
-The signal error variable has two components: the first one is the
-normalization error of the signal, here it is $0.1\%$. The second component
-signifies the energy calibration error of the signal, whose definition 
-can be found in section~\ref{sec:energy}. The background variable
-specifies the composition of the beam background and this case is
-given by the channel named {\tt \#disappearance}, \ie\ this are
-the muons whose charge has been mis-identified. They enter the calculation
-with an efficiency of $1\cdot 10^{-5}$. The background center variable allows
-to rescale the total background contribution which is especially useful if
-there are many background components. 
-The background error variable is defined in the same way as the signal error
-variable, \ie\ we have a $5\%$ background uncertainty.
-The error dimension variable selects how
-the systematical errors are treated according to table~\ref{tab:error_dim}.
-
-The above defined experiment represents a first version of a neutrino factory
-experiment. It still lacks the correct energy dependence of the efficiencies
-and of course the channels and rules for the operation with $\mu^-$ stored
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Features of the  AEDL}
+The {\tt @signal} refers to the ``signal'' in our experiment. We use the
+above defined channel named {\tt \#appearance} with an constant overall
+efficiency of $0.45$. The signal error variable has two components: 
+The first one is the normalization error of the signal, here $0.1\%$. The second 
+one refers to the energy calibration error of the signal, which is defined 
+in \Sec~\ref{sec:energy}. The background variable
+specifies the composition of the beam background. In this (simplified) case, we
+use the fraction $1\cdot 10^{-5}$ of the channel named {\tt \#disappearance}, \ie , the muon neutrinos with a mis-identified charge.
+The background center variable allows to rescale the total background contribution from all background components
+simultaneously.
+It is only useful if there is more than one background component, otherwise it is usually $1$.
+The background error variable is defined such as the signal error variable, \ie\ we have a $5\%$ background uncertainty and a very small energy calibration error.
+The ``error dimension variable'' {\tt @errordim} selects how the systematical errors are treated (\cf, \Tab~\ref{tab:error_dim}).
+
+The here defined experiment represents a first simplified version of a neutrino factory experiment.
+It still lacks the correct energy dependence of the efficiencies, the antineutrino disappearance channel, and the channels and rules for the symmetric operation with $\mu^-$ stored.
+However, it may serve as a simple, introductory example.
+In the next chapter, we will demonstrate that the AEDL is much more powerful than illustrated here.
+
+% continue here (WW):
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Syntax and types of  AEDL}
+\section{Introduction to the syntax of AEDL}
 \label{sec:syntax}
@@
  the end of input, alternatively you can put a comment at the end.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\chapter{Experiment definition with AEDL}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2004-05-11 Documentation/part2.tex
*** empty log message ***

@@
 
-% continue here (WW):
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+We now give a short introduction to the syntax of AEDL.
  The first eight characters have to be {\tt \%!GLoBES}
- this serves to protect from parsing MBytes of chunk
- and cluttering the screen with error messages.
-
-Comments
- are entered like this one -- everything between the 
- delimiters is ignored. 
+in order to avoid parsing megabytes of chunk
+ and producing thousands of error messages. 
+%
+Comments can be used such as in C:
 \begin{quote}
@@
 \end{quote}
-thus the comments in AEDL work like in C
-
-There are so called  named quantities, they all start with {\tt \$} 
-and are predefined. They are checked for being within range, \eg\ {\tt 
-\$bins} can be only in between 0 and 150. If you give a {\tt float} where
- an {\tt int} is expected the {\tt float} gets converted to an {\tt int} 
-in the usual way in C.  We have named scalar quantities
+There are pre-defined variables which all start with {\tt \$}. Their range
+is also checked. For example,  {\tt 
+\$bins} can be only between $0$ and $150$.
+If one uses a {\tt float} quantity where  an {\tt int} is expected, the {\tt float} will be converted to an {\tt int} in the same way as in C.
+For example, we have scalar variables
 \begin{quote}
@@
 \end{quote}
-as well as named simple lists
+and simple lists
 \begin{quote}
@@
 \end{quote}
-
-Since there are some very special data to parse, some special
-provisions have been made so called named environments. This is illustrated 
-with the channel definition part
+%
+Since there are often groups of data which we want to refer to later,
+environments can be used. This is illustrated 
+with the channel definition part:
 \begin{quote}
@@
 \end{quote}
-
-The first part tells the type of environment, here it is {\tt channel}. 
-There are the following types 
+The first part is the type of environment, which is {\tt channel} here. 
+There are the following types of environment in AEDL:
 \begin{quote}
@@
 \end{quote}
-Besides the type there is a user-defined name beginning with {\tt \#}, in the
-above example {\tt \#ch1}, which can be used later on to refer to the 
-channel defined in between {\tt <$\ldots$>}. Those names are so called
-automatic variables. They start all with {\tt \#} and do not behave as 
-usual variables. New names can only be
-created within the above construction. Only names which have been created
-can be refered to. Using the same name for different types of environments 
-does not work.  Each type has its own counter, \eg\ for channel the counter is 
-{\tt numofchannels}. The counter keeps track of how many different names 
-there are for one type, \ie\ they count the number of channels, 
-the number of rules, the number of energy resolution functions asf. 
-The value of those automatic variables is the number which can be used 
-to refer to them in your C code. You can create a variable for later use by
+Besides the environment type, there is a user-defined name 
+beginning with {\tt \#}
+in the above example: {\tt \#ch1}. It can be used later to refer to the 
+channel defined in {\tt <$\ldots$>}. Those names are so-called 
+``automatic variables'' and have to start with {\tt \#}. Note that these names have to be unique and can only be refered to after their definition.
+However, similar to C, one can give a header definition before:
 \begin{quote}
@@
 \end{quote}
-Now you can refer to {\tt \#ch2} without any trouble and you now could define
-a rule using {\tt \#ch2}. Later on of course you still have to specify the
-actual channel as described in section~\ref{sec:channel}.
-
-Variables beginning with {\tt @} are named and can be used only within the 
-appropriate type of environment, since we have to keep track where 
-they belong to. Moreover many of them have a special syntax, like 
-{\tt @channel}.
-
-You can have several experiments in one file, just separate the different
+Now one can refer to the name {\tt \#ch2}, while the actual channel definition comes later. The internal representation of this automatic
+variable is a number, which obtains its value from a counter for each type of environment.
+For example, for {\tt channel} the counter is {\tt numofchannels}.
+The counter keeps track of how many different names
+there are for one type of environment, which means that it counts the number of channels, rules, energy resolution functions \etc . Thus, the automatic
+variables are numbered in the order of their definition, and the number
+can later be used to refer to them in the C code (from $0$ to {\tt numof...}$-1$). Within each environment type, there are several 
+variables beginning with {\tt @}, which can only be used within the 
+appropriate type of environment. In many cases, 
+they have a special syntax, such as {\tt @channel}.
+
+If you want to have several experiments in one file, separate the different
  experiments by 
@@
 This command resets the counters for channels, rules and energy resolution
-functions. All variables have their scope only within a region delimited
+functions. All variables have their scope limited 
 by either {\tt \%!GLoBES, \#NEXT\#} or {\tt EOF}. The only exception are 
-the counters defined with {\tt cross} and {\tt flux}. They are valid for 
-the complete session, since you never must overload the flux and 
-cross section information. This ensures  a consistent treatment of various 
-experiments in one file - there is no  shadowing of any variables.
-
-Using multiple files to keep things clean and clear is supported by 
- an {\tt include} command. Includes can be nested up 
- to {\tt MAX\_INCLUSION\_DEPTH}, which is set to 10. Error reporting works 
- also for nested includes. The included file is not required to begin 
- with {\tt \%!GLoBES} to facilitate cut and paste
+the counters defined by {\tt cross} and {\tt flux}. They are valid for 
+the complete session, since one does not want too overload the flux and 
+cross section information when using more than one experiment. This allows 
+a consistent treatment of various experiments in one file.
+
+As another feature of AEDL one can use include files with the {\tt include} command. Includes can be nested up to {\tt MAX\_INCLUSION\_DEPTH}, which is currently set to $10$. Error reporting works 
+ for nested includes, too. The included file is not required to begin 
+ with {\tt \%!GLoBES} to facilitate cut and paste:
 \begin{quote}
@@
 \end{quote}
-
-Thus it is possible and advised to use constructs as the following
+With this include mechanism, one can use constructions such as 
 \begin{quote}
@@
 \end{quote}
-to define a combined analysis of the experiments defined in the files 
-{\tt NuFact.gls} and {\tt JHFHK.gls}. Even if you use the named 
-variable {\tt \#CC} in both experiments, but the cross section data used 
-is different (\eg\ because of different target nuclei), the correct 
-cross section data is used for each of the experiments, since {\tt \#CC} 
-in {\tt NuFact.gls} has the value $0$ and
-in {\tt JHFHK.gls} it has the value $1$, thus no shadowing occurs.
-
-Furthermore you can store numbers in variables
+in order to be able to initialize a combined analysis of the experiments defined in the files {\tt NuFact.gls} and {\tt JHFHK.gls} later. 
+Even if one uses the
+automatic variable {\tt \#CC} in both experiments 
+but the cross section data are different (for example, because of different target nuclei), the correct 
+cross section data will be applied to each of the experiments in the 
+application software. In this case, {\tt \#CC} in {\tt NuFact.gls} has 
+the value $0$ and in {\tt JHFHK.gls} it has the value $1$, 
+which means that no shadowing occurs.
+
+Furthermore, one can define constants such as
 \begin{quote}
@@
 \end{quote}
-And additionally we have some simple calculator functionality like
+and use some simple algebraic manipulations such as
 \begin{quote}
@@
 \end{quote}
-The following mathematical functions from {\tt <math.h>} are available 
+The following mathematical functions from {\tt <math.h>} are available: 
 {\tt sin}, {\tt cos}, {\tt atan}, {\tt ln}, {\tt exp}, {\tt sqrt}. 
-Those calculations work everywhere, where
-a simple number could stand. However threading over lists does not
- work, \eg\ {\tt sin(\{1,2,3\})} will not work.
-
-One nasty thing is that a line feed character \verb+\n+ is necessary at
- the end of input, alternatively you can put a comment at the end.
+These functions can be used everywhere, where
+otherwise only a scalar number would appear. However, they can not be
+applied to lists, such as {\tt sin(\{1,2,3\})} will not work.
+
+Finally, note that a line feed character \verb+\n+ is necessary at
+ the end of the input -- alternatively you can put a comment at the end.
+
+% Continue here (WW):
 
@@
 
+In this chapter, we give a detailed definition of the AEDL features.
+We also show the underlying mathematical concepts in parallel, where applicable.
+We do not exactly follow the separation of source, oscillation, and detection properties, since most issues more or less involve the detection.
+Instead,
+we illustrate many of the features of the \GLOBES\ simulation by successively
+evolving the underlying statistical picture, and demonstrating how it translates into AEDL .
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Source properties}
+\section{Source properties and luminosity}
 \label{sec:source}
@@
 given.
+
+The most obvious property of the neutrino detection is the total
+active mass of the target as already defined in section~\ref{sec:source} and
+its value is defined by 
+\begin{quote}
+{\tt \$target\_mass = 50.0 }
+\end{quote}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Oscillation properties}
-
-\subsection{Baseline \& matter profile}
+\section{Baseline and matter density profile}
+
 The purpose of this module is to specify the path the neutrinos
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Detection properties}
-
-The most obvious property of the neutrino detection is the total
-active mass of the target as already defined in section~\ref{sec:source} and
-its value is defined by 
-\begin{quote}
-{\tt \$target\_mass = 50.0 }
-\end{quote}
-
-\subsection{Cross section}
+
+\section{Cross sections}
 \label{sec:cross_section}
@@
 
-\subsection{Channels}
+\section{Oscillation channels}
 \label{sec:channel}
@@
 flavor and IT as seen by the detector, which we call a ``channel''.
+
 The master formula for the differential event rate for each channel 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Rules}
-\label{sec:rules}
-
-The set of rules\index{Rule} for an experiment is the final 
-link between the event rate
-computation and the statistical analysis. The information given by the rules
-specifies how the $\chi^2$ is computed based on the raw event rates as 
-described by the channels including possible systematical errors. 
-Therefore a rule has two parts: the first part describes how signal and 
-background events are composed out of the channels and the second part
-specifies which systematical errors are considered and what the size
-of the relevant errors is.
-
-Based on the definition of channels it is possible to assemble the 
-signal and background event numbers for each bin.  
-The signal $s_i$ in one bin for a rule now can be composed out of 
-several channels by
-\begin{equation}
-s_i=\epsilon_{c_1}\cdot n_i^{c_1}\,+\,\epsilon_{c_2}\cdot n_i^{c_2}\,+\,\ldots
-\end{equation}
-where the $\epsilon$'s are weight factors determined by the properties
-of the detector.
-Similarly the background in one bin $i$ can be specified by
-\begin{equation}
-b_i=\epsilon_{c_1}\cdot n_i^{c_1}\,+\,\epsilon_{c_2}\cdot n_i^{c_2}\,+\,\ldots
-\end{equation}
-This basic building blocks of a rule are specified by
-\begin{quote}
-{\tt \tb @signal = 0.5 @ \#channel\_1\\
-\tb @background = 0.001 @ \#channel\_2 :  0.005 @ \#channel\_3
-}
-\end{quote}
-
-Quite often it is not sufficient to have constant efficiencies and therefore
-it is possible to define efficiencies for each bin and channel $\alpha_i^c$
-and thus the event rates in one bin become
-\begin{equation}
-s_i=\epsilon_{c_1}\alpha_i^{c_1}\cdot n_i^{c_1}\,+
-\,\epsilon_{c_2}\alpha_i^{c_2}\cdot n_i^{c_2}\,+\,\ldots\,,
-\end{equation}
-and analogous for $b_i$.
-
-The set of all $s_i$'s and $b_i$'s now can be composed in a form that takes
-some types of systematical error into account. The two most important and
-most easily parameterized systematics are a normalization and an energy
-calibration error. These error are considered separately for the signal events
-and the background events. The implementation of the normalization error
-is straight forward
-\begin{equation}
-s_i(a):=a\cdot s_i
-\end{equation} 
-with an analogous definition for the background events.
-
-For the parameterization of an energy calibration error two possibilities
-are implemented. The first one being somewhat simpler, whereas the second one
-is more accurate but requires a careful choice of parameters. The first 
-option is
-\begin{equation}
-s_i(a,b):=s_i(a)+b\cdot s_i\, E_i/(E_\mathrm{max}-E_\mathrm{min})
-\end{equation}
-and it is refered to as tilt since it describes a linear distortion, a tilt
-of the event rate spectrum. The second option is closer to an actual energy
-calibration error, which basically amounts to replacing the reconstructed 
-energy $E'$ by $(1+b)\cdot E'$ and we use following approximation
-\begin{eqnarray}
-s_i(b)&=& (1+b)\cdot(s_{k+1}-s_k)\cdot\delta+s_i\,,\\
-\delta&=&b\cdot(i+t_0+ 1/2)+i\,,\nonumber\\
-k&=&\mod(\delta,1)\,,\nonumber\\
-t_0&=&E_\mathrm{min}/\Delta E_0\,.\nonumber
-\end{eqnarray}
-Special care is required at the boundaries when $k<1$ or $k+1>N_\mathrm{bins}$,
-since there $s_k$ or $s_{k+1}$ may not have been calculated. One possibility
-to deal with those cases is to assume that for those values of $k$ $s_k$ is
-zero. This is the default. In cases where $s$ at the boundary is however still
-sizeable this may lead to intolerable errors and subsequently to a wrong
-estimate of the impact of a calibration error. Therefore in those cases it
-is advisable to truncate the analysis range by a few bins at the boundary
-and therefore ensure in this way that only those $s_i$ are used whose index
-$k$ is within the range $1,\ldots, N_\mathrm{bins}-1$.
-
-In order to perform this fine tuning, it is possible to constrain 
-the energy range from which the data are taken for the 
-analysis\index{Energy window}. This is achieved by setting the energy window
-variable. 
-\begin{quote}
-{\tt 
-\tb @energy\_window = 4.0 : 50.0 
-}
-\end{quote}
-The purpose is mainly to allow a fine tuning of the description
-of an energy calibration error. The default setting is that the energy window
-coincides with the range defined by the minimal and maximal reconstructed 
-energy. For the error dimension using the calibration method (C) it is 
-advisable to reduce the energy window by two or three bins at both the lower
-and upper end. As a rule of
-thumb consider the three standard deviations error on the calibration and
-evaluate the corresponding shift in $E'$ at the lower and upper energy limit 
-and reduce the analysis range by this amount.
-
-Thus the total event rate $x_i$ in a bin $i$ is given by
-\begin{equation}
-x_i(a,b,c,d)=s_i(a,b)+b_i(c,d)
-\end{equation}
-and thus a function of four parameters. The central values for all of
-the four parameters are always needed. They are called signal normalization
-($a$), signal tilt/calibration ($b$), background  normalization ($c$) and
-background tilt/calibration ($d$). The default values are
-\begin{equation}
-a=1\,,\quad b=0\,,\quad c={\tt NaN}\,,\quad d=0\,,
-\end{equation}
-thus for the background normalization $c$ a value has to be specified in 
-\emph{all} cases. The central values for the normalization and the 
-corresponding central value of tilt/calibration are regarded
-as a pair. The errors are treated in the same way.
-\begin{quote}
-{\tt
-\tb @signalerror =       0.001  :       0.01\\
-\tb @backgroundcenter =  0.1 :       0.0\\
-\tb @backgrounderror =   0.001 :       0.01
-}
-\end{quote}
-There is no {\tt @signalcenter} since by default the central value for the
-signal normalization is $1$ and the central value for the tilt/calibration 
-is $0$.
-
-The four parameters $a,b,c,d$ have been introduced in order to describe
-systematical uncertainties, they are so called nuisance parameters.
-For the analysis of the systematical errors the so called 
- pull method is 
-used~\cite{Fogli:2002pt}\footnote{In fact the pull method
-was employed already in~\cite{Huber:2002mx} before 
-\cite{Fogli:2002pt} appeared.}. 
-In the  pull method $k$ systematical errors are included by introducing 
-$k$ additional variables $\zeta_k$, which will  be called 
-nuisance parameters in the following. 
-The nuisance parameters describe the dependence of the event rates on the 
-various sources of systematical errors, \eg\ an error on the total 
-normalization is included by multiplying the expected number of events in 
-each bin by a factor $(1+\zeta_1)$. The variation of $\zeta_1$ in the fit 
-is constrained by adding a penalty $p_1$ to the $\chi^2$-function. In case 
-of a  Gau\ss ian distributed systematical error this penalty is 
-given by
-\begin{equation}
-\label{eq:penalty}
-p_i=\frac{(\zeta_i-\zeta_i^0)^2}{\sigma_{\zeta_i}^2}\,,
-\end{equation}
-where $\zeta_i^0$ denotes the mean and $\sigma_{\zeta_i}$ the standard 
-deviation of the corresponding nuisance parameter, \ie\ the amount of 
-systematical uncertainty. 
-The resulting $\chi^2$ is then minimized with respect to all nuisance 
-parameters $\zeta_i$ and this yields $\chi^2_\mathrm{pull}$
-\begin{equation}
-\chi^2_\mathrm{pull}(\boldsymbol{\lambda}):=\min_{\{\zeta_i\} } \,\, \left( 
-\chi^2(\boldsymbol{\lambda},
-\zeta_1, \ldots, \zeta_k)+ \sum_{j=1}^{k} p_j(\zeta_j)\right)\,,
-\end{equation}
-where $\boldsymbol{\lambda}$ denotes the oscillation parameters 
-including the matter density
-$\rho$. One advantage of the pull method is that whenever the number $N$ of 
-data points is much larger than $k$, it is numerically easier to compute 
-$\chi^2_\mathrm{pull}$ than to invert the $N\times N$ covariance matrix. For
-the experiments considered here $N$ is typically $20$ and $k\sim 4$, thus
-the pull method is numerically much faster. Moreover it is more flexible and 
-allows the inclusion of systematical errors also for a 
-Poissonian $\chi^2$-function.
-In~\cite{Fogli:2002pt} it was shown that the pull method and the covariance
-based approach are equivalent for a Gau\ss ian and linear model. In general
-there is a separate $(\chi^2_\mathrm{pull})^\alpha$ for each rule $\alpha$, 
-\ie\ pair of signal and background spectra, with a separate set of 
-nuisance parameters $\zeta_i^\alpha$. In this case $\chi^2_\mathrm{pull}$
-is the sum of all individual  $(\chi^2_\mathrm{pull})^\alpha$.
-In this way the dependence on the $k$ nuisance parameters has been eliminated 
-from $\chi^2_\mathrm{pull}$.
-
-The user has the possibility to choose the set $\{\zeta_i\}$ of nuisance 
-parameters which are marginalized. This achieved with the error dimension 
-variable\index{Error dimension}, and the different possibilities are shown in
-table~\ref{tab:error_dim}. If a parameter is designated with $+$, it will
-be marginalized and therefore the corresponding error needs to have a non-zero
-value. In the cases labeled (number 4 and 8) ``total rates'' 
-the summation over the bins is performed \emph{before} computing 
-the $\chi^2$. The case labeled (number 7) ``spectrum only'' leaves the 
-normalization free ($\sigma_a=\sigma_c=\infty$) and therefore counts only the 
-spectral information in the data. As a consequence
-the settings for the normalization error will be ignored. It is furthermore
-possible to extend the set of error dimensions and thus the set of possible 
-systematical errors to be studied, but this requires proper C programing, 
-compiling and linking with \GLOBES. The interface for user defined error 
-dimensions is described in detail in section~\ref{sec:ud_error_dim}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{center}
-\begin{table}[hbt!]
-\begin{center}
-\begin{tabular}[h]{|c|cccc|c|c|}
-\hline
-Error dimension&$a$&$b$&$c$&$d$&Tilt/Calibration&Remarks\\
-\hline
-\hline
-0&+&+&+&+&T&\\
-2&-&-&-&-&-&\\
-4&+&+&+&+&T&total rates\\
-7&o&-&o&-&-&spectrum only\\
-8&-&-&-&-&-&total rates\\
-9&+&+&+&+&C&\\
-\hline
-\end{tabular}
-\caption[Table of error dimensions]{\label{tab:error_dim}
-Values of the error dimension variable and their meaning.
- }
-\end{center} 
-\end{table} 
-\end{center}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The error dimensions can be chosen for each rule by
-\begin{quote}
-{\tt 
-\tb @errordim = 1
-}
-\end{quote}
-
-Next we can assemble the various pieces into one rule
-\begin{quote}
-{\tt rule(\#rule\_1)<\\
-\tb @signal = 0.5 @ \#channel\_1\\
-\tb @background = 0.001 @ \#channel\_2 :  0.005 @ \#channel\_3\\
-\tb @signalerror =       0.001  :       0.01\\
-\tb @backgroundcenter =  0.1 :       0.0\\
-\tb @backgrounderror =   0.001 :       0.01\\
-\tb @errordim = 1\\
-\tb @energy\_window = 4.0 : 50.0\\ 
->}
-\end{quote}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Energy resolution function}
+\section{Energy resolution function}
 \label{sec:energy}
@@
 \end{quote}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Rules and the treatment of systematics}
+\label{sec:rules}
+
+The set of rules\index{Rule} for an experiment is the final 
+link between the event rate
+computation and the statistical analysis. The information given by the rules
+specifies how the $\chi^2$ is computed based on the raw event rates as 
+described by the channels including possible systematical errors. 
+Therefore a rule has two parts: the first part describes how signal and 
+background events are composed out of the channels and the second part
+specifies which systematical errors are considered and what the size
+of the relevant errors is.
+
+Based on the definition of channels it is possible to assemble the 
+signal and background event numbers for each bin.  
+The signal $s_i$ in one bin for a rule now can be composed out of 
+several channels by
+\begin{equation}
+s_i=\epsilon_{c_1}\cdot n_i^{c_1}\,+\,\epsilon_{c_2}\cdot n_i^{c_2}\,+\,\ldots
+\end{equation}
+where the $\epsilon$'s are weight factors determined by the properties
+of the detector.
+Similarly the background in one bin $i$ can be specified by
+\begin{equation}
+b_i=\epsilon_{c_1}\cdot n_i^{c_1}\,+\,\epsilon_{c_2}\cdot n_i^{c_2}\,+\,\ldots
+\end{equation}
+This basic building blocks of a rule are specified by
+\begin{quote}
+{\tt \tb @signal = 0.5 @ \#channel\_1\\
+\tb @background = 0.001 @ \#channel\_2 :  0.005 @ \#channel\_3
+}
+\end{quote}
+
+Quite often it is not sufficient to have constant efficiencies and therefore
+it is possible to define efficiencies for each bin and channel $\alpha_i^c$
+and thus the event rates in one bin become
+\begin{equation}
+s_i=\epsilon_{c_1}\alpha_i^{c_1}\cdot n_i^{c_1}\,+
+\,\epsilon_{c_2}\alpha_i^{c_2}\cdot n_i^{c_2}\,+\,\ldots\,,
+\end{equation}
+and analogous for $b_i$.
+
+The set of all $s_i$'s and $b_i$'s now can be composed in a form that takes
+some types of systematical error into account. The two most important and
+most easily parameterized systematics are a normalization and an energy
+calibration error. These error are considered separately for the signal events
+and the background events. The implementation of the normalization error
+is straight forward
+\begin{equation}
+s_i(a):=a\cdot s_i
+\end{equation} 
+with an analogous definition for the background events.
+
+For the parameterization of an energy calibration error two possibilities
+are implemented. The first one being somewhat simpler, whereas the second one
+is more accurate but requires a careful choice of parameters. The first 
+option is
+\begin{equation}
+s_i(a,b):=s_i(a)+b\cdot s_i\, E_i/(E_\mathrm{max}-E_\mathrm{min})
+\end{equation}
+and it is refered to as tilt since it describes a linear distortion, a tilt
+of the event rate spectrum. The second option is closer to an actual energy
+calibration error, which basically amounts to replacing the reconstructed 
+energy $E'$ by $(1+b)\cdot E'$ and we use following approximation
+\begin{eqnarray}
+s_i(b)&=& (1+b)\cdot(s_{k+1}-s_k)\cdot\delta+s_i\,,\\
+\delta&=&b\cdot(i+t_0+ 1/2)+i\,,\nonumber\\
+k&=&\mod(\delta,1)\,,\nonumber\\
+t_0&=&E_\mathrm{min}/\Delta E_0\,.\nonumber
+\end{eqnarray}
+Special care is required at the boundaries when $k<1$ or $k+1>N_\mathrm{bins}$,
+since there $s_k$ or $s_{k+1}$ may not have been calculated. One possibility
+to deal with those cases is to assume that for those values of $k$ $s_k$ is
+zero. This is the default. In cases where $s$ at the boundary is however still
+sizeable this may lead to intolerable errors and subsequently to a wrong
+estimate of the impact of a calibration error. Therefore in those cases it
+is advisable to truncate the analysis range by a few bins at the boundary
+and therefore ensure in this way that only those $s_i$ are used whose index
+$k$ is within the range $1,\ldots, N_\mathrm{bins}-1$.
+
+In order to perform this fine tuning, it is possible to constrain 
+the energy range from which the data are taken for the 
+analysis\index{Energy window}. This is achieved by setting the energy window
+variable. 
+\begin{quote}
+{\tt 
+\tb @energy\_window = 4.0 : 50.0 
+}
+\end{quote}
+The purpose is mainly to allow a fine tuning of the description
+of an energy calibration error. The default setting is that the energy window
+coincides with the range defined by the minimal and maximal reconstructed 
+energy. For the error dimension using the calibration method (C) it is 
+advisable to reduce the energy window by two or three bins at both the lower
+and upper end. As a rule of
+thumb consider the three standard deviations error on the calibration and
+evaluate the corresponding shift in $E'$ at the lower and upper energy limit 
+and reduce the analysis range by this amount.
+
+Thus the total event rate $x_i$ in a bin $i$ is given by
+\begin{equation}
+x_i(a,b,c,d)=s_i(a,b)+b_i(c,d)
+\end{equation}
+and thus a function of four parameters. The central values for all of
+the four parameters are always needed. They are called signal normalization
+($a$), signal tilt/calibration ($b$), background  normalization ($c$) and
+background tilt/calibration ($d$). The default values are
+\begin{equation}
+a=1\,,\quad b=0\,,\quad c={\tt NaN}\,,\quad d=0\,,
+\end{equation}
+thus for the background normalization $c$ a value has to be specified in 
+\emph{all} cases. The central values for the normalization and the 
+corresponding central value of tilt/calibration are regarded
+as a pair. The errors are treated in the same way.
+\begin{quote}
+{\tt
+\tb @signalerror =       0.001  :       0.01\\
+\tb @backgroundcenter =  0.1 :       0.0\\
+\tb @backgrounderror =   0.001 :       0.01
+}
+\end{quote}
+There is no {\tt @signalcenter} since by default the central value for the
+signal normalization is $1$ and the central value for the tilt/calibration 
+is $0$.
+
+The four parameters $a,b,c,d$ have been introduced in order to describe
+systematical uncertainties, they are so called nuisance parameters.
+For the analysis of the systematical errors the so called 
+ pull method is 
+used~\cite{Fogli:2002pt}\footnote{In fact the pull method
+was employed already in~\cite{Huber:2002mx} before 
+\cite{Fogli:2002pt} appeared.}. 
+In the  pull method $k$ systematical errors are included by introducing 
+$k$ additional variables $\zeta_k$, which will  be called 
+nuisance parameters in the following. 
+The nuisance parameters describe the dependence of the event rates on the 
+various sources of systematical errors, \eg\ an error on the total 
+normalization is included by multiplying the expected number of events in 
+each bin by a factor $(1+\zeta_1)$. The variation of $\zeta_1$ in the fit 
+is constrained by adding a penalty $p_1$ to the $\chi^2$-function. In case 
+of a  Gau\ss ian distributed systematical error this penalty is 
+given by
+\begin{equation}
+\label{eq:penalty}
+p_i=\frac{(\zeta_i-\zeta_i^0)^2}{\sigma_{\zeta_i}^2}\,,
+\end{equation}
+where $\zeta_i^0$ denotes the mean and $\sigma_{\zeta_i}$ the standard 
+deviation of the corresponding nuisance parameter, \ie\ the amount of 
+systematical uncertainty. 
+The resulting $\chi^2$ is then minimized with respect to all nuisance 
+parameters $\zeta_i$ and this yields $\chi^2_\mathrm{pull}$
+\begin{equation}
+\chi^2_\mathrm{pull}(\boldsymbol{\lambda}):=\min_{\{\zeta_i\} } \,\, \left( 
+\chi^2(\boldsymbol{\lambda},
+\zeta_1, \ldots, \zeta_k)+ \sum_{j=1}^{k} p_j(\zeta_j)\right)\,,
+\end{equation}
+where $\boldsymbol{\lambda}$ denotes the oscillation parameters 
+including the matter density
+$\rho$. One advantage of the pull method is that whenever the number $N$ of 
+data points is much larger than $k$, it is numerically easier to compute 
+$\chi^2_\mathrm{pull}$ than to invert the $N\times N$ covariance matrix. For
+the experiments considered here $N$ is typically $20$ and $k\sim 4$, thus
+the pull method is numerically much faster. Moreover it is more flexible and 
+allows the inclusion of systematical errors also for a 
+Poissonian $\chi^2$-function.
+In~\cite{Fogli:2002pt} it was shown that the pull method and the covariance
+based approach are equivalent for a Gau\ss ian and linear model. In general
+there is a separate $(\chi^2_\mathrm{pull})^\alpha$ for each rule $\alpha$, 
+\ie\ pair of signal and background spectra, with a separate set of 
+nuisance parameters $\zeta_i^\alpha$. In this case $\chi^2_\mathrm{pull}$
+is the sum of all individual  $(\chi^2_\mathrm{pull})^\alpha$.
+In this way the dependence on the $k$ nuisance parameters has been eliminated 
+from $\chi^2_\mathrm{pull}$.
+
+The user has the possibility to choose the set $\{\zeta_i\}$ of nuisance 
+parameters which are marginalized. This achieved with the error dimension 
+variable\index{Error dimension}, and the different possibilities are shown in
+table~\ref{tab:error_dim}. If a parameter is designated with $+$, it will
+be marginalized and therefore the corresponding error needs to have a non-zero
+value. In the cases labeled (number 4 and 8) ``total rates'' 
+the summation over the bins is performed \emph{before} computing 
+the $\chi^2$. The case labeled (number 7) ``spectrum only'' leaves the 
+normalization free ($\sigma_a=\sigma_c=\infty$) and therefore counts only the 
+spectral information in the data. As a consequence
+the settings for the normalization error will be ignored. It is furthermore
+possible to extend the set of error dimensions and thus the set of possible 
+systematical errors to be studied, but this requires proper C programing, 
+compiling and linking with \GLOBES. The interface for user defined error 
+dimensions is described in detail in section~\ref{sec:ud_error_dim}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{center}
+\begin{table}[hbt!]
+\begin{center}
+\begin{tabular}[h]{|c|cccc|c|c|}
+\hline
+Error dimension&$a$&$b$&$c$&$d$&Tilt/Calibration&Remarks\\
+\hline
+\hline
+0&+&+&+&+&T&\\
+2&-&-&-&-&-&\\
+4&+&+&+&+&T&total rates\\
+7&o&-&o&-&-&spectrum only\\
+8&-&-&-&-&-&total rates\\
+9&+&+&+&+&C&\\
+\hline
+\end{tabular}
+\caption[Table of error dimensions]{\label{tab:error_dim}
+Values of the error dimension variable and their meaning.
+ }
+\end{center} 
+\end{table} 
+\end{center}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The error dimensions can be chosen for each rule by
+\begin{quote}
+{\tt 
+\tb @errordim = 1
+}
+\end{quote}
+
+Next we can assemble the various pieces into one rule
+\begin{quote}
+{\tt rule(\#rule\_1)<\\
+\tb @signal = 0.5 @ \#channel\_1\\
+\tb @background = 0.001 @ \#channel\_2 :  0.005 @ \#channel\_3\\
+\tb @signalerror =       0.001  :       0.01\\
+\tb @backgroundcenter =  0.1 :       0.0\\
+\tb @backgrounderror =   0.001 :       0.01\\
+\tb @errordim = 1\\
+\tb @energy\_window = 4.0 : 50.0\\ 
+>}
+\end{quote}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{User-defined systematics}
-\label{sec:ud_error_dim}.
+%\section{User-defined systematics}
+%\label{sec:ud_error_dim}.
 

### 2004-05-12 Documentation/part2.tex
*** empty log message ***

@@
 \end{quote}
-The {\tt energy} command opens the energy environmen, which has the name 
+The {\tt energy} command starts the energy environment, which has the name 
 {\tt \#MINOS} here. Out of several possibilities, it uses algorithm one,
@@
 \end{quote}
-in order to be able to initialize a combined analysis of the experiments defined in the files {\tt NuFact.gls} and {\tt JHFHK.gls} later. 
+in order to initialize a combined analysis of the experiments defined in the files {\tt NuFact.gls} and {\tt JHFHK.gls}. 
 Even if one uses the
-automatic variable {\tt \#CC} in both experiments 
+automatic variable {\tt \#CC} in both experiments, 
 but the cross section data are different (for example, because of different target nuclei), the correct 
@@
 the value $0$ and in {\tt JHFHK.gls} it has the value $1$, 
-which means that no shadowing occurs.
+which means that no shadowing occurs. Note that, alternatively, one can 
+also load both files successively.
 
@@
 
-% Continue here (WW):
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 In this chapter, we give a detailed definition of the AEDL features.
-We also show the underlying mathematical concepts in parallel, where applicable.
+We also show the underlying mathematical concepts, where applicable.
 We do not exactly follow the separation of source, oscillation, and detection properties, since most issues more or less involve the detection.
 Instead,
-we illustrate many of the features of the \GLOBES\ simulation by successively
-evolving the underlying statistical picture, and demonstrating how it translates into AEDL .
+we illustrate many of the features of the \GLOBES\ simulation successively
+in the logical order of their definition, and demonstrate how they translate into AEDL .
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Source properties and luminosity}
+\section{Source properties and integrated luminosity}
 \label{sec:source}
 
-\GLOBES\ handles only point sources, thus it is not possible
-to study effects from the finite size of the neutrino production
-region like in the Sun. Furthermore only one source for each experiment
-is possible, this make it impossible\footnote{at least rather tricky} 
-to simulate multiple source experiments like KamLAND. All neutrino
-sources usable within \GLOBES\ therefore can be specified by the
-flux spectra for each flavour and CP sign and the total luminosity
-of the source. There are two principal ways to provide a flux:
-use a built-in source or a file provided by the user. In both cases
-a flux is defined by a named environment.
+As we have disussed before, \GLOBES\ can only deal with point sources.
+Thus,  it is not possible to study effects from the finite size of the neutrino production region, such as in the Sun or in reactor experiments with many
+neutrino sources (\eg, KamLAND).
+Therefore, a neutrino source in \GLOBES\ can, in general, be characterized by the flux spectrum for each neutrino flavor, the CP sign (neutrinos or antineutrinos), and the total luminosity
+of the source.
+
+Before we come to the definition of the source properties, let us discuss
+the total integrated luminosity of the experiment. In \GLOBES , the total number of events is in general proportional to the product of
+\begin{equation}
+\mathrm{Fid.~detector~mass}\,\left[\mathrm{kt/t}\right]\times 
+\mathrm{Running~time} \,\left[\mathrm{yr}\right]\times\left\{ \begin{array}{c}
+\mathrm{Source~power}\,\left[\mathrm{MW/GW}\right]\\
+\mathrm{Useful~muon~decays}\,\left[\mathrm{yr}^{-1}\right]
+\end{array}\right.\,.
+\end{equation}
+Thus, the source power corresponds to either the amount of energy produced per time frame in the target (such as for nuclear reactors or sources based on pion decay), or the useful muon decays per time frame (neutrino factories).
+In addition, the definition of the source power makes only sense together with the flux normalization, the running time, and fiducial detector mass in order to define the total integrated luminosity.
+Therefore, one can, in principle, use arbitrary units for these components as long as their product gives the wanted neutrino flux.
+However, it is
+recommended to use normalizations such that the source power units are $\mathrm{MW}$ for a proton-based beam, and $\mathrm{GW}_\mathrm{thermal}$ for a reactor experiment.
+Correspondingly, the detector mass units should be kilotons for a proton-based beam, and tons for a reactor experiment.
+
+The quantity which can be used to scale the overall integrated luminosity of an experiment, is the fiducial detector mass. For example,
+\begin{quote}
+{\tt \$target\_mass = 50.0 }
+\end{quote}
+defines a $50 \, \mathrm{kt}$ detector for a neutrino factory.
+
+There are two principal ways to initialize a neutrino flux: 
+Either one can use 
+a built-in source, or one can provide a file. In both cases,
+a flux is defined by the environment {\tt flux}, such as
 \begin{quote}
@@
 \tb $\ldots$\\
+\tb @time = 8.0 \\
 >}
 \end{quote}
-
-In case a built-in source is used the user has to specify which
-built-in spectrum is used. Currently two built-in fluxes are available:
-$\mu^+$-decay ({\tt @builtin = 1}) and $\mu^-$-decay ({\tt @builtin = 2}).
-In any case  the energy of the parent particle has to be specified
-and the number of useful decays per year. Thus the sequence for setting
-up a neutrino factory is
+with a running time of $8$ years. Note that the running time is used within
+the {\tt flux} environment. This feature can be used to load the neutrino and antineutrino fluxes separately, in order to combine them with different
+running times within one experiment. The name of the flux {\tt \#name} will later be refered to in the channel definitions.
+
+For a built-in neutrino source, one has to specify which
+built-in spectrum has to be used, as well as its parameters. The software
+will then automatically calculate the neutrino spectrum. Note that in this
+case, there is no degree of freedom in the choice of the source units.
+Currently, two built-in fluxes are available: $\mu^+$-decay ({\tt @builtin = 1}) and $\mu^-$-decay ({\tt @builtin = 2}).
+In these cases, the muon energy (enery of the parent particle) has to be specified together with the number of useful decays  per year.
+Thus, an example to set up a neutrino factory flux is
+\begin{quote}
 {\tt flux(\#mu\_plus)<\\
@@
 \tb  @time = 8.0\\
->}\\
-
-Alternatively user defined fluxes may used like this
+>}
+\end{quote}
+%
+For a user-defined flux, one has to give it the file name:
 \begin{quote}
@@
 \end{quote}
-
-The norm variable allows to set the conversion factor from the fluxes in
-the file to the units in \GLOBES. The power variable allows to
-specify a target or reactor power. The units used for power are determined
-by the units in the flux file and by the value of the norm variable. It is
-recommended to use $\mathrm{MW}$ in case a proton based beam is used and
-$\mathrm{GW}_\mathrm{thermal}$ for the reactor experiments.
-
-For sources based on $\pi$-decay and nuclear reactors the total power
-involved defines how many neutrinos are produced therefore this variable is
-used to define the luminosity. Furthermore it is useful to define 
-the operation time of an experiment in years. Note however that the 
-number of events used in analysis only depends on the product of
-\begin{equation}
-\mathrm{mass}\,\left[10^6\,\mathrm{kg}\right]\times \mathrm{time}
-\,\left[\mathrm{years}\right]\times\left\{ \begin{array}{c}
-\mathrm{source~power}\,\left[\mathrm{MWy}\right]\\
-\mathrm{useful~decays}\,\left[\mathrm{year}^{-1}\right]
-\end{array}\right.\,.
-\end{equation}
-
-The format of the flux files is seven rows and
-501 lines.
+In this case, the {\tt @norm} variable is an overall normalization which defines a conversion factor from the fluxes in the file to the units in \GLOBES .
+In general, there are many ways to give the source power of a
+neutrino source, such as neutrinos per proton on target per area per time frame. Right now, each flux has its own normalization factor, which is
+not always straightfoward to calculate. Often, one has to take into account
+many things, such as the number of target particles per unit mass. 
+In addition, the fluxes will be rescaled by $1/L^2$, which means that the
+normalization must contain a factor $L_0^2$.
+Here $L_0$ is the distance from the source for which the flux is given to the actual neutrino production region.
+At the end, it is left to the user to ensure that the
+numbers in the flux file give, after the multiplication with {\tt @norm}, 
+the proper numbers of produced neutrinos corresponding to the chosen target power {\tt @power}.
+Usually this adjustment of {\tt @norm} is performed by comparison with known energy spectra for a specific experiment.
+
+The software assumes that the given flux file has seven columns and
+501 lines with equidistant energies. The format is:
 \begin{quotation}
@@
 \end{quotation}
-where the energy values are evenly spaced. 
-For accessing fluxes at 
-arbitrary energies linear interpolation is used. Once the energy leaves the
-range of values given in the file $0.0$ is returned. The units used for the
-fluxes are not specified and right now each flux has its own normalization 
-factor as specified by {\tt norm}. The problem is that
-in computing the normalization one has to know things like number of target
-particles per unit mass asf. which is not always straightforward. The fluxes
-will be rescaled by $1/L^2$, thus the normalization must contain a factor
-$L_0^2$, when $L_0$ is the distance from the source for which the flux is
-given.
-
-The most obvious property of the neutrino detection is the total
-active mass of the target as already defined in section~\ref{sec:source} and
-its value is defined by 
-\begin{quote}
-{\tt \$target\_mass = 50.0 }
-\end{quote}
+In order to access fluxes at arbitrary energies, linear interpolation 
+is used by the software.
+In general, it is advisable to provide the flux between {\tt \$sampling\_min} and {\tt \$sampling\_max} (\cf, \Sec~\ref{sec:energy}), since these values are used by the software.
+However, if the energy leaves the range of values given in the file, zero is returned.
+Note that unused (and therefore often unknown) fluxes can be initialized with zeros, but can not just be omitted.
+For example, if one wants to define an experiment with different running times with neutrinos and antineutrinos, one can just use the same flux file in two different {\tt flux} environments with two different running times {\tt @time}.
+The unused fluxes in each case will then just be ignored.
+
+% Continue here (WW)
 
@@
 
-The purpose of this module is to specify the path the neutrinos
-are traveling. The central piece of information is the total length
-of the path - the baseline\index{Baseline}
-\begin{equation}
-L\,\left[\mathrm{km}\right]
-\end{equation} 
-The baseline is set by 
+The baseline and matter density profile determine, besides energy and
+involved flavors, the neutrino oscillation physics at the experiment
+description level. All of the neutrino oscillation parameters are defined at running time.
+
+The baseline\index{Baseline} is given by
 \begin{quote}
@@
 \end{quote}
-
-Furthermore there is
-the matter profile encountered along the baseline which can be set in
-several ways. The simplest way is to use a constant density corresponding
-to the average density along the baseline calculated using the 
-PREM\cite{Stacey} onion shell model of the 
-Earth\index{PREM}\index{Earth matter density}.
-\begin{quote}
-TBD
-\end{quote}
-
-Another possibility is to use several layers of constant density each in 
-order to improve the approximation to the actual Earth matter profile. In this
-case the user specifies the total baseline and the number of layers. The
-time consumed in computing the oscillation probabilities is directly 
+Note that baseline lengths are always assumed to be in
+kilometers.
+
+\begin{table}[t!]
+\begin{tabular}{|clp{7cm}|}
+\hline
+{\tt \$profiletype} & Additional variables & Description \\ 
+\hline
+$1$ & - & Average density (constant) \\
+$2$ & {\tt \$densitysteps} & PREM profile with given number of equidistant steps \\
+$3$ & {\tt \$lengthtab},  {\tt \$densitytab} & Arbitrary profile (table of layer thicknesses, table of densities) \\
+\hline
+\end{tabular}
+\caption{\label{tab:profiletypes} Different matter density profiles which can be used with \GLOBES .}
+\end{table}
+
+Furthermore, the matter density profile along the baseline
+has to be specified.
+The simplest matter density profile is a constant matter density profile with the average matter density from the PREM\cite{Stacey} onion shell model of the Earth\index{PREM}\index{Earth matter density}:
+\begin{quote}
+{\tt \$profiletype=1 }
+\end{quote}
+%
+For a better approximation of the realistic Earth matter density profile, one can use an arbitrary number of equidistant steps of the PREM profile:
+\begin{quote}
+{\tt \$profiletype=2 } \\
+{\tt \$densitysteps=20 }
+\end{quote}
+Note that the value of {\tt \$densitysteps} is time-critical, since the
+computation time of oscillation probabilities is directly 
 proportional to the number of layers.
-\begin{quote}
-TBD
-\end{quote}
-
-The third possibility to specify the matter profile is to provide a list
-of thicknesses and densities for the layers. 
-\begin{quote}
-{\tt \$densitytab=\{ 2.8, 3.5\}}\\
-{\tt \$lengthtab=\{ 1000.0, 2000.0\}}\\
-\end{quote}
-It is important that both list are of equal length and that the sum
-of the thicknesses given in  {\tt \$lengthtab} add up to the value of
-{\tt \$baseline}.
-
-For any type of profile a central value $a$ and a corresponding error 
-$\delta a$ have to be specified\index{Density error}\index{Density center}.
+%
+As a  third possibility, one can specify the matter density profile 
+manually with a list of thicknesses and densities of the matter density layers. This example uses two density steps with two different densities:
+\begin{quote}
+{\tt \$profiletype=3 } \\
+{\tt \$densitytab=\{2.8, 3.5\}}\\
+{\tt \$lengthtab=\{1000.0, 2000.0\}}\\
+\end{quote}
+It is important that both lists habe the same length and that the  thicknesses given in  {\tt \$lengthtab} add up to the value of
+{\tt \$baseline}. In addition, matter densities are always given in $g/cm^3$.
+%
+This approach can also be used for a constant matter density profile with
+a specific matter density:
+\begin{quote}
+{\tt \$profiletype=3 } \\
+{\tt \$densitytab=\{3.5\}}\\
+{\tt \$lengthtab=\{3000.0\}}\\
+\end{quote}
+The possible options for matter density profiles are summarized in \Tab~\ref{tab:profiletypes}.
+
+In addition to the matter density profile, \GLOBES\ implements a simple treatment of matter density uncertainties by a scaling factor $\hat{\rho}$
+of the matter density profile:
 \begin{equation}
 \label{eq:density_error}
-\rho(x)=a\cdot\rho_0(x)\,\left[\frac{\mathrm{g}}{\mathrm{cm}^2}\right]\,
-\end{equation}
-where $\rho_0(x)$ is the matter profile as chosen by the user and $\rho(x)$ is
-the one used in the calculation of the oscillation probabilities. The default
-value for $a$ is $1$ and a typical value for $\delta a = 0.05$.
+\rho(x)=\hat{\rho} \cdot\rho_0(x)\,\left[\frac{\mathrm{g}}{\mathrm{cm}^3}\right] \, .
+\end{equation}
+Here $\rho_0(x)$ is the matter profile chosen by the user, and $\rho(x)$ is
+the one used in the calculation of the oscillation probabilities.
+As it is described in \Sec~\ref{sec:externalinput}, the matter density uncertainty is implemented by using $\hat{\rho}$ as an additional oscillation parameter for each experiment, and by assuming to know $\hat{\rho}$ with an external precision.
+Thus, the central value of $\hat{\rho}$ (default value $1.0$) and its error $\delta \hat{\rho}$ have to be defined.
+The following code assumes a matter density uncertainty of $5\%$:\index{Density error}\index{Density center}
+\begin{quote}
+ {\tt \$densitycenter=1.0} \\
+ {\tt \$densityerror=0.05}
+\end{quote}
+In most cases, this is in combination with a constant profile a very good approximation to take into account the realistic structure of the profile as well as its uncertainty~\cite{Ohlsson:2003ip}. 
+ If one does not want to take into account  matter density uncertainties, 
+one can either a very small value of $\hat{\rho}$, or the minimizer {\tt ChiNP} with $\hat{\rho}$ fixed in the application software.
 
@@
 
-Cross sections\index{Cross sections} are used as part of the 
-channel definition 
-(see section~\ref{sec:channel}) and have to be provided by the user in form
-of a file. Cross section are in general given as differential cross section
-divided by the energy
+Cross sections\index{Cross sections} will later be used as part of the 
+channel definition (see \Sec~\ref{sec:channel}). Similar to the source 
+fluxes, they are provided by the user as a data file:
+\begin{quote}
+{\tt cross(\#name)<}\\
+{\tt \tb @cross\_file =user\_file\_1.dat}\\
+{\tt >}
+\end{quote}  
+This cross section can later be refered to by {\tt \#name}.
+
+Cross sections are in \GLOBES\ given as differential cross section per energy:
 \begin{equation}
@@
 \end{equation}
-The cross section files have seven rows and $1001$ lines.
+The software assumes that the cross section files are text files with 
+seven columns and $1001$ lines of the form
 \begin{quotation}
-$log_{10} E\quad
+$\mathrm{log}_{10} E\quad
 \hat\sigma_{\nu_e}\quad
@@
 \end{quotation}
-where the logarithm of the energy values is evenly spaced. 
-For accessing cross section at 
-arbitrary energies linear interpolation is used. Once the energy leaves the
-range of values given in the file $0.0$ is returned.
-
-A cross section file is read by 
-\begin{quote}
-{\tt flux(\#name)<}\\
-{\tt \tb @cross\_file =user\_file\_1.dat}\\
-{\tt >}
-\end{quote}  
-and can later be refered to by {\tt \#name}.
+Here the logarithms of the energy values have to be equidistant. For 
+arbitrary energies, linear interpolation is used. If the energy leaves the
+range of values given in the file, $0.0$ will be assumed.
+In general, it is advisable to provide the cross sections in the range between {\tt \$sampling\_min} and {\tt \$sampling\_max} (\cf, \Sec~\ref{sec:energy}).
+Unused cross sections have to be filled with zeros, and can not be just omitted.
+
+% WW: CONTINUE HERE:
 

### 2004-05-13 Documentation/part2.tex
*** empty log message ***

@@
 
-% WW: CONTINUE HERE:
-
 \section{Oscillation channels}
@@
 
-Channels\index{Channel} represent an intermediate level in between 
-the pure oscillation 
-physics given by the oscillation probability $P_{\alpha\beta}$ and the
-detected signal and background. A channel basically describes the path from one
-initial state in the source to one interaction type (IT) in  the detector.  
-Therefore a channel consists of the description of the initial state given 
-by the polarity 
-of the source, \eg\ muons or anti-muons in a neutrino factory, by the CP 
-eigenvalue of the initial state, \ie\ neutrino or anti-neutrino and by the 
-initial flavour. A channel furthermore consists of the interaction type in the
-detector given by the final flavour of the neutrino, the interaction cross
-section and the energy resolution function.
-
-For the calculation of event rates, the first step is to compute the number of
-events for each IT in the fiducial mass of the detector for each 
-neutrino flavor and energy bin. The second step is to include the detector 
-effects coming from
-the insufficient knowledge used in the event reconstruction. We combine these
-two steps in order to obtain the differential event rate spectrum for each
-flavor and IT as seen by the detector, which we call a ``channel''.
-
-The master formula for the differential event rate for each channel 
-is given by
+Channels\index{Channel} in \GLOBES\ represent an intermediate level 
+between the pure oscillation physics given by the oscillation probability
+$P_{\alpha\beta}$, and the total event rates composed of signal and 
+background. A channel describes the path from one
+initial neutrino flavor in the source to the event rates in the detector for one specific interaction type (IT) and final flavor.  
+Therefore, a channel contains the description of the 
+initial neutrino flavor, its CP eigenvalue (neutrino or antineutrino), 
+the detected neutrino flavor, the interaction cross sections for the chosen interaction type, and the energy resolution function of the detector.
+
+Before we come to the definition of the channel in AEDL, we introduce the general concept for the calculation of event rates. The first step is to
+compute the number of events for each IT in the detector for each 
+initial and final neutrino flavor and energy bin.
+The second step is to include the detector effects coming from the insufficient knowledge in the event reconstruction.
+These two steps combined lead to the differential event rate spectrum for each initial and final flavor and IT as seen by the detector, which we call the ``channel''.
+In this section, we focus on the
+ first step, \ie , we discuss the definition of the energy resolution function in the next section, since this is a rather comprehensive issue.
+
+The differential event rate for each channel is given by
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{eq:master_event}
-\frac{dn_{\beta}^{\text{IT}}}{dE'}=&&N\,\int \int dE\,d\hat{E}\quad
+\frac{dn_{\beta}^{\text{IT}}}{dE'}=&&N\,\int\limits_0^\infty \int\limits_0^\infty dE\,d\hat{E}\quad
 \underbrace{\Phi_{\alpha} (E)}_{\mathrm{Production}} \times \nonumber\\
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-where $\alpha$ is the initial flavor of the neutrino and 
-$\beta$ is the final flavour, 
-$E$ is the incident neutrino energy, $\Phi_{\alpha} (E)$ is the flux of the 
+where $\alpha$ is the initial flavor of the neutrino, 
+$\beta$ is the final flavor, $\Phi_{\alpha} (E)$ is the flux of the 
 initial flavor at the
 source, $L$ is the baseline length, $N$ is a normalization factor, and 
-$\rho$ is the matter density. The interaction term is composed of 
+$\rho$ is the matter density. The energies in this formula are given as follows:
+\begin{itemize}
+\item
+ $E$ is the incident neutrino energy, \ie, the actual energy of the 
+incoming neutrino (which is not directly accessible to the experiment)
+\item
+ $\hat{E}$ is the energy of the secondary particle
+\item
+ $E'$ is the reconstructed neutrino energy, \ie, the measured
+neutrino energy as obtained from the experiment
+\end{itemize}
+The interaction term is composed of 
 two factors, which are the total cross section 
@@
 the interaction type IT, and the energy distribution of the 
-secondary particle $k_\beta^{\text{IT}}(E-\hat{E})$ with 
-$\hat{E}$ the energy of the secondary particle. The detector properties are 
+secondary particle $k_\beta^{\text{IT}}(E-\hat{E})$.
+The detector properties are 
 modeled by the threshold function $T_\beta(\hat{E})$, coming from the the 
 limited resolution or the cuts in the analysis, and the energy resolution 
-function $V_\beta(\hat{E}-E')$ of the secondary particle. Thus, $E'$ is the 
-{\em reconstructed} neutrino energy.
-
-Since it is rather cumbersome to numerically solve this double integral,
-we split the two integrations. We first evaluate the integral over
+function $V_\beta(\hat{E}-E')$ of the secondary particle.
+
+Since it is a lot of effort to solve this double integral numerically,
+we split up the two integrations. First, we evaluate the integral over
 $\hat{E}$, where the only terms containing $\hat{E}$ are
 $k_\beta^{\text{IT}}(E-\hat{E})$,  $ T_\beta(\hat{E})$, and 
-$ V_\beta(\hat{E}-E')$, and define:
+$ V_\beta(\hat{E}-E')$. We define:
 \begin{eqnarray}
@@
  \equiv
-\int d\hat{E} \quad T_\beta(\hat{E})\,k_\beta^{\text{IT}}(E-\hat{E})
-\,V_\beta(\hat{E}-E')\,, 
-e\end{eqnarray}
-
-Thus $R_\beta^{\text{IT}}(E,E')$ describes the energy response of 
-the detector, \ie\ a neutrino with a (true) energy of $E$ has a probability
-of $R_\beta^{\text{IT}}(E,E') dE'$ to have a reconstructed energy 
-in between $E'$ and $E'+dE'$. The function $R(E,E')$ is also refered to
-as energy resolution function or smearing matrix in its actual representation
-in the program\index{Energy resolution}\index{Smear matrix}. The detailed
-definition and initialization of the energy resolution function is described
-in section~\ref{sec:energy}.
-
-The number of events in one channel and  in one bin\index{Bin} $i$ is given by
+\int\limits_0^\infty d\hat{E} \quad T_\beta(\hat{E})\,k_\beta^{\text{IT}}(E-\hat{E})
+\,V_\beta(\hat{E}-E')\,. 
+\end{eqnarray}
+Thus, $R_\beta^{\text{IT}}(E,E')$ describes the energy response of 
+the detector, \ie , a neutrino with a (true) energy $E$ is reconstructed
+with an energy between $E'$ and $E'+dE'$ with a probability
+$R_\beta^{\text{IT}}(E,E') dE'$. The function $R(E,E')$ is also often called ``energy resolution function''. Actually, its internal representation
+in the software is a smearing matrix\index{Energy resolution}\index{Smear matrix}.
+The function $\epsilon_\beta^{\text{IT}}(E')$ will later be refered to as ``post-smearing efficiencies'', since it will allow us to define cuts and threshold functions {\em after} the smearing is performed, \ie, as function of $E'$.
+The detailed definition and initialization of the energy resolution function is described in \Sec~\ref{sec:energy}.
+
+Eventually, we can write down the number of events per bin $i$ \index{Bin} and channel $c$ as
 \begin{equation}
 \label{eq:channel}
-n_i^c=\int_{E_i-\Delta E/2}^{E_i+\Delta E/2} dE' \quad
-\frac{dn_{\beta}^{\text{IT}}}{dE'} (E')
-\end{equation}
-where $c$ is the channel index.
+n_i^c=\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' \quad
+\frac{dn_{\beta}^{\text{IT}}}{dE'} (E').
+\end{equation}
+This means that one has to solve this integral
+\begin{eqnarray}
+\label{eq:events_bin}
+n_i^c=N/L^2\,\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
+\quad \int\limits_0^\infty dE \,\, \Phi^c(E)\,
+P^c(E)\,
+\sigma^c(E)\,
+R^c(E,E')\,
+\epsilon^c(E')\,.
+\end{eqnarray} 
 Note that the events are binned according to their \emph{reconstructed} energy.
-In order to perform the computation \GLOBES\ needs to know the flux,
-the CP-sign of the initial state, the initial flavour, the final flavour,
-the cross section and the energy resolution function. Employing the notation
-of section~\ref{sec:energy} the full formula is
-
-Thus the actual definition of a channel looks like
+
+A simple channel definition in \GLOBES\ consists of the flux,
+the CP-sign of the initial state, the initial flavor, the final flavor,
+the cross sections, and the energy resolution function. In order to refer to
+flux, cross sections, and energy resolution functions, they have to be 
+defined before with their {\tt \#name} in the respective environments. 
+Thus, a simple definition of a channel is
 \begin{quote}
@@
 \end{quote}
-Furthermore is is possible to define energy dependent efficiencies and
-backgrounds (non-beam). These additional components can be either added
-before the convolution in with $R(E,E')$ is done, \ie\ the integration with
-respect to $E$ in equation~\ref{eq:e_res}. This is called 
-{\tt @pre\_smearing\_$\ldots$}. The pre-smearing efficiencies and backgrounds
-have to have as many elements as there are sampling points 
-(see section~\ref{sec:energy}). In same cases the number of sampling points
-is determined automatically by \GLOBES and the number of elements in the 
-pre-smearing vectors has to be adjusted in the second run.
-Of course, it is possible after all
-integrations been performed to have similar factors  and this is called  
-{\tt @post\_smearing\_$\ldots$}. All post-smearing factors have to have as 
-many elements as there are bins. Efficiencies are multiplicative and their
-default value is $1$ whereas backgrounds are additive and thus their default
-value is $0$. Thus a more elaborate channel looks like
-\begin{quote}
-{\tt channel(\#channel\_1)<\\
-\tb @channel = \#flux : $+$: muon: muon: \#cross: \#energy\\
-\tb @pre\_smearing\_background = \{1,2,3,4,5,6,7,8,9,10\}\\
-\tb @post\_smearing\_efficiencies = \{0.1,0.2,0.3,0.4,0.5\}\\
->}
-\end{quote}
-This experiment obviously uses $10$ sampling points and $5$ bins.
+Note that the energy environment will be described in the next section.
+
+% WW: HIER WEITER
 
@@
 \label{sec:energy}
+
+The definition of the energy resolution function is rather sophisticated in 
+\GLOBES . In particular, since this aspect is rather time-critical, several
+algorithms are implemented. The choice of the proper algorithm depends on the
+experiment and the frequencies of the involved neutrino oscillations.
+
+In this section, we first discuss the principles of the energy smearing, where it is assumed that the reader is familiar with \Sec~\ref{sec:channel}.
+Then we introduce an automatic energy smearing algorithm, which is fairly simple to understand and applicable to most beam-based experiments.
+In most cases, the reader may want to continue to the next section after reading these two subsections.
+In the third subsection, we describe a more elaborate (and slower) smearing algorithm, which can be used together with rather fast neutrino oscillations compared to the bin size, such as for reactor experiments.
+Eventually, we show how one can use a manual smearing matrix instead of using one of the implemented algorithms.
+
+\subsection{Introduction and principles}
+
+\begin{figure}[t!]
+\begin{center}
+\includegraphics[width=15cm]{energies}
+\end{center}
+\caption{\label{fig:energies} The different evaluation levels for the energy smearing in \GLOBES .}
+\end{figure}
+
+DESCRIBE PICTURE!
 
@@
 
-In order to obtain the number of events 
-$n_i^c$ from one channel $c$ in one bin $i$ one has to solve this integral
-%
-\begin{eqnarray}
-\label{eq:events_bin}
-n_i^c=N/L^2\,\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
-\quad \int dE \,\, \Phi^c(E)\,
-P^c(E)\,
-\sigma^c(E)\,
-R^c(E,E')\,
-\epsilon^c(E')\,.
-\end{eqnarray} 
-%
-
 Instead of using directly equation~\ref{eq:e_res} we will use a different
@@
 \end{quote}
+
+ADJUST TO SECTION AND REPLACE:
+
+In combination with the channels, is is possible to define energy 
+dependent efficiencies and (non-beam) backgrounds.
+Efficiencies are multiplicative factors, whereas backgrounds are added to the event rates.
+These components can be introduced before or after the convolution in \eq~(\ref{eq:e_res}) is done.
+If they are introduced before, we call them
+{\tt @pre\_smearing\_efficiencies} or {\tt @pre\_smearing\_background}. 
+If they are introduced after, we call them {\tt @post\_smearing\_efficiencies} or {\tt @post\_smearing\_background}.
+Note that pre-smearing components are always afunction of the incident neutrino energy $E$.
+Thus, there have to be as  many elements as there are sampling points (see \Sec~\ref{sec:energy}).
+In same cases, the number of sampling points is determined automatically by \GLOBES , and the number of elements in the  pre-smearing vectors has to be adjusted in the second run.
+Examples for pre-smearing quantities are non-beam backgrounds, such as from geophysical neutrinos.
+The post-smearing components are always a function of the reconstructed neutrino energy $E'$, such as the post-smearing efficiencies $\epsilon_\beta^{\text{IT}}(E')$ in \eq~(\ref{eq:e_res}).
+Examples for post-smearing efficiencies are cuts and detection threshold functions.
+All post-smearing components have to have as
+many elements as there are energy bins. Efficiencies are multiplicative 
+and their default value is $1$, whereas backgrounds are additive and their default value is $0$. Thus, a more elaborate channel can be defined as
+\begin{quote}
+{\tt channel(\#channel\_1)<\\
+\tb @channel = \#flux : $+$: muon: muon: \#cross: \#energy\\
+\tb @pre\_smearing\_background = \{1,2,3,4,5,6,7,8,9,10\}\\
+\tb @post\_smearing\_efficiencies = \{0.1,0.2,0.3,0.4,0.5\}\\
+>}
+\end{quote}
+This experiment uses $10$ sampling points and $5$ bins.
+
+\subsection{Bin based automatic energy smearing}
 
@@
 \end{quote}
+
+\subsection{Low-pass filter based automatic energy smearing}
 
@@
 
+\subsection{Manual energy smearing}
+
 Finally the user has the possibility to directly specify the smearing data.

### 2004-05-17 Documentation/part2.tex
*** empty log message ***

@@
 
-\section{A simple example for the AEDL}
+\section{A simple example for AEDL}
 
@@
 n_i^c=\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' \quad
-\frac{dn_{\beta}^{\text{IT}}}{dE'} (E').
-\end{equation}
-This means that one has to solve this integral
+\frac{dn_{\beta}^{\text{IT}}}{dE'} (E') \,
+\end{equation}
+where $\Delta E_i$ is the bin size of the $i$th energy bin.
+This means that one has to solve the integral
 \begin{eqnarray}
@@
 
-% WW: HIER WEITER
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-The definition of the energy resolution function is rather sophisticated in 
-\GLOBES . In particular, since this aspect is rather time-critical, several
-algorithms are implemented. The choice of the proper algorithm depends on the
-experiment and the frequencies of the involved neutrino oscillations.
+The definition and implementation of the energy resolution function is 
+rather sophisticated in \GLOBES . In particular, since this aspect is 
+ time-critical, several algorithms are implemented. The choice of the 
+proper algorithm depends on the experiment and the frequencies of the
+ involved neutrino oscillations.
 
@@
 \subsection{Introduction and principles}
+
+The energy resolution function $R^c(E,E')$ has been already introduced in 
+\Sec~\ref{sec:channel}, where a definition
+has been given in \eq~(\ref{eq:e_res}). Instead of using
+ \eq~(\ref{eq:e_res}) directly, we apply a slightly different
+definition of the post-smearing efficiencies $\epsilon(E')$. 
+In general, $\epsilon(E')$ has to be
+determined by means of a Monte Carlo simulation of the experiment. 
+This usually involves a binning of the simulated events in the 
+reconstructed energy $E'$. Therefore, one simplify \eq~(\ref{eq:events_bin}) by
+\begin{equation}
+\label{eq:post_smearing}
+\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' \, R^c(E,E') \, \epsilon^c(E') = \hat\epsilon_i^c \cdot \int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
+\quad R^c(E,E')\,.
+\end{equation}
+Here the  $\hat\epsilon_i^c$ are the 
+``post-smearing'' efficiencies, which will be set within the corresponding {\tt channel} environment (see below).
+From \eq~(\ref{eq:events_bin}) it is obvious that the integration with respect to the reconstructed energy $E'$ can be
+performed independently of the oscillation parameters. We define
+the ``bin kernel'' $K_i^c$ for the $i$th bin as
+\begin{equation}
+\label{eq:kernel}
+K_i^c(E) \equiv \int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
+\quad R^c(E,E')\,.
+\end{equation}
+With this definition, \eq~(\ref{eq:events_bin}) can be re-written as
+\begin{equation}
+\label{eq:simple_int}
+n_i^c=N/L^2 \int\limits_0^\infty dE\quad  \underbrace{\Phi^c(E)\,
+P^c(E)\,
+\sigma^c(E)\,
+K_i^c(E)\,}_{f(E)}. 
+\end{equation}
+
+There is no principle reason why one should not evaluate this integral directly by the usual numerical methods. However, it turns out that this 
+is very slow in many cases. Therefore, we will introduce two different approximation schemes for different applications in the next two subsections.
+In either case, the integrand in \eq~(\ref{eq:simple_int}) has to be evaluated at fixed ``sampling points''. These sampling points can be 
+either defined by the user, or be computed by
+\GLOBES . The choice between these options also depends on 
+the algorithm which is used for the calculation, as it is explained in the following subsections.
 
@@
 
-DESCRIBE PICTURE!
-
-The origin and meaning of the energy resolution function 
-$R^c(E,E')$ was already discussed in 
-section~\ref{sec:channel} and  a mathematical definition
-was given in equation~\ref{eq:e_res}. Here we deal with
-how to tell \GLOBES\ which energy resolution function it should
-use. This also naturally involves the issue how the actual convolution
-of the differential event rate spectrum with $R_\beta^{\text{IT}}(E,E')$
-is performed. First we will discuss the mathematical structure of
-the problem and then the various algorithms and possibilities which
-\GLOBES\ offers are introduced.
-
-Before going into the details we define that the range from $E_\mathrm{min}$
-to $E_\mathrm{max}$ is divided in a certain number of bins $N_\mathrm{bins}$.
-These bins form a partition, which not necessarily is equidistant. The 
-partition is any case completely defined by specifying $E_\mathrm{min}$
-and $E_\mathrm{max}$ and the size of each bin  $\Delta E_i$. It is also
-straightforward to compute the midpoint of each bin $E_i$. This information
-has to be provided by the user. Equidistant bins are obtained by
+Before we come to the calculation algorithms, it is useful to understand
+the general evaluation algorithm. As it is illustrated in \figu{energies}, 
+\GLOBES\ uses several levels with respect to the energy ranges:
+\begin{description}
+\item[Sampling point level]
+ This level is used internally to evaluate the integrand in \eq~(\ref{eq:simple_int}) at all sampling points.
+The energy scale is the actual incident neutrino energy $E$.
+Note that the sampling range should exceed the energy bin range by about the energy resolution error, since these sampling points need to be evaluated by the algorithm.
+However, this is only
+necessary to avoid aliasing effects if the integrand in \eq~(\ref{eq:simple_int}) is large at the limits of the sampling range.
+For a manual definition of the sampling points (depends on algorithm), use
+\begin{quote}
+{\tt
+\$sampling\_points = 20\\
+\$sampling\_min =          4.0\\
+\$sampling\_max =         50.0
+}
+\end{quote}
+for equidistant sampling points. Arbitrary sampling points can be specified analogously to the bins (see below).
+\item[Bin level] This level is determined by the experiment. Note than energy
+bin sizes much smaller than the energy resolution will not improve the results.
+The energy bin range and the number of energy bins do always have to be specified.
+For the case of large values of the integrand in \eq~(\ref{eq:simple_int}) at the energy range limits, it is recommended to exceed the analysis energy window by about the energy calibration error in order to avoid aliasing effects.
+
+In order to define a range between $E_\mathrm{min}$
+and $E_\mathrm{max}$ divided by a certain number of equidistant bins,
+use
 \begin{quote}
@@
 \end{quote}
-Another possibility is to specify arbitrary bin sizes and to leave out
-the number of bins itself, by
+For arbitrary bins, use $E_\mathrm{min}$
+and $E_\mathrm{max}$ and the size of each bin  $\Delta E_i$:
 \begin{quote}
@@
 \end{quote}
-
-Instead of using directly equation~\ref{eq:e_res} we will use a different
-definition concerning $\epsilon(E')$. In general  $\epsilon(E')$ has to be
-determined by means of a Monte Carlo simulation of the experiment. This
-usually involves anyhow a binning of the simulated events in the reconstructed
-energy $E'$, therefore one more easily can obtain  $\hat\epsilon_i$ defined
-by
-\begin{equation}
-\label{eq:post_smearing}
-\hat\epsilon_i \cdot \int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
-\quad R^c(E,E')\,.
-\end{equation}
-These  $\hat\epsilon_i$ are refered to as so called 
-``post-smearing'' efficiencies and set within the corresponding {\tt channel}
-environment.
-The integration with respect to the reconstructed energy $E'$ can be
-performed independently of the oscillation parameters and we define
-the bin kernel for the $i$th bin $K_i^c$
-\begin{equation}
-\label{eq:kernel}
-K_i^c(E):=\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
-\quad R^c(E,E')\,.
-\end{equation}
-With this definition equation~\ref{eq:events_bin} can be rewritten as
-\begin{equation}
-\label{eq:simple_int}
-n_i^c=N/L^2 \int dE\quad  \underbrace{\Phi^c(E)\,
-P^c(E)\,
-\sigma^c(E)\,
-K_i^c(E)\,}_{f(E)}. 
-\end{equation}
-
-There is no basic reason why one could not evaluate this integral directly
-by the usual numerical methods. However it turns out that this may be rather
-slow, therefore we introduce two different approximation schemes which have
-a varying degree of accuracy and speed. In any case each scheme to evaluate this
-integral involves a the evaluation of the integrand at fixed sampling points.
-These sampling points can be either defined by the user or be computed by
-\GLOBES, the choice between these possibilities also depends on the algorithm
-which is used for the actual calculation, as explained in the following.
-If the user chooses to set the sampling points the procedure is analogous
-to the one setting the bins. For equidistant sampling points use
-\begin{quote}
-{\tt
-\$simbins = 20\\
-\$simtresh =          4.0\\
-\$simbeam =         50.0
-}
-\end{quote}
-The advantage in setting the sampling points in this way is that in case many
-different energy resolution functions are used the user has the control
-how the different sampling requirements are fulfilled. In case a user-defined
-smear matrix is used this data is mandatory.
-
-All energy resolution function are defined within an {\tt energy} environment
-and can be refered to by {\tt \#name}.
-\begin{quote}
-  {\tt energy(\#name)<\\
-\tb $\ldots$\\
->}
-\end{quote}
-
-ADJUST TO SECTION AND REPLACE:
-
-In combination with the channels, is is possible to define energy 
+The number of bins will be automatically computed by \GLOBES . Note that the
+bin sizes have to add up to the energy range {\tt \$emax}-{\tt \$emin}.
+%
+It is also
+straightforward to compute the midpoint of each bin $E_i$. This information
+has to be provided by the user ???? HOW ????. 
+\item[Analysis level]
+On the analysis level, an energy window can be defined within each rule. For
+details, see next chapter.
+\end{description}
+In general, the energy smearing happens between the sampling point and bin levels, which means that the energy smearing matrix will have {\tt \$sampling\_points} columns and {\tt \$bins} rows.
+
+As illustrated in the figure, an interesting feature in combination with the
+channels are pre- and post-smearing effects. Pre-smearing effects are taken into account on the sampling point level, and post-smearing effects on the bin level. Examples for these effects are energy 
 dependent efficiencies and (non-beam) backgrounds.
 Efficiencies are multiplicative factors, whereas backgrounds are added to the event rates.
-These components can be introduced before or after the convolution in \eq~(\ref{eq:e_res}) is done.
-If they are introduced before, we call them
+These components can be introduced before or after the convolution in \eq~(\ref{eq:post_smearing}) is done.
+If they are introduced before,
+we call them
 {\tt @pre\_smearing\_efficiencies} or {\tt @pre\_smearing\_background}. 
@@
 This experiment uses $10$ sampling points and $5$ bins.
+
+In the following subsections we will define the energy resolution function.
+All energy resolution functions are defined within an {\tt energy} environment and can be refered to by {\tt \#name}.
+\begin{quote}
+  {\tt energy(\#name)<\\
+\tb $\ldots$\\
+>}
+\end{quote}
+
+%% WW: CONTINUE HERE
 

### 2004-05-17 Documentation/part2.tex
*** empty log message ***

@@
 Note that the energy environment will be described in the next section. 
+In addition, one can define pre- and post-smearing effects together
+with the channels, which will also be introduced together with the
+energy resolution function in the next section.
 
@@
 Here the  $\hat\epsilon_i^c$ are the 
-``post-smearing'' efficiencies, which will be set within the corresponding {\tt channel} environment (see below).
+binned ``post-smearing'' efficiencies, which will be set within the corresponding {\tt channel} environment (see below).
 From \eq~(\ref{eq:events_bin}) it is obvious that the integration with respect to the reconstructed energy $E'$ can be
@@
 \label{eq:simple_int}
-n_i^c=N/L^2 \int\limits_0^\infty dE\quad  \underbrace{\Phi^c(E)\,
+n_i^c=N/L^2 \,
+\hat\epsilon_i^c \, \int\limits_0^\infty dE\quad  \underbrace{\Phi^c(E)\,
 P^c(E)\,
-\sigma^c(E)\,
-K_i^c(E)\,}_{f(E)}. 
+\sigma^c(E) \, K_i^c(E)\,}_{f(E)}. 
 \end{equation}
@@
 The energy scale is the actual incident neutrino energy $E$.
-Note that the sampling range should exceed the energy bin range by about the energy resolution error, since these sampling points need to be evaluated by the algorithm.
-However, this is only
-necessary to avoid aliasing effects if the integrand in \eq~(\ref{eq:simple_int}) is large at the limits of the sampling range.
+Note that the sampling range should exceed the energy bin range by about the energy resolution error, since the corresponding sampling points have to be evaluated by the algorithm.
+However, this exceeding is only
+necessary to avoid aliasing effects, if the integrand in \eq~(\ref{eq:simple_int}) is large at the limits of the sampling range.
 For a manual definition of the sampling points (depends on algorithm), use
@@
 \end{quote}
-for equidistant sampling points. Arbitrary sampling points can be specified analogously to the bins (see below).
+for equidistant sampling points. Arbitrarily spaced sampling points can 
+be specified analogously to the energy bins below. 
+Note that the algorithm expects to 
+have at least as many sampling points as energy bins, since otherwise 
+the number of (unphysical) sampling constrains the experiment performance.
 \item[Bin level] This level is determined by the experiment. Note than energy
@@
 bin sizes have to add up to the energy range {\tt \$emax}-{\tt \$emin}.
-%
-It is also
-straightforward to compute the midpoint of each bin $E_i$. This information
-has to be provided by the user ???? HOW ????. 
 \item[Analysis level]
@@
 Efficiencies are multiplicative factors, whereas backgrounds are added to the event rates.
-These components can be introduced before or after the convolution in \eq~(\ref{eq:post_smearing}) is done.
+These components can be introduced before or after the integration in \eq~(\ref{eq:simple_int}) is done.
 If they are introduced before,
@@
 If they are introduced after, we call them {\tt @post\_smearing\_efficiencies} or {\tt @post\_smearing\_background}.
-Note that pre-smearing components are always afunction of the incident neutrino energy $E$.
-Thus, there have to be as  many elements as there are sampling points (see \Sec~\ref{sec:energy}).
-In same cases, the number of sampling points is determined automatically by \GLOBES , and the number of elements in the  pre-smearing vectors has to be adjusted in the second run.
+Note that pre-smearing components are always a function of the incident neutrino energy $E$.
+Thus, there have to be as  many elements as there are sampling points.
+For some algorithms, the number of sampling points is determined automatically by \GLOBES , and the number of elements in the  pre-smearing vectors has to be adjusted in the second run.
 Examples for pre-smearing quantities are non-beam backgrounds, such as from geophysical neutrinos.
@@
 \end{quote}
-
-%% WW: CONTINUE HERE
-
-\subsection{Bin based automatic energy smearing}
-
-First we discuss the builtin algorithms. The simplest, fastest and
-least accurate is algorithm A. The key idea is to choose a set of assumptions
-which allow to evaluate the integrand of equation~\ref{eq:simple_int}
-at fixed values of the energy $E$ and to use the same number of points $N$ in
-the calculation as there are bins. Thus equation~\ref{eq:simple_int} is reduced
-to
+The individual parameters of the environment will be defined below and depend on the algorithm used.
+
+\subsection{Bin-based automatic energy smearing}
+
+This algorithm is the simplest of the built-in algorithms for the evaluation
+of \eq~(\ref{eq:simple_int}). It is applicable to most of the
+ accelerator-based experiments which can be simulated with \GLOBES .
+
+The key idea is to use a ``flat'' model with sampling points identical to 
+the bins (\cf, \figu{energies}). This is a good approximation as long as
+\begin{itemize}
+\item
+ The number of bins is larger than the energy resolution.
+\item
+ The integrand in \eq~(\ref{eq:simple_int}) is small at the bin range limits
+{\tt \$emin} and {\tt \$emax} (in order to avoid aliasing effects).
+\item
+ The neutrino oscillations are slow on a scale of the bin distance.
+\end{itemize}
+In this case, \eq~(\ref{eq:simple_int}) is reduced to
 \begin{equation}
 \label{eq:algo_one}
-n_i^c=N/L^2 \Delta E_i \sum_{j=1}^N \quad  \Phi^c(E_j)\,
+n_i^c=N/L^2 \, \sum_{j=1}^N \,  \Phi^c(E_j)\,
 P^c(E_j)\,
 \sigma^c(E_j)\,
-K_i^c(E_j)\,.
-\end{equation}
-The advantage of using equation~\ref{eq:algo_one} is that all factors
-independent of the oscillation parameters have to be evaluated only at 
-values of $E$ which are known in advance, \ie\ they can be put 
-into a look-up table. Another advantage is that also the probability
-has to be evaluated only at previously known values of the energy, which
+K_i^c(E_j) \, \Delta E_j \,.
+\end{equation}
+The advantages of this algorithm are obvious: All factors
+independent of the oscillation parameters have to be only evaluated once
+at values of $E$ which are known in advance, which means that they can be put 
+into a look-up table. In addition, the probability
+has to be only evaluated at previously known values of the energy, which
 makes it possible to compute the transition amplitudes for all channels
-in one step. The key assumption is that all involved factors are piece-wise
-constant, \ie\ they do not change within one bin. This assumption seems to be
-very restrictive, note however whenever one is analyzing simulated 
-data\footnote{simulated with the same algorithm} the errors
-made by violating this assumption nearly cancel between the simulated 
-data and the fitted data. This algorithm is the fastest which \GLOBES\ offers
-and is suitable whenever there are no fast oscillations. Otherwise severe
-aliasing effects may occur which usually lead to a failure of one of the 
-minimizing routines. This algorithm is selected by
+simultaneously. One assumption is that all involved factors are piece-wise
+constant, \ie, they hardly change within each bin. This 
+assumption seems to be very restrictive. Note, however, if one analyzes simulated data (which are simulated with the same algorithm), the errors
+will cancel between the simulated and fitted data. 
+%
+Furthermore, we will demonstrate in the next subsection that the assumption
+is rather good. ??? COMES THERE ???
+%
+This algorithm is the fastest which \GLOBES\ offers, and is selected by
 \begin{quote}
@@
 \end{quote}
+within the {\tt \#energy} environment.
 The computation of the bin kernel $K_i^c$ is performed
-by \GLOBES. To this end it requires  the number of bins
+by \GLOBES. Thus, it requires  the number of bins
  $N_\mathrm{bins}$
-and the minimum $E_\mathrm{min}$ and maximum energy $E_\mathrm{max}$. 
-Furthermore following parameterization
-for $R^c(E,E')$ is used
-\begin{equation}
-R^c(E,E')=\frac{1}{\sigma(E)\,\sqrt{2\pi}}\,e^{-\frac{(E-E')^2}{2\sigma^2(E)}}
-\end{equation} 
-There are several energy resolution functions available and by default
-{\tt \#standard} is used. They can be chosen
-by
+and the minimum $E_\mathrm{min}$ and maximum energy $E_\mathrm{max}$
+for equidistant bins.
+
+As far as the parameterization for the energy resolution function 
+$R^c(E,E')$ in \eq~(\ref{eq:kernel}) is concerned, the algorithm uses
+the Gaussian
+\begin{equation}
+R^c(E,E')=\frac{1}{\sigma(E)\,\sqrt{2\pi}}\,e^{-\frac{(E-E')^2}{2\sigma^2(E)}} \, .
+\end{equation} 
+CHECK FACTOR OF TWO IN DENOMINATOR WITH SOFTWARE (IMPORTANT)!!! (NORMALIZATION IS OK!)
+There are several energy resolution functions available, where by default
+{\tt \#standard} is used:
 \begin{quote}
@@
 \end{quote}
-where the energy resolution function named {\tt \#standard} is defined by
+The energy resolution function {\tt \#standard} is defined by
 \begin{equation}
 \label{eq:sigma_e}
-\sigma(E)=\alpha\cdot E + \beta \cdot \sqrt{E} +\gamma\,.
-\end{equation}
-The parameters $\alpha, \beta$ and $\gamma$ are provided by the user.
-\begin{quote}
-{\tt \tb @sigma\_e = \{ 0.15, 0.0, 0.0\}}
-\end{quote}
-The other possible choice for {\tt @sigma\_function} is {\tt \#inverse\_beta}
-which has only one parameter $\alpha$ and is defined by
+\sigma(E)=\alpha\cdot E + \beta \cdot \sqrt{E} +\gamma\, ,
+\end{equation}
+where the parameters $\alpha, \beta$ and $\gamma$ are provided by the user:
+\begin{quote}
+{\tt \tb @sigma\_e = \{0.15, 0.0, 0.0\}}
+\end{quote}
+Currently, another possible choice for {\tt @sigma\_function} is {\tt \#inverse\_beta},
+which only uses the parameter $\alpha$. It is defined by
 \begin{equation}
@@
 \end{equation}
-
-In the actual implementation 
-the sum in equation one only extends over those $E_j$ where $K(E_j)$ 
-is above a certain threshold, by default $10^{-7}$. This threshold is 
-defined at compile time. Thus a complete energy resolution definition of this 
-type looks like
+In the actual implementation of the algorithm,  the sum in \eq~(\ref{eq:algo_one}) is only computed for the $E_j$'s where $K(E_j)$ 
+is above a certain threshold, which is by default $10^{-7}$. 
+This threshold is defined at the compiling time.
+
+Eventually,  a complete energy resolution definition of this 
+algorithm is, for example,
 \begin{quote}
@@
 \tb @sigma\_function = \#standard\\
-\tb @sigma\_e = \{ 0.15 ,0.0 ,0.0\}\\
+\tb @sigma\_e = \{0.15 ,0.0 ,0.0\}\\
 >
@@
 \subsection{Low-pass filter based automatic energy smearing}
+
+THIS SUBSECTION WILL BE RE-WRITTEN BY PATRICK (WW)
 
@@
 
-Finally the user has the possibility to directly specify the smearing data.
-The format requires to provide $k_l^i$ and $k_u^i$ for each bin, the number of
-bins, the number of sampling points, the lower and upper limit of the sampling
-range. Furthermore the numbers $K_i(E_j)^c$ have to provide for all values
-in between $k_l^i$ and $k_u^i$. The convolution is then performed according to
-equation~\ref{eq:algo_one}. Thus a complete example for this case would be
-\begin{quote}
-{\tt
-\$simbins = 20\\
-\$simtresh =          4.0\\
-\$simbeam =         50.0\\
-}
-
+In some cases, one may use the output of a detector Monte Carlo simulation
+directly. Then one can use manual energy smearing instead of one of the
+automatic energy smearing algorithms.
+
+The energy smearing matrix $K_{ij}$ has {\tt \$bins} rows and {\tt \$sampling\_points} columns, which are numbered from $0$ to {\tt \$bins}$-1$ or {\tt \$sampling\_points}$-1$.
+It is equivalent to the
+bin- and sampling-point-based kernel in \eq~(\ref{eq:kernel}):
+\begin{equation}
+K_{ij} = K_i^c(E) |_{E=E_j},
+\label{equ:ematrix}
+\end{equation}
+where $E_j$ is the energy of the $j$th sampling point.
+In general, many of the entries in this matrix are zero, which means that it is convenient to evaluate the integrand in \eq~(\ref{eq:simple_int}) only at positions where
+$K_{ij}$ is non-zero. The corresponding ``sampling range'' range of non-zero matrix entries  in $K_{ij}$ for the $i$th energy bin is defined to run from
+column $k_l^i$ (``lower index'') to column $k_u^i$ (``upper index'').
+An example for a smearing matrix is
+\begin{equation}
+K_{ij} =   \underbrace{ \left( \begin{array}{cccccccccc} 
+a_{00} & a_{01} & a_{02} & a_{03} &  &  &  &  &  & \\
+a_{10} & a_{11} & a_{12} & a_{13} & a_{14} &  &  &  &  &  \\
+ & a_{21} & a_{22} & a_{23} & a_{24} & a_{25} &  &  &  &  \\
+ &  & a_{32} & a_{33} & a_{34} & a_{35} & a_{36} &  &  &  \\
+ &  &  & a_{43} & a_{44} & a_{45} & a_{46} & a_{47} &  &  \\
+& & & \uparrow  & & \ddots & & \uparrow & & \\
+& & & k_l^i & & & & k_u^i & & \\ 
+\end{array}\right) }_{\mathtt{\$sampling\_points}  \, \, \mathrm{columns}} \quad \leftarrow \quad \footnotesize{\mathtt{\$bins} \, \, \mathrm{rows}} \, ,
+\end{equation}
+where the unshown entries are zero. Thus, the values of $K_{ij}$ have to be specified between $k_l^i$ and $k_u^i$ in the form $\{ k_l^i,k_u^i, K_{i \, k_l^i}, K_{i k_l^i+1} , \hdots , K_{i k_u^i} \}$:
+\begin{quote}
 {\tt 
@@
 \end{quote}
+Note that the sum of all entries in each {\em column} should be equal 
+to unity,
+since all of the incoming neutrinos should be assigned to energy bins. In many practical cases, however, the definition of the energy smearing can
+lead to sums smaller than unity, such as in the case of truncated Gaussian
+distributions. The sum of entries in each {\em row} is not defined, since the events might be unevently distributed into the energy bins
+according to the energy resolution function.
 

### 2004-05-19 Documentation/part2.tex
*** empty log message ***

@@
 The set of rules\index{Rule} for an experiment is the final 
-link between the event rate
-computation and the statistical analysis. The information given by the rules
-specifies how the $\chi^2$ is computed based on the raw event rates as 
-described by the channels including possible systematical errors. 
-Therefore a rule has two parts: the first part describes how signal and 
-background events are composed out of the channels and the second part
-specifies which systematical errors are considered and what the size
-of the relevant errors is.
+link between the event rate computation and the statistical analysis. 
+The information in the rules
+specifies how the $\chi^2$ is computed based upon the raw event rates 
+given by the channels and possible systematical errors. 
+Therefore a rule has two parts: The first part describes how signal and 
+background events are composed out of the channels, and the second part
+specifies which systematical errors are considered, as well as their values.
+%
+For a rule, the splitting
+in signal and background is useful for simplicity and the treatment of systematics, as we will se later. Each rule will lead to a $\chi^2$-value,
+which means that all $\chi^2$'s of the different rules will be added
+for the whole experiment. Within each rule, the event rates are added, and
+the systematics is considered to be independent of the other rules.
+Thus, it is convenient to combine the above defined channels for different
+oscillation patterns and interaction types into one logical construction,
+which is the rule. For example, a superbeam usually has two rules: One for
+the $\nu_e$-appearance rates, and one for the $\nu_\mu$-disappearance rates.
+ In each case, contributions of several interaction types, as well as from
+ the $\nu_e$-contamination of the beam will lead to a number of contributing signal and background event channels.
+
+% WEITER:
 
@@
 The signal $s_i$ in one bin for a rule now can be composed out of 
-several channels by
+several channels by (OFTEN ONLY ONE) 
 \begin{equation}

### 2004-05-19 Documentation/part2.tex
*** empty log message ***

@@
 {\tt \$target\_mass = 50}\\
-{\tt \$numofbins = 20}\\
+{\tt \$bins = 20}\\
 {\tt \$emin = 4.0}\\
@@
 
-% WEITER:
-
-Based on the definition of channels it is possible to assemble the 
-signal and background event numbers for each bin.  
-The signal $s_i$ in one bin for a rule now can be composed out of 
-several channels by (OFTEN ONLY ONE) 
-\begin{equation}
-s_i=\epsilon_{c_1}\cdot n_i^{c_1}\,+\,\epsilon_{c_2}\cdot n_i^{c_2}\,+\,\ldots
-\end{equation}
-where the $\epsilon$'s are weight factors determined by the properties
-of the detector.
-Similarly the background in one bin $i$ can be specified by
-\begin{equation}
-b_i=\epsilon_{c_1}\cdot n_i^{c_1}\,+\,\epsilon_{c_2}\cdot n_i^{c_2}\,+\,\ldots
-\end{equation}
-This basic building blocks of a rule are specified by
+For each rule, the signal event rate $s_i$ in the $i$th bin can be composed 
+out of one or more channels by 
+\begin{equation}
+s_i=\alpha_{c_{s1}}\cdot n_i^{c_{s1}}\,+\,\alpha_{c_{s2}}\cdot n_i^{c_{s2}}\,+\,\ldots
+\end{equation}
+where the $\alpha$'s are overall normalization factors/efficiencies
+ determined by the properties of the detector. Note that bin-based (energy-dependent) efficiencies can be defined with the post-smearing efficiencies in the last section. 
+In addition, note that in most cases, it makes sense to have only one 
+signal channel and to assign all sorts of perturbations to the background. 
+%
+Similarly, the background event rate $b_i$ in the $i$th bin can be composed
+out of one or more channels
+\begin{equation}
+b_i=\beta_{c_{b1}}\cdot n_i^{c_{b1}}\,+\,\beta_{c_{b2}}\cdot n_i^{c_{b2}}\,+\,\ldots \, ,
+\end{equation}
+where the channels can be any combination of the ones in the signal rate and 
+additional ones. The background normalization factors very often have
+a specific meaning. For example, they may correspond to a fraction
+of mis-identified events (charge or flavor mis-identification).
+%
+These basic building blocks of each rule are, within the rule environment,
+for example defined by
 \begin{quote}
@@
 
-Quite often it is not sufficient to have constant efficiencies and therefore
-it is possible to define efficiencies for each bin and channel $\alpha_i^c$
-and thus the event rates in one bin become
-\begin{equation}
-s_i=\epsilon_{c_1}\alpha_i^{c_1}\cdot n_i^{c_1}\,+
-\,\epsilon_{c_2}\alpha_i^{c_2}\cdot n_i^{c_2}\,+\,\ldots\,,
-\end{equation}
-and analogous for $b_i$.
-
-The set of all $s_i$'s and $b_i$'s now can be composed in a form that takes
-some types of systematical error into account. The two most important and
-most easily parameterized systematics are a normalization and an energy
-calibration error. These error are considered separately for the signal events
-and the background events. The implementation of the normalization error
-is straight forward
+For the analysis of the systematical errors, the so called 
+ ``pull method'' is used~\cite{Fogli:2002pt}\footnote{In fact the pull 
+method was employed already in \Ref~\cite{Huber:2002mx} before 
+\Ref~\cite{Fogli:2002pt} appeared.}. 
+For the pull method, $k$ systematical errors are included by introducing 
+$k$ additional variables $\zeta_k$, which are the 
+so-called ``nuisance parameters''. 
+The nuisance parameters describe the dependence 
+of the event rates on the various sources of systematical errors, 
+such as an error on the total 
+normalization is included by multiplying the expected number of events in 
+each bin by a factor $(1+\zeta_1)$. The variation of $\zeta_1$ is in the fit 
+constrained by adding a penalty $p_1$ to the $\chi^2$-function. In case 
+of a  Gau\ss ian distributed systematical error, this penalty is 
+given by
+\begin{equation}
+\label{eq:penalty}
+p_i=\frac{(\zeta_i-\zeta_i^0)^2}{\sigma_{\zeta_i}^2}\,,
+\end{equation}
+where $\zeta_i^0$ denotes the mean and $\sigma_{\zeta_i}$ the standard 
+deviation of the corresponding nuisance parameter. We further on also refer
+to the mean as the ``central value'', and to the standard deviation as
+the ``error''. The latter corresponds to the actual systematical uncertainty. 
+The resulting $\chi^2$ is then minimized with respect to all nuisance 
+parameters $\zeta_i$, which leads to $\chi^2_\mathrm{pull}$
+\begin{equation}
+\chi^2_\mathrm{pull}(\boldsymbol{\lambda}):=\min_{\{\zeta_i\} } \,\, \left( 
+\chi^2(\boldsymbol{\lambda},
+\zeta_1, \ldots, \zeta_k)+ \sum_{j=1}^{k} p_j(\zeta_j)\right)\,.
+\end{equation}
+Here $\boldsymbol{\lambda}$ refers to the oscillation parameters 
+including the matter density
+$\rho$. One advantage of the pull method is that whenever the number $N$ of 
+data points is much larger than $k$, it is numerically easier to compute 
+$\chi^2_\mathrm{pull}$ than to invert the $N\times N$ covariance matrix. For
+the experiments considered here, $N$ is typically $20$ and $k\sim 4$, 
+which means that the pull method is numerically much faster. Moreover,
+ it is more flexible and  allows the inclusion of systematical errors 
+also for a  Poissonian $\chi^2$-function.
+In~\cite{Fogli:2002pt}, it has been demonstrated that the pull method 
+and the covariance based approach are equivalent for a Gau\ss ian and 
+linear model. In general,
+there is a separate $(\chi^2_\mathrm{pull})^r$ for each rule $r$, 
+\ie , pair of signal and background spectra, with a separate set of 
+nuisance parameters $\zeta_i^\alpha$. Thus, $\chi^2_\mathrm{pull}$
+is the sum of all individual  $(\chi^2_\mathrm{pull})^r$'s.
+By the minimization, the dependence on the $k$ nuisance parameters has been eliminated  from $\chi^2_\mathrm{pull}$.
+
+Now we can introduce the different systematical errors. 
+The two most important and
+most easily parameterized systematical errors are the normalization 
+and an energy calibration errors. These errors are assumed to be independent between the signal events and the background events, which means that
+this systematics treatment defines the grouping into signal or background.
+The implementation of the normalization error
+is straightforward:
 \begin{equation}
@@
 \end{equation} 
-with an analogous definition for the background events.
-
-For the parameterization of an energy calibration error two possibilities
-are implemented. The first one being somewhat simpler, whereas the second one
-is more accurate but requires a careful choice of parameters. The first 
-option is
-\begin{equation}
-s_i(a,b):=s_i(a)+b\cdot s_i\, E_i/(E_\mathrm{max}-E_\mathrm{min})
-\end{equation}
-and it is refered to as tilt since it describes a linear distortion, a tilt
-of the event rate spectrum. The second option is closer to an actual energy
-calibration error, which basically amounts to replacing the reconstructed 
-energy $E'$ by $(1+b)\cdot E'$ and we use following approximation
+with an analogous definition for the background events. Here, $a$ is the``nuisance'' parameter, which will be minimized over later.
+
+For the parameterization of an energy calibration error, two possibilities
+are implemented. The first one (method ``T'') is somewhat simpler, 
+whereas the second one (method ``C'')
+is more accurate, but it requires a careful choice of parameters. 
+The first option (method ``T'') is
+\begin{equation}
+s_i(a,b) \equiv s_i(a)+b\cdot s_i\, E'_i/(E'_\mathrm{max}-E'_\mathrm{min}),
+\end{equation}
+where $E'_{\mathrm{min}}$ and $E'_{\mathrm{max}}$ correspond to {\tt \$emin}
+and {\tt \$emax}, and $E'_i$ is the mean (reconstructed) 
+energy of the $i$th bin. It is often refered to as a ``tilt'' of the
+spectrum, since it describes a linear distortion 
+of the event rate spectrum. Note that this definition of a tilt makes
+only sense in combination with a large enough normalization error, since
+the tilt also affects the normalization.
+%
+The second option (method ``C'') is closer to an actual energy
+calibration error, which means that one should test this option whenever
+one suspects a large impact of this systematical error.
+It is based upon replacing the events in the $i$th bin by the ones at
+the energy $(1+b)\cdot E'_i$. If this target energy does not exactly hit
+a (discrete) bin energy $E_k$, linear interpolation is used. We use the following approximation:
 \begin{eqnarray}
-s_i(b)&=& (1+b)\cdot(s_{k+1}-s_k)\cdot\delta+s_i\,,\\
+s_i(a,b)&=& (1+b)\cdot \left[ \left( s_{k+1}(a)-s_k(a) \right)\cdot (\delta-k) +s_k(a) \right] \,,\\
 \delta&=&b\cdot(i+t_0+ 1/2)+i\,,\nonumber\\
-k&=&\mod(\delta,1)\,,\nonumber\\
-t_0&=&E_\mathrm{min}/\Delta E_0\,.\nonumber
+k&=& \mathrm{div}(\delta,1)\,,\nonumber\\
+t_0&=&E'_\mathrm{min}/\Delta E_0\,.\nonumber
 \end{eqnarray}
-Special care is required at the boundaries when $k<1$ or $k+1>N_\mathrm{bins}$,
-since there $s_k$ or $s_{k+1}$ may not have been calculated. One possibility
-to deal with those cases is to assume that for those values of $k$ $s_k$ is
-zero. This is the default. In cases where $s$ at the boundary is however still
-sizeable this may lead to intolerable errors and subsequently to a wrong
-estimate of the impact of a calibration error. Therefore in those cases it
-is advisable to truncate the analysis range by a few bins at the boundary
+Here $\Delta E_0$ is the bin width ({\tt \$emax}-{\tt \$emin})/{\tt \$bins}
+and ``div'' refers to the integer part of the division. Note that the
+factor $(1+b)$ in the first equation comes from a renormalization of
+the bin width, since also the bin width is altered by the replacement of
+the energies. Furthermore, special care has to be payed to the limits
+ $k<1$ or $k+1>N_\mathrm{bins}$, since there $s_k$ or $s_{k+1}$ may not have been calculated. By default, it is assumed that $s_k$ is
+zero in those cases. However, if the event rates are still large at
+the limits, errors will be introduced leading to a wrong
+estimate of the impact of the calibration error. In this case,
+one should truncate the analysis range by a few bins at the boundaries
 and therefore ensure in this way that only those $s_i$ are used whose index
-$k$ is within the range $1,\ldots, N_\mathrm{bins}-1$.
-
-In order to perform this fine tuning, it is possible to constrain 
-the energy range from which the data are taken for the 
-analysis\index{Energy window}. This is achieved by setting the energy window
-variable. 
+$k$ is within the range $0,\ldots, N_\mathrm{bins}-1$ (\cf, \figu{energies}). 
+Thus, it is possible to constrain the analysis energy range 
+with each rule to an energy window\index{Energy window}:
 \begin{quote}
@@
 \end{quote}
-The purpose is mainly to allow a fine tuning of the description
-of an energy calibration error. The default setting is that the energy window
-coincides with the range defined by the minimal and maximal reconstructed 
-energy. For the error dimension using the calibration method (C) it is 
-advisable to reduce the energy window by two or three bins at both the lower
-and upper end. As a rule of
-thumb consider the three standard deviations error on the calibration and
-evaluate the corresponding shift in $E'$ at the lower and upper energy limit 
-and reduce the analysis range by this amount.
-
-Thus the total event rate $x_i$ in a bin $i$ is given by
-\begin{equation}
-x_i(a,b,c,d)=s_i(a,b)+b_i(c,d)
-\end{equation}
-and thus a function of four parameters. The central values for all of
-the four parameters are always needed. They are called signal normalization
-($a$), signal tilt/calibration ($b$), background  normalization ($c$) and
+The default energy window is given between the minimal and maximal reconstructed energy. To be on the save side, reduce analysis window
+compared to the bin range on each side by about three standard deviations 
+of the calibration error.
+
+Eventually, the total event rate $x_i$ in a bin $i$ is given by
+\begin{equation}
+x_i(a,b,c,d)=s_i(a,b)+b_i(c,d) \, ,
+\end{equation}
+and is thus a function of four parameters. 
+The four parameters $a,b,c,d$ have been introduced in order to describe
+systematical uncertainties and are the nuisance parameters.
+Each of the four parameters has a central value and systematical error.
+The central values for all of
+the four parameters have to be always defined. They are called 
+signal normalization ($a$), signal tilt/calibration ($b$), 
+background  normalization ($c$) and
 background tilt/calibration ($d$). The default values are
 \begin{equation}
-a=1\,,\quad b=0\,,\quad c={\tt NaN}\,,\quad d=0\,,
-\end{equation}
-thus for the background normalization $c$ a value has to be specified in 
-\emph{all} cases. The central values for the normalization and the 
-corresponding central value of tilt/calibration are regarded
-as a pair. The errors are treated in the same way.
+a=1\,,\quad b=0\,,\quad c={\tt not \, \, assigned}\,,\quad d=0\,.
+\end{equation}
+Thus, for the background normalization $c$, the value has to be specified 
+in \emph{either} case. The values for the normalization and the 
+values of tilt/calibration are always regarded
+as a pairs, \ie, they are given in the form {\tt normalization : tilt}. The errors are treated in the same way. For example, we have
 \begin{quote}
@@
 \end{quote}
-There is no {\tt @signalcenter} since by default the central value for the
+There is no {\tt @signalcenter} in this definition, 
+since by default the central value for the
 signal normalization is $1$ and the central value for the tilt/calibration 
@@
 
-The four parameters $a,b,c,d$ have been introduced in order to describe
-systematical uncertainties, they are so called nuisance parameters.
-For the analysis of the systematical errors the so called 
- pull method is 
-used~\cite{Fogli:2002pt}\footnote{In fact the pull method
-was employed already in~\cite{Huber:2002mx} before 
-\cite{Fogli:2002pt} appeared.}. 
-In the  pull method $k$ systematical errors are included by introducing 
-$k$ additional variables $\zeta_k$, which will  be called 
-nuisance parameters in the following. 
-The nuisance parameters describe the dependence of the event rates on the 
-various sources of systematical errors, \eg\ an error on the total 
-normalization is included by multiplying the expected number of events in 
-each bin by a factor $(1+\zeta_1)$. The variation of $\zeta_1$ in the fit 
-is constrained by adding a penalty $p_1$ to the $\chi^2$-function. In case 
-of a  Gau\ss ian distributed systematical error this penalty is 
-given by
-\begin{equation}
-\label{eq:penalty}
-p_i=\frac{(\zeta_i-\zeta_i^0)^2}{\sigma_{\zeta_i}^2}\,,
-\end{equation}
-where $\zeta_i^0$ denotes the mean and $\sigma_{\zeta_i}$ the standard 
-deviation of the corresponding nuisance parameter, \ie\ the amount of 
-systematical uncertainty. 
-The resulting $\chi^2$ is then minimized with respect to all nuisance 
-parameters $\zeta_i$ and this yields $\chi^2_\mathrm{pull}$
-\begin{equation}
-\chi^2_\mathrm{pull}(\boldsymbol{\lambda}):=\min_{\{\zeta_i\} } \,\, \left( 
-\chi^2(\boldsymbol{\lambda},
-\zeta_1, \ldots, \zeta_k)+ \sum_{j=1}^{k} p_j(\zeta_j)\right)\,,
-\end{equation}
-where $\boldsymbol{\lambda}$ denotes the oscillation parameters 
-including the matter density
-$\rho$. One advantage of the pull method is that whenever the number $N$ of 
-data points is much larger than $k$, it is numerically easier to compute 
-$\chi^2_\mathrm{pull}$ than to invert the $N\times N$ covariance matrix. For
-the experiments considered here $N$ is typically $20$ and $k\sim 4$, thus
-the pull method is numerically much faster. Moreover it is more flexible and 
-allows the inclusion of systematical errors also for a 
-Poissonian $\chi^2$-function.
-In~\cite{Fogli:2002pt} it was shown that the pull method and the covariance
-based approach are equivalent for a Gau\ss ian and linear model. In general
-there is a separate $(\chi^2_\mathrm{pull})^\alpha$ for each rule $\alpha$, 
-\ie\ pair of signal and background spectra, with a separate set of 
-nuisance parameters $\zeta_i^\alpha$. In this case $\chi^2_\mathrm{pull}$
-is the sum of all individual  $(\chi^2_\mathrm{pull})^\alpha$.
-In this way the dependence on the $k$ nuisance parameters has been eliminated 
-from $\chi^2_\mathrm{pull}$.
-
 The user has the possibility to choose the set $\{\zeta_i\}$ of nuisance 
-parameters which are marginalized. This achieved with the error dimension 
-variable\index{Error dimension}, and the different possibilities are shown in
-table~\ref{tab:error_dim}. If a parameter is designated with $+$, it will
-be marginalized and therefore the corresponding error needs to have a non-zero
-value. In the cases labeled (number 4 and 8) ``total rates'' 
-the summation over the bins is performed \emph{before} computing 
-the $\chi^2$. The case labeled (number 7) ``spectrum only'' leaves the 
-normalization free ($\sigma_a=\sigma_c=\infty$) and therefore counts only the 
-spectral information in the data. As a consequence
-the settings for the normalization error will be ignored. It is furthermore
-possible to extend the set of error dimensions and thus the set of possible 
-systematical errors to be studied, but this requires proper C programing, 
-compiling and linking with \GLOBES. The interface for user defined error 
-dimensions is described in detail in section~\ref{sec:ud_error_dim}.
+parameters which are minimized over. This choice is specified with the 
+error dimension variable\index{Error dimension}, and the different possibilities are shown in
+\Tab~\ref{tab:error_dim}. The error dimensions can be chosen for each rule 
+by {\tt @errordim}. For example, 
+\begin{quote}
+{\tt 
+\tb @errordim = 1
+}
+\end{quote}
+In general, it is possible to extend the set of error dimensions or the 
+set of possible systematical errors, but this requires C programming, 
+compiling and linking with \GLOBES. 
+%The interface for user defined error 
+%dimensions is described in detail in section~\ref{sec:ud_error_dim}.
 
@@
 \begin{center}
-\begin{table}[hbt!]
+\begin{table}[t!]
 \begin{center}
-\begin{tabular}[h]{|c|cccc|c|c|}
+\begin{tabular}[h]{|c|cccc|c|l|}
 \hline
@@
 \hline
-0&+&+&+&+&T&\\
-2&-&-&-&-&-&\\
-4&+&+&+&+&T&total rates\\
-7&o&-&o&-&-&spectrum only\\
-8&-&-&-&-&-&total rates\\
-9&+&+&+&+&C&\\
+0&+&+&+&+&T&Systematics with tilt\\
+2&-&-&-&-&-&No systematics\\
+4&+&+&+&+&T&Total rates\\
+7& $\infty$ &-& $\infty$ &-&-&Spectrum only\\
+8&-&-&-&-&-&Total rates, No systematics\\
+9&+&+&+&+&C&Systematics with calibration\\
 \hline
@@
 \caption[Table of error dimensions]{\label{tab:error_dim}
-Values of the error dimension variable and their meaning.
+Possible values of the error dimensions variable in \GLOBES\ and their meaning. If a parameter is designated with $+$, it will
+be marginalized over, and therefore the corresponding error needs to 
+have a non-zero value. If the cases with ``total rates'' in the remarks, 
+the summation over the bins is performed \emph{before} computing 
+the $\chi^2$, \ie, no spectral information is used. 
+The error dimension~7 ``spectrum only'' leaves the 
+normalization free ($\sigma_a=\sigma_c=\infty$), and therefore 
+only the spectral information is used. As a consequence,
+the settings for the normalization error will be ignored (designated 
+with the symbol $\infty$). 
  }
@@
 
-The error dimensions can be chosen for each rule by
-\begin{quote}
-{\tt 
-\tb @errordim = 1
-}
-\end{quote}
-
-Next we can assemble the various pieces into one rule
+Eventually, a rule looks like
 \begin{quote}

### 2004-06-24 Documentation/part1.tex
*** empty log message ***

@@
 \section{The treatment of external input}
+\label{sec:externalinput}
 
@@
 
+\example{The impact of systematics, correlations, and degeneracies}{
+\label{ex:corrproj}
+\index{Bar charts}
+\index{Systematics on/off}
+
+Here it is shown how one can successively include systematics, correlations,
+and degeneracies at the example of the $\stheta$ sensitivity limit.
+An important part of this example is how two switch the systematics off,
+\ie, how to obtain the sensitivity limit from statistics only. Since
+this example is very advanced, we only show the respective part of the code:
+\begin{quote}
+{\tt 
+Here comes the code excerpt.
+}
+\end{quote}
+The complete code can be found as ``example4.c'' with the software.
+
+The returned lists of data then represent $\Delta \chi^2$ 
+as function of the fit value of $\stheta$. The intersections of these
+curves with the line $\Delta \chi^2 = 9$ give the $\stheta$ sensitivity
+limits, where we do not include the $\mathrm{sgn}(\ldm)$- and $(\deltacp,\theta_{13})$-degeneracies in the sensitivity limit with
+correlations only (green bar):
+\begin{center}
+\colorbox{white}{\includegraphics[width=7cm]{barsex}}
+
+\end{center}
+}
+
 \chapter{Changing experiment parameters at running time}
@@
 \item
+ Systematics
+\item
  Threshold function, efficiencies etc.

### 2004-06-24 Documentation/example4.c

@@ -1,0 +1,193 @@
+/* Example: Creating bar charts
+ * Copyright 2004 GLoBES collaboration
+ * Compile with ``make example4''
+ */ 
+
+#include <stdio.h>
+#include <stdlib.h>
+#include <math.h>
+#include <string.h>
+
+#include <globes/globes.h>   /* GLoBES library */
+#include "myio.h"             /* my input-output routines */
+
+/* If filename given, write to file; for empty filename write to screen */
+char MYFILE1[]="test4a.dat";
+char MYFILE2[]="test4b.dat";
+char MYFILE3[]="test4c.dat";
+char MYFILE4[]="test4d.dat";
+
+double theta12,theta13,theta23,deltacp,sdm,ldm;
+
+double Sign(double n)
+{
+ if(n>=0) return +1.0;
+ else return -1.0;   
+}
+
+void SetOscParams(double theldm)
+{
+ glb_params true_values = alloc_params();
+ true_values = 
+   GLBDefineParams(true_values,theta12,theta13,theta23,deltacp,sdm,theldm);
+ GLBSetOscillationParameters(true_values);
+ MSetRates();
+ free_params(true_values);  
+}
+
+double CalcSystematics(double theldm,double thex)
+{
+  SetOscParams(theldm);
+  double thetheta13=asin(sqrt(pow(10,thex)))/2;
+  glb_params test_values = alloc_params();
+  test_values=
+    GLBDefineParams(test_values,theta12,thetheta13,theta23,0.0,sdm,theldm);
+  double res=GLBChiSys(test_values,GLB_ALL,GLB_ALL);
+  free_params(test_values);  
+  return res;
+}
+
+double CalcNoSystematics(double theldm,double thex)
+{
+ int rules=GLBGetNumberOfRules(0);
+ int j;
+  /* Later with systematics off/on */
+ for(j=0;j<rules;j++) MSetErrorDim(2,0,j);
+ double res=CalcSystematics(theldm,thex);
+ for(j=0;j<rules;j++) MSetErrorDim(0,0,j);
+ return res; 
+}
+
+double CalcProjection(double theldm,double thex,glb_params start_vector)
+{
+  SetOscParams(theldm);
+
+  double thetheta13=asin(sqrt(pow(10,thex)))/2;
+  double thesign=Sign(get_osc_params(start_vector,5));
+
[…]

### 2004-06-28 Documentation/Manual.tex
*** empty log message ***
two files added for the GPL and FDL
latest version of part2

@@
 
+\usepackage{verbatim}
 \usepackage{cite}
+\usepackage[dvips]{color}
 \usepackage{graphicx}
@@
 \usepackage[footnotesize,bf]{caption}
+\usepackage{makeidx}
 \usepackage{ae,aecompl} % only for Type 1 fonts
+
+\newcommand{\capdef}{}
+\newcommand{\mycaption}[2][\capdef]{\renewcommand{\capdef}{#2}%
+        \caption[#1]{{ %\itshape
+ {\footnotesize #2}}}}
+\makeatletter
+\renewcommand{\fnum@table}{\textbf{\tablename~\thetable}}
+\renewcommand{\fnum@figure}{\textbf{\figurename~\thefigure}}
+\makeatother
+
+% Index: use "makeindex Manual" to create .ind file
+\makeindex
 
@@
 \newcommand{\nn}{\nonumber}
-\newcommand{\ldm}{$\Delta m_{31}^2$}
-\newcommand{\sdm}{$\Delta m_{21}^2$}
-\newcommand{\deltacp}{$\delta_{\mathrm{CP}}$}
-\newcommand{\stheta}{$\sin^2 2 \theta_{13}$}
+\newcommand{\ldm}{\Delta m_{31}^2}
+\newcommand{\sdm}{\Delta m_{21}^2}
+\newcommand{\deltacp}{\delta_{\mathrm{CP}}}
+\newcommand{\stheta}{\sin^2 2 \theta_{13}}
 
@@
 \newcommand{\Secs}{Secs.}
+\newcommand{\Chapt}{Chapter}
+\newcommand{\Chapts}{Chapters}
+\newcommand{\Part}{Part}
 \newcommand{\App}{the Appendix}
@@
 \newcommand{\NuFactII}{{\sc NuFact-II}}
+\newcommand{\SPL}{{\sc SPL}}
+\newcommand{\Beta}{{\sc $\beta$-Beam}}
 
 \newcommand{\GLOBES}{{\sf GLoBES}}
+\newcommand{\AEDL}{{\sf AEDL}}
 \newcommand{\EDM}{{\sf EDM}}
@@
 \newcommand{\figu}[1]{\fig~\ref{fig:#1}}
+\newcommand{\tabl}[1]{\Tab~\ref{tab:#1}}
+\newcommand{\tb}{\hspace*{3ex}}
+
+\definecolor{light}{gray}{0.93}
+\definecolor{heavy}{gray}{0.35}
+\setlength{\fboxrule}{0.1cm}
+\newcommand{\example}[2]{
+\begin{table}[p]
+\begin{center}
+\fcolorbox{heavy}{light}{
+\begin{minipage}[t][\textheight][t]{15.5cm}
+{\em \underline{Example:} #1}
+
+\vspace*{0.3cm}
+
+#2
+
+\end{minipage}
+}
+\end{center}
+\end{table}
+}
+
+\newtheorem{function}{Function}[chapter]
 
@@
 \GLOBES\ 1.X \\
-Global Long Baseline Experiment Simulator \\ }
+General Long Baseline Experiment Simulator \\ }
 \vspace*{0.5cm}
@@
 \begin{center}
-Here comes our logo!
+\colorbox{black}{\includegraphics[width=7cm]{earthint}}
+
+\vspace*{-1cm}
+
+{\Huge \textcolor{yellow}{GLoBES}}
 \end{center}
@@
 
-\chapter*{What is \GLOBES\ good for?}
+\chapter*{What is \GLOBES ?}
 
-Here comes a maximum of one page what \GLOBES\ is good for and what main features it has.
+\GLOBES\ (``General Long Baseline Experiment Simulator'') is a flexible
+software package to simulate neutrino long baseline experiments. On the
+one hand, it contains a comprehensive abstract experiment definition
+language, which allows to describe most classes of long baseline experinents
+at an abstract level. On the other hand, it provides a C-library to 
+process the experiment information in order to obtain oscillation
+probabilities, rate vectors, and $\Delta \chi^2$-values. Currently, 
+\GLOBES\ is available for Linux. Since the source code is included,
+the modifications to other operating systems should be doable.
+
+\GLOBES\ allows to simulate experiments with stationary neutrino point sources, where each experiment is assumed to have only one neutrino source.
+Such experiments are neutrino beam experiments and reactor experiments. 
+Geometrical effects of a source distribution, such as in the sun or the 
+atmosphere, can not be described. In addition, sources with a physically 
+significant time dependencies  can not be studied, such as  supernov\ae. It 
+is, however, possible to simulate beams with bunch structure, since the 
+time dependence of the neutrino source is physically only important to suppress backgrounds.
+
+On the experiment definition side, either built-in neutrino fluxes
+(\eg, neutrino factory) or arbitrary fluxes can be used. Similarly,
+arbitrary cross sections, energy dependent efficiencies, the
+energy resolution function, the considered oscillation channels, and
+many other features can be specified. For the systematics, energy
+normalization and calibration errors can be simulated. Note that
+the energy ranges and windows, as well as the bin widths can be
+(almost) arbitrarily chosen, which means that variable bin
+widths are allowed. Together with \GLOBES\ comes a number of
+pre-defined experiments in order to demonstrate the capabilities
+of \GLOBES\ and to provide prototypes for new experiments.
+
+With the C-library, one can extract the $\Delta \chi^2$ for all defined 
+oscillation channels for an experiment or any combination of experiments.
+Of course, also low-level information, such as oscillation
+probabilities or event rates, can be obtained. \GLOBES\ includes the
+simulation of neutrino oscillations in matter with arbitrary matter 
+density profiles, as well as it allows to simulate the matter density
+uncertainty. As one of the most
+advanced features of \GLOBES , it provides the technology to 
+project the $\Delta \chi^2$, which is a function of all oscillation
+parameters, onto any subspace of parameters by local minimization. 
+This approach allows the inclusion of multi-parameter-correlations,
+where external input (\eg, from solar parameters) can be imposed, too.
+Applications of the projection mechanism include the projections onto the $\stheta$-axis and the $\stheta$-$\deltacp$-plane.
+In addition, all oscillation parameters can be kept free to precicely localize
+degenerate solutions.
+
 \cleardoublepage
@@
 \addcontentsline{toc}{chapter}{How to use this manual}
+
+As it is illustrated in \figu{GLOBES}, \GLOBES\ consists 
+of several modules.
+%
+\begin{figure}[bht]
+\begin{center}
+\includegraphics[width=16cm]{GLOBES}
+\end{center}
+\caption{\label{fig:GLOBES} Different modules in \GLOBES .}
+\end{figure}
+%
+\AEDL (``Abstract Experiment Definition Language'') is a language
+to define experiments in form of normal text files. One or more of 
+the resulting \AEDL\ files can then be processed together with supporting 
+flux or cross section files by the user interface. The user interface
+is a C-library, which loads one or more \AEDL\ file(s)
+containing the experiment definition(s). The user interface is linked 
+against the application software, which calls the user interface functions
+for the intended experiment simulation.
+
+The application 
+software is, except from some example files, not part of \GLOBES , since
+the evaluation of the experiment performance is often a matter of taste
+and definition. In addition, the algorithms depend, especially for
+high-precision instruments, very much on the oscillation parameters.
+In general, it is quite simple to simulate superbeams and reactor
+experiments. However, because of the more complicated topology, the
+simulation of neutrino factories is much more difficult. In order
+to demonstrate some of these difficulties, we show in this manual only
+examples with neutrino factories. These examples can be found in
+\Part~\ref{part:1} within the boxed pages. As complete files, they
+are also available in the \GLOBES\ software package.
+
+The \GLOBES\ software may have two target groups: 
+Physicists, who are mainly interested in optimizing the potential
+of specific experimental setups, and others, who are mainly
+interested in the physics potential of different experiment types
+from a theoretical point of view.
+For the first group, \AEDL\ could be the most interesting aspect of
+\GLOBES , where the user interface is only a tool to obtain specific
+parameter sensitivities. In this case, \GLOBES\ could, serve as a
+unified tool for the comparison and optimization of different experiment
+ setups on equal footing, where
+it is the primary objective to simulate the experiments as accurate
+as possible. In addition, changes in experimental parameters, such as
+efficiences or the energy resolutions, can quickly be tested.
+%
+For the second user group, the pre-defined 
+experiment definition files might already be sufficient to test
+new conceptual approaches, and the user interface is the most interesting
+aspect for sophisticated applications including correlations,
+degeneracies, and multi-experiment setups. In either case, the \GLOBES\
+software could serve as a platform for the exchange of experiment
+definitions, and for an efficient splitting of work between
+experimentalists and theorists.
+
+The user interface functions are described in \Part~\ref{part:1} of 
+this manual, which is the ``user's manual''. In there, first of all a 
+short \GLOBES\ tour is given in \Chapt~\ref{chapter:tour} in order to 
+have an overview over \GLOBES . 
+After that, the user
+interface is successively introduced from very basic to more sophisticated
+function. Eventually, it is demonstrated how one can change many
+experiment parameters at running time (such as baseline or target mass), and how one can ontain low-level
+information. We recommend that everybody interested in \GLOBES\ should
+become familiar at least with the concepts in \Chapt~\ref{chapter:tour}
+ and some of the examples on the boxed pages. The examples can be 
+ directly compiled 
+ from the respective directory in the \GLOBES\ software package.
+
+In \Part~\ref{part:2} of the manual, \AEDL\ is described. After an
+introductory chapter, all functions are defined in greater detail.
+This part might be more interesting for the experimental users who
+want to modify or create \AEDL\ files. A useful tool in this context
+is the software program ``globes'', which returns event rates and other
+information for individual \AEDL\ files without further programming. 
+For example, flux normalizations can with this tool be easily adjusted 
+to reproduce the event rates of a specific experiment.
+
+Finally, \Part~\ref{part:3} of this manual serves as reference manual,
+and the appendix contains further supplementary information.
 
@@
 
+% Bibliography:
+
+\bibliographystyle{./unsrt}
+\bibliography{references}
+
+% Index:
+
+\printindex
+
 \end{document}

### 2004-06-29 Documentation/Manual.tex
*** empty log message ***

@@
 \newcommand{\Part}{Part}
-\newcommand{\App}{the Appendix}
+\newcommand{\App}{Appendix}
 \newcommand{\Apps}{Appendices}
@@
 \newcommand{\tb}{\hspace*{3ex}}
+
+% Set as GLoBES command and add to index:
+% Always use GLB when mentioning a command in text!
+\newcommand{\GLB}[1]{{\tt #1} \index{{\tt #1}}}
 
@@
 arbitrary cross sections, energy dependent efficiencies, the
-energy resolution function, the considered oscillation channels, and
-many other features can be specified. For the systematics, energy
+energy resolution function, the considered oscillation channels, 
+backgrounds, and many other features can be specified. 
+For the systematics, energy
 normalization and calibration errors can be simulated. Note that

### 2004-06-29 Documentation/part1.tex
*** empty log message ***

@@
 \part{User's manual}
+\label{part:1}
+
 \chapter{A short \GLOBES\ tour}
+\label{chapter:tour}
 
@@
 overview of main functions in \tabl{stdfunctions}.
-Maybe: in form of long example which involves everything (for example: our famous bar plot, which involves
-systematics, correlations, degeneracies and thus the full set of \GLOBES\ functions without being too
-complicated in the application software part).
+Maybe: in form of long example which involves everything
+and can be compiled.
 
@@
 \begin{center}
-\begin{tabular}{p{1.8cm}p{4.5cm}p{8.6cm}}
+\begin{tabular}{p{3.8cm}p{3.8cm}p{7cm}}
 \hline
@@
 \hline
-{\tt Chi} & $\chi^2$ with systematics only \newline (all initialized exps.) & ($\{ \theta_{12}, \theta_{13}, \theta_{23}, \deltacp , \sdm , \ldm, \hat{\rho_1},...
-, \hat{\rho_n} \}$)  \newline \ra\  $\chi^2$ \\[0.1cm]
-{\tt SingleChi} & $\chi^2$ with systematics only \newline (only one experiment) & $(\{ \theta_{12}, \theta_{13}, \theta_{23}, \deltacp , \sdm , \ldm, \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}} )$   \newline \ra\ $\chi^2$ \\[0.1cm]
-{\tt ChiTheta} & $\chi^2$ with systematics and correlations: Projection onto $\theta_{13}$-axis (all exps.) &  ($ \theta_{13}, \, \{ \theta_{12}, \theta_{23}, \deltacp , \sdm , \ldm, \hat{\rho_1}, ...
-, \hat{\rho_n} \}$) \newline \ra\  $\{ \chi^2, \theta_{12}, \theta_{23}, \deltacp , \sdm , \ldm, \hat{\rho_1}, ...
-, \hat{\rho_n} , N_{\mathrm{Iter}} \}$ \\[0.1cm]
-{\tt Single-} \newline {\tt ChiTheta} & $\chi^2$ with systematics and correlations: Projection onto $\theta_{13}$-axis (one exp.) &  ($ \theta_{13}, \, \{ \theta_{12}, \theta_{23}, \deltacp , \sdm , \ldm,  \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}}$) \newline \ra\  $\{ \chi^2, \theta_{12}, \theta_{23}, \deltacp , \sdm , \ldm, \hat{\rho}_{N_{\mathrm{exp}}} , N_{\mathrm{Iter}} \}$ \\[0.1cm]
-{\tt ChiDelta} & $\chi^2$ with systematics and correlations: Projection onto $\deltacp$-axis (all exps.) &  ($\deltacp, \, \{ \theta_{12}, \theta_{13}, \theta_{23},  \sdm , \ldm, \hat{\rho_1}, ...
-, \hat{\rho_n} \}$) \newline \ra\  $\{ \chi^2, \theta_{12}, \theta_{13}, \theta_{23}, \sdm , \ldm, \hat{\rho_1}, ...
-, \hat{\rho_n} , N_{\mathrm{Iter}} \}$ \\[0.1cm]
-{\tt Single-} \newline {\tt ChiDelta} & $\chi^2$ with systematics and correlations: Projection onto $\deltacp$-axis (one exp.) &  ($ \deltacp, \, \{ \theta_{12}, \theta_{13}, \theta_{23}, \sdm , \ldm,  \hat{\rho}_{N_{\mathrm{exp}}}  \}, \, N_{\mathrm{exp}}$) \newline \ra\  $\{ \chi^2, \theta_{12}, \theta_{13}, \theta_{23},  \sdm , \ldm, \hat{\rho}_{N_{\mathrm{exp}}} , N_{\mathrm{Iter}} \}$ \\[0.1cm]
-{\tt ChiTheta-} {\tt Delta} & $\chi^2$ with systematics and correlations: Projection onto $\deltacp$-$\theta_{13}$-plane (all exps.) &  ($\theta_{13}, \, \deltacp, \, \{ \theta_{12}, \theta_{23},  \sdm , \ldm,  \hat{\rho_1}, ...
-, \hat{\rho_n} \}$) \newline \ra\  $\{ \chi^2, \theta_{12}, \theta_{23}, \sdm , \ldm,  \hat{\rho_1}, ...
-, \hat{\rho_n} , N_{\mathrm{Iter}} \}$ \\[0.1cm]
-{\tt SingleChi-} {\tt Theta- } {\tt Delta} & $\chi^2$ with systematics and correlations: Projection onto $\deltacp$-$\theta_{13}$-plane (one exp.) &  ($ \theta_{13}, \, \deltacp, \, \{ \theta_{12}, \theta_{23}, \sdm , \ldm,  \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}}$) \newline \ra\  $\{ \chi^2, \theta_{12},  \theta_{23},  \sdm , \ldm, \hat{\rho}_{N_{\mathrm{exp}}} , N_{\mathrm{Iter}} \}$ \\[0.1cm]
-{\tt ChiNP} & $\chi^2$ with systematics and correlations: Projection onto $N$-parameter hyper-plane (one exp.) & ??? to be defined \\[0.1cm]
-{\tt Single-} {\tt ChiNP} & $\chi^2$ with systematics and correlations: Projection onto $N$-parameter hyper-plane (one exp.) & ??? to be defined \\[0.1cm]
-{\tt ChiAll} & Lokal minimum of $\chi^2$ with respect to all parameters (all exps.) &
-($ \{ \theta_{13}, \theta_{12}, \theta_{23}, \deltacp , \sdm , \ldm,  \hat{\rho_1}, ...
-, \hat{\rho_n} \}$) \newline \ra\  $\{ \chi^2, \theta_{13}, \theta_{12}, \theta_{23},$ \newline \hspace*{1.4cm} $ \deltacp , \sdm , \ldm,  \hat{\rho_1}, ...
-, \hat{\rho_n} , N_{\mathrm{Iter}} \}$
-\\[0.1cm]
-{\tt Single-} {\tt ChiAll} & Lokal minimum of $\chi^2$ with respect to all parameters (one exp.) &  ($ \{ \theta_{13}, \theta_{12}, \theta_{23}, \deltacp , \sdm , \ldm,  \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}}$) \newline \ra\  $\{ \chi^2, \theta_{13}, \theta_{12}, \theta_{23}, \deltacp , \sdm , \ldm, \hat{\rho}_{N_{\mathrm{exp}}} , N_{\mathrm{Iter}} \}$ \\[0.1cm]
+\multicolumn{3}{l}{{\bf Systematics only:}} \\
+\GLB{glbChiSys} & $\chi^2$ with systematics only  & ({\tt glb\_params in, int exp, int rule}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
+\multicolumn{3}{l}{{\bf Projections onto axes:}} \\
+\GLB{glbChiTheta} & Projection onto $\theta_{13}$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
+\GLB{glbChiDelta} & Projection onto $\deltacp$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
+\GLB{glbChiTheta23} & Projection onto $\theta_{23}$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
+\GLB{glbChiDm} & Projection onto $\ldm$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
+\GLB{glbChiDms} & Projection onto $\sdm$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
+\multicolumn{3}{l}{{\bf Projection onto plane:}} \\
+\GLB{glbChiThetaDelta} & Projection onto $\theta_{13}$-$\deltacp$-plane  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
+\multicolumn{3}{l}{{\bf Projection onto any hyper-plane:}} \\
+\GLB{glbChiNP} & Projection onto any $n$-dimensional hyper-plane  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \newline
+Needs \GLB{glbSetProjection} before! \\[0.2cm]
+\multicolumn{3}{l}{{\bf Localization of degeneracies:}} \\
+\GLB{glbChiAll} & (Local) Minimization over all parameters  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\
 \hline
@@
 \end{center}
-\caption{\label{tab:stdfunctions} \index{Standard functions (table)} The \GLOBES\ standard function to obtain a $\chi^2$-value for all or one of the initialized experiments.
-The curly brackets refer to the parameters to be transferred in form of a list.
-The parameter $\hat{\rho}_i \equiv \rho_i/\bar{\rho}_i$ refers to the matter density scaling factor, which is $1.0$ for leaving the matter density profile unchanged.
-Note that all functions but {\tt Chi} and {\tt SingleChi} are using minimizers which have to be initialized with {\tt SetInputErrors} and {\tt SetStartingValues} first.}
+\caption{\label{tab:stdfunctions} \index{Standard functions (table)} The \GLOBES\ standard function to obtain a $\chi^2$-value with systematics only or systematics and correlations.
+The parameters {\tt rule} and {\tt exp}
+can either be \GLB{GLB\_ALL} for all initialized experiment or the
+experiment number ($0$ to \GLB{glb\_num\_of\_exps}-1) for a specific experiment.
+The format of \GLB{glb\_params} is discussed in detail in \Chapt~\ref{chapt:gettingstarted}.
+Note that all functions but {\tt ChiSys}
+  are using minimizers which have to be initialized with \GLB{glbSetInputErrors} and \GLB{glbSetStartingValues} first.}
 \end{table}
@@
 \chapter{Getting started with \GLOBES }
+\label{chapt:gettingstarted}
 
-In this first chapter of the user's manual, we assume that the \GLOBES\ software is readily installed on your computer system.
-We demonstrate how to load pre-defined experiments and re-obtain information about them.
-However, we only secondarily discuss the usage of \GLOBES\ in your specific programming language, such as C, Mathematica, or others.
-Thus, you should be familiar of how to load \GLOBES\ on your computer system before reading this chapter.
-An example of how to use \GLOBES\ with C can be found on page~\pageref{ex:c}.
+In this first chapter of the user's manual, we assume that the \GLOBES\ software is readily installed on your computer system. For the installation,
+see \App~\ref{app:installation} and the {\tt INSTALL} file in the
+software package. We demonstrate how to load pre-defined experiments 
+and re-obtain important information about them. In addition, we discuss
+some standard concepts of the \GLOBES\ user interface. We do not go
+into details of the programming language, which means that standard parts
+of the program code common to all examples are, in general, omitted.
+An example of a minimal \GLOBES\ program in C can be found on page~\pageref{ex:c}.
+Furthermore, the files of the examples in this part can be found in the {\tt Example} subdirectory of your \GLOBES\ distribution.
+\index{Examples} Since it depends on your installed
+configuration, we refer to the {\tt README} file in the \GLOBES\ main directory, and the comments in the Makefile of the {\tt Example} subdirectory for how to complile the example files.
 
@@
 \index{C-Code}
+Here comes the C-code sceleton, which is
+(more or less) common to all of our \GLOBES\ examples:
+\begin{quote}
+{\tt {\footnotesize
+\#include <stdio.h> \\
+\#include <stdlib.h> \\
+\#include <math.h> \\
+\#include <string.h> \\
+\\
+\#include <globes/globes.h> \hspace*{0.5cm} /* Include GLoBES library */ \\
+\\
+\#include "myio.h" \hspace*{0.5cm} /* Include "housemade" I/O-routines */ \\
+\\
+/* If filename given, write to file; if empty, to screen: */ \\
+char MYFILE[]="testX.dat"; \\
+\\
+int main(int argc, char *argv[]) \\
+\{  \\
+\\
+ \hspace*{0.5cm} GLBInit(argv[0]); \hspace*{0.5cm} /* Initialize GLoBES library */ \\
+\\  
+  \hspace*{0.5cm} glbInitExperiment("NuFact.glb",\&glb\_experiment\_list[0], \\
+  \hspace*{1cm} \&glb\_num\_of\_exps); \hspace*{0.5cm} /* Initialize experiment NuFact.glb */\\
+\\  
+  \hspace*{0.5cm}  /* Initialize housemade output function */\\
+  \hspace*{0.5cm} 
+   InitOutput(MYFILE,"Format: ... ... ... $\backslash$n"); \\
+\\  
+  \hspace*{0.5cm} /* Initialize parameter vector(s) */ \\
+  \hspace*{0.5cm} glb\_params true\_values = glbAllocParams(); \\
+  \hspace*{0.5cm} /* ... */ \\
+\\  
+ \hspace*{0.5cm} /* Assign: theta12,theta13,theta23,deltacp,dm2solar,dm2atm */ \\
+   \hspace*{0.5cm}  
+     glbDefineParams(true\_values,\\
+     \hspace*{1.5cm}asin(sqrt(0.8))/2,asin(sqrt(0.001))/2,M\_PI/4,M\_PI/2,7e-5,2e-3); \\
 
-Here comes some complete C-code with a very simple example of how to use \GLOBES .
+     \hspace*{0.5cm}  
+   /* The simulated data are computed */ \\
+   \hspace*{0.5cm} glbSetOscillationParameters(true\_values); \\
+   \hspace*{0.5cm} glbSetRates(); \\
+
+  \hspace*{0.5cm} /* ... CODE ... */ \\
+
+\hspace*{0.5cm}  /* Destroy parameter vector(s) */ \\
+\hspace*{0.5cm}  glbFreeParams(true\_values); \\
+\hspace*{0.5cm} /* ... */ \\
+
+  \hspace*{0.5cm}   exit(0); \\
+\} 
+}}
+\end{quote}
 
@@
 \begin{center}
-\begin{tabular}{lll}
+\begin{tabular}{llp{7cm}}
 \hline
@@
 Energies & $E_\nu$ & $\mathrm{GeV}$ \\  
-Fiducial masses & $m_{\mathrm{Det}}$ & $\mathrm{kt}$ \\
+Fiducial masses & $m_{\mathrm{Det}}$ & $\mathrm{t}$ (reactor exp.) or $\mathrm{kt}$ (accelerator exp.), \newline  depends on experiment definition \\
 Time intervals & $t_{\mathrm{run}}$ & $\mathrm{yr}$ \\
-Source powers & $P_{\mathrm{Source}}$ & ??? \\
+Source powers & $P_{\mathrm{Source}}$ & Useful parent particle decays/$\mathrm{yr}$ \newline (Neutrino factory, $\beta$-Beam), \newline $\mathrm{GW}$ thermal power (reactor exps.), \newline or
+$\mathrm{MW}$ target power (superbeams); \newline
+depends on flux definition
+ \\
 % Integrated luminosities & $m_{\mathrm{Det}} \, t_{\mathrm{run}}$ & $\mathrm{kt \cdot yr}$ \\
-Cross sections & $\sigma_{\mathrm{CC}}$ &  ???? \\
+Cross sections/E & $\sigma_{\mathrm{CC}}/E$ & $10^{-38} \, \mathrm{cm^2}/\mathrm{GeV^2}$ \\
 \hline
@@
 \end{table}
+
+WW CONTINUE HERE:
 
@@
 \begin{function}
-\index{{\tt LoadExperiment}}
 \index{Experiment initialization}
-{\tt LoadExperiment}$($``filename'' $)$ adds a single experiment to the list of currently loaded experiments.
+\GLB{LoadExperiment}$($``filename'' $)$ adds a single experiment to the list of currently loaded experiments.
 All currently loaded experiments are evaluated simultaneously (if not explicitely stated otherwise), \ie , their $\chi^2$-values are added.
@@
 \begin{function}
-\index{{\tt ClearExp}}
-{\tt ClearExp}$()$ removes all experiments from the evaluation list.   
+\GLB{ClearExp}$()$ removes all experiments from the evaluation list.   
 \end{function}
@@
 \begin{function}
-\index{{\tt SetRunningTime}}
-{\tt SetRunningTime}$(N_{\mathrm{exp}},t_{\mathrm{run}})$ sets the running time of experiment number $N_{\mathrm{exp}}$ to $t_{\mathrm{run}}$ years.
+\GLB{SetRunningTime}$(N_{\mathrm{exp}},t_{\mathrm{run}})$ sets the running time of experiment number $N_{\mathrm{exp}}$ to $t_{\mathrm{run}}$ years.
 \end{function}
  \begin{function}
-\index{{\tt SetTargetMass}}
-{\tt SetTargetMass}$(N_{\mathrm{exp}},m_{\mathrm{Det}})$ sets the fiducial mass of experiment number $N_{\mathrm{exp}}$ to $m_{\mathrm{Det}}$ kilotons.
+\GLB{SetTargetMass}$(N_{\mathrm{exp}},m_{\mathrm{Det}})$ sets the fiducial mass of experiment number $N_{\mathrm{exp}}$ to $m_{\mathrm{Det}}$ kilotons.
 \end{function}
 \begin{function}
-\index{{\tt SetSourcePower}}
-{\tt SetSourcePower}$(N_{\mathrm{exp}},P_{\mathrm{Source}})$ sets the source power of experiment number $N_{\mathrm{exp}}$ to $P_{\mathrm{source}}$.
+\GLB{SetSourcePower}$(N_{\mathrm{exp}},P_{\mathrm{Source}})$ sets the source power of experiment number $N_{\mathrm{exp}}$ to $P_{\mathrm{source}}$.
 The definition of the source power depends on the experiment type: ...
@@
 \begin{function}
-\index{{\tt InfoExperiment}}
-{\tt InfoExperiment}$()$ prints a list of the initialized functions with their experiment numbers and their most important parameters to the standard output.
+\GLB{InfoExperiment}$()$ prints a list of the initialized functions with their experiment numbers and their most important parameters to the standard output.
 \end{function} 
@@
 \begin{function}
-\index{{\tt ShowCannels}}
-{\tt ShowChannels}$(N_{\mathrm{exp}})$ prints the information about the oscillation channels of the experiment with the number $N_{\mathrm{exp}}$ to the standard output.
+\GLB{ShowChannels}$(N_{\mathrm{exp}})$ prints the information about the oscillation channels of the experiment with the number $N_{\mathrm{exp}}$ to the standard output.
 \end{function}
@@
 \begin{function}
-\index{{\tt SetVacuumParameters}}
-{\tt SetVacuumParameters}$(\{\theta_{12}, \theta_{13}, \theta_{23}, \deltacp , \sdm , \ldm \})$ sets the neutrino oscillation parameters to be used to compute the reference rate vector in vacuum.
+\GLB{SetVacuumParameters}$(\{\theta_{12}, \theta_{13}, \theta_{23}, \deltacp , \sdm , \ldm \})$ sets the neutrino oscillation parameters to be used to compute the reference rate vector in vacuum.
 \end{function}
 \begin{function}
-\index{{\tt SetRates}}
-{\tt SetRates}$( )$ computes the reference rate vector for the neutrino oscillation parameters set with {\tt SetVacuumParameters}. 
+\GLB{SetRates}$( )$ computes the reference rate vector for the neutrino oscillation parameters set with {\tt SetVacuumParameters}. 
 \end{function}
@@
 \begin{function} 
-\index{{\tt Chi}}
-{\tt Chi}$(\{ \theta_{12}, \theta_{13}, \theta_{23}, \deltacp , \sdm , \ldm, \hat{\rho_1}, \hdots , \hat{\rho_n} \})$  returns the total added $\chi^2$ of all loaded experiments.
+\GLB{Chi}$(\{ \theta_{12}, \theta_{13}, \theta_{23}, \deltacp , \sdm , \ldm, \hat{\rho_1}, \hdots , \hat{\rho_n} \})$  returns the total added $\chi^2$ of all loaded experiments.
 \end{function}
 \begin{function}
-\index{{\tt SingleChi}}
- {\tt SingleChi}$(\{ \theta_{12}, \theta_{13}, \theta_{23}, \deltacp , \sdm , \ldm, \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}} )$  returns the total $\chi^2$ of the experiment number $N_{\mathrm{exp}}$.
+ \GLB{SingleChi}$(\{ \theta_{12}, \theta_{13}, \theta_{23}, \deltacp , \sdm , \ldm, \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}} )$  returns the total $\chi^2$ of the experiment number $N_{\mathrm{exp}}$.
 \end{function}
@@
 \begin{function}
-\index{{\tt SetSystematics}}
-{\tt SetSystematics}$(N_{mathrm{exp}}, \, sys)$ switches the systematics minimization for experiment $N_{\mathrm{exp}}$ on (sys=1) or off (sys=0).
+\GLB{SetSystematics}$(N_{mathrm{exp}}, \, sys)$ switches the systematics minimization for experiment $N_{\mathrm{exp}}$ on (sys=1) or off (sys=0).
 \end{function}
@@
 \begin{function}
-\index{{\tt SetStartingValues}}
-{\tt SetStartingValues}$(\{\rho^0_{\theta_{12}}, \, \rho^0_{\theta_{13}}, \, \rho^0_{\theta_{23}}, \, \rho^0_{\deltacp}, \, \rho^0_{\sdm}, \, \rho^0_{\ldm}, \, \rho^0_{\hat\rho} \})$ sets the starting values for all of the following minimizer calls.
+\GLB{SetStartingValues}$(\{\rho^0_{\theta_{12}}, \, \rho^0_{\theta_{13}}, \, \rho^0_{\theta_{23}}, \, \rho^0_{\deltacp}, \, \rho^0_{\sdm}, \, \rho^0_{\ldm}, \, \rho^0_{\hat\rho} \})$ sets the starting values for all of the following minimizer calls.
 \end{function}
 \begin{function}
-\index{{\tt SetInputErrors}}
-{\tt SetInputErrors}$(\{\sigma_{\theta_{12}}, \, \sigma_{\theta_{13}}, \, \sigma_{\theta_{23}}, \, \sigma_{\deltacp}, \, \sigma_{\sdm}, \, \sigma_{\ldm}, \, \sigma_{\hat\rho} \})$ sets the input errors for all of the following minimizer calls.
+\GLB{SetInputErrors}$(\{\sigma_{\theta_{12}}, \, \sigma_{\theta_{13}}, \, \sigma_{\theta_{23}}, \, \sigma_{\deltacp}, \, \sigma_{\sdm}, \, \sigma_{\ldm}, \, \sigma_{\hat\rho} \})$ sets the input errors for all of the following minimizer calls.
 \end{function}
@@
 \begin{function}
-\index{{\tt ChiTheta}}
-{\tt ChiTheta}$( \theta_{13}, \, \{ \theta_{12}, \, \theta_{23}, \, \deltacp , \, \sdm , \, \ldm, \, \hat{\rho_1}, \,  \hdots ,  \,\hat{\rho_n} \})$ returns the projection onto the $\theta_{13}$-axis for all experiments.
+\GLB{ChiTheta}$( \theta_{13}, \, \{ \theta_{12}, \, \theta_{23}, \, \deltacp , \, \sdm , \, \ldm, \, \hat{\rho_1}, \,  \hdots ,  \,\hat{\rho_n} \})$ returns the projection onto the $\theta_{13}$-axis for all experiments.
 The parameters are the fixed value of $\theta_{13}$ and the starting point of the minimizer including all starting values for the matter density scaling factors (without $\theta_{13}$, which is fixed).
@@
 \begin{function}
-\index{{\tt SingleChiTheta}}
-{\tt SingleChiTheta}$( \theta_{13}, \, \{ \theta_{12}, \, \theta_{23}, \, \deltacp , \, \sdm , \, \ldm,  \, \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}})$ works similar to {\tt ChiTheta}, but only for one of the loaded experiments $N_{\mathrm{exp}}$.
+\GLB{SingleChiTheta}$( \theta_{13}, \, \{ \theta_{12}, \, \theta_{23}, \, \deltacp , \, \sdm , \, \ldm,  \, \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}})$ works similar to {\tt ChiTheta}, but only for one of the loaded experiments $N_{\mathrm{exp}}$.
 It returns the list $\{ \chi^2, \, \theta_{12}, \,  \theta_{23}, \, \deltacp , \, \sdm , \, \ldm, \, \hat{\rho}_{N_{\mathrm{exp}}} , \, N_{\mathrm{Iter}} \}$.
@@
 \begin{function}
-\index{{\tt ChiDelta}}
-{\tt ChiDelta}$( \deltacp, \, \{ \theta_{12}, \, \theta_{13}, \, \theta_{23}, \, \sdm , \, \ldm, \, \hat{\rho_1}, \,  \hdots ,  \,\hat{\rho_n} \})$ returns the projection onto the $\deltacp$-axis for all experiments.
+\GLB{ChiDelta}$( \deltacp, \, \{ \theta_{12}, \, \theta_{13}, \, \theta_{23}, \, \sdm , \, \ldm, \, \hat{\rho_1}, \,  \hdots ,  \,\hat{\rho_n} \})$ returns the projection onto the $\deltacp$-axis for all experiments.
 The parameters are the fixed value of $\deltacp$ and the starting point of the minimizer including all starting values for the matter density scaling factors (without $\deltacp$, which is fixed).
@@
 \begin{function}
-\index{{\tt SingleChiDelta}}
-{\tt SingleChiDelta}$( \deltacp, \, \{ \theta_{12}, \, \theta_{13}, \, \theta_{23}, \, \sdm , \, \ldm, \,  \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}})$ works similar to {\tt ChiDelta}, but only for one of the loaded experiments $N_{\mathrm{exp}}$.
+\GLB{SingleChiDelta}$( \deltacp, \, \{ \theta_{12}, \, \theta_{13}, \, \theta_{23}, \, \sdm , \, \ldm, \,  \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}})$ works similar to {\tt ChiDelta}, but only for one of the loaded experiments $N_{\mathrm{exp}}$.
 It returns the list $\{ \chi^2, \, \theta_{12}, \,  \theta_{13}, \,  \theta_{23}, \, \sdm , \, \ldm, \, \hat{\rho}_{N_{\mathrm{exp}}} , \, N_{\mathrm{Iter}} \}$.
@@
 \begin{function}
-\index{{\tt ChiThetaDelta}}
-{\tt ChiThetaDelta}$(\theta_{13}, \, \deltacp, \, \{ \theta_{12}, \, \theta_{23},  \, \sdm , \, \ldm, \,   \hat{\rho_1}, ...
+\GLB{ChiThetaDelta}$(\theta_{13}, \, \deltacp, \, \{ \theta_{12}, \, \theta_{23},  \, \sdm , \, \ldm, \,   \hat{\rho_1}, ...
 , \hat{\rho_n} \})$ returns the list  $\{ \chi^2, \, \theta_{12}, \, \theta_{23}, \, \sdm , \, \ldm,  \, \hat{\rho_1}, \, \hdots , \, \hat{\rho_n} , \, N_{\mathrm{Iter}} \}$ for the projection into the $\theta_{13}$-$\deltacp$-plane for all experiments.
@@
 \begin{function}
-\index{{\tt SingleChiThetaDelta}}
-{\tt SingleChiThetaDelta}$( \theta_{13}, \, \deltacp, \, \{ \theta_{12}, \, \theta_{23}, \, \sdm , \, \ldm, \,   \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}})$ returns the list $\{ \chi^2, \, \theta_{12}, \,   \theta_{23},  \, \sdm , \, \ldm, \, \hat{\rho}_{N_{\mathrm{exp}}} ,  \, N_{\mathrm{Iter}} \}$ for the projection into the $\theta_{13}$-$\deltacp$-plane for one experiment $N_{\mathrm{exp}}$.
+\GLB{SingleChiThetaDelta}$( \theta_{13}, \, \deltacp, \, \{ \theta_{12}, \, \theta_{23}, \, \sdm , \, \ldm, \,   \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}})$ returns the list $\{ \chi^2, \, \theta_{12}, \,   \theta_{23},  \, \sdm , \, \ldm, \, \hat{\rho}_{N_{\mathrm{exp}}} ,  \, N_{\mathrm{Iter}} \}$ for the projection into the $\theta_{13}$-$\deltacp$-plane for one experiment $N_{\mathrm{exp}}$.
 \end{function}
@@
 \begin{function}
-\index{{\tt ChiAll}}
-{\tt ChiAll}$(\{ \theta_{13}, \, \theta_{12}, \, \theta_{23}, \, \deltacp , \, \sdm , \, \ldm,  \, \hat{\rho_1}, \, \hdots , \, \hat{\rho_n} \})$ finds the local minimum close to the given starting point (list) for all initialized experiments.
+\GLB{ChiAll}$(\{ \theta_{13}, \, \theta_{12}, \, \theta_{23}, \, \deltacp , \, \sdm , \, \ldm,  \, \hat{\rho_1}, \, \hdots , \, \hat{\rho_n} \})$ finds the local minimum close to the given starting point (list) for all initialized experiments.
 It returns the list  $\{ \chi^2, \, \theta_{13}, \, \theta_{12}, \, \theta_{23}, \, \deltacp , \, \sdm , \,  \ldm,  \, \hat{\rho_1}, \, \hdots , \hat{\rho_n} , \, N_{\mathrm{Iter}} \}$ with the minimum $\chi^2$-value, the actual position of the local minimum, and the number of iterations $N_{\mathrm{Iter}}$ used.
@@
 \begin{function}
-\index{{\tt SingleChiAll}} 
-{\tt SingleChiAll}$( \{ \theta_{13}, \, \theta_{12}, \, \theta_{23}, \, \deltacp , \, \sdm , \, \ldm, \,   \hat{\rho}_{N_{\mathrm{exp}}} \}, \, \, N_{\mathrm{exp}})$ finds the local minimum close to the given starting point (list) for the experiment $N_{\mathrm{exp}}$ only.
+\GLB{SingleChiAll}$( \{ \theta_{13}, \, \theta_{12}, \, \theta_{23}, \, \deltacp , \, \sdm , \, \ldm, \,   \hat{\rho}_{N_{\mathrm{exp}}} \}, \, \, N_{\mathrm{exp}})$ finds the local minimum close to the given starting point (list) for the experiment $N_{\mathrm{exp}}$ only.
 It returns the list

### 2004-06-29 Documentation/part2.tex
*** empty log message ***

@@
 \part{The Abstract Experiment Definition Language -- AEDL}
+\label{part:2}
 
@@
 
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=13cm]{AEDL1}
+\end{center}
+\caption{\label{fig:channel} General concept of a ``channel''.}
+\end{figure}
+
 In order to implement a experiment definition including various
@@
 The first level is the so-called ``channel'', which is the link between 
-the oscillation physics and the detection properties for a specfific oscillation pattern.
+the oscillation physics and the detection properties for a specfific oscillation pattern (\cf, \figu{channel}).
 A channel specifies the mapping of a specific neutrino flavor produced by the source onto a reconstructed neutrino flavor.
@@
 
-The channels are the building blocks for the so-called ``rules''.
-In general, a rule consists of one or more ``signal'' and ``background'' oscillation channels, which are normalized with efficiencies.
-The event numbers from these channels are added {\em before} the $\chi^2$-value is calculated.
-In addition, each rule implements an {\em independent} systematics, such as signal and background normalization errors.
-Eventually, each rule gives a $\chi^2$-value, and the total $\chi^2$ is obtained by adding the $\chi^2$'s of all rules.
- An example for a rule could look like this: We want to detect electron neutrino appearance (``signal''), where the overall efficiency for quasi-elastics electron neutrino events is $0.4$.
-There is a fraction of $0.01$ of all neutral current events which are mis-identified as quasi-elastic electron neutrino events (``background'').
-The neutral current fraction is only known within $10\%$ (``background uncertainty'') and there is
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=8cm]{SignalBackground}
+\end{center}
+\caption{\label{fig:rule} General concept of a ``rule''.}
+\end{figure}
+
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=9cm]{Rules}
+\end{center}
+\caption{\label{fig:experiment} General concept of an ``experiment''.}
+\end{figure}
+
+The channels are the building blocks for the so-called ``rules''. In general, a rule consists of one or more ``signal'' and ``background'' oscillation channels, which are normalized with efficiencies
+(\cf, \figu{rule}). The event numbers from these channels are added {\em before} the $\Delta \chi^2$-value is calculated.\footnote{Note that
+in this manual, we $\chi^2$ and $\Delta \chi^2$ are equal, since
+for simulated data $\Delta \chi^2 = 0$ at the best-fit point. Thus, we
+are using $\chi^2$ and $\Delta \chi^2$ as equal quantities.} In addition, each rule implements an {\em independent} systematics, such as signal and background normalization errors.
+Eventually, each rule gives a $\Delta \chi^2$-value, and the total $\Delta \chi^2$ of one experiment is obtained by adding the $\Delta \chi^2$'s of all rules (\cf, \figu{experiment}).
+ An example for a rule could look like this: We want to detect electron
+  neutrino appearance (``signal''), where the overall efficiency for 
+  quasi-elastics electron neutrino events is $0.4$. There is a fraction of 
+  $0.01$ of all neutral current events which are mis-identified as 
+  quasi-elastic electron neutrino events (``background''). The neutral 
+  current fraction is only known within $10\%$ (``background uncertainty'') 
+  and there is
 an energy scale uncertainty of $100\,\mathrm{MeV}$ (``energy calibration error'').
-All this systematics is independent of the other rules.  Thus, a rule connects the event rates to the calculation of a $\chi^2$ which properly includes 
-systematical errors. The resulting $\chi^2$ is then the starting point for the
-oscillation physics analysis. Note again that
+All this systematics is independent of the other rules.
+Thus, a rule connects the event rates to the calculation of a $\Delta \chi^2$ which properly includes systematical errors.
+The resulting $\Delta \chi^2$ is then the starting point for the oscillation physics analysis.
+Note again that
 \begin{itemize}
@@
 \item
- For each rule the $\chi^2$ is computed; the $\chi^2$'s from all rules are added.
+ For each rule the $\Delta \chi^2$ is computed; the $\Delta \chi^2$'s from all rules are added.
 \end{itemize}

### 2004-06-30 Documentation/part1.tex
*** empty log message ***

@@
 \\
- \hspace*{0.5cm} GLBInit(argv[0]); \hspace*{0.5cm} /* Initialize GLoBES library */ \\
+ \hspace*{0.5cm} glbInit(argv[0]); \hspace*{0.5cm} /* Initialize GLoBES library */ \\
 \\  
@@
 \label{ex:corrth13dcp}
-
 \index{Two-parameter correlation}
-A typical application for {\tt Chi} and {\tt SingleChi} is the visualization of two-parameter correlations.
-For example, to show the correlation between $\stheta$ and $\deltacp$ at a large neutrino factory with four years of running time for each polarity, one can use the following code:
+A typical and fast application for \GLB{glbChiSys} is the visualization of
+ two-parameter correlations. For example, to calculate the two-parameter 
+ correlation between $\stheta$ and $\deltacp$ at a neutrino factory, one
+ can use the following code excerpt from {\tt example1.c}:
 \begin{quote}
-{\tt
-/\% Experiment initialization: \%/ \\
-ClearExp(); \\
- InitExperiment("{\tt NuFact2.exp}"); \\
- SetRunningTime(0,4.0); \\
- MInfoExperiment(); \\
+{\tt {\footnotesize
+  /* Initialize parameter vector(s) and compute simulated data */ \\
+  glbDefineParams(true\_values,theta12,theta13,theta23,deltacp,sdm,ldm); \\
+  glbDefineParams(test\_values,theta12,theta13,theta23,deltacp,sdm,ldm);  \\
+  glbSetOscillationParameters(true\_values); glbSetRates(); \\
 \\
-/\% Set reference rate vector: \%/ \\
-th12=arcsin(sqrt(0.8))/2; sdm=7e-5; \\
-th23=3.14/4; ldm=2e-3; \\
-th13=arcsin(sqrt(0.001))/2; dcp=3.14/2; \\
-\mbox{SetVacuumParameters(\{th12, \, th13, \, th23, \, dcp, \, sdm, \, ldm\}); }\\
- SetRates(); \\
+  /* Iteration over all values to be computed */ \\
+  double x,y,res; \\   
+  for(x=-4.0;x<-2.0+0.01;x=x+2.0/50) \\
+  for(y=0.0;y<200.0+0.01;y=y+200.0/50) \\
+  \{ \\
+\hspace*{0.5cm} /* Set parameters in vector of test values */ \\
+\hspace*{0.5cm} glbSetOscParams(test\_values,asin(sqrt(pow(10,x)))/2,GLB\_THETA\_13); \\
+\hspace*{0.5cm} glbSetOscParams(test\_values,y*M\_PI/180.0,GLB\_DELTA\_CP); \\
   \\
-/\% Compute chi-square matrix: \%/ \\
-for (float x=-4; x<-2; x=x+2/50) \\
-\hspace*{0.5cm} for (float y=0; y<200; y=y+200/50) \{ \\
-\hspace*{1cm} theth13 = arcsin(sqrt(10$\hat{\, \, \,}$x))/2; \\
-\mbox{\hspace*{1cm} Print(\{x,y,Chi(\{th12, theth13,  th23,  y, sdm, ldm, 1.0\}) \}) \};}
-}
+\hspace*{0.5cm} /* Compute Chi2 for all loaded experiments and all rules */ \\
+\hspace*{0.5cm} res=glbChiSys(test\_values,GLB\_ALL,GLB\_ALL); \\
+\hspace*{0.5cm} AddToOutput(x,y,res); \\
+  \} \\
+} }
 \end{quote}
 
-The resulting matrix can then be plotted as a contour plot (2 d.o.f.):
+The resulting data can then be plotted as a contour plot (2 d.o.f.):
 \begin{center}
@@
 
-This example demonstrates how to use  {\tt ChiTheta} to project the fit manifold onto the $\stheta$-axis, \ie, how one can include correlations.
-We compute two sets of data: one for keeping all parameters but $\deltacp$ fixed (two-parameter correlations), and one for keeping all parameters free (multi-parameter correlation), but some external input for the solar parameters and $\rho$.
-Note that fixing all parameters but $\deltacp$ corresponds to imposing external knowledge on these parameters.
+This example demonstrates how to project the fit manifold onto the $\stheta$-axis, \ie, how one can include correlations.
+We compute two sets of data: one for keeping all parameters but $\deltacp$ fixed (two-parameter correlation), and one for keeping all parameters free (multi-parameter correlation).
+However, we impose external precisions for the solar parameters and the matter density.
+The following code excerpt is from
+{\tt example2.c}:
 
 \begin{quote}
-{\tt {\footnotesize
-/\% Experiment initialization, set reference rate vector: \%/ \\
-\mbox{ClearExp(); InitExperiment("{\tt NuFact2.exp}"); SetRunningTime(0,4.0); }\\
-th12=arcsin(sqrt(0.8))/2; sdm=7e-5; th23=3.14/4; ldm=2e-3; \\
-th13=arcsin(sqrt(0.001))/2; dcp=3.14/2; \\
-\mbox{SetVacuumParameters(\{th12, th13, th23, dcp, sdm, ldm\}); SetRates();}\\
+{\tt {\scriptsize
+  /* Set starting values and input errors for all projections */  \\
+  glbDefineParams(input\_errors,theta12*0.1,10,10,10,sdm*0.1,ldm);\\  
+  glbSetDensityParams(input\_errors,0.05,GLB\_ALL);\\
+  glbSetStartingValues(true\_values);\\
+  glbSetInputErrors(input\_errors);\\
 \\
-/\%  Compute chi-square list with all but dcp fixed: \%/ \\
-SetStartingValues(\{th12, th13, th23, dcp, sdm, ldm, 1.0\}); \\
-\mbox{SetInputErrors(\{th12*0.001, 10, th23*0.001, 10, sdm*0.001, ldm*0.001, 0.001\});} \\
-for (float x=-4; x<-2; x=x+2/20) \\
-\mbox{\hspace*{0.5cm} Print(\{x,ChiDelta(x,\{th12, th13, th23, dcp, sdm, ldm, 1.0\}) \}); }\\
+  /* Define two-parameter projection: Only deltacp is free! */ \\
+  glbDefineProjection(th13\_projection,GLB\_FIXED,GLB\_FIXED,GLB\_FIXED,GLB\_FREE,GLB\_FIXED,GLB\_FIXED); \\
+  glbSetProjection(th13\_projection); \\ 
 \\
-/\% Compute chi-square list with all parameters free: \%/ \\
-\mbox{/\% (but: 10\% prec. for solar params, 5\% for matter density) \%/} \\
-SetStartingValues(\{th12, th13, th23, dcp, sdm, ldm, 1.0\}); \\
-SetInputErrors(\{th12*0.1, 10, 10, 10, sdm*0.1, ldm, 0.05\}); \\
-for (float x=-4; x<-2; x=x+2/20) \\
-\mbox{\hspace*{0.5cm} Print(\{x,ChiDelta(x,\{th12, th13, th23, dcp, sdm, ldm, 1.0\}) \});}
+  /* Iteration over all values to be computed */ \\
+  double x,res1,res2;     \\
+  for(x=-4;x<-2.0+0.001;x=x+2.0/50) \\
+  \{ \\
+\hspace*{0.5cm}  /* Set fit value of stheta */ \\
+\hspace*{0.5cm} glbSetOscParams(test\_values,asin(sqrt(pow(10,x)))/2,1); \\
+    \\
+\hspace*{0.5cm} /* Guess fit value for deltacp in order to safely find minimum */ \\
+\hspace*{0.5cm} glbSetOscParams(test\_values,200.0/2*(x+4)*M\_PI/180,3); \\
+ \\
+\hspace*{0.5cm} /* Compute Chi2 for user-defined two-parameter correlation */ \\
+\hspace*{0.5cm} res1=glbChiNP(test\_values,NULL,GLB\_ALL); \\
+      \\
+\hspace*{0.5cm} /* Compute Chi2 for full correlation: minimize over all but theta13 */ \\
+\hspace*{0.5cm} res2=glbChiTheta(test\_values,NULL,GLB\_ALL); \\
+  \\
+\hspace*{0.5cm} AddToOutput(x,res1,res2);\\
+  \} \\
+
 }}
@@
 
-In many cases, one can find the $\mathrm{sgn}(\ldm)$-degeneracy with {\tt ChiAll}: One lets the local minimizer start at the position of the suspected solution at the opposite sign of $\ldm$.
-The following example corresponds to finding the degenerate solution for the example on page~\pageref{ex:corrth13dcp}.
+In many cases, one can find the exact position of the $\mathrm{sgn}(\ldm)$-degeneracy with \GLB{glbChiAll}, where one starts 
+the local minimizer at the suspected position 
+and let it converge into the minimum.  The following code excerpt  corresponds to finding the degenerate solution for the example on
+  page~\pageref{ex:corrth13dcp}, and it is from {\tt example3.c}: 
 \begin{quote}
@@
 {\footnotesize
-/\% Experiment initialization: \%/ \\
-ClearExp(); InitExperiment("{\tt NuFact2.exp}"); SetRunningTime(0,4.0); \\
+  /* Set starting vales to suspected position at opposite sign of ldm */ \\
+  \mbox{glbDefineParams(starting\_values,theta12,theta13,theta23,deltacp,sdm,-ldm);} \\
   \\
-/\% Set reference rate vector: \%/ \\
-th12=arcsin(sqrt(0.8))/2; sdm=7e-5; th23=3.14/4; ldm=2e-3; \\
-th13=arcsin(sqrt(0.001))/2; dcp=3.14/2; \\
-\mbox{SetVacuumParameters(\{th12, \, th13, \, th23, \, dcp, \, sdm, \, ldm\});  SetRates();} \\
+  /* Set input errors for external input, where some information on ldm \\
+   \hspace*{0.5cm} is imposed in order to avoid falling into the right-sign solution */ \\
+  glbDefineParams(input\_errors,theta12*0.1,10,10,10,sdm*0.1,ldm/3); \\ 
+  glbSetDensityParams(input\_errors,0.05,GLB\_ALL); \\
   \\
-/\% Find negative-sign solution: \%/ \\
-SetStartingValues(\{th12, \, th13, \, th23, \, dcp, \, sdm, \, -ldm, \, 1.0\}); \\
-SetInputErrors(\{th12*0.1, \, 10, \, 10, \, 10, \, sdm*0.1, \, ldm/3, \, 0.05\}); \\
-d = ChiAll(\{th12, \, th13, \, th23, \, dcp, \, sdm, \, ldm, \, 1.0\}); \\
+  glbSetStartingValues(starting\_values); \\
+  glbSetInputErrors(input\_errors); \\
   \\
-/\% Compute chi-square matrix: \%/ \\
-for (float x=-4; x<-2; x=x+2/50) \\
-\hspace*{0.5cm} for (float y=0; y<200; y=y+200/50) \{ \\
-\hspace*{1cm} theth13 = arcsin(sqrt(10$\hat{\, \, \,}$x))/2; \\
-\mbox{\hspace*{1cm} Print(\{x,y,Chi(\{d[2], theth13,  d[4],  y, d[6], d[7], d[8]\}) \}) \};}
-}
-}
+  /* Localize degenerate solution by minimzation over all parameters */ \\
+  double CL=glbChiAll(starting\_values,deg\_pos,GLB\_ALL); \\
+   \\
+  \mbox{/* Now: CL is the chi2 of the deg.~solution and deg\_pos the position */} 
+}}
 \end{quote}
 
-The resulting matrix can then be plotted as a contour plot in addition to the original solution (2 d.o.f., gray contours):
+Using {\tt ent\_pos} to obtain a section of the degeneracy in the
+$\stheta$-$\deltacp$-plane (\cf, {\tt example3.c}), one can plot it as a contour plot in addition to the original solution (2 d.o.f., gray curves):
 \begin{center}
@@
 \index{Systematics on/off}
-
-Here it is shown how one can successively include systematics, correlations,
-and degeneracies at the example of the $\stheta$ sensitivity limit.
+Here it is demonstrated how one can successively include systematics, correlations, and degeneracies at the example of the $\stheta$-sensitivity limit.
 An important part of this example is how two switch the systematics off,
 \ie, how to obtain the sensitivity limit from statistics only. Since
-this example is very advanced, we only show the respective part of the code:
+this example is very advanced, we only show the respective function of the code:
 \begin{quote}
 {\tt 
-Here comes the code excerpt.
+/* Calculate chi2 with statistics only */ \\
+double CalcNoSystematics(double theldm,double thex) \\
+\{ \\
+\hspace*{0.5cm} /* Switch systematics off for all exps and all rules */ \\
+\hspace*{0.5cm} glbSwitchSystematics(GLB\_ALL,GLB\_ALL,GLB\_OFF); \\
+\\
+\hspace*{0.5cm} /* Calculate Chi2-list as if systematics were on */ \\
+\hspace*{0.5cm} double res=CalcSystematics(theldm,thex); \\
+\\
+\hspace*{0.5cm} /* Switch systematics on for all exps and all rules */ \\
+\hspace*{0.5cm} glbSwitchSystematics(GLB\_ALL,GLB\_ALL,GLB\_ON); \\
+\\
+\hspace*{0.5cm} return res; \\
+\} \\
 }
 \end{quote}
-The complete code can be found as ``example4.c'' with the software.
 
-The returned lists of data then represent $\Delta \chi^2$ 
+\vspace*{-0.5cm}
+
+The complete code can be found as {\tt example4.c} with the software,
+which consists of many of the applications from the earlier examples.
+In addition, it uses a little trick: It avoids falling into the wrong
+ solution with {\tt glbChiTheta}
+by using the fit value of $\deltacp$ from the step before as prediction
+of the position of the current calculation.
+
+The returned lists of data from thie example represent $\chi^2$ 
 as function of the fit value of $\stheta$. The intersections of these
-curves with the line $\Delta \chi^2 = 9$ give the $\stheta$ sensitivity
-limits, where we do not include the $\mathrm{sgn}(\ldm)$- and $(\deltacp,\theta_{13})$-degeneracies in the sensitivity limit with
-correlations only (green bar):
+curves with the line $\chi^2 = 9$ give the $\stheta$ sensitivity
+limits at the $3 \sigma$ confidence level, where we do not include the
+ $\mathrm{sgn}(\ldm)$- and $(\deltacp,\theta_{13})$-degeneracies in the sensitivity limit with correlations only (green bar):
 \begin{center}
-\colorbox{white}{\includegraphics[width=7cm]{barsex}}
-
+\colorbox{white}{\includegraphics[width=6.5cm]{barsex}}
 \end{center}

### 2004-06-30 Documentation/example3.c

@@ -1,5 +1,27 @@
-/* Example: Finding the sgn(dm31^2)-degeneracy
- * Copyright 2004 GLoBES collaboration
+/* GLoBES -- General LOng Baseline Experiment Simulator
+ * (C) 2002 - 2004,  The GLoBES Team
+ *
+ * GLoBES is mainly intended for academic purposes. Proper
+ * credit must be given if you use GLoBES or parts of it. Please
+ * read the section 'Credit' in the README file.
+ *
+ * This program is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU General Public License as published by
+ * the Free Software Foundation; either version 2 of the License, or
+ * (at your option) any later version.
+ *
+ * This program is distributed in the hope that it will be useful,
+ * but WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ * GNU General Public License for more details.
+ *
+ * You should have received a copy of the GNU General Public License
+ * along with this program; if not, write to the Free Software
+ * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
+ */
+ 
+ /* 
+ * Example: Find the sgn(dm31^2)-degeneracy and compute shape with glbChiSys
  * Compile with ``make example3''
  */ 
 
@@ -17,10 +39,10 @@ char MYFILE[]="test3.dat";
 int main(int argc, char *argv[])
 { 
   /* Initialize libglobes */
-  GLBInit(argv[0]); 
+  glbInit(argv[0]); 
 
   /* Initialize experiment NuFact.glb */
-  GLBInitExperiment("NuFact.glb",&ExpList[0],&numofexps); 
+  glbInitExperiment("NuFact.glb",&glb_experiment_list[0],&glb_num_of_exps); 
 
   /* Intitialize output */
   InitOutput(MYFILE,"Format: Log(10,s22th13)   deltacp   chi^2 \n"); 
@@ -34,34 +56,26 @@ int main(int argc, char *argv[])
   double ldm = 2e-3;
   
 	/* Initialize parameter vector(s) */
-  glb_params true_values = alloc_params();
-  glb_params starting_values = alloc_params();
[…]

### 2004-06-30 Documentation/part1.tex
*** empty log message ***

@@
 \begin{center}
-\begin{tabular}{llp{7cm}}
-\hline
-Quantities & Examples & Units \\
-\hline
-Angles & $\theta_{13}$, $\theta_{12}$, $\theta_{23}$, $\deltacp$ & Radians  \\
-Mass squared differences & $\sdm$, $\ldm$ & $\mathrm{eV}^2$ \\
-Matter densitities & $\rho_i$ & $\mathrm{g}/\mathrm{cm}^3$ \\
-Baseline lengths & $L_i$ & $\mathrm{km}$ \\
-Energies & $E_\nu$ & $\mathrm{GeV}$ \\  
-Fiducial masses & $m_{\mathrm{Det}}$ & $\mathrm{t}$ (reactor exp.) or $\mathrm{kt}$ (accelerator exp.), \newline  depends on experiment definition \\
-Time intervals & $t_{\mathrm{run}}$ & $\mathrm{yr}$ \\
-Source powers & $P_{\mathrm{Source}}$ & Useful parent particle decays/$\mathrm{yr}$ \newline (Neutrino factory, $\beta$-Beam), \newline $\mathrm{GW}$ thermal power (reactor exps.), \newline or
-$\mathrm{MW}$ target power (superbeams); \newline
-depends on flux definition
- \\
-% Integrated luminosities & $m_{\mathrm{Det}} \, t_{\mathrm{run}}$ & $\mathrm{kt \cdot yr}$ \\
-Cross sections/E & $\sigma_{\mathrm{CC}}/E$ & $10^{-38} \, \mathrm{cm^2}/\mathrm{GeV^2}$ \\
-\hline
-\end{tabular}
-\mycaption{\label{tab:units} \index{Units in \GLOBES } Quantities used in \GLOBES , examples of these quantities, and their standard units in the application software.}
-\end{center}
-\end{table}
-
-WW CONTINUE HERE:
-
-Throughout the programming interface of \GLOBES , the software needs to transfer parameters to and from the software core.
-Unless the Experiment Definition Module \EDM , the programming interface only uses one set of units for each type of quantity in order to avoid confusion about the definition of individual parameters.
-\tabl{units} summarizes the units of the most important quantities used in \GLOBES .
-
-\begin{table}[t]
-\begin{center}
 \begin{tabular}{llp{7cm}c}
@@
 \end{center}
-\mycaption{\label{tab:experiments} \index{Experiment data files (table)} Different pre-defined experiments, their filenames (to be used in {\tt LoadExperiment}), their short description, and the references in which they are defined.
+\mycaption{\label{tab:experiments} \index{Experiment data files (table)} NEEDS TO BE ADJUSTED WHEN PROTOTYPES ARE AVAILABLE!
+Different pre-defined experiments, their filenames (to be used in {\tt LoadExperiment}), their short description, and the references in which they are defined.
 Note that all experiments use in their standard configurations one year of running time.
@@
 
-In principle, \GLOBES\ can handle any number of different long-baseline experiments simultaneously.
-This means that their $\chi^2$-values are added {\em after} the minimization over the independent systematics parameters and {\em before} any minimization over the oscillation parameters.
-Though the simplest case of only one experiment may be most often used, more experiments are useful in many cases.
-For example, running a superbeam some years in the neutrino mode and some years in the antineutrino mode can be, to a first approximation, simulated by the combination of two such experiments.\footnote{Note that in this case the systematics parameters are minimized over independently, which means that this approach does not allow correlations among the systematics parameters.
-Therefore, the neutrino factory with the symmetric operation of both polarities is encapsulated into a single experiment.} Another example is the test of synergetic effects among different experiment types.
-Thus, \GLOBES\ has an internal (initially empty) list of currently initialized experiments.
-To add a pre-defined experiment to this list, one can use the function {\tt LoadExperiment}:
+Before one can use \GLOBES , one has to initialize the \GLOBES\
+library \GLB{libglobes}:
+\begin{function}
+\index{Initialization of {\tt libglobes}}
+\index{glbInit} {\tt void glbInit(char *name)} initializes the library {\tt libglobes} and has
+to be called in the beginning of each \GLOBES\ program. It takes the
+name {\tt name} of the program as a string to initialize the error handling
+functions. In many cases, it is sufficient to use the first
+argument from the command line as the program name (such as in example on page~\pageref{ex:c}).
+\end{function}
+
+\index{Number of experiments}
+In principle, the \GLOBES\ user interface can currently handle up to 32 of different long-baseline experiments simultaneously, where the number
+of existing experiment definition files can, of course, be unlimited.
+This means that their $\Delta \chi^2$-values are added {\em after} the minimization over the systematics parameters, and {\em before} any minimization over the oscillation parameters.
+Note that each experiment
+assumes a specific matter density profile for all of its rules, which means
+that it makes sense to simulate different operation modes within one
+experiment definition, and physically different baselines in different
+definitions. For details of the rate computation and
+simulation techniques, we refer at this place to \Part~\ref{part:2}. Though
+ the simplest case of simulating one experiment may be most often used, 
+ using more than one experiments are useful in many cases. For example, combinations of experiments can be tested for
+complementarity and competitiveness by equal means within one program.
+In general, many \GLOBES\ functions take the experiment number as
+a parameter, which runs from $0$ to \GLB{glb\_num\_of\_exps}-1 in the order of their initialization in the program.\footnote{Note that
+the global variable {\tt glb\_num\_of\_exps} must not be modified by the
+user.} In addition, using the parameter value \GLB{GLB\_ALL} as
+experiment number initiates a combined analysis of all loaded experiments.
+
+For storing the experiments, \GLOBES\ uses the initially empty list of experiments \GLB{glb\_experiment\_list}.
+To add a pre-defined experiment to this list, one can use the function {\tt glbInitExperiment}:
 \begin{function}
 \index{Experiment initialization}
-\GLB{LoadExperiment}$($``filename'' $)$ adds a single experiment to the list of currently loaded experiments.
-All currently loaded experiments are evaluated simultaneously (if not explicitely stated otherwise), \ie , their $\chi^2$-values are added.
+\index{glbInitExperiment}
+{\tt int glbInitExperiment(char *inf, glb\_exp *in, int *counter)}
+ adds a single experiment with the filename {\tt inf} to the list of currently loaded experiments. The {\tt counter} is a pointer to the 
+ variable containg the number of experiments, and the experiment {\tt in}
+ points to the beginning of the experiment list. The function returns
+ zero if it was successful. 
 \end{function}
-A list of pre-defined experiment types, their filenames, their short descriptions, and the references of their definitions can be found in \tabl{experiments}.
-To remove all experiments from the evaluation list, one uses {\tt ClearExp}:
+Normally, a typical  call of {\tt glbInitExperiment} is 
+\begin{quote}
+{\tt glbInitExperiment("NuFact.glb",\&glb\_experiment\_list[0],\\  \hspace*{8cm} \&glb\_num\_of\_exps); }
+\end{quote}
+In this case, the experiment in the file {\tt NuFact.glb} is added to the internal global list of experiments, and the experiment counter is increased. The
+experiment then has the number {\tt glb\_num\_of\_exps}-1. The elements
+of the experiment list have the type \GLB{glb\_exp}, which the
+user will not need to access directly in most cases. The experiment definition files, which usually end with {\tt .glb}, and any
+supporting files, are first of all searched in the current directory, and then in the global path variable \GLB{GLB\_PATH} set in your environment.
+A list of pre-defined experiment prototypes, their filenames, their short descriptions, and the references of their definitions can be found in \tabl{experiments}.
+If the program cannot find these files, or some of them are syntactially not correct, it will break at this place.
+
+One can also remove all experiments from the evaluation list at running
+ time:
 \begin{function}
-\GLB{ClearExp}$()$ removes all experiments from the evaluation list.   
+\index{glbClearExperimentList}
+{\tt void ClearExperimentList()} removes all experiments from the internal list and resets all counters.   
 \end{function}
-Both functions do not return anything.
-Thus, one can either add an experiment to the internal evaluation list, or remove all experiments from this list.
-After adding an experiment, it gets an internal experiment number $N_{\mathrm{exp}}$ assigned in the order of the addition, which is starting from zero and running to the number of experiments minus one.
-Therefore, one will be able to access the individual experiment by its number later.
+Note that changing the number of experiments requires a new initialization
+of all parameters of the types \GLB{glb\_params} and \GLB{glb\_projection}
+if the number of experiments changes, since these parameter structures internally carry lists for the matter densities of all experiments.
+Similarly, once should never call {\tt glbAlloc...} before the
+experiment initialization.
+
+\begin{table}[t]
+\begin{center}
+\begin{tabular}{llp{7cm}}
+\hline
+Quantities & Examples & Units \\
+\hline
+Angles & $\theta_{13}$, $\theta_{12}$, $\theta_{23}$, $\deltacp$ & Radians  \\
+Mass squared differences & $\sdm$, $\ldm$ & $\mathrm{eV}^2$ \\
+Matter densitities & $\rho_i$ & $\mathrm{g}/\mathrm{cm}^3$ \\
+Baseline lengths & $L_i$ & $\mathrm{km}$ \\
+Energies & $E_\nu$ & $\mathrm{GeV}$ \\  
+Fiducial masses & $m_{\mathrm{Det}}$ & $\mathrm{t}$ (reactor exp.) or $\mathrm{kt}$ (accelerator exp.), \newline  depends on experiment definition \\
+Time intervals & $t_{\mathrm{run}}$ & $\mathrm{yr}$ \\
+Source powers & $P_{\mathrm{Source}}$ & Useful parent particle decays/$\mathrm{yr}$ \newline (Neutrino factory, $\beta$-Beam), \newline $\mathrm{GW}$ thermal power (reactor exps.), \newline or
+$\mathrm{MW}$ target power (superbeams); \newline
+depends on flux definition
+ \\
+% Integrated luminosities & $m_{\mathrm{Det}} \, t_{\mathrm{run}}$ & $\mathrm{kt \cdot yr}$ \\
+Cross sections/E & $\sigma_{\mathrm{CC}}/E$ & $10^{-38} \, \mathrm{cm^2}/\mathrm{GeV^2}$ \\
+\hline
+\end{tabular}
+\mycaption{\label{tab:units} \index{Units in \GLOBES } Quantities used in \GLOBES , examples of these quantities, and their standard units in the application software.}
+\end{center}
+\end{table}
+
+While interacting with the user interface of \GLOBES , parameters are transferred to and from the \GLOBES\ library. In \GLOBES , one set of units 
+for each type of quantity is used in order to avoid confusion about the definition of individual parameters.
+\tabl{units} summarizes the units of the most important quantities.
+In principle, the event rates are
+proportional to the product of source power $\times$ target mass $\times$
+ running time, which we call ``integrated luminosity''. Since especially the
+ definition of the source power depends on the experiment type, the quantities of the three luminosity components
+ are not unique and depend on the experiment definition. Usually,
+ one uses detector masses in kilotons for beam experiments,
+ and detector masses in tons for reactor experiments. Running times
+ are normally given in years, where it is often assumed that the 
+ experiment runs $100\%$ of the year. Thus, for shorter running periods,
+ the running times need to be renormalized. Source powers are
+ usually useful parent particle decays per year (neutrino factories,
+ $\beta$-beams), target power in megawatts (superbeams), or thermal
+ reactor power in gigawatts (reactor experiments).
+
+WW: WEITER FROM HERE:
 

### 2004-07-01 Documentation/part1.tex
*** empty log message ***

@@
 
-\chapter{Getting started with \GLOBES }
+\chapter{\GLOBES\ basics}
 \label{chapt:gettingstarted}
@@
 software package. We demonstrate how to load pre-defined experiments 
-and re-obtain important information about them. In addition, we discuss
-some standard concepts of the \GLOBES\ user interface. We do not go
+and introduce the basic concepts of \GLOBES . We do not go
 into details of the programming language, which means that standard parts
-of the program code common to all examples are, in general, omitted.
+of the program code common to all of the examples in the following chapters are, in general, omitted.
 An example of a minimal \GLOBES\ program in C can be found on page~\pageref{ex:c}.
@@
 
-\begin{table}[t]
+\begin{table}[tbp]
 \begin{center}
@@
 \index{Initialization of {\tt libglobes}}
-\index{glbInit} {\tt void glbInit(char *name)} initializes the library {\tt libglobes} and has
+\GLBNS{glbInit} {\tt void glbInit(char *name)} initializes the library {\tt libglobes} and has
 to be called in the beginning of each \GLOBES\ program. It takes the
@@
 \index{Experiment initialization}
-\index{glbInitExperiment}
+\GLBNS{glbInitExperiment}
 {\tt int glbInitExperiment(char *inf, glb\_exp *in, int *counter)}
@@
 \begin{function}
-\index{glbClearExperimentList}
-{\tt void ClearExperimentList()} removes all experiments from the internal list and resets all counters.   
+\GLBNS{glbClearExperimentList}
+{\tt void glbClearExperimentList()} removes all experiments from the internal list and resets all counters.   
 \end{function}
@@
 
-\begin{table}[t]
+\begin{table}[tb]
 \begin{center}
@@
  reactor power in gigawatts (reactor experiments).
-
-WW: WEITER FROM HERE:
-
-Since the pre-defined experiments in \tabl{experiments} are given for one year running time, specific target masses, and specific source powers, it is useful to change these parameters of the individual experiments:
+%
+Since the pre-defined experiments in \tabl{experiments} are given for specfific luminosities, it is useful to change these parameters of the
+individual experiments:
 \begin{function}
-\GLB{SetRunningTime}$(N_{\mathrm{exp}},t_{\mathrm{run}})$ sets the running time of experiment number $N_{\mathrm{exp}}$ to $t_{\mathrm{run}}$ years.
+\GLBNS{glbSetSourcePower}
+{\tt void glbSetSourcePower(double power, int exp, int fluxno)}
+ sets the source power of experiment number {\tt exp} and flux number
+ {\tt fluxno} to {\tt power}. The definition of the source power depends on the experiment type as described above.
 \end{function}
 \begin{function}
-\GLB{SetTargetMass}$(N_{\mathrm{exp}},m_{\mathrm{Det}})$ sets the fiducial mass of experiment number $N_{\mathrm{exp}}$ to $m_{\mathrm{Det}}$ kilotons.
+\GLBNS{glbSetRunningTime}
+{\tt void glbSetRunningTime(double time, int exp, int fluxno)}
+ sets the running time of experiment number {\tt exp} and flux number
+ {\tt fluxno} to {\tt time} years.
 \end{function}
 \begin{function}
-\GLB{SetSourcePower}$(N_{\mathrm{exp}},P_{\mathrm{Source}})$ sets the source power of experiment number $N_{\mathrm{exp}}$ to $P_{\mathrm{source}}$.
-The definition of the source power depends on the experiment type: ...
-(MISSING).
+ \GLBNS{glbSetTargetMass}
+{\tt void glbSetTargetMass(double mass,int exp)}
+sets the fiducial detector mass of experiment number {\tt exp} to {\tt mass} tons or kilotons (depending on the experiment definition).
 \end{function}
-Thus, these functions also demonstrate how to use the assigned experiment number.
+Thus, these functions also demonstrate how to use the assigned experiment number and others.
+These numbers run from $0$ to the number of experiments-1, fluxes-1, \etc, where the individual elements are numbered in the order of their appearance.
+ Note that the source power and running time are quantities defined
+together with the neutrino flux, whereas the target mass scales the whole
+experiment.
 
-A useful function to re-obtain the information about the initialized experiments is the function {\tt InfoExperiment}:
+Before we can set the simulated event rates or access any oscillation
+parameters, we need to become familiar with the concept \GLOBES\ uses
+for oscillation parameters. In order to transfer sets of oscillation
+parameter vectors $(\theta_{12}, \theta_{13}, \theta_{23}, \deltacp, \sdm, \ldm)$, the parameter type \GLB{glb\_params\_type} is used.
+In general, pointers to this structure are of the type \GLB{glb\_params}, and are often
+transferred to and from \GLOBES\ functions. Therefore, the memory space
+for these vectors has to be reserved before they can be used, and it has
+to be returned afterwards. \GLOBES\ functions usually use the pointers
+of the type {\tt glb\_params} for the input or output to the functions. As
+an input parameter, the pointer has to be created and point towards a valid
+parameter structure, where the oscillation parameters are read from. As
+an output parameter, the pointer has to be created, too, and point towards
+a structure which will contain the return values will be written to.
+This parameter transfer concept seems to be very sophisticated, but, as we
+will see in the next chapters, it hides a lot of complicated parameter 
+mappings which otherwise need to be done by the user. For example, not
+only the oscillation parameters are stored in the pointer structure, but also information on the
+densities of all of the initialized experiments. Since \GLOBES\ treats
+the matter density as a free parameter known with some external precision
+to include matter density uncertainties,
+the minimizers also use fit values and external errors for the matter
+densities of all loaded experiments. More precisely, the matter density
+profile of each experiment $i$ is multiplied by a scaling factor
+$\hat{\rho_i}$, \index{Density scaling factor} 
+which is stored in the density information of {\tt glb\_params}. Each of
+these scaling factors has $1.0$ as pre-defined value. Since it
+is in most cases not necessary to change this value, the user does
+not need to take care of these values.\footnote{For more information,
+we refer to the following chapters.}
+Note that the {\tt glb\_params}
+structures must not be initialized before all experiments are loaded,
+since the number of matter densities can only be determined after the
+experiments are initialized. 
+Similarly, any change in the number of experiments requires that the
+parameter structures be re-initialized, \ie, destroyed and created again.
+Another piece of information returned from the minimizers and transferred
+into the {\tt glb\_params} structure is the number
+of iterations used for the minimization, which is proportional to the
+running time of the minimizer. In general, the user does not need to
+access the elements in {\tt glb\_params} directly. A number of functions
+is provided to handle these parameter structures:
 \begin{function}
-\GLB{InfoExperiment}$()$ prints a list of the initialized functions with their experiment numbers and their most important parameters to the standard output.
+\GLBNS{glbAllocParams}
+{\tt glb\_params glbAllocParams()} allocates the memory space
+needed for a parameter vector and returns a pointer to it.
 \end{function}
-Especially, after changing individual parameters, such as baseline or target mass, this information can be useful to check the changes.
-Another useful function is {\tt ShowChannels}, which prints the initialized oscillation channels for a specified experiment:
 \begin{function}
-\GLB{ShowChannels}$(N_{\mathrm{exp}})$ prints the information about the oscillation channels of the experiment with the number $N_{\mathrm{exp}}$ to the standard output.
+\GLBNS{glbAllocParams}
+{\tt void glbFreeParams(glb\_params stale)} destroys the memory space
+needed for a parameter vector {\tt stale} and sets the pointer to {\tt NULL}.
 \end{function}
+\begin{function}
+\GLBNS{glbDefineParams}
+{\tt glb\_params glbDefineParams(glb\_params in, double theta12, double theta13,double theta23, double delta, double dms, double dma)} assigns
+the complete set of oscillation parameters to the vector {\tt in}, which has
+to be allocated before. The return value is the pointer to {\tt in} if 
+the assignment was successful, and {\tt NULL} otherwise.
+\end{function}
+\begin{function}
+ \GLBNS{glbCopyParams}
+ {\tt glb\_params glbCopyParams(const glb\_params source, glb\_params dest)}
+ copies the vector {\tt source} to the vector {\tt destination}. The return
+ value is {\tt NULL} if the assignment was not successful.
+\end{function}
+\begin{function}
+\GLBNS{glbPrintParams}
+{\tt void glbPrintParams(FILE *stream, const glb\_params in)} prints the
+parameters in {\tt in} to the file {\tt stream}. The oscillation
+parameters, all density values, and the number of iterations are
+printed as pretty output.
+\end{function}
+In addition to these basic functions, there are functions to access
+the individual parameters within the parameter vectors:
+\begin{function}
+\GLBNS{glbSetOscParams}
+{\tt glb\_params glbSetOscParams(glb\_params in, double osc, int which)}
+sets the oscillation parameter {\tt which} in the structure {\tt in}
+to the value {\tt osc}. If the assignment was unsuccessful, the 
+function returns {\tt NULL}.
+\end{function}
+\begin{function}
+\GLBNS{glbGetOscParams}
+{\tt double glbGetOscParams(glb\_params in, int which)}
+returns the value of the oscillation parameter {\tt which} in the 
+structure {\tt in}.
+\end{function}
+In both of these functions, the parameter {\tt which} runs from
+$0$ to $5$, where the parameters in \GLOBES\ always have the order $\theta_{12}$, $\theta_{13}$, $\theta_{23}$, $\deltacp$, $\sdm$, $\ldm$.
+Alternatively to the number, the constants \GLB{GLB\_THETA\_12}, \GLB{GLB\_THETA\_13}, \GLB{GLB\_THETA\_23}, \GLB{GLB\_DELTA\_CP}, 
+\GLB{GLB\_DM\_SOL}, or \GLB{GLB\_DM\_ATM} can be used.
 
-Compared to an existing experiment, which uses real data, a future experiment uses simulated data.
-Thus, the {\em true parameter values} and their results in form of the reference rate vectors are simulated.
+Similarly, the density parameters or iterations can be accessed:
+\begin{function}
+\GLBNS{glbSetDensityParams}
+{\tt glb\_params glbSetDensityParams(glb\_params in, double dens, int which)} sets the density parameter {\tt which} in the structure {\tt in}
+to the value {\tt dens}. If the assignment was unsuccessful, the 
+function returns {\tt NULL}.
+\end{function}
+\begin{function}
+\GLBNS{glbGetDensityParams}
+{\tt double glbGetDensityParams(glb\_params in, int which)}
+returns the value of the density parameter {\tt which} in the 
+structure {\tt in}.
+\end{function}
+\begin{function}
+\GLBNS{glbSetIteration}
+{\tt glb\_params glbSetIteration(glb\_params in, int iter)} sets the number of iterations in the structure {\tt in}
+to the value {\tt iter}. If the assignment was unsuccessful, the 
+function returns {\tt NULL}.
+\end{function}
+\begin{function}
+\GLBNS{glbGetIteration}
+{\tt int glbGetIteration(glb\_params in)}
+returns the value of the number of iterations in the 
+structure {\tt in}.
+\end{function}
+In total, the parameter vector handling in a program normally has the
+following order:
+\begin{quote}
+{\tt
+glbInitExperiment(...); \\
+/* ... more initializations ... */ \\
+\\
+glb\_params vector1 = glbAllocParams(); \\
+/* ... more vectors allocated ... */ \\
+\\
+/* Program code: assign and use vectors */ \\
+\\
+glbFreeParams(vector1); \\
+/* ... more vectors destroyed ... */ \\
+\\
+/* ... end of program or glbClearExperimentList ... */ 
+}
+\end{quote}
+Compared to existing experiments, which use real data, future experiments
+uses simulated data.
+Thus, the {\em true parameter values} and their results in form of the reference event rate vectors are simulated.
 After setting the true parameter values, the {\em fit parameter values} can be varied in order to obtain information on the measurement performance for the given set of true parameter values.
 Therefore, it is often useful to show the results of a future measurement as function of the true parameter values for which the reference rate vectors are computed -- at least within the currently allowed ranges.
-The true parameter values for the vacuum neutrino oscillation parameters have to be set by the functions {\tt SetVacuumParameters} and {\tt SetRates} {\em before} any evaluation function is used and {\em after} the experiments have been initialized and the experiment parameters have been adjusted which could change the rates (such as baseline or target mass).
-Any matter effects are then included automatically depending on the experiment definitions.
+The true parameter values for the vacuum neutrino oscillation parameters have to be set by the functions \GLB{glbSetOscillationParameters} and \GLB{glbSetRates} {\em before} any evaluation function is used and {\em after} the experiments have been initialized and the experiment parameters have been adjusted which could change the rates (such as baseline or target mass).
+This means that
+after any change of an experiment parameter, {\tt glbSetRates} has to be called.
+Matter effects are automatically included as specified in the experiment definition.
+We have the following functions to assign and
+read out the vacuum oscillation parameters:
 \begin{function}
-\GLB{SetVacuumParameters}$(\{\theta_{12}, \theta_{13}, \theta_{23}, \deltacp , \sdm , \ldm \})$ sets the neutrino oscillation parameters to be used to compute the reference rate vector in vacuum.
+\GLBNS{glbSetOscillationParameters}
+{\tt int glbSetOscillationParameters(const glb\_params in)} sets the
+vacuum oscillation parameters to the ones in the vector {\tt in}.
 \end{function}
 \begin{function}
-\GLB{SetRates}$( )$ computes the reference rate vector for the neutrino oscillation parameters set with {\tt SetVacuumParameters}. 
+\GLBNS{glbGetOscillationParameters}
+{\tt int glbGetOscillationParameters(glb\_params out)} returns the
+vacuum oscillation parameters in the vector {\tt out}. The result of
+the function is $0$ if the call was successful.
 \end{function}
-Finally, an initialization sequence for \GLOBES\ could look like this:
-\begin{quote}
-{\tt
-ClearExp(); \\
-InitExperiment("JHFHK.exp"); \\
-SetRunningTime(0, 2.0); \\
-InitExperiment("JHFHKanti.exp"); \\
-SetRunningTime(1, 6.0);\\
-MInfoExperiment(); \\
-SetVacuumParameters(\{0.55, 0.16, 3.14/4, 3.14/2, 7e-5, 2e-3\}); \\
-SetRates();
-} 
-\end{quote}
-This piece of code initializes the JHF (J-PARC) to Hyper-Kamiokande superbeam upgrade with two years of neutrino running and six years of antineutrino running, \ie, an overall running time of eight years.
-The final configuration is then printed to the standard output and the reference rate vector is set to the chosen parameter values.
+The reference rate vector is then computed with:
+\begin{function}
+\GLBNS{glbSetRates}
+{\tt void glbSetRates()} computes the reference rate vector for the neutrino oscillation parameters set by {\tt glbSetOscillationParameters}. 
+\end{function}
+A complete example for a minimal \GLOBES\ program can be found on
+Page~\pageref{ex:c}.
 
@@
 
+\chapter{Changing experiment parameters at running time}
+
 \example{The impact of systematics, correlations, and degeneracies}{
@@
 
-\chapter{Changing experiment parameters at running time}
-
 \bi

### 2004-07-02 Documentation/part1.tex
*** empty log message ***

@@
 %
-Since the pre-defined experiments in \tabl{experiments} are given for specfific luminosities, it is useful to change these parameters of the
-individual experiments:
+Since the pre-defined experiments in \tabl{experiments} are given for specfific luminosities, it is useful to read out and change these
+ parameters of the individual experiments:
 \begin{function}
@@
  {\tt fluxno} to {\tt power}. The definition of the source power depends on the experiment type as described above.
+\end{function}
+\begin{function}
+\GLBNS{glbGetSourcePower}
+{\tt double glbGetSourcePower(int exp, int fluxno)}
+ returns the source power of experiment number {\tt exp} and flux number
+ {\tt fluxno}. 
 \end{function}
@@
 \begin{function}
+\GLBNS{glbGetRunningTime}
+{\tt double glbGetRunningTime(int exp, int fluxno)}
+ returns the running time of experiment number {\tt exp} and flux number
+ {\tt fluxno}.
+\end{function}
+\begin{function}
  \GLBNS{glbSetTargetMass}
@@
 sets the fiducial detector mass of experiment number {\tt exp} to {\tt mass} tons or kilotons (depending on the experiment definition).
+\end{function}
+\begin{function}
+ \GLBNS{glbGetTargetMass}
+{\tt double glbGetTargetMass(int exp)}
+returns the fiducial detector mass of experiment number {\tt exp}.
 \end{function}
@@
 parameters, all density values, and the number of iterations are
-printed as pretty output.
+printed as pretty output. Use {\tt stdout} for {\tt stream} if you want
+to print to the screen.
 \end{function}

### 2004-07-02 Documentation/example2.c

@@ -85,10 +85,10 @@ int main(int argc, char *argv[])
   {
       /* Set vector of test=fit values */
       thetheta13=asin(sqrt(pow(10,x)))/2;
-      glbSetOscParams(test_values,thetheta13,1);
+      glbSetOscParams(test_values,thetheta13,GLB_THETA_13);
      
       /* Guess fit value for deltacp in order to safely find minimum */
-      glbSetOscParams(test_values,200.0/2*(x+4)*M_PI/180,3);
+      glbSetOscParams(test_values,200.0/2*(x+4)*M_PI/180,GLB_DELTA_CP);
  
       /* Compute Chi^2 for two-parameter correlation: minimize over deltacp only */
       res1=glbChiNP(test_values,NULL,GLB_ALL);

### 2004-07-05 Documentation/part1.tex
Prior handling updated

@@
 
+\section{Initialization of \GLOBES}
+
 Before one can use \GLOBES , one has to initialize the \GLOBES\
@@
 
+\section{Units in \GLOBES\ and the integrated luminosity}
+
 While interacting with the user interface of \GLOBES , parameters are transferred to and from the \GLOBES\ library. In \GLOBES , one set of units 
@@
 
+\section{Handling oscillation parameter vectors}
+
 Before we can set the simulated event rates or access any oscillation
@@
 mappings which otherwise need to be done by the user. For example, not
-only the oscillation parameters are stored in the pointer structure, but also information on the
+only the oscillation parameters are stored in the pointer structure, but also information on the matter
 densities of all of the initialized experiments. Since \GLOBES\ treats
@@
 profile of each experiment $i$ is multiplied by a scaling factor
-$\hat{\rho_i}$, \index{Density scaling factor} 
-which is stored in the density information of {\tt glb\_params}. Each of
+$\hat{\rho_i}$, \index{Matter density scaling factor} 
+which is stored in the density information of {\tt glb\_params}. 
+Each of
 these scaling factors has $1.0$ as pre-defined value. Since it
 is in most cases not necessary to change this value, the user does
-not need to take care of these values.\footnote{For more information,
-we refer to the following chapters.}
+not need to take care of it.
+For a constant matter density, it is simply the ratio of the matter density and the average matter density specified in the experiment definition, \ie , $\hat{\rho_i} \equiv \rho_i/\bar{\rho}_i$.
+For a matter density profile, it acts as an overall normalization factor: The matter density in each layer is multiplied by this factor.
+In most cases one wants to take a scaling factor of $1.0$ here, which simply means taking the matter density profile as it is given in the experiment definition.
+For the treatment of correlations, however,
+an external precision of the scaling factor might be used to include the
+correlations with the matter density uncertainty. For more details, we
+refer to the following sections.
 Note that the {\tt glb\_params}
@@
 parameter structures be re-initialized, \ie, destroyed and created again.
-Another piece of information returned from the minimizers and transferred
+
+Another piece of information will be returned from the minimizers 
+(\cf, \Chapt~\ref{chapt:correlations}) and transferred
 into the {\tt glb\_params} structure is the number
@@
 \end{quote}
+
+\section{Computing the simulated data}
+
 Compared to existing experiments, which use real data, future experiments
@@
 Calculating a $\chi^2$-value with or without systematics, but no correlations and degeneracies, is the simplest and fastest possibility to obtain high-level information on an experiment.
-I general, \GLOBES\ uses the six independent oscillation parameters $\theta_{12}$, $\theta_{13}$, $\theta_{23}$, $\deltacp$, $\sdm$, $\ldm$, as well as the matter density of each experiment.
-Thus, there are six plus the number of experiments parameters determining the rate vectors.
-Using the matter densities in addition to the oscillation parameters will allow the simulation of matter density uncertainties: In this approach, the matter densities can be treated as parameters to be measured by the experiments within certain limits, where the limits are given by the observed precision of the matter density profile.
-Defining a independent matter density for each experiment is necessary if the baselines are completely uncorrelated.
-For correlated or even identical baselines, one may rather want to encapsulate the experiments into a single experiment definition.
+In general, \GLOBES\ uses the six independent oscillation parameters $\theta_{12}$, $\theta_{13}$, $\theta_{23}$, $\deltacp$, $\sdm$, $\ldm$, as well as the matter density
+scaling factor $\hat{\rho}$ of each experiment.\index{Matter density
+scaling factor} Thus, there are six plus the number of experiments parameters determining the rate vectors.
+Using the matter density scaling factors  in addition to the oscillation parameters will allow the simulation of the correlations with matter density uncertainties: In this
+approach, the matter density profile normalization $\hat{\rho}$ can 
+be treated
+as parameter to be measured by the experiment, where an external precision
+given by observations is imposed (typically up to $5\%$). 
+ For this section, it is important to keep in mind
+that there are more parameters than just the oscillation parameters
+determining the simple $\chi^2$. However, as we have described in the
+last section, the mechanism for the matter density scaling factors
+is hidden in the defintion of \GLB{glb\_params}: Each of the scaling
+factors is initially set to $1.0$. Therefore, for the calculation of
+$\chi^2$ with systematics only, we do not have to care about the
+matter density scaling factors.
 
@@
 A typical and fast application for \GLB{glbChiSys} is the visualization of
- two-parameter correlations. For example, to calculate the two-parameter 
+ two-parameter correlations using systematics only. For example, to
+  calculate the two-parameter 
  correlation between $\stheta$ and $\deltacp$ at a neutrino factory, one
@@
 
+\vspace*{-0.4cm}
+
 The resulting data can then be plotted as a contour plot (2 d.o.f.):
@@
 
-Keeping all oscillation parameters and matter densities fixed, one can use the following functions to obtain the total $\chi^2$ of all specified oscillation channels including systematics:
+Keeping all oscillation parameters and matter density scaling factors fixed,
+ one can use the following functions to obtain the total $\chi^2$ of all 
+ specified oscillation channels including systematics:
 \begin{function} 
-\GLB{Chi}$(\{ \theta_{12}, \theta_{13}, \theta_{23}, \deltacp , \sdm , \ldm, \hat{\rho_1}, \hdots , \hat{\rho_n} \})$  returns the total added $\chi^2$ of all loaded experiments.
+\GLBNS{glbChiSys}
+{\tt double glbChiSys(const glb\_params in,int exp, int rule)} returns
+the $\chi^2$ for the (fixed) oscillation parameters {\tt in}, the
+experiment number {\tt exp}, and the rule number {\tt rule}. For all
+experiments or rules, use \GLB{GLB\_ALL} as parameter value.
 \end{function}
-\begin{function}
- \GLB{SingleChi}$(\{ \theta_{12}, \theta_{13}, \theta_{23}, \deltacp , \sdm , \ldm, \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}} )$  returns the total $\chi^2$ of the experiment number $N_{\mathrm{exp}}$.
-\end{function}
-Note that the result of {\tt Chi} corresponds to the sum of all of the {\tt SingleChi}'s of the loaded experiment.
+Note that the result of {\tt glbChiSys} for all experiments or rules
+corresponds to the sum of all of the individual {\tt glbChiSys} calls. 
 This equality will not hold for the minimizors in the next sections anymore. 
-In both functions, one has to give the matter density scaling factor $\hat{\rho_i}$ for each of the used experiments.
-The effect of this scaling factor depends on the type of the matter density profile, which is given in the  experiment definition.
-For a constant matter density, it is simply the ratio of the matter density and the average matter density specified in the experiment definition, \ie , $\hat{\rho_i} \equiv \rho_i/\bar{\rho}_i$.\index{Matter scaling factor} For a matter density profile, it is an overall scaling factor: The matter density in each layer is multiplied by this factor.
-In most cases one wants to take a scaling factor of $1.0$ here, which simply means taking the matter density profile as it is given in the experiment definition.
-Moreover, note that its effect is in general small for short baselines.
-An example of how to use  {\tt Chi} (or {\tt SingleChi}) can be found on page~\pageref{ex:corrth13dcp}.
+ An example how to use  {\tt glbChiSys} can be found on page~\pageref{ex:corrth13dcp}.
 
 \index{Systematics}
-The treatment of systematics is done by the usage of auxiliary systematics parameters, which are taken completely uncorrelated among different oscillation channels, and treated with simple Gaussian statistics.
-One such example is the signal normalization error, \ie, an error to the overall normalization of the signal.
+The treatment of systematics in \GLOBES\ is performed by the so-call
+{\em pull method} with the help of auxiliary systematics parameters.
+They are taken completely uncorrelated among different rules, and treated with simple Gaussian statistics. In general, a rule is a set 
+of signal and background event rates coming from different oscillation
+channels, where the event rates of all rule contributions are added.
+For more details of the rule concept, see \Part~\ref{part:2} of this manual,
+and for the treatment of systematics, see \Sec~\ref{sec:rules}.
+
+ One example for a systematics parameter the signal normalization error, \ie, an error to the overall normalization of the signal.
 For illustration, we assume that the signal event rate in the $i$th bin $s_i^0$ of one oscillation channel is altered by the overall normalization auxiliary parameter\index{Auxiliary parameter} of this channel, \ie ,
@@
 \ee 
-This minimization is done independently for all auxiliary parameters of this oscillation channel.
-The total $\chi^2$ for the considered experiment is finally obtained by repreating this procedure for all oscillation channels and adding their $\chi^2$-values.
+This minimization is done independently for all auxiliary parameters of the rule.
+The total $\chi^2$ for the considered experiment is finally obtained by repreating this procedure for all rules and adding their $\chi^2$-values.
 In general, the situation is more complicated because of the usage of many systematical errors.
-More details about systematics parameters and the definition of signal, background, and oscillation channels can be found in the \EDM\ part of this book, too.
+More details about systematics parameters and the definition of signal, background, and oscillation channels can be found in \Sec~\ref{sec:rules}, too.
 
-The systematics minimization of an experiment can be easily switched on and off, \ie, one can also compute the $\chi^2$ without even taking into account systematics:
-\begin{function}
-\GLB{SetSystematics}$(N_{mathrm{exp}}, \, sys)$ switches the systematics minimization for experiment $N_{\mathrm{exp}}$ on (sys=1) or off (sys=0).
-\end{function}
-This function can be especially useful for the test of the impact of systematics.
-??? Wie einzelne Systematik-Parameter ein und ausschalten? Sonst individuell nicht testbar! ??? Verweis auf appendix?
+The systematics minimization of an experiment can be easily switched on and off with \GLB{glbSwitchSystematics}, \ie, one can also compute the $\chi^2$ with statistics only.
+In addition, several options for
+systematics are available, such as only using total event rates without
+spectral information. For details, we refer to \Chapt~\ref{chapt:running}.
 
 \chapter[Calculating $\chi^2$-projections: how one can include correlations]{Calculating $\boldsymbol{\chi^2}$-projections: how one can include correlations}
+\label{chapt:correlations}
 
 \index{Multi-parameter correlation}
-This chapter deals with the rather complicated issue of $n$-parameter correlations.
-Since it has before this software  not been possible to include the full $n$-parameter correlations in the high-dimensional parameter space with reasonable effort, it is the core part of this software  -- as well as its strength.
+This chapter deals with the rather complicated issue of $n$-parameter correlations. It is one of the greatest strengths of this software 
+to include the full $n$-parameter correlation in the high-dimensional parameter space with reasonable effort.
 Of course, calculating $\chi^2$-projections is somewhat more complicated than using systematics only.
@@
 \index{Projection of manifold}
-In principle, the precision of an individual parameter measurement including correlations can be obtained as the projection of the $n$-dimensional fit manifold onto the respective axis.
-Similarly, one can project the fit manifold onto a plane, such as the $\stheta$-$\deltacp$-plane, if one wants to explicitely show this correlation with all the other parameter correlations included.
-In practice, this projection is very difficult: a grid-based method would need $(N_{\mathrm{grid}})^n$ function calls of {\tt Chi} or {\tt SingleChi} to calculate the precision including the full $n$-parameter correlation, where $N_{\mathrm{grid}}$ is the number of points in each direction of the lattice.
-For example, taking only $N_{\mathrm{grid}}=20$ and $n=7$ (six oscillation parameters and matter density) would mean more than one billion function calls of {\tt Chi} or {\tt SingleChi}.
+In principle, the precision of an individual parameter measurement including correlations in the $\chi^2$-approach can be obtained as the projection of the $n$-dimensional fit manifold onto the respective axis.
+Similarly, one can project the fit manifold onto a plane, such as the $\stheta$-$\deltacp$-plane, if one wants to explicitely show the allowed
+region in this plane with all the other parameter correlations included. 
+In practice, this projection is very difficult: a grid-based method would need $(N_{\mathrm{grid}})^n$ function calls of {\tt glbChiSys} to calculate the precision including the full $n$-parameter correlation, where $N_{\mathrm{grid}}$ is the number of points in each direction of the lattice.
+For example, taking only $N_{\mathrm{grid}}=20$ and $n=7$ (six oscillation parameters and matter density) would mean more than one billion function calls of {\tt glbChiSys}.
 One can easily image that this is too much for any sophisticated application.
@@
 \index{Minimizer}
-The solution to this problem is using a local $n$-dimensional minimizer instead of a grid-based method.
-It turns out that such a minimizer can include a full $6$-parameter correlation with of the order of $1\, 000$ function calls of {\tt Chi} or {\tt ChiNew}.
-It is a standard method which can be found in every good book for standard numerical calculation routines.
+The solution to this problem is using a local $n$-dimensional (local) minimizer for the procection instead of a grid-based method, where we will
+illustrate this minimization process later. It turns out
+that such a minimizer can include a full $6$-parameter correlation with of the order of $1\, 000$ function calls of {\tt glbChiSys}.
+It is a standard method which can be found in every good book for numerical calculation routines.
 Thus, for each point on the projection axis/plane, one can obtain a result within about $10$ to $30$ seconds on a modern computer, which means that the complete measurement precision for one fixed true parameter set can be obtained in as much as $10$ to $15$ minutes.
@@
 
-This approach also has a major disadvantage: One can not simply program a robust grid-based code and let it run, since using a local minimizer always means that one may end up in an unwanted local minimum and not in the investigated one.
+This approach also has one major disadvantage: One can not simply program a robust grid-based code and let it run, since using a local minimizer always means that one may end up in an unwanted local minimum and not in the investigated global one.
 Thus, one has to use some (analytical or numerical) knowledge on the topology of the fit manifold and start the local minimizer close enough to the investigated solution.
 Fortunately, this can be done quite straightfoward in most cases, since the structure of the neutrino oscillation formulas does not cause very complicated topologies of the fit manifolds.
-Especially, the are plenty of analytical discussions of this issue, which means that one can implicitely use this knowledge to obtain better predictions for the measurement performances.
-Note that, since one can easily find the global fit minimum at the best-fit values, any solution found with the local minimizer makes the measurement performance worse.
-Thus, one can only run the danger to obtain a too optimistic solution if one does not find the other local minima below the chosen confidence level.
+Especially the simulation of reactor experiments and conventional beams or superbeams is rather simple with purely numerical
+approaches. Neutrino factories have, especially for small values of
+$\theta_{13}$, a much more complicated topology. In this case, results
+of the many analytical discussions of this issue can be used. This means
+ that one can implicitely use the analytical knowledge to obtain better predictions for the measurement performances, since the minimizer has
+ to be started only close enough to the exact solution by an ``educated
+ guess''. One can easily imagine that the used methods then also depend
+ on the region of the parameter space. In this manual, we only use
+ examples with a neutrino factory, since some of these complications
+ can be illustrated there.
+
+Now, some words of warning with respect to the minimization process: 
+Since one can easily find the global fit minimum at the best-fit/simulated values, any additional solution found by the local minimizer makes the measurement performance worse.
+However, for the real result, the complete
+topology has to be taken into account, which means that {\em all} results
+fitting the simulated data have to be included, and not only the best-fit
+point.
+Thus, one can only run the danger to obtain a too optimistic solution if one does not find the other local minima appearing below the chosen confidence level.
+Thus, with this approach and proper usage, it should not possible to produce a too pessimistic solution.
+However, if one is not careful enough to locate
+all local minima, one can easily produce too optimistic solutions.
+This danger can be summarized as follows:
+\begin{equation}
+\mathrm{Too \, pessimistic \, result} \, < \, \underbrace{\mathrm{Real \, result}}_{\mathrm{Located \, by \, careful \, usage}} \, \le \, \mathrm{GLoBES \, result} \,  < \,  \mathrm{Too \, optimisitic \, result} \nonumber
+\end{equation}
 
@@
 Thus, using the projection method on the axis/plane of interest is a reasonable approach.
-The inclusion of external input in \GLOBES\ is done by the use of Gaussian {\em priors}: We assume that an external measurement has determined the measured parameter to be at the central value (called {\em starting value}) with a $1 \sigma$ Gaussian error (called {\em input error}).
+The inclusion of external input in \GLOBES\ is done by the use of Gaussian {\em priors}: We assume that an external measurement has determined the measured parameter to be at the central value (which we call {\em starting value}) with a $1 \sigma$ Gaussian error (which we call {\em input error}).
 The explicit definition of these priors will be shown in the next section.
@@
 \chapter{Changing experiment parameters at running time}
+\label{chapt:running}
 

### 2004-07-05 Documentation/example-tour.c

@@ -1,0 +1,254 @@
+/* GLoBES -- General LOng Baseline Experiment Simulator
+ * (C) 2002 - 2004,  The GLoBES Team
+ *
+ * GLoBES is mainly intended for academic purposes. Proper
+ * credit must be given if you use GLoBES or parts of it. Please
+ * read the section 'Credit' in the README file.
+ *
+ * This program is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU General Public License as published by
+ * the Free Software Foundation; either version 2 of the License, or
+ * (at your option) any later version.
+ *
+ * This program is distributed in the hope that it will be useful,
+ * but WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ * GNU General Public License for more details.
+ *
+ * You should have received a copy of the GNU General Public License
+ * along with this program; if not, write to the Free Software
+ * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
+ */
+ 
+ /* 
+ * Example: GLoBES tour
+ * Compile with ``make example-tour''
+ */ 
+
+#include <stdio.h>
+#include <stdlib.h>
+#include <math.h>
+#include <string.h>
+
+#include <globes/globes.h>   /* GLoBES library */
+
+int main(int argc, char *argv[])
+{ 
+  char* MYFILE="";
+  /* char* MYFILE="gl-tour.dat"; /* if empty, write to screen */
+  FILE* stream;
+  
+  if(strlen(MYFILE)>0) stream=fopen(MYFILE, "w");
+  else stream = stdout;
+  
+  /* Initialize libglobes */
+  glbInit(argv[0]); 
+
+  /* Define my standard oscillation parameters */
+  double theta12 = asin(sqrt(0.8))/2;
+  double theta13 = asin(sqrt(0.001))/2;
+  double theta23 = M_PI/4;
+  double deltacp = M_PI/2;
+  double sdm = 7e-5;
+  double ldm = 2e-3;
+ 
+  /* Initialize one experiment NuFact.glb */
+  glbInitExperiment("NuFact.glb",&glb_experiment_list[0],&glb_num_of_exps); 
+   
[…]

### 2004-07-05 Documentation/example2.c

@@ -69,7 +69,7 @@ int main(int argc, char *argv[])
   glbSetRates();
 
   /* Set starting values and input errors for all projections */  
-  glbDefineParams(input_errors,theta12*0.1,10,10,10,sdm*0.1,ldm);  
+  glbDefineParams(input_errors,theta12*0.1,0,0,0,sdm*0.1,0);  
   glbSetDensityParams(input_errors,0.05,GLB_ALL);
   glbSetStartingValues(true_values);
   glbSetInputErrors(input_errors);

### 2004-07-05 Documentation/example3.c

@@ -68,7 +68,7 @@ int main(int argc, char *argv[])
 
   /* Find sgn-degeneracy */  
   glbDefineParams(starting_values,theta12,theta13,theta23,deltacp,sdm,-ldm);  
-  glbDefineParams(input_errors,theta12*0.1,10,10,10,sdm*0.1,ldm/3);  
+  glbDefineParams(input_errors,theta12*0.1,0,0,0,sdm*0.1,ldm/3);  
   glbSetDensityParams(input_errors,0.05,GLB_ALL);
   glbSetStartingValues(starting_values);
   glbSetInputErrors(input_errors);

### 2004-07-05 Documentation/example4.c

@@ -94,7 +94,7 @@ double CalcProjection(double theldm,double thex,glb_params start_vector)
   glb_params starting_values = glbAllocParams();
   glb_params minimum = glbAllocParams();
 
-  glbDefineParams(input_errors,theta12*0.1,10,10,10,sdm*0.1,theldm/3);  
+  glbDefineParams(input_errors,theta12*0.1,0,0,0,sdm*0.1,theldm/3);  
   glbSetDensityParams(input_errors,0.05,GLB_ALL);
   /* Set starting values to respective +-dm31^2 to avoid falling into unwanted solution;
      Note that the error on dm31^2 should not be too small in order to avoid large
@@ -127,7 +127,7 @@ double FindDeg(glb_params deg_pos,double theldm)
   glb_params starting_values = glbAllocParams();
 
   glbDefineParams(starting_values,theta12,theta13,theta23,deltacp,sdm,-theldm);  
-  glbDefineParams(input_errors,theta12*0.1,10,10,10,sdm*0.1,theldm/3);  
+  glbDefineParams(input_errors,theta12*0.1,0,0,0,sdm*0.1,theldm/3);  
   glbSetDensityParams(input_errors,0.05,GLB_ALL);
   glbSetStartingValues(starting_values);
   glbSetInputErrors(input_errors);

### 2004-07-05 Documentation/part1.tex
*** empty log message ***

@@
 
-\begin{table}[t]
+\begin{table}[tpb]
 \begin{center}
@@
 to the value {\tt dens}. If the assignment was unsuccessful, the 
-function returns {\tt NULL}.
+function returns {\tt NULL}. If {\tt GLB\_ALL} is used for {\tt which},
+the density parameters of all experiments will be set accordingly.
 \end{function}
@@
 \index{External input}
-It is one of the strengths of the \GLOBES\ software to use external input to reduce the fit manifold with the knowledge from external (eariler) measurements.
-The treatment of external input is done by the addition of Gaussian {\em priors} to the final\footnote{After the systematics minimization and after all $\chi^2$'s are added for all channels} $\chi^2$-function  and the minimization over the respective parameters.
-For example, for the matter density, one obtains as the minimized $\chi^2_F$ after minimzation over $\rho$
+It is one of the strengths of the \GLOBES\ software to use external input in order to reduce the fit manifold with the knowledge from external
+(eariler) measurements.
+The treatment of external input is done by the addition of Gaussian so-called {\em priors} to the systematics-minimized $\chi^2$-function.
+For example, for the matter density, one obtains as the final projected $\chi^2_F$ after minimzation over the matter density
+scaling factor \index{Matter density scaling factor} $\hat{\rho}$
 \be
- \chi^2_F = \underset{\rho}{\mathrm{min}} \left( \chi^2(\rho) + \frac{(\rho - \rho^0)^2}{\sigma_\rho^2} \right).
+ \chi^2_F = \underset{\rho}{\mathrm{min}} \left( \chi^2(\hat{\rho}) + \frac{(\hat{\rho} - \hat{\rho}^0)^2}{\sigma_{\hat{\rho}}^2} \right).
 \label{equ:priors}
 \ee
-In practice, this minimzation is done simultaneously over all priors and free oscillation parameters, but this example serves as a very simple illustration.
-In \equ{priors}, $\rho^0$ is the {\em starting value} of the prior, and $\sigma_\rho$ the $1 \sigma$ absolute {\em input error}.
-Thus, it is assumed that an external measurement has determined the matter density with a precision (input error) $\sigma_\rho$ at the central value $\rho^0$.
-Usually, the starting value corresponds to the best-fit value and the input error to the $1 \sigma$ half width of the external measurement.
-For the matter density, $\rho^0$ can be set to the average matter density $\bar{\rho}$ and $\sigma_\rho$ to the matter density uncertainty.
-Since \GLOBES\ may also use a matter density profile, it actually uses a scaling factor $\hat\rho$ instead of the average matter density (see last section), \ie, $\rho^0=1.0$ is the central value in either case.
-Thus, $\sigma_\rho$ directly correponds to the uncertainty of this scaling factor.
-For instance, using the PREM (``Preliminary Reference Earth Model'') profile, $\rho^0 = 1.0$ and $\sigma_\rho = 0.05$ is a conservative estimate of the PREM profile uncertainty.
-\index{PREM profile}
+This example is a very simple one, since in fact the
+minimzation is simultaneously performed over all priors and free oscillation parameters.
+In \equ{priors}, $\hat{\rho}^0$ is the {\em starting value} of the prior, and $\sigma_{\hat{\rho}}$ the $1 \sigma$ absolute (half width) {\em input error}.
+Thus, it is assumed that an external measurement has determined the matter density with a precision (input error) $\sigma_{\hat{\rho}}$ at the central value $\hat{\rho}^0$.
+Usually, the starting value is fixed at the best-fit value, and the input error to the $1 \sigma$ half width of the external measurement.
+For the matter density, $\hat{\rho}^0$  is usually set to $1.0$ corresponding to
+the actual matter density profile such as given by the experiment
+ definition file, and $\sigma_{\hat{\rho}}$ to the 
+relative matter density uncertainty (\eg, $0.05$ for 5\% uncertainty).
 
 In principle, one can set the priors for the matter density and all oscillation parameters.
-For example, if the disappearance channels of the experiment determine the leading oscillation parameters, once can set very large values for the input errors ($\sigma$'s).
+For example, if the disappearance channels of the experiment determine the leading oscillation parameters with unprecedented
+precisions, one can omit the respective input errors. In \GLOBES , a
+value of $0$ correponds to neglecting the prior.
 If, however, earlier external measurements provide better information, one can set their absolute precisions with the input errors.
-The starting values are usually set to the best-fit values.
-In some cases, it may be necessary to adjust them, such as for $\ldm$ and the negative mass hierarchy to $\rho^0_{\ldm} = - |\ldm|$ if some external precision on $| \ldm |$ is imposed and one is investigating the negative-sign solution.
-In other cases, minor modifications of the starting values can cause a faster convergence of the algorithm.
-In either case, two function have to be called {\em before the usage of any minimizer}:
-\begin{function}
-\GLB{SetStartingValues}$(\{\rho^0_{\theta_{12}}, \, \rho^0_{\theta_{13}}, \, \rho^0_{\theta_{23}}, \, \rho^0_{\deltacp}, \, \rho^0_{\sdm}, \, \rho^0_{\ldm}, \, \rho^0_{\hat\rho} \})$ sets the starting values for all of the following minimizer calls.
-\end{function}
-\begin{function}
-\GLB{SetInputErrors}$(\{\sigma_{\theta_{12}}, \, \sigma_{\theta_{13}}, \, \sigma_{\theta_{23}}, \, \sigma_{\deltacp}, \, \sigma_{\sdm}, \, \sigma_{\ldm}, \, \sigma_{\hat\rho} \})$ sets the input errors for all of the following minimizer calls.
-\end{function}
-Both functions may take one matter density parameter or as many as there are experiments.
-Since the assumptions about the matter density profile are usually similar for the loaded experiments, using only one matter density parameter is a useful tool to apply the same starting value  and input error to all loaded experiments.
-Eventually, a typical initialization of the external input may look like this:
+The starting values are usually chosen to be the best-fit values of this external input, such as
+for the precisions of the solar experiments. In some cases, it may be
+necessary to adjust them, such as for artificial contraints to the
+oscillation parameters. In other cases, minor modifications of the starting values can cause a faster convergence of the algorithm.
+For example, for the investigation of the
+opposite-sign solution, one can use the prior to constrain $\ldm$
+in order to force the minimizer not to  fall into the (unwanted) best-fit solution.
+In this case, the starting value of $\ldm$ would be set to $\rho^0_{\ldm} = - \ldm$, and a $\sigma_{\ldm}$ of the order of $\ldm$ would be imposed.
+For the algorithm, it would then be rather difficult
+to converge into the unwanted best-fit solution.
+However, note that one should in this case check that the actually determined value for $\ldm$ after minimization is close enough to the guessed values $-\ldm$ in order to avoid significant artifical contributions of the priors to the
+final $\chi^2$.
+
+In order to set the starting values and input errors, two function have to
+ be called {\em before the usage of any minimizer}:
+\begin{function}
+\GLBNS{glbSetStartingValues}
+{\tt int glbSetStartingValues(const glb\_params in)} sets the starting values for all of the following minimizer calls to {\tt in}.
+\end{function}
+\begin{function}
+\GLBNS{glbSetInputErrors}
+{\tt int glbSetInputErrors(const glb\_params in)}
+ sets the input errors for all of the following minimizer calls to {\tt in}.
+ An input error of $0$ correponds to not taking into account the
+ respective prior.
+\end{function}
+Accordingly, there are functions to return the actually set starting values
+and input errors:
+\begin{function}
+\GLBNS{glbGetStartingValues}
+{\tt int glbGetStartingValues(glb\_params out)} writes the currently
+set starting values to {\tt out}.
+\end{function}
+\begin{function}
+\GLBNS{glbGetInputErrors}
+{\tt int glbGetInputErrors(glb\_params out)}
+ writes the currently set input errors to {\tt out}.
+\end{function}
+All functions take or return as many matter density parameters as there are initialized experiments. In addition, they return $-1$ if the operation
+was not successful.
+
+Eventually, a typical initialization of the external input with
+$10\%$ external precisions for the solar parameters\footnote{In fact, accelerator-based long-baseline experiments are primarily sensitive to the product $\sin 2 \theta_{12} \cdot \sdm$, which means that these errors effectively add up to an error of this product of about $15\%$.},
+and $5\%$ matter density uncertainties for all experiments looks like this:
 \begin{quote}
 {\tt
-SetStartingValues(\{th12, th13, th23, dcp, sdm, ldm, 1.0\}); \\
-SetInputErrors(\{th12*0.1, 10, 10, 10, sdm*0.1, ldm, 0.05\}); 
+ glbDefineParams(input\_errors,theta12*0.1,0,0,0,sdm*0.1,0);\\  
+ glbSetDensityParams(input\_errors,0.05,GLB\_ALL);\\
+ glbSetStartingValues(true\_values);\\
+ glbSetInputErrors(input\_errors);\\
 }
 \end{quote}
-In this example, the starting values are set to the best-fit values.
-The input errors for $\theta_{13}$, $\theta_{23}$, $\deltacp$, and $\ldm$ are kept large, since the experiment measures these parameters itself.
-However, some precision on $\ldm$ is imposed to avoid ending up in the negative-sign solution.
-The solar parameters $\theta_{12}$ and $\sdm$ are assumed to be known with $10\%$ precision each.\footnote{In fact, accelerator-based long-baseline experiments are only sensitive to the product $\sin 2 \theta_{12} \cdot \sdm$, which means that these errors effectively add up to an error of this product.} Finally, the matter density uncertainties are assumed to have an amplitude of $5\% \,  \bar\rho$.
+In this example, the starting values are set to the best-fit (simulated) values. Remember that initially the matter density scaling factors are 
+all $1.0$, which means that they do not need to be adjusted for the
+starting values.
+
+Though the priors are an elegant way to treat external input, there are also some complications with priors. The following hints are for the
+more advanced \GLOBES\ user:
+\begin{enumerate}
+\item
+ The priors are only added once to the final $\chi^2$, no matter how
+ many experiments there are simulated. This is already one reason
+ (besides the minimization) why the sum of all projected $\chi^2$'s of 
+ the indivual experiments 
+ cannot correspond to the $\chi^2$ of the combination of all experiments.
+\item
+ Priors are not used for parameters which are not minimized over, \ie, 
+ kept fixed. This will be important together with arbitrary projections
+ using \GLB{glbChiNP}. A more subtile consequence is the comparison
+ of fit manifold sections and projections for the solutions where
+ the absolute minimum $\chi^2$ is larger than zero, \ie, degeneracies
+ other than the best-fit solution. In this case, the sections and
+ projections are not  comparable if not corrected by the prior contributions, where the
+ correction can be obtained as the $\chi^2$-difference at the minimum.
+ For example, projecting the $\mathrm{sgn}(\ldm)$-degeneracy onto
+ the $\theta_{13}$-$\deltacp$-plane and comparing it with the section
+ (all other parameters fixed), the section region would in many cases be 
+ larger than the projection region if the priors are not added to the
+ section. At the best-fit solution, this problem usually does not occur
+ because the prior contributions are close to zero.    
+\item
+Currently, \GLOBES\ only supports Gaussian priors for the individual
+oscillation parameters. Especially for the solar parameters, this is only
+ an approximation, since they are imposed on $\theta_{12}$ and not
+ on $\sin 2 \theta_{12}$, $\sin 2 \theta_{12} \cdot \sdm$, or $\sin  
+ \theta_{12}$. Later versions of  \GLOBES\ may include more alternatives.
+ \end{enumerate}
 
@@
   /* Set starting values and input errors for all projections */  \\
-  glbDefineParams(input\_errors,theta12*0.1,10,10,10,sdm*0.1,ldm);\\  
+  glbDefineParams(input\_errors,theta12*0.1,0,0,0,sdm*0.1,0);\\  
   glbSetDensityParams(input\_errors,0.05,GLB\_ALL);\\
@@
 We illustrate this method at the example of the projection of the two-dimensional manifold in the $\stheta$-$\deltacp$-plane onto the $\stheta$-axis in \figu{projex}.
-In this figure, the left-hand plot shows the correlation in the $\stheta$-$\deltacp$-plane computed with {\tt Chi} or {\tt SingleChi}.
-The right-hand plot shows the projection of this two-dimensional manifold onto the $\stheta$ axis by minimizing $\chi^2$ over $\deltacp$.
+In this figure, the left-hand plot shows the correlation in the $\stheta$-$\deltacp$-plane computed with {\tt glbChiSys}.
+The right-hand plot illustrates the projection of this two-dimensional manifold onto the $\stheta$ axis by minimizing $\chi^2$ over $\deltacp$.
 In this simple example, the minimization is done along the vertical gray lines in the left hand plot.
@@
 In general, the full $n$-parameter correlation is treated similarly by the simultaneous (local) minimization over all free fit parameters.
-The example on page~\pageref{ex:corrproj} demonstrates how one can obtain \figu{projex} (right) with keeping all parameters but $\deltacp$ fixed, as well as how one can inlcude the full $n$-parameter correlation with external input.
-It also demonstrates how these two compare to each other.
-
-One mentioned disadvantage of the local minimizer is the possibility to end up in a local minimum.
-Usually it works very well to start the minimizer close to the best-fit values.
-In the example in \figu{projex}, the minimizer is always started at $\deltacp=90^\circ$ and finds its way to the appropriate minimum.
-In some cases, however, the topology may be more difficult and the true minimum is not correctly found.
-A strong indication for such a situation are discontinuities in the projected $\chi^2$-function, where the minimizer jumps from one minimum to the other.
-In such a case, the starting point of the minimizer has to be adjusted to help it find the true minimum.
-In other cases, the convergence of the minimizer may be infinitively small.
-Then it usually helps to start the minimizer somewhat off the best-fit values.
-The following functions are the simplest minimizers provided by \GLOBES :
-\begin{function}
-\GLB{ChiTheta}$( \theta_{13}, \, \{ \theta_{12}, \, \theta_{23}, \, \deltacp , \, \sdm , \, \ldm, \, \hat{\rho_1}, \,  \hdots ,  \,\hat{\rho_n} \})$ returns the projection onto the $\theta_{13}$-axis for all experiments.
-The parameters are the fixed value of $\theta_{13}$ and the starting point of the minimizer including all starting values for the matter density scaling factors (without $\theta_{13}$, which is fixed).
-The return value is a list $\{ \chi^2, \, \theta_{12}, \, \theta_{23}, \, \deltacp , \, \sdm , \, \ldm, \, \hat{\rho_1}, \, \hdots , \, \hat{\rho_n} , \, N_{\mathrm{Iter}} \}$ with the minimum $\chi^2$ found, the coordinates of the local minimum (without $\theta_{13}$), and the number of iterations $N_{\mathrm{Iter}}$ used by the minimizer (number of function calls of {\tt Chi}).
-\end{function}
-\begin{function}
-\GLB{SingleChiTheta}$( \theta_{13}, \, \{ \theta_{12}, \, \theta_{23}, \, \deltacp , \, \sdm , \, \ldm,  \, \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}})$ works similar to {\tt ChiTheta}, but only for one of the loaded experiments $N_{\mathrm{exp}}$.
-It returns the list $\{ \chi^2, \, \theta_{12}, \,  \theta_{23}, \, \deltacp , \, \sdm , \, \ldm, \, \hat{\rho}_{N_{\mathrm{exp}}} , \, N_{\mathrm{Iter}} \}$.
-\end{function}
-\begin{function}
-\GLB{ChiDelta}$( \deltacp, \, \{ \theta_{12}, \, \theta_{13}, \, \theta_{23}, \, \sdm , \, \ldm, \, \hat{\rho_1}, \,  \hdots ,  \,\hat{\rho_n} \})$ returns the projection onto the $\deltacp$-axis for all experiments.
-The parameters are the fixed value of $\deltacp$ and the starting point of the minimizer including all starting values for the matter density scaling factors (without $\deltacp$, which is fixed).
-The return value is a list $\{ \chi^2, \, \theta_{12}, \, \theta_{13}, \, \theta_{23}, \, \sdm , \, \ldm, \, \hat{\rho_1}, \, \hdots , \, \hat{\rho_n} , \, N_{\mathrm{Iter}} \}$ with the minimum $\chi^2$ found, the coordinates of the local minimum (without $\deltacp$), and the number of iterations $N_{\mathrm{Iter}}$ used by the minimizer (number of function calls of {\tt Chi}).
-\end{function}
-\begin{function}
-\GLB{SingleChiDelta}$( \deltacp, \, \{ \theta_{12}, \, \theta_{13}, \, \theta_{23}, \, \sdm , \, \ldm, \,  \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}})$ works similar to {\tt ChiDelta}, but only for one of the loaded experiments $N_{\mathrm{exp}}$.
-It returns the list $\{ \chi^2, \, \theta_{12}, \,  \theta_{13}, \,  \theta_{23}, \, \sdm , \, \ldm, \, \hat{\rho}_{N_{\mathrm{exp}}} , \, N_{\mathrm{Iter}} \}$.
-\end{function}
-All of these function have the same parameter structure: The fixed parameter is not transferred in the list, but as a the first separate parameter.
-The number of matter density scaling factors corresponds to the number of experiments used, since each experiment may face other matter density conditions.
-Note that before any of these function calls, {\tt SetStartingValues} and {\tt SetInputErrors} have to be used at least once.
-In addition, note that the resulting $\chi^2$ of {\tt ChiTheta} (or {\tt ChiDelta}) is not the sum of the $\chi^2$-values over all {\tt Single}-functions of all experiments anymore.
+
+The following functions are some of the simplest minimizers 
+provided by \GLOBES :
+\begin{function}
+\GLBNS{glbChiTheta}
+{\tt double glbChiTheta(const glb\_params in, glb\_params out, int exp)}
+returns the projected $\chi^2$ onto the $\theta_{13}$-axis for the  experiment {\tt exp}. For the simulation of all initialized experiments,
+use {\tt GLB\_ALL} for {\tt exp}.
+The values in {\tt in} are the guessed fit values for the minimizer (all parameters other than $\theta_{13}$) and the fixed fit value of $\theta_{13}$.
+The actually determined parameters at the minimum are returned in {\tt out}, where $\theta_{13}$ is still at
+its fixed value. If {\tt out} is set to {\tt NULL}, this information will not be returned.
+\end{function}
+\begin{function}
+\GLBNS{glbChiDelta}
+{\tt double glbChiDelta(const glb\_params in, glb\_params out, int exp)}
+returns the projected $\chi^2$ onto the $\deltacp$-axis for the  experiment {\tt exp}. For the simulation of all initialized experiments,
+use {\tt GLB\_ALL} for {\tt exp}.
+The values in {\tt in} are the guessed fit values for the minimizer (all parameters other than $\deltacp$) and the fixed fit value of $\deltacp$.
+The actually determined parameters at the minimum are returned in {\tt out}, where $\deltacp$ is still at its fixed value.
+If {\tt out} is set to {\tt NULL}, this information will not be returned.
+\end{function}
+
+All of the minimization functions have a similar parameter structure: The fixed fit parameter value and the guessed starting point of the minimizer, \ie, the guessed position of the minimum, are transferred in the list {\tt in}.
+Part of this list are the matter density
+scaling factors of all experiments, which are also minimized over. 
+The minimizer is then started at the guessed point and runs into
+the local minimum, where the fit parameter of the projection
+axis is fixed. For the best-fit solution, it is usually sufficient to start the minimizer at the best-fit point. However, the convergence speed might
+be better by starting it slightly off this point. In addition, there
+are problems in many cases with more complicated topologies, which means
+that better guesses for the position of the minimum are needed.
+The position of the minimum is then returned in {\tt out}
+together with the number of iterations used for the minimization.
+It is very often useful to print the output of the minimization with
+\GLB{glbPrintParams} in order to check that the minimum is the
+appropriate one. For example, if the mininizer ends up in the wong-sign
+solution in $\ldm$, priors can be used to force it into the tested
+minimum. In addition, the number of iterations used allows an optimization
+of the convergence speed. 
+% 
+Note that before any minimization, \GLB{glbSetStartingValues} 
+and \GLB{glbSetInputErrors} have to be used at least once.
+In addition, note that the resulting $\chi^2$ of {\tt glbChiTheta} (or {\tt glbChiDelta}) for the combination of more than one experiment is not
+equal to the sum of the individual $\chi^2$-values anymore.
 This has two reasons: First, the topology of the fit manifold is altered by the addition of $\chi^2$-values of different experiments.
@@
 Second, the priors for the external knowledge on the parameters are only added once -- independent of the number of experiments.
-A simple application of {\tt ChiTheta} can be found in the example on page~\pageref{ex:corrproj}.
-
-The return values of these functions do not only contain the minimum $\chi^2$-value found, but also the position of the minimum.
-This information can often be valuable, since one can often quickly locate irregularities: If one of the parameters is far off the best-fit value, the minimzer may have ended up in a local unwanted minimum.
-For instance, a negative value of $\ldm$ (with a positive best-fit value) indicates that the minimizer ended up in the opposite-sign solution.
-Another possibility is that the experiment may not be able to determine the measured parameter at all without external knowledge about the parameter far off its best-fit value.
-In both cases, an externally imposed precision on the run-off parameter can help to solve the problem.
+
+\index{Matter density scaling factor}
+The the output of the minimizer in {\tt out} carries as many matter density scaling factors as there are experiments. Either
+one (for the simulation of one experiment) or all (for the simulation of
+all experiments) are different from $1.0$ if matter density uncertainties
+are present, since each experiment may face other matter density conditions. The minimizer for indivdual experiments ``know'' which
+experiment they are currently treating, which means that only the returned
+matter density scaling factor of the appropriate experiment is modified.
+For example, calculating {\tt glbChiTheta} for the last experiment
+number, the last density values will be modified.
+This approach turns out to be extremely useful together with the
+simulation of more than one experiment. One can, for instance, locate the
+degeneracies of all individual experiments. In order to test if
+these degeneracies are still present in the combination of all experiments (which has a very different topology), one can test the combination of experiments with the output {\tt out} from the individual experiments.
+In this case, even the correct matter density scaling factor
+output is used.
+
+The example on page~\pageref{ex:corrproj} demonstrates how one can obtain \figu{projex} (right) with keeping all parameters but $\deltacp$ fixed, as well as how one can include the full $n$-parameter correlation with external input.
+It also demonstrates how these two compare to each other.
+One can easily read off this example that
+there is a substantial impact of the
+correlation with oscillation parameters other than $\deltacp$. 
+Note that it uses the function
+\GLB{glbChiNP} for arbitrary projections from the next section for the minimization over $\deltacp$.
+In addition, there is one interesting feature in guessing the
+oscillation parameters in this example: In order to avoid falling
+into the wrong minimum, the fit value of $\deltacp$ is guessed from \figu{projex} (left). This quite sophisticated ``guessing'' is typical
+for neutrino factories because of the $(\deltacp, \theta_{13})$-degeneracy, whereas it is for superbeams often sufficient
+to use the best-fit values. A strong indication for a wrong guessing 
+are discontiuous jumps in the projected $\chi^2$-function, where the minimizer jumps from one minimum to another.
+In such cases, the starting point of the minimizer has to be adjusted to help it find the true minimum.
+
+Other examples for projections onto a parameter axis while keeping exactly one parameter fixed are \GLB{glbChiTheta23}, \GLB{glbChiDm}, and \GLB{glbChiDms}, which
+can be found in \Tab~\ref{tab:stdfunctions} on page~\pageref{tab:stdfunctions}.
 
@@
 One such example is the projection of the fit manifold onto the $\stheta$-$\deltacp$-plane, \ie, $k=2$ here.
-This projection can be performed with the following functions:
-\begin{function}
-\GLB{ChiThetaDelta}$(\theta_{13}, \, \deltacp, \, \{ \theta_{12}, \, \theta_{23},  \, \sdm , \, \ldm, \,   \hat{\rho_1}, ...
-, \hat{\rho_n} \})$ returns the list  $\{ \chi^2, \, \theta_{12}, \, \theta_{23}, \, \sdm , \, \ldm,  \, \hat{\rho_1}, \, \hdots , \, \hat{\rho_n} , \, N_{\mathrm{Iter}} \}$ for the projection into the $\theta_{13}$-$\deltacp$-plane for all experiments.
-\end{function}
-\begin{function}
-\GLB{SingleChiThetaDelta}$( \theta_{13}, \, \deltacp, \, \{ \theta_{12}, \, \theta_{23}, \, \sdm , \, \ldm, \,   \hat{\rho}_{N_{\mathrm{exp}}} \}, \, N_{\mathrm{exp}})$ returns the list $\{ \chi^2, \, \theta_{12}, \,   \theta_{23},  \, \sdm , \, \ldm, \, \hat{\rho}_{N_{\mathrm{exp}}} ,  \, N_{\mathrm{Iter}} \}$ for the projection into the $\theta_{13}$-$\deltacp$-plane for one experiment $N_{\mathrm{exp}}$.
-\end{function}
-These functions work analogously to the ones in the last section.
-They can, for example, be used to obtain a figure similar to \figu{projex}, left.
-However, the result would not be a two-dimensional cut through the fit-manifold, but a two-dimensional projection.
-Though the running time for one call of these functions is somewhat shorter than the one for the $\stheta$- or $\deltacp$-projections, one has to compute a two-dimensional array for such a figure (instead of a one-dimensional list).
-Therefore, the overall computational effort is much higher.
+In this case, the two
+parameters $\stheta$ and $\deltacp$ are kept fixed, and the others are
+minimized over. 
+The corresponding function is
+\begin{function}
+\GLBNS{glbChiThetaDelta}
+{\tt double glbChiThetaDelta(const glb\_params in, glb\_params out, int exp)} returns the projected $\chi^2$ onto the $\theta_{13}$-$\deltacp$-plane for the  experiment {\tt exp}.
+For the simulation of all initialized experiments,
+use {\tt GLB\_ALL} for {\tt exp}.
+The values in {\tt in} are the guessed fit values for the minimizer (all parameters other than $\theta_{13}$ and $\deltacp$) and the fixed fit values of $\theta_{13}$ and $\deltacp$.
+The actually determined parameters at the minimum are returned in {\tt out}, where $\theta_{13}$ and $\deltacp$ are still at their fixed values.
+If {\tt out} is set to {\tt NULL}, this information will not be returned.
+\end{function}
+This function works analogously to the ones in the last section. They can, for example, be used to obtain a figure similar to \figu{projex}, left.
+The example on page~\pageref{ex:corrproj} illustrates then the difference
+between the projections of the ``eggs'' within the 
+$\stheta$-$\deltacp$-plane onto the $\theta_{13}$-axis. 
+Though the running time for one call of these functions is somewhat 
+shorter than the one for the $\stheta$- or $\deltacp$-projections, one 
+has to compute a two-dimensional array for such a figure (instead of a 
+one-dimensional list). Therefore, the overall computational effort is 
+much higher, \ie, in the order of hours. In many cases, it is therefore
+convenient to run {\tt glbChiSys} first to obtain a picture of
+the manifold and to adjust the parameter ranges. Then, one can run
+{\tt glbChiThetaDelta} for a complete evaluation of the problem
+including correlations.
+
+\begin{table}[tp]
+\begin{tabular}{p{4.2cm}p{5.5cm}p{5.1cm}}
+\hline
+Function & Purpose & Parameters $\rightarrow$ Result\\
+\hline
+\GLB{glbAllocProjection} & Create projection vector & {\tt ()} \\
+\GLB{glbFreeProjection} & Destroy projection vector {\tt stale} & {\tt (glb\_projection stale)} \\
+\GLB{glbDefineProjection} & Assign projection vector {\tt in} & {\tt (glb\_projection in, int theta12, int theta13, int theta23, int delta, int dms, int dma)} \\ 
+\GLB{glbCopyProjection} & Copy vector {\tt source} to {\tt dest} & {\tt (const glb\_projection source, glb\_projection dest)} \\
+\GLB{glbPrintProjection} & Print vector {\tt in} to file {\tt stream} & {\tt (FILE* stream, const glb\_projection in)} \\
+\GLB{glbSetProjectionFlag} & Set flag for oscillation parameter {\tt which} in vector {\tt in} to value {\tt flag}. & {\tt (glb\_projection in, int flag, int which)} \\
+\GLB{glbGetProjectionFlag} & Return flag for oscillation parameter {\tt which} in vector {\tt in}. & {\tt (const glb\_projection in, int which)} $\rightarrow$ {\tt int} flag \\
+\GLBNS{glbSetDensityProjectionFlag} {\tt glbSetDensity-} \newline {\tt ProjectionFlag}  & Set flag for density parameter {\tt which} in vector {\tt in} to value {\tt flag}.
+& {\tt (glb\_projection in, int flag, int which)} \\
+\GLBNS{glbGetDensityProjectionFlag} {\tt glbGetDensity-} \newline {\tt ProjectionFlag} & Return flag for density parameter {\tt which} in vector {\tt in}.
+& {\tt (const glb\_projection in, int which)} $\rightarrow$ {\tt int} flag \\
+\hline
+\end{tabular}
+\caption{\label{tab:defprojection} Different functions handling the
+\GLB{glb\_projection} type. Flags are either \GLB{GLB\_FIXED} or \GLB{GLB\_FREE}.}
+\end{table}
 
@@
 In addition, one may want to use any other set of parameters for single- or two-parameter projections.
-The functions {\tt ChiNP} and {\tt SingleChiNP} are designed for this purpose:
-
-FUNCTIONS TO BE DEFINED. (WORK IN PROGRESS).
+The function \GLB{glbChiNP} is designed for this purpose.
+
+WW: CONTINUE HERE
+
+glb\_projection type: exactly similar to glb\_params (return values)
+
+Functions in table: komische Aufteilung
+
+Hint on example.
 
@@
    \hspace*{0.5cm} is imposed in order to avoid falling into the right-sign solution */ \\
-  glbDefineParams(input\_errors,theta12*0.1,10,10,10,sdm*0.1,ldm/3); \\ 
+  glbDefineParams(input\_errors,theta12*0.1,0,0,0,sdm*0.1,ldm/3); \\ 
   glbSetDensityParams(input\_errors,0.05,GLB\_ALL); \\

### 2004-07-06 Documentation/part1.tex
*** empty log message ***

@@
 
-\begin{table}[tp]
+\begin{table}[t!]
 \begin{tabular}{p{4.2cm}p{5.5cm}p{5.1cm}}
@@
 \GLB{glbGetProjectionFlag} & Return flag for oscillation parameter {\tt which} in vector {\tt in}. & {\tt (const glb\_projection in, int which)} $\rightarrow$ {\tt int} flag \\
-\GLBNS{glbSetDensityProjectionFlag} {\tt glbSetDensity-} \newline {\tt ProjectionFlag}  & Set flag for density parameter {\tt which} in vector {\tt in} to value {\tt flag}.
+ {\tt glbSetDensity\-ProjectionFlag} \GLBNS{glbSetDensityProjectionFlag} & Set flag for density parameter {\tt which} in vector {\tt in} to value {\tt flag}.
 & {\tt (glb\_projection in, int flag, int which)} \\
-\GLBNS{glbGetDensityProjectionFlag} {\tt glbGetDensity-} \newline {\tt ProjectionFlag} & Return flag for density parameter {\tt which} in vector {\tt in}.
+{\tt glbGetDensity\-ProjectionFlag} \GLBNS{glbGetDensityProjectionFlag} & Return flag for density parameter {\tt which} in vector {\tt in}.
 & {\tt (const glb\_projection in, int which)} $\rightarrow$ {\tt int} flag \\
@@
 \caption{\label{tab:defprojection} Different functions handling the
-\GLB{glb\_projection} type. Flags are either \GLB{GLB\_FIXED} or \GLB{GLB\_FREE}.}
+\GLB{glb\_projection} type.
+Flags are either \GLB{GLB\_FIXED} or \GLB{GLB\_FREE}.
+The (unshown) return values of the {\tt Set-} and {\tt Define-} functions point either to the assigned vector if successful, or they are {\tt NULL} if unsuccessful.}
 \end{table}
@@
 In principle, one can also use three- or more-dimensional projections.
-In addition, one may want to use any other set of parameters for single- or two-parameter projections.
-The function \GLB{glbChiNP} is designed for this purpose.
-
-WW: CONTINUE HERE
-
-glb\_projection type: exactly similar to glb\_params (return values)
-
-Functions in table: komische Aufteilung
-
-Hint on example.
+In addition, one may want to use a different set of parameters for single- or two-parameter projections.
+The very flexible function \GLB{glbChiNP} is
+designed for this purpose. However, because of its flexibility, it 
+involves more sophistication.
+
+In order to define arbitrary projections, we introduce the vector
+\GLB{glb\_projection}, which is very similar to the
+oscillation parameter vector {\tt glb\_params}.
+Normally, the user does not need to access this type directly:
+A set of function similar to the ones for {\tt glb\_params} is
+provided. The purpose of {\tt glb\_projection} is to tell \GLOBES\ 
+what parameters are fixed, and what are minimized over. Thus, in
+comparison to {\tt glb\_params}, it does not take values for the
+parameters, but flags \GLB{GLB\_FIXED} or \GLB{GLB\_FREE}.
+For example, the projection onto the $\theta_{13}$-axis {\tt glbChiTheta}
+is nothing else than a special case of {\tt glbChiNP} with $\theta_{13}$
+fixed and all the other parameters free. Similar to {\tt glb\_params},
+the type {\tt glb\_projection} has to be created first, and destroyed
+later. The access functions for {\tt glb\_projection} are summarized in \Tab~\ref{tab:defprojection}. Since the complete set is very similar to
+the one for {\tt glb\_params}, we do not go into greater details here.
+
+As soon as we have defined a projection, we can assign it:
+\begin{function}
+\GLBNS{glbSetProjection}
+{\tt int glbSetProjection(const glb\_projection in)} sets the projection
+to {\tt in}. The return value is $0$ if successful, and $-1$ if
+unsuccessful.
+\end{function}
+Similarly, the currenty assigned projection can be returned with:
+\begin{function}
+\GLBNS{glbGetProjection}
+{\tt int glbGetProjection(glb\_projection out)} writes the currently
+set projection to {\tt out}. The return value is $0$ if successful, and $-1$ if unsuccessful.
+\end{function}
+After setting the starting values, input errors, and the projection, 
+we can run the minimizer:
+\begin{function}
+\GLBNS{glbChiNP}
+{\tt double glbChiNP(const glb\_params in, glb\_params out, int exp)} 
+returns the projected $\chi^2$ onto the hyperplane specified by 
+{\tt glbSetProjection} for the  experiment {\tt exp}. 
+For the simulation of all initialized experiments,
+use {\tt GLB\_ALL} for {\tt exp}. The values in {\tt in} are the guessed fit values for the minimizer (all free parameters) and the fit values
+on the hyperplane (all fixed parameters).
+The actually determined parameters at the minimum are returned in {\tt out}, where the fixed parameters are still at their input values.
+If {\tt out} is set to {\tt NULL}, this information will not be returned.
+\end{function}
+As an example, the projection sequence for a minimization over
+$\deltacp$ only looks like this:
+\begin{quote}
+{\tt
+  glb\_projection th13\_projection = glbAllocProjection(); \\
+  glbDefineProjection(th13\_projection,GLB\_FIXED,GLB\_FIXED,GLB\_FIXED,\\
+  \hspace*{2cm} GLB\_FREE,GLB\_FIXED,GLB\_FIXED); \\
+  glbSetProjection(th13\_projection); \\ 
+  res1=glbChiNP(test\_values,NULL,GLB\_ALL); \\
+  glbFreeProjection(th13\_projection);
+}
+\end{quote}
+In this case, only the correlation with $\deltacp$ is taken into account.
+Note that in the  example on page~\pageref{ex:corrproj} this projection
+is compared with the result including the full multi-parameter correlation.
 

### 2004-07-06 Documentation/part1.tex
GLoBES tour added

@@
 
-\chapter{A short \GLOBES\ tour}
+\chapter{A \GLOBES\ tour}
 \label{chapter:tour}
 
-To be written later: Short introduction such as ``With the following lines we obtain ...''
-without detailed description of parameters. Should contain all most important functions of \GLOBES\ --
-overview of main functions in \tabl{stdfunctions}.
-Maybe: in form of long example which involves everything
-and can be compiled.
+In this first chapter, we show a \GLOBES\ tour illustrating the
+main features of \GLOBES . The complete example  
+can be found as {\tt example-tour.c} in the example subdirectory of your \GLOBES\ distribution.
+The output is written to {\tt stream}, which can be either {\tt stdout},
+or a file. Details about how to use \GLOBES\ with C can found in \Chapt~\ref{chapt:gettingstarted} and the following chapters.
+You can also find a summary of the most important \GLOBES\ $\chi^2$-functions in \tabl{stdfunctions}. Note that this chapter
+can be skipped without loss of relevant information.
 
@@
 \end{table}
+
+\vspace*{0.5cm}
+
+\noindent Initialize the \GLOBES\ library:
+\gq{
+glbInit(argv[0]);
+} 
+Define my standard oscillation parameters:
+\gq{
+double theta12 = asin(sqrt(0.8))/2; \\
+double theta13 = asin(sqrt(0.001))/2;\\
+double theta23 = M\_PI/4;\\
+double deltacp = M\_PI/2;\\
+double sdm = 7e-5;\\
+double ldm = 2e-3;
+}
+Load one neutrino factory experiment:
+\gq{
+glbInitExperiment("NuFact.glb",\&glb\_experiment\_list[0], \\ \hspace*{4cm} \&glb\_num\_of\_exps);
+} 
+Initialize a number of parameter vectors we are going to use later:
+\gq{
+glb\_params true\_values = glbAllocParams();\\
+glb\_params fit\_values = glbAllocParams();\\
+glb\_params starting\_values = glbAllocParams();\\
+glb\_params input\_errors = glbAllocParams();\\
+glb\_params minimum = glbAllocParams();
+}
+Assign values to our standard oscillation parameters:
+\gq{
+ glbDefineParams(true\_values,theta12,theta13,theta23,deltacp,sdm,ldm);
+}
+Compute the simulated data with our standard parameters:
+\gq{
+ glbSetOscillationParameters(true\_values); \\
+ glbSetRates();
+}
+Return the oscillation probabilities in vacuum and matter for the
+electron neutrino as initial flavor:
+\gq{
+ int i; \\
+ fprintf(stream,"$\backslash$nOscillation probabilities in vacuum: ");\\
+ for(i=1;i<4;i++) fprintf(stream,"1->\%i: \%g",i, \\
+ \hspace*{4cm} glbVacuumProbability(1,i,+1,50,3000)); \\ 
+ fprintf(stream,"$\backslash$nOscillation probabilities in matter: ");\\
+ for(i=1;i<4;i++) fprintf(stream,"1->\%i: \%g ",i,\\ \hspace*{4cm} glbProfileProbability(1,i,+1,50)); \\
+}
+\go{
+Oscillation probabilities in vacuum: 1->1: 0.999953 1->2: 2.69441e-05 1->3: 1.98019e-05 \\
+Oscillation probabilities in matter: 1->1: 0.999965 1->2: 2.02573e-05 1->3: 1.49021e-05 
+}
+Now assign fit values, where we will test the fit value $\stheta=0.0015$:
+\gq{
+ glbCopyParams(true\_values,fit\_values); \\
+ glbSetOscParams(fit\_values,asin(sqrt(0.0015))/2,GLB\_THETA\_13);
+}
+Compute $\chi^2$ with systematics only for all experiments and rules:
+\gq{
+  chi2 = glbChiSys(fit\_values,GLB\_ALL,GLB\_ALL); \\
+  fprintf(stream,"chi2 with systematics only: \%g$\backslash$n$\backslash$n",chi2);
+}
+\go{
+chi2 with systematics only: 22.3984
+}
+This we would obtain from the first appearance channel only:
+\gq{
+ chi2 = glbChiSys(fit\_values,0,0);\\
+ fprintf(stream,"This we would have from the CP-even appearance channel only: \%g$\backslash$n$\backslash$n",chi2);
+}
+\go{
+This we would have from the CP-even appearance channel only: 21.6223
+}
+The sum over all rules again gives:
+\gq{
+ chi2 = glbChiSys(fit\_values,GLB\_ALL,0)+ 
+        glbChiSys(fit\_values,GLB\_ALL,1)+ \\  
+\hspace*{1.3cm} glbChiSys(fit\_values,GLB\_ALL,2)+          glbChiSys(fit\_values,GLB\_ALL,3); \\ 
+\mbox{fprintf(stream,"The sum over all rules gives again: \%g$\backslash$n$\backslash$n",chi2);}
+}
+\go{
+The sum over all rules gives again: 22.3984
+}
+Let's prepare the minimizers for taking into account correlations.
+Set errors for external parameters, too: 10\% for each of the solar parameters, and 5\% for the matter density. 
+\gq{
+ glbDefineParams(input\_errors,theta12*0.1,0,0,0,sdm*0.1,0); \\
+ glbSetDensityParams(input\_errors,0.05,GLB\_ALL); \\
+ glbSetStartingValues(true\_values); \\
+ glbSetInputErrors(input\_errors);
+}
+Then we can calculate $\chi^2$ including the full multi-parameter
+correlation, and show where \GLOBES\ actually found the minimum
+(note that this takes somewhat longer than systematics only). 
+This corresponds to a projection onto the $\stheta$-axis:
+\gq{
+ chi2 = glbChiTheta(fit\_values,minimum,GLB\_ALL); \\ 
+ fprintf(stream,"chi2 with correlations: \%g $\backslash$n",chi2); \\
+ fprintf(stream,"Position of minimum: theta12, theta13, theta23, \\ \hspace*{0.5cm} delta, sdm, ldm, rho$\backslash$n"); \\
+ glbPrintParams(stream,minimum); \\ 
+ fprintf(stream,"Note that s22theta13 is unchanged/kept fixed: \\
+ \hspace*{0.5cm} \%g! $\backslash$n$\backslash$n",    pow(sin(2*glbGetOscParams(minimum,GLB\_THETA\_13)),2));
+}
+\go{
+chi2 with correlations: 2.1038 \\
+Position of minimum: theta12,theta13,theta23,delta,sdm,ldm,rho \\
+0.542002 0.0193698 0.747915 1.77688 6.66156e-05 0.00200817 \\
+1.00434 \\
+Iterations: 1693 \\
+Note that s22theta13 is unchanged/kept fixed: 0.0015! 
+}
+Instead of including the full correlation, we can take the
+correlation with every parameter except from $\deltacp$, \ie,
+we keep (in addition to $\theta_{13}$) $\deltacp$ fixed.
+This corresponds to projection onto the $\stheta$-$\deltacp$-plane:
+\gq{
+ chi2 = glbChiThetaDelta(fit\_values,minimum,GLB\_ALL); \\
+fprintf(stream,"chi2 with correlations other than with deltacp: \%g $\backslash$n$\backslash$n",chi2); 
+}
+\go{
+chi2 with correlations other than with deltacp: 4.32831 
+}
+Similarly, we can only take into account the correlation with $\deltacp$.
+For this, we need to define our own (user-defined) projection, where
+only $\deltacp$ is a free parameter:
+\gq{
+ glb\_projection myprojection = glbAllocProjection(); \\
+ glbDefineProjection(myprojection,GLB\_FIXED, GLB\_FIXED, GLB\_FIXED, \\
+ \hspace*{0.5cm} GLB\_FREE, GLB\_FIXED, GLB\_FIXED); \\
+ glbSetProjection(myprojection); \\
+ chi2 = glbChiNP(fit\_values,minimum,GLB\_ALL); \\
+ fprintf(stream,"chi2 with correlation only with deltacp: \\
+ \hspace*{0.5cm} \%g $\backslash$n$\backslash$n",chi2); \\
+ glbFreeProjection(myprojection); 
+}  
+\go{
+chi2 with correlation only with deltacp: 2.80651 
+}
+We can also switch of the systematics and compute the
+statistics $\chi^2$ only:
+\gq{
+ glbSwitchSystematics(GLB\_ALL,GLB\_ALL,GLB\_OFF); \\   
+ chi2 = glbChiSys(fit\_values,GLB\_ALL,GLB\_ALL); \\
+ glbSwitchSystematics(GLB\_ALL,GLB\_ALL,GLB\_ON); \\   
+ fprintf(stream,"chi2 with statistics only: \\ 
+ \hspace*{0.5cm} \%g$\backslash$n$\backslash$n",chi2);
+}
+\go{
+chi2 with statistics only: 39.143
+}
+Let us now locate the exact position of the sgn-degeneracy:
+\gq{
+  glbDefineParams(input\_errors,theta12*0.1,0,0,0,sdm*0.1,ldm/3); \\
+  glbDefineParams(starting\_values,theta12,theta13,theta23, \\
+  \hspace*{0.5cm} deltacp,sdm,-ldm); \\
+  glbSetDensityParams(input\_errors,0.05,GLB\_ALL); \\
+  glbSetStartingValues(starting\_values);\\
+  glbSetInputErrors(input\_errors);\\
+  chi2=glbChiAll(starting\_values,minimum,GLB\_ALL);\\ 
+  fprintf(stream,"chi2 at minimum: \%g $\backslash$n",chi2);\\
+  fprintf(stream,"Position of minimum: \\ \hspace*{0.5cm} theta12,theta13,theta23,delta,sdm,ldm,rho$\backslash$n"); \\
+  glbPrintParams(stream,minimum); 
+}
+\go{
+chi2 at minimum: 6.20025 \\
+Position of minimum: theta12,theta13,theta23,delta,sdm,ldm,rho \\
+0.591812 0.0264717 0.72763 1.08709 8.0004e-05 -0.00206094 \\
+0.970685  \\
+Iterations: 1946 \\
+}
+After testing these functions with only one experiment, let us now
+go to a two-experiment setup with two different neutrino factory baselines.
+Since the \GLOBES\ parameter vectors depend on the number of experiments,
+we have to destroy them first:
+\gq{
+ glbFreeParams(true\_values); \\
+ glbFreeParams(fit\_values); \\ 
+ glbFreeParams(starting\_values);\\
+ glbFreeParams(input\_errors);\\
+ glbFreeParams(minimum);
+}
+Then we clear the experiment list and load the new experiments:
+\gq{
+ fprintf(stream,"$\backslash$nNOW: TWO-EXPERIMENT SETUP \\
+ \hspace*{0.5cm} NuFact at 3000km+NuFact at 7500km$\backslash$n$\backslash$n"); \\
+ \\
+  glbClearExperimentList(); \\
+\\
+ glbInitExperiment("NuFact.glb",\&glb\_experiment\_list[0], \\
+ \hspace*{0.5cm} \&glb\_num\_of\_exps);\\
+  glbInitExperiment("NuFact.glb",\&glb\_experiment\_list[0], \\
+ \hspace*{0.5cm} \&glb\_num\_of\_exps);
+}
+\go{
+NOW: TWO-EXPERIMENT SETUP NuFact at 3000km+NuFact at 7500km
+}
+Then we need to change the baseline of the second experiment, where
+we set the density to the average density of this baseline:
+\gq{
+  double* lengths;  \\
+  double* densities; \\
+ glbAverageDensityProfile(7500,\&lengths,\&densities);\\
+  fprintf(stream,"Magic baseline length: \%g, \\
+  \hspace*{0.5cm} Density: \%g$\backslash$n$\backslash$n",lengths[0],densities[0]); \\
+ glbSetProfileDataInExperiment(1,1,lengths,densities); \\
+   free(lengths); \\
+   free(densities);
+}
+\go{
+ Magic baseline length: 7500, Density: 4.25286
+}
+Now we can re-initialize our parameter vectors again:
+\gq{
+ true\_values = glbAllocParams();\\
+ fit\_values = glbAllocParams(); \\
+ starting\_values = glbAllocParams(); \\
+ input\_errors = glbAllocParams(); \\
+ minimum = glbAllocParams(); \\
+ glb\_params minimum2 = glbAllocParams(); 
+}
+In addition, we repeat the procedure for the simulated rates and
+the fit parameter vector:
+\gq{
+  glbDefineParams(true\_values,theta12,theta13,theta23,deltacp,sdm,ldm); \\
+  glbSetOscillationParameters(true\_values); \\
+  glbSetRates();\\
+  \\
+  glbCopyParams(true\_values,fit\_values);\\
+  glbSetOscParams(fit\_values,asin(sqrt(0.0015))/2,GLB\_THETA\_13);
+} 
+Here comes the $\chi^2$ with systematics only for all experiments and
+rules:
+\gq{
+ chi2 = glbChiSys(fit\_values,GLB\_ALL,GLB\_ALL); \\
+ fprintf(stream,"chi2 with systematics for all exps: \\ 
+ \hspace*{0.5cm} \%g$\backslash$n",chi2); 
+}
+\go{
+ chi2 with systematics for all exps: 31.0797
+}
+Compute $\chi^2$ for each experiment and compute the sum:
+\gq{
+  chi2 = glbChiSys(fit\_values,0,GLB\_ALL); \\
+ fprintf(stream,"chi2 with systematics for 3000km: \%g$\backslash$n",chi2); \\
+ chi2b = glbChiSys(fit\_values,1,GLB\_ALL); \\
+ fprintf(stream,"chi2 with systematics for 7500km: \%g$\backslash$n",chi2b);
+ \\
+  fprintf(stream,"The two add again to: \newline
+  \hspace*{0.5cm} \%g$\backslash$n$\backslash$n",chi2+chi2b);
+}
+\go{
+chi2 with systematics for 3000km: 22.3984 \\
+chi2 with systematics for 7500km: 8.68131 \\
+The two add again to: 31.0797
+}
+Similarly, compute the $\chi^2$ with correlations for each experiment
+and their combination. Compare it to the $\chi^2$ for all experiments:
+the sum of the individual results is not equal to the $\chi^2$ of the
+combination anymore. Note that there are now two densities in the
+output vectors. 
+\gq{
+ glbDefineParams(input\_errors,theta12*0.1,0,0,0,sdm*0.1,0); \\
+ glbSetDensityParams(input\_errors,0.05,GLB\_ALL); \\
+ glbSetStartingValues(true\_values); \\
+ glbSetInputErrors(input\_errors); \\
+ chi2 = glbChiTheta(fit\_values,minimum,0);\\ 
+  fprintf(stream,"chi2 with correlations for 3000km: \%g $\backslash$n",chi2); \\
+ glbPrintParams(stream,minimum); \\ 
+  chi2b = glbChiTheta(fit\_values,minimum,1); \\ 
+ fprintf(stream,"$\backslash$nchi2 with correlations for 7500km: \\
+  \hspace*{0.5cm}  \%g $\backslash$n",chi2b); \\
+  glbPrintParams(stream,minimum); \\ 
+ chi2sum = glbChiTheta(fit\_values,minimum,GLB\_ALL); \\ 
+ fprintf(stream,"$\backslash$nchi2 with correlations for combination: \\
+  \hspace*{0.5cm} \%g $\backslash$n",chi2sum); \\
+ glbPrintParams(stream,minimum); \\ 
+  fprintf(stream,"$\backslash$nThe sum of the two chi2s is \%g,  \\
+  \hspace*{0.5cm} whereas the total chi2 is \%g !$\backslash$n$\backslash$n",chi2+chi2b,chi2sum); 
+}
+\go{
+chi2 with correlations for 3000km: 2.1038 \\
+0.542002 0.0193698 0.747915 1.77688 6.66156e-05 0.00200817 \\ 
+1.00434 1 \\
+Iterations: 1693 \\
+\\
+chi2 with correlations for 7500km: 1.08421 \\
+0.557356 0.0193698 0.771359 4.77751 7.00762e-05 0.00200105 \\
+1 1.01517 \\
+Iterations: 661 \\
+\\
+chi2 with correlations for combination: 3.90835  \\
+0.544432 0.0193698 0.770175 1.78502 6.61621e-05 0.00200303 \\ 
+1.00431 1.03679 \\
+Iterations: 1636 \\
+\\
+The sum of the two chi2s is 3.18801, whereas the total chi2 is 3.90835!
+}
+Now find the $\mathrm{sgn}(\ldm)$-degeneracies for both individual
+experiments and test if they are still there in the combination of the
+experiments.
+Note that a minimum at a negative value of $\theta_{13}$ is unphysical.
+However, if there can be no degeneracy found at a positive value, there is probably none at a low confidence level.
+\gq{
+ glbDefineParams(input\_errors,theta12*0.1,theta13,theta23, \\
+ \hspace*{0.5cm} deltacp,sdm*0.1,ldm/3); \\
+glbDefineParams(starting\_values,theta12,theta13,theta23, \\
+\hspace*{0.5cm} deltacp,sdm,-ldm);\\
+  glbSetDensityParams(input\_errors,0.05,GLB\_ALL);\\
+  glbSetStartingValues(starting\_values);\\
+  glbSetInputErrors(input\_errors);\\
+  \\
+  chi2=glbChiAll(starting\_values,minimum,0); \\ 
+  fprintf(stream,"chi2 at minimum, L=3000km: \%g $\backslash$n",chi2); \\
+  glbPrintParams(stream,minimum);   \\
+  \\
+  chi2b=glbChiAll(starting\_values,minimum2,1);  \\
+  \mbox{fprintf(stream,"$\backslash$nchi2 at minimum, L=7500km: \%g \\ $\backslash$n",chi2b);} \\
+  glbPrintParams(stream,minimum2);  \\
+  \\
+  chi2=glbChiAll(minimum,minimum,GLB\_ALL); \\
+  fprintf(stream,"$\backslash$nchi2 for combination at minimum of Exp. 1:\\ \hspace*{0.5cm} \%g $\backslash$n",chi2); \\
+  glbPrintParams(stream,minimum);  \\
+  \\
+  chi2b=glbChiAll(minimum2,minimum2,GLB\_ALL); \\
+  fprintf(stream,"$\backslash$nchi2 for combination at minimum of Exp. 2: \\
+  \hspace*{0.5cm} \%g $\backslash$n",chi2b); \\
+  glbPrintParams(stream,minimum2);  
+}
+\go{
+chi2 at minimum, L=3000km: 6.71794 \\
+0.591497 0.0257396 0.729058 1.11537 7.98867e-05 -0.00206005 \\
+0.970499 1 \\
+Iterations: 2104\\
+\\
+chi2 at minimum, L=7500km: 47.1013 \\
+0.590347 0.0018489 0.768372 0.984827 8.23415e-05 -0.00204588 \\
+1 0.780995 \\
+Iterations: 1270\\
+\\
+chi2 for combination at minimum of Exp. 1: 70.6353 \\
+0.607988 0.0165985 0.767682 1.41422 8.44573e-05 -0.00204853 \\
+0.96147 1.1831 \\
+Iterations: 1549 \\
+\\
+chi2 for combination at minimum of Exp. 2: 70.6357 \\
+0.608454 0.0165823 0.767757 1.41481 8.43864e-05 -0.00204853 \\
+0.961129 1.18304 \\
+Iterations: 1447 
+}
+Finally, we have to destroy the parameter vectors again:
+\gq{
+ glbFreeParams(true\_values); \\
+ glbFreeParams(fit\_values); \\
+ glbFreeParams(starting\_values); \\
+ glbFreeParams(input\_errors); \\
+ glbFreeParams(minimum); \\
+ glbFreeParams(minimum2); \\
+}
 
@@
 
-\chapter{Finding degenerate solutions}
+\chapter{Locating degenerate solutions}
 \index{Degenerate solutions}
@@
 Similarly, one can minimize over {\em all} $n$ parameters to find the local minimum close to any starting point.
-This approach is very useful for the exact numerical location of a degenerate if its approximate position is known.
-For the determination of the approximate positon, one can use analytical approaches, such as in \Refs~TO BE ENTERED, or an educated guess.
+This approach is very useful for the exact numerical location of a degeneracy if its approximate position is known.
+For the determination of the approximate positon, one can use analytical approaches or an educated guess.
 Though the usage of the all-parameter minimizers is quite simple, one should keep in mind that they are local minimizers.
-Therefore, one may need a very sophisticated application software to actually find all degenerate solutions.
+Therefore, one may need a very sophisticated application software to find all degenerate solutions.
 
@@
 \index{All-parameter minimization}
-The functions to perform the all-parameter minimization are {\tt ChiAll} and {\tt SingleChiAll}:
-\begin{function}
-\GLB{ChiAll}$(\{ \theta_{13}, \, \theta_{12}, \, \theta_{23}, \, \deltacp , \, \sdm , \, \ldm,  \, \hat{\rho_1}, \, \hdots , \, \hat{\rho_n} \})$ finds the local minimum close to the given starting point (list) for all initialized experiments.
-It returns the list  $\{ \chi^2, \, \theta_{13}, \, \theta_{12}, \, \theta_{23}, \, \deltacp , \, \sdm , \,  \ldm,  \, \hat{\rho_1}, \, \hdots , \hat{\rho_n} , \, N_{\mathrm{Iter}} \}$ with the minimum $\chi^2$-value, the actual position of the local minimum, and the number of iterations $N_{\mathrm{Iter}}$ used.
-\end{function}
-\begin{function}
-\GLB{SingleChiAll}$( \{ \theta_{13}, \, \theta_{12}, \, \theta_{23}, \, \deltacp , \, \sdm , \, \ldm, \,   \hat{\rho}_{N_{\mathrm{exp}}} \}, \, \, N_{\mathrm{exp}})$ finds the local minimum close to the given starting point (list) for the experiment $N_{\mathrm{exp}}$ only.
-It returns the list
- $\{ \chi^2, \, \theta_{13}, \, \theta_{12}, \, \theta_{23}, \, \deltacp , \, \sdm , \, \ldm, \, \hat{\rho}_{N_{\mathrm{exp}}} ,  \, N_{\mathrm{Iter}} \}$ with the minimum $\chi^2$-value, the actual position of the local minimum, and the number of iterations $N_{\mathrm{Iter}}$ used.
+The function to perform the all-parameter minimization is {\tt glbChiAll}:
+\begin{function}
+\GLBNS{glbChiAll} 
+{\tt double glbChiAll(const glb\_params in, glb\_params out, int exp)}
+  returns the minimized $\chi^2$ over all parameters for the  experiment {\tt exp}. For the simulation of all initialized experiments,
+use {\tt GLB\_ALL} for {\tt exp}.
+The values in {\tt in} are the guessed fit values for the minimizer.
+The actually determined parameters at the minimum are returned in {\tt out}.
+If {\tt out} is set to {\tt NULL}, this information will not be returned.
 \end{function}
 %
-Both functions take the suspected position of the local minimum and return its true position.
-The application of  {\tt SingleChiAll} can be especially usefull for many different experiments evaluated simultaneously: It turns out to be useful to find the positions of the degeneracies for the individual experiments first, and then test all of these with the combination of all experiments in order not to miss a degenerate solution.
+This function takes the suspected position of the local minimum and returns its actual position in {\tt out}, as well as the $\chi^2$-value 
+at the minimum as return value. Thus, the return value can be immediately
+used to judge wether the located degeneracy appears at the chosen
+confidence level.
+
 The example on page~\pageref{ex:sgndeg} illustrates how to locate the $\mathrm{sgn}(\ldm)$-degeneracy and show the corresponding degenerate solution in the $\stheta$-$\deltacp$-plane together with the original solution.
@@
 Thus, the degeneracy may not even appear at the chosen confidence level on the plane, but it does appear at the true minimum.
-The two cuts\footnote{The discussed figure on page~\pageref{ex:sgndeg} is produced by  {\tt Chi} and thus only represents a cut through the fit manifold.
-For the projection including correlations, one may rather want to use {\tt ChiThetaDelta}.} through the fit manifold shown in the figure on page~\pageref{ex:sgndeg} therefore do not have the same oscillation parameter values (except from the ones shown in the figure).
-
-In practice, a number of tricks can be useful for the treatment of degenerate solutions:
+The two sections\footnote{The discussed figure on page~\pageref{ex:sgndeg} is produced by  {\tt glbChiSys} and thus only represents a section through the fit manifold.
+For the projection including correlations, one may rather want to use {\tt glbChiThetaDelta}.} through the fit manifold shown in the figure on page~\pageref{ex:sgndeg} therefore do not appear at the same oscillation parameter values (except from the ones shown in the figure).
+
+For the more advanced reader, a number of tricks can be useful for the numerical localization of degenerate solutions:
 \begin{description}
@@
 This saves a lot of computation time.
-\item[Locating degeneracies with more complicated topologies.] For more complicated topologies, such as for neutrino factories, it is often useful to use multi-step location procedures or analytical knowledge.
-For example, for a numerical procedure, one may first of all switch off the systematics and keep $\stheta$ or $\deltacp$ fixed, \ie, use {\tt SingleChiTheta} or {\tt SingleChiDelta}, where $\stheta$ or $\deltacp$ is fixed to the best-fit value.
-The result can then be used as a starting point for {\tt SingleChiAll} with the systematics switched on again.
-In addition to switching off the systematics, it can be useful to reduce the input errors during the first steps in order to make the minimizer not to run away too much from the true solution.
-\item[Finding degeneracies with multiple experiments.] For multiple experiments, the degeneracies should be located for each of the individual experiments first.
-Then, all of the found degeneracies below the threshold can be tested for the combination of experiments.
+\item[Locating degeneracies in more complicated topologies.] For more complicated topologies, such as for neutrino factories, it is often useful to use multi-step location procedures or analytical knowledge.
+For example, for a numerical procedure, one may first of all switch off the systematics and keep $\stheta$ or $\deltacp$ fixed, \ie, use {\tt glbChiTheta}, where $\stheta$ or $\deltacp$ is fixed at the best-fit value.
+The result can then be used as a starting point for {\tt glbChiAll} for the individual experiments with the systematics switched on again.
+\item[Forcing the minimizer into the targeted solution.]
+In addition to switching off the systematics, it can be useful to reduce the input errors during some steps of the localization process in order to make the minimizer not to run away too much from the targeted solution.
+The example on page~\pageref{ex:sgndeg} illustrates this with the
+input error for $\ldm$: Since the guessed starting point might be
+quite far away from the real degeneracy, the algorithm may in some cases
+find the original solution instead of the degeneracy (which can
+be immediately seen from the output vector). The input error
+for $\ldm$ gives the algorithm a ``bias'' against the original solution.
+However, note that the input error must not be too small in order
+to avoid a significant contribution of the prior to the final $\chi^2$.
+Alternatively, one could run {\tt glbChiAll} again with the located minimum
+as {\tt in} vector and $\ldm$ kept free.
+\item[Finding degeneracies with multiple experiments.] For multiple experiments, it turns out to be useful to locate the degeneracies for individual experiments first.
+Then, all of the found degeneracies below the threshold can be tested for the combination of all experiments.
 \end{description}
 Finally, note that any degenerate solution below the confidence level threshold which can not be located makes the result appear better than it actually is.
-Thus, one should be careful with the determination of the degenerate solutions.
+Thus, one should be careful with the determination of the degenerate solutions in order to find all of them.
 

### 2004-07-06 Documentation/part2.tex
GLoBES tour added

@@
 \end{equation} 
-CHECK FACTOR OF TWO IN DENOMINATOR WITH SOFTWARE (IMPORTANT)!!! (NORMALIZATION IS OK!)
 There are several energy resolution functions available, where by default

### 2004-07-07 Documentation/part1.tex
*** empty log message ***

@@
 
+\section{Version control}
+\index{Version control}
+
+In order to keep track of the used version of \GLOBES , the software
+provides a number of functions to check the \GLOBES\ and experiment
+versions. It is up to the user to implement mechanisms into the
+program and \AEDL\ files to check wether
+\begin{itemize}
+\item
+ The program should only run with this specific version of \GLOBES
+\item
+ The program can only run with a minimum version of \GLOBES
+\item
+ The program can only run up to a certain \GLOBES\ version.
+\end{itemize}
+The same holds for \AEDL\ files: For example, 
+some features may not be supported by earlier versions of \GLOBES\ anymore.
+The program can then check the version of the \AEDL\ file and break if it is
+too old.
+
+The functions in \GLOBES\ for version control are:
+\begin{function}
+\GLBNS{glbTestReleaseVersion}
+{\tt int glbTestReleaseVersion(const char *version)} returns $0$ if
+the {\tt version} string of the format ``X.Y.Z'' is exactly the
+used \GLOBES\ version, $1$ if it is newer, and $-1$ if it is older. 
+\end{function}
+\begin{function}
+\GLBNS{glbTestLibraryVersion}
+{\tt int glbTestLibraryVersion(const char *version)} returns $0$ if
+the {\tt version} string of the format ``X.Y.Z'' is exactly the
+used \GLOBES\ version, $1$ if it is newer, and $-1$ if it is older.
+Note that the library and \GLOBES\ versions are not the same. 
+\end{function}
+\begin{function}
+\GLBNS{glbTestLibraryVersion}
+{\tt int glbTestLibraryVersion(const char *version)} returns $0$ if
+the {\tt version} string of the format ``X.Y.Z'' is exactly the
+used \GLOBES\ version, $1$ if it is newer, and $-1$ if it is older.
+Note that the library and \GLOBES\ versions are not the same. 
+\end{function}
+\begin{function}
+\GLBNS{glbVersionOfExperiment} 
+{\tt const char* glbVersionOfExperiment(int experiment)} returns
+the version string of the experiment number {\tt experiment}.
+The version string points to the experiment structure, which means
+that it cannot be altered and does not have to be destroyed by the
+user.
+\end{function}
+
 \chapter[Calculating $\chi^2$ with systematics only]{Calculating $\boldsymbol{\chi^2}$ with systematics only}
@@
 
-\bi
-\item
- Obtaining rate vectors
-\item
- Obtaining fluxes/cross sections etc.
-\item
- ``Check''-functions
-\ei
+In this chapter, we discuss possibilities to obtain low-level information
+in \GLOBES , \ie, about oscillation probabilities, rates, and other
+information lower than on the $\chi^2$-level.
+
+\section{Oscillation probabilities}
+\index{Oscillation probabilities}
+
+\GLOBES\ can compute the probabilities in vacuum
+with the following function:
+\begin{function}
+\GLBNS{glbVacuumProbability}
+{\tt double glbVacuumProbability(int l, int m, int panti,double E, double L)} returns the neutrino oscillation probability $\nu_l \rightarrow \nu_m$ for the energy {\tt E} and the baseline {\tt L} in vacuum.
+The parameter
+{\tt panti} is $+1$ for neutrinos and $-1$ for antineutrinos. 
+\end{function}
+In addition, the oscillation probabilities in matter can be obtained
+with:
+\begin{function}
+\GLBNS{glbProfileProbability}
+{\tt double glbProfileProbability(int l, int m, int panti,double E)}
+ returns the neutrino oscillation probability $\nu_l \rightarrow \nu_m$ for the energy {\tt E} in matter. The parameter
+{\tt panti} is $+1$ for neutrinos and $-1$ for antineutrinos.
+The matter density profile including baseline is the one from the last
+evaluated experiment. 
+\end{function}
+THIS FUNCTION MAY OBTAIN AN EXPERIMENT INDEX (WW).
+\section{Event rates}
+\index{Event rates}
+One can also return event rates in \GLOBES , but this feature
+requires some knowledge about the experiment definition. \GLOBES\
+currently supports rule-based and channel-based event rate functions,
+where the information is in most cases written into a file.
+The rule-based functions are:
+\begin{function}
+\GLBNS{glbShowRuleRates}
+{\tt int glbShowRuleRates(FILE* stream, int exp, int rule, int pos,
+int effi, int bgi, int coeffi, int signal)} prints the binned rule rates as
+list with energy and event rate to the file {\tt stream} (either an
+open file, or {\tt stdout}). A specific experiment {\tt exp} and a 
+specific rule {\tt rule} have to be chosen, as well as the signal
+or background rate {\tt signal} (either \GLB{GLB\_SIG} or \GLB{GLB\_BG}).
+The position {\tt pos} refers to the component within the signal or 
+background, and can also be {\tt GLB\_ALL}. The function may return
+the rates with (\GLB{GLB\_W\_COEFF}) or without (\GLB{GLB\_WO\_COEFF})
+overall efficieny coefficient, as it is specified by {\tt coeffi}. 
+In addition, it may contain the post-smearing efficiencies (set
+{\tt effi} to \GLB{GLB\_W\_EFF} or \GLB{GLB\_WO\_EFF}), and the
+post-smearing backgrounds (set
+{\tt bgi} to \GLB{GLB\_W\_BG} or \GLB{GLB\_WO\_BG}). The pre-smearing
+efficiencies and backgrounds cannot be accessed at the rule level.
+The return value
+is $0$ if successful, and $-1$ if unsuccessful.
+\end{function}
+\begin{function}
+\GLBNS{glbShowTotalRuleRates}
+{\tt double glbShowTotalRuleRates(int exp, int rule, int pos,
+int effi, int bgi, int coeffi, int signal)} returns the total rates with
+the same parameters as {\tt glbShowRuleRates}.
+\end{function}
+(THIS SYNTAX NEEDS TO BE CHECKED, BECAUSE FUNCTION IS NOT YET IMPLEMENTED - WW)
+
+The function {\tt glbShowTotalRuleRates} is especially useful if
+one wants to draw bi-rate graphs with total event rates, or look
+for the $(\deltacp,\theta_{13})$-degeneracy by the intersection of 
+neutrino and antineutrino constant event rate curves.
+
+In addition, \GLOBES\ has a channel-based rate function:
+\begin{function}
+\GLBNS{glbShowChannelRates}
+{\tt int glbShowChannelRates(FILE *stream,
+int exp, int channel, int smearing, int effi, int bgi)}
+prints the binned channel rates as
+list with energy and event rate to the file {\tt stream} (either an
+open file, or {\tt stdout}). A specific experiment {\tt exp} and a 
+specific channel {\tt channel} have to be chosen.
+The function may return
+the rates before (\GLB{GLB\_PRE}) or after (\GLB{GLB\_POST})
+the energy smearing, as it is specified by {\tt smearing}.
+In addition, it may contain the pre- and post-smearing efficiencies (set
+{\tt effi} to \GLB{GLB\_W\_EFF} or \GLB{GLB\_WO\_EFF}), and the
+pre- and post-smearing backgrounds (set
+{\tt bgi} to \GLB{GLB\_W\_BG} or \GLB{GLB\_WO\_BG}). Note that
+the post-smearing efficiencies and backgrounds cannot be taken into
+account if the rates are returned before the energy smearing.
+The return value
+is $0$ if successful, and $-1$ if unsuccessful.
+\end{function}
+
+\section{Fluxes and cross sections}
+\index{Flux} \index{Cross section}
+
+Another piece of low-level information, which can be returned by \GLOBES ,
+are the numbers from the loaded fluxes and cross sections.
+The following functions interpolate on the loaded fluxes and cross 
+sections, \ie, any value in the allowed energy range can be given as input:
+\begin{function}
+\GLBNS{glbFlux}
+{\tt double glbFlux(int exp, int ident, 
+double E, double distance, int l, int anti)} returns
+the flux of flux number {\tt ident} of the experiment {\tt exp}
+for the flavor $\nu_l$ 
+and polarity {\tt anti} ($+1$: neutrinos, $-1$: antineutinos) at the energy {\tt E} and distance {\tt distance}.
+\end{function}
+
+\begin{function}
+\GLBNS{glbXSection}
+{\tt double glbXSection(int ident, double E, int l, int anti)} returns
+the cross section of cross section number {\tt ident} for the flavor $\nu_l$ 
+and polarity {\tt anti} ($+1$: neutrinos, $-1$: antineutinos) at the energy {\tt E}.
+\end{function}
+Note that all flux or flavor numbers run from $0$ to $N-1$ in these functions.
+
+(THIS FUNCTION MAY GET AN EXPERIMENT INDEX IF NOT THROUGH-COUNTED IN AEDL!)
 

### 2004-07-08 Documentation/part1.tex
*** empty log message ***

@@
 \section{Units in \GLOBES\ and the integrated luminosity}
+\label{sec:luminosity}
 
@@
 One can also return event rates in \GLOBES , but this feature
-requires some knowledge about the experiment definition. \GLOBES\
-currently supports rule-based and channel-based event rate functions,
-where the information is in most cases written into a file.
-The rule-based functions are:
+requires some knowledge about the experiment definition. 
+In fact, many of these functions are very advanced, which means
+that the reader who wants to use them should be familiar with
+\Secs~\ref{sec:channel} and~\Sec~\ref{sec:rules} of the \AEDL\ manual.
+
+\GLOBES\ currently supports rule-based and channel-based event rate functions, where the information is in written into a file
+or returned in a list. The rule-based functions are:
 \begin{function}
@@
 \begin{function}
-\GLBNS{glbShowTotalRuleRates}
-{\tt double glbShowTotalRuleRates(int exp, int rule, int pos,
+\GLBNS{glbTotalRuleRate}
+{\tt double glbTotalRuleRate(int exp, int rule, int pos,
 int effi, int bgi, int coeffi, int signal)} returns the total rates with
@@
 \end{function}
-(THIS SYNTAX NEEDS TO BE CHECKED, BECAUSE FUNCTION IS NOT YET IMPLEMENTED - WW)
-
-The function {\tt glbShowTotalRuleRates} is especially useful if
+The function {\tt glbTotalRuleRate} is especially useful if
 one wants to draw bi-rate graphs with total event rates, or look
@@
 neutrino and antineutrino constant event rate curves.
-
-In addition, \GLOBES\ has a channel-based rate function:
+%
+In order to obtain information on the structure of the rules, 
+a number of additional functions are provided:
+\begin{function}
+\GLBNS{glbGetNumberOfRules}
+{\tt int glbGetNumberOfRules(int exp)} returns the number of
+rules in experiment {\tt exp}.
+\end{function}
+\begin{function}
+\GLBNS{glbGetLengthOfRule}
+{\tt int glbGetLengthOfRule(int exp, int rule, int signal)} returns
+the length of rule {\tt rule} in experiment {\tt exp}. The parameter
+{\tt signal} can be either \GLB{GLB\_SIG} for the number of signal
+components or \GLB{GLB\_BG} for the number of background components.
+\end{function}
+\begin{function}
+\GLBNS{glbGetNormalizationInRule}
+{\tt double glbGetNormalizationInRule(int exp, int rule, int signal)}
+returns the normalization (normalization or background center)
+in rule {\tt rule} of the experiment {\tt exp}.
+The parameter {\tt signal} refers to signal (\GLB{GLB\_SIG}) or background
+(\GLB{GLB\_BG}).
+\end{function}
+\begin{function}
+\GLBNS{glbGetChannelInRule}
+{\tt int glbGetChannelInRule(int exp, int rule, int pos, int signal)}
+returns the channel number in rule {\tt rule} at the position {\tt pos}
+of the experiment {\tt exp}.
+The parameter {\tt signal} refers to signal (\GLB{GLB\_SIG}) or background
+(\GLB{GLB\_BG}).
+\end{function}
+\begin{function}
+\GLBNS{glbGetCoefficientInRule}
+{\tt double glbGetCoefficientInRule(int exp, int rule, int pos, int signal)}
+returns the coefficient of the component {\tt pos} in rule {\tt rule} 
+of the experiment {\tt exp}.
+The parameter {\tt signal} refers to signal (\GLB{GLB\_SIG}) or background
+(\GLB{GLB\_BG}).
+\end{function}
+
+In addition, \GLOBES\ has channel-based rate functions:
 \begin{function}
@@
 \end{function}
+\begin{function}
+\GLBNS{glbGetChannelRates}
+{\tt int glbGetChannelRates(double** data,
+size\_t* length, int exp, int channel, int smearing)}
+writes the binned raw channel rates to the list {\tt data}
+and the length of this list to {\tt length}. 
+A specific experiment {\tt exp} and a 
+specific channel {\tt channel} have to be chosen.
+The function may return
+the rates before ({\tt smearing} is \GLB{GLB\_PRE}) or after ({\tt smearing} is \GLB{GLB\_POST})
+the energy smearing, where no user-defined data
+(pre-/post-smearing efficiencies or backgrounds) are taken into account.
+The return value is $-1$ if unsuccessful.
+\end{function}
+\begin{function}
+\GLBNS{glbGetUserData}
+{\tt int glbGetUserData(double** data,
+size\_t* length, int exp, int channel, int smearing, int bgeff)}
+writes the binned user-defined backgrounds or efficiencies 
+to the list {\tt data} and the length of this list to {\tt length}. 
+A specific experiment {\tt exp} and a 
+specific channel {\tt channel} have to be chosen.
+The function may return
+the pre- ({\tt smearing} is \GLB{GLB\_PRE}) or post- ({\tt smearing} is \GLB{GLB\_POST}) smearing backgrounds ({\tt bgeff} is \GLB{GLB\_BG}) 
+or efficiencies ({\tt bgeff} is \GLB{GLB\_EFF}). 
+The return value is $-1$ if unsuccessful.
+\end{function}
+Since \GLOBES\ reserved the memory for the lists returned in these
+functions, which it allocates on an internal stack, one has to reset
+the stack at the end of the rates access with
+\begin{function}
+\GLBNS{glbResetRateStack}
+{\tt void glbResetRateStack()} resets the rate stack used for the
+lists returned from {\tt glbGetChannelRates} or {\tt glbGetUserData}.
+\end{function}
+A code excerpt to show the channel rates may look like this:
+\begin{quote}
+{\tt
+ double* rates; \\
+  size\_t length; \\
+  glbGetChannelRates(\&rates,\&length,0,0,GLB\_PRE); \\
+  int i; \\
+  for(i=0;i<length;i++) printf("\%g $\backslash$n",rates[i]); \\
+  glbResetRateStack(); \\
+}
+\end{quote}
+Finally, one can find the number of channels of an experiment:
+\begin{function}
+\GLBNS{glbGetNumberOfChannels}
+{\tt int glbGetNumberOfChannels(int exp)} returns the number of 
+channels of experiment {\tt exp}.
+\end{function}
 
@@
 \GLBNS{glbXSection}
-{\tt double glbXSection(int ident, double E, int l, int anti)} returns
-the cross section of cross section number {\tt ident} for the flavor $\nu_l$ 
-and polarity {\tt anti} ($+1$: neutrinos, $-1$: antineutinos) at the energy {\tt E}.
+{\tt double glbXSection(int exp, 
+int ident, double E, int l, int anti)} returns
+the cross section of experiment {\tt exp}, 
+cross section number {\tt ident} for the flavor $\nu_l$ and polarity {\tt anti} ($+1$: neutrinos, $-1$: antineutinos) at the energy {\tt E}.
 \end{function}
 Note that all flux or flavor numbers run from $0$ to $N-1$ in these functions.
-
-(THIS FUNCTION MAY GET AN EXPERIMENT INDEX IF NOT THROUGH-COUNTED IN AEDL!)
 
@@
 
+Many of the parameters in experiment definitions can be changed
+at running time. For example, we have introduced in 
+\Sec~\ref{sec:luminosity} possibilities to change the integrated
+luminosity, which consists of source power, running time, and target mass.
+In this chapter, we discuss more sophisticated
+experiment changes. However, since \GLOBES\ computes a lot of
+information only once when an experiment is loaded, many
+parameters can not be changed at running time. For example,
+the energy resolution function or the number of bins are used
+to compute the smearing matrix already at the initialization
+of the experiment, which saves a lot of computation
+time for most applications. In \Sec~\ref{sec:aedlparams}, we 
+introduce a mechanism how one can even change these \AEDL\ parameters
+during running time.
+
+\section{Baseline and matter density profile}
+
+In order to change the baseline of an experiment, it is important
+to keep in mind that each experiment has a profile type defined
+in the \AEDL\ file (average density, PREM profile with a given
+number of steps, or arbitrary profile). One can check the currently
+used profile type with
+\begin{function}
+\GLBNS{glbGetProfileType}
+{\tt int glbGetProfileType(int exp)} returns the matter density profile
+type of experiment {\tt exp}.
+\end{function}
+For each profile type, one can easily change the baseline with {\tt glbSetBaselineInExperiment},
+where the average density or the PREM profile are re-computed, or the
+steps in the arbitrary profile are re-scaled. Of course, though this
+function is sufficient in most cases, it does often not make sense
+for arbitrary matter density profiles.
+\begin{function}
+\GLBNS{glbSetBaselineInExperiment}
+{\tt int glbSetBaselineInExperiment(int exp, double baseline)}
+sets the baseline length in experimente {\tt exp} to {\tt baseline}.
+The function returns $-1$ if it was not successful.
+\end{function}
+Note that {\tt glbSetBaselineInExperiment} does not change the
+profile type in the experiment. The counterpart of this function is:
+\begin{function}
+\GLBNS{glbGetBaselineInExperiment} 
+{\tt double glbGetBaselineInExperiment(int exp)} returns the
+baseline length currently used for experiment {\tt exp}.
+\end{function}
+One can not change the profile type of an experiment manually
+during running time. However, one can change the matter density
+profile, where the profile type is automatically changed to three
+(arbitrary matter density profile). In addition, a number of functions 
+are provided to compute possible matter density profiles (average density,
+PREM profile). In general, a matter density profile in \GLOBES\ with
+$N$ layers is represented by a list of lengths 
+\begin{equation}
+\mathrm{Lengths} = (x_1,x_2, \hdots, x_N) 
+\end{equation}
+and a list of densities
+\begin{equation}
+\mathrm{Densities} = (\rho_1,\rho_2, \hdots, \rho_N), 
+\end{equation}
+where the baseline is given by
+\begin{equation}
+L = \sum\limits_{i=1}^N x_i.
+\end{equation}
+In C, lists are represented as pointers to the first element:
+\begin{quote}
+{\tt  
+  double* lengths; \\
+  double* densities;
+}
+\end{quote}
+Many of the \GLOBES\ baseline functions take and return
+such lists as parameters, and are therefore more sophisticated
+to handle. In general, any function
+{\em returning} lists reserves the memory space for them.
+It is then up to the user to return the allocated memory!
+In addition, they normally return the length of the lists $N$
+in form of a pointer, where the corresponding memory space 
+has to be provided by the user (normally, it is enough to declare
+a variable of the type {\tt size\_t} corrsponding to positive
+integer numbers).
+The following functions return matter density profiles:
+\begin{function}
+\GLBNS{glbLoadProfileData}
+{\tt int glbLoadProfileData(const char* filename, size\_t *layers, double **lengths, double **densities)} loads a density file from the file
+{\tt filename}. It returns the number of layers {\tt layers}, the
+list of lengths {\tt lengths}, and the list of densities {\tt densities}.
+\end{function}
+The file should contain in each line a length and density for one layer,
+which are separated by an empty space.
+\begin{function}
+\GLBNS{glbStaceyProfile}
+{\tt int glbStaceyProfile(double baseline, size\_t layers, double **lengths, double **densities)} creates a PREM/Stacey matter density profile with a
+number of {\tt layers} steps for the baseline {\tt baseline}. The list of lengths {\tt lengths} and the list of densities {\tt densities} are returned.
+\end{function}
+\begin{function}
+\GLBNS{glbAverageDensityProfile}
+{\tt glbAverageDensityProfile(double baseline, double **lengths, 
+double **densities)} creates a average matter density profile from the PREM/Stacey profile with one step for the baseline {\tt baseline}.
+The list of lengths {\tt lengths} and the list of densities {\tt densities} are returned.
+\end{function}
+\begin{function}
+\GLBNS{glbStaceyProfile}
+{\tt int glbGetProfileDataInExperiment(int exp,size\_t *layers, double** lengths, double** densities)} returns the matter density profile 
+currently used for experiment {\tt exp}. The number of layers {\tt layers}, the list of lengths {\tt lengths}, and the list of densities {\tt densities} are returned.
+\end{function}
+All these functions return $-1$ if they were not successful.
+
+The counterpart of these functions to assign a specific matter density
+profile to an experiment is
+\begin{function}
+\GLBNS{glbSetProfileDataInExperiment}
+{\tt int glbSetProfileDataInExperiment(int exp, size\_t layers,const double* lengths, const double* densities)} sets the matter density of experiment
+{\tt exp} to an arbitrary profile with {\tt layers} steps. The density
+layers are specified by the lists {\tt lengths} and {\tt densities}.
+The function returns $-1$ if it was not successful.
+\end{function}
+Note that this function requires that the memory space for the lists
+be reserved already.
+
+Finally, let us take a look at two examples. This example changes
+the baseline length to $7\,500 \, \mathrm{km}$, where the average 
+matter density is manually computed:
+\begin{quote}
+{\tt  
+  double* lengths; \\
+  double* densities; \\
+  glbAverageDensityProfile(7500,\&lengths,\&densities); \\
+  glbSetProfileDataInExperiment(0,1,lengths,densities); \\
+  free(lengths); \\
+  free(densities); \\
+}
+\end{quote}
+In the second example, we change the baseline to a PREM profile with
+$100$ matter density steps and print them:
+\begin{quote}
+{\tt
+  double* lengths; \\
+  double* densities; \\
+  glbStaceyProfile(7500,100,\&lengths,\&densities); \\
+  int i; \\
+  for(i=0;i<100;i++) printf("\%g \%g $\backslash$n",lengths[i],densities[i]); \\
+  glbSetProfileDataInExperiment(0,100,lengths,densities); \\
+  free(lengths);\\
+  free(densities);\\
+}
+\end{quote}
+
 \example{The impact of systematics, correlations, and degeneracies}{
-\label{ex:corrproj}
+\label{ex:barcharts}
 \index{Bar charts}
@@
 
+\section{Systematics}
+
+Changing the systematics at running time can be useful to investigate
+the impact factors affecting the measurement.
+In \GLOBES , the sytematics is defined rule-based, \ie, each rule
+has its own systematics. In addition, \AEDL\ requires that it has to be
+defined in each rule what ``Systematics on'' and ``Systematics off'' means.
+Therefore, it is usually very simple to switch the systematics on and off:
+\begin{function}
+\GLBNS{glbSwitchSystematics}
+{\tt int glbSwitchSystematics(int exp, int rule, int on\_off)}
+switches the systematics in experiment {\tt exp} and rule {\tt rule}
+on ({\tt on\_off} is \GLB{GLB\_ON}) or off ({\tt on\_off} is \GLB{GLB\_OFF}). For the experiment or
+rule index, one can also use \GLB{GLB\_ALL}. 
+\end{function}
+In the example on page~\pageref{ex:barcharts}, the application of
+{\tt glbSwitchSystematics} is demonstrated to show the impact of
+systematics, correlations, and degeneracies.
+
+The error dimension (for the definition, see \Sec~\ref{sec:rules}) can also be accessed directly with 
+\begin{function}
+\GLBNS{glbSetErrorDim}
+{\tt int glbSetErrorDim(int exp, int rule, int on\_off, int value)}
+sets the error dimension for systematics on ({\tt on\_off} is \GLB{GLB\_ON}) or off ({\tt on\_off} is \GLB{GLB\_OFF}) of experiment {\tt exp} and rule {\tt rule} to the value {\tt value}.
+The function returns
+$-1$ if not successful.
+\end{function}
+\begin{function}
+\GLBNS{glbGetErrorDim}
+{\tt int glbGetErrorDim(int exp, int rule, int on\_off)}
+returns the error dimension for systematics on ({\tt on\_off} is \GLB{GLB\_ON}) or off ({\tt on\_off} is \GLB{GLB\_OFF}) of experiment {\tt exp} and rule {\tt rule}.
+\end{function}
+
+Except from the general treatment of systematics, one can read out
+and write the signal and background errors, as well as the background
+centers. For the definitions of these quantities, see \Sec~\ref{sec:rules}.
+\begin{function}
+\GLBNS{glbSetSignalErrors}
+{\tt int glbSetSignalErrors(int exp, int rule, double norm, double tilt)}
+sets the signal errors of experiment {\tt exp} and rule {\tt rule}
+to {\tt norm} (normalization error) and {\tt tilt} (tilt/calibration error).
+\end{function}
+\begin{function}
+\GLBNS{glbGetSignalErrors}
+{\tt int glbGetSignalErrors(int exp, int rule, double* norm, double* tilt)}
+writes the signal errors of experiment {\tt exp} and rule {\tt rule}
+to {\tt norm} (normalization error) and {\tt tilt} (tilt/calibration error).
+\end{function}
+\begin{function}
+\GLBNS{glbSetBGErrors}
+{\tt int glbSetBGErrors(int exp, int rule, double norm, double tilt)}
+sets the background errors of experiment {\tt exp} and rule {\tt rule}
+to {\tt norm} (normalization error) and {\tt tilt} (tilt/calibration error).
+\end{function}
+\begin{function}
+\GLBNS{glbGetBGErrors}
+{\tt int glbGetBGErrors(int exp, int rule, double* norm, double* tilt)}
+writes the background errors of experiment {\tt exp} and rule {\tt rule}
+to {\tt norm} (normalization error) and {\tt tilt} (tilt/calibration error).
+\end{function}
+\begin{function}
+\GLBNS{glbSetBGCenters}
+{\tt int glbSetBGCenters(int exp, int rule, double norm, double tilt)}
+sets the background centers of experiment {\tt exp} and rule {\tt rule}
+to {\tt norm} (normalization center) and {\tt tilt} (tilt/calibrationcenter).
+\end{function}
+\begin{function}
+\GLBNS{glbGetBGCenters}
+{\tt int glbGetBGCenters(int exp, int rule, double* norm, double* tilt)}
+writes the background centers of experiment {\tt exp} and rule {\tt rule}
+to {\tt norm} (normalization center) and {\tt tilt} (tilt/calibration center).
+\end{function}
+As usual, all these functions return $-1$ if they were not successful.
+
+\section{External parameters in \AEDL\ files}
+\label{sec:aedlparams}
+
 \bi
@@
 \ei
+
+\section{Algorithm parameters: Filter functions}
+
+Here come the filter functions

### 2004-07-09 Documentation/part1.tex
*** empty log message ***

@@
 configuration, we refer to the {\tt README} file in the \GLOBES\ main directory, and the comments in the Makefile of the {\tt Example} subdirectory for how to complile the example files.
+Make sure that
+the data files (\AEDL\ and supporting files) in the {\tt data} 
+subdirectory can be found by the
+examples. In case of doubt, you can simply copy the necessary files
+into the working directory where the examples are executed.
+
+We will in this part not go into details of the experiment
+definition. The pre-defined experiment prototypes in the {\tt data}
+subdirectory are summarized in \tabl{experiments}. They correspond
+(except from minor modifications) to the experiments in the
+respective references in the table.
 
@@
 \begin{center}
-\begin{tabular}{llp{7cm}c}
+\begin{tabular}{llp{7.5cm}c}
 \hline
-Experiment & Filename & Short description & Refs. \\
+Experiment & Filename & Short description &  Ref. \\
 \hline 
-\multicolumn{3}{l}{\underline{Conventional beams:}} \\
-??? & & \\[0.1cm]
-
-\multicolumn{3}{l}{\underline{First-generation superbeams:}} \\
-\JHFSK\ ($\nu$) & {\tt JHFSK.exp} & JHF (J-PARC) to Super-Kamiokande, neutrino running &  \cite{Huber:2002mx,Huber:2002rs} \\
-\JHFSK\ ($\bar\nu$)& {\tt JHFSKanti.exp} & JHF (J-PARC) to Super-Kamiokande, antineutrino running &  \cite{Huber:2002rs} \\
-\NUMI\  ($\nu$), OA $9 \, \mathrm{km}$ & {\tt NUMI9.exp} & NuMI with off-axis angle of $9 \, \mathrm{km}$ for $L=712 \, \mathrm{km}$, neutrino running & \cite{Huber:2002rs} \\
-\NUMI\  ($\bar{\nu}$), OA $9 \, \mathrm{km}$ & {\tt NUMI9anti.exp} & NuMI with off-axis angle of $9 \, \mathrm{km}$ for $L=712 \, \mathrm{km}$, antineutrino running & \cite{Huber:2002rs} \\
-\NUMI\  ($\nu$), OA $12 \, \mathrm{km}$ & {\tt NUMI12.exp} & NuMI with off-axis angle of $12 \, \mathrm{km}$ for $L=712 \, \mathrm{km}$, neutrino running & \cite{Huber:2002rs} \\
-\NUMI\  ($\bar{\nu}$), OA $12 \, \mathrm{km}$ & {\tt NUMI12anti.exp} & NuMI with off-axis angle of $12 \, \mathrm{km}$ for $L=712 \, \mathrm{km}$, antineutrino running & \cite{Huber:2002rs} \\
-\SPL\  ($\nu$) & {\tt SPL.exp} & SPL (CERN), neutrino running &  ??? \\
-\SPL\  ($\bar\nu$) & {\tt SPLanti.exp} & SPL (CERN), antineutrino running & ??? \\[0.1cm]
-
-\multicolumn{3}{l}{\underline{Superbeam upgrades:}} \\
-\JHFHK\ ($\nu$) & {\tt JHFHK.exp} & JHF (J-PARC) to Hyper-Kamiokande superbeam upgrade, neutrino running &  \cite{Huber:2002mx,Huber:2002rs} \\
-\JHFHK\ ($\bar\nu$)& {\tt JHFHKanti.exp} & JHF (J-PARC) to Hyper-Kamiokande superbeam upgrade, antineutrino running &  \cite{Huber:2002mx,Huber:2002rs} \\[0.1cm]
-
-\multicolumn{3}{l}{\underline{Neutrino factories:}} \\
-\NuFactI\ & {\tt NuFact.exp} & Initial stage neutrino factory, symmetric operation in both polarities & \cite{Huber:2002mx} \\
-\NuFactII\  & {\tt NuFact2.exp} & Advanced stage neutrino factory, symmetric operation in both polarities & \cite{Huber:2002mx,Huber:2003ak} \\[0.1cm]
-
-\multicolumn{3}{l}{\underline{Reactor experiments:}} \\
-\ReactorI\ & {\tt Reactor.exp} & Small reactor experiment with identical near and far detectors & \cite{Huber:2003pm} \\
-\ReactorII\ & {\tt Reactor2.exp} & Large reactor experiment with identical near and far detectors & \cite{Huber:2003pm} \\[0.1cm]
-
-\multicolumn{3}{l}{\underline{$\beta$-Beams:}} \\
-\Beta\ ($\nu$) & {\tt BETA.exp} & $\beta$-Beam, neutrino running & ??? \\
-\Beta\ ($\bar\nu$) & {\tt BETAanti.exp} & $\beta$-Beam, antineutrino running & ??? \\
+\multicolumn{4}{l}{\underline{Conventional beams:}} \\
+MINOS & {\tt MINOS.glb} & MINOS exp., 5~yr running time & \cite{Huber:2004ug}\\
+OPERA & {\tt OPERA.glb} & OPERA exp., 5~yr running time & \cite{Huber:2004ug}\\
+ICARUS & {\tt ICARUS.glb} & ICARUS exp., 5~yr running time & \cite{Huber:2004ug}\\[0.1cm]
+
+\multicolumn{4}{l}{\underline{First-generation superbeams:}} \\
+T2K & {\tt JHFSKnew.glb} & J-PARC to Super-K, 5~yr $\nu$-running &  \cite{Huber:2002rs} \\
+ & {\tt JHFSKantinew.glb} & J-PARC to Super-K, 5~yr $\bar{\nu}$-running &  \cite{Huber:2002rs} \\
+ & {\tt JHFSKcomb.glb} & Same, but 1.25~yr $\nu$- and 3.75~yr $\bar{\nu}$-running &  \cite{Huber:2002rs} \\[0.1cm]
+NO$\nu$A & {\tt NUMI9.glb} & NuMI OA $9 \, \mathrm{km}/712 \, \mathrm{km}$, 5~yr $\nu$-running & \cite{Huber:2002rs} \\
+& {\tt NUMI9anti.glb} & NuMI OA $9 \, \mathrm{km}/712 \, \mathrm{km}$, 5~yr $\bar{\nu}$-running & \cite{Huber:2002rs} \\
+& {\tt NUMI9comb.glb} & NuMI OA $9 \, \mathrm{km}/712 \, \mathrm{km}$, 1.43~yr $\nu$- and 3.57~yr $\bar{\nu}$-running & \cite{Huber:2002rs} \\
+& {\tt NUMI12.glb} & NuMI OA $12 \, \mathrm{km}/712 \, \mathrm{km}$, 5~yr $\nu$-running & \cite{Huber:2002rs} \\
+& {\tt NUMI12anti.glb} & NuMI OA $12 \, \mathrm{km}/712 \, \mathrm{km}$, 5~yr $\bar{\nu}$-running & \cite{Huber:2002rs} \\
+& {\tt NUMI12comb.glb} & NuMI OA $12 \, \mathrm{km}/712 \, \mathrm{km}$, 1.43~yr $\nu$- and 3.57~yr $\bar{\nu}$-running & \cite{Huber:2002rs} \\[0.1cm]
+
+\multicolumn{4}{l}{\underline{Superbeam upgrade:}} \\
+J-PARC-HK & {\tt JHFHKAll.glb} & J-PARC to Hyper-K, 2~yr $\nu$- and 6~yr $\bar{\nu}$-running &  \cite{Huber:2002mx} \\[0.1cm]
+
+\multicolumn{4}{l}{\underline{Neutrino factories:}} \\
+\NuFactI\ & {\tt NuFact1.glb} & Initial stage NF, 2$\times$2.5~yr running time (each pol.), $m_{\mathrm{Det}} = 10 \, \mathrm{kt}$,
+$P_{\mathrm{Tg}} \simeq 0.75 \, \mathrm{MW}$ & \cite{Huber:2002mx} \\
+\NuFactII\  & {\tt NuFact2.glb} & Advanced stage NF, 2$\times$4~yr running time (each pol.), $m_{\mathrm{Det}} = 50 \, \mathrm{kt}$,
+$P_{\mathrm{Tg}} \simeq  4 \, \mathrm{MW}$ & \cite{Huber:2002mx} \\[0.1cm]
+
+\multicolumn{4}{l}{\underline{Reactor experiments:}} \\
+\ReactorI\ & {\tt Reactor1.glb} & Small reactor exp., $\mathcal{L} = 400 \, \mathrm{t} \, \mathrm{GW} \, \mathrm{yr}$ & \cite{Huber:2003pm} \\
+\ReactorII\ & {\tt Reactor2.glb} & Large reactor exp., $\mathcal{L} = 8\, 000 \, \mathrm{t} \, \mathrm{GW} \, \mathrm{yr}$ & \cite{Huber:2003pm} \\[0.1cm]
+
+\underline{$\beta$-Beams:} & \multicolumn{3}{l}{In preparation} \\[0.1cm]
 \hline
@@
 \end{center}
-\mycaption{\label{tab:experiments} \index{Experiment data files (table)} NEEDS TO BE ADJUSTED WHEN PROTOTYPES ARE AVAILABLE!
-Different pre-defined experiments, their filenames (to be used in {\tt LoadExperiment}), their short description, and the references in which they are defined.
-Note that all experiments use in their standard configurations one year of running time.
-Details about the experiment parameters can be obtained with {\tt InfoExperiment}(Experiment number) after they have been loaded.}
+\mycaption{\label{tab:experiments} \index{Experiment data files (table)} Pre-defined experiment prototypes, their filenames (to be used in \GLB{glbInitExperiment}), their short descriptions, and the references in which they are originally used and discussed (except from minor modifications, such as a
+different implementation of the energy threshold function). Note that
+some of these experiments are outdated in terms of integrated luminosities,
+baseline, fluxes, efficiencies, or other factors.}
 \end{table}
@@
 the {\tt version} string of the format ``X.Y.Z'' is exactly the
-used \GLOBES\ version, $1$ if it is newer, and $-1$ if it is older. 
+used \GLOBES\ version, $1$ if it is older, and $-1$ if it is newer. 
 \end{function}
@@
 the {\tt version} string of the format ``X.Y.Z'' is exactly the
-used \GLOBES\ version, $1$ if it is newer, and $-1$ if it is older.
+used \GLOBES\ version, $1$ if it is older, and $-1$ if it is newer.
 Note that the library and \GLOBES\ versions are not the same. 
@@
 
-\bi
-\item
- Baseline, Target mass, Source power, running time
-\item
- Matter density profile
-\item
- Systematics
-\item
- Threshold function, efficiencies etc.
-\ei
+Using external parameters in \AEDL\ files is a very powerful feature
+to change experiment parameters at running time which require that
+the experiment be re-initialized. For example, one can change the
+energy resolution function or the number of energy bins. However,
+in some cases there might be complications, such that the number
+of pre- or post-smearing efficiencies does not correspond to the number
+of energy bins anymore. Therefore, this feature needs to be 
+used with care.
+
+In order to use external parameters in \AEDL\ files, one simply
+introduces them. For example, an energy resolution function
+\begin{quote}
+{\tt
+energy(\#EnergyResolution1)< \\
+\hspace*{1cm} type = 1 \\
+\hspace*{1cm} @sigma\_e = \{ myres ,0,0 \} \\
+> \\
+}
+\end{quote}
+might be defined in \AEDL , where the energy resolution is proportional
+to {\tt myres} $\times$ energy.
+
+In order to use the user-defined variable, one has to assign it 
+with {\tt glbDefineAEDLVariable} {\em before} the experiment is initialized with \GLB{glbInitExperiment}:
+\begin{function}
+\GLBNS{glbDefineAEDLVariable}
+{\tt void glbDefineAEDLVariable(const char* name, double value)}
+assigns the value {\tt value} to the \AEDL\ variable {\tt name}.
+\end{function}
+In our energy resolution example, one could now loop over the
+energy resolution such as with
+\begin{quote}
+{\tt
+int i; \\
+for(i=5;i<20;i++) \\
+\{ \\	 
+\hspace*{1cm} glbClearExperimentList(); \\
+\hspace*{1cm} glbDefineAEDLVariable("myres",0.01*i); \\
+\hspace*{1cm} glbInitExperiment(...); \\
+\\
+\hspace*{1cm} /* do something */ \\
+\}
+}
+\end{quote}
+Note that one does not have do re-initialize the oscillation
+parameter vectors every time within the loop as long as the
+number of experiments does not change.
+
+In order to clear the external variable stack if one is
+excessively using it, one can use
+\begin{function}
+\GLBNS{glbClearAEDLVariables}
+{\tt void glbClearAEDLVariables()}
+clears the \AEDL\ variable list.
+\end{function}
+This function is called automatically upon exit of the program.
 
@@
 
-Here come the filter functions
+The oscillation frequency filters to filter fast oscillations
+can also be accessed by the user interface. For details of
+the filter functions, we refer to \Sec~\ref{sec:energy} of
+the \AEDL\ manual.
+
+In particular, there are a number of functions:
+\begin{function}
+\GLBNS{glbSetFilterState}
+{\tt int glbSetFilterState(int on\_off)} sets the filter state to
+on (\GLB{GLB\_ON}) or off (\GLB{GLB\_OFF}).
+\end{function}
+\begin{function}
+\GLBNS{glbGetFilterState}
+{\tt int glbGetFilterState()} returns the filter state.
+\end{function}
+\begin{function}
+\GLBNS{glbSetFilterStateInExperiment}
+{\tt int glbSetFilterStateInExperiment(int exp, int on\_off)} sets the filter state in experiment {\tt exp} to on (\GLB{GLB\_ON}) or off (\GLB{GLB\_OFF}).
+\end{function}
+\begin{function}
+\GLBNS{glbGetFilterStateInExperiment}
+{\tt int glbGetFilterStateInExperiment(int exp)} returns the filter state of experiment {\tt exp}.
+\end{function}
+Analogously, the filter value can be accessed:
+\begin{function}
+\GLBNS{glbSetFilter}
+{\tt int glbSetFilter(double filter)} sets the filter to the
+value {\tt filter}.
+\end{function}
+\begin{function}
+\GLBNS{glbGetFilter}
+{\tt double glbGetFilter()} returns the filter value.
+\end{function}
+\begin{function}
+\GLBNS{glbSetFilterInExperiment}
+{\tt int glbSetFilterInExperiment(int exp, double filter)} sets the filter  in experiment {\tt exp} to the value {\tt value}.
+\end{function}
+\begin{function}
+\GLBNS{glbGetFilterInExperiment}
+{\tt double glbGetFilterInExperiment(int exp)} returns the filter value of experiment {\tt exp}.
+\end{function}
+The return value of all {\tt Set-} functions is $-1$ if they were not successful.

### 2004-07-09 Documentation/Manual.tex
*** empty log message ***
GLoBES tour added

@@
 \usepackage{makeidx}
-\usepackage{ae,aecompl} % only for Type 1 fonts
+\usepackage{ae,aecompl} % only for Type 1 fonts for readable pdf convert
 
@@
 % Always use GLB when mentioning a command in text!
-\newcommand{\GLB}[1]{{\tt #1} \index{{\tt #1}}}
+\newcommand{\GLB}[1]{{\tt #1}\index{{\tt #1}}}
+% Always use GLBNS when adding a command to the index without showing it!
+\newcommand{\GLBNS}[1]{\index{{\tt #1}}}
+\newcommand{\gq}[1]{\begin{quote} {\tt #1} \end{quote}}
+\newcommand{\go}[1]{$\rightarrow$ Output: \begin{quote} {\sf #1} \end{quote}}
 
@@
 {\large \bf User's and reference manual }
+
+\vspace*{1.5cm}
+{\large \bf !!!DRAFT IN PROGRESS!!! }
 \end{center}
@@
 one hand, it contains a comprehensive abstract experiment definition
-language, which allows to describe most classes of long baseline experinents
+language, which allows to describe most classes of long baseline experiments
 at an abstract level. On the other hand, it provides a C-library to 
@@
 Applications of the projection mechanism include the projections onto the $\stheta$-axis and the $\stheta$-$\deltacp$-plane.
-In addition, all oscillation parameters can be kept free to precicely localize
+In addition, all oscillation parameters can be kept free to precisely localize
 degenerate solutions.
@@
 as possible. In addition, changes in experimental parameters, such as
-efficiences or the energy resolutions, can quickly be tested.
+efficiencies or the energy resolutions, can quickly be tested.
 %
@@
 function. Eventually, it is demonstrated how one can change many
-experiment parameters at running time (such as baseline or target mass), and how one can ontain low-level
+experiment parameters at running time (such as baseline or target mass), and how one can obtain low-level
 information. We recommend that everybody interested in \GLOBES\ should
@@
 
-Finally, \Part~\ref{part:3} of this manual serves as reference manual,
-and the appendix contains further supplementary information.
+%Finally, \Part~\ref{part:3} of this manual serves as reference manual,
+%and the appendix contains further supplementary information.
 
@@
 
-\input{part3}
+% \input{part3}
 

### 2004-07-09 Documentation/part1.tex
*** empty log message ***

@@
 \label{chapter:tour}
+\index{GLoBES tour}
 
@@
 
+\index{Installation}
 In this first chapter of the user's manual, we assume that the \GLOBES\ software is readily installed on your computer system. For the installation,
@@
 \index{Examples} Since it depends on your installed
-configuration, we refer to the {\tt README} file in the \GLOBES\ main directory, and the comments in the Makefile of the {\tt Example} subdirectory for how to complile the example files.
-Make sure that
+configuration, we refer to the {\tt README} file in the \GLOBES\ main directory, and the comments in the Makefile of the {\tt Example} subdirectory for how to compile the example files. Make sure that
 the data files (\AEDL\ and supporting files) in the {\tt data} 
@@
 \index{C-Code}
-Here comes the C-code sceleton, which is
+\index{Program}
+Here comes the C-code skeleton, which is
 (more or less) common to all of our \GLOBES\ examples:
@@
 \end{center}
-\mycaption{\label{tab:experiments} \index{Experiment data files (table)} Pre-defined experiment prototypes, their filenames (to be used in \GLB{glbInitExperiment}), their short descriptions, and the references in which they are originally used and discussed (except from minor modifications, such as a
+\mycaption{\label{tab:experiments} \index{Experiment files (table)}
+\index{{\tt glb}-files} 
+Pre-defined experiment prototypes, their filenames (to be used in \GLB{glbInitExperiment}), their short descriptions, and the references in which they are originally used and discussed (except from minor modifications, such as a
 different implementation of the energy threshold function). Note that
@@
 \section{Initialization of \GLOBES}
+\index{Initialization}
 
 Before one can use \GLOBES , one has to initialize the \GLOBES\
-library \GLB{libglobes}:
+library \GLB{libglobes}: \index{Initialization: \GLOBES\ library}
 \begin{function}
@@
 
+\index{Experiment list}
+\index{Initialization: Experiments}
 For storing the experiments, \GLOBES\ uses the initially empty list of experiments \GLB{glb\_experiment\_list}.
@@
  adds a single experiment with the filename {\tt inf} to the list of currently loaded experiments. The {\tt counter} is a pointer to the 
- variable containg the number of experiments, and the experiment {\tt in}
+ variable containing the number of experiments, and the experiment {\tt in}
  points to the beginning of the experiment list. The function returns
@@
 A list of pre-defined experiment prototypes, their filenames, their short descriptions, and the references of their definitions can be found in \tabl{experiments}.
-If the program cannot find these files, or some of them are syntactially not correct, it will break at this place.
-
+If the program cannot find these files, or some of them are syntactically not correct, it will break at this place.
+
+\index{Clear experiment list}
+\index{Delete experiments}
 One can also remove all experiments from the evaluation list at running
@@
 Mass squared differences & $\sdm$, $\ldm$ & $\mathrm{eV}^2$ \\
-Matter densitities & $\rho_i$ & $\mathrm{g}/\mathrm{cm}^3$ \\
+Matter densities & $\rho_i$ & $\mathrm{g}/\mathrm{cm}^3$ \\
 Baseline lengths & $L_i$ & $\mathrm{km}$ \\
@@
 \end{tabular}
-\mycaption{\label{tab:units} \index{Units in \GLOBES } Quantities used in \GLOBES , examples of these quantities, and their standard units in the application software.}
+\mycaption{\label{tab:units} \index{Units in \GLOBES\ (table)} Quantities used in \GLOBES , examples of these quantities, and their standard units in the application software.}
 \end{center}
@@
 \label{sec:luminosity}
+\index{Integrated luminosity}
+\index{Detector mass} \index{Source power} \index{Running time}
 
@@
  usually useful parent particle decays per year (neutrino factories,
- $\beta$-beams), target power in megawatts (superbeams), or thermal
- reactor power in gigawatts (reactor experiments).
+ $\beta$-beams), target power in mega watts (superbeams), or thermal
+ reactor power in giga watts (reactor experiments).
 %
-Since the pre-defined experiments in \tabl{experiments} are given for specfific luminosities, it is useful to read out and change these
+Since the pre-defined experiments in \tabl{experiments} are given for specific luminosities, it is useful to read out and change these
  parameters of the individual experiments:
@@
 \section{Handling oscillation parameter vectors}
+\index{Oscillation parameter vectors}
 
@@
 $\hat{\rho_i}$, \index{Matter density scaling factor}
+\index{Matter density profile} 
 which is stored in the density information of {\tt glb\_params}. 
@@
 \begin{function}
-\GLBNS{glbSetIteration}
+\GLBNS{glbSetIteration} \index{Iterations}
 {\tt glb\_params glbSetIteration(glb\_params in, int iter)} sets the number of iterations in the structure {\tt in}
@@
 In total, the parameter vector handling in a program normally has the
-following order:
+following order:\index{Parameter vector handling}
 \begin{quote}
@@
 \section{Computing the simulated data}
+\index{True values} \index{Simulated data} \index{Reference rate vector}
+\index{Set oscillation parameters}
 
@@
 versions. It is up to the user to implement mechanisms into the
-program and \AEDL\ files to check wether
+program and \AEDL\ files to check whether
 \begin{itemize}
@@
 \chapter[Calculating $\chi^2$ with systematics only]{Calculating $\boldsymbol{\chi^2}$ with systematics only}
-
-\index{Systematics}
+\index{Systematics $\chi^2$}
+
 Calculating a $\chi^2$-value with or without systematics, but no correlations and degeneracies, is the simplest and fastest possibility to obtain high-level information on an experiment.
@@
 given by observations is imposed (typically up to $5\%$). 
+\index{Matter density uncertainty}
  For this section, it is important to keep in mind
@@
 last section, the mechanism for the matter density scaling factors
-is hidden in the defintion of \GLB{glb\_params}: Each of the scaling
+is hidden in the definition of \GLB{glb\_params}: Each of the scaling
 factors is initially set to $1.0$. Therefore, for the calculation of
@@
 corresponds to the sum of all of the individual {\tt glbChiSys} calls. 
-This equality will not hold for the minimizors in the next sections anymore. 
+This equality will not hold for the minimizers in the next sections anymore. 
  An example how to use  {\tt glbChiSys} can be found on page~\pageref{ex:corrth13dcp}.
 
-\index{Systematics}
+\index{Systematics} \index{Pull method}
 The treatment of systematics in \GLOBES\ is performed by the so-call
@@
 This minimization is done independently for all auxiliary parameters of the rule.
-The total $\chi^2$ for the considered experiment is finally obtained by repreating this procedure for all rules and adding their $\chi^2$-values.
+The total $\chi^2$ for the considered experiment is finally obtained by repeating this procedure for all rules and adding their $\chi^2$-values.
 In general, the situation is more complicated because of the usage of many systematical errors.
@@
 \label{chapt:correlations}
+\index{Correlations $\chi^2$}
 
@@
 \index{Minimizer}
-The solution to this problem is using a local $n$-dimensional (local) minimizer for the procection instead of a grid-based method, where we will
+The solution to this problem is using a local $n$-dimensional (local) minimizer for the projection instead of a grid-based method, where we will
 illustrate this minimization process later. It turns out
@@
 Thus, one has to use some (analytical or numerical) knowledge on the topology of the fit manifold and start the local minimizer close enough to the investigated solution.
-Fortunately, this can be done quite straightfoward in most cases, since the structure of the neutrino oscillation formulas does not cause very complicated topologies of the fit manifolds.
+Fortunately, this can be done quite straightforward in most cases, since the structure of the neutrino oscillation formulas does not cause very complicated topologies of the fit manifolds.
 Especially the simulation of reactor experiments and conventional beams or superbeams is rather simple with purely numerical
@@
 of the many analytical discussions of this issue can be used. This means
- that one can implicitely use the analytical knowledge to obtain better predictions for the measurement performances, since the minimizer has
+ that one can implicitly use the analytical knowledge to obtain better predictions for the measurement performances, since the minimizer has
  to be started only close enough to the exact solution by an ``educated
@@
 It is one of the strengths of the \GLOBES\ software to use external input in order to reduce the fit manifold with the knowledge from external
-(eariler) measurements.
+(earlier) measurements.
 The treatment of external input is done by the addition of Gaussian so-called {\em priors} to the systematics-minimized $\chi^2$-function.
-For example, for the matter density, one obtains as the final projected $\chi^2_F$ after minimzation over the matter density
+For example, for the matter density, one obtains as the final projected $\chi^2_F$ after minimization over the matter density
 scaling factor \index{Matter density scaling factor} $\hat{\rho}$
@@
 This example is a very simple one, since in fact the
-minimzation is simultaneously performed over all priors and free oscillation parameters.
+minimization is simultaneously performed over all priors and free oscillation parameters.
 In \equ{priors}, $\hat{\rho}^0$ is the {\em starting value} of the prior, and $\sigma_{\hat{\rho}}$ the $1 \sigma$ absolute (half width) {\em input error}.
@@
 precisions, one can omit the respective input errors. In \GLOBES , a
-value of $0$ correponds to neglecting the prior.
+value of $0$ corresponds to neglecting the prior.
 If, however, earlier external measurements provide better information, one can set their absolute precisions with the input errors.
@@
 for the precisions of the solar experiments. In some cases, it may be
-necessary to adjust them, such as for artificial contraints to the
+necessary to adjust them, such as for artificial constraints to the
 oscillation parameters. In other cases, minor modifications of the starting values can cause a faster convergence of the algorithm.
@@
 
+\index{Input errors} \index{Starting values}
 In order to set the starting values and input errors, two function have to
@@
  sets the input errors for all of the following minimizer calls to {\tt in}.
- An input error of $0$ correponds to not taking into account the
+ An input error of $0$ corresponds to not taking into account the
  respective prior.
@@
 
+\index{External precision}
 Eventually, a typical initialization of the external input with
@@
 
+\index{Priors}
+\index{Advanced tricks}
 Though the priors are an elegant way to treat external input, there are also some complications with priors. The following hints are for the
@@
  (besides the minimization) why the sum of all projected $\chi^2$'s of 
- the indivual experiments 
+ the individual experiments 
  cannot correspond to the $\chi^2$ of the combination of all experiments.
@@
 
+\index{Minimizer}
 The following functions are some of the simplest minimizers 
@@
 \GLB{glbPrintParams} in order to check that the minimum is the
-appropriate one. For example, if the mininizer ends up in the wong-sign
+appropriate one. For example, if the minimizer ends up in the wrong-sign
 solution in $\ldm$, priors can be used to force it into the tested
@@
 all experiments) are different from $1.0$ if matter density uncertainties
-are present, since each experiment may face other matter density conditions. The minimizer for indivdual experiments ``know'' which
+are present, since each experiment may face other matter density conditions. The minimizer for individual experiments ``know'' which
 experiment they are currently treating, which means that only the returned
@@
 to use the best-fit values. A strong indication for a wrong guessing 
-are discontiuous jumps in the projected $\chi^2$-function, where the minimizer jumps from one minimum to another.
+are discontinuous jumps in the projected $\chi^2$-function, where the minimizer jumps from one minimum to another.
 In such cases, the starting point of the minimizer has to be adjusted to help it find the true minimum.
@@
 minimized over. 
-The corresponding function is
+The corresponding function is \index{Projection onto $\theta_{13}$-$\deltacp$-plane}
 \begin{function}
@@
 \end{tabular}
-\caption{\label{tab:defprojection} Different functions handling the
+\caption{\label{tab:defprojection}\index{Projection type}
+Different functions handling the
 \GLB{glb\_projection} type.
 Flags are either \GLB{GLB\_FIXED} or \GLB{GLB\_FREE}.
-The (unshown) return values of the {\tt Set-} and {\tt Define-} functions point either to the assigned vector if successful, or they are {\tt NULL} if unsuccessful.}
+The (un-shown) return values of the {\tt Set-} and {\tt Define-} functions point either to the assigned vector if successful, or they are {\tt NULL} if unsuccessful.}
 \end{table}
@@
 
+\index{Projection: Define}
 As soon as we have defined a projection, we can assign it:
@@
 \end{function}
-Similarly, the currenty assigned projection can be returned with:
+Similarly, the currently assigned projection can be returned with:
 \begin{function}
@@
 \index{Degenerate solutions}
+\index{Degeneracies $\chi^2$}
 
@@
 This approach is very useful for the exact numerical location of a degeneracy if its approximate position is known.
-For the determination of the approximate positon, one can use analytical approaches or an educated guess.
+For the determination of the approximate position, one can use analytical approaches or an educated guess.
 Though the usage of the all-parameter minimizers is quite simple, one should keep in mind that they are local minimizers.
@@
 \label{ex:sgndeg}
+\index{Degeneracies: sgn($\ldm$)-degeneracy}
 
@@
   \\
-  /* Localize degenerate solution by minimzation over all parameters */ \\
+  /* Localize degenerate solution by minimization over all parameters */ \\
   double CL=glbChiAll(starting\_values,deg\_pos,GLB\_ALL); \\
@@
 at the minimum as return value. Thus, the return value can be immediately
-used to judge wether the located degeneracy appears at the chosen
+used to judge whether the located degeneracy appears at the chosen
 confidence level.
@@
 
+\index{Advanced tricks}
 For the more advanced reader, a number of tricks can be useful for the numerical localization of degenerate solutions:
@@
 \chapter{Obtaining low-level information}
+\index{Low-level information}
 
@@
 
+\index{Event rates}
 \GLOBES\ currently supports rule-based and channel-based event rate functions, where the information is in written into a file
@@
 the rates with (\GLB{GLB\_W\_COEFF}) or without (\GLB{GLB\_WO\_COEFF})
-overall efficieny coefficient, as it is specified by {\tt coeffi}. 
+overall efficiency coefficient, as it is specified by {\tt coeffi}. 
 In addition, it may contain the post-smearing efficiencies (set
@@
 for the flavor $\nu_l$ 
-and polarity {\tt anti} ($+1$: neutrinos, $-1$: antineutinos) at the energy {\tt E} and distance {\tt distance}.
+and polarity {\tt anti} ($+1$: neutrinos, $-1$: antineutrinos) at the energy {\tt E} and distance {\tt distance}.
 \end{function}
@@
 \label{chapt:running}
+\index{Experiment parameters}
 
@@
 \section{Baseline and matter density profile}
+\index{Baseline: Change}
+\index{Matter density profile: Change}
 
@@
 {\tt int glbSetBaselineInExperiment(int exp, double baseline)}
-sets the baseline length in experimente {\tt exp} to {\tt baseline}.
+sets the baseline length in experiment {\tt exp} to {\tt baseline}.
 The function returns $-1$ if it was not successful.
@@
 has to be provided by the user (normally, it is enough to declare
-a variable of the type {\tt size\_t} corrsponding to positive
+a variable of the type {\tt size\_t} corresponding to positive
 integer numbers).
@@
 
-The returned lists of data from thie example represent $\chi^2$ 
+The returned lists of data from the example represent $\chi^2$ 
 as function of the fit value of $\stheta$. The intersections of these
@@
 \section{Systematics}
+\index{Systematics}
+\index{Systematics on/off}
 
@@
 the impact factors affecting the measurement.
-In \GLOBES , the sytematics is defined rule-based, \ie, each rule
+In \GLOBES , the systematics is defined rule-based, \ie, each rule
 has its own systematics. In addition, \AEDL\ requires that it has to be
@@
 
+\index{Error dimension}
 The error dimension (for the definition, see \Sec~\ref{sec:rules}) can also be accessed directly with 
@@
 
+\index{Signal errors}
+\index{Background errors}
+\index{Background centers}
 Except from the general treatment of systematics, one can read out
@@
 sets the background centers of experiment {\tt exp} and rule {\tt rule}
-to {\tt norm} (normalization center) and {\tt tilt} (tilt/calibrationcenter).
+to {\tt norm} (normalization center) and {\tt tilt} (tilt/calibration center).
 \end{function}
@@
 \label{sec:aedlparams}
+\index{External \AEDL\ parameters}
 
@@
 \section{Algorithm parameters: Filter functions}
+\index{Filter functions}
 

### 2004-07-12 Documentation/part2.tex
Some changes in part II: simbinXXX -> samplingXXX Type 2 smearing algorithm decsription replaced by section 'Low pass filter' Checked that the commands in the documentation are the same than in the code.

@@
 
-\subsection{Low-pass filter based automatic energy smearing}
-
-THIS SUBSECTION WILL BE RE-WRITTEN BY PATRICK (WW)
-
-The second available algorithm is more complicated, a little slower but
-much more accurate and much less susceptible to aliasing effects
-\index{Aliasing}. The starting
-point again is equation~\ref{eq:simple_int}. The idea now is to
- have a good approximation to the integral but still keeping the fixed sampling
-point in $E$. To this end we rewrite the integrand of 
-equation~\ref{eq:simple_int} using the sampling theorem~\cite{NRC,Rybicki}
-\begin{equation}
-\label{eq:sampling}
-f(E)=\sum_{k=-\infty}^{\infty}\,f(E_k) 
-\times \mathrm{sinc}\left(\frac{\pi}{h}(E-E_k)\right)+e(E)
-\end{equation}
-with $\mathrm{sinc}(x):=\sin(x)/x$ and $E_k=E^0+h\cdot k$. 
-This the so called sampling 
-representation of the integrand $f(E)$ and $e(E)$ is an error term.
-The error term vanishes if the Fourier transform $\hat f(\omega)$ of $f(E)$
-is zero outside the interval $[-\pi/h,\pi/h]$. Thus one suspects that
-the error term is small once the power beyond $\pi/h$ is low. Furthermore
-the sum in equation~\ref{eq:sampling} can be truncated to a few terms if
-$f(E)$ has a compact support, \ie\ it is different from zero only 
-on a finite interval. Thus the sampling representation is most useful
-for functions which have both a compact support in real \emph{and} Fourier
-space. This of course is impossible, however the function which comes
-most closely to this, is well known -- it is a Gau\ss ian. Our $f(E)$ is not
-exactly a Gau\ss ian  but the convolution of a top hat and a Gau\ss ian, thus
-its real and Fourier space components still behave rather Gau\ss ian 
-especially for large values of $E$ and $\omega$ respectively. The advantage of
-using the sampling representation is that the integration with respect to $E$
-is easily performed analytically
-  \begin{eqnarray}
-\label{eq:int_sampling}
-\int dE \quad f(E)&\simeq& \sum_{k=k_{l}}^{k_{u}}\,f(E_k) 
-\times \mathrm{Si}\left(\frac{\pi}{h}(E-E_k)\right)\,,\\
-\mathrm{Si}(x)&:=&\int_{0}^x dx'\quad \sin(x')/x'\,,\nonumber 
-\end{eqnarray}
-where $k_l$ and $k_u$ are suitable chosen bounds for the summation.
-
-In order to have a good compromise between speed an accuracy the choice
-of $h$, $k_l$ and $k_u$ is crucial. The choice of $h$ is governed by the
-need to sample $f(E)$ at a sufficient rate in order not to produce aliasing.
-For this purpose it useful to consider some properties of the power spectrum
-of $f(E)$. Provided the cross section and flux is smooth enough  the only
-high power components in $\hat f(\omega)$ are due to a fast oscillating
-$P(E)$. We can estimate the largest frequency $\omega_m$ quite easily, it is
-given by the largest $\Delta m^2_m$ and the lowest energy $E_l$
-\begin{equation}
-\omega_m \simeq \frac{\Delta m^2_m \cdot L}{E_l^2}
-\end{equation}
-On the other hand the finite energy resolution of any real detector anyhow
-will act as a low-pass filter eliminating any power beyond $\sim 1/\sigma(E)$.
-For now we assume that there are no frequencies which are not resolved by
-the detector thus the sampling frequency is determined by the energy 
-resolution of the detector. In that case the correct sampling 
-width $h$ is approximately
-\begin{equation}
-h\simeq 1/\sigma(E_l)\,.
-\end{equation}
-
-The range for the summation $k_l^i$ and $k_u^i$ is given by the range of 
-integration $a_i$ and $b_i$ in equation~\ref{eq:simple_int}, the integration 
-limits there are $0$ to $\infty$. For practical purposes however the range
-can be reduced to $a_i$ and $b_i$. They are  determined by requiring
-that the area under the bin kernel $K_i^c$ has reached a specific 
-value close to one,
-the so called confidence level $c$ (new name would be better). $1-c$ is a 
-small number and specifies the fraction of  events in bin which lie outside
-the summation range, usually $1-c=10^{-3}$ is sufficient. Furthermore the
-value of the bin kernel at the lower and upper boundary should
-be the same
-\begin{eqnarray}
-\int_{a_i}^{b_i} dE\quad K_i^c(E) &=& c\,,\\
-K_i^c(a_i)&=&K_i^c(b_i)\,.
-\end{eqnarray} 
-Knowing $a_i$ and $b_i$ it is easy to obtain $k_l$ and $k_u$ by requiring
-that the summation extends one bin beyond  each $a_i$ and $b_i$. Furthermore
-it is possible to specify an offset which extends the index range on each side
-by its value.
-
-Since the $a_i$'s and $b_i$'s are not known in advance it could be possible
-that $a_0$ and/or $b_{N_\mathrm{bins}}$ lie outside the computational domain,
-\eg\ no cross section or flux is defined at those energies. To remedy this 
-problem the computational domain can be specified. No calculation in \GLOBES\
-will leave this domain.
-
-In order to ensure that fast oscillating probabilities do not lead to aliasing
+\subsection{Low-pass filter}
+
+In order to ensure that fast oscillating probabilities do not lead to 
+aliasing\index{Aliasing}
 it is desirable to impose a low-pass filter already during the calculation
@@
 see~\cite{Kiers:1996zj,Giunti:2003ax}. This feature works \emph{only} 
-for vacuum and constant densities and controlled
-by the filer state variable and $\sigma_f^0$ is set by the filter value 
-variable. Assuming that $P_{\alpha\beta}$ still is the source for the highest
-frequencies one can split the energy resolution function $\sigma(E)^c$ in
+for vacuum and constant densities and is controlled
+by the filer state variable and $\sigma_e$ is set by the filter value 
+variable:
+\begin{quote}
+{\tt
+\$filter\_state = 1\\
+\$filter\_value = 2.0\\
+}
+\end{quote}
+would switch the filter feature on and set the width to $2.0\,\mathrm{GeV}$.
+The setting of {\tt \$filter\_state} is ignored whenever a density profile
+with more than one layer is used.
+
+With a type 1 ({\tt @type = 1}) energy resolution 
+function, $\sigma_e$ adds on to the energy resolution function 
+of the detector $\sigma_c(E)$ like
+\begin{equation}
+\sigma_{\mathrm{eff}}(E)^2\simeq \sigma_e^2 + \sigma_c(E)^2\,.
+\end{equation} 
+Sometimes this behaviour is undesirable and therefore one can try to 
+'subtract' the filtering from the energy resolution function by splitting 
+the energy resolution function $\sigma(E)_{\mathrm{eff}}$ in
 two parts by
 \begin{equation}
-\sigma_c^2(E)=\underbrace{\sigma_c^2(E)-\sigma_c^2(E_\mathrm{min})}_
-{\tilde\sigma^2_c(E)}+\underbrace{\sigma_c^2(E_\mathrm{min})}_{\sigma_e^2}\,,
-\end{equation}
-where $\tilde\sigma_c(E)$ is used instead of $\sigma_c(E)$ in computing the
-smearing data and $\sigma_e$ is used to determine $\sigma_f^0$.
+\sigma_{\mathrm{eff}}(E)^2=\underbrace{\sigma_c(E)^2-\sigma_e^2}_
+{\tilde\sigma^2_c(E)}+\sigma_e^2\,,
+\end{equation}
+where the truncated energy resolution function $\tilde\sigma_c(E)$ 
+is used instead of $\sigma_c(E)$ in computing the
+smearing data. Thus one obtains that
+\begin{equation}
+\sigma_{\mathrm{eff}}(E)^2\simeq \sigma_c(E)^2\,.
+\end{equation} 
+This scheme is enabled by choosing as type for the energy resolution
+\begin{quote}
+{\tt
+@type = 2
+}
+\end{quote}
 

### 2004-07-12 Documentation/Paper/paper.tex
*** empty log message ***

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\NeedsTeXFormat{LaTeX2e}
\documentclass[12pt,a4paper]{article}

%-- used packages ------------------------------------------------------

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{cite}
%\usepackage{mcite}
%-- page parameters -------------------------------------------------

\jot = 1.5ex
\parskip 5pt plus 1pt
\parindent 0pt
\evensidemargin -0.1in   \oddsidemargin  -0.1in
\textwidth  6.45in       \textheight 9.1in
\topmargin -1.0cm        \headsep    1.0cm

%-- command (re)definitions -----------------------------------------

\newcommand{\capdef}{}
%\newcommand{\mycaption}[2][\capdef]{\renewcommand{\capdef}{#2}%
%       \caption[#1]{{\itshape #2}}}
\newcommand{\mycaption}[2][\capdef]{\renewcommand{\capdef}{#2}%
       \caption[#1]{{\footnotesize #2}}}
\makeatletter
\renewcommand{\fnum@table}{\textbf{\tablename~\thetable}}
\renewcommand{\fnum@figure}{\textbf{\figurename~\thefigure}}
\makeatother
\def\ltap{\ \raisebox{-.4ex}{\rlap{$\sim$}} \raisebox{.4ex}{$<$}\ }
\def\gtap{\ \raisebox{-.4ex}{\rlap{$\sim$}} \raisebox{.4ex}{$>$}\ }

\newcounter{myenumi}
\newcommand{\myitem}{\refstepcounter{myenumi}\item}
\renewcommand{\themyenumi}{\roman{myenumi}}
\newenvironment{mylist}{%
        \setcounter{myenumi}{0}
        \begin{list}{\textit{\themyenumi)}}{%
        \setlength{\topsep}{0.2\baselineskip}%
        \setlength{\partopsep}{-\topsep}%
        \setlength{\itemsep}{\topsep}%
        \setlength{\parsep}{0\baselineskip}%
        \setlength{\leftmargin}{0em}%
        \setlength{\listparindent}{\parindent}%
        \setlength{\itemindent}{2.5em}%
        \setlength{\labelwidth}{1.5em}%
        \setlength{\labelsep}{0.75em}}}%
{\end{list}}

\newlength{\myem}
\settowidth{\myem}{m}
\newcommand{\sep}[1]{#1}
\newcounter{mysubequation}[equation]
\renewcommand{\themysubequation}{\alph{mysubequation}}
\newcommand{\mytag}{\stepcounter{mysubequation}%
\tag{\theequation\protect\sep{\themysubequation}}}
\newcommand{\globallabel}[1]{\refstepcounter{equation}\label{#1}}

\makeatletter
\renewcommand{\section}{\@startsection{section}{1}{0em}{-\baselineskip}%
{\baselineskip}{\normalfont\large\bfseries}}
\renewcommand{\subsection}%
{\@startsection{subsection}{2}{0em}{-0.7\baselineskip}%
{0.7\baselineskip}{\normalfont\bfseries}}
\makeatother

\newcommand{\ie}{{\it i.e.}}
\newcommand{\Ie}{{\it I.e.}}
\newcommand{\eg}{{\it e.g.}}
\newcommand{\Eg}{{\it E.g.}}
\newcommand{\cf}{{\it cf.}}
\newcommand{\etc}{{\it etc.}}
\newcommand{\eq}{Eq.}
\newcommand{\eqs}{Eqs.}
\newcommand{\Def}{Definition}
\newcommand{\fig}{Fig.}
\newcommand{\Fig}{Fig.}
\newcommand{\figs}{Figs.}
\newcommand{\Figs}{Figs.}
\newcommand{\Ref}{Ref.}
\newcommand{\Refs}{Refs.}
\newcommand{\Sec}{Sec.}
\newcommand{\Secs}{Secs.}
\newcommand{\Chapt}{Chapter}
\newcommand{\Chapts}{Chapters}
\newcommand{\Part}{Part}
\newcommand{\App}{Appendix}
\newcommand{\Apps}{Appendices}
\newcommand{\Tab}{Table}
\newcommand{\Tabs}{Tables}

\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\ra}{$\rightarrow$}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}
\newcommand{\nn}{\nonumber}
\newcommand{\ldm}{\Delta m_{31}^2}
\newcommand{\sdm}{\Delta m_{21}^2}
\newcommand{\deltacp}{\delta_{\mathrm{CP}}}
\newcommand{\stheta}{\sin^2 2 \theta_{13}}

\newcommand{\JHFSK}{{\sc JHF-SK}}
\newcommand{\NUMI }{{\sc NuMI}}
\newcommand{\ReactorI}{{\sc Reactor-I}}
\newcommand{\ReactorII}{{\sc Reactor-II}}
\newcommand{\JHFHK}{{\sc JHF-HK}}
\newcommand{\NuFactI}{{\sc NuFact-I}}
\newcommand{\NuFactII}{{\sc NuFact-II}}

\newcommand{\GLOBES}{{\sf GLoBES}}
\newcommand{\AEDL}{{\sf AEDL}}
\newcommand{\EDM}{{\sf EDM}}

\newcommand{\equ}[1]{\eq~(\ref{equ:#1})}
\newcommand{\figu}[1]{\fig~\ref{fig:#1}}
\newcommand{\tabl}[1]{\Tab~\ref{tab:#1}}
\newcommand{\tb}{\hspace*{3ex}}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%                     Title-page                              %%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{titlepage}

% the footnote symbols are only redefined for the title page !
\renewcommand{\thefootnote}{\alph{footnote}}

\vspace*{-3.cm}
\begin{flushright}
TUM-HEP-5XX/04\\
MPP-2004-YY\\
%hep-ph/
\end{flushright}

\vspace*{0.5cm}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\setcounter{footnote}{-1}

{\begin{center}
{\Large\bf GLoBES}
\end{center}}
\renewcommand{\thefootnote}{\alph{footnote}}

\vspace*{.8cm}
%\vspace*{.3cm}
{\begin{center} {\large{\sc
                P.~Huber\footnote[1]{\makebox[1.cm]{Email:}
                phuber@ph.tum.de},~
                M.~Lindner\footnote[2]{\makebox[1.cm]{Email:}
                lindner@ph.tum.de},~and~
                W.~Winter\footnote[5]{\makebox[1.cm]{Email:}
                wwinter@ph.tum.de}
                }}
\end{center}}
\vspace*{0cm}
{\it
\begin{center}

\footnotemark[1]${}^,$\footnotemark[2]${}^,$\footnotemark[3]%
       Institut f\"ur theoretische Physik, Physik--Department,\\
       Technische Universit\"at M\"unchen,
       James--Franck--Strasse, D--85748 Garching, Germany

\end{center}}

\vspace*{1cm}

\begin{abstract}
This is the abstract.
\end{abstract}

\vspace*{.5cm}

\end{titlepage}

\newpage

\renewcommand{\thefootnote}{\arabic{footnote}}
\setcounter{footnote}{0}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                     Introduction                                  %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%             References                         %%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\bibliographystyle{./apsrev}
\bibliography{./references}

\end{document}

### 2004-07-12 Documentation/Paper/paper.tex
*** empty log message ***

@@
 {\begin{center}
-{\Large\bf GLoBES}
+{\Large\bf Simulation of long-baseline neutrino oscillation
+experiments with GLoBES}
+\end{center}}
+{\begin{center}
+{\large\bf (General Long Baseline Experiment Simulator)}
 \end{center}}
@@
 
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=16cm]{GLOBES}
+\end{center}
+\caption{\label{fig:GLOBES} General concept of the \GLOBES\ package.}
+\end{figure}
+
+\bi
+ \item 
+  \figu{GLOBES} and main features
+ \item
+  Physics basics: For what purpose? What are difficulties
+ \item
+  Detector simulation: What happens there?
+ \item
+  References for work done with \GLOBES\
+ \item
+  Parts of \GLOBES\
+ \item 
+  \figu{GLOBES} and main features
+\ei
+
+\section{The computation of raw event rates}
+
+In this section, we sketch the computation of 
+the event rates, which is on of the core parts of the \GLOBES\
+software. However, because of the compecity of this issue,
+we refer to the \GLOBES\ manual~\cite{Manual} for
+details.
+
+The differential event rate for each channel is given by
+\begin{eqnarray}
+\label{equ:master_event}
+\frac{dn_{\beta}^{\text{IT}}}{dE'}=&&N\,\int\limits_0^\infty \int\limits_0^\infty dE\,d\hat{E}\quad
+\underbrace{\Phi_{\alpha} (E)}_{\mathrm{Production}} \times \nonumber\\
+&&\underbrace{\frac{1}{L^2} P_{(\alpha\rightarrow\beta)}(E,L,\rho;\theta_{23},
+\theta_{12},\theta_{13},
+\Delta m^2_{31},\Delta m^2_{21},\deltacp)}_{\mathrm{Propagation}}
+\times \nonumber \\ &&\underbrace{\sigma^{\text{IT}}_f(E)
+k_f^{\text{IT}}(E-\hat{E})}_{\mathrm{Interaction}} \times \nonumber \\
+&&\underbrace{ T_f(\hat{E}) V_f(\hat{E}-E')}_{\mathrm{Detection}}\,,
+\end{eqnarray}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+where $\alpha$ is the initial flavor of the neutrino, 
+$\beta$ is the final flavor, $\Phi_{\alpha} (E)$ is the flux of the 
+initial flavor at the
+source, $L$ is the baseline length, $N$ is a normalization factor, and 
+$\rho$ is the matter density. The energies in this formula are given as follows:
+\begin{itemize}
+\item
+ $E$ is the incident neutrino energy, \ie, the actual energy of the 
+incoming neutrino (which is not directly accessible to the experiment)
+\item
+ $\hat{E}$ is the energy of the secondary particle
+\item
+ $E'$ is the reconstructed neutrino energy, \ie, the measured
+neutrino energy as obtained from the experiment
+\end{itemize}
+The interaction term is composed of 
+two factors, which are the total cross section 
+$\sigma^{\text{IT}}_\beta(E)$ for the flavor $f$ and
+the interaction type IT, and the energy distribution of the 
+secondary particle $k_\beta^{\text{IT}}(E-\hat{E})$.
+The detector properties are 
+modeled by the threshold function $T_\beta(\hat{E})$, coming from the the 
+limited resolution or the cuts in the analysis, and the energy resolution 
+function $V_\beta(\hat{E}-E')$ of the secondary particle.
+
+Since it is a lot of effort to solve this double integral numerically,
+we split up the two integrations. First, we evaluate the integral over
+$\hat{E}$, where the only terms containing $\hat{E}$ are
+$k_\beta^{\text{IT}}(E-\hat{E})$,  $ T_\beta(\hat{E})$, and 
+$ V_\beta(\hat{E}-E')$. We define:
+\begin{eqnarray}
+\label{equ:e_res} 
+R_\beta^{\text{IT}}(E,E')\,\epsilon_\beta^{\text{IT}}(E')
+ \equiv
+\int\limits_0^\infty d\hat{E} \quad T_\beta(\hat{E})\,k_\beta^{\text{IT}}(E-\hat{E})
+\,V_\beta(\hat{E}-E')\,. 
+\end{eqnarray}
+Thus, $R_\beta^{\text{IT}}(E,E')$ describes the energy response of 
+the detector, \ie , a neutrino with a (true) energy $E$ is reconstructed
+with an energy between $E'$ and $E'+dE'$ with a probability
+$R_\beta^{\text{IT}}(E,E') dE'$. The function $R(E,E')$ is also often called ``energy resolution function''. Actually, its internal representation
+in the software is a smearing matrix.
+The function $\epsilon_\beta^{\text{IT}}(E')$ will be refered to as ``post-smearing efficiencies'', since it will allow us to define cuts and threshold functions {\em after} the smearing is performed, \ie, as function of $E'$.
+In addition, \GLOBES\ uses ``pre-smearing efficiencies'', which are 
+evaluated {\em before} the smearing is performed, \ie, as function of $E$.
+Similarly, constant event rates can be added to the event rates before or after the energy smearing, which are refered to as ``pre-smearing backgrounds'' and ``post-smearing beackground''.
+These types of
+efficiencies an backgrounds allow a very accurate modeling of
+many factors, such as geoneutrino backgrounds, energy cuts, or
+energy threshold functions.
+
+Eventually, we can write down the number of events per bin $i$ \index{Bin} and channel $c$ as
+\begin{equation}
+\label{equ:channel}
+n_i^c=\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' \quad
+\frac{dn_{\beta}^{\text{IT}}}{dE'} (E') \,
+\end{equation}
+where $\Delta E_i$ is the bin size of the $i$th energy bin.
+This means that one has to solve the integral
+\begin{eqnarray}
+\label{equ:events_bin}
+n_i^c=N/L^2\,\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
+\quad \int\limits_0^\infty dE \,\, \Phi^c(E)\,
+P^c(E)\,
+\sigma^c(E)\,
+R^c(E,E')\,
+\epsilon^c(E')\,,
+\end{eqnarray} 
+which gives the raw event rates of the channel $c$ in the $i$th bin.
+Note that the events are binned according to their \emph{reconstructed} energy.
+
+Core part of the event rate computation is the energy smearing algorithm
+to evaluate \equ{events_bin}, where either a particular energy resolution
+ function can be chosen for automated energy smearing, or the discretized
+smearing matrix $R^c(E,E') = R^c_{ij} $ can be given manually. In addition, it is possible
+to define a low-pass filter to avoid aliasing effects for neutrino
+oscillations faster that the sampling width.
+
+\section{Definition of experiments with \AEDL\ (Abstract Experiment 
+Defintion Language)}
+
+\begin{figure}[t]
+\begin{center}
+\begin{tabular}{p{5cm}p{5cm}p{5cm}}
+\includegraphics[width=5cm]{AEDL1} &
+\includegraphics[width=5cm]{SignalBackground} &
+\includegraphics[width=5cm]{Rules} \\
+\end{tabular}
+\end{center}
+\caption{\label{fig:aedl} The most important components of \AEDL :
+Channels, rules, and experiments.}
+\end{figure}
+
+In order to define experiments, \GLOBES\ uses \AEDL\ (``Abstract
+Experiment Definition Language''). An experiment normally corresponds
+to an \AEDL\ file, which is a text file written in \AEDL\ syntax.
+
+The main parts of any \AEDL\ experiment definition are channels,
+rules, and experiments (\cf, \figu{aedl}).
+A channel corresponds to a neutrino oscillation channel including flux, cross section (for one specific interaction type), energy resolution function, initial and final neutrino flavors, their polarity (neutrinos or antineutrinos), and efficiencies.
+The channel raw event rates are computated according to \equ{events_bin}.
+Each channel leads to the raw event rates for a
+specfifc interaction type. In \AEDL , many of the channel components
+have to be defined or loaded from files before, such as fluxes, 
+cross sections, or the energy resolution function. For the
+different options, see the \GLOBES\ manual~\cite{Manual}.
+
+For a ``rule'', the raw event rates of one or more signal channels and
+one or more background channels are added. The splitting in signal and
+background is arbitrary, but all of the signal and background components
+are defined to have a common systematics. The event rates of all 
+signal and background components add up to the total event rate of the
+rule, which leads to a $(\Delta \chi^2)_r$. The signal or background
+within each rule allows the specification of signal and background normalization errors and energy tilt or calibration errors. 
+These systematical errors will be evaluated with the ``pull method''
+with the help of nuisance parameters REFERENCE. In addition to these
+systematics errors, an overall evaluation strategy is assigned 
+to each rule, which specifies the type of systematics (tilt or calibration
+error), and the use of spectral information or total event rates.
+
+Finally, one ore more rules add up to an experiment, where the
+total $\Delta \chi^2$ is obtained as the sum of the $\Delta \chi^2$
+of all rules. This approach allows the definition of 
+appearance and disappearance channels, neutrino and antineutrino running, or interaction types with different systematics (spectral information
+versus counting rate) within one experiment. The \GLOBES\ user interface
+allows the simulation of one or more experiment simultaneously, which
+means that one could also use different experiments for different
+oscillation channels. However, there is still one component missing
+in the experiment definition, which is the matter density profile.
+For an experiment, an arbitrary matter density profiles can be specified, which is evaluated with the evolution operator method REFERENCES.
+In addition, the matter density profile is multiplied by the matter
+density scaling factor, which is treated as an independent parameter
+with a relative (matter density) uncertainty. With this
+approach, one can take into account that the matter density profile along a specific baseline is only known to about $3\%-5\%$. Therefore, it makes
+sense to combine the rules of one baseline in one experiment,
+since for all of these rules the matter density profile is equal.
+
+\section{Simulation of experiments: The C user's interface}
+
+Minimizers?
+
+\section{Complexity and computational issues}
+
+\bi
+\item
+ Complexity of projection: $\mathcal{O}(n^7)$ (incl. matter density) API
+ $\rightarrow$ Polynomical, but large effort;
+ Approach with Local minimizer \ra\ $\mathcal{O}(1000)$ steps
+\item
+ Running time of detector simulation: (each step above)
+ bins times sampling points + sampling points times matter profile steps.
+ Normally dominated by first summand; for many matter profile steps
+ second dominates
+\item
+ Computational effort: Estimates (time)
+\item
+ Pre-computation of smearing matrices, Spezielles
+ Format bei dem Nullen nicht mitgerechnet;
+  oscillation probabilities nur einmal gleichzeitig fuer alls
+  Oszillationskanaele und alle ENergien berechnet
+\ei
+
+\section{Use of program}
+
+\bi
+\item
+ Whole package
+\item
+ Shared library, globes command, Examples, Experiment protoypes
+\item
+ Manual \cite{Manual}
+\ei
+
+\section{Summary and conclusions}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2004-07-12 Documentation/part2.tex
extended the \GLB command to handle a leading '@' (this is meta character for makeindex and thus has to be escaped by '"'). Added a label for section 7.2

@@
 functions. All variables have their scope limited 
-by either {\tt \%!GLoBES, \#NEXT\#} or {\tt EOF}. The only exception are 
-the counters defined by {\tt cross} and {\tt flux}. They are valid for 
-the complete session, since one does not want too overload the flux and 
-cross section information when using more than one experiment. This allows 
+by either {\tt \%!GLoBES, \#NEXT\#} or {\tt EOF}.  This allows 
 a consistent treatment of various experiments in one file.
@@
 but the cross section data are different (for example, because of different target nuclei), the correct 
-cross section data will be applied to each of the experiments in the 
-application software. In this case, {\tt \#CC} in {\tt NuFact.gls} has 
-the value $0$ and in {\tt JHFHK.gls} it has the value $1$, 
-which means that no shadowing occurs. Note that, alternatively, one can 
+cross section data will be applied to each of the experiments. 
+Note that, alternatively, one can 
 also load both files successively.
@@
 
-In addition to the matter density profile, \GLOBES\ implements a simple treatment of matter density uncertainties by a scaling factor $\hat{\rho}$
-of the matter density profile:
-\begin{equation}
-\label{eq:density_error}
-\rho(x)=\hat{\rho} \cdot\rho_0(x)\,\left[\frac{\mathrm{g}}{\mathrm{cm}^3}\right] \, .
-\end{equation}
-Here $\rho_0(x)$ is the matter profile chosen by the user, and $\rho(x)$ is
-the one used in the calculation of the oscillation probabilities.
-As it is described in \Sec~\ref{sec:externalinput}, the matter density uncertainty is implemented by using $\hat{\rho}$ as an additional oscillation parameter for each experiment, and by assuming to know $\hat{\rho}$ with an external precision.
-Thus, the central value of $\hat{\rho}$ (default value $1.0$) and its error $\delta \hat{\rho}$ have to be defined.
-The following code assumes a matter density uncertainty of $5\%$:\index{Density error}\index{Density center}
-\begin{quote}
- {\tt \$densitycenter=1.0} \\
- {\tt \$densityerror=0.05}
-\end{quote}
-In most cases, this is in combination with a constant profile a very good approximation to take into account the realistic structure of the profile as well as its uncertainty~\cite{Ohlsson:2003ip}. 
- If one does not want to take into account  matter density uncertainties, 
-one can either a very small value of $\hat{\rho}$, or the minimizer {\tt ChiNP} with $\hat{\rho}$ fixed in the application software.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 In general, it is advisable to provide the cross sections in the range between {\tt \$sampling\_min} and {\tt \$sampling\_max} (\cf, \Sec~\ref{sec:energy}).
-Unused cross sections have to be filled with zeros, and can not be just omitted.
+Unused cross sections have to be filled with zeros, and can not
+ be just omitted.
 
@@
 for equidistant sampling points. Arbitrarily spaced sampling points can 
-be specified analogously to the energy bins below. 
+be specified with \GLB{\$sampling\_stepsize}
+\begin{quote}
+{\tt
+\$sampling\_stepsize=\{1.0,2.0,3.0,4.0,5.0,...\}
+}
+\end{quote}
+
 Note that the algorithm expects to 
@@
 x>1.8\cdot10^{-3}\\
-5\cdot10^{-5} \,,&\mathrm{for}\,\, x \leq 1.8\cdot10^{-3}
+\alpha\cdot10^{-3} \,,&\mathrm{for}\,\, x \leq 1.8\cdot10^{-3}
 \end{array} \right.
@@
 parameters which are minimized over. This choice is specified with the 
-error dimension variable\index{Error dimension}, and the different possibilities are shown in
-\Tab~\ref{tab:error_dim}. The error dimensions can be chosen for each rule 
-by {\tt @errordim}. For example, 
+error dimension variable\index{Error dimension}, and the different 
+possibilities are shown in
+\Tab~\ref{tab:error_dim}. Since the error dimension defines the treatement
+of systematics it is useful to have define a matched pair of error dimensions
+for each rule, where on value describes how the event rate is computed
+for \emph{no} systematics and the other one is with systematics 
+(see also~\Sec~\ref{sec:systematics}).
+The error dimensions for the case of no systematics is set for each rule 
+by the value of \GLB{@errordim\_sys\_off}, whereas the error dimension
+for systematics on is given by \GLB{@errordim\_sys\_on}. Thus, for example,
+the complete error dimension definition could look like  
 \begin{quote}
 {\tt 
-\tb @errordim = 1
-}
-\end{quote}
-In general, it is possible to extend the set of error dimensions or the 
-set of possible systematical errors, but this requires C programming, 
-compiling and linking with \GLOBES. 
-%The interface for user defined error 
-%dimensions is described in detail in section~\ref{sec:ud_error_dim}.
-
+\tb @errordim\_sys\_off = 2\\
+\tb @errordim\_sys\_on = 0
+}
+\end{quote}
+It is foreseen to add the possiblity to extend the set 
+of error dimensions or the set of possible systematical errors in one
+of the next versions of \GLOBES.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2004-07-12 Documentation/Paper/paper.tex
*** empty log message ***

@@
  \item 
-  \figu{GLOBES} and main features
+  \figu{GLOBES} and main features:
+  \bi
+   \item
+    3-flavor analysis
+   \item
+    Matter density profile+uncertainty
+   \item
+    Full detector simulation
+   \item
+    Up to 32 experiments simultaneously
+   \item
+    Full n-parameter correlation
+   \item
+    Localization of degeneracies
+  \ei
 \ei
@@
 
-Minimizers?
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=16cm]{projex}
+\end{center}
+\mycaption{\label{fig:projex} Left plot: Example for a  correlation between $\stheta$ and $\deltacp$ (for 1 d.o.f.
+only).
+Right plot: The $\chi^2$-value of the projection onto the $\stheta$-axis as function of $\stheta$.
+The projection onto the  $\stheta$-axis is obtained by finding the minimum $\chi^2$-value for each fixed value of $\stheta$ in the left-hand plot, \ie, along the gray vertical lines.
+The thick gray curve marks the position of these minima in the left-hand plot.
+The arrows mark the obtained fit ranges for $\stheta$ at the $3 \sigma$ confidence level (1 d.o.f.), \ie , the precision of $\stheta$.}
+\end{figure}
+
+In order to use \GLOBES , a C~library is linked with the application
+software. This library provides the user's interface functions.
+It allows to load \AEDL\ files, initialize the experiments, and
+have access to various $\Delta \chi^2$-functions including any combination
+of systematics and correlations. In addition, it provides low-level
+access to oscillation probabilities and event rates, and allows
+the readout and modification of many experiment parameters at running time.
+
+The most sophisticated feature of the user's interface is the
+possibility to include the full multi-parameter correlation among
+all oscillation parameters. Together with the matter densities
+of the $N$ loaded experiments ($N \ge 1$), the neutrino oscillation parameter space in \GLOBES\ has $6+N$ dimensions. In the $\chi^2$-approach,
+the allowed region of $k$ ($1\le k<6+N$) fit parameters (at the chosen confidence level) is obtained as the projection of the $6+N$-dimensional fit manifold onto the $k$-dimensional hyperplane.
+For example,
+for the precision of the parameter $\stheta$ only, the fit manifold
+needs to be projected onto the $\stheta$-axis. One can easily imagine
+that the computation of the full multi-parameter correlation is
+very expensive with a grid-based method. However, the topology
+of the neutrino oscillation parameter space is well investigated and
+rather fortunate. Therefore, we use local minimization for each
+solution in the parameter space at the chosen confidence level,
+where tools for the location of (disconnected) degenerate solutions 
+are provided. For this process, the $6+N$-dimensional local minimizer
+is started at the ``guessed'' solution (such as from analytical 
+knowledge) and runs into the local minimum. In \figu{projex},
+this minimization is illustrated at the simple example of a 
+two-parameter correlation at a neutrino factory.
+
+As we have discussed in the last section, the matter density of
+each experiment is treated as an addition parameter, which is known
+from external measurement with a certain precision. In principle, 
+the \GLOBES\ user's interface allows such externally imposed
+precisions for all oscillation parameters. This feature turns out
+to be especially useful for input from the solar neutrino experiments,
+since they provide better precisions for the solar neutrino
+osillation parameters $\sdm$ and $\theta_{13}$ than the long-baseline
+ experiments themselves.
 

### 2004-07-12 Documentation/part2.tex
Minor changes

@@
 {\tt cross(\#CC)<}\\
-{\tt \tb @cross\_file = XCC.dat}\\
+{\tt \tb @cross\_file = "XCC.dat"}\\
 {\tt >}
@@
 is also checked. For example,  {\tt 
-\$bins} can be only between $0$ and $150$.
+\$bins} can be only between $0$ and $500$~\footnote{The upper limit is 
+there only for safety reasons, the memory is allocated dynamically.}.
 If one uses a {\tt float} quantity where  an {\tt int} is expected, the {\tt float} will be converted to an {\tt int} in the same way as in C.
@@
 \begin{quote}
-{\tt include ./file\_1}
+{\tt include "./file\_1"}
 \end{quote}
@@
 \begin{quote}
-{\tt    include NuFact.gls\\
+{\tt    include  "NuFact.gls"\\
         \#NEXT\#\\
-        include JHFHK.gls
-}
-\end{quote}
-in order to initialize a combined analysis of the experiments defined in the files {\tt NuFact.gls} and {\tt JHFHK.gls}. 
+        include   "JHFHK.gls"
+}
+\end{quote}
+in order to initialize a combined analysis of the experiments defined in the files {\tt NuFact.gls} and {\tt JHFHK.gls}. As one sees from the example above
+file names are always quoted in \AEDL.\index{File names}
 Even if one uses the
@@
 \end{quote}
-and use some simple algebraic manipulations such as
+These constants can not only be defined within one \AEDL\ file but also
+by the calling C program, which allows to use a simple but powerful variable
+substitution mechanism as described in~\Sec~\ref{sec:aedlparams}.
+\index{External \AEDL\ parameters}
+
+Alos some simple algebraic manipulations are possible, such as
 \begin{quote}
@@
 The following mathematical functions from {\tt <math.h>} are available: 
-{\tt sin}, {\tt cos}, {\tt atan}, {\tt ln}, {\tt exp}, {\tt sqrt}. 
+{\tt sin}, {\tt cos}, {\tt tan}, {\tt asin}, {\tt acos}, {\tt atan}, 
+{\tt log}, {\tt log10}, {\tt exp}, {\tt sqrt}. 
 These functions can be used everywhere, where
@@
 {\tt flux(\#user)<}\\
-{\tt \tb @flux\_file = user\_file\_1.dat\\
+{\tt \tb @flux\_file = "user\_file\_1.dat"\\
 \tb @time = 2.0\\
@@
 {\tt cross(\#name)<}\\
-{\tt \tb @cross\_file =user\_file\_1.dat}\\
+{\tt \tb @cross\_file ="user\_file\_1.dat"}\\
 {\tt >}
@@
 \end{quote}
+
+\index{Oscillation!switching off}
+It is also possible to define a channel as no-oscillation by using the
+prefix \GLB{NOSC\_} in either the initial  flavour or 
+the final flavour, like this
+\begin{quote}
+{\tt channel(\#channel\_1)<\\
+\tb @channel = \#flux : $+$: NOSC\_muon: muon: \#cross: \#energy\\
+>}
+\end{quote}
+
+ In that case
+all diagonal probabilities $P_{\alpha\alpha}$ are unity and all off-diagonal
+probablities  $P_{\alpha\beta}$ are zero. This is \eg\ especially useful for
+neutral current events since they do not depend on any oscillation 
+parameters\footnote{at least in the absence of sterile neutrinos}. The channels
+marked as \GLB{NOSC\_} are only computed by a call to \GLB{glbSetRates}
+and not recomputed in the subsequent fit (which calls \GLB{glbSetNewRates}).
+Therefore this feature can be used to speed up the rate 
+computation considerably, especially in cases where a large set
+of channels exist which are only used for the computation of backgrounds.
+Usually it is an excellent approximation to treat backgrounds as if they
+would not be affected by oscillations.
+
 Note that the energy environment will be described in the next section. 
@@
 \tb @backgrounderror =   0.001 :       0.01\\
-\tb @errordim = 1\\
+\tb @errordim\_sys\_off = 2\\
+\tb @errordim\_sys\_on = 0\\
 \tb @energy\_window = 4.0 : 50.0\\ 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{User-defined systematics}
-%\label{sec:ud_error_dim}.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Versioning of \AEDL\ files}
+\label{sec:aedl_versioning}
+\index{Version control}
+
+In order to provide a handle on problems which may arise due to different
+versions of \GLOBES\ and \AEDL\ files, it is possible to stamp each 
+\AEDL\ file with the corresponding version number of the \GLOBES\ package with
+\GLB{\$version} by
+\begin{quote}
+{\tt \$version="1.8.1"}
+\end{quote}
+This information can be accesed by the versioning functions as described in~\Sec~\ref{sec:versioning}.
 

### 2004-07-13 Documentation/Paper/paper.tex
*** empty log message ***

@@
 
-\newcommand{\JHFSK}{{\sc JHF-SK}}
-\newcommand{\NUMI }{{\sc NuMI}}
-\newcommand{\ReactorI}{{\sc Reactor-I}}
-\newcommand{\ReactorII}{{\sc Reactor-II}}
-\newcommand{\JHFHK}{{\sc JHF-HK}}
-\newcommand{\NuFactI}{{\sc NuFact-I}}
-\newcommand{\NuFactII}{{\sc NuFact-II}}
+\newcommand{\MINOS}{{\sf MINOS}}
+\newcommand{\ICARUS}{{\sf ICARUS}}
+\newcommand{\OPERA}{{\sf OPERA}}
+\newcommand{\JHFSK}{{\sf JHF-SK}}
+\newcommand{\NUMI }{{\sf NuMI}}
+\newcommand{\ReactorI}{{\sf Reactor-I}}
+\newcommand{\ReactorII}{{\sf Reactor-II}}
+\newcommand{\JHFHK}{{\sf JHF-HK}}
+\newcommand{\NuFactI}{{\sf NuFact-I}}
+\newcommand{\NuFactII}{{\sf NuFact-II}}
 
@@
 \begin{abstract}
-This is the abstract.
+We present the \GLOBES\ (``General Long Baseline Experiment Simulator'')
+software package, which allows the simulation of future long-baseline
+neutrino oscillation experiments. One part of the software is the
+abstract experiment definition language to define experiments with
+beam and full detector simulations as accurate as possible. For the
+definition of experiments, a new transparent building block concept is introduced. In addition, a binary provides the possibility to test
+new experiment definitions quickly.
+%
+Another
+part of \GLOBES\ is the user's interface, which provides probability, rate, and $\Delta \chi^2$ information for a given experiment or any
+combination of up to 32 experiments. Especially, the
+$\Delta \chi^2$ functions allow a simulation with statistics only,
+systematics, correlations, and degeneracies. In particular, \GLOBES\
+can handle the full multi-parameter correlation among the oscillation
+parameters, external input, and matter density uncertainties. Eventually,
+many systematics options are provided, such as normalization and energy
+calibration errors, or the choice between spectral or total rate 
+information. 
 \end{abstract}
@@
 
+Neutrino oscillations are now established as the leading flavor
+transition mechanism for neutrinos. By the existing experiments, 
+two oscillation
+frequencies have been found: The fast ``atmospheric'' and the
+slow ``solar'' oscillations, which are driven by the respective
+mass squared differences.  In addition, there could be interference
+effects between these two oscillations, provided that the coupling given by the small mixing angle $\theta_{13}$ is large enough. Such interference
+effects include, for example, leptonic CP violation.  In order to
+test the unknown oscillation parameters, \ie, the mixing angle $\theta_{13}$, the leptonic CP phase, and the
+neutrino mass hierarchy, future long-baseline experiments are proposed.
+These experiments send an artificial neutrino beam to a detector, or
+detect the neutrinos produced by a nuclear fission reactor. However,
+instrinsic degenerate solutions in the neutrino oscillation
+probabilities affect these measurements and require optimization
+strategies for the experiments to maximally exploit complementary
+information. Therefore, a modern complete experiment simulation
+does not only need to have a beam and detector simulations, which are
+as accurate as possible, but also powerful tools to analyze correlations
+and degeneracies. In this study, we present the \GLOBES\ software package
+as such a tool.
+
+\GLOBES\ (``General Long Baseline Experiment Simulator'') is a flexible
+software tool to simulate neutrino oscillation long-baseline experiments using a complete three-flavor analysis. On the
+one hand, it contains a comprehensive abstract experiment definition
+language, which allows to describe most classes of long baseline experiments
+at an abstract level. On the other hand, it provides a C-library to 
+process the experiment information in order to obtain oscillation
+probabilities, rate vectors, and $\Delta \chi^2$-values (\cf, \figu{GLOBES}). In addition, it provides a binary to test experiment
+definitions very quickly, before they are used by the application software.
+Currently, \GLOBES\ is available for Linux. Since the source code is included, the modifications to other operating systems should be doable.
+
+\GLOBES\ allows to simulate experiments with stationary neutrino point sources, where each experiment is assumed to have only one neutrino source.
+Such experiments are neutrino beam experiments and reactor experiments. 
+Geometrical effects of a source distribution, such as in the sun or the 
+atmosphere, can not be described. In addition, sources with a physically 
+significant time dependencies  can not be studied, such as  supernov\ae. It 
+is, however, possible to simulate beams with bunch structure, since the 
+time dependence of the neutrino source is physically only important to suppress backgrounds.
+
 \begin{figure}[t]
@@
 
-\bi
- \item 
-  \figu{GLOBES} and main features
- \item
-  Physics basics: For what purpose? What are difficulties
- \item
-  Detector simulation: What happens there?
- \item
-  References for work done with \GLOBES\
- \item
-  Parts of \GLOBES\
- \item 
-  \figu{GLOBES} and main features:
-  \bi
-   \item
-    3-flavor analysis
-   \item
-    Matter density profile+uncertainty
-   \item
-    Full detector simulation
-   \item
-    Up to 32 experiments simultaneously
-   \item
-    Full n-parameter correlation
-   \item
-    Localization of degeneracies
-  \ei
-\ei
+On the experiment definition side, either built-in neutrino fluxes
+(\eg, neutrino factory) or arbitrary fluxes can be used. Similarly,
+arbitrary cross sections, energy dependent efficiencies, the
+energy resolution function, the considered oscillation channels, 
+backgrounds, and many other features can be specified. 
+For the systematics, energy
+normalization and calibration errors can be simulated. Note that
+the energy ranges and windows, as well as the bin widths can be
+(almost) arbitrarily chosen, which means that variable bin
+widths are allowed. Together with \GLOBES\ comes a number of
+pre-defined experiments in order to demonstrate the capabilities
+of \GLOBES\ and to provide prototypes for new experiments.
+Examples for these prototypes are the \MINOS , \ICARUS , and
+\OPERA\ simulations from \Ref~\cite{Huber:2004ug}, the
+\JHFSK\ and \NUMI\ superbeam simulations from \Refs~\cite{Huber:2002mx,
+Huber:2002rs}, the \JHFHK\ superbeam upgrade simulation from
+\Ref~\cite{Huber:2002mx}, the neutrino factory simulations from
+\Ref~\cite{Huber:2002mx}, and the reactor experiment simulations from
+\Ref~\cite{Huber:2003pm}. Other projects using the \GLOBES\ 
+software include \Refs~\cite{Huber:2003ak,Ohlsson:2003ip,Winter:2003ye,Antusch:2004yx}.
+
+With the C-library, one can extract the $\Delta \chi^2$ for all defined 
+oscillation channels for an experiment or any combination of experiments.
+Of course, also low-level information, such as oscillation
+probabilities or event rates, can be obtained. \GLOBES\ includes the
+simulation of neutrino oscillations in matter with arbitrary matter 
+density profiles, as well as it allows to simulate the matter density
+uncertainty. As one of the most
+advanced features of \GLOBES , it provides the technology to 
+project the $\Delta \chi^2$, which is a function of all oscillation
+parameters, onto any subspace of parameters by local minimization. 
+This approach allows the inclusion of multi-parameter-correlations,
+where external input (\eg, from solar parameters) can be imposed, too.
+Applications of the projection mechanism include the projections onto the $\stheta$-axis and the $\stheta$-$\deltacp$-plane.
+In addition, all oscillation parameters can be kept free to precisely localize
+degenerate solutions.
 
@@
 \begin{center}
-\begin{tabular}{p{5cm}p{5cm}p{5cm}}
-\includegraphics[width=5cm]{AEDL1} &
-\includegraphics[width=5cm]{SignalBackground} &
-\includegraphics[width=5cm]{Rules} \\
+\begin{tabular}{p{6.3cm}p{4.3cm}p{4.3cm}}
+\includegraphics[width=6.5cm]{AEDL1} &
+\includegraphics[width=4.5cm]{SignalBackground} &
+\includegraphics[width=4.5cm]{Rules} \\
 \end{tabular}
@@
 
-\bi
-\item
- Complexity of projection: $\mathcal{O}(n^7)$ (incl. matter density) API
- $\rightarrow$ Polynomical, but large effort;
- Approach with Local minimizer \ra\ $\mathcal{O}(1000)$ steps
-\item
- Running time of detector simulation: (each step above)
- bins times sampling points + sampling points times matter profile steps.
- Normally dominated by first summand; for many matter profile steps
- second dominates
-\item
- Computational effort: Estimates (time)
-\item
- Pre-computation of smearing matrices, Spezielles
- Format bei dem Nullen nicht mitgerechnet;
-  oscillation probabilities nur einmal gleichzeitig fuer alls
-  Oszillationskanaele und alle ENergien berechnet
-\ei
+The \GLOBES\ software only uses polynomial algorithms, which are,
+however, suffering from the large number of dimensions.
+The $\Delta \chi^2$ computation basically consists of two steps:
+The systematics $(\Delta \chi^2)_{\mathrm{sys}}$ computation with {\tt glbChiSys},
+and the projection algorithm which uses {\tt glbChiSys} many times. 
+The overall computation time is obtained as the product of these two steps.
 
-\section{Use of program}
+The running time of the systematics $(\Delta \chi^2)_{\mathrm{sys}}$ 
+is 
+\begin{equation}
+\mathrm{t_{sys}} = \mathcal{O} \left( N_{\mathrm{Sampling}} \times N_{\mathrm{Bins}}  \, + \, N_{\mathrm{Sampling}} \times N_{\mathrm{Layers}} \right) ,
+\end{equation}
+since the oscillation probabilities for all channels and energies
+are only computed once and stored in a list. Here $N_{\mathrm{Sampling}}$ is the number of sampling points for the evaluation of \equ{events_bin}, $N_{\mathrm{Bins}}$ is the number of energy bins,
+and $N_{\mathrm{Layers}}$ is the number of matter density layers.\footnote{Note that we are here using the \AEDL\ format for these
+parameters.}. Thus, whenever the
+number of matter density steps is small, the product of sampling points
+and bins dominates. In practice, the $N_{\mathrm{Sampling}} \times N_{\mathrm{Bins}}$ energy smearing matrix is already computed at 
+the experiment initialization in order to save running time. In addition,
+since it contains a lot of zeros, \GLOBES\ uses a special format to
+avoid looping over the zeros.
 
-\bi
-\item
- Whole package
-\item
- Shared library, globes command, Examples, Experiment protoypes
-\item
- Manual \cite{Manual}
-\ei
+The running time for the projected $\Delta \chi^2$ is, for a grid based method,
+\begin{equation}
+ \mathrm{t_{proj}^{Grid}} = \mathcal{O}\left( n^{6+N_{\mathrm{exp}}}\right) \, \times \, \mathrm{t_{sys}},
+\end{equation}
+where $n$ is the number of steps to be evaluated for each oscillation
+parameter, and $N_{\mathrm{exp}}$ is the number of experiments
+(since the matter densities of all experiments appear as parameters). 
+Though it is, in principle, possible to use such a method with \GLOBES ,
+it is impractiable in most cases, since it would involve at least 
+billions of
+steps for the complete multi-parameter correlation. The local $6+N_{\mathrm{exp}}$-dimensional minimization reduces this effort
+to 
+\begin{equation}
+\mathrm{t_{proj}^{Minimizer}} = \mathcal{O}(1000) \, \times \,
+N_{\mathrm{Deg}} \,  \times \, \mathrm{t_{sys}}
+\end{equation}
+steps, where $N_{\mathrm{Deg}}$ is the number of (disconnected) 
+local minima at the chosen confidence level, \ie, the number of
+degeneracies.
+
+Eventually, \GLOBES\ needs about $10$ to $15$ seconds to compute
+a projected $\Delta \chi^2$ on a modern Pentium machine. Therefore,
+more sophisticated applications, such as a two dimensional plot
+for CP violation as function of the simulated values of $\stheta$ and
+$\deltacp$ can be obtained in several hours.
+
+\section{Use of package}
+
+The \GLOBES\ package~\cite{Globes} is a tar-Ball which consists of the
+C-library for the user's interface, a binary to test
+\AEDL\ files, example files, 
+experiment prototypes with supporting flux and cross 
+section files, and other 
+supporting files for installation and compilation.
+Currently, the C-library is only provided for Linux.
+In addition, the user's interface and \AEDL\ manual~\cite{Manual} 
+can be obtained separately. Upon extracting the tar-ball,
+the \GLOBES\ user's interface library needs to be configured, 
+compiled, and installed as 
+described in the {\tt INSTALL} file. In principle, it can be
+compiled as a static library or a dynamically-linked shared library
+(which can be installed with superuser rights or only used in one's
+home directory). Then it can be linked against any C program
+containing the application software, such as the examples
+from the manual. \AEDL\ files, such as the experiment protoypes,
+can be edited with any text editor, in order to be loaded by the
+users's interface later. In addition, a binary is provided to
+test and create experiment descriptions. For example, event rate
+information can be quickly provided to adjust the neutrino flux
+normalization. For further information, we refer
+at this place to the \GLOBES\ manual~\cite{Manual}.
 
 \section{Summary and conclusions}
+
+In summary, the \GLOBES\ software package provides powerful tools
+to do a full three-flavor analysis of future long-baseline experiments
+including systematics, correlations, and degeneracies. In addition,
+the abstract experiment definition language allows to define
+experiments at an abstract level as accurate as possible. 
+Some of the major strengths of \GLOBES\ are the ability to quickly
+define new experiments, the potential to take into account the
+full multi-parameter correlation, and the possibilities to include
+external input and matter density uncertainties.
+
+We conclude that the \GLOBES\ software has two major target groups:
+Experimentalists, who want to quickly evaluate the physics potential
+of their setups, and theorists, who want to test new ideas or
+strategies with pre-defined experiments. Especially, the separation
+between experiment descriptions as simple text files (together
+with supporting files) and the application software should allow an
+efficient and lively interaction between those two target groups.
+
+\subsection*{Acknowledgments}
+
+We would like to thank Martin Freund and Andrea Romanino ???
+for initiating this project many years ago with one of the first
+neutrino factory simulations. In addition, we would like to thank
+Marc Rolinec for the help to translate the experiment descriptions
+into \AEDL , and for creating the illustrations in this manuscript.
+Finally, thanks to all the people who have been pushing this project
+for many years, to the ones who have been continuing asking for the publication of the software, and the referees of several of our
+papers for suggestions which lead to improvements in the software.
 

### 2004-07-13 Documentation/Manual.tex
extended the \GLB command to handle a leading '@' (this is meta character for makeindex and thus has to be escaped by '"'). Added a label for section 7.2
added section about globes put in the fdl part on page two, made the sections of fdl disappear from toc added multiple index support changed the \GLB and \GLBNS commands such, that they prepend 'glb' in the text and add their argument to the api index. makeindex has not be called four times: 	makeindex norm.idx 	makeindex api.idx 	makeindex aedl.idx 	makeindex constants.idx
Adding entries now works with \index{index_name}{index_entry}, e.g.
\

@@
 \usepackage{makeidx}
+%Preparations for multiple indices
+\usepackage{multind}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \usepackage{ae,aecompl} % only for Type 1 fonts for readable pdf convert
@@
 % Index: use "makeindex Manual" to create .ind file
-\makeindex
+%\makeindex
+
+\makeindex{norm}
+\makeindex{constants}
+\makeindex{api}
+\makeindex{aedl}
 
@@
 % Always use GLB when mentioning a command in text!
-\newcommand{\GLB}[1]{{\tt #1}\index{{\tt #1}}}
+\newcommand{\GLB}[1]{{\tt glb#1}\index{api}{#1@{\tt #1}}}
 % Always use GLBNS when adding a command to the index without showing it!
-\newcommand{\GLBNS}[1]{\index{{\tt #1}}}
+\newcommand{\GLBNS}[1]{\index{api}{#1@{\tt #1}}}
+
+% Set as GLoBES command and add to index:
+% Always use GLB when mentioning a command in text!
+\newcommand{\GLBC}[1]{{\tt #1}\index{constants}{{\tt  #1}}}
+% Always use GLBNS when adding a command to the index without showing it!
+\newcommand{\GLBNSC}[1]{\index{constants}{{\tt #1}}}
 \newcommand{\gq}[1]{\begin{quote} {\tt #1} \end{quote}}
@@
 \thispagestyle{empty}
+\bigskip
+\begin{quote}
+    Copyright \copyright  2004  The GLoBES Team.
+    Permission is granted to copy, distribute and/or modify this document
+    under the terms of the GNU Free Documentation License, Version 1.2
+    or any later version published by the Free Software Foundation;
+    with no Invariant Sections, no Front-Cover Texts, and no Back-Cover Texts.
+    A copy of the license is included in the section entitled "GNU
+    Free Documentation License".
+\end{quote}
+\bigskip
 
@@
 
-\printindex
+\printindex{api}{API functions}
+\printindex{constants}{API constants \& macros}
+\printindex{aedl}{AEDL reference}
+\printindex{norm}{Index}
 

### 2004-07-13 Documentation/fdl.tex
added section about globes put in the fdl part on page two, made the sections of fdl disappear from toc added multiple index support changed the \GLB and \GLBNS commands such, that they prepend 'glb' in the text and add their argument to the api index. makeindex has not be called four times: 	makeindex norm.idx 	makeindex api.idx 	makeindex aedl.idx 	makeindex constants.idx
Adding entries now works with \index{index_name}{index_entry}, e.g.
\index{norm}{Blau} adds 'Blau' to the usual index.

@@
 {\Large\bf 1. APPLICABILITY AND DEFINITIONS}
-\addcontentsline{toc}{section}{1. APPLICABILITY AND DEFINITIONS}
+%\addcontentsline{toc}{section}{1. APPLICABILITY AND DEFINITIONS}
 \end{center}
@@
 {\Large\bf 2. VERBATIM COPYING}
-\addcontentsline{toc}{section}{2. VERBATIM COPYING}
+%\addcontentsline{toc}{section}{2. VERBATIM COPYING}
 \end{center}
@@
 {\Large\bf 3. COPYING IN QUANTITY}
-\addcontentsline{toc}{section}{3. COPYING IN QUANTITY}
+%\addcontentsline{toc}{section}{3. COPYING IN QUANTITY}
 \end{center}
@@
 {\Large\bf 4. MODIFICATIONS}
-\addcontentsline{toc}{section}{4. MODIFICATIONS}
+%\addcontentsline{toc}{section}{4. MODIFICATIONS}
 \end{center}
@@
 {\Large\bf 5. COMBINING DOCUMENTS}
-\addcontentsline{toc}{section}{5. COMBINING DOCUMENTS}
+%\addcontentsline{toc}{section}{5. COMBINING DOCUMENTS}
 \end{center}
@@
 {\Large\bf 6. COLLECTIONS OF DOCUMENTS}
-\addcontentsline{toc}{section}{6. COLLECTIONS OF DOCUMENTS}
+%\addcontentsline{toc}{section}{6. COLLECTIONS OF DOCUMENTS}
 \end{center}
@@
 {\Large\bf 7. AGGREGATION WITH INDEPENDENT WORKS}
-\addcontentsline{toc}{section}{7. AGGREGATION WITH INDEPENDENT WORKS}
+%\addcontentsline{toc}{section}{7. AGGREGATION WITH INDEPENDENT WORKS}
 \end{center}
@@
 {\Large\bf 8. TRANSLATION}
-\addcontentsline{toc}{section}{8. TRANSLATION}
+%\addcontentsline{toc}{section}{8. TRANSLATION}
 \end{center}
@@
 {\Large\bf 9. TERMINATION}
-\addcontentsline{toc}{section}{9. TERMINATION}
+%\addcontentsline{toc}{section}{9. TERMINATION}
 \end{center}
@@
 {\Large\bf 10. FUTURE REVISIONS OF THIS LICENSE}
-\addcontentsline{toc}{section}{10. FUTURE REVISIONS OF THIS LICENSE}
+%\addcontentsline{toc}{section}{10. FUTURE REVISIONS OF THIS LICENSE}
 \end{center}
@@
 
-\begin{center}
-{\Large\bf ADDENDUM: How to use this License for your documents}
-\addcontentsline{toc}{section}{ADDENDUM: How to use this License for your documents}
-\end{center}
-
-To use this License in a document you have written, include a copy of
-the License in the document and put the following copyright and
-license notices just after the title page:
-
-\bigskip
-\begin{quote}
-    Copyright \copyright  YEAR  YOUR NAME.
-    Permission is granted to copy, distribute and/or modify this document
-    under the terms of the GNU Free Documentation License, Version 1.2
-    or any later version published by the Free Software Foundation;
-    with no Invariant Sections, no Front-Cover Texts, and no Back-Cover Texts.
-    A copy of the license is included in the section entitled "GNU
-    Free Documentation License".
-\end{quote}
-\bigskip
-
-If you have Invariant Sections, Front-Cover Texts and Back-Cover Texts,
-replace the "with...Texts." line with this:
-
-\bigskip
-\begin{quote}
-    with the Invariant Sections being LIST THEIR TITLES, with the
-    Front-Cover Texts being LIST, and with the Back-Cover Texts being LIST.
-\end{quote}
-\bigskip
-
-If you have Invariant Sections without Cover Texts, or some other
-combination of the three, merge those two alternatives to suit the
-situation.
-
-If your document contains nontrivial examples of program code, we
-recommend releasing these examples in parallel under your choice of
-free software license, such as the GNU General Public License,
-to permit their use in free software.
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "Manual"
+%%% End:

### 2004-07-13 Documentation/gpl.tex
added section about globes put in the fdl part on page two, made the sections of fdl disappear from toc added multiple index support changed the \GLB and \GLBNS commands such, that they prepend 'glb' in the text and add their argument to the api index. makeindex has not be called four times: 	makeindex norm.idx 	makeindex api.idx 	makeindex aedl.idx 	makeindex constants.idx
Adding entries now works with \index{index_name}{index_entry}, e.g.
\index{norm}{Blau} adds 'Blau' to the usual index.

\chapter{The GNU General Public License}

\begin{center}
{\parindent 0in

Version 2, June 1991

Copyright \copyright\ 1989, 1991 Free Software Foundation, Inc.

\bigskip

59 Temple Place - Suite 330, Boston, MA  02111-1307, USA

\bigskip

Everyone is permitted to copy and distribute verbatim copies
of this license document, but changing it is not allowed.
}
\end{center}

\begin{center}
{\bf\large Preamble}
\end{center}

The licenses for most software are designed to take away your freedom to
share and change it.  By contrast, the GNU General Public License is
intended to guarantee your freedom to share and change free software---to
make sure the software is free for all its users.  This General Public
License applies to most of the Free Software Foundation's software and to
any other program whose authors commit to using it.  (Some other Free
Software Foundation software is covered by the GNU Library General Public
License instead.)  You can apply it to your programs, too.

When we speak of free software, we are referring to freedom, not price.
Our General Public Licenses are designed to make sure that you have the
freedom to distribute copies of free software (and charge for this service
if you wish), that you receive source code or can get it if you want it,
that you can change the software or use pieces of it in new free programs;
and that you know you can do these things.

To protect your rights, we need to make restrictions that forbid anyone to
deny you these rights or to ask you to surrender the rights.  These
restrictions translate to certain responsibilities for you if you
distribute copies of the software, or if you modify it.

For example, if you distribute copies of such a program, whether gratis or
for a fee, you must give the recipients all the rights that you have.  You
must make sure that they, too, receive or can get the source code.  And
you must show them these terms so they know their rights.

We protect your rights with two steps: (1) copyright the software, and (2)
offer you this license which gives you legal permission to copy,
distribute and/or modify the software.

Also, for each author's protection and ours, we want to make certain that
everyone understands that there is no warranty for this free software.  If
the software is modified by someone else and passed on, we want its
recipients to know that what they have is not the original, so that any
problems introduced by others will not reflect on the original authors'
reputations.

Finally, any free program is threatened constantly by software patents.
We wish to avoid the danger that redistributors of a free program will
individually obtain patent licenses, in effect making the program
proprietary.  To prevent this, we have made it clear that any patent must
be licensed for everyone's free use or not licensed at all.

The precise terms and conditions for copying, distribution and
modification follow.

\begin{center}
{\Large \sc Terms and Conditions For Copying, Distribution and
  Modification}
\end{center}

%\renewcommand{\theenumi}{\alpha{enumi}}
\begin{enumerate}

\addtocounter{enumi}{-1}

\item

This License applies to any program or other work which contains a notice
placed by the copyright holder saying it may be distributed under the
terms of this General Public License.  The ``Program'', below, refers to
any such program or work, and a ``work based on the Program'' means either
the Program or any derivative work under copyright law: that is to say, a
work containing the Program or a portion of it, either verbatim or with
modifications and/or translated into another language.  (Hereinafter,
translation is included without limitation in the term ``modification''.)
Each licensee is addressed as ``you''.

Activities other than copying, distribution and modification are not
covered by this License; they are outside its scope.  The act of
running the Program is not restricted, and the output from the Program
is covered only if its contents constitute a work based on the
Program (independent of having been made by running the Program).
Whether that is true depends on what the Program does.

\item You may copy and distribute verbatim copies of the Program's source
  code as you receive it, in any medium, provided that you conspicuously
  and appropriately publish on each copy an appropriate copyright notice
  and disclaimer of warranty; keep intact all the notices that refer to
  this License and to the absence of any warranty; and give any other
  recipients of the Program a copy of this License along with the Program.

You may charge a fee for the physical act of transferring a copy, and you
may at your option offer warranty protection in exchange for a fee.

\item

You may modify your copy or copies of the Program or any portion
of it, thus forming a work based on the Program, and copy and
distribute such modifications or work under the terms of Section 1
above, provided that you also meet all of these conditions:

\begin{enumerate}

\item

You must cause the modified files to carry prominent notices stating that
you changed the files and the date of any change.

\item

You must cause any work that you distribute or publish, that in
whole or in part contains or is derived from the Program or any
part thereof, to be licensed as a whole at no charge to all third
parties under the terms of this License.

\item
If the modified program normally reads commands interactively
when run, you must cause it, when started running for such
interactive use in the most ordinary way, to print or display an
announcement including an appropriate copyright notice and a
notice that there is no warranty (or else, saying that you provide
a warranty) and that users may redistribute the program under
these conditions, and telling the user how to view a copy of this
License.  (Exception: if the Program itself is interactive but
does not normally print such an announcement, your work based on
the Program is not required to print an announcement.)

\end{enumerate}

These requirements apply to the modified work as a whole.  If
identifiable sections of that work are not derived from the Program,
and can be reasonably considered independent and separate works in
themselves, then this License, and its terms, do not apply to those
sections when you distribute them as separate works.  But when you
distribute the same sections as part of a whole which is a work based
on the Program, the distribution of the whole must be on the terms of
this License, whose permissions for other licensees extend to the
entire whole, and thus to each and every part regardless of who wrote it.

Thus, it is not the intent of this section to claim rights or contest
your rights to work written entirely by you; rather, the intent is to
exercise the right to control the distribution of derivative or
collective works based on the Program.

In addition, mere aggregation of another work not based on the Program
with the Program (or with a work based on the Program) on a volume of
a storage or distribution medium does not bring the other work under
the scope of this License.

\item
You may copy and distribute the Program (or a work based on it,
under Section 2) in object code or executable form under the terms of
Sections 1 and 2 above provided that you also do one of the following:

\begin{enumerate}

\item

Accompany it with the complete corresponding machine-readable
source code, which must be distributed under the terms of Sections
1 and 2 above on a medium customarily used for software interchange; or,

\item

Accompany it with a written offer, valid for at least three
years, to give any third party, for a charge no more than your
cost of physically performing source distribution, a complete
machine-readable copy of the corresponding source code, to be
distributed under the terms of Sections 1 and 2 above on a medium
customarily used for software interchange; or,

\item

Accompany it with the information you received as to the offer
to distribute corresponding source code.  (This alternative is
allowed only for noncommercial distribution and only if you
received the program in object code or executable form with such
an offer, in accord with Subsection b above.)

\end{enumerate}

The source code for a work means the preferred form of the work for
making modifications to it.  For an executable work, complete source
code means all the source code for all modules it contains, plus any
associated interface definition files, plus the scripts used to
control compilation and installation of the executable.  However, as a
special exception, the source code distributed need not include
anything that is normally distributed (in either source or binary
form) with the major components (compiler, kernel, and so on) of the
operating system on which the executable runs, unless that component
itself accompanies the executable.

If distribution of executable or object code is made by offering
access to copy from a designated place, then offering equivalent
access to copy the source code from the same place counts as
distribution of the source code, even though third parties are not
compelled to copy the source along with the object code.

\item
You may not copy, modify, sublicense, or distribute the Program
except as expressly provided under this License.  Any attempt
otherwise to copy, modify, sublicense or distribute the Program is
void, and will automatically terminate your rights under this License.
However, parties who have received copies, or rights, from you under
this License will not have their licenses terminated so long as such
parties remain in full compliance.

\item
You are not required to accept this License, since you have not
signed it.  However, nothing else grants you permission to modify or
distribute the Program or its derivative works.  These actions are
prohibited by law if you do not accept this License.  Therefore, by
modifying or distributing the Program (or any work based on the
Program), you indicate your acceptance of this License to do so, and
all its terms and conditions for copying, distributing or modifying
the Program or works based on it.

\item
Each time you redistribute the Program (or any work based on the
Program), the recipient automatically receives a license from the
original licensor to copy, distribute or modify the Program subject to
these terms and conditions.  You may not impose any further
restrictions on the recipients' exercise of the rights granted herein.
You are not responsible for enforcing compliance by third parties to
this License.

\item
If, as a consequence of a court judgment or allegation of patent
infringement or for any other reason (not limited to patent issues),
conditions are imposed on you (whether by court order, agreement or
otherwise) that contradict the conditions of this License, they do not
excuse you from the conditions of this License.  If you cannot
distribute so as to satisfy simultaneously your obligations under this
License and any other pertinent obligations, then as a consequence you
may not distribute the Program at all.  For example, if a patent
license would not permit royalty-free redistribution of the Program by
all those who receive copies directly or indirectly through you, then
the only way you could satisfy both it and this License would be to
refrain entirely from distribution of the Program.

If any portion of this section is held invalid or unenforceable under
any particular circumstance, the balance of the section is intended to
apply and the section as a whole is intended to apply in other
circumstances.

It is not the purpose of this section to induce you to infringe any
patents or other property right claims or to contest validity of any
such claims; this section has the sole purpose of protecting the
integrity of the free software distribution system, which is
implemented by public license practices.  Many people have made
generous contributions to the wide range of software distributed
through that system in reliance on consistent application of that
system; it is up to the author/donor to decide if he or she is willing
to distribute software through any other system and a licensee cannot
impose that choice.

This section is intended to make thoroughly clear what is believed to
be a consequence of the rest of this License.

\item
If the distribution and/or use of the Program is restricted in
certain countries either by patents or by copyrighted interfaces, the
original copyright holder who places the Program under this License
may add an explicit geographical distribution limitation excluding
those countries, so that distribution is permitted only in or among
countries not thus excluded.  In such case, this License incorporates
the limitation as if written in the body of this License.

\item
The Free Software Foundation may publish revised and/or new versions
of the General Public License from time to time.  Such new versions will
be similar in spirit to the present version, but may differ in detail to
address new problems or concerns.

Each version is given a distinguishing version number.  If the Program
specifies a version number of this License which applies to it and ``any
later version'', you have the option of following the terms and conditions
either of that version or of any later version published by the Free
Software Foundation.  If the Program does not specify a version number of
this License, you may choose any version ever published by the Free Software
Foundation.

\item
If you wish to incorporate parts of the Program into other free
programs whose distribution conditions are different, write to the author
to ask for permission.  For software which is copyrighted by the Free
Software Foundation, write to the Free Software Foundation; we sometimes
make exceptions for this.  Our decision will be guided by the two goals
of preserving the free status of all derivatives of our free software and
of promoting the sharing and reuse of software generally.

\begin{center}
{\Large\sc
No Warranty
}
\end{center}

\item
{\sc Because the program is licensed free of charge, there is no warranty
for the program, to the extent permitted by applicable law.  Except when
otherwise stated in writing the copyright holders and/or other parties
provide the program ``as is'' without warranty of any kind, either expressed
or implied, including, but not limited to, the implied warranties of
merchantability and fitness for a particular purpose.  The entire risk as
to the quality and performance of the program is with you.  Should the
program prove defective, you assume the cost of all necessary servicing,
repair or correction.}

\item
{\sc In no event unless required by applicable law or agreed to in writing
will any copyright holder, or any other party who may modify and/or
redistribute the program as permitted above, be liable to you for damages,
including any general, special, incidental or consequential damages arising
out of the use or inability to use the program (including but not limited
to loss of data or data being rendered inaccurate or losses sustained by
you or third parties or a failure of the program to operate with any other
programs), even if such holder or other party has been advised of the
possibility of such damages.}

\end{enumerate}

\begin{center}
{\Large\sc End of Terms and Conditions}
\end{center}

### 2004-07-13 Documentation/part1.tex
extended the \GLB command to handle a leading '@' (this is meta character for makeindex and thus has to be escaped by '"'). Added a label for section 7.2
Minor changes
added section about globes put in the fdl part on page two, made the sections of fdl disappear from toc added multiple index support changed the \GLB and \GLBNS commands such, that they prepend 'glb' in the text and add their argument to the api index. makeindex has not be called four times: 	makeindex norm.idx 	makeindex api.idx 	makeindex aedl.idx 	makeindex constants.idx
Adding entries now works with \index{index_name}{index_

@@
 \label{chapter:tour}
-\index{GLoBES tour}
+\index{norm}{GLoBES tour}
 
@@
 \multicolumn{3}{l}{{\bf Systematics only:}} \\
-\GLB{glbChiSys} & $\chi^2$ with systematics only  & ({\tt glb\_params in, int exp, int rule}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
+\GLB{ChiSys} & $\chi^2$ with systematics only  & ({\tt glb\_params in, int exp, int rule}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
 \multicolumn{3}{l}{{\bf Projections onto axes:}} \\
-\GLB{glbChiTheta} & Projection onto $\theta_{13}$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
-\GLB{glbChiDelta} & Projection onto $\deltacp$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
-\GLB{glbChiTheta23} & Projection onto $\theta_{23}$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
-\GLB{glbChiDm} & Projection onto $\ldm$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
-\GLB{glbChiDms} & Projection onto $\sdm$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
+\GLB{ChiTheta} & Projection onto $\theta_{13}$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
+\GLB{ChiDelta} & Projection onto $\deltacp$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
+\GLB{ChiTheta23} & Projection onto $\theta_{23}$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
+\GLB{ChiDm} & Projection onto $\ldm$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
+\GLB{ChiDms} & Projection onto $\sdm$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
 \multicolumn{3}{l}{{\bf Projection onto plane:}} \\
-\GLB{glbChiThetaDelta} & Projection onto $\theta_{13}$-$\deltacp$-plane  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
+\GLB{ChiThetaDelta} & Projection onto $\theta_{13}$-$\deltacp$-plane  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
 \multicolumn{3}{l}{{\bf Projection onto any hyper-plane:}} \\
-\GLB{glbChiNP} & Projection onto any $n$-dimensional hyper-plane  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \newline
-Needs \GLB{glbSetProjection} before! \\[0.2cm]
+\GLB{ChiNP} & Projection onto any $n$-dimensional hyper-plane  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \newline
+Needs \GLB{SetProjection} before! \\[0.2cm]
 \multicolumn{3}{l}{{\bf Localization of degeneracies:}} \\
-\GLB{glbChiAll} & (Local) Minimization over all parameters  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\
+\GLB{ChiAll} & (Local) Minimization over all parameters  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\
 \hline
@@
 \end{center}
-\caption{\label{tab:stdfunctions} \index{Standard functions (table)} The \GLOBES\ standard function to obtain a $\chi^2$-value with systematics only or systematics and correlations.
+\caption{\label{tab:stdfunctions} \index{norm}{Standard functions (table)} The \GLOBES\ standard function to obtain a $\chi^2$-value with systematics only or systematics and correlations.
 The parameters {\tt rule} and {\tt exp}
-can either be \GLB{GLB\_ALL} for all initialized experiment or the
-experiment number ($0$ to \GLB{glb\_num\_of\_exps}-1) for a specific experiment.
-The format of \GLB{glb\_params} is discussed in detail in \Chapt~\ref{chapt:gettingstarted}.
+can either be \GLBC{GLB\_ALL} for all initialized experiment or the
+experiment number ($0$ to \GLB{\_num\_of\_exps}-1) for a specific experiment.
+The format of \GLB{\_params} is discussed in detail in \Chapt~\ref{chapt:gettingstarted}.
 Note that all functions but {\tt ChiSys}
-  are using minimizers which have to be initialized with \GLB{glbSetInputErrors} and \GLB{glbSetStartingValues} first.}
+  are using minimizers which have to be initialized with \GLB{SetInputErrors} and \GLB{SetStartingValues} first.}
 \end{table}
@@
 
-\index{Installation}
+\index{norm}{Installation}
 In this first chapter of the user's manual, we assume that the \GLOBES\ software is readily installed on your computer system. For the installation,
@@
 Furthermore, the files of the examples in this part can be found in the {\tt Example} subdirectory of your \GLOBES\ distribution.
-\index{Examples} Since it depends on your installed
+\index{norm}{Examples} Since it depends on your installed
 configuration, we refer to the {\tt README} file in the \GLOBES\ main directory, and the comments in the Makefile of the {\tt Example} subdirectory for how to compile the example files. Make sure that
@@
 \example{Using \GLOBES\ with C}{\label{ex:c}
-\index{C-Code}
-\index{Program}
+\index{norm}{C-Code}
+\index{norm}{Program}
 Here comes the C-code skeleton, which is
@@
 \end{center}
-\mycaption{\label{tab:experiments} \index{Experiment files (table)}
-\index{{\tt glb}-files} 
-Pre-defined experiment prototypes, their filenames (to be used in \GLB{glbInitExperiment}), their short descriptions, and the references in which they are originally used and discussed (except from minor modifications, such as a
+\mycaption{\label{tab:experiments} \index{norm}{Experiment files (table)}
+\index{norm}{{\tt glb}-files} 
+Pre-defined experiment prototypes, their filenames (to be used in \GLB{InitExperiment}), their short descriptions, and the references in which they are originally used and discussed (except from minor modifications, such as a
 different implementation of the energy threshold function). Note that
@@
 \section{Initialization of \GLOBES}
-\index{Initialization}
+\index{norm}{Initialization}
 
 Before one can use \GLOBES , one has to initialize the \GLOBES\
-library \GLB{libglobes}: \index{Initialization: \GLOBES\ library}
-\begin{function}
-\index{Initialization of {\tt libglobes}}
-\GLBNS{glbInit} {\tt void glbInit(char *name)} initializes the library {\tt libglobes} and has
+library \index{norm}{libglobes@{\tt libglobes}}: \index{norm}{Initialization!\GLOBES\ library}
+\begin{function}
+\index{norm}{Initialization!libglobes@{\tt libglobes}}
+\GLBNS{Init} {\tt void glbInit(char *name)} initializes the library {\tt libglobes} and has
 to be called in the beginning of each \GLOBES\ program. It takes the
@@
 
-\index{Number of experiments}
+\index{norm}{Experiment!number of}
 In principle, the \GLOBES\ user interface can currently handle up to 32 of different long-baseline experiments simultaneously, where the number
@@
 In general, many \GLOBES\ functions take the experiment number as
-a parameter, which runs from $0$ to \GLB{glb\_num\_of\_exps}-1 in the order of their initialization in the program.\footnote{Note that
+a parameter, which runs from $0$ to \GLB{\_num\_of\_exps}-1 in the order of their initialization in the program.\footnote{Note that
 the global variable {\tt glb\_num\_of\_exps} must not be modified by the
-user.} In addition, using the parameter value \GLB{GLB\_ALL} as
+user.} In addition, using the parameter value \GLBC{GLB\_ALL} as
 experiment number initiates a combined analysis of all loaded experiments.
 
-\index{Experiment list}
-\index{Initialization: Experiments}
-For storing the experiments, \GLOBES\ uses the initially empty list of experiments \GLB{glb\_experiment\_list}.
+\index{norm}{Experiment!list}
+\index{norm}{Initialization!experiments}
+For storing the experiments, \GLOBES\ uses the initially empty list of experiments \GLB{\_experiment\_list}.
 To add a pre-defined experiment to this list, one can use the function {\tt glbInitExperiment}:
 \begin{function}
-\index{Experiment initialization}
-\GLBNS{glbInitExperiment}
+\index{norm}{Experiment initialization}
+\GLBNS{InitExperiment}
 {\tt int glbInitExperiment(char *inf, glb\_exp *in, int *counter)}
@@
 experiment then has the number {\tt glb\_num\_of\_exps}-1. The elements
-of the experiment list have the type \GLB{glb\_exp}, which the
+of the experiment list have the type \GLB{\_exp}, which the
 user will not need to access directly in most cases. The experiment definition files, which usually end with {\tt .glb}, and any
-supporting files, are first of all searched in the current directory, and then in the global path variable \GLB{GLB\_PATH} set in your environment.
+supporting files, are first of all searched in the current directory, and then in the global path variable \GLBC{GLB\_PATH} set in your environment.
 A list of pre-defined experiment prototypes, their filenames, their short descriptions, and the references of their definitions can be found in \tabl{experiments}.
@@
 
-\index{Clear experiment list}
-\index{Delete experiments}
+\index{norm}{Clear experiment list}
+\index{norm}{Delete experiments}
 One can also remove all experiments from the evaluation list at running
@@
 \begin{function}
-\GLBNS{glbClearExperimentList}
+\GLBNS{ClearExperimentList}
 {\tt void glbClearExperimentList()} removes all experiments from the internal list and resets all counters.   
@@
 Note that changing the number of experiments requires a new initialization
-of all parameters of the types \GLB{glb\_params} and \GLB{glb\_projection}
+of all parameters of the types \GLB{\_params} and \GLB{\_projection}
 if the number of experiments changes, since these parameter structures internally carry lists for the matter densities of all experiments.
@@
 \end{tabular}
-\mycaption{\label{tab:units} \index{Units in \GLOBES\ (table)} Quantities used in \GLOBES , examples of these quantities, and their standard units in the application software.}
+\mycaption{\label{tab:units} \index{norm}{Units in \GLOBES\ (table)} Quantities used in \GLOBES , examples of these quantities, and their standard units in the application software.}
 \end{center}
@@
 \label{sec:luminosity}
-\index{Integrated luminosity}
-\index{Detector mass} \index{Source power} \index{Running time}
+\index{norm}{Integrated luminosity}
+\index{norm}{Detector mass} \index{norm}{Source power} \index{norm}{Running time}
 
@@
 \begin{function}
-\GLBNS{glbSetSourcePower}
+\GLBNS{SetSourcePower}
 {\tt void glbSetSourcePower(double power, int exp, int fluxno)}
@@
 \begin{function}
-\GLBNS{glbGetSourcePower}
+\GLBNS{GetSourcePower}
 {\tt double glbGetSourcePower(int exp, int fluxno)}
@@
 \begin{function}
-\GLBNS{glbSetRunningTime}
+\GLBNS{SetRunningTime}
 {\tt void glbSetRunningTime(double time, int exp, int fluxno)}
@@
 \begin{function}
-\GLBNS{glbGetRunningTime}
+\GLBNS{GetRunningTime}
 {\tt double glbGetRunningTime(int exp, int fluxno)}
@@
 \begin{function}
- \GLBNS{glbSetTargetMass}
+ \GLBNS{SetTargetMass}
 {\tt void glbSetTargetMass(double mass,int exp)}
@@
 \begin{function}
- \GLBNS{glbGetTargetMass}
+ \GLBNS{GetTargetMass}
 {\tt double glbGetTargetMass(int exp)}
@@
 \section{Handling oscillation parameter vectors}
-\index{Oscillation parameter vectors}
+\index{norm}{Oscillation!parameter vectors}
 
@@
 for oscillation parameters. In order to transfer sets of oscillation
-parameter vectors $(\theta_{12}, \theta_{13}, \theta_{23}, \deltacp, \sdm, \ldm)$, the parameter type \GLB{glb\_params\_type} is used.
-In general, pointers to this structure are of the type \GLB{glb\_params}, and are often
+parameter vectors $(\theta_{12}, \theta_{13}, \theta_{23}, \deltacp, \sdm, \ldm)$, the parameter type \GLB{\_params\_type} is used.
+In general, pointers to this structure are of the type \GLB{\_params}, and are often
 transferred to and from \GLOBES\ functions. Therefore, the memory space
@@
 profile of each experiment $i$ is multiplied by a scaling factor
-$\hat{\rho_i}$, \index{Matter density scaling factor}
-\index{Matter density profile} 
+$\hat{\rho_i}$, \index{norm}{Matter density scaling factor}
+\index{norm}{Matter density profile} 
 which is stored in the density information of {\tt glb\_params}. 
@@
 \begin{function}
-\GLBNS{glbAllocParams}
+\GLBNS{AllocParams}
 {\tt glb\_params glbAllocParams()} allocates the memory space
@@
 \begin{function}
-\GLBNS{glbAllocParams}
+\GLBNS{AllocParams}
 {\tt void glbFreeParams(glb\_params stale)} destroys the memory space
@@
 \begin{function}
-\GLBNS{glbDefineParams}
+\GLBNS{DefineParams}
 {\tt glb\_params glbDefineParams(glb\_params in, double theta12, double theta13,double theta23, double delta, double dms, double dma)} assigns
@@
 \begin{function}
- \GLBNS{glbCopyParams}
+ \GLBNS{CopyParams}
  {\tt glb\_params glbCopyParams(const glb\_params source, glb\_params dest)}
@@
 \begin{function}
-\GLBNS{glbPrintParams}
+\GLBNS{PrintParams}
 {\tt void glbPrintParams(FILE *stream, const glb\_params in)} prints the
@@
 \begin{function}
-\GLBNS{glbSetOscParams}
+\GLBNS{SetOscParams}
 {\tt glb\_params glbSetOscParams(glb\_params in, double osc, int which)}
@@
 \begin{function}
-\GLBNS{glbGetOscParams}
+\GLBNS{GetOscParams}
 {\tt double glbGetOscParams(glb\_params in, int which)}
@@
 $0$ to $5$, where the parameters in \GLOBES\ always have the order $\theta_{12}$, $\theta_{13}$, $\theta_{23}$, $\deltacp$, $\sdm$, $\ldm$.
-Alternatively to the number, the constants \GLB{GLB\_THETA\_12}, \GLB{GLB\_THETA\_13}, \GLB{GLB\_THETA\_23}, \GLB{GLB\_DELTA\_CP}, 
-\GLB{GLB\_DM\_SOL}, or \GLB{GLB\_DM\_ATM} can be used.
+Alternatively to the number, the constants \GLBC{GLB\_THETA\_12}, \GLBC{GLB\_THETA\_13}, \GLBC{GLB\_THETA\_23}, \GLBC{GLB\_DELTA\_CP}, 
+\GLBC{GLB\_DM\_SOL}, or \GLBC{GLB\_DM\_ATM} can be used.
 
@@
 \begin{function}
-\GLBNS{glbSetDensityParams}
+\GLBNS{SetDensityParams}
 {\tt glb\_params glbSetDensityParams(glb\_params in, double dens, int which)} sets the density parameter {\tt which} in the structure {\tt in}
@@
 \begin{function}
-\GLBNS{glbGetDensityParams}
+\GLBNS{GetDensityParams}
 {\tt double glbGetDensityParams(glb\_params in, int which)}
@@
 \begin{function}
-\GLBNS{glbSetIteration} \index{Iterations}
+\GLBNS{SetIteration} \index{norm}{Iterations}
 {\tt glb\_params glbSetIteration(glb\_params in, int iter)} sets the number of iterations in the structure {\tt in}
@@
 \begin{function}
-\GLBNS{glbGetIteration}
+\GLBNS{GetIteration}
 {\tt int glbGetIteration(glb\_params in)}
@@
 In total, the parameter vector handling in a program normally has the
-following order:\index{Parameter vector handling}
+following order:\index{norm}{Parameter vector handling}
 \begin{quote}
@@
 \section{Computing the simulated data}
-\index{True values} \index{Simulated data} \index{Reference rate vector}
-\index{Set oscillation parameters}
+\index{norm}{True values} \index{norm}{Simulated data} \index{norm}{Reference rate vector}
+\index{norm}{Set oscillation parameters}
 
@@
 Therefore, it is often useful to show the results of a future measurement as function of the true parameter values for which the reference rate vectors are computed -- at least within the currently allowed ranges.
-The true parameter values for the vacuum neutrino oscillation parameters have to be set by the functions \GLB{glbSetOscillationParameters} and \GLB{glbSetRates} {\em before} any evaluation function is used and {\em after} the experiments have been initialized and the experiment parameters have been adjusted which could change the rates (such as baseline or target mass).
+The true parameter values for the vacuum neutrino oscillation parameters have to be set by the functions \GLB{SetOscillationParameters} and \GLB{SetRates} {\em before} any evaluation function is used and {\em after} the experiments have been initialized and the experiment parameters have been adjusted which could change the rates (such as baseline or target mass).
 This means that
@@
 \begin{function}
-\GLBNS{glbSetOscillationParameters}
+\GLBNS{SetOscillationParameters}
 {\tt int glbSetOscillationParameters(const glb\_params in)} sets the
@@
 \begin{function}
-\GLBNS{glbGetOscillationParameters}
+\GLBNS{GetOscillationParameters}
 {\tt int glbGetOscillationParameters(glb\_params out)} returns the
@@
 \begin{function}
-\GLBNS{glbSetRates}
+\GLBNS{SetRates}
 {\tt void glbSetRates()} computes the reference rate vector for the neutrino oscillation parameters set by {\tt glbSetOscillationParameters}. 
@@
 \section{Version control}
-\index{Version control}
+\label{sec:versioning}
+\index{norm}{Version control}
 
@@
 \begin{function}
-\GLBNS{glbTestReleaseVersion}
+\GLBNS{TestReleaseVersion}
 {\tt int glbTestReleaseVersion(const char *version)} returns $0$ if
@@
 \begin{function}
-\GLBNS{glbTestLibraryVersion}
+\GLBNS{TestLibraryVersion}
 {\tt int glbTestLibraryVersion(const char *version)} returns $0$ if
@@
 \begin{function}
-\GLBNS{glbVersionOfExperiment} 
+\GLBNS{VersionOfExperiment} 
 {\tt const char* glbVersionOfExperiment(int experiment)} returns
@@
 \chapter[Calculating $\chi^2$ with systematics only]{Calculating $\boldsymbol{\chi^2}$ with systematics only}
-\index{Systematics $\chi^2$}
+\index{norm}{Systematics $\chi^2$}
 
@@
 In general, \GLOBES\ uses the six independent oscillation parameters $\theta_{12}$, $\theta_{13}$, $\theta_{23}$, $\deltacp$, $\sdm$, $\ldm$, as well as the matter density
-scaling factor $\hat{\rho}$ of each experiment.\index{Matter density
+scaling factor $\hat{\rho}$ of each experiment.\index{norm}{Matter density
 scaling factor} Thus, there are six plus the number of experiments parameters determining the rate vectors.
@@
 given by observations is imposed (typically up to $5\%$). 
-\index{Matter density uncertainty}
+\index{norm}{Matter density uncertainty}
  For this section, it is important to keep in mind
@@
 last section, the mechanism for the matter density scaling factors
-is hidden in the definition of \GLB{glb\_params}: Each of the scaling
+is hidden in the definition of \GLB{\_params}: Each of the scaling
 factors is initially set to $1.0$. Therefore, for the calculation of
@@
 \label{ex:corrth13dcp}
-\index{Two-parameter correlation}
-A typical and fast application for \GLB{glbChiSys} is the visualization of
+\index{norm}{Two-parameter correlation}
+A typical and fast application for \GLB{ChiSys} is the visualization of
  two-parameter correlations using systematics only. For example, to
@@
 \begin{function} 
-\GLBNS{glbChiSys}
+\GLBNS{ChiSys}
 {\tt double glbChiSys(const glb\_params in,int exp, int rule)} returns
@@
 experiment number {\tt exp}, and the rule number {\tt rule}. For all
-experiments or rules, use \GLB{GLB\_ALL} as parameter value.
+experiments or rules, use \GLBC{GLB\_ALL} as parameter value.
 \end{function}
@@
 
-\index{Systematics} \index{Pull method}
+\index{norm}{Systematics} \index{norm}{Pull method}
 The treatment of systematics in \GLOBES\ is performed by the so-call
@@
  One example for a systematics parameter the signal normalization error, \ie, an error to the overall normalization of the signal.
-For illustration, we assume that the signal event rate in the $i$th bin $s_i^0$ of one oscillation channel is altered by the overall normalization auxiliary parameter\index{Auxiliary parameter} of this channel, \ie ,
+For illustration, we assume that the signal event rate in the $i$th bin $s_i^0$ of one oscillation channel is altered by the overall normalization auxiliary parameter\index{norm}{Auxiliary parameter} of this channel, \ie ,
 \be
@@
 
-The systematics minimization of an experiment can be easily switched on and off with \GLB{glbSwitchSystematics}, \ie, one can also compute the $\chi^2$ with statistics only.
+The systematics minimization of an experiment can be easily switched on and off with \GLB{SwitchSystematics}, \ie, one can also compute the $\chi^2$ with statistics only.
 In addition, several options for
@@
 \label{chapt:correlations}
-\index{Correlations $\chi^2$}
-
-\index{Multi-parameter correlation}
+\index{norm}{Correlations $\chi^2$}
+
+\index{norm}{Multi-parameter correlation}
 This chapter deals with the rather complicated issue of $n$-parameter correlations. It is one of the greatest strengths of this software 
@@
 
-\index{Projection of manifold}
+\index{norm}{Projection of manifold}
 In principle, the precision of an individual parameter measurement including correlations in the $\chi^2$-approach can be obtained as the projection of the $n$-dimensional fit manifold onto the respective axis.
@@
 
-\index{Minimizer}
+\index{norm}{Minimizer}
 The solution to this problem is using a local $n$-dimensional (local) minimizer for the projection instead of a grid-based method, where we will
@@
 
-\index{Priors} \index{Input errors} \index{Starting values}
+\index{norm}{Priors} \index{norm}{Input errors} \index{norm}{Starting values}
 In many cases, the fit manifold is restricted by the knowledge from earlier experiments.
@@
 
-\index{External input}
+\index{norm}{External input}
 It is one of the strengths of the \GLOBES\ software to use external input in order to reduce the fit manifold with the knowledge from external
@@
 For example, for the matter density, one obtains as the final projected $\chi^2_F$ after minimization over the matter density
-scaling factor \index{Matter density scaling factor} $\hat{\rho}$
+scaling factor \index{norm}{Matter density scaling factor} $\hat{\rho}$
 \be
@@
 
-\index{Input errors} \index{Starting values}
+\index{norm}{Input errors} \index{norm}{Starting values}
 In order to set the starting values and input errors, two function have to
@@
 \begin{function}
-\GLBNS{glbSetStartingValues}
+\GLBNS{SetStartingValues}
 {\tt int glbSetStartingValues(const glb\_params in)} sets the starting values for all of the following minimizer calls to {\tt in}.
@@
 \begin{function}
-\GLBNS{glbSetInputErrors}
+\GLBNS{SetInputErrors}
 {\tt int glbSetInputErrors(const glb\_params in)}
@@
 \begin{function}
-\GLBNS{glbGetStartingValues}
+\GLBNS{GetStartingValues}
 {\tt int glbGetStartingValues(glb\_params out)} writes the currently
@@
 \begin{function}
-\GLBNS{glbGetInputErrors}
+\GLBNS{GetInputErrors}
 {\tt int glbGetInputErrors(glb\_params out)}
@@
 
-\index{External precision}
+\index{norm}{External precision}
 Eventually, a typical initialization of the external input with
@@
 
-\index{Priors}
-\index{Advanced tricks}
+\index{norm}{Priors}
+\index{norm}{Advanced tricks}
 Though the priors are an elegant way to treat external input, there are also some complications with priors. The following hints are for the
@@
  kept fixed. This will be important together with arbitrary projections
- using \GLB{glbChiNP}. A more subtile consequence is the comparison
+ using \GLB{ChiNP}. A more subtile consequence is the comparison
  of fit manifold sections and projections for the solutions where
@@
 \label{ex:corrproj}
-\index{Projection of manifold}
-\index{Two-parameter correlation}
-\index{Multi-parameter correlation}
+\index{norm}{Projection of manifold}
+\index{norm}{Two-parameter correlation}
+\index{norm}{Multi-parameter correlation}
 
@@
 \section[Projection onto the $\stheta$-axis or $\deltacp$-axis]{Projection onto the $\boldsymbol{\stheta}$- or $\boldsymbol{\deltacp}$-axis}
-\index{Projection onto axis}
+\index{norm}{Projection onto axis}
 
@@
 
-\index{Minimizer}
+\index{norm}{Minimizer}
 The following functions are some of the simplest minimizers 
@@
 \begin{function}
-\GLBNS{glbChiTheta}
+\GLBNS{ChiTheta}
 {\tt double glbChiTheta(const glb\_params in, glb\_params out, int exp)}
@@
 \begin{function}
-\GLBNS{glbChiDelta}
+\GLBNS{ChiDelta}
 {\tt double glbChiDelta(const glb\_params in, glb\_params out, int exp)}
@@
 It is very often useful to print the output of the minimization with
-\GLB{glbPrintParams} in order to check that the minimum is the
+\GLB{PrintParams} in order to check that the minimum is the
 appropriate one. For example, if the minimizer ends up in the wrong-sign
@@
 % 
-Note that before any minimization, \GLB{glbSetStartingValues} 
-and \GLB{glbSetInputErrors} have to be used at least once.
+Note that before any minimization, \GLB{SetStartingValues} 
+and \GLB{SetInputErrors} have to be used at least once.
 In addition, note that the resulting $\chi^2$ of {\tt glbChiTheta} (or {\tt glbChiDelta}) for the combination of more than one experiment is not
@@
 
-\index{Matter density scaling factor}
+\index{norm}{Matter density scaling factor}
 The the output of the minimizer in {\tt out} carries as many matter density scaling factors as there are experiments. Either
@@
 Note that it uses the function
-\GLB{glbChiNP} for arbitrary projections from the next section for the minimization over $\deltacp$.
+\GLB{ChiNP} for arbitrary projections from the next section for the minimization over $\deltacp$.
 In addition, there is one interesting feature in guessing the
@@
 
-Other examples for projections onto a parameter axis while keeping exactly one parameter fixed are \GLB{glbChiTheta23}, \GLB{glbChiDm}, and \GLB{glbChiDms}, which
+Other examples for projections onto a parameter axis while keeping exactly one parameter fixed are \GLB{ChiTheta23}, \GLB{ChiDm}, and \GLB{ChiDms}, which
 can be found in \Tab~\ref{tab:stdfunctions} on page~\pageref{tab:stdfunctions}.
@@
 \section[Projection onto any hyperplane]{Projection onto any hyperplane}
-\index{Projection onto hyperplane}
+\index{norm}{Projection onto hyperplane}
 
@@
 minimized over. 
-The corresponding function is \index{Projection onto $\theta_{13}$-$\deltacp$-plane}
-\begin{function}
-\GLBNS{glbChiThetaDelta}
+The corresponding function is \index{norm}{Projection onto $\theta_{13}$-$\deltacp$-plane}
+\begin{function}
+\GLBNS{ChiThetaDelta}
 {\tt double glbChiThetaDelta(const glb\_params in, glb\_params out, int exp)} returns the projected $\chi^2$ onto the $\theta_{13}$-$\deltacp$-plane for the  experiment {\tt exp}.
@@
 \hline
-\GLB{glbAllocProjection} & Create projection vector & {\tt ()} \\
-\GLB{glbFreeProjection} & Destroy projection vector {\tt stale} & {\tt (glb\_projection stale)} \\
-\GLB{glbDefineProjection} & Assign projection vector {\tt in} & {\tt (glb\_projection in, int theta12, int theta13, int theta23, int delta, int dms, int dma)} \\ 
-\GLB{glbCopyProjection} & Copy vector {\tt source} to {\tt dest} & {\tt (const glb\_projection source, glb\_projection dest)} \\
-\GLB{glbPrintProjection} & Print vector {\tt in} to file {\tt stream} & {\tt (FILE* stream, const glb\_projection in)} \\
-\GLB{glbSetProjectionFlag} & Set flag for oscillation parameter {\tt which} in vector {\tt in} to value {\tt flag}. & {\tt (glb\_projection in, int flag, int which)} \\
-\GLB{glbGetProjectionFlag} & Return flag for oscillation parameter {\tt which} in vector {\tt in}. & {\tt (const glb\_projection in, int which)} $\rightarrow$ {\tt int} flag \\
- {\tt glbSetDensity\-ProjectionFlag} \GLBNS{glbSetDensityProjectionFlag} & Set flag for density parameter {\tt which} in vector {\tt in} to value {\tt flag}.
+\GLB{AllocProjection} & Create projection vector & {\tt ()} \\
+\GLB{FreeProjection} & Destroy projection vector {\tt stale} & {\tt (glb\_projection stale)} \\
+\GLB{DefineProjection} & Assign projection vector {\tt in} & {\tt (glb\_projection in, int theta12, int theta13, int theta23, int delta, int dms, int dma)} \\ 
+\GLB{CopyProjection} & Copy vector {\tt source} to {\tt dest} & {\tt (const glb\_projection source, glb\_projection dest)} \\
+\GLB{PrintProjection} & Print vector {\tt in} to file {\tt stream} & {\tt (FILE* stream, const glb\_projection in)} \\
+\GLB{SetProjectionFlag} & Set flag for oscillation parameter {\tt which} in vector {\tt in} to value {\tt flag}. & {\tt (glb\_projection in, int flag, int which)} \\
+\GLB{GetProjectionFlag} & Return flag for oscillation parameter {\tt which} in vector {\tt in}. & {\tt (const glb\_projection in, int which)} $\rightarrow$ {\tt int} flag \\
+ {\tt glbSetDensity\-ProjectionFlag} \GLBNS{SetDensityProjectionFlag} & Set flag for density parameter {\tt which} in vector {\tt in} to value {\tt flag}.
 & {\tt (glb\_projection in, int flag, int which)} \\
-{\tt glbGetDensity\-ProjectionFlag} \GLBNS{glbGetDensityProjectionFlag} & Return flag for density parameter {\tt which} in vector {\tt in}.
+{\tt glbGetDensity\-ProjectionFlag} \GLBNS{GetDensityProjectionFlag} & Return flag for density parameter {\tt which} in vector {\tt in}.
 & {\tt (const glb\_projection in, int which)} $\rightarrow$ {\tt int} flag \\
@@
 \end{tabular}
-\caption{\label{tab:defprojection}\index{Projection type}
+\caption{\label{tab:defprojection}\index{norm}{Projection type}
 Different functions handling the
-\GLB{glb\_projection} type.
-Flags are either \GLB{GLB\_FIXED} or \GLB{GLB\_FREE}.
+\GLB{\_projection} type.
+Flags are either \GLBC{GLB\_FIXED} or \GLBC{GLB\_FREE}.
 The (un-shown) return values of the {\tt Set-} and {\tt Define-} functions point either to the assigned vector if successful, or they are {\tt NULL} if unsuccessful.}
@@
 In addition, one may want to use a different set of parameters for single- or two-parameter projections.
-The very flexible function \GLB{glbChiNP} is
+The very flexible function \GLB{ChiNP} is
 designed for this purpose. However, because of its flexibility, it 
@@
 In order to define arbitrary projections, we introduce the vector
-\GLB{glb\_projection}, which is very similar to the
+\GLB{\_projection}, which is very similar to the
 oscillation parameter vector {\tt glb\_params}.
@@
 comparison to {\tt glb\_params}, it does not take values for the
-parameters, but flags \GLB{GLB\_FIXED} or \GLB{GLB\_FREE}.
+parameters, but flags \GLBC{GLB\_FIXED} or \GLBC{GLB\_FREE}.
 For example, the projection onto the $\theta_{13}$-axis {\tt glbChiTheta}
@@
 
-\index{Projection: Define}
+\index{norm}{Projection: Define}
 As soon as we have defined a projection, we can assign it:
 \begin{function}
-\GLBNS{glbSetProjection}
+\GLBNS{SetProjection}
 {\tt int glbSetProjection(const glb\_projection in)} sets the projection
@@
 \begin{function}
-\GLBNS{glbGetProjection}
+\GLBNS{GetProjection}
 {\tt int glbGetProjection(glb\_projection out)} writes the currently
@@
 \begin{function}
-\GLBNS{glbChiNP}
+\GLBNS{ChiNP}
 {\tt double glbChiNP(const glb\_params in, glb\_params out, int exp)} 
@@
 \chapter{Locating degenerate solutions}
-\index{Degenerate solutions}
-\index{Degeneracies $\chi^2$}
+\index{norm}{Degenerate solutions}
+\index{norm}{Degeneracies $\chi^2$}
 
@@
 \label{ex:sgndeg}
-\index{Degeneracies: sgn($\ldm$)-degeneracy}
-
-In many cases, one can find the exact position of the $\mathrm{sgn}(\ldm)$-degeneracy with \GLB{glbChiAll}, where one starts 
+\index{norm}{Degeneracies: sgn($\ldm$)-degeneracy}
+
+In many cases, one can find the exact position of the $\mathrm{sgn}(\ldm)$-degeneracy with \GLB{ChiAll}, where one starts 
 the local minimizer at the suspected position 
@@
 
-\index{All-parameter minimization}
+\index{norm}{All-parameter minimization}
 The function to perform the all-parameter minimization is {\tt glbChiAll}:
 \begin{function}
-\GLBNS{glbChiAll} 
+\GLBNS{ChiAll} 
 {\tt double glbChiAll(const glb\_params in, glb\_params out, int exp)}
@@
 
-\index{Advanced tricks}
+\index{norm}{Advanced tricks}
 For the more advanced reader, a number of tricks can be useful for the numerical localization of degenerate solutions:
@@
 \chapter{Obtaining low-level information}
-\index{Low-level information}
+\index{norm}{Low-level information}
 
@@
 \section{Oscillation probabilities}
-\index{Oscillation probabilities}
+\index{norm}{Oscillation!probabilities}
 
@@
 \begin{function}
-\GLBNS{glbVacuumProbability}
+\GLBNS{VacuumProbability}
 {\tt double glbVacuumProbability(int l, int m, int panti,double E, double L)} returns the neutrino oscillation probability $\nu_l \rightarrow \nu_m$ for the energy {\tt E} and the baseline {\tt L} in vacuum.
@@
 \begin{function}
-\GLBNS{glbProfileProbability}
+\GLBNS{ProfileProbability}
 {\tt double glbProfileProbability(int l, int m, int panti,double E)}
@@
 THIS FUNCTION MAY OBTAIN AN EXPERIMENT INDEX (WW).
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Event rates}
-\index{Event rates}
+\index{norm}{Event rates}
+\label{sec:event_rates}
 One can also return event rates in \GLOBES , but this feature
@@
 
-\index{Event rates}
+\index{norm}{Event rates}
 \GLOBES\ currently supports rule-based and channel-based event rate functions, where the information is in written into a file
@@
 \begin{function}
-\GLBNS{glbShowRuleRates}
+\GLBNS{ShowRuleRates}
 {\tt int glbShowRuleRates(FILE* stream, int exp, int rule, int pos,
@@
 specific rule {\tt rule} have to be chosen, as well as the signal
-or background rate {\tt signal} (either \GLB{GLB\_SIG} or \GLB{GLB\_BG}).
+or background rate {\tt signal} (either \GLBC{GLB\_SIG} or \GLBC{GLB\_BG}).
 The position {\tt pos} refers to the component within the signal or 
 background, and can also be {\tt GLB\_ALL}. The function may return
-the rates with (\GLB{GLB\_W\_COEFF}) or without (\GLB{GLB\_WO\_COEFF})
+the rates with (\GLBC{GLB\_W\_COEFF}) or without (\GLBC{GLB\_WO\_COEFF})
 overall efficiency coefficient, as it is specified by {\tt coeffi}. 
 In addition, it may contain the post-smearing efficiencies (set
-{\tt effi} to \GLB{GLB\_W\_EFF} or \GLB{GLB\_WO\_EFF}), and the
+{\tt effi} to \GLBC{GLB\_W\_EFF} or \GLBC{GLB\_WO\_EFF}), and the
 post-smearing backgrounds (set
-{\tt bgi} to \GLB{GLB\_W\_BG} or \GLB{GLB\_WO\_BG}). The pre-smearing
+{\tt bgi} to \GLBC{GLB\_W\_BG} or \GLBC{GLB\_WO\_BG}). The pre-smearing
 efficiencies and backgrounds cannot be accessed at the rule level.
@@
 \begin{function}
-\GLBNS{glbTotalRuleRate}
+\GLBNS{TotalRuleRate}
 {\tt double glbTotalRuleRate(int exp, int rule, int pos,
@@
 \begin{function}
-\GLBNS{glbGetNumberOfRules}
+\GLBNS{GetNumberOfRules}
 {\tt int glbGetNumberOfRules(int exp)} returns the number of
@@
 \begin{function}
-\GLBNS{glbGetLengthOfRule}
+\GLBNS{GetLengthOfRule}
 {\tt int glbGetLengthOfRule(int exp, int rule, int signal)} returns
 the length of rule {\tt rule} in experiment {\tt exp}. The parameter
-{\tt signal} can be either \GLB{GLB\_SIG} for the number of signal
-components or \GLB{GLB\_BG} for the number of background components.
-\end{function}
-\begin{function}
-\GLBNS{glbGetNormalizationInRule}
+{\tt signal} can be either \GLBC{GLB\_SIG} for the number of signal
+components or \GLBC{GLB\_BG} for the number of background components.
+\end{function}
+\begin{function}
+\GLBNS{GetNormalizationInRule}
 {\tt double glbGetNormalizationInRule(int exp, int rule, int signal)}
@@
 in rule {\tt rule} of the experiment {\tt exp}.
-The parameter {\tt signal} refers to signal (\GLB{GLB\_SIG}) or background
-(\GLB{GLB\_BG}).
-\end{function}
-\begin{function}
-\GLBNS{glbGetChannelInRule}
+The parameter {\tt signal} refers to signal (\GLBC{GLB\_SIG}) or background
+(\GLBC{GLB\_BG}).
+\end{function}
+\begin{function}
+\GLBNS{GetChannelInRule}
 {\tt int glbGetChannelInRule(int exp, int rule, int pos, int signal)}
@@
 of the experiment {\tt exp}.
-The parameter {\tt signal} refers to signal (\GLB{GLB\_SIG}) or background
-(\GLB{GLB\_BG}).
-\end{function}
-\begin{function}
-\GLBNS{glbGetCoefficientInRule}
+The parameter {\tt signal} refers to signal (\GLBC{GLB\_SIG}) or background
+(\GLBC{GLB\_BG}).
+\end{function}
+\begin{function}
+\GLBNS{GetCoefficientInRule}
 {\tt double glbGetCoefficientInRule(int exp, int rule, int pos, int signal)}
@@
 of the experiment {\tt exp}.
-The parameter {\tt signal} refers to signal (\GLB{GLB\_SIG}) or background
-(\GLB{GLB\_BG}).
+The parameter {\tt signal} refers to signal (\GLBC{GLB\_SIG}) or background
+(\GLBC{GLB\_BG}).
 \end{function}
@@
 \begin{function}
-\GLBNS{glbShowChannelRates}
+\GLBNS{ShowChannelRates}
 {\tt int glbShowChannelRates(FILE *stream,
@@
 The function may return
-the rates before (\GLB{GLB\_PRE}) or after (\GLB{GLB\_POST})
+the rates before (\GLBC{GLB\_PRE}) or after (\GLBC{GLB\_POST})
 the energy smearing, as it is specified by {\tt smearing}.
 In addition, it may contain the pre- and post-smearing efficiencies (set
-{\tt effi} to \GLB{GLB\_W\_EFF} or \GLB{GLB\_WO\_EFF}), and the
+{\tt effi} to \GLBC{GLB\_W\_EFF} or \GLBC{GLB\_WO\_EFF}), and the
 pre- and post-smearing backgrounds (set
-{\tt bgi} to \GLB{GLB\_W\_BG} or \GLB{GLB\_WO\_BG}). Note that
+{\tt bgi} to \GLBC{GLB\_W\_BG} or \GLBC{GLB\_WO\_BG}). Note that
 the post-smearing efficiencies and backgrounds cannot be taken into
@@
 \begin{function}
-\GLBNS{glbGetChannelRates}
+\GLBNS{GetChannelRates}
 {\tt int glbGetChannelRates(double** data,
@@
 The function may return
-the rates before ({\tt smearing} is \GLB{GLB\_PRE}) or after ({\tt smearing} is \GLB{GLB\_POST})
+the rates before ({\tt smearing} is \GLBC{GLB\_PRE}) or after ({\tt smearing} is \GLBC{GLB\_POST})
 the energy smearing, where no user-defined data
@@
 \begin{function}
-\GLBNS{glbGetUserData}
+\GLBNS{GetUserData}
 {\tt int glbGetUserData(double** data,
@@
 The function may return
-the pre- ({\tt smearing} is \GLB{GLB\_PRE}) or post- ({\tt smearing} is \GLB{GLB\_POST}) smearing backgrounds ({\tt bgeff} is \GLB{GLB\_BG}) 
-or efficiencies ({\tt bgeff} is \GLB{GLB\_EFF}). 
+the pre- ({\tt smearing} is \GLBC{GLB\_PRE}) or post- ({\tt smearing} is \GLBC{GLB\_POST}) smearing backgrounds ({\tt bgeff} is \GLBC{GLB\_BG}) 
+or efficiencies ({\tt bgeff} is \GLBC{GLB\_EFF}). 
 The return value is $-1$ if unsuccessful.
@@
 \begin{function}
-\GLBNS{glbResetRateStack}
+\GLBNS{ResetRateStack}
 {\tt void glbResetRateStack()} resets the rate stack used for the
@@
 \begin{function}
-\GLBNS{glbGetNumberOfChannels}
+\GLBNS{GetNumberOfChannels}
 {\tt int glbGetNumberOfChannels(int exp)} returns the number of 
@@
 \section{Fluxes and cross sections}
-\index{Flux} \index{Cross section}
+\index{norm}{Flux} \index{norm}{Cross section}
 
@@
 \begin{function}
-\GLBNS{glbFlux}
+\GLBNS{Flux}
 {\tt double glbFlux(int exp, int ident, 
@@
 \begin{function}
-\GLBNS{glbXSection}
+\GLBNS{XSection}
 {\tt double glbXSection(int exp, 
@@
 \label{chapt:running}
-\index{Experiment parameters}
+\index{norm}{Experiment parameters}
 
@@
 \section{Baseline and matter density profile}
-\index{Baseline: Change}
-\index{Matter density profile: Change}
+\index{norm}{Baseline: Change}
+\index{norm}{Matter density profile: Change}
 
@@
 \begin{function}
-\GLBNS{glbGetProfileType}
+\GLBNS{GetProfileType}
 {\tt int glbGetProfileType(int exp)} returns the matter density profile
@@
 \begin{function}
-\GLBNS{glbSetBaselineInExperiment}
+\GLBNS{SetBaselineInExperiment}
 {\tt int glbSetBaselineInExperiment(int exp, double baseline)}
@@
 \begin{function}
-\GLBNS{glbGetBaselineInExperiment} 
+\GLBNS{GetBaselineInExperiment} 
 {\tt double glbGetBaselineInExperiment(int exp)} returns the
@@
 \begin{function}
-\GLBNS{glbLoadProfileData}
+\GLBNS{LoadProfileData}
 {\tt int glbLoadProfileData(const char* filename, size\_t *layers, double **lengths, double **densities)} loads a density file from the file
@@
 \begin{function}
-\GLBNS{glbStaceyProfile}
+\GLBNS{StaceyProfile}
 {\tt int glbStaceyProfile(double baseline, size\_t layers, double **lengths, double **densities)} creates a PREM/Stacey matter density profile with a
@@
 \begin{function}
-\GLBNS{glbAverageDensityProfile}
+\GLBNS{AverageDensityProfile}
 {\tt glbAverageDensityProfile(double baseline, double **lengths, 
@@
 \begin{function}
-\GLBNS{glbStaceyProfile}
+\GLBNS{StaceyProfile}
 {\tt int glbGetProfileDataInExperiment(int exp,size\_t *layers, double** lengths, double** densities)} returns the matter density profile 
@@
 \begin{function}
-\GLBNS{glbSetProfileDataInExperiment}
+\GLBNS{SetProfileDataInExperiment}
 {\tt int glbSetProfileDataInExperiment(int exp, size\_t layers,const double* lengths, const double* densities)} sets the matter density of experiment
@@
 \label{ex:barcharts}
-\index{Bar charts}
-\index{Systematics on/off}
+\index{norm}{Bar charts}
+\index{norm}{Systematics on/off}
 Here it is demonstrated how one can successively include systematics, correlations, and degeneracies at the example of the $\stheta$-sensitivity limit.
@@
 \section{Systematics}
-\index{Systematics}
-\index{Systematics on/off}
+\label{sec:systematics}
+\index{norm}{Systematics}
+\index{norm}{Systematics on/off}
 
@@
 \begin{function}
-\GLBNS{glbSwitchSystematics}
+\GLBNS{SwitchSystematics}
 {\tt int glbSwitchSystematics(int exp, int rule, int on\_off)}
 switches the systematics in experiment {\tt exp} and rule {\tt rule}
-on ({\tt on\_off} is \GLB{GLB\_ON}) or off ({\tt on\_off} is \GLB{GLB\_OFF}). For the experiment or
-rule index, one can also use \GLB{GLB\_ALL}. 
+on ({\tt on\_off} is \GLBC{GLB\_ON}) or off ({\tt on\_off} is \GLBC{GLB\_OFF}). For the experiment or
+rule index, one can also use \GLBC{GLB\_ALL}. 
 \end{function}
@@
 
-\index{Error dimension}
+\index{norm}{Error dimension}
 The error dimension (for the definition, see \Sec~\ref{sec:rules}) can also be accessed directly with 
 \begin{function}
-\GLBNS{glbSetErrorDim}
+\GLBNS{SetErrorDim}
 {\tt int glbSetErrorDim(int exp, int rule, int on\_off, int value)}
-sets the error dimension for systematics on ({\tt on\_off} is \GLB{GLB\_ON}) or off ({\tt on\_off} is \GLB{GLB\_OFF}) of experiment {\tt exp} and rule {\tt rule} to the value {\tt value}.
+sets the error dimension for systematics on ({\tt on\_off} is \GLBC{GLB\_ON}) or off ({\tt on\_off} is \GLBC{GLB\_OFF}) of experiment {\tt exp} and rule {\tt rule} to the value {\tt value}.
 The function returns
@@
 \begin{function}
-\GLBNS{glbGetErrorDim}
+\GLBNS{GetErrorDim}
 {\tt int glbGetErrorDim(int exp, int rule, int on\_off)}
-returns the error dimension for systematics on ({\tt on\_off} is \GLB{GLB\_ON}) or off ({\tt on\_off} is \GLB{GLB\_OFF}) of experiment {\tt exp} and rule {\tt rule}.
-\end{function}
-
-\index{Signal errors}
-\index{Background errors}
-\index{Background centers}
+returns the error dimension for systematics on ({\tt on\_off} is \GLBC{GLB\_ON}) or off ({\tt on\_off} is \GLBC{GLB\_OFF}) of experiment {\tt exp} and rule {\tt rule}.
+\end{function}
+
+\index{norm}{Signal errors}
+\index{norm}{Background errors}
+\index{norm}{Background centers}
 Except from the general treatment of systematics, one can read out
@@
 \begin{function}
-\GLBNS{glbSetSignalErrors}
+\GLBNS{SetSignalErrors}
 {\tt int glbSetSignalErrors(int exp, int rule, double norm, double tilt)}
@@
 \begin{function}
-\GLBNS{glbGetSignalErrors}
+\GLBNS{GetSignalErrors}
 {\tt int glbGetSignalErrors(int exp, int rule, double* norm, double* tilt)}
@@
 \begin{function}
-\GLBNS{glbSetBGErrors}
+\GLBNS{SetBGErrors}
 {\tt int glbSetBGErrors(int exp, int rule, double norm, double tilt)}
@@
 \begin{function}
-\GLBNS{glbGetBGErrors}
+\GLBNS{GetBGErrors}
 {\tt int glbGetBGErrors(int exp, int rule, double* norm, double* tilt)}
@@
 \begin{function}
-\GLBNS{glbSetBGCenters}
+\GLBNS{SetBGCenters}
 {\tt int glbSetBGCenters(int exp, int rule, double norm, double tilt)}
@@
 \begin{function}
-\GLBNS{glbGetBGCenters}
+\GLBNS{GetBGCenters}
 {\tt int glbGetBGCenters(int exp, int rule, double* norm, double* tilt)}
@@
 \label{sec:aedlparams}
-\index{External \AEDL\ parameters}
+\index{norm}{External \AEDL\ parameters}
 
@@
 In order to use the user-defined variable, one has to assign it 
-with {\tt glbDefineAEDLVariable} {\em before} the experiment is initialized with \GLB{glbInitExperiment}:
-\begin{function}
-\GLBNS{glbDefineAEDLVariable}
+with {\tt glbDefineAEDLVariable} {\em before} the experiment is initialized with \GLB{InitExperiment}:
+\begin{function}
+\GLBNS{DefineAEDLVariable}
 {\tt void glbDefineAEDLVariable(const char* name, double value)}
@@
 \begin{function}
-\GLBNS{glbClearAEDLVariables}
+\GLBNS{ClearAEDLVariables}
 {\tt void glbClearAEDLVariables()}
@@
 \section{Algorithm parameters: Filter functions}
-\index{Filter functions}
+\index{norm}{Filter functions}
 
@@
 \begin{function}
-\GLBNS{glbSetFilterState}
+\GLBNS{SetFilterState}
 {\tt int glbSetFilterState(int on\_off)} sets the filter state to
-on (\GLB{GLB\_ON}) or off (\GLB{GLB\_OFF}).
-\end{function}
-\begin{function}
-\GLBNS{glbGetFilterState}
+on (\GLBC{GLB\_ON}) or off (\GLBC{GLB\_OFF}).
+\end{function}
+\begin{function}
+\GLBNS{GetFilterState}
 {\tt int glbGetFilterState()} returns the filter state.
@@
 \begin{function}
-\GLBNS{glbSetFilterStateInExperiment}
-{\tt int glbSetFilterStateInExperiment(int exp, int on\_off)} sets the filter state in experiment {\tt exp} to on (\GLB{GLB\_ON}) or off (\GLB{GLB\_OFF}).
-\end{function}
-\begin{function}
-\GLBNS{glbGetFilterStateInExperiment}
+\GLBNS{SetFilterStateInExperiment}
+{\tt int glbSetFilterStateInExperiment(int exp, int on\_off)} sets the filter state in experiment {\tt exp} to on (\GLBC{GLB\_ON}) or off (\GLBC{GLB\_OFF}).
+\end{function}
+\begin{function}
+\GLBNS{GetFilterStateInExperiment}
 {\tt int glbGetFilterStateInExperiment(int exp)} returns the filter state of experiment {\tt exp}.
@@
 \begin{function}
-\GLBNS{glbSetFilter}
+\GLBNS{SetFilter}
 {\tt int glbSetFilter(double filter)} sets the filter to the
@@
 \begin{function}
-\GLBNS{glbGetFilter}
+\GLBNS{GetFilter}
 {\tt double glbGetFilter()} returns the filter value.
@@
 \begin{function}
-\GLBNS{glbSetFilterInExperiment}
+\GLBNS{SetFilterInExperiment}
 {\tt int glbSetFilterInExperiment(int exp, double filter)} sets the filter  in experiment {\tt exp} to the value {\tt value}.
@@
 \begin{function}
-\GLBNS{glbGetFilterInExperiment}
+\GLBNS{GetFilterInExperiment}
 {\tt double glbGetFilterInExperiment(int exp)} returns the filter value of experiment {\tt exp}.
@@
 The return value of all {\tt Set-} functions is $-1$ if they were not successful.
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: Manual.tex
+%%% End:

### 2004-07-13 Documentation/part2.tex
added section about globes put in the fdl part on page two, made the sections of fdl disappear from toc added multiple index support changed the \GLB and \GLBNS commands such, that they prepend 'glb' in the text and add their argument to the api index. makeindex has not be called four times: 	makeindex norm.idx 	makeindex api.idx 	makeindex aedl.idx 	makeindex constants.idx
Adding entries now works with \index{index_name}{index_entry}, e.g.
\index{norm}{Blau} adds 'Blau' to the usual index.

@@
 sources of systematical errors, we use several abstraction levels. 
-The first level is the so-called ``channel'', which is the link between 
+The first level is the so-called ``channel'', \index{norm}{Channel}
+ which is the link between 
 the oscillation physics and the detection properties for a specfific oscillation pattern (\cf, \figu{channel}).
@@
 
-The channels are the building blocks for the so-called ``rules''. In general, a rule consists of one or more ``signal'' and ``background'' oscillation channels, which are normalized with efficiencies
+The channels are the building blocks for the so-called ``rules''.
+\index{norm}{Rule} In general, a rule consists of one or more ``signal'' and ``background'' oscillation channels, which are normalized with efficiencies
 (\cf, \figu{rule}). The event numbers from these channels are added {\em before} the $\Delta \chi^2$-value is calculated.\footnote{Note that
@@
 
+\index{aedl}{NEXT@{\tt \#NEXT\#}}
 If you want to have several experiments in one file, separate the different
@@
 
+\index{aedl}{inlcude@{\tt include}}
 As another feature of AEDL one can use include files with the {\tt include} command. Includes can be nested up to {\tt MAX\_INCLUSION\_DEPTH}, which is currently set to $10$. Error reporting works 
@@
 in order to initialize a combined analysis of the experiments defined in the files {\tt NuFact.gls} and {\tt JHFHK.gls}. As one sees from the example above
-file names are always quoted in \AEDL.\index{File names}
+file names are always quoted in \AEDL.\index{norm}{File names}
 Even if one uses the
@@
 substitution mechanism as described in~\Sec~\ref{sec:aedlparams}.
-\index{External \AEDL\ parameters}
+\index{norm}{External \AEDL\ parameters}
 
@@
 {\tt log}, {\tt log10}, {\tt exp}, {\tt sqrt}.
+\index{aedl}{sin@{\tt sin}}
+\index{aedl}{cos@{\tt cos}}
+\index{aedl}{tan@{\tt tan}}
+\index{aedl}{asin@{\tt asin}}
+\index{aedl}{acos@{\tt acos}}
+\index{aedl}{atan@{\tt atan}}
+\index{aedl}{log@{\tt log}}
+\index{aedl}{log10@{\tt log10}}
+\index{aedl}{exp@{\tt exp}}
+\index{aedl}{sqrt@{\tt sqrt}}
+
 These functions can be used everywhere, where
@@
 
+\index{aedl}{target mass@{\tt \$target\_mass}}
 The quantity which can be used to scale the overall integrated luminosity of an experiment, is the fiducial detector mass. For example,
@@
 
+\index{aedl}{{\tt flux}!time@{\tt "@time}}
 There are two principal ways to initialize a neutrino flux: 
@@
 a flux is defined by the environment {\tt flux}, such as
+\index{aedl}{{\tt flux}}
 \begin{quote}
@@
 Thus, an example to set up a neutrino factory flux is
+\index{aedl}{{\tt flux}!builtin@{\tt "@builtin}}
+\index{aedl}{{\tt flux}!parent energy@{\tt "@parent\_energy}}
+\index{aedl}{{\tt flux}!stored muons@{\tt "@stored\_muons}}
 \begin{quote}
@@
 %
+\index{aedl}{{\tt flux}!power@{\tt "@power}}
+\index{aedl}{{\tt flux}!norm@{\tt "@norm}}
+\index{aedl}{{\tt flux}!flux file@{\tt "@flux\_file}}
 For a user-defined flux, one has to give it the file name:
@@
 
+\index{aedl}{{\tt flux}!flux file@{\tt "@flux\_file}}
 The software assumes that the given flux file has seven columns and
@@
 
-The baseline\index{Baseline} is given by
+\index{aedl}{baseline@{\tt \$baseline}}
+The baseline\index{norm}{Baseline} is given by
 \begin{quote}
@@
 
+\index{aedl}{profiletype@{\tt \$profiletype}}
+\index{aedl}{densitysteps@{\tt \$densitysteps}}
+\index{aedl}{lengthtab@{\tt \$lengthtab}}
+\index{aedl}{densitytab@{\tt \$densitytab}}
+
 Furthermore, the matter density profile along the baseline
 has to be specified.
-The simplest matter density profile is a constant matter density profile with the average matter density from the PREM\cite{Stacey} onion shell model of the Earth\index{PREM}\index{Earth matter density}:
+The simplest matter density profile is a constant matter density profile with the average matter density from the PREM\cite{Stacey} onion shell model of the Earth\index{norm}{PREM}\index{norm}{Earth matter density}:
 \begin{quote}
@@
 
-Cross sections\index{Cross sections} will later be used as part of the 
+Cross sections\index{norm}{Cross sections} will later be used as part of the 
 channel definition (see \Sec~\ref{sec:channel}). Similar to the source 
 fluxes, they are provided by the user as a data file:
+\index{aedl}{{\tt cross}}
+\index{aedl}{{\tt cross}!cross file@{\tt  "@cross\_file}}
 \begin{quote}
@@
 seven columns and $1001$ lines of the form
+\index{aedl}{{\tt cross}!cross file@{\tt  "@cross\_file}}
 \begin{quotation}
@@
 
-Channels\index{Channel} in \GLOBES\ represent an intermediate level 
+\index{aedl}{{\tt channel}|(}
+Channels\index{norm}{Channel} in \GLOBES\ represent an intermediate level 
 between the pure oscillation physics given by the oscillation probability
@@
 $R_\beta^{\text{IT}}(E,E') dE'$. The function $R(E,E')$ is also often called ``energy resolution function''. Actually, its internal representation
-in the software is a smearing matrix\index{Energy resolution}\index{Smear matrix}.
+in the software is a smearing matrix\index{norm}{Energy resolution}\index{norm}{Smear matrix}.
 The function $\epsilon_\beta^{\text{IT}}(E')$ will later be refered to as ``post-smearing efficiencies'', since it will allow us to define cuts and threshold functions {\em after} the smearing is performed, \ie, as function of $E'$.
@@
 
-Eventually, we can write down the number of events per bin $i$ \index{Bin} and channel $c$ as
+Eventually, we can write down the number of events per bin $i$ \index{norm}{Bin} and channel $c$ as
 \begin{equation}
@@
 defined before with their {\tt \#name} in the respective environments. 
-Thus, a simple definition of a channel is
+Thus, a simple definition of a channel\index{aedl}{{\tt channel}} is
 \begin{quote}
@@
 
-\index{Oscillation!switching off}
+\index{norm}{Oscillation!switching off}
 It is also possible to define a channel as no-oscillation by using the
-prefix \GLB{NOSC\_} in either the initial  flavour or 
+prefix {\tt NOSC\_}\index{aedl}{{\tt channel}!{\tt NOSC\_}} 
+in either the initial  flavour or 
 the final flavour, like this
@@
 parameters\footnote{at least in the absence of sterile neutrinos}. The channels
-marked as \GLB{NOSC\_} are only computed by a call to \GLB{glbSetRates}
-and not recomputed in the subsequent fit (which calls \GLB{glbSetNewRates}).
+marked as {\tt NOSC\_} are only computed by a call to \GLB{SetRates}
+and not recomputed in the subsequent fit (which calls \GLB{SetNewRates}).
 Therefore this feature can be used to speed up the rate 
@@
 energy resolution function in the next section.
+\index{aedl}{{\tt channel}|)}
 
@@
 
+\index{aedl}{{\tt energy}|(}
 The definition and implementation of the energy resolution function is 
@@
 For a manual definition of the sampling points (depends on algorithm), use
+\index{aedl}{sampling points@{\tt \$sampling\_points}}
+\index{aedl}{sampling min@{\tt \$sampling\_min}}
+\index{aedl}{sampling max@{\tt \$sampling\_max}}
 \begin{quote}
@@
 for equidistant sampling points. Arbitrarily spaced sampling points can 
-be specified with \GLB{\$sampling\_stepsize}
+be specified with 
+{\tt \$sampling\_stepsize}\index{aedl}{sampling stepsize@{\tt \$sampling\_stepsize}}
 \begin{quote}
@@
 use
+\index{aedl}{emax@{\tt \$emax}}
+\index{aedl}{emin@{\tt \$emin}}
+\index{aedl}{bins@{\tt \$bins}}
 \begin{quote}
@@
 \end{quote}
+\index{aedl}{binsize@{\tt \$binsize}}
 The number of bins will be automatically computed by \GLOBES . Note that the
@@
 These components can be introduced before or after the integration in \eq~(\ref{eq:simple_int}) is done.
+%
+\index{aedl}{{\tt channel}!pre smearing efficiencies@{\tt 
+"@pre\_smearing\_efficiencies}}
+\index{aedl}{{\tt channel}!pre smearing background@{\tt 
+"@pre\_smearing\_background}}
+\index{aedl}{{\tt channel}!post smearing efficiencies@{\tt 
+"@post\_smearing\_efficiencies}}
+\index{aedl}{{\tt channel}!post smearing background@{\tt 
+"@post\_smearing\_background}}
+%
 If they are introduced before, 
@@
 \end{quote}
+%
+\index{aedl}{{\tt channel}!pre smearing efficiencies@{\tt 
+"@pre\_smearing\_efficiencies}}
+\index{aedl}{{\tt channel}!pre smearing background@{\tt 
+"@pre\_smearing\_background}}
+\index{aedl}{{\tt channel}!post smearing efficiencies@{\tt 
+"@post\_smearing\_efficiencies}}
+\index{aedl}{{\tt channel}!post smearing background@{\tt 
+"@post\_smearing\_background}}
+%
 This experiment uses $10$ sampling points and $5$ bins.
@@
 This algorithm is the fastest which \GLOBES\ offers, and is selected by
+%
+\index{aedl}{{\tt energy}!type@{\tt "@type}}
+%
 \begin{quote}
@@
 
+%
+\index{norm}{Energy!resolution}
+\index{norm}{Energy!resolution function}
+%
 As far as the parameterization for the energy resolution function 
@@
 \end{equation} 
+%
+\index{aedl}{{\tt energy}!sigma function@{\tt "@sigma\_function}}
+%
 There are several energy resolution functions available, where by default
@@
 \end{quote}
+%
+\index{aedl}{{\tt energy}!standard@{\tt \#standard}}
+%
 The energy resolution function {\tt \#standard} is defined by
@@
 Currently, another possible choice for {\tt @sigma\_function} is {\tt \#inverse\_beta},
+%
+\index{aedl}{{\tt energy}!inverse beta@{\tt \#inverse\_beta}}
+%
 which only uses the parameter $\alpha$. It is defined by
@@
 In order to ensure that fast oscillating probabilities do not lead to 
-aliasing\index{Aliasing}
+aliasing\index{norm}{Aliasing}
 it is desirable to impose a low-pass filter already during the calculation
 of the probabilities itself. This is possible and implemented has a highly
-experimental feature called ``filter''\index{Filter}. 
+experimental feature called ``filter''\index{norm}{Filter}. 
 The calculation of oscillation
@@
 variable:
+\index{aedl}{filter state@{\tt \$filter\_state}}
+\index{aedl}{filter value@{\tt \$filter\_value}}
 \begin{quote}
@@
 
+\index{aedl}{{\tt energy}!type@{\tt  "@type}}
 With a type 1 ({\tt @type = 1}) energy resolution 
@@
 where the unshown entries are zero. Thus, the values of $K_{ij}$ have to be specified between $k_l^i$ and $k_u^i$ in the form $\{ k_l^i,k_u^i, K_{i \, k_l^i}, K_{i k_l^i+1} , \hdots , K_{i k_u^i} \}$:
+\index{aedl}{{\tt energy}!energy@{\tt "@energy}}
 \begin{quote}
@@
 
+\index{aedl}{{\tt energy}|)}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-The set of rules\index{Rule} for an experiment is the final 
+\index{aedl}{{\tt rule}|(}
+The set of rules\index{norm}{Rule} for an experiment is the final 
 link between the event rate computation and the statistical analysis. 
@@
 of mis-identified events (charge or flavor mis-identification).
+%
+\index{aedl}{{\tt rule}!signal@{\tt "@signal}}
+\index{aedl}{{\tt rule}!background@{\tt "@background}}
 %
@@
 Thus, it is possible to constrain the analysis energy range 
-with each rule to an energy window\index{Energy window}:
+with each rule to an energy window\index{norm}{Energy window}:
+\index{aedl}{{\tt rule}!energy window@{\tt "@energy\_window}}
 \begin{quote}
@@
 parameters which are minimized over. This choice is specified with the 
-error dimension variable\index{Error dimension}, and the different 
+error dimension variable\index{norm}{Error dimension}, and the different 
 possibilities are shown in
@@
 The error dimensions for the case of no systematics is set for each rule 
-by the value of \GLB{@errordim\_sys\_off}, whereas the error dimension
-for systematics on is given by \GLB{@errordim\_sys\_on}. Thus, for example,
+by the value of {\tt @errordim\_sys\_off}\index{aedl}{{\tt rule}!{\tt "@errordim\_sys\_off}}, whereas the error dimension
+for systematics on is given by {\tt @errordim\_sys\_on}
+\index{aedl}{{\tt rule}!{\tt "@errordim\_sys\_on}}. Thus, for example,
 the complete error dimension definition could look like  
@@
 \end{quote}
-
+\index{aedl}{{\tt rule}|)}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{sec:aedl_versioning}
-\index{Version control}
+\index{norm}{Version control}
 
@@
 \AEDL\ file with the corresponding version number of the \GLOBES\ package with
-\GLB{\$version} by
+{\tt \$version}\index{aedl}{version@{\tt \$version}} by
 \begin{quote}
@@
 \end{quote}
-This information can be accesed by the versioning functions as described in~\Sec~\ref{sec:versioning}.
+This information can be accessed by the versioning functions as described in~\Sec~\ref{sec:versioning}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\chapter{Testing \& debugging of \AEDL\ files}
+\label{chap:exp_def}
+
+\AEDL\ is a powerful language to describe a vast variety 
+of different experiments.
+This chapter demonstrates how to test whether a given \AEDL\ file describes
+a given experiment. To this end the \GLOBES\ package also features a program
+called {\tt globes}\index{norm}{{\tt globes}}, which can be seen either as a 
+crude \AEDL\ debugger 
+or a simple command line oriented tool to convert the rather 
+abstract \AEDL\ experiment description into more accessible event rates.
+
+\section{Basic usage of {\tt globes}}
+\label{sec:globes_basics}
+
+The {\tt globes}\index{norm}{{\tt globes}} binary is installed along 
+with the library \index{norm}{{\tt libglobes}}, but
+into the directory {\tt \$prefix/bin/}. In order to use the {\tt globes} 
+utility this directory\footnote{This is automatically the case, when
+no options have been given to {\tt configure}, and {\tt make install} was 
+executed with root-privilege, \ie\ a standard installation was done.} has to 
+be in the path of the shell used to invoke it.
+
+First {\tt globes} parses the {\tt .glb}-file which was given as argument,
+prints any warnings and errors which occured during parsing. Then it
+uses the experiment description in the file to compute the event rates
+at a certain point in parameters space. And finally it displays the result
+with a varying degree of detail dependent on the options used to invoke
+{\tt globes}. The option parsing follows the GNU standards and thus there
+is a {\tt \verb+--+help} option which displays all other options 
+along with a short description.
+
+\index{norm}{{\tt globes}!oscillation parameters}
+Calling {\tt globes} without any options and with a {\tt .glb}-file as argument produces an event summary at rule level, taking into account 
+the full description in the file, \ie\ all efficiencies, backgrounds and resolution effects. Thus, these are the event rates which are actually 
+used to compute the $\chi^2$. By default the oscillation parameters used to
+calculate the transition probability are
+\begin{eqnarray}
+\label{eq:globes_params}
+\sin^22\theta_{12}=0.8&\quad&\Delta m^2_{21}=7\cdot10^{-5}
+\,\mathrm{eV}^2\,,\nonumber\\
+\sin^22\theta_{23}=1.0&\quad&\Delta m^2_{31}=3\cdot10^{-3}
+\,\mathrm{eV}^2\,,\nonumber\\
+\delta=0&\quad&\sin^22\theta_{13}=0.1\,.\
+\end{eqnarray}
+Of course, it is possible to change these default values either by using the
+option {\tt -p} on a call by call basis or by setting the environment variable
+{\tt GLB\_CENTRAL\_VALUES}\index{norm}{{\tt  GLB\_CENTRAL\_VALUES}}.
+\begin{quote}
+{\tt
+globes -p'0.55,0,0.785,0,0.0008,0.0025'\\
+globes \verb+--+parameters='0.55,0,0.785,0,0.0008,0.0025'\\
+}
+\end{quote}
+or by defining within the shell session or in the shell profile:
+\begin{quote}
+{\tt
+export GLB\_CENTRAL\_VALUES='0.55,0,0.785,0,0.0008,0.0025'
+}
+\end{quote}
+Furthermore it is possible to switch off oscillations altogether with t
+he {\tt -N} option
+and to switch them on again with {\tt -O} (the default). The effect of {\tt -N}
+is the same than to prefix all channels with {\tt NOSC\_}.
+
+\index{norm}{{\tt globes}!warnings}\index{norm}{{\tt globes}!errors}
+\index{norm}{{\tt globes}!verbosity}
+The \AEDL\ parser and interpreter has basically three levels of messages to
+th user: warnings, errors and fatal errors. Fatal errors always are reported
+and the program exits with status '1'. Usually only errors and no warnings
+are reported. The verbosity level can be chosen by the {\tt -v} option, where
+{\tt -v1} is default, \ie\ only errors and fatal errors are reported. {-v0}
+corresponds to reporting fatal errors only, whereas {\tt -v2} will
+print in addition to fatal errors and errors also warnings. It is recommended
+to test any new {\tt .glb}-file with {\tt -v2} to see the warnings at least 
+once and to decide whether there is a problem and in case fix it. With {\tt -v3}
+all files read by {\tt globes} are displayed together with their path and
+with  {\tt -v4} all files which have been attempted to be read are shown.
+These two setting are useful to clarify path resolution issues and shadowing
+of file names.
+
+\section{Testing \AEDL\ files}
+\label{sec:globes_test}
+
+In the process of defining a new experiment the default output of {\tt globes}
+at rule level is the final step but in order to arrive there it is often
+necessary to view the intermediate steps in the event rate calculation. The 
+{\tt globes} utility offers many possibilities to do so based on the 
+rate access functions as described in \Sec~\ref{sec:event_rates}.
+
+\index{norm}{{\tt globes}!total rates}
+\index{norm}{{\tt globes}!spectral rates}
+By default {\tt globes} returns total rates corresponding to the 
+{\tt -t} option, this can be changed to
+to a full spectrum by using {\tt -s}. The spectral rates are shown in a
+table where the first columns always gives the central energy of the
+corresponding bin.
+
+If there are more than one experiment in a file, \ie\ at least one 
+{\tt \#NEXT\#}
+command, then only the event rates for one experiment are shown.  This
+one experiment can be choose with the {\tt -e} option which takes 
+as mandatory argument the number of the experiment (starting with zero), 
+where {\tt -e0} is the default.
+
+\subsection*{Channel level}
+\index{norm}{{\tt globes}!channel rates}
+
+As a first step one often wants to check whether each channel produces the 
+desired output. Channel rates are returned when the {\tt -c} option is used.
+This option takes as an optional argument the channel number 
+(starting at zero), if no argument is given all channels are displayed.
+By default the sum of the event rates in each channel is shown. Where each 
+column has as first line the same channel name as in the file.
+
+It is also possible to take out one detector effect after the other. First
+one can switch off the post-smearing efficiencies ({\tt -f}) and the 
+post-smearing backgrounds ({\tt -g}). Next, one can switch off the 
+energy resolution function with ({\tt -b}) and view the rates before
+smearing. In case the {\tt -s} options is also used, the number of
+lines in the output now is given by the setting of {\tt \$sampling\_points}.
+Another effect of the {\tt -b} option is that the post-smearing efficiencies
+and backgrounds are no longer taken into account. Therefore the {\tt -g} 
+and {\tt -f} options now apply to the \emph{pre}-smearing efficiencies 
+and the \emph{pre}-smearing backgrounds. Thus 
+\begin{quote}
+{\tt
+globes -c -b -g -f FILE
+}
+\end{quote}
+produces the raw event rate corresponding to the convolution of flux, 
+probability and cross section, neglecting all detector effects.
+
+\subsection*{Rule level}
+\index{norm}{{\tt globes}!rule rates}
+
+The logical step after channels is to check the rule rates. Rule rates 
+are returned with {\tt -r}.
+This option takes as an optional argument the rule number 
+(starting at zero), if no argument is given all rules are displayed.
+By default the sum of the event rates in each rule is shown, as well
+as for each part of the rule. Where each 
+rule is preceeded by a line with the same rule name as in the file.
+
+It is also possible for rules to take out one detector effect after 
+the other, however with the limitation that rules make only sense
+after the energy resolution function has been applied to each channel.
+Therefore it is \emph{not} possible to use {\tt -b} together with {\tt -r} or
+to switch off any pre-smearing efficiencies or backgrounds. 
+One can, however, switch off the post-smearing efficiencies ({\tt -f}) 
+and the post-smearing backgrounds ({\tt -g}) for each channel. Since
+the definition of a rule also contains so called coefficients it is
+possible to switch them off with {\tt -i}. This options also deactivates
+any setting of {\tt @backgroundcenter}.
+
+\subsection*{Output}
+\index{norm}{{\tt globes}!output}
+
+The default output stream is {\tt stdout}. The output can be redirected to 
+a file using the {\tt -o} option which takes as mandatory argument 
+the file name. The default output format aims at maximal readability for
+a human eye, in many cases however, the output of {\tt globes} is intended
+for processing by other programs. There is a bunch of features which 
+should facilitate this. Usually one would like to omit the channel and rule 
+names by using simple printing {\tt -S} instead of pretty printing {\tt -P}.
+
+There are special options for certain special formats: {\tt -m} produces
+Mathematica\footnote{Mathematica is a trademark of Wolfram Inc.} list
+output, which can be directly visualized using {\tt MultipleListPlot}.
+The option {\tt -u} uses the same principal formatting than {\tt -m} does, but
+it allows to specify the left, middle and right delimiters in constructing
+the list, \eg\
+\begin{quote}
+left\\
+left 1 middle 2 middle 3 right\\
+middle\\
+left\\
+left 1 middle 2 middle 3 right\\
+right\\
+\end{quote}
+is, with left = '\{', middle = ',' and right = '\}',
+equivalent to the list $\{\{1,2,3\},\{1,2,3\}\}$. 
+The delimiters can be set by {\tt -L}, {\tt -M} and {\tt -R} as in the 
+following example
+\begin{quote}
+{\tt
+globes -Su -R\$'\verb+\n+' \verb+--+Middle=" " -L" " ...
+}
+\end{quote}
+where {\tt \$'\verb+\n+'} is the escape sequence in the shell for ANSI C
+like characters like linefeed '\verb+\n+'. The above example
+produces a a two columns style file like this
+\begin{quote}
+1.0 0.12\\
+1.2 0.14\\
+1.3 0.18\\
+...
+\end{quote}
+where the first columns is the central energy of the bin and
+the second columns gives the event rate. The output is a concatenation
+of many such two columns tables, each rule part or channel has its
+own table. Thus one can, by using
+{\tt -u} and user-defined delimiters, construct many different  
+output formats.
+
+\subsection*{\AEDL\ variable substitution}
+\index{norm}{{\tt globes}!variable substitution}
+
+Some {\tt .glb}-files use \AEDL\ variable substitution in order to allow
+special purpose studies like of baseline dependence. If no special care is
+taken, any \AEDL\ variable which is not defined within a given {\tt .glb}-file
+is interpreted by the parser as zero. Thus it would be impossible to properly
+parse with {\tt globes} any files which contain such undefined variables. 
+Hence, there is the possibility to define \AEDL\ variables by using the define
+option {\tt -D}. A call like this
+\begin{quote}
+{\tt
+globes -DBASELINE=3000 ...
+}
+\end{quote}
+would define the \AEDL\ variable {\tt BASELINE} to be $3000$.
 

### 2004-07-13 Documentation/part1.tex
Minor changes only The spaces in index entries with \_ remains, it seems to be rooted in the way how \index processes its arguments -> we have to edit the idn-files at last by hand :-<

@@
 user will not need to access directly in most cases. The experiment definition files, which usually end with {\tt .glb}, and any
-supporting files, are first of all searched in the current directory, and then in the global path variable \GLBC{GLB\_PATH} set in your environment.
+supporting files, are first of all searched in the current directory, and then in the global path variable 
+%
+\index{norm}{GLB PATH@{\tt GLB\_PATH}}
+\index{norm}{Environment variables!GLB PATH@{\tt GLB\_PATH}}
+%
+set in your environment.
+\index{norm}{Path resolution}
 A list of pre-defined experiment prototypes, their filenames, their short descriptions, and the references of their definitions can be found in \tabl{experiments}.

### 2004-07-13 Documentation/part2.tex
Minor changes only The spaces in index entries with \_ remains, it seems to be rooted in the way how \index processes its arguments -> we have to edit the idn-files at last by hand :-<

@@
 as a pairs, \ie, they are given in the form {\tt normalization : tilt}. The errors are treated in the same way. For example, we have
+\index{aedl}{{\tt rule}!signal error@{\tt "@signalerror}}
+\index{aedl}{{\tt rule}!background error@{\tt "@backgrounderror}}
+\index{aedl}{{\tt rule}!background center@{\tt "@backgroundcenter}}
 \begin{quote}
@@
 The error dimensions for the case of no systematics is set for each rule 
-by the value of {\tt @errordim\_sys\_off}\index{aedl}{{\tt rule}!{\tt "@errordim\_sys\_off}}, whereas the error dimension
+by the value of {\tt @errordim\_sys\_off}\index{aedl}{{\tt rule}!errordim 
+sys off@{\tt "@errordim\_sys\_off}}, whereas the error dimension
 for systematics on is given by {\tt @errordim\_sys\_on}
-\index{aedl}{{\tt rule}!{\tt "@errordim\_sys\_on}}. Thus, for example,
+\index{aedl}{{\tt rule}!errordim sys on@{\tt "@errordim\_sys\_on}}. Thus,
+ for example,
 the complete error dimension definition could look like  
@@
 with the symbol $\infty$).
+
  }
 \end{center}
+\index{aedl}{{\tt rule}!errordim@{\tt "@errordim}} 
 \end{table} 
@@
 a given experiment. To this end the \GLOBES\ package also features a program
-called {\tt globes}\index{norm}{{\tt globes}}, which can be seen either as a 
+called {\tt globes}\index{norm}{globes@{\tt globes}}, which can be seen either as a 
 crude \AEDL\ debugger 
@@
 
-The {\tt globes}\index{norm}{{\tt globes}} binary is installed along 
-with the library \index{norm}{{\tt libglobes}}, but
+The {\tt globes}\index{norm}{globes@{\tt globes}} binary is installed along 
+with the library \index{norm}{libglobes@{\tt libglobes}}, but
 into the directory {\tt \$prefix/bin/}. In order to use the {\tt globes} 
@@
 
-\index{norm}{{\tt globes}!oscillation parameters}
+\index{norm}{globes@{\tt globes}!oscillation parameters}
 Calling {\tt globes} without any options and with a {\tt .glb}-file as argument produces an event summary at rule level, taking into account 
@@
 option {\tt -p} on a call by call basis or by setting the environment variable
-{\tt GLB\_CENTRAL\_VALUES}\index{norm}{{\tt  GLB\_CENTRAL\_VALUES}}.
+{\tt GLB\_CENTRAL\_VALUES}
+%
+\index{norm}{GLB CENTRAL VALUES@{\tt GLB\_CENTRAL\_VALUES}} 
+\index{norm}{Environment variables!GLB CENTRAL VALUES@{\tt GLB\_CENTRAL\_VALUES}}
+%
 \begin{quote}
@@
 
-\index{norm}{{\tt globes}!warnings}\index{norm}{{\tt globes}!errors}
-\index{norm}{{\tt globes}!verbosity}
+\index{norm}{globes@{\tt globes}!warnings}
+\index{norm}{globes@{\tt globes}!errors}
+\index{norm}{globes@{\tt globes}!verbosity}
 The \AEDL\ parser and interpreter has basically three levels of messages to
@@
 
-\index{norm}{{\tt globes}!total rates}
-\index{norm}{{\tt globes}!spectral rates}
+\index{norm}{globes@{\tt globes}!total rates}
+\index{norm}{globes@{\tt globes}!spectral rates}
 By default {\tt globes} returns total rates corresponding to the 
@@
 \subsection*{Channel level}
-\index{norm}{{\tt globes}!channel rates}
+\index{norm}{globes@{\tt globes}!channel rates}
 
@@
 \subsection*{Rule level}
-\index{norm}{{\tt globes}!rule rates}
+\index{norm}{globes@{\tt globes}!rule rates}
 
@@
 \subsection*{Output}
-\index{norm}{{\tt globes}!output}
+\index{norm}{globes@{\tt globes}!output}
 
@@
 \subsection*{\AEDL\ variable substitution}
-\index{norm}{{\tt globes}!variable substitution}
+\index{norm}{globes@{\tt globes}!variable substitution}
 

### 2004-07-14 Documentation/appendix.tex
two files added for the GPL and FDL
*** empty log message ***
GLoBES tour added
added the installation section in app A added targte 'fast' to Makefile

@@
 \chapter{\GLOBES\ installation}
+\label{app:installation}
+\index{norm}{Installation|(}
 
-\chapter{Statistics and \GLOBES }
+The installation of \GLOBES\ is highly automated and there should not be
+any problems on a decently up-to-date GNU/Linux system. The installation
+has however only been tested on a limited number of platforms, mainly
+running with various versions of SuSE Linux. Thus we would appreciate
+to know your experiences with installation on different platforms, please
+send e-mail to <\bugs>.
 
-Here comes a statistics summary (similar to Superbeams appendix):
-\bi
-\item
- What $\chi^2$ is used
-\item
- How channels/rules are treated
-\item
- How does the flux-folding work
-\item
- ...
-\ei
+\section*{Prerequisites for installation of \GLOBES}
+\index{norm}{Installation!prerequisites}
 
-\chapter{Licence agreement}
+Besides the usual things like a working libc you need to have
+\begin{itemize}
+\item[gcc] The GNU compiler collection\\
+\verb+gcc.gnu.org+
+\item[BLAS] Basic Linear Algebra Subprograms\\
+\verb+www.netlib.org/blas/+    
+\item[LAPACK] Linear Algebra PACKage\\
+\verb+www.netlib.org/lapack/+
+\item[f2c] Fortran to C\\ 
+\verb+www.netlib.org/f2c/+
+\item[GSL] The GNU Scientific Library\\
+\verb+www.gnu.org/software/gsl/+
+\end{itemize}
+The library \verb+libglobes+ should in principle compile with any
+ANSI C/C++ compiler but the \verb+globes+ binary uses the 
+\verb+argp+ facility of \verb+glibc+ to parse its command line options.
+
+All those libraries are also available as rpm's from the various
+distributors of GNU/Linux, see their web sites for downloads. Due to
+the rather difficult nature of the build process of BLAS and LAPACK
+it is recommended to use a suitable rpm instead. Chances are that
+gcc, f2c and GSL are already part of your installation. Furthmore you need
+a working \verb+make+ to build and install \GLOBES.
+
+In future releases it is planed to at least get rid of LAPACK, because
+then also BLAS and f2c would drop out. Furthemore there is no prinicipal
+reason why not also to get rid of any C++ parts. This will greatly simplify
+the build process and ease the requirements for installation.
+
+\section*{Installation Instructions}
+
+\GLOBES\ follows the standard GNU installation procedure.  To compile \GLOBES\
+you will need gcc.  After unpacking the distribution
+the Makefiles can be prepared using the configure command,
+\begin{quote}
+{\tt
+./configure
+}
+\end{quote}
+You can then build the library by typing,
+\begin{quote}
+{\tt
+make
+}
+\end{quote}
+A shared  version of the library will be compiled by
+default.
+
+The library can be installed using the command,
+\begin{quote}
+{\tt
+make install
+}
+\end{quote}
+The default install directory prefix is \verb+/usr/local+.  Consult the
+"Further Information" section below for instructions on installing the
+library in another location or changing other default compilation
+options.
+
+The install target also will install a program with name \verb+globes+ to
+\verb+$prefix+.
+
+If you are not using \verb+make install+ you will find the static libary
+at \verb+source/.lib/libglobes.a+ which you can copy to any destination.
+However keep in mind that the linking command will be somewhat different,
+\ie\ you have to specify all the dynamically linked objects besides libglobes,
+which are
+\begin{quote}
+{\tt
+         -lm -lgsl -lgslcblas -lf2c -llapack -lblas
+}
+\end{quote}
+
+In general it is preferable to use the shared libraries, in case you don't
+have root privileges see the corresponding section of this file.
+
+\section*{Basic Installation}
+
+   The \verb+configure+ shell script attempts to guess correct values for
+various system-dependent variables used during compilation.  It uses
+those values to create a \verb+Makefile+ in each directory of the package.
+Finally, it creates a shell script \verb+config.status+ that
+you can run in the future to recreate the current configuration, a file
+\verb+config.cache+ that saves the results of its tests to speed up
+reconfiguring, and a file \verb+config.log+ containing compiler output
+(useful mainly for debugging \verb+configure+).
+
+   If you need to do unusual things to compile the package, please try
+to figure out how \verb+configure+ could check whether to do them, and mail
+diffs or instructions to \bugs\ so they can
+be considered for the next release.  If at some point \verb+config.cache+
+contains results you don't want to keep, you may remove or edit it.
+
+   The file \verb+configure.in+ is used to create \verb+configure+ by a program
+called \verb+autoconf+.  You only need \verb+configure.in+ 
+if you want to change it or regenerate \verb+configure+ using a newer 
+version of \verb+autoconf+.
+
+The simplest way to compile this package is:
+\begin{enumerate}
+\item \verb+cd+ to the directory containing the package's source code and type
+     \verb+./configure+ to configure the package for your system.  If you're
+     using \verb+csh+ on an old version of System V, you might need to type
+     \verb+sh ./configure+ instead to prevent \verb+csh+ from trying to execute
+     \verb+configure+ itself.
+
+     Running \verb+configure+ takes awhile.  While running, it prints some
+     messages telling which features it is checking for. It also prints
+     a reminder for things to do after installation.
+
+\item Type \verb+make+ to compile the package.
+
+\item Type \verb+make install+ to install the programs and any data files and
+     documentation.
+
+\item You can remove the program binaries and object files from the
+     source code directory by typing \verb+make clean+.  To also remove the
+     files that \verb+configure+ created (so you can compile the package for
+     a different kind of computer), type \verb+make distclean+.  There is
+     also a \verb+make maintainer-clean+ target, but that is intended mainly
+     for the package's developers.  If you use it, you may have to get
+     all sorts of other programs in order to regenerate files that came
+     with the distribution.
+
+\item Since you've installed a library don't forget to run \verb+ldconfig+ !
+\end{enumerate}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Installation without root privilege}
+\index{norm}{Installation!w/o root privilege}
+
+Install \GLOBES\ to a directory of your choice \verb+GLB_DIR+. This is done by
+\begin{quote}
+\verb+configure --prefix=GLB_DIR+
+\end{quote}
+and then follow the usual installation guide. The only remaining problem
+is that you have to tell the compiler where to find the header files, and 
+the linker where to find the library. Furthermore you have to make sure 
+that the shared object files are found during execution. Running 
+\verb+configure+
+also produces a \verb+Makefile+ in the \verb+examples+ 
+subdirectory which can serve as
+a template for the compilation and linking process, since all necessary
+flags are correctly filled in.
+
+Another solution is to set the environment variable \verb+LD_RUN_PATH+
+during linking to \verb+GLB_DIR/lib+ . Best thing is to add this to your
+shell dot-file (e.g. \verb+.bashrc+). Then you can use:
+A typical compiler command like
+\begin{quote}
+\verb+gcc -c my_program.c -IGLB_DIR/include/+
+\end{quote}
+and a typical linker command like
+\begin{quote}
+\verb#g++ my_program.o -lglobes -LGLB_DIR/lib/ -o my_executable#
+\end{quote}
+More information on this issue can be obtained by having a look 
+into the aforementioned  \verb+Makefile+ in \verb+examples+.
+
+\begin{itemize}
+\item[CAVEAT] It is in principle possible to have many installations 
+on one machine,
+especially the situation of having an installation by root and by a user at
+the same time might occur. However it is strictly warned against this 
+possibility since it is \emph{extremely} likely to create some 
+versioning problem at some time!
+\end{itemize}
+
+\section*{Advanced topics}
+
+\subsection*{Compilers and Options}
+
+   Some systems require unusual options for compilation or linking that
+the \verb+configure+ script does not know about.  You can give \verb+configure+
+initial values for variables by setting them in the environment.  Using
+a Bourne-compatible shell, you can do that on the command line like
+this
+\begin{quote}
+\verb+CC=c89 CFLAGS=-O2 LIBS=-lposix ./configure+
+\end{quote}
+Or on systems that have the `env' program, you can do it like this
+\begin{quote}
+\verb+env CPPFLAGS=-I/usr/local/include LDFLAGS=-s ./configure+
+\end{quote}
+
+\subsection*{Compiling For Multiple Architectures}
+
+   You can compile the package for more than one kind of computer at the
+same time, by placing the object files for each architecture in their
+own directory.  To do this, you must use a version of \verb+make+ that
+supports the \verb+VPATH+ variable, such as GNU \verb+make+.  `cd' to the
+directory where you want the object files and executables to go and run
+the \verb+configure+ script.  \verb+configure+ automatically checks for the
+source code in the directory that \verb+configure+ is in and in `..'.
+
+   If you have to use a \verb+make+ that does not supports the \verb+VPATH+
+variable, you have to compile the package for one architecture at a time
+in the source code directory.  After you have installed the package for
+one architecture, use `make distclean' before reconfiguring for another
+architecture.
+
+\subsection*{Installation Names}
+
+By default, \verb+make install+ will install the package's files in
+\verb+/usr/local/bin+, \verb+/usr/local/include+, etc.  You can specify an
+installation prefix other than \verb+/usr/local+ by giving \verb+configure+ the
+option \verb+--prefix=PATH+.
+
+\subsection*{Specifying the System Type}
+
+   There may be some features \verb+configure+ can not figure out
+automatically, but needs to determine by the type of host the package
+will run on.  Usually \verb+configure+ can figure that out, but if it prints
+a message saying it can not guess the host type, give it the
+\verb+--host=TYPE+ option.  TYPE can either be a short name for the system
+type, such as `sun4', or a canonical name with three fields: 
+\verb+CPU-COMPANY-SYSTEM+
+
+See the file \verb+config.sub+ for the possible values of each field.
+
+   If you are building compiler tools for cross-compiling, you can also
+use the \verb+--target=TYPE+ option to select the type of system they will
+produce code for and the \verb+--build=TYPE+ option to select the type of
+system on which you are compiling the package.
+
+\subsection*{Sharing Defaults}
+   If you want to set default values for \verb+configure+ scripts to share,
+you can create a site shell script called \verb+config.site+ that gives
+default values for variables like \verb+CC+,  \verb+cache_file+, and 
+\verb+prefix+.
+\verb+configure+ looks for \verb+PREFIX/share/config.site+ if it exists, then
+\verb+PREFIX/etc/config.site+ if it exists.  Or, you can set the
+\verb+CONFIG_SITE+ environment variable to the location of the site script.
+A warning: not all \verb+configure+ scripts look for a site script.
+
+\index{norm}{Installation|)}
+
+\input{gpl}
+
+\input{fdl}
 
@@
 \end{appendix}
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: Manual.tex
+%%% End:

### 2004-07-15 Documentation/Paper/paper.tex
Meine Version ...

@@
 TUM-HEP-5XX/04\\
-MPP-2004-YY\\
 %hep-ph/
@@
 We present the \GLOBES\ (``General Long Baseline Experiment Simulator'')
-software package, which allows the simulation of future long-baseline
+software package, which allows the simulation of long-baseline and reactor
 neutrino oscillation experiments. One part of the software is the
 abstract experiment definition language to define experiments with
-beam and full detector simulations as accurate as possible. For the
-definition of experiments, a new transparent building block concept is introduced. In addition, a binary provides the possibility to test
-new experiment definitions quickly.
+beam and full detector descriptions as accurate as possible. For the
+definition of experiments, a new transparent building block concept is 
+introduced. In addition, an additional program provides the possibility 
+to develop and test new experiment definitions quickly.
 %
 Another
-part of \GLOBES\ is the user's interface, which provides probability, rate, and $\Delta \chi^2$ information for a given experiment or any
-combination of up to 32 experiments. Especially, the
+part of \GLOBES\ is the user's interface, which provides probability, rate, 
+and $\Delta \chi^2$ information for a given experiment or any
+combination of up to 32 experiments in C. Especially, the
 $\Delta \chi^2$ functions allow a simulation with statistics only,
@@
 Neutrino oscillations are now established as the leading flavor
-transition mechanism for neutrinos. By the existing experiments, 
+transition mechanism for neutrinos. In a long history of many experiments, see
+\eg~\cite{Barger:2003qi}, 
 two oscillation
-frequencies have been found: The fast ``atmospheric'' and the
+frequencies have been identified: The fast ``atmospheric'' and the
 slow ``solar'' oscillations, which are driven by the respective
 mass squared differences.  In addition, there could be interference
-effects between these two oscillations, provided that the coupling given by the small mixing angle $\theta_{13}$ is large enough. Such interference
+effects between these two oscillations, provided that the coupling given by 
+the small mixing angle $\theta_{13}$ is large enough. Such interference
 effects include, for example, leptonic CP violation.  In order to
-test the unknown oscillation parameters, \ie, the mixing angle $\theta_{13}$, the leptonic CP phase, and the
-neutrino mass hierarchy, future long-baseline experiments are proposed.
+test the unknown oscillation parameters, \ie, the mixing angle $\theta_{13}$,
+ the leptonic CP phase, and the
+neutrino mass hierarchy, new long-baseline and reactor experiments 
+are proposed.
 These experiments send an artificial neutrino beam to a detector, or
 detect the neutrinos produced by a nuclear fission reactor. However,
-instrinsic degenerate solutions in the neutrino oscillation
-probabilities affect these measurements and require optimization
-strategies for the experiments to maximally exploit complementary
-information. Therefore, a modern complete experiment simulation
-does not only need to have a beam and detector simulations, which are
-as accurate as possible, but also powerful tools to analyze correlations
-and degeneracies. In this study, we present the \GLOBES\ software package
+the presence of multiple solutions which are intrinsic to
+ neutrino oscillation probabilities~\cite{Fogli:1996pv,
+Burguet-Castell:2001ez,Minakata:2001qm,Barger:2001yr} 
+affect these measurements.
+Thus optimization strategies are required which maximally exploit 
+complementarity between experiments. Therefore, a modern complete experiment 
+simulation  and analysis tool does not only need to have 
+a highly accurate beam and detector simulation but also powerful 
+means to analyze correlations and degeneracies, especially for the combination
+of several experiments.
+In this paper, we present the \GLOBES\ software package
 as such a tool.
@@
 \GLOBES\ (``General Long Baseline Experiment Simulator'') is a flexible
-software tool to simulate neutrino oscillation long-baseline experiments using a complete three-flavor analysis. On the
+software tool to simulate and analyze neutrino oscillation 
+long-baseline and reactor experiments using a 
+complete three-flavor description. On the
 one hand, it contains a comprehensive abstract experiment definition
-language, which allows to describe most classes of long baseline experiments
+language (AEDL), which allows to describe 
+most classes of long baseline and reactor experiments
 at an abstract level. On the other hand, it provides a C-library to 
 process the experiment information in order to obtain oscillation
-probabilities, rate vectors, and $\Delta \chi^2$-values (\cf, \figu{GLOBES}). In addition, it provides a binary to test experiment
+probabilities, rate vectors, and $\Delta \chi^2$-values (\cf, \figu{GLOBES}). 
+In addition, it provides a binary program to test experiment
 definitions very quickly, before they are used by the application software.
-Currently, \GLOBES\ is available for Linux. Since the source code is included, the modifications to other operating systems should be doable.
+Currently, \GLOBES\ is available for GNU/Linux. Since the source code is 
+included, the modifications to other operating systems should be doable.
 
-\GLOBES\ allows to simulate experiments with stationary neutrino point sources, where each experiment is assumed to have only one neutrino source.
+\GLOBES\ allows to simulate experiments with stationary neutrino point 
+sources, where each experiment is assumed to have only one neutrino source.
 Such experiments are neutrino beam experiments and reactor experiments. 
@@
 atmosphere, can not be described. In addition, sources with a physically 
-significant time dependencies  can not be studied, such as  supernov\ae. It 
-is, however, possible to simulate beams with bunch structure, since the 
-time dependence of the neutrino source is physically only important to suppress backgrounds.
+significant time dependencies  can not be studied, such as  supernov\ae.
 
@@
 On the experiment definition side, either built-in neutrino fluxes
-(\eg, neutrino factory) or arbitrary fluxes can be used. Similarly,
-arbitrary cross sections, energy dependent efficiencies, the
-energy resolution function, the considered oscillation channels, 
-backgrounds, and many other features can be specified. 
+(\eg, neutrino factory) or arbitrary, user defined fluxes can be used. 
+Similarly,
+arbitrary cross sections, energy dependent efficiencies,
+energy resolution functions as well as the considered oscillation channels 
+backgrounds, and many other properties can be specified. 
 For the systematics, energy
 normalization and calibration errors can be simulated. Note that
-the energy ranges and windows, as well as the bin widths can be
-(almost) arbitrarily chosen, which means that variable bin
-widths are allowed. Together with \GLOBES\ comes a number of
+energy ranges and windows and bin widths can be
+(almost) arbitrarily chosen, including bins of different
+widths. Together with \GLOBES\ comes a number of
 pre-defined experiments in order to demonstrate the capabilities
@@
 This approach allows the inclusion of multi-parameter-correlations,
-where external input (\eg, from solar parameters) can be imposed, too.
-Applications of the projection mechanism include the projections onto the $\stheta$-axis and the $\stheta$-$\deltacp$-plane.
-In addition, all oscillation parameters can be kept free to precisely localize
+where external constraints (\eg, on the solar parameters) can be imposed, too.
+Applications of the projection mechanism include the projections onto 
+the $\stheta$-axis and the $\stheta$-$\deltacp$-plane. In addition, 
+all oscillation parameters can be kept free to precisely localize 
 degenerate solutions.
@@
 the event rates, which is on of the core parts of the \GLOBES\
-software. However, because of the compecity of this issue,
+software. However, because of the complexity of this issue,
 we refer to the \GLOBES\ manual~\cite{Manual} for
@@
 evaluated {\em before} the smearing is performed, \ie, as function of $E$.
-Similarly, constant event rates can be added to the event rates before or after the energy smearing, which are refered to as ``pre-smearing backgrounds'' and ``post-smearing beackground''.
-These types of
-efficiencies an backgrounds allow a very accurate modeling of
-many factors, such as geoneutrino backgrounds, energy cuts, or
+Similarly, constant\footnote{with respect to the oscillation parameters, 
+not the energy} event rates can be added to the event rates before or 
+after the energy smearing, which are refered to as 
+``pre-smearing backgrounds'' and ``post-smearing background''. These types of 
+efficiencies and backgrounds allow a very accurate modeling of
+many factors, such as atmospheric or geoneutrino backgrounds, energy cuts, or
 energy threshold functions.
@@
 
-\section{Definition of experiments with \AEDL\ (Abstract Experiment 
-Defintion Language)}
+\section{Definition of experiments with \AEDL\\ (Abstract Experiment 
+Definition Language)}
 
@@
 Experiment Definition Language''). An experiment normally corresponds
-to an \AEDL\ file, which is a text file written in \AEDL\ syntax.
+to an \AEDL\ file, which is a human readable text file written 
+in \AEDL\ syntax.
+
+The key concept of \AEDL\ is to regard a detector as an abstract system
+which maps the true properties of a neutrino into the reconstructed properties
+of the neutrino which subsequently are used in the fit of the oscillation
+parameters. Within \GLOBES\ only the energy and flavour are used as 
+observables, which is
+sufficient for long baseline and reactor experiments. In  other
+experiments, like \eg\ atmospheric experiments, more observables like
+direction may have to be considered.
 
@@
 A channel corresponds to a neutrino oscillation channel including flux, cross section (for one specific interaction type), energy resolution function, initial and final neutrino flavors, their polarity (neutrinos or antineutrinos), and efficiencies.
-The channel raw event rates are computated according to \equ{events_bin}.
+The channel raw event rates are computed according to \equ{events_bin}.
 Each channel leads to the raw event rates for a
-specfifc interaction type. In \AEDL , many of the channel components
+specific interaction type. In \AEDL , many of the channel components
 have to be defined or loaded from files before, such as fluxes, 
@@
 rule, which leads to a $(\Delta \chi^2)_r$. The signal or background
-within each rule allows the specification of signal and background normalization errors and energy tilt or calibration errors. 
-These systematical errors will be evaluated with the ``pull method''
-with the help of nuisance parameters REFERENCE. In addition to these
+within each rule allows the specification of signal and background 
+normalization errors and energy tilt or calibration errors. 
+These systematical errors are evaluated with the ``pull method'' like 
+in~\cite{Huber:2002mx,Fogli:2002pt}.  
+In addition to these
 systematics errors, an overall evaluation strategy is assigned 
@@
 of all rules. This approach allows the definition of 
-appearance and disappearance channels, neutrino and antineutrino running, or interaction types with different systematics (spectral information
+appearance and disappearance channels, neutrino and antineutrino running, 
+or interaction types with different systematics (spectral information
 versus counting rate) within one experiment. The \GLOBES\ user interface
@@
 in the experiment definition, which is the matter density profile.
-For an experiment, an arbitrary matter density profiles can be specified, which is evaluated with the evolution operator method REFERENCES.
-In addition, the matter density profile is multiplied by the matter
-density scaling factor, which is treated as an independent parameter
+For an experiment, an arbitrary matter density profiles can be specified, 
+which is evaluated with the evolution operator method REFERENCES.
+In addition, the matter density profile is multiplied by a scaling factor, 
+which is treated as an independent parameter
 with a relative (matter density) uncertainty. With this
-approach, one can take into account that the matter density profile along a specific baseline is only known to about $3\%-5\%$. Therefore, it makes
-sense to combine the rules of one baseline in one experiment,
-since for all of these rules the matter density profile is equal.
+approach, one can take into account that the matter density profile along a 
+specific baseline is only known to about $3\%-5\%$.
 
-\section{Simulation of experiments: The C user's interface}
+\section{Analysis of experiments: The C user's interface}
 
@@
 
-In order to use \GLOBES , a C~library is linked with the application
+In order to use \GLOBES , a C-library is linked with the application
 software. This library provides the user's interface functions.
@@
 of the neutrino oscillation parameter space is well investigated and
-rather fortunate. Therefore, we use local minimization for each
-solution in the parameter space at the chosen confidence level,
-where tools for the location of (disconnected) degenerate solutions 
-are provided. For this process, the $6+N$-dimensional local minimizer
-is started at the ``guessed'' solution (such as from analytical 
-knowledge) and runs into the local minimum. In \figu{projex},
+rather smooth. Therefore, it is feasible to use local minimization for each
+solution in the parameter space.  For this process, 
+a $6+N$-dimensional local minimizer
+is started at the ``guessed'' position for the solution
+(such as from analytical knowledge) and then runs into a
+local minimum. This exercise has to be repeated till all local 
+minima, \ie\ degenerate solutions, are found. In \figu{projex},
 this minimization is illustrated at the simple example of a 
@@
 As we have discussed in the last section, the matter density of
-each experiment is treated as an addition parameter, which is known
+each experiment is treated as an additional parameter, which is known
 from external measurement with a certain precision. In principle, 
@@
 since they provide better precisions for the solar neutrino
-osillation parameters $\sdm$ and $\theta_{13}$ than the long-baseline
+oscillation parameters $\sdm$ and $\theta_{12}$ than long-baseline
  experiments themselves.
@@
 Though it is, in principle, possible to use such a method with \GLOBES ,
-it is impractiable in most cases, since it would involve at least 
+it is impracticable in most cases, since it would involve at least 
 billions of
@@
 steps, where $N_{\mathrm{Deg}}$ is the number of (disconnected) 
-local minima at the chosen confidence level, \ie, the number of
+local minima, \ie, the number of
 degeneracies.
@@
 
-The \GLOBES\ package~\cite{Globes} is a tar-Ball which consists of the
-C-library for the user's interface, a binary to test
-\AEDL\ files, example files, 
-experiment prototypes with supporting flux and cross 
-section files, and other 
+The \GLOBES\ package~\cite{Globes} is a tar-ball which consists of the
+source code for the C-library providing the user's interface, the source code
+ for a program to test
+\AEDL\ files, example C-files illustrating the use of the library, 
+experiment prototypes in \AEDL\ with supporting flux and cross 
+section files, and the usual 
 supporting files for installation and compilation.
-Currently, the C-library is only provided for Linux.
-In addition, the user's interface and \AEDL\ manual~\cite{Manual} 
-can be obtained separately. Upon extracting the tar-ball,
-the \GLOBES\ user's interface library needs to be configured, 
-compiled, and installed as 
-described in the {\tt INSTALL} file. In principle, it can be
-compiled as a static library or a dynamically-linked shared library
-(which can be installed with superuser rights or only used in one's
-home directory). Then it can be linked against any C program
-containing the application software, such as the examples
-from the manual. \AEDL\ files, such as the experiment protoypes,
+Currently, \GLOBES\ is only provided for GNU/Linux.
+In addition, an extensive manual covering the user's interface and 
+\AEDL~\cite{Manual} 
+can be obtained from the web-site. \GLOBES\ is free software and as such 
+licensed under the GNU Public License.
+
+The installation of \GLOBES\ is highly automated by the use of the
+Autotools family and follows the usual procedures. By default
+a shared library {\tt libglobes.so} is installed but 
+a static version is available
+also. Additionally an executable with name {\tt globes} is installed. 
+Provisions have been made to install \GLOBES\ without
+root privilege into a user's home directory. During the automatic
+configuration process also an example Makefile is produced which can
+serve as a template for compiling and linking own applications against
+{\tt libglobes.so}. This Makefile can be used to directly compile and
+link the examples
+from the manual.
+
+\AEDL\ files, such as the experiment prototypes,
 can be edited with any text editor, in order to be loaded by the
-users's interface later. In addition, a binary is provided to
-test and create experiment descriptions. For example, event rate
+users's interface later. In addition, the {\tt globes} binary allows to
+develop and test experiment descriptions. For example, event rate
 information can be quickly provided to adjust the neutrino flux
@@
 In summary, the \GLOBES\ software package provides powerful tools
-to do a full three-flavor analysis of future long-baseline experiments
+to do a full three-flavor analysis of future long-baseline and reactor
+ experiments
 including systematics, correlations, and degeneracies. In addition,
 the abstract experiment definition language allows to define
-experiments at an abstract level as accurate as possible. 
+experiments at an abstract level in a highly accurate fashion. 
 Some of the major strengths of \GLOBES\ are the ability to quickly
 define new experiments, the potential to take into account the
-full multi-parameter correlation, and the possibilities to include
-external input and matter density uncertainties.
+full multi-parameter correlation, the possibilities to include
+external input and matter density uncertainties and last but not least
+the ability to jointly analyze several experiments.
 
@@
 
+For the future it is planed to further extend the flexibility of
+\GLOBES\ by providing support for dynamically loadable modules, to
+allow completely user defined systematics or non-standard oscillation
+probabilities.
+
 \subsection*{Acknowledgments}
 
-We would like to thank Martin Freund and Andrea Romanino ???
-for initiating this project many years ago with one of the first
-neutrino factory simulations. In addition, we would like to thank
-Marc Rolinec for the help to translate the experiment descriptions
+We would like to thank Martin Freund, who wrote the very first
+version of a three-flavour matter profile treatment many years ago.
+PH is especially thankful for the invaluable advice of Thomas Fischbacher on
+many design issues. 
+In addition, we would like to thank
+Mark Rolinec for his help to translate the experiment descriptions
 into \AEDL , and for creating the illustrations in this manuscript.
 Finally, thanks to all the people who have been pushing this project
-for many years, to the ones who have been continuing asking for the publication of the software, and the referees of several of our
+for many years, to the ones who have been continuing asking for the 
+publication of the software, and the referees of several of our
 papers for suggestions which lead to improvements in the software.
+Parts of this works have been supported by the ``Sonderforschungsbereich
+375 f\"ur Astro-Teilchenphysik der Deutschen Forschungsgemeinschaft''.
 

### 2004-07-15 Documentation/Manual.tex
Minor changes only The spaces in index entries with \_ remains, it seems to be rooted in the way how \index processes its arguments -> we have to edit the idn-files at last by hand :-<
added the installation section in app A added targte 'fast' to Makefile
*** empty log message ***

@@
 % Always use GLBNS when adding a command to the index without showing it!
-\newcommand{\GLBNS}[1]{\index{api}{#1@{\tt #1}}}
+\newcommand{\GLBNS}[1]{\index{api}{#1@{\tt \protect{#1}}}}
 
@@
 % Always use GLB when mentioning a command in text!
-\newcommand{\GLBC}[1]{{\tt #1}\index{constants}{{\tt  #1}}}
+\newcommand{\GLBC}[1]{{\tt #1}\index{constants}{\protect{\tt{#1}}}}
 % Always use GLBNS when adding a command to the index without showing it!
-\newcommand{\GLBNSC}[1]{\index{constants}{{\tt #1}}}
+\newcommand{\GLBNSC}[1]{\index{constants}{{\protect{\tt #1}}}}
 \newcommand{\gq}[1]{\begin{quote} {\tt #1} \end{quote}}
 \newcommand{\go}[1]{$\rightarrow$ Output: \begin{quote} {\sf #1} \end{quote}}
-
+\newcommand{\bugs}{{\tt globes@ph.tum.de}}
 \definecolor{light}{gray}{0.93}
@@
 \vspace*{0.5cm}
-{\large \bf User's and reference manual }
+{\large \bf User's and experiment definition manual }
 
-\vspace*{1.5cm}
-{\large \bf !!!DRAFT IN PROGRESS!!! }
+%\vspace*{1.5cm}
+%{\large \bf !!!DRAFT IN PROGRESS!!! }
 \end{center}
@@
 \GLOBES\ (``General Long Baseline Experiment Simulator'') is a flexible
-software package to simulate neutrino long baseline experiments. On the
+software package to simulate neutrino oscillation 
+long baseline experiments. On the
 one hand, it contains a comprehensive abstract experiment definition
@@
 \AEDL (``Abstract Experiment Definition Language'') is a language
-to define experiments in form of normal text files. One or more of 
+to define experiments in form of ordinary text files. One or more of 
 the resulting \AEDL\ files can then be processed together with supporting 
@@
 containing the experiment definition(s). The user interface is linked 
-against the application software, which calls the user interface functions
+against the application software, and provides the user interface functions
 for the intended experiment simulation.
@@
 simulation of neutrino factories is much more difficult. In order
-to demonstrate some of these difficulties, we show in this manual only
+to demonstrate some of these difficulties, we present in this manual only
 examples with neutrino factories. These examples can be found in
@@
 interface is successively introduced from very basic to more sophisticated
-function. Eventually, it is demonstrated how one can change many
+functions. Eventually, it is demonstrated how one can change many
 experiment parameters at running time (such as baseline or target mass), and how one can obtain low-level
@@
 For example, flux normalizations can with this tool be easily adjusted 
-to reproduce the event rates of a specific experiment.
+to reproduce the event rates of a specific experiment. It is described
+in the last chapter of \Part~\ref{part:2}.
 

### 2004-07-15 Documentation/appendix.tex
*** empty log message ***

@@
 any problems on a decently up-to-date GNU/Linux system. The installation
-has however only been tested on a limited number of platforms, mainly
+has, however, only been tested on a limited number of platforms, mainly
 running with various versions of SuSE Linux. Thus we would appreciate
-to know your experiences with installation on different platforms, please
-send e-mail to <\bugs>.
+to know your experiences with the installation on different platforms.
+Please send an e-mail to <\bugs>.
 
-\section*{Prerequisites for installation of \GLOBES}
+\section*{Prerequisites for the installation of \GLOBES}
 \index{norm}{Installation!prerequisites}
 
-Besides the usual things like a working libc you need to have
+Besides the usual things such as a working libc, you need to have
 \begin{itemize}
@@
 The library \verb+libglobes+ should in principle compile with any
-ANSI C/C++ compiler but the \verb+globes+ binary uses the 
+ANSI C/C++ compiler, but the \verb+globes+ binary uses the 
 \verb+argp+ facility of \verb+glibc+ to parse its command line options.
@@
 All those libraries are also available as rpm's from the various
-distributors of GNU/Linux, see their web sites for downloads. Due to
-the rather difficult nature of the build process of BLAS and LAPACK
-it is recommended to use a suitable rpm instead. Chances are that
-gcc, f2c and GSL are already part of your installation. Furthmore you need
+distributors of GNU/Linux: See their web sites for downloads. Due to
+the rather difficult nature of the build process of BLAS and LAPACK,
+it is recommended that one use a suitable rpm instead. In addition,
+there is a good chance that
+gcc, f2c and GSL are already part of your installation. Furthmore, you need
 a working \verb+make+ to build and install \GLOBES.
 
-In future releases it is planed to at least get rid of LAPACK, because
-then also BLAS and f2c would drop out. Furthemore there is no prinicipal
-reason why not also to get rid of any C++ parts. This will greatly simplify
-the build process and ease the requirements for installation.
+In future releases, it is planned to at least make LAPACK obsolete, because
+then also BLAS and f2c would drop out. Furthemore, there is no priniciple
+reason why one should not get rid of any C++ parts. 
+This will greatly simplify
+the build process and reduce the requirements for installation.
 
@@
 
-In general it is preferable to use the shared libraries, in case you don't
-have root privileges see the corresponding section of this file.
+In general it is advisable to use the shared libraries. If you don't
+have root privileges, see the corresponding section of this file.
+If you install \GLOBES\ with root privileges, do not forget to run 
+{\tt ldconfig} after installation.
 
@@
 various system-dependent variables used during compilation.  It uses
-those values to create a \verb+Makefile+ in each directory of the package.
+these values to create a \verb+Makefile+ in each directory of the package.
 Finally, it creates a shell script \verb+config.status+ that
@@
 More information on this issue can be obtained by having a look 
-into the aforementioned  \verb+Makefile+ in \verb+examples+.
+into the mentioned  \verb+Makefile+ in \verb+examples+.
 
@@
 \item[CAVEAT] It is in principle possible to have many installations 
-on one machine,
-especially the situation of having an installation by root and by a user at
-the same time might occur. However it is strictly warned against this 
-possibility since it is \emph{extremely} likely to create some 
-versioning problem at some time!
+on one machine.
+Especially the situation of having an installation by root and by a user at the same time might occur.
+However, it is strictly warned against this possibility, since it is \emph{extremely} likely to create some versioning problem at some time!
 \end{itemize}
@@
 use the \verb+--target=TYPE+ option to select the type of system they will
-produce code for and the \verb+--build=TYPE+ option to select the type of
+produce code for, and the \verb+--build=TYPE+ option to select the type of
 system on which you are compiling the package.

### 2004-07-15 Documentation/part1.tex
*** empty log message ***

@@
 The format of \GLB{\_params} is discussed in detail in \Chapt~\ref{chapt:gettingstarted}.
-Note that all functions but {\tt ChiSys}
-  are using minimizers which have to be initialized with \GLB{SetInputErrors} and \GLB{SetStartingValues} first.}
+Note that all functions but {\tt glbChiSys} are using minimizers which have to be initialized with \GLB{SetInputErrors} and \GLB{SetStartingValues} first.}
 \end{table}
@@
 experiments.
-Note that a minimum at a negative value of $\theta_{13}$ is unphysical.
-However, if there can be no degeneracy found at a positive value, there is probably none at a low confidence level.
 \gq{
@@
 Note that each experiment
-assumes a specific matter density profile for all of its rules, which means
+assumes a specific matter density profile, which means
 that it makes sense to simulate different operation modes within one
@@
 together with the neutrino flux, whereas the target mass scales the whole
-experiment.
+experiment. Thus, if one has, for instance, a neutrino and an antineutrino
+running mode, one can scale them independently.
 
@@
 
-Similarly, the density parameters or iterations can be accessed:
+Similarly, the density parameters or number of iterations 
+(returned by the minimizers) can be accessed:
 \begin{function}
@@
 \begin{function}
-\GLBNS{TestLibraryVersion}
-{\tt int glbTestLibraryVersion(const char *version)} returns $0$ if
-the {\tt version} string of the format ``X.Y.Z'' is exactly the
-used \GLOBES\ version, $1$ if it is older, and $-1$ if it is newer.
-Note that the library and \GLOBES\ versions are not the same. 
-\end{function}
-\begin{function}
 \GLBNS{VersionOfExperiment} 
@@
 the version string of the experiment number {\tt experiment}.
-The version string points to the experiment structure, which means
-that it cannot be altered and does not have to be destroyed by the
+The version string is allocated within the experiment structure, which means
+that it cannot be altered and must not be destroyed by the
 user.
@@
 In principle, the precision of an individual parameter measurement including correlations in the $\chi^2$-approach can be obtained as the projection of the $n$-dimensional fit manifold onto the respective axis.
-Similarly, one can project the fit manifold onto a plane, such as the $\stheta$-$\deltacp$-plane, if one wants to explicitely show the allowed
+Similarly, one can project the fit manifold onto a plane, such as the $\stheta$-$\deltacp$-plane, if one wants to show the allowed
 region in this plane with all the other parameter correlations included. 
@@
 For example, taking only $N_{\mathrm{grid}}=20$ and $n=7$ (six oscillation parameters and matter density) would mean more than one billion function calls of {\tt glbChiSys}.
-One can easily image that this is too much for any sophisticated application.
+One can easily imagine that these
+would be too many for any reasonable application.
 
@@
 to converge into the unwanted best-fit solution.
-However, note that one should in this case check that the actually determined value for $\ldm$ after minimization is close enough to the guessed values $-\ldm$ in order to avoid significant artifical contributions of the priors to the
+However, note that one should in this case check that the actually determined value for $\ldm$ after minimization is close enough to the guessed value $-\ldm$ in order to avoid significant artifical contributions of the priors to the
 final $\chi^2$.
@@
 \\
-  /* Define two-parameter projection: Only deltacp is free! */ \\
+  /* Define my own two-parameter projection for glbChiNP: Only deltacp is free! */ \\
   glbDefineProjection(th13\_projection,GLB\_FIXED,GLB\_FIXED,GLB\_FIXED,GLB\_FREE,GLB\_FIXED,GLB\_FIXED); \\
@@
 \index{norm}{Matter density scaling factor}
-The the output of the minimizer in {\tt out} carries as many matter density scaling factors as there are experiments. Either
+The output of the minimizer in {\tt out} carries as many matter density scaling factors as there are experiments. Either
 one (for the simulation of one experiment) or all (for the simulation of
 all experiments) are different from $1.0$ if matter density uncertainties
-are present, since each experiment may face other matter density conditions. The minimizer for individual experiments ``know'' which
+are present, since each experiment may face other matter density conditions. The minimizers of individual experiments ``know'' which
 experiment they are currently treating, which means that only the returned
@@
 For example, calculating {\tt glbChiTheta} for the last experiment
-number, the last density values will be modified.
+number, the last density value will be modified.
 This approach turns out to be extremely useful together with the
@@
 
-In the last chapter, we introduced the projection of any set of $k$ parameter onto any $n-k$ dimensional hyperplane, which was done by the minimization over the $k$ free fit parameters.
+In the last chapter, we introduced the projection of any set of $k$ parameters onto any $n-k$ dimensional hyperplane, which was performed by the minimization over the $k$ free fit parameters.
 Similarly, one can minimize over {\em all} $n$ parameters to find the local minimum close to any starting point.
@@
 the local minimizer at the suspected position 
-and let it converge into the minimum.  The following code excerpt  corresponds to finding the degenerate solution for the example on
+and let it run into the minimum.  The following code excerpt  corresponds to finding the degenerate solution for the example on
   page~\pageref{ex:corrth13dcp}, and it is from {\tt example3.c}: 
@@
 The example on page~\pageref{ex:sgndeg} illustrates how to locate the $\mathrm{sgn}(\ldm)$-degeneracy and show the corresponding degenerate solution in the $\stheta$-$\deltacp$-plane together with the original solution.
-In this case, the position of the degeneracy can be easily guessed to be at the best-fit parameter values but the $\ldm$ inverted.
-The minimizer then runs off the plane of the best-fit parameters into the local minimum.
-It is very important to take into account the position of the degeneracy off this plane, since the actual $\chi^2$ in the minimum is certainly lower than on the plane of the best-fit parameter values.
+In this case, the position of the degeneracy can be easily guessed to be at the best-fit parameter values but the inverted $\ldm$.
+The minimizer then runs off the plane of these parameters into the local minimum.
+It is very important to take into account the position of the degeneracy off this plane, since the actual $\chi^2$ in the minimum is probably lower than on the plane of the best-fit parameter values.
 Thus, the degeneracy may not even appear at the chosen confidence level on the plane, but it does appear at the true minimum.
-The two sections\footnote{The discussed figure on page~\pageref{ex:sgndeg} is produced by  {\tt glbChiSys} and thus only represents a section through the fit manifold.
-For the projection including correlations, one may rather want to use {\tt glbChiThetaDelta}.} through the fit manifold shown in the figure on page~\pageref{ex:sgndeg} therefore do not appear at the same oscillation parameter values (except from the ones shown in the figure).
+The two sections through the fit manifold shown in the figure on page~\pageref{ex:sgndeg} therefore do not appear at the same oscillation parameter values (except from the ones shown in the
+figure).\footnote{The discussed figure on page~\pageref{ex:sgndeg} is produced by  {\tt glbChiSys} and thus only represents a section through the fit manifold.
+For the projection including correlations, one may rather want to use {\tt glbChiThetaDelta}.}
 
@@
 to avoid a significant contribution of the prior to the final $\chi^2$.
-Alternatively, one could run {\tt glbChiAll} again with the located minimum
-as {\tt in} vector and $\ldm$ kept free.
+Alternatively, one could once again run {\tt glbChiAll} with the located 
+minimum as {\tt in} vector, and $\ldm$ kept free.
 \item[Finding degeneracies with multiple experiments.] For multiple experiments, it turns out to be useful to locate the degeneracies for individual experiments first.
@@
 \GLBNS{ProfileProbability}
-{\tt double glbProfileProbability(int l, int m, int panti,double E)}
- returns the neutrino oscillation probability $\nu_l \rightarrow \nu_m$ for the energy {\tt E} in matter. The parameter
+{\tt double glbProfileProbability(int exp, int l, int m, int panti,double E)} returns the neutrino oscillation probability $\nu_l \rightarrow \nu_m$ for the energy {\tt E} in matter, where the matter density profile
+is the one of experiment {\tt exp}. The parameter
 {\tt panti} is $+1$ for neutrinos and $-1$ for antineutrinos.
@@
 \end{function}
-THIS FUNCTION MAY OBTAIN AN EXPERIMENT INDEX (WW).
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{function}
-Since \GLOBES\ reserved the memory for the lists returned in these
+Since \GLOBES\ reserves the memory for the lists returned in these
 functions, which it allocates on an internal stack, one has to reset
@@
 \end{function}
-\begin{function}
-\GLBNS{StaceyProfile}
+The average matter density $\bar{\rho}(L)$ for a matter density profile $\rho(x)$ along the baseline $L$ baseline is defined as
+\be
+\bar{\rho}(L) = \frac{1}{L} \int\limits_{0}^L \rho(x) dx = 
+\frac{1}{L} \int\limits_{0}^L \tilde{\rho}\left(d(x) \right) dx \, , 
+\ee
+where $\tilde{\rho}(d)$ is the PREM matter density as function of the
+distance $d$ to the Earth's core, and $d(x)=\sqrt{x^2+R^2-2 x R \cos \theta }$ is the purely geometrical relationship between $d$ and $x$ with the
+Earth radius $R$ and the nadir angle $\cos \theta = L/(2 R)$.
+\begin{function}
+\GLBNS{GetProfileDataInExperiment}
 {\tt int glbGetProfileDataInExperiment(int exp,size\_t *layers, double** lengths, double** densities)} returns the matter density profile 
@@
 \GLBNS{SetFilterState}
-{\tt int glbSetFilterState(int on\_off)} sets the filter state to
-on (\GLBC{GLB\_ON}) or off (\GLBC{GLB\_OFF}).
+{\tt int glbSetFilterState(int on\_off)} sets the currently
+used filter state to on (\GLBC{GLB\_ON}) or off (\GLBC{GLB\_OFF}).
 \end{function}
@@
 \GLBNS{GetFilterState}
-{\tt int glbGetFilterState()} returns the filter state.
+{\tt int glbGetFilterState()} returns the currently used filter state.
 \end{function}
@@
 \GLBNS{SetFilter}
-{\tt int glbSetFilter(double filter)} sets the filter to the
+{\tt int glbSetFilter(double filter)} sets the currently used filter to the
 value {\tt filter}.
@@
 \GLBNS{GetFilter}
-{\tt double glbGetFilter()} returns the filter value.
+{\tt double glbGetFilter()} returns the currently used filter value.
 \end{function}

### 2004-07-15 Documentation/part2.tex
*** empty log message ***

@@
 Finally, the transition probability is obtained by the absolute square of the total neutrino transition amplitudes.
-Depending on the precision of the studied experiment, this approach turns out to be precise enough in Earth matter even for a small number matter density steps.
+Depending on the precision of the studied experiment, this approach turns out to be precise enough in Earth matter even if only a small number matter density steps is used.
 Since we allow an uncertainty of the matter density profile, it is, in fact, in most cases sufficient to consider only one density step with the average matter density together with a matter density uncertainty~\cite{Ohlsson:2003ip}.
@@
 and a set $D(F,F')$ for the flavor variable $F$.
-In general, $D(E,E')$ also depends on the true flavor $F$, as well as $D(F,F')$ depends on the true energy $E$.
+In general, $D(E,E',F)$ also depends on the incident neutrino flavor $F$, as well as $D(F,F',E)$ depends on the incident neutrino energy $E$.
 These sets of mapping functions usually are obtained from a
@@
 (\cf, \figu{rule}). The event numbers from these channels are added {\em before} the $\Delta \chi^2$-value is calculated.\footnote{Note that
-in this manual, we $\chi^2$ and $\Delta \chi^2$ are equal, since
+in this manual, the $\chi^2$ and $\Delta \chi^2$ are equal, since
 for simulated data $\Delta \chi^2 = 0$ at the best-fit point. Thus, we
@@
 Currently, a number of pre-defined experiment definition files are provided with \GLOBES , which have to be modified manually in order to define new experiments.
-An IDE (Integrated Developement Environment) for AEDL files is currently in preparation.
+An IDE (Integrated Developement Environment) for AEDL files may follow in one of the next
+versions.
 The application software then uses this text file to initialize the experiment, where other secondary files might read for source fluxes, cross sections \etc .
@@
 %
-Of course, the baseline and matter profile have to be specified, too:
+Of course, the baseline and matter profile have to be specified, too, where
+we use an arbitrary matter density profile here:
 \begin{quote}
 {\tt /* baseline */}\\
-{\tt \$baseline = 3000.0}\\
+{\tt \$profiletype = 3}\\
 {\tt \$densitytab = \{3.5\}}\\
 {\tt \$lengthtab = \{3000.0\}}\\
-{\tt \$density\_error = 0.05}
 \end{quote}
@@
 {\tt \$lengthtab} refer to a list of numbers. Here, the lists contain only
-one element, because we only use one density layer: We initialize a baseline length of $3000 \, \mathrm{km}$ with a constant matter density of $3.5 \, \mathrm{g/cm^3}$ and a matter density uncertainty of $5\%$ ($1 \sigma$ relative error).
-Note that the elements in {\tt \$lengthtab} have to add up to the baseline length.
+one element, because we only use one density layer: We initialize a baseline length of $3000 \, \mathrm{km}$ with a constant matter density of $3.5 \, \mathrm{g/cm^3}$. 
 %
@@
 \end{quote}
-The first element is the name of the flux, which we have already defined above as defined above. The second element $\pm$ determines whether 
+The first element is the name of the flux, which we have defined above. 
+The second element ``$\pm$'' determines whether 
 neutrinos or anti-neutrinos are taken from the flux table (two different polarities allowed). The third position defines the initial flavor,
@@
 is also checked. For example,  {\tt 
-\$bins} can be only between $0$ and $500$~\footnote{The upper limit is 
-there only for safety reasons, the memory is allocated dynamically.}.
-If one uses a {\tt float} quantity where  an {\tt int} is expected, the {\tt float} will be converted to an {\tt int} in the same way as in C.
+\$bins} can be only between $0$ and $500$.\footnote{The upper limit is 
+only there for safety reasons, the memory is allocated dynamically.} If one uses a {\tt float} quantity where  an {\tt int} is expected, the {\tt float} will be converted to an {\tt int} in the same way as in C.
 For example, we have scalar variables
@@
 \end{quote}
-This command resets the counters for channels, rules and energy resolution
-functions. All variables have their scope limited 
+This command resets the counters for channels, rules, fluxes, and energy resolution environments. All variables have their scope limited 
 by either {\tt \%!GLoBES, \#NEXT\#} or {\tt EOF}.  This allows 
@@
 \end{quote}
-in order to initialize a combined analysis of the experiments defined in the files {\tt NuFact.gls} and {\tt JHFHK.gls}. As one sees from the example above
-file names are always quoted in \AEDL.\index{norm}{File names}
+in order to initialize a combined analysis of the experiments defined in the files {\tt NuFact.gls} and {\tt JHFHK.gls}. Note that one has 
+to use quotation marks for filenames in \AEDL.\index{norm}{File names}
 Even if one uses the
@@
 \end{quote}
-These constants can not only be defined within one \AEDL\ file but also
+These constants can not only be defined within one \AEDL\ file, but also
 by the calling C program, which allows to use a simple but powerful variable
@@
 
-Alos some simple algebraic manipulations are possible, such as
+In addition, some simple algebraic manipulations are possible, such as
 \begin{quote}
@@
 \hline
-$1$ & - & Average density (constant) \\
-$2$ & {\tt \$densitysteps} & PREM profile with given number of equidistant steps \\
+$1$ & {\tt \$baseline} & Average density (constant) \\
+$2$ & {\tt \$baseline}, {\tt \$densitysteps} & PREM profile with given number of equidistant steps \\
 $3$ & {\tt \$lengthtab},  {\tt \$densitytab} & Arbitrary profile (table of layer thicknesses, table of densities) \\
@@
 {\tt \$densitytab=\{2.8, 3.5\}}\\
-{\tt \$lengthtab=\{1000.0, 2000.0\}}\\
-\end{quote}
-It is important that both lists habe the same length and that the  thicknesses given in  {\tt \$lengthtab} add up to the value of
-{\tt \$baseline}. In addition, matter densities are always given in $g/cm^3$.
+{\tt \$lengthtab=\{1000.0, 2000.0\}}
+\end{quote}
+It is important that both lists habe the same length and that the  thicknesses given in  {\tt \$lengthtab} add up to the length of
+the baseline, which does not have to be explicitely specified anymore. In addition, matter densities are always given in $g/cm^3$.
 %
@@
 {\tt \$densitytab=\{3.5\}}\\
-{\tt \$lengthtab=\{3000.0\}}\\
+{\tt \$lengthtab=\{3000.0\}}
 \end{quote}
@@
 \end{quote}
-
+%
 \index{norm}{Oscillation!switching off}
@@
 \end{quote}
-
- In that case
-all diagonal probabilities $P_{\alpha\alpha}$ are unity and all off-diagonal
-probablities  $P_{\alpha\beta}$ are zero. This is \eg\ especially useful for
-neutral current events since they do not depend on any oscillation 
-parameters\footnote{at least in the absence of sterile neutrinos}. The channels
-marked as {\tt NOSC\_} are only computed by a call to \GLB{SetRates}
-and not recomputed in the subsequent fit (which calls \GLB{SetNewRates}).
+%
+In this case all diagonal probabilities $P_{\alpha\alpha}$ are unity, and all off-diagonal probablities  $P_{\alpha\beta}$ are zero. This is,
+for instance, useful for
+neutral current events, since these do not depend on any oscillation 
+parameters\footnote{At least in the absence of sterile neutrinos}. The channels
+marked as {\tt NOSC\_} are already computed by \GLB{SetRates}
+and do not have to be recomputed in the subsequent fit (which calls the undocumented function \GLB{SetNewRates}).
 Therefore this feature can be used to speed up the rate 
@@
 Usually it is an excellent approximation to treat backgrounds as if they
-would not be affected by oscillations.
+were not affected by oscillations.
 
@@
 In most cases, the reader may want to continue to the next section after reading these two subsections.
-In the third subsection, we describe a more elaborate (and slower) smearing algorithm, which can be used together with rather fast neutrino oscillations compared to the bin size, such as for reactor experiments.
-Eventually, we show how one can use a manual smearing matrix instead of using one of the implemented algorithms.
+In the third subsection, we describe a more elaborate (and slower) smearing algorithm, which can be used together with rather fast neutrino oscillations compared to the bin size, such as for
+(solar) reactor experiments to avoid aliasing effects. Eventually, we show how one can use a manual smearing matrix instead of using one of the implemented algorithms.
 
@@
 However, this exceeding is only
-necessary to avoid aliasing effects, if the integrand in \eq~(\ref{eq:simple_int}) is large at the limits of the sampling range.
+necessary to avoid aliasing effects if the integrand in \eq~(\ref{eq:simple_int}) is large at the limits of the sampling range.
 For a manual definition of the sampling points (depends on algorithm), use
@@
 will cancel between the simulated and fitted data. 
-%
-Furthermore, we will demonstrate in the next subsection that the assumption
-is rather good. ??? COMES THERE ???
-%
 This algorithm is the fastest which \GLOBES\ offers, and is selected by
@@
 The computation of the bin kernel $K_i^c$ is performed
-by \GLOBES. Thus, it requires  the number of bins
- $N_\mathrm{bins}$
-and the minimum $E_\mathrm{min}$ and maximum energy $E_\mathrm{max}$
-for equidistant bins.
-
+by \GLOBES. Thus, it requires that the number of bins
+{\tt \$bins} and the minimum energy {\tt \$emin} and maximum energy {\tt \$emax} for equidistant bins are given.
 %
@@
 In order to ensure that fast oscillating probabilities do not lead to 
-aliasing\index{norm}{Aliasing}
-it is desirable to impose a low-pass filter already during the calculation
-of the probabilities itself. This is possible and implemented has a highly
+aliasing,\index{norm}{Aliasing}
+it is possible to impose a low-pass filter already during the calculation
+of the probabilities itself. This filter is implemented has a highly
 experimental feature called ``filter''\index{norm}{Filter}. 
 The calculation of oscillation
-probabilities basically is a computation of phase differences. Restricting
-the maximum admissible size of those phase difference effectively filters
+probabilities is, in principle, a computation of phase differences. Restricting the maximum admissible size of those phase differences effectively filters
 the high frequency component of the oscillation probability. This idea is
@@
 $\sigma_f(E):=\sigma_f^0 \cdot E$ ensures that this filter behaves 
-approximately like an energy resolution function with constant width 
+approximately such as an energy resolution function with constant width 
 $\sigma_e=\sqrt{2}/\sigma_f^0$, \ie\
@@
 \end{equation}
-The equivalence of equation~\ref{eq:filter_a} and equation~\ref{eq:filter_b}
-is not obvious and connected to the properties of $P_{\alpha\beta}$, 
-see~\cite{Kiers:1996zj,Giunti:2003ax}. This feature works \emph{only} 
+The relationship between \eqs~\ref{eq:filter_a} and~\ref{eq:filter_b}
+is not obvious and connected to the properties of $P_{\alpha\beta}$: 
+see \Refs~\cite{Kiers:1996zj,Giunti:2003ax}. This feature works \emph{only} 
 for vacuum and constant densities and is controlled
-by the filer state variable and $\sigma_e$ is set by the filter value 
-variable:
+by the filer state variable. In addition, $\sigma_e$ is set by the filter value variable:
 \index{aedl}{filter state@{\tt \$filter\_state}}
@@
 function, $\sigma_e$ adds on to the energy resolution function 
-of the detector $\sigma_c(E)$ like
+of the detector $\sigma_c(E)$ such as
 \begin{equation}
@@
 \end{equation} 
-Sometimes this behaviour is undesirable and therefore one can try to 
+Sometimes this behavior is unwanted, and therefore one can try to 
 'subtract' the filtering from the energy resolution function by splitting 
@@
 is used instead of $\sigma_c(E)$ in computing the
-smearing data. Thus one obtains that
+smearing data. Thus one obtains as effective energy resolution
 \begin{equation}
@@
 \end{equation} 
-This scheme is enabled by choosing as type for the energy resolution
+This scheme is used by choosing as type for the energy resolution
 \begin{quote}
@@
 For a rule, the splitting
-in signal and background is useful for simplicity and the treatment of systematics, as we will se later. Each rule will lead to a $\chi^2$-value,
-which means that all $\chi^2$'s of the different rules will be added
+in signal and background is useful for the treatment of systematics, as we will se later. Each rule will lead to a $\Delta \chi^2$-value,
+which means that all $\Delta \chi^2$'s of the different rules will be added
 for the whole experiment. Within each rule, the event rates are added, and
@@
 also for a  Poissonian $\chi^2$-function.
-In~\cite{Fogli:2002pt}, it has been demonstrated that the pull method 
+In \Ref~\cite{Fogli:2002pt}, it has been demonstrated that the pull method 
 and the covariance based approach are equivalent for a Gau\ss ian and 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Versioning of \AEDL\ files}
+\section{Version control in \AEDL\ files}
 \label{sec:aedl_versioning}
@@
 
-In order to provide a handle on problems which may arise due to different
-versions of \GLOBES\ and \AEDL\ files, it is possible to stamp each 
-\AEDL\ file with the corresponding version number of the \GLOBES\ package with
+In order to avoid problems which come from different
+versions of \GLOBES\ and \AEDL\ files, it is possible to use in each 
+\AEDL\ file a version number. For example, it may correspond to the  minimum version number of the \GLOBES\ package with which it works. Set the
 {\tt \$version}\index{aedl}{version@{\tt \$version}} by
@@
 
-\AEDL\ is a powerful language to describe a vast variety 
+\AEDL\ is a powerful language to describe a variety 
 of different experiments.
-This chapter demonstrates how to test whether a given \AEDL\ file describes
-a given experiment. To this end the \GLOBES\ package also features a program
-called {\tt globes}\index{norm}{globes@{\tt globes}}, which can be seen either as a 
-crude \AEDL\ debugger 
-or a simple command line oriented tool to convert the rather 
-abstract \AEDL\ experiment description into more accessible event rates.
-
-\section{Basic usage of {\tt globes}}
+This chapter demonstrates how to test an \AEDL\ file in order to check
+if they really describes given experiment. 
+For this application, the \GLOBES\ package contains the program
+{\tt globes}\index{norm}{globes@{\tt globes}}. It can either be 
+regarded as an \AEDL\ debugger, or as a simple command-line oriented tool to convert the rather  abstract \AEDL\ experiment description into more accessible event rates.
+
+\section{Basic usage of the {\tt globes} binary}
 \label{sec:globes_basics}
 
-The {\tt globes}\index{norm}{globes@{\tt globes}} binary is installed along 
-with the library \index{norm}{libglobes@{\tt libglobes}}, but
+The {\tt globes}\index{norm}{globes@{\tt globes}} binary is installed 
+together with the library \index{norm}{libglobes@{\tt libglobes}}, but
 into the directory {\tt \$prefix/bin/}. In order to use the {\tt globes} 
-utility this directory\footnote{This is automatically the case, when
-no options have been given to {\tt configure}, and {\tt make install} was 
-executed with root-privilege, \ie\ a standard installation was done.} has to 
-be in the path of the shell used to invoke it.
-
-First {\tt globes} parses the {\tt .glb}-file which was given as argument,
-prints any warnings and errors which occured during parsing. Then it
-uses the experiment description in the file to compute the event rates
-at a certain point in parameters space. And finally it displays the result
-with a varying degree of detail dependent on the options used to invoke
-{\tt globes}. The option parsing follows the GNU standards and thus there
-is a {\tt \verb+--+help} option which displays all other options 
-along with a short description.
+utility, this directory has to be in the path of the shell used to call the program.\footnote{This is automatically done if
+no options are given to {\tt configure}, and {\tt make install} was 
+executed with root-privilege, \ie, a standard installation was done.}
+
+As an argument, {\tt globes} takes a {\tt .glb}-file. While parsing it,
+it prints any warnings and errors which have occured during reading the file.
+Then it uses the experiment description in the file to compute the event rates at a certain point in parameter space.
+Finally, it displays the result based on the options used to call {\tt globes}.
+The options of {\tt globes} follow the GNU standard. Thus, there
+is a {\tt \verb+--+help} option to display all other options 
+together with short descriptions.
 
 \index{norm}{globes@{\tt globes}!oscillation parameters}
-Calling {\tt globes} without any options and with a {\tt .glb}-file as argument produces an event summary at rule level, taking into account 
-the full description in the file, \ie\ all efficiencies, backgrounds and resolution effects. Thus, these are the event rates which are actually 
-used to compute the $\chi^2$. By default the oscillation parameters used to
-calculate the transition probability are
+Calling {\tt globes} without any options and with a {\tt .glb}-file as argument produces an event summary at rule level. In this case,  
+the full experiment description in the file is taken into account, \ie, all efficiencies, backgrounds, and energy resolution effects. 
+Thus, the returned event rates are the ones which will be actually 
+used to compute the $\chi^2$ later. By default, the oscillation parameters used to calculate the transition probability are
 \begin{eqnarray}
@@
 Of course, it is possible to change these default values either by using the
-option {\tt -p} on a call by call basis or by setting the environment variable
-{\tt GLB\_CENTRAL\_VALUES}
+option {\tt -p} on a call by call basis, or by setting the environment variable
+{\tt GLB\_CENTRAL\_VALUES}:
 %
@@
 globes -p'0.55,0,0.785,0,0.0008,0.0025'\\
-globes \verb+--+parameters='0.55,0,0.785,0,0.0008,0.0025'\\
-}
-\end{quote}
-or by defining within the shell session or in the shell profile:
+globes \verb+--+parameters='0.55,0,0.785,0,0.0008,0.0025'
+}
+\end{quote}
+For example, {\tt GLB\_CENTRAL\_VALUES} can be defined
+within the shell session or in the shell profile:
 \begin{quote}
@@
 \end{quote}
-Furthermore it is possible to switch off oscillations altogether with t
-he {\tt -N} option
-and to switch them on again with {\tt -O} (the default). The effect of {\tt -N}
-is the same than to prefix all channels with {\tt NOSC\_}.
+Furthermore it is possible to switch off oscillations with the {\tt -N} option and to switch them on again with {\tt -O} (the default).
+The effect of {\tt -N} is the same as to use {\tt NOSC\_} in all oscillation channels.
+Thi feature is useful if one wants to normalize an expriment flux if
+the number of un-oscillated events is given.
 
@@
 The \AEDL\ parser and interpreter has basically three levels of messages to
-th user: warnings, errors and fatal errors. Fatal errors always are reported
-and the program exits with status '1'. Usually only errors and no warnings
-are reported. The verbosity level can be chosen by the {\tt -v} option, where
-{\tt -v1} is default, \ie\ only errors and fatal errors are reported. {-v0}
-corresponds to reporting fatal errors only, whereas {\tt -v2} will
-print in addition to fatal errors and errors also warnings. It is recommended
-to test any new {\tt .glb}-file with {\tt -v2} to see the warnings at least 
-once and to decide whether there is a problem and in case fix it. With {\tt -v3}
-all files read by {\tt globes} are displayed together with their path and
-with  {\tt -v4} all files which have been attempted to be read are shown.
+the user: Warnings, errors and fatal errors. 
+Fatal errors are always reported and lead to a 
+program exit with status '1'. Usually only errors and no warnings
+are reported.
+The verbosity level can be chosen by the {\tt -v} option, where {\tt -v1} is default, \ie, only errors and fatal errors are reported.
+The level {-v0} corresponds to reporting fatal errors only, and {\tt -v2} will print warnings in addition to fatal errors.
+It is recommended
+to test any new {\tt .glb}-file with {\tt -v2} to check the warnings at least once, and to decide whether there is a problem to be fixed.
+With {\tt -v3} all files read by {\tt globes} are displayed together with their path, and with  {\tt -v4} all files which have been attempted to be read are shown.
 These two setting are useful to clarify path resolution issues and shadowing
@@
 
-In the process of defining a new experiment the default output of {\tt globes}
-at rule level is the final step but in order to arrive there it is often
-necessary to view the intermediate steps in the event rate calculation. The 
-{\tt globes} utility offers many possibilities to do so based on the 
-rate access functions as described in \Sec~\ref{sec:event_rates}.
+In the process of defining a new experiment, the default output of {\tt globes} at rule level is the final step. However, in order to arrive 
+at this level it is often necessary to review the intermediate steps in the event rate calculation.
+The {\tt globes} utility offers many possibilities to do this based on the rate access functions as described in \Sec~\ref{sec:event_rates}.
 
@@
 \index{norm}{globes@{\tt globes}!spectral rates}
-By default {\tt globes} returns total rates corresponding to the 
-{\tt -t} option, this can be changed to
+By default, {\tt globes} returns total rates corresponding to the 
+{\tt -t} option. This can be changed to
 to a full spectrum by using {\tt -s}. The spectral rates are shown in a
@@
 
-If there are more than one experiment in a file, \ie\ at least one 
-{\tt \#NEXT\#}
-command, then only the event rates for one experiment are shown.  This
-one experiment can be choose with the {\tt -e} option which takes 
-as mandatory argument the number of the experiment (starting with zero), 
-where {\tt -e0} is the default.
+If there is more than one experiment in a file, \ie, there is at least one 
+{\tt \#NEXT\#} command, only the event rates for one experiment will be shown.  This experiment can be chosen with the {\tt -e} option, which takes 
+as a mandatory argument the number of the experiment (starting with zero).
+The default is {\tt -e0}.
 
@@
 
-As a first step one often wants to check whether each channel produces the 
-desired output. Channel rates are returned when the {\tt -c} option is used.
+As a first step, one may want to check if each channel produces the 
+anticipated output. Channel rates are returned if the {\tt -c} option 
+is used.
 This option takes as an optional argument the channel number 
-(starting at zero), if no argument is given all channels are displayed.
-By default the sum of the event rates in each channel is shown. Where each 
+(starting at zero). If no argument is given all channels are displayed.
+By default, the sum of the event rates in each channel is shown. Each 
 column has as first line the same channel name as in the file.
 
-It is also possible to take out one detector effect after the other. First
-one can switch off the post-smearing efficiencies ({\tt -f}) and the 
+It is also possible to switch off one detector effect after the other. 
+First, one can switch off the post-smearing efficiencies ({\tt -f}) and the 
 post-smearing backgrounds ({\tt -g}). Next, one can switch off the 
 energy resolution function with ({\tt -b}) and view the rates before
-smearing. In case the {\tt -s} options is also used, the number of
-lines in the output now is given by the setting of {\tt \$sampling\_points}.
+smearing. If the {\tt -s} option is also used, the number of
+lines in the output will be given by {\tt \$sampling\_points}.
 Another effect of the {\tt -b} option is that the post-smearing efficiencies
-and backgrounds are no longer taken into account. Therefore the {\tt -g} 
+and backgrounds are no longer taken into account. Therefore, the {\tt -g} 
 and {\tt -f} options now apply to the \emph{pre}-smearing efficiencies 
-and the \emph{pre}-smearing backgrounds. Thus 
+and the \emph{pre}-smearing backgrounds. Thus, 
 \begin{quote}
@@
 produces the raw event rate corresponding to the convolution of flux, 
-probability and cross section, neglecting all detector effects.
+probability, and cross section, which is neglecting all detector effects.
 
@@
 
-The logical step after channels is to check the rule rates. Rule rates 
-are returned with {\tt -r}.
+The next logical step after checking the channel rates is to investigate
+ the rule rates. The rule rates are returned with the option {\tt -r}.
 This option takes as an optional argument the rule number 
-(starting at zero), if no argument is given all rules are displayed.
-By default the sum of the event rates in each rule is shown, as well
-as for each part of the rule. Where each 
+(starting at zero). If no argument is given, all rules will be displayed.
+By default, the sum of the event rates in each rule is shown, as well
+as for each component within the rule. Each 
 rule is preceeded by a line with the same rule name as in the file.
 
-It is also possible for rules to take out one detector effect after 
-the other, however with the limitation that rules make only sense
+It is also possible for rules to switch off one detector effect after 
+the other -- with the limitation that rules only make sense
 after the energy resolution function has been applied to each channel.
-Therefore it is \emph{not} possible to use {\tt -b} together with {\tt -r} or
-to switch off any pre-smearing efficiencies or backgrounds. 
+Therefore, it is \emph{not} possible to use {\tt -b} together with {\tt -r}, or to switch off any pre-smearing efficiencies or backgrounds. 
 One can, however, switch off the post-smearing efficiencies ({\tt -f}) 
 and the post-smearing backgrounds ({\tt -g}) for each channel. Since
-the definition of a rule also contains so called coefficients it is
+the definition of a rule also contains so-called ``coefficients'', it is
 possible to switch them off with {\tt -i}. This options also deactivates
@@
 
-The default output stream is {\tt stdout}. The output can be redirected to 
-a file using the {\tt -o} option which takes as mandatory argument 
+The default output stream is {\tt stdout}. The output can be re-directed to 
+a file using the {\tt -o} option, which takes as mandatory argument 
 the file name. The default output format aims at maximal readability for
-a human eye, in many cases however, the output of {\tt globes} is intended
-for processing by other programs. There is a bunch of features which 
-should facilitate this. Usually one would like to omit the channel and rule 
+a human eye. In many cases however, the output of {\tt globes} is produced
+as input for other programs. There are some features to adjust the
+output format. Usually one would like to omit the channel and rule 
 names by using simple printing {\tt -S} instead of pretty printing {\tt -P}.
@@
 Mathematica\footnote{Mathematica is a trademark of Wolfram Inc.} list
-output, which can be directly visualized using {\tt MultipleListPlot}.
-The option {\tt -u} uses the same principal formatting than {\tt -m} does, but
-it allows to specify the left, middle and right delimiters in constructing
-the list, \eg\
+output, which can be directly visualized by {\tt MultipleListPlot}.
+The option {\tt -u} uses the same principal formatting as {\tt -m}, but
+it allows to specify the left, middle, and right delimiters in constructing
+the list, such as
 \begin{quote}
@@
 \end{quote}
-is, with left = '\{', middle = ',' and right = '\}',
+This is, with left = '\{', middle = ',' and right = '\}',
 equivalent to the list $\{\{1,2,3\},\{1,2,3\}\}$. 
 The delimiters can be set by {\tt -L}, {\tt -M} and {\tt -R} as in the 
-following example
+following example:
 \begin{quote}
@@
 \end{quote}
-where {\tt \$'\verb+\n+'} is the escape sequence in the shell for ANSI C
-like characters like linefeed '\verb+\n+'. The above example
-produces a a two columns style file like this
+Here {\tt \$'\verb+\n+'} is the escape sequence in the shell for ANSI C-like characters, such as linefeed '\verb+\n+'. The above example
+produces a a two column file such as
 \begin{quote}
@@
 \end{quote}
-where the first columns is the central energy of the bin and
-the second columns gives the event rate. The output is a concatenation
-of many such two columns tables, each rule part or channel has its
-own table. Thus one can, by using
+where the first columns is the central energy of the bin, and
+the second columns gives the event rate. Usually, the output is a concatenation of many such two columns tables, where 
+each rule part or channel part has its own table. Thus one can, by using
 {\tt -u} and user-defined delimiters, construct many different  
@@
 
-\subsection*{\AEDL\ variable substitution}
+\subsection*{\AEDL\ external variable substitution}
 \index{norm}{globes@{\tt globes}!variable substitution}
 
-Some {\tt .glb}-files use \AEDL\ variable substitution in order to allow
-special purpose studies like of baseline dependence. If no special care is
-taken, any \AEDL\ variable which is not defined within a given {\tt .glb}-file
-is interpreted by the parser as zero. Thus it would be impossible to properly
-parse with {\tt globes} any files which contain such undefined variables. 
-Hence, there is the possibility to define \AEDL\ variables by using the define
-option {\tt -D}. A call like this
+Some {\tt .glb}-files use external \AEDL\ variables in order to allow
+special purpose studies (such as the energy resolution-dependence). If 
+the external variables are not explicitely specified, they are
+interpreted by the parser as zeros. Thus, it is impossible to properly
+parse any files with {\tt globes} which contain such undefined variables. 
+Hence, there is the possibility to define \AEDL\ variables by using the define option {\tt -D}. A call such as 
 \begin{quote}

### 2004-07-16 Documentation/Manual.tex
part 1 finis

@@
 % Always use GLBNS when adding a command to the index without showing it!
-\newcommand{\GLBNS}[1]{\index{api}{#1@{\tt \protect{#1}}}}
+\newcommand{\GLBNS}[1]{\index{api}{#1@{\tt #1}}}
 
@@
 % Always use GLB when mentioning a command in text!
-\newcommand{\GLBC}[1]{{\tt #1}\index{constants}{\protect{\tt{#1}}}}
+\newcommand{\GLBC}[1]{{\tt #1}\index{constants}{\tt{#1}}}
 % Always use GLBNS when adding a command to the index without showing it!
-\newcommand{\GLBNSC}[1]{\index{constants}{{\protect{\tt #1}}}}
+\newcommand{\GLBNSC}[1]{\index{constants}{{\tt #1}}}
 \newcommand{\gq}[1]{\begin{quote} {\tt #1} \end{quote}}
@@
 
+\frontmatter
 \pagenumbering{Roman}
@@
     or any later version published by the Free Software Foundation;
-    with no Invariant Sections, no Front-Cover Texts, and no Back-Cover Texts.
+    with the invariant Sections ``Terms of usage of \GLOBES" 
+    and ``Acknowledgments'', no Front-Cover Texts, and no Back-Cover Texts.
     A copy of the license is included in the section entitled "GNU
@@
 software package to simulate neutrino oscillation 
-long baseline experiments. On the
+long baseline and reactor experiments. On the
 one hand, it contains a comprehensive abstract experiment definition
-language, which allows to describe most classes of long baseline experiments
+language (AEDL), which allows to describe most classes 
+of long baseline experiments
 at an abstract level. On the other hand, it provides a C-library to 
@@
 probabilities, rate vectors, and $\Delta \chi^2$-values. Currently, 
-\GLOBES\ is available for Linux. Since the source code is included,
-the modifications to other operating systems should be doable.
+\GLOBES\ is available for GNU/Linux. Since the source code is included,
+the port to other operating systems is in principle possible.
 
@@
 degenerate solutions.
+
+\chapter*{Terms of usage of \GLOBES}
+
+\GLOBES\ is mainly developed for academic use. Thus The \GLOBES\ Team would
+appreciate being given academic credit for it. Whenever you use \GLOBES\
+to produce a publication or a talk indicate that you have used \GLOBES\ and
+please cite the following reference
+\begin{quote}
+P. Huber, M. Lindner and W. Winter ... 
+\end{quote}
+and \emph{not} this manual. This manual does not constitute a scientific 
+publication and will not be submitted to any scientific journal. 
+It will evolve during time since it is intendend for 
+regular revision.
+Besides, many of the data which is used by \GLOBES\ and distributed together
+with it should be correctly referenced where possible. For details see the
+???.
+
+Apart from that, \GLOBES\ is free software and open source, as such it is 
+licensed under the GNU Public License.
 
@@
 want to modify or create \AEDL\ files. A useful tool in this context
-is the software program ``globes'', which returns event rates and other
+is the software program \verb+globes+, which returns event rates and other
 information for individual \AEDL\ files without further programming. 
@@
 
+\paragraph{Note:} All examples for application software in C do require
+a C++ compiler to be properly compiled. For pedagogical reasons, variable
+declarations are done at that place where the variable is needed for the
+first time, which is at variance with C syntax but not with C++ syntax.
+That is the only way in which the examples deviate form ISO C.
+
 %Finally, \Part~\ref{part:3} of this manual serves as reference manual,
@@
 
+\mainmatter
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\backmatter
+\chapter*{Acknowledgments}
+\addcontentsline{toc}{chapter}{Acknowledgments}
+
+We would like to thank Martin Freund, who wrote the very first
+version of a three-flavour matter profile treatment many years ago.
+PH is especially thankful for the invaluable advice of Thomas Fischbacher on
+many design issues. 
+In addition, we would like to thank
+Mark Rolinec for his help to translate the experiment descriptions
+into \AEDL , and for creating the illustrations in this manuscript.
+Finally, thanks to all the people who have been pushing this project
+for many years, to the ones who have been continuing asking for the 
+publication of the software, and the referees of several of our
+papers for suggestions which lead to improvements in the software.
+
 \input{appendix}
@@
 
+\chapter{Indices}
 \printindex{api}{API functions}
+
 \printindex{constants}{API constants \& macros}
+
 \printindex{aedl}{AEDL reference}
+
 \printindex{norm}{Index}

### 2004-07-16 Documentation/Paper/paper.tex
added the one missing reference
part 1 finis

@@
 For an experiment, an arbitrary matter density profiles can be specified, 
-which is evaluated with the evolution operator method REFERENCES.
+which is evaluated with the evolution operator method~\cite{Ohlsson:1999um}.
 In addition, the matter density profile is multiplied by a scaling factor, 
@@
 allow completely user defined systematics or non-standard oscillation
-probabilities.
+probabilities. Additionally we plan to provide interfaces to various script
+languages like Perl.
 

### 2004-07-16 Documentation/part1.tex
part 1 finis

@@
 main features of \GLOBES . The complete example  
-can be found as {\tt example-tour.c} in the example subdirectory of your \GLOBES\ distribution.
+can be found as {\tt example-tour.c} in the \verb+example+ subdirectory 
+of your \GLOBES\ distribution.
 The output is written to {\tt stream}, which can be either {\tt stdout},
@@
  fprintf(stream,"$\backslash$nOscillation probabilities in matter: ");\\
- for(i=1;i<4;i++) fprintf(stream,"1->\%i: \%g ",i,\\ \hspace*{4cm} glbProfileProbability(1,i,+1,50)); \\
+ for(i=1;i<4;i++) fprintf(stream,"1->\%i: \%g ",i,\\ \hspace*{4cm} glbProfileProbability(0,1,i,+1,50)); \\
 }
@@
 Since the \GLOBES\ parameter vectors depend on the number of experiments,
-we have to destroy them first:
+we have to free them first:
 \gq{
@@
 }
-Finally, we have to destroy the parameter vectors again:
+Finally, we have to free the parameter vectors again:
 \gq{
@@
 }
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \chapter{\GLOBES\ basics}
@@
 \index{norm}{Installation}
-In this first chapter of the user's manual, we assume that the \GLOBES\ software is readily installed on your computer system. For the installation,
+In this first chapter of the user's manual, we assume that the \GLOBES\ 
+software is readily installed on your computer system. For the installation,
 see \App~\ref{app:installation} and the {\tt INSTALL} file in the
@@
 into details of the programming language, which means that standard parts
-of the program code common to all of the examples in the following chapters are, in general, omitted.
-An example of a minimal \GLOBES\ program in C can be found on page~\pageref{ex:c}.
-Furthermore, the files of the examples in this part can be found in the {\tt Example} subdirectory of your \GLOBES\ distribution.
-\index{norm}{Examples} Since it depends on your installed
-configuration, we refer to the {\tt README} file in the \GLOBES\ main directory, and the comments in the Makefile of the {\tt Example} subdirectory for how to compile the example files. Make sure that
-the data files (\AEDL\ and supporting files) in the {\tt data} 
-subdirectory can be found by the
-examples. In case of doubt, you can simply copy the necessary files
-into the working directory where the examples are executed.
+of the program code common to all of the examples in the following chapters 
+are, in general, omitted.
+An example of a minimal \GLOBES\ program in C can be found on 
+page~\pageref{ex:c}. Furthermore, the files of the examples in this part 
+can be found in the \verb+example+ subdirectory of your \GLOBES\ distribution.
+ \index{norm}{Examples} 
+After the installation of \GLOBES, they can be compiled using 
+the \verb+Makefile+ in the \verb+examples+
+ directory. The  \verb+Makefile+ has been correctly setup by the 
+\verb+configure+
+script to take into account details of the installation on your system. Thus
+you've just to type \verb+make+ and you're done.
+\footnote{The  data files (\AEDL\ and supporting files) needed by the
+examples are already in place.} This \verb+Makefile+ very well serves
+as a template for your own applications.
+\index{norm}{Compilation!of application programs}
+\index{norm}{Build process| \see{Compilation}}
 
@@
 
-\hspace*{0.5cm}  /* Destroy parameter vector(s) */ \\
+\hspace*{0.5cm}  /* Free parameter vector(s) */ \\
 \hspace*{0.5cm}  glbFreeParams(true\_values); \\
@@
 experiment number initiates a combined analysis of all loaded experiments.
+\index{norm}{GLB ALL@{\tt GLB\_ALL}}
+\index{constants}{{\tt GLB\_ALL}}
+
+In general \GLBC{GLB\_ALL} can be used in many cases where there is an argument
+selecting `$i$ out of $N$', \eg\ the 1st experiment out of 5, or the 5th
+rule of 20. In those cases  using \GLBC{GLB\_ALL} is equivalent to
+calling the corresponding function for \emph{all} $i$ in $N$  and `add' 
+the effect of each
+invocation, like in
+\begin{quote}
+\verb&for (i=0;i<N;i++) result += some_function(i);&\\
+is the same as\\
+\verb&result = some_function(GLB_ALL);&
+\end{quote}
+Here the meaning of `add' is, that whatever the desired result of calling
+\verb+some_function+ is, this result is obtained for each $i$ in $N$, \eg\ 
+setting the baseline in all experiments to a certain value or
+or compute the $\chi^2$ for each experiment and return the total result.
+There are however some functions where the action performed or the result
+is so complex that is not possible or sensible to perform this for all $i$
+in $N$. Calling these functions with  \GLBC{GLB\_ALL} as argument will 
+in any case result in an exit status indicating
+failure and the function will produce an 
+error message\footnote{if the verbosity
+level is set accordingly}.
 
@@
 user will not need to access directly in most cases. The experiment definition files, which usually end with {\tt .glb}, and any
-supporting files, are first of all searched in the current directory, and then in the global path variable 
+supporting files, are first of all searched in the current directory, and then in the path given in the environment variable \verb+GLB_PATH+.
 %
@@
 %
-set in your environment.
+
 \index{norm}{Path resolution}
@@
 for oscillation parameters. In order to transfer sets of oscillation
-parameter vectors $(\theta_{12}, \theta_{13}, \theta_{23}, \deltacp, \sdm, \ldm)$, the parameter type \GLB{\_params\_type} is used.
-In general, pointers to this structure are of the type \GLB{\_params}, and are often
-transferred to and from \GLOBES\ functions. Therefore, the memory space
-for these vectors has to be reserved before they can be used, and it has
-to be returned afterwards. \GLOBES\ functions usually use the pointers
+parameter vectors $(\theta_{12}, \theta_{13}, \theta_{23}, \deltacp, \sdm, \ldm)$, the parameter type \GLB{\_params} is used. In general,
+ this type is often
+transferred to and from \GLOBES\ functions. Therefore, the memory 
+for these vectors has to be reserved (allocated) before they can be used, 
+and it has
+to be returned (freed) afterwards. \GLOBES\ functions usually use the pointers
 of the type {\tt glb\_params} for the input or output to the functions. As
@@
 profile of each experiment $i$ is multiplied by a scaling factor
-$\hat{\rho_i}$, \index{norm}{Matter density scaling factor}
-\index{norm}{Matter density profile} 
+$\hat{\rho_i}$, \index{norm}{Matter density!scaling factor}
+\index{norm}{Matter density!profile} 
 which is stored in the density information of {\tt glb\_params}. 
@@
 an external precision of the scaling factor might be used to include the
-correlations with the matter density uncertainty. For more details, we
-refer to the following sections.
+correlations with the matter density uncertainty.
 Note that the {\tt glb\_params}
@@
 Similarly, any change in the number of experiments requires that the
-parameter structures be re-initialized, \ie, destroyed and created again.
+parameter structures be re-initialized, \ie, freed and allocated again.
 
@@
 \GLBNS{AllocParams}
-{\tt void glbFreeParams(glb\_params stale)} destroys the memory space
+{\tt void glbFreeParams(glb\_params stale)} frees the memory
 needed for a parameter vector {\tt stale} and sets the pointer to {\tt NULL}.
@@
 glbFreeParams(vector1); \\
-/* ... more vectors destroyed ... */ \\
+/* ... more vectors freed ... */ \\
 \\
@@
 \section{Computing the simulated data}
-\index{norm}{True values} \index{norm}{Simulated data} \index{norm}{Reference rate vector}
+\index{norm}{True values} 
+\index{norm}{Simulated data} 
+\index{norm}{Reference rate vector}
 \index{norm}{Set oscillation parameters}
@@
 Compared to existing experiments, which use real data, future experiments
-uses simulated data.
-Thus, the {\em true parameter values} and their results in form of the reference event rate vectors are simulated.
-After setting the true parameter values, the {\em fit parameter values} can be varied in order to obtain information on the measurement performance for the given set of true parameter values.
-Therefore, it is often useful to show the results of a future measurement as function of the true parameter values for which the reference rate vectors are computed -- at least within the currently allowed ranges.
-The true parameter values for the vacuum neutrino oscillation parameters have to be set by the functions \GLB{SetOscillationParameters} and \GLB{SetRates} {\em before} any evaluation function is used and {\em after} the experiments have been initialized and the experiment parameters have been adjusted which could change the rates (such as baseline or target mass).
-This means that
-after any change of an experiment parameter, {\tt glbSetRates} has to be called.
-Matter effects are automatically included as specified in the experiment definition.
-We have the following functions to assign and
+use simulated data. Thus, the {\em true parameter values} and their results in 
+form of the reference event rate vectors are simulated. After setting the 
+true parameter values, the {\em fit parameter values} can be varied in order to 
+obtain information on the measurement performance for the given set of true 
+parameter values. Therefore, it is often useful to show the results of a 
+future measurement as function of the true parameter values for which the 
+reference rate vectors are computed -- at least within the currently allowed 
+ranges. The true parameter values for the vacuum neutrino oscillation 
+parameters have to be set by the function \GLB{SetOscillationParameters} 
+and the reference rate vector, \ie\ the data, has to be computed by a call 
+to \GLB{SetRates}. This has to be done {\em before} any evaluation 
+function is used and  {\em after} the experiments have been initialized and 
+also the experiment 
+parameters have been adjusted which could change the rates (such as baseline
+ or target mass). This means that
+after any change of an experiment parameter, \GLB{SetRates} has to be 
+called. Matter effects are automatically included as specified in the 
+experiment definition. We have the following functions to assign and
 read out the vacuum oscillation parameters:
@@
 The version string is allocated within the experiment structure, which means
-that it cannot be altered and must not be destroyed by the
+that it cannot be altered and must not be freed by the
 user.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \chapter[Calculating $\chi^2$ with systematics only]{Calculating $\boldsymbol{\chi^2}$ with systematics only}
@@
 In general, \GLOBES\ uses the six independent oscillation parameters $\theta_{12}$, $\theta_{13}$, $\theta_{23}$, $\deltacp$, $\sdm$, $\ldm$, as well as the matter density
-scaling factor $\hat{\rho}$ of each experiment.\index{norm}{Matter density
+scaling factor $\hat{\rho}$ of each experiment.\index{norm}{Matter density!
 scaling factor} Thus, there are six plus the number of experiments parameters determining the rate vectors.
@@
 given by observations is imposed (typically up to $5\%$). 
-\index{norm}{Matter density uncertainty}
+\index{norm}{Matter density!uncertainty}
  For this section, it is important to keep in mind
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \chapter[Calculating $\chi^2$-projections: how one can include correlations]{Calculating $\boldsymbol{\chi^2}$-projections: how one can include correlations}
@@
 
-\index{norm}{Multi-parameter correlation}
+\index{norm}{Correlations!multi-parameter}
 This chapter deals with the rather complicated issue of $n$-parameter correlations. It is one of the greatest strengths of this software 
@@
 
-\index{norm}{Projection of manifold}
-In principle, the precision of an individual parameter measurement including correlations in the $\chi^2$-approach can be obtained as the projection of the $n$-dimensional fit manifold onto the respective axis.
-Similarly, one can project the fit manifold onto a plane, such as the $\stheta$-$\deltacp$-plane, if one wants to show the allowed
+\index{norm}{Projection!of manifold}
+In principle, the precision of an individual parameter measurement including
+ correlations in the $\chi^2$-approach can be obtained as the projection of 
+the $n$-dimensional fit manifold onto the respective axis. Similarly, one can 
+project the fit manifold onto a plane, such as the $\stheta$-$\deltacp$-plane,
+ if one wants to show the allowed
 region in this plane with all the other parameter correlations included. 
-In practice, this projection is very difficult: a grid-based method would need $(N_{\mathrm{grid}})^n$ function calls of {\tt glbChiSys} to calculate the precision including the full $n$-parameter correlation, where $N_{\mathrm{grid}}$ is the number of points in each direction of the lattice.
-For example, taking only $N_{\mathrm{grid}}=20$ and $n=7$ (six oscillation parameters and matter density) would mean more than one billion function calls of {\tt glbChiSys}.
-One can easily imagine that these
-would be too many for any reasonable application.
+In practice, this projection is very difficult: a grid-based method would 
+need $(N_{\mathrm{grid}})^n$ function calls of {\tt glbChiSys} to 
+calculate the projection onto one dimension 
+including the full $n$-parameter correlation, where 
+$N_{\mathrm{grid}}$ is the number of points in each direction of the 
+lattice. For example, taking only $N_{\mathrm{grid}}=20$ and 
+$n=7$ (six oscillation parameters and matter density) would mean more 
+than one billion function calls of {\tt glbChiSys}. One can easily imagine 
+that these would be too many for any reasonable application.
 
 \index{norm}{Minimizer}
-The solution to this problem is using a local $n$-dimensional (local) minimizer for the projection instead of a grid-based method, where we will
+The solution to this problem is using a $n$-dimensional, local 
+minimizer for the projection instead of a grid-based method, where we will
 illustrate this minimization process later. It turns out
-that such a minimizer can include a full $6$-parameter correlation with of the order of $1\, 000$ function calls of {\tt glbChiSys}.
-It is a standard method which can be found in every good book for numerical calculation routines.
+that such a minimizer can include a full $6$-parameter correlation 
+with of the order of $1\, 000$ function calls of {\tt glbChiSys}. For the 
+minimization we use a derivative free method due to Powell in a 
+modified~\cite{Brent:1973} version\footnote{Not needing derivatives is
+highly desired since the event rate depends in a non-linear way on the
+oscillation parameters, thus there is no easy analytical way
+to obtain derivatives of the $\chi^2$ function.}.
+
 Thus, for each point on the projection axis/plane, one can obtain a result within about $10$ to $30$ seconds on a modern computer, which means that the complete measurement precision for one fixed true parameter set can be obtained in as much as $10$ to $15$ minutes.
@@
 
-This approach also has one major disadvantage: One can not simply program a robust grid-based code and let it run, since using a local minimizer always means that one may end up in an unwanted local minimum and not in the investigated global one.
-Thus, one has to use some (analytical or numerical) knowledge on the topology of the fit manifold and start the local minimizer close enough to the investigated solution.
-Fortunately, this can be done quite straightforward in most cases, since the structure of the neutrino oscillation formulas does not cause very complicated topologies of the fit manifolds.
-Especially the simulation of reactor experiments and conventional beams or superbeams is rather simple with purely numerical
+This approach also has one major disadvantage: There is \emph{no} such
+thing as a global minimization algorithm or even an algorithm which guarantees
+to find \emph{all} local minima of a function. In practice this means using 
+ a local minimizer, one may end up in an unwanted local minimum and not in 
+the investigated (possibly global) one or 
+one may miss a local minimum which affects the results\footnote{NB -- 
+Implementing a grid-based method which guarantee to find all local minima is
+not straightforward either, to say the least.}. 
+The only way out of this dilemma is to use some heuristic approach, \ie\
+although one can not guarantee anything one can use schemes which work
+in most cases and announce their failure loudly. In order to use such a
+heuristic some (analytical or numerical) knowledge on the topology of 
+the fit manifold is necessary. With this knowledge it is possible
+to obtain an approximate position for each local minimum 
+and thus to start the local minimizer 
+close enough to the investigated minimum.
+ Fortunately, this can be done quite straightforward in most cases, since 
+the structure of the neutrino oscillation formulas does not cause very 
+complicated topologies of the fit manifolds. Especially the simulation of 
+reactor experiments and conventional beams or superbeams is rather simple 
+with purely numerical
 approaches. Neutrino factories have, especially for small values of
@@
 of the many analytical discussions of this issue can be used. This means
- that one can implicitly use the analytical knowledge to obtain better predictions for the measurement performances, since the minimizer has
- to be started only close enough to the exact solution by an ``educated
- guess''. One can easily imagine that the used methods then also depend
+ that one can implicitly use the analytical knowledge to obtain 
+better predictions for the location of a minimum. One can easily imagine 
+that the used methods then also depend
 on the region of the parameter space. In this manual, we only use
 examples with a neutrino factory, since some of these complications
- can be illustrated there.
-
-Now, some words of warning with respect to the minimization process: 
-Since one can easily find the global fit minimum at the best-fit/simulated values, any additional solution found by the local minimizer makes the measurement performance worse.
-However, for the real result, the complete
-topology has to be taken into account, which means that {\em all} results
-fitting the simulated data have to be included, and not only the best-fit
-point.
-Thus, one can only run the danger to obtain a too optimistic solution if one does not find the other local minima appearing below the chosen confidence level.
-Thus, with this approach and proper usage, it should not possible to produce a too pessimistic solution.
+can be illustrated there. Albeit the methods described here are neither
+complete nor will they work everywhere in the parameter space. It is
+in any case up to the user to make sure that the results are what he/she
+thinks.
+
+Some more words of warning with respect to results obtained by projecting
+the $\chi^2$:
+The results obtained this way are always only a \emph{upper} bound on the
+value of the projected $\chi^2$ function, \ie\ an undiscovered minimum 
+decreases the value of the the projected $\chi^2$ function. If the value
+of the $\chi^2$ function in the missed minimum is larger than the
+previously found ones it will not influence the value of the projected 
+value. 
+Thus, one can only run the danger to obtain a too optimistic solution 
+if one does not find the other local minima appearing 
+below the chosen confidence level. Thus, with this approach and proper usage,
+ it should not possible to produce a too pessimistic solution. 
 However, if one is not careful enough to locate
@@
 
-\index{norm}{Priors} \index{norm}{Input errors} \index{norm}{Starting values}
-In many cases, the fit manifold is restricted by the knowledge from earlier experiments.
-For example, the knowledge on the solar parameters will in most cases be supplied by the solar neutrino experiments.
-If the external precision of a parameter is at the time of the measurement better than the one of the experiment itself, one has to impose some external knowledge on this parameter.
-This external knowledge may reduce the $n$-dimensional fit manifold in the respective direction.
-In the most extreme case, keeping all parameters but the measured one fixed in the analysis means that all parameters are determined externally with infinitively high precisions.
-Thus, using the projection method on the axis/plane of interest is a reasonable approach.
-The inclusion of external input in \GLOBES\ is done by the use of Gaussian {\em priors}: We assume that an external measurement has determined the measured parameter to be at the central value (which we call {\em starting value}) with a $1 \sigma$ Gaussian error (which we call {\em input error}).
-The explicit definition of these priors will be shown in the next section.
+\index{norm}{Priors} 
+\index{norm}{Input errors} 
+\index{norm}{Starting values}
+In many cases, the fit manifold is restricted by the knowledge 
+from earlier experiments. For example, the knowledge on the solar 
+parameters will in most cases be supplied by the solar neutrino experiments. 
+If the external precision of a parameter is at the time of the 
+measurement better than the one of the experiment itself, one usually will use
+this external, better knowledge and impose a corresponding constraint
+ on this parameter. This external knowledge may reduce the extension of the 
+$n$-dimensional fit manifold in the respective direction. In the most 
+extreme case, keeping all parameters but the measured one fixed 
+in the analysis is equivalent to the assumption that all parameters 
+are determined externally with infinitively high precisions. As this 
+is a quite strong assumption, one should \emph{always} check the 
+consequences of relaxing it and using realistic errors. Only if such
+a test has demonstrated that the impact of the uncertainty on a 
+given fit parameter is negligible it can be assumed as fixed safely. 
+ The inclusion of external input in \GLOBES\ is done by the use of 
+Gaussian {\em priors}: We assume that an external measurement has 
+determined the measured parameter to be at the central value 
+(which we call {\em starting value}) with a $1 \sigma$ Gaussian error 
+(which we call {\em input error}). The explicit definition of these 
+priors will be shown in the next section.
 
@@
 \index{norm}{External input}
-It is one of the strengths of the \GLOBES\ software to use external input in order to reduce the fit manifold with the knowledge from external
-(earlier) measurements.
-The treatment of external input is done by the addition of Gaussian so-called {\em priors} to the systematics-minimized $\chi^2$-function.
-For example, for the matter density, one obtains as the final projected $\chi^2_F$ after minimization over the matter density
-scaling factor \index{norm}{Matter density scaling factor} $\hat{\rho}$
+It is one of the strengths of the \GLOBES\ software to use external input 
+in order to reduce the extension of the fit manifold with the knowledge 
+from external
+(earlier) measurements. The treatment of external input is done by the 
+addition of Gaussian so-called {\em priors} to the systematics-minimized 
+$\chi^2$-function. For example, for the matter density, one obtains as the 
+final projected $\chi^2_F$ after minimization over the matter density
+scaling factor \index{norm}{Matter density!scaling factor} $\hat{\rho}$
 \be
- \chi^2_F = \underset{\rho}{\mathrm{min}} \left( \chi^2(\hat{\rho}) + \frac{(\hat{\rho} - \hat{\rho}^0)^2}{\sigma_{\hat{\rho}}^2} \right).
+ \chi^2_F = \underset{\rho}{\mathrm{min}} \left( \chi^2(\hat{\rho}) +
+ \frac{(\hat{\rho} - \hat{\rho}^0)^2}{\sigma_{\hat{\rho}}^2} \right).
 \label{equ:priors}
@@
 This example is a very simple one, since in fact the
-minimization is simultaneously performed over all priors and free oscillation parameters.
-In \equ{priors}, $\hat{\rho}^0$ is the {\em starting value} of the prior, and $\sigma_{\hat{\rho}}$ the $1 \sigma$ absolute (half width) {\em input error}.
-Thus, it is assumed that an external measurement has determined the matter density with a precision (input error) $\sigma_{\hat{\rho}}$ at the central value $\hat{\rho}^0$.
-Usually, the starting value is fixed at the best-fit value, and the input error to the $1 \sigma$ half width of the external measurement.
-For the matter density, $\hat{\rho}^0$  is usually set to $1.0$ corresponding to
+minimization is simultaneously performed over all priors and free oscillation 
+parameters. In \equ{priors}, $\hat{\rho}^0$ is the {\em starting value} of the 
+prior, and $\sigma_{\hat{\rho}}$ the $1 \sigma$ absolute (half width) 
+{\em input error}. Thus, it is assumed that an external measurement has 
+determined the matter density with a precision (input error) 
+$\sigma_{\hat{\rho}}$ at the central value $\hat{\rho}^0$. Usually, 
+the starting value is fixed at the best-fit value, and the input error 
+to the $1 \sigma$ half width of the external measurement. For the matter 
+density, $\hat{\rho}^0$  is usually set to $1.0$ corresponding to
 the actual matter density profile such as given by the experiment
@@
 precisions, one can omit the respective input errors. In \GLOBES , a
-value of $0$ corresponds to neglecting the prior.
-If, however, earlier external measurements provide better information, one can set their absolute precisions with the input errors.
-The starting values are usually chosen to be the best-fit values of this external input, such as
-for the precisions of the solar experiments. In some cases, it may be
+value of $0$ corresponds\footnote{to be precise, a value for the error
+in between $-10^{-12}$ and $+10^{-12}$} to neglecting the prior. If, however,
+ earlier external measurements provide better information, one can set their 
+absolute precisions with the input errors. The starting values are usually 
+chosen to be the best-fit values of this external experiments, such as
+for the input from solar experiments. In some cases, it may be
 necessary to adjust them, such as for artificial constraints to the
-oscillation parameters. In other cases, minor modifications of the starting values can cause a faster convergence of the algorithm.
+oscillation parameters. In other cases, minor modifications of the 
+starting values can cause a faster convergence of the algorithm.
 For example, for the investigation of the
 opposite-sign solution, one can use the prior to constrain $\ldm$
-in order to force the minimizer not to  fall into the (unwanted) best-fit solution.
-In this case, the starting value of $\ldm$ would be set to $\rho^0_{\ldm} = - \ldm$, and a $\sigma_{\ldm}$ of the order of $\ldm$ would be imposed.
-For the algorithm, it would then be rather difficult
-to converge into the unwanted best-fit solution.
-However, note that one should in this case check that the actually determined value for $\ldm$ after minimization is close enough to the guessed value $-\ldm$ in order to avoid significant artifical contributions of the priors to the
-final $\chi^2$.
-
-\index{norm}{Input errors} \index{norm}{Starting values}
+in order to force the minimizer not to  fall into the (unwanted) 
+true-sign solution. In this case, the starting value of $\ldm$ would 
+be set to $\rho^0_{\ldm} = - \ldm$, and a $\sigma_{\ldm}$ of the order 
+of $\ldm$ would be imposed. For the algorithm, it would 
+then be rather difficult 
+to converge into the unwanted true-sign solution. However, note that 
+one should in this case check that the actually determined value 
+for $\ldm$ after minimization is close enough to the guessed value $-\ldm$ 
+in order to avoid significant artifical contributions of the priors to the 
+final $\chi^2$. Alternatively one could re-run the minimizer with the
+position of the previously found minimum as starting position but now
+with switching off the constraint on $\ldm$.
+
+\index{norm}{Input errors} 
+\index{norm}{Starting values}
 In order to set the starting values and input errors, two function have to
@@
 Eventually, a typical initialization of the external input with
-$10\%$ external precisions for the solar parameters\footnote{In fact, accelerator-based long-baseline experiments are primarily sensitive to the product $\sin 2 \theta_{12} \cdot \sdm$, which means that these errors effectively add up to an error of this product of about $15\%$.},
+$10\%$ external precisions for the solar 
+parameters\footnote{In fact, accelerator-based long-baseline experiments 
+are primarily sensitive to the product $\sin 2 \theta_{12} \cdot \sdm$, 
+which means that these errors effectively add up to an error of this 
+product of about $15\%$.}, 
 and $5\%$ matter density uncertainties for all experiments looks like this:
@@
 \end{quote}
-In this example, the starting values are set to the best-fit (simulated) values. Remember that initially the matter density scaling factors are 
+In this example, the starting values are set to 
+the true (simulated) values. Remember that initially the 
+matter density scaling factors are 
 all $1.0$, which means that they do not need to be adjusted for the
@@
 \index{norm}{Advanced tricks}
-Though the priors are an elegant way to treat external input, there are also some complications with priors. The following hints are for the
+Though the priors are an elegant way to treat external input, 
+there are also some complications with priors. The following hints are for the
 more advanced \GLOBES\ user:
@@
  the absolute minimum $\chi^2$ is larger than zero, \ie, degeneracies
- other than the best-fit solution. In this case, the sections and
+ other than the true solution. In this case, the sections and
  projections are not  comparable if not corrected by the prior contributions, where the
@@
  larger than the projection region if the priors are not added to the
- section. At the best-fit solution, this problem usually does not occur
+ section. At the true solution, this problem usually does not occur
  because the prior contributions are close to zero.    
@@
 \label{ex:corrproj}
-\index{norm}{Projection of manifold}
-\index{norm}{Two-parameter correlation}
-\index{norm}{Multi-parameter correlation}
+\index{norm}{Projection!of manifold}
+\index{norm}{Correlation!two-parameter}
+\index{norm}{Correlation!multi-parameter}
 
@@
 }
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section[Projection onto the $\stheta$-axis or $\deltacp$-axis]{Projection onto the $\boldsymbol{\stheta}$- or $\boldsymbol{\deltacp}$-axis}
-\index{norm}{Projection onto axis}
+\index{norm}{Projection!axis}
 
@@
 the local minimum, where the fit parameter of the projection
-axis is fixed. For the best-fit solution, it is usually sufficient to start the minimizer at the best-fit point. However, the convergence speed might
+axis is fixed. For the true solution, it is usually sufficient to start 
+the minimizer at the true parameter values. However, the convergence 
+speed might
 be better by starting it slightly off this point. In addition, there
@@
 
-\index{norm}{Matter density scaling factor}
+\index{norm}{Matter density!scaling factor}
 The output of the minimizer in {\tt out} carries as many matter density scaling factors as there are experiments. Either
@@
 all experiments) are different from $1.0$ if matter density uncertainties
-are present, since each experiment may face other matter density conditions. The minimizers of individual experiments ``know'' which
-experiment they are currently treating, which means that only the returned
-matter density scaling factor of the appropriate experiment is modified.
+are present, since each experiment may face other matter density conditions. 
+The minimizers of individual experiments ``know'' which
+experiment they are currently treating, which means that they only returned the
+matter density scaling factor of the appropriate experiment.
 For example, calculating {\tt glbChiTheta} for the last experiment
-number, the last density value will be modified.
+number, the last density value will be returned.
 This approach turns out to be extremely useful together with the
@@
 for neutrino factories because of the $(\deltacp, \theta_{13})$-degeneracy, whereas it is for superbeams often sufficient
-to use the best-fit values. A strong indication for a wrong guessing 
+to use the true values. A strong indication for a wrong guessing 
 are discontinuous jumps in the projected $\chi^2$-function, where the minimizer jumps from one minimum to another.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section[Projection onto any hyperplane]{Projection onto any hyperplane}
-\index{norm}{Projection onto hyperplane}
+\index{norm}{Projection!hyperplane}
 
@@
 minimized over. 
-The corresponding function is \index{norm}{Projection onto $\theta_{13}$-$\deltacp$-plane}
+The corresponding function is 
+\index{norm}{Projection!$\theta_{13}$-$\deltacp$-plane}
 \begin{function}
@@
 \hline
-\GLB{AllocProjection} & Create projection vector & {\tt ()} \\
-\GLB{FreeProjection} & Destroy projection vector {\tt stale} & {\tt (glb\_projection stale)} \\
+\GLB{AllocProjection} & Allocate projection vector & {\tt ()} \\
+\GLB{FreeProjection} & Free projection vector {\tt stale} & {\tt (glb\_projection stale)} \\
 \GLB{DefineProjection} & Assign projection vector {\tt in} & {\tt (glb\_projection in, int theta12, int theta13, int theta23, int delta, int dms, int dma)} \\ 
@@
 fixed and all the other parameters free. Similar to {\tt glb\_params},
-the type {\tt glb\_projection} has to be created first, and destroyed
+the type {\tt glb\_projection} has to be allocated first, and freed
 later. The access functions for {\tt glb\_projection} are summarized in \Tab~\ref{tab:defprojection}. Since the complete set is very similar to
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \chapter{Locating degenerate solutions}
@@
 The example on page~\pageref{ex:sgndeg} illustrates how to locate the $\mathrm{sgn}(\ldm)$-degeneracy and show the corresponding degenerate solution in the $\stheta$-$\deltacp$-plane together with the original solution.
-In this case, the position of the degeneracy can be easily guessed to be at the best-fit parameter values but the inverted $\ldm$.
+In this case, the position of the degeneracy can be easily guessed to be 
+at the true parameter values but with inverted $\ldm$.
 The minimizer then runs off the plane of these parameters into the local minimum.
-It is very important to take into account the position of the degeneracy off this plane, since the actual $\chi^2$ in the minimum is probably lower than on the plane of the best-fit parameter values.
-Thus, the degeneracy may not even appear at the chosen confidence level on the plane, but it does appear at the true minimum.
+It is very important to take into account the position of the degeneracy off this plane, since the actual $\chi^2$ in the minimum is probably lower than in the plane.
+Thus, the degeneracy may not even appear at the chosen confidence level in the plane, but it does appear at the real minimum.
 The two sections through the fit manifold shown in the figure on page~\pageref{ex:sgndeg} therefore do not appear at the same oscillation parameter values (except from the ones shown in the
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \chapter{Obtaining low-level information}
@@
 evaluated experiment. 
+\end{function}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{\AEDL\ names}
+\index{norm}{AEDL!names}
+\label{sec:aedl_names}
+
+Since in \AEDL\ rules, cross section, fluxes etc. carry a `name' by
+which they can be refered to and in C they cary only an integer as
+index it is sometimes difficult to figure out the correct correspondence.
+Therefore the information about this correspondence obtained during
+parsing is stored and can be accessed within C by these two functions.
+\begin{function}
+\GLBNS{NameToValue}
+{\tt int glbNameToValue(int exp, const char* context, const char *name)}
+Converts an \AEDL\ name given as argument {\tt name} into the corresponding
+C index. The variable {\tt context} describes wether this name belongs
+to a rule, channel, flux, energy or cross type environment. {\tt exp}
+is the number of the experiment and can not be {\tt GLB\_ALL}. It returns
+either the index in case of success or {\tt -1} in case the name was not
+found. 
+\end{function}
+
+\begin{function}
+\GLBNS{ValueToName}
+{\tt const char *glbValueToName(int exp,const char* context, int value)}
+Converts a C index given as argument {\tt value} into the corresponding 
+\AEDL\ name. The variable {\tt context} describes wether the index belongs
+to a rule, channel, flux, energy or cross type environment. {\tt exp}
+is the number of the experiment and can not be {\tt GLB\_ALL}. It returns
+either the name in case of success or {\tt NULL} in case the name was not
+found. The returned string must not be modified.
 \end{function}
@@
 \end{function}
-Note that all flux or flavor numbers run from $0$ to $N-1$ in these functions.
-
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \chapter{Changing experiment parameters at running time}
@@
 \section{Baseline and matter density profile}
-\index{norm}{Baseline: Change}
-\index{norm}{Matter density profile: Change}
+\index{norm}{Baseline!change}
+\index{norm}{Matter density!change profile}
 
@@
 where the average density or the PREM profile are re-computed, or the
-steps in the arbitrary profile are re-scaled. Of course, though this
-function is sufficient in most cases, it does often not make sense
-for arbitrary matter density profiles.
+steps in the arbitrary profile are re-scaled. If this behaviour is 
+not the desired one, one has to use {\tt glbSetProfilDataInExperiment}
+as explained below.
 \begin{function}
@@
 during running time. However, one can change the matter density
-profile, where the profile type is automatically changed to three
-(arbitrary matter density profile). In addition, a number of functions 
+profile, where the profile type is automatically changed to 3, \ie\ 
+arbitrary matter density profile. In addition, a number of functions 
 are provided to compute possible matter density profiles (average density,
@@
 to handle. In general, any function
-{\em returning} lists reserves the memory space for them.
-It is then up to the user to return the allocated memory!
-In addition, they normally return the length of the lists $N$
-in form of a pointer, where the corresponding memory space 
-has to be provided by the user (normally, it is enough to declare
-a variable of the type {\tt size\_t} corresponding to positive
-integer numbers).
+{\em returning} lists allocates the memory for them.
+It is then up to the user to free this memory!
+In addition, they normally provide the length of the lists $N$
+by means of an additinal argument which is a pointer to {\tt size\_t}. 
+Normally, it is enough to declare
+a variable of the type {\tt size\_t} and to give its address to the function.
 The following functions return matter density profiles:

### 2004-07-19 Documentation/Manual.tex
Editing done ;-)

@@
 
-\input{tumlogo}
-
-\begin{minipage}{7cm}
-\begin{center}
-{\bf Technische Universit{\"a}t M{\"u}nchen} \\
-Physik-Department \\
-Institut f{\"u}r Theoretische Physik, T30d \\
-\end{center}
-\end{minipage}
-
-\vspace{-2cm}
-
-\hfill \oTUM{3.5cm}
-
-\vspace{3cm}
+\vspace*{3cm}
 
@@
 { \Large \bf
-\GLOBES\ 1.X \\
+\GLOBES\ 2.0 \\
 General Long Baseline Experiment Simulator \\ }
@@
 to produce a publication or a talk indicate that you have used \GLOBES\ and
-please cite the following reference
+please cite the following reference~\cite{globes_paper}
 \begin{quote}
@@
 with it should be correctly referenced where possible. For details see the
-???.
+corresponding sub-section of the bibliography.
 
@@
 
+\addcontentsline{toc}{chapter}{Bibliography}
 \bibliographystyle{./unsrt}
 \bibliography{references}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Referencing the data in \GLOBES}
+\label{ref_data}
+
+\index{norm}{Referencing!data in \GLOBES}
+\GLOBES\ wouldn't be of much use without having high quality input data.
+Much of this input data has been published somewhere and the authors
+of those publication would appreciate to be cited by the user
+when he/she is using their work. It is solely the user's responsibility 
+to make sure that he/she understand where a certain input for \GLOBES\ comes
+from and wether additional work has to be cited in addition to the 
+\GLOBES\ paper~\cite{globes_paper}. To help this task we provide here the
+necessary information for the data which is shiped together with \GLOBES.
+
+When using the built-in earth matter profile, the original source is 
+reference~\cite{Stacey}.
+
+All files ending with \verb+.dat+ or \verb+.glb+ 
+in the \verb+data+ subdirectory of the \GLOBES\ tar-ball have at their
+beginning a comment field which clearly indicates which works should be
+cited when using a certain file. Make sure that dependencies are correctly
+tracked, \ie\ in some cases files included by other files have to checked
+also. One can use the \verb+-v3+ option to \verb+globes+ to see which files
+are included.
+
+It is advisable to use the same style for your own input files, since in
+case they are distributed, everbody will know how to correctly reference your
+work.
 

### 2004-07-19 Documentation/appendix.tex
part 1 finis
Editing done ;-)

@@
 the rather difficult nature of the build process of BLAS and LAPACK,
-it is recommended that one use a suitable rpm instead. In addition,
+it is recommended that one uses a suitable rpm instead. In addition,
 there is a good chance that
@@
 The install target also will install a program with name \verb+globes+ to
-\verb+$prefix+.
+\verb+$prefix/bin+.
 
@@
 In general it is advisable to use the shared libraries. If you don't
-have root privileges, see the corresponding section of this file.
+have root privileges, see the corresponding section 
+(page~\pageref{inst_noroot}).
 If you install \GLOBES\ with root privileges, do not forget to run 
@@
 \index{norm}{Installation!w/o root privilege}
+\label{inst_noroot}
 
@@
 
-\chapter*{Acknowledgments}
-\addcontentsline{toc}{chapter}{Acknowledgments}
-
 \end{appendix}

### 2004-07-19 Documentation/part1.tex
Editing done ;-)

@@
 \label{chapter:tour}
-\index{norm}{GLoBES tour}
+\index{norm}{GLOBES tour@\GLOBES~tour}
 
@@
 \mycaption{\label{tab:experiments} \index{norm}{Experiment files (table)}
-\index{norm}{{\tt glb}-files} 
+\index{norm}{glb files@{\tt glb}-files} 
 Pre-defined experiment prototypes, their filenames (to be used in \GLB{InitExperiment}), their short descriptions, and the references in which they are originally used and discussed (except from minor modifications, such as a
@@
 
-\index{norm}{Clear experiment list}
-\index{norm}{Delete experiments}
+\index{norm}{Experiment!list!clear}
+\index{norm}{Experiment!delete}
 One can also remove all experiments from the evaluation list at running
@@
 \begin{function}
-\GLBNS{SetIteration} \index{norm}{Iterations}
+\GLBNS{SetIteration} \index{norm}{Minimizer!iterations}
 {\tt glb\_params glbSetIteration(glb\_params in, int iter)} sets the number of iterations in the structure {\tt in}
@@
 \chapter[Calculating $\chi^2$ with systematics only]{Calculating $\boldsymbol{\chi^2}$ with systematics only}
-\index{norm}{Systematics $\chi^2$}
+\index{norm}{Systematics!$\chi^2$}
 
@@
 \label{ex:corrth13dcp}
-\index{norm}{Two-parameter correlation}
+\index{norm}{Correlation!two-parameter}
 A typical and fast application for \GLB{ChiSys} is the visualization of
@@
 {\em pull method} with the help of auxiliary systematics parameters.
-They are taken completely uncorrelated among different rules, and treated with simple Gaussian statistics. In general, a rule is a set 
+They are taken completely uncorrelated among different rules, and treated with simple Gau\ss ian statistics. In general, a rule is a set 
 of signal and background event rates coming from different oscillation
@@
 \label{chapt:correlations}
-\index{norm}{Correlations $\chi^2$}
-
-\index{norm}{Correlations!multi-parameter}
+\index{norm}{Correlation!and $\Delta\chi^2$}
+
+\index{norm}{Correlation!multi-parameter}
 This chapter deals with the rather complicated issue of $n$-parameter correlations. It is one of the greatest strengths of this software 
@@
 
-\index{norm}{Priors} 
-\index{norm}{Input errors} 
-\index{norm}{Starting values}
+\index{norm}{External information!priors} 
+\index{norm}{External information!input errors} 
+\index{norm}{External information!starting values}
 In many cases, the fit manifold is restricted by the knowledge 
@@
  The inclusion of external input in \GLOBES\ is done by the use of 
-Gaussian {\em priors}: We assume that an external measurement has 
+Gau\ss ian {\em priors}: We assume that an external measurement has 
 determined the measured parameter to be at the central value 
-(which we call {\em starting value}) with a $1 \sigma$ Gaussian error 
+(which we call {\em starting value}) with a $1 \sigma$ Gau\ss ian error 
 (which we call {\em input error}). The explicit definition of these 
@@
 
-\index{norm}{External input}
+\index{norm}{External information}
+\index{norm}{External input| \see{External information}}
 It is one of the strengths of the \GLOBES\ software to use external input 
@@
 (earlier) measurements. The treatment of external input is done by the 
-addition of Gaussian so-called {\em priors} to the systematics-minimized 
+addition of Gau\ss ian so-called {\em priors} to the systematics-minimized 
 $\chi^2$-function. For example, for the matter density, one obtains as the 
@@
 
-\index{norm}{Input errors} 
-\index{norm}{Starting values}
+\index{norm}{External information!input errors} 
+\index{norm}{External information!starting values}
 In order to set the starting values and input errors, two function have to
@@
 
-\index{norm}{External precision}
+\index{norm}{External information!precision}
 Eventually, a typical initialization of the external input with
@@
 
-\index{norm}{Priors}
+\index{norm}{Minimizer!priors}
+\index{norm}{External information!priors}
 \index{norm}{Advanced tricks}
@@
 \item
-Currently, \GLOBES\ only supports Gaussian priors for the individual
+Currently, \GLOBES\ only supports Gau\ss ian priors for the individual
 oscillation parameters. Especially for the solar parameters, this is only
@@
 \end{tabular}
-\caption{\label{tab:defprojection}\index{norm}{Projection type}
+\caption{\label{tab:defprojection}\index{norm}{Projection!type}
 Different functions handling the
@@
 
-\index{norm}{Projection: Define}
+\index{norm}{Projection! definition}
 As soon as we have defined a projection, we can assign it:
@@
 \chapter{Locating degenerate solutions}
-\index{norm}{Degenerate solutions}
-\index{norm}{Degeneracies $\chi^2$}
+\index{norm}{Degeneracies|(}
+\index{norm}{Degeneracies!multiple solutions}
+\index{norm}{Degeneracies!and $\Delta\chi^2$}
 
@@
 \label{ex:sgndeg}
-\index{norm}{Degeneracies: sgn($\ldm$)-degeneracy}
+\index{norm}{Degeneracies!sgn($\ldm$)-degeneracy}
 
@@
 
-\index{norm}{All-parameter minimization}
+\index{norm}{Minimization!all-parameter}
 The function to perform the all-parameter minimization is {\tt glbChiAll}:
@@
 Thus, one should be careful with the determination of the degenerate solutions in order to find all of them.
+\index{norm}{Degeneracies|)}
 
@@
 \section{\AEDL\ names}
-\index{norm}{AEDL!names}
+\index{norm}{AEDL@\AEDL!names}
 \label{sec:aedl_names}
@@
 \label{ex:barcharts}
-\index{norm}{Bar charts}
-\index{norm}{Systematics on/off}
+\index{norm}{Bar plots}
+\index{norm}{Systematics!on/off}
 Here it is demonstrated how one can successively include systematics, correlations, and degeneracies at the example of the $\stheta$-sensitivity limit.
@@
 \index{norm}{Systematics}
-\index{norm}{Systematics on/off}
+\index{norm}{Systematics!on/off}
 
@@
 
-\index{norm}{Signal errors}
-\index{norm}{Background errors}
-\index{norm}{Background centers}
+\index{norm}{Signal!errors}
+\index{norm}{Background!errors}
+\index{norm}{Background!centers}
 Except from the general treatment of systematics, one can read out
@@
 \label{sec:aedlparams}
-\index{norm}{External \AEDL\ parameters}
+\index{norm}{AEDL@\AEDL!external parameters}
 
@@
 \section{Algorithm parameters: Filter functions}
-\index{norm}{Filter functions}
+\index{norm}{Filter!functions}
 

### 2004-07-19 Documentation/part2.tex
Editing done ;-)

@@
 
+\index{norm}{AEDL@\AEDL|(}
 \part{The Abstract Experiment Definition Language -- AEDL}
@@
 representation of very different setups within one data structure, and thus implements universal rate and $\chi^2$ computation methods.
-For experiment simulations, a new piece of code is usually written and compiled
+For experiment simulations, usually a new piece of code is written and compiled
 for each different experiment. In many cases, even parameter changes, such as
-the number of bins, requires the recompilation of the source code. 
+the number of bins, require the recompilation of the source code. 
 However, such a technique soon reaches its limits when the simulated experiments are rather complex, or more than one type of experiment is studied simultaneously.
@@
 \item
- Sources with a physically significant time dependencies  can not be studied, such as  supernov\ae. It is, however, possible
+ Sources with a physically significant time dependency  can not be studied, such as  supernov\ae. It is, however, possible
 to study beams with bunch structure, since the time dependence of the
@@
 The description of the neutrino oscillation physics is, at least numerically, relatively simple.
-We use the {\em evolution operator method} to compute the neutrino oscillation probabilities and divide the matter density profile into layers with constant matter densities.
+We use the {\em evolution operator method}~\cite{Ohlsson:1999um}  to compute the neutrino oscillation probabilities and divide the matter density profile into layers of constant matter density.
 For each of these layers, the Hamiltonian in matter is diagonalized in order to propagate the neutrino transition amplitudes.
@@
 Currently, a number of pre-defined experiment definition files are provided with \GLOBES , which have to be modified manually in order to define new experiments.
-An IDE (Integrated Developement Environment) for AEDL files may follow in one of the next
-versions.
 The application software then uses this text file to initialize the experiment, where other secondary files might read for source fluxes, cross sections \etc .
@@
 First, we instruct \GLOBES\ to use the built-in source flux for a neutrino factory
-originating from stored $\mu^+$'s. Therefore, we set the {\tt @builtin} variable to $1$. Next, we specify the muon energy to be $50\,\mathrm{GeV}$ by the {\tt @parent\_energy} variable. We assume 
+originating from stored $\mu^+$'s.
+This achieved by setting the {\tt @builtin} variable to $1$.
+Next, we specify the muon energy to be $50\,\mathrm{GeV}$ by the {\tt @parent\_energy} variable.
+We assume
 that there will be $5.33\cdot 10^{20}$ useful muon decays per year
@@
 {\tt \tb @backgrounderror = 0.05 : 0.0001}\\
-{\tt \tb @errordim = 0}\\
+{\tt \tb @errordim\_sys\_on = 0}\\
+{\tt \tb @errordim\_sys\_off = 2}\\
 {\tt \tb @energy\_window = 4.0 : 50.0}\\
@@
 The background error variable is defined such as the signal error variable, \ie\ we have a $5\%$ background uncertainty and a very small energy calibration error.
-The ``error dimension variable'' {\tt @errordim} selects how the systematical errors are treated (\cf, \Tab~\ref{tab:error_dim}).
+The ``error dimension variable'' {\tt @errordim\_sys\_X} selects how the systematical errors are treated (\cf, \Tab~\ref{tab:error_dim}).
 
@@
 ``automatic variables'' and have to start with {\tt \#}. Note that these names have to be unique and can only be refered to after their definition.
-However, similar to C, one can give a header definition before:
+However, similar to C, one can give a declaration without definition before:
 \begin{quote}
@@
 variables are numbered in the order of their definition, and the number
-can later be used to refer to them in the C code (from $0$ to {\tt numof...}$-1$). Within each environment type, there are several 
+can later be used to refer to them in the C code (from $0$ to {\tt numof...}$-1$). In order to facilitate the the mapping from names in AEDL to indices
+in C there are two functions \GLB{NameToValue} and \GLB{ValueToName} which
+make this transition (see \Sec~\ref{sec:aedl_names}, 
+page~\pageref{sec:aedl_names}).
+
+Within each environment type, there are several 
 variables beginning with {\tt @}, which can only be used within the 
@@
 \end{quote}
-This command resets the counters for channels, rules, fluxes, and energy resolution environments. All variables have their scope limited 
+This command resets the counters for channels, rules, fluxes, cross section 
+and energy resolution environments. All variables have their scope limited 
 by either {\tt \%!GLoBES, \#NEXT\#} or {\tt EOF}.  This allows 
@@
 Note that, alternatively, one can 
-also load both files successively.
+also load both files successively by two separate calls of 
+\GLB{InitExperiment}.
 
@@
 substitution mechanism as described in~\Sec~\ref{sec:aedlparams}.
-\index{norm}{External \AEDL\ parameters}
+\index{norm}{AEDL@\AEDL!external parameters}
 
@@
 As we have disussed before, \GLOBES\ can only deal with point sources.
-Thus,  it is not possible to study effects from the finite size of the neutrino production region, such as in the Sun or in reactor experiments with many
+Thus,  it is not possible to study effects from the finite size of the neutrino production region, such as in the sun or in reactor experiments with many
 neutrino sources (\eg, KamLAND).
@@
 Correspondingly, the detector mass units should be kilotons for a proton-based beam, and tons for a reactor experiment.
+In any case it is a good
+idea to document the choices made by the user by corresponding comments
+in \AEDL.
 
@@
 
-\index{aedl}{{\tt flux}!flux file@{\tt "@flux\_file}}
+\index{aedl}{{\tt flux}!flux file@{\tt "@flux\_file}}\index{norm}{Flux!file}
 The software assumes that the given flux file has seven columns and
@@
 However, if the energy leaves the range of values given in the file, zero is returned.
-Note that unused (and therefore often unknown) fluxes can be initialized with zeros, but can not just be omitted.
-For example, if one wants to define an experiment with different running times with neutrinos and antineutrinos, one can just use the same flux file in two different {\tt flux} environments with two different running times {\tt @time}.
-The unused fluxes in each case will then just be ignored.
-
-% Continue here (WW)
+The the columns for the fluxes for unused flavours have to be filled all the 
+same,\eg\ with zeros.
+
+The flux files accept one-line comments, which start
+with {\tt \#} and end with the linefeed character `\verb+\n+', they are
+not counted as a line and their content is discarded. This comments
+are useful to provide meta information about the fluxes like units
+or the origin of the information. This is also the default method
+to point the user to the references he/she should to cite when using
+a particular flux.\index{norm}{Flux!file!comments in}
+\index{norm}{Referencing!flux data}
 
@@
 has to be specified.
-The simplest matter density profile is a constant matter density profile with the average matter density from the PREM\cite{Stacey} onion shell model of the Earth\index{norm}{PREM}\index{norm}{Earth matter density}:
+The simplest matter density profile is a constant matter density profile with the average matter density from the PREM~\cite{Stacey} onion shell model of the earth\index{norm}{PREM| \see{Matter density}}
+\index{norm}{Matter density!of the earth}:
 \begin{quote}
@@
 %
-For a better approximation of the realistic Earth matter density profile, one can use an arbitrary number of equidistant steps of the PREM profile:
+If your using this option please cite reference~\cite{Stacey}\index{norm}{Referencing!matter profile data}.
+
+For a better approximation of the realistic earth matter density profile, one can use an arbitrary number of equidistant steps of the PREM profile:
 \begin{quote}
@@
 \end{quote}
-It is important that both lists habe the same length and that the  thicknesses given in  {\tt \$lengthtab} add up to the length of
+It is important that both lists have the same length and that the  thicknesses given in  {\tt \$lengthtab} add up to the length of
 the baseline, which does not have to be explicitely specified anymore. In addition, matter densities are always given in $g/cm^3$.
@@
 
-Cross sections\index{norm}{Cross sections} will later be used as part of the 
+Cross sections\index{norm}{Cross section} will later be used as part of the 
 channel definition (see \Sec~\ref{sec:channel}). Similar to the source 
@@
 seven columns and $1001$ lines of the form
+\index{norm}{Cross section!file}
 \index{aedl}{{\tt cross}!cross file@{\tt  "@cross\_file}}
@@
 
+Like the flux files, the cross section files 
+accept one-line comments, which start
+with {\tt \#} and end with the linefeed character `\verb+\n+', they are
+not counted as a line and their content is discarded. This comments
+are useful to provide meta information about the cross sections like units
+or the origin of the information. This is also the default method
+to point the user to the references he/she should to cite when using
+a particular cross section.\index{norm}{Cross section!file!comments in}
+\index{norm}{Referencing!cross section data}
+
 \section{Oscillation channels}
@@
 Therefore, a channel contains the description of the 
-initial neutrino flavor, its CP eigenvalue (neutrino or antineutrino), 
+initial neutrino flavor, its CP eigenvalue 
+(neutrino or antineutrino)\footnote{Currently \GLOBES\ does not support lepton
+number violating transitions, \ie\ no transitions from neutrino to 
+antineutrino (or vice versa) are considered.}, 
 the detected neutrino flavor, the interaction cross sections for the chosen interaction type, and the energy resolution function of the detector.
@@
 $R_\beta^{\text{IT}}(E,E') dE'$. The function $R(E,E')$ is also often called ``energy resolution function''. Actually, its internal representation
-in the software is a smearing matrix\index{norm}{Energy resolution}\index{norm}{Smear matrix}.
+in the software is a smearing matrix\index{norm}{Energy!resolution}\index{norm}{Smear matrix}.
 The function $\epsilon_\beta^{\text{IT}}(E')$ will later be refered to as ``post-smearing efficiencies'', since it will allow us to define cuts and threshold functions {\em after} the smearing is performed, \ie, as function of $E'$.
@@
 Usually it is an excellent approximation to treat backgrounds as if they
-were not affected by oscillations.
+were not affected by oscillations\footnote{In case, the backgrounds have
+a sizeable dependence on the oscillation parameters they carry information
+on the oscillation parameters and are therefore more like a signal.}.
 
@@
 \index{aedl}{{\tt energy}|(}
+\index{norm}{Energy!resolution|(}
 The definition and implementation of the energy resolution function is 
-rather sophisticated in \GLOBES . In particular, since this aspect is 
- time-critical, several algorithms are implemented. The choice of the 
-proper algorithm depends on the experiment and the frequencies of the
- involved neutrino oscillations.
+rather sophisticated in \GLOBES . In particular, the choice of the 
+proper parameters depends on the experiment and the frequencies of the
+ involved neutrino oscillations. This choice also greatly influences the
+speed of the calculation.
 
@@
 \label{eq:post_smearing}
-\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' \, R^c(E,E') \, \epsilon^c(E') = \hat\epsilon_i^c \cdot \int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
+\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' \, R^c(E,E') \, \epsilon^c(E')
+ \simeq \hat\epsilon_i^c \cdot \int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
 \quad R^c(E,E')\,.
@@
 is very slow in many cases. Therefore, we will introduce two different approximation schemes for different applications in the next two subsections.
-In either case, the integrand in \eq~(\ref{eq:simple_int}) has to be evaluated at fixed ``sampling points''. These sampling points can be 
-either defined by the user, or be computed by
-\GLOBES . The choice between these options also depends on 
-the algorithm which is used for the calculation, as it is explained in the following subsections.
+In either case, the integrand in \eq~(\ref{eq:simple_int}) has to be evaluated at fixed ``sampling points''. These sampling points have to directly 
+or indirectly be defined by the user.
 
@@
 \item[Sampling point level]
- This level is used internally to evaluate the integrand in \eq~(\ref{eq:simple_int}) at all sampling points.
-The energy scale is the actual incident neutrino energy $E$.
-Note that the sampling range should exceed the energy bin range by about the energy resolution error, since the corresponding sampling points have to be evaluated by the algorithm.
-However, this exceeding is only
-necessary to avoid aliasing effects if the integrand in \eq~(\ref{eq:simple_int}) is large at the limits of the sampling range.
-For a manual definition of the sampling points (depends on algorithm), use
+ This level is used internally to evaluate the integrand in \eq~(\ref{eq:simple_int}) at all sampling points. The energy scale is the actual incident neutrino energy $E$. 
+For a manual definition of the sampling points, use
 \index{aedl}{sampling points@{\tt \$sampling\_points}}
@@
 \end{quote}
-for equidistant sampling points. Arbitrarily spaced sampling points can 
+for equidistant sampling points. If no values are given for these variables
+they are assumed to be equal to their corresponding counter part at bin level,
+\ie\ \verb+$sampling_points = $bins+, \verb+$sampling_min = $emin+ and
+\verb+$sampling_max = $emax+.
+
+Arbitrarily spaced sampling points can 
 be specified with 
@@
 \end{quote}
-
-Note that the algorithm expects to 
-have at least as many sampling points as energy bins, since otherwise 
-the number of (unphysical) sampling constrains the experiment performance.
-\item[Bin level] This level is determined by the experiment. Note than energy
+The choice of the sampling point configuration strongly depends on the 
+experiment and required accuracy. Ideally the integrand 
+of~\eq~\ref{eq:simple_int} is zero outside the sampling range, if this cannot
+be achieved it is usually sufficient that the sampling range is by about
+three times the energy resolution\footnote{evaluated at {\tt \$emin} and
+{\tt \$emax} respectively} larger than the bin range. 
+The spacing of the sampling points should be somewhat smaller (a factor 
+$\simeq 2$ usually is more than ebough) than the 
+finest details of the integrand.
+
+\item[Bin level] This level is determined by the experiment and its analysis. 
+Note that energy
 bin sizes much smaller than the energy resolution will not improve the results.
 The energy bin range and the number of energy bins do always have to be specified.
-For the case of large values of the integrand in \eq~(\ref{eq:simple_int}) at the energy range limits, it is recommended to exceed the analysis energy window by about the energy calibration error in order to avoid aliasing effects.
+For the case of large values of the integrand in \eq~(\ref{eq:simple_int}) at the energy range limits, it is recommended to exceed the analysis energy window by about three times the energy calibration error in order
+to avoid cutoff effects.
 
@@
 bin sizes have to add up to the energy range {\tt \$emax}-{\tt \$emin}.
+
+The choices at bin level are mainly determined by optimizing the
+performance of the experiment.
+
 \item[Analysis level]
@@
 \end{description}
+
 In general, the energy smearing happens between the sampling point and bin levels, which means that the energy smearing matrix will have {\tt \$sampling\_points} columns and {\tt \$bins} rows.
@@
 If they are introduced after, we call them {\tt @post\_smearing\_efficiencies} or {\tt @post\_smearing\_background}.
-Note that pre-smearing components are always a function of the incident neutrino energy $E$.
-Thus, there have to be as  many elements as there are sampling points.
-For some algorithms, the number of sampling points is determined automatically by \GLOBES , and the number of elements in the  pre-smearing vectors has to be adjusted in the second run.
+Note that pre-smearing components are always a function of the incident neutrino energy $E$. Thus, there have to be as  many elements as there are sampling points. 
 Examples for pre-smearing quantities are non-beam backgrounds, such as from geophysical neutrinos.
@@
 of \eq~(\ref{eq:simple_int}). It is applicable to most of the
- accelerator-based experiments which can be simulated with \GLOBES .
-
-The key idea is to use a ``flat'' model with sampling points identical to 
-the bins (\cf, \figu{energies}). This is a good approximation as long as
+experiments which can be simulated with \GLOBES .
+
+The key idea is to use a ``flat'' model, \ie\ the integrand 
+of~\eq~\ref{eq:simple_int}  is well approximated by being piecewise
+constant in each sampling step. This is a good approximation as long as
 \begin{itemize}
 \item
- The number of bins is larger than the energy resolution.
+ No details are lost, \ie\ the spacing of sampling points 
+is smaller than the energy resolution.
 \item
- The integrand in \eq~(\ref{eq:simple_int}) is small at the bin range limits
-{\tt \$emin} and {\tt \$emax} (in order to avoid aliasing effects).
+The edges are treated correctly.
 \item
- The neutrino oscillations are slow on a scale of the bin distance.
+ The neutrino oscillations are slow on a scale of the smapling point distance.
 \end{itemize}
@@
 constant, \ie, they hardly change within each bin. This 
-assumption seems to be very restrictive. Note, however, if one analyzes simulated data (which are simulated with the same algorithm), the errors
-will cancel between the simulated and fitted data. 
-This algorithm is the fastest which \GLOBES\ offers, and is selected by
+assumption seems to be very restrictive, which is however not quite correct.
+First of all, if one analyzes simulated data (which are simulated with the same algorithm), the errors
+will cancel between the simulated and fitted data. Second, and more important,
+this algorithm is just a very basic integration routine\footnote{It is planed 
+for to have something like a Gau\ss-Kronrod scheme as an alternative here.} and converges to
+the true result for decreasing step size. Thus if the number of sampling
+points is large enough this algorithm is very accurate. 
+This algorithm is selected by
 %
@@
 by \GLOBES. Thus, it requires that the number of bins
-{\tt \$bins} and the minimum energy {\tt \$emin} and maximum energy {\tt \$emax} for equidistant bins are given.
+{\tt \$bins} and the minimum energy {\tt \$emin} and maximum energy {\tt \$emax} are given in case of equidistant bins.
 %
@@
 $R^c(E,E')$ in \eq~(\ref{eq:kernel}) is concerned, the algorithm uses
-the Gaussian
+a Gau\ss ian
 \begin{equation}
@@
 \end{equation}
+The somewhat complicated form is due to the fact that inverse $\beta$-decay
+has a neutrino threshold of $1.8\,\mathrm{MeV}$ and that a neutrino
+at threshold already produces $\simeq 1\,\mathrm{MeV}$ visible energy in
+the detector (for more details see \eg~\cite{Huber:2003pm}).
+
 In the actual implementation of the algorithm,  the sum in \eq~(\ref{eq:algo_one}) is only computed for the $E_j$'s where $K(E_j)$ 
@@
 In some cases, one may use the output of a detector Monte Carlo simulation
-directly. Then one can use manual energy smearing instead of one of the
+directly. Then one can use ''manual'' energy smearing instead  of the
 automatic energy smearing algorithms.
@@
 since all of the incoming neutrinos should be assigned to energy bins. In many practical cases, however, the definition of the energy smearing can
-lead to sums smaller than unity, such as in the case of truncated Gaussian
+lead to sums smaller than unity, such as in the case of truncated Gau\ss ian
 distributions. The sum of entries in each {\em row} is not defined, since the events might be unevently distributed into the energy bins
@@
 \index{aedl}{{\tt energy}|)}
-
+\index{norm}{Energy!resolution|)}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Thus, it is possible to constrain the analysis energy range 
-with each rule to an energy window\index{norm}{Energy window}:
+with each rule to an energy window\index{norm}{Energy!window}:
 \index{aedl}{{\tt rule}!energy window@{\tt "@energy\_window}}
@@
 The default energy window is given between the minimal and maximal reconstructed energy. To be on the save side, reduce analysis window
-compared to the bin range on each side by about three standard deviations 
-of the calibration error.
+compared to the bin range on each side by about three times the
+energy calibration error.
 
@@
 This information can be accessed by the versioning functions as described in~\Sec~\ref{sec:versioning}.
-
+\index{norm}{AEDL@\AEDL|)}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 This chapter demonstrates how to test an \AEDL\ file in order to check
-if they really describes given experiment. 
+if it really describes a given experiment. 
 For this application, the \GLOBES\ package contains the program
@@
 The {\tt globes}\index{norm}{globes@{\tt globes}} binary is installed 
-together with the library \index{norm}{libglobes@{\tt libglobes}}, but
+together with the library\index{norm}{libglobes@{\tt libglobes}}, but
 into the directory {\tt \$prefix/bin/}. In order to use the {\tt globes} 
-utility, this directory has to be in the path of the shell used to call the program.\footnote{This is automatically done if
+utility, this directory has to be in the path of the shell used to call the program.\footnote{This is automatically the case if
 no options are given to {\tt configure}, and {\tt make install} was 
@@
 \index{norm}{globes@{\tt globes}!verbosity}
-The \AEDL\ parser and interpreter has basically three levels of messages to
+The \AEDL\ parser and interpreter have basically three levels of messages to
 the user: Warnings, errors and fatal errors. 
@@
 to a full spectrum by using {\tt -s}. The spectral rates are shown in a
-table where the first columns always gives the central energy of the
-corresponding bin.
+table where the first column always gives the central energy of the
+corresponding bin or the sampling point.
 
@@
 \begin{quote}
+{\tt
 left\\
@@
 right\\
-\end{quote}
-This is, with left = '\{', middle = ',' and right = '\}',
+}
+\end{quote}
+This is, with {\tt left} = '\{', {\tt middle} = ',' and {\tt right} = '\}',
 equivalent to the list $\{\{1,2,3\},\{1,2,3\}\}$. 
@@
 \end{quote}
-where the first columns is the central energy of the bin, and
-the second columns gives the event rate. Usually, the output is a concatenation of many such two columns tables, where 
+where the first column is the central energy of the bin or the sampling point,
+ and
+the second column gives the event rate. Usually, the output is a concatenation of many such two columns tables, where 
 each rule part or channel part has its own table. Thus one can, by using

### 2004-07-20 Documentation/Manual.tex
Changes Meeting P+W made

--- 2004-07-16 Documentation/Manual.tex
@@
 
-\input{tumlogo}
-
-\begin{minipage}{7cm}
-\begin{center}
-{\bf Technische Universit{\"a}t M{\"u}nchen} \\
-Physik-Department \\
-Institut f{\"u}r Theoretische Physik, T30d \\
-\end{center}
-\end{minipage}
-
-\vspace{-2cm}
-
-\hfill \oTUM{3.5cm}
-
-\vspace{3cm}
+\vspace*{1.5cm}
 
@@
 { \Large \bf
-\GLOBES\ 1.X \\
+{\Huge \GLOBES\ 2.0} \\
 General Long Baseline Experiment Simulator \\ }
@@
 
-\vspace{1cm}
+\vspace{0.5cm}
 
 \begin{center}
-{\large Patrick Huber, Manfred Lindner, Walter Winter}
+{\large Patrick Huber$^a$, Manfred Lindner$^b$, Walter Winter$^c$}
+\end{center}
+
+\vspace{0.3cm}
+
+\begin{center}
+
+Current affiliation(s):
+
+\vspace*{0.5cm}
+
+{\em $^{\mathrm{a},\mathrm{b},\mathrm{c}}$%
+       Institut f\"ur Theoretische Physik, Physik--Department,\\
+       Technische Universit\"at M\"unchen,
+       James--Franck--Strasse, D--85748 Garching, Germany}
+
 \end{center}
@@
 
-\GLOBES\ is mainly developed for academic use. Thus The \GLOBES\ Team would
+\subsection*{Referencing the \GLOBES\ software}
+
+\GLOBES\ is developed for academic use. Thus, the \GLOBES\ Team would
 appreciate being given academic credit for it. Whenever you use \GLOBES\
 to produce a publication or a talk indicate that you have used \GLOBES\ and
-please cite the following reference
+please cite the reference~\cite{globes_paper}
 \begin{quote}
@@
 \end{quote}
-and \emph{not} this manual. This manual does not constitute a scientific 
-publication and will not be submitted to any scientific journal. 
-It will evolve during time since it is intendend for 
-regular revision.
-Besides, many of the data which is used by \GLOBES\ and distributed together
-with it should be correctly referenced where possible. For details see the
-???.
+but \emph{not} this manual. This manual itself is not scientific 
+publication and will not be submitted to a scientific journal. 
+It will evolve during time since it is intended for 
+regular revision. Besides that, many of the data which are used by \GLOBES\ and distributed together with it should be properly referenced. 
+For details see below.
 
-Apart from that, \GLOBES\ is free software and open source, as such it is 
+Apart from that, \GLOBES\ is free software and open source, \ie, it is 
 licensed under the GNU Public License.
+
+\subsection*{Referencing the data in \GLOBES}
+\label{ref_data}
+
+\index{norm}{Referencing!data in \GLOBES}
+\GLOBES\ wouldn't be useful without having high quality input data.
+Much of these input data have been published elsewhere and the authors
+of those publications would appreciate to be cited whenever their work
+is used. It is solely the user's responsibility 
+to make sure that he understands where the input material for \GLOBES\ comes
+from and if additional work has to be cited in addition to the 
+\GLOBES\ paper~\cite{globes_paper}. To assist with this task, we provide  the necessary information for the data coming along together with \GLOBES.
+
+When using the built-in Earth matter density profile, the 
+original source is \Ref~\cite{Stacey}.
+
+All files ending with \verb+.dat+ or \verb+.glb+ 
+in the \verb+data+ subdirectory of the \GLOBES\ tar-ball have on top a comment field which clearly indicates which works should be
+cited when using a certain file. Make sure that dependencies are correctly
+tracked, \ie , in some cases files included by other files need to be 
+checked, too (for example, cross section or flux files). One can use 
+the \verb+-v3+ option to \verb+globes+ to see which files
+are included.
+
+It is recommended that you use the same style for your own input files, since, in case they are distributed, everbody will know how to correctly reference your work.
 
@@
 
+\addcontentsline{toc}{chapter}{Bibliography}
 \bibliographystyle{./unsrt}
 \bibliography{references}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2004-07-20 Documentation/Paper/paper.tex
Changes Meeting P+W made

--- 2004-07-15 Documentation/Paper/paper.tex
@@
                 lindner@ph.tum.de},~and~
-                W.~Winter\footnote[5]{\makebox[1.cm]{Email:}
+                W.~Winter\footnote[3]{\makebox[1.cm]{Email:}
                 wwinter@ph.tum.de}
@@
 \footnotemark[1]${}^,$\footnotemark[2]${}^,$\footnotemark[3]%
-       Institut f\"ur theoretische Physik, Physik--Department,\\
+       Institut f\"ur Theoretische Physik, Physik--Department,\\
        Technische Universit\"at M\"unchen,
@@
 abstract experiment definition language to define experiments with
-beam and full detector descriptions as accurate as possible. For the
+beam and full detector descriptions as accurate as possible. Many systematics options are provided, such as normalization and energy
+calibration errors, or the choice between spectral or total rate 
+information. For the 
 definition of experiments, a new transparent building block concept is 
@@
 can handle the full multi-parameter correlation among the oscillation
-parameters, external input, and matter density uncertainties. Eventually,
-many systematics options are provided, such as normalization and energy
-calibration errors, or the choice between spectral or total rate 
-information. 
+parameters, external input, and matter density uncertainties. 
 \end{abstract}
@@
 Thus optimization strategies are required which maximally exploit 
-complementarity between experiments. Therefore, a modern complete experiment 
+complementarity between experiments. Therefore, a modern, complete experiment 
 simulation  and analysis tool does not only need to have 
-a highly accurate beam and detector simulation but also powerful 
+a highly accurate beam and detector simulation, but also powerful 
 means to analyze correlations and degeneracies, especially for the combination
@@
 \begin{center}
-\includegraphics[width=16cm]{GLOBES}
+\includegraphics[width=14cm]{GLOBES}
 \end{center}
@@
 arbitrary cross sections, energy dependent efficiencies,
-energy resolution functions as well as the considered oscillation channels 
+energy resolution functions as well as the considered oscillation channels, 
 backgrounds, and many other properties can be specified. 
@@
 of \GLOBES\ and to provide prototypes for new experiments.
+In addition, they can be used to test new physics ideas with
+complete experiment simulations.
 Examples for these prototypes are the \MINOS , \ICARUS , and
@@
 \Ref~\cite{Huber:2002mx}, and the reactor experiment simulations from
-\Ref~\cite{Huber:2003pm}. Other projects using the \GLOBES\ 
-software include \Refs~\cite{Huber:2003ak,Ohlsson:2003ip,Winter:2003ye,Antusch:2004yx}.
+\Ref~\cite{Huber:2003pm}. Other projects using earlier versions
+of the \GLOBES\ software include \Refs~\cite{Huber:2003ak,Ohlsson:2003ip,Winter:2003ye,Antusch:2004yx}.
 
@@
 In this section, we sketch the computation of 
-the event rates, which is on of the core parts of the \GLOBES\
+the event rates, which is one of the core parts of the \GLOBES\
 software. However, because of the complexity of this issue,
@@
 evaluated {\em before} the smearing is performed, \ie, as function of $E$.
-Similarly, constant\footnote{with respect to the oscillation parameters, 
-not the energy} event rates can be added to the event rates before or 
+Similarly, constant\footnote{With respect to the oscillation parameters, 
+not the energy.} background rates can be added to the event rates before or 
 after the energy smearing, which are refered to as 
-``pre-smearing backgrounds'' and ``post-smearing background''. These types of 
+``pre-smearing backgrounds'' and ``post-smearing backgrounds''. These types of 
 efficiencies and backgrounds allow a very accurate modeling of
@@
 to define a low-pass filter to avoid aliasing effects for neutrino
-oscillations faster that the sampling width.
+oscillations faster than the sampling width.
 
@@
 The key concept of \AEDL\ is to regard a detector as an abstract system
-which maps the true properties of a neutrino into the reconstructed properties
-of the neutrino which subsequently are used in the fit of the oscillation
-parameters. Within \GLOBES\ only the energy and flavour are used as 
+which maps the true properties of a neutrino onto the reconstructed properties
+of the neutrino. The latter are subsequently used in the fit of the oscillation
+parameters. Within \GLOBES , only energy and flavor are  
 observables, which is
-sufficient for long baseline and reactor experiments. In  other
-experiments, like \eg\ atmospheric experiments, more observables like
-direction may have to be considered.
+sufficient for long-baseline and reactor experiments. In  other
+experiments, such as atmospheric neutrino experiments, more observables 
+(\eg , direction) may have to be considered.
 
-The main parts of any \AEDL\ experiment definition are channels,
-rules, and experiments (\cf, \figu{aedl}).
+The main components of any \AEDL\ experiment definition are ``channels'',
+``rules'', and ``experiments'' (\cf, \figu{aedl}).
 A channel corresponds to a neutrino oscillation channel including flux, cross section (for one specific interaction type), energy resolution function, initial and final neutrino flavors, their polarity (neutrinos or antineutrinos), and efficiencies.
@@
 normalization errors and energy tilt or calibration errors. 
-These systematical errors are evaluated with the ``pull method'' like 
-in~\cite{Huber:2002mx,Fogli:2002pt}.  
+These systematical errors are evaluated with the ``pull method'' such as 
+in \Refs~\cite{Huber:2002mx,Fogli:2002pt}.  
 In addition to these
@@
 Finally, one ore more rules add up to an experiment, where the
-total $\Delta \chi^2$ is obtained as the sum of the $\Delta \chi^2$
+total $\Delta \chi^2$ is obtained as the sum of the $(\Delta \chi^2)_r$
 of all rules. This approach allows the definition of 
@@
 in the experiment definition, which is the matter density profile.
-For an experiment, an arbitrary matter density profiles can be specified, 
-which is evaluated with the evolution operator method REFERENCES.
+For an experiment, an arbitrary matter density profile can be specified, 
+which is evaluated with the evolution operator method~\cite{Ohlsson:1999um}.
 In addition, the matter density profile is multiplied by a scaling factor, 
@@
 approach, one can take into account that the matter density profile along a 
-specific baseline is only known to about $3\%-5\%$.
+specific baseline is only known to about $3\%-5\%$. Thus, for one
+particular baseline, all rules should be defined within one experiment.
 
@@
 \end{center}
-\mycaption{\label{fig:projex} Left plot: Example for a  correlation between $\stheta$ and $\deltacp$ (for 1 d.o.f.
-only).
+\mycaption{\label{fig:projex} Left plot: Example for a  correlation between $\stheta$ and $\deltacp$ (for 1 d.o.f. only, un-shown oscillation
+parameters fixed).
 Right plot: The $\chi^2$-value of the projection onto the $\stheta$-axis as function of $\stheta$.
@@
 access to oscillation probabilities and event rates, and allows
-the readout and modification of many experiment parameters at running time.
+the readout and modification of many experiment parameters at run time.
 
@@
 all oscillation parameters. Together with the matter densities
-of the $N$ loaded experiments ($N \ge 1$), the neutrino oscillation parameter space in \GLOBES\ has $6+N$ dimensions. In the $\chi^2$-approach,
-the allowed region of $k$ ($1\le k<6+N$) fit parameters (at the chosen confidence level) is obtained as the projection of the $6+N$-dimensional fit manifold onto the $k$-dimensional hyperplane.
+of the $N_{\mathrm{exp}}$ loaded experiments ($N_{\mathrm{exp}} \ge 1$), the neutrino oscillation parameter space in \GLOBES\ has $6+N_{\mathrm{exp}}$ dimensions. In the $\chi^2$-approach,
+the allowed region of $k$ ($1\le k<6+N_{\mathrm{exp}}$) fit parameters (at the chosen confidence level) is obtained as the projection of the $6+N_{\mathrm{exp}}$-dimensional fit manifold onto the $k$-dimensional hyperplane.
 For example,
@@
 solution in the parameter space.  For this process, 
-a $6+N$-dimensional local minimizer
+a $6+N_{\mathrm{exp}}$-dimensional local minimizer
 is started at the ``guessed'' position for the solution
@@
 
-As we have discussed in the last section, the matter density of
+As we have indicated above, the matter density of
 each experiment is treated as an additional parameter, which is known
@@
 
-The running time of the systematics $(\Delta \chi^2)_{\mathrm{sys}}$ 
+The run time of the systematics $(\Delta \chi^2)_{\mathrm{sys}}$ 
 is 
@@
 since the oscillation probabilities for all channels and energies
-are only computed once and stored in a list. Here $N_{\mathrm{Sampling}}$ is the number of sampling points for the evaluation of \equ{events_bin}, $N_{\mathrm{Bins}}$ is the number of energy bins,
-and $N_{\mathrm{Layers}}$ is the number of matter density layers.\footnote{Note that we are here using the \AEDL\ format for these
-parameters.}. Thus, whenever the
+are only computed once and stored in a list.
+Here $N_{\mathrm{Sampling}}$ is the number of sampling points for the evaluation of the right integral in \equ{events_bin}, $N_{\mathrm{Bins}}$ is the number of energy bins,
+and $N_{\mathrm{Layers}}$ is the number of matter density layers. Thus, whenever the
 number of matter density steps is small, the product of sampling points
 and bins dominates. In practice, the $N_{\mathrm{Sampling}} \times N_{\mathrm{Bins}}$ energy smearing matrix is already computed at 
-the experiment initialization in order to save running time. In addition,
+the experiment initialization in order to save run time. In addition,
 since it contains a lot of zeros, \GLOBES\ uses a special format to
@@
 
-The running time for the projected $\Delta \chi^2$ is, for a grid based method,
+The run time for the projected $\Delta \chi^2$ would, for a grid based method, be
 \begin{equation}
@@
 \end{equation}
-where $n$ is the number of steps to be evaluated for each oscillation
+where $n$ is the number of grid points to be evaluated for each oscillation
 parameter, and $N_{\mathrm{exp}}$ is the number of experiments
@@
 local minima, \ie, the number of
-degeneracies.
+degeneracies (typically $\mathcal{O}(10)$).
 
@@
 Autotools family and follows the usual procedures. By default
-a shared library {\tt libglobes.so} is installed but 
-a static version is available
-also. Additionally an executable with name {\tt globes} is installed. 
-Provisions have been made to install \GLOBES\ without
+a shared library {\tt libglobes.so} is installed, but 
+a static version is available, too. In addition, an executable named {\tt globes} is installed. It is another possibility to install \GLOBES\ without
 root privilege into a user's home directory. During the automatic
@@
 full multi-parameter correlation, the possibilities to include
-external input and matter density uncertainties and last but not least
-the ability to jointly analyze several experiments.
+external input and matter density uncertainties and 
+the ability to analyze several experiments simultaneously.
 
@@
 
-For the future it is planed to further extend the flexibility of
-\GLOBES\ by providing support for dynamically loadable modules, to
-allow completely user defined systematics or non-standard oscillation
-probabilities.
-
 \subsection*{Acknowledgments}
@@
 We would like to thank Martin Freund, who wrote the very first
-version of a three-flavour matter profile treatment many years ago.
+version of a three-flavor matter profile treatment many years ago.
 PH is especially thankful for the invaluable advice of Thomas Fischbacher on
@@
 papers for suggestions which lead to improvements in the software.
-Parts of this works have been supported by the ``Sonderforschungsbereich
+This work has been supported by the ``Sonderforschungsbereich
 375 f\"ur Astro-Teilchenphysik der Deutschen Forschungsgemeinschaft''.

### 2004-07-26 Documentation/Manual.tex
last changes am Manual

@@
 
+\newcommand{\glbxxx}{hep-ph/0407xxx}
 \newcommand{\capdef}{}
@@
 { \Large \bf
-{\Huge \GLOBES\ 2.0} \\
+{\Huge \GLOBES} \\
 General Long Baseline Experiment Simulator \\ }
@@
 
-Current affiliation(s):
-
-\vspace*{0.5cm}
-
-{\em $^{\mathrm{a},\mathrm{b},\mathrm{c}}$%
-       Institut f\"ur Theoretische Physik, Physik--Department,\\
-       Technische Universit\"at M\"unchen,
-       James--Franck--Strasse, D--85748 Garching, Germany}
-
 \end{center}
@@
 \begin{center}
-Version from \today
+Version from \today\ for \GLOBES\ 2.0
 \end{center}
+
+\vspace*{0.5cm}
+\vfill
+\hrule
+
+{\em\small $^{\mathrm{a},\mathrm{b},\mathrm{c}}$%
+       Institut f\"ur Theoretische Physik, Physik--Department,\\
+       Technische Universit\"at M\"unchen,
+       James--Franck--Strasse, D--85748 Garching, Germany}
 
@@
 
+\vspace*{-2ex}
 \GLOBES\ (``General Long Baseline Experiment Simulator'') is a flexible
@@
 \begin{quote}
-P. Huber, M. Lindner and W. Winter ... 
+P. Huber, M. Lindner and W. Winter\\
+Simulation of long baseline neutrino oscillation experiments with \GLOBES\\
+arXiv:\glbxxx.  
 \end{quote}
-but \emph{not} this manual. This manual itself is not scientific 
+but \emph{not} this manual. This manual itself is not a scientific 
 publication and will not be submitted to a scientific journal. 
 It will evolve during time since it is intended for 
-regular revision. Besides that, many of the data which are used by \GLOBES\ and distributed together with it should be properly referenced. 
+regular revision. Besides that, many of the data which are used by \GLOBES\ 
+and distributed together with it should be properly referenced. 
 For details see below.
@@
 
-It is recommended that you use the same style for your own input files, since, in case they are distributed, everbody will know how to correctly reference your work.
+It is recommended that you use the same style for your own input files, since, in case they are distributed, everybody will know how to correctly reference your work.
 
@@
 first time, which is at variance with C syntax but not with C++ syntax.
-That is the only way in which the examples deviate form ISO C.
-
-%Finally, \Part~\ref{part:3} of this manual serves as reference manual,
-%and the appendix contains further supplementary information.
+That is the only way in which the examples deviate form ISO C. Moreover
+the actual numerical values of the results of the examples may be different
+from the ones in this manual.
 
@@
 Mark Rolinec for his help to translate the experiment descriptions
-into \AEDL , and for creating the illustrations in this manuscript.
+into \AEDL, and for creating the illustrations in this manuscript.
 Finally, thanks to all the people who have been pushing this project
@@
 papers for suggestions which lead to improvements in the software.
+
+This work and the development of \GLOBES\ have been supported by:
+\begin{itemize}
+\item Physik Department der Technischen Universit\"at M\"unchen
+\item Max-Planck-Institut f\"ur Physik 
+\item Sonderforschungsbereich 375 f\"ur Astro-Teilchenphysik\\ 
+der Deutschen Forschungsgemeinschaft
+\item Studienstiftung des Deutschen Volkes 
+\end{itemize}
 
@@
 \addcontentsline{toc}{chapter}{Bibliography}
-\bibliographystyle{./unsrt}
+\bibliographystyle{./OurBibTeX}
 \bibliography{references}

### 2004-07-26 Documentation/part1.tex
last changes am Manual

@@
 }
-Let us now locate the exact position of the sgn-degeneracy:
+Let us now locate the exact position\footnote{For a exact definition of 
+inverted hierarchy, see page~\pageref{mass_ordering}.} of the sgn-degeneracy:
 \gq{
@@
 (except from minor modifications) to the experiments in the
-respective references in the table.
+respective references in the table. These files are installed to
+the directory \verb+${prefix}/share/globes+ which usually defaults
+to  \verb+/usr/local/share/globes+. It is useful to add this path
+to the value of \verb+GLB_PATH+.
+\index{norm}{glb files@{\tt glb}-files!installation}
 
@@
 some of these experiments are outdated in terms of integrated luminosities,
-baseline, fluxes, efficiencies, or other factors.}
+baseline, fluxes, efficiencies, or other factors. In any case these file are installed along with \GLOBES.}
 \end{table}
@@
 
+\label{mass_ordering}
+\index{norm}{Mass hierarchy}
+\paragraph{Note:} Inverting the mass hierarchy is not
+precisely the same than to change from $\Delta m^2_{31}
+\rightarrow -\Delta m^2_{31}$.
+In this case the absolute value of $\Delta m^2_{32}$ changes also, which
+introduces a new frequency to the problem. Therefore, if we assume
+normal hierarchy  whenever $\Delta m^2_{31}>0$, the corresponding
+point in parameters space for inverted hierarchy is given by $\Delta 
+m^2_{31}\rightarrow - \Delta m^2_{31} + \Delta m^2_{21}$, because
+with this defintion the absolute value of  $\Delta m^2_{32}$ is unchanged
+and no new frequency is introduced.\\
+
 Another piece of information will be returned from the minimizers 
@@
 and the reference rate vector, \ie\ the data, has to be computed by a call 
-to \GLB{SetRates}. This has to be done {\em before} any evaluation 
+to \GLB{SetRates}.
+This has to be done {\em before} any evaluation 
 function is used and  {\em after} the experiments have been initialized and 
@@
 \index{norm}{Degeneracies!sgn($\ldm$)-degeneracy}
-
-In many cases, one can find the exact position of the $\mathrm{sgn}(\ldm)$-degeneracy with \GLB{ChiAll}, where one starts 
+\index{norm}{Mass hierarchy}
+
+In many cases, one can find the exact position of the 
+$\mathrm{sgn}(\ldm)$-degeneracy with \GLB{ChiAll}, where one starts 
 the local minimizer at the suspected position 
@@
 In this case, the position of the degeneracy can be easily guessed to be 
-at the true parameter values but with inverted $\ldm$.
+at the true parameter values but with inverted\footnote{For a exact 
+definition of 
+inverted hierarchy, see page~\pageref{mass_ordering}.} $\ldm$.
 The minimizer then runs off the plane of these parameters into the local minimum.

### 2004-07-27 Documentation/Paper/submission/paper.tex
Final version of paper

--- 2004-07-20 Documentation/Paper/paper.tex
@@
 \begin{flushright}
-TUM-HEP-5XX/04\\
+TUM-HEP-553/04\\
 %hep-ph/
@@
 
-\newpage
-\bibliographystyle{./apsrev}
-\bibliography{./references}
+\begin{thebibliography}{10}
+\expandafter\ifx\csname bibnamefont\endcsname\relax
+  \def\bibnamefont#1{#1}\fi
+\expandafter\ifx\csname bibfnamefont\endcsname\relax
+  \def\bibfnamefont#1{#1}\fi
+\expandafter\ifx\csname url\endcsname\relax
+  \def\url#1{\texttt{#1}}\fi
+\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
+\providecommand{\bibinfo}[2]{#2}
+\providecommand{\eprint}[2][]{\url{#2}}
+
+\bibitem{Barger:2003qi}
+\bibinfo{author}{\bibfnamefont{V.}~\bibnamefont{Barger}},
+  \bibinfo{author}{\bibfnamefont{D.}~\bibnamefont{Marfatia}}, \bibnamefont{and}
+  \bibinfo{author}{\bibfnamefont{K.}~\bibnamefont{Whisnant}},
+  \bibinfo{journal}{Int. J. Mod. Phys.} \textbf{\bibinfo{volume}{E12}},
+  \bibinfo{pages}{569} (\bibinfo{year}{2003}), \bibinfo{note}{and references
+  therein}, \eprint{hep-ph/0308123}.
+
+\bibitem{Fogli:1996pv}
+\bibinfo{author}{\bibfnamefont{G.~L.} \bibnamefont{Fogli}} \bibnamefont{and}
+  \bibinfo{author}{\bibfnamefont{E.}~\bibnamefont{Lisi}},
+  \bibinfo{journal}{Phys. Rev.} \textbf{\bibinfo{volume}{D54}},
+  \bibinfo{pages}{3667} (\bibinfo{year}{1996}), \eprint{hep-ph/9604415}.
+
+\bibitem{Burguet-Castell:2001ez}
+\bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Burguet-Castell}},
+  \bibinfo{author}{\bibfnamefont{M.~B.} \bibnamefont{Gavela}},
+  \bibinfo{author}{\bibfnamefont{J.~J.} \bibnamefont{Gomez-Cadenas}},
+  \bibinfo{author}{\bibfnamefont{P.}~\bibnamefont{Hernandez}},
+  \bibnamefont{and} \bibinfo{author}{\bibfnamefont{O.}~\bibnamefont{Mena}},
+  \bibinfo{journal}{Nucl. Phys.} \textbf{\bibinfo{volume}{B608}},
+  \bibinfo{pages}{301} (\bibinfo{year}{2001}),
+  \eprint[http://arXiv.org/abs]{hep-ph/0103258}.
+
+\bibitem{Minakata:2001qm}
+\bibinfo{author}{\bibfnamefont{H.}~\bibnamefont{Minakata}} \bibnamefont{and}
+  \bibinfo{author}{\bibfnamefont{H.}~\bibnamefont{Nunokawa}},
+  \bibinfo{journal}{JHEP} \textbf{\bibinfo{volume}{10}}, \bibinfo{pages}{001}
+  (\bibinfo{year}{2001}), \eprint[http://arXiv.org/abs]{hep-ph/0108085}.
+
+\bibitem{Barger:2001yr}
+\bibinfo{author}{\bibfnamefont{V.}~\bibnamefont{Barger}},
+  \bibinfo{author}{\bibfnamefont{D.}~\bibnamefont{Marfatia}}, \bibnamefont{and}
+  \bibinfo{author}{\bibfnamefont{K.}~\bibnamefont{Whisnant}},
+  \bibinfo{journal}{Phys. Rev.} \textbf{\bibinfo{volume}{D65}},
+  \bibinfo{pages}{073023} (\bibinfo{year}{2002}),
+  \eprint[http://arXiv.org/abs]{hep-ph/0112119}.
+
+\bibitem{Huber:2004ug}
+\bibinfo{author}{\bibfnamefont{P.}~\bibnamefont{Huber}},
+  \bibinfo{author}{\bibfnamefont{M.}~\bibnamefont{Lindner}},
+  \bibinfo{author}{\bibfnamefont{M.}~\bibnamefont{Rolinec}},
+  \bibinfo{author}{\bibfnamefont{T.}~\bibnamefont{Schwetz}}, \bibnamefont{and}
+  \bibinfo{author}{\bibfnamefont{W.}~\bibnamefont{Winter}}
+  \eprint{hep-ph/0403068}.
+
+\bibitem{Huber:2002mx}
+\bibinfo{author}{\bibfnamefont{P.}~\bibnamefont{Huber}},
+  \bibinfo{author}{\bibfnamefont{M.}~\bibnamefont{Lindner}}, \bibnamefont{and}
+  \bibinfo{author}{\bibfnamefont{W.}~\bibnamefont{Winter}},
+  \bibinfo{journal}{Nucl. Phys.} \textbf{\bibinfo{volume}{B645}},
+  \bibinfo{pages}{3} (\bibinfo{year}{2002}), \eprint{hep-ph/0204352}.
+
+\bibitem{Huber:2002rs}
+\bibinfo{author}{\bibfnamefont{P.}~\bibnamefont{Huber}},
+  \bibinfo{author}{\bibfnamefont{M.}~\bibnamefont{Lindner}}, \bibnamefont{and}
+  \bibinfo{author}{\bibfnamefont{W.}~\bibnamefont{Winter}},
+  \bibinfo{journal}{Nucl. Phys.} \textbf{\bibinfo{volume}{B654}},
+  \bibinfo{pages}{3} (\bibinfo{year}{2003}), \eprint{hep-ph/0211300}.
+
+\bibitem{Huber:2003pm}
+\bibinfo{author}{\bibfnamefont{P.}~\bibnamefont{Huber}},
+  \bibinfo{author}{\bibfnamefont{M.}~\bibnamefont{Lindner}},
+  \bibinfo{author}{\bibfnamefont{T.}~\bibnamefont{Schwetz}}, \bibnamefont{and}
+  \bibinfo{author}{\bibfnamefont{W.}~\bibnamefont{Winter}},
+  \bibinfo{journal}{Nucl. Phys.} \textbf{\bibinfo{volume}{B665}},
+  \bibinfo{pages}{487} (\bibinfo{year}{2003}), \eprint{hep-ph/0303232}.
+
+\bibitem{Huber:2003ak}
+\bibinfo{author}{\bibfnamefont{P.}~\bibnamefont{Huber}} \bibnamefont{and}
+  \bibinfo{author}{\bibfnamefont{W.}~\bibnamefont{Winter}},
+  \bibinfo{journal}{Phys. Rev.} \textbf{\bibinfo{volume}{D68}},
+  \bibinfo{pages}{037301} (\bibinfo{year}{2003}), \eprint{hep-ph/0301257}.
+
+\bibitem{Ohlsson:2003ip}
+\bibinfo{author}{\bibfnamefont{T.}~\bibnamefont{Ohlsson}} \bibnamefont{and}
+  \bibinfo{author}{\bibfnamefont{W.}~\bibnamefont{Winter}},
+  \bibinfo{journal}{Phys. Rev.} \textbf{\bibinfo{volume}{D68}},
+  \bibinfo{pages}{073007} (\bibinfo{year}{2003}), \eprint{hep-ph/0307178}.
+
+\bibitem{Winter:2003ye}
+\bibinfo{author}{\bibfnamefont{W.}~\bibnamefont{Winter}},
+  \bibinfo{journal}{Phys. Rev.} \textbf{\bibinfo{volume}{D}} (\bibinfo{year}{to
+  be published}), \eprint{hep-ph/0310307}.
+
+\bibitem{Antusch:2004yx}
+\bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Antusch}},
+  \bibinfo{author}{\bibfnamefont{P.}~\bibnamefont{Huber}},
+  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Kersten}},
+  \bibinfo{author}{\bibfnamefont{T.}~\bibnamefont{Schwetz}}, \bibnamefont{and}
+  \bibinfo{author}{\bibfnamefont{W.}~\bibnamefont{Winter}}
+  (\bibinfo{year}{2004}), \eprint{hep-ph/0404268}.
+
+\bibitem{Manual}
+\emph{\bibinfo{title}{\GLOBES\ manual}} (\bibinfo{year}{2004}),
+  \bibinfo{note}{{\tt http://www.ph.tum.de/$^\sim$globes}}.
+
+\bibitem{Fogli:2002pt}
+\bibinfo{author}{\bibfnamefont{G.~L.} \bibnamefont{Fogli}},
+  \bibinfo{author}{\bibfnamefont{E.}~\bibnamefont{Lisi}},
+  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Marrone}},
+  \bibinfo{author}{\bibfnamefont{D.}~\bibnamefont{Montanino}},
+  \bibnamefont{and} \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Palazzo}},
+  \bibinfo{journal}{Phys. Rev.} \textbf{\bibinfo{volume}{D66}},
+  \bibinfo{pages}{053010} (\bibinfo{year}{2002}), \eprint{hep-ph/0206162}.
+
+\bibitem{Ohlsson:1999um}
+\bibinfo{author}{\bibfnamefont{T.}~\bibnamefont{Ohlsson}} \bibnamefont{and}
+  \bibinfo{author}{\bibfnamefont{H.}~\bibnamefont{Snellman}},
+  \bibinfo{journal}{Phys. Lett.} \textbf{\bibinfo{volume}{B474}},
+  \bibinfo{pages}{153} (\bibinfo{year}{2000}), \bibinfo{note}{erratum ibidem
+  {\bf B480}, 419(E) (2000)}, \eprint{hep-ph/9912295}.
+
+\bibitem{Globes}
+\bibinfo{author}{\bibfnamefont{P.}~\bibnamefont{Huber}},
+  \bibinfo{author}{\bibfnamefont{M.}~\bibnamefont{Lindner}}, \bibnamefont{and}
+  \bibinfo{author}{\bibfnamefont{W.}~\bibnamefont{Winter}},
+  \emph{\bibinfo{title}{{GLoBES (Global Long Baseline Experiment Simulator)
+  software}}}, \bibinfo{note}{{\tt http://www.ph.tum.de/$^\sim$globes}}.
+
+\end{thebibliography}
 

### 2004-08-09 Documentation/part2.tex
Added the the information how to terminate a user-defined smearin matrix in AEDL (page 81)

@@
 \tb\tb $\ldots$\\
+\tb\tb \{40,42, 0.1507103, 0.6965592, 0.1507103\};\\
 >
@@
 \end{quote}
+The last line has to be terminated by a semicolon `;'.
 Note that the sum of all entries in each {\em column} should be equal 

### 2004-09-13 Documentation/Manual.tex
Fixed a minor bug in XCC_CNGS.dat (it contained a spurious newline)
Affiliation updated

@@
 
-\newcommand{\glbxxx}{hep-ph/0407xxx}
+\newcommand{\glbxxx}{hep-ph/0407333}
 \newcommand{\capdef}{}
@@
 
-{\em\small $^{\mathrm{a},\mathrm{b},\mathrm{c}}$%
+{\em\small $^{\mathrm{a},\mathrm{b}}$%
        Institut f\"ur Theoretische Physik, Physik--Department,\\
@@
        James--Franck--Strasse, D--85748 Garching, Germany}
+
+{\em\small $^{\mathrm{c}}$%
+       Institute for Advanced Study, 
+       Einstein Drive, Princeton, NJ 08540, USA}
 
@@
 
-\vspace*{-2ex}
+\vspace*{-4ex}
 \GLOBES\ (``General Long Baseline Experiment Simulator'') is a flexible
@@
 \GLOBES\ is available for GNU/Linux. Since the source code is included,
-the port to other operating systems is in principle possible.
+the port to other operating systems is in principle possible. The software
+as well as up-to-date versions of this manual can be found at this URL:
+ \verb+http://www.ph.tum.de/~globes+
 

### 2004-09-22 Documentation/part3.tex
*** empty log message ***
PH -- Added pdf support. make pdf produces a file Manual.pdf

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% PART III: Reference manual
+% PART III: Module Support
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\part{Reference manual}
+\part{\GLOBES\ modules}
+\label{part:3}
+~\\
+=================================\\
+This is a very raw version of the modules part, it is not clear
+wether it will be an own part or just a chapter somewhere else.
+In any case it should serve a reference in using modules in
+Version 2.1\\
+=================================\\
 
-Here comes a list of all functions and their function in alphabetical order.
+The ability of \GLOBES\ to use dynamically loadable modules makes it easy
+to extend the functionality of \GLOBES\ at run-time and to adapt it to
+new tasks in a transparent way. This part of the manual describes how 
+modules are used, how they are written and compiled. Also some details
+on static linking of modules will be discussed.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\chapter{Using modules}
+\label{chap:using_modules}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\chapter{Writting own modules}
+\label{chap:own_modules}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\chapter{Compiling modules}
+\label{chap:comp_modules}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Static linking}
+\label{sec:static_modules}

### 2004-11-10 Documentation/part1.tex
Parameter order sourcepower etc. corrected (WW)

@@
 \GLBNS{SetSourcePower}
-{\tt void glbSetSourcePower(double power, int exp, int fluxno)}
+{\tt void glbSetSourcePower(int exp, int fluxno, double power)}
  sets the source power of experiment number {\tt exp} and flux number
@@
 \GLBNS{SetRunningTime}
-{\tt void glbSetRunningTime(double time, int exp, int fluxno)}
+{\tt void glbSetRunningTime(int exp, int fluxno, double time)}
  sets the running time of experiment number {\tt exp} and flux number
@@
  \GLBNS{SetTargetMass}
-{\tt void glbSetTargetMass(double mass,int exp)}
+{\tt void glbSetTargetMass(int exp, double mass)}
 sets the fiducial detector mass of experiment number {\tt exp} to {\tt mass} tons or kilotons (depending on the experiment definition).

### 2005-01-13 Documentation/Paper/paper.tex
last changes am Manual
Ref. report changes incorporated

@@
 abstract experiment definition language to define experiments with
-beam and full detector descriptions as accurate as possible. Many systematics options are provided, such as normalization and energy
+beam and full detector descriptions as accurately as possible. Many systematics options are provided, such as normalization and energy
 calibration errors, or the choice between spectral or total rate
@@
 
-\section{Use of package}
+\section{Use of the package}
 
@@
 a shared library {\tt libglobes.so} is installed, but 
-a static version is available, too. In addition, an executable named {\tt globes} is installed. It is another possibility to install \GLOBES\ without
+a static version is available, too. In addition, an executable named {\tt globes} is installed. It possible to install \GLOBES\ without
 root privilege into a user's home directory. During the automatic

### 2005-03-14 Documentation/Manual.tex
PH -- Added pdf support. make pdf produces a file Manual.pdf
PH -- moving some files
Fig. earthint added - for some reason disappeared [WW]
*** empty log message ***
PH prepared the manual for the release of 2.0.11 and removed part 3 (modules) again since it becomes pertinent only for version 2.3

@@
 \setlength{\unitlength}{1cm}
-
+% Do not change following line pdfmanual depends on it!
+\input{pdfswitch}
+%\usepackage{pdfcolmk}
 \usepackage{verbatim}
 \usepackage{cite}
-\usepackage[dvips]{color}
+
 \usepackage{graphicx}
@@
 \begin{center}
-\colorbox{black}{\includegraphics[width=7cm]{earthint}}
+\colorbox{black}{\includegraphics[width=7cm]{.save/earthint}}
 
@@
 \begin{center}
-Version from \today\ for \GLOBES\ 2.0
+Version from \today\ for \GLOBES\ 2.0.11
 \end{center}
@@
 
-{\em\small $^{\mathrm{a},\mathrm{b}}$%
+{\em\small $^{\mathrm{a}}$%
+       Physics Department,
+       University of Wisconsin,
+       1150 University Av., Madison, WI 53706, USA}
+
+{\em\small $^{\mathrm{b}}$%
        Institut f\"ur Theoretische Physik, Physik--Department,\\
@@
 \begin{quote}
-    Copyright \copyright  2004  The GLoBES Team.
+    Copyright \copyright  2004,2005  The GLoBES Team.
     Permission is granted to copy, distribute and/or modify this document
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% PART III: Reference manual
+% PART III: Module Support
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% \input{part3}
+% not yet!
+%\input{part3}
 
@@
 \item Studienstiftung des Deutschen Volkes
+\item Institute for Advanced Study, Princeton
+\item Physics Department of the University of Wisconsin, Madison
 \end{itemize}

### 2005-03-14 Documentation/appendix.tex
last changes am Manual
PH prepared the manual for the release of 2.0.11 and removed part 3 (modules) again since it becomes pertinent only for version 2.3

@@
 any problems on a decently up-to-date GNU/Linux system. The installation
-has, however, only been tested on a limited number of platforms, mainly
-running with various versions of SuSE Linux. Thus we would appreciate
-to know your experiences with the installation on different platforms.
+has, however, only been tested on a limited number of platforms. The 
+current version has been tested for SuSE, Red Hat, Fedora Core and
+a 64bit version of SuSE. It should also compile and work on Mac OS. We 
+would appreciate to know your experiences with the installation on different platforms.
 Please send an e-mail to <\bugs>.
@@
 
-Besides the usual things such as a working libc, you need to have
+Besides the usual things such as a working \verb`libc`, you need to have
 \begin{itemize}
@@
 \verb+gcc.gnu.org+
-\item[BLAS] Basic Linear Algebra Subprograms\\
-\verb+www.netlib.org/blas/+    
-\item[LAPACK] Linear Algebra PACKage\\
-\verb+www.netlib.org/lapack/+
-\item[f2c] Fortran to C\\ 
-\verb+www.netlib.org/f2c/+
 \item[GSL] The GNU Scientific Library\\
@@
 \verb+argp+ facility of \verb+glibc+ to parse its command line options.
+However, on platforms where \verb+argp+ is lacking \GLOBES\ has replacement code, 
+thus it should also work there.
 
 All those libraries are also available as rpm's from the various
-distributors of GNU/Linux: See their web sites for downloads. Due to
-the rather difficult nature of the build process of BLAS and LAPACK,
-it is recommended that one uses a suitable rpm instead. In addition,
-there is a good chance that
-gcc, f2c and GSL are already part of your installation. Furthmore, you need
+distributors of GNU/Linux, see their web sites for downloads. Chances are 
+that \verb+gcc+ and GSL are already part of your installation.  For building \GLOBES\ 
+from source, however, not only working libaries for above packages are 
+needed but also the headers,  especially for GSL. Depending on your
+installation, eg. on RedHat/Fedora, of GSL this may require to additionally 
+install a rpm-package named \verb+gsl-devel+. If GSL has been installed from the 
+tar-ball as provided by \verb`gnu.org` no problems should occur. Furthmore you need 
 a working \verb+make+ to build and install \GLOBES.
-
-In future releases, it is planned to at least make LAPACK obsolete, because
-then also BLAS and f2c would drop out. Furthemore, there is no priniciple
-reason why one should not get rid of any C++ parts. 
-This will greatly simplify
-the build process and reduce the requirements for installation.
 
@@
 The install target also will install a program with name \verb+globes+ to
-\verb+$prefix/bin+.
+\verb+$prefix/bin+ and the files in the \verb+data+ directory of the tar-ball
+to \verb+$prefix/share/globes+.
+\index{norm}{glb files@{\tt glb}-files!installation}
 
@@
 However keep in mind that the linking command will be somewhat different,
-\ie\ you have to specify all the dynamically linked objects besides libglobes,
-which are
-\begin{quote}
-{\tt
-         -lm -lgsl -lgslcblas -lf2c -llapack -lblas
-}
-\end{quote}
+\ie\ you have to specify all the dynamically linked objects besides libglobes.
 
-In general it is advisable to use the shared libraries. If you don't
-have root privileges, see the corresponding section 
-(page~\pageref{inst_noroot}).
-If you install \GLOBES\ with root privileges, do not forget to run 
-{\tt ldconfig} after installation.
+The default install directory prefix is \verb+/usr/local+.  Consult the
+"Further Information" section below for instructions on installing the
+library in another location or changing other default compilation
+options.
+
+Moreover a config-script called \verb+globes-config+ will be installed. This
+script displays all information necessary to link any program with GLoBES.
+For building static libaries and linking against them see the corresponding
+section of this file.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{LAPACK \& BLAS requirements}
+
+GLoBES uses one single routine out of LAPACK (\verb^www.netlib.org/lapack/^) named
+\verb^zgeev^ to diagonalize the oscillation Hamiltonian in matter. Since LAPACK
+is not part of the standard installation of many distributions and we need
+only one function, \GLOBES\ comes with a convenience version of the required
+routines, which are taken from LAPACK and BLAS (\verb^www.netlib.org/blas^) and have
+been translated to C with \verb^f2c^. In case your system has a working installation
+of BLAS/LAPACK \GLOBES\ will use it. However, if your system is lacking BLAS/
+LAPACK the convenience version will be compiled and linked into \GLOBES. If
+for some reason BLAS/LAPACK is installed on the system and one still wishes
+to use the convience version use \verb^--enable-lapack-convenience^. Using Fortran
+code which has been translated to C with \verb^f2c^ requires a run-time library
+called \verb^libf2c^ (on some systems it is called \verb^libg2c^). The configury of 
+\GLOBES\ detects the presence of either of the two and takes care of it.
+If none of the two is present, a convience library is used. If
+for some reason \verb^libf2c^ is installed on the system and one still wishes
+to use the convience version use \verb^--enable-libf2c-convenience^.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{GSL requirements}
+
+Sometimes the GNU scientific library is not available or is installed in a non-standard
+location. This situation can arise in an installation without root privileges. In this 
+case one can specify \verb^--with-gsl-prefix=path_to_gsl^ as option to the \verb^configure^ script.
+If one wants to use a shared version of \verb^libgsl^ then one has to make sure that the
+linker finds the library at run-time. This can be achieved by setting the environment variable
+\verb^LD_LIBRARY_PATH^ correctly, \ie\ (in bash)
+\begin{quote} 
+\verb^export LD_LIBRARY_PATH=path_to_gsl^
+\end{quote}
+You also can use a static version of GSL by either building \GLOBES\ with \verb^LDFLAG=-all-static^
+or by configuring GSL with \verb^--disable-shared^. In both cases no further actions like setting
+any environment variables is necessary.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Distributions}
+
+\subsection*{Fedora Core 3}
+
+FC3 is shipped with a non working version of LAPACK. Thus to
+make \GLOBES\ work on FC3, configure \GLOBES\ with \verb'--enable-lapack-convenience', which
+replaces the defunct LAPACK with \GLOBES' own version thereof. You can diagnose a 
+non-working LAPACK by compiling one of the examples in the \verb^examples^ directory and 
+trying to run it. It will hang forever in the function \GLB{SetRates()}.
+
+\subsection*{RedHat -- all versions}
+
+The standard rpm-based installation of GSL does not provide any header files for GSL, 
+which are however needed to compile \GLOBES. You have to install an additional rpm-package
+called \verb'gsl-devel'. Altnernatively you can install GSL from a tar-ball and use the 
+verb'--with-gsl-prefix' option to the configure script of \GLOBES.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section*{Advanced topics}
+
+\subsection*{Building and Using static versions of \GLOBES}
+
+In certain circumstances it may be useful to use a static version of libglobes
+or any of the binaries, \eg\ when running on a cluster.
+
+   The \verb+configure+ script accepts the option \verb+--disable-shared+, in which
+case only static objects are built, i.e. only a static version of libglobes. 
+In case your system does not support shared libraries the \verb+configure+ script 
+recognizes this. If you give no options to \verb+configure+, both shared and static 
+versions are built and will be installed. All binaries, however, will use dynamic 
+linking. If you want to build static binaries, use \verb+LDFLAGS=-all-static+ for building
+them.
+
+   Sometimes it is convenient, eg. for debugging purposes, to have a statically
+linked version of a program using \GLOBES\, which is easiest achieved by just linking
+with \verb+libglobes.a+. If you need a completely statically linked version, please, have
+a look at the Makefile in the \verb+examples+ directory.
+\begin{quote}
+\verb#make example-static#
+\end{quote}
+produces a statically linked program that should in principle run on most Linuxes.
+It should be straightforward to adapt this example to your needs.
+
+All these options rely on \verb+libtool+ and a working \verb`gcc` installation. It seems that \verb`gcc 3.x`
+is broken in subtle way which makes it necessary to add a symbolic link in the \verb`gcc` library
+directory. The diagnostic for this requirement is that building static programs fails with
+the error message \verb+cannot find -lgcc_s+. In those cases, find \verb+libgcc.a+ and add a symbolic 
+link in the same directory where you found it (this requires probably root privileges)
+\begin{quote}
+\verb#ln -s libgcc.a libgcc_s.a#
+\end{quote}
+
+If you can not write to this directory just use the following work around. Add the same link  
+as above to the directory where you installed \GLOBES\ into
+\begin{quote}
+\verb#cd prefix/lib#
+\verb#ln -s path_to_libgcc.a/libgcc.a libgcc_s.a#
+\end{quote}
+and then change back into the \verb+examples+ directory and type
+\begin{quote}
+\verb#make LDFLAGS=-Lprefix/lib example-static#
+\end{quote}
+and you are done.
+
+If you don't have libtool on your system either, then you just can use the libtool script in
+the \GLOBES\ tar-ball instead. Make sure it is in a place where it is found by \verb+make+.
 
@@
 
+{\footnotesize
 \input{gpl}
@@
 \input{fdl}
+}
 
@@
 %%% TeX-master: Manual.tex
+%%% TeX-master: "Manual"
 %%% End:

### 2005-07-27 Distribution/doc/howto.tex
PH -- added the GNUlib support to the configury and to the makefiles. All the nice feature of 2.0.11 should now work with 2.3.7. Still has to be tested.

\documentclass{article}
\usepackage{verbatim}
\newcommand{\glb}{{\sf GLoBES}}
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Purpose}
This document describes how to install \glb\ in an adverse
environment, ie. as user without root privileges. This 
situation may either arise if one just wants to quickly 
evaluate wether \glb\ is suitable for a certain task or
when the system administration proves to be un-cooperrative.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Prerequisites}

The dependencies of \glb\ have been reduced as far as 
possible and the only remaining hard prerequisite is the
GNU Scientific library, which itself is useful for many
projects and thus should be installed at any rate. However,
there are occasions where even this minimal requirement 
cannot be met. The simplest solution in these cases is, to
download the tar-ball from {\tt www.gnu.org/software/gsl/}
to unpack it, to change into the GSL directory and to run
\begin{quote}
\begin{verbatim}
./configure --prefix=path_to_gsl --disable-shared
make
make install
\end{verbatim}
\end{quote}
where \verb^path_to_gsl^ points to a location where the user
has write permission.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Building \& Installing \glb}

Once GSL is made available the next step is to download
the \glb\ tar-ball, unpack it and to change into the \glb\
directory. There the following commands have to be executed
\begin{quote}
\begin{verbatim}
./configure --prefix=path_to_globes --with-gsl-prefix=path_to_gsl
make
make install
\end{verbatim}
\end{quote}
Now you can change into the \verb^examples^ subdirectory
and build the first example with
\begin{quote}
\begin{verbatim}
make
\end{verbatim}
\end{quote}
typing
\begin{quote}
\begin{verbatim}
./example-tour
\end{verbatim}
\end{quote}
shoud produce something like
\begin{quote}
{\small
\begin{verbatim}
Oscillation probabilities in vacuum: 1->1: 0.999955 1->2: 2.59018e-05 1->3: 1.92391e-05
Oscillation probabilities in matter: 1->1: 0.999965 1->2: 2.01594e-05 1->3: 1.49783e-05
...
\end{verbatim}
}
\end{quote}
in which case you have succeded to install \glb.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Using {\tt libglobes}}

The \verb^examples^ subdirectory does not only contain example C-code, but also a \verb^Makefile^ which can be used as a basis for building own projects. You just have to replace the
source and object files by the ones you are going to use. Moreover there is a script
called \verb^globes-config^ installed which can be used to find the necessary linker
flags by
\begin{quote}
\begin{verbatim}
globes-config --libs
\end{verbatim}
\end{quote}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Building static binaries}

Especially for CPU intensive task having the possibility to use many CPUs is desirable. This is much easier if one can buil static binaries, i.e. binaries which can be compiled on
one machine providing a reasonable build environment and at
the same can be run on another machine 
(or many other machines). Again the \verb^Makefile^ in the
\verb^examples^ subdirectory contains the necessary information. Change into \verb^examples^ and type
\begin{quote}
\begin{verbatim}
make example-static
\end{verbatim}
\end{quote}
The resulting binary \verb^example-static^ should run on most Linuxes and the corresponding parts of the \verb^Makefile^ should be easily adaptible to your problem. Note, that this approach uses \verb^libtool^ to do the linking. If \verb^libtool^ should not be available on your system, one can use the
\verb^libtool^ script in the \glb\ build directory as replacement. Just make sure it is in the path of the invocation of
\verb^make^.

All these options rely on \verb^libtool^ and a working gcc installation. It seems that \verb^gcc 3.x^
is broken in subtle way which makes it necessary to add a symbolic link in the gcc library
directory. The diagnostic for this requirement is that building static programs fails with
the error message \verb^cannot find -lgcc_s^. In those cases, find \verb^libgcc.a^ and add a symbolic 
link in the same directory where you found it (this requires probably root privileges)
\begin{quote}
\begin{verbatim}
ln -s libgcc.a libgcc_s.a
\end{verbatim}
\end{quote}

If you can not write to this directory just use the following work around. Add the same link  as above to 
the directory where you installed \verb^globes^ into
\begin{quote}
\begin{verbatim}
cd prefix/lib       
ln -s path_to_libgcc.a/libgcc.a libgcc_s.a
\end{verbatim}
\end{quote}
and then change back into the \verb^examples^ directory and type
 \begin{quote}
\begin{verbatim}
make LDFLAGS=-Lprefix/lib example-static
\end{verbatim}
\end{quote}
and you are done.

\end{document}

### 2005-09-27 Presentations/RingbergYSW2005/globes.tex
Added talk of A. Merle and J. Kopp at the 2005 Young Scientists Workshop, Ringbaerg Castle

% Simulaing Neutrino Oscillation Experiments with GLoBES
% -----------------------------------------------------------------------------
% Talk given at the 2005 Ringberg Young Scientists Workshop on Hot
% Topics in Particle and Astroparticle Physics
% -----------------------------------------------------------------------------
\documentclass{beamer}

% -----------------------------------------------------------------------------
\title[Neutrino Oscillations and GLoBES]
      {Simulating Neutrino Oscillation Experiments with GLoBES}
\author[J. Kopp, A. Merle]
       {\parbox{9 cm}{Part 1: Alexander Merle (\texttt{amerle@ph.tum.de}) \\
                      -- Theory of Neutrino Oscillations       \\
                      -- State of the Art in Neutrino Physics  \\ \\
                      Part 2: Joachim Kopp (\texttt{jkopp@ph.tum.de})     \\
                      -- The GLoBES Software Package           \\
                      -- Simulation of Future Experiments}}
\institute[TUM]
          {TU M\"unchen, Lehrstuhl T30d f\"ur Theoretische Elementarteilchenphysik (Prof. Lindner)}
\date[21.07.2005]
     {Young Scientists Workshop on Hot Topics in Particle and Astroparticle Physics \\
      18.07. - 22.07.2005, Ringberg Castle}
% -----------------------------------------------------------------------------
\usepackage[english]{babel}
\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{graphicx}

% Set up beamer package
\mode<presentation>
{
  \usetheme{Madrid}
  \setbeamercovered{transparent}
}
\setbeamertemplate{navigation symbols}{}

% Some commands to simplify things
\newcommand{\twovec}[2]{\ensuremath{ \left( \begin{array}{c} {#1}\\{#2} \end{array} \right) }}
\newcommand{\twomat}[4]{\ensuremath{ \left( \begin{array}{lr} {#1} & {#2} \\ {#3} & {#4} \end{array} \right) }}
\newcommand{\threevec}[3]{\ensuremath{ \left( \begin{array}{c} {#1}\\{#2}\\{#3} \end{array} \right) }}
\newcommand{\threemat}[9]{\ensuremath{ \left( \begin{array}{lcr} {#1} & {#2} & {#3} \\
                                                                {#4} & {#5} & {#6} \\
                                                                {#7} & {#8} & {#9} \end{array} \right) }}
\newcommand{\degrees}{\ensuremath{^\circ}}
\newcommand{\reference}[1]{\tiny #1 \normalsize}

% Show table of content before each section
\AtBeginSection[]
{
    \begin{frame}
      \frametitle{Outline}
      \tableofcontents[currentsection]
    \end{frame}
}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}
  \frametitle{Outline}
  \tableofcontents
\end{frame}

% =============================================================================
\section{The Neutrino Oscillation Framework}
% =============================================================================

%..............................................................................
\subsection{2-Flavour Oscillations in Vacuum}
% -----------------------------------------------------------------------------

\begin{frame}
  \frametitle{Mass and flavour eigenstates}

  \begin{itemize}
    \item For neutrinos, the eigenstates of the mass operator are not equal to
    the ones of the weak interaction:

   \begin{figure}
    \begin{center}
      \includegraphics[width=7.5cm]{fig/osc1.jpg}
    \end{center}
  \end{figure}

     \item  Massless neutrinos: the mass eigenstates can be arbitrarily chosen $\Rightarrow$ can be defined as identical to the flavour eigenstates $\Rightarrow$ no oscillations! $\Rightarrow$ oscillations are the evidence for non-zero neutrino mass!

   \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{The formalism}

  \begin{itemize}
    \item Let ($|\nu_{e}\rangle$,$|\nu_{\mu}\rangle$) be the flavour and ($|\nu_1\rangle$,$|\nu_2\rangle$) the mass eigenstates $\Rightarrow$ every flavour-state can be written as a linear combination of mass eigenstates (since it is a rotation one can use a rotation matrix):

$$
\left(\begin{array}{c}
| \nu_{e} \rangle \\
| \nu_{\mu} \rangle \\
\end{array}\right)
=
\left(\begin{array}{cc}
\cos \theta & \sin \theta\\
-\sin \theta	& \cos\theta\\
\end{array}\right)
\cdot
\left(\begin{array}{c}
| \nu_{1} \rangle \\
| \nu_{2} \rangle \\
\end{array}\right)
$$

	\item Time evolution: $$|\nu_{e} (t) \rangle \ =\ \cos \theta  \ e^{-iE_1 t}|\nu_1 (t=0) \rangle \ + \ \sin \theta \ e^{-iE_2 t}|\nu_2 (t=0) \rangle $$

	\item Energy momentum relation: $$E_{i}^2=p_{i}^2+m_{i}^2 \Rightarrow E_{i}=\sqrt{p_{i}^2 +m_{i}^2}\approx p_{i}+\frac{m_{i}^2}{2p_{i}}$$

   \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Diasappearance Probability}

  \begin{itemize}
    \item Assumption: both of the mass eigenstates have the same momentum p (in fact, this does not have to be true, but the longer calculation gives the same result) $\Rightarrow$ with $p=E_{\nu}$: $E_2=E_1-\frac{\Delta m^2}{2E_{\nu}}$ where $\Delta m^2 =m_1^2 - m_2^2$

	\item E. g. for the muon-neutrino we get: $$|\nu_{\mu} \rangle =e^{-i(p+\frac{m_{1}^2}{2p})t} \cdot \left[ -\sin\theta \ |\nu_1(0) \rangle +\cos\theta \ e^{\frac{i\Delta m^2}{2E_{\nu}}} |\nu_2(0) \rangle \right]$$

	\item With a simple calculation (trigonometric identities, etc.) one can construct the probability for detecting a $\nu_{\mu}$ as a $\nu_e$ at a time t: $$P_{\nu_{\mu}\rightarrow \nu_{e}}(t)=| \langle \nu_{e}(0)|\nu_{\mu}(t) \rangle |^2=|(\cos\theta \ \langle \nu_1(0)| + \sin \theta \ \langle \nu_2(0)| ) \cdot |\nu_{\mu}(t) \rangle |^2$$ $=\sin^2 (2\theta) \cdot \sin^2 (\frac{t \Delta m^2}{4E_{\nu}} ) \Rightarrow$ oscillates with time!

   \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Survival Probability}

  \begin{itemize}
    \item Probability to see the $\nu_{\mu}$ again as a $\nu_{\mu}$: $$P_{\nu_{\mu}\rightarrow \nu_{\mu}}(t)=1-P_{\nu_{\mu}\rightarrow \nu_{e}}(t)=1-\sin^2 (2\theta) \cdot \sin^2 (\frac{t \Delta m^2}{4E_{\nu}} )$$

	\item Neutrinos travel (nearly) with the speed of light $\Rightarrow$ one can set t=x, where x is the distance to the place where the neutrino was born

	\item Oscillation length: distance after which the $\nu_{\mu}$ is again a 100 \% $\nu_{\mu}$ $$ \sin^2(\frac{x \Delta m^2}{4E_{\nu}})\equiv \sin^2(\frac{\pi x}{L_0}) \Rightarrow L_0=\frac{4\pi E_{\nu}}{\Delta m^2} \propto \frac{E_{\nu}}{\Delta m^2}$$ 
   $\rightarrow$ important quantity for experiments

	\item Notice, that only the squared mass difference $\Delta m^2$ can be determined by oscillation experiments!

   \end{itemize}
\end{frame}

% -----------------------------------------------------------------------------
\subsection{Matter Effects}
% -----------------------------------------------------------------------------

\begin{frame}
  \frametitle{Matter Effects}

  \begin{itemize}
 	\item The Hamiltonian in the basis of the mass eigenstates is diagonal $\Rightarrow$ Schr\"odinger's equation (one can skip the first p from $E=p+\frac{m^2}{2p}$ since it just gives a global phase): 
$$
i\frac{d}{dt}
\left(\begin{array}{c}
| \nu_{1} \rangle \\
| \nu_{2} \rangle \\
\end{array}\right)=
H_{M}
\left(\begin{array}{c}
| \nu_{1} \rangle \\
| \nu_{2} \rangle \\
\end{array}\right)=
\frac{1}{2p}\cdot
\left(\begin{array}{cc}
m_1^2 & 0\\
0 & m_2^2\\
\end{array}\right)
\left(\begin{array}{c}
| \nu_{1} \rangle \\
| \nu_{2} \rangle \\
\end{array}\right)$$

	\item To go into the flavour basis, one has to rotate the states as well as the Hamiltonian by a unitary matrix U, which leeds to the Hamiltonian $H_{F}$ in the flavour basis:
$$
H_{F}=
\frac{1}{2p}\cdot
\left(\begin{array}{cc}
m_{ee}^2 & m_{e \mu}^2\\
m_{e \mu}^2 & m_{\mu \mu}^2\\
\end{array}\right)=
U H_{M} U^{\dagger}=
$$
$$
=\frac{1}{2p} \cdot
\left(\begin{array}{cc}
\cos \theta & \sin \theta\\
-\sin \theta & \cos \theta\\
\end{array}\right)
\cdot
\left(\begin{array}{cc}
m_1^2 & 0\\
0 & m_2^2\\
\end{array}\right)
\cdot
\left(\begin{array}{cc}
\cos \theta & -\sin \theta\\
\sin \theta & \cos \theta\\
\end{array}\right)=
$$

  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Matter Effects}

$$
=\frac{1}{4p}\left[ \Sigma 
\left(\begin{array}{cc}
1 & 0\\
0 & 1\\
\end{array}\right)
+D
\left(\begin{array}{cc}
-\cos(2 \theta) & \sin(2 \theta)\\
\sin(2 \theta) & \cos(2 \theta)\\
\end{array}\right)
\right]
$$

\hspace{2cm} where $\Sigma=m_1^2+m_2^2$ and $D=m_2^2-m_1^2$

  \begin{itemize}

	\item Since it is diagonal, the first term just gives a common phase $\Rightarrow$ has no physical meaning

	\item Coherent forward scattering in matter: $\nu_e$ and $\overline{\nu}_e$ can - different from all other flavours - interact with the electrons via CC-reactions $\Rightarrow$ change of the oscillation probabilities!

	\item The Hamiltonian in the flavour basis has to be changed: 
$$H_{F} \rightarrow H_{F, mat}=H_{F}+\frac{1}{2p}
\left(\begin{array}{cc}
A & 0\\
0 & 0\\
\end{array}\right)$$
where $A=2\sqrt{2}G_{F}N_{e}p$ ($G_F$: Fermi's constant, $N_e$: electron density)

  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Matter Effects}

  \begin{itemize}
    \item Back to the mass eigenstates $\Rightarrow$ after a lengthy calculation we get:
$$
H_{M,mat}=
\frac{1}{2p} \cdot
\left(\begin{array}{cc}
m_1^2+A \cos^2(\theta) & A \sin(\theta) \cos(\theta)\\
A \sin(\theta) \cos(\theta) & m_2^2+A \sin^2(\theta)\\
\end{array}\right)
$$

$\Rightarrow$ this Hamiltonian is not diagonal anymore $\Rightarrow$ the mass eigenstates of the vacuum are not eigenstates in matter

	\item So one has to diagonalize again to get the new mass eigenstates $|\nu_{1m}\rangle$ and $|\nu_{2m}\rangle$ in matter, which gives

$$m_{1,2m}^{2}=\frac{1}{2}\cdot \left[ (\Sigma + A) \pm \sqrt{(A-D \cos(2 \theta))^2+ D^2 \sin^2(2\theta)} \right]$$

and so the squared mass difference in matter is (remember that $D=m_2^2-m_1^2$)

$$D_{m}=m_{2m}^{2}-m_{1m}^{2}= D \cdot \sqrt{(\frac{A}{D}-\cos(2 \theta))^2+\sin^2(2 \theta)}$$

   \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Matter Effects}

  \begin{itemize}
    \item So we also have in matter a new mixing angle $\theta_{mat}$
and so a new transition probability
$$P_{mat, \nu_e \rightarrow \nu_{\mu}}(x)= \sin^2(2\theta_{mat}) \sin^2\left(\frac{x D_{m}}{4p}\right) $$

	\item The oscillation amplitude is

$$\sin^2(2\theta_{mat})=\frac{\sin^2(2 \theta)}{(\frac{A}{D}-\cos(2 \theta))^2+\sin^2(2 \theta)}=f(A/D)=f(E_{\nu}N_e)$$
$\Rightarrow$ the ratio $\frac{A}{D}$ determines the oscillation probabilities in matter $\Rightarrow$ for $\frac{A}{D}\approx\cos(2\theta)$ we have a resonance effect $\Rightarrow$ \textbf{a muon-neutrino can at certain distances be a 100\% electron-neutrino!!!}

   \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Flavour-Flipping/The MSW-Effect}

(\underline{M}ikheyev-\underline{S}mirnov-\underline{W}olfenstein)

\begin{figure}
    \begin{center}
      \includegraphics[width=7.5cm]{fig/resonant.jpg}
    \end{center}
  \end{figure}

e.g. the first mass eigenstate ($\nu_{1m}$-curve) is most probably a $\nu_e$ below and most probably a $\nu_{\mu}$ above the resonant matter density $A_R$

\end{frame}

%..............................................................................
\subsection{3-Flavour Oscillations }
%..............................................................................

\begin{frame}
  \frametitle{3-Flavour Oscillations}
	\begin{itemize}

	\item For 3 neutrino-flavours the mixing looks like:
$$
\left(\begin{array}{c}
| \nu_{e} \rangle \\
| \nu_{\mu} \rangle \\
| \nu_{\tau} \rangle \\
\end{array}\right)=
U\cdot
\left(\begin{array}{c}
| \nu_{1} \rangle \\
| \nu_{2} \rangle \\
| \nu_{3} \rangle \\
\end{array}\right)
$$
with $U^{\dagger}=U^{-1}$ and
$$
U=
\left(\begin{array}{ccc}
c_{12}c_{13} & s_{12}c_{13}&s_{13}e^{-i\delta}\\
-s_{12}c_{23}-c_{12}s_{23}s_{13}e^{i\delta} & c_{12}c_{23}-s_{12}s_{23}s_{13}e^{i\delta} & s_{23}c_{13}\\
s_{12}s_{23}-c_{12}c_{23}s_{13}e^{i\delta} & -c_{12}s_{23}-s_{12}c_{23}s_{13}e^{i\delta} & c_{23}c_{13}\\
\end{array}\right)
$$
where c$_{ij}=\cos(\theta_{ij})$ and s$_{ij}=\sin(\theta_{ij})$

\item Parameters: 3 mixing angles ($\theta_{12}$, $\theta_{23}$ and $\theta_{13}$), 1 CP-violating phase ($\delta$) and 2 squared mass differences ($(\Delta m^2)_{sol}=(\Delta m^2)_{21}$ and $(\Delta m^2)_{atm}=(\Delta m^2)_{31}$)

	\end{itemize}  
\end{frame}

\begin{frame}
	\frametitle{3-Neutrino-Mixing}

\begin{figure}
    \begin{center}
      \includegraphics[width=12cm]{fig/angles.pdf}
    \end{center}
 \end{figure}

\end{frame}

% =============================================================================
\section{The State of the Art in Neutrino Physics}
% =============================================================================

%..............................................................................
\subsection{The Oscillation Parameters}
% -----------------------------------------------------------------------------
\begin{frame}
	\frametitle{The Oscillation Parameters}

\begin{block}{}
    \begin{tabular}{lcc}
      Parameter         &  Best fit value and $\pm 3 \sigma$ range & Experiments \\ \hline \\ 
	$\theta_{12}\approx 32.6\degrees $		&  $\sin^2(\theta_{12})=0.29^{+0.08}_{-0.06}$	 & solar/reactor neutrinos \\ 
	$\theta_{23}\approx 45.0\degrees$ 		&  $\sin^2(\theta_{23})=0.50\pm 0.16$		&   atmospheric/accelerator\\
	$\theta_{13}\lesssim 9.6\degrees$ 		& $\sin^2(\theta_{13})=0.000+0.047$ & CHOOZ-experiment\\
	\hspace{0.8cm}$\delta$			 & ????? & several proposals\\
	\hspace{0.3cm}$(\Delta m^2)_{21}$  & $(8.1^{+1.0}_{-0.8})\cdot 10^{-5}$ eV$^2$ & solar/reactor\\
	\hspace{0.3cm}$(\Delta m^2)_{31}$  & $(2.2^{+1.1}_{-0.8}) \cdot 10^{-3}$ eV$^{2}$ & atmospheric/accelerator\\
	\end{tabular}
\begin{flushright}
\tiny{taken from Maltoni, Schwetz et al., New J. Phys. 6 (2004)\\
see also: Bahcall et al., JHEP 08 (2004) or Bandyopadhyay et al., Phys. Lett. B608 (2005)}
\end{flushright}
	 \end{block}

\begin{itemize}
    \item
    Two large mixing angles and one small mixing angle ($\neq$CKM!).

    \item
    No information about CP violation.

    \item
    Precise measurement of neutrino mixing parameters is essential for a deeper
    theoretical understanding of flavour physics and physics beyond the Standard Model.
  \end{itemize}

\end{frame}

%..............................................................................
\subsection{Mass Hierarchies}
%..............................................................................
\begin{frame}
	\frametitle{Mass Hierarchies}

Problem: since oscillation only measures squared mass differences, and we do not know the sign of $(\Delta
m^2)_{atm}$ yet, there could be different hierarchies

\begin{figure}
    \begin{center}
      \includegraphics[width=8cm]{fig/hierarchy.pdf}
    \end{center}
 \end{figure}

$\Rightarrow$ from the oscillations we get only the information, that neutrinos do have mass, but nothing about the absolute values

\end{frame}

% -----------------------------------------------------------------------------
\subsection{The Past and the Future}
% -----------------------------------------------------------------------------

\begin{frame}
	\frametitle{Where do we know all that from?}

\begin{itemize}

\item Super-Kamiokande (atmospheric $\nu$'s), K2K,... $\Rightarrow$ $\theta_{23}$, $(\Delta m^2)_{31}$
\begin{figure}
    \begin{center}
      \includegraphics[width=3.8cm]{fig/atm1.jpg}
    \end{center}
 \end{figure}

\item Solar neutrinos, KamLAND, ... $\Rightarrow$ $\theta_{12}$, $(\Delta m^2)_{21}$
\begin{figure}
    \begin{center}
      \includegraphics[width=3cm]{fig/sun.jpg}
    \end{center}
 \end{figure}

\item CHOOZ $\Rightarrow$ $\theta_{13}$

\end{itemize}

\end{frame}

\begin{frame}
\frametitle{The next steps}

\begin{itemize}
\item Achieving a better precision for the known mixing parameters
\item Measurement of the small mixing angle $\theta_{13}$
\item Measurement of the CP-phase $\delta$
\end{itemize}

\vspace{1.5cm}

\large{AND (what we do): Making Simulations with GLoBES }
\end{frame}

% =============================================================================
\section{GLoBES - The General Long Baseline Experiment Simulator}
% =============================================================================

% -----------------------------------------------------------------------------
\subsection{GLoBES Basics}
% -----------------------------------------------------------------------------

\begin{frame}
  \frametitle{The GLoBES software package}

  \begin{columns}
    \begin{column}{6 cm}
      \begin{figure}
        \includegraphics[width=5 cm]{fig/globes_logo.jpg}
      \end{figure}
    \end{column}
    \begin{column}{6 cm}
      \begin{itemize}
        \item C library for simulating neutrino oscillation experiments
        \item Provides sophisticated functions for high level and low level access
              to the simulation
        \item AEDL --- Abstract Experiment Definition Language for defining experimental
              setups
        \item Developed at TUM by P.~Huber, M.~Lindner and W.~Winter \\
      \end{itemize}
    \end{column}
  \end{columns}
  \begin{flushright}
    \reference{Comput. Phys. Commun. 167:195, 2005, hep-ph/0407333}
  \end{flushright}
\end{frame}

\begin{frame}
  \frametitle{Components of a GLoBES simulation}

  \begin{columns}
    \begin{column}{5.5 cm}
      \begin{block}{Neutrino Sources}
        \parbox[t][2.6 cm]{5.2 cm}{  
        \begin{itemize}
          \item Energy and flavour spectrum
          \item Signal and Background
        \end{itemize}}
      \end{block}
      \begin{block}{Detector}
        \parbox[t][2.7 cm]{5.2 cm}{  
        \begin{itemize}
          \item Interaction cross sections
          \item Energy resolution
          \item Systematical errors
        \end{itemize}}
      \end{block}
    \end{column}
    \begin{column}{5.5 cm}
      \begin{block}{Oscillations}
        \parbox[t][2.6 cm]{5.2 cm}{  
        \begin{itemize}
          \item Fundamental parameters: $\theta_{12}$, $\theta_{13}$, $\theta_{23}$,
                $\delta$, $\Delta m_{21}^2$, $\Delta m_{31}^2$.
          \item Baseline and matter density profile
        \end{itemize}}
      \end{block}
      \begin{block}{Analysis}
        \parbox[t][2.7 cm]{5.2 cm}{  
        \begin{itemize}
          \item Number and width of energy bins
          \item Fitting parameters to the simulated data
          \item $\chi^2$ analysis
        \end{itemize}}
      \end{block}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Calculation of event rates}
  \begin{block}{}
    \vspace{-0.2 cm}
    \begin{eqnarray*}
      \frac{dn_{f}^{\mathrm{IT}}}{dE'} &\propto&  \sum_{f_i}\int \int dE\,d\hat{E}\quad
                              \underbrace{\Phi_{f_i} (E)}_{\mathrm{Production}} \times \\
            &&\underbrace{\frac{1}{L^2} P_{(f_i\rightarrow f)}(E,L,\rho;\theta_{23},\theta_{12},\theta_{13},
                     \Delta m^2_{31},\Delta m^2_{21},\delta)}_{\mathrm{Propagation}}
                     \times \nonumber \\ &&\underbrace{\sigma^{\mathrm{IT}}_f(E)
                         k_f^{\mathrm{IT}}(E-\hat{E})}_{\mathrm{Interaction}} \times 
              \underbrace{ T_f(\hat{E}) V_f(\hat{E}-E')}_{\mathrm{Detection}}
    \end{eqnarray*}
    \begin{flushright}
      \reference{P. Huber, M. Lindner, W. Winter, Nucl. Phys. B645:3-48, 2002, hep-ph/0204352}
    \end{flushright}
  \end{block}
  \small
  \begin{tabular}{l@{\ }c@{\ }ll@{\ }c@{\ }l}
    $E$              &=& Neutrino energy       & $P_{(f_i\rightarrow f)}(\ldots)$ &=& Oscillation probability      \\
    $\hat{E}$        &=& Charged Lepton energy & $\sigma^{\mathrm{IT}}_f(E)$      &=& Cross sections               \\
    $E'$             &=& Reconstructed energy  & $k_f^{\mathrm{IT}}(E-\hat{E})$   &=& Charged lepton distribution  \\
    $\Phi_{f_i} (E)$ &=& Source flux           & $T_f(\hat{E})$                   &=& Threshold function           \\
    $L$              &=& Baseline              & $V_f(\hat{E}-E')$                &=& Energy resolution function
  \end{tabular}
\end{frame}

\begin{frame}
  \frametitle{Sensitivity of a reactor experiment to $\sin^2 (2 \theta_{13})$}

  \begin{enumerate}
    \item Define what you want to measure: What 90 \% C.L. limit can the experiment set on
          $\sin^2 (2 \theta_{13})$, assuming the true value is 0?
    \item Define the setup using AEDL (Abstract Experiment Definition Language)
    \item Calculate the ``true'' event rates.
    \item Now calculate the expected event rates for nonzero values of $\sin^2 (2 \theta_{13})$.
    \item Perform a $\chi^2$ analysis to find out which values produce event rates that
          are still compatible with $\sin^2 (2 \theta_{13}) = 0$ at the chosen C.L.
  \end{enumerate}
\end{frame}

% -----------------------------------------------------------------------------
\subsection{$\chi^2$ Analysis}
% -----------------------------------------------------------------------------

\begin{frame}
  \frametitle{$\chi^2$ Analysis in GLoBES}

  \begin{tabular}{ll}
    \begin{minipage}[b]{6 cm}
      \begin{itemize}
        \item
        If $X_i$ are normally distributed with $\langle X_i \rangle = 0$ and $\sigma = 1$,
        then $\chi^2 = \sum_{i=1}^n X_i^2$ is distributed according to the $\chi^2$
        distribution with $n$ degrees of freedom.
      \end{itemize}
      \vspace{0.2 cm}
    \end{minipage}    &
    \includegraphics[height=3 cm]{fig/chi2_dist.pdf} \\

    \multicolumn{2}{l}{
    \begin{minipage}{10 cm}
      \begin{itemize}
        \item
        For $n$ energy bins with event rates following the Poisson distribution
        \vspace{-0.3 cm}
        \begin{equation*}
          \chi^2 = \sum_{i=1}^n 2 (N_i(\vec{\Lambda}) - N_i(\vec{\Lambda}_0))
                          + 2 N_i(\vec{\Lambda}) \log \frac{N_i(\vec{\Lambda})}{N_i(\vec{\Lambda}_0)}
          \vspace{-0.2 cm}
        \end{equation*}
        follows a $\chi^2$ distribution with $\dim \vec{\Lambda}$ degrees of freedom. Here: \\
        \vspace{0.3 cm}
        \begin{tabular}{lcl}
          $N_i(\vec{\Lambda}_0)$ &=& Event rates for true parameter values $\vec{\Lambda}_0$.  \\
          $N_i(\vec{\Lambda})$   &=& Event rates for hypothesised parameters $\vec{\Lambda}$.
        \end{tabular}
      \end{itemize}
    \end{minipage}}
  \end{tabular}
\end{frame}

\begin{frame}
  \frametitle{$\chi^2$ Analysis in GLoBES}
  \begin{block}{Inclusion of systematics and external input}
    \begin{eqnarray*}
      \chi^2 &=& \underbrace{\sum_{i=1}^n 
                    2 (N_i(\vec{\Lambda},\vec{x}) - N_i(\vec{\Lambda}_0,\vec{x}_0))
                          + 2 N_i(\vec{\Lambda},\vec{x})
                                   \log \frac{N_i(\vec{\Lambda},\vec{x})}{N_i(\vec{\Lambda}_0,\vec{x}_0)}}
                           _{\textnormal{Statistical part}}   \\
          && + \ \ \ \underbrace{\sum_{j=1}^{\dim{\vec{\Lambda}}}
                   \frac{(\vec{\Lambda}_j - \vec{\Lambda}_{0j})^2}{\sigma_{\vec{\Lambda}j}^2}}
                           _{\textnormal{External input}} \ \ \
             + \ \ \ \underbrace{\sum_{k=1}^{\dim{\vec{x}}} \frac{(\vec{x}_k - \vec{x}_{0k})^2}{\sigma_{\vec{x}k}^2}}
                           _{\textnormal{Systematics}}
    \end{eqnarray*}
  \end{block}
  \small
  \begin{tabular}{lcl}
    $\vec{\Lambda}, \vec{\Lambda}_0$  &=& True and hypothesised oscillation parameters             \\
    $\vec{x}, \vec{x}_0$              &=& Systematics parameters                                   \\
    $N_i(\vec{\Lambda},\vec{x})$      &=& Expected event rates                                     \\
    $\sigma_{\vec{\Lambda}j}$         &=& Standard errors of external parameter input              \\
    $\sigma_{\vec{x}k}$               &=& Standard errors of systematics parameters 
  \end{tabular}
\end{frame}

\begin{frame}
  \frametitle{$\chi^2$ Analysis in GLoBES}

  \begin{figure}
    \raisebox{0.5 cm}{\includegraphics[width=6 cm]{fig/chi2_1dim.pdf} \hspace{0.5 cm}}
    \includegraphics[width=5 cm]{fig/chi2_2dim.pdf}
  \end{figure}
  \begin{flushright}
    \reference{P. Huber, M. Lindner, W. Winter, Nucl. Phys. B645:3-48, 2002, hep-ph/0204352}
  \end{flushright}
\end{frame}

% =============================================================================
\section{Simulation of Future Oscillation Experiments}
% =============================================================================

% -----------------------------------------------------------------------------
\subsection{The R2D2 Setup}
% -----------------------------------------------------------------------------

\begin{frame}
  \frametitle{The $\bar{\nu}_e$ disappearance channel}

  \begin{figure}
    \includegraphics[width=12 cm]{fig/P_ee.pdf}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{The R2D2 Setup}

  \begin{figure}
    \includegraphics[width=11 cm]{fig/R2D2_geometry.pdf}
  \end{figure}
  \begin{flushright}
    \reference{P. Huber, M. Lindner, T. Schwetz, JHEP 0502:029, 2005, hep-ph/0411166}
  \end{flushright}
  \begin{itemize}
    \item Near and far detectors $\longrightarrow$ Elimination of flux normalization error
    \item Two reactors $\longrightarrow$ Elimination of detector systematics
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{The R2D2 Setup}

  \begin{figure}
    \includegraphics[height=7 cm]{fig/R2D2_s22th13.pdf}
  \end{figure}
  \begin{flushright}
    \reference{P. Huber, M. Lindner, T. Schwetz, JHEP 0502:029, 2005, hep-ph/0411166}
  \end{flushright}
\end{frame}

% -----------------------------------------------------------------------------
\subsection{Superbeams and Neutrino Factories}
% -----------------------------------------------------------------------------

\begin{frame}
  \frametitle{Superbeams}
  \begin{figure}
    \includegraphics[height=5 cm]{fig/cern_superbeam.jpg}
  \end{figure}
  \begin{flushright}
    \reference{J. Bouchez, CERN Courier 42, No. 5}
  \end{flushright}
  \begin{itemize}
    \item High neutrino energy $\longrightarrow$ Easy to detect
    \item $\nu_\mu$ beam with $\nu_e$, $\bar{\nu}_e$ and $\bar{\nu}_\mu$ contamination
          $\longrightarrow$ limited physics potential
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Neutrino Factories}
  \begin{figure}
    \vspace{-0.8 cm}
    \includegraphics[height=8 cm,angle=270]{fig/cern_nufact.pdf}
    \vspace{-0.2 cm}
  \end{figure}
  \begin{itemize}
    \item High neutrino energy $\longrightarrow$ Easy to detect
    \item Long Baselines $\longrightarrow$ Exploiting matter effects
    \item Beam contains only $\nu_\mu$ and $\bar{\nu}_e$
          $\longrightarrow$ ``Wrong sign muon'' signature:
          Detection of $\mu^+$ indicates oscillation
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{CP Violation in Superbeams and Neutrino Factories}
  \begin{figure}
    \includegraphics[height=4.5 cm]{fig/cpviolation.pdf}
  \end{figure}
  \begin{flushright}
    \reference{P. Huber, M. Lindner, W. Winter, Nucl. Phys. B645:3-48, 2002, hep-ph/0204352}
  \end{flushright}
  \begin{itemize}
    \item JHF-HK: $E_p$ = 50~GeV, L = 295 km, Target Power 4~MW, 1~Mt~Water \v{C}erenkov detector
    \item Nu-Fact II: $E_\mu$ = 50~GeV, L = 3000~km, Target Power: 4~MW,
                      $5 \cdot 10^{20} \frac{\mu}{\textnormal{year}}$, 50~kt~magnetized iron calorimeter
  \end{itemize}
\end{frame}

% -----------------------------------------------------------------------------
\subsection{Beta Beams}
% -----------------------------------------------------------------------------

\begin{frame}
  \frametitle{Beta Beams}
  \begin{figure}
    \includegraphics[height=5 cm]{fig/cern_betabeam.pdf}
  \end{figure}
  \begin{flushright}
    \reference{J. Bouchez, M. Lindroos, M. Mezzetto, AIP Conf. Proc. 721:37-47, 2004, hep-ex/0310059}
  \end{flushright}
  \begin{itemize}
    \item High neutrino energy $\longrightarrow$ Easy to detect
    \item Long Baselines $\longrightarrow$ Exploiting matter effects
    \item Pure $\nu_e$ or $\bar{\nu}_e$ beam
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Limitations of Neutrino Oscillation Experiments}
  \begin{figure}
    \includegraphics[height=5 cm]{fig/barplots.pdf}
  \end{figure}
  \begin{flushright}
    \reference{C. Albright et.al., physics/0411123}
  \end{flushright}
  \begin{itemize}
    \item Correlations: Sensitivity only to combination of parameters
    \item Degeneracies: Distinct regions in parameter space can explain the experiment
  \end{itemize}
\end{frame}

% =============================================================================
\section*{Conclusions}
% =============================================================================

\begin{frame}
  \frametitle{Conclusions}
  \begin{itemize}
    \item Neutrino oscillations are theoretically well understood and experimentally clearly
          verified.
    \item A precise determination of the oscillation parameters is desirable to achieve a deeper
          understanding of particle physics (flavour structure, $\ldots$).
    \item GLoBES is a tool for simulating future neutrino oscillation experiments and determine their
          physics potential.
    \item Nuclear reactors, Superbeams, Neutrino Factories and Beta Beams are excellent tools
          for solving many open questions.
  \end{itemize}
\end{frame}

\end{document}

### 2005-10-17 Documentation/part1.tex
- Added documentation for user-defined systematics functions (these changes   affect the chapters "Calculating chi^2 with systematics only" and   "Experiment definitions with AEDL -> Rules and the treatment of systematics" - Changed the names of several functions in user_chi.c to give them more   unambiguous names (glbGetSignalBin -> glbGetSignalBinInRule, glbGetBackgroundBin   -> glbGetBackgroundBinInRule, glbGetMeasuredBin -> glbGetObservedBinInRule). JK

@@
 
+\section{Built-in systematics}
+\index{norm}{Systematics!builtin functions}
+
 \example{Correlation between $\stheta$ and $\deltacp$}{
@@
 Keeping all oscillation parameters and matter density scaling factors fixed,
- one can use the following functions to obtain the total $\chi^2$ of all 
+ one can use the following builtin functions to obtain the total $\chi^2$ of all 
  specified oscillation channels including systematics:
@@
 spectral information. For details, we refer to \Chapt~\ref{chapt:running}.
+
+\section{User-defined systematics calculation}
+\index{norm}{Systematics!user-defined}
+\label{sec:userchi}
+
+For some experimental setups the built-in systematics functions are not sufficient
+because they cannot handle systematical errors that are correlated between different
+parts of a multi-detector or multi-source setup. Therefore \GLOBES\ allows the user
+to override the default $\chi^2$ calculation. The reader who wants to use this
+feature should be familiar with \Sec~\ref{sec:channel} and~\Sec~\ref{sec:rules}
+of the \AEDL\ manual.
+
+To specify a user-defined $\chi^2$ algorithm, the following function can be used:
+\begin{function}
+\GLBNS{SetUserChi}
+{\tt void glbSetUserChi(int exp,int rule,double (*chi\_func)(double x[]),int dim,%
+double centers[],double errors[],char *info)} tells \GLOBES\ to use {\tt chi\_func}
+to calculate $\chi^2$ for experiment {\tt exp} and rule {\tt rule} (these parameters
+can take the value \GLBC{GLB\_ALL} to specify all experiments or all rules). {\tt dim} is
+the number of variables parametrizing the systematical errors, while {\tt centers[]}
+and {\tt errors[]} give their central values and standard deviations, respectively.
+When \GLOBES\ calls {\tt chi\_func}, the parameter array {\tt x[]} will contain
+the values of the systematics parameters for which $\chi^2$ should be calculated.
+Deviating from the standard conventions, the entries of {\tt x[]} will start at
+index 1.
+\end{function}
+
+In addition to calling {\tt glbSetUserChi}, {\tt @errordim\_sys\_on} has to be set
+to 20 in the \AEDL\ file (see~\Sec~\ref{sec:rules}).
+
+Note that {\tt chi\_func} will be called many times by the \GLOBES\ minimizers.
+Therefore the function should be as efficient as possible. The function
+will need low-level access to internal \GLOBES\ data. Therefore the following
+routines are provided:
+\begin{function}
+\GLBNS{GetCurrentExp}
+{\tt int glbGetCurrentExp()} gives the number of the experiment for which
+{\tt chi\_func} has been called.
+\end{function}
+\begin{function}
+\GLBNS{GetCurrentRule}
+{\tt int glbGetCurrentRule()} gives the number of the rule in the current experiment
+for which {\tt chi\_func} has been called.
+\end{function}
+\begin{function}
+\GLBNS{GetObservedBinInRule}
+{\tt double glbGetObservedBinInRule(int exp,int rule,int bin)} returns the number of
+signal events for experiment {\tt exp}, rule {\tt rule} and bin {\tt bin}, as calculated
+in the preceding call to {\tt glbSetRates}.
+\end{function}
+\begin{function}
+\GLBNS{GetSignalBinInRule}
+{\tt double glbGetSignalBinInRule(int exp,int rule,int bin)} returns the number of
+signal events for experiment {\tt exp}, rule {\tt rule} and bin {\tt bin}, calculated
+for the oscillation parameter values that are currently being tested.
+\end{function}
+\begin{function}
+\GLBNS{GetBackgroundBinInRule}
+{\tt double glbGetBackgroundBinInRule(int exp,int rule,int bin)} returns the number of
+background events for experiment {\tt exp}, rule {\tt rule} and bin {\tt bin}, calculated
+for the oscillation parameter values that are currently being tested.
+\end{function}
+\begin{function}
+\GLBNS{GetChannelBin}
+{\tt double glbGetChannelBin(int exp,int channel,int bin)} returns the number of
+signal events for experiment {\tt exp}, channel {\tt channel} and bin {\tt bin}, calculated
+for the oscillation parameter values that are currently being tested.
+\end{function}
+\begin{function}
+\GLBNS{ShiftSignalEnergyScale}
+{\tt int glbShiftSignalEnergyScale(int exp,int rule,double amount)} applies an energy
+calibration error of magnitude {\tt amount} to the signal event rates in experiment
+{\tt exp} and rule {\tt rule}. Subsequent calls to {\tt glbGetSignalBinInRule} will
+return the modified event rates. For details on energy calibration errors, see
+\Sec~\ref{sec:rules}.
+\end{function}
+\begin{function}
+\GLBNS{ShiftBackgroundEnergyScale}
+{\tt int glbShiftBackgroundEnergyScale(int exp,int rule,double amount)} applies an energy
+calibration error of magnitude {\tt amount} to the background event rates in experiment
+{\tt exp} and rule {\tt rule}. Subsequent calls to {\tt glbGetBackgroundBinInRule} will
+return the modified event rates. For details on energy calibration errors, see
+\Sec~\ref{sec:rules}.
+\end{function}
+
+Typical applications for user-defined systematics are reactor neutrino experiments,
+where systematics play a crucial role. For a setup with near and far detectors,
+the formula for $\chi^2$ is
+\begin{align}
+  \chi^2 =& \sum_{i=1}^{\textrm{\# of bins}} \sum_{d = N,F}2
+     \left( O_{d,i} - (1 + a_R + a_d) T_{d,i} +
+       (1 + a_R + a_d) T_{d,i} \log \frac{(1 + a_R + a_d) T_{d,i}}{O_{d,i}}  \right) \nonumber\\
+         & \ \ + \frac{a_R^2}{\sigma_R^2} + \frac{a_N^2}{\sigma_N^2} + \frac{a_F^2}{\sigma_F^2}
+\label{equ:chi2reactor}
+\end{align}
+where $O_{N,i}$ and $O_{F,i}$ are the event rates for the $i$-th bin in the near and
+far detector, calculated for the assumed ``true'' values of the oscillation parameters.
+$T_{d,i}$ are the event rates for the parameter values that are currently being tested.
+$a_R$, $a_N$ and $a_F$ parametrize the small uncertainties in the reactor flux and the fiducial
+mass of the two detectors. In this example their central values are assumed to be zero,
+while their standard deviations are $\sigma_R$, $\sigma_N$ and $\sigma_F$.
+
+The first line of \equ{chi2reactor} is the standard $\chi^2$ expression for
+the Poisson distribution, while the terms in the second line are penalties
+for deviations of the systematics parameters from their central values.
+
+\example{Systematics for a reactor setup with near and far detector}{
+The following code fragment calculates $\chi^2$ for a reactor setup with
+a far detector (experiment 0) and a near detector (experiment 1),
+including six different systematical errors: the flux normalization
+of the reactor ({\tt x[1]}), the fiducial mass uncertainties of the far
+({\tt x[2]}) and near ({\tt x[3]}) detector, the normalization of the
+background ({\tt x[4]}) and energy calibration errors for the far
+({\tt x[5]}) and near ({\tt x[6]}) detectors. We assume that the central
+values and errors of these parameters are stored in the global arrays
+{\tt sys\_centers} and {\tt sys\_errors}. The calculation follows
+\equ{chi2reactor}.
+
+\begin{quote}{\tt \footnotesize
+  double chi2nearfar(double *x)                                                 \\
+  \{                                                                            \\
+\hspace*{0.5 cm}      double chi2 = 0;                                          \\
+\hspace*{0.5 cm}      double chirate, measured\_rate;                           \\
+\hspace*{0.5 cm}                                                                \\
+\hspace*{0.5 cm}      glbShiftSignalEnergyScale(0, 0, x[5]);   /* Apply E calib. error */ \\
+\hspace*{0.5 cm}      glbShiftSignalEnergyScale(1, 0, x[6]);                    \\
+\hspace*{0.5 cm}      glbShiftBackgroundEnergyScale(0, 0, x[5]);                \\
+\hspace*{0.5 cm}      glbShiftBackgroundEnergyScale(1, 0, x[6]);                \\
+\hspace*{0.5 cm}      for (int i=0; i < NBINS; i++) \{                          \\
+\hspace*{1.0 cm}        /* Statistical part of chi\textasciicircum2 for far detector */ \\
+\hspace*{1.0 cm}        chirate = (x[1]+x[2]) * glbGetSignalBinInRule(0,0,i) +  \\
+\hspace*{2.0 cm}                        (x[4]+x[3]) * glbGetBackgroundBinInRule(0,0,i);\\
+\hspace*{1.0 cm}        measured\_rate = glbGetObservedBinInRule(0,0,i);        \\
+\hspace*{1.0 cm}        chi2 += 2 * (measured\_rate - chirate);                 \\
+\hspace*{1.0 cm}        if (measured\_rate > 0  \&\&  chirate > 0)              \\
+\hspace*{1.5 cm}          chi2 += 2 * chirate * log(chirate / measured\_rate);  \\
+\hspace*{1.0 cm}                                                                \\
+\hspace*{1.0 cm}        /* Statistical part of chi\textasciicircum2 for near detector */ \\
+\hspace*{1.0 cm}        chirate = (x[1]+x[3]) * glbGetSignalBinInRule(1,0,i) +  \\
+\hspace*{2.0 cm}                        (x[4]+x[3]) * glbGetBackgroundBinInRule(1,0,i);\\
+\hspace*{1.0 cm}        measured\_rate = glbGetObservedBinInRule(1,0,i);        \\
+\hspace*{1.0 cm}        chi2 += 2 * (measured\_rate - chirate);                 \\
+\hspace*{1.0 cm}        if (measured\_rate > 0  \&\&  chirate > 0)              \\
+\hspace*{1.5 cm}          chi2 += 2 * chirate * log(chirate / measured\_rate);  \\
+\hspace*{0.5 cm}      \}                                                        \\
+\hspace*{0.5 cm}                                                                \\
+\hspace*{0.5 cm}      for (int j=0; j < 6; j++)  /* Penalties */                \\
+\hspace*{1.0 cm}        chi2 += square((x[j+1] - sys\_centers[j]) / sys\_errors[j]);\\
+\hspace*{0.5 cm}                                                                \\
+\hspace*{0.5 cm}      /* The function is called for each exp ->
+                                       Assign chi\textasciicircum2/2 to each */ \\
+\hspace*{0.5 cm}      return chi2 / 2;                                          \\
+  \}                                                                            
+}
+\end{quote}
+}
 

### 2005-10-17 Documentation/part2.tex
- Added documentation for user-defined systematics functions (these changes   affect the chapters "Calculating chi^2 with systematics only" and   "Experiment definitions with AEDL -> Rules and the treatment of systematics" - Changed the names of several functions in user_chi.c to give them more   unambiguous names (glbGetSignalBin -> glbGetSignalBinInRule, glbGetBackgroundBin   -> glbGetBackgroundBinInRule, glbGetMeasuredBin -> glbGetObservedBinInRule). JK

@@
 \end{quote}
-It is foreseen to add the possiblity to extend the set 
-of error dimensions or the set of possible systematical errors in one
-of the next versions of \GLOBES.
+If none of the built-in algorithms listed in \Tab~\ref{tab:error_dim} is
+suitable for a certain setup, the user can override them by setting
+{\tt @errordim\_sys\_on} to 20 and using the \GLB{SetUserChi}
+function to specify a specialized function for the calculation
+of $\chi^2$. This function will then be called by the \GLOBES\ minimizers.
+For details, see \Sec~\ref{sec:userchi}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 7& $\infty$ &-& $\infty$ &-&-&Spectrum only\\
-8&-&-&-&-&-&Total rates, No systematics\\
+8&-&-&-&-&-&Total rates, no systematics\\
 9&+&+&+&+&C&Systematics with calibration\\
+20&?&?&?&?&?&User-defined systematics\\
 \hline

### 2005-10-17 Distribution/globes/globes_perl.c

@@ -1,6 +1,6 @@
 /* ----------------------------------------------------------------------------
  * This file was automatically generated by SWIG (http://www.swig.org).
- * Version 1.3.21
+ * Version 1.3.25
  * 
  * This file is not intended to be easily readable and contains a number of 
  * coding conventions designed to improve portability and efficiency. Do not make
@@ -8,70 +8,112 @@
  * interface file instead. 
  * ----------------------------------------------------------------------------- */
 
-/*************************************************************** -*- c -*-
- * perl5/precommon.swg
+/***********************************************************************
  *
- * Rename all exported symbols from common.swg, to avoid symbol
- * clashes if multiple interpreters are included
+ *  This section contains generic SWIG labels for method/variable
+ *  declarations/attributes, and other compiler dependent labels.
  *
  ************************************************************************/
 
-#define SWIG_TypeRegister    SWIG_Perl_TypeRegister
-#define SWIG_TypeCheck       SWIG_Perl_TypeCheck
-#define SWIG_TypeCast        SWIG_Perl_TypeCast
-#define SWIG_TypeDynamicCast SWIG_Perl_TypeDynamicCast
-#define SWIG_TypeName        SWIG_Perl_TypeName
-#define SWIG_TypeQuery       SWIG_Perl_TypeQuery
-#define SWIG_TypeClientData  SWIG_Perl_TypeClientData
-#define SWIG_PackData        SWIG_Perl_PackData 
-#define SWIG_UnpackData      SWIG_Perl_UnpackData 
+/* template workaround for compilers that cannot correctly implement the C++ standard */
+#ifndef SWIGTEMPLATEDISAMBIGUATOR
+#  if defined(__SUNPRO_CC) && (__SUNPRO_CC <= 0x560)
+#    define SWIGTEMPLATEDISAMBIGUATOR template
+#  else
+#    define SWIGTEMPLATEDISAMBIGUATOR 
+#  endif
+#endif
+
+/* inline attribute */
+#ifndef SWIGINLINE
+# if defined(__cplusplus) || (defined(__GNUC__) && !defined(__STRICT_ANSI__))
+#   define SWIGINLINE inline
+# else
+#   define SWIGINLINE
+# endif
+#endif
+
[…]

### 2005-10-21 Documentation/Paper2/paper2.tex
New paper draft added [WW]

--- 2004-07-20 Documentation/Paper/paper.tex
@@
 \begin{flushright}
-TUM-HEP-5XX/04\\
+TUM-HEP-YYY/0X\\
 %hep-ph/
@@
 {\begin{center}
-{\Large\bf Simulation of long-baseline neutrino oscillation
-experiments with GLoBES}
+{\Large\bf New features in the simulation of long baseline experiments with GLoBES 3.0 (working title)}
 \end{center}}
@@
 {\begin{center} {\large{\sc
-                P.~Huber\footnote[1]{\makebox[1.cm]{Email:}
-                phuber@ph.tum.de},~
-                M.~Lindner\footnote[2]{\makebox[1.cm]{Email:}
-                lindner@ph.tum.de},~and~
-                W.~Winter\footnote[3]{\makebox[1.cm]{Email:}
-                wwinter@ph.tum.de}
+                P.~Huber\footnotemark[1],~
+                J.~Kopp\footnotemark[2],~
+                M.~Lindner\footnotemark[2],~
+                M.~Rolinec\footnotemark[2],~
+                W.~Winter\footnotemark[3]
                 }}
+
+\footnote{All correspondence should be addressed to {\tt globes@ph.tum.de}}
+
 \end{center}}
@@
 
-\footnotemark[1]${}^,$\footnotemark[2]${}^,$\footnotemark[3]%
-       Institut f\"ur Theoretische Physik, Physik--Department,\\
-       Technische Universit\"at M\"unchen,
-       James--Franck--Strasse, D--85748 Garching, Germany
+\footnotemark[1]%${}^,$\footnotemark[2]%
+       Department of Physics, University of Wisconsin, \\
+       1150 University Avenue, Madison, WI 53706, USA
+
+\vspace*{1mm}
+
+\footnotemark[2]%
+       Physik--Department, Technische Universit\"at M\"unchen, \\
+       James--Franck--Strasse, 85748 Garching, Germany
+
+\vspace*{1mm}
+
+\footnotemark[3]%
+       School of Natural Sciences, Institute for Advanced Study, \\
+       Einstein Drive, Princeton, NJ 08540, USA
 
@@
 \begin{abstract}
-We present the \GLOBES\ (``General Long Baseline Experiment Simulator'')
-software package, which allows the simulation of long-baseline and reactor
-neutrino oscillation experiments. One part of the software is the
-abstract experiment definition language to define experiments with
-beam and full detector descriptions as accurate as possible. Many systematics options are provided, such as normalization and energy
-calibration errors, or the choice between spectral or total rate 
-information. For the 
-definition of experiments, a new transparent building block concept is 
-introduced. In addition, an additional program provides the possibility 
-to develop and test new experiment definitions quickly.
-%
-Another
-part of \GLOBES\ is the user's interface, which provides probability, rate, 
-and $\Delta \chi^2$ information for a given experiment or any
-combination of up to 32 experiments in C. Especially, the
-$\Delta \chi^2$ functions allow a simulation with statistics only,
-systematics, correlations, and degeneracies. In particular, \GLOBES\
-can handle the full multi-parameter correlation among the oscillation
-parameters, external input, and matter density uncertainties. 
+Here comes the abstract.
 \end{abstract}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%                     Introduction                                  %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \section{Introduction}
 
-Neutrino oscillations are now established as the leading flavor
-transition mechanism for neutrinos. In a long history of many experiments, see
-\eg~\cite{Barger:2003qi}, 
-two oscillation
-frequencies have been identified: The fast ``atmospheric'' and the
-slow ``solar'' oscillations, which are driven by the respective
-mass squared differences.  In addition, there could be interference
-effects between these two oscillations, provided that the coupling given by 
-the small mixing angle $\theta_{13}$ is large enough. Such interference
-effects include, for example, leptonic CP violation.  In order to
-test the unknown oscillation parameters, \ie, the mixing angle $\theta_{13}$,
- the leptonic CP phase, and the
-neutrino mass hierarchy, new long-baseline and reactor experiments 
-are proposed.
-These experiments send an artificial neutrino beam to a detector, or
-detect the neutrinos produced by a nuclear fission reactor. However,
-the presence of multiple solutions which are intrinsic to
- neutrino oscillation probabilities~\cite{Fogli:1996pv,
-Burguet-Castell:2001ez,Minakata:2001qm,Barger:2001yr} 
-affect these measurements.
-Thus optimization strategies are required which maximally exploit 
-complementarity between experiments. Therefore, a modern, complete experiment 
-simulation  and analysis tool does not only need to have 
-a highly accurate beam and detector simulation, but also powerful 
-means to analyze correlations and degeneracies, especially for the combination
-of several experiments.
-In this paper, we present the \GLOBES\ software package
-as such a tool.
+\nocite{Huber:2004ka}
 
-\GLOBES\ (``General Long Baseline Experiment Simulator'') is a flexible
-software tool to simulate and analyze neutrino oscillation 
-long-baseline and reactor experiments using a 
-complete three-flavor description. On the
-one hand, it contains a comprehensive abstract experiment definition
-language (AEDL), which allows to describe 
-most classes of long baseline and reactor experiments
-at an abstract level. On the other hand, it provides a C-library to 
-process the experiment information in order to obtain oscillation
-probabilities, rate vectors, and $\Delta \chi^2$-values (\cf, \figu{GLOBES}). 
-In addition, it provides a binary program to test experiment
-definitions very quickly, before they are used by the application software.
-Currently, \GLOBES\ is available for GNU/Linux. Since the source code is 
-included, the modifications to other operating systems should be doable.
+\section{User-defined priors and systematics}
 
-\GLOBES\ allows to simulate experiments with stationary neutrino point 
-sources, where each experiment is assumed to have only one neutrino source.
-Such experiments are neutrino beam experiments and reactor experiments. 
-Geometrical effects of a source distribution, such as in the sun or the 
-atmosphere, can not be described. In addition, sources with a physically 
-significant time dependencies  can not be studied, such as  supernov\ae.
+\bi
+ \item
+  User-defined priors
+ \item
+  User-defined systematics
+\ei
 
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=14cm]{GLOBES}
-\end{center}
-\caption{\label{fig:GLOBES} General concept of the \GLOBES\ package.}
-\end{figure}
+\section{Simulation of non-standard physics with GLoBES}
 
-On the experiment definition side, either built-in neutrino fluxes
-(\eg, neutrino factory) or arbitrary, user defined fluxes can be used. 
-Similarly,
-arbitrary cross sections, energy dependent efficiencies,
-energy resolution functions as well as the considered oscillation channels, 
-backgrounds, and many other properties can be specified. 
-For the systematics, energy
-normalization and calibration errors can be simulated. Note that
-energy ranges and windows and bin widths can be
-(almost) arbitrarily chosen, including bins of different
-widths. Together with \GLOBES\ comes a number of
-pre-defined experiments in order to demonstrate the capabilities
-of \GLOBES\ and to provide prototypes for new experiments.
-In addition, they can be used to test new physics ideas with
-complete experiment simulations.
-Examples for these prototypes are the \MINOS , \ICARUS , and
-\OPERA\ simulations from \Ref~\cite{Huber:2004ug}, the
-\JHFSK\ and \NUMI\ superbeam simulations from \Refs~\cite{Huber:2002mx,
-Huber:2002rs}, the \JHFHK\ superbeam upgrade simulation from
-\Ref~\cite{Huber:2002mx}, the neutrino factory simulations from
-\Ref~\cite{Huber:2002mx}, and the reactor experiment simulations from
-\Ref~\cite{Huber:2003pm}. Other projects using earlier versions
-of the \GLOBES\ software include \Refs~\cite{Huber:2003ak,Ohlsson:2003ip,Winter:2003ye,Antusch:2004yx}.
+HOWTO
 
-With the C-library, one can extract the $\Delta \chi^2$ for all defined 
-oscillation channels for an experiment or any combination of experiments.
-Of course, also low-level information, such as oscillation
-probabilities or event rates, can be obtained. \GLOBES\ includes the
-simulation of neutrino oscillations in matter with arbitrary matter 
-density profiles, as well as it allows to simulate the matter density
-uncertainty. As one of the most
-advanced features of \GLOBES , it provides the technology to 
-project the $\Delta \chi^2$, which is a function of all oscillation
-parameters, onto any subspace of parameters by local minimization. 
-This approach allows the inclusion of multi-parameter-correlations,
-where external constraints (\eg, on the solar parameters) can be imposed, too.
-Applications of the projection mechanism include the projections onto 
-the $\stheta$-axis and the $\stheta$-$\deltacp$-plane. In addition, 
-all oscillation parameters can be kept free to precisely localize 
-degenerate solutions.
+\section{Simulation of new experiment features}
 
-\section{The computation of raw event rates}
+\bi
+\item 
+ Built-in $\beta$-beam fluxes
+\item
+ New AEDL features
+\item
+ New AEDL files in standard GLoBES delivery + changes to existing ones
+\ei
 
-In this section, we sketch the computation of 
-the event rates, which is one of the core parts of the \GLOBES\
-software. However, because of the complexity of this issue,
-we refer to the \GLOBES\ manual~\cite{Manual} for
-details.
+\section{New platform and interpreter support}
 
-The differential event rate for each channel is given by
-\begin{eqnarray}
-\label{equ:master_event}
-\frac{dn_{\beta}^{\text{IT}}}{dE'}=&&N\,\int\limits_0^\infty \int\limits_0^\infty dE\,d\hat{E}\quad
-\underbrace{\Phi_{\alpha} (E)}_{\mathrm{Production}} \times \nonumber\\
-&&\underbrace{\frac{1}{L^2} P_{(\alpha\rightarrow\beta)}(E,L,\rho;\theta_{23},
-\theta_{12},\theta_{13},
-\Delta m^2_{31},\Delta m^2_{21},\deltacp)}_{\mathrm{Propagation}}
-\times \nonumber \\ &&\underbrace{\sigma^{\text{IT}}_f(E)
-k_f^{\text{IT}}(E-\hat{E})}_{\mathrm{Interaction}} \times \nonumber \\
-&&\underbrace{ T_f(\hat{E}) V_f(\hat{E}-E')}_{\mathrm{Detection}}\,,
-\end{eqnarray}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-where $\alpha$ is the initial flavor of the neutrino, 
-$\beta$ is the final flavor, $\Phi_{\alpha} (E)$ is the flux of the 
-initial flavor at the
-source, $L$ is the baseline length, $N$ is a normalization factor, and 
-$\rho$ is the matter density. The energies in this formula are given as follows:
-\begin{itemize}
+\bi
 \item
- $E$ is the incident neutrino energy, \ie, the actual energy of the 
-incoming neutrino (which is not directly accessible to the experiment)
+ PERL?
 \item 
- $\hat{E}$ is the energy of the secondary particle
+ Windows?
 \item
- $E'$ is the reconstructed neutrino energy, \ie, the measured
-neutrino energy as obtained from the experiment
-\end{itemize}
-The interaction term is composed of 
-two factors, which are the total cross section 
-$\sigma^{\text{IT}}_\beta(E)$ for the flavor $f$ and
-the interaction type IT, and the energy distribution of the 
-secondary particle $k_\beta^{\text{IT}}(E-\hat{E})$.
-The detector properties are 
-modeled by the threshold function $T_\beta(\hat{E})$, coming from the the 
-limited resolution or the cuts in the analysis, and the energy resolution 
-function $V_\beta(\hat{E}-E')$ of the secondary particle.
+ Better RedHat support
+\item
+ Changes in installation scheme
+\ei
 
-Since it is a lot of effort to solve this double integral numerically,
-we split up the two integrations. First, we evaluate the integral over
-$\hat{E}$, where the only terms containing $\hat{E}$ are
-$k_\beta^{\text{IT}}(E-\hat{E})$,  $ T_\beta(\hat{E})$, and 
-$ V_\beta(\hat{E}-E')$. We define:
-\begin{eqnarray}
-\label{equ:e_res} 
-R_\beta^{\text{IT}}(E,E')\,\epsilon_\beta^{\text{IT}}(E')
- \equiv
-\int\limits_0^\infty d\hat{E} \quad T_\beta(\hat{E})\,k_\beta^{\text{IT}}(E-\hat{E})
-\,V_\beta(\hat{E}-E')\,. 
-\end{eqnarray}
-Thus, $R_\beta^{\text{IT}}(E,E')$ describes the energy response of 
-the detector, \ie , a neutrino with a (true) energy $E$ is reconstructed
-with an energy between $E'$ and $E'+dE'$ with a probability
-$R_\beta^{\text{IT}}(E,E') dE'$. The function $R(E,E')$ is also often called ``energy resolution function''. Actually, its internal representation
-in the software is a smearing matrix.
-The function $\epsilon_\beta^{\text{IT}}(E')$ will be refered to as ``post-smearing efficiencies'', since it will allow us to define cuts and threshold functions {\em after} the smearing is performed, \ie, as function of $E'$.
-In addition, \GLOBES\ uses ``pre-smearing efficiencies'', which are 
-evaluated {\em before} the smearing is performed, \ie, as function of $E$.
-Similarly, constant\footnote{With respect to the oscillation parameters, 
-not the energy.} background rates can be added to the event rates before or 
-after the energy smearing, which are refered to as 
-``pre-smearing backgrounds'' and ``post-smearing backgrounds''. These types of 
-efficiencies and backgrounds allow a very accurate modeling of
-many factors, such as atmospheric or geoneutrino backgrounds, energy cuts, or
-energy threshold functions.
+\section{New ways to improve performance}
 
-Eventually, we can write down the number of events per bin $i$ \index{Bin} and channel $c$ as
-\begin{equation}
-\label{equ:channel}
-n_i^c=\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' \quad
-\frac{dn_{\beta}^{\text{IT}}}{dE'} (E') \,
-\end{equation}
-where $\Delta E_i$ is the bin size of the $i$th energy bin.
-This means that one has to solve the integral
-\begin{eqnarray}
-\label{equ:events_bin}
-n_i^c=N/L^2\,\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
-\quad \int\limits_0^\infty dE \,\, \Phi^c(E)\,
-P^c(E)\,
-\sigma^c(E)\,
-R^c(E,E')\,
-\epsilon^c(E')\,,
-\end{eqnarray} 
-which gives the raw event rates of the channel $c$ in the $i$th bin.
-Note that the events are binned according to their \emph{reconstructed} energy.
-
-Core part of the event rate computation is the energy smearing algorithm
-to evaluate \equ{events_bin}, where either a particular energy resolution
- function can be chosen for automated energy smearing, or the discretized
-smearing matrix $R^c(E,E') = R^c_{ij} $ can be given manually. In addition, it is possible
-to define a low-pass filter to avoid aliasing effects for neutrino
-oscillations faster than the sampling width.
-
-\section{Definition of experiments with \AEDL\\ (Abstract Experiment 
-Definition Language)}
-
-\begin{figure}[t]
-\begin{center}
-\begin{tabular}{p{6.3cm}p{4.3cm}p{4.3cm}}
-\includegraphics[width=6.5cm]{AEDL1} &
-\includegraphics[width=4.5cm]{SignalBackground} &
-\includegraphics[width=4.5cm]{Rules} \\
-\end{tabular}
-\end{center}
-\caption{\label{fig:aedl} The most important components of \AEDL :
-Channels, rules, and experiments.}
-\end{figure}
-
-In order to define experiments, \GLOBES\ uses \AEDL\ (``Abstract
-Experiment Definition Language''). An experiment normally corresponds
-to an \AEDL\ file, which is a human readable text file written 
-in \AEDL\ syntax.
-
-The key concept of \AEDL\ is to regard a detector as an abstract system
-which maps the true properties of a neutrino onto the reconstructed properties
-of the neutrino. The latter are subsequently used in the fit of the oscillation
-parameters. Within \GLOBES , only energy and flavor are  
-observables, which is
-sufficient for long-baseline and reactor experiments. In  other
-experiments, such as atmospheric neutrino experiments, more observables 
-(\eg , direction) may have to be considered.
-
-The main components of any \AEDL\ experiment definition are ``channels'',
-``rules'', and ``experiments'' (\cf, \figu{aedl}).
-A channel corresponds to a neutrino oscillation channel including flux, cross section (for one specific interaction type), energy resolution function, initial and final neutrino flavors, their polarity (neutrinos or antineutrinos), and efficiencies.
-The channel raw event rates are computed according to \equ{events_bin}.
-Each channel leads to the raw event rates for a
-specific interaction type. In \AEDL , many of the channel components
-have to be defined or loaded from files before, such as fluxes, 
-cross sections, or the energy resolution function. For the
-different options, see the \GLOBES\ manual~\cite{Manual}.
-
-For a ``rule'', the raw event rates of one or more signal channels and
-one or more background channels are added. The splitting in signal and
-background is arbitrary, but all of the signal and background components
-are defined to have a common systematics. The event rates of all 
-signal and background components add up to the total event rate of the
-rule, which leads to a $(\Delta \chi^2)_r$. The signal or background
-within each rule allows the specification of signal and background 
-normalization errors and energy tilt or calibration errors. 
-These systematical errors are evaluated with the ``pull method'' such as 
-in \Refs~\cite{Huber:2002mx,Fogli:2002pt}.  
-In addition to these
-systematics errors, an overall evaluation strategy is assigned 
-to each rule, which specifies the type of systematics (tilt or calibration
-error), and the use of spectral information or total event rates.
-
-Finally, one ore more rules add up to an experiment, where the
-total $\Delta \chi^2$ is obtained as the sum of the $(\Delta \chi^2)_r$
-of all rules. This approach allows the definition of 
-appearance and disappearance channels, neutrino and antineutrino running, 
-or interaction types with different systematics (spectral information
-versus counting rate) within one experiment. The \GLOBES\ user interface
-allows the simulation of one or more experiment simultaneously, which
-means that one could also use different experiments for different
-oscillation channels. However, there is still one component missing
-in the experiment definition, which is the matter density profile.
-For an experiment, an arbitrary matter density profile can be specified, 
-which is evaluated with the evolution operator method~\cite{Ohlsson:1999um}.
-In addition, the matter density profile is multiplied by a scaling factor, 
-which is treated as an independent parameter
-with a relative (matter density) uncertainty. With this
-approach, one can take into account that the matter density profile along a 
-specific baseline is only known to about $3\%-5\%$. Thus, for one
-particular baseline, all rules should be defined within one experiment.
-
-\section{Analysis of experiments: The C user's interface}
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=16cm]{projex}
-\end{center}
-\mycaption{\label{fig:projex} Left plot: Example for a  correlation between $\stheta$ and $\deltacp$ (for 1 d.o.f. only, un-shown oscillation
-parameters fixed).
-Right plot: The $\chi^2$-value of the projection onto the $\stheta$-axis as function of $\stheta$.
-The projection onto the  $\stheta$-axis is obtained by finding the minimum $\chi^2$-value for each fixed value of $\stheta$ in the left-hand plot, \ie, along the gray vertical lines.
-The thick gray curve marks the position of these minima in the left-hand plot.
-The arrows mark the obtained fit ranges for $\stheta$ at the $3 \sigma$ confidence level (1 d.o.f.), \ie , the precision of $\stheta$.}
-\end{figure}
-
-In order to use \GLOBES , a C-library is linked with the application
-software. This library provides the user's interface functions.
-It allows to load \AEDL\ files, initialize the experiments, and
-have access to various $\Delta \chi^2$-functions including any combination
-of systematics and correlations. In addition, it provides low-level
-access to oscillation probabilities and event rates, and allows
-the readout and modification of many experiment parameters at run time.
-
-The most sophisticated feature of the user's interface is the
-possibility to include the full multi-parameter correlation among
-all oscillation parameters. Together with the matter densities
-of the $N_{\mathrm{exp}}$ loaded experiments ($N_{\mathrm{exp}} \ge 1$), the neutrino oscillation parameter space in \GLOBES\ has $6+N_{\mathrm{exp}}$ dimensions. In the $\chi^2$-approach,
-the allowed region of $k$ ($1\le k<6+N_{\mathrm{exp}}$) fit parameters (at the chosen confidence level) is obtained as the projection of the $6+N_{\mathrm{exp}}$-dimensional fit manifold onto the $k$-dimensional hyperplane.
-For example,
-for the precision of the parameter $\stheta$ only, the fit manifold
-needs to be projected onto the $\stheta$-axis. One can easily imagine
-that the computation of the full multi-parameter correlation is
-very expensive with a grid-based method. However, the topology
-of the neutrino oscillation parameter space is well investigated and
-rather smooth. Therefore, it is feasible to use local minimization for each
-solution in the parameter space.  For this process, 
-a $6+N_{\mathrm{exp}}$-dimensional local minimizer
-is started at the ``guessed'' position for the solution
-(such as from analytical knowledge) and then runs into a
-local minimum. This exercise has to be repeated till all local 
-minima, \ie\ degenerate solutions, are found. In \figu{projex},
-this minimization is illustrated at the simple example of a 
-two-parameter correlation at a neutrino factory.
-
-As we have indicated above, the matter density of
-each experiment is treated as an additional parameter, which is known
-from external measurement with a certain precision. In principle, 
-the \GLOBES\ user's interface allows such externally imposed
-precisions for all oscillation parameters. This feature turns out
-to be especially useful for input from the solar neutrino experiments,
-since they provide better precisions for the solar neutrino
-oscillation parameters $\sdm$ and $\theta_{12}$ than long-baseline
- experiments themselves.
-
-\section{Complexity and computational issues}
-
-The \GLOBES\ software only uses polynomial algorithms, which are,
-however, suffering from the large number of dimensions.
-The $\Delta \chi^2$ computation basically consists of two steps:
-The systematics $(\Delta \chi^2)_{\mathrm{sys}}$ computation with {\tt glbChiSys},
-and the projection algorithm which uses {\tt glbChiSys} many times. 
-The overall computation time is obtained as the product of these two steps.
-
-The run time of the systematics $(\Delta \chi^2)_{\mathrm{sys}}$ 
-is 
-\begin{equation}
-\mathrm{t_{sys}} = \mathcal{O} \left( N_{\mathrm{Sampling}} \times N_{\mathrm{Bins}}  \, + \, N_{\mathrm{Sampling}} \times N_{\mathrm{Layers}} \right) ,
-\end{equation}
-since the oscillation probabilities for all channels and energies
-are only computed once and stored in a list.
-Here $N_{\mathrm{Sampling}}$ is the number of sampling points for the evaluation of the right integral in \equ{events_bin}, $N_{\mathrm{Bins}}$ is the number of energy bins,
-and $N_{\mathrm{Layers}}$ is the number of matter density layers. Thus, whenever the
-number of matter density steps is small, the product of sampling points
-and bins dominates. In practice, the $N_{\mathrm{Sampling}} \times N_{\mathrm{Bins}}$ energy smearing matrix is already computed at 
-the experiment initialization in order to save run time. In addition,
-since it contains a lot of zeros, \GLOBES\ uses a special format to
-avoid looping over the zeros.
-
-The run time for the projected $\Delta \chi^2$ would, for a grid based method, be
-\begin{equation}
- \mathrm{t_{proj}^{Grid}} = \mathcal{O}\left( n^{6+N_{\mathrm{exp}}}\right) \, \times \, \mathrm{t_{sys}},
-\end{equation}
-where $n$ is the number of grid points to be evaluated for each oscillation
-parameter, and $N_{\mathrm{exp}}$ is the number of experiments
-(since the matter densities of all experiments appear as parameters). 
-Though it is, in principle, possible to use such a method with \GLOBES ,
-it is impracticable in most cases, since it would involve at least 
-billions of
-steps for the complete multi-parameter correlation. The local $6+N_{\mathrm{exp}}$-dimensional minimization reduces this effort
-to 
-\begin{equation}
-\mathrm{t_{proj}^{Minimizer}} = \mathcal{O}(1000) \, \times \,
-N_{\mathrm{Deg}} \,  \times \, \mathrm{t_{sys}}
-\end{equation}
-steps, where $N_{\mathrm{Deg}}$ is the number of (disconnected) 
-local minima, \ie, the number of
-degeneracies (typically $\mathcal{O}(10)$).
-
-Eventually, \GLOBES\ needs about $10$ to $15$ seconds to compute
-a projected $\Delta \chi^2$ on a modern Pentium machine. Therefore,
-more sophisticated applications, such as a two dimensional plot
-for CP violation as function of the simulated values of $\stheta$ and
-$\deltacp$ can be obtained in several hours.
-
-\section{Use of package}
-
-The \GLOBES\ package~\cite{Globes} is a tar-ball which consists of the
-source code for the C-library providing the user's interface, the source code
- for a program to test
-\AEDL\ files, example C-files illustrating the use of the library, 
-experiment prototypes in \AEDL\ with supporting flux and cross 
-section files, and the usual 
-supporting files for installation and compilation.
-Currently, \GLOBES\ is only provided for GNU/Linux.
-In addition, an extensive manual covering the user's interface and 
-\AEDL~\cite{Manual} 
-can be obtained from the web-site. \GLOBES\ is free software and as such 
-licensed under the GNU Public License.
-
-The installation of \GLOBES\ is highly automated by the use of the
-Autotools family and follows the usual procedures. By default
-a shared library {\tt libglobes.so} is installed, but 
-a static version is available, too. In addition, an executable named {\tt globes} is installed. It is another possibility to install \GLOBES\ without
-root privilege into a user's home directory. During the automatic
-configuration process also an example Makefile is produced which can
-serve as a template for compiling and linking own applications against
-{\tt libglobes.so}. This Makefile can be used to directly compile and
-link the examples
-from the manual.
-
-\AEDL\ files, such as the experiment prototypes,
-can be edited with any text editor, in order to be loaded by the
-users's interface later. In addition, the {\tt globes} binary allows to
-develop and test experiment descriptions. For example, event rate
-information can be quickly provided to adjust the neutrino flux
-normalization. For further information, we refer
-at this place to the \GLOBES\ manual~\cite{Manual}.
+\bi
+\item 
+ GLoBES on clusters, Condor; parallized code; static links
+\item 
+ 64 bit support
+\item 
+ New minimization algorithm!?
+\ei
 
@@
 
-In summary, the \GLOBES\ software package provides powerful tools
-to do a full three-flavor analysis of future long-baseline and reactor
- experiments
-including systematics, correlations, and degeneracies. In addition,
-the abstract experiment definition language allows to define
-experiments at an abstract level in a highly accurate fashion. 
-Some of the major strengths of \GLOBES\ are the ability to quickly
-define new experiments, the potential to take into account the
-full multi-parameter correlation, the possibilities to include
-external input and matter density uncertainties and 
-the ability to analyze several experiments simultaneously.
+\begin{appendix}
+\section{To do list for new feature release (early spring?); internal}
 
-We conclude that the \GLOBES\ software has two major target groups:
-Experimentalists, who want to quickly evaluate the physics potential
-of their setups, and theorists, who want to test new ideas or
-strategies with pre-defined experiments. Especially, the separation
-between experiment descriptions as simple text files (together
-with supporting files) and the application software should allow an
-efficient and lively interaction between those two target groups.
+\subsection*{Programming [PH+JK?]}
+
+\bi
+\item
+ Implement user-defined priors+ systematics (done?)
+\item
+ Clean up linking scheme: No dynamic linking of modules (physical module
+ support) anymore; support static binaries, support Condor etc.;
+ problem with append mode (files) [PH]
+\item
+ New algoriihm for minimzer [JK]? In this context: Is there any
+ efficient algorithm for more than six parameters?
+\item
+ Can globes-command (!) take more than six parameters if GLB\_OSCP $>$ 6?
+ Adjust, if necessary.
+\item
+ Implement Walter's or any modified bugfix for glb\_min\_sup.c
+ (we had some negative feedback about that, and for more than
+ six parameters ITMAX is far away from being sufficient)
+\ei
+
+\subsection*{Testing [MR+All?]}
+
+\bi
+\item
+ Does release-version support Condor and more than six parameters?
+\item 
+ Do beta-beam fluxes etc. work?
+\item
+ Do examples produce same results as before?
+\item
+ Are AEDL files consistent?
+\ei
+
+\subsection*{Documentation [WW with some help?]}
+
+\bi
+\item
+ Document all new features and make examples; 
+ especially user-defined priors+systematics, Non-standard-Support
+\item
+ New examples files for new examples in manual
+\item
+ Describe what papers have to be cited
+\ei
+
+\subsection*{Paper [WW/ML?]}
+
+\bi
+ \item
+  Write new release paper (this draft)
+\ei
+
+\subsection*{AEDL files [MR?]}
+
+\bi
+\item
+ Provide new AEDL files for some/any finished (=published) projects!?
+ \bi
+  \item
+   $\beta$-Beams
+  \item
+   New NuFact (from scoping study); our new NuFact; silver channels?
+  \item
+   Double Chooz
+  \item
+   NOvA, BNL beam?
+ \ei
+\item
+ Cleaned up old AEDL files (e.g., 2.5\% background error NuFact etc.);
+ document what changes
+\item
+ New AEDL file forum on webpage
+\item
+ Clear description in each AEDL file which papers have to be cited
+\ei
+
+\subsection*{Website [ML/JK?]}
+
+\bi
+ \item
+  New layout with new release
+ \item
+  New FAQ section (e.g., about normalization in AEDL fluxes)
+ \item
+  New AEDL file forum
+ \item
+  Figure forum, at least from current projects (download)
+\ei
+
+\end{appendix}
 
@@
 
-We would like to thank Martin Freund, who wrote the very first
-version of a three-flavor matter profile treatment many years ago.
-PH is especially thankful for the invaluable advice of Thomas Fischbacher on
-many design issues. 
-In addition, we would like to thank
-Mark Rolinec for his help to translate the experiment descriptions
-into \AEDL , and for creating the illustrations in this manuscript.
-Finally, thanks to all the people who have been pushing this project
-for many years, to the ones who have been continuing asking for the 
-publication of the software, and the referees of several of our
-papers for suggestions which lead to improvements in the software.
-This work has been supported by the ``Sonderforschungsbereich
-375 f\"ur Astro-Teilchenphysik der Deutschen Forschungsgemeinschaft''.
+We would like to thank ...
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%             References                         %%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+WW would like to
+acknowledge support from the W.~M.~Keck Foundation and NSF grant PHY-0503584.
 
-\newpage
-\bibliographystyle{./apsrev}
-\bibliography{./references}
+\bibliographystyle{apsrev}
+\bibliography{references}
 

### 2005-11-21 Documentation/part1.tex
*** empty log message ***

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\chapter{Simulating non-standard physics}
+\label{chapt:nsphysics}
+\index{norm}{Non-standard physics}
+
+In this chapter, we discuss how to simulate non-standard physics with \GLOBES , \ie, physics
+beyond three-flavor neutrino oscillations. For the first time, this feature was used in \Ref~\cite{Blennow:2005yk},
+and it has become a standard feature of \GLOBES\ starting from version 3.0 (CHECK). Since the 
+computation of oscilltion probabilities, and therefore the flavor transition probabilities, is the core basic
+element of \GLOBES , the introduction of non-standard physics requires that one produce a new version branch
+of the software, \ie, the software has to be re-compiled after the modification of the probability engine.
+Below, we will demonstrate how and where to do the necessary mofifications, and how to simulate non-standard
+physics in the application software using the ``new'' \GLOBES .
+
+\section{Modification of \GLOBES }
+
+In order to modify \GLOBES , first install a current version of GLOBES (version larger than X.X.X CHECK) as
+a new branch, \ie, in a new directory, and rename the directory \GLOBES\ expanded into. 
+Using multiple versions of \GLOBES , it is, in general, advisable
+to specify the installation directory of all used versions manually (see ``Installation without
+root privilege'' in {\tt INSTALL} file). For instance, if the current ``standard'' version of \GLOBES\ is
+installed in {\tt GLOBES/Versions}, you may use {\tt GLOBES/NSVersion} for the non-standard version
+specified in {\tt configure}. Do not forget to rename the directory \GLOBES\ tar-ball expanded into: For instance,
+change the name from {\tt globes-3.0.0} to {\tt globes-3.0.0nsi}. 
+Then you can modify the source code in the sub-directories of {\tt globes-3.0.0nsi}, such as {\tt globes-3.0.0nsi/source}.
+In particular, we will be interested in modifying {\tt glb\_probability.cc}, which contains the
+Hamiltonian in the routine {\tt ComputeMixingMatrix}.
+%
+In general, the steps to introduce non-standard physics as follows:
+\begin{enumerate}
+\item
+ Change the number of oscillation parameters in {\tt GLB\_OSCP} from six (standard) to the
+ new number $6+N$, where $N$ is the number of non-standard parameters. 
+ This change has to be done in the header file {\tt globes.h} in the sub-directory {\tt globes}
+ where the tar-ball was extracted into,  {\em as well as} in the {\tt include/globes} sub-directory
+ where the non-standard version was installed into (such as {\tt GLOBES/NSVersion/include/globes}).
+\item 
+ Modify {\tt glb\_probability.cc} to include the non-standard effects. The parameters
+ can be accessed by {\tt nsparams[0]}, {\tt nsparams[1]}, ..., {\tt nsparams[N-1]}.
+\item
+ Re-compile the software using \\
+ {\tt make} \\
+ {\tt make install}
+\end{enumerate}
+Now your version of \GLOBES\ uses $6+N$ transition parameters, the six standard oscillation parameters,
+and $N$ additional ones.
+
+\section{Using a non-standard version of \GLOBES }
+
+Using more than six transition parameters in \GLOBES , you will have to maintain the additional parameters. 
+For example, there are no standard routines in \GLOBES\ which define projections including more than six parameters, 
+which means that you should only use {\tt glbChiNP} further on. In addition, you can still use {\tt glbDefineParams}, but this 
+will only access the six standard parameters. You will need to set the additional ones manually using {\tt glbSetOscParams}. 
+Do not forget to maintain your non-standard parameters, since any negligence will be punished by un-predicted behavior!
+
+In order to use the non-standard parameters, \GLOBES\ creates any set of oscillation parameters (type {\tt glb\_params})
+with the additional parameters. To access the non-standard parameters, you can use the following functions as usual:
+\begin{itemize}
+\item
+{\tt glbSetOscParams}
+\item
+{\tt glbGetOscParams}
+\item
+{\tt glbSetProjectionFlag}
+\item
+{\tt glbGetProjectionFlag}
+\end{itemize}
+In all these cases, the last parameter {\tt which} can now run from $0$ to {\tt GLB\_OSCP}$-1$. For instance,
+\begin{quote}
+{\tt
+ glbDefineParams(true\_values,0.55,0.15,0.78,0.0,0.000082,0.0022); \\
+ glbSetOscParams(true\_values,0.0,6);
+}
+\end{quote}
+sets all of the oscillation parameters if {\tt GLB\_OSCP}=7. In addition, the function {\tt glbCopyParams}
+can be used to copy all parameters inluding the non-standard ones. For projections,
+do not use any of the {\tt glbChi...} functions anymore except from {\tt glbChiNP} in order to have a
+predictive behavior of the extra dimenstions. Similar to the oscillation parameters, you can define
+the marginalization over the extra parameter(s) by {\tt glbSetProjectionFlag}. You can find a simple
+example using non-standard physics on page~\pageref{ex:nsphysics}.
+
+\example{Non-standard process on Hamiltonian level}{
+\label{ex:nsphysics}
+\index{norm}{Hamiltonian}
+Let us discuss an example of a very simple Hamiltonian-level non-standard flavor-changing effect which 
+couples $\nu_\mu$ to $\nu_\tau$ in flavor base. This effect can be parameterized
+relative to the standard matter effect acting on $\nu_e$ by an additional
+parameter, which we adopt as our 7th oscillation parameter. In order
+to incorporate this effect, we change a few lines in the function {\tt ComputeMixingMatrix}
+in {\tt glb\_probability.cc}:
+\begin{quote}
+{\tt {\footnotesize
+double msw = 0.1*1.5*anti\_neutrinos*(10E-4)*neutrino\_energy*0.5*matter\_density; \\
+md[0][0] += msw; \\
+md[1][2] += msw*nsparams[0]; md[2][1] += msw*nsparams[0];
+}}
+\end{quote}
+Here we have assume a real contribution which has to be added in a Hermitian way (the Hamiltonian
+has to be Hermitian), where the first line is the standard matter effect acting on $\nu_e$ only 
+($2 \sqrt{2} G_F n_e E$ expressed in our units).
+
+In the application software, we can now use this additional parameter. Let us assume that we
+want to compute the projection onto the $\stheta$-axis when we do not know the size of the  non-standard
+effect, \ie, we keep it completely free but assume it to be zero as null hypothesis:
+\begin{quote}
+{\tt {\footnotesize
+ \#define MY\_FV 6 \\
+ ... \\
+ glbDefineParams(true\_values,theta12,theta13,theta23,deltacp,sdm,ldm); \\
+ glbSetOscParams(true\_values,0.0,MY\_FV); \\
+ glbDefineParams(input\_errors,0.1*theta12,0,0,0,0.1*sdm,0); \\
+ glbSetOscParams(input\_errors,0,MY\_FV);  \\
+ glbSetDensityParams(input\_errors,0.05,GLB\_ALL); \\
+ glbSetOscillationParameters(); glbSetRates(); \\
+ glbSetStartingValues(true\_values);glbSetInputErrors(input\_errors); \\
+ \\
+ glbDefineProjection(my\_proj,GLB\_FREE,GLB\_FIXED,GLB\_FREE,\\
+  \hspace*{2cm} GLB\_FREE,GLB\_FREE,GLB\_FREE); \\
+ glbSetProjectionFlag(my\_proj,GLB\_FREE,MY\_FV); \\
+ glbSetDensityProjectionFlag(my\_proj,GLB\_FREE,GLB\_ALL); \\
+ glbSetProjection(my\_proj);
+ \\
+ glbCopyParams(true\_values,fit\_values); \\
+ glbSetOscParams(fit\_values,0.1,GLB\_THETA\_13); \\
+ res=glbChiNP(fit\_values,minimum,GLB\_ALL); 
+}}
+\end{quote}
+This code fragment illustrates the implementation of a {\tt glbChiTheta} function
+including the non-standard effect, which is being marginalized over.
+Alternatively, one could define a {\tt glbChiTheta} fixing the non-standard
+parameter, or include external constrainys to limit its size. Therefore, 
+there is no ``standard'' treatment of new physics effects incorporated
+in \GLOBES .
+}
+
 %%% Local Variables: 

### 2005-11-21 Documentation/tutorials/norm.tex
PH -- firs version of norm.tex

--- 2004-07-20 Documentation/Paper/paper.tex
@@
 \NeedsTeXFormat{LaTeX2e}
-\documentclass[12pt,a4paper]{article}
+\documentclass[12pt]{article}
 
@@
 
-\begin{titlepage}
+%\begin{titlepage}
 
@@
 \begin{flushright}
-TUM-HEP-5XX/04\\
+\today
+%TUM-HEP-553/04\\
 %hep-ph/
@@
 {\begin{center}
-{\Large\bf Simulation of long-baseline neutrino oscillation
-experiments with GLoBES}
-\end{center}}
-{\begin{center}
-{\large\bf (General Long Baseline Experiment Simulator)}
+{\Large\bf Units for flux files in GLoBES}
 \end{center}}
@@
                 P.~Huber\footnote[1]{\makebox[1.cm]{Email:}
-                phuber@ph.tum.de},~
-                M.~Lindner\footnote[2]{\makebox[1.cm]{Email:}
-                lindner@ph.tum.de},~and~
-                W.~Winter\footnote[3]{\makebox[1.cm]{Email:}
-                wwinter@ph.tum.de}
+                phuber@physics.wisc.edu}
+             %   M.~Lindner\footnote[2]{\makebox[1.cm]{Email:}
+             %   lindner@ph.tum.de},~and~
+             %   W.~Winter\footnote[3]{\makebox[1.cm]{Email:}
+             %   wwinter@ph.tum.de}
                 }}
@@
 
-\footnotemark[1]${}^,$\footnotemark[2]${}^,$\footnotemark[3]%
-       Institut f\"ur Theoretische Physik, Physik--Department,\\
-       Technische Universit\"at M\"unchen,
-       James--Franck--Strasse, D--85748 Garching, Germany
+\footnotemark[1]${}^,$% 
+      Department of Physics, University of Wisconsin,\\ 
+ 1150 University
+      Avenue, Madison ,WI 53706, USA
 
@@
 \begin{abstract}
-We present the \GLOBES\ (``General Long Baseline Experiment Simulator'')
-software package, which allows the simulation of long-baseline and reactor
-neutrino oscillation experiments. One part of the software is the
-abstract experiment definition language to define experiments with
-beam and full detector descriptions as accurate as possible. Many systematics options are provided, such as normalization and energy
-calibration errors, or the choice between spectral or total rate 
-information. For the 
-definition of experiments, a new transparent building block concept is 
-introduced. In addition, an additional program provides the possibility 
-to develop and test new experiment definitions quickly.
-%
-Another
-part of \GLOBES\ is the user's interface, which provides probability, rate, 
-and $\Delta \chi^2$ information for a given experiment or any
-combination of up to 32 experiments in C. Especially, the
-$\Delta \chi^2$ functions allow a simulation with statistics only,
-systematics, correlations, and degeneracies. In particular, \GLOBES\
-can handle the full multi-parameter correlation among the oscillation
-parameters, external input, and matter density uncertainties. 
+A common issue with \GLOBES\ is confusion about the proper units for
+the input flux files for use in \AEDL\ experiment descriptions. 
 \end{abstract}
@@
 
-\end{titlepage}
-
-\newpage
+%\end{titlepage}
 
@@
 \setcounter{footnote}{0}
+
+\section{Problem}
+
+One problem for the design of \AEDL\ was initially that meaningful
+units for flux data strongly depend on the given type of experiment,
+but also on relatively arbitrary decisions. For accelerator beams
+based on $\pi$-decay one frequently defines the beam luminosity in
+{\it protons on target (pot)} since this number has a one-to-one
+correspondence with the number of neutrinos produced. Another sensible
+unit could be {\it megawatt on target (MW)}, again this number is directly
+correlated with the number of neutrinos and moreover there is a unique
+relation to {\it pot} for a given accelerator. Of course,
+what matters is the integrated luminosity. Thus in some  cases the
+neutrino flux is given per $10^7\,\mathrm{s}$. However,
+most experiments will run for several years, hence also this number
+has to enter somewhere. For neutrino factories the proper number is
+useful muon decays per unit time and for reactor experiments it is
+the thermal power of the reactor {\it asf}. This shows that there is
+not much of a  point to declare one unit for the input fluxes.
+
+\section{Solution}
+
+In understanding how one still can figure out what the correct units
+are for each case, it is a good starting point to look at what
+\GLOBES\ does with the input files. The cross section in the file is given as
+differential cross section divided by energy $x=\sigma/E$, and the flux file
+gives $f$. The differential number of events per GeV $n$ as computed in
+\GLOBES\ without oscillation and efficiencies is given by
+\begin{eqnarray}
+n &=& 5.20034\times x\times E\times f\times\\
+&&\mathtt{@norm}\times\mathtt{@power}\times\mathtt{@stored\_muons}\times\mathtt{@time}\times\mathtt{\$target\_mass}\times(\mathtt{\$baseline})^{-2}\nonumber
+\end{eqnarray}
+ Beware: $5.2\ldots$ is a undocumented fudge factor!
+
+It is the sole responsibility of the author of the \AEDL\ file and its
+supporting files, to ensure that the result makes actually sense. In
+principle it is possible to divide {\tt @time}  by
+{\it eg} $\pi$ and fix that by redefining the flux file by multiplying
+it with $\pi$. Things like that happen and have happened, and worse,
+many of them are not properly commented.
+
+\section{Writing \AEDL\ files}
+
+The task is to choose the value of  {\tt @norm} such that all the
+variables in the \AEDL\ file have the proper units {\it eg} {\tt
+  @time} has proper unit years.
+
+\GLOBES\ assume that the cross
+section $x$ is given in $10^{-38}\,\mathrm{cm}^2$, that all fluxes are
+given at a distance of $1\,\mathrm{km}$ and the proper accounting for
+the number of target nuclei $\tau$ (or protons or whatever applies to the
+given cross section) per unit target mass $m_u$ (which usually is {\it
+  kt}), is dealt with in the
+definition of the flux inside \AEDL.  Assuming that in the flux file
+the data is given as number of neutrinos per unit area $A$ and
+energy bin of width $\Delta E$ at a distance $L$ from the source, one
+obtains
+
+\begin{eqnarray}
+(\mathtt{@norm})^{-1}=5.20034\left(\frac{\Delta E}{\mathrm{GeV}}\right)\left(\frac{A}{\mathrm{cm}^2}\right)\left(\frac{\mathrm{km}}{L}\right)^2\left(\frac{\tau}{m_u}\right)\times10^{-38}\times\left(\frac{\mathcal{L}}{\mathcal{L}_u}\right)
+\end{eqnarray}
+where the $\mathcal{L}$ absorbs all factors in the flux file related to
+the integrated luminosity and $\mathcal{L_u}$ is the unit chosen for
+it.  The concept of integrated luminosity is
+nicely described in the \GLOBES\ manual:
+
+\begin{quote}
+ In \GLOBES , the total number of events is in general proportional to the product of
+\begin{equation}
+\mathrm{Fid.~detector~mass}\,\left[\mathrm{kt/t}\right]\times 
+\mathrm{Running~time} \,\left[\mathrm{yr}\right]\times\left\{ \begin{array}{c}
+\mathrm{Source~power}\,\left[\mathrm{MW/GW}\right]\\
+\mathrm{Useful~muon~decays}\,\left[\mathrm{yr}^{-1}\right]
+\end{array}\right.\,.
+\end{equation}
+Thus, the source power corresponds to either the amount of energy produced per time frame in the target (such as for nuclear reactors or sources based on pion decay), or the useful muon decays per time frame (neutrino factories).
+In addition, the definition of the source power makes only sense together with the flux normalization, the running time, and fiducial detector mass in order to define the total integrated luminosity.
+Therefore, one can, in principle, use arbitrary units for these components as long as their product gives the wanted neutrino flux.
+However, it is
+recommended to use normalizations such that the source power units are $\mathrm{MW}$ for a proton-based beam, and $\mathrm{GW}_\mathrm{thermal}$ for a reactor experiment.
+Correspondingly, the detector mass units should be kilotons for a proton-based beam, and tons for a reactor experiment.
+In any case it is a good
+idea to document the choices made by the user by corresponding comments
+in \AEDL.
+\end{quote}
+
+A little example may illustrate this:
+The flux is given for $10^{21}\, \mathrm{pot}\,\mathrm{y}^{-1}$ of
+$10\,\mathrm{GeV}$ protons, thus a good choice for the units $\mathcal{L}_u$
+is $\mathrm{MW}\,\mathrm{y}^{-1}$, which means that
+$\mathcal{L}/\mathcal{L}_u$ is given by (assuming a $10^7\,\mathrm{s}$
+year)
+\begin{equation}
+\frac{\mathcal{L}}{\mathcal{L}_u}=\frac{10\,\mathrm{GeV}\,10^{21}\,\mathrm{pot}\,\mathrm{y}}{10^7\,\mathrm{s}}\times(\mathrm{MW}\,\mathrm{y}^{-1})^{-1}=0.160218\ldots
+\end{equation}
 
@@
 
-\section{Introduction}
-
-Neutrino oscillations are now established as the leading flavor
-transition mechanism for neutrinos. In a long history of many experiments, see
-\eg~\cite{Barger:2003qi}, 
-two oscillation
-frequencies have been identified: The fast ``atmospheric'' and the
-slow ``solar'' oscillations, which are driven by the respective
-mass squared differences.  In addition, there could be interference
-effects between these two oscillations, provided that the coupling given by 
-the small mixing angle $\theta_{13}$ is large enough. Such interference
-effects include, for example, leptonic CP violation.  In order to
-test the unknown oscillation parameters, \ie, the mixing angle $\theta_{13}$,
- the leptonic CP phase, and the
-neutrino mass hierarchy, new long-baseline and reactor experiments 
-are proposed.
-These experiments send an artificial neutrino beam to a detector, or
-detect the neutrinos produced by a nuclear fission reactor. However,
-the presence of multiple solutions which are intrinsic to
- neutrino oscillation probabilities~\cite{Fogli:1996pv,
-Burguet-Castell:2001ez,Minakata:2001qm,Barger:2001yr} 
-affect these measurements.
-Thus optimization strategies are required which maximally exploit 
-complementarity between experiments. Therefore, a modern, complete experiment 
-simulation  and analysis tool does not only need to have 
-a highly accurate beam and detector simulation, but also powerful 
-means to analyze correlations and degeneracies, especially for the combination
-of several experiments.
-In this paper, we present the \GLOBES\ software package
-as such a tool.
-
-\GLOBES\ (``General Long Baseline Experiment Simulator'') is a flexible
-software tool to simulate and analyze neutrino oscillation 
-long-baseline and reactor experiments using a 
-complete three-flavor description. On the
-one hand, it contains a comprehensive abstract experiment definition
-language (AEDL), which allows to describe 
-most classes of long baseline and reactor experiments
-at an abstract level. On the other hand, it provides a C-library to 
-process the experiment information in order to obtain oscillation
-probabilities, rate vectors, and $\Delta \chi^2$-values (\cf, \figu{GLOBES}). 
-In addition, it provides a binary program to test experiment
-definitions very quickly, before they are used by the application software.
-Currently, \GLOBES\ is available for GNU/Linux. Since the source code is 
-included, the modifications to other operating systems should be doable.
-
-\GLOBES\ allows to simulate experiments with stationary neutrino point 
-sources, where each experiment is assumed to have only one neutrino source.
-Such experiments are neutrino beam experiments and reactor experiments. 
-Geometrical effects of a source distribution, such as in the sun or the 
-atmosphere, can not be described. In addition, sources with a physically 
-significant time dependencies  can not be studied, such as  supernov\ae.
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=14cm]{GLOBES}
-\end{center}
-\caption{\label{fig:GLOBES} General concept of the \GLOBES\ package.}
-\end{figure}
-
-On the experiment definition side, either built-in neutrino fluxes
-(\eg, neutrino factory) or arbitrary, user defined fluxes can be used. 
-Similarly,
-arbitrary cross sections, energy dependent efficiencies,
-energy resolution functions as well as the considered oscillation channels, 
-backgrounds, and many other properties can be specified. 
-For the systematics, energy
-normalization and calibration errors can be simulated. Note that
-energy ranges and windows and bin widths can be
-(almost) arbitrarily chosen, including bins of different
-widths. Together with \GLOBES\ comes a number of
-pre-defined experiments in order to demonstrate the capabilities
-of \GLOBES\ and to provide prototypes for new experiments.
-In addition, they can be used to test new physics ideas with
-complete experiment simulations.
-Examples for these prototypes are the \MINOS , \ICARUS , and
-\OPERA\ simulations from \Ref~\cite{Huber:2004ug}, the
-\JHFSK\ and \NUMI\ superbeam simulations from \Refs~\cite{Huber:2002mx,
-Huber:2002rs}, the \JHFHK\ superbeam upgrade simulation from
-\Ref~\cite{Huber:2002mx}, the neutrino factory simulations from
-\Ref~\cite{Huber:2002mx}, and the reactor experiment simulations from
-\Ref~\cite{Huber:2003pm}. Other projects using earlier versions
-of the \GLOBES\ software include \Refs~\cite{Huber:2003ak,Ohlsson:2003ip,Winter:2003ye,Antusch:2004yx}.
-
-With the C-library, one can extract the $\Delta \chi^2$ for all defined 
-oscillation channels for an experiment or any combination of experiments.
-Of course, also low-level information, such as oscillation
-probabilities or event rates, can be obtained. \GLOBES\ includes the
-simulation of neutrino oscillations in matter with arbitrary matter 
-density profiles, as well as it allows to simulate the matter density
-uncertainty. As one of the most
-advanced features of \GLOBES , it provides the technology to 
-project the $\Delta \chi^2$, which is a function of all oscillation
-parameters, onto any subspace of parameters by local minimization. 
-This approach allows the inclusion of multi-parameter-correlations,
-where external constraints (\eg, on the solar parameters) can be imposed, too.
-Applications of the projection mechanism include the projections onto 
-the $\stheta$-axis and the $\stheta$-$\deltacp$-plane. In addition, 
-all oscillation parameters can be kept free to precisely localize 
-degenerate solutions.
-
-\section{The computation of raw event rates}
-
-In this section, we sketch the computation of 
-the event rates, which is one of the core parts of the \GLOBES\
-software. However, because of the complexity of this issue,
-we refer to the \GLOBES\ manual~\cite{Manual} for
-details.
-
-The differential event rate for each channel is given by
-\begin{eqnarray}
-\label{equ:master_event}
-\frac{dn_{\beta}^{\text{IT}}}{dE'}=&&N\,\int\limits_0^\infty \int\limits_0^\infty dE\,d\hat{E}\quad
-\underbrace{\Phi_{\alpha} (E)}_{\mathrm{Production}} \times \nonumber\\
-&&\underbrace{\frac{1}{L^2} P_{(\alpha\rightarrow\beta)}(E,L,\rho;\theta_{23},
-\theta_{12},\theta_{13},
-\Delta m^2_{31},\Delta m^2_{21},\deltacp)}_{\mathrm{Propagation}}
-\times \nonumber \\ &&\underbrace{\sigma^{\text{IT}}_f(E)
-k_f^{\text{IT}}(E-\hat{E})}_{\mathrm{Interaction}} \times \nonumber \\
-&&\underbrace{ T_f(\hat{E}) V_f(\hat{E}-E')}_{\mathrm{Detection}}\,,
-\end{eqnarray}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-where $\alpha$ is the initial flavor of the neutrino, 
-$\beta$ is the final flavor, $\Phi_{\alpha} (E)$ is the flux of the 
-initial flavor at the
-source, $L$ is the baseline length, $N$ is a normalization factor, and 
-$\rho$ is the matter density. The energies in this formula are given as follows:
-\begin{itemize}
-\item
- $E$ is the incident neutrino energy, \ie, the actual energy of the 
-incoming neutrino (which is not directly accessible to the experiment)
-\item
- $\hat{E}$ is the energy of the secondary particle
-\item
- $E'$ is the reconstructed neutrino energy, \ie, the measured
-neutrino energy as obtained from the experiment
-\end{itemize}
-The interaction term is composed of 
-two factors, which are the total cross section 
-$\sigma^{\text{IT}}_\beta(E)$ for the flavor $f$ and
-the interaction type IT, and the energy distribution of the 
-secondary particle $k_\beta^{\text{IT}}(E-\hat{E})$.
-The detector properties are 
-modeled by the threshold function $T_\beta(\hat{E})$, coming from the the 
-limited resolution or the cuts in the analysis, and the energy resolution 
-function $V_\beta(\hat{E}-E')$ of the secondary particle.
-
-Since it is a lot of effort to solve this double integral numerically,
-we split up the two integrations. First, we evaluate the integral over
-$\hat{E}$, where the only terms containing $\hat{E}$ are
-$k_\beta^{\text{IT}}(E-\hat{E})$,  $ T_\beta(\hat{E})$, and 
-$ V_\beta(\hat{E}-E')$. We define:
-\begin{eqnarray}
-\label{equ:e_res} 
-R_\beta^{\text{IT}}(E,E')\,\epsilon_\beta^{\text{IT}}(E')
- \equiv
-\int\limits_0^\infty d\hat{E} \quad T_\beta(\hat{E})\,k_\beta^{\text{IT}}(E-\hat{E})
-\,V_\beta(\hat{E}-E')\,. 
-\end{eqnarray}
-Thus, $R_\beta^{\text{IT}}(E,E')$ describes the energy response of 
-the detector, \ie , a neutrino with a (true) energy $E$ is reconstructed
-with an energy between $E'$ and $E'+dE'$ with a probability
-$R_\beta^{\text{IT}}(E,E') dE'$. The function $R(E,E')$ is also often called ``energy resolution function''. Actually, its internal representation
-in the software is a smearing matrix.
-The function $\epsilon_\beta^{\text{IT}}(E')$ will be refered to as ``post-smearing efficiencies'', since it will allow us to define cuts and threshold functions {\em after} the smearing is performed, \ie, as function of $E'$.
-In addition, \GLOBES\ uses ``pre-smearing efficiencies'', which are 
-evaluated {\em before} the smearing is performed, \ie, as function of $E$.
-Similarly, constant\footnote{With respect to the oscillation parameters, 
-not the energy.} background rates can be added to the event rates before or 
-after the energy smearing, which are refered to as 
-``pre-smearing backgrounds'' and ``post-smearing backgrounds''. These types of 
-efficiencies and backgrounds allow a very accurate modeling of
-many factors, such as atmospheric or geoneutrino backgrounds, energy cuts, or
-energy threshold functions.
-
-Eventually, we can write down the number of events per bin $i$ \index{Bin} and channel $c$ as
-\begin{equation}
-\label{equ:channel}
-n_i^c=\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' \quad
-\frac{dn_{\beta}^{\text{IT}}}{dE'} (E') \,
-\end{equation}
-where $\Delta E_i$ is the bin size of the $i$th energy bin.
-This means that one has to solve the integral
-\begin{eqnarray}
-\label{equ:events_bin}
-n_i^c=N/L^2\,\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
-\quad \int\limits_0^\infty dE \,\, \Phi^c(E)\,
-P^c(E)\,
-\sigma^c(E)\,
-R^c(E,E')\,
-\epsilon^c(E')\,,
-\end{eqnarray} 
-which gives the raw event rates of the channel $c$ in the $i$th bin.
-Note that the events are binned according to their \emph{reconstructed} energy.
-
-Core part of the event rate computation is the energy smearing algorithm
-to evaluate \equ{events_bin}, where either a particular energy resolution
- function can be chosen for automated energy smearing, or the discretized
-smearing matrix $R^c(E,E') = R^c_{ij} $ can be given manually. In addition, it is possible
-to define a low-pass filter to avoid aliasing effects for neutrino
-oscillations faster than the sampling width.
-
-\section{Definition of experiments with \AEDL\\ (Abstract Experiment 
-Definition Language)}
-
-\begin{figure}[t]
-\begin{center}
-\begin{tabular}{p{6.3cm}p{4.3cm}p{4.3cm}}
-\includegraphics[width=6.5cm]{AEDL1} &
-\includegraphics[width=4.5cm]{SignalBackground} &
-\includegraphics[width=4.5cm]{Rules} \\
-\end{tabular}
-\end{center}
-\caption{\label{fig:aedl} The most important components of \AEDL :
-Channels, rules, and experiments.}
-\end{figure}
-
-In order to define experiments, \GLOBES\ uses \AEDL\ (``Abstract
-Experiment Definition Language''). An experiment normally corresponds
-to an \AEDL\ file, which is a human readable text file written 
-in \AEDL\ syntax.
-
-The key concept of \AEDL\ is to regard a detector as an abstract system
-which maps the true properties of a neutrino onto the reconstructed properties
-of the neutrino. The latter are subsequently used in the fit of the oscillation
-parameters. Within \GLOBES , only energy and flavor are  
-observables, which is
-sufficient for long-baseline and reactor experiments. In  other
-experiments, such as atmospheric neutrino experiments, more observables 
-(\eg , direction) may have to be considered.
-
-The main components of any \AEDL\ experiment definition are ``channels'',
-``rules'', and ``experiments'' (\cf, \figu{aedl}).
-A channel corresponds to a neutrino oscillation channel including flux, cross section (for one specific interaction type), energy resolution function, initial and final neutrino flavors, their polarity (neutrinos or antineutrinos), and efficiencies.
-The channel raw event rates are computed according to \equ{events_bin}.
-Each channel leads to the raw event rates for a
-specific interaction type. In \AEDL , many of the channel components
-have to be defined or loaded from files before, such as fluxes, 
-cross sections, or the energy resolution function. For the
-different options, see the \GLOBES\ manual~\cite{Manual}.
-
-For a ``rule'', the raw event rates of one or more signal channels and
-one or more background channels are added. The splitting in signal and
-background is arbitrary, but all of the signal and background components
-are defined to have a common systematics. The event rates of all 
-signal and background components add up to the total event rate of the
-rule, which leads to a $(\Delta \chi^2)_r$. The signal or background
-within each rule allows the specification of signal and background 
-normalization errors and energy tilt or calibration errors. 
-These systematical errors are evaluated with the ``pull method'' such as 
-in \Refs~\cite{Huber:2002mx,Fogli:2002pt}.  
-In addition to these
-systematics errors, an overall evaluation strategy is assigned 
-to each rule, which specifies the type of systematics (tilt or calibration
-error), and the use of spectral information or total event rates.
-
-Finally, one ore more rules add up to an experiment, where the
-total $\Delta \chi^2$ is obtained as the sum of the $(\Delta \chi^2)_r$
-of all rules. This approach allows the definition of 
-appearance and disappearance channels, neutrino and antineutrino running, 
-or interaction types with different systematics (spectral information
-versus counting rate) within one experiment. The \GLOBES\ user interface
-allows the simulation of one or more experiment simultaneously, which
-means that one could also use different experiments for different
-oscillation channels. However, there is still one component missing
-in the experiment definition, which is the matter density profile.
-For an experiment, an arbitrary matter density profile can be specified, 
-which is evaluated with the evolution operator method~\cite{Ohlsson:1999um}.
-In addition, the matter density profile is multiplied by a scaling factor, 
-which is treated as an independent parameter
-with a relative (matter density) uncertainty. With this
-approach, one can take into account that the matter density profile along a 
-specific baseline is only known to about $3\%-5\%$. Thus, for one
-particular baseline, all rules should be defined within one experiment.
-
-\section{Analysis of experiments: The C user's interface}
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=16cm]{projex}
-\end{center}
-\mycaption{\label{fig:projex} Left plot: Example for a  correlation between $\stheta$ and $\deltacp$ (for 1 d.o.f. only, un-shown oscillation
-parameters fixed).
-Right plot: The $\chi^2$-value of the projection onto the $\stheta$-axis as function of $\stheta$.
-The projection onto the  $\stheta$-axis is obtained by finding the minimum $\chi^2$-value for each fixed value of $\stheta$ in the left-hand plot, \ie, along the gray vertical lines.
-The thick gray curve marks the position of these minima in the left-hand plot.
-The arrows mark the obtained fit ranges for $\stheta$ at the $3 \sigma$ confidence level (1 d.o.f.), \ie , the precision of $\stheta$.}
-\end{figure}
-
-In order to use \GLOBES , a C-library is linked with the application
-software. This library provides the user's interface functions.
-It allows to load \AEDL\ files, initialize the experiments, and
-have access to various $\Delta \chi^2$-functions including any combination
-of systematics and correlations. In addition, it provides low-level
-access to oscillation probabilities and event rates, and allows
-the readout and modification of many experiment parameters at run time.
-
-The most sophisticated feature of the user's interface is the
-possibility to include the full multi-parameter correlation among
-all oscillation parameters. Together with the matter densities
-of the $N_{\mathrm{exp}}$ loaded experiments ($N_{\mathrm{exp}} \ge 1$), the neutrino oscillation parameter space in \GLOBES\ has $6+N_{\mathrm{exp}}$ dimensions. In the $\chi^2$-approach,
-the allowed region of $k$ ($1\le k<6+N_{\mathrm{exp}}$) fit parameters (at the chosen confidence level) is obtained as the projection of the $6+N_{\mathrm{exp}}$-dimensional fit manifold onto the $k$-dimensional hyperplane.
-For example,
-for the precision of the parameter $\stheta$ only, the fit manifold
-needs to be projected onto the $\stheta$-axis. One can easily imagine
-that the computation of the full multi-parameter correlation is
-very expensive with a grid-based method. However, the topology
-of the neutrino oscillation parameter space is well investigated and
-rather smooth. Therefore, it is feasible to use local minimization for each
-solution in the parameter space.  For this process, 
-a $6+N_{\mathrm{exp}}$-dimensional local minimizer
-is started at the ``guessed'' position for the solution
-(such as from analytical knowledge) and then runs into a
-local minimum. This exercise has to be repeated till all local 
-minima, \ie\ degenerate solutions, are found. In \figu{projex},
-this minimization is illustrated at the simple example of a 
-two-parameter correlation at a neutrino factory.
-
-As we have indicated above, the matter density of
-each experiment is treated as an additional parameter, which is known
-from external measurement with a certain precision. In principle, 
-the \GLOBES\ user's interface allows such externally imposed
-precisions for all oscillation parameters. This feature turns out
-to be especially useful for input from the solar neutrino experiments,
-since they provide better precisions for the solar neutrino
-oscillation parameters $\sdm$ and $\theta_{12}$ than long-baseline
- experiments themselves.
-
-\section{Complexity and computational issues}
-
-The \GLOBES\ software only uses polynomial algorithms, which are,
-however, suffering from the large number of dimensions.
-The $\Delta \chi^2$ computation basically consists of two steps:
-The systematics $(\Delta \chi^2)_{\mathrm{sys}}$ computation with {\tt glbChiSys},
-and the projection algorithm which uses {\tt glbChiSys} many times. 
-The overall computation time is obtained as the product of these two steps.
-
-The run time of the systematics $(\Delta \chi^2)_{\mathrm{sys}}$ 
-is 
-\begin{equation}
-\mathrm{t_{sys}} = \mathcal{O} \left( N_{\mathrm{Sampling}} \times N_{\mathrm{Bins}}  \, + \, N_{\mathrm{Sampling}} \times N_{\mathrm{Layers}} \right) ,
-\end{equation}
-since the oscillation probabilities for all channels and energies
-are only computed once and stored in a list.
-Here $N_{\mathrm{Sampling}}$ is the number of sampling points for the evaluation of the right integral in \equ{events_bin}, $N_{\mathrm{Bins}}$ is the number of energy bins,
-and $N_{\mathrm{Layers}}$ is the number of matter density layers. Thus, whenever the
-number of matter density steps is small, the product of sampling points
-and bins dominates. In practice, the $N_{\mathrm{Sampling}} \times N_{\mathrm{Bins}}$ energy smearing matrix is already computed at 
-the experiment initialization in order to save run time. In addition,
-since it contains a lot of zeros, \GLOBES\ uses a special format to
-avoid looping over the zeros.
-
-The run time for the projected $\Delta \chi^2$ would, for a grid based method, be
-\begin{equation}
- \mathrm{t_{proj}^{Grid}} = \mathcal{O}\left( n^{6+N_{\mathrm{exp}}}\right) \, \times \, \mathrm{t_{sys}},
-\end{equation}
-where $n$ is the number of grid points to be evaluated for each oscillation
-parameter, and $N_{\mathrm{exp}}$ is the number of experiments
-(since the matter densities of all experiments appear as parameters). 
-Though it is, in principle, possible to use such a method with \GLOBES ,
-it is impracticable in most cases, since it would involve at least 
-billions of
-steps for the complete multi-parameter correlation. The local $6+N_{\mathrm{exp}}$-dimensional minimization reduces this effort
-to 
-\begin{equation}
-\mathrm{t_{proj}^{Minimizer}} = \mathcal{O}(1000) \, \times \,
-N_{\mathrm{Deg}} \,  \times \, \mathrm{t_{sys}}
-\end{equation}
-steps, where $N_{\mathrm{Deg}}$ is the number of (disconnected) 
-local minima, \ie, the number of
-degeneracies (typically $\mathcal{O}(10)$).
-
-Eventually, \GLOBES\ needs about $10$ to $15$ seconds to compute
-a projected $\Delta \chi^2$ on a modern Pentium machine. Therefore,
-more sophisticated applications, such as a two dimensional plot
-for CP violation as function of the simulated values of $\stheta$ and
-$\deltacp$ can be obtained in several hours.
-
-\section{Use of package}
-
-The \GLOBES\ package~\cite{Globes} is a tar-ball which consists of the
-source code for the C-library providing the user's interface, the source code
- for a program to test
-\AEDL\ files, example C-files illustrating the use of the library, 
-experiment prototypes in \AEDL\ with supporting flux and cross 
-section files, and the usual 
-supporting files for installation and compilation.
-Currently, \GLOBES\ is only provided for GNU/Linux.
-In addition, an extensive manual covering the user's interface and 
-\AEDL~\cite{Manual} 
-can be obtained from the web-site. \GLOBES\ is free software and as such 
-licensed under the GNU Public License.
-
-The installation of \GLOBES\ is highly automated by the use of the
-Autotools family and follows the usual procedures. By default
-a shared library {\tt libglobes.so} is installed, but 
-a static version is available, too. In addition, an executable named {\tt globes} is installed. It is another possibility to install \GLOBES\ without
-root privilege into a user's home directory. During the automatic
-configuration process also an example Makefile is produced which can
-serve as a template for compiling and linking own applications against
-{\tt libglobes.so}. This Makefile can be used to directly compile and
-link the examples
-from the manual.
-
-\AEDL\ files, such as the experiment prototypes,
-can be edited with any text editor, in order to be loaded by the
-users's interface later. In addition, the {\tt globes} binary allows to
-develop and test experiment descriptions. For example, event rate
-information can be quickly provided to adjust the neutrino flux
-normalization. For further information, we refer
-at this place to the \GLOBES\ manual~\cite{Manual}.
-
-\section{Summary and conclusions}
-
-In summary, the \GLOBES\ software package provides powerful tools
-to do a full three-flavor analysis of future long-baseline and reactor
- experiments
-including systematics, correlations, and degeneracies. In addition,
-the abstract experiment definition language allows to define
-experiments at an abstract level in a highly accurate fashion. 
-Some of the major strengths of \GLOBES\ are the ability to quickly
-define new experiments, the potential to take into account the
-full multi-parameter correlation, the possibilities to include
-external input and matter density uncertainties and 
-the ability to analyze several experiments simultaneously.
-
-We conclude that the \GLOBES\ software has two major target groups:
-Experimentalists, who want to quickly evaluate the physics potential
-of their setups, and theorists, who want to test new ideas or
-strategies with pre-defined experiments. Especially, the separation
-between experiment descriptions as simple text files (together
-with supporting files) and the application software should allow an
-efficient and lively interaction between those two target groups.
-
-\subsection*{Acknowledgments}
-
-We would like to thank Martin Freund, who wrote the very first
-version of a three-flavor matter profile treatment many years ago.
-PH is especially thankful for the invaluable advice of Thomas Fischbacher on
-many design issues. 
-In addition, we would like to thank
-Mark Rolinec for his help to translate the experiment descriptions
-into \AEDL , and for creating the illustrations in this manuscript.
-Finally, thanks to all the people who have been pushing this project
-for many years, to the ones who have been continuing asking for the 
-publication of the software, and the referees of several of our
-papers for suggestions which lead to improvements in the software.
-This work has been supported by the ``Sonderforschungsbereich
-375 f\"ur Astro-Teilchenphysik der Deutschen Forschungsgemeinschaft''.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%             References                         %%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newpage
-\bibliographystyle{./apsrev}
-\bibliography{./references}
-
 \end{document}
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2005-11-22 Documentation/tutorials/norm.tex
PH -- changed the author

@@
 \begin{flushright}
-\today
+\today\\
+GLoBES-FAQ-1/05
 %TUM-HEP-553/04\\
@@
 {\begin{center} {\large{\sc
-                P.~Huber\footnote[1]{\makebox[1.cm]{Email:}
-                phuber@physics.wisc.edu}
+                The \GLOBES\ Team\footnote[1]{\makebox[1.cm]{Email:}
+                globes@ph.tum.de}
              %   M.~Lindner\footnote[2]{\makebox[1.cm]{Email:}
@@
 \begin{center}
-
-\footnotemark[1]${}^,$% 
-      Department of Physics, University of Wisconsin,\\ 
- 1150 University
-      Avenue, Madison ,WI 53706, USA
 
@@
 
-\GLOBES\ assume that the cross
+\GLOBES\ assumes that the cross
 section $x$ is given in $10^{-38}\,\mathrm{cm}^2$, that all fluxes are
@@
 where the $\mathcal{L}$ absorbs all factors in the flux file related to
-the integrated luminosity and $\mathcal{L_u}$ is the unit chosen for
+the integrated luminosity and $\mathcal{L}_u$ is the unit chosen for
 it.  The concept of integrated luminosity is

### 2006-03-24 Documentation/part1.tex
Minor changes to user-defined systematics part

@@
 double centers[],double errors[],char *info)} tells \GLOBES\ to use {\tt chi\_func}
-to calculate $\chi^2$ for experiment {\tt exp} and rule {\tt rule} (these parameters
-can take the value \GLBC{GLB\_ALL} to specify all experiments or all rules). {\tt dim} is
+instead of the builtin functions to calculate $\chi^2$ for experiment {\tt exp} and
+rule {\tt rule}. These parameters can take the value \GLBC{GLB\_ALL} to specify that
+{\tt chi\_func} should be called once for each experiment or each rule. {\tt dim} is
 the number of variables parametrizing the systematical errors, while {\tt centers[]}
@@
 the values of the systematics parameters for which $\chi^2$ should be calculated.
-Deviating from the standard conventions, the entries of {\tt x[]} will start at
-index 1.
+Deviating from the standard conventions, the entries of {\tt x[]} will start at index 1
+(The indices for {\tt params} are {\tt errors} in the call to {\tt glbSetUserChi} start
+at 0 as usual).
 \end{function}
@@
 \hspace*{0.5 cm}      double chirate, measured\_rate;                           \\
+\hspace*{0.5 cm}                                                                \\
+\hspace*{0.5 cm}      /* Assign complete chi\textasciicircum2 to first experiment */ \\
+\hspace*{0.5 cm}      if (glbGetCurrentExp() != 0)  return 0;                   \\
 \hspace*{0.5 cm}                                                                \\
@@
 \hspace*{0.5 cm}      \}                                                        \\
-\hspace*{0.5 cm}                                                                \\
 \hspace*{0.5 cm}      for (int j=0; j < 6; j++)  /* Penalties */                \\
 \hspace*{1.0 cm}        chi2 += square((x[j+1] - sys\_centers[j]) / sys\_errors[j]);\\
-\hspace*{0.5 cm}                                                                \\
-\hspace*{0.5 cm}      /* The function is called for each exp ->
-                                       Assign chi\textasciicircum2/2 to each */ \\
-\hspace*{0.5 cm}      return chi2 / 2;                                          \\
+\hspace*{0.5 cm}      return chi2;                                              \\
   \}                                                                            

### 2007-01-03 Documentation/Manual.tex
PH channged credits
Part of manual updated

@@
 \newcommand{\glbxxx}{hep-ph/0407333}
+\newcommand{\glbxxxtwo}{hep-ph/XXYYZZZ}
 \newcommand{\capdef}{}
@@
 \begin{center}
-{\large Patrick Huber$^a$, Manfred Lindner$^b$, Walter Winter$^c$}
+{\large Patrick Huber$^a$, Joachim Kopp$^b$, Manfred Lindner$^b$, \\ Mark Rolinec$^c$, Walter Winter$^d$}
 \end{center}
@@
 
-\vspace{1cm}
+\vspace{0.5cm}
 
@@
 \begin{center}
-Version from \today\ for \GLOBES\ 2.0.11
+Version from \today\ for \GLOBES\ 3.0
 \end{center}
@@
 
+\vspace*{0.1cm}
+
 {\em\small $^{\mathrm{a}}$%
-       Physics Department,
-       University of Wisconsin,
+       University of Wisconsin,  Physics Department,
        1150 University Av., Madison, WI 53706, USA}
@@
 {\em\small $^{\mathrm{b}}$%
+       Max--Planck--Institut f\"ur Kernphysik, 
+       Postfach 10~39~80, D--69029 Heidelberg, Germany}
+
+{\em\small $^{\mathrm{c}}$%
+       Technische Universit\"at M\"unchen,
        Institut f\"ur Theoretische Physik, Physik--Department,\\
-       Technische Universit\"at M\"unchen,
        James--Franck--Strasse, D--85748 Garching, Germany}
 
-{\em\small $^{\mathrm{c}}$%
-       Institute for Advanced Study, 
-       Einstein Drive, Princeton, NJ 08540, USA}
+{\em\small $^{\mathrm{d}}$%
+       Universit\"at W\"urzburg, 
+       Lehrstuhl f\"ur theoretische Physik II, \\
+       Institut f\"ur theoretische Physik und Astrophysik, 
+       Am Hubland,
+       D-97074 W\"urzburg, Germany}
 
@@
 \begin{quote}
-    Copyright \copyright  2004,2005  The GLoBES Team.
+    Copyright \copyright  2004--2007  The GLoBES Team.
     Permission is granted to copy, distribute and/or modify this document
@@
 On the experiment definition side, either built-in neutrino fluxes
-(\eg, neutrino factory) or arbitrary fluxes can be used. Similarly,
+(\eg, neutrino factory, $\beta$-Beam) or arbitrary fluxes can be used. Similarly,
 arbitrary cross sections, energy dependent efficiencies, the
@@
 For the systematics, energy
-normalization and calibration errors can be simulated. Note that
+normalization and calibration errors can be simulated in a straightforward way,
+or the systematics can be completely user-defined (Version 3.0 and higher). Note that
 the energy ranges and windows, as well as the bin widths can be
@@
 to produce a publication or a talk indicate that you have used \GLOBES\ and
-please cite the reference~\cite{globes_paper}
+please cite the following references~\cite{globes_paper,globes_paper_two}
 \begin{quote}
@@
 Simulation of long baseline neutrino oscillation experiments with \GLOBES\\
-arXiv:\glbxxx.  
+Comput. Phys. Commun. 167 (2005) 195, arXiv:\glbxxx.   
+\end{quote}
+
+\begin{quote}
+P. Huber, J. Kopp, M. Lindner, M. Rolinec, and W. Winter\\
+TO BE DEFINED, NEEDS TO BE UPDATED [WW]\\
+arXiv:\glbxxxtwo.   
 \end{quote}
@@
 from and if additional work has to be cited in addition to the 
-\GLOBES\ paper~\cite{globes_paper}. To assist with this task, we provide  the necessary information for the data coming along together with \GLOBES.
+\GLOBES\ papers~\cite{globes_paper,globes_paper_two}. To assist with this task, we provide  the necessary information for the data coming along together with \GLOBES.
 
@@
 It is recommended that you use the same style for your own input files, since, in case they are distributed, everybody will know how to correctly reference your work.
+
+\chapter*{What is new in \GLOBES ?}
+
+TO BE UPDATED [WW].
+
+Here we briefly summarize the main changes of the new GLoBES version. For details, please
+refer to the respective parts of the manual. Please note that any new \GLOBES\ version is
+compatible with older versions, \ie, old application software and AEDL files should also run with the new version.
+EXCEPTION: VERSION NUMBER IN AEDL REQUIRES?
+[WW] However, some function names and features will
+evolve during time, which means that outdated features may not be documented anymore.
+
+\section*{Version 3.0}
+\index{norm}{Version 3.0}
+
+Here comes a summary of the most important changes in this version for older users of GLoBES.
+
+\subsection*{New features}
+
+\begin{itemize}
+ \item
+  User-defined systematics to simulate experiments with correlated systematics \etc\ (such as
+reactor experiments); see \Sec~?
+ \item
+  User-defined priors to include arbitrary external information in the $\chi^2$ before
+marginalization over the oscillation parameters (such as to prevent minimizer from
+running into the wrong octant \etc); see \Sec~?
+\item
+  Non-standard physics support; see \Sec~?
+\item
+  Beta beam fluxes available as built-in fluxes; see \Sec~?
+\item
+  Enhanced support for parallelization, such as Condor; see \Sec~?
+\item
+  Updated AEDL files, new beta beam files; see \Sec~?
+\item
+ Faster probability engine
+\item
+  ...
+\end{itemize}
+
+\subsection*{Major changes}
+
+Most of the modifications below should not require that old pieces of software be changed. 
+However, the following changes are necessary:
+\begin{itemize}
+\item
+ AEDL requires now that {\tt \$version} be used to define the minimum version number this AEDL
+file is to be used with. With this requirement one can easily avoid that new AEDL
+files with new features be used with old versions of \GLOBES\ which may not recognize these features.
+\item
+ Some of the earlier AEDL files have been updated, changed names, or have been removed.
+In addition, new files have been added.
+Although old AEDL files will run as usual for compatibility, they will not be
+supported by the \GLOBES\ team anymore. You should make sure to keep these files when updating \GLOBES . 
+\item
+ ...
+\end{itemize}
+The following changes should not affect older software:
+\begin{itemize}
+\item
+ {\tt glbChi...} functions renamend for consistency
+\item
+ {\tt glbChiSetStartingValues} and {\tt glbChiGetStartingValues} renamed into 
+{\tt glbChiSetCentralValues} and {\tt glbChiGetCentralValues}
+\item
+ Systematics concept changed, concept of error dimensions removed
+\item
+ Some functions changed (such as for direct rate access)
+\item
+ ..
+\end{itemize}
 
@@
 simulation of neutrino factories is much more difficult. In order
-to demonstrate some of these difficulties, we present in this manual only
+to demonstrate some of these difficulties, we present in this manual mainly
 examples with neutrino factories. These examples can be found in
@@
 the actual numerical values of the results of the examples may be different
-from the ones in this manual.
+from the ones in this manual.  (IS THAT STILL SO? [WW])
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% PART III: Module Support
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% not yet!
-%\input{part3}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Appendix
@@
 We would like to thank Martin Freund, who wrote the very first
-version of a three-flavour matter profile treatment many years ago.
+version of a three-flavor matter profile treatment many years ago,
+and Thomas Schwetz, who has been pushing the software to the edge in the 
+past few years. Furthermore, we would like to thank Toshihiko Ota and WHO ELSE? [WW]
+for using and testing unpublished new features of the software.
 PH is especially thankful for the invaluable advice of Thomas Fischbacher on
-many design issues.
-In addition, we would like to thank
-Mark Rolinec for his help to translate the experiment descriptions
-into \AEDL, and for creating the illustrations in this manuscript.
+many design issues in the early stage of the project.
+% In addition, we would like to thank
+% Mark Rolinec for his help to translate the experiment descriptions
+% into \AEDL, and for creating the illustrations in this manuscript.
 Finally, thanks to all the people who have been pushing this project
@@
 
-This work and the development of \GLOBES\ have been supported by:
+This work and the development of \GLOBES\ have over the years been supported by:
 \begin{itemize}
-\item Physik Department der Technischen Universit\"at M\"unchen
-\item Max-Planck-Institut f\"ur Physik
+\item Technische Universit\"at M\"unchen [All authors]
+\item Max-Planck-Institut f\"ur Physik, M\"unchen [PH]
 \item Sonderforschungsbereich 375 f\"ur Astro-Teilchenphysik\\
-der Deutschen Forschungsgemeinschaft
-\item Studienstiftung des Deutschen Volkes
-\item Institute for Advanced Study, Princeton
-\item Physics Department of the University of Wisconsin, Madison
+der Deutschen Forschungsgemeinschaft [All authors]
+\item Studienstiftung des Deutschen Volkes [JK, WW]
+\item Institute for Advanced Study, Princeton [WW]
+\item W. M. Keck Foundation [WW]
+\item National Science Foundation [WW]
+\item University of Wisconsin, Madison [PH]
+\item Max-Planck-Institut f\"ur Kernphysik, Heidelberg [JK, ML]
+\item Emmy Noether-Programm der Deutschen Forschungsgemeinschaft [WW]
+\item Universit\"at W\"urzburg [WW]
 \end{itemize}

### 2007-01-03 Documentation/part1.tex
Part of manual updated

@@
 \index{norm}{GLOBES tour@\GLOBES~tour}
+
+TOUR ITSELF (+EXAMPLE FILE) TO BE UPDATED. JK: PLEASE LET ME KNOW WHEN {\tt example-tour.c} IS FIXED. MAYBE WE SHOULD  ALSO UPDATE THE PARAMTER VALUES [WW].
 
@@
 \multicolumn{3}{l}{{\bf Projections onto axes:}} \\
-\GLB{ChiTheta} & Projection onto $\theta_{13}$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
+\GLB{ChiTheta13} & Projection onto $\theta_{13}$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
 \GLB{ChiDelta} & Projection onto $\deltacp$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
 \GLB{ChiTheta23} & Projection onto $\theta_{23}$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
-\GLB{ChiDm} & Projection onto $\ldm$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
-\GLB{ChiDms} & Projection onto $\sdm$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
+\GLB{ChiDm31} & Projection onto $\ldm$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
+\GLB{ChiDm21} & Projection onto $\sdm$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
 \multicolumn{3}{l}{{\bf Projection onto plane:}} \\
-\GLB{ChiThetaDelta} & Projection onto $\theta_{13}$-$\deltacp$-plane  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
+\GLB{ChiTheta13Delta} & Projection onto $\theta_{13}$-$\deltacp$-plane  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
 \multicolumn{3}{l}{{\bf Projection onto any hyper-plane:}} \\
@@
 The format of \GLB{\_params} is discussed in detail in \Chapt~\ref{chapt:gettingstarted}.
-Note that all functions but {\tt glbChiSys} are using minimizers which have to be initialized with \GLB{SetInputErrors} and \GLB{SetStartingValues} first.}
+Note that all functions but {\tt glbChiSys} are using minimizers which have to be initialized with \GLB{SetInputErrors} and \GLB{SetCentralValues} first.}
 \end{table}
@@
  glbSetDensityParams(input\_errors,0.05,GLB\_ALL); \\
- glbSetStartingValues(true\_values); \\
+ glbSetCentralValues(true\_values); \\
  glbSetInputErrors(input\_errors);
@@
 \gq{
- chi2 = glbChiTheta(fit\_values,minimum,GLB\_ALL); \\ 
+ chi2 = glbChiTheta13(fit\_values,minimum,GLB\_ALL); \\ 
  fprintf(stream,"chi2 with correlations: \%g $\backslash$n",chi2); \\
@@
 \gq{
- chi2 = glbChiThetaDelta(fit\_values,minimum,GLB\_ALL); \\
+ chi2 = glbChiTheta13Delta(fit\_values,minimum,GLB\_ALL); \\
 fprintf(stream,"chi2 with correlations other than with deltacp: \%g $\backslash$n$\backslash$n",chi2); 
@@
   glbSetDensityParams(input\_errors,0.05,GLB\_ALL); \\
-  glbSetStartingValues(starting\_values);\\
+  glbSetCentralValues(starting\_values);\\
   glbSetInputErrors(input\_errors);\\
@@
  glbSetDensityParams(input\_errors,0.05,GLB\_ALL); \\
- glbSetStartingValues(true\_values); \\
+ glbSetCentralValues(true\_values); \\
  glbSetInputErrors(input\_errors); \\
- chi2 = glbChiTheta(fit\_values,minimum,0);\\ 
+ chi2 = glbChiTheta13(fit\_values,minimum,0);\\ 
   fprintf(stream,"chi2 with correlations for 3000km: \%g $\backslash$n",chi2); \\
  glbPrintParams(stream,minimum); \\ 
-  chi2b = glbChiTheta(fit\_values,minimum,1); \\ 
+  chi2b = glbChiTheta13(fit\_values,minimum,1); \\ 
  fprintf(stream,"$\backslash$nchi2 with correlations for 7500km: \\
@@
   glbPrintParams(stream,minimum); \\ 
- chi2sum = glbChiTheta(fit\_values,minimum,GLB\_ALL); \\ 
+ chi2sum = glbChiTheta13(fit\_values,minimum,GLB\_ALL); \\ 
  fprintf(stream,"$\backslash$nchi2 with correlations for combination: \\
@@
   glbSetDensityParams(input\_errors,0.05,GLB\_ALL);\\
-  glbSetStartingValues(starting\_values);\\
+  glbSetCentralValues(starting\_values);\\
   glbSetInputErrors(input\_errors);\\
@@
 script to take into account details of the installation on your system. Thus
-you've just to type \verb+make+ and you're done.
-\footnote{The  data files (\AEDL\ and supporting files) needed by the
+you've just to type \verb+make+ and you're done.\footnote{The  
+data files (\AEDL\ and supporting files) needed by the
 examples are already in place.} This \verb+Makefile+ very well serves
@@
 \\  
-  \hspace*{0.5cm} glbInitExperiment("NuFact.glb",\&glb\_experiment\_list[0], \\
-  \hspace*{1cm} \&glb\_num\_of\_exps); \hspace*{0.5cm} /* Initialize experiment NuFact.glb */\\
+  \hspace*{0.5cm} glbInitExperiment("NFstandard.glb",\&glb\_experiment\_list[0], \\
+  \hspace*{1cm} \&glb\_num\_of\_exps); \hspace*{0.5cm} /* Initialize experiment NFstandard.glb */\\
 \\  
@@
 \index{norm}{glb files@{\tt glb}-files} 
+TO BE UPDATED BY MR; POSSIBLY MORE DETAILED DESCRIPTION IN APPENDIX [WW].
 Pre-defined experiment prototypes, their filenames (to be used in \GLB{InitExperiment}), their short descriptions, and the references in which they are originally used and discussed (except from minor modifications, such as a
@@
 user.} In addition, using the parameter value \GLBC{GLB\_ALL} as
-experiment number initiates a combined analysis of all loaded experiments.
+experiment number (for example, in the {\tt glbChi...} functions) initiates a combined analysis of all loaded experiments.
 \index{norm}{GLB ALL@{\tt GLB\_ALL}}
@@
 
-In general \GLBC{GLB\_ALL} can be used in many cases where there is an argument
-selecting `$i$ out of $N$', \eg\ the 1st experiment out of 5, or the 5th
-rule of 20. In those cases  using \GLBC{GLB\_ALL} is equivalent to
-calling the corresponding function for \emph{all} $i$ in $N$  and `add' 
-the effect of each
-invocation, like in
-\begin{quote}
-\verb&for (i=0;i<N;i++) result += some_function(i);&\\
-is the same as\\
-\verb&result = some_function(GLB_ALL);&
-\end{quote}
-Here the meaning of `add' is, that whatever the desired result of calling
-\verb+some_function+ is, this result is obtained for each $i$ in $N$, \eg\ 
-setting the baseline in all experiments to a certain value or
-or compute the $\chi^2$ for each experiment and return the total result.
-There are however some functions where the action performed or the result
-is so complex that is not possible or sensible to perform this for all $i$
-in $N$. Calling these functions with  \GLBC{GLB\_ALL} as argument will 
-in any case result in an exit status indicating
-failure and the function will produce an 
-error message\footnote{if the verbosity
-level is set accordingly}.
+% THIS PART IS CONFUSING, BECAUSE THE EXPLAINED BEHAVIOR IS DIFFERENT FROM THE glbChi-functions
+% (THE MARGINALIZATION IS PERFORMED AFTER THE CHI2 IS ADDED! ONE MAY GET A DIFFERENT IMPRESSION
+% FROM THIS DESCRIPTION). THEREFORE, I HAVE REMOVED IT [WW].
+%
+% In general \GLBC{GLB\_ALL} can be used in many cases where there is an argument
+% selecting `$i$ out of $N$', \eg\ the 1st experiment out of 5, or the 5th
+% rule of 20. In those cases  using \GLBC{GLB\_ALL} is equivalent to
+% calling the corresponding function for \emph{all} $i$ in $N$  and `add' 
+% the effect of each
+% invocation, such as in
+% \begin{quote}
+% \verb&for (i=0;i<N;i++) result += some_function(i);&\\
+% is the same as\\
+% \verb&result = some_function(GLB_ALL);&
+% \end{quote}
+% Here the meaning of `add' is, that whatever the desired result of calling
+% \verb+some_function+ is, this result is obtained for each $i$ in $N$, \eg\ 
+% setting the baseline in all experiments to a certain value or
+% or compute the $\chi^2$ for each experiment and return the total result.
+% There are however some functions where the action performed or the result
+% is so complex that is not possible or sensible to perform this for all $i$
+% in $N$. Calling these functions with  \GLBC{GLB\_ALL} as argument will 
+% in any case result in an exit status indicating
+% failure and the function will produce an 
+% error message\footnote{if the verbosity
+% level is set accordingly}.
 
@@
 \GLBNS{InitExperiment}
-{\tt int glbInitExperiment(char *inf, glb\_exp *in, int *counter)}
- adds a single experiment with the filename {\tt inf} to the list of currently loaded experiments. The {\tt counter} is a pointer to the 
- variable containing the number of experiments, and the experiment {\tt in}
+{\tt int glbInitExperiment(char *infile, glb\_exp *ptr, int *counter)}
+ adds a single experiment with the filename {\tt infile} to the list of currently loaded experiments. The {\tt counter} is a pointer to the 
+ variable containing the number of experiments, and the experiment {\tt ptr}
  points to the beginning of the experiment list. The function returns
@@
 \begin{quote}
-{\tt glbInitExperiment("NuFact.glb",\&glb\_experiment\_list[0],\\  \hspace*{8cm} \&glb\_num\_of\_exps); }
+{\tt glbInitExperiment("NFstandard.glb",\&glb\_experiment\_list[0],\\  \hspace*{8cm} \&glb\_num\_of\_exps); }
 \end{quote}
-In this case, the experiment in the file {\tt NuFact.glb} is added to the internal global list of experiments, and the experiment counter is increased. The
+In this case, the experiment in the file {\tt NFstandard.glb} is added to the internal global list of experiments, and the experiment counter is increased. The
 experiment then has the number {\tt glb\_num\_of\_exps}-1. The elements
@@
 for oscillation parameters. In order to transfer sets of oscillation
-parameter vectors $(\theta_{12}, \theta_{13}, \theta_{23}, \deltacp, \sdm, \ldm)$, the parameter type \GLB{\_params} is used. In general,
+parameter vectors $(\theta_{12}, \theta_{13}, \theta_{23}, \deltacp, \sdm, \ldm)$
+as well as some other information, the parameter type \GLB{\_params} is used. In general,
  this type is often
@@
 an output parameter, the pointer has to be created, too, and point towards
-a structure which will contain the return values will be written to.
+a structure which the return values will be written to.
 This parameter transfer concept seems to be very sophisticated, but, as we
@@
 
-\label{mass_ordering}
-\index{norm}{Mass hierarchy}
-\paragraph{Note:} Inverting the mass hierarchy is not
-precisely the same than to change from $\Delta m^2_{31}
-\rightarrow -\Delta m^2_{31}$.
-In this case the absolute value of $\Delta m^2_{32}$ changes also, which
-introduces a new frequency to the problem. Therefore, if we assume
-normal hierarchy  whenever $\Delta m^2_{31}>0$, the corresponding
-point in parameters space for inverted hierarchy is given by $\Delta 
-m^2_{31}\rightarrow - \Delta m^2_{31} + \Delta m^2_{21}$, because
-with this defintion the absolute value of  $\Delta m^2_{32}$ is unchanged
-and no new frequency is introduced.\\
-
 Another piece of information will be returned from the minimizers 
@@
 \GLBNS{DefineParams}
-{\tt glb\_params glbDefineParams(glb\_params in, double theta12, double theta13,double theta23, double delta, double dms, double dma)} assigns
+{\tt glb\_params glbDefineParams(glb\_params in, double theta12, double theta13,double theta23, double delta, double dm21, double dm31)} assigns
 the complete set of oscillation parameters to the vector {\tt in}, which has
@@
 In both of these functions, the parameter {\tt which} runs from
-$0$ to $5$, where the parameters in \GLOBES\ always have the order $\theta_{12}$, $\theta_{13}$, $\theta_{23}$, $\deltacp$, $\sdm$, $\ldm$.
+$0$ to $5$ (or the number of oscillation parameters-1), where the parameters in \GLOBES\ always have the order $\theta_{12}$, $\theta_{13}$, $\theta_{23}$, $\deltacp$, $\sdm$, $\ldm$.
 Alternatively to the number, the constants \GLBC{GLB\_THETA\_12}, \GLBC{GLB\_THETA\_13}, \GLBC{GLB\_THETA\_23}, \GLBC{GLB\_DELTA\_CP}, 
@@
 \item
- The program should only run with this specific version of \GLOBES
-\item
- The program can only run with a minimum version of \GLOBES
+ The program should only run with this specific version of \GLOBES .
 \item
  The program can only run up to a certain \GLOBES\ version.
+\item
+ The program can only run with a minimum version of \GLOBES .
+\item
+ The program and AEDL file versions are compatible.
 \end{itemize}
-The same holds for \AEDL\ files: For example, 
-some features may not be supported by earlier versions of \GLOBES\ anymore.
-The program can then check the version of the \AEDL\ file and break if it is
-too old.
-
+However, note that GLOBES~3.0 and higher requires that the \GLOBES\ version be at least as new as the
+version the AEDL file was written for.
+%
 The functions in \GLOBES\ for version control are:
@@
 \end{function}
+WHAT IS THE DIFFERENCE BETWEEN THE LIBRARY AND GLOBES VERSION???? [WW]
 \begin{function}
@@
 {\tt const char* glbVersionOfExperiment(int experiment)} returns
-the version string of the experiment number {\tt experiment}.
+the version string of the experiment number {\tt experiment} (set by
+{\tt \$version} in AEDL).
 The version string is allocated within the experiment structure, which means
@@
 \begin{function}
-\GLBNS{SetStartingValues}
-{\tt int glbSetStartingValues(const glb\_params in)} sets the starting values for all of the following minimizer calls to {\tt in}.
+\GLBNS{SetCentralValues}
+{\tt int glbSetCentralValues(const glb\_params in)} sets the starting values for all of the following minimizer calls to {\tt in}.
 \end{function}
@@
 \begin{function}
-\GLBNS{GetStartingValues}
-{\tt int glbGetStartingValues(glb\_params out)} writes the currently
+\GLBNS{GetCentralValues}
+{\tt int glbGetCentralValues(glb\_params out)} writes the currently
 set starting values to {\tt out}.
@@
  glbSetDensityParams(input\_errors,0.05,GLB\_ALL);\\
- glbSetStartingValues(true\_values);\\
+ glbSetCentralValues(true\_values);\\
  glbSetInputErrors(input\_errors);\\
@@
   glbSetDensityParams(input\_errors,0.05,GLB\_ALL);\\
-  glbSetStartingValues(true\_values);\\
+  glbSetCentralValues(true\_values);\\
   glbSetInputErrors(input\_errors);\\
@@
 \hspace*{0.5cm} /* Compute Chi2 for full correlation: minimize over all but theta13 */ \\
-\hspace*{0.5cm} res2=glbChiTheta(test\_values,NULL,GLB\_ALL); \\
+\hspace*{0.5cm} res2=glbChiTheta13(test\_values,NULL,GLB\_ALL); \\
   \\
@@
 \begin{function}
-\GLBNS{ChiTheta}
-{\tt double glbChiTheta(const glb\_params in, glb\_params out, int exp)}
+\GLBNS{ChiTheta13}
+{\tt double glbChiTheta13(const glb\_params in, glb\_params out, int exp)}
 returns the projected $\chi^2$ onto the $\theta_{13}$-axis for the  experiment {\tt exp}. For the simulation of all initialized experiments,
@@
 % 
-Note that before any minimization, \GLB{SetStartingValues} 
+Note that before any minimization, \GLB{SetCentralValues} 
 and \GLB{SetInputErrors} have to be used at least once.
-In addition, note that the resulting $\chi^2$ of {\tt glbChiTheta} (or {\tt glbChiDelta}) for the combination of more than one experiment is not
+In addition, note that the resulting $\chi^2$ of {\tt glbChiTheta13} (or {\tt glbChiDelta}) for the combination of more than one experiment is not
 equal to the sum of the individual $\chi^2$-values anymore.
@@
 matter density scaling factor of the appropriate experiment.
-For example, calculating {\tt glbChiTheta} for the last experiment
+For example, calculating {\tt glbChiTheta13} for the last experiment
 number, the last density value will be returned.
@@
 
-Other examples for projections onto a parameter axis while keeping exactly one parameter fixed are \GLB{ChiTheta23}, \GLB{ChiDm}, and \GLB{ChiDms}, which
+Other examples for projections onto a parameter axis while keeping exactly one parameter fixed are \GLB{ChiTheta23}, \GLB{ChiDm31}, and \GLB{ChiDm21}, which
 can be found in \Tab~\ref{tab:stdfunctions} on page~\pageref{tab:stdfunctions}.
@@
 \begin{function}
-\GLBNS{ChiThetaDelta}
-{\tt double glbChiThetaDelta(const glb\_params in, glb\_params out, int exp)} returns the projected $\chi^2$ onto the $\theta_{13}$-$\deltacp$-plane for the  experiment {\tt exp}.
+\GLBNS{ChiTheta13Delta}
+{\tt double glbChiTheta13Delta(const glb\_params in, glb\_params out, int exp)} returns the projected $\chi^2$ onto the $\theta_{13}$-$\deltacp$-plane for the  experiment {\tt exp}.
 For the simulation of all initialized experiments,
@@
 the manifold and to adjust the parameter ranges. Then, one can run
-{\tt glbChiThetaDelta} for a complete evaluation of the problem
+{\tt glbChiTheta13Delta} for a complete evaluation of the problem
 including correlations.
@@
 parameters, but flags \GLBC{GLB\_FIXED} or \GLBC{GLB\_FREE}.
-For example, the projection onto the $\theta_{13}$-axis {\tt glbChiTheta}
+For example, the projection onto the $\theta_{13}$-axis {\tt glbChiTheta13}
 is nothing else than a special case of {\tt glbChiNP} with $\theta_{13}$
@@
   \\
-  glbSetStartingValues(starting\_values); \\
+  glbSetCentralValues(starting\_values); \\
   glbSetInputErrors(input\_errors); \\
@@
 In this case, the position of the degeneracy can be easily guessed to be 
-at the true parameter values but with inverted\footnote{For a exact 
-definition of 
-inverted hierarchy, see page~\pageref{mass_ordering}.} $\ldm$.
+at the true parameter values but with inverted  $\ldm$.
 The minimizer then runs off the plane of these parameters into the local minimum.
@@
 Thus, the degeneracy may not even appear at the chosen confidence level in the plane, but it does appear at the real minimum.
-The two sections through the fit manifold shown in the figure on page~\pageref{ex:sgndeg} therefore do not appear at the same oscillation parameter values (except from the ones shown in the
-figure).\footnote{The discussed figure on page~\pageref{ex:sgndeg} is produced by  {\tt glbChiSys} and thus only represents a section through the fit manifold.
-For the projection including correlations, one may rather want to use {\tt glbChiThetaDelta}.}
+The two sections through the fit manifold shown in the figure on page~\pageref{ex:sgndeg} therefore do not appear at the same oscillation parameter values.\footnote{The discussed figure on page~\pageref{ex:sgndeg} is produced by  {\tt glbChiSys} and thus only represents a section through the fit manifold.
+For the projection including correlations, one may rather want to use {\tt glbChiTheta13Delta}.}
+
+\label{mass_ordering}
+\index{norm}{Mass hierarchy}
+\paragraph{Note:} Inverting the mass hierarchy is not
+precisely the same than to change from $\Delta m^2_{31}
+\rightarrow -\Delta m^2_{31}$.
+In this case the absolute value of $\Delta m^2_{32}$ changes also, which
+introduces a new frequency to the problem. Therefore, if we assume
+normal hierarchy  whenever $\Delta m^2_{31}>0$, the corresponding
+point in parameters space for inverted hierarchy is given by $\Delta 
+m^2_{31}\rightarrow - \Delta m^2_{31} + \Delta m^2_{21}$, because
+with this defintion the absolute value of  $\Delta m^2_{32}$ is unchanged
+and no new frequency is introduced.\\
 
@@
 \item[Locating degeneracies in more complicated topologies.] For more complicated topologies, such as for neutrino factories, it is often useful to use multi-step location procedures or analytical knowledge.
-For example, for a numerical procedure, one may first of all switch off the systematics and keep $\stheta$ or $\deltacp$ fixed, \ie, use {\tt glbChiTheta}, where $\stheta$ or $\deltacp$ is fixed at the best-fit value.
+For example, for a numerical procedure, one may first of all switch off the systematics and keep $\stheta$ or $\deltacp$ fixed, \ie, use {\tt glbChiTheta13}, where $\stheta$ or $\deltacp$ is fixed at the best-fit value.
 The result can then be used as a starting point for {\tt glbChiAll} for the individual experiments with the systematics switched on again.
@@
 In addition, it uses a little trick: It avoids falling into the wrong
- solution with {\tt glbChiTheta}
+ solution with {\tt glbChiTheta13}
 by using the fit value of $\deltacp$ from the step before as prediction
@@
  glbSetOscillationParameters(); glbSetRates(); \\
- glbSetStartingValues(true\_values);glbSetInputErrors(input\_errors); \\
+ glbSetCentralValues(true\_values);glbSetInputErrors(input\_errors); \\
  \\
@@
 \end{quote}
-This code fragment illustrates the implementation of a {\tt glbChiTheta} function
+This code fragment illustrates the implementation of a {\tt glbChiTheta13} function
 including the non-standard effect, which is being marginalized over.
-Alternatively, one could define a {\tt glbChiTheta} fixing the non-standard
+Alternatively, one could define a {\tt glbChiTheta13} fixing the non-standard
 parameter, or include external constrainys to limit its size. Therefore, 

### 2007-01-03 Documentation/part2.tex
Corrected the description of the cross section files
Part of manual updated

@@
 
-\index{aedl}{inlcude@{\tt include}}
+\index{aedl}{include@{\tt include}}
 As another feature of AEDL one can use include files with the {\tt include} command. Includes can be nested up to {\tt MAX\_INCLUSION\_DEPTH}, which is currently set to $10$. Error reporting works 
@@
 
-Cross sections are in \GLOBES\ given as differential cross section per energy:
+Cross sections in \GLOBES\ are given as total cross section divided by energy:
 \begin{equation}
 \hat\sigma(E)=\sigma(E)/E\,\left[ 10^{-38}\,
-\frac{\mathrm{cm}^2}{\mathrm{GeV}^2} \right]
+\frac{\mathrm{cm}^2}{\mathrm{GeV}} \right]
 \end{equation}

### 2007-01-04 Documentation/part1.tex
Systematics changed

@@
 
-For some experimental setups the built-in systematics functions are not sufficient
+\example{Systematics for a reactor setup with near and far detector}{\label{ex:syst}
+\index{norm}{Systematics!Example, user-defined}
+
+The following code fragment from the much more extensive {\tt example5.c}
+calculates $\chi^2$ for a reactor setup with two detectors
+including five different systematical errors: the flux normalization
+of the reactor ({\tt x[0]}), the fiducial mass uncertainties of the far
+({\tt x[1]}) and near ({\tt x[2]}) detector, and energy calibration errors for the far
+({\tt x[3]}) and near ({\tt x[4]}) detectors.  The calculation follows
+\equ{chi2reactor}, but includes the energy calibration.
+
+\begin{quote}{\tt \footnotesize
+
+double likelihood(double true\_rate, double fit\_rate, double sqr\_sigma) \\
+\{ if (sqr\_sigma > 0) return square(true\_rate - fit\_rate) / sqr\_sigma; else return 0.0; \}
+
+double chiDCNorm(int exp, int rule, int np, double *x, double *errors) \\
+\{ \\
+  \hspace*{0.3cm} const EXP\_FAR = 0; const EXP\_NEAR = 1; \\
+  \hspace*{0.3cm} int n\_bins = glbGetNumberOfBins(EXP\_FAR); \\
+  \hspace*{0.3cm} double *true\_rates\_N = glbGetRuleRatePtr(EXP\_NEAR, 0); \\
+  \hspace*{0.3cm} double *true\_rates\_F = glbGetRuleRatePtr(EXP\_FAR, 0); \\
+  \hspace*{0.3cm} double signal\_fit\_rates\_N[n\_bins]; double signal\_fit\_rates\_F[n\_bins]; \\
+  \hspace*{0.3cm} double signal\_norm\_N, signal\_norm\_F; \\
+  \hspace*{0.3cm} int ew\_low, ew\_high, i; \\
+  \hspace*{0.3cm} double emin, emax, fit\_rate;  double chi2 = 0.0; \\
+  \\
+   \hspace*{0.3cm} /* Request simulated energy interval and analysis energy window */ \\
+   \hspace*{0.3cm} glbGetEminEmax(exp, \&emin, \&emax); \\
+   \hspace*{0.3cm} glbGetEnergyWindowBins(exp, rule, \&ew\_low, \&ew\_high);
+
+  \hspace*{0.3cm} /* Apply energy calibration error */ \\
+  \hspace*{0.3cm} glbShiftEnergyScale(x[3], glbGetSignalFitRatePtr(EXP\_FAR, 0), \\
+  \hspace*{0.9cm}                    signal\_fit\_rates\_F, n\_bins, emin, emax); \\
+  \hspace*{0.3cm} glbShiftEnergyScale(x[4], glbGetSignalFitRatePtr(EXP\_NEAR, 0), \\
+  \hspace*{0.9cm}                    signal\_fit\_rates\_N, n\_bins, emin, emax);
+
+  \hspace*{0.3cm} /* Loop over all bins in energy window */ \\
+  \hspace*{0.3cm} signal\_norm\_F = 1.0 + x[0] + x[1]; \\
+  \hspace*{0.3cm} signal\_norm\_N = 1.0 + x[0] + x[2]; \\
+  \hspace*{0.3cm} for (i=ew\_low; i <= ew\_high; i++) \\
+  \hspace*{0.3cm} \{ \\
+  \hspace*{0.6cm}  /* Statistical part of chi2 for far detector */ \\
+  \hspace*{0.6cm}  fit\_rate  = signal\_norm\_F * signal\_fit\_rates\_F[i]; \\
+  \hspace*{0.6cm}  chi2 += likelihood(true\_rates\_F[i], fit\_rate, true\_rates\_F[i]); \\
+   \\
+  \hspace*{0.6cm}  /* Statistical part of chi2 for near detector  */ \\
+  \hspace*{0.6cm}  fit\_rate  = signal\_norm\_N * signal\_fit\_rates\_N[i]; \\
+  \hspace*{0.6cm}  chi2 += likelihood(true\_rates\_N[i], fit\_rate, true\_rates\_N[i]); \\
+  \hspace*{0.3cm} \}
+
+  \hspace*{0.3cm} /* Systematical part of chi2 (= priors) */ \\
+  \hspace*{0.3cm} for (i=0; i < np; i++)  chi2 += square(x[i]/errors[i]);
+
+ \hspace*{0.3cm} return chi2; \\
+\} 
+}
+\end{quote}
+}
+
+For some experimental setups, the built-in systematics functions are not sufficient
 because they cannot handle systematical errors that are correlated between different
-parts of a multi-detector or multi-source setup. Therefore \GLOBES\ allows the user
+parts of a multi-detector setup. Therefore \GLOBES\ allows the user
 to override the default $\chi^2$ calculation. The reader who wants to use this
-feature should be familiar with \Sec~\ref{sec:channel} and~\Sec~\ref{sec:rules}
-of the \AEDL\ manual.
-
-To specify a user-defined $\chi^2$ algorithm, the following function can be used:
-\begin{function}
-\GLBNS{SetUserChi}
-{\tt void glbSetUserChi(int exp,int rule,double (*chi\_func)(double x[]),int dim,%
-double centers[],double errors[],char *info)} tells \GLOBES\ to use {\tt chi\_func}
-instead of the builtin functions to calculate $\chi^2$ for experiment {\tt exp} and
-rule {\tt rule}. These parameters can take the value \GLBC{GLB\_ALL} to specify that
-{\tt chi\_func} should be called once for each experiment or each rule. {\tt dim} is
-the number of variables parametrizing the systematical errors, while {\tt centers[]}
-and {\tt errors[]} give their central values and standard deviations, respectively.
-When \GLOBES\ calls {\tt chi\_func}, the parameter array {\tt x[]} will contain
-the values of the systematics parameters for which $\chi^2$ should be calculated.
-Deviating from the standard conventions, the entries of {\tt x[]} will start at index 1
-(The indices for {\tt params} are {\tt errors} in the call to {\tt glbSetUserChi} start
-at 0 as usual).
-\end{function}
-
-In addition to calling {\tt glbSetUserChi}, {\tt @errordim\_sys\_on} has to be set
-to 20 in the \AEDL\ file (see~\Sec~\ref{sec:rules}).
-
-Note that {\tt chi\_func} will be called many times by the \GLOBES\ minimizers.
-Therefore the function should be as efficient as possible. The function
-will need low-level access to internal \GLOBES\ data. Therefore the following
-routines are provided:
-\begin{function}
-\GLBNS{GetCurrentExp}
-{\tt int glbGetCurrentExp()} gives the number of the experiment for which
-{\tt chi\_func} has been called.
-\end{function}
-\begin{function}
-\GLBNS{GetCurrentRule}
-{\tt int glbGetCurrentRule()} gives the number of the rule in the current experiment
-for which {\tt chi\_func} has been called.
-\end{function}
-\begin{function}
-\GLBNS{GetObservedBinInRule}
-{\tt double glbGetObservedBinInRule(int exp,int rule,int bin)} returns the number of
-signal events for experiment {\tt exp}, rule {\tt rule} and bin {\tt bin}, as calculated
-in the preceding call to {\tt glbSetRates}.
-\end{function}
-\begin{function}
-\GLBNS{GetSignalBinInRule}
-{\tt double glbGetSignalBinInRule(int exp,int rule,int bin)} returns the number of
-signal events for experiment {\tt exp}, rule {\tt rule} and bin {\tt bin}, calculated
-for the oscillation parameter values that are currently being tested.
-\end{function}
-\begin{function}
-\GLBNS{GetBackgroundBinInRule}
-{\tt double glbGetBackgroundBinInRule(int exp,int rule,int bin)} returns the number of
-background events for experiment {\tt exp}, rule {\tt rule} and bin {\tt bin}, calculated
-for the oscillation parameter values that are currently being tested.
-\end{function}
-\begin{function}
-\GLBNS{GetChannelBin}
-{\tt double glbGetChannelBin(int exp,int channel,int bin)} returns the number of
-signal events for experiment {\tt exp}, channel {\tt channel} and bin {\tt bin}, calculated
-for the oscillation parameter values that are currently being tested.
-\end{function}
-\begin{function}
-\GLBNS{ShiftSignalEnergyScale}
-{\tt int glbShiftSignalEnergyScale(int exp,int rule,double amount)} applies an energy
-calibration error of magnitude {\tt amount} to the signal event rates in experiment
-{\tt exp} and rule {\tt rule}. Subsequent calls to {\tt glbGetSignalBinInRule} will
-return the modified event rates. For details on energy calibration errors, see
-\Sec~\ref{sec:rules}.
-\end{function}
-\begin{function}
-\GLBNS{ShiftBackgroundEnergyScale}
-{\tt int glbShiftBackgroundEnergyScale(int exp,int rule,double amount)} applies an energy
-calibration error of magnitude {\tt amount} to the background event rates in experiment
-{\tt exp} and rule {\tt rule}. Subsequent calls to {\tt glbGetBackgroundBinInRule} will
-return the modified event rates. For details on energy calibration errors, see
-\Sec~\ref{sec:rules}.
-\end{function}
-
-Typical applications for user-defined systematics are reactor neutrino experiments,
-where systematics play a crucial role. For a setup with near and far detectors,
+feature should be familiar with \Secs~\ref{sec:event_rates},~\ref{sec:channel}, and~\ref{sec:rules}
+of the manual. Note that this is an advanced topic, which requires \GLOBES\ 3.0 or
+higher. It can be omitted in a first reading of the manuscript.
+
+User-defined systematics is, compared to the built-in systematics, a compound between
+the AEDL definition and some source code defining the behavior. In the simplest case,
+the AEDL file defines user-defined systematics for a specific rule and gives it a 
+name/identifier,  as will as it defines the systematical errors. The source code has then to match this AEDL
+definition -- therefore, make sure to use only unique identifiers. 
+Ones needs to register the user-defined systematics in the 
+source code (typically after {\tt glbInit}) with
+\begin{function}
+\GLBNS{DefineChiFunction}
+{\tt int glbDefineChiFunction(glb\_chi\_function chi\_func, int dim, const char *name)}
+ tells \GLOBES\ to register the user-defined systematics $\chi^2$ function {\tt chi\_func} 
+identified by the string {\tt name} with {\tt dim} systematics parameters. Returns zero, if successful.
+\end{function}
+The user-defined systematics function itself has to have the pre-defined type 
+\begin{function}
+\GLBNS{\_chi\_function}
+{\tt double (*glb\_chi\_function)(int exp, int rule, int dim, double *params, double *errors)}.
+In this function type, {\tt exp} is the experiment number, {\tt rule} is the rule number, {\tt dim} is the number of systematics parameters, {\tt params} is an array of the systematics auxiliary parameters themselves, and {\tt errors} is an array with the systematical errors.
+Note the the central values for {\tt params} are $0$, which means that $0$ corresponds to the un-modified rates.
+The array indices are running from $0$ to {\tt dim}-1.
+\end{function}
+Note that this function will be called many times by the \GLOBES\ minimizers.
+Therefore the function should be as efficient as possible.
+In many cases, it is also recommendable to use different functions for different rules, since the fewer auxiliary parameters in the minimization, the faster the routine will be.
+In addition, note that complicated systematics with many parameters may introduce complicated (unknown) topologies for the minimization, which means that the minimizers may
+end up in a local minimum instead of the global minimum. \GLOBES\ provides the function \GLB{SetSysStartingValuesList} to change the starting values of the minimizer in the case of
+ convergence problems (see below).
+A typical implementation may look like the following code, where ``{\tt chiDCNorm}'' is the identifier
+assigned in the AEDL definition:
+\begin{quote}
+{\tt
+  double chiDCNorm(int exp, int rule, int dim, double *x, \\
+  \hspace*{0.5cm} double *errors) \\
+  \{ \\
+   \hspace*{0.5cm} double chi2 = 0.0; \\
+   \hspace*{0.5cm} int i; \\
+   \\
+   \hspace*{0.5cm} ... /* Some code to calculate the chi2 */ \\
+   \\
+    \hspace*{0.5cm} /* Here the systematics priors (penalties) are added: */ \\
+    \hspace*{0.5cm} for(i=0;i<dim;i++) chi2 += (x[i]*x[i]/errors[i]/errors[i]);
+    \hspace*{0.5cm} return chi2; \\
+  \} \\
+   \\
+  int main(int argc,char*argv[]) \\
+ \{ \\ 
+  \hspace*{0.5cm} ... \\
+  \hspace*{0.5cm} glbInit(agrv[0]); \\
+  \hspace*{0.5cm} glbDefineChiFunction(\&chiDCNorm,5,"chiDCNorm"); \\
+  \hspace*{0.5cm} ... \\
+ \} \\
+}
+\end{quote}
+During running time, it is also possible to change the user-defined systematics of an experiment or
+rule (as compared to the systematics assigned in the AEDL file or earlier) with the following function:
+\begin{function}
+\GLBNS{SetChiFunction}
+{\tt int glbSetChiFunction(int exp, int rule, int on\_off, const char *name, double *errors)}
+ tells \GLOBES\ to use the registered user-defined systematics identified by the string {\tt name}
+instead of the built-in systematics
+to calculate $\chi^2$ for the experiment {\tt exp} and the rule {\tt rule}. 
+Both of the parameters {\tt exp} and {\tt rule} can take the value \GLBC{GLB\_ALL} to specify that the
+user-defined systematics can be used for each experiment or each rule. The parameter 
+{\tt on\_off} determines if the systematics function should be used for systematics on
+{\tt GLB\_ON} or systematics off {\tt GLB\_OFF}.
+The array {\tt errors} sets the systematical errors in the order they
+are defined in the AEDL file (running from $0$ to the number of systematics parameters-1). 
+Returns zero, if successful.
+\end{function}
+Note that the new user-defined systematics has to be registered with {\tt glbDefineChiFunction} first.
+
+The $\chi^2$ function {\tt glb\_chi\_function} may use a number of helper functions for
+for the $\chi^2$ calculation. Some of these functions may not be directly related to user-defined
+systematics, but may be mainly used in this context:
+\begin{function}
+\GLBNS{GetEminEmax}
+{\tt int glbGetEminEmax(int experiment, double *emin, double *emax)} returns the
+energy range [{\tt emin},{\tt emax}] covered by the binning as defined in the AEDL file for the experiment
+{\tt experiment}.
+\end{function}
+\begin{function}
+\GLBNS{GetEnergyWindow}
+{\tt int glbGetEnergyWindow(int exp, int rule, double *low, double *high)} returns the
+energy window [{\tt low},{\tt high}] defined in the AEDL file for the experiment
+{\tt exp} and rule {\tt rule}.
+\end{function}
+\begin{function}
+\GLBNS{GetEnergyWindowBins}
+{\tt int glbGetEnergyWindowBins(int exp, int rule, int *lowbin, int *highbin)} returns the
+energy window in terms of bin numbers [{\tt lowbin},{\tt highbin}] corresponding to the energy window defined in the AEDL file for the experiment {\tt exp} and rule {\tt rule}.
+\end{function}
+\begin{function}
+\GLBNS{GetNumberOfBins}
+{\tt int glbGetNumberOfBins(int exp)} returns the number of bins for the experiment {\tt exp}.
+\end{function}
+\begin{function}
+\GLBNS{GetNumberOfSimBins}
+{\tt int glbGetNumberOfSimBins(int exp)} returns the number of sampling points for the experiment {\tt exp}.
+\end{function}
+SHOULD THIS NOT BE CALLED {\tt glbGetNumberOfSamplingPoints}??? [WW]
+
+A very useful function to include energy calibration errors is
+\begin{function}
+\GLBNS{ShiftEnergyScale}
+{\tt glbShiftEnergyScale(double b, double *rates\_in, double *rates\_out,int bins, double emin, double emax)}
+shifts the energy scale in {\tt rates\_in} by the amount {\tt b} and stores the result in 
+{\tt rates\_out}. The parameters {\tt emin} and {\tt emax} denote the minimal and maximal energy
+as obtained with {\tt glbGetEminEmax}.
+\end{function}
+DOES THAT ALSO WORK WITH VARIABLE BINS? WHAT IS THE MEAING OF EMIN AND EMAX? [WW]
+For details on energy calibration errors, see \Sec~\ref{sec:rules}.
+%
+In addition, in order to implement user-defined systematics, one usually needs low-level access to the event rate vectors using \GLB{GetRuleRatePtr}, \GLB{GetSignalFitRatePtr}, \GLB{GetBGFitRatePtr} and possibly other functions described in \Sec~\ref{sec:event_rates}.
+We recommend that you familiarize yourself with these functions at this point.
+
+Let us now discuss a simple example.
+Typical applications for user-defined systematics are reactor neutrino experiments, where systematics play a crucial role.
+For a setup with near and far detectors,
 the formula for $\chi^2$ is
@@
 while their standard deviations are $\sigma_R$, $\sigma_N$ and $\sigma_F$.
-
+%
 The first line of \equ{chi2reactor} is the standard $\chi^2$ expression for
@@
 for deviations of the systematics parameters from their central values. 
-
-\example{Systematics for a reactor setup with near and far detector}{
-The following code fragment calculates $\chi^2$ for a reactor setup with
-a far detector (experiment 0) and a near detector (experiment 1),
-including six different systematical errors: the flux normalization
-of the reactor ({\tt x[1]}), the fiducial mass uncertainties of the far
-({\tt x[2]}) and near ({\tt x[3]}) detector, the normalization of the
-background ({\tt x[4]}) and energy calibration errors for the far
-({\tt x[5]}) and near ({\tt x[6]}) detectors. We assume that the central
-values and errors of these parameters are stored in the global arrays
-{\tt sys\_centers} and {\tt sys\_errors}. The calculation follows
-\equ{chi2reactor}.
-
-\begin{quote}{\tt \footnotesize
-  double chi2nearfar(double *x)                                                 \\
-  \{                                                                            \\
-\hspace*{0.5 cm}      double chi2 = 0;                                          \\
-\hspace*{0.5 cm}      double chirate, measured\_rate;                           \\
-\hspace*{0.5 cm}                                                                \\
-\hspace*{0.5 cm}      /* Assign complete chi\textasciicircum2 to first experiment */ \\
-\hspace*{0.5 cm}      if (glbGetCurrentExp() != 0)  return 0;                   \\
-\hspace*{0.5 cm}                                                                \\
-\hspace*{0.5 cm}      glbShiftSignalEnergyScale(0, 0, x[5]);   /* Apply E calib. error */ \\
-\hspace*{0.5 cm}      glbShiftSignalEnergyScale(1, 0, x[6]);                    \\
-\hspace*{0.5 cm}      glbShiftBackgroundEnergyScale(0, 0, x[5]);                \\
-\hspace*{0.5 cm}      glbShiftBackgroundEnergyScale(1, 0, x[6]);                \\
-\hspace*{0.5 cm}      for (int i=0; i < NBINS; i++) \{                          \\
-\hspace*{1.0 cm}        /* Statistical part of chi\textasciicircum2 for far detector */ \\
-\hspace*{1.0 cm}        chirate = (x[1]+x[2]) * glbGetSignalBinInRule(0,0,i) +  \\
-\hspace*{2.0 cm}                        (x[4]+x[3]) * glbGetBackgroundBinInRule(0,0,i);\\
-\hspace*{1.0 cm}        measured\_rate = glbGetObservedBinInRule(0,0,i);        \\
-\hspace*{1.0 cm}        chi2 += 2 * (measured\_rate - chirate);                 \\
-\hspace*{1.0 cm}        if (measured\_rate > 0  \&\&  chirate > 0)              \\
-\hspace*{1.5 cm}          chi2 += 2 * chirate * log(chirate / measured\_rate);  \\
-\hspace*{1.0 cm}                                                                \\
-\hspace*{1.0 cm}        /* Statistical part of chi\textasciicircum2 for near detector */ \\
-\hspace*{1.0 cm}        chirate = (x[1]+x[3]) * glbGetSignalBinInRule(1,0,i) +  \\
-\hspace*{2.0 cm}                        (x[4]+x[3]) * glbGetBackgroundBinInRule(1,0,i);\\
-\hspace*{1.0 cm}        measured\_rate = glbGetObservedBinInRule(1,0,i);        \\
-\hspace*{1.0 cm}        chi2 += 2 * (measured\_rate - chirate);                 \\
-\hspace*{1.0 cm}        if (measured\_rate > 0  \&\&  chirate > 0)              \\
-\hspace*{1.5 cm}          chi2 += 2 * chirate * log(chirate / measured\_rate);  \\
-\hspace*{0.5 cm}      \}                                                        \\
-\hspace*{0.5 cm}      for (int j=0; j < 6; j++)  /* Penalties */                \\
-\hspace*{1.0 cm}        chi2 += square((x[j+1] - sys\_centers[j]) / sys\_errors[j]);\\
-\hspace*{0.5 cm}      return chi2;                                              \\
-  \}                                                                            
+%
+We use two AEDL files for this experiment, one for the far detector and one for the
+near detector.
+The corresponding rule definition ($\nu_e$ disappearance) in the far detector AEDL-file is
+\begin{quote}
+{\tt
+rule(\#rule0)< \\
+\hspace*{0.5cm}        @signal     = 1.0@\#nu\_e\_disappearance\_CC \\
+\hspace*{0.5cm}        @background = 0.0@\#nu\_e\_disappearance\_CC   /* No background */ \\
+\hspace*{0.5cm} @energy\_window = 0.0015 : 0.01 \\
+\\
+\hspace*{0.5cm}        @sys\_off\_function = "chiNoSysSpectrum" \\
+\hspace*{0.5cm}        @sys\_off\_errors   = \{ \} \\
+\hspace*{0.5cm}        @sys\_on\_function  = "chiDCNorm" \\
+\hspace*{0.5cm}        @sys\_on\_errors    = \{ 0.02,    0.006,        0.006,       0.005,     0.005   \} \\
+\hspace*{0.5cm}        \mbox{ // \{ Flux, Fid. mass FD, Fid. mass ND, Energy FD, Energy ND \} } \\
+> 
 }
 \end{quote}
-}
+In this case, the systematics is called ``{\tt chiDCNorm}'' for systematics on, whereas systematics
+off is computed with spectral information, but without systematical errors.
+For the near detector AEDL-file, we have instead
+\begin{quote}
+{\tt
+rule(\#rule0)< \\
+\hspace*{0.5cm}        @signal     = 1.0@\#nu\_e\_disappearance\_CC \\
+\hspace*{0.5cm}        @background = 0.0@\#nu\_e\_disappearance\_CC   /* No background */ \\
+\hspace*{0.5cm} @energy\_window = 0.0015 : 0.01 \\
+\\
+\hspace*{0.5cm}        @sys\_off\_function = "chiNoSysSpectrum" \\
+\hspace*{0.5cm}        @sys\_off\_errors   = \{ \} \\
+\hspace*{0.5cm}        @sys\_on\_function  = "chiZero" \\
+\hspace*{0.5cm}        @sys\_on\_errors    = \{ \} \\
+> 
+}
+\end{quote}
+In this case, the systematics {\tt chiZero} is used for systematics on, which means that there will
+be no active $\chi^2$ calculation in this rule. With this definition, the user-defined systematics
+will only be called once for the far detector. However, the rates from the near detector will be 
+passively provided for the common $\chi^2$ function.
+You can find the corresponding files {\tt dchooz-near.glb} and {\tt dchooz-far.glb}, you will need for {\tt example5.c}, in the example directory. 
+%
+We show the implementation of the $\chi^2$ function in the example on page~\pageref{ex:syst},
+where we in addition include an uncorrelated energy calibration error.
+MAYBE: ADDITIONAL PLOT TO SHOW RESULT? [WW]
+
+Further functions useful in the context of user-defined systematics are:
+\begin{function}
+\GLBNS{SetSysErrorsList} 
+{\tt int glbSetSysErrorsList(int exp, int rule, int on\_off, const double *sys\_list)}
+changes the systematical errors defined in the AEDL file in experiment {\tt exp} and 
+rule {\tt rule} to the values in {\tt sys\_list}. This change can be perfomed for
+systematics on or systematics off by using {\tt GLB\_ON} or {\tt GLB\_OFF} for {\tt on\_off}. 
+\end{function}
+\begin{function}
+\GLBNS{SetSysErrorsListPtr} 
+{\tt double* glbSetSysErrorsListPtr(int exp, int rule, int on\_off)}
+returns a pointer to the list of systematical errors as defined in the AEDL file in experiment {\tt exp} and 
+rule {\tt rule}. Choose
+systematics on or systematics off by using {\tt GLB\_ON} or {\tt GLB\_OFF} for {\tt on\_off}. 
+\end{function}
+In addition, in some cases the systematics minimization may not converge very fast or not at all, because it end up in a local minimum.
+In this case, one can change the starting values of the minimizer, \ie, the
+starting point from which the local minimizer rolls into the local minimum.
+\begin{function}
+\GLBNS{SetSysStartingValuesList} 
+{\tt int glbSetSysStartingValuesList(int exp, int rule, int on\_off, const double *sys\_list)}
+sets the starting values of the local systematics minimizer in experiment {\tt exp} and 
+rule {\tt rule} to the values in {\tt sys\_list}. This change can be perfomed for
+systematics on or systematics off by using {\tt GLB\_ON} or {\tt GLB\_OFF} for {\tt on\_off}. 
+Usually, zero is used for all starting values. 
+\end{function}
+IS THAT CORRECT (THAT ZERO IS USUALLY USED)? [WW]
+\begin{function}
+\GLBNS{GetSysStartingValuesListPtr} 
+{\tt double *glbGetSysStartingValuesListPtr(int exp, int rule, int on\_off, const double *sys\_list)}
+returns the starting values of the local systematics minimizer in experiment {\tt exp} and 
+rule {\tt rule}. This list can be obtained for
+systematics on or systematics off by using {\tt GLB\_ON} or {\tt GLB\_OFF} for {\tt on\_off}.
+\end{function}
+A useful trick is often to use the minimum from the last run as starting values for the next one
+if the input parameters are only slighltly changed. One can obtain the minimum from the last run 
+from the last call of the $\chi^2$ function (\cf, {\tt example5.c}).
 

### 2007-01-04 Documentation/part1.tex
Prior chapter added

@@
   \hspace*{0.3cm} int ew\_low, ew\_high, i; \\
-  \hspace*{0.3cm} double emin, emax, fit\_rate;  double chi2 = 0.0; \\
-  \\
+  \hspace*{0.3cm} double emin, emax, fit\_rate;  double chi2 = 0.0;
+
    \hspace*{0.3cm} /* Request simulated energy interval and analysis energy window */ \\
@@
   \hspace*{0.6cm}  chi2 += likelihood(true\_rates\_F[i], fit\_rate, true\_rates\_F[i]); \\
-   \\
+
   \hspace*{0.6cm}  /* Statistical part of chi2 for near detector  */ \\
@@
 region in this plane with all the other parameter correlations included. 
-In practice, this projection is very difficult: a grid-based method would 
+In practice, this projection (or marginalization) is very difficult: a grid-based method would 
 need $(N_{\mathrm{grid}})^n$ function calls of {\tt glbChiSys} to 
@@
 minimization we use a derivative free method due to Powell in a 
-modified~\cite{Brent:1973} version\footnote{Not needing derivatives is
-highly desired since the event rate depends in a non-linear way on the
-oscillation parameters, thus there is no easy analytical way
+modified~\cite{Brent:1973} version\footnote{Not to need derivatives is
+highly desired, since the event rate depends in a non-linear way on the
+oscillation parameters. Thus, there is no easy analytical way
 to obtain derivatives of the $\chi^2$ function.}.
@@
 one may miss a local minimum which affects the results\footnote{NB -- 
-Implementing a grid-based method which guarantee to find all local minima is
+Implementing a grid-based method which guarantees to find all local minima is
 not straightforward either, to say the least.}. 
-The only way out of this dilemma is to use some heuristic approach, \ie\
-although one can not guarantee anything one can use schemes which work
+The only way out of this dilemma is to use some heuristic approach, \ie , 
+one can use schemes which work
 in most cases and announce their failure loudly. In order to use such a
@@
 that the used methods then also depend
-on the region of the parameter space. In this manual, we only use
+on the region of the parameter space. In this manual, we mainly use
 examples with a neutrino factory, since some of these complications
-can be illustrated there. Albeit the methods described here are neither
-complete nor will they work everywhere in the parameter space. It is
+can be illustrated there. Note that the methods described here are neither
+complete, nor will they work everywhere in the parameter space. It is
 in any case up to the user to make sure that the results are what he/she
@@
 if one does not find the other local minima appearing 
-below the chosen confidence level. Thus, with this approach and proper usage,
+below the chosen confidence level. With this approach and proper usage,
  it should not possible to produce a too pessimistic solution. 
@@
 \index{norm}{External information!input errors} 
-\index{norm}{External information!starting values}
+\index{norm}{External information!central values}
 In many cases, the fit manifold is restricted by the knowledge 
@@
 a test has demonstrated that the impact of the uncertainty on a 
-given fit parameter is negligible it can be assumed as fixed safely. 
+given fit parameter is negligible, it can be assumed as fixed safely. 
  The inclusion of external input in \GLOBES\ is done by the use of 
@@
 determined the measured parameter to be at the central value 
-(which we call {\em starting value}) with a $1 \sigma$ Gau\ss ian error 
+(which we call {\em central value}) with a $1 \sigma$ Gau\ss ian error 
 (which we call {\em input error}). The explicit definition of these 
@@
 minimization is simultaneously performed over all priors and free oscillation 
-parameters. In \equ{priors}, $\hat{\rho}^0$ is the {\em starting value} of the 
+parameters. In \equ{priors}, $\hat{\rho}^0$ is the {\em central value} of the 
 prior, and $\sigma_{\hat{\rho}}$ the $1 \sigma$ absolute (half width) 
@@
 $\sigma_{\hat{\rho}}$ at the central value $\hat{\rho}^0$. Usually, 
-the starting value is fixed at the best-fit value, and the input error 
+the central value is fixed at the best-fit value, and the input error 
 to the $1 \sigma$ half width of the external measurement. For the matter 
@@
 precisions, one can omit the respective input errors. In \GLOBES , a
-value of $0$ corresponds\footnote{to be precise, a value for the error
+value of $0$ corresponds\footnote{To be precise, a value for the error
 in between $-10^{-12}$ and $+10^{-12}$} to neglecting the prior. If, however,
  earlier external measurements provide better information, one can set their 
-absolute precisions with the input errors. The starting values are usually 
+absolute precisions with the input errors. The central values are usually 
 chosen to be the best-fit values of this external experiments, such as
@@
 necessary to adjust them, such as for artificial constraints to the
-oscillation parameters. In other cases, minor modifications of the 
-starting values can cause a faster convergence of the algorithm.
+oscillation parameters. 
 For example, for the investigation of the
@@
 in order to force the minimizer not to  fall into the (unwanted) 
-true-sign solution. In this case, the starting value of $\ldm$ would 
+true-sign solution. In this case, the central value of $\ldm$ would 
 be set to $\rho^0_{\ldm} = - \ldm$, and a $\sigma_{\ldm}$ of the order 
@@
 \index{norm}{External information!input errors} 
-\index{norm}{External information!starting values}
-In order to set the starting values and input errors, two function have to
+\index{norm}{External information!central values}
+In order to set the central values and input errors, two function have to
  be called {\em before the usage of any minimizer}:
@@
 \GLBNS{SetCentralValues}
-{\tt int glbSetCentralValues(const glb\_params in)} sets the starting values for all of the following minimizer calls to {\tt in}.
+{\tt int glbSetCentralValues(const glb\_params in)} sets the central values for all of the following minimizer calls to {\tt in}.
 \end{function}
@@
 \end{function}
-Accordingly, there are functions to return the actually set starting values
+Accordingly, there are functions to return the actually set central values
 and input errors:
@@
 {\tt int glbGetCentralValues(glb\_params out)} writes the currently
-set starting values to {\tt out}.
+set central values to {\tt out}.
 \end{function}
@@
 \end{quote}
-In this example, the starting values are set to 
+In this example, the central values are set to 
 the true (simulated) values. Remember that initially the 
@@
 all $1.0$, which means that they do not need to be adjusted for the
-starting values.
+central values.
 
@@
  kept fixed. This will be important together with arbitrary projections
- using \GLB{ChiNP}. A more subtile consequence is the comparison
+ using \GLB{ChiNP}. A more subtle consequence is the comparison
  of fit manifold sections and projections for the solutions where
@@
  (all other parameters fixed), the section region would in many cases be 
- larger than the projection region if the priors are not added to the
+ larger than the projection region if the priors were not added to the
  section. At the true solution, this problem usually does not occur
@@
  on $\sin 2 \theta_{12}$, $\sin 2 \theta_{12} \cdot \sdm$, or $\sin  
- \theta_{12}$. Later versions of  \GLOBES\ may include more alternatives.
+ \theta_{12}$. \GLOBES\ 3.0 and higher provide an alternative
+to that by the concept of user-defined priors.
  \end{enumerate}
@@
 {\tt {\scriptsize
-  /* Set starting values and input errors for all projections */  \\
+  /* Set central values and input errors for all projections */  \\
   glbDefineParams(input\_errors,theta12*0.1,0,0,0,sdm*0.1,0);\\  
@@
 \end{function}
-After setting the starting values, input errors, and the projection, 
+After setting the central values, input errors, and the projection, 
 we can run the minimizer:
@@
 is compared with the result including the full multi-parameter correlation.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{User-defined priors}
+\index{norm}{User-defined priors}
+
+\example{Non-Gaussian external solar input}{
+\label{ex:userprior}
+
+This examples demonstrates how to use a non-Gaussian error for the
+external solar mixing angle input instead of a Gaussian error.
+The user-defined prior function reads as follows and is very similar to the standard prior:
+\begin{quote}
+{\tt {\footnotesize
+ double my\_prior(const glb\_params in) \\
+\{ \\
+  \hspace*{0.5cm} glb\_params central\_values = glbAllocParams(); \\
+  \hspace*{0.5cm} glb\_params input\_errors = glbAllocParams(); \\
+  \hspace*{0.5cm}  glb\_projection p = glbAllocProjection(); \\
+  \hspace*{0.5cm}  glbGetCentralValues(central\_values); \\
+  \hspace*{0.5cm}  glbGetInputErrors(input\_errors); \\
+  \hspace*{0.5cm}  glbGetProjection(p); \\
+  \hspace*{0.5cm}  int i; \\
+  \hspace*{0.5cm}  double pv = 0.0; \\
+  \hspace*{0.5cm}  double fitvalue,centralvalue,inputerror;
+
+ \hspace*{0.5cm} /* Add oscillation parameter priors */ \\
+  \hspace*{0.5cm}  for(i=0;i<6;i++) \\
+  \hspace*{0.5cm} if(glbGetProjectionFlag(p,i)==GLB\_FREE) \\
+  \hspace*{0.5cm} \{ \\
+  \hspace*{1cm} fitvalue=glbGetOscParams(in,i); \\
+  \hspace*{1cm} centralvalue=glbGetOscParams(central\_values,i); \\
+  \hspace*{1cm} inputerror=glbGetOscParams(input\_errors,i); \\
+  \hspace*{1cm} if(inputerror>1e-12) 
+  \hspace*{1cm} \{ \\
+  \hspace*{1.5cm} if(i==GLB\_THETA\_12) /* Prior on $\sin^2 \theta_{12}$ */ \\ 
+\hspace*{2.0cm} pv+=square((square(sin(startvalue))-fitvalue)/inputerror);\\ 
+  \hspace*{1.5cm} else \\ 
+  \hspace*{2.0cm}  pv+=square((startvalue-fitvalue)/inputerror); \\
+  \hspace*{1cm} \} \\
+  \hspace*{0.5cm} \}
+
+\hspace*{0.5cm} /* Add matter parameter priors */ \\
+  \hspace*{0.5cm} for(i=0;i<glb\_num\_of\_exps;i++) \\
+  \hspace*{0.5cm} if(glbGetDensityProjectionFlag(p,i)==GLB\_FREE) \\
+  \hspace*{0.5cm} \{ \\
+  \hspace*{1cm}  fitvalue=glbGetDensityParams(in,i); \\
+  \hspace*{1cm}  centralvalue=1.0; \\
+  \hspace*{1cm}  inputerror=glbGetDensityParams(input\_errors,i); \\
+  \hspace*{1cm}  if(inputerror>1e-12) \\
+  \hspace*{1.5cm}   pv+=square((centralvalue-fitvalue)/inputerror); \\
+  \hspace*{0.5cm} \}
+
+  \hspace*{0.5cm} glbFreeParams(central\_values); \\
+  \hspace*{0.5cm} glbFreeParams(input\_errors); \\
+  \hspace*{0.5cm} glbFreeProjection(p); \\
+  \hspace*{0.5cm} return pv; \\
+\}
+}
+}
+
+Note that this prior interprets the central values and input errors in terms of  $\sin^2 \theta_{12}$
+instead of $\theta_{12}$.
+\end{quote}
+}
+
+THIS CHAPTER MAY NEED SOME MODIFICATIONS, SINCENOT YET FINALLY IMPLEMENTED! [WW]
+
+User-defined priors are an advanced concept of \GLOBES\ 3.0 and higher. Therefore,
+this section can be jumped over in a first reading of the manual.
+They allow for arbitrary priors being added to the $\chi^2$ after the systematics
+minimization and before the marginalization over the oscillation parameters (\cf, \equ{priors}).
+Therefore, compared to user-defined systematics, they depend on the oscillation
+parameters only, but not on the systematics parameters. Examples for applications
+are non-Gaussian external input, the combination with externally simulated experiments, and the
+constraint to certain parameter-subspaces (such as a specific octant).
+
+The following function replaces the standard priors by user defined ones and has to be used
+after {\tt glbInit}:
+\begin{function}
+\GLBNS{glbRegisterPriorFunction}
+{\tt 
+glbRegisterPriorFunction(double (*prior)(const glb\_params), int (*starting)(const glb\_params), int (*error)(const glb\_params))} registers a user-defined prior function {\tt prior}. 
+\end{function}
+DO WE REALLY NEED THE {\tt starting} AND {\tt error} VARIABLES? WHAT ARE THEY GOOD FOR? 
+DOES THAT MAKE SOMTHING FASTER? [WW]
+
+The prior function has to be any function using one parameter {\tt glb\_params} containing the fit values.
+We show an example for a user-defined prior on page~\pageref{ex:userprior}. This prior
+will simply be registered with
+\begin{quote}
+{\tt
+ glbInit(argv[0]); \\
+ glbRegisterPriorFunction(my\_prior,NULL,NULL);
+}
+\end{quote}
+and replaces the standard prior of \GLOBES . It behaves exactly as the standard prior, but
+the solar mixing angle is interpreted as $\sin^2 \theta_{12}$ instead of $\theta_{12}$. Therefore,
+a Gaussian error is imposed on $\sin^2 \theta_{12}$ instead of $\theta_{12}$, and the central
+values and input errors are interpreted in terms of $\sin^2 \theta_{12}$ as well. Note that
+one does not need to know if the prior was called for one experiment only or all experiments,
+since this information is implicitely given by the different density projection flags being set
+accordingly.
 

### 2007-01-05 Documentation/part1.tex
*** empty log message ***

@@
 Note that the new user-defined systematics has to be registered with {\tt glbDefineChiFunction} first. 
+One can also return the systematics function name by 
+\begin{function}
+\GLBNS{GetChiFunction}
+{\tt int glbGetChiFunction(int exp, int rule, int on\_off, char *sys\_id)}
+returns the name of the systematics $\chi^2$ function of a given experiment {\tt exp} and rule {\tt rule} for
+systematics on or off as given by {\tt on\_off}. The user must 
+ensure that the return string {\tt sys\_id} is large enough to hold the name of the function.
+\end{function}
+Similarly, a pointer to the systematics function can be returned either by name, or by experiment and rule selection:
+\begin{function}
+\GLBNS{GetChiFunctionPtr}
+{\tt glb\_chi\_function glbGetChiFunctionPtr(const char *name)}
+returns a pointer to the systematics $\chi^2$ function with a specified name {\tt name}.
+\end{function}
+\begin{function}
+\GLBNS{GetChiFunctionPtrInExperiment}
+{\tt glb\_chi\_function glbGetChiFunctionPtrInExperiment(int exp, int rule, int on\_off)}
+returns a pointer to the systematics $\chi^2$ function of experiment {\tt exp} and rule {\tt rule}.
+Systematics on or off can be accessed by {\tt on\_off}.
+\end{function}
 
 The $\chi^2$ function {\tt glb\_chi\_function} may use a number of helper functions for
-for the $\chi^2$ calculation. Some of these functions may not be directly related to user-defined
-systematics, but may be mainly used in this context:
-\begin{function}
-\GLBNS{GetEminEmax}
-{\tt int glbGetEminEmax(int experiment, double *emin, double *emax)} returns the
-energy range [{\tt emin},{\tt emax}] covered by the binning as defined in the AEDL file for the experiment
-{\tt experiment}.
-\end{function}
-\begin{function}
-\GLBNS{GetEnergyWindow}
-{\tt int glbGetEnergyWindow(int exp, int rule, double *low, double *high)} returns the
-energy window [{\tt low},{\tt high}] defined in the AEDL file for the experiment
-{\tt exp} and rule {\tt rule}.
-\end{function}
-\begin{function}
-\GLBNS{GetEnergyWindowBins}
-{\tt int glbGetEnergyWindowBins(int exp, int rule, int *lowbin, int *highbin)} returns the
-energy window in terms of bin numbers [{\tt lowbin},{\tt highbin}] corresponding to the energy window defined in the AEDL file for the experiment {\tt exp} and rule {\tt rule}.
-\end{function}
-\begin{function}
-\GLBNS{GetNumberOfBins}
-{\tt int glbGetNumberOfBins(int exp)} returns the number of bins for the experiment {\tt exp}.
-\end{function}
-\begin{function}
-\GLBNS{GetNumberOfSimBins}
-{\tt int glbGetNumberOfSimBins(int exp)} returns the number of sampling points for the experiment {\tt exp}.
-\end{function}
-SHOULD THIS NOT BE CALLED {\tt glbGetNumberOfSamplingPoints}??? [WW]
-
+for the $\chi^2$ calculation. 
 A very useful function to include energy calibration errors is
@@
 {\tt rates\_out}. The parameters {\tt emin} and {\tt emax} denote the minimal and maximal energy
-as obtained with {\tt glbGetEminEmax}.
+as obtained with \GLB{glbGetEminEmax}.
 \end{function}
@@
 In addition, in order to implement user-defined systematics, one usually needs low-level access to the event rate vectors using \GLB{GetRuleRatePtr}, \GLB{GetSignalFitRatePtr}, \GLB{GetBGFitRatePtr} and possibly other functions described in \Sec~\ref{sec:event_rates}.
+Additionally, the
+functions \GLB{GetEminEmax}, \GLB{GetEnergyWindow}, \GLB{GetEnergyWindowBins}, and \GLB{GetNumberOfBins} from \Sec~\ref{sec:aedl_names} are useful.
 We recommend that you familiarize yourself with these functions at this point.
@@
 \index{norm}{User-defined priors}
+\label{sec:userdefined}
 
@@
 Then, all of the found degeneracies below the threshold can be tested for the combination of all experiments.
+\item[Tracking algorithm.] If one scans a large portion of the parameter space with different input values, it is often useful to use the output from the previous minimization as educated guess for the next minimization.
+This works often very well if the degeneracy can be located in a part of the parameter space and the input parameter values change slowly enough (adiabatic transformation of the fit manifold).
+\item[Pre-scanning the parameter space.] In some cases, a very fast procedure can be a pre-scan 
+of the relevant parameters using the very fast {\tt glbChiSys}.
+For example, for the intrinsic degeneracy, the location of the degeneracy in the $(\deltacp,\theta_{13})$ plane can easily and quickly found with {\tt glbChiSys} while keeping the other parameters fixed.
+Use this location then as a starting point for {\tt glbChiAll}.
+\item[Manual scan of a subspace.] In some cases, the minimizer easily ends up in the wrong/unwanted solution, which is in most cases the already known best-fit solution.
+For example, for the octant degeneracy location, it is difficult to prevent the minimizer from running into the best-fit octant.
+In this case, one can scan one parameter (such as $\theta_{23}$) on a grid by choosing only valid values, and use {\tt glbChiNP} to marginalize over the other ones (while keeping \eg\ $\theta_{23}$ fixed).
+Therefore, the 6-parameter minimization is split into a 1-parameter manual minimization and a 5-parameter simultaneous minimization.
+Advantage: Runs with all \GLOBES\ versions.
+Disadvantage: Very slow.
+\item[Using Schwetz-priors (\GLOBES\ 3.0 and higher).] An elegant and fast method to prevent the minimizer from running into parts of the parameter space is to use user-defined priors and add a penalty as soon as a taboo zone is entered (\cf, \Sec~\ref{sec:userdefined}).
+This method was initially suggested by Thomas Schwetz.
+\item[Using total rates.] In order to systematically locate all degeneracies (the eight-fold degeneracy), relatively reliable methods use the total appearance neutrino and antineutrino event rates of the experiment to determine educated guesses.
+The method goes as follows: Plot the curves with equal total rate in the $\sin^2 2 \theta_{13}$-$\delta_{\mathrm{CP}}$-plane for both neutrinos and antineutrinos using the same rates as in the best-fit point.
+The curves will intersect at the intrinsic degeneracy (if it all).
+Plot the same curves for the same rates with $\mathrm{sgn}(\Delta m_{31}^2)$ flipped.
+Again you will find a maximum of two intersection points.
+Now do the same for $\pi/2 - \theta_{23}$ flipped, and for  $\mathrm{sgn}(\Delta m_{31}^2)$ combined with $\pi/2 - \theta_{23}$ flipped (mixed degeneracy).
+You will find at most two more intersection points in each case.
+The results should look somewhat like this, where the best-fit point is not marked (the marks correspond to the points where the discrete degeneracies are located according to this specific algorithm):
+\begin{center}
+\includegraphics[width=0.95\textwidth]{ircurves}
+\end{center}
+ Altogether, there is a maximum of eight intersection points in the $\sin^2 2 \theta_{13}$-$\delta_{\mathrm{CP}}$-plane, one of which is the best-fit point.
+These points can be using as starting points of the minimizer to locate the eight-fold degeneracy.
+Note that similar methods using the $\chi^2$ instead of the total rates have also been successfully used in the past.
 \end{description}
-Finally, note that any degenerate solution below the confidence level threshold which can not be located makes the result appear better than it actually is.
+Finally, note that any degenerate solution below the confidence level threshold which cannot be located makes the result appear better than it actually is.
 Thus, one should be careful with the determination of the degenerate solutions in order to find all of them.
@@
 \end{function}
+For a constant matter density profile, it is sufficient to specify the oscillation channel with
+\begin{function}
+\GLBNS{ConstantDensityProbability}
+{\tt double glbConstantDensityProbability(int l, int m, int panti, double E, double L, double rho)} returns the neutrino oscillation probability $\nu_l \rightarrow \nu_m$ for the energy {\tt E} in constant matter, where the matter density profile
+has the constant density {\tt rho} and the baseline is {\tt L}. The parameter
+{\tt panti} is $+1$ for neutrinos and $-1$ for antineutrinos.
+\end{function}
+If one in addition wants to use the low-pass filter feature in \GLOBES\ (see \Sec~\ref{sec:energy}), can can use
+\begin{function}
+\GLBNS{FilteredConstantDensityProbability}
+{\tt double glbFilteredConstantDensityProbability(int exp, int l, int m, int panti, double E)} returns the neutrino oscillation probability $\nu_l \rightarrow \nu_m$ for the energy {\tt E} in constant matter, where the matter density and baseline, as well as the filter properties are taken from experiment {\tt exp}.
+The parameter {\tt panti} is $+1$ for neutrinos and $-1$ for antineutrinos.
+\end{function}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{\AEDL\ names}
+\section{Information from \AEDL\ files}
+\label{sec:aedl_names}
+
+In some cases, it is necessary to obtain information from the \AEDL\
+files. Very basic information is supplied with
+\begin{function}
+\GLBNS{GetEminEmax}
+{\tt int glbGetEminEmax(int experiment, double *emin, double *emax)} returns the
+energy range [{\tt emin},{\tt emax}] covered by the binning as defined in the AEDL file for the experiment
+{\tt experiment}.
+\end{function}
+\begin{function}
+\GLBNS{GetEnergyWindow}
+{\tt int glbGetEnergyWindow(int exp, int rule, double *low, double *high)} returns the
+energy window [{\tt low},{\tt high}] defined in the AEDL file for the experiment
+{\tt exp} and rule {\tt rule}.
+\end{function}
+\begin{function}
+\GLBNS{GetEnergyWindowBins}
+{\tt int glbGetEnergyWindowBins(int exp, int rule, int *lowbin, int *highbin)} returns the
+energy window in terms of bin numbers [{\tt lowbin},{\tt highbin}] corresponding to the energy window defined in the AEDL file for the experiment {\tt exp} and rule {\tt rule}.
+\end{function}
+\begin{function}
+\GLBNS{GetNumberOfBins}
+{\tt int glbGetNumberOfBins(int exp)} returns the number of bins for the experiment {\tt exp}.
+\end{function}
+\begin{function}
+\GLBNS{GetNumberOfSimBins}
+{\tt int glbGetNumberOfSimBins(int exp)} returns the number of sampling points for the experiment {\tt exp}.
+\end{function}
+SHOULD THIS NOT BE CALLED {\tt glbGetNumberOfSamplingPoints}??? [WW]
+
+In order to obtain information on the structure of the rules, 
+a number of functions are provided:
+\begin{function}
+\GLBNS{GetNumberOfRules}
+{\tt int glbGetNumberOfRules(int exp)} returns the number of
+rules in experiment {\tt exp}.
+\end{function}
+\begin{function}
+\GLBNS{GetLengthOfRule}
+{\tt int glbGetLengthOfRule(int exp, int rule, int signal)} returns
+the length of rule {\tt rule} in experiment {\tt exp}. The parameter
+{\tt signal} can be either \GLBC{GLB\_SIG} for the number of signal
+components or \GLBC{GLB\_BG} for the number of background components.
+\end{function}
+\begin{function}
+\GLBNS{GetNormalizationInRule}
+{\tt double glbGetNormalizationInRule(int exp, int rule, int signal)}
+returns the normalization (normalization or background center)
+in rule {\tt rule} of the experiment {\tt exp}.
+The parameter {\tt signal} refers to signal (\GLBC{GLB\_SIG}) or background
+(\GLBC{GLB\_BG}).
+\end{function}
+\begin{function}
+\GLBNS{GetChannelInRule}
+{\tt int glbGetChannelInRule(int exp, int rule, int pos, int signal)}
+returns the channel number in rule {\tt rule} at the position {\tt pos}
+of the experiment {\tt exp}.
+The parameter {\tt signal} refers to signal (\GLBC{GLB\_SIG}) or background
+(\GLBC{GLB\_BG}).
+\end{function}
+\begin{function}
+\GLBNS{GetCoefficientInRule}
+{\tt double glbGetCoefficientInRule(int exp, int rule, int pos, int signal)}
+returns the coefficient of the component {\tt pos} in rule {\tt rule} 
+of the experiment {\tt exp}.
+The parameter {\tt signal} refers to signal (\GLBC{GLB\_SIG}) or background
+(\GLBC{GLB\_BG}).
+\end{function}
+%
+Similarly to the rules, one can find the number of channels of an experiment:
+\begin{function}
+\GLBNS{GetNumberOfChannels}
+{\tt int glbGetNumberOfChannels(int exp)} returns the number of 
+channels of experiment {\tt exp}.
+\end{function}
+
 \index{norm}{AEDL@\AEDL!names}
-\label{sec:aedl_names}
-
 Since in \AEDL\ rules, cross section, fluxes etc. carry a `name' by
@@
 \Secs~\ref{sec:channel} and~\Sec~\ref{sec:rules} of the \AEDL\ manual.
-
-\index{norm}{Event rates}
-\GLOBES\ currently supports rule-based and channel-based event rate functions, where the information is in written into a file
-or returned in a list. The rule-based functions are:
+Note that some of the event rate access has changed in \GLOBES\ 3.0,
+because user-defined systematics requires very fast access which was
+not possible with the old method.
+
+A very simple function is for the total rate
+\begin{function}
+\GLBNS{TotalRuleRate}
+{\tt double glbTotalRuleRate(int exp, int rule, int pos,
+int effi, int bgi, int coeffi, int signal)} returns the total rates.
+A specific experiment {\tt exp} and a 
+specific rule {\tt rule} have to be chosen, as well as the signal
+or background rate {\tt signal} (either \GLBC{GLB\_SIG} or \GLBC{GLB\_BG}).
+The position {\tt pos} refers to the component within the signal or 
+background, and can also be {\tt GLB\_ALL}. The function may return
+the rates with (\GLBC{GLB\_W\_COEFF}) or without (\GLBC{GLB\_WO\_COEFF})
+overall efficiency coefficient, as it is specified by {\tt coeffi}. 
+In addition, it may contain the post-smearing efficiencies (set
+{\tt effi} to \GLBC{GLB\_W\_EFF} or \GLBC{GLB\_WO\_EFF}), and the
+post-smearing backgrounds (set
+{\tt bgi} to \GLBC{GLB\_W\_BG} or \GLBC{GLB\_WO\_BG}). The pre-smearing
+efficiencies and backgrounds cannot be accessed at the rule level.
+\end{function}
+The function {\tt glbTotalRuleRate} is especially useful if
+one wants to draw bi-rate graphs with total event rates, or look
+for the $(\deltacp,\theta_{13})$-degeneracy by the intersection of 
+neutrino and antineutrino constant event rate curves.
+
+There are several functions to directly print or save the event rate
+information:
 \begin{function}
@@
 \begin{function}
-\GLBNS{TotalRuleRate}
-{\tt double glbTotalRuleRate(int exp, int rule, int pos,
-int effi, int bgi, int coeffi, int signal)} returns the total rates with
-the same parameters as {\tt glbShowRuleRates}.
-\end{function}
-The function {\tt glbTotalRuleRate} is especially useful if
-one wants to draw bi-rate graphs with total event rates, or look
-for the $(\deltacp,\theta_{13})$-degeneracy by the intersection of 
-neutrino and antineutrino constant event rate curves.
-%
-In order to obtain information on the structure of the rules, 
-a number of additional functions are provided:
-\begin{function}
-\GLBNS{GetNumberOfRules}
-{\tt int glbGetNumberOfRules(int exp)} returns the number of
-rules in experiment {\tt exp}.
-\end{function}
-\begin{function}
-\GLBNS{GetLengthOfRule}
-{\tt int glbGetLengthOfRule(int exp, int rule, int signal)} returns
-the length of rule {\tt rule} in experiment {\tt exp}. The parameter
-{\tt signal} can be either \GLBC{GLB\_SIG} for the number of signal
-components or \GLBC{GLB\_BG} for the number of background components.
-\end{function}
-\begin{function}
-\GLBNS{GetNormalizationInRule}
-{\tt double glbGetNormalizationInRule(int exp, int rule, int signal)}
-returns the normalization (normalization or background center)
-in rule {\tt rule} of the experiment {\tt exp}.
-The parameter {\tt signal} refers to signal (\GLBC{GLB\_SIG}) or background
-(\GLBC{GLB\_BG}).
-\end{function}
-\begin{function}
-\GLBNS{GetChannelInRule}
-{\tt int glbGetChannelInRule(int exp, int rule, int pos, int signal)}
-returns the channel number in rule {\tt rule} at the position {\tt pos}
-of the experiment {\tt exp}.
-The parameter {\tt signal} refers to signal (\GLBC{GLB\_SIG}) or background
-(\GLBC{GLB\_BG}).
-\end{function}
-\begin{function}
-\GLBNS{GetCoefficientInRule}
-{\tt double glbGetCoefficientInRule(int exp, int rule, int pos, int signal)}
-returns the coefficient of the component {\tt pos} in rule {\tt rule} 
-of the experiment {\tt exp}.
-The parameter {\tt signal} refers to signal (\GLBC{GLB\_SIG}) or background
-(\GLBC{GLB\_BG}).
-\end{function}
-
-In addition, \GLOBES\ has channel-based rate functions:
-\begin{function}
 \GLBNS{ShowChannelRates}
@@
 \end{function}
-\begin{function}
-\GLBNS{GetChannelRates}
-{\tt int glbGetChannelRates(double** data,
-size\_t* length, int exp, int channel, int smearing)}
-writes the binned raw channel rates to the list {\tt data}
-and the length of this list to {\tt length}. 
-A specific experiment {\tt exp} and a 
-specific channel {\tt channel} have to be chosen.
-The function may return
-the rates before ({\tt smearing} is \GLBC{GLB\_PRE}) or after ({\tt smearing} is \GLBC{GLB\_POST})
-the energy smearing, where no user-defined data
-(pre-/post-smearing efficiencies or backgrounds) are taken into account.
-The return value is $-1$ if unsuccessful.
-\end{function}
-\begin{function}
-\GLBNS{GetUserData}
-{\tt int glbGetUserData(double** data,
-size\_t* length, int exp, int channel, int smearing, int bgeff)}
-writes the binned user-defined backgrounds or efficiencies 
-to the list {\tt data} and the length of this list to {\tt length}. 
-A specific experiment {\tt exp} and a 
-specific channel {\tt channel} have to be chosen.
-The function may return
-the pre- ({\tt smearing} is \GLBC{GLB\_PRE}) or post- ({\tt smearing} is \GLBC{GLB\_POST}) smearing backgrounds ({\tt bgeff} is \GLBC{GLB\_BG}) 
-or efficiencies ({\tt bgeff} is \GLBC{GLB\_EFF}). 
-The return value is $-1$ if unsuccessful.
-\end{function}
-Since \GLOBES\ reserves the memory for the lists returned in these
-functions, which it allocates on an internal stack, one has to reset
-the stack at the end of the rates access with
-\begin{function}
-\GLBNS{ResetRateStack}
-{\tt void glbResetRateStack()} resets the rate stack used for the
-lists returned from {\tt glbGetChannelRates} or {\tt glbGetUserData}.
-\end{function}
-A code excerpt to show the channel rates may look like this:
+
+\index{norm}{Event rates}
+For rate vectors, \GLOBES\ currently supports rule-based and channel-based event rate functions, where typically pointers on the rate vectors are returned.
+The following pointer-based functions are currently supported:
+\begin{function}
+\GLBNS{GetChannelRatePtr}
+{\tt double *glbGetChannelRatePtr(int exp, int ch, int pre\_post)} returns a 
+pointer on the simulated rate vector of experiment {\tt exp} and channel {\tt ch}. Either
+pre-smearing ({\tt pre\_post} is {\tt GLB\_PRE}) or post-smearing ({\tt pre\_post} is {\tt GLB\_POST})
+rates can be accessed.
+\end{function}
+\begin{function}
+\GLBNS{GetRuleRatePtr}
+{\tt double *glbGetRuleRatePtr(int exp, int rule)} returns a 
+pointer on the simulated rate vector of experiment {\tt exp} and rule {\tt rule}.
+\end{function}
+\begin{function}
+\GLBNS{GetSignalRatePtr}
+{\tt double *glbGetSignalRatePtr(int exp, int rule)} returns a 
+pointer on the simulated signal rate vector of experiment {\tt exp} and rule {\tt rule}.
+\end{function}
+\begin{function}
+\GLBNS{GetBGRatePtr}
+{\tt double *glbGetBGRatePtr(int exp, int rule)} returns a 
+pointer on the simulated background rate vector of experiment {\tt exp} and rule {\tt rule}.
+\end{function}
+\begin{function}
+\GLBNS{GetChannelFitRatePtr}
+{\tt double *glbGetChannelFitRatePtr(int exp, int ch, int pre\_post)} returns a 
+pointer on the fit rate vector of experiment {\tt exp} and channel {\tt ch}. Either
+pre-smearing ({\tt pre\_post} is {\tt GLB\_PRE}) or post-smearing ({\tt pre\_post} is {\tt GLB\_POST})
+rates can be accessed.
+\end{function}
+\begin{function}
+\GLBNS{GetSignalFitRatePtr}
+{\tt double *glbGetSignalFitRatePtr(int exp, int rule)} returns a 
+pointer on the fit signal rate vector of experiment {\tt exp} and rule {\tt rule}.
+\end{function}
+\begin{function}
+\GLBNS{GetBGFitRatePtr}
+{\tt double *glbGetBGFitRatePtr(int exp, int rule)} returns a 
+pointer on the fit background rate vector of experiment {\tt exp} and rule {\tt rule}.
+\end{function}
+A simple example how to use these functions to print a rate vector is
 \begin{quote}
 {\tt
- double* rates; \\
-  size\_t length; \\
-  glbGetChannelRates(\&rates,\&length,0,0,GLB\_PRE); \\
  int i; \\
-  for(i=0;i<length;i++) printf("\%g $\backslash$n",rates[i]); \\
-  glbResetRateStack(); \\
+   int n\_bins = glbGetNumberOfBins(EXP\_FAR); \\
+  double *true\_rates\_N = glbGetRuleRatePtr(0, 0); \\
+  printf("Simulated rates, experiment 0, rule 0: $\backslash$n"); \\
+  for(i=0;i<n\_bins;i++) printf("\% g ",true\_rates\_N[i]); \\
+  printf("$\backslash$n");
 }
 \end{quote}
-Finally, one can find the number of channels of an experiment:
-\begin{function}
-\GLBNS{GetNumberOfChannels}
-{\tt int glbGetNumberOfChannels(int exp)} returns the number of 
-channels of experiment {\tt exp}.
-\end{function}
+
+% OLD FUNCTIONS:
+% \begin{function}
+% \GLBNS{GetChannelRates}
+% {\tt int glbGetChannelRates(double** data,
+% size\_t* length, int exp, int channel, int smearing)}
+% writes the binned raw channel rates to the list {\tt data}
+% and the length of this list to {\tt length}. 
+% A specific experiment {\tt exp} and a 
+% specific channel {\tt channel} have to be chosen.
+% The function may return
+% the rates before ({\tt smearing} is \GLBC{GLB\_PRE}) or after ({\tt smearing} is \GLBC{GLB\_POST})
+% the energy smearing, where no user-defined data
+% (pre-/post-smearing efficiencies or backgrounds) are taken into account.
+% The return value is $-1$ if unsuccessful.
+% \end{function}
+% \begin{function}
+% \GLBNS{GetUserData}
+% {\tt int glbGetUserData(double** data,
+% size\_t* length, int exp, int channel, int smearing, int bgeff)}
+% writes the binned user-defined backgrounds or efficiencies 
+% to the list {\tt data} and the length of this list to {\tt length}. 
+% A specific experiment {\tt exp} and a 
+% specific channel {\tt channel} have to be chosen.
+% The function may return
+% the pre- ({\tt smearing} is \GLBC{GLB\_PRE}) or post- ({\tt smearing} is \GLBC{GLB\_POST}) smearing backgrounds ({\tt bgeff} is \GLBC{GLB\_BG}) 
+% or efficiencies ({\tt bgeff} is \GLBC{GLB\_EFF}). 
+% The return value is $-1$ if unsuccessful.
+% \end{function}
+% Since \GLOBES\ reserves the memory for the lists returned in these
+% functions, which it allocates on an internal stack, one has to reset
+% the stack at the end of the rates access with
+% \begin{function}
+% \GLBNS{ResetRateStack}
+% {\tt void glbResetRateStack()} resets the rate stack used for the
+% lists returned from {\tt glbGetChannelRates} or {\tt glbGetUserData}.
+% \end{function}
+% A code excerpt to show the channel rates may look like this:
+% \begin{quote}
+% {\tt
+%  double* rates; \\
+%   size\_t length; \\
+%   glbGetChannelRates(\&rates,\&length,0,0,GLB\_PRE); \\
+%   int i; \\
+%   for(i=0;i<length;i++) printf("\%g $\backslash$n",rates[i]); \\
+%   glbResetRateStack(); \\
+% }
+% \end{quote}
 
@@
 cross section number {\tt ident} for the flavor $\nu_l$ and polarity {\tt anti} ($+1$: neutrinos, $-1$: antineutinos) at the energy {\tt E}.
+\end{function}
+%
+The number of fluxes can be obtained with
+\begin{function}
+\GLBNS{GetNumberOfFluxes}
+{\tt int glbGetNumberOfFluxes(int exp)} returns the number of 
+fluxes defined in experiment {\tt exp}.
 \end{function}
@@
 \begin{function}
-\GLBNS{GetProfileType}
-{\tt int glbGetProfileType(int exp)} returns the matter density profile
+\GLBNS{GetProfileTypeInExperiment}
+{\tt int glbGetProfileTypeInExperiment(int exp)} returns the matter density profile
 type of experiment {\tt exp}.
@@
 In \GLOBES , the systematics is defined rule-based, \ie, each rule
-has its own systematics. In addition, \AEDL\ requires that it has to be
+has its own systematics. 
+In addition, \GLOBES\ supports dual systematics, \ie, \AEDL\ requires that it has to be
 defined in each rule what ``Systematics on'' and ``Systematics off'' means.
-Therefore, it is usually very simple to switch the systematics on and off:
+In principle, these two sets correspond to two completely different systematics
+implementations, and it is up to the \AEDL\ authors to define what that means.
+From the point of view of the API, it is very simple to switch the systematics on and off, \ie,
+to switch between the two systematics modes:
 \begin{function}
@@
 \end{function}
+One can also return the systematics state with
+\begin{function}
+\GLBNS{GetSysOnOffState}
+{\tt int glbGetSysOnOffState(int exp, int rule)}
+returns the systematics state in experiment {\tt exp} and rule {\tt rule}. 
+\end{function}
 In the example on page~\pageref{ex:barcharts}, the application of
@@
 systematics, correlations, and degeneracies.
-
-\index{norm}{Error dimension}
-The error dimension (for the definition, see \Sec~\ref{sec:rules}) can also be accessed directly with 
-\begin{function}
-\GLBNS{SetErrorDim}
-{\tt int glbSetErrorDim(int exp, int rule, int on\_off, int value)}
-sets the error dimension for systematics on ({\tt on\_off} is \GLBC{GLB\_ON}) or off ({\tt on\_off} is \GLBC{GLB\_OFF}) of experiment {\tt exp} and rule {\tt rule} to the value {\tt value}.
+Note that also the systematics function itself may be changed during running time,
+as described in \Sec~\ref{sec:userchi}.
+
+% \index{norm}{Error dimension}
+% The error dimension (for the definition, see \Sec~\ref{sec:rules}) can also be accessed directly with 
+% \begin{function}
+% \GLBNS{SetErrorDim}
+% {\tt int glbSetErrorDim(int exp, int rule, int on\_off, int value)}
+% sets the error dimension for systematics on ({\tt on\_off} is \GLBC{GLB\_ON}) or off ({\tt on\_off} is \GLBC{GLB\_OFF}) of experiment {\tt exp} and rule {\tt rule} to the value {\tt value}.
 The function returns
-$-1$ if not successful.
-\end{function}
-\begin{function}
-\GLBNS{GetErrorDim}
-{\tt int glbGetErrorDim(int exp, int rule, int on\_off)}
-returns the error dimension for systematics on ({\tt on\_off} is \GLBC{GLB\_ON}) or off ({\tt on\_off} is \GLBC{GLB\_OFF}) of experiment {\tt exp} and rule {\tt rule}.
-\end{function}
+% $-1$ if not successful.
+% \end{function}
+% \begin{function}
+% \GLBNS{GetErrorDim}
+% {\tt int glbGetErrorDim(int exp, int rule, int on\_off)}
+% returns the error dimension for systematics on ({\tt on\_off} is \GLBC{GLB\_ON}) or off ({\tt on\_off} is \GLBC{GLB\_OFF}) of experiment {\tt exp} and rule {\tt rule}.
+% \end{function}
 
@@
 Except from the general treatment of systematics, one can read out
-and write the signal and background errors, as well as the background
-centers. For the definitions of these quantities, see \Sec~\ref{sec:rules}.
+and write the signal and background errors for standard pre-defined systematics.
+For user-defined systematics, see \Sec~\ref{sec:userchi}, and 
+for the definitions of these quantities, see \Sec~\ref{sec:rules}.
 \begin{function}
@@
 \end{function}
-\begin{function}
-\GLBNS{SetBGCenters}
-{\tt int glbSetBGCenters(int exp, int rule, double norm, double tilt)}
-sets the background centers of experiment {\tt exp} and rule {\tt rule}
-to {\tt norm} (normalization center) and {\tt tilt} (tilt/calibration center).
-\end{function}
-\begin{function}
-\GLBNS{GetBGCenters}
-{\tt int glbGetBGCenters(int exp, int rule, double* norm, double* tilt)}
-writes the background centers of experiment {\tt exp} and rule {\tt rule}
-to {\tt norm} (normalization center) and {\tt tilt} (tilt/calibration center).
-\end{function}
+% ELIMINATED FROM DOC, BECAUSE WE WANT TO AVOID THE BG CENTERS IN THE FUTURE (CONFUSION) [WW]
+%
+% \begin{function}
+% \GLBNS{SetBGCenters}
+% {\tt int glbSetBGCenters(int exp, int rule, double norm, double tilt)}
+% sets the background centers of experiment {\tt exp} and rule {\tt rule}
+% to {\tt norm} (normalization center) and {\tt tilt} (tilt/calibration center).
+% \end{function}
+% \begin{function}
+% \GLBNS{GetBGCenters}
+% {\tt int glbGetBGCenters(int exp, int rule, double* norm, double* tilt)}
+% writes the background centers of experiment {\tt exp} and rule {\tt rule}
+% to {\tt norm} (normalization center) and {\tt tilt} (tilt/calibration center).
+% \end{function}
 As usual, all these functions return $-1$ if they were not successful.
@@
 \index{norm}{AEDL@\AEDL!external parameters}
+
+TO BE EXTENDED BY LIST HANDLING (TRANSFER LISTS IN AEDL VARIABLES). STATUS? [WW].
 
@@
 In particular, there are a number of functions:
-\begin{function}
-\GLBNS{SetFilterState}
-{\tt int glbSetFilterState(int on\_off)} sets the currently
-used filter state to on (\GLBC{GLB\_ON}) or off (\GLBC{GLB\_OFF}).
-\end{function}
-\begin{function}
-\GLBNS{GetFilterState}
-{\tt int glbGetFilterState()} returns the currently used filter state.
-\end{function}
+% \begin{function}
+% \GLBNS{SetFilterState}
+% {\tt int glbSetFilterState(int on\_off)} sets the currently
+% used filter state to on (\GLBC{GLB\_ON}) or off (\GLBC{GLB\_OFF}).
+% \end{function}
+% \begin{function}
+% \GLBNS{GetFilterState}
+% {\tt int glbGetFilterState()} returns the currently used filter state.
+% \end{function}
 \begin{function}
@@
 Analogously, the filter value can be accessed:
-\begin{function}
-\GLBNS{SetFilter}
-{\tt int glbSetFilter(double filter)} sets the currently used filter to the
-value {\tt filter}.
-\end{function}
-\begin{function}
-\GLBNS{GetFilter}
-{\tt double glbGetFilter()} returns the currently used filter value.
-\end{function}
+% \begin{function}
+% \GLBNS{SetFilter}
+% {\tt int glbSetFilter(double filter)} sets the currently used filter to the
+% value {\tt filter}.
+% \end{function}
+% \begin{function}
+% \GLBNS{GetFilter}
+% {\tt double glbGetFilter()} returns the currently used filter value.
+% \end{function}
 \begin{function}

### 2007-01-08 Documentation/part1.tex
*** empty log message ***

@@
 
-TOUR ITSELF (+EXAMPLE FILE) TO BE UPDATED. JK: PLEASE LET ME KNOW WHEN {\tt example-tour.c} IS FIXED. MAYBE WE SHOULD  ALSO UPDATE THE PARAMTER VALUES [WW].
+TOUR ITSELF (+EXAMPLE FILE) TO BE UPDATED (VALUES MAY HAVE CHANGED DUE TO AEDL FILE DEFINITION
+CHANGES ETC. MAYBE ONLY COMMENT ABOUT THAT [WW].
 
@@
 \gq{
-glbInitExperiment("NuFact.glb",\&glb\_experiment\_list[0], \\ \hspace*{4cm} \&glb\_num\_of\_exps);
+glbInitExperiment("NFstandard.glb",\&glb\_experiment\_list[0], \\ \hspace*{4cm} \&glb\_num\_of\_exps);
 } 
@@
 \\
- glbInitExperiment("NuFact.glb",\&glb\_experiment\_list[0], \\
+ glbInitExperiment("NFstandard.glb",\&glb\_experiment\_list[0], \\
  \hspace*{0.5cm} \&glb\_num\_of\_exps);\\
-  glbInitExperiment("NuFact.glb",\&glb\_experiment\_list[0], \\
+  glbInitExperiment("NFstandard.glb",\&glb\_experiment\_list[0], \\
  \hspace*{0.5cm} \&glb\_num\_of\_exps);
@@
 {\tt rates\_out}. The parameters {\tt emin} and {\tt emax} denote the minimal and maximal energy
-as obtained with \GLB{glbGetEminEmax}.
-\end{function}
-DOES THAT ALSO WORK WITH VARIABLE BINS? WHAT IS THE MEAING OF EMIN AND EMAX? [WW]
+as obtained with \GLB{glbGetEminEmax}. Requires constant energy bin widths!
+\end{function}
 For details on energy calibration errors, see \Sec~\ref{sec:rules}.
@@
 \begin{function}
-\GLBNS{SetSysErrorsListPtr} 
-{\tt double* glbSetSysErrorsListPtr(int exp, int rule, int on\_off)}
+\GLBNS{GetSysErrorsListPtr} 
+{\tt double* glbGetSysErrorsListPtr(int exp, int rule, int on\_off)}
 returns a pointer to the list of systematical errors as defined in the AEDL file in experiment {\tt exp} and 
@@
 \end{function}
-IS THAT CORRECT (THAT ZERO IS USUALLY USED)? [WW]
 \begin{function}
@@
 In such cases, the starting point of the minimizer has to be adjusted to help it find the true minimum.
-
+%
 Other examples for projections onto a parameter axis while keeping exactly one parameter fixed are \GLB{ChiTheta23}, \GLB{ChiDm31}, and \GLB{ChiDm21}, which
 can be found in \Tab~\ref{tab:stdfunctions} on page~\pageref{tab:stdfunctions}.
+
+Since the number of different parameter vectors used by \GLOBES\ may be a bit confusing, we summarize the most common parameter sets used in the calculations:
+\begin{description}
+\item[The simulated/true values] are the values set in \GLB{SetRates}.
+\item[The fit values] are the values in the first parameter {\tt in} of {\tt glbChiTheta13} (and the other minimization functions) which are {\em fixed} by definition (such as $\theta_{13}$ for {\tt glbChiTheta13}).
+\item[The starting point/educated guess for the minimizer] are the values in the first parameter {\tt in} of {\tt glbChiTheta13} (and the other minimization functions) which are {\em free} by definition (such as all but $\theta_{13}$ for {\tt glbChiTheta13}).
+\item[The minimization result] of the marginalization process are the values in the second parameter {\tt out} of {\tt glbChiTheta13} (and the other minimization functions) which are {\em free} by definition (such as all but $\theta_{13}$ for {\tt glbChiTheta13}).
+The other values in {\tt out} still correspond to the fit values.
+\item[The position of external priors,] \ie, the best-fit values of external input, is set by \GLB{SetCenralValues}.
+\item[The magnitude of external errors,] \ie, the errors themselves of external input, is set by \GLB{SetInputErrors}.
+\end{description}
 
@@
 \begin{function}
-\GLBNS{GetNumberOfSimBins}
-{\tt int glbGetNumberOfSimBins(int exp)} returns the number of sampling points for the experiment {\tt exp}.
-\end{function}
-SHOULD THIS NOT BE CALLED {\tt glbGetNumberOfSamplingPoints}??? [WW]
+\GLBNS{GetNumberOfSamplingPoints}
+{\tt int glbGetNumberOfSamplingPoints(int exp)} returns the number of sampling points for the experiment {\tt exp}.
+\end{function}
 

### 2007-01-08 Documentation/part1.tex
*** empty log message ***

@@
 {\tt rates\_out}. The parameters {\tt emin} and {\tt emax} denote the minimal and maximal energy
-as obtained with \GLB{glbGetEminEmax}. Requires constant energy bin widths!
+as obtained with \GLB{GetEminEmax}. Requires constant energy bin widths!
 \end{function}
@@
 \begin{function}
-\GLBNS{glbRegisterPriorFunction}
+\GLBNS{RegisterPriorFunction}
 {\tt 
@@
 In this chapter, we discuss how to simulate non-standard physics with \GLOBES , \ie, physics
-beyond three-flavor neutrino oscillations. For the first time, this feature was used in \Ref~\cite{Blennow:2005yk},
-and it has become a standard feature of \GLOBES\ starting from version 3.0 (CHECK). Since the 
-computation of oscilltion probabilities, and therefore the flavor transition probabilities, is the core basic
-element of \GLOBES , the introduction of non-standard physics requires that one produce a new version branch
-of the software, \ie, the software has to be re-compiled after the modification of the probability engine.
-Below, we will demonstrate how and where to do the necessary mofifications, and how to simulate non-standard
-physics in the application software using the ``new'' \GLOBES .
+beyond three-flavor neutrino oscillations. For the first time, this feature was used as experimental 
+feature in \Ref~\cite{Blennow:2005yk},
+and it has become a standard feature of \GLOBES\ starting from version 3.0. Since the 
+computation of oscillation probabilities, and therefore the flavor transition probabilities, is the core basic
+element of \GLOBES , the introduction of non-standard physics requires fimiliarity with the
+probability calculation in \GLOBES . It is therefore an advanced topic.
+Below, we will demonstrate how and where to do the necessary modifications, and how to simulate non-standard
+physics in the application software.
 
@@
 
-In order to modify \GLOBES , first install a current version of GLOBES (version larger than X.X.X CHECK) as
-a new branch, \ie, in a new directory, and rename the directory \GLOBES\ expanded into. 
-Using multiple versions of \GLOBES , it is, in general, advisable
-to specify the installation directory of all used versions manually (see ``Installation without
-root privilege'' in {\tt INSTALL} file). For instance, if the current ``standard'' version of \GLOBES\ is
-installed in {\tt GLOBES/Versions}, you may use {\tt GLOBES/NSVersion} for the non-standard version
-specified in {\tt configure}. Do not forget to rename the directory \GLOBES\ tar-ball expanded into: For instance,
-change the name from {\tt globes-3.0.0} to {\tt globes-3.0.0nsi}. 
-Then you can modify the source code in the sub-directories of {\tt globes-3.0.0nsi}, such as {\tt globes-3.0.0nsi/source}.
-In particular, we will be interested in modifying {\tt glb\_probability.cc}, which contains the
-Hamiltonian in the routine {\tt ComputeMixingMatrix}.
-%
-In general, the steps to introduce non-standard physics as follows:
-\begin{enumerate}
-\item
- Change the number of oscillation parameters in {\tt GLB\_OSCP} from six (standard) to the
- new number $6+N$, where $N$ is the number of non-standard parameters. 
- This change has to be done in the header file {\tt globes.h} in the sub-directory {\tt globes}
- where the tar-ball was extracted into,  {\em as well as} in the {\tt include/globes} sub-directory
- where the non-standard version was installed into (such as {\tt GLOBES/NSVersion/include/globes}).
-\item 
- Modify {\tt glb\_probability.cc} to include the non-standard effects. The parameters
- can be accessed by {\tt nsparams[0]}, {\tt nsparams[1]}, ..., {\tt nsparams[N-1]}.
-\item
- Re-compile the software using \\
- {\tt make} \\
- {\tt make install}
-\end{enumerate}
-Now your version of \GLOBES\ uses $6+N$ transition parameters, the six standard oscillation parameters,
-and $N$ additional ones.
-
-\section{Using a non-standard version of \GLOBES }
+\GLOBES\ 3.0 and higher does not require a re-compilation of the software to simulate non-standard physics.
+However, the probability engine has to be changed, which one would usually do by copying the respective parts from {\tt glb\_probability.c} (in the {\tt source} sub-directory).
+Therefore, for sophisticated applications, familiarity with this file is required.
+A number of comments are provided in this file to keep it self-explanatory.
+The key function types are
+\begin{function}
+\GLBNS{\_set\_oscillation\_parameters\_function}
+{\tt int (*glb\_set\_oscillation\_parameters\_function)(glb\_params p)} reads the fundamental oscillation parameters
+obtained from {\tt p} and stores them into global variables accessible to {\tt glb\_probability\_matrix\_function}.
+In addition, it can be used to pre-compute quantities which are dependent on the oscillation paramaters only for efficiency.
+\end{function}
+\begin{function}
+\GLBNS{\_get\_oscillation\_parameters\_function}
+{\tt int (*glb\_get\_oscillation\_parameters\_function)(glb\_params p)} returns the fundamental oscillation parameters
+in {\tt p} which are obtained from the global variables accessible to {\tt glb\_probability\_matrix\_function}.
+\end{function}
+\begin{function}
+\GLBNS{\_probability\_matrix\_function}
+{\tt int (*glb\_probability\_matrix\_function)(double P[3][3], int cp\_sign, double E,
+                  int psteps, const double *lengths, const double *densities,
+                  double filter\_sigma)} calculates the neutrino oscillation probability matrix to be returned in {\tt P}.
+The {\tt cp\_sign} is +1 for neutrinos and -1 for antineutrinos.
+In addition, the matter density profile to be used is characterized by the number of steps {\tt psteps}, the lengths of the matter density layers in the list {\tt lengths}, and the corresponding densities in the list {\tt densities}.
+The parameter {\tt filter\_sigma} defines the width of the low-pass filter to be used, or no filter usage if negative.
+\end{function}
+These three function types correspond to the standard implementation functions \GLB{\_set\_oscillation\_parameters}, \GLB{\_get\_oscillation\_parameters}, and \GLB{\_probability\_matrix} in {\tt glb\_probability.c} of the source code of \GLOBES , where one can find the standard behavior and use it by cut and paste (including the variables and code used by these functions).
+
+In order to use non-standard physics in \GLOBES , one needs re-define the three above functions and register them after {\tt glbInit} with
+\begin{function}
+\GLBNS{RegisterProbabilityEngine}
+{\tt int glbRegisterProbabilityEngine(int n\_parameters,
+                 glb\_probability\_matrix\_function prob\_func, \\
+                 glb\_set\_oscillation\_parameters\_function set\_params\_func, \\
+                 glb\_get\_oscillation\_parameters\_function get\_params\_func)} registers a probability
+engine for the simulation of non-standard physics with {\tt n\_parameters} oscillation parameters ({\tt n\_parameters} $\ge6$).
+The three functions {\tt prob\_func}, {\tt set\_params\_func}, and {get\_params\_func}
+will be used with the function types defined above.         
+\end{function}
+The number of oscillation parameters can, at any time, be obtained with
+\begin{function}
+\GLBNS{GetNumOfOscParams}
+{\tt  int glbGetNumOfOscParams()} returns the number of oscillation parameters.
+\end{function}
+
+HIER BIN ICH GERADE [WW].
+
+Example part here for non-standard implementation? [WW]
+
+\section{Using non-standard physics in the application software}
 

### 2007-01-09 Documentation/Paper2/paper2.tex
JK - Added paragraph about numerical efficiency in the paper

@@
 
+Sophisticated \GLOBES\ simulations with several experiments and complicated
+parameter correlations can take several days to finish even on modern computers.
+To mitigate this problem as far as possible, we have taken several steps to
+optimize the performance of the multi-dimensional minimizations, which are at
+the heart of each \GLOBES\ $\chi^2$ analysis.
+
+Firstly, the \GLOBES\ minimizers distinguish between minimization over
+oscillation parameters and minimization over systematics parameters.
+This is necessary because each step in the oscillation parameter space
+requires a time-consuming re-computation of oscillation probabilities,
+while a modification of the systematics parameters does not. The standard
+procedure is to perform a full minimization in systematics space after each
+step in oscillation space. This avoids interference of the two categories
+of parameters, and is known to have excellent convergence properties.
+The numerical method that does the minimization is the Powell
+algorithm~\cite{Press:NumRecip}.
+
+In \GLOBES\ 3.0, we provide an alternative method, which does the full minimization
+in one go and is therefore several times faster in most cases. In this algorithm,
+each Powell iteration over oscillation parameters is followed by one iteration
+over systematics parameters. Since there may be some pathological cases where this
+interleaved minimization strategy behaves differently than the old method,
+it is by default not used in \GLOBES\ 3.0. This ensures maximal compatibility
+with old application programs, but it requires users who wish to benefit from the
+new algorithm to explicitly select it by an API function call.
+
+Besides optimizing the minimization functions, it is also mandatory to
+optimize the routines which are called by them to calculate $\chi^2$
+for one specific set of parameters. In particular, the computation of
+oscillation probabilities, which has to be done every time the oscillation
+parameters have changed, should be as efficient as possible. The most expensive
+step here is the diagonalization of the hermitian $3\times3$ neutrino Hamilton
+operator in matter. The LAPACK rotuines~\cite{Anderson:LAPACK}, which were
+used to solve this eigenproblem in previous versions of \GLOBES, turned out
+not to be the optimal choice because they are mainly optimized for very large
+matrices~\cite{Kopp:2006wp}. In the case of the small and usually well-conditioned
+matrices appearing in \GLOBES, they spend too much time evaluating their parameter
+list and planning the optimal diagonalization strategy. Therefore, \GLOBES\ now
+relies an algorithm which has been designed specifically for the diagonalization
+of hermitian $3\times3$ matrices. It is essentially the extremely fast analytical
+algorithm discussed in~\cite{Kopp:2006wp}, but for the rare cases where this may become
+inaccurate, it falls back to the QL method from~\cite{Press:NumRecip}.
+
 \section{Summary and conclusions}

### 2007-01-09 Documentation/part1.tex
*** empty log message ***

@@
 TOUR ITSELF (+EXAMPLE FILE) TO BE UPDATED (VALUES MAY HAVE CHANGED DUE TO AEDL FILE DEFINITION
-CHANGES ETC. MAYBE ONLY COMMENT ABOUT THAT [WW].
+CHANGES ETC. ALSO EXPERIMENT (OLD) NUFACT DOES NOT EXIST ANYMORE. AND: SOME ERROR IN THIS OLD
+NUFACT-FILE [WW].
+
+CHECK: GLBSETPROJECTION REALLY USED WITH GLB\_FIXED FOR DENSITY PARAMS? IS NFSTANDARD USED?
 
@@
 }
-Assign values to our standard oscillation parameters:
+Assign values to our standard oscillation parameters and the standard matter density scaling factors:
 \gq{
- glbDefineParams(true\_values,theta12,theta13,theta23,deltacp,sdm,ldm);
+ glbDefineParams(true\_values,theta12,theta13,theta23,deltacp,sdm,ldm); \\
+ glbSetDensityParams(true\_values,1.0,GLB\_ALL);
 }
@@
 For this, we need to define our own (user-defined) projection, where
-only $\deltacp$ is a free parameter:
+only $\deltacp$ (and the matter density) is a free parameter:
 \gq{
@@
  \hspace*{0.5cm} GLB\_FREE, GLB\_FIXED, GLB\_FIXED); \\
+  glbSetDensityProjectionFlag(myprojection,GLB\_FREE,GLB\_ALL); \\
  glbSetProjection(myprojection); \\
@@
   glbSetDensityParams(input\_errors,0.05,GLB\_ALL); \\
+  glbSetDensityParams(starting\_errors,1.0,GLB\_ALL); \\
   glbSetCentralValues(starting\_values);\\
@@
   glbDefineParams(true\_values,theta12,theta13,theta23,deltacp,sdm,ldm); \\
+  glbSetDensityParams(true\_values,1.0,GLB\_ALL); \\
   glbSetOscillationParameters(true\_values); \\
@@
   glbSetDensityParams(input\_errors,0.05,GLB\_ALL);\\
+  glbSetDensityParams(starting\_values,1.0,GLB\_ALL); \\
   glbSetCentralValues(starting\_values);\\
@@
      \hspace*{1.5cm}asin(sqrt(0.8))/2,asin(sqrt(0.001))/2,M\_PI/4,M\_PI/2,7e-5,2e-3); \\
+    \hspace*{0.5cm}
+     glbSetDensityParams(true\_values,1.0,GLB\_ALL); // Matter scaling \\
 
@@
 which is stored in the density information of {\tt glb\_params}.
-Each of
-these scaling factors has $1.0$ as pre-defined value. Since it
-is in most cases not necessary to change this value, the user does
-not need to take care of it.
+% CHANGED IN GLOBES 3.0: 
+% Each of
+% these scaling factors has $1.0$ as pre-defined value. Since it
+% is in most cases not necessary to change this value, the user does
+% not need to take care of it. 
 For a constant matter density, it is simply the ratio of the matter density and the average matter density specified in the experiment definition, \ie , $\hat{\rho_i} \equiv \rho_i/\bar{\rho}_i$.
@@
 {\tt glb\_params glbAllocParams()} allocates the memory space
-needed for a parameter vector and returns a pointer to it.
+needed for a parameter vector and returns a pointer to it. All parameter values are
+initially set to {\tt nan} (not a number).
 \end{function}
@@
 \end{function}
-\begin{function}
-\GLBNS{TestLibraryVersion}
-{\tt int glbTestLibraryVersion(const char *version)} returns $0$ if
-the {\tt version} string of the format ``X.Y.Z'' is exactly the
-used \GLOBES\ version, $1$ if it is older, and $-1$ if it is newer.
-Note that the library and \GLOBES\ versions are not the same. 
-\end{function}
-WHAT IS THE DIFFERENCE BETWEEN THE LIBRARY AND GLOBES VERSION???? [WW]
+% UNDOCUMENTED FUNCTION, SINCE VERY DIFICULT TO EXPLAIN DIFFERENCE:
+% \begin{function}
+% \GLBNS{TestLibraryVersion}
+% {\tt int glbTestLibraryVersion(const char *version)} returns $0$ if
+% the {\tt version} string of the format ``X.Y.Z'' is exactly the
+% used \GLOBES\ library version, $1$ if it is older, and $-1$ if it is newer.
+% \end{function}
+% Note that the library and \GLOBES\ versions are not the same.
+
 \begin{function}
@@
   glbDefineParams(true\_values,theta12,theta13,theta23,deltacp,sdm,ldm); \\
-  glbDefineParams(test\_values,theta12,theta13,theta23,deltacp,sdm,ldm);  \\
+  glbSetDensityParams(true\_values,1.0,GLB\_ALL); \\
+  glbCopyParams(true\_values,test\_values);  \\
   glbSetOscillationParameters(true\_values); glbSetRates(); \\
@@
   /* Iteration over all values to be computed */ \\
-  double x,y,res; \\   
+%  double x,y,res; \\   
   for(x=-4.0;x<-2.0+0.01;x=x+2.0/50) \\
@@
 \GLBNS{DefineChiFunction}
-{\tt int glbDefineChiFunction(glb\_chi\_function chi\_func, int dim, const char *name)}
+{\tt int glbDefineChiFunction(glb\_chi\_function chi\_func, int dim, const char *name, void* user\_data)}
  tells \GLOBES\ to register the user-defined systematics $\chi^2$ function {\tt chi\_func} 
-identified by the string {\tt name} with {\tt dim} systematics parameters. Returns zero, if successful.
+identified by the string {\tt name} with {\tt dim} systematics parameters. The parameter {\tt user\_data}
+is an arbitrary pointer being transferred to {\tt chi\_func}. It can, for instance, be used to avoid
+global variables. Returns zero, if successful.
 \end{function}
@@
 \GLBNS{\_chi\_function}
-{\tt double (*glb\_chi\_function)(int exp, int rule, int dim, double *params, double *errors)}.
+{\tt double (*glb\_chi\_function)(int exp, int rule, int dim, double *params, double *errors, void* user\_data)}.
 In this function type, {\tt exp} is the experiment number, {\tt rule} is the rule number, {\tt dim} is the number of systematics parameters, {\tt params} is an array of the systematics auxiliary parameters themselves, and {\tt errors} is an array with the systematical errors.
-Note the the central values for {\tt params} are $0$, which means that $0$ corresponds to the un-modified rates.
-The array indices are running from $0$ to {\tt dim}-1.
+The parameter {\tt user\_data} is set as defined with
+{\tt glbDefineChiFunction}.
+Note the the central values for {\tt params} are $0$, which means that $0$ corresponds to the un-modified rates. The array indices are running from $0$ to {\tt dim}-1.
 \end{function}
@@
 We show the implementation of the $\chi^2$ function in the example on page~\pageref{ex:syst},
-where we in addition include an uncorrelated energy calibration error.
-MAYBE: ADDITIONAL PLOT TO SHOW RESULT? [WW]
+where we in addition include an uncorrelated energy calibration error. See \figu{reactor} for the result
+of this example.
+%
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=10cm]{reactor}
+\end{center}
+\caption{\label{fig:reactor} The result from {\tt example5.c} (figure similar in \Ref~\cite{Huber:2006vr}). The fat curve corresponds to the systematics on page~\pageref{ex:syst}.}
+\end{figure}
 
@@
   glbSetDensityParams(input\_errors,0.05,GLB\_ALL);\\
-  glbSetCentralValues(true\_values);\\
+  glbSetCentralValues(true\_values);
   glbSetInputErrors(input\_errors);\\
@@
   glbDefineProjection(th13\_projection,GLB\_FIXED,GLB\_FIXED,GLB\_FIXED,GLB\_FREE,GLB\_FIXED,GLB\_FIXED); \\
+  glbSetDensityProjectionFlag(th13\_projection,GLB\_FREE,GLB\_ALL); \\
   glbSetProjection(th13\_projection); \\ 
@@
 As an example, the projection sequence for a minimization over
-$\deltacp$ only looks like this:
+$\deltacp$ only (and the matter density parameters) looks like this:
 \begin{quote}
@@
   \hspace*{2cm} GLB\_FREE,GLB\_FIXED,GLB\_FIXED); \\
+  glbSetDensityProjectionFlag(t13\_projection,GLB\_FREE,GLB\_ALL); \\
   glbSetProjection(th13\_projection); \\ 
@@
 {\tt {\footnotesize
- double my\_prior(const glb\_params in) \\
+ double my\_prior(const glb\_params in, void* user\_data) \\
 \{ \\
@@
 
-THIS CHAPTER MAY NEED SOME MODIFICATIONS, SINCENOT YET FINALLY IMPLEMENTED! [WW]
-
 User-defined priors are an advanced concept of \GLOBES\ 3.0 and higher. Therefore,
@@
 {\tt 
-glbRegisterPriorFunction(double (*prior)(const glb\_params), int (*starting)(const glb\_params), int (*error)(const glb\_params))} registers a user-defined prior function {\tt prior}. 
-\end{function}
-DO WE REALLY NEED THE {\tt starting} AND {\tt error} VARIABLES? WHAT ARE THEY GOOD FOR? 
-DOES THAT MAKE SOMTHING FASTER? [WW]
+glbRegisterPriorFunction(double (*prior)(const glb\_params, void *user\_data), int (*central)(const glb\_params, void *user\_data), int (*error)(const glb\_params, void *user\_data),  void *user\_data)} registers a user-defined prior function {\tt prior}.
+In addition, it is possible to register
+two functions {\tt central} and {\tt error} being called everytime \GLB{SetCentralValues} or \GLB{SetInputErrors} are called with the same argument.
+For example, the central values and input errors can be stored in
+global variables for a faster access.
+These pointers can also be {\tt NULL}.
+The pointer {\tt user\_data} can be optionally used to circumvent the use of global variables.
+It is set by {\tt glbRegisterPriorFunction} and transferred to the registered functions whenever called.
+\end{function}
 
@@
  glbInit(argv[0]); \\
- glbRegisterPriorFunction(my\_prior,NULL,NULL);
+ glbRegisterPriorFunction(my\_prior,NULL,NULL,NULL);
 }
@@
   \mbox{glbDefineParams(starting\_values,theta12,theta13,theta23,deltacp,sdm,-ldm);} \\
+  glbSetDensityParams(starting\_values,1.0,GLB\_ALL); \\
   \\
@@
 \end{function}
+Note that for this and the other probability functions $1 \le$ {\tt l}, {\tt m} $\le 3$.
 In addition, the oscillation probabilities in matter can be obtained
@@
 \section{Baseline and matter density profile}
+\label{sec:baselinemd}
 \index{norm}{Baseline!change}
@@
 
-TO BE EXTENDED BY LIST HANDLING (TRANSFER LISTS IN AEDL VARIABLES). STATUS? [WW].
-
 Using external parameters in \AEDL\ files is a very powerful feature
@@
 
+Similar to a simple \AEDL\ variable, one can transfer the value of an
+\AEDL\ list with (\cf, \Sec~\ref{sec:syntax})
+\begin{function}
+\GLBNS{DefineAEDLList}
+{\tt glbDefineAEDLList(const char *name, double *list, size\_t length)} assigns the list {\tt list}
+of length {\tt length} to the \AEDL\ variable {\tt name}.
+\end{function}
+See \Sec~\ref{sec:baselinemd} for how to use such lists.
+%
 In order to clear the external variable stack if one is
@@
 \GLBNS{\_set\_oscillation\_parameters\_function}
-{\tt int (*glb\_set\_oscillation\_parameters\_function)(glb\_params p)} reads the fundamental oscillation parameters
+{\tt int (*glb\_set\_oscillation\_parameters\_function)(glb\_params p, void* user\_data)} reads the fundamental oscillation parameters
 obtained from {\tt p} and stores them into global variables accessible to {\tt glb\_probability\_matrix\_function}.
  In addition, it can be used to pre-compute quantities which are dependent on the oscillation paramaters only for efficiency. 
+In order to circumvent global variables, an arbitrary pointer can be used by {\tt user\_data} which is set by {\tt glbRegisterProbabilityEngine}.
 \end{function}
@@
 \GLBNS{\_get\_oscillation\_parameters\_function}
-{\tt int (*glb\_get\_oscillation\_parameters\_function)(glb\_params p)} returns the fundamental oscillation parameters
+{\tt int (*glb\_get\_oscillation\_parameters\_function)(glb\_params p, void* user\_data)} returns the fundamental oscillation parameters
 in {\tt p} which are obtained from the global variables accessible to {\tt glb\_probability\_matrix\_function}.
+In order to circumvent global variables, an arbitrary pointer can be used by {\tt user\_data} which is set by {\tt glbRegisterProbabilityEngine}.
 \end{function}
@@
                   int psteps, const double *lengths, const double *densities,
-                  double filter\_sigma)} calculates the neutrino oscillation probability matrix to be returned in {\tt P}.
+                  double filter\_sigma, void* user\_data)} calculates the neutrino oscillation probability matrix to be returned in {\tt P}.
 The {\tt cp\_sign} is +1 for neutrinos and -1 for antineutrinos.
@@
 The parameter {\tt filter\_sigma} defines the width of the low-pass filter to be used, or no filter usage if negative.
+In order to circumvent global variables, an arbitrary pointer can be used by {\tt user\_data} which is set by {\tt glbRegisterProbabilityEngine}.
 \end{function}
@@
                  glb\_set\_oscillation\_parameters\_function set\_params\_func, \\
-                 glb\_get\_oscillation\_parameters\_function get\_params\_func)} registers a probability
+                 glb\_get\_oscillation\_parameters\_function get\_params\_func, \\
+                 void* user\_data)} registers a probability
 engine for the simulation of non-standard physics with {\tt n\_parameters} oscillation parameters ({\tt n\_parameters} $\ge6$).
 The three functions {\tt prob\_func}, {\tt set\_params\_func}, and {get\_params\_func}
-will be used with the function types defined above.         
+will be used with the function types defined above. In order to circumvent global variables, an arbitrary pointer can be used by {\tt user\_data} being set by this function.
 \end{function}

### 2007-01-09 Documentation/part1.tex
User's manual complete

@@
 
-HIER BIN ICH GERADE [WW].
-
-Example part here for non-standard implementation? [WW]
+Let us now illustrate the implementation of non-standard physics with a simple example, which can
+be found as {\tt example6.c} in the example directory.
+This example is a simplified version of \Sec~4 of 
+\Ref~\cite{Blennow:2005yk}.
+It uses an analytical probability calculation for a reactor experiment with a baseline being treated
+in vacuum for simplicity. The non-standard effect is the loss of coherence because of wave packet decoherence
+or any other such effect, \ie, one more oscillation parameter. We need to define two functions to access a set of 
+global oscillation parameters
+\begin{quote}
+{\tt 
+double th12, th13, th23, deltacp, sdm, ldm, sigma\_E; \\
+\\
+int my\_set\_oscillation\_parameters(glb\_params p, void *user\_data) \\
+\{ \\
+ \hspace*{0.5cm} th12    = glbGetOscParams(p, GLB\_THETA\_12); \\
+ \hspace*{0.5cm} th13    = glbGetOscParams(p, GLB\_THETA\_13); \\
+ \hspace*{0.5cm} th23    = glbGetOscParams(p, GLB\_THETA\_23); \\
+ \hspace*{0.5cm} deltacp = glbGetOscParams(p, GLB\_DELTA\_CP); \\
+ \hspace*{0.5cm} sdm     = glbGetOscParams(p, GLB\_DM\_21) * 1.0e-18;  \\
+ \hspace*{0.5cm} ldm     = glbGetOscParams(p, GLB\_DM\_31) * 1.0e-18;  \\
+ \hspace*{0.5cm} sigma\_E = glbGetOscParams(p, GLB\_SIGMA\_E); \\
+\\
+  return 0; \\
+\} \\
+\\
+int my\_get\_oscillation\_parameters(glb\_params p, void *user\_data) \\
+\{ \\
+  \hspace*{0.5cm} glbSetOscParams(p, th12, GLB\_THETA\_12); \\
+  \hspace*{0.5cm} glbSetOscParams(p, th13, GLB\_THETA\_13); \\
+  \hspace*{0.5cm} glbSetOscParams(p, th23, GLB\_THETA\_23); \\
+  \hspace*{0.5cm} glbSetOscParams(p, deltacp, GLB\_DELTA\_CP); \\
+  \hspace*{0.5cm} glbSetOscParams(p, sdm*1.0e18, GLB\_DM\_21);   \\
+  \hspace*{0.5cm} glbSetOscParams(p, ldm*1.0e18, GLB\_DM\_31);   \\
+  \hspace*{0.5cm} glbSetOscParams(p, sigma\_E, GLB\_SIGMA\_E); \\
+\\
+  \hspace*{0.5cm} return 0; \\
+\} 
+}
+\end{quote}
+Furthermore, we need to define the probability matrix calculation itself -- in this case,
+we perform an analytical calculation:
+\begin{quote}
+{\tt
+int my\_probability\_matrix(double P[3][3], int cp\_sign, double E, int psteps,
+                          const double *length, const double *density,
+                          double filter\_sigma, void *user\_data) \\
+\{ \\
+ \hspace*{0.5cm}  int i, j; \\
+ \hspace*{0.5cm} double L, Delta21, Delta31, Delta32; \\
+ \hspace*{0.5cm} double D21, D31, D32, s13, c13, s12, c12, t; \\
+  \\
+  \hspace*{0.5cm} /* Set all probabilities to zero initially */ \\
+  \hspace*{0.5cm} for (i=0; i < 3; i++) for (j=0; j < 3; j++) P[i][j] = 0.0; \\
+\\
+ \hspace*{0.5cm} /* Calculate total baseline */ \\
+ \hspace*{0.5cm} L = 0.0; for (i=0; i < psteps; i++) L += length[i]; \\
+ \hspace*{0.5cm} L = KM\_TO\_EV(L) * 1.0e9;      /* Convert to GeV$^{-1}$ */ \\
+\\
+ \hspace*{0.5cm} /* Compute P\_ee */ \\
+  \hspace*{0.5cm} s12 = sin(th12); c12 = cos(th12); \\
+  \hspace*{0.5cm} s13 = sin(th13); c13 = cos(th13); \\
+  \hspace*{0.5cm} t = L / (4.0 * E); \\ 
+  \hspace*{0.5cm} Delta21 = sdm * t; Delta31 = ldm * t; \\
+  \hspace*{0.5cm} Delta32 = Delta31 - Delta21; \\
+  \hspace*{0.5cm} t = M\_SQRT2 * sigma\_E / E; \\
+  \hspace*{0.5cm} D21 = exp(-square( Delta21 * t )); \\
+  \hspace*{0.5cm} D31 = exp(-square( Delta31 * t )); \\
+  \hspace*{0.5cm} D32 = exp(-square( Delta32 * t )); \\
+  \hspace*{0.5cm} P[0][0] = square(square(c13)) * ( 1 - 2.0*square(s12*c12)* \\
+  \hspace*{1cm} (1 - D21*cos(2.0*Delta21)) ) 
+  \hspace*{1cm} + 2.0*square(s13*c13) * \\
+  \hspace*{1cm} ( D31*square(c12)*cos(2.0*Delta31) \\
+  \hspace*{1cm} + D32*square(s12)*cos(2.0*Delta32) )  + square(square(s13)); \\
+\\
+  \hspace*{0.5cm} return 0; \\
+\}
+}
+\end{quote}
+Now we can register the probability engine after {\tt glbInit}:
+\begin{quote}
+{\tt
+  glbInit(argv[0]); \\
+  glbRegisterProbabilityEngine(7,      /* Number of parameters */ \\
+  \hspace*{1cm}                             \&my\_probability\_matrix, \\
+  \hspace*{1cm}                             \&my\_set\_oscillation\_parameters, \\
+  \hspace*{1cm}                             \&my\_get\_oscillation\_parameters, \\
+  \hspace*{1cm}                             NULL); 
+}
+\end{quote}
+We will demonstrate in the next section how to access the non-standard physics 
+parameter in the application software.
+
+\example{Decoherence in neutrino oscillations}{
+\label{ex:nsphysics}
+\index{norm}{Decoherence}
+The following code fragment from {\tt example6.c} calculates 
+a fit region on $\stheta$-$\sigma_E$-space, where $\sigma_E$ is the
+non-standard decoherence parameter (\cf, \Ref~\cite{Blennow:2005yk}):
+\begin{quote}
+{\tt {\footnotesize
+  const GLB\_SIGMA\_E = 6; \\ \\
+  /* Fix theta13 and sigma\_E, and marginalize over all other parameters except deltaCP and th23, which do not enter P\_ee */ \\
+  glbDefineProjection(myproj, GLB\_FREE, GLB\_FIXED, GLB\_FIXED,
+    GLB\_FIXED, \\ \hspace*{0.5cm} GLB\_FREE, GLB\_FREE); \\
+  glbSetDensityProjectionFlag(myproj, GLB\_FIXED, GLB\_ALL); \\
+  glbSetProjectionFlag(myproj, GLB\_FIXED, GLB\_SIGMA\_E);  \\
+  glbSetProjection(myproj); \\ 
+\\
+  for(x=0; x < 0.05+0.001; x+=0.005)      /* th13 loop */ \\
+  for(y=0.0; y < 0.010+0.001; y+=0.001)   /* sigma\_E loop */ \\
+    \{ \\
+   \hspace*{0.5cm}    /* Set vector of test=fit values */ \\
+   \hspace*{0.5cm}    thetheta13=asin(sqrt(x))/2.0; \\
+   \hspace*{0.5cm}    glbSetOscParams(test\_values,thetheta13,GLB\_THETA\_13); \\
+   \hspace*{0.5cm}    glbSetOscParams(test\_values,y,GLB\_SIGMA\_E); \\
+     \\
+   \hspace*{0.5cm}    /* Compute Chi2 with correlations */ \\
+   \hspace*{0.5cm}    res=glbChiNP(test\_values,NULL,GLB\_ALL); \\
+  \\
+   \hspace*{0.5cm}    AddToOutput(x,y,res); \\
+    \} 
+}}
+\end{quote}
+
+The result is represented by the dark curves
+ in the following figure (similar to \Ref~\cite{Blennow:2005yk}):
+\begin{center}
+\colorbox{white}{\includegraphics[width=6cm]{reactorNSI}}
+\end{center}
+}
 
@@
 
-Using more than six transition parameters in \GLOBES , you will have to maintain the additional parameters. 
+Using more than six parameters in \GLOBES , you will have to maintain the additional parameters. 
 For example, there are no standard routines in \GLOBES\ which define projections including more than six parameters, 
@@
 \item
-{\tt glbSetOscParams}
+\GLB{SetOscParams}
 \item
-{\tt glbGetOscParams}
+\GLB{GetOscParams}
 \item
-{\tt glbSetProjectionFlag}
+\GLB{SetProjectionFlag}
 \item
-{\tt glbGetProjectionFlag}
+\GLB{GetProjectionFlag}
 \end{itemize}
-In all these cases, the last parameter {\tt which} can now run from $0$ to {\tt GLB\_OSCP}$-1$. For instance,
+In all these cases, the last parameter {\tt which} can now run from $0$ to \GLB{GetNumOfOscParams}{\tt ()-1}. For instance,
 \begin{quote}
@@
  glbDefineParams(true\_values,0.55,0.15,0.78,0.0,0.000082,0.0022); \\
+ glbSetDensityParams(true\_values,1.0,GLB\_ALL); \\
  glbSetOscParams(true\_values,0.0,6);
@@
 \end{quote}
-sets all of the oscillation parameters if {\tt GLB\_OSCP}=7. In addition, the function {\tt glbCopyParams}
+sets all of the oscillation parameters if you have one additional parameter. 
+In addition, the function {\tt glbCopyParams}
 can be used to copy all parameters inluding the non-standard ones. For projections,
@@
 the marginalization over the extra parameter(s) by {\tt glbSetProjectionFlag}. You can find a simple
-example using non-standard physics on page~\pageref{ex:nsphysics}.
-
-\example{Non-standard process on Hamiltonian level}{
-\label{ex:nsphysics}
-\index{norm}{Hamiltonian}
-Let us discuss an example of a very simple Hamiltonian-level non-standard flavor-changing effect which 
-couples $\nu_\mu$ to $\nu_\tau$ in flavor base. This effect can be parameterized
-relative to the standard matter effect acting on $\nu_e$ by an additional
-parameter, which we adopt as our 7th oscillation parameter. In order
-to incorporate this effect, we change a few lines in the function {\tt ComputeMixingMatrix}
-in {\tt glb\_probability.cc}:
-\begin{quote}
-{\tt {\footnotesize
-double msw = 0.1*1.5*anti\_neutrinos*(10E-4)*neutrino\_energy*0.5*matter\_density; \\
-md[0][0] += msw; \\
-md[1][2] += msw*nsparams[0]; md[2][1] += msw*nsparams[0];
-}}
-\end{quote}
-Here we have assume a real contribution which has to be added in a Hermitian way (the Hamiltonian
-has to be Hermitian), where the first line is the standard matter effect acting on $\nu_e$ only 
-($2 \sqrt{2} G_F n_e E$ expressed in our units).
-
-In the application software, we can now use this additional parameter. Let us assume that we
-want to compute the projection onto the $\stheta$-axis when we do not know the size of the  non-standard
-effect, \ie, we keep it completely free but assume it to be zero as null hypothesis:
-\begin{quote}
-{\tt {\footnotesize
- \#define MY\_FV 6 \\
- ... \\
- glbDefineParams(true\_values,theta12,theta13,theta23,deltacp,sdm,ldm); \\
- glbSetOscParams(true\_values,0.0,MY\_FV); \\
- glbDefineParams(input\_errors,0.1*theta12,0,0,0,0.1*sdm,0); \\
- glbSetOscParams(input\_errors,0,MY\_FV);  \\
- glbSetDensityParams(input\_errors,0.05,GLB\_ALL); \\
- glbSetOscillationParameters(); glbSetRates(); \\
- glbSetCentralValues(true\_values);glbSetInputErrors(input\_errors); \\
- \\
- glbDefineProjection(my\_proj,GLB\_FREE,GLB\_FIXED,GLB\_FREE,\\
-  \hspace*{2cm} GLB\_FREE,GLB\_FREE,GLB\_FREE); \\
- glbSetProjectionFlag(my\_proj,GLB\_FREE,MY\_FV); \\
- glbSetDensityProjectionFlag(my\_proj,GLB\_FREE,GLB\_ALL); \\
- glbSetProjection(my\_proj);
- \\
- glbCopyParams(true\_values,fit\_values); \\
- glbSetOscParams(fit\_values,0.1,GLB\_THETA\_13); \\
- res=glbChiNP(fit\_values,minimum,GLB\_ALL); 
-}}
-\end{quote}
-This code fragment illustrates the implementation of a {\tt glbChiTheta13} function
-including the non-standard effect, which is being marginalized over.
-Alternatively, one could define a {\tt glbChiTheta13} fixing the non-standard
-parameter, or include external constrainys to limit its size. Therefore, 
-there is no ``standard'' treatment of new physics effects incorporated
-in \GLOBES .
-}
+example using non-standard physics with the code from the last section on page~\pageref{ex:nsphysics}.
+
+\chapter{Experimental features}
+
+Here we describe experimental features currently being implemented in \GLOBES .
+These features have not yet been tested extensively and should be used with care.
+They may evolve into standard features in the future, or they may not be supported anymore at a certain point.
+In general, the number of experiment features is small in a new release version, and increases towards a new version number.
+
+\subsection*{\GLOBES\ 3.0 and higher}
+
+One can change the minimization algorithm in \GLOBES\ with
+\begin{function}
+\GLBNS{SelectMinimizer}
+{\tt int glbSelectMinimizer(int minimizer\_ID)} selects the minimizer {\tt GLB\_MIN\_NESTED\_POWELL}, {\tt GLB\_MIN\_DEFAULT}, or the hybrid minimzer {\tt GLB\_MIN\_POWELL} by the parameter {\tt minimizer\_ID}.
+{\tt GLB\_MIN\_NESTED\_POWELL} is the currently implemented standard algorithm, {\tt GLB\_MIN\_DEFAULT} chooses the standard algorithm at any given time, and {\tt GLB\_MIN\_POWELL} is a faster, yet rarely tested hybrid
+minimizer.
+\end{function}
+Compared to the standard minimization which is performed for systematics first and then for the
+oscillation parameters, the hybrid minimizer mixes the systematics and oscillation parameter minimizations.
+This method is much faster, but new (unknown) correlations between systematics parameters and oscillation parameters may emerge.
+Since one can change the minimizer at any time, it is recommended that one cross check the minimization for a particular experiment.
 

### 2007-01-09 Documentation/part2.tex
*** empty log message ***

@@
 The description of the neutrino oscillation physics is, at least numerically, relatively simple.
-We use the {\em evolution operator method}~\cite{Ohlsson:1999um}  to compute the neutrino oscillation probabilities and divide the matter density profile into layers of constant matter density.
+We use the {\em evolution operator method} (see, \eg, \Ref~\cite{Ohlsson:1999um})  to compute the neutrino oscillation probabilities and divide the matter density profile into layers of constant matter density.
 For each of these layers, the Hamiltonian in matter is diagonalized in order to propagate the neutrino transition amplitudes.
@@
 in order not to confuse it with some other file format.
+In addition, \GLOBES\ 3.0 and higher requires the identification of the
+minimum \GLOBES\ version the \AEDL\ file can be used with:
+\begin{quote}
+{\tt \$version="3.0.0"}
+\end{quote}
+
 %
@@
 {\tt \tb @signal = 0.45 @ \#appearance}\\
-{\tt \tb @signalerror = 0.001 : 0.0001}\\
+{\tt \tb @signalerror = 0.025 : 0.0001}\\
 {\tt \tb @background = 1.0e-05 @ \#disappearance}\\
-{\tt \tb @backgroundcenter = 1 : 0.0}\\
-{\tt \tb @backgrounderror = 0.05 : 0.0001}\\
-{\tt \tb @errordim\_sys\_on = 0}\\
-{\tt \tb @errordim\_sys\_off = 2}\\
+%{\tt \tb @backgroundcenter = 1 : 0.0}\\
+{\tt \tb @backgrounderror = 0.2 : 0.0001}\\
+{\tt \tb @sys\_on\_function = "chiSpectrumTilt"}\\
+{\tt \tb @sys\_off\_function = "chiNoSysSpectrum"}\\
 {\tt \tb @energy\_window = 4.0 : 50.0}\\
@@
 above defined channel named {\tt \#appearance} with an constant overall
-efficiency of $0.45$. The signal error variable has two components: 
-The first one is the normalization error of the signal, here $0.1\%$. The second 
+efficiency of $0.45$ (in a more realistic simulation, one would introduce an
+energy threshold function). The signal error variable has two components: 
+The first one is the normalization error of the signal, here $2.5\%$. The second 
 one refers to the energy calibration error of the signal, which is defined 
@@
 use the fraction $1\cdot 10^{-5}$ of the channel named {\tt \#disappearance}, \ie , the muon neutrinos with a mis-identified charge. 
-The background center variable allows to rescale the total background contribution from all background components
-simultaneously.
-It is only useful if there is more than one background component, otherwise it is usually $1$.
-The background error variable is defined such as the signal error variable, \ie\ we have a $5\%$ background uncertainty and a very small energy calibration error.
-The ``error dimension variable'' {\tt @errordim\_sys\_X} selects how the systematical errors are treated (\cf, \Tab~\ref{tab:error_dim}).
+%The background center variable allows to rescale the total background contribution from all background components
+%simultaneously. It is only useful if there is more than one background component, otherwise it is usually $1$. 
+The background error variable is defined such as the signal error variable, \ie , we have a $20\%$ background uncertainty and a very small energy calibration error.
+In addition, the systematics treatment is specified
+in {\tt @sys\_on\_function} (for systematics switched on) and {\tt @sys\_off\_function} (for systematics switched off) -- see \Tab~\ref{tab:error_dim}.
 
@@
 in order to avoid parsing megabytes of chunk
- and producing thousands of error messages. 
+ and producing thousands of error messages.  In addition, the version of the minimum
+\GLOBES\ version the \AEDL\ file is running with has to be defined by {\tt \$version}, such as
+{\tt \$version="3.0.0"}.
+\index{aedl}{version@{\tt \$version}}
 %
@@
 {\tt /* This starts a comment\\
- and here the comment ends */
+ and here the comment ends */ \\
+// Another comment
 }
@@
 The first part is the type of environment, which is {\tt channel} here. 
-There are the following types of environment in AEDL:
+There are the following types of environments in AEDL:
 \begin{quote}
@@
 \begin{quote}
-{\tt    include  "NuFact.gls"\\
+{\tt    include  "Exp1.glb"\\
         \#NEXT\#\\
-        include   "JHFHK.gls"
-}
-\end{quote}
-in order to initialize a combined analysis of the experiments defined in the files {\tt NuFact.gls} and {\tt JHFHK.gls}. Note that one has 
+        include   "Exp2.glb"
+}
+\end{quote}
+in order to initialize a combined analysis of the experiments defined in the files {\tt Exp1.glb} and {\tt Exp2.glb}. Note that one has 
 to use quotation marks for filenames in \AEDL.\index{norm}{File names}
@@
 \verb+Pi^2+\\
-sin(Pi/2)\\
+sin(Pi/2)
 }
@@
 \index{aedl}{sqrt@{\tt sqrt}}
-
+% 
 These functions can be used everywhere, where
@@
  the end of the input -- alternatively you can put a comment at the end.
+
+\section{More advanced AEDL features}
+
+In \GLOBES\ 3.0 and higher, a number of new features can be used. The
+features in this section are slighlty more advanced and can be jumped over at a first reading of
+the manual. The
+most important one are lists as variables in \AEDL . They start with {\tt \%}, such as
+\begin{quote}
+{\tt
+\%effs = \{ 0.2, 0.4, 0.6, 0.8, 1.0, 1.0, 1.0 \}
+}
+\end{quote}
+Functions can be threaded over lists, \ie, by applying the function to each element
+and returning a list. Note that the original list will be destroyed by this
+process. Therefore, it is necessary to create a copy of your list if you want to
+use the original and the threading result. For that purpose, the {\tt copy}
+funcion is provided:
+\begin{quote}
+{\tt
+ listb = copy(lista) \\
+ listb := lista // Alternative method 
+}
+\end{quote}
+\index{aedl}{copy@{\tt copy}}
+You will also need to use {\tt copy} when you assign a list to an experiment structure (see below).
+
+Two helper functions {\tt bincenter()} 
+\index{aedl}{bincenter@{\tt bincenter}} and {\tt simbincenter} \index{aedl}{simbincenter@{\tt simbincenter}} return lists with the central energies of the bins or sampling points, respectively.
+For example,
+\begin{quote}
+{\tt
+ \%bc=bincenter() 
+}
+\end{quote}
+A very useful new features is an interpolation function which can directly interpolate a number of
+points and evaluate them at a different set of places. For example,
+\begin{quote}
+{\tt
+ \%energ = \{ 4.0,20.0,50.0 \} \\
+ \%effs = \{ 0.0,1.0,1.0 \} \\
+ \%ires = interpolation(\%energ,\%effs,1,\%bc) 
+}
+\end{quote}
+\index{aedl}{interpolation@{\tt interpolation}}
+interpolates the points with x-values {\tt \%energ} and y-values {\tt \%effs} with the interpolation order one (linear interpolation, third parameter) and evaluates the interpolation result at the bin centers obtained above, \ie, it returns a list of the y-values at the places specified by the last parameter.
+The only allowed interpolation orders are ``1'' (linear) and ``2'' (cubic splines).
+This example creates
+a neutrino factory energy threshold function linearly climbing from 0 to 1 between 4~GeV and 20~GeV. It can be directly used in a channel definition, \eg, 
+\begin{quote}
+{\tt
+ channel(\#nu\_mu\_appearance)< \\
+\hspace*{0.5cm}	@channel = \#mu\_plus: +: electron: muon: \#CC: \#MINOS\ \\
+\hspace*{0.5cm}	@post\_smearing\_efficiencies = copy(\%ires) \\
+>
+}
+\end{quote}
+
+For debugging purposes of lists and numbers, one can \GLOBES\ now supports output directly
+from \AEDL\ files:
+\begin{quote}
+{\tt 
+ R = 1.15 \\
+ echo(R) // Print without line feed \\
+ line(2) // Two line feeds \\
+ echon(ires) // Print with line feed 
+}
+\end{quote}
+\index{aedl}{echo@{\tt echo}}
+\index{aedl}{echon@{\tt echon}}
+\index{aedl}{line@{\tt line}}
 

### 2007-01-10 Documentation/appendix.tex
Part of manual updated
*** empty log message ***

@@
 \index{norm}{Installation|(}
+
+{\bf NEEDS TO BE UPDATED [WW] -> PH? ADD SECTION ON CONDOR/CLUSTER SUPPORT?}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\chapter{Flux normalization in \GLOBES }
+\index{norm}{Normalization of fluxes}
+\label{app:flux}
+
+A common issue with \GLOBES\ is confusion about the proper units for
+the input flux files for use in \AEDL\ experiment descriptions. Source of
+the confusion is an undocumented factor 5.2 in \GLOBES\ versions older
+than 3.0 the fluxes are multiplied with (see below). In Version 3.0 and 
+higher, the alternative flux environment {\tt nuflux} is provided, 
+which does not contain this factor. The following material is
+based on the old environment {\tt flux}. For the use of {\tt nuflux},
+replace the factor 5.20034 by unity.
+
+\section*{Historical problem}
+
+One problem for the design of \AEDL\ was initially that meaningful
+units for flux data strongly depend on the given type of experiment,
+but also on relatively arbitrary decisions. For accelerator beams
+based on pion decay, one frequently defines the beam luminosity in
+{\it protons on target (pot)} since this number has a one-to-one
+correspondence with the number of neutrinos produced. Another sensible
+unit could be {\it megawatt on target (MW)}, again this number is directly
+correlated with the number of neutrinos and moreover there is a unique
+relation to {\it pot} for a given accelerator. Of course,
+what matters is the integrated luminosity. In some  cases the
+neutrino flux is given per $10^7\,\mathrm{s}$. However,
+most experiments will run for several years, hence also this number
+has to enter somewhere. For neutrino factories the proper number is
+useful muon decays per unit time, and for reactor experiments it is
+the thermal power of the reactor {\it asf}. This demonstrates that it is
+reasonable to keep the flux definition flexible.
+
+\section*{Implementation in \GLOBES }
+
+In understanding how one still can figure out what the correct units
+are for each case, it is a good starting point to look at what
+\GLOBES\ does with the input files. The cross section in the file is given as
+differential cross section divided by energy $x=\sigma/E$, and the flux file
+gives $f$. The differential number of events per GeV $n$ as computed in
+\GLOBES\ without oscillation and efficiencies is given by
+\begin{eqnarray}
+n &=& 5.20034\times x\times E\times f\times \nonumber \\
+&&\mathtt{@norm}\times\mathtt{@power}\times\mathtt{@stored\_muons}\times\mathtt{@time}\times\mathtt{\$target\_mass}\times(\mathtt{\$baseline})^{-2} \nonumber
+\end{eqnarray}
+Note that $5.20034$ is a undocumented fudge factor!
+
+It is the sole responsibility of the author of the \AEDL\ file and its
+supporting files, to ensure that the result makes sense. In
+principle, it is possible to divide, for example, {\tt @time}  by
+$\pi$ and fix that by redefining the flux file by multiplying
+it with $\pi$. Modifications like that have happened in the past
+and still happen, and many of them are not properly commented.
+
+\section*{Writing \AEDL\ files}
+
+The task is to choose the value of  {\tt @norm} such that all the
+variables in the \AEDL\ file have the proper units, \eg , {\tt
+  @time} has proper unit years.
+
+\GLOBES\ assumes that the cross
+section $x$ is given in $10^{-38}\,\mathrm{cm}^2$ and that all fluxes are
+given at a distance of $1\,\mathrm{km}$. In addition, it assumes that
+the number of target nuclei $\tau$ (or protons or whatever applies to the
+given cross section) per unit target mass $m_u$ (which usually is $kt$)
+ are properly accounted for.\footnote{Note that the cross sections which are delivered with
+  \GLOBES\ always are per nucleon.}  Assuming that in the flux file
+the data is given as number of neutrinos per unit area $A$ and
+energy bin of width $\Delta E$ at a distance $L$ from the source, one
+obtains
+\begin{eqnarray}
+\mathtt{@norm}=\frac{1}{5.20034}\left(\frac{\mathrm{GeV}}{\Delta
+      E}\right)
+\left(\frac{\mathrm{cm}^2}{A}\right)\left(\frac{L}{\mathrm{km}}\right)^2\left(\frac{\tau}{m_u}\right)\times10^{-38}\times\left(\frac{\mathcal{L}_u}{\mathcal{L}}\right)
+\end{eqnarray}
+where $\mathcal{L}$ absorbs all factors in the flux file related to
+the integrated luminosity, and $\mathcal{L}_u$ is the unit chosen for
+it.  The concept of integrated luminosity is
+nicely described in the \GLOBES\ manual in \Sec~\ref{sec:source}.
+A little example illustrates this concept:
+The flux is given for $10^{21}\, \mathrm{pot}\,\mathrm{y}^{-1}$ of
+$10\,\mathrm{GeV}$ protons, thus a good choice for the units $\mathcal{L}_u$
+is $\mathrm{MW}\,\mathrm{y}^{-1}$, which means that
+$\mathcal{L}/\mathcal{L}_u$ is given by (assuming a $10^7\,\mathrm{s}$
+year)
+\begin{equation}
+\frac{\mathcal{L}}{\mathcal{L}_u}=\frac{10\,\mathrm{GeV}\,10^{21}\,\mathrm{pot}\,\mathrm{y}}{10^7\,\mathrm{s}}\times(\mathrm{MW}\,\mathrm{y}^{-1})^{-1}=0.160218\ldots
+\end{equation}
+
 {\footnotesize

### 2007-01-10 Documentation/part1.tex
*** empty log message ***

@@
 \end{center}
-\caption{\label{tab:stdfunctions} \index{norm}{Standard functions (table)} The \GLOBES\ standard function to obtain a $\chi^2$-value with systematics only or systematics and correlations.
+\mycaption{\label{tab:stdfunctions} \index{norm}{Standard functions (table)} The \GLOBES\ standard function to obtain a $\chi^2$-value with systematics only or systematics and correlations.
 The parameters {\tt rule} and {\tt exp}
@@
 \end{center}
-\caption{\label{fig:reactor} The result from {\tt example5.c} (figure similar in \Ref~\cite{Huber:2006vr}). The fat curve corresponds to the systematics on page~\pageref{ex:syst}.}
+\mycaption{\label{fig:reactor} The result from {\tt example5.c} (figure similar in \Ref~\cite{Huber:2006vr}). The fat curve corresponds to the systematics on page~\pageref{ex:syst}.}
 \end{figure}
@@
 \end{tabular}
-\caption{\label{tab:defprojection}\index{norm}{Projection!type}
+\mycaption{\label{tab:defprojection}\index{norm}{Projection!type}
 Different functions handling the

### 2007-01-10 Documentation/part2.tex
*** empty log message ***

@@
 \end{center}
-\caption{\label{fig:distro} A detector maps a true parameter value onto
+\mycaption{\label{fig:distro} A detector maps a true parameter value onto
 a distribution of reconstructed parameter values. This is illustrated here for there energy.}
@@
 \end{center}
-\caption{\label{fig:channel} General concept of a ``channel''.}
+\mycaption{\label{fig:channel} General concept of a ``channel''.}
 \end{figure}
@@
 \end{center}
-\caption{\label{fig:rule} General concept of a ``rule''.}
+\mycaption{\label{fig:rule} General concept of a ``rule''.}
 \end{figure}
@@
 \end{center}
-\caption{\label{fig:experiment} General concept of an ``experiment''.}
+\mycaption{\label{fig:experiment} General concept of an ``experiment''.}
 \end{figure}
@@
 {\tt /* beam */}\\
-{\tt flux(\#mu\_plus)<\\
+{\tt nuflux(\#mu\_plus)<\\
 \tb  @builtin = 1\\
@@
 \begin{quote}
-{\tt flux\\
+{\tt nuflux\\
 cross\\
@@
 
-In this chapter, we give a detailed definition of the AEDL features.
+In this chapter, we give a detailed definition of the \AEDL\ features.
 We also show the underlying mathematical concepts, where applicable.
@@
 we illustrate many of the features of the \GLOBES\ simulation successively
-in the logical order of their definition, and demonstrate how they translate into AEDL .
+in the logical order of their definition, and demonstrate how they translate into \AEDL .
 
@@
 \mathrm{Source~power}\,\left[\mathrm{MW/GW}\right]\\
-\mathrm{Useful~muon~decays}\,\left[\mathrm{yr}^{-1}\right]
+\mathrm{Useful~parent~decays}\,\left[\mathrm{yr}^{-1}\right]
 \end{array}\right.\,.
 \end{equation}
-Thus, the source power corresponds to either the amount of energy produced per time frame in the target (such as for nuclear reactors or sources based on pion decay), or the useful muon decays per time frame (neutrino factories).
-In addition, the definition of the source power makes only sense together with the flux normalization, the running time, and fiducial detector mass in order to define the total integrated luminosity.
+Thus, the source power corresponds to either the amount of energy produced per time frame in the target (such as for nuclear reactors or sources based on pion decay), or the useful parent particle decays per time frame (neutrino factories, beta beams).
+In addition, the definition of the source power makes only sense together with the flux normalization, the running time, and the fiducial detector mass in order to define the total integrated luminosity.
 Therefore, one can, in principle, use arbitrary units for these components as long as their product gives the wanted neutrino flux.
@@
 idea to document the choices made by the user by corresponding comments
-in \AEDL.
+in \AEDL. For more details on the luminosity implementation, see appendix on page~\pageref{app:flux}.
 
@@
 
-\index{aedl}{{\tt flux}!time@{\tt "@time}}
+\index{aedl}{{\tt nuflux}!time@{\tt "@time}}
 There are two principal ways to initialize a neutrino flux: 
@@
 a built-in source, or one can provide a file. In both cases,
-a flux is defined by the environment {\tt flux}, such as
-\index{aedl}{{\tt flux}}
-\begin{quote}
-  {\tt flux(\#name)<\\
+a flux is defined by the environment {\tt nuflux}, such as
+\index{aedl}{{\tt nuflux}}
+\begin{quote}
+  {\tt nuflux(\#name)<\\
 \tb $\ldots$\\
@@
 with a running time of $8$ years. Note that the running time is used within
-the {\tt flux} environment. This feature can be used to load the neutrino and antineutrino fluxes separately, in order to combine them with different
-running times within one experiment. The name of the flux {\tt \#name} will later be refered to in the channel definitions.
+the {\tt nuflux} environment. This feature can be used to load the neutrino and antineutrino fluxes separately, in order to combine them with different
+running times within one experiment.
+The name of the flux {\tt \#name} will later be referred to in the channel definitions.
+Note that \GLOBES\ versions older than 3.0 use the still supported {\tt flux} environment, which is different from {\tt nuflux} by an undocumented normalization factor 5.2 for user-defined fluxes.
+This difference is explained in the appendix on page~\pageref{app:flux}.
+
+\index{aedl}{{\tt flux}}
+
+\begin{table}[t]
+\begin{center}
+\begin{tabular}{|cllc|}
+\hline
+{\tt @builtin} & Description & Parameters & Min. version \\
+\hline
+1 & Neutrino factory $\mu^+$ decay & {\tt @parent\_energy} [GeV], & 2.0 \\
+& & {\tt @stored\_muons} &  \\
+2 & Neutrino factory $\mu^-$ decay & {\tt @parent\_energy} [GeV], & 2.0 \\
+& & {\tt @stored\_muons} &  \\
+3 & Beta beam, inv. beta decay & {\tt @end\_point} [GeV], & 3.0 \\
+& & {\tt @stored\_ions}, &  \\
+& & {\tt @gamma} & \\
+4 & Beta beam, beta decay & {\tt @end\_point} [GeV], & 3.0 \\
+& & {\tt @stored\_ions}, &  \\
+& & {\tt @gamma} & \\
+\hline
+\end{tabular}
+\end{center}
+\mycaption{\label{tab:binfluxes} Built-in fluxes currently supported by \GLOBES . For details on
+the beta beams, see \Ref~\cite{Huber:2005jk}.
+MR ABOUT BETA BEAMS: DOES THIS REALLY WORK FOR ALL ISOTOPES?
+IS THE DIFFERENCE BETWEEN 3 AND 4 CORRECTLY DESCRIBED?
+WHY IS THERE A NORMALIZATION FACTOR 50.9 IN THE FILES?}
+\end{table}
 
@@
 case, there is no degree of freedom in the choice of the source units.
-Currently, two built-in fluxes are available: $\mu^+$-decay ({\tt @builtin = 1}) and $\mu^-$-decay ({\tt @builtin = 2}).
+The currently available fluxes are described in \Tab~\ref{tab:binfluxes}.
+For example, two built-in neutrino factory fluxes are available: $\mu^+$-decay ({\tt @builtin = 1}) and $\mu^-$-decay ({\tt @builtin = 2}).
 In these cases, the muon energy (enery of the parent particle) has to be specified together with the number of useful decays  per year.
 Thus, an example to set up a neutrino factory flux is
-\index{aedl}{{\tt flux}!builtin@{\tt "@builtin}}
-\index{aedl}{{\tt flux}!parent energy@{\tt "@parent\_energy}}
-\index{aedl}{{\tt flux}!stored muons@{\tt "@stored\_muons}}
-\begin{quote}
-{\tt flux(\#mu\_plus)<\\
+\index{aedl}{{\tt nuflux}!builtin@{\tt "@builtin}}
+\index{aedl}{{\tt nuflux}!parent energy@{\tt "@parent\_energy}}
+\index{aedl}{{\tt nuflux}!stored muons@{\tt "@stored\_muons}}
+\index{aedl}{{\tt nuflux}!end point@{\tt "@end\_point}}
+\index{aedl}{{\tt nuflux}!stored ions@{\tt "@stored\_ions}}
+\index{aedl}{{\tt nuflux}!gamma@{\tt "@gamma}}
+\begin{quote}
+{\tt nuflux(\#mu\_plus)<\\
 \tb  @builtin = 1\\
@@
 %
-\index{aedl}{{\tt flux}!power@{\tt "@power}}
-\index{aedl}{{\tt flux}!norm@{\tt "@norm}}
-\index{aedl}{{\tt flux}!flux file@{\tt "@flux\_file}}
+\index{aedl}{{\tt nuflux}!power@{\tt "@power}}
+\index{aedl}{{\tt nuflux}!norm@{\tt "@norm}}
+\index{aedl}{{\tt nuflux}!flux file@{\tt "@flux\_file}}
 For a user-defined flux, one has to give it the file name:
 \begin{quote}
-{\tt flux(\#user)<}\\
+{\tt nuflux(\#user)<}\\
 {\tt \tb @flux\_file = "user\_file\_1.dat"\\
@@
 Usually this adjustment of {\tt @norm} is performed by comparison with known energy spectra for a specific experiment.
+For more details on the flux definition, see appendix on page~\pageref{app:flux}.
 
@@
 with {\tt \#} and end with the linefeed character `\verb+\n+', they are
-not counted as a line and their content is discarded. This comments
-are useful to provide meta information about the fluxes like units
+not counted as a line and their content is discarded. These comments
+are useful to provide meta information about the fluxes such as units
 or the origin of the information. This is also the default method
-to point the user to the references he/she should to cite when using
+to point the user to the references one should to cite when using
 a particular flux.\index{norm}{Flux!file!comments in}
@@
 \end{tabular}
-\caption{\label{tab:profiletypes} Different matter density profiles which can be used with \GLOBES .}
+\mycaption{\label{tab:profiletypes} Different matter density profiles which can be used with \GLOBES .}
 \end{table}
@@
 \end{center}
-\caption{\label{fig:energies} The different evaluation levels for the energy smearing in \GLOBES .}
+\mycaption{\label{fig:energies} The different evaluation levels for the energy smearing in \GLOBES .}
 \end{figure}
@@
 \end{tabular}
-\caption[Table of error dimensions]{\label{tab:error_dim}
+\mycaption[Table of error dimensions]{\label{tab:error_dim}
 Possible values of the error dimensions variable in \GLOBES\ and their meaning. If a parameter is designated with $+$, it will

### 2007-01-10 Documentation/part2.tex
*** empty log message ***

@@
 has to be specified.
-The simplest matter density profile is a constant matter density profile with the average matter density from the PREM~\cite{Stacey} onion shell model of the earth\index{norm}{PREM| \see{Matter density}}
-\index{norm}{Matter density!of the earth}:
+The simplest matter density profile is a constant matter density profile with the average matter density from the PREM~\cite{Stacey} onion shell model of the earth:\index{norm}{PREM| \see{Matter density}}
+\index{norm}{Matter density!of the earth}
 \begin{quote}
@@
 Usually it is an excellent approximation to treat backgrounds as if they
-were not affected by oscillations\footnote{In case, the backgrounds have
-a sizeable dependence on the oscillation parameters they carry information
-on the oscillation parameters and are therefore more like a signal.}.
+were not affected by oscillations\footnote{In the case that the backgrounds have
+a sizeable dependence on the oscillation parameters, they carry information
+on the oscillation parameters, and therefore behave more like a signal.}.
 
@@
 This usually involves a binning of the simulated events in the 
-reconstructed energy $E'$. Therefore, one simplify \eq~(\ref{eq:events_bin}) by
+reconstructed energy $E'$. Therefore, one can simplify \eq~(\ref{eq:events_bin}) by
 \begin{equation}
@@
 \end{quote}
-for equidistant sampling points. If no values are given for these variables
-they are assumed to be equal to their corresponding counter part at bin level,
-\ie\ \verb+$sampling_points = $bins+, \verb+$sampling_min = $emin+ and
+for equidistant sampling points. If no values are given for these variables,
+they are assumed to be equal to their corresponding counter parts at the bin level,
+\ie , \verb+$sampling_points = $bins+, \verb+$sampling_min = $emin+ and
 \verb+$sampling_max = $emax+.
@@
 The choice of the sampling point configuration strongly depends on the 
-experiment and required accuracy. Ideally the integrand 
-of~\eq~\ref{eq:simple_int} is zero outside the sampling range, if this cannot
-be achieved it is usually sufficient that the sampling range is by about
-three times the energy resolution\footnote{evaluated at {\tt \$emin} and
-{\tt \$emax} respectively} larger than the bin range. 
-The spacing of the sampling points should be somewhat smaller (a factor 
-$\simeq 2$ usually is more than ebough) than the 
-finest details of the integrand.
+experiment and required accuracy. Ideally, the integrand 
+of~\eq~\ref{eq:simple_int} is zero outside the sampling range. If this cannot
+be achieved, it is usually sufficient that the sampling range is by about
+three times the energy resolution (evaluated at {\tt \$emin} and
+{\tt \$emax}, respectively) larger than the bin range. 
+The spacing of the sampling points should be somewhat smaller than the 
+finest details of the integrand (a factor 
+$\simeq 2$ usually is more than enough).
 
@@
 In the following subsections we will define the energy resolution function.
-All energy resolution functions are defined within an {\tt energy} environment and can be refered to by {\tt \#name}.
+All energy resolution functions are defined within an {\tt energy} environment and can be refered to by {\tt \#name}:
 \begin{quote}
@@
 
-In the actual implementation of the algorithm,  the sum in \eq~(\ref{eq:algo_one}) is only computed for the $E_j$'s where $K(E_j)$ 
-is above a certain threshold, which is by default $10^{-7}$. 
+In the actual implementation of the algorithm,  the sum in \eq~(\ref{eq:algo_one}) is only evaluated
+ for the $E_j$'s with $K(E_j)$ 
+above a certain threshold, which is, by default, $10^{-7}$. 
 This threshold is defined at the compiling time.
@@
 \end{equation} 
-with an analogous definition for the background events. Here, $a$ is the``nuisance'' parameter, which will be minimized over later.
+with an analogous definition for the background events. Here, $a$ is the ``nuisance'' parameter, which will be minimized over later.
 
@@
 is more accurate, but it requires a careful choice of parameters. 
-The first option (method ``T'') is
+The first option (method ``T'') is (JK: IS THAT CORRECT? AND ALSO THE NEXT EQUATION? (THERE HAVE BEEN
+SOME CHANGES) [WW])
 \begin{equation}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{table}[t!]
+\begin{center}
+\begin{tabular}[h]{|lccccccl|}
+\hline
+Systematics function &$a$&$b$&$c$&$d$&Tilt/Calibr.& Dim. & Remarks\\
+\hline
+\hline
+\multicolumn{8}{|l|}{{\bf Standard systematics:}} \\
+{\tt chiSpectrumTilt} &+&+&+&+&T&0&Systematics with tilt\\
+{\tt chiNoSysSpectrum} &-&-&-&-&-&2&No systematics\\
+{\tt chiTotalRatesTilt} &+&+&+&+&T&4&Total rates\\
+{\tt chiSpectrumOnly} & $\infty$ &-& $\infty$ &-&-&7&Spectrum only\\
+{\tt chiNoSysTotalRates} &-&-&-&-&-&8&Total rates, no syst.\\
+{\tt chiSpectrumCalib} &+&+&+&+&C&9&ystematics with calibr.\\
+\hline\multicolumn{8}{|l|}{{\bf User-defined systematics:}} \\
+{\tt chiZero} & - & - & - & - & - & n/a & Passive rule\\
+ & & & & & & & ($\chi^2$ returns zero) \\
+Any other name & \multicolumn{5}{c}{User-defined behavior} & n/a & User-defined syst. \\
+\hline
+\end{tabular}
+\mycaption[Table of error dimensions]{\label{tab:error_dim}
+Possible systematics $\chi^2$ functions  in \GLOBES\ and their meaning. 
+If a parameter is designated with $+$, it will
+be marginalized over, and therefore the corresponding error needs to 
+have a non-zero value. If the cases with ``total rates'' in the remarks, 
+the summation over the bins is performed \emph{before} computing 
+the $\chi^2$, \ie, no spectral information is used. 
+The function {\tt chiSpectrumOnly} leaves the 
+normalization free ($\sigma_a=\sigma_c=\infty$), and therefore 
+only the spectral information is used. As a consequence,
+the settings for the normalization error will be ignored (designated 
+with the symbol $\infty$). In addition, the corresponding error
+dimension from earlier versions of \GLOBES\ is shown in the column ``Dim.''.
+ }
+\end{center}
+\index{aedl}{{\tt rule}!errordim@{\tt "@errordim}} 
+\index{norm}{Systematics function} 
+\end{table} 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 Eventually, the total event rate $x_i$ in a bin $i$ is given by
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{center}
-\begin{table}[t!]
-\begin{center}
-\begin{tabular}[h]{|c|cccc|c|l|}
-\hline
-Error dimension&$a$&$b$&$c$&$d$&Tilt/Calibration&Remarks\\
-\hline
-\hline
-0&+&+&+&+&T&Systematics with tilt\\
-2&-&-&-&-&-&No systematics\\
-4&+&+&+&+&T&Total rates\\
-7& $\infty$ &-& $\infty$ &-&-&Spectrum only\\
-8&-&-&-&-&-&Total rates, no systematics\\
-9&+&+&+&+&C&Systematics with calibration\\
-20&?&?&?&?&?&User-defined systematics\\
-\hline
-\end{tabular}
-\mycaption[Table of error dimensions]{\label{tab:error_dim}
-Possible values of the error dimensions variable in \GLOBES\ and their meaning. If a parameter is designated with $+$, it will
-be marginalized over, and therefore the corresponding error needs to 
-have a non-zero value. If the cases with ``total rates'' in the remarks, 
-the summation over the bins is performed \emph{before} computing 
-the $\chi^2$, \ie, no spectral information is used. 
-The error dimension~7 ``spectrum only'' leaves the 
-normalization free ($\sigma_a=\sigma_c=\infty$), and therefore 
-only the spectral information is used. As a consequence,
-the settings for the normalization error will be ignored (designated 
-with the symbol $\infty$).
-
- }
-\end{center}
-\index{aedl}{{\tt rule}!errordim@{\tt "@errordim}} 
-\end{table} 
-\end{center}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 Eventually, a rule looks like
@@
 \index{aedl}{{\tt rule}|)}
+
+% THIS PART HAS GAINED IMPORTANCE AND THEREFORE BEEN EXPLAINED ALREADY IN THE LAST CHAPTER!
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Version control in \AEDL\ files}
-\label{sec:aedl_versioning}
-\index{norm}{Version control}
-
-In order to avoid problems which come from different
-versions of \GLOBES\ and \AEDL\ files, it is possible to use in each 
-\AEDL\ file a version number. For example, it may correspond to the  minimum version number of the \GLOBES\ package with which it works. Set the
-{\tt \$version}\index{aedl}{version@{\tt \$version}} by
-\begin{quote}
-{\tt \$version="1.8.1"}
-\end{quote}
-This information can be accessed by the versioning functions as described in~\Sec~\ref{sec:versioning}.
-\index{norm}{AEDL@\AEDL|)}
+% \section{Version control in \AEDL\ files}
+% \label{sec:aedl_versioning}
+% \index{norm}{Version control}
+% 
+% In order to avoid problems which come from different
+% versions of \GLOBES\ and \AEDL\ files, it is possible to use in each 
+% \AEDL\ file a version number. For example, it may correspond to the  minimum version number of the \GLOBES\ package with which it works. Set the
+% {\tt \$version}\index{aedl}{version@{\tt \$version}} by
+% \begin{quote}
+% {\tt \$version="1.8.1"}
+% \end{quote}
+% This information can be accessed by the versioning functions as described in~\Sec~\ref{sec:versioning}.
+% \index{norm}{AEDL@\AEDL|)}
 

### 2007-01-10 Documentation/part2.tex
*** empty log message ***

@@
 for the whole experiment. Within each rule, the event rates are added, and
-the systematics is considered to be independent of the other rules.
+the systematics is considered to be independent of the other rules (unless user-defined
+systematics specifies a dependence).
 Thus, it is convenient to combine the above defined channels for different
@@
 \label{eq:penalty}
-p_i=\frac{(\zeta_i-\zeta_i^0)^2}{\sigma_{\zeta_i}^2}\,,
-\end{equation}
-where $\zeta_i^0$ denotes the mean and $\sigma_{\zeta_i}$ the standard 
+p_i=\frac{\zeta_i^2}{\sigma_{\zeta_i}^2}\,,
+\end{equation}
+where $\sigma_{\zeta_i}$ is the standard 
 deviation of the corresponding nuisance parameter. We further on also refer
-to the mean as the ``central value'', and to the standard deviation as
-the ``error''. The latter corresponds to the actual systematical uncertainty. 
+to the  standard deviation as
+the ``error'', it corresponds to the actual systematical uncertainty.
+Note that the central values of all penalties are zero in \GLOBES\ 3.0 and higher. 
 The resulting $\chi^2$ is then minimized with respect to all nuisance 
@@
 \begin{equation}
-s_i(a):=a\cdot s_i
+s_i(a):=(1+a)\cdot s_i
 \end{equation} 
@@
 only sense in combination with a large enough normalization error, since
-the tilt also affects the normalization.
+the tilt may also affect the normalization. 
 %
@@
 {\tt chiSpectrumTilt} &+&+&+&+&T&0&Systematics with tilt\\
-{\tt chiNoSysSpectrum} &-&-&-&-&-&2&No systematics\\
+{\tt chiNoSysSpectrum} &-&-&-&-&-&2&No systematics, but\\
+ & & & & & & & spectral information \\
 {\tt chiTotalRatesTilt} &+&+&+&+&T&4&Total rates\\
@@
 {\tt chiNoSysTotalRates} &-&-&-&-&-&8&Total rates, no syst.\\
-{\tt chiSpectrumCalib} &+&+&+&+&C&9&ystematics with calibr.\\
+{\tt chiSpectrumCalib} &+&+&+&+&C&9&Systematics with calibr.\\
 \hline\multicolumn{8}{|l|}{{\bf User-defined systematics:}} \\
@@
 \end{center}
-\index{aedl}{{\tt rule}!errordim@{\tt "@errordim}} 
+%\index{aedl}{{\tt rule}!errordim@{\tt "@errordim}} 
 \index{norm}{Systematics function} 
@@
 systematical uncertainties and are the nuisance parameters.
-Each of the four parameters has a central value and systematical error.
-The central values for all of
-the four parameters have to be always defined. They are called 
-signal normalization ($a$), signal tilt/calibration ($b$), 
+Each of the four parameters has a corresponding systematical error.
+% The central values for all of
+% the four parameters have to be always defined for built-in systematics. 
+They are called  signal normalization ($a$), signal tilt/calibration ($b$), 
 background  normalization ($c$) and
-background tilt/calibration ($d$). The default values are
-\begin{equation}
-a=1\,,\quad b=0\,,\quad c={\tt not \, \, assigned}\,,\quad d=0\,.
-\end{equation}
-Thus, for the background normalization $c$, the value has to be specified 
-in \emph{either} case. The values for the normalization and the 
+background tilt/calibration ($d$). Their default (central) values are zero.
+% \begin{equation}
+% a=1\,,\quad b=0\,,\quad c={\tt not \, \, assigned}\,,\quad d=0\,.
+% \end{equation}
+% Thus, for the background normalization $c$, the value has to be specified 
+% in \emph{either} case. 
+The errors for the normalization and the 
 values of tilt/calibration are always regarded
-as a pairs, \ie, they are given in the form {\tt normalization : tilt}. The errors are treated in the same way. For example, we have
+as a pairs, \ie, they are given in the form {\tt normalization : tilt}.
+For example, we have\footnote{The old parameter {\tt @backgroundcenter} should not be used anymore.
+A background normalization center of 1.0 will be
+interpreted as zero central value. IS THAT CORRECT [WW]?}
 \index{aedl}{{\tt rule}!signal error@{\tt "@signalerror}}
 \index{aedl}{{\tt rule}!background error@{\tt "@backgrounderror}}
-\index{aedl}{{\tt rule}!background center@{\tt "@backgroundcenter}}
+%\index{aedl}{{\tt rule}!background center@{\tt "@backgroundcenter}}
 \begin{quote}
@@
 \tb @signalerror =       0.001  :       0.01\\
-\tb @backgroundcenter =  0.1 :       0.0\\
+%\tb @backgroundcenter =  0.1 :       0.0\\
 \tb @backgrounderror =   0.001 :       0.01
@@
 \end{quote}
-There is no {\tt @signalcenter} in this definition, 
-since by default the central value for the
-signal normalization is $1$ and the central value for the tilt/calibration 
-is $0$.
-
+% There is no {\tt @signalcenter} in this definition, 
+% since by default the central value for the
+% signal normalization is $1$ and the central value for the tilt/calibration 
+% is $0$.  
+%
 The user has the possibility to choose the set $\{\zeta_i\}$ of nuisance 
 parameters which are minimized over. This choice is specified with the 
-error dimension variable\index{norm}{Error dimension}, and the different 
-possibilities are shown in
-\Tab~\ref{tab:error_dim}. Since the error dimension defines the treatement
-of systematics it is useful to have define a matched pair of error dimensions
-for each rule, where on value describes how the event rate is computed
-for \emph{no} systematics and the other one is with systematics 
-(see also~\Sec~\ref{sec:systematics}).
-The error dimensions for the case of no systematics is set for each rule 
-by the value of {\tt @errordim\_sys\_off}\index{aedl}{{\tt rule}!errordim 
-sys off@{\tt "@errordim\_sys\_off}}, whereas the error dimension
-for systematics on is given by {\tt @errordim\_sys\_on}
-\index{aedl}{{\tt rule}!errordim sys on@{\tt "@errordim\_sys\_on}}. Thus,
- for example,
-the complete error dimension definition could look like  
-\begin{quote}
-{\tt 
-\tb @errordim\_sys\_off = 2\\
-\tb @errordim\_sys\_on = 0
-}
-\end{quote}
-If none of the built-in algorithms listed in \Tab~\ref{tab:error_dim} is
-suitable for a certain setup, the user can override them by setting
-{\tt @errordim\_sys\_on} to 20 and using the \GLB{SetUserChi}
-function to specify a specialized function for the calculation
-of $\chi^2$. This function will then be called by the \GLOBES\ minimizers.
-For details, see \Sec~\ref{sec:userchi}.
-
-Eventually, a rule looks like
+systematics functions {\tt sys\_on\_function} and {\tt sys\_off\_function}
+corresponding two systematics modes ``systematics on'' and ``systematics off''.\footnote{In earlier
+versions before \GLOBES\ 3.0 the ``error dimension'' was used. The corresponding
+parameters {\tt errordim\_sys\_on} and {\tt errordim\_sys\_off} are still supported
+but should not be used anymore.}
+\index{aedl}{{\tt rule}!sys on function@{\tt "@sys\_on\_function}}
+\index{aedl}{{\tt rule}!sys off function@{\tt "@sys\_off\_function}}
+\index{norm}{Systematics function}
+The different  possibilities are shown in
+\Tab~\ref{tab:error_dim}. Since the dual systematics modes define the experiment
+behavior for systematics on and off, it is useful to have a matching pair of 
+systematics functions for each rule 
+(see also~\Sec~\ref{sec:systematics}).. The signal and background errors specified
+by {\tt @signalerror} and  {\tt @backgrounderror} will then be used, if applicable.
+For example, one may define 
+\begin{quote}
+{\tt
+\tb @signalerror =       0.001  :       0.01\\
+\tb @backgrounderror =   0.001 :       0.01 \\
+\tb @sys\_on\_function = "chiSpectrumTilt"  \\
+\tb @sys\_off\_function = "chiNoSysSpectrum"  
+}
+\end{quote}
+
+For user-defined systematics \index{norm}{User-defined systematics}
+(see, \Sec~\ref{sec:userchi}), one can use arbitrary names
+for the systematics functions which are not pre-defined. In this case,
+one would specify the systematical errors in lists, such as
+\index{aedl}{{\tt rule}!sys on errors@{\tt "@sys\_on\_errors}}
+\index{aedl}{{\tt rule}!sys off errors@{\tt "@sys\_off\_errors}}
+\begin{quote}
+{\tt
+\tb @sys\_on\_function = "chiMySystematics"  \\
+\tb @sys\_on\_errors = \{ 0.2, 0.3, 0.5 \}~// Uses three syst. errors \\
+\tb @sys\_off\_function = "chiNoSysSpectrum"  \\
+\tb @sys\_off\_errors = \{ \}
+}
+\end{quote}
+WHICH SYSTEMATICAL ERRORS WILL A FUNCTION USE IF THERE IS A MIXED DECLARATION,
+SUCH AS WITH SYS\_ON\_ERRORS IN SYS ON AND SIGNALERROR/BACKGROUNDERROR in SYS OFF?
+DOES THAT WORK? [WW]
+Note that here the systematics on and systematics off models can be used 
+as fully functional different systematics modes with different systematical
+errors. The interpretation of the systematical errors in {\tt @sys\_on\_errors} and
+{\tt @sys\_off\_errors} is left to the application software code and the user. 
+In addition, the application software has to register the user-defined systematics $\chi^2$
+function with \GLB{DefineChiFunction} and uniquely identify the name
+given by {\tt @sys\_on\_function} or {\tt @sys\_off\_function} to avoid
+confusing different systematics routines. In addition, it is possible to define
+a rule with passive systematics using {\tt glbChiZero}. Such passive systematics
+is not directly used or called, but the corresponding rate vectors will be calculated
+and provided for indirect access. For example, a reactor experiment with correlated
+systematics between near and far detectors may define the user-defined systematics
+{\tt chiReactor} for the far detector and use {\tt chiZero} for the near detector.
+The routine assigned to the far detector will then perform the $\chi^2$ calculation,
+which, of course, also involves the rates of the near detector. However, defining
+{\tt chiReactor} in both the near and far detectors would result in a double call
+of the $\chi^2$ function, \ie, the program will be a factor two slower and the
+resulting $\chi^2$ will be a factor two too large.
+
+Eventually, a complete rule may look like this:
 \begin{quote}
@@
 \tb @signalerror =       0.001  :       0.01\\
-\tb @backgroundcenter =  0.1 :       0.0\\
 \tb @backgrounderror =   0.001 :       0.01\\
-\tb @errordim\_sys\_off = 2\\
-\tb @errordim\_sys\_on = 0\\
-\tb @energy\_window = 4.0 : 50.0\\ 
+\tb @sys\_on\_function = "chiSpectrumTilt"  \\
+\tb @sys\_off\_function = "chiNoSysSpectrum"  \\
+\tb @energy\_window = 4.0 : 50.0
 >}
@@
 \end{quote}
+Note that in the case of additional non-standard parameters, these will just to 
+be included at the end of these lists.
 Furthermore it is possible to switch off oscillations with the {\tt -N} option and to switch them on again with {\tt -O} (the default).
@@
 the definition of a rule also contains so-called ``coefficients'', it is
-possible to switch them off with {\tt -i}. This options also deactivates
-any setting of {\tt @backgroundcenter}.
+possible to switch them off with {\tt -i}. 
+% This options also deactivates
+% any setting of {\tt @backgroundcenter}.
 
@@
 parse any files with {\tt globes} which contain such undefined variables. 
-Hence, there is the possibility to define \AEDL\ variables by using the define option {\tt -D}. A call such as 
-\begin{quote}
-{\tt
-globes -DBASELINE=3000 ...
-}
-\end{quote}
-would define the \AEDL\ variable {\tt BASELINE} to be $3000$.
+Hence, there is the possibility to define \AEDL\ variables by using the define option {\tt -D}. The example
+\begin{quote}
+{\tt
+globes -DBASELINE=3000 -D\%BLUE=$\backslash$\{8, 15$\backslash$\} ...
+}
+\end{quote}
+defines the \AEDL\ variable {\tt BASELINE} to be $3000$ and the \AEDL\ variable list
+{\tt \%BLUE} to be \{8,15\} (please note the syntax for the brackets!).
 

### 2007-01-10 Documentation/appendix.tex
PH -- updated the installation instructions in the appendix

@@
 
-{\bf NEEDS TO BE UPDATED [WW] -> PH? ADD SECTION ON CONDOR/CLUSTER SUPPORT?}
+\section*{Prerequisites for installation of \GLOBES}
 
-The installation of \GLOBES\ is highly automated and there should not be
-any problems on a decently up-to-date GNU/Linux system. The installation
-has, however, only been tested on a limited number of platforms. The 
-current version has been tested for SuSE, Red Hat, Fedora Core and
-a 64bit version of SuSE. It should also compile and work on Mac OS. We 
-would appreciate to know your experiences with the installation on different platforms.
-Please send an e-mail to <\bugs>.
-
-\section*{Prerequisites for the installation of \GLOBES}
-\index{norm}{Installation!prerequisites}
-
-Besides the usual things such as a working \verb`libc`, you need to have
+Besides the usual things like a working libc you need to have
  \begin{itemize}
@@
 \end{itemize}
-The library \verb+libglobes+ should in principle compile with any
-ANSI C/C++ compiler, but the \verb+globes+ binary uses the 
-\verb+argp+ facility of \verb+glibc+ to parse its command line options.
-However, on platforms where \verb+argp+ is lacking \GLOBES\ has replacement code, 
-thus it should also work there.
+The library \verb+libglobes+ should in principle compile with any C/C++
+compiler but the \verb+globes+ binary uses the \verb+argp+ facility of \verb+glibc+
+to parse its command line options. However, on platforms where \verb+argp+
+is lacking \GLOBES\ has replacement code, thus it should also work
+there. \GLOBES\ is, however, using the C99 standard in order to handle
+complex numbers, but that is the only feature of C99 used.
 
-All those libraries are also available as rpm's from the various
-distributors of GNU/Linux, see their web sites for downloads. Chances are 
-that \verb+gcc+ and GSL are already part of your installation.  For building \GLOBES\ 
-from source, however, not only working libaries for above packages are 
+GSL is also available as rpm+s from the various distributors of
+GNU/Linux, see their web sites for downloads. Chances are that gcc and
+GSL are already part of your installation.  For building \GLOBES\ from
+source, however, not only working libraries for above packages are
 needed but also the headers, especially for GSL. Depending on your
-installation, eg. on RedHat/Fedora, of GSL this may require to additionally 
-install a rpm-package named \verb+gsl-devel+. If GSL has been installed from the 
-tar-ball as provided by \verb`gnu.org` no problems should occur. Furthmore you need 
-a working \verb+make+ to build and install \GLOBES.
+installation, eg. on RedHat/Fedora, of GSL this may require to
+additionally install a rpm-package named \verb+gsl-devel+. If GSL has been
+installed from the tar-ball as provided by gnu.org no problems should
+occur. Furthermore you need a working \verb+make+ to build and install
+\GLOBES.
 
@@
 
-\GLOBES\ follows the standard GNU installation procedure.  To compile \GLOBES\
-you will need gcc.  After unpacking the distribution
-the Makefiles can be prepared using the configure command,
-\begin{quote}
-{\tt
-./configure
-}
-\end{quote}
-You can then build the library by typing,
-\begin{quote}
-{\tt
-make
-}
-\end{quote}
+\GLOBES\ follows the standard GNU installation procedure.  To compile
+\GLOBES\ you will need an ANSI C-compiler.  After unpacking the
+distribution the Makefiles can be prepared using the configure
+command,\\
+\verb+  ./configure+\\
+You can then build the library by typing,\\
+\verb+  make+\\
 A shared  version of the library will be compiled by
@@
 
-The library can be installed using the command,
-\begin{quote}
-{\tt
-make install
-}
-\end{quote}
-The default install directory prefix is \verb+/usr/local+.  Consult the
-"Further Information" section below for instructions on installing the
-library in another location or changing other default compilation
-options.
-
-The install target also will install a program with name \verb+globes+ to
-\verb+$prefix/bin+ and the files in the \verb+data+ directory of the tar-ball
-to \verb+$prefix/share/globes+.
-\index{norm}{glb files@{\tt glb}-files!installation}
-
-If you are not using \verb+make install+ you will find the static libary
-at \verb+source/.lib/libglobes.a+ which you can copy to any destination.
-However keep in mind that the linking command will be somewhat different,
-\ie\ you have to specify all the dynamically linked objects besides libglobes.
+The libraries and modules can be installed using the command,\\
+\verb+  make install+\\
+The install target also will install a program with name \verb+globes+
+ to \verb+/usr/local/bin+
 
@@
 
-Moreover a config-script called \verb+globes-config+ will be installed. This
-script displays all information necessary to link any program with GLoBES.
-For building static libaries and linking against them see the corresponding
-section of this file.
+Moreover a config-script called \verb+globes-config+ will be
+installed. This script displays all information necessary to link any
+program with \GLOBES.  For building static libraries and linking against
+them see the corresponding section of this file.
 
@@
    The \verb+configure+ shell script attempts to guess correct values for
-various system-dependent variables used during compilation.  It uses
-these values to create a \verb+Makefile+ in each directory of the package.
-Finally, it creates a shell script \verb+config.status+ that
-you can run in the future to recreate the current configuration, a file
-\verb+config.cache+ that saves the results of its tests to speed up
-reconfiguring, and a file \verb+config.log+ containing compiler output
-(useful mainly for debugging \verb+configure+).
+various system-dependent variables used during compilation.  It
+uses those values to create a \verb+Makefile+ in each directory of the
+package.  It may also create one or more \verb+.h+ files containing
+system-dependent definitions.  Finally, it creates a shell script
+\verb+config.status+ that you can run in the future to recreate the
+current configuration, a file \verb+config.cache+ that saves the results
+of its tests to speed up reconfiguring, and a file \verb+config.log+
+containing compiler output (useful mainly for debugging
+\verb+configure+).
 
@@
 to figure out how \verb+configure+ could check whether to do them, and mail
-diffs or instructions to \bugs\ so they can
+diffs or instructions to the address given in the \verb+README+ so they can
 be considered for the next release.  If at some point \verb+config.cache+
@@
    The file \verb+configure.in+ is used to create \verb+configure+ by a program
-called \verb+autoconf+.  You only need \verb+configure.in+ 
-if you want to change it or regenerate \verb+configure+ using a newer 
-version of \verb+autoconf+.
+called \verb+autoconf+.  You only need \verb+configure.in+ if you want to change
+it or regenerate \verb+configure+ using a newer version of \verb+autoconf+.
 
@@
      Running \verb+configure+ takes awhile.  While running, it prints some
-     messages telling which features it is checking for. It also prints
-     a reminder for things to do after installation.
-
+     messages telling which features it is checking for.
 \item Type \verb+make+ to compile the package.
@@
 
-\item Since you've installed a library don't forget to run \verb+ldconfig+ !
+\item Since you have installed a library don't forget to run \verb+ldconfig+!
 \end{enumerate}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section*{Installation without root privilege}
-\index{norm}{Installation!w/o root privilege}
-\label{inst_noroot}
 
-Install \GLOBES\ to a directory of your choice \verb+GLB_DIR+. This is done by
+Install \GLOBES\ to a directory of your choice \verb+GLB_DIR+. This is done by\\
+\verb+  configure --prefix=GLB_DIR+\\ 
+and then follow the usual
+installation guide. The only remaining problem is that you
+have to tell the compiler where to find the header files, and
+the linker where to find the library. Furthermore you have to
+make sure that the shared object files are found during
+execution. Running \verb+configure+ also produces a \verb+Makefile+ in
+the examples subdirectory which can serve as a template for
+the compilation and linking process, since all necessary flags
+are correctly filled in. Another solution is to set the
+environment variable \verb+LD_RUN_PATH+ during linking to
+\verb+GLB_DIR/lib/+. Best thing is to add this to your shell dot-file
+(e.g. \verb+.bashrc+). Then you can use: A typical compiler command
+like\\
+\verb+  gcc -c my_program.c -IGLB_DIR/include/+\\
+and a typical linker command like\\
+\verb+  gcc my_program.o -lglobes -LGLB_DIR/lib/ -o my_executable+\\
+More information on this issue can be obtained by having a look into
+the output of make install.
+
+{\bf CAVEAT}: It is in principle possible to have many installations on one
+machine, especially the situation of having an installation by root
+and by a user at the same time might occur. However it is strictly
+warned against this possibility since it is \emph{extremely} likely to
+create some versioning problem at some time!
+
+\section*{Building and Using static versions of \GLOBES}
+
+Under certain circumstances it may be useful to use a static version of
+libglobes or any of the binaries, e.g. when running on a cluster.
+
+The \verb+configure+ script accepts the option \verb+--disable-shared+, in
+which case only static objects are built, i.e. only a static
+version of libglobes. In case your system does not support shared
+libraries the \verb+configure+ script recognizes this. If you give no
+options to \verb+configure+, both shared and static versions are built
+and will be installed. All binaries, however, will use dynamic
+linking. If you want to build static binaries, use
+LDFLAGS=+-all-static+ for building them.
+
+Sometimes it is convenient, eg. for debugging purposes, to have a
+statically linked version of a program using \GLOBES\, which is easiest
+achieved by just linking with \verb+libglobes.a+. If you need a completely
+statically linked version, please, have a look at the Makefile in the
+\verb+examples+ directory.\\
+\verb+  make example-static+\\ 
+produces a statically linked program that should in principle run on
+most Linuxes.  It should be straightforward to adapt this example to
+your needs.
+
+All these options rely on a working \verb+gcc+ installation. It seems that
+\verb+gcc 3.x+ is broken in subtle way which makes it necessary to add a
+symbolic link in the gcc library directory. The diagnostic for this
+requirement is that building static programs fails with the error
+message \verb+cannot find -lgcc_s+. In those cases, find \verb+libgcc.a+ and add
+a symbolic link in the same directory where you found it (this
+requires probably root privileges)\\
+\verb+  ln -s libgcc.a libgcc_s.a+
+
+If you can not write to this directory just use the following work
+around. Add the same link as above to the directory where you
+installed \GLOBES\ into\\
+\verb+  cd prefix/lib+\\       
+\verb+  ln -s path_to_libgcc.a/libgcc.a libgcc_s.a+\\
+and then change back into the \verb+examples+ directory and type\\
+\verb+  make LDFLAGS=-Lprefix/lib example-static+\\
+and you are done.
+
+\section*{\GLOBES\ and Condor}
+\index{norm}{Condor|(}
 \begin{quote}
-\verb+configure --prefix=GLB_DIR+
+  Condor is a specialized workload management system for
+  compute-intensive jobs. Like other full-featured batch systems,
+  Condor provides a job queuing mechanism, scheduling policy,
+  priority scheme, resource monitoring, and resource management.
 \end{quote}
-and then follow the usual installation guide. The only remaining problem
-is that you have to tell the compiler where to find the header files, and 
-the linker where to find the library. Furthermore you have to make sure 
-that the shared object files are found during execution. Running 
-\verb+configure+
-also produces a \verb+Makefile+ in the \verb+examples+ 
-subdirectory which can serve as
-a template for the compilation and linking process, since all necessary
-flags are correctly filled in.
 
-Another solution is to set the environment variable \verb+LD_RUN_PATH+
-during linking to \verb+GLB_DIR/lib+ . Best thing is to add this to your
-shell dot-file (e.g. \verb+.bashrc+). Then you can use:
-A typical compiler command like
-\begin{quote}
-\verb+gcc -c my_program.c -IGLB_DIR/include/+
-\end{quote}
-and a typical linker command like
-\begin{quote}
-\verb#g++ my_program.o -lglobes -LGLB_DIR/lib/ -o my_executable#
-\end{quote}
-More information on this issue can be obtained by having a look 
-into the mentioned  \verb+Makefile+ in \verb+examples+.
+A Condor (\verb+www.cs.wisc.edu/condor/+) cluster is very well suited
+to run large \GLOBES-based computation. The nature of the problems
+addressed with \GLOBES\ is such that one typically ends up with a so
+called 'embarrassingly parallel' program. That means, that one repeats
+the same task $N$ times, where each execution is independent of the
+other $N-1$. Therefore, this execution should become $M$ times faster
+if one uses $M$ processors. For this class of problems running on a
+dedicated cluster will not improve performance (but may reduce latency
+and such).
 
-\begin{itemize}
-\item[CAVEAT] It is in principle possible to have many installations 
-on one machine.
-Especially the situation of having an installation by root and by a user at the same time might occur.
-However, it is strictly warned against this possibility, since it is \emph{extremely} likely to create some versioning problem at some time!
-\end{itemize}
+In order to fully exploit the functionality offered by Condor one
+should submit the jobs into the so called 'standard universe'. To do
+this, it is necessary to re-link the application with the
+Condor-library (this assumes that Condor is installed)\\ 
+\verb+  condor_compile gcc your_object_files -static `globes-config --libs`+\\
+It may be necessary to prefix the call of \verb+globes-config+ with
+the path to it, in case that location is not in \verb+$PATH+.
+\index{norm}{Condor|)}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{LAPACK \& BLAS requirements}
-
-GLoBES uses one single routine out of LAPACK (\verb^www.netlib.org/lapack/^) named
-\verb^zgeev^ to diagonalize the oscillation Hamiltonian in matter. Since LAPACK
-is not part of the standard installation of many distributions and we need
-only one function, \GLOBES\ comes with a convenience version of the required
-routines, which are taken from LAPACK and BLAS (\verb^www.netlib.org/blas^) and have
-been translated to C with \verb^f2c^. In case your system has a working installation
-of BLAS/LAPACK \GLOBES\ will use it. However, if your system is lacking BLAS/
-LAPACK the convenience version will be compiled and linked into \GLOBES. If
-for some reason BLAS/LAPACK is installed on the system and one still wishes
-to use the convience version use \verb^--enable-lapack-convenience^. Using Fortran
-code which has been translated to C with \verb^f2c^ requires a run-time library
-called \verb^libf2c^ (on some systems it is called \verb^libg2c^). The configury of 
-\GLOBES\ detects the presence of either of the two and takes care of it.
-If none of the two is present, a convience library is used. If
-for some reason \verb^libf2c^ is installed on the system and one still wishes
-to use the convience version use \verb^--enable-libf2c-convenience^.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section*{GSL requirements}
 
-Sometimes the GNU scientific library is not available or is installed in a non-standard
-location. This situation can arise in an installation without root privileges. In this 
-case one can specify \verb^--with-gsl-prefix=path_to_gsl^ as option to the \verb^configure^ script.
-If one wants to use a shared version of \verb^libgsl^ then one has to make sure that the
-linker finds the library at run-time. This can be achieved by setting the environment variable
-\verb^LD_LIBRARY_PATH^ correctly, \ie\ (in bash)
-\begin{quote} 
-\verb^export LD_LIBRARY_PATH=path_to_gsl^
-\end{quote}
-You also can use a static version of GSL by either building \GLOBES\ with \verb^LDFLAG=-all-static^
-or by configuring GSL with \verb^--disable-shared^. In both cases no further actions like setting
-any environment variables is necessary.
+Sometimes the GNU scientific library is not available or is installed
+in a non-standard location. This situation can arise in an
+installation without root privileges. In this case one can specify
+\verb+--with-gsl-prefix=path_to_gsl+ as option to the \verb+configure+ script.
+If one wants to use a shared version of \verb+libgsl+ then one has to make
+sure that the linker find the library at run-time. This can be
+achieved by setting the environment variable \verb+LD_LIBRARY_PATH+
+correctly, i.e. (in bash)\\
+\verb+  export LD_LIBRARY_PATH='path_to_gsl'+\\
+You also can use a static version of GSL by either building \GLOBES\
+with \verb+LDFLAG='-all-static'+ or by configuring GSL with
+\verb+--disable-shared+. In both cases no further actions like setting any
+environment variables is necessary.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section*{Distributions}
 
-\subsection*{Fedora Core 3}
+\subsection*{RedHat (all versions)}
 
-FC3 is shipped with a non working version of LAPACK. Thus to
-make \GLOBES\ work on FC3, configure \GLOBES\ with \verb'--enable-lapack-convenience', which
-replaces the defunct LAPACK with \GLOBES' own version thereof. You can diagnose a 
-non-working LAPACK by compiling one of the examples in the \verb^examples^ directory and 
-trying to run it. It will hang forever in the function \GLB{SetRates()}.
+The standard rpm-based installation of GSL does not provide any header
+files for GSL, which are however needed to compile \GLOBES\. You have to
+install an additional rpm-package called \verb+gsl-devel+. Alternatively
+you can install GSL from a tar-ball and use the \verb+--with-gsl-prefix+
+option to the configure script of \GLOBES.
 
-\subsection*{RedHat -- all versions}
+\section*{Platforms}
 
-The standard rpm-based installation of GSL does not provide any header files for GSL, 
-which are however needed to compile \GLOBES. You have to install an additional rpm-package
-called \verb'gsl-devel'. Altnernatively you can install GSL from a tar-ball and use the 
-verb'--with-gsl-prefix' option to the configure script of \GLOBES.
+\GLOBES\ builds and installs on 64bit Linux systems.
+\GLOBES\ should work on Mac OS.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Advanced topics}
+\subsection*{Windows}
 
-\subsection*{Building and Using static versions of \GLOBES}
+Currently \GLOBES\ is only able to work under Cygwin \verb+www.cygwin.com+.
+Inside Cygwin \GLOBES\ needs to be built with these commands\\
+\verb+  configure+\\
+\verb+  make LDFLAGS=-no-undefined'+
 
-In certain circumstances it may be useful to use a static version of libglobes
-or any of the binaries, \eg\ when running on a cluster.
-
-   The \verb+configure+ script accepts the option \verb+--disable-shared+, in which
-case only static objects are built, i.e. only a static version of libglobes. 
-In case your system does not support shared libraries the \verb+configure+ script 
-recognizes this. If you give no options to \verb+configure+, both shared and static 
-versions are built and will be installed. All binaries, however, will use dynamic 
-linking. If you want to build static binaries, use \verb+LDFLAGS=-all-static+ for building
-them.
-
-   Sometimes it is convenient, eg. for debugging purposes, to have a statically
-linked version of a program using \GLOBES\, which is easiest achieved by just linking
-with \verb+libglobes.a+. If you need a completely statically linked version, please, have
-a look at the Makefile in the \verb+examples+ directory.
-\begin{quote}
-\verb#make example-static#
-\end{quote}
-produces a statically linked program that should in principle run on most Linuxes.
-It should be straightforward to adapt this example to your needs.
-
-All these options rely on \verb+libtool+ and a working \verb`gcc` installation. It seems that \verb`gcc 3.x`
-is broken in subtle way which makes it necessary to add a symbolic link in the \verb`gcc` library
-directory. The diagnostic for this requirement is that building static programs fails with
-the error message \verb+cannot find -lgcc_s+. In those cases, find \verb+libgcc.a+ and add a symbolic 
-link in the same directory where you found it (this requires probably root privileges)
-\begin{quote}
-\verb#ln -s libgcc.a libgcc_s.a#
-\end{quote}
-
-If you can not write to this directory just use the following work around. Add the same link  
-as above to the directory where you installed \GLOBES\ into
-\begin{quote}
-\verb#cd prefix/lib#
-\verb#ln -s path_to_libgcc.a/libgcc.a libgcc_s.a#
-\end{quote}
-and then change back into the \verb+examples+ directory and type
-\begin{quote}
-\verb#make LDFLAGS=-Lprefix/lib example-static#
-\end{quote}
-and you are done.
-
-If you don't have libtool on your system either, then you just can use the libtool script in
-the \GLOBES\ tar-ball instead. Make sure it is in a place where it is found by \verb+make+.
-
-\subsection*{Compilers and Options}
+\section*{Compilers and Options}
 
@@
 a Bourne-compatible shell, you can do that on the command line like
-this
-\begin{quote}
-\verb+CC=c89 CFLAGS=-O2 LIBS=-lposix ./configure+
-\end{quote}
-Or on systems that have the `env' program, you can do it like this
-\begin{quote}
-\verb+env CPPFLAGS=-I/usr/local/include LDFLAGS=-s ./configure+
-\end{quote}
+this\\
+\verb+  CC=c89 CFLAGS=-O2 LIBS=-lposix ./configure+\\
 
-\subsection*{Compiling For Multiple Architectures}
+Or on systems that have the \verb+env+ program, you can do it like this\\
+\verb+  env CPPFLAGS=-I/usr/local/include LDFLAGS=-s ./configure+
+
+\section*{Compiling For Multiple Architectures}
 
@@
 own directory.  To do this, you must use a version of \verb+make+ that
-supports the \verb+VPATH+ variable, such as GNU \verb+make+.  `cd' to the
+supports the \verb+VPATH+ variable, such as GNU \verb+make+.  \verb+cd+ to the
 directory where you want the object files and executables to go and run
 the \verb+configure+ script.  \verb+configure+ automatically checks for the
-source code in the directory that \verb+configure+ is in and in `..'.
+source code in the directory that \verb+configure+ is in and in \verb+..+.
 
@@
 in the source code directory.  After you have installed the package for
-one architecture, use `make distclean' before reconfiguring for another
+one architecture, use \verb+make distclean+ before reconfiguring for another
 architecture.
 
-\subsection*{Installation Names}
+\section*{Installation Names}
 
-By default, \verb+make install+ will install the package's files in
-\verb+/usr/local/bin+, \verb+/usr/local/include+, etc.  You can specify an
+   By default, \verb+make install+ will install the package+s files in
+\verb+/usr/local/bin+, \verb+/usr/local/man+, etc.  You can specify an
 installation prefix other than \verb+/usr/local+ by giving \verb+configure+ the
@@
 
-\subsection*{Specifying the System Type}
+   You can specify separate installation prefixes for
+architecture-specific files and architecture-independent files.  If you
+give \verb+configure+ the option \verb+--exec-prefix=PATH+, the package will use
+PATH as the prefix for installing programs and libraries.
+Documentation and other data files will still use the regular prefix.
+
+   In addition, if you use an unusual directory layout you can give
+options like \verb+--bindir=PATH+ to specify different values for particular
+kinds of files.  Run \verb+configure --help+ for a list of the directories
+you can set and what kinds of files go in them.
+
+   If the package supports it, you can cause programs to be installed
+with an extra prefix or suffix on their names by giving \verb+configure+ the
+option \verb+--program-prefix=PREFIX+ or \verb+--program-suffix=SUFFIX+.
+
+\section*{Optional Features}
+
+   Some packages pay attention to \verb+--enable-FEATURE+ options to
+\verb+configure+, where FEATURE indicates an optional part of the package.
+They may also pay attention to \verb+--with-PACKAGE+ options, where PACKAGE
+is something like \verb+gnu-as+ or \verb+x+ (for the X Window System).  The
+\verb+README+ should mention any \verb+--enable-+ and \verb+--with-+ options that the
+package recognizes.
+
+   For packages that use the X Window System, \verb+configure+ can usually
+find the X include and library files automatically, but if it doesn't,
+you can use the \verb+configure+ options \verb+--x-includes=DIR+ and
+\verb+--x-libraries=DIR+ to specify their locations.
+
+\subsection*{Building a perl extension}
+
+This feature is experimental and your mileage my vary!
+
+This feature allows to build a perl binding of \GLOBES\, i.e. you will
+in the end have a perl module from which you can use \GLOBES\ from within
+any perl program.
+
+If(!) everything works as intended, all you have to do is to provide
+\verb+--enable-perl+ to \verb+configure+ and type \verb+make install+. Now have a
+look at \verb+globes/example.pl+ and you should see how that works in
+principle.
+
+The trick here, is that we use SWIG (\verb+www.swig.org+) to generate
+a wrapper file for \GLOBES. The wrapper file is part of the \GLOBES\
+tar-ball (\verb+globes/globes_perl.c+) and hence you should not need SWIG to
+be installed on your system.
+
+All the tricks employed to get perl extension working should in some
+form be applicable to building other extensions, like python. If
+you want to try that you will need SWIG.
+
+\subsection*{Building RPMs}
+
+This feature is experimental and your mileage my vary!
+
+Many people find binary RPMs useful, therefore we provide an optional
+feature \verb+--enable-rpm-rules+ which should produce all the necessary
+makefile rules for RPM building. To actually build RPMs requires that
+your system is properly setup for that. How you can do that you can
+learn at \verb+http://www.rpm.org+. You then can use \verb+make rpm+, most likely
+you will need to be root to do that (\verb+sudo+ won't work!).
+
+{\bf NOTE} to people packaging \GLOBES\ RPMs: Please, use the provided spec
+file and do include the headers!
+
+\section*{Specifying the System Type}
 
@@
 \verb+--host=TYPE+ option.  TYPE can either be a short name for the system
-type, such as `sun4', or a canonical name with three fields: 
-\verb+CPU-COMPANY-SYSTEM+
+type, such as \verb+sun4+, or a canonical name with three fields\\
+\verb+  CPU-COMPANY-SYSTEM+
 
-See the file \verb+config.sub+ for the possible values of each field.
+See the file \verb+config.sub+ for the possible values of each field.  If
+\verb+config.sub+ isn't included in this package, then this package doesn't
+need to know the host type.
 
@@
 use the \verb+--target=TYPE+ option to select the type of system they will
-produce code for, and the \verb+--build=TYPE+ option to select the type of
+produce code for and the \verb+--build=TYPE+ option to select the type of
 system on which you are compiling the package.
 
-\subsection*{Sharing Defaults}
+\section*{Sharing Defaults}
+
    If you want to set default values for \verb+configure+ scripts to share,
 you can create a site shell script called \verb+config.site+ that gives
-default values for variables like \verb+CC+,  \verb+cache_file+, and 
-\verb+prefix+.
+default values for variables like \verb+CC+, \verb+cache_file+, and \verb+prefix+.
 \verb+configure+ looks for \verb+PREFIX/share/config.site+ if it exists, then

### 2007-01-11 Documentation/Manual.tex
*** empty log message ***

@@
 
-\paragraph{Note:} All examples for application software in C do require
-a C++ compiler to be properly compiled. For pedagogical reasons, variable
-declarations are done at that place where the variable is needed for the
-first time, which is at variance with C syntax but not with C++ syntax.
-That is the only way in which the examples deviate form ISO C. Moreover
-the actual numerical values of the results of the examples may be different
-from the ones in this manual.  (IS THAT STILL SO? [WW])
+% THE FOLLOWING HAS CHANGED: C99, COMPILED WITH GCC; MAYBE ONLY PROBLEMS WITH C89
+%
+% \paragraph{Note:} All examples for application software in C do require
+% a C++ compiler to be properly compiled. For pedagogical reasons, variable
+% declarations are done at that place where the variable is needed for the
+% first time, which is at variance with C syntax but not with C++ syntax.
+% That is the only way in which the examples deviate form ISO C. Moreover
+% the actual numerical values of the results of the examples may be different
+% from the ones in this manual.
 

### 2007-01-11 Documentation/Manual.tex
*** empty log message ***

@@
 
-TO BE UPDATED [WW].
-
 Here we briefly summarize the main changes of the new GLoBES version. For details, please
 refer to the respective parts of the manual. Please note that any new \GLOBES\ version is
-compatible with older versions, \ie, old application software and AEDL files should also run with the new version.
-EXCEPTION: VERSION NUMBER IN AEDL REQUIRES?
-[WW] However, some function names and features will
+compatible with older versions, \ie, old application software and \AEDL\ files should, with minor
+modifications, run with the new version as well.  However, some function names and features will
 evolve during time, which means that outdated features may not be documented anymore.
@@
 
-Here comes a summary of the most important changes in this version for older users of GLoBES.
+Here comes a summary of the most important changes in this version for users of earlier
+versions of \GLOBES .
 
@@
  \item
-  User-defined systematics to simulate experiments with correlated systematics \etc\ (such as
-reactor experiments); see \Sec~?
+  User-defined systematics, which can be used to simulate reactor experiments \etc ; see \Sec~\ref{sec:userchi}
+and~\ref{sec:rules}
  \item
   User-defined priors to include arbitrary external information in the $\chi^2$ before
-marginalization over the oscillation parameters (such as to prevent minimizer from
-running into the wrong octant \etc); see \Sec~?
+marginalization over the oscillation parameters; see \Sec~\ref{sec:userdefined} 
 \item
-  Non-standard physics support; see \Sec~?
+  Non-standard physics support; see \Chapt~\ref{chapt:nsphysics}
 \item
-  Beta beam fluxes available as built-in fluxes; see \Sec~?
+  Beta beam fluxes available as built-in fluxes; see \Sec~\ref{sec:source}
 \item
-  Enhanced support for parallelization, such as Condor; see \Sec~?
+  Enhanced support for parallelization, such as Condor; see, \eg, page~\pageref{sec:condor}
 \item
-  Updated AEDL files, new beta beam files; see \Sec~?
+  Updated AEDL files; see \Tab~\ref{tab:experiments}
 \item
- Faster probability engine
+ New \AEDL\ features, such as the support of lists as variables and an interpolation
+routine; see \Sec~\ref{sec:advaedl}
 \item
-  ...
+ Clean-up of inconsistencies, such as an overall (internal) normalization factor in the flux files;
+see, \eg, page~\pageref{app:flux}
+\item
+  Faster probability engine, easier installation (internal changes)
+\item
+  Experimental feature: Alternative minimizer provided, which is usually faster than
+the standard minimizer; see \Chapt~\ref{chapt:experimental}
 \end{itemize}
@@
 
-Most of the modifications below should not require that old pieces of software be changed. 
-However, the following changes are necessary:
+Most of the modifications should not require that old pieces of software be changed. 
+However, the following changes are relevant:
 \begin{itemize}
@@
 supported by the \GLOBES\ team anymore. You should make sure to keep these files when updating \GLOBES . 
-\item
- ...
 \end{itemize}
-The following changes should not affect older software:
+
+\subsection*{Minor changes}
+
+Here we document the most important changes which should not affect older software:
 \begin{itemize}
 \item
- {\tt glbChi...} functions renamend for consistency
+ Functions and constants renamed for consistency:
+\begin{itemize}
 \item
- {\tt glbChiSetStartingValues} and {\tt glbChiGetStartingValues} renamed into 
-{\tt glbChiSetCentralValues} and {\tt glbChiGetCentralValues}
+ {\tt glbChiTheta} $\rightarrow$ {\tt glbChiTheta13}
 \item
- Systematics concept changed, concept of error dimensions removed
+ {\tt glbChiThetaDelta} $\rightarrow$ {\tt glbChiTheta13Delta}
 \item
- Some functions changed (such as for direct rate access)
+ {\tt glbChiDms} $\rightarrow$ {\tt glbChiDm21}
 \item
- ..
+ {\tt glbChiDm} $\rightarrow$ {\tt glbChiDm31}
+\item
+ {\tt GLB\_DM\_SOL} $\rightarrow$ {\tt GLB\_DM\_21}
+\item
+ {\tt GLB\_DM\_ATM} $\rightarrow$ {\tt GLB\_DM\_31}
+\item
+ {\tt glbSetStartingValues} $\rightarrow$ {\tt glbSetCentralValues}
+\item
+ {\tt glbGetStartingValues} $\rightarrow$ {\tt glbGetCentralValues}
+\item
+ {\tt glbGetProfileData} $\rightarrow$ {\tt glbGetProfileDataInExperiment}
+\end{itemize}
+\item
+ Rate access changed; see \Sec~\ref{sec:event_rates}
+\item
+ Systematics concept changed, concept of error dimensions removed; {\tt @backgroundcenter} removed, central values for all systematics parameters
+now zero; see \Sec~\ref{sec:rules}
 \end{itemize}

### 2007-01-11 Documentation/appendix.tex
*** empty log message ***

@@
 \section*{\GLOBES\ and Condor}
+\label{sec:condor}
 \index{norm}{Condor|(}
@@
 based on the old environment {\tt flux}. For the use of {\tt nuflux},
-replace the factor 5.20034 by unity.
+replace the factor 5.2 by unity.
 
@@
 \begin{eqnarray}
-n &=& 5.20034\times x\times E\times f\times \nonumber \\
+n &=& 5.2\times x\times E\times f\times \nonumber \\
 &&\mathtt{@norm}\times\mathtt{@power}\times\mathtt{@stored\_muons}\times\mathtt{@time}\times\mathtt{\$target\_mass}\times(\mathtt{\$baseline})^{-2} \nonumber
 \end{eqnarray}
-Note that $5.20034$ is a undocumented fudge factor!
+Note that $5.2$ is a undocumented fudge factor!
 
@@
 \begin{eqnarray}
-\mathtt{@norm}=\frac{1}{5.20034}\left(\frac{\mathrm{GeV}}{\Delta
+\mathtt{@norm}=\frac{1}{5.2}\left(\frac{\mathrm{GeV}}{\Delta
       E}\right)
@@
 \begin{equation}
-\frac{\mathcal{L}}{\mathcal{L}_u}=\frac{10\,\mathrm{GeV}\,10^{21}\,\mathrm{pot}\,\mathrm{y}}{10^7\,\mathrm{s}}\times(\mathrm{MW}\,\mathrm{y}^{-1})^{-1}=0.160218\ldots
+\frac{\mathcal{L}}{\mathcal{L}_u}=\frac{10\,\mathrm{GeV}\,10^{21}\,\mathrm{pot}\,\mathrm{y}}{10^7\,\mathrm{s}}\times(\mathrm{MW}\,\mathrm{y}^{-1})^{-1}=0.16\ldots
 \end{equation}
+
+\section*{Moving from {\tt flux} to {\tt nuflux}}
+
+In order to change the older {\tt flux} environment to the
+new {\tt nuflux} (\GLOBES\ 3.0 and higher), replace all user-defined fluxes, such as
+\begin{quote}
+{\tt flux(\#user)<}\\
+{\tt \tb @flux\_file = "user\_file\_1.dat"\\
+\tb @time = 2.0\\
+\tb @power = 4.0\\
+\tb @norm = 1e+8}\\
+{\tt >}
+\end{quote}
+by
+\begin{quote}
+{\tt FF=5.1989} \\
+\\
+{\tt nuflux(\#user)<}\\
+{\tt \tb @flux\_file = "user\_file\_1.dat"\\
+\tb @time = 2.0\\
+\tb @power = 4.0\\
+\tb @norm = FF*1e+8}\\
+{\tt >}
+\end{quote}
+
+This replacement is not necessary for neutrino factory built-in fluxes,
+and built-in beta beams fluxes where not supported by earlier versions
+of \GLOBES .
+
+IS THAT CORRECT? MR: WHAT HAPPENS TO OLD BETA BEAM FILES [WW]?
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2007-01-11 Documentation/part1.tex
*** empty log message ***

@@
 Alternatively to the number, the constants \GLBC{GLB\_THETA\_12}, \GLBC{GLB\_THETA\_13}, \GLBC{GLB\_THETA\_23}, \GLBC{GLB\_DELTA\_CP}, 
-\GLBC{GLB\_DM\_SOL}, or \GLBC{GLB\_DM\_ATM} can be used.
+\GLBC{GLB\_DM\_21}, or \GLBC{GLB\_DM\_31} can be used.
 
@@
 
-\section{Version control}
+\section{Version control and debugging}
 \label{sec:versioning}
@@
 \end{function}
+%
+A useful function to debug \GLOBES\ source code is 
+\begin{function}
+\GLBNS{SetVerbosityLevel}
+{\tt int glbSetVerbosityLevel(int level)} sets the verbosity level
+for \GLOBES\ messages.
+\end{function}
+The following verbosity levels are currently used:
+\begin{description}
+\item[0] Display no messages
+\item[1] (Standard) Display error messages
+\item[2] Display warnings
+\item[3] Display file access history
+\item[4] Display search paths
+\end{description}
+Note that always messages with the chosen verbosity level and lower
+are displayed.
 
@@
 systematics, correlations, and degeneracies.
-Note that also the systematics function itself may be changed during running time,
-as described in \Sec~\ref{sec:userchi}.
+Note that also the systematics function itself may be changed during running time
+using \GLB{SetChiFunction}, as it is described in \Sec~\ref{sec:userchi}.
 
@@
 \chapter{Experimental features}
+\label{chapt:experimental}
 

### 2007-01-11 Documentation/part2.tex
*** empty log message ***

@@
 \section{More advanced AEDL features}
+\label{sec:advaedl}
 
@@
 
-\index{aedl}{{\tt nuflux}!time@{\tt "@time}}
+\index{aedl}{nuflux@{\tt nuflux}!time@{\tt "@time}}
 There are two principal ways to initialize a neutrino flux: 
@@
 a flux is defined by the environment {\tt nuflux}, such as
-\index{aedl}{{\tt nuflux}}
+\index{aedl}{nuflux@{\tt nuflux}}
 \begin{quote}
@@
 
-\index{aedl}{{\tt flux}}
+\index{aedl}{nuflux@{\tt nuflux}}
 
@@
 Thus, an example to set up a neutrino factory flux is
-\index{aedl}{{\tt nuflux}!builtin@{\tt "@builtin}}
-\index{aedl}{{\tt nuflux}!parent energy@{\tt "@parent\_energy}}
-\index{aedl}{{\tt nuflux}!stored muons@{\tt "@stored\_muons}}
-\index{aedl}{{\tt nuflux}!end point@{\tt "@end\_point}}
-\index{aedl}{{\tt nuflux}!stored ions@{\tt "@stored\_ions}}
-\index{aedl}{{\tt nuflux}!gamma@{\tt "@gamma}}
+\index{aedl}{nuflux@{\tt nuflux}!builtin@{\tt "@builtin}}
+\index{aedl}{nuflux@{\tt nuflux}!parent energy@{\tt "@parent\_energy}}
+\index{aedl}{nuflux@{\tt nuflux}!stored muons@{\tt "@stored\_muons}}
+\index{aedl}{nuflux@{\tt nuflux}!end point@{\tt "@end\_point}}
+\index{aedl}{nuflux@{\tt nuflux}!stored ions@{\tt "@stored\_ions}}
+\index{aedl}{nuflux@{\tt nuflux}!gamma@{\tt "@gamma}}
 \begin{quote}
@@
 \end{quote}
-%
-\index{aedl}{{\tt nuflux}!power@{\tt "@power}}
-\index{aedl}{{\tt nuflux}!norm@{\tt "@norm}}
-\index{aedl}{{\tt nuflux}!flux file@{\tt "@flux\_file}}
+MR: SOME MORE ELABORATION AND AN EXAMPLE FOR THE BETA BEAMS MAY BE APPROPRIATE HERE [WW].
+%
+\index{aedl}{nuflux@{\tt nuflux}!power@{\tt "@power}}
+\index{aedl}{nuflux@{\tt nuflux}!norm@{\tt "@norm}}
+\index{aedl}{nuflux@{\tt nuflux}!flux file@{\tt "@flux\_file}}
 For a user-defined flux, one has to give it the file name:
@@
 
-\index{aedl}{{\tt flux}!flux file@{\tt "@flux\_file}}\index{norm}{Flux!file}
+\index{aedl}{nuflux@{\tt nuflux}!flux file@{\tt "@flux\_file}}\index{norm}{Flux!file}
 The software assumes that the given flux file has seven columns and
@@
 fluxes, they are provided by the user as a data file:
-\index{aedl}{{\tt cross}}
-\index{aedl}{{\tt cross}!cross file@{\tt  "@cross\_file}}
+\index{aedl}{cross@{\tt cross}}
+\index{aedl}{cross@{\tt cross}!cross file@{\tt  "@cross\_file}}
 \begin{quote}
@@
 \index{norm}{Cross section!file}
-\index{aedl}{{\tt cross}!cross file@{\tt  "@cross\_file}}
+\index{aedl}{cross@{\tt cross}!cross file@{\tt  "@cross\_file}}
 \begin{quotation}
@@
 
-\index{aedl}{{\tt channel}|(}
+\index{aedl}{channel@{\tt channel}|(}
 Channels\index{norm}{Channel} in \GLOBES\ represent an intermediate level 
@@
 defined before with their {\tt \#name} in the respective environments. 
-Thus, a simple definition of a channel\index{aedl}{{\tt channel}} is
+Thus, a simple definition of a channel\index{aedl}{channel@{\tt channel}} is
 \begin{quote}
@@
 It is also possible to define a channel as no-oscillation by using the
-prefix {\tt NOSC\_}\index{aedl}{{\tt channel}!{\tt NOSC\_}} 
+prefix {\tt NOSC\_}\index{aedl}{channel@{\tt channel}!{\tt NOSC\_}} 
 in either the initial  flavour or 
@@
 energy resolution function in the next section.
-\index{aedl}{{\tt channel}|)}
+\index{aedl}{channel@{\tt channel}|)}
 
@@
 
-\index{aedl}{{\tt energy}|(}
+\index{aedl}{energy@{\tt energy}|(}
 \index{norm}{Energy!resolution|(}
@@
 %
-\index{aedl}{{\tt channel}!pre smearing efficiencies@{\tt 
+\index{aedl}{channel@{\tt channel}!pre smearing efficiencies@{\tt 
 "@pre\_smearing\_efficiencies}}
-\index{aedl}{{\tt channel}!pre smearing background@{\tt 
+\index{aedl}{channel@{\tt channel}!pre smearing background@{\tt 
 "@pre\_smearing\_background}}
-\index{aedl}{{\tt channel}!post smearing efficiencies@{\tt 
+\index{aedl}{channel@{\tt channel}!post smearing efficiencies@{\tt 
 "@post\_smearing\_efficiencies}}
-\index{aedl}{{\tt channel}!post smearing background@{\tt 
+\index{aedl}{channel@{\tt channel}!post smearing background@{\tt 
 "@post\_smearing\_background}}
@@
 %
-\index{aedl}{{\tt energy}!type@{\tt "@type}}
+\index{aedl}{energy@{\tt energy}!type@{\tt "@type}}
 %
@@
 %
-\index{aedl}{{\tt energy}!sigma function@{\tt "@sigma\_function}}
+\index{aedl}{energy@{\tt energy}!sigma function@{\tt "@sigma\_function}}
 %
@@
 %
-\index{aedl}{{\tt energy}!standard@{\tt \#standard}}
+\index{aedl}{energy@{\tt energy}!standard@{\tt \#standard}}
 %
@@
 %
-\index{aedl}{{\tt energy}!inverse beta@{\tt \#inverse\_beta}}
+\index{aedl}{energy@{\tt energy}!inverse beta@{\tt \#inverse\_beta}}
 %
@@
 
-\index{aedl}{{\tt energy}!type@{\tt  "@type}}
+\index{aedl}{energy@{\tt energy}!type@{\tt  "@type}}
 With a type 1 ({\tt @type = 1}) energy resolution 
@@
 where the unshown entries are zero. Thus, the values of $K_{ij}$ have to be specified between $k_l^i$ and $k_u^i$ in the form $\{ k_l^i,k_u^i, K_{i \, k_l^i}, K_{i k_l^i+1} , \hdots , K_{i k_u^i} \}$:
-\index{aedl}{{\tt energy}!energy@{\tt "@energy}}
+\index{aedl}{energy@{\tt energy}!energy@{\tt "@energy}}
 \begin{quote}
@@
 
-\index{aedl}{{\tt energy}|)}
+\index{aedl}{energy@{\tt energy}|)}
 \index{norm}{Energy!resolution|)}
@@
 
-\index{aedl}{{\tt rule}|(}
+\index{aedl}{rule@{\tt rule}|(}
 The set of rules\index{norm}{Rule} for an experiment is the final 
@@
 %
-\index{aedl}{{\tt rule}!signal@{\tt "@signal}}
-\index{aedl}{{\tt rule}!background@{\tt "@background}}
+\index{aedl}{rule@{\tt rule}!signal@{\tt "@signal}}
+\index{aedl}{rule@{\tt rule}!background@{\tt "@background}}
 %
@@
 with each rule to an energy window\index{norm}{Energy!window}:
-\index{aedl}{{\tt rule}!energy window@{\tt "@energy\_window}}
+\index{aedl}{rule@{\tt rule}!energy window@{\tt "@energy\_window}}
 \begin{quote}
@@
 interpreted as zero central value. IS THAT CORRECT [WW]?}
-\index{aedl}{{\tt rule}!signal error@{\tt "@signalerror}}
-\index{aedl}{{\tt rule}!background error@{\tt "@backgrounderror}}
+\index{aedl}{rule@{\tt rule}!signal error@{\tt "@signalerror}}
+\index{aedl}{rule@{\tt rule}!background error@{\tt "@backgrounderror}}
 %\index{aedl}{{\tt rule}!background center@{\tt "@backgroundcenter}}
@@
 but should not be used anymore.}
-\index{aedl}{{\tt rule}!sys on function@{\tt "@sys\_on\_function}}
-\index{aedl}{{\tt rule}!sys off function@{\tt "@sys\_off\_function}}
+\index{aedl}{rule@{\tt rule}!sys on function@{\tt "@sys\_on\_function}}
+\index{aedl}{rule@{\tt rule}!sys off function@{\tt "@sys\_off\_function}}
 \index{norm}{Systematics function}
@@
 one would specify the systematical errors in lists, such as
-\index{aedl}{{\tt rule}!sys on errors@{\tt "@sys\_on\_errors}}
-\index{aedl}{{\tt rule}!sys off errors@{\tt "@sys\_off\_errors}}
+\index{aedl}{rule@{\tt rule}!sys on errors@{\tt "@sys\_on\_errors}}
+\index{aedl}{rule@{\tt rule}!sys off errors@{\tt "@sys\_off\_errors}}
 \begin{quote}
@@
 \end{quote}
-\index{aedl}{{\tt rule}|)}
+\index{aedl}{rule@{\tt rule}|)}
 

### 2007-01-11 Documentation/part2.tex
JK - Modified discussion of tilt errors

@@
 is more accurate, but it requires a careful choice of parameters. 
-The first option (method ``T'') is (JK: IS THAT CORRECT? AND ALSO THE NEXT EQUATION? (THERE HAVE BEEN
-SOME CHANGES) [WW])
-\begin{equation}
-s_i(a,b) \equiv s_i(a)+b\cdot s_i\, E'_i/(E'_\mathrm{max}-E'_\mathrm{min}),
-\end{equation}
-where $E'_{\mathrm{min}}$ and $E'_{\mathrm{max}}$ correspond to {\tt \$emin}
-and {\tt \$emax}, and $E'_i$ is the mean (reconstructed) 
-energy of the $i$th bin. It is often refered to as a ``tilt'' of the
+The first option (method ``T'') is\footnote{Note that this behaviour
+has slightly changed compared to previous \GLOBES\ releases.}
+\begin{equation}
+s_i(a,b) \equiv s_i(a)+b\cdot s_i \cdot i / N_\mathrm{bins}.
+\end{equation}
+It is often refered to as a ``tilt'' of the
 spectrum, since it describes a linear distortion 
@@
 one suspects a large impact of this systematical error.
-It is based upon replacing the events in the $i$th bin by the ones at
-the energy $(1+b)\cdot E'_i$. If this target energy does not exactly hit
+It is based upon replacing the events in the $i$th bin (mean reconstructed energy $E'_i$)
+by the ones at the energy $(1+b)\cdot E'_i$. If the target energy does not exactly hit
 a (discrete) bin energy $E_k$, linear interpolation is used. We use the following approximation:
@@
 \end{eqnarray}
-Here $\Delta E_0$ is the bin width ({\tt \$emax}-{\tt \$emin})/{\tt \$bins}
+Here, $E'_\mathrm{min} = {\tt \$emin}$,  $\Delta E_0$ is the bin width
+({\tt \$emax}-{\tt \$emin})/{\tt \$bins},
 and ``div'' refers to the integer part of the division. Note that the

### 2007-01-11 Documentation/part1.tex
*** empty log message ***

@@
 
-TOUR ITSELF (+EXAMPLE FILE) TO BE UPDATED (VALUES MAY HAVE CHANGED DUE TO AEDL FILE DEFINITION
+FOR WW MYSELF: TOUR ITSELF (+EXAMPLE FILE) TO BE UPDATED (VALUES MAY HAVE CHANGED DUE TO AEDL FILE DEFINITION
 CHANGES ETC. ALSO EXPERIMENT (OLD) NUFACT DOES NOT EXIST ANYMORE. AND: SOME ERROR IN THIS OLD
@@
 CHECK: GLBSETPROJECTION REALLY USED WITH GLB\_FIXED FOR DENSITY PARAMS? IS NFSTANDARD USED?
+(QUESTIONS FOR WW MYSELF!)
 
@@
 oscillation parameters, the hybrid minimizer mixes the systematics and oscillation parameter minimizations. 
-This method is much faster, but new (unknown) correlations between systematics parameters and oscillation parameters may emerge.
+This method is much faster,
+but correlations between systematics parameters and oscillation parameters might lead to a different convergence behaviour in some situations.
+Thus, when switching to the hybrid minimizer in an application program, one has to reconsider the question whether all degeneracies are found.
 Since one can change the minimizer at any time, it is recommended that one cross check the minimization for a particular experiment.
+The danger of modifying the convergence behaviour in existing application programs is another reason why the new, faster minimizer is still declared as experimental and not used by default
 

### 2007-01-11 Documentation/part2.tex
*** empty log message ***

@@
 A background normalization center of 1.0 will be
-interpreted as zero central value. IS THAT CORRECT [WW]?}
+interpreted as zero central value.}
 \index{aedl}{rule@{\tt rule}!signal error@{\tt "@signalerror}}
@@
 systematics functions for each rule 
-(see also~\Sec~\ref{sec:systematics}).. The signal and background errors specified
+(see also~\Sec~\ref{sec:systematics}). The signal and background errors specified
 by {\tt @signalerror} and  {\tt @backgrounderror} will then be used, if applicable.
@@
 \end{quote}
-WHICH SYSTEMATICAL ERRORS WILL A FUNCTION USE IF THERE IS A MIXED DECLARATION,
-SUCH AS WITH SYS\_ON\_ERRORS IN SYS ON AND SIGNALERROR/BACKGROUNDERROR in SYS OFF?
-DOES THAT WORK? [WW]
 Note that here the systematics on and systematics off models can be used 
@@
 resulting $\chi^2$ will be a factor two too large.
+Note that mixed declarations of {\tt @sys\_on\_errors}, {\tt @sys\_off\_errors}, {\tt @signalerror}, and
+{\tt @backgrounderror} are possible. In this case, {\tt @sys\_on\_errors} and {\tt @sys\_off\_errors}
+have priority. If for one or both of these no systematical errors are defined, {\tt @signalerror} or
+{\tt @backgrounderror} will be used.
 

### 2007-01-11 Documentation/Paper2/paper2.tex
*** empty log message ***

@@
 {\begin{center}
-{\Large\bf New features in the simulation of long baseline experiments with GLoBES 3.0 (working title)}
+{\Large\bf New features in the simulation of neutrino oscillation experiments with GLoBES 3.0}
 \end{center}}
@@
 {\begin{center} {\large{\sc
-                P.~Huber\footnotemark[1],~
-                J.~Kopp\footnotemark[2],~
-                M.~Lindner\footnotemark[2],~
-                M.~Rolinec\footnotemark[2],~
-                W.~Winter\footnotemark[3]
+                Patrick~Huber\footnotemark[1],~
+                Joachim~Kopp\footnotemark[2],~
+                Manfred~Lindner\footnotemark[2], \\
+                Mark~Rolinec\footnotemark[3],~
+                Walter~Winter\footnotemark[4]
                 }}
 
-\footnote{All correspondence should be addressed to {\tt globes@ph.tum.de}}
+\footnote{All correspondence should be addressed to {\tt globes@mpi-hd.mpg.de}}
 
@@
 \footnotemark[2]%
+       Max--Planck--Institut f\"ur Kernphysik,  \\
+       Postfach 10~39~80, D--69029 Heidelberg, Germany
+
+\vspace*{1mm}
+
+\footnotemark[3]%
        Physik--Department, Technische Universit\"at M\"unchen, \\
@@
 
-\footnotemark[3]%
-       School of Natural Sciences, Institute for Advanced Study, \\
-       Einstein Drive, Princeton, NJ 08540, USA
+\footnotemark[4]%
+       Universit\"at W\"urzburg, 
+       Institut f\"ur theoretische Physik und Astrophysik, \\
+       Am Hubland, D-97074 W\"urzburg, Germany
 
@@
 \begin{abstract}
-Here comes the abstract.
+We present Version 3.0 of the GLoBES (``General Long Baseline Experiment Simulator'') software, which is
+a simulation tool for short- and long-baseline neutrino oscillation experiments.
+As new features, \GLOBES\ 3.0 allows the simulation of experiments with multiple detectors by the concept of user-defined systematics.
+In addition, the combination with external information, such as from different experiment classes, is 
+simplified by the concept of user-defined priors. As far as the probability calculation is concerned, \GLOBES\ 
+now provides an interface for the inclusion of non-standard physics without re-compliation of the software. 
+The set of experiment descriptions coming with \GLOBES\ has been updated. For example, built-in fluxes are now
+provided for the simulation of beta beams.
 \end{abstract}
@@
 
-\nocite{Huber:2004ka}
+Neutrino oscillations are now established as the leading flavor
+transition mechanism for neutrinos in a long history of many experiments, see
+\eg , \Ref~\cite{Barger:2003qi}. Future facilities, using accelerator-based
+ neutrino beams or a nuclear reactors as neutrino sources, are proposed
+for precision measurements of the neutrino oscillation parameters. In the
+simulation of these experiments, 
+the presence of multiple solutions which are intrinsic to
+ neutrino oscillation probabilities~\cite{Fogli:1996pv,Burguet-Castell:2001ez,Minakata:2001qm,Barger:2001yr} 
+affect the performance.
+Thus optimization strategies are required which maximally exploit 
+complementarity between experiments. The \GLOBES\ software package~\cite{Huber:2004ka}
+is a  modern experiment  simulation  and analysis tool for
+a highly accurate beam and detector simulation. In addition, it provides powerful 
+means to analyze correlations and degeneracies, especially for the combination
+of several experiments. Compared to a Monte Carlo simulation, which yields a different
+result in each run, it simulates the performance of the {\em average} experiment
+(see \Ref~\cite{Schwetz:2006md} for a discussion of the meaning of ``average'').
+The advantage of such an average prediction is a tremendous performance gain,
+which can be used for systematical parameter space scans. In addition, it 
+simplified the direct comparison of experiments.
 
-\section{User-defined priors and systematics}
+The \GLOBES\ software has, in the past, been used for many studies, some of them
+are referred to in the paper. However, recent developments have indicated that extensions
+and improvements are necessary. In this work, we present the most important new
+features and changes in \GLOBES\ 3.0. For experimentalists, \GLOBES\ now allows
+for the implementation of arbitrary systematics, which can also be used for
+the simulation of multi-detector experiments. In addition, several extensions
+have been included in \AEDL\ (``Abstract Experiment Description Language''), such as
+built-in beta beam fluxes and the possibility to use lists as variables.
+For phenomenologists, user-defined
+priors provide a flexible interface to include external information, such as from
+different experiments. In addition, \GLOBES\ now can be used for the simulation
+of non-standard physics without re-compiling the \GLOBES\ software.
 
-\bi
- \item
-  User-defined priors
- \item
-  User-defined systematics
-\ei
+\section{Concept of \GLOBES }
+
+\GLOBES\ (``General Long Baseline Experiment Simulator'') is a flexible
+software tool to simulate and analyze neutrino oscillation 
+short- and long-baseline experiments using a 
+complete three-flavor description. On the
+one hand, it contains a comprehensive abstract experiment definition
+language (\AEDL ), which allows to describe 
+most classes of long baseline and reactor experiments
+at an abstract level. On the other hand, it provides a C-library to 
+process the experiment information in order to obtain oscillation
+probabilities, rate vectors, and $\Delta \chi^2$-values (\cf, \figu{GLOBES}). 
+In addition, it provides a binary program to test experiment
+definitions very quickly, before they are used by the application software.
+Currently, \GLOBES\ is available for GNU/Linux.
+
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=14cm]{GLOBES}
+\end{center}
+\caption{\label{fig:GLOBES} General concept of the \GLOBES\ package.}
+\end{figure}
+
+\GLOBES\ allows to simulate experiments with stationary neutrino point 
+sources, where each experiment is assumed to have only one neutrino source.
+Such experiments are neutrino beam experiments and reactor experiments. 
+Geometrical effects of a continuous source distribution, such as in the sun or the 
+atmosphere, can not be described. In addition, sources with a physically 
+significant time dependencies can not be studied, such as  supernov\ae. 
+However, in \GLOBES\ 3.0 and higher, new flexibility is introduced by the
+concept of user-defined systematics. This new feature allows the cross-definition
+of systematical errors over different experiments. In principle, 
+this mechanism can be used for the combination of several discrete sources
+and one detector, or several detectors and one source, or several sources and
+several detectors. In addition, already implemented
+concepts of \GLOBES\ have been used for indirect simulations of geometrical
+effects. For example, the mapping of the detector location on the neutrino
+energy has been simulated in \Ref~\cite{Rolinec:2006xr} by the use of
+variable bin widths.
+
+On the experiment definition side, either built-in neutrino fluxes
+(\eg, neutrino factory, beta beam) or arbitrary, user-defined fluxes can be used. 
+Similarly,
+arbitrary cross sections, energy dependent efficiencies,
+energy resolution functions as well as the considered oscillation channels, 
+backgrounds, and many other properties can be specified. 
+For the systematics, energy
+normalization and calibration errors can be simulated, or the
+systematics can be completely user-defined. Note that
+energy ranges and windows and bin widths can be
+(almost) arbitrarily chosen, including bins of different
+widths. Together with \GLOBES\ comes a number of
+pre-defined experiments in order to demonstrate the capabilities
+of \GLOBES\ and to provide prototypes for new experiments.
+In addition, they can be used to test new physics ideas with
+complete experiment simulations.
+% Examples for these prototypes are the \MINOS , \ICARUS , and
+% \OPERA\ simulations from \Ref~\cite{Huber:2004ug}, the
+% \JHFSK\ and \NUMI\ superbeam simulations from \Refs~\cite{Huber:2002mx,
+% Huber:2002rs}, the \JHFHK\ superbeam upgrade simulation from
+% \Ref~\cite{Huber:2002mx}, the neutrino factory simulations from
+% \Ref~\cite{Huber:2002mx}, and the reactor experiment simulations from
+% \Ref~\cite{Huber:2003pm}. Other projects using earlier versions
+% of the \GLOBES\ software include \Refs~\cite{Huber:2003ak,Ohlsson:2003ip,Winter:2003ye,Antusch:2004yx}.
+
+With the C-library, one can extract the $\Delta \chi^2$ for all defined 
+oscillation channels for an experiment or any combination of experiments.
+Of course, also low-level information, such as oscillation
+probabilities or event rates, can be obtained. \GLOBES\ includes the
+simulation of neutrino oscillations in matter with arbitrary matter 
+density profiles. In addition, it allows to simulate the matter density
+uncertainty (see, \eg, \Refs~\cite{Huber:2002mx,Ohlsson:2003ip}) and to 
+extract the precision on the matter density (see, \eg, \Refs~\cite{Winter:2005we,Gandhi:2006gu}).
+ As one of the most
+advanced features of \GLOBES , it provides the technology to 
+project the $\Delta \chi^2$, which is a function of all oscillation
+parameters, onto any subspace of parameters by local minimization. 
+This approach allows the inclusion of multi-parameter-correlations,
+where external constraints (\eg, on the solar parameters) can be imposed, too.
+Applications of the projection mechanism include the projections onto 
+the $\stheta$-axis and the $\stheta$-$\deltacp$-plane. In addition, 
+all oscillation parameters can be kept free to precisely localize 
+degenerate solutions.
+
+\section{User-defined systematics and changes in the systematics implementation}
+
+How defined, what good for (physics examples: T2KK, Reactor-paper,).
+Compare to standard systematics (nuisance parameters)
+
+\section{User-defined priors}
+
+How defined, what good for (physics examples: AstroNus, Schwetz priors, atm. neutrinos, octant degeneracy,
+).
 
@@
 
-HOWTO
+Eg. Damping, Hamiltonian, MAVANS, Toshi etc.
 
-\section{Simulation of new experiment features}
+\section{Simulation of new experiment features and \AEDL\ changes}
 
@@
 
-\section{New platform and interpreter support}
+Beta beams paper!
+
+\section{New platform and interpreter support, new ways to improve the performance}
 
@@
 \ei
-
-\section{New ways to improve performance}
 
@@
 
-\begin{appendix}
-\section{To do list for new feature release (early spring?); internal}
-
-\subsection*{Programming [PH+JK?]}
-
-\bi
-\item
- Implement user-defined priors+ systematics (done?)
-\item
- Clean up linking scheme: No dynamic linking of modules (physical module
- support) anymore; support static binaries, support Condor etc.;
- problem with append mode (files) [PH]
-\item
- New algoriihm for minimzer [JK]? In this context: Is there any
- efficient algorithm for more than six parameters?
-\item
- Can globes-command (!) take more than six parameters if GLB\_OSCP $>$ 6?
- Adjust, if necessary.
-\item
- Implement Walter's or any modified bugfix for glb\_min\_sup.c
- (we had some negative feedback about that, and for more than
- six parameters ITMAX is far away from being sufficient)
-\ei
-
-\subsection*{Testing [MR+All?]}
-
-\bi
-\item
- Does release-version support Condor and more than six parameters?
-\item 
- Do beta-beam fluxes etc. work?
-\item
- Do examples produce same results as before?
-\item
- Are AEDL files consistent?
-\ei
-
-\subsection*{Documentation [WW with some help?]}
-
-\bi
-\item
- Document all new features and make examples; 
- especially user-defined priors+systematics, Non-standard-Support
-\item
- New examples files for new examples in manual
-\item
- Describe what papers have to be cited
-\ei
-
-\subsection*{Paper [WW/ML?]}
-
-\bi
- \item
-  Write new release paper (this draft)
-\ei
-
-\subsection*{AEDL files [MR?]}
-
-\bi
-\item
- Provide new AEDL files for some/any finished (=published) projects!?
- \bi
-  \item
-   $\beta$-Beams
-  \item
-   New NuFact (from scoping study); our new NuFact; silver channels?
-  \item
-   Double Chooz
-  \item
-   NOvA, BNL beam?
- \ei
-\item
- Cleaned up old AEDL files (e.g., 2.5\% background error NuFact etc.);
- document what changes
-\item
- New AEDL file forum on webpage
-\item
- Clear description in each AEDL file which papers have to be cited
-\ei
-
-\subsection*{Website [ML/JK?]}
-
-\bi
- \item
-  New layout with new release
- \item
-  New FAQ section (e.g., about normalization in AEDL fluxes)
- \item
-  New AEDL file forum
- \item
-  Figure forum, at least from current projects (download)
-\ei
-
-\end{appendix}
-
 \subsection*{Acknowledgments}
 
-We would like to thank ...
-
-WW would like to
-acknowledge support from the W.~M.~Keck Foundation and NSF grant PHY-0503584.
+For the current version of \GLOBES , we would like to 
+thank Thomas Schwetz, who has been pushing the software to the edge in the 
+past few years. Furthermore, we would like to thank Tommy Ohlsson,
+Toshihiko Ota, and Julian Skrotzki for using and testing unpublished new features of the software.
+%
+The development of \GLOBES\ is currently being supported by the Studienstiftung des Deutschen Volkes [JK] and the
+ Emmy Noether-Program of Deutsche Forschungsgemeinschaft [WW].
+%
+In the past, the following institutions and organizations have supported 
+members of the \GLOBES\ team: Technische Universit\"at M\"unchen,
+Max-Planck-Institut f\"ur Physik, SFB 375 for astro-particle physics
+of Deutsche Forschungsgemeinschaft, Studienstiftung des Deutschen Volkes,
+Institute for Advanced Study, Princeton, W. M. Keck Foundation, and National Science Foundation.
 

### 2007-01-11 Documentation/part1.tex
JK   - Documented new functions which return pointers to arrays of bin widths and     bin central energies   - Adjusted documentation of tilt   - Added warning in the section about energy calibration errors, saying that it     will not work with variable bin widths.

@@
 \end{function}
+\begin{function}
+\GLBNS{GetBinSizeListPtr}
+{\tt double *glbGetBinSizeListPtr(int exp)} returns a pointer to the array of bin widths
+for the experiment {\tt exp}.
+\end{function}
+\begin{function}
+\GLBNS{GetSamplingStepsizeListPtr}
+{\tt double *glbGetSamplingStepsizeListPtr(int exp)} returns a pointer to the array of
+sampling step sizes for the experiment {\tt exp}.
+\end{function}
+\begin{function}
+\GLBNS{GetBinCentersListPtr}
+{\tt double *glbGetBinCentersListPtr(int exp)} returns a pointer to the array of
+mean bin energies for the experiment {\tt exp}.
+\end{function}
+\begin{function}
+\GLBNS{GetSamplingPointsListPtr}
+{\tt double *glbGetSamplingPointsListPtr(int exp)} returns a pointer to the array of
+sampling points for the experiment {\tt exp}.
+\end{function}
 

### 2007-01-11 Documentation/part2.tex
JK   - Documented new functions which return pointers to arrays of bin widths and     bin central energies   - Adjusted documentation of tilt   - Added warning in the section about energy calibration errors, saying that it     will not work with variable bin widths.

@@
 \begin{equation}
-s_i(a,b) \equiv s_i(a)+b\cdot s_i \cdot i / N_\mathrm{bins}.
-\end{equation}
-It is often refered to as a ``tilt'' of the
+  s_i(a,b) \equiv s_i(a)+b\cdot s_i \, (E'_i - \bar{E'}) / (E'_\mathrm{max}-E'_\mathrm{min}),
+\end{equation}
+where $E'_\mathrm{min}$ and $E'_\mathrm{max}$ correspond to {\tt \$emin} and
+{\tt \$emax}, $\bar{E'} = \tfrac{1}{2} (E'_\mathrm{max}+E'_\mathrm{min})$ is
+the median of this energy interval, and $E'_i$ is the mean (reconstructed)
+energy of the $i$th bin.
+This method is often refered to as a ``tilt'' of the
 spectrum, since it describes a linear distortion 
-of the event rate spectrum. Note that this definition of a tilt makes
-only sense in combination with a large enough normalization error, since
-the tilt may also affect the normalization. 
+of the event rate spectrum.
+%Note that this definition of a tilt makes
+%only sense in combination with a large enough normalization error, since
+%the tilt may also affect the normalization. 
 %
@@
 one suspects a large impact of this systematical error.
-It is based upon replacing the events in the $i$th bin (mean reconstructed energy $E'_i$)
+It is based upon replacing the events in the $i$th bin
 by the ones at the energy $(1+b)\cdot E'_i$. If the target energy does not exactly hit
@@
 \begin{eqnarray}
-s_i(a,b)&=& (1+b)\cdot \left[ \left( s_{k+1}(a)-s_k(a) \right)\cdot (\delta-k) +s_k(a) \right] \,,\\
+s_i(a,b)&=& (1+b)\cdot \left[ \left( s_{k+1}(a)-s_k(a) \right)\cdot (\delta-k) +s_k(a) \right] \,,
+  \label{eq:Ecalib} \\
 \delta&=&b\cdot(i+t_0+ 1/2)+i\,,\nonumber\\
@@
 \end{eqnarray}
-Here, $E'_\mathrm{min} = {\tt \$emin}$,  $\Delta E_0$ is the bin width
-({\tt \$emax}-{\tt \$emin})/{\tt \$bins},
-and ``div'' refers to the integer part of the division. Note that the
-factor $(1+b)$ in the first equation comes from a renormalization of
+Here, $\Delta E_0$ is the bin width ({\tt \$emax}-{\tt \$emin})/{\tt \$bins},
+and ``div'' refers to the integer part of the division. It is important
+to keep in mind that this definition of the energy calibration error 
+makes sense only for constant bin width, so the corresponding $\chi^2$
+functions should not be used in conjunction with the {\tt \$binwidth} directive.
+The factor $(1+b)$ in \eq~\eqref{eq:Ecalib} comes from a renormalization of
 the bin width, since also the bin width is altered by the replacement of

### 2007-01-11 Documentation/appendix.tex
*** empty log message ***

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\chapter{Catalogue of {\sf AEDL}-Files}
+\label{app:aedlfiles}
+Along with the \GLOBES\ package comes a catalogue of pre-defined experiment \AEDL\ files for different
+future experiments and different beam and detector technologies. These include the planned superbeam experiments
+and their possible upgrades, different reactor experiment setups, different $\beta$-beam setups, and different 
+neutrino factory setups. A complete list
+of all pre-defined experiment files an be found in \tabl{experiments}. More detailed descriptions of the
+corresponding files, the assumptions, requirements, and references are given in the following.
+
+\section*{Superbeam experiments}
+\subsection*{T2K -- {\tt T2K.glb}}
+
+The \TtoK\ experiment can be simulated with the file {\tt T2K.glb}. This file tries to approximate as closely as
+possible the LOI \cite{Itow:2001ee} and the basic version was used within \cite{Huber:2002mx}. These references
+should be cited if the file {\tt T2K.glb} is used for a scientific publication or a talk. For calculations that
+involve {\tt T2K.glb}, the following additional files are required: 
+\bi
+\item {\tt JHFplus.dat} (neutrino flux from J-PARC -- $\nu_\mu$)
+\item {\tt JHFminus.dat} (neutrino flux from J-PARC -- $\bar{\nu}_\mu$)
+\item {\tt XCC.dat} (charged current cross sections)
+\item {\tt XNC.dat} (neutral current cross sections)
+\item {\tt XQE.dat} (quasi elastic cross sections)
+\ei
+The \TtoK\ neutrino beam is produced at J-PARC and directed towards the Super-Kamiokande detector. The target
+power is 0.77~MW and 2~years $\nu$-running and 6~years $\bar{\nu}$-running is assumed. The fiducial mass of the
+Super-Kamiokande Water Cerenkov detector is taken to be $\mathrm{m_{det} = 2.5 \,kt}$ at a baseline of
+L~=~295~km. The appearance measurement involves the total rates data from all CC events and the spectral data
+from the QE sample with a free normalization at an energy resolution of $\mathrm{\sigma_e=0.085\, GeV}$ due to the
+Fermi Motion as is described in more detail in~\cite{Huber:2002mx}. The treatment of systematics is similar to
+\cite{Ishitsuka:2005qi}. The following rules are defined within {\tt T2K.glb}:  
+\begin{center}
+\begin{tabular}{|l|ll|c|c|}
+\hline \hline
+\multicolumn{3}{|l|}{Disappearance (+)} & $\sigma_\mathrm{norm}$ & $\sigma_\mathrm{cal}$ \\ \hline 
+Signal & $0.9 \, \otimes \, (\nu_\mu\rightarrow\nu_\mu)_{\mathrm{QE}}$ & & 0.025 & $10^{-4}$\\
+ & &  & &\\
+Background & $0.0056 \, \otimes \, (\nu_\mu \rightarrow \nu_x)_\mathrm{NC}$ & & 0.2 & $10^{-4}$ \\ \hline \hline 
+\multicolumn{3}{|l|}{Appearance (+) -- Spectrum}  & & \\ \hline
+Signal & $0.505 \, \otimes \, (\nu_\mu \rightarrow \nu_e)_\mathrm{QE}$ & & 10.0 & $10^{-4}$\\
+ & & & & \\
+Background & $0.0056 \, \otimes \, (\nu_\mu \rightarrow \nu_x)_\mathrm{NC}$ & $3.3\cdot 10^{-4} \, \otimes \, (\nu_\mu\rightarrow\nu_\mu)_{\mathrm{CC}}$ & 0.05 & 0.05\\
+Beam background & $0.505 \, \otimes \, (\nu_e\rightarrow \nu_e)_\mathrm{CC}$ & $0.505 \, \otimes \, (\bar{\nu}_e\rightarrow \bar{\nu}_e)_\mathrm{CC}$  & 0.05 & 0.05\\ \hline \hline
+\multicolumn{3}{|l|}{Appearance (+) -- Total Rates}  & & \\ \hline
+Signal & $0.505 \, \otimes \, (\nu_\mu \rightarrow \nu_e)_\mathrm{CC}$ & & 0.05 & $10^{-4}$\\
+ & & & & \\
+Background & $0.0056 \, \otimes \, (\nu_\mu \rightarrow \nu_x)_\mathrm{NC}$ & $3.3\cdot 10^{-4} \, \otimes \, (\nu_\mu\rightarrow\nu_\mu)_{\mathrm{CC}}$ & 0.05 & $10^{-4}$\\
+Beam background & $0.505 \, \otimes \, (\nu_e\rightarrow \nu_e)_\mathrm{CC}$ & $0.505 \, \otimes \, (\bar{\nu}_e\rightarrow \bar{\nu}_e)_\mathrm{CC}$  & 0.05 & $10^{-4}$\\ \hline \hline
+\multicolumn{3}{|l|}{Disappearance (--)} & & \\ \hline 
+Signal & $0.9 \, \otimes \, (\nu_\mu\rightarrow\nu_\mu)_{\mathrm{QE}}$ & & 0.025 & $10^{-4}$\\
+ & &  & &\\
+Background & $0.0056 \, \otimes \, (\nu_\mu \rightarrow \nu_x)_\mathrm{NC}$ & & 0.2 & $10^{-4}$ \\ \hline \hline 
+\multicolumn{3}{|l|}{Appearance (--) -- Spectrum}  & & \\ \hline
+Signal & $0.505 \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_e)_\mathrm{QE}$ & & 10.0 & $10^{-4}$\\
+ & & & & \\
+Background & $0.0056 \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_x)_\mathrm{NC}$ & $3.3\cdot 10^{-4} \, \otimes \, (\bar{\nu}_\mu\rightarrow\bar{\nu}_\mu)_{\mathrm{CC}}$ & 0.05 & 0.05\\
+Beam background & $0.505 \, \otimes \, (\bar{\nu}_e\rightarrow \bar{\nu}_e)_\mathrm{CC}$ & $0.505 \, \otimes \, (\nu_e\rightarrow \nu_e)_\mathrm{CC}$  & 0.05 & 0.05\\ \hline \hline
+\multicolumn{3}{|l|}{Appearance (--) -- Total Rates}  & & \\ \hline
+Signal & $0.505 \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_e)_\mathrm{CC}$ & & 0.05 & $10^{-4}$\\
+ & & & & \\
+Background & $0.0056 \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_x)_\mathrm{NC}$ & $3.3\cdot 10^{-4} \, \otimes \, (\bar{\nu}_\mu\rightarrow\bar{\nu}_\mu)_{\mathrm{CC}}$ & 0.05 & $10^{-4}$\\
+Beam background & $0.505 \, \otimes \, (\bar{\nu}_e\rightarrow \bar{\nu}_e)_\mathrm{CC}$ & $0.505 \, \otimes \, (\nu_e\rightarrow \nu_e)_\mathrm{CC}$  & 0.05 & $10^{-4}$\\ \hline \hline
+\end{tabular}
+\end{center}
+
+\subsection*{T2HK -- {\tt T2HK.glb}}
+
+\TtoHK\ is the superbeam upgrade of the \TtoK\ experiment and can be simulated with the file {\tt T2HK.glb}. 
+The target power is 4~MW and 
+4~years $\nu$-running and 4~years $\bar{\nu}$-running is assumed.
+The fiducial mass of the Water Cerenkov detector is taken to be $\mathrm{m_{det} = 500 \,kt}$ at the same
+baseline as the \TtoK\ experiment. Besides these changes, the file {\tt T2HK.glb} is similar to {\tt T2K.glb}
+and the same additional files are required. 
+The basic version was used within \cite{Huber:2002mx} which should be cited if the file {\tt T2K.glb} is 
+used for a scientific publication or a talk.
+
+\subsection*{NO$\nu$A -- {\tt NOvA.glb}}
+
+The \NOVA\ experiment can be simulated with the file {\tt NOvA.glb}. The description of the disapperance 
+channels is taken from \cite{Yang_2004} and the description of the appearance channels follows 
+the proposal \cite{Ambats:2004js}. These references
+should be cited if the file {\tt NOvA.glb} is used for a scientific publication or a talk. For calculations that
+involve {\tt NOvA.glb}, the following additional files are required: 
+\bi
+\item {\tt NOvAplus.dat} (NuMI neutrino flux -- $\nu_\mu$)
+\item {\tt NOvAminus.dat} (NuMI neutrino flux -- $\bar{\nu}_\mu$)
+\item {\tt XCC.dat} (charged current cross sections)
+\item {\tt XNC.dat} (neutral current cross sections)
+\ei
+The \NOVA\ experiment uses a neutrino beam from the Fermilab NuMI beamline. The target power is 
+1.2~MV which results in $10^{21}$~pot~$\mathrm{yr^{-1}}$ and 3~years $\nu$-running and 3~years $\bar{\nu}$-running 
+is assumed. The fiducial mass of the Totally Liquid Scintillator Detector (TASD) is taken to be 
+$\mathrm{m_{det} = 25 \,kt}$ at a baseline of L~=~812~km approximately 12~km off-axis to the beamline. The
+energy resolution is $\mathrm{\sigma_e=10 \% \cdot \sqrt{E}}$ for electrons and  $\mathrm{\sigma_e=5 \% \cdot \sqrt{E}}$
+for muons. The following rules are defined within {\tt NOvA.glb}: 
+\begin{center}
+\begin{tabular}{|l|ll|c|c|}
+\hline \hline
+\multicolumn{3}{|l|}{Disappearance (+)} & $\sigma_\mathrm{norm}$ & $\sigma_\mathrm{cal}$ \\ \hline
+Signal & $0.8 \, \otimes \, (\nu_\mu\rightarrow\nu_\mu)_{\mathrm{CC}}$ & & 0.05 & 0.025 \\
+ & & & & \\
+Background & $0.0015 \, \otimes \, (\nu_\mu \rightarrow \nu_x)_\mathrm{NC}$ & & 0.05 & 0.025 \\ \hline \hline 
+\multicolumn{3}{|l|}{Appearance (+)} & & \\ \hline
+Signal & $0.24 \, \otimes \, (\nu_\mu \rightarrow \nu_e)_\mathrm{CC}$ & & 0.05 & 0.025\\
+ & & & & \\
+Background & $0.0015 \, \otimes \, (\nu_\mu \rightarrow \nu_x)_\mathrm{NC}$ & $1.0\cdot 10^{-4} \, \otimes \,
+(\nu_\mu\rightarrow\nu_\mu)_{\mathrm{CC}}$ & 0.05 & 0.025 \\
+Beam background & $0.12 \, \otimes \, (\nu_e\rightarrow \nu_e)_\mathrm{CC}$ & & 0.05 & 0.025 \\ \hline \hline
+\multicolumn{3}{|l|}{Disappearance (--)} &  &  \\ \hline
+Signal & $0.8 \, \otimes \, (\bar{\nu}_\mu\rightarrow\bar{\nu}_\mu)_{\mathrm{CC}}$ & & 0.05 & 0.025 \\
+ & & & & \\
+Background & $0.0015 \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_x)_\mathrm{NC}$ & & 0.05 & 0.025 \\ \hline \hline 
+\multicolumn{3}{|l|}{Appearance (--)} & & \\ \hline
+Signal & $0.37 \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_e)_\mathrm{CC}$ & & 0.05 & 0.025\\
+ & & & & \\
+Background & $0.0037 \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_x)_\mathrm{NC}$ & $1.0\cdot 10^{-4} \, \otimes \,
+(\bar{\nu}_\mu\rightarrow\bar{\nu}_\mu)_{\mathrm{CC}}$ & 0.05 & 0.025 \\
+Beam background & $0.12 \, \otimes \, (\bar{\nu}_e\rightarrow \bar{\nu}_e)_\mathrm{CC}$ & & 0.05 & 0.025 \\ \hline \hline
+\end{tabular}
+\end{center}
+
+\subsection*{SPL - {\tt SPL.glb}}
+
+The \SPL\ experiment can be simulated with the file {\tt SPL.glb}. This file was used in \cite{Campagne:2006yx}
+and follows the experiment description from \cite{Mezzetto:2003mm, Campagne:2004wt}. These references
+should be cited if the file {\tt SPL.glb} is used for a scientific publication or a talk. For calculations that
+involve {\tt SPL.glb}, the following additional files are required: 
+\bi
+\item {\tt SPLplus.dat} (neutrino flux from CERN -- $\nu_\mu$)
+\item {\tt SPLminus.dat} (neutrino flux from CERN -- $\bar{\nu}_\mu$)
+\item {\tt Mig\_WC\_numu.dat} (migration matrix -- $\nu_\mu$)
+\item {\tt Mig\_WC\_numubar.dat} (migration matrix -- $\bar{\nu}_\mu$)
+\item {\tt Mig\_WC\_nue.dat} (migration matrix -- $\nu_e$)
+\item {\tt Mig\_WC\_nuebar.dat} (migration matrix -- $\bar{\nu}_e$)
+\item {\tt XCC\_spl.dat} (charged current cross sections)
+\item {\tt XNC\_spl.dat} (neutral current cross sections)
+\ei
+The \SPL\ experiment uses a neutrino beam from the CERN to Fr\'{e}jus. The target power is 
+4~MV and 2~years $\nu$-running and 8~years $\bar{\nu}$-running 
+is assumed. The fiducial mass of the Water Cerenkov detector is taken to be 
+$\mathrm{m_{det} = 500 \,kt}$ at a baseline of L~=~130~km. The energy resolution is introduced manually by
+four migration matrices (for $\nu_e , \bar{\nu}_e ,\nu_{\mu} ,\bar{\nu}_\mu$) that describe energy smearing mainly
+from Fermi Motion. The following rules are defined within {\tt SPL.glb}: 
+\begin{center}
+\begin{tabular}{|l|ll|c|c|}
+\hline \hline
+\multicolumn{3}{|l|}{Disappearance (+)} & $\sigma_\mathrm{norm}$ & $\sigma_\mathrm{cal}$ \\ \hline 
+Signal & $1.0 \, \otimes \, (\nu_\mu\rightarrow\nu_\mu)_{\mathrm{CC}}$ & (energy dep. efficiency) & 0.02 & $10^{-4}$\\
+ & &  & &\\
+Background & $4.3\cdot 10^{-5} \, \otimes \, (\nu_\mu \rightarrow \nu_\mu)_\mathrm{CC}$ & & 0.02 & $10^{-4}$ \\ \hline \hline 
+\multicolumn{3}{|l|}{Appearance (+)} &  & \\ \hline
+Signal & $0.707 \, \otimes \, (\nu_\mu\rightarrow \nu_e)_\mathrm{CC}$ & & 0.02 & $10^{-4}$\\
+&&&&\\
+Background & $6.5\cdot 10^{-4} \, \otimes \, (\nu_\mu \rightarrow \nu_x)_\mathrm{NC}$ & $5.4\cdot 10^{-4} \, \otimes \, (\nu_\mu \rightarrow \nu_\mu)_\mathrm{CC}$ & 0.02 & $10^{-4}$ \\
+ & $0.7 \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_e )_\mathrm{CC}$ & & 0.02 &  $10^{-4}$ \\
+Beam Background & $0.677 \, \otimes \, (\bar{\nu}_e \rightarrow \bar{\nu}_e )_\mathrm{CC}$ & $0.707 \, \otimes \, (\nu_e \rightarrow \nu_e )_\mathrm{CC}$ & 0.02 & $10^{-4}$ \\ \hline \hline
+\end{tabular}
+\end{center}
+\begin{center}
+\begin{tabular}{|l|ll|c|c|}
+\hline \hline
+\multicolumn{3}{|l|}{Disappearance (--)} & & \\ \hline 
+Signal & $1.0 \, \otimes \, (\bar{\nu}_\mu\rightarrow\bar{\nu}_\mu)_{\mathrm{CC}}$ & (energy dep. efficiency) & 0.02 & $10^{-4}$\\
+ & &  & &\\
+Background & $4.3\cdot 10^{-5} \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_\mu)_\mathrm{CC}$ & & 0.02 & $10^{-4}$ \\ \hline \hline 
+\multicolumn{3}{|l|}{Appearance (--)} &  & \\ \hline
+Signal & $0.677 \, \otimes \, (\bar{\nu}_\mu\rightarrow \bar{\nu}_e)_\mathrm{CC}$ & & 0.02 & $10^{-4}$\\
+&&&&\\
+Background & $0.0025 \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_x)_\mathrm{NC}$ & $5.4\cdot 10^{-4} \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_\mu)_\mathrm{CC}$ & 0.02 & $10^{-4}$ \\
+ & $0.7 \, \otimes \, (\nu_\mu \rightarrow \nu_e )_\mathrm{CC}$ & & 0.02 &  $10^{-4}$ \\
+Beam Background & $0.677 \, \otimes \, (\bar{\nu}_e \rightarrow \bar{\nu}_e )_\mathrm{CC}$ & $0.707 \, \otimes \, (\nu_e \rightarrow \nu_e )_\mathrm{CC}$ & 0.02 & $10^{-4}$ \\ \hline \hline
+\end{tabular}
+\end{center}
+
+\section*{Reactor experiments}
+\subsection*{Small reactor experiment -- {\tt Reactor1.glb}}
+
+The file {\tt Reactor1.glb} allows to simulate a small $\bar{\nu}_e$-disappearance reactor experiment. The basic
+version of this file was used within \cite{Huber:2003pm} which should be cited if the file {\tt Reactor1.glb} is
+used for a scientific publication or a talk. For calculations that involve {\tt Reactor1.glb}, the following
+additional files are required:
+\bi
+\item {\tt Reactor.dat} (neutrino flux from reactor)
+\item {\tt XCCreactor.dat} (charged current cross sections for low energies)
+\ei
+The neutrino source is the core of a nuclear power reactor. The integrated luminosity is assumed to be
+$\mathrm{\mathcal{L} = 400 \,t \, GW\, yr}$, \eg\ a 20~t detector, a reactor with a thermal power of 4~GW
+and a running period of 5 years. As detector technology a liquid scintillator 
+detector is assumed, a far detector at a baseline of L~=~1.7~km and a near detector which is
+assumed to be identical to the far detector (maybe apart from the size) in order to minimize the impact of systematical uncertainties. The normalization error
+used in the file {\tt Reactor1.glb} has to be considered as an effective error, receiving contributions from
+individual uncertainties (see~\cite{Huber:2003pm}). The energy resolution is
+$\mathrm{\sigma_e=5\%\cdot\sqrt{E_{vis}}}$ and the choice for {\tt sigma\_function} is {\tt \#inverse\_beta}. The following rules are defined within {\tt Reactor1.glb}: 
+\begin{center}
+\begin{tabular}{|l|ll|c|c|}
+\hline \hline
+\multicolumn{3}{|l|}{Disappearance} & $\sigma_\mathrm{norm}$ & $\sigma_\mathrm{cal}$ \\ \hline 
+Signal & $1.0 \, \otimes \, (\bar{\nu}_e\rightarrow\bar{\nu}_e)_{\mathrm{CC}}$ & \hspace{5.5cm} & 0.008 & 0.005\\
+ & &  & &\\
+Background & $5.8\cdot 10^{-5} \, \otimes \, (\bar{\nu}_e\rightarrow\bar{\nu}_e)_\mathrm{CC}$ & \hspace{5.5cm} &
+$10^{-6}$ & $10^{-6}$ \\ \hline \hline 
+\end{tabular}
+\end{center}
+
+\subsection*{Large reactor experiment -- {\tt Reactor2.glb}}
+
+The file {\tt Reactor2.glb} allows to simulate a large $\bar{\nu}_e$-disappearance reactor experiment. The basic
+version of this file was used within \cite{Huber:2003pm} which should be cited if the file {\tt Reactor2.glb} is
+used for a scientific publication or a talk. The integrated luminosity is assumed to be
+$\mathrm{\mathcal{L} = 8000 \,t \, GW\, yr}$, \eg\ a 100~t detector, a reactor with a thermal power of 10~GW
+and a running period of 8 years. Besides the higher integrated luminosity, the attributes of {\tt Reactor2.glb}
+are similar to the ones of {\tt Reactor1.glb}.
+
+\subsection*{DoubleCHOOZ -- {\tt Dchooz-near.glb} and {\tt Dchooz-far.glb}}
+
+The files {\tt Dchooz-near.glb} and {\tt Dchooz-far.glb} allow to simulate the \DC\ reactor experiment in France. 
+The basic version of these files were used within \cite{Huber:2006vr} which should be cited if the files 
+{\tt Dchooz-near.glb} and {\tt Dchooz-far.glb} are
+used for a scientific publication or a talk. For calculations that involve {\tt Dchooz-near.glb} and/or {\tt Dchooz-far.glb}, the following
+additional files are required:
+\bi
+\item {\tt Reactor.dat} (neutrino flux from reactor)
+\item {\tt XCCreactor.dat} (charged current cross sections for low energies)
+\ei
+The \DC\ experiment is located at the Chooz reactor complex and the two reactor cores serve as $\bar{\nu}_e$
+neutrino source, so the thermal power is $\mathrm{2\cdot4.2\,GW}$. Two identical liquid scintillator detectors
+with a fiducial mass of $\mathrm{m_{det} = 10.16 \,t}$ are used as near and far detector. The far detector
+is planned to be located in the {\sc Chooz} cavern at a baseline of L~=~1.05~km from the two reactor cores and
+the near detector is assumed to be located at a distance of 0.1~km to the cores. The total running time of the
+experiment is assumed to be 5 years. So, the integrated luminosity at the far detector yields $\mathrm{\mathcal{L} \approx
+427 \,t \, GW\, yr}$.
+Here, the total running time of 5 years is assumed within {\tt Dchooz-near.glb} and {\tt Dchooz-far.glb}, so near
+and far detector are assumed to start the mode of operation simultaneously. For the simulation of \DC\ and
+considering a delayed start of data taking at the near detector the file {\tt Dchooz-near.glb} has to be
+modified. The cancellation of systematical uncertainties is considered by the manual definition of a $\chi^2$
+as described in \cite{Huber:2006vr} with the treatment of user-defined systematics as described in \Sec~\ref{sec:userchi}   
+The energy resolution is
+$\mathrm{\sigma_e=5\%\cdot\sqrt{E_{vis}}}$ and the choice for {\tt sigma\_function} is {\tt \#inverse\_beta}. The following rules are defined within {\tt Dchooz-near.glb} and {\tt Dchooz-far.glb}: 
+\begin{center}
+\begin{tabular}{|l|ll|c|c|c|c|c|}
+\hline \hline
+\multicolumn{3}{|l|}{Disappearance} & $\sigma_\mathrm{flux}$ & $\sigma^N_\mathrm{fid}$ & $\sigma^F_\mathrm{fid}$ &
+$\sigma^N_\mathrm{cal}$ & $\sigma^F_\mathrm{cal}$ \\ \hline 
+Signal & $1.0 \, \otimes \, (\bar{\nu}_e\rightarrow\bar{\nu}_e)_{\mathrm{CC}}$ & & 0.02 & 0.006 & 0.006 & 0.005 &
+0.005 \\
+ & & & & & & & \\
+Background & neglected & (sys. dominates) & -- & -- & -- & -- & -- \\ \hline \hline 
+\end{tabular}
+\end{center}
+
+\section*{Beta beam experiments}
+\subsection*{CERN-Fr\'{e}jus baseline scenario -- {\tt BB\_100.glb}}
+
+The $\gamma=100$ $\beta$-beam baseline scenario from CERN to Fr\'{e}jus can be simulated with the file {\tt
+BB\_100.glb}. The basic version of this file was used within \cite{Campagne:2006yx}. This reference
+should be cited if the file {\tt BB\_100.glb} is used for a scientific publication or a talk. For calculations that
+involve {\tt BB\_100.glb}, the following additional files are required: 
+\bi
+\item {\tt BB100flux\_Ne.dat} ($\beta$-beam neutrino flux -- $^{18}$Ne stored at $\gamma=100$)
+\item {\tt BB100flux\_He.dat} ($\beta$-beam neutrino flux -- $^{6}$He stored at $\gamma=100$)
+\item {\tt BeamBckg\_100.dat} (beam background)
+\item {\tt AtmBckg\_100.dat} (atmospheric background)
+\item {\tt Mig\_WC\_numu.dat} (migration matrix -- $\nu_\mu$)
+\item {\tt Mig\_WC\_numubar.dat} (migration matrix -- $\bar{\nu}_\mu$)
+\item {\tt Mig\_WC\_nue.dat} (migration matrix -- $\nu_e$)
+\item {\tt Mig\_WC\_nuebar.dat} (migration matrix -- $\bar{\nu}_e$)
+\item {\tt XCC\_Nuance.dat} (charged current cross sections)
+\item {\tt XNC\_Nuance.dat} (neutral current cross sections)
+\item {\tt Null.dat} (auxiliary file)
+\ei
+The neutrino beam is produced at CERN and directed towards a a megaton Water Cerenkov detector at Fr\'{e}jus. 
+The neutrinos originate from the decays of accelerated isotopes $^{18}$Ne ($\nu_e$) and $^6$He ($\bar{\nu}_e$).
+The acceleration factor is $\gamma=100$ for both types of isotopes and $2.2\cdot10^{18}$ $^{18}$Ne decays per
+year and $5.8\cdot10^{18}$ $^{6}$He decays per year are assumed. The CERN-Fr\'{e}jus baseline is L~=~130~km, the
+fiducial mass of the detector is $\mathrm{m_{det} = 500 \,kt}$ and 4~years $\nu$-running and 4~years
+$\bar{\nu}$-running are assumed. The energy resolution is introduced manually by
+four migration matrices (for $\nu_e , \bar{\nu}_e ,\nu_{\mu} ,\bar{\nu}_\mu$) that describes energy smearing mainly
+from Fermi Motion. The following rules are defined within {\tt BB\_100.glb}:
+\begin{center}
+\begin{tabular}{|l|ll|c|c|}
+\hline \hline
+\multicolumn{3}{|l|}{Disappearance -- $^{18}$Ne stored} & $\sigma_\mathrm{norm}$ & $\sigma_\mathrm{cal}$ \\ \hline
+Signal & $0.707 \, \otimes \, (\nu_e\rightarrow\nu_e)_{\mathrm{CC}}$ & & 0.02 & $10^{-4}$ \\
+ & & & & \\
+Background & $4.3\cdot 10^{-5} \, \otimes \, (\nu_e \rightarrow \nu_e)_\mathrm{CC}$ & & 0.02 & $10^{-4}$ \\ \hline \hline 
+\multicolumn{3}{|l|}{Appearance -- $^{18}$Ne stored} & & \\ \hline
+Signal & $1.0 \, \otimes \, (\nu_e \rightarrow \nu_\mu)_\mathrm{CC}$ & (energy dep. efficiency) & 0.02 & $10^{-4}$ \\
+ & & & & \\
+Background & $1.0 \, \otimes \, (\nu_e \rightarrow \nu_x)_\mathrm{NC}$ & (from external file) & 0.02 & $10^{-4}$ \\
+Atm. background & & (from external file) & 0.02 & $10^{-4}$ \\ \hline \hline
+\end{tabular}
+\end{center}
+\begin{center}
+\begin{tabular}{|l|ll|c|c|}
+\hline \hline
+\multicolumn{3}{|l|}{Disappearance -- $^6$He stored} & &  \\ \hline
+Signal & $0.677 \, \otimes \, (\bar{\nu}_e\rightarrow\bar{\nu}_e)_{\mathrm{CC}}$ & & 0.02 & $10^{-4}$ \\
+ & & & & \\
+Background & $4.3\cdot 10^{-5} \, \otimes \, (\bar{\nu}_e \rightarrow \nu_e)_\mathrm{CC}$ & & 0.02 & $10^{-4}$ \\ \hline \hline 
+\multicolumn{3}{|l|}{Appearance -- $^6$He stored} & & \\ \hline
+Signal & $1.0 \, \otimes \, (\bar{\nu}_e \rightarrow \bar{\nu}_\mu)_\mathrm{CC}$ & (energy dep. efficiency) & 0.02 & $10^{-4}$ \\
+ & & & & \\
+Background & $1.0 \, \otimes \, (\bar{\nu}_e \rightarrow \bar{\nu}_x)_\mathrm{NC}$ & (from external file) & 0.02 & $10^{-4}$ \\
+Atm. background & & (from external file) & 0.02 & $10^{-4}$ \\ \hline \hline
+\end{tabular}
+\end{center}
+
+\subsection*{Higher gamma scenario -- {\tt BB\_350.glb}}
+
+A $\gamma=350$ medium gamma $\beta$-beam green-field scenario involving a megaton Water Cerenkov detector can be simulated with the file {\tt
+BB\_350.glb}. This file tries to approximate as closely as possible the scenario {\sf Setup III} from
+\cite{Burguet-Castell:2005pa}. This reference
+should be cited if the file {\tt BB\_350.glb} is used for a scientific publication or a talk. For calculations that
+involve {\tt BB\_350.glb}, the following additional files are required: 
+\bi
+\item {\tt BB350flux.dat} ($\beta$-beam neutrino flux -- $\gamma=350$ )
+\item {\tt NeEffMig350.dat} (migration matrix -- $\nu_\mu$)
+\item {\tt NeBckgRej350.dat} (migration matrix -- background)
+\item {\tt NeDisEff350.dat} (migration matrix -- $\nu_e$)
+\item {\tt HeEffMig350.dat} (migration matrix -- $\bar{\nu}_\mu$)
+\item {\tt HeBckgRej350.dat} (migration matrix -- background)
+\item {\tt HeDisEff350.dat} (migration matrix -- $\bar{\nu}_e$)
+\item {\tt XCC.dat} (charged current cross sections)
+\item {\tt XNC.dat} (neutral current cross sections)
+\ei
+The neutrinos originate from the decays of accelerated isotopes $^{18}$Ne ($\nu_e$) and $^6$He ($\bar{\nu}_e$).
+The acceleration factor is $\gamma=350$ for both types of isotopes and $2.2\cdot10^{18}$ $^{18}$Ne decays per
+year and $5.8\cdot10^{18}$ $^{6}$He decays per year are assumed. The baseline is L~=~730~km, the
+fiducial mass of the detector is $\mathrm{m_{det} = 500 \,kt}$ and 4~years $\nu$-running and 4~years
+$\bar{\nu}$-running are assumed. The energy resolution is introduced manually by
+six migration matrices (for $\nu_e , \bar{\nu}_e ,\nu_{\mu} ,\bar{\nu}_\mu$, and the background from NC events
+for $^{18}$Ne and $^6$He) that describes energy smearing. These migration matrices also already include energy
+dependend efficiencies and background rejection factors. They are taken from the appendix of
+\cite{Burguet-Castell:2005pa}. The following rules are defined within {\tt BB\_350.glb}:
+\begin{center}
+\begin{tabular}{|l|ll|c|c|}
+\hline \hline
+\multicolumn{3}{|l|}{Disappearance -- $^{18}$Ne stored} & $\sigma_\mathrm{norm}$ & $\sigma_\mathrm{cal}$ \\ \hline
+Signal & $1.0 \, \otimes \, (\nu_e\rightarrow\nu_e)_{\mathrm{CC}}$ & (migration matrix) & 0.025 & $10^{-4}$ \\
+ & & & & \\
+Background & neglected & (systematic uncertainty dominates) & -- & -- \\ \hline \hline 
+\multicolumn{3}{|l|}{Appearance -- $^{18}$Ne stored} & & \\ \hline
+Signal & $1.0 \, \otimes \, (\nu_e \rightarrow \nu_\mu)_\mathrm{CC}$ & (migration matrix) & 0.025 & $10^{-4}$ \\
+ & & & & \\
+Background & $1.0 \, \otimes \, (\nu_e \rightarrow \nu_x)_\mathrm{NC}$ & (migration matrix) & 0.05 & $10^{-4}$
+\\ \hline \hline
+\multicolumn{3}{|l|}{Disappearance -- $^6$He stored} & &  \\ \hline
+Signal & $1.0 \, \otimes \, (\bar{\nu}_e\rightarrow\bar{\nu}_e)_{\mathrm{CC}}$ & (migration matrix) & 0.025 & $10^{-4}$ \\
+ & & & & \\
+Background & neglected & (systematic uncertainty dominates) & -- & -- \\ \hline \hline 
+\multicolumn{3}{|l|}{Appearance -- $^6$He stored} & & \\ \hline
+Signal & $1.0 \, \otimes \, (\bar{\nu}_e \rightarrow \bar{\nu}_\mu)_\mathrm{CC}$ & (migration matrix)  &
+0.025 & $10^{-4}$ \\
+ & & & & \\
+Background & $1.0 \, \otimes \, (\bar{\nu}_e \rightarrow \bar{\nu}_x)_\mathrm{NC}$ & (migration matrix) & 0.05 & $10^{-4}$ \\ \hline \hline
+\end{tabular}
+\end{center}
+
+\subsection*{Variable beta beam (Water Cerenkov) -- {\tt BBvar\_WC.glb}}
+
+A variable $\beta$-beam scenario involving a megaton Water Cerenkov detector can be simulated with the file 
+{\tt BBvar\_WC.glb}. The basic file was used within \cite{Huber:2005jk}. This reference
+should be cited if the file {\tt BBvar\_WC.glb} is used for a scientific publication or a talk. For calculations that
+involve {\tt BBvar\_WC.glb}, the following additional files are required:
+\bi
+\item {\tt BckgMig\_var.dat} (migration matrix -- background)
+\item {\tt XCC.dat} (charged current cross sections)
+\item {\tt XNC.dat} (neutral current cross sections)
+\item {\tt XQE.dat} (quasi elastic cross sections)
+\ei
+and the values of the following {\sf AEDL}-Variables have to be set:
+\bi
+\item {\tt gammafactor} (acceleration factor $\gamma$)
+\item {\tt EXP\_FACTOR} (parameter of ion decay scaling)
+\item {\tt baselinefactor} (baseline parameter L/$\gamma\,\left[\mathrm{km}\right]$)
+\ei
+The neutrinos originate from the decays of accelerated isotopes $^{18}$Ne ($\nu_e$) and $^6$He ($\bar{\nu}_e$).
+The acceleration factor is $\gamma=${\tt gammafactor} for both types of isotopes and
+$(100/\gamma)^\alpha\cdot2.2\cdot10^{18}$ $^{18}$Ne decays per
+year and $(60/\gamma)^\alpha\cdot5.8\cdot10^{18}$ $^{6}$He decays per year are assumed where $\alpha$~=~{\tt
+EXP\_FACTOR} is a parameter that describes ion decay scaling. As default value {\tt EXP\_FACTOR=0} should be chosen.
+See \cite{Huber:2005jk} for a detailed discussion of this parameter. The baseline is
+L~=~{\tt baselinefactor}$\cdot\gamma$~km, the
+fiducial mass of the detector is $\mathrm{m_{det} = 500 \,kt}$ and 4~years $\nu$-running and 4~years
+$\bar{\nu}$-running are assumed. The appearance measurement involves the total rates data from all CC events and the spectral data
+from the QE sample with a free normalization at an energy resolution of $\mathrm{\sigma_e=0.085GeV}$ due to the
+Fermi Motion identical to the treatment of systematics within the \TtoK\ and \TtoHK\ files. The following rules are defined within {\tt BBvar\_WC.glb}:
+\begin{center}
+\begin{tabular}{|l|ll|c|c|}
+\hline \hline
+\multicolumn{3}{|l|}{Disappearance -- $^{18}$Ne stored} & $\sigma_\mathrm{norm}$ & $\sigma_\mathrm{cal}$ \\ \hline
+Signal & $0.55 \, \otimes \, (\nu_e\rightarrow\nu_e)_{\mathrm{QE}}$ & \hspace{6cm} & 0.025 & $10^{-4}$ \\
+ & & & & \\
+Background & $0.003 \, \otimes \, (\nu_e \rightarrow \nu_x)_\mathrm{NC}$ & & 0.05 & $10^{-4}$ \\ \hline \hline 
+\multicolumn{3}{|l|}{Appearance -- $^{18}$Ne stored -- Spectrum} & & \\ \hline
+Signal & $0.55 \, \otimes \, (\nu_e \rightarrow \nu_\mu)_\mathrm{QE}$ & & 10.0 & $10^{-4}$ \\
+ & & & & \\
+Background & $0.003 \, \otimes \, (\nu_e \rightarrow \nu_x)_\mathrm{NC}$ & & 0.05 & $10^{-4}$
+\\ \hline \hline
+\multicolumn{3}{|l|}{Appearance -- $^{18}$Ne stored -- Total Rates} & & \\ \hline
+Signal & $0.55 \, \otimes \, (\nu_e \rightarrow \nu_\mu)_\mathrm{CC}$ & & 0.025 & $10^{-4}$ \\
+ & & & & \\
+Background & $0.003 \, \otimes \, (\nu_e \rightarrow \nu_x)_\mathrm{NC}$ & & 0.05 & $10^{-4}$
+\\ \hline \hline\multicolumn{3}{|l|}{Disappearance -- $^6$He stored} & &  \\ \hline
+Signal & $0.75 \, \otimes \, (\bar{\nu}_e\rightarrow\bar{\nu}_e)_{\mathrm{QE}}$ & & 0.025 & $10^{-4}$ \\
+ & & & & \\
+Background & $0.0025 \, \otimes \, (\bar{\nu}_e \rightarrow \bar{\nu}_x)_\mathrm{NC}$ & & 0.05 & $10^{-4}$ \\ \hline \hline 
+\multicolumn{3}{|l|}{Appearance -- $^6$He stored -- Spectrum} & & \\ \hline
+Signal & $0.75 \, \otimes \, (\bar{\nu}_e \rightarrow \bar{\nu}_\mu)_\mathrm{QE}$ & &
+10.0 & $10^{-4}$ \\
+ & & & & \\
+Background & $0.0025 \, \otimes \, (\bar{\nu}_e \rightarrow \bar{\nu}_x)_\mathrm{NC}$ & & 0.05 & $10^{-4}$ \\ \hline \hline
+\multicolumn{3}{|l|}{Appearance -- $^6$He stored -- Total Rates} & & \\ \hline
+Signal & $0.75 \, \otimes \, (\bar{\nu}_e \rightarrow \bar{\nu}_\mu)_\mathrm{CC}$ & &
+0.025 & $10^{-4}$ \\
+ & & & & \\
+Background & $0.0025 \, \otimes \, (\bar{\nu}_e \rightarrow \bar{\nu}_x)_\mathrm{NC}$ & & 0.05 & $10^{-4}$ \\ \hline \hline
+\end{tabular}
+\end{center}
+
+\subsection*{Variable beta beam (TASD) -- {\tt BBvar\_TASD.glb}}
+
+A variable $\beta$-beam scenario involving a \NOVA-like TASD detector can be simulated with the file 
+{\tt BBvar\_TASD.glb}. The basic file was used within \cite{Huber:2005jk}. This reference
+should be cited if the file {\tt BBvar\_TASD.glb} is used for a scientific publication or a talk. For calculations that
+involve {\tt BBvar\_TASD.glb}, the following additional files are required:
+\bi
+\item {\tt XCC.dat} (charged current cross sections)
+\item {\tt XNC.dat} (neutral current cross sections)
+\ei
+and the values of the following {\sf AEDL}-Variables have to be set:
+\bi
+\item {\tt gammafactor} (acceleration factor $\gamma$)
+\item {\tt EXP\_FACTOR} (parameter of ion decay scaling)
+\item {\tt baselinefactor} (baseline parameter L/$\gamma\,\left[\mathrm{km}\right]$)
+\ei
+The neutrinos originate from the decays of accelerated isotopes $^{18}$Ne ($\nu_e$) and $^6$He ($\bar{\nu}_e$).
+The acceleration factor is $\gamma=${\tt gammafactor} for both types of isotopes and
+$(100/\gamma)^\alpha\cdot2.2\cdot10^{18}$ $^{18}$Ne decays per
+year and $(60/\gamma)^\alpha\cdot5.8\cdot10^{18}$ $^{6}$He decays per year are assumed where $\alpha$~=~{\tt
+EXP\_FACTOR} is a parameter that describes ion decay scaling. As default value {\tt EXP\_FACTOR=0} should be chosen.
+See \cite{Huber:2005jk} for a detailed discussion of this parameter. The baseline is L~=~{\tt baselinefactor}$\cdot\gamma$~km, the
+fiducial mass of the detector is $\mathrm{m_{det} = 50 \,kt}$ and 4~years $\nu$-running and 4~years
+$\bar{\nu}$-running are assumed. The energy resolution is $\mathrm{\sigma_e=6 \% \cdot \sqrt{E}}$ for electrons and  $\mathrm{\sigma_e=3 \% \cdot \sqrt{E}}$
+for muons.  The following rules are defined within {\tt BBvar\_TASD.glb}:
+\begin{center}
+\begin{tabular}{|l|ll|c|c|}
+\hline \hline
+\multicolumn{3}{|l|}{Disappearance -- $^{18}$Ne stored} & $\sigma_\mathrm{norm}$ & $\sigma_\mathrm{cal}$ \\ \hline
+Signal & $0.2 \, \otimes \, (\nu_e\rightarrow\nu_e)_{\mathrm{CC}}$ & \hspace{6cm} & 0.025 & $10^{-4}$ \\
+ & & & & \\
+Background & $0.001 \, \otimes \, (\nu_e \rightarrow \nu_x)_\mathrm{NC}$ & & 0.05 & $10^{-4}$ \\ \hline \hline 
+\multicolumn{3}{|l|}{Appearance -- $^{18}$Ne stored} & & \\ \hline
+Signal & $0.8 \, \otimes \, (\nu_e \rightarrow \nu_\mu)_\mathrm{CC}$ & & 0.025 & $10^{-4}$ \\
+ & & & & \\
+Background & $0.001 \, \otimes \, (\nu_e \rightarrow \nu_x)_\mathrm{NC}$ & & 0.05 & $10^{-4}$
+\\ \hline \hline
+\multicolumn{3}{|l|}{Disappearance -- $^6$He stored} & &  \\ \hline
+Signal & $0.2 \, \otimes \, (\bar{\nu}_e\rightarrow\bar{\nu}_e)_{\mathrm{CC}}$ & & 0.025 & $10^{-4}$ \\
+ & & & & \\
+Background & $0.001 \, \otimes \, (\bar{\nu}_e \rightarrow \bar{\nu}_x)_\mathrm{NC}$ & & 0.05 & $10^{-4}$  \\ \hline \hline 
+\multicolumn{3}{|l|}{Appearance -- $^6$He stored} & & \\ \hline
+Signal & $0.8 \, \otimes \, (\bar{\nu}_e \rightarrow \bar{\nu}_\mu)_\mathrm{CC}$ & &
+0.025 & $10^{-4}$ \\
+ & & & & \\
+Background & $0.001 \, \otimes \, (\bar{\nu}_e \rightarrow \bar{\nu}_x)_\mathrm{NC}$ & & 0.05 & $10^{-4}$ \\ \hline \hline
+\end{tabular}
+\end{center}
+
+\section*{Neutrino factory experiments}
+\subsection*{Standard neutrino factory -- {\tt NFstandard.glb}}
+
+A standard neutrino factory scenario can be simulated with the file {\tt NFstandard.glb}. The basic version was
+used within \cite{Huber:2002mx} (NuFact-II scenario). This reference should be cited if the file {\tt
+NFstandard.glb} is used for a scientific publication or a talk. For calculations that involve {\tt
+NFstandard.glb}, the following additional files are required:
+\bi
+\item {\tt XCC.dat} (charged current cross sections)
+\item {\tt XNC.dat} (neutral current cross sections)
+\ei
+The neutrino beam is produced by the decay of muons stored in a storage ring at a parent energy of
+$\mathrm{E_\mu = 50\,GeV}$. A number of $1.06\cdot10^{21}$ useful muon decays per year are assumed in each
+polarity (corresponding to $5.3\cdot10^{20}$ useful muon decays per year and polarity for simultaneous operation
+with both polarities) and 4~years $\nu$-running and 4~years $\bar{\nu}$-running is
+assumed. The fiducial mass of the MID detector is taken to be $\mathrm{m_{det} = 50 \,kt}$ at a baseline of
+L~=~3000~km. The energy resolution is $\mathrm{\sigma_e=15\%\cdot E}$. The following rules are defined within
+{\tt NFstandard.glb}:
+\begin{center}
+\begin{tabular}{|l|ll|c|c|}
+\hline \hline
+\multicolumn{3}{|l|}{Disappearance -- $\mu^+$-stored} & $\sigma_\mathrm{norm}$ & $\sigma_\mathrm{cal}$ \\ \hline
+Signal & $0.35 \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_\mu)_\mathrm{CC}$ & & 0.001 & $10^{-4}$ \\
+ & & & & \\
+Background & $1.0\cdot 10^{-5} \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_x)_\mathrm{NC}$ & & 0.2 & $10^{-4}$ \\ \hline \hline
+\multicolumn{3}{|l|}{Appearance -- $\mu^+$-stored} & & \\ \hline
+Signal &  $0.45 \, \otimes \, (\nu_e \rightarrow \nu_\mu)_\mathrm{CC}$ & & 0.001 & $10^{-4}$ \\
+ & & & & \\
+Background &  $5.0\cdot 10^{-6} \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_x)_\mathrm{NC}$ &  $5.0\cdot
+10^{-6} \, \otimes \, (\bar{\nu}_\mu \rightarrow\bar{\nu}_\mu)_\mathrm{CC}$ & 0.2& $10^{-4}$\\ \hline \hline
+\multicolumn{3}{|l|}{Disappearance -- $\mu^-$-stored} & & \\ \hline
+Signal &  $0.45 \, \otimes \, (\nu_\mu \rightarrow \nu_\mu)_\mathrm{CC}$ & & 0.001& $10^{-4}$\\
+ & & & & \\
+Background &  $1.0\cdot 10^{-5} \, \otimes \, (\nu_\mu \rightarrow \nu_x)_\mathrm{NC}$ & & 0.2& $10^{-4}$\\ \hline \hline
+\multicolumn{3}{|l|}{Appearance -- $\mu^-$-stored} & & \\ \hline
+Signal & $0.35 \, \otimes \, (\bar{\nu}_e \rightarrow \bar{\nu}_\mu)_\mathrm{CC}$  & & 0.001& $10^{-4}$\\
+ & & & & \\
+Background &  $5.0\cdot 10^{-6} \, \otimes \, (\nu_\mu \rightarrow \nu_x)_\mathrm{NC}$ & $5.0\cdot 10^{-6} \, \otimes \, (\nu_\mu \rightarrow
+\nu_\mu)_\mathrm{CC}$  & 0.2& $10^{-4}$\\ \hline \hline
+\end{tabular}
+\end{center}
+
+\subsection*{Variable neutrino factory -- {\tt NFvar.glb}}
+
+A variable neutrino factory scenario can be simulated with the file {\tt NFvar.glb}. The basic version was
+used within \cite{Huber:2006wb} and follows the neutrino factory scenarios from \cite{Huber:2002mx}. These
+references should be cited if the file {\tt
+NFvar.glb} is used for a scientific publication or a talk. For calculations that involve {\tt
+NFvar.glb}, the following additional files are required:
+\bi
+\item {\tt ThresholdEnergies.dat} (list of energies)
+\item {\tt CIDthreshold.dat} (list for CID threshold)
+\item {\tt MINOSthreshold.dat} (list for MINOS threshold)
+\item {\tt XCC.dat} (charged current cross sections)
+\item {\tt XNC.dat} (neutral current cross sections)
+\ei
+and the values of the following {\sf AEDL}-Variables have to be set:
+\bi
+\item {\tt emax} (parent energy of the stored muons $\left[\mathrm{km}\right]$)
+\item {\tt BASELINE} (experiment baseline $\left[\mathrm{GeV}\right]$)
+\ei
+The neutrino beam is produced by the decay of muons stored in a storage ring at a parent energy of
+$\mathrm{E_\mu =}$~{\tt emax}. The parent energy of the muons can be set in the range
+$\mathrm{10\,GeV\lesssim E_\mu\lesssim 80\, GeV}$. The baseline of the scenario is L~=~{\tt BASELINE}. 
+Besides these settings the other attributes of {\tt NFvar.glb} are similar to the ones from {\tt
+NFstandard.glb}. Only the treatment of the disappearance channels is different. Whereas the CID threshold is
+also applied in the disappearance channels of {\tt NFstandard.glb} this is not the case for the disapperance
+channel definition within {\tt NFvar.glb} where no CID is applied and a threshold similar to the threshold of
+the MINOS experiment \cite{Ables:1995wq} is applied. The following rules are defined within
+{\tt NFvar.glb}:
+\begin{center}
+\begin{tabular}{|l|ll|c|c|}
+\hline \hline
+\multicolumn{3}{|l|}{Disappearance -- $\mu^+$-stored} & $\sigma_\mathrm{norm}$ & $\sigma_\mathrm{cal}$ \\ \hline
+Signal & $0.9 \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_\mu)_\mathrm{CC}$ & $0.9 \, \otimes \, (\nu_e \rightarrow \nu_\mu)_\mathrm{CC}$& 0.025 & $10^{-4}$ \\
+ & & & & \\
+Background & $1.0\cdot 10^{-5} \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_x)_\mathrm{NC}$ & & 0.2 & $10^{-4}$ \\ \hline \hline
+\multicolumn{3}{|l|}{Appearance -- $\mu^+$-stored} & & \\ \hline
+Signal &  $0.45 \, \otimes \, (\nu_e \rightarrow \nu_\mu)_\mathrm{CC}$ & & 0.025 & $10^{-4}$ \\
+ & & & & \\
+Background &  $5.0\cdot 10^{-6} \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_x)_\mathrm{NC}$ &  $5.0\cdot
+10^{-6} \, \otimes \, (\bar{\nu}_\mu \rightarrow\bar{\nu}_\mu)_\mathrm{CC}$ & 0.2& $10^{-4}$\\ \hline \hline
+\multicolumn{3}{|l|}{Disappearance -- $\mu^-$-stored} & & \\ \hline
+Signal &  $0.9 \, \otimes \, (\nu_\mu \rightarrow \nu_\mu)_\mathrm{CC}$ & $0.9 \, \otimes \, (\bar{\nu}_e
+\rightarrow \bar{\nu}_\mu)_\mathrm{CC}$& 0.025& $10^{-4}$\\
+ & & & & \\
+Background &  $1.0\cdot 10^{-5} \, \otimes \, (\nu_\mu \rightarrow \nu_x)_\mathrm{NC}$ & & 0.2& $10^{-4}$\\ \hline \hline
+\multicolumn{3}{|l|}{Appearance -- $\mu^-$-stored} & & \\ \hline
+Signal & $0.35 \, \otimes \, (\bar{\nu}_e \rightarrow \bar{\nu}_\mu)_\mathrm{CC}$  & & 0.025& $10^{-4}$\\
+ & & & & \\
+Background &  $5.0\cdot 10^{-6} \, \otimes \, (\nu_\mu \rightarrow \nu_x)_\mathrm{NC}$ & $5.0\cdot 10^{-6} \, \otimes \, (\nu_\mu \rightarrow
+\nu_\mu)_\mathrm{CC}$  & 0.2& $10^{-4}$\\ \hline \hline
+\end{tabular}
+\end{center}
+
+\subsection*{Variable neutrino factory with Silver Channel -- {\tt NFvar\_GoldSilver.glb}}
+
+A variable neutrino factory scenario that includes the golden and silver appearance channels can be 
+simulated with the file {\tt NFvar\_GoldSilver.glb}. The basic version was
+used within \cite{Huber:2006wb} and the golden channel follows the neutrino factory scenarios from \cite{Huber:2002mx}
+and the description of the silver channel follows \cite{Autiero:2003fu}. These
+references should be cited if the file {\tt
+NFvar\_GoldSilver.glb} is used for a scientific publication or a talk. For calculations that involve {\tt
+NFvar\_GoldSilver.glb}, the following additional files are required:
+\bi
+\item {\tt ThresholdEnergies.dat} (list of energies)
+\item {\tt CIDthreshold.dat} (list for CID threshold)
+\item {\tt MINOSthreshold.dat} (list for MINOS threshold)
+\item {\tt ECCthreshold.dat} (list for ECC threshold)
+\item {\tt XCC.dat} (charged current cross sections)
+\item {\tt XNC.dat} (neutral current cross sections)
+\ei
+and the values of the following {\sf AEDL}-Variables have to be set:
+\bi
+\item {\tt emax} (parent energy of the stored muons $\left[\mathrm{km}\right]$)
+\item {\tt BASELINE} (experiment baseline $\left[\mathrm{GeV}\right]$)
+\ei
+The beam and golden channel attributes are similar to {\tt NFvar.glb}. The parent energy of the muons can be set
+in the range $\mathrm{10\,GeV\lesssim E_\mu\lesssim 80\, GeV}$. The baseline is set by the {\sf
+AEDL}-Variable {\tt BASELINE}. For the silver channel an additional ECC detector with a fiducial mass
+$\mathrm{m_{ECC} = 5\,kt}$ is assumed to be located at the same baseline as the MID detector. The energy
+resolution of the silver channel is set to $\mathrm{\sigma_e=20\%\cdot E}$. The following additional rule
+compared to {\tt NFvar.glb} is introduced to {\tt NFvar\_GoldSilver.glb}:
+\begin{center}
+\begin{tabular}{|l|ll|c|c|}
+\hline \hline
+\multicolumn{3}{|l|}{$\tau$-Appearance -- $\mu^+$-stored} & $\sigma_\mathrm{norm}$ & $\sigma_\mathrm{cal}$ \\ \hline
+Signal & $0.096 \, \otimes \, (\nu_e \rightarrow \nu_\tau)_\mathrm{CC}$ & & 0.15 & $10^{-4}$ \\
+ & & & & \\
+Background & $3.1\cdot 10^{-8} \, \otimes \, (\nu_e \rightarrow \nu_e)_\mathrm{CC}$ & $2.0\cdot 10^{-8} \,
+\otimes \, (\nu_e \rightarrow \nu_\mu)_\mathrm{CC}$& 0.2 & $10^{-4}$ \\ 
+ & $3.7\cdot 10^{-6} \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_\mu)_\mathrm{CC}$ & $1.0\cdot 10^{-3} \,
+ \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_\tau)_\mathrm{CC}$& 0.2 & $10^{-4}$ \\ 
+ & $7.0 \cdot 10^{-7} \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_x)_\mathrm{NC}$ & $7.0\cdot 10^{-7} \, \otimes \, (\nu_e \rightarrow \nu_x)_\mathrm{NC}$& 0.2 & $10^{-4}$ \\ \hline \hline
+\end{tabular}
+\end{center}
+
+\subsection*{High Resolution/Low Threshold neutrino factory -- \\{\tt NFvar\_hR\_lT.glb}}
+
+A variable neutrino factory hybrid detector scenario can be simulated with the file {\tt NFvar\_hR\_lT.glb}. The basic version was
+used within \cite{Huber:2006wb} and follows the neutrino factory scenarios from \cite{Huber:2002mx}. These
+references should be cited if the file {\tt NFvar\_hR\_lT.glb} is used for a scientific publication or a talk.
+For calculations that involve {\tt NFvar\_hR\_lT.glb}, the same additional files as for {\tt NFvar.glb} are
+required and the {\sf AEDL}-Variables {\tt emax} and {\tt BASELINE} have to be set. {\tt NFvar\_hR\_lT.glb} 
+implements a lower threshold ($\mathrm{\sim\,1\,GeV}$) at a higher energy resolution
+$\mathrm{\sigma_e=15\%\cdot E + 0.085\, MeV}$ where the constant term represents 
+the effects from Fermi Motion. The background rejection is introduced energy dependent as 
+$10^{-3}/\mathrm{E^2}$ and matches {\tt NFvar.glb} at higher energies. The following rules are defined within {\tt NFvar\_hR\_lT.glb}: 
+\begin{center}
+\begin{tabular}{|l|ll|c|c|}
+\hline \hline
+\multicolumn{3}{|l|}{Disappearance -- $\mu^+$-stored} & $\sigma_\mathrm{norm}$ & $\sigma_\mathrm{cal}$ \\ \hline
+Signal & $0.9 \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_\mu)_\mathrm{CC}$ & $0.9 \, \otimes \, (\nu_e \rightarrow \nu_\mu)_\mathrm{CC}$& 0.025 & $10^{-4}$ \\
+ & & & & \\
+Background & $1.0 \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_x)_\mathrm{NC}$ & & 0.2 & $10^{-4}$ \\ \hline \hline
+\multicolumn{3}{|l|}{Appearance -- $\mu^+$-stored} & & \\ \hline
+Signal &  $0.5 \, \otimes \, (\nu_e \rightarrow \nu_\mu)_\mathrm{CC}$ & & 0.025 & $10^{-4}$ \\
+ & & & & \\
+Background &  $1.0 \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_x)_\mathrm{NC}$ & (energy dep. rejection)
+& 0.2 & $10^{-4}$\\
+ &  $1.0 \, \otimes \, (\bar{\nu}_\mu \rightarrow\bar{\nu}_\mu)_\mathrm{CC}$ & (energy dep. rejection)
+&0.2& $10^{-4}$\\ \hline \hline
+\multicolumn{3}{|l|}{Disappearance -- $\mu^-$-stored} & & \\ \hline
+Signal &  $0.9 \, \otimes \, (\nu_\mu \rightarrow \nu_\mu)_\mathrm{CC}$ & $0.9 \, \otimes \, (\bar{\nu}_e
+\rightarrow \bar{\nu}_\mu)_\mathrm{CC}$& 0.025& $10^{-4}$\\
+ & & & & \\
+Background &  $1.0\cdot 10^{-5} \, \otimes \, (\nu_\mu \rightarrow \nu_x)_\mathrm{NC}$ & & 0.2& $10^{-4}$\\ \hline \hline
+\multicolumn{3}{|l|}{Appearance -- $\mu^-$-stored} & & \\ \hline
+Signal & $0.5 \, \otimes \, (\bar{\nu}_e \rightarrow \bar{\nu}_\mu)_\mathrm{CC}$  & & 0.025& $10^{-4}$\\
+ & & & & \\
+Background &  $1.0 \, \otimes \, (\nu_\mu \rightarrow \nu_x)_\mathrm{NC}$ & (energy dep. rejection)
+& 0.2 & $10^{-4}$\\
+ & $1.0 \, \otimes \, (\nu_\mu \rightarrow
+\nu_\mu)_\mathrm{CC}$  & (energy dep. rejection)
+&0.2& $10^{-4}$\\ \hline \hline
+\end{tabular}
+\end{center}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \chapter{Flux normalization in \GLOBES }

### 2007-01-11 Documentation/part1.tex
*** empty log message ***

@@
 \hline
-Experiment & Filename & Short description &  Ref. \\
+Experiment & Filename & Short description &  Ref. \\[0.2cm]
 \hline 
-\multicolumn{4}{l}{\underline{Conventional beams:}} \\
-MINOS & {\tt MINOS.glb} & MINOS exp., 5~yr running time & \cite{Huber:2004ug}\\
-OPERA & {\tt OPERA.glb} & OPERA exp., 5~yr running time & \cite{Huber:2004ug}\\
-ICARUS & {\tt ICARUS.glb} & ICARUS exp., 5~yr running time & \cite{Huber:2004ug}\\[0.1cm]
-
-\multicolumn{4}{l}{\underline{First-generation superbeams:}} \\
-T2K & {\tt JHFSKnew.glb} & J-PARC to Super-K, 5~yr $\nu$-running &  \cite{Huber:2002rs} \\
- & {\tt JHFSKantinew.glb} & J-PARC to Super-K, 5~yr $\bar{\nu}$-running &  \cite{Huber:2002rs} \\
- & {\tt JHFSKcomb.glb} & Same, but 1.25~yr $\nu$- and 3.75~yr $\bar{\nu}$-running &  \cite{Huber:2002rs} \\[0.1cm]
-NO$\nu$A & {\tt NUMI9.glb} & NuMI OA $9 \, \mathrm{km}/712 \, \mathrm{km}$, 5~yr $\nu$-running & \cite{Huber:2002rs} \\
-& {\tt NUMI9anti.glb} & NuMI OA $9 \, \mathrm{km}/712 \, \mathrm{km}$, 5~yr $\bar{\nu}$-running & \cite{Huber:2002rs} \\
-& {\tt NUMI9comb.glb} & NuMI OA $9 \, \mathrm{km}/712 \, \mathrm{km}$, 1.43~yr $\nu$- and 3.57~yr $\bar{\nu}$-running & \cite{Huber:2002rs} \\
-& {\tt NUMI12.glb} & NuMI OA $12 \, \mathrm{km}/712 \, \mathrm{km}$, 5~yr $\nu$-running & \cite{Huber:2002rs} \\
-& {\tt NUMI12anti.glb} & NuMI OA $12 \, \mathrm{km}/712 \, \mathrm{km}$, 5~yr $\bar{\nu}$-running & \cite{Huber:2002rs} \\
-& {\tt NUMI12comb.glb} & NuMI OA $12 \, \mathrm{km}/712 \, \mathrm{km}$, 1.43~yr $\nu$- and 3.57~yr $\bar{\nu}$-running & \cite{Huber:2002rs} \\[0.1cm]
-
-\multicolumn{4}{l}{\underline{Superbeam upgrade:}} \\
-J-PARC-HK & {\tt JHFHKAll.glb} & J-PARC to Hyper-K, 2~yr $\nu$- and 6~yr $\bar{\nu}$-running &  \cite{Huber:2002mx} \\[0.1cm]
-
-\multicolumn{4}{l}{\underline{Neutrino factories:}} \\
-\NuFactI\ & {\tt NuFact1.glb} & Initial stage NF, 2$\times$2.5~yr running time (each pol.), $m_{\mathrm{Det}} = 10 \, \mathrm{kt}$,
-$P_{\mathrm{Tg}} \simeq 0.75 \, \mathrm{MW}$ & \cite{Huber:2002mx} \\
-\NuFactII\  & {\tt NuFact2.glb} & Advanced stage NF, 2$\times$4~yr running time (each pol.), $m_{\mathrm{Det}} = 50 \, \mathrm{kt}$,
-$P_{\mathrm{Tg}} \simeq  4 \, \mathrm{MW}$ & \cite{Huber:2002mx} \\[0.1cm]
-
-\multicolumn{4}{l}{\underline{Reactor experiments:}} \\
-\ReactorI\ & {\tt Reactor1.glb} & Small reactor exp., $\mathcal{L} = 400 \, \mathrm{t} \, \mathrm{GW} \, \mathrm{yr}$ & \cite{Huber:2003pm} \\
-\ReactorII\ & {\tt Reactor2.glb} & Large reactor exp., $\mathcal{L} = 8\, 000 \, \mathrm{t} \, \mathrm{GW} \, \mathrm{yr}$ & \cite{Huber:2003pm} \\[0.1cm]
-
-\underline{$\beta$-Beams:} & \multicolumn{3}{l}{In preparation} \\[0.1cm]
+\multicolumn{4}{l}{\underline{Superbeam experiments:}} \\[0.2cm]
+
+T2K & {\tt T2K.glb} & J-PARC to Super-K, 2~yr $\nu$- and 6~yr 
+$\bar{\nu}$-running &  \cite{Itow:2001ee,Huber:2002mx} \\[0.2cm]
+
+T2HK & {\tt T2HK.glb} & J-PARC to Hyper-K, 4~yr $\nu$- and 4~yr 
+$\bar{\nu}$-running (T2K upgrade) &  \cite{Itow:2001ee,Huber:2002mx} \\[0.2cm]
+
+NO$\nu$A & {\tt NOvA.glb} & FermiLab NuMI beamline off-axis, 3~yr $\nu$- and 
+3~yr $\bar{\nu}$-running  & \cite{Ambats:2004js,Yang_2004} \\[0.2cm]
+
+SPL & {\tt SPL.glb} & CERN to Fr\'{e}jus, 2~yr $\nu$- and 8~yr 
+$\bar{\nu}$-running & \cite{Campagne:2006yx,Campagne:2004wt,Mezzetto:2003mm} \\[0.2cm]
+
+\multicolumn{4}{l}{\underline{Reactor experiments:}} \\[0.2cm]
+
+{\sc Reactor-I} & {\tt Reactor1.glb} & Small reactor exp., 
+$\mathcal{L} = 400 \, \mathrm{t} \, \mathrm{GW} \, \mathrm{yr}$
+& \cite{Huber:2003pm} \\[0.2cm]
+
+{\sc Reactor-II} & {\tt Reactor2.glb} & Large reactor exp., 
+$\mathcal{L} = 8\, 000 \, \mathrm{t} \, \mathrm{GW} \, \mathrm{yr}$ 
+& \cite{Huber:2003pm} \\[0.2cm]
+
+Double{\sc Chooz} & {\tt D-Chooz\_near.glb} & DoubleCHOOZ near 
+detector, 5 years data taking & \cite{Huber:2006vr} \\ 
+ & {\tt D-Chooz\_far.glb} & DoubleCHOOZ near detector, 5 years data taking & \\[0.2cm]
+
+\multicolumn{4}{l}{\underline{$\beta$-Beams:}} \\[0.2cm]
+
+low $\gamma$ & {\tt BB\_100.glb} & $\gamma=100$ CERN to Fr\'{e}jus baseline 
+scenario, 4~yr $\nu$- and 4~yr $\bar{\nu}$-running & \cite{Campagne:2006yx} \\[0.2cm]
+
+medium $\gamma$ & {\tt BB\_350.glb} & $\gamma=350$ frogreen field scenario, 4~yr $\nu$- and
+4~yr $\bar{\nu}$-running & \cite{Burguet-Castell:2005pa} \\[0.2cm]
+
+Variable & {\tt BBvar\_WC.glb} & Variable $\beta$-beam, 4~yr $\nu$- and 
+4~yr $\bar{\nu}$-running, {\sf AEDL}-Variables {\tt gammafactor}, 
+{\tt EXP\_FACTOR} and {\tt baselinefactor}, 500~kt WC detector & \cite{Huber:2005jk} \\[0.2cm]
+
+	& {\tt BBvar\_TASD.glb} & like {\tt BBvar\_WC.glb} but with a 50~kt TASD detector  
+& \cite{Huber:2005jk} \\[0.2cm]
+
+\multicolumn{4}{l}{\underline{Neutrino factories:}} \\[0.2cm]
+
+Standard & {\tt NFstandard.glb} & Standard neutrino factory, 
+4~yr $\nu$- and 4~yr $\bar{\nu}$-running & \cite{Huber:2002mx} \\[0.2cm]
+
+Variable  & {\tt NFvar.glb} & Variable neutrino factory, 4~yr 
+$\nu$- and 4~yr $\bar{\nu}$-running, {\sf AEDL}-Variables {\tt emax} and {\tt BASELINE}, disappearance channels
+{\em without} CID & \cite{Huber:2002mx,Huber:2006wb}  \\[0.2cm]
+
+Gold + Silver & {\tt NF\_Goldsilver.glb} & as {\tt NFvar.glb} 
+plus 5~kt ECC detector for Silver Channel
+measurement & \cite{Huber:2002mx,Huber:2006wb,Autiero:2003fu} \\[0.2cm]
+
+Hybrid detector & {\tt NF\_hR\_lT.glb} & like {\tt NFvar.glb} 
+with lower threshold and higher energy resolution & \cite{Huber:2002mx,Huber:2006wb} \\[0.2cm]
+
 \hline
@@
 \index{norm}{glb files@{\tt glb}-files} 
-TO BE UPDATED BY MR; POSSIBLY MORE DETAILED DESCRIPTION IN APPENDIX [WW].
-Pre-defined experiment prototypes, their filenames (to be used in \GLB{InitExperiment}), their short descriptions, and the references in which they are originally used and discussed (except from minor modifications, such as a
-different implementation of the energy threshold function). Note that
-some of these experiments are outdated in terms of integrated luminosities,
-baseline, fluxes, efficiencies, or other factors. In any case these file are installed along with \GLOBES.}
+Pre-defined experiment prototypes, their filenames (to be used in \GLB{InitExperiment}), 
+their short descriptions, and the references in which they are originally used and 
+discussed (except from minor modifications, such as a different implementation of 
+the energy threshold function). Note that some of these experiments are outdated 
+in terms of integrated luminosities, baseline, fluxes, efficiencies, or other factors. 
+In any case these file are installed along with \GLOBES.}
 \end{table}

### 2007-01-12 Documentation/Paper2/paper2.tex
*** empty log message ***

@@
 \newcommand{\GLOBES}{{\sf GLoBES}}
+\newcommand{\GLOBESN}{{\sf GLoBES~3.0}}
 \newcommand{\AEDL}{{\sf AEDL}}
-\newcommand{\EDM}{{\sf EDM}}
+%\newcommand{\EDM}{{\sf EDM}}
 
@@
 {\Large\bf New features in the simulation of neutrino oscillation experiments with GLoBES 3.0}
+
+(or: ``GLoBES 3.0 release: From user-defined systematics to the simulation of non-standard physics'')
 \end{center}}
@@
 We present Version 3.0 of the GLoBES (``General Long Baseline Experiment Simulator'') software, which is
-a simulation tool for short- and long-baseline neutrino oscillation experiments.
-As new features, \GLOBES\ 3.0 allows the simulation of experiments with multiple detectors by the concept of user-defined systematics.
+a simulation tool for short- and long-baseline neutrino oscillation experiments. As new features, \GLOBESN\ allows the simulation of experiments with multiple discrete sources and detectors by the 
+concept of user-defined systematics.
 In addition, the combination with external information, such as from different experiment classes, is 
@@
 and improvements are necessary. In this work, we present the most important new
-features and changes in \GLOBES\ 3.0. For experimentalists, \GLOBES\ now allows
+features and changes in \GLOBESN . For experimentalists, \GLOBES\ now allows
 for the implementation of arbitrary systematics, which can also be used for
@@
 significant time dependencies can not be studied, such as  supernov\ae. 
-However, in \GLOBES\ 3.0 and higher, new flexibility is introduced by the
+However, in \GLOBESN\ and higher, new flexibility is introduced by the
 concept of user-defined systematics. This new feature allows the cross-definition
@@
 complete experiment simulations.
-% Examples for these prototypes are the \MINOS , \ICARUS , and
-% \OPERA\ simulations from \Ref~\cite{Huber:2004ug}, the
-% \JHFSK\ and \NUMI\ superbeam simulations from \Refs~\cite{Huber:2002mx,
-% Huber:2002rs}, the \JHFHK\ superbeam upgrade simulation from
-% \Ref~\cite{Huber:2002mx}, the neutrino factory simulations from
-% \Ref~\cite{Huber:2002mx}, and the reactor experiment simulations from
-% \Ref~\cite{Huber:2003pm}. Other projects using earlier versions
-% of the \GLOBES\ software include \Refs~\cite{Huber:2003ak,Ohlsson:2003ip,Winter:2003ye,Antusch:2004yx}.
 
@@
 
-\section{User-defined systematics and changes in the systematics implementation}
+\section{Systematics implementation and user-defined systematics}
 
-How defined, what good for (physics examples: T2KK, Reactor-paper,).
-Compare to standard systematics (nuisance parameters)
+\GLOBES\ supports four types of systematical errors as a standard: Signal and background
+normalization errors, as well as signal and background tilt or energy calibration errors.
+The tilt (T) is impemented as a linear distortion of the spectrum around the center,
+and the energy calibration (C) as a distortion (stretching) of the reconstructed energy scale.
+For the implementation of these systematical errors, the  ``pull method'' is used~\cite{Fogli:2002pt}, which introduces  so-called nuisance parameters $\zeta_i$. 
+For example, for the signal and background normalization errors, the rates in
+each bin $i$ are multiplied with
+\begin{equation}
+s_i(\zeta_1) := (1+\zeta_1)\cdot s_i \, , \quad b_i(\zeta_2) : =(1+\zeta_2)\cdot b_i \, ,
+\end{equation} 
+\ie, the rates are scaled by these parameters. The systematics $\chi^2$ is then minimized over
+these parameters:
+\begin{equation}
+\chi^2_\mathrm{pull}(\boldsymbol{\lambda}):=\min_{\{\zeta_i\} } \,\, \left( \,
+\chi^2(\boldsymbol{\lambda},
+\zeta_1, \ldots, \zeta_k)+ \sum_{j=1}^{k} \frac{\zeta_j^2}{\sigma_{\zeta_j}^2}
+\, \right)\,.
+\label{equ:chipull}
+\end{equation}
+Here $\chi^2(\boldsymbol{\lambda}, \zeta_1, \ldots, \zeta_k)$ is the usual Possonian $\chi^2$
+depending on the neutrino oscillation parameters $\boldsymbol{\lambda}$ and the nuisance
+parameters $\zeta_i$ (which, for instance, scale the signal or background rates).
+In addition, Gaussian penalties $\zeta_j^2/\sigma_{\zeta_j}^2$ are added, where 
+$\sigma_{\zeta_j}$ corresponds to the actual systematical error.
+Core part of \equ{chipull} is the definition of $\chi^2$. For example, for a background-free measurement and
+a signal normalization error only, it reads
+\begin{equation}
+  \chi^2 (\boldsymbol{\lambda},a) =  \sum_{i=1}^{\textrm{\# of bins}}2
+     \left( O_{i} - (1 + a) T_{i} +
+       (1 + a) T_{i} \log \frac{(1 + a) T_{i}}{O_{i}}  \right) + \frac{a^2}{\sigma_a^2} 
+\label{equ:chistandard}
+\end{equation}
+(JK/PH: CHECK THIS EQUATION IF THIS IS THE ACTUAL IMPLEMENTATION. HERE THE SYSTEMATICS IS\\
+IMPLEMENTED IN THE SIMULATED RATES. IN ADDITION, CHECK IF THEORETICAL AND EXP. RATES
+ARE CORRECTLY GIVEN. [WW])
 
-\section{User-defined priors}
+Let us assume that we have two detectors, such as for a reactor experiment with a near and
+far detector. Using standard systematics in \GLOBES , one has
+\begin{equation}
+\chi^2_\mathrm{pull}(\boldsymbol{\lambda}) = \min_{a } \,\, \left( \,
+\chi_1^2(\boldsymbol{\lambda},a)+ \frac{a^2}{\sigma_{a}^2} \, \right)
++\min_{ b } \,\, \left( \,
+\chi_2^2(\boldsymbol{\lambda},b)+ \frac{b^2}{\sigma_{b}^2}
+\, \right)\,.
+\label{equ:chipullex}
+\end{equation}
+In this case, the normalization between the two detectors will be uncorrelated.
+
+Using \GLOBESN\ and higher, one can modify the standard $\chi^2$ function in
+\GLOBES\ which corresponds to \equ{chistandard}. In addition, the systematics
+concept of ``error dimensions'' is replaced by a concept directly related to
+this $\chi^2$ function. As an example, consider a reactor experiment with near ($N$)
+and far ($F$) detectors which are described by a reactor flux uncertainty $\sigma_R$
+and two fiducial volume errors of the near $\sigma_N$ and far $\sigma_F$ detectors.
+Then \equ{chistandard} will be replaced by
+\begin{align}
+  \chi^2(\boldsymbol{\lambda},a_N,a_F,a_R) = & \sum_{i=1}^{\textrm{\# of bins}} \sum_{d = N,F} 2
+     \left( O_{d,i} - (1 + a_R + a_d) T_{d,i} + \right. \nonumber \\
+& \left. + (1 + a_R + a_d) T_{d,i} \log \frac{(1 + a_R + a_d) T_{d,i}}{O_{d,i}}  \right) 
+          + \frac{a_R^2}{\sigma_R^2} + \frac{a_N^2}{\sigma_N^2} + \frac{a_F^2}{\sigma_F^2}
+\label{equ:chi2reactor}
+\end{align}
+HOW CAN ONE MAKE THE LEFT BRACKET LARGER? [WW]
+This function can be defined with user-defined systenatics. Similarly, calibration errors, shape
+errors, uncorrelated bin-to-bin errors, \etc , can be introduced.
+Furthermore, more complicated setups using multiple discrete sources and detectors can be simulated.
+For details, we refer to the manual~\cite{Manual}.
+In the past, user-defined systematics has (as experimental feature) been used in \Ref~\cite{Huber:2006vr} to
+simulate Double Chooz and in \Ref~\cite{Barger:2006kp} to simulate T2KK. ANYTHING ELSE? [WW]
+
+In summary, \GLOBES\ currently supports the following systematics $\chi^2$ functions:
+\begin{description}
+\item[{\tt chiSpectrumTilt}]
+ Spectral information, signal and background normalization errors and tilts 
+\item[{\tt chiNoSysSpectrum}]
+ No systematical errors, but spectral information
+\item[{\tt chiTotalRatesTilt}]
+ Total rates only, signal and background normalization errors and tilts 
+\item[{\tt chiSpectrumOnly}]
+ Spectrum only, no other systematics 
+\item[{\tt chiNoSysTotalRates}]
+ Total rates only, no systematics
+\item[{\tt chiSpectrumCalib}]
+ Spectral information, signal and background normalization errors and energy calibration errors 
+\item[{\tt chiZero}]
+ Passive systematics (returns $\chi^2=0$, only provides the event rates for the access by other
+rules or experiments)
+\item[{\tt Any other name}]
+ User-defined systematics
+\end{description}
+
+\section{Adding external information with user-defined priors}
 
@@
 
-\section{Simulation of non-standard physics with GLoBES}
+\section{Simulation of non-standard physics}
 
@@
 
-In \GLOBES\ 3.0, we provide an alternative method, which does the full minimization
+In \GLOBESN , we provide an alternative method, which does the full minimization
 in one go and is therefore several times faster in most cases. In this algorithm,
@@
 interleaved minimization strategy behaves differently than the old method,
-it is by default not used in \GLOBES\ 3.0. This ensures maximal compatibility
+it is by default not used in \GLOBESN . This ensures maximal compatibility
 with old application programs, but it requires users who wish to benefit from the

### 2007-01-12 Documentation/Paper2/paper2.tex
*** empty log message ***

@@
 
+\section{Simulation of non-standard physics}
+
+The probability calculation in \GLOBES\ 
+Eg. Damping, Hamiltonian, MAVANS, Toshi etc.
+
 \section{Systematics implementation and user-defined systematics}
@@
 
-How defined, what good for (physics examples: AstroNus, Schwetz priors, atm. neutrinos, octant degeneracy,
-).
+In order to include correlations and degeneracies, \GLOBES\ provides
+(local) maginalization routines to project the fit manifold onto
+a subspace. For example, if one wants to compute the $\stheta$ precision,
+one scans the $\stheta$ direction and marginalizes for each fixed $\stheta$
+over the other oscillation parameters. This marginalization is performed
+after the systematics $\chi^2_{\mathrm{pull}}$ in \equ{chipull} (or similar) 
+has been determined:
+\begin{equation}
+ \chi^2_{\mathrm{proj}}(\boldsymbol{\lambda_1}) := \min_{\{ \boldsymbol{\lambda_2} \} } \left( \, \chi^2_{\mathrm{pull}} (\boldsymbol{\lambda_1} , \boldsymbol{\lambda_2} ) \, \right) \, .
+\label{equ:margin}
+\end{equation}
+In this case, the parameter space of oscillation parameters 
+$\boldsymbol{\lambda}$ is projected onto the subspace $\boldsymbol{\lambda_1}$.
+If one wants to include external input which only depends on the oscillation
+parameters and not on systematics, one can do this at this level. \GLOBES\ provides
+the pre-defined possibility to include external Gaussian constraints on the 
+oscillation parameters (and the matter density, which is treated as another
+oscillation parameter). For example, an external constraint on $\theta_{12}$
+is added before the marginalition in \equ{margin} by the replacement after the
+systematics minimization
+\begin{equation}
+ \chi^2_{\mathrm{pull}}(\boldsymbol{\lambda}) \rightarrow \chi^2_{\mathrm{pull}}(\boldsymbol{\lambda})  +
+ \frac{(\theta_{12} - \theta_{12}^0)^2}{\sigma_{\theta_{12}}^2}
+\label{equ:priors}
+\end{equation}
+with the central value $\theta_{12}^0$, where the input is added, and
+the $1\sigma$ Gaussian error $\sigma_{\theta_{12}}$. This corresponds to an external
+measurement of the solar mixing angle at the best-fit value $\theta_{12}^0$
+with the error $\sigma_{\theta_{12}}$.
 
-\section{Simulation of non-standard physics}
+In practice, the external error on $\theta_{12}$ may not be Gaussian or one may
+want to add other external information. Therefore,
+\GLOBESN\ and higher provides the concept of user-defined priors  $f(\boldsymbol{\lambda})$. 
+Instead of \equ{priors}, we then have
+\begin{equation}
+ \chi^2_{\mathrm{pull}}(\boldsymbol{\lambda}) \rightarrow \chi^2_{\mathrm{pull}}(\boldsymbol{\lambda})  + f(\boldsymbol{\lambda}) \, .
+\label{equ:userpriors}
+\end{equation}
+One can easily imagine that this simple concept allows for a high degree of freedom. Examples where
+this feature has been used as an experimental feature are the combination of terrestrial neutrino
+data with the information from neutrino telescopes~\cite{Winter:2006ce} and the combination of long-baseline data with atmospheric neutrino data~\cite{Huber:2005ep,Campagne:2006yx} (CORRECT?).
+Another very interesting option is the use of penalties in degeneracy localization, as suggested by Thomas
+Schwetz. For example, if the minimizer runs into the wrong octant, user-defined priors can be
+used to add a penalty and prevent it from doing so.
 
-Eg. Damping, Hamiltonian, MAVANS, Toshi etc.
-
-\section{Simulation of new experiment features and \AEDL\ changes}
+\section{\AEDL\ changes and new experiment prototypes}
 

### 2007-01-12 Documentation/Paper2/paper2.tex
*** empty log message ***

@@
 
-\section{Simulation of non-standard physics}
+\section{Oscillation probabilities and the simulation of non-standard physics}
 
-The probability calculation in \GLOBES\ 
-Eg. Damping, Hamiltonian, MAVANS, Toshi etc.
+The probability calculation in \GLOBES\ is based on the diagonalization of the
+Hamiltonian in layers of constant matter density using the standard three flavor
+scenario of neutrino oscillations. In flavor base, we have
+\begin{equation}
+\mathcal{H}(n_e) = \frac{1}{2E} \, U \,
+\left(
+\begin{array}{ccc}
+0 & 0 & 0 \\
+0  & \sdm & 0 \\
+0  & 0 & \ldm \\
+\end{array}
+\right) \,
+U^\dagger + 
+\left(
+\begin{array}{ccc}
+ \pm \sqrt{2} G_F n_e & 0 & 0 \\
+0 & 0 & 0 \\
+0 & 0 & 0 
+\end{array}
+\right)
+\label{equ:ham}
+\end{equation}
+with $U$ the mixing matrix (described by four parameters $\theta_{12}$, $\theta_{13}$, $\theta_{23}$, and $\deltacp$), and $n_e$ the constant electron density
+in that matter density layer. The electron density is related to the matter density $\rho$ by 
+$n_e \simeq 0.5 \, \rho/m_N$ with $m_N$ being the nucleon mass. The sign in the second term depends on if one
+uses neutrinos or antineutrinos. For $k$ matter density layers with densities
+$n^k_e$ and thicknesses $x_i$, the oscillation probability then evaluates to
+\begin{equation}
+P(\nu_\alpha \rightarrow \nu_\beta) = \left| \langle \nu_\beta | \mathcal{E}(n^k_e, x_k) \hdots  \mathcal{E}(n^1_e, x_1) | \nu_\alpha \rangle \right|^2
+\label{equ:probs}
+\end{equation}
+with the evolution operators (CORRECT REFERENCE?)
+\begin{equation}
+  \mathcal{E}(n^k_e, x_k) = \exp \left( -i \mathcal{H}(n^k_e) x_k \right) = \frac{1}{2E} \, \hat{U}(n^k_e) \, \left( \begin{array}{ccc}
+e^{-i \, m_1^k \, x_k} & 0 & 0 \\
+0 &  e^{-i \, m_2^k \, x_k} & 0 \\
+0 & 0 & e^{-i \, m_3^k \, x_k} 
+\end{array}
+\right) \, \hat{U}(n^k_e)^\dagger \, .
+\end{equation}
+Here the Hamiltonian $\mathcal{H}(n^k_e)$ is diagonalized by the unitary mixing matrix in matter $\hat{U}(n^k_e)$
+with the eigenvalues $m_1^k$, $m_2^k$, and $m_3^k$. The probabilities in \equ{probs} are used for the
+event rate compution, which is described in greater detail in \Ref~\cite{Huber:2004ka}. In the next section,
+we will demonstrate how systematics is implemented thereafter.
+
+\GLOBESN\ allows the modification of this Hamiltonian and the whole probability engine without
+re-compilation of the \GLOBES\ software. This is implemented
+by using pointers for the functions calling the standard probability engine, which can be changed if
+a different probability engine is registered. For example, a user may just copy the standard probability
+engine of \GLOBES , and modify it, and register it. An important prerequisite for the ability to handle
+more than the standard six oscillation parameters (plus matter density) $\boldsymbol{\lambda}=(\theta_{12},\theta_{13}, \theta_{23}, \deltacp, \sdm, \ldm, \rho)$, \ie,
+\begin{equation}
+\boldsymbol{\lambda}=(\theta_{12},\theta_{13}, \theta_{23}, \deltacp, \sdm, \ldm, \rho, \eta_1, \hdots, \eta_n)
+\end{equation}
+with $n$ non-standard parameters $\eta_j$. These parameters can be accessed in the usual way, see \Ref~\cite{Manual}.
+For example, non-standard Hamiltonian effects in the $e$-$\tau$-sector may be introduced by modifying \equ{ham} to
+\begin{equation}
+ \mathcal{\hat{H}} = \mathcal{H}(n_e) + 
+\left(
+\begin{array}{ccc}
+0 & 0 & \epsilon_{e \tau} \\
+0 & 0 & 0 \\
+\epsilon_{e \tau}^* & 0 & 0 
+\end{array}
+\right)
+\end{equation}
+In this case, there are two more real parameters, such as the absolute value and phase of $\epsilon_{e \tau}$.
+In the literature, the non-standard physics feature in \GLOBES\ as experimental feature has been used for damping effects (such as neutrino decoherence, decay, \etc) in \Ref~\cite{Blennow:2005yk}, Hamiltonian-level effects (such as non-standard matter effects) in \Ref~\cite{Blennow:2005qj}, and mass-varying neutrinos in \Ref~\cite{Schwetz:2005fy} by implementing an environment dependence of neutrino mass.
+ANY OTHERS (PH!)?
 
@@
 For the implementation of these systematical errors, the  ``pull method'' is used~\cite{Fogli:2002pt}, which introduces  so-called nuisance parameters $\zeta_i$. 
-For example, for the signal and background normalization errors, the rates in
-each bin $i$ are multiplied with
+For example, for the signal and background normalization errors, the respective event rates 
+$s_i$ and $b_i$ in
+each bin $i$ are multiplied with\footnote{For details on the event rate calculation, see \Ref~\cite{Huber:2004ka}.}
 \begin{equation}
@@
 
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=9cm]{reactor}
+\end{center}
+\mycaption{\label{fig:reactor} Luminosity scaling of the $\stheta$ sensitivity for different systematical error assumptions (90\% confidence level).
+The thick curve corresponds to \equ{chi2reactor} for $\sigma_R=\sigma_{\mathrm{corr}}=2.8\%$, $\sigma_N=\sigma_F=\sigma_{\mathrm{uncorr}}=0.6\%$ and an additional calibration error.
+Figure similar to \Ref~\cite{Huber:2006vr}.
+MISLEADING: WHY IS THE LUMINOSITY
+OF THE ``FAR DETECTOR'' ON THE X-AXIS? SHOULD THIS NOT BE THE SAME AS THE ONE FOR THE NEAR DETECTOR? 
+AT LEAST IF I INTERPRET THE CODE CORRECTLY [WW]}
+\end{figure}
+
 Using \GLOBESN\ and higher, one can modify the standard $\chi^2$ function in
@@
 HOW CAN ONE MAKE THE LEFT BRACKET LARGER? [WW]
-This function can be defined with user-defined systenatics. Similarly, calibration errors, shape
-errors, uncorrelated bin-to-bin errors, \etc , can be introduced.
-Furthermore, more complicated setups using multiple discrete sources and detectors can be simulated.
-For details, we refer to the manual~\cite{Manual}.
+This function can be defined with user-defined systematics. Similarly, calibration errors, shape
+errors, uncorrelated bin-to-bin errors, \etc , can be introduced. We show in \figu{reactor} the result of
+such a calculation, where the thick curve corresponds to \equ{chi2reactor} plus an additional 
+energy calibration error.
+Furthermore, more complicated setups using multiple discrete sources and detectors can be simulated. For details, we refer to the manual~\cite{Manual}.
 In the past, user-defined systematics has (as experimental feature) been used in \Ref~\cite{Huber:2006vr} to
@@
 \section{\AEDL\ changes and new experiment prototypes}
+
+TABLE FROM MANUAL?
 

### 2007-01-12 Documentation/Paper2/paper2.tex
*** empty log message ***

@@
 
-(or: ``GLoBES 3.0 release: From user-defined systematics to the simulation of non-standard physics'')
+(Alternative: ``GLoBES 3.0: From the simulation of non-standard physics to user-defined systematics'')
+
+(Alternative: ``GLoBES 3.0: Simulation of non-standard physics and multi-source multi-detector
+setups'')
+
 \end{center}}
@@
 \end{center}
-\caption{\label{fig:GLOBES} General concept of the \GLOBES\ package.}
+\mycaption{\label{fig:GLOBES} General concept of the \GLOBES\ package.}
 \end{figure}
@@
 
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=8cm]{reactorNSI}
+\end{center}
+\mycaption{\label{fig:reactorNSI} Fit in the $\stheta$-$\sigma_E$-plane for the simulated values
+$\stheta=\sigma_E=0$ (figure similar to \Ref~\cite{Blennow:2005yk}).
+Here $\sigma_E$ is an additional non-standard parameter describing the wave packet width of the neutrino oscillation wave packet.
+The figure is computed with
+{\tt example6.c} coming with the \GLOBESN\ distribution (and higher).}
+\end{figure}
+
 \GLOBESN\ allows the modification of this Hamiltonian and the whole probability engine without
@@
 \right)
+\label{equ:nsi}
 \end{equation}
@@
 ANY OTHERS (PH!)?
+We show in \figu{reactorNSI} a result from \Ref~\cite{Blennow:2005yk} provided as {\tt example6.c} in the \GLOBES\
+distribution. Compared to \equ{nsi}, the underlying physics takes place at the probability level, \ie, 
+\equ{probs} (for details, see \Ref~\cite{Blennow:2005yk}). In {\tt example6.c}, an analytical implementation
+of this physics is used for simplicity.
 
@@
 \end{center}
-\mycaption{\label{fig:reactor} Luminosity scaling of the $\stheta$ sensitivity for different systematical error assumptions (90\% confidence level).
+\mycaption{\label{fig:reactor} Luminosity scaling of the $\stheta$ sensitivity for different systematical error assumptions (90\% confidence level, figure similar to \Ref~\cite{Huber:2006vr}).
 The thick curve corresponds to \equ{chi2reactor} for $\sigma_R=\sigma_{\mathrm{corr}}=2.8\%$, $\sigma_N=\sigma_F=\sigma_{\mathrm{uncorr}}=0.6\%$ and an additional calibration error.
-Figure similar to \Ref~\cite{Huber:2006vr}.
+The figure is computed with {\tt example5.c} provided with the \GLOBES\ software.
  MISLEADING: WHY IS THE LUMINOSITY

### 2007-01-13 Documentation/Manual.tex
*** empty log message ***
JK - Some typos in Manual.tex corrected

@@
 \newcommand{\Beta}{{\sc $\beta$-Beam}}
+\newcommand{\NOVA}{{\sc NO$\nu$A}}
+\newcommand{\TtoK}{{\sc T2K}}
+\newcommand{\TtoHK}{{\sc T2HK}}
+\newcommand{\DC}{{Double\sc Chooz}}
 
@@
 as well as up-to-date versions of this manual can be found at this URL:
- \verb+http://www.ph.tum.de/~globes+
+ \verb+http://www.mpi-hd.mpg.de/~globes+
 
@@
 atmosphere, can not be described. In addition, sources with a physically 
-significant time dependencies  can not be studied, such as  supernov\ae. It 
+significant time dependence, such as supernov\ae, can not be studied. It 
 is, however, possible to simulate beams with bunch structure, since the 
@@
 This approach allows the inclusion of multi-parameter-correlations,
-where external input (\eg, from solar parameters) can be imposed, too.
+where external input (\eg, from solar experiments) can be imposed, too.
 Applications of the projection mechanism include the projections onto the $\stheta$-axis and the $\stheta$-$\deltacp$-plane.
@@
 Simulation of long baseline neutrino oscillation experiments with \GLOBES\\
-Comput. Phys. Commun. 167 (2005) 195, arXiv:\glbxxx.   
+Comput. Phys. Commun. 167 (2005) 195, arXiv:\glbxxx,
 \end{quote}
@@
 TO BE DEFINED, NEEDS TO BE UPDATED [WW]\\
-arXiv:\glbxxxtwo.   
+arXiv:\glbxxxtwo,
 \end{quote}
@@
 
-It is recommended that you use the same style for your own input files, since, in case they are distributed, everybody will know how to correctly reference your work.
+It is recommended that you use the same style for your own input files, since then, in case they are distributed, everybody will know how to correctly reference your work.
 
@@
 of several modules.
-%
+
 \begin{figure}[bht]
@@
 \AEDL (``Abstract Experiment Definition Language'') is a language
-to define experiments in form of ordinary text files. One or more of 
+to define experiments in the form of ordinary text files. One or more of 
 the resulting \AEDL\ files can then be processed together with supporting 
@@
 The application 
-software is, except from some example files, not part of \GLOBES , since
+software is, except for some example files, not part of \GLOBES , since
 the evaluation of the experiment performance is often a matter of taste
@@
 \GLOBES , where the user interface is only a tool to obtain specific
-parameter sensitivities. In this case, \GLOBES\ could, serve as a
+parameter sensitivities. In this case, \GLOBES\ could serve as a
 unified tool for the comparison and optimization of different experiment
@@
 as possible. In addition, changes in experimental parameters, such as
-efficiencies or the energy resolutions, can quickly be tested.
+efficiencies or energy resolutions, can quickly be tested.
 %
@@
 want to modify or create \AEDL\ files. A useful tool in this context
-is the software program \verb+globes+, which returns event rates and other
+is the executable program \verb+globes+, which returns event rates and other
 information for individual \AEDL\ files without further programming. 
@@
 % from the ones in this manual.
+%
+% Comment by JK: A pedantic C89 compile will reject the example files (it would
+% also reject GLoBES itself). With its default settings, gcc is relatively
+% tolerant, so everything works with gcc. I assume that also the Intel
+% compiler would be OK, but maybe Borland, Microsoft, or other compilers would
+% produce problems.
 
@@
 
-\backmatter
+%\backmatter
 \chapter*{Acknowledgments}
@@
 and Thomas Schwetz, who has been pushing the software to the edge in the 
-past few years. Furthermore, we would like to thank Toshihiko Ota and WHO ELSE? [WW]
-for using and testing unpublished new features of the software.
+past few years. Furthermore, we would like to thank Tommy Ohlsson,
+Toshihiko Ota, and Julian Skrotzki for using and testing unpublished new features of the software.
 PH is especially thankful for the invaluable advice of Thomas Fischbacher on

### 2007-01-14 Documentation/part1.tex
JK - Minor corrections in chapters 1 - 3

@@
 Instead of including the full correlation, we can take the
-correlation with every parameter except from $\deltacp$, \ie,
+correlation with every parameter except for $\deltacp$, \ie,
 we keep (in addition to $\theta_{13}$) $\deltacp$ fixed.
-This corresponds to projection onto the $\stheta$-$\deltacp$-plane:
+This corresponds to a projection onto the $\stheta$-$\deltacp$-plane:
 \gq{
@@
 }
-Let us now locate the exact position\footnote{For a exact definition of 
+Let us now locate the exact position\footnote{For an exact definition of 
 inverted hierarchy, see page~\pageref{mass_ordering}.} of the sgn-degeneracy:
@@
 functions. In many cases, it is sufficient to use the first
-argument from the command line as the program name (such as in example on page~\pageref{ex:c}).
+argument from the command line as the program name (such as in the example on page~\pageref{ex:c}).
 \end{function}
@@
 that it makes sense to simulate different operation modes within one
-experiment definition, and physically different baselines in different
+experiment definition, and physically different baselines, in different
 definitions. For details of the rate computation and
@@
  the simplest case of simulating one experiment may be most often used, 
- using more than one experiments are useful in many cases. For example, combinations of experiments can be tested for
+ using several experiments is useful in many cases. For example, combinations of experiments can be tested for
 complementarity and competitiveness by equal means within one program.
@@
 for these vectors has to be reserved (allocated) before they can be used, 
-and it has
-to be returned (freed) afterwards. \GLOBES\ functions usually use the pointers
-of the type {\tt glb\_params} for the input or output to the functions. As
-an input parameter, the pointer has to be created and point towards a valid
+and it has to be returned (freed) afterwards.
+\GLOBES\ functions usually use pointers
+of the type {\tt glb\_params} for the input or output to the functions.
+As an input parameter, the pointer has to be initialized with the address of a valid
 parameter structure, where the oscillation parameters are read from. As
-an output parameter, the pointer has to be created, too, and point towards
+an output parameter, it has to be initialized with the address of
 a structure which the return values will be written to.
@@
 mappings which otherwise need to be done by the user. For example, not
-only the oscillation parameters are stored in the pointer structure, but also information on the matter
+only the oscillation parameters are stored in the \GLB{\_params} structure, but also information on the matter
 densities of all of the initialized experiments. Since \GLOBES\ treats
@@
 
-Another piece of information will be returned from the minimizers 
+Another piece of information that will be returned from the minimizers 
 (\cf, \Chapt~\ref{chapt:correlations}) and transferred
@@
 ranges. The true parameter values for the vacuum neutrino oscillation 
-parameters have to be set by the function \GLB{SetOscillationParameters} 
+parameters have to be set by the function \GLB{SetOscillationParameters}, 
 and the reference rate vector, \ie\ the data, has to be computed by a call 
@@
 \index{norm}{Systematics} \index{norm}{Pull method}
-The treatment of systematics in \GLOBES\ is performed by the so-call
+The treatment of systematics in \GLOBES\ is performed by the so-called
 {\em pull method} with the help of auxiliary systematics parameters.
-They are taken completely uncorrelated among different rules, and treated with simple Gau\ss ian statistics. In general, a rule is a set 
-of signal and background event rates coming from different oscillation
-channels, where the event rates of all rule contributions are added.
-For more details of the rule concept, see \Part~\ref{part:2} of this manual,
+They are taken to be completely uncorrelated among different rules,
+and treated with simple Gau\ss ian statistics.
+In general, a rule is a prescription for summing up experimentally indistinguishable
+signal and background events from different oscillation channels.
+For more details on the rule concept, see \Part~\ref{part:2} of this manual,
 and for the treatment of systematics, see \Sec~\ref{sec:rules}.
 
- One example for a systematics parameter the signal normalization error, \ie, an error to the overall normalization of the signal.
+ One example for a systematics parameter is the signal normalization error, \ie, an error on the overall normalization of the signal.
 For illustration, we assume that the signal event rate in the $i$th bin $s_i^0$ of one oscillation channel is altered by the overall normalization auxiliary parameter\index{norm}{Auxiliary parameter} of this channel, \ie ,
@@
 the AEDL definition and some source code defining the behavior. In the simplest case,
-the AEDL file defines user-defined systematics for a specific rule and gives it a 
-name/identifier,  as will as it defines the systematical errors. The source code has then to match this AEDL
+the AEDL file defines user-defined systematics for a specific rule, gives it a 
+name/identifier, and defines the systematical errors. The source code has then to match this AEDL
 definition -- therefore, make sure to use only unique identifiers. 
@@
 {\tt glbDefineChiFunction}.
-Note the the central values for {\tt params} are $0$, which means that $0$ corresponds to the un-modified rates. The array indices are running from $0$ to {\tt dim}-1.
+Note the the central values for {\tt params} are $0$, which means that $0$ corresponds to the un-modified rates. The array indices run from $0$ to {\tt dim}-1.
 \end{function}
@@
 Therefore the function should be as efficient as possible.
-In many cases, it is also recommendable to use different functions for different rules, since the fewer auxiliary parameters in the minimization, the faster the routine will be.
+%JK - THIS IS INCORRECT: ONE MINIMIZATION OVER, SAY, 8 PARAMETERS IS FASTER THAN
+%2 MINIMIZATIONS OVER 4 PARAMETERS EACH-
+%In many cases, it is also recommendable to use different functions for different
+%rules, since the fewer auxiliary parameters in the minimization, the faster the routine will be.
 In addition, note that complicated systematics with many parameters may introduce complicated (unknown) topologies for the minimization, which means that the minimizers may
@@
     \hspace*{0.5cm} /* Here the systematics priors (penalties) are added: */ \\
-    \hspace*{0.5cm} for(i=0;i<dim;i++) chi2 += (x[i]*x[i]/errors[i]/errors[i]);
+    \hspace*{0.5cm} for(i=0;i<dim;i++) chi2 += (x[i]*x[i])/(errors[i]*errors[i]);
     \hspace*{0.5cm} return chi2; \\
@@
 \end{quote}
-During running time, it is also possible to change the user-defined systematics of an experiment or
+During running time, it is also possible to change the systematics of an experiment or
 rule (as compared to the systematics assigned in the AEDL file or earlier) with the following function:
@@
 Both of the parameters {\tt exp} and {\tt rule} can take the value \GLBC{GLB\_ALL} to specify that the
-user-defined systematics can be used for each experiment or each rule. The parameter 
-{\tt on\_off} determines if the systematics function should be used for systematics on
-{\tt GLB\_ON} or systematics off {\tt GLB\_OFF}.
-The array {\tt errors} sets the systematical errors in the order they
-are defined in the AEDL file (running from $0$ to the number of systematics parameters-1). 
-Returns zero, if successful.
-\end{function}
-Note that the new user-defined systematics has to be registered with {\tt glbDefineChiFunction} first. 
-One can also return the systematics function name by 
+given systematics function should be used for each experiment or each rule. The parameter 
+{\tt on\_off} determines if the systematics function should be used when systematical errors
+are switched on ({\tt GLB\_ON}), or when they are switched off ({\tt GLB\_OFF}).
+The array {\tt errors} sets the systematical errors in the order in which they
+are expected by the systematics function (indices run from $0$ to
+the number of systematics parameters-1). The function returns zero, if successful.
+\end{function}
+Note that new user-defined systematics functions have to be registered with {\tt glbDefineChiFunction} first. 
+One can also request the systematics function name by 
 \begin{function}
@@
 {\tt glbShiftEnergyScale(double b, double *rates\_in, double *rates\_out,int bins, double emin, double emax)}
-shifts the energy scale in {\tt rates\_in} by the amount {\tt b} and stores the result in 
+shifts the energy scale in {\tt rates\_in} by the relative amount {\tt b} and stores the result in 
 {\tt rates\_out}. The parameters {\tt emin} and {\tt emax} denote the minimal and maximal energy
@@
 $a_R$, $a_N$ and $a_F$ parametrize the small uncertainties in the reactor flux and the fiducial
-mass of the two detectors. In this example their central values are assumed to be zero,
+mass of the two detectors. In this example, their central values are assumed to be zero,
 while their standard deviations are $\sigma_R$, $\sigma_N$ and $\sigma_F$.
@@
 passively provided for the common $\chi^2$ function.
-You can find the corresponding files {\tt dchooz-near.glb} and {\tt dchooz-far.glb}, you will need for {\tt example5.c}, in the example directory. 
+You can find the corresponding files {\tt dchooz-near.glb} and {\tt dchooz-far.glb} you will need for {\tt example5.c}, in the example directory. 
 %
@@
 \end{center}
-\mycaption{\label{fig:reactor} The result from {\tt example5.c} (figure similar in \Ref~\cite{Huber:2006vr}). The fat curve corresponds to the systematics on page~\pageref{ex:syst}.}
+\mycaption{\label{fig:reactor} The result from {\tt example5.c} (figure similar to the one
+in \Ref~\cite{Huber:2006vr}). The thick curve corresponds to the systematics on page~\pageref{ex:syst}.}
 \end{figure}
@@
 \end{function}
-In addition, in some cases the systematics minimization may not converge very fast or not at all, because it end up in a local minimum.
-In this case, one can change the starting values of the minimizer, \ie, the
+In cases where the systematics minimization does not converge fast enough or ends
+up in a local minimum, one can change the starting values of the minimizer, \ie, the
 starting point from which the local minimizer rolls into the local minimum.

### 2007-01-14 Documentation/part1.tex
JK - Minor corrections in chapter 4

@@
 close enough to the investigated minimum.
- Fortunately, this can be done quite straightforward in most cases, since 
+ Fortunately, this can be done quite straightforwardly in most cases, since 
 the structure of the neutrino oscillation formulas does not cause very 
@@
 the $\chi^2$:
-The results obtained this way are always only a \emph{upper} bound on the
+The results obtained this way are always only an \emph{upper} bound on the
 value of the projected $\chi^2$ function, \ie\ an undiscovered minimum 
@@
 below the chosen confidence level. With this approach and proper usage,
- it should not possible to produce a too pessimistic solution. 
+ it should not be possible to produce a too pessimistic solution. 
 However, if one is not careful enough to locate
@@
 \begin{equation}
-\mathrm{Too \, pessimistic \, result} \, < \, \underbrace{\mathrm{Real \, result}}_{\mathrm{Located \, by \, careful \, usage}} \, \le \, \mathrm{GLoBES \, result} \,  < \,  \mathrm{Too \, optimisitic \, result} \nonumber
+\mathrm{Too \, pessimistic \, result} \, < \, \underbrace{\mathrm{Real \, result}}_{\mathrm{Located \, by \, careful \, usage}} \, \le \, \mathrm{\GLOBES \, result} \,  < \,  \mathrm{Too \, optimisitic \, result} \nonumber
 \end{equation}
@@
 parameters will in most cases be supplied by the solar neutrino experiments. 
-If the external precision of a parameter is at the time of the 
-measurement better than the one of the experiment itself, one usually will use
-this external, better knowledge and impose a corresponding constraint
- on this parameter. This external knowledge may reduce the extension of the 
+If, at the time of the measurement, the external precision of a parameter is
+better than the one of the considered experiment itself, one usually will use
+this better, external knowledge and impose a corresponding constraint
+ on this parameter. The external knowledge may reduce the extension of the 
 $n$-dimensional fit manifold in the respective direction. In the most 
@@
 in the analysis is equivalent to the assumption that all parameters 
-are determined externally with infinitively high precisions. As this 
-is a quite strong assumption, one should \emph{always} check the 
+are determined externally with infinitively high precision. As this 
+is quite a strong assumption, one should \emph{always} check the 
 consequences of relaxing it and using realistic errors. Only if such
 a test has demonstrated that the impact of the uncertainty on a 
-given fit parameter is negligible, it can be assumed as fixed safely. 
+given fit parameter is negligible, it can safely be assumed as fixed. 
  The inclusion of external input in \GLOBES\ is done by the use of 
 Gau\ss ian {\em priors}: We assume that an external measurement has 
-determined the measured parameter to be at the central value 
-(which we call {\em central value}) with a $1 \sigma$ Gau\ss ian error 
-(which we call {\em input error}). The explicit definition of these 
-priors will be shown in the next section.
+determined the measured parameter to be at the \emph{central value}
+with a $1 \sigma$ Gau\ss ian error (which we call {\em input error}).
+The explicit definition of these priors will be shown in the next section.
 
@@
 the central value is fixed at the best-fit value, and the input error 
-to the $1 \sigma$ half width of the external measurement. For the matter 
-density, $\hat{\rho}^0$  is usually set to $1.0$ corresponding to
+is chosen as the $1 \sigma$ half width of the external measurement. For the matter 
+density, $\hat{\rho}^0$  is usually set to $1.0$, corresponding to
 the actual matter density profile such as given by the experiment
@@
 In this example, the central values are set to 
-the true (simulated) values. Remember that initially the 
-matter density scaling factors are 
-all $1.0$, which means that they do not need to be adjusted for the
-central values.
+the true (simulated) values.
+
+%JK - NO LONGER TRUE IN GLOBES 3.0
+%Remember that initially the 
+%matter density scaling factors are 
+%all $1.0$, which means that they do not need to be adjusted for the
+%central values.
 
@@
 The projection onto the $\stheta$- (or $\deltacp$-) axis is performed by fixing $\stheta$ (or $\deltacp$) and minimizing the $\chi^2$-function over all free fit parameters and the matter densities.
-We illustrate this method at the example of the projection of the two-dimensional manifold in the $\stheta$-$\deltacp$-plane onto the $\stheta$-axis in \figu{projex}.
+We illustrate this method by the example of the projection of the two-dimensional manifold in the $\stheta$-$\deltacp$-plane onto the $\stheta$-axis in \figu{projex}.
 In this figure, the left-hand plot shows the correlation in the $\stheta$-$\deltacp$-plane computed with {\tt glbChiSys}.
@@
 In this simple example, the minimization is done along the vertical gray lines in the left hand plot.
-The obtained minima are located on the thick gray curve, which means the the right-hand plot represents the $\chi^2$-value along this curve.
+The obtained minima are located on the thick gray curve, which means that the right-hand plot represents the $\chi^2$-value along this curve.
 In fact, one can easily see that one obtains the correct projected $3 \sigma$ errors in this example (\cf, arrows).
@@
 The minimizers of individual experiments ``know'' which
-experiment they are currently treating, which means that they only returned the
+experiment they are currently treating, which means that they only return the
 matter density scaling factor of the appropriate experiment.
@@
 The other values in {\tt out} still correspond to the fit values.
-\item[The position of external priors,] \ie, the best-fit values of external input, is set by \GLB{SetCenralValues}.
-\item[The magnitude of external errors,] \ie, the errors themselves of external input, is set by \GLB{SetInputErrors}.
+\item[The position of external priors,] \ie, the best-fit values of external input, is set by \GLB{SetCentralValues}.
+\item[The magnitude of external errors,] \ie, the errors of external input, is set by \GLB{SetInputErrors}.
 \end{description}
@@
 \end{function}
-This function works analogously to the ones in the last section. They can, for example, be used to obtain a figure similar to \figu{projex}, left.
+This function works analogously to the ones in the last section. It can, for example, be used to obtain a figure similar to \figu{projex}, left.
 The example on page~\pageref{ex:corrproj} illustrates then the difference
@@
 provided. The purpose of {\tt glb\_projection} is to tell \GLOBES\ 
-what parameters are fixed, and what are minimized over. Thus, in
+which parameters are fixed, and which are minimized over. Thus, in
 comparison to {\tt glb\_params}, it does not take values for the
@@
 User-defined priors are an advanced concept of \GLOBES\ 3.0 and higher. Therefore,
-this section can be jumped over in a first reading of the manual.
+this section can be skipped in a first reading of the manual.
 They allow for arbitrary priors being added to the $\chi^2$ after the systematics
@@
 
-The prior function has to be any function using one parameter {\tt glb\_params} containing the fit values.
+The prior function should expect the parameter structre {\tt glb\_params} containing the
+current fit values and the {\tt void}-pointer {\tt user\_data} as arguments.
 We show an example for a user-defined prior on page~\pageref{ex:userprior}. This prior
-will simply be registered with
+is simply registered with
 \begin{quote}

### 2007-01-14 Documentation/part1.tex
JK - Minor corrections in chapters 5 and 6

@@
 the local minimizer at the suspected position 
-and let it run into the minimum.  The following code excerpt  corresponds to finding the degenerate solution for the example on
+and lets it run into the minimum.  The following code excerpt  corresponds to finding the degenerate solution for the example on
   page~\pageref{ex:corrth13dcp}, and it is from {\tt example3.c}: 
@@
 
-Using {\tt ent\_pos} to obtain a section of the degeneracy in the
+Using {\tt deg\_pos} to obtain a section of the degeneracy in the
 $\stheta$-$\deltacp$-plane (\cf, {\tt example3.c}), one can plot it as a contour plot in addition to the original solution (2 d.o.f., gray curves):
@@
 \paragraph{Note:} Inverting the mass hierarchy is not
-precisely the same than to change from $\Delta m^2_{31}
+precisely the same as changing from $\Delta m^2_{31}
 \rightarrow -\Delta m^2_{31}$.
@@
 \item[Forcing the minimizer into the targeted solution.]
-In addition to switching off the systematics, it can be useful to reduce the input errors during some steps of the localization process in order to make the minimizer not to run away too much from the targeted solution.
+In addition to switching off the systematics, it can be useful to reduce the input errors during some steps of the localization process in order to prevent the minimizer from running away too far from the targeted solution.
 The example on page~\pageref{ex:sgndeg} illustrates this with the
@@
 \item[Manual scan of a subspace.] In some cases, the minimizer easily ends up in the wrong/unwanted solution, which is in most cases the already known best-fit solution.
-For example, for the octant degeneracy location, it is difficult to prevent the minimizer from running into the best-fit octant.
+For example, when locating the octant degeneracy, it is difficult to prevent the minimizer from running into the best-fit octant.
 In this case, one can scan one parameter (such as $\theta_{23}$) on a grid by choosing only valid values, and use {\tt glbChiNP} to marginalize over the other ones (while keeping \eg\ $\theta_{23}$ fixed).
@@
  Altogether, there is a maximum of eight intersection points in the $\sin^2 2 \theta_{13}$-$\delta_{\mathrm{CP}}$-plane, one of which is the best-fit point.
-These points can be using as starting points of the minimizer to locate the eight-fold degeneracy.
+These points can be used as starting points for the minimizer to locate the eight-fold degeneracy.
 Note that similar methods using the $\chi^2$ instead of the total rates have also been successfully used in the past.
 \end{description}
-Finally, note that any degenerate solution below the confidence level threshold which cannot be located makes the result appear better than it actually is.
+Finally, note that any degenerate solution below the confidence level threshold, which cannot be located, makes the result appear better than it actually is.
 Thus, one should be careful with the determination of the degenerate solutions in order to find all of them.
@@
 In this chapter, we discuss possibilities to obtain low-level information
-in \GLOBES , \ie, about oscillation probabilities, rates, and other
+in \GLOBES , \ie, oscillation probabilities, rates, and other
 information lower than on the $\chi^2$-level.
@@
 {\tt panti} is $+1$ for neutrinos and $-1$ for antineutrinos.
-The matter density profile including baseline is the one from the last
-evaluated experiment. 
 \end{function}
@@
 \end{function}
-If one in addition wants to use the low-pass filter feature in \GLOBES\ (see \Sec~\ref{sec:energy}), can can use
+If one in addition wants to use the low-pass filter feature in \GLOBES\ (see \Sec~\ref{sec:energy}), one can can use
 \begin{function}
@@
 {\tt int glbGetEminEmax(int experiment, double *emin, double *emax)} returns the
-energy range [{\tt emin},{\tt emax}] covered by the binning as defined in the AEDL file for the experiment
+energy range [{\tt emin}, {\tt emax}] covered by the binning as defined in the AEDL file for the experiment
 {\tt experiment}.
@@
 {\tt int glbGetEnergyWindow(int exp, int rule, double *low, double *high)} returns the
-energy window [{\tt low},{\tt high}] defined in the AEDL file for the experiment
+energy window [{\tt low}, {\tt high}] defined in the AEDL file for the experiment
 {\tt exp} and rule {\tt rule}.
@@
 {\tt int glbGetEnergyWindowBins(int exp, int rule, int *lowbin, int *highbin)} returns the
-energy window in terms of bin numbers [{\tt lowbin},{\tt highbin}] corresponding to the energy window defined in the AEDL file for the experiment {\tt exp} and rule {\tt rule}.
+energy window in terms of bin numbers [{\tt lowbin}, {\tt highbin}] corresponding to the energy window defined in the AEDL file for the experiment {\tt exp} and rule {\tt rule}.
 \end{function}
@@
 \index{norm}{AEDL@\AEDL!names}
-Since in \AEDL\ rules, cross section, fluxes etc. carry a `name' by
-which they can be refered to and in C they cary only an integer as
-index it is sometimes difficult to figure out the correct correspondence.
+Since in \AEDL\, rules, cross section, fluxes etc. carry a `name' by
+which they can be refered to, while in C they carry only an integer
+index, it is sometimes difficult to figure out the correct correspondence.
 Therefore the information about this correspondence obtained during
-parsing is stored and can be accessed within C by these two functions.
+parsing is stored and can be accessed within C by the following two functions.
 \begin{function}
@@
 C index. The variable {\tt context} describes wether this name belongs
-to a rule, channel, flux, energy or cross type environment. {\tt exp}
+to a rule, channel, flux, energy, or cross type environment. {\tt exp}
 is the number of the experiment and can not be {\tt GLB\_ALL}. It returns
@@
 \Secs~\ref{sec:channel} and~\Sec~\ref{sec:rules} of the \AEDL\ manual.
-Note that some of the event rate access has changed in \GLOBES\ 3.0,
-because user-defined systematics requires very fast access which was
+Note that parts of the event rate access have changed in \GLOBES\ 3.0,
+because user-defined systematics require very fast access, which was
 not possible with the old method.
@@
 int effi, int bgi, int coeffi, int signal)} prints the binned rule rates as
-list with energy and event rate to the file {\tt stream} (either an
+a list with energy and event rate to the file {\tt stream} (either an
 open file, or {\tt stdout}). A specific experiment {\tt exp} and a 
@@
 prints the binned channel rates as
-list with energy and event rate to the file {\tt stream} (either an
+a list with energy and event rate to the file {\tt stream} (either an
 open file, or {\tt stdout}). A specific experiment {\tt exp} and a 
@@
 {\tt double *glbGetChannelRatePtr(int exp, int ch, int pre\_post)} returns a 
-pointer on the simulated rate vector of experiment {\tt exp} and channel {\tt ch}. Either
+pointer to the simulated rate vector of experiment {\tt exp} and channel {\tt ch}. Either
 pre-smearing ({\tt pre\_post} is {\tt GLB\_PRE}) or post-smearing ({\tt pre\_post} is {\tt GLB\_POST})
@@
 {\tt double *glbGetRuleRatePtr(int exp, int rule)} returns a 
-pointer on the simulated rate vector of experiment {\tt exp} and rule {\tt rule}.
+pointer to the simulated rate vector of experiment {\tt exp} and rule {\tt rule}.
 \end{function}
@@
 {\tt double *glbGetSignalRatePtr(int exp, int rule)} returns a 
-pointer on the simulated signal rate vector of experiment {\tt exp} and rule {\tt rule}.
+pointer to the simulated signal rate vector of experiment {\tt exp} and rule {\tt rule}.
 \end{function}
@@
 {\tt double *glbGetBGRatePtr(int exp, int rule)} returns a 
-pointer on the simulated background rate vector of experiment {\tt exp} and rule {\tt rule}.
+pointer to the simulated background rate vector of experiment {\tt exp} and rule {\tt rule}.
 \end{function}
@@
 {\tt double *glbGetChannelFitRatePtr(int exp, int ch, int pre\_post)} returns a 
-pointer on the fit rate vector of experiment {\tt exp} and channel {\tt ch}. Either
+pointer to the fit rate vector of experiment {\tt exp} and channel {\tt ch}. Either
 pre-smearing ({\tt pre\_post} is {\tt GLB\_PRE}) or post-smearing ({\tt pre\_post} is {\tt GLB\_POST})
@@
 {\tt double *glbGetSignalFitRatePtr(int exp, int rule)} returns a 
-pointer on the fit signal rate vector of experiment {\tt exp} and rule {\tt rule}.
+pointer to the fit signal rate vector of experiment {\tt exp} and rule {\tt rule}.
 \end{function}
@@
 {\tt double *glbGetBGFitRatePtr(int exp, int rule)} returns a 
-pointer on the fit background rate vector of experiment {\tt exp} and rule {\tt rule}.
+pointer to the fit background rate vector of experiment {\tt exp} and rule {\tt rule}.
 \end{function}

### 2007-01-14 Documentation/part1.tex
JK - Minor corrections in chapter 7

@@
 \index{norm}{Systematics!on/off}
-Here it is demonstrated how one can successively include systematics, correlations, and degeneracies at the example of the $\stheta$-sensitivity limit.
-An important part of this example is how two switch the systematics off,
-\ie, how to obtain the sensitivity limit from statistics only. Since
-this example is very advanced, we only show the respective function of the code:
+Here, we demonstrate how systematics, correlations, and degeneracies
+can be succesively included in the calculation of the $\stheta$-sensitivity limit.
+The following code fragment shows how systematics can be switched off
+in order to compute the sensitivity limit from statistics only:
 \begin{quote}
@@
 
-The complete code can be found as {\tt example4.c} with the software,
-which consists of many of the applications from the earlier examples.
-In addition, it uses a little trick: It avoids falling into the wrong
- solution with {\tt glbChiTheta13}
-by using the fit value of $\deltacp$ from the step before as prediction
-of the position of the current calculation.
+The complete code is very advanced and can be found in {\tt example4.c}.
+It includes many concepts from earlier examples, and, in addition,
+it uses a little trick: It avoids falling into the wrong solution with
+{\tt glbChiTheta13} by using the fit value of $\deltacp$ from the step
+before as prediction for the position of the minimum in the current calculation.
 
@@
 curves with the line $\chi^2 = 9$ give the $\stheta$ sensitivity
-limits at the $3 \sigma$ confidence level, where we do not include the
- $\mathrm{sgn}(\ldm)$- and $(\deltacp,\theta_{13})$-degeneracies in the sensitivity limit with correlations only (green bar):
+limits at the $3 \sigma$ confidence level. Note that, in the following
+plot, the $\mathrm{sgn}(\ldm)$- and $(\deltacp,\theta_{13})$-degeneracies
+are not included in the sensitivity limit with correlations only (green bar),
+but only in the limit with degeneracies (yellow bar):
 \begin{center}
@@
 Using external parameters in \AEDL\ files is a very powerful feature
-to change experiment parameters at running time which require that
+to change experiment parameters at running time, which requires, however, that
 the experiment be re-initialized. For example, one can change the
 energy resolution function or the number of energy bins. However,
-in some cases there might be complications, such that the number
+in some cases, there might be complications, such that the number
 of pre- or post-smearing efficiencies does not correspond to the number
@@
 \end{quote}
-Note that one does not have do re-initialize the oscillation
+Note that one does not have to re-initialize the oscillation
 parameter vectors every time within the loop as long as the

### 2007-01-14 Documentation/part1.tex
JK - Minor changes in chapters 8 - 9

@@
 \GLBNS{\_set\_oscillation\_parameters\_function}
-{\tt int (*glb\_set\_oscillation\_parameters\_function)(glb\_params p, void* user\_data)} reads the fundamental oscillation parameters
-obtained from {\tt p} and stores them into global variables accessible to {\tt glb\_probability\_matrix\_function}.
- In addition, it can be used to pre-compute quantities which are dependent on the oscillation paramaters only for efficiency. 
-In order to circumvent global variables, an arbitrary pointer can be used by {\tt user\_data} which is set by {\tt glbRegisterProbabilityEngine}.
+{\tt int (*glb\_set\_oscillation\_parameters\_function)(glb\_params p, void* user\_data)} is
+used to pass the fundamental oscillation parameters {\tt p} to the probability engine.
+The function should store them into variables accessible to {\tt glb\_probability\_matrix\_function}.
+In addition, it can be used to pre-compute quantities which are dependent on the oscillation paramaters for efficiency. 
+In order to circumvent global variables, arbitrary additional parameters can be
+passed to the function in {\tt user\_data}, which is set by {\tt glbRegisterProbabilityEngine}.
 \end{function}
@@
 \GLBNS{\_get\_oscillation\_parameters\_function}
-{\tt int (*glb\_get\_oscillation\_parameters\_function)(glb\_params p, void* user\_data)} returns the fundamental oscillation parameters
-in {\tt p} which are obtained from the global variables accessible to {\tt glb\_probability\_matrix\_function}.
-In order to circumvent global variables, an arbitrary pointer can be used by {\tt user\_data} which is set by {\tt glbRegisterProbabilityEngine}.
+{\tt int (*glb\_get\_oscillation\_parameters\_function)(glb\_params p, void* user\_data)}
+reads the fundamental oscillation parameters from the internal variables of the probability
+engine, and writes them into {\tt p}.
+In order to circumvent global variables, arbitrary additional parameters can be
+passed to the function in {\tt user\_data}, which is set by {\tt glbRegisterProbabilityEngine}.
 \end{function}
@@
                   int psteps, const double *lengths, const double *densities,
-                  double filter\_sigma, void* user\_data)} calculates the neutrino oscillation probability matrix to be returned in {\tt P}.
-The {\tt cp\_sign} is +1 for neutrinos and -1 for antineutrinos.
-In addition, the matter density profile to be used is characterized by the number of steps {\tt psteps}, the lengths of the matter density layers in the list {\tt lengths}, and the corresponding densities in the list {\tt densities}.
-The parameter {\tt filter\_sigma} defines the width of the low-pass filter to be used, or no filter usage if negative.
-In order to circumvent global variables, an arbitrary pointer can be used by {\tt user\_data} which is set by {\tt glbRegisterProbabilityEngine}.
+                  double filter\_sigma, void* user\_data)}
+calculates the neutrino oscillation probability matrix and returns it in {\tt P}.
+The {\tt cp\_sign} is +1 for neutrinos and -1 for antineutrinos. In addition, the matter
+density profile to be used is characterized by the number of steps {\tt psteps}, the
+lengths of the matter density layers in the list {\tt lengths}, and the corresponding
+densities in the list {\tt densities}. The parameter {\tt filter\_sigma} defines
+the width of the low-pass filter to be used, or no filter usage if negative.
+In order to circumvent global variables, arbitrary additional parameters can be
+passed to the function in {\tt user\_data}, which is set by {\tt glbRegisterProbabilityEngine}.
 \end{function}
@@
 
-In order to use non-standard physics in \GLOBES , one needs re-define the three above functions and register them after {\tt glbInit} with
+In order to use non-standard physics in \GLOBES , one needs to re-define the three above functions and register them after {\tt glbInit} with
 \begin{function}
@@
 engine for the simulation of non-standard physics with {\tt n\_parameters} oscillation parameters ({\tt n\_parameters} $\ge6$).
-The three functions {\tt prob\_func}, {\tt set\_params\_func}, and {get\_params\_func}
-will be used with the function types defined above. In order to circumvent global variables, an arbitrary pointer can be used by {\tt user\_data} being set by this function.
+The three functions {\tt prob\_func}, {\tt set\_params\_func}, and {\tt get\_params\_func}
+will be used with the function types defined above.
+In order to circumvent global variables, an arbitrary pointer {\tt user\_data} can be defined, that will be passed to the probability engine in each function call.
 \end{function}
@@
 in vacuum for simplicity. The non-standard effect is the loss of coherence because of wave packet decoherence
-or any other such effect, \ie, one more oscillation parameter. We need to define two functions to access a set of 
+or any other such effect, \ie, we have one additional oscillation parameter. We need to define two functions to access a set of 
 global oscillation parameters
@@
 {\tt {\footnotesize
-  const GLB\_SIGMA\_E = 6; \\ \\
+  const int GLB\_SIGMA\_E = 6; \\ \\
   /* Fix theta13 and sigma\_E, and marginalize over all other parameters except deltaCP and th23, which do not enter P\_ee */ \\
@@
 They may evolve into standard features in the future, or they may not be supported anymore at a certain point.
-In general, the number of experiment features is small in a new release version, and increases towards a new version number.
+In general, the number of experimental features is small in a new release version, and increases towards a new version number.
 

### 2007-01-15 Documentation/Paper2/paper2.tex
*** empty log message ***

@@
 
-\section{\AEDL\ changes and new experiment prototypes}
+\section{\AEDL\ changes and experiment prototypes}
 
-TABLE FROM MANUAL?
+\begin{table}[t!]
+{\small
+\begin{center}
+\begin{tabular}{llp{7.0cm}c}
+\hline
+Experiment & Filename & Short description &  \Refs \\
+\hline 
+\multicolumn{4}{l}{\underline{Superbeam experiments:}} \\
+T2K & {\tt T2K.glb} & J-PARC to Super-Kamiokande &  \cite{Itow:2001ee,Huber:2002mx} \\
+T2HK & {\tt T2HK.glb} & J-PARC to Hyper-Kamiokande &  \cite{Itow:2001ee,Huber:2002mx} \\
+NO$\nu$A & {\tt NOvA.glb} & Fermilab NuMI beamline off-axis  & \cite{Ambats:2004js,Yang_2004} \\
+SPL & {\tt SPL.glb} & CERN to Fr\'{e}jus & \cite{Campagne:2006yx,Campagne:2004wt,Mezzetto:2003mm} \\[0.2cm]
+\multicolumn{4}{l}{\underline{Reactor experiments:}} \\
+{\sc Reactor-I} & {\tt Reactor1.glb} & Small reactor exp., 
+$\mathcal{L} = 400 \, \mathrm{t} \, \mathrm{GW} \, \mathrm{yr}$
+& \cite{Huber:2003pm} \\
+{\sc Reactor-II} & {\tt Reactor2.glb} & Large reactor exp., 
+$\mathcal{L} = 8\, 000 \, \mathrm{t} \, \mathrm{GW} \, \mathrm{yr}$ 
+& \cite{Huber:2003pm} \\
+Double{\sc Chooz} & {\tt D-Chooz\_near.glb} & Double Chooz near 
+detector & \cite{Huber:2006vr} \\ 
+ & {\tt D-Chooz\_far.glb} & Double Chooz far detector & \\[0.2cm]
+\multicolumn{4}{l}{\underline{$\beta$-Beams:}} \\
+Low $\gamma$ & {\tt BB\_100.glb} & $\gamma=100$ CERN to Fr\'{e}jus 
+scenario & \cite{Campagne:2006yx} \\
+Medium $\gamma$ & {\tt BB\_350.glb} & $\gamma=350$ green field scenario 
+& \cite{Burguet-Castell:2005pa} \\
+Variable $\gamma$ & {\tt BBvar\_WC.glb} & Variable $\gamma$ beta beam with water Cherenkov detector, $50 \lesssim \gamma \lesssim 500$ & \cite{Huber:2005jk} \\[0.2cm]
+	& {\tt BBvar\_TASD.glb} & Variable $\gamma$ beta beam with totally active scint. detector (TASD), 
+$100 \lesssim \gamma \lesssim 3\, 000$ & \cite{Huber:2005jk} \\[0.4cm]
+\multicolumn{4}{l}{\underline{Neutrino factories:}} \\
+Standard & {\tt NFstandard.glb} & Standard neutrino factory, $50 \, \mathrm{kt}$ magnetized iron calorimeter, 
+$E_\mu = 50 \, \mathrm{GeV}$ & \cite{Huber:2002mx} \\
+Variable $E_\mu$ & {\tt NFvar.glb} & Variable neutrino factory, disapp. channels
+{\em without} CID; $10 \, \mathrm{GeV} \lesssim E_\mu \lesssim 80 \, \mathrm{GeV}$ & \cite{Huber:2002mx,Huber:2006wb}  \\
+Gold + Silver & {\tt NF\_GoldSilver.glb} & As {\tt NFvar.glb} 
+plus 5~kt ECC detector for Silver Channel
+measurement & \cite{Huber:2002mx,Huber:2006wb,Autiero:2003fu} \\
+Hybrid detector & {\tt NF\_hR\_lT.glb} & As {\tt NFvar.glb}, but 
+lower threshold and better energy resolution & \cite{Huber:2002mx,Huber:2006wb} \\
+\hline
+\end{tabular}
+\end{center}
+} % small
+\mycaption{\label{tab:experiments} 
+Pre-defined experiment prototypes coming with \GLOBESN : The columns represent the \AEDL\ filename, 
+a short description, and the references in which they are originally used and 
+discussed (except from minor modifications, such as a different implementation of 
+the energy threshold function). 
+%
+IS THE GAMMA=350 OPTION REALLY GREEN-FIELD? I THOUGHT THAT THIS MAY BE THE MAXIMUM
+ONE COULD GET OUT OF THE SPS WITH ACCELERATOR MODIFICATIONS? [WW]
+}
+\end{table}
 
-\bi
-\item 
- Built-in $\beta$-beam fluxes
-\item
- New AEDL features
-\item
- New AEDL files in standard GLoBES delivery + changes to existing ones
-\ei
+\AEDL\ (Abstract Experiment Definition Language) is a special syntax developed for the description of
+experiments in \GLOBES\ represented by simple text files. Core part of \AEDL\ are the following constructions
+(for details, see \Ref~\cite{Manual}):
+\begin{description}
+\item[Channel] A channel represents an oscillation channel from the source flux, over the initial and final
+flavors and the use of neutrinos or antineutrinos, to the interaction type and cross section. The result
+is an event rate vector.
+\item[Rule] A rule combines one or more signal channels with one or more background channels. The rates
+from these channels are added, and the signal and background are implemented with specific systematics, \ie,
+systematics is connected to the concept of a rule.
+The result of a rule is a systematics $\chi^2$ corresponding to \equ{chipull}.
+\item[Experiment] An experiment is a combination of one or more rules, which may correspond to different
+appearance and disappearance channels, neutrino and antineutrino operations modes, \etc.
+An experiment has a common detector and baseline, but may use more different fluxes (such as for neutrinos and antineutrinos).
+Therefore,
+the matter density profile is the same for all rules of an experiment. The result of an experiment
+is again a systematics $\chi^2$, but summed over all rules.
+\item[Experiment combination] An experiment combination consists of different experiments with different detectors, baselines, and (uncorreleted) matter density profiles.
+The result of an experiment combination
+is again a systematics $\chi^2$, but summed over all experiments.
+\end{description}
+Note that after the $\chi^2$ is summed over all rules or experiments as specified by the user in the application software, \equ{margin} is applied, and the oscillation parameters are marginalized over.
 
-Beta beams paper!
+In \GLOBESN , this construction of experiments is extended by user-defined systematics, which allows
+the user to break up the strict coupling of systematics with rules. For example, systematics can be
+correlated between different rules or experiments, such as for an reactor experiment with identical
+near and far detectors. In addition, one can introduce user-defined systematical errors.
+For that purpose, the former error dimension concept has been changed and
+extended. For example, one has to give each user-defined systematics a name, and this name has
+to be matched by the application software. 
+%
+Further changes in \AEDL\ are the requirement of a minimum \GLOBES\ version the \AEDL\ is running with,
+and the extension of lists as \AEDL\ variables and an interpolation feature.
 
-\section{New platform and interpreter support, new ways to improve the performance}
+As far as the experiment prototypes are concerned, a list of the ones provided with \GLOBESN\ 
+is given in \Tab~\ref{tab:experiments}. These prototypes can be modified for the definition of
+one's own experiments (experimentalist), or they can be used for the test of new physics concepts
+(phenomenologist/theorist).
+Note that some of these experiments may contain 
+integrated luminosities, baseline, fluxes, efficiencies, or other factors, which the user
+may not agree with and which can be easily adjusted. Some of the \AEDL\ files have been
+changed or updated from eariler versions, such as for T2K and T2HK, or the standard neutrino
+factory. Other files represent the results from recent developments, such as the variable energy
+beta beam or neutrino factory files. They push \AEDL\ to the edge in terms of advanced features,
+such as variable bin widths, use of the new interpolation feature, threshold function implementations, \etc.
+In particular, note that Double Chooz requires user-defined
+systematics, and the beta beams built-in beta beam fluxes. Both of these features are only
+available in \GLOBESN\ and higher. Some of the files require that \AEDL\ variables be pre-set by the
+user, such as the muon energy for the variable energy neutrino factory. These variables are
+described in the corresponding file headers. Old \AEDL\ prototypes should still run, but will not
+be maintained by the \GLOBES\ team anymore.
+
+\section{Internal changes and improved performance}
+
+New platform and interpreter support, new ways to improve the performance
 

### 2007-01-15 Documentation/Paper2/paper2.tex
*** empty log message ***

@@
 simplified by the concept of user-defined priors. As far as the probability calculation is concerned, \GLOBES\ 
-now provides an interface for the inclusion of non-standard physics without re-compliation of the software. 
+now provides an interface for the inclusion of non-standard physics without re-compilation of the software. 
 The set of experiment descriptions coming with \GLOBES\ has been updated. For example, built-in fluxes are now
@@
 with the eigenvalues $m_1^k$, $m_2^k$, and $m_3^k$. The probabilities in \equ{probs} are used for the
-event rate compution, which is described in greater detail in \Ref~\cite{Huber:2004ka}. In the next section,
+event rate computation, which is described in greater detail in \Ref~\cite{Huber:2004ka}. In the next section,
 we will demonstrate how systematics is implemented thereafter.
@@
 re-compilation of the \GLOBES\ software. This is implemented
-by using pointers for the functions calling the standard probability engine, which can be changed if
+by using pointers to the functions of the standard probability engine, which can be changed if
 a different probability engine is registered. For example, a user may just copy the standard probability
-engine of \GLOBES , and modify it, and register it. An important prerequisite for the ability to handle
+engine of \GLOBES , modify it, and register it. An important prerequisite for the ability to handle
 more than the standard six oscillation parameters (plus matter density) $\boldsymbol{\lambda}=(\theta_{12},\theta_{13}, \theta_{23}, \deltacp, \sdm, \ldm, \rho)$, \ie,
@@
 normalization errors, as well as signal and background tilt or energy calibration errors.
-The tilt (T) is impemented as a linear distortion of the spectrum around the center,
+The tilt (T) is implemented as a linear distortion of the spectrum around the center,
 and the energy calibration (C) as a distortion (stretching) of the reconstructed energy scale.
@@
 In order to include correlations and degeneracies, \GLOBES\ provides
-(local) maginalization routines to project the fit manifold onto
+(local) marginalization routines to project the fit manifold onto
 a subspace. For example, if one wants to compute the $\stheta$ precision,
@@
 oscillation parameter). For example, an external constraint on $\theta_{12}$
-is added before the marginalition in \equ{margin} by the replacement after the
-systematics minimization
+is added before the marginalization in \equ{margin} by the replacement
 \begin{equation}
@@
 is again a systematics $\chi^2$, but summed over all rules.
-\item[Experiment combination] An experiment combination consists of different experiments with different detectors, baselines, and (uncorreleted) matter density profiles.
+\item[Experiment combination] An experiment combination consists of different experiments with different detectors, baselines, and (uncorrelated) matter density profiles.
 The result of an experiment combination
@@
 In \GLOBESN , this construction of experiments is extended by user-defined systematics, which allows
-the user to break up the strict coupling of systematics with rules. For example, systematics can be
+the user to break up the strict coupling between systematics and rules. For example, systematics can be
 correlated between different rules or experiments, such as for an reactor experiment with identical
@@
 %
-Further changes in \AEDL\ are the requirement of a minimum \GLOBES\ version the \AEDL\ is running with,
-and the extension of lists as \AEDL\ variables and an interpolation feature.
+Further changes in \AEDL\ are the requirement to specify the minimum \GLOBES\ version the \AEDL\ is running with,
+and the extensions by lists as \AEDL\ variables and by an interpolation feature.
 
@@
 may not agree with and which can be easily adjusted. Some of the \AEDL\ files have been
-changed or updated from eariler versions, such as for T2K and T2HK, or the standard neutrino
+changed or updated from earlier versions, such as for T2K and T2HK, or the standard neutrino
 factory. Other files represent the results from recent developments, such as the variable energy
@@
 In particular, note that Double Chooz requires user-defined
-systematics, and the beta beams built-in beta beam fluxes. Both of these features are only
+systematics, and the beta beams require built-in beta beam fluxes. Both of these features are only
 available in \GLOBESN\ and higher. Some of the files require that \AEDL\ variables be pre-set by the
@@
 \section{Internal changes and improved performance}
-
-New platform and interpreter support, new ways to improve the performance
-
-\bi
-\item
- PERL?
-\item 
- Windows?
-\item
- Better RedHat support
-\item
- Changes in installation scheme
-\ei
-
-\bi
-\item 
- GLoBES on clusters, Condor; parallized code; static links
-\item 
- 64 bit support
-\item 
- New minimization algorithm!?
-\ei
 
@@
 The numerical method that does the minimization is the Powell
-algorithm~\cite{Press:NumRecip}.
+algorithm HERE ORIGINAL POWELL REFERENCE [WW].
 
 In \GLOBESN , we provide an alternative method, which does the full minimization
-in one go and is therefore several times faster in most cases. In this algorithm,
+in one go, and is therefore several times faster in most cases. In this algorithm,
 each Powell iteration over oscillation parameters is followed by one iteration
@@
 step here is the diagonalization of the hermitian $3\times3$ neutrino Hamilton
-operator in matter. The LAPACK rotuines~\cite{Anderson:LAPACK}, which were
+operator in matter. The LAPACK routines~\cite{Anderson:LAPACK}, which were
 used to solve this eigenproblem in previous versions of \GLOBES, turned out
@@
 of hermitian $3\times3$ matrices. It is essentially the extremely fast analytical
-algorithm discussed in~\cite{Kopp:2006wp}, but for the rare cases where this may become
-inaccurate, it falls back to the QL method from~\cite{Press:NumRecip}.
+algorithm discussed in \Ref~\cite{Kopp:2006wp}, but for the rare cases where this may become
+inaccurate, it falls back to the QL (WRITE OUT!)
+method from~\cite{Press:NumRecip} REPLACE BY ORIGINAL REFERENCE!.
+
+As far as the installation of \GLOBES\ on different systems is concerned, \GLOBES\
+has been improved in the past, such as to handle different Linux distributions.
+In addition, for high-throughput computing, \GLOBES\ has been used or tested
+on 64 Bit systems, Condor clusters, and parallel clusters. An important prerequisite
+for the parallelization is the option to build static binaries, which is now
+illustrated in the examples. With these improvements, the performance of \GLOBES\
+has been pushed to the edge. \Ref~\cite{Huber:2006wb} is probably the currently
+most demanding product of \GLOBES , for which most of these features
+have been developed.
+
+PH, JK: IS THERE ANYTHING ELSE YOU WANT TO SAY? ABOUT NEW PLATFORMS,
+PERL, WINDOWS, PERFORMANCE, 64BIT, INSTALLATION?
 
 \section{Summary and conclusions}
+
+We have presented \GLOBESN , which is a simulation software for 
+future terrestrial neutrino oscillation facilities. \GLOBES\ is now no further
+restricted to one source-one detector configurations, but it can simulate
+the systematics of arbitrary combinations of discrete sets of sources and
+detectors. In addition, user-defined systematical errors can be introduced.
+The new concept of user-defined priors allows the inclusion of external
+information at a different level. For example, constraints from atmospheric
+neutrino experiments or neutrino telescopes can be taken into account. In addition,
+they provide a powerful handle on complicated degeneracy localization.
+On the probability level, \GLOBES\ now supports the simulation of non-standard
+physics beyond three-flavor neutrino oscillations without a re-compilation of
+the software. A very important prerequisite for this feature is the ability
+of \GLOBESN\ to carry more than six oscillation parameters. New experiment
+prototypes make extensive use these new features, as well as they are illustrated by
+examples in the manual. One prominent example is the Double Chooz experiment,
+which requires user-defined systematics to simulate correlated errors between
+the near and far detectors.
+
+In conclusion, \GLOBESN\ provides a number of new improvements, such as by
+ including more flexibility. At the end, the developers of a software can never
+exactly foresee what the users are up for, and therefore flexible concepts
+allow for more space in problem realizations. In this version of \GLOBES ,
+three new such flexible concepts have been introduced: User-defined systematics,
+user-defined priors, and user-defined oscillation probabilities. These approaches
+will make an easy adjustment of the software possible for both experimentalists and theorists.
+Finally, note that the past has shown that it is often more than one possibility 
+how a specific problem can be implemented in \GLOBES .
 

### 2007-01-15 Documentation/Paper2/paper2.tex
*** empty log message ***

@@
 \hline
-Experiment & Filename & Short description &  \Refs \\
+Experiment & File name & Short description &  \Refs \\
 \hline 
@@
  & {\tt D-Chooz\_far.glb} & Double Chooz far detector & \\[0.2cm]
-\multicolumn{4}{l}{\underline{$\beta$-Beams:}} \\
+\multicolumn{4}{l}{\underline{Beta beams:}} \\
 Low $\gamma$ & {\tt BB\_100.glb} & $\gamma=100$ CERN to Fr\'{e}jus 

### 2007-01-15 Documentation/part2.tex
JK - Minor corrections in chapter 10

@@
 
-Here the general concept of the AEDL is described and illustrated by an example. In addition, a short introduction to the syntax of the AEDL is given.
+Here, the general concept of the AEDL is described and illustrated by an example. In addition, a short introduction to the syntax of the AEDL is given.
 
@@
 For each of these layers, the Hamiltonian in matter is diagonalized in order to propagate the neutrino transition amplitudes.
-Finally, the transition probability is obtained by the absolute square of the total neutrino transition amplitudes.
-Depending on the precision of the studied experiment, this approach turns out to be precise enough in Earth matter even if only a small number matter density steps is used.
+Finally, the transition probability is obtained as the absolute square of the total neutrino transition amplitudes.
+Depending on the precision of the studied experiment, this approach turns out to be precise enough in Earth matter even if only a small number of matter density steps is used.
 Since we allow an uncertainty of the matter density profile, it is, in fact, in most cases sufficient to consider only one density step with the average matter density together with a matter density uncertainty~\cite{Ohlsson:2003ip}.
@@
 The same, in principle, applies to the nature of the neutrino flavor.
-However, in this case only discrete values are applicable.
+However, in this case, only discrete values are applicable.
 Note that the reconstructed neutrino energy and the neutrino flavor are the only observables in \GLOBES .
@@
 
-In order to implement a experiment definition including various
+In order to implement an experiment definition including various
 sources of systematical errors, we use several abstraction levels. 
@@
 for simulated data $\Delta \chi^2 = 0$ at the best-fit point. Thus, we
-are using $\chi^2$ and $\Delta \chi^2$ as equal quantities.} In addition, each rule implements an {\em independent} systematics, such as signal and background normalization errors.
+are using $\chi^2$ and $\Delta \chi^2$ as equal quantities.} In addition, each rule implements {\em independent} systematics, such as signal and background normalization errors.
 Eventually, each rule gives a $\Delta \chi^2$-value, and the total $\Delta \chi^2$ of one experiment is obtained by adding the $\Delta \chi^2$'s of all rules (\cf, \figu{experiment}).
@@
 \item
- Within each rule the event numbers are added.
+ Within each rule, the event numbers are added.
 \item
- Within each rule the systematics is treated independently from the other rules.
+ Within each rule, the systematics is treated independently from the other rules.
 \item
- For each rule the $\Delta \chi^2$ is computed; the $\Delta \chi^2$'s from all rules are added.
+ For each rule, the $\Delta \chi^2$ is computed; the $\Delta \chi^2$'s from all rules are added.
 \end{itemize}
@@
 
-Experiments are in \GLOBES\ defined by the Abstract Experiment Definition Language (AEDL).
+Experiments in \GLOBES\ are defined by the Abstract Experiment Definition Language (AEDL).
 The experiment definition is written into a text file using the AEDL syntax.
 Currently, a number of pre-defined experiment definition files are provided with \GLOBES , which have to be modified manually in order to define new experiments.
-The application software then uses this text file to initialize the experiment, where other secondary files might read for source fluxes, cross sections \etc .
+The application software then uses this text file to initialize the experiment, where other secondary files might be read for source fluxes, cross sections \etc .
 In this section, we show the definition of a very simple neutrino factory in AEDL, where we do not go into details.
@@
 originating from stored $\mu^+$'s.
-This achieved by setting the {\tt @builtin} variable to $1$.
+This is achieved by setting the {\tt @builtin} variable to $1$.
 Next, we specify the muon energy to be $50\,\mathrm{GeV}$ by the {\tt @parent\_energy} variable.
@@
 \end{quote}
-Then we specify the file which contains the cross sections we want to 
+Then, we specify the file which contains the cross sections we want to 
 use:
@@
 {\tt \$lengthtab} refer to a list of numbers. Here, the lists contain only
-one element, because we only use one density layer: We initialize a baseline length of $3000 \, \mathrm{km}$ with a constant matter density of $3.5 \, \mathrm{g/cm^3}$. 
+one element each, because we only use one density layer: We specify a
+baseline length of $3000 \, \mathrm{km}$ with a constant matter density of
+$3.5 \, \mathrm{g/cm^3}$. 
 %
@@
 The {\tt @signal} refers to the ``signal'' in our experiment. We use the
-above defined channel named {\tt \#appearance} with an constant overall
+above defined channel named {\tt \#appearance} with a constant overall
 efficiency of $0.45$ (in a more realistic simulation, one would introduce an
@@
 %simultaneously. It is only useful if there is more than one background component, otherwise it is usually $1$. 
-The background error variable is defined such as the signal error variable, \ie , we have a $20\%$ background uncertainty and a very small energy calibration error.
+The background error variable is defined in the same way as the signal error variable, \ie , we have a $20\%$ background uncertainty and a very small energy calibration error.
 In addition, the systematics treatment is specified
@@
 
-The here defined experiment represents a first simplified version of a neutrino factory experiment.
+The experiment defined here represents a first simplified version of a neutrino factory experiment.
 It still lacks the correct energy dependence of the efficiencies, the antineutrino disappearance channel, and the channels and rules for the symmetric operation with $\mu^-$ stored.
 However, it may serve as a simple, introductory example.
-In the next chapter, we will demonstrate that the AEDL is much more powerful than illustrated here.
+In the next chapter, we will demonstrate that AEDL is much more powerful than illustrated here.
 
@@
 in order to avoid parsing megabytes of chunk
- and producing thousands of error messages.  In addition, the version of the minimum
-\GLOBES\ version the \AEDL\ file is running with has to be defined by {\tt \$version}, such as
+ and producing thousands of error messages.  In addition, the minimum
+\GLOBES\ version that the \AEDL\ file is supposed to run with has to be
+defined by a {\tt \$version} statement, such as
 {\tt \$version="3.0.0"}.
@@
 %
-Comments can be used such as in C:
+Comments can be used in the same way as in C:
 \begin{quote}
@@
 The counter keeps track of how many different names
-there are for one type of environment, which means that it counts the number of channels, rules, energy resolution functions \etc . Thus, the automatic
+there are for one type of environment, which means that it counts the number of channels, rules, energy resolution functions, \etc . Thus, the automatic
 variables are numbered in the order of their definition, and the number
@@
 \index{aedl}{include@{\tt include}}
-As another feature of AEDL one can use include files with the {\tt include} command. Includes can be nested up to {\tt MAX\_INCLUSION\_DEPTH}, which is currently set to $10$. Error reporting works 
+As another feature of AEDL one, can use include files with the {\tt include} command. Includes can be nested up to {\tt MAX\_INCLUSION\_DEPTH}, which is currently set to $10$. Error reporting works 
  for nested includes, too. The included file is not required to begin 
@@
 {\tt
-Pi = 3.141
+Pi = 3.14159
 }
@@
 {\tt log}, {\tt log10}, {\tt exp}, {\tt sqrt}.
-\index{aedl}{sin@{\tt sin}}
-\index{aedl}{cos@{\tt cos}}
-\index{aedl}{tan@{\tt tan}}
-\index{aedl}{asin@{\tt asin}}
-\index{aedl}{acos@{\tt acos}}
-\index{aedl}{atan@{\tt atan}}
-\index{aedl}{log@{\tt log}}
-\index{aedl}{log10@{\tt log10}}
-\index{aedl}{exp@{\tt exp}}
-\index{aedl}{sqrt@{\tt sqrt}}
+\index{aedl}{sin@{\tt sin}}%
+\index{aedl}{cos@{\tt cos}}%
+\index{aedl}{tan@{\tt tan}}%
+\index{aedl}{asin@{\tt asin}}%
+\index{aedl}{acos@{\tt acos}}%
+\index{aedl}{atan@{\tt atan}}%
+\index{aedl}{log@{\tt log}}%
+\index{aedl}{log10@{\tt log10}}%
+\index{aedl}{exp@{\tt exp}}%
+\index{aedl}{sqrt@{\tt sqrt}}%
 % 
@@
 otherwise only a scalar number would appear. However, they can not be
-applied to lists, such as {\tt sin(\{1,2,3\})} will not work.
+applied to lists, \ie, expressions like {\tt sin(\{1,2,3\})} will not work.
 
@@
 In \GLOBES\ 3.0 and higher, a number of new features can be used. The
-features in this section are slighlty more advanced and can be jumped over at a first reading of
+features in this section are slighlty more advanced and can be skipped in a first reading of
 the manual. The
@@
 \end{quote}
-Functions can be threaded over lists, \ie, by applying the function to each element
-and returning a list. Note that the original list will be destroyed by this
+Functions can be threaded over lists, \ie, they will be applied to each element
+of list, and return a list. Note that the original list will be destroyed by this
 process. Therefore, it is necessary to create a copy of your list if you want to
@@
 Two helper functions {\tt bincenter()} 
-\index{aedl}{bincenter@{\tt bincenter}} and {\tt simbincenter} \index{aedl}{simbincenter@{\tt simbincenter}} return lists with the central energies of the bins or sampling points, respectively.
+\index{aedl}{bincenter@{\tt bincenter}} and {\tt samplingbincenter} \index{aedl}{samplingbincenter@{\tt samplingbincenter}} return lists with the central energies of the bins or sampling points, respectively.
 For example,
@@
 
-For debugging purposes of lists and numbers, one can \GLOBES\ now supports output directly
+To simplify debugging of lists and numbers, \GLOBES\ now supports output directly
 from \AEDL\ files:

### 2007-01-15 Documentation/part2.tex
JK - Minor corrections in chapter 11

@@
 
-In this chapter, we give a detailed definition of the \AEDL\ features.
-We also show the underlying mathematical concepts, where applicable.
+In this chapter, we give a detailed description of the \AEDL\ features.
+We also show the underlying mathematical concepts where applicable.
 We do not exactly follow the separation of source, oscillation, and detection properties, since most issues more or less involve the detection.
@@
 As we have disussed before, \GLOBES\ can only deal with point sources.
-Thus,  it is not possible to study effects from the finite size of the neutrino production region, such as in the sun or in reactor experiments with many
+Thus,  it is not possible to study effects of the finite size of the neutrino production region, such as in the sun or in reactor experiments with many
 neutrino sources (\eg, KamLAND).
@@
 with a running time of $8$ years. Note that the running time is used within
-the {\tt nuflux} environment. This feature can be used to load the neutrino and antineutrino fluxes separately, in order to combine them with different
-running times within one experiment.
+the {\tt nuflux} environment. This feature can be used to load the neutrino
+and antineutrino fluxes in an accelerator experiment separately, in order
+to combine them with different running times in the respective operation modes.
 The name of the flux {\tt \#name} will later be referred to in the channel definitions.
@@
 For a built-in neutrino source, one has to specify which
-built-in spectrum has to be used, as well as its parameters. The software
+built-in spectrum should be used, as well as its parameters. The software
 will then automatically calculate the neutrino spectrum. Note that in this
@@
 \index{aedl}{nuflux@{\tt nuflux}!flux file@{\tt "@flux\_file}}
-For a user-defined flux, one has to give it the file name:
+For a user-defined flux, one has to specify the file name:
 \begin{quote}
@@
 Usually this adjustment of {\tt @norm} is performed by comparison with known energy spectra for a specific experiment.
-For more details on the flux definition, see appendix on page~\pageref{app:flux}.
+For more details on the flux definition, see page~\pageref{app:flux} in the appendix.
 
@@
 In order to access fluxes at arbitrary energies, linear interpolation 
-is used by the software.
-In general, it is advisable to provide the flux between {\tt \$sampling\_min} and {\tt \$sampling\_max} (\cf, \Sec~\ref{sec:energy}), since these values are used by the software.
-However, if the energy leaves the range of values given in the file, zero is returned.
-The the columns for the fluxes for unused flavours have to be filled all the 
-same,\eg\ with zeros.
+is used by \GLOBES. In general, it is advisable to provide the flux
+between {\tt \$sampling\_min} and {\tt \$sampling\_max} (\cf, \Sec~\ref{sec:energy}),
+since this is the energy range considered in the simulation. However,
+if part of this interval is omitted in the flux file, zero will be
+used there. If some neutrino flavours are not used in the simulation,
+the corresponding columns in the flux file have to be filled nevertheless,
+\eg\ with zeros.
 
@@
 or the origin of the information. This is also the default method
-to point the user to the references one should to cite when using
-a particular flux.\index{norm}{Flux!file!comments in}
+to point the user to the references that should be cited when using
+a particular flux file.\index{norm}{Flux!file!comments in}
 \index{norm}{Referencing!flux data}
@@
 
-The baseline and matter density profile determine, besides energy and
-involved flavors, the neutrino oscillation physics at the experiment
-description level. All of the neutrino oscillation parameters are defined at running time.
+Besides the energy and the involved flavors, the neutrino oscillation
+physics depends on the baseline and the matter density profile.
+All neutrino oscillation parameters are defined at running time.
 
@@
 has to be specified.
-The simplest matter density profile is a constant matter density profile with the average matter density from the PREM~\cite{Stacey} onion shell model of the earth:\index{norm}{PREM| \see{Matter density}}
+The simplest profile is a constant matter density equal to the average matter density from the PREM~\cite{Stacey} onion shell model of the earth:\index{norm}{PREM| \see{Matter density}}
 \index{norm}{Matter density!of the earth}
@@
 %
-If your using this option please cite reference~\cite{Stacey}\index{norm}{Referencing!matter profile data}.
+If you are using this option please cite reference~\cite{Stacey}\index{norm}{Referencing!matter profile data}.
 
@@
 In general, it is advisable to provide the cross sections in the range between {\tt \$sampling\_min} and {\tt \$sampling\_max} (\cf, \Sec~\ref{sec:energy}).
-Unused cross sections have to be filled with zeros, and can not
- be just omitted.
+Cross sections for unused neutrino flavours have to be filled with zeros, and
+can not just be omitted.
 
@@
 accept one-line comments, which start
-with {\tt \#} and end with the linefeed character `\verb+\n+', they are
-not counted as a line and their content is discarded. This comments
+with {\tt \#} and end with the linefeed character `\verb+\n+'; they are
+not counted as a line and their content is discarded. These comments
 are useful to provide meta information about the cross sections like units
 or the origin of the information. This is also the default method
-to point the user to the references he/she should to cite when using
-a particular cross section.\index{norm}{Cross section!file!comments in}
+to point the user to the references he/she should cite when using
+a particular cross section file.\index{norm}{Cross section!file!comments in}
 \index{norm}{Referencing!cross section data}
@@
 between the pure oscillation physics given by the oscillation probability
-$P_{\alpha\beta}$, and the total event rates composed of signal and 
+$P_{\alpha\beta}$, and the final event rates composed of signal and 
 background. A channel describes the path from one
@@
 initial neutrino flavor, its CP eigenvalue 
-(neutrino or antineutrino)\footnote{Currently \GLOBES\ does not support lepton
+(neutrino or antineutrino)\footnote{Currently, \GLOBES\ does not support lepton
 number violating transitions, \ie\ no transitions from neutrino to 
@@
 
-Before we come to the definition of the channel in AEDL, we introduce the general concept for the calculation of event rates. The first step is to
+Before we come to the definition of channels in AEDL, we introduce the general concept for the calculation of event rates. The first step is to
 compute the number of events for each IT in the detector for each 
-initial and final neutrino flavor and energy bin.
-The second step is to include the detector effects coming from the insufficient knowledge in the event reconstruction.
-These two steps combined lead to the differential event rate spectrum for each initial and final flavor and IT as seen by the detector, which we call the ``channel''.
+initial and final neutrino flavor and energy bin. The second step is to include the detector effects coming from the insufficient knowledge in the event reconstruction.
+The combination of these two steps leads to the differential event rate spectrum for each initial and final flavor and IT as seen by the detector, which we call the ``channel''.
 In this section, we focus on the
@@
 \underbrace{\Phi_{\alpha} (E)}_{\mathrm{Production}} \times \nonumber\\
-&&\underbrace{\frac{1}{L^2} P_{(\alpha\rightarrow\beta)}(E,L,\rho;\theta_{23},
-\theta_{12},\theta_{13},
+&&\underbrace{\frac{1}{L^2} P_{(\alpha\rightarrow\beta)}(E,L,\rho;\theta_{12},
+\theta_{13},\theta_{23},
 \Delta m^2_{31},\Delta m^2_{21},\deltacp)}_{\mathrm{Propagation}}
@@
 
-Since it is a lot of effort to solve this double integral numerically,
-we split up the two integrations. First, we evaluate the integral over
-$\hat{E}$, where the only terms containing $\hat{E}$ are
-$k_\beta^{\text{IT}}(E-\hat{E})$,  $ T_\beta(\hat{E})$, and 
-$ V_\beta(\hat{E}-E')$. We define:
+Since it is computationally very expensive to solve this double integral
+numerically, we split up the two integrations. The integral over $\hat{E}$
+depends only on the terms containing $\hat{E}$, i.e., on $k_\beta^{\text{IT}}(E-\hat{E})$,
+$ T_\beta(\hat{E})$, and $V_\beta(\hat{E}-E')$. These terms do not depend
+on the oscillation parameters, so they will not vary during the fit,
+and the $\hat{E}$ integral can be pre-computed in the initialization phase.
+We define:
 \begin{eqnarray}
@@
 with an energy between $E'$ and $E'+dE'$ with a probability
-$R_\beta^{\text{IT}}(E,E') dE'$. The function $R(E,E')$ is also often called ``energy resolution function''. Actually, its internal representation
+$R_\beta^{\text{IT}}(E,E') dE'$. The function $R(E,E')$ is often called ``energy resolution function''. Actually, its internal representation
 in the software is a smearing matrix\index{norm}{Energy!resolution}\index{norm}{Smear matrix}.
@@
 the cross sections, and the energy resolution function. In order to refer to
-flux, cross sections, and energy resolution functions, they have to be 
-defined before with their {\tt \#name} in the respective environments. 
-Thus, a simple definition of a channel\index{aedl}{channel@{\tt channel}} is
+the fluxes, cross sections, and energy resolution functions, they have to be 
+defined first with their {\tt \#name} in the respective environments. 
+A simple definition of a channel\index{aedl}{channel@{\tt channel}} is
 \begin{quote}
@@
 of channels exist which are only used for the computation of backgrounds.
-Usually it is an excellent approximation to treat backgrounds as if they
+Usually, it is an excellent approximation to treat backgrounds as if they
 were not affected by oscillations\footnote{In the case that the backgrounds have
@@
 Then we introduce an automatic energy smearing algorithm, which is fairly simple to understand and applicable to most beam-based experiments.
-In most cases, the reader may want to continue to the next section after reading these two subsections.
-In the third subsection, we describe a more elaborate (and slower) smearing algorithm, which can be used together with rather fast neutrino oscillations compared to the bin size, such as for
-(solar) reactor experiments to avoid aliasing effects. Eventually, we show how one can use a manual smearing matrix instead of using one of the implemented algorithms.
+In most cases, the reader may want to proceed to the next section after reading these two subsections.
+In the third subsection, we describe a more elaborate (and slower) smearing algorithm, which is useful to avoid aliasing effects if the neutrino oscillations are rather fast compared to the bin size, as is the case for
+(solar) reactor experiments. Eventually, we show how one can use a manual smearing matrix instead of using one of the implemented algorithms.
 
@@
 
-The energy resolution function $R^c(E,E')$ has been already introduced in 
+The energy resolution function $R^c(E,E')$ and the post-smearing
+efficiencies $\epsilon(E')$ have already been introduced in 
 \Sec~\ref{sec:channel}, where a definition
@@
 definition of the post-smearing efficiencies $\epsilon(E')$. 
-In general, $\epsilon(E')$ has to be
+In general, they have to be
 determined by means of a Monte Carlo simulation of the experiment. 
@@
 for equidistant sampling points. If no values are given for these variables,
-they are assumed to be equal to their corresponding counter parts at the bin level,
+they are assumed to be equal to their corresponding counterparts at the bin level,
 \ie , \verb+$sampling_points = $bins+, \verb+$sampling_min = $emin+ and
@@
 bin sizes much smaller than the energy resolution will not improve the results.
-The energy bin range and the number of energy bins do always have to be specified.
+The energy bin range and the number of energy bins always have to be specified.
 For the case of large values of the integrand in \eq~(\ref{eq:simple_int}) at the energy range limits, it is recommended to exceed the analysis energy window by about three times the energy calibration error in order
@@
 
-As illustrated in the figure, an interesting feature in combination with the
+As illustrated in \figu{energies}, an interesting feature in combination with the
 channels are pre- and post-smearing effects. Pre-smearing effects are taken into account on the sampling point level, and post-smearing effects on the bin level. Examples for these effects are energy 
@@
 These components can be introduced before or after the integration in \eq~(\ref{eq:simple_int}) is done.
+%
+\index{aedl}{channel@{\tt channel}!pre smearing efficiencies@{\tt 
+"@pre\_smearing\_efficiencies}}%
+\index{aedl}{channel@{\tt channel}!pre smearing background@{\tt 
+"@pre\_smearing\_background}}%
+\index{aedl}{channel@{\tt channel}!post smearing efficiencies@{\tt 
+"@post\_smearing\_efficiencies}}%
+\index{aedl}{channel@{\tt channel}!post smearing background@{\tt 
+"@post\_smearing\_background}}%
+%
+If they are introduced before, 
+we call them
+{\tt @pre\_smearing\_efficiencies} or {\tt @pre\_smearing\_background}. 
+If they are introduced afterwards, we call them {\tt @post\_smearing\_efficiencies} or {\tt @post\_smearing\_background}.
+Note that pre-smearing components are always a function of the incident neutrino energy $E$. Thus, there have to be as  many elements as there are sampling points. 
+Examples for pre-smearing quantities are non-beam backgrounds, such as from geophysical neutrinos.
+The post-smearing components are always a function of the reconstructed neutrino energy $E'$, such as the post-smearing efficiencies $\epsilon_\beta^{\text{IT}}(E')$ in \eq~(\ref{eq:e_res}).
+Examples for post-smearing efficiencies are cuts and detection threshold functions.
+All post-smearing components have to have as
+many elements as there are energy bins. Efficiencies are multiplicative 
+and their default value is $1$, whereas backgrounds are additive and their default value is $0$. Thus, a more elaborate channel can be defined as
+\begin{quote}
+{\tt channel(\#channel\_1)<\\
+\tb @channel = \#flux : $+$: muon: muon: \#cross: \#energy\\
+\tb @pre\_smearing\_background = \{1,2,3,4,5,6,7,8,9,10\}\\
+\tb @post\_smearing\_efficiencies = \{0.1,0.2,0.3,0.4,0.5\}\\
+>}
+\end{quote}
 %
@@
 %
-If they are introduced before, 
-we call them
-{\tt @pre\_smearing\_efficiencies} or {\tt @pre\_smearing\_background}. 
-If they are introduced after, we call them {\tt @post\_smearing\_efficiencies} or {\tt @post\_smearing\_background}.
-Note that pre-smearing components are always a function of the incident neutrino energy $E$. Thus, there have to be as  many elements as there are sampling points. 
-Examples for pre-smearing quantities are non-beam backgrounds, such as from geophysical neutrinos.
-The post-smearing components are always a function of the reconstructed neutrino energy $E'$, such as the post-smearing efficiencies $\epsilon_\beta^{\text{IT}}(E')$ in \eq~(\ref{eq:e_res}).
-Examples for post-smearing efficiencies are cuts and detection threshold functions.
-All post-smearing components have to have as
-many elements as there are energy bins. Efficiencies are multiplicative 
-and their default value is $1$, whereas backgrounds are additive and their default value is $0$. Thus, a more elaborate channel can be defined as
-\begin{quote}
-{\tt channel(\#channel\_1)<\\
-\tb @channel = \#flux : $+$: muon: muon: \#cross: \#energy\\
-\tb @pre\_smearing\_background = \{1,2,3,4,5,6,7,8,9,10\}\\
-\tb @post\_smearing\_efficiencies = \{0.1,0.2,0.3,0.4,0.5\}\\
->}
-\end{quote}
-%
-\index{aedl}{channel@{\tt channel}!pre smearing efficiencies@{\tt 
-"@pre\_smearing\_efficiencies}}
-\index{aedl}{channel@{\tt channel}!pre smearing background@{\tt 
-"@pre\_smearing\_background}}
-\index{aedl}{channel@{\tt channel}!post smearing efficiencies@{\tt 
-"@post\_smearing\_efficiencies}}
-\index{aedl}{channel@{\tt channel}!post smearing background@{\tt 
-"@post\_smearing\_background}}
-%
 This experiment uses $10$ sampling points and $5$ bins.
 
-In the following subsections we will define the energy resolution function.
+In the following subsections we will explain the energy resolution function.
 All energy resolution functions are defined within an {\tt energy} environment and can be refered to by {\tt \#name}:
@@
 will cancel between the simulated and fitted data. Second, and more important,
-this algorithm is just a very basic integration routine\footnote{It is planed 
-for to have something like a Gau\ss-Kronrod scheme as an alternative here.} and converges to
-the true result for decreasing step size. Thus if the number of sampling
-points is large enough this algorithm is very accurate. 
-This algorithm is selected by
+this algorithm is just a very basic integration routine\footnote{It is planned 
+for the future to implement something like a Gau\ss-Kronrod scheme as an
+alternative here.} and converges to the true result for decreasing step size.
+Thus if the number of sampling points is large enough this algorithm is very
+accurate. Bin-based energy smearing is selected by
 %
@@
 above a certain threshold, which is, by default, $10^{-7}$. 
-This threshold is defined at the compiling time.
-
-Eventually,  a complete energy resolution definition of this 
-algorithm is, for example,
+This threshold is defined at compilation time.
+
+Eventually,  a complete energy resolution definition with bin-based
+automatic energy smearing is, for example,
 \begin{quote}
@@
 
-In order to ensure that fast oscillating probabilities do not lead to 
+In order to ensure that very fast oscillations do not lead to 
 aliasing,\index{norm}{Aliasing}
 it is possible to impose a low-pass filter already during the calculation
-of the probabilities itself. This filter is implemented has a highly
-experimental feature called ``filter''\index{norm}{Filter}. 
+of the probabilities itself. This highly
+experimental feature will be called ``filter''\index{norm}{Filter} in the following. 
 The calculation of oscillation
@@
 $\sigma_f(E):=\sigma_f^0 \cdot E$ ensures that this filter behaves 
-approximately such as an energy resolution function with constant width 
+approximately as an energy resolution function with constant width 
 $\sigma_e=\sqrt{2}/\sigma_f^0$, \ie\
@@
 \end{equation}
-The relationship between \eqs~\ref{eq:filter_a} and~\ref{eq:filter_b}
+The relationship between \eqs~\eqref{eq:filter_a} and~\eqref{eq:filter_b}
 is not obvious and connected to the properties of $P_{\alpha\beta}$: 
@@
 With a type 1 ({\tt @type = 1}) energy resolution 
-function, $\sigma_e$ adds on to the energy resolution function 
-of the detector $\sigma_c(E)$ such as
+function, $\sigma_e$ contributes to the energy resolution function 
+of the detector $\sigma_c(E)$ according to
 \begin{equation}
@@
 'subtract' the filtering from the energy resolution function by splitting 
-the energy resolution function $\sigma(E)_{\mathrm{eff}}$ in
+the energy resolution function $\sigma(E)_{\mathrm{eff}}$ into
 two parts by
@@
 
-In some cases, one may use the output of a detector Monte Carlo simulation
+In some cases, one may want to use the output of a detector Monte Carlo simulation
 directly. Then one can use ''manual'' energy smearing instead  of the
@@
 
-The energy smearing matrix $K_{ij}$ has {\tt \$bins} rows and {\tt \$sampling\_points} columns, which are numbered from $0$ to {\tt \$bins}$-1$ or {\tt \$sampling\_points}$-1$.
+The energy smearing matrix $K_{ij}$ has {\tt \$bins} rows and {\tt \$sampling\_points} columns, which are numbered from $0$ to {\tt \$bins}$-1$ resp.\ {\tt \$sampling\_points}$-1$.
 It is equivalent to the
@@
 In general, many of the entries in this matrix are zero, which means that it is convenient to evaluate the integrand in \eq~(\ref{eq:simple_int}) only at positions where
-$K_{ij}$ is non-zero. The corresponding ``sampling range'' range of non-zero matrix entries  in $K_{ij}$ for the $i$th energy bin is defined to run from
+$K_{ij}$ is non-zero. The corresponding ``sampling range'' of non-zero matrix entries  in $K_{ij}$ for the $i$th energy bin is defined to run from
 column $k_l^i$ (``lower index'') to column $k_u^i$ (``upper index'').
@@
 For a rule, the splitting
-in signal and background is useful for the treatment of systematics, as we will se later. Each rule will lead to a $\Delta \chi^2$-value,
+into signal and background is useful for the treatment of systematics, as we will se later. Each rule will lead to a $\Delta \chi^2$-value,
 which means that all $\Delta \chi^2$'s of the different rules will be added
@@
 systematics specifies a dependence).
-Thus, it is convenient to combine the above defined channels for different
+Thus, it is convenient to combine the previously defined channels for different
 oscillation patterns and interaction types into one logical construction,
@@
 For each rule, the signal event rate $s_i$ in the $i$th bin can be composed 
-out of one or more channels by 
+out of one or more channels according to 
 \begin{equation}
@@
 Similarly, the background event rate $b_i$ in the $i$th bin can be composed
-out of one or more channels
+out of one or more channels:
 \begin{equation}
@@
 The nuisance parameters describe the dependence 
-of the event rates on the various sources of systematical errors, 
-such as an error on the total 
+of the event rates on the various sources of systematical errors. 
+For example, an error on the total 
 normalization is included by multiplying the expected number of events in 
-each bin by a factor $(1+\zeta_1)$. The variation of $\zeta_1$ is in the fit 
+each bin by a factor $(1+\zeta_1)$. The variation of $\zeta_1$ in the fit is
 constrained by adding a penalty $p_1$ to the $\chi^2$-function. In case 
@@
 where $\sigma_{\zeta_i}$ is the standard 
-deviation of the corresponding nuisance parameter. We further on also refer
-to the  standard deviation as
-the ``error'', it corresponds to the actual systematical uncertainty.
+deviation of the corresponding nuisance parameter. In the following,
+we will refer to the  standard deviation as
+the ``error'', since it corresponds to the actual systematical uncertainty.
 Note that the central values of all penalties are zero in \GLOBES\ 3.0 and higher. 
@@
 \ie , pair of signal and background spectra, with a separate set of 
-nuisance parameters $\zeta_i^\alpha$. Thus, $\chi^2_\mathrm{pull}$
+nuisance parameters $\zeta_i^r$. Thus, $\chi^2_\mathrm{pull}$
 is the sum of all individual  $(\chi^2_\mathrm{pull})^r$'s.
-By the minimization, the dependence on the $k$ nuisance parameters has been eliminated  from $\chi^2_\mathrm{pull}$.
-
-Now we can introduce the different systematical errors. 
+By the minimization, the dependence on the $k$ nuisance parameters is eliminated  from $\chi^2_\mathrm{pull}$.
+
+Now, we can introduce the different systematical errors. 
 The two most important and
 most easily parameterized systematical errors are the normalization 
-and an energy calibration errors. These errors are assumed to be independent between the signal events and the background events, which means that
+and energy calibration errors. These errors are assumed to be independent between the signal events and the background events, which means that
 this systematics treatment defines the grouping into signal or background.
@@
 to keep in mind that this definition of the energy calibration error 
-makes sense only for constant bin width, so the corresponding $\chi^2$
+makes sense only for constant bin widths, so the corresponding $\chi^2$
 functions should not be used in conjunction with the {\tt \$binwidth} directive.
@@
 the bin width, since also the bin width is altered by the replacement of
-the energies. Furthermore, special care has to be payed to the limits
+the energies. Furthermore, special attention has to be given to the limits
  $k<1$ or $k+1>N_\mathrm{bins}$, since there $s_k$ or $s_{k+1}$ may not have been calculated. By default, it is assumed that $s_k$ is
 zero in those cases. However, if the event rates are still large at
-the limits, errors will be introduced leading to a wrong
+the limits, this will introduce errors, leading to a wrong
 estimate of the impact of the calibration error. In this case,
 one should truncate the analysis range by a few bins at the boundaries
-and therefore ensure in this way that only those $s_i$ are used whose index
-$k$ is within the range $0,\ldots, N_\mathrm{bins}-1$ (\cf, \figu{energies}). 
-Thus, it is possible to constrain the analysis energy range 
+and thus ensure that only those $s_i$, whose index
+$k$ is within the range $0,\ldots, N_\mathrm{bins}-1$ (\cf, \figu{energies}), are used.
+Therefore, it is possible to constrain the analysis energy range 
 with each rule to an energy window\index{norm}{Energy!window}:
@@
 \end{quote}
-The default energy window is given between the minimal and maximal reconstructed energy. To be on the save side, reduce analysis window
-compared to the bin range on each side by about three times the
-energy calibration error.
+The default energy window is given by the minimal and maximal reconstructed
+energies {\tt \$emin} and {\tt \$emax}. To be on the safe side, one should
+reduce the analysis window compared to the bin range on each side by about
+three times the energy calibration error.
 
@@
 be marginalized over, and therefore the corresponding error needs to 
-have a non-zero value. If the cases with ``total rates'' in the remarks, 
+have a non-zero value. In the cases with ``total rates'' in the remarks, 
 the summation over the bins is performed \emph{before} computing 
@@
 background  normalization ($c$) and
-background tilt/calibration ($d$). Their default (central) values are zero.
+background tilt/calibration ($d$). Their default (central) values are
+zero.\footnote{The old parameter {\tt @backgroundcenter} should not be used anymore.
+A background normalization center of 1.0 will be interpreted as zero central value.}
 % \begin{equation}
@@
 values of tilt/calibration are always regarded
-as a pairs, \ie, they are given in the form {\tt normalization : tilt}.
-For example, we have\footnote{The old parameter {\tt @backgroundcenter} should not be used anymore.
-A background normalization center of 1.0 will be
-interpreted as zero central value.}
+as pairs, \ie, they are given in the form {\tt normalization :\ tilt}.
+For example, we have
 \index{aedl}{rule@{\tt rule}!signal error@{\tt "@signalerror}}
@@
 The different  possibilities are shown in
-\Tab~\ref{tab:error_dim}. Since the dual systematics modes define the experiment
-behavior for systematics on and off, it is useful to have a matching pair of 
+\Tab~\ref{tab:error_dim}. Since the dual systematics modes define the behavior
+of the experiment for systematics on and off, it is useful to have a matching pair of 
 systematics functions for each rule 
@@
 \end{quote}
-Note that here the systematics on and systematics off models can be used 
-as fully functional different systematics modes with different systematical
+Note that here, the systematics on and systematics off models can be used 
+as two different, fully functional systematics modes with different systematical
 errors. The interpretation of the systematical errors in {\tt @sys\_on\_errors} and
@@
 confusing different systematics routines. In addition, it is possible to define
-a rule with passive systematics using {\tt glbChiZero}. Such passive systematics
-is not directly used or called, but the corresponding rate vectors will be calculated
-and provided for indirect access. For example, a reactor experiment with correlated
+a rule with passive systematics using {\tt glbChiZero}. In this case, the
+contribution to $\chi^2$ from this rule will always be set to zero, but
+the corresponding rate vectors will be calculated and provided for indirect access
+by other systematics functions. For example, a reactor experiment with correlated
 systematics between near and far detectors may define the user-defined systematics
@@
 {\tt chiReactor} in both the near and far detectors would result in a double call
-of the $\chi^2$ function, \ie, the program will be a factor two slower and the
-resulting $\chi^2$ will be a factor two too large.
+of the $\chi^2$ function, \ie, the resulting $\chi^2$ would be too large by a factor two
+and the program would be slower by a factor of two.
 Note that mixed declarations of {\tt @sys\_on\_errors}, {\tt @sys\_off\_errors}, {\tt @signalerror}, and
 {\tt @backgrounderror} are possible. In this case, {\tt @sys\_on\_errors} and {\tt @sys\_off\_errors}
-have priority. If for one or both of these no systematical errors are defined, {\tt @signalerror} or
+have priority. If one or both of these are not present, {\tt @signalerror} or
 {\tt @backgrounderror} will be used.
@@
 \tb @signal = 0.5 @ \#channel\_1\\
-\tb @background = 0.001 @ \#channel\_2 :  0.005 @ \#channel\_3\\
-\tb @signalerror =       0.001  :       0.01\\
-\tb @backgrounderror =   0.001 :       0.01\\
+\tb @background = 0.001 @ \#channel\_2 :\ 0.005 @ \#channel\_3\\
+\tb @signalerror =       0.001  :\ 0.01\\
+\tb @backgrounderror =   0.001 :\ 0.01\\
 \tb @sys\_on\_function = "chiSpectrumTilt"  \\
 \tb @sys\_off\_function = "chiNoSysSpectrum"  \\
-\tb @energy\_window = 4.0 : 50.0
+\tb @energy\_window = 4.0 :\ 50.0 \\
 >}

### 2007-01-15 Documentation/part1.tex
*** empty log message ***

@@
 
-Gold + Silver & {\tt NF\_Goldsilver.glb} & as {\tt NFvar.glb} 
+Gold + Silver & {\tt NF\_GoldSilver.glb} & as {\tt NFvar.glb} 
 plus 5~kt ECC detector for Silver Channel

### 2007-01-15 Documentation/appendix.tex
*** empty log message ***

@@
 The \NOVA\ experiment uses a neutrino beam from the Fermilab NuMI beamline. The target power is 
-1.2~MV which results in $10^{21}$~pot~$\mathrm{yr^{-1}}$ and 3~years $\nu$-running and 3~years $\bar{\nu}$-running 
+1.12~MV which results in $10^{21}$~pot~$\mathrm{yr^{-1}}$ and 3~years $\nu$-running and 3~years $\bar{\nu}$-running 
 is assumed. The fiducial mass of the Totally Liquid Scintillator Detector (TASD) is taken to be 
@@
 
-\subsection*{DoubleCHOOZ -- {\tt Dchooz-near.glb} and {\tt Dchooz-far.glb}}
-
-The files {\tt Dchooz-near.glb} and {\tt Dchooz-far.glb} allow to simulate the \DC\ reactor experiment in France. 
+\subsection*{DoubleCHOOZ -- {\tt D-Chooz\_near.glb} and {\tt D-Chooz\_far.glb}}
+
+The files {\tt D-Chooz\_near.glb} and {\tt D-Chooz\_far.glb} allow to simulate the \DC\ reactor experiment in France. 
 The basic version of these files were used within \cite{Huber:2006vr} which should be cited if the files 
-{\tt Dchooz-near.glb} and {\tt Dchooz-far.glb} are
-used for a scientific publication or a talk. For calculations that involve {\tt Dchooz-near.glb} and/or {\tt Dchooz-far.glb}, the following
+{\tt D-Chooz\_near.glb} and {\tt D-Chooz\_far.glb} are
+used for a scientific publication or a talk. For calculations that involve {\tt D-Chooz\_near.glb} and/or {\tt D-Chooz\_far.glb}, the following
 additional files are required:
@@
 427 \,t \, GW\, yr}$.
-Here, the total running time of 5 years is assumed within {\tt Dchooz-near.glb} and {\tt Dchooz-far.glb}, so near
+Here, the total running time of 5 years is assumed within {\tt D-Chooz\_near.glb} and {\tt D-Chooz\_far.glb}, so near
 and far detector are assumed to start the mode of operation simultaneously. For the simulation of \DC\ and
-considering a delayed start of data taking at the near detector the file {\tt Dchooz-near.glb} has to be
+considering a delayed start of data taking at the near detector the file {\tt D-Chooz\_near.glb} has to be
 modified. The cancellation of systematical uncertainties is considered by the manual definition of a $\chi^2$
@@
 The energy resolution is
-$\mathrm{\sigma_e=5\%\cdot\sqrt{E_{vis}}}$ and the choice for {\tt sigma\_function} is {\tt \#inverse\_beta}. The following rules are defined within {\tt Dchooz-near.glb} and {\tt Dchooz-far.glb}: 
+$\mathrm{\sigma_e=5\%\cdot\sqrt{E_{vis}}}$ and the choice for {\tt sigma\_function} is {\tt \#inverse\_beta}.
+The following rules are defined within {\tt D-Chooz\_near.glb} and {\tt D-Chooz\_far.glb}:
 \begin{center}
@@
 
-\subsection*{Variable neutrino factory with Silver Channel -- {\tt NFvar\_GoldSilver.glb}}
+\subsection*{Variable neutrino factory with Silver Channel -- {\tt NF\_GoldSilver.glb}}
 
 A variable neutrino factory scenario that includes the golden and silver appearance channels can be 
-simulated with the file {\tt NFvar\_GoldSilver.glb}. The basic version was
+simulated with the file {\tt NF\_GoldSilver.glb}. The basic version was
 used within \cite{Huber:2006wb} and the golden channel follows the neutrino factory scenarios from \cite{Huber:2002mx}
@@
 resolution of the silver channel is set to $\mathrm{\sigma_e=20\%\cdot E}$. The following additional rule
-compared to {\tt NFvar.glb} is introduced to {\tt NFvar\_GoldSilver.glb}:
+compared to {\tt NFvar.glb} is introduced to {\tt NF\_GoldSilver.glb}:
 \begin{center}
@@
 
-\subsection*{High Resolution/Low Threshold neutrino factory -- \\{\tt NFvar\_hR\_lT.glb}}
-
-A variable neutrino factory hybrid detector scenario can be simulated with the file {\tt NFvar\_hR\_lT.glb}. The basic version was
+\subsection*{High Resolution/Low Threshold neutrino factory -- \\{\tt NF\_hR\_lT.glb}}
+
+A variable neutrino factory hybrid detector scenario can be simulated with the file {\tt NF\_hR\_lT.glb}. The basic version was
 used within \cite{Huber:2006wb} and follows the neutrino factory scenarios from \cite{Huber:2002mx}. These
-references should be cited if the file {\tt NFvar\_hR\_lT.glb} is used for a scientific publication or a talk.
-For calculations that involve {\tt NFvar\_hR\_lT.glb}, the same additional files as for {\tt NFvar.glb} are
-required and the {\sf AEDL}-Variables {\tt emax} and {\tt BASELINE} have to be set. {\tt NFvar\_hR\_lT.glb} 
+references should be cited if the file {\tt NF\_hR\_lT.glb} is used for a scientific publication or a talk.
+For calculations that involve {\tt NF\_hR\_lT.glb}, the same additional files as for {\tt NFvar.glb} are
+required and the {\sf AEDL}-Variables {\tt emax} and {\tt BASELINE} have to be set. {\tt NF\_hR\_lT.glb} 
 implements a lower threshold ($\mathrm{\sim\,1\,GeV}$) at a higher energy resolution
@@
 the effects from Fermi Motion. The background rejection is introduced energy dependent as 
-$10^{-3}/\mathrm{E^2}$ and matches {\tt NFvar.glb} at higher energies. The following rules are defined within {\tt NFvar\_hR\_lT.glb}: 
+$10^{-3}/\mathrm{E^2}$ and matches {\tt NFvar.glb} at higher energies. The following rules are defined within {\tt NF\_hR\_lT.glb}: 
 \begin{center}
@@
 
-IS THAT CORRECT? MR: WHAT HAPPENS TO OLD BETA BEAM FILES [WW]?
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2007-01-15 Documentation/appendix.tex
*** empty log message ***

@@
 power is 0.77~MW and 2~years $\nu$-running and 6~years $\bar{\nu}$-running is assumed. The fiducial mass of the
-Super-Kamiokande Water Cerenkov detector is taken to be $\mathrm{m_{det} = 2.5 \,kt}$ at a baseline of
+Super-Kamiokande Water Cerenkov detector is taken to be $\mathrm{m_{det} = 22.5 \,kt}$ at a baseline of
 L~=~295~km. The appearance measurement involves the total rates data from all CC events and the spectral data
@@
 
-\subsection*{Variable neutrino factory with Silver Channel -- {\tt NF\_GoldSilver.glb}}
+\subsection*{Variable neutrino factory with Silver Channel -- \\{\tt NF\_GoldSilver.glb}}
 

### 2007-01-15 Documentation/part1.tex
JK - Minor corrections chapter 11, modified documentation of glbGetChiFunction

@@
 \GLBNS{GetChiFunction}
-{\tt int glbGetChiFunction(int exp, int rule, int on\_off, char *sys\_id)}
+{\tt int glbGetChiFunction(int exp, int rule, int on\_off, char *sys\_id, size\_t max\_len)}
 returns the name of the systematics $\chi^2$ function of a given experiment {\tt exp} and rule {\tt rule} for
-systematics on or off as given by {\tt on\_off}. The user must 
-ensure that the return string {\tt sys\_id} is large enough to hold the name of the function.
+systematics on or off as given by {\tt on\_off}. The name is copied to the string
+{\tt sys\_id}, the maximum length of which is specified by {\tt max\_len}. If
+{\tt max\_len} is too small, or if any other error occurs, the return value is $< 0$.
 \end{function}

### 2007-01-15 Documentation/part2.tex
JK - Minor corrections chapter 11, modified documentation of glbGetChiFunction

@@
 As an argument, {\tt globes} takes a {\tt .glb}-file. While parsing it,
-it prints any warnings and errors which have occured during reading the file.
+it prints any warnings and errors which occur while reading the file.
 Then it uses the experiment description in the file to compute the event rates at a certain point in parameter space.
@@
 \end{quote}
-Note that in the case of additional non-standard parameters, these will just to 
+Note that in the case of additional non-standard parameters, these will have to 
 be included at the end of these lists.
@@
 The effect of {\tt -N} is the same as to use {\tt NOSC\_} in all oscillation channels.
-Thi feature is useful if one wants to normalize an expriment flux if
-the number of un-oscillated events is given.
+This feature is useful if one wants to normalize the flux in an experimental to
+a given number of un-oscillated events.
 
@@
 at this level it is often necessary to review the intermediate steps in the event rate calculation.
-The {\tt globes} utility offers many possibilities to do this based on the rate access functions as described in \Sec~\ref{sec:event_rates}.
+The {\tt globes} utility offers many possibilities to do this based on the rate access functions described in \Sec~\ref{sec:event_rates}.
 
@@
 This option takes as an optional argument the channel number 
-(starting at zero). If no argument is given all channels are displayed.
+(starting at zero). If no argument is given, all channels are displayed.
 By default, the sum of the event rates in each channel is shown. Each 
@@
 produces the raw event rate corresponding to the convolution of flux, 
-probability, and cross section, which is neglecting all detector effects.
+probability, and cross section, neglecting all detector effects.
 
@@
 (starting at zero). If no argument is given, all rules will be displayed.
-By default, the sum of the event rates in each rule is shown, as well
-as for each component within the rule. Each 
-rule is preceeded by a line with the same rule name as in the file.
-
-It is also possible for rules to switch off one detector effect after 
+By default, the signal and background rates in each rule are shown
+separately, together with their decomposition into the different
+contributing channels. Each rule is preceeded by a line with the same
+rule name as in the file.
+
+Also for the rules, it is possible to switch off one detector effect after 
 the other -- with the limitation that rules only make sense
@@
 output, which can be directly visualized by {\tt MultipleListPlot}.
-The option {\tt -u} uses the same principal formatting as {\tt -m}, but
+The option {\tt -u} uses in principle the same formatting as {\tt -m}, but
 it allows to specify the left, middle, and right delimiters in constructing

### 2007-01-15 Documentation/appendix.tex
*** empty log message ***
JK - Minor modifications in the appendices (except appendix B)

@@
 to parse its command line options. However, on platforms where \verb+argp+
-is lacking \GLOBES\ has replacement code, thus it should also work
+is lacking, \GLOBES\ has replacement code, thus it should also work
 there. \GLOBES\ is, however, using the C99 standard in order to handle
@@
 
-GSL is also available as rpm+s from the various distributors of
+GSL is also available as rpm's from the various distributors of
 GNU/Linux, see their web sites for downloads. Chances are that gcc and
 GSL are already part of your installation.  For building \GLOBES\ from
-source, however, not only working libraries for above packages are
-needed but also the headers, especially for GSL. Depending on your
-installation, eg. on RedHat/Fedora, of GSL this may require to
+source, however, not only working libraries for the above packages are
+needed, but also the headers, especially for GSL. For some installations
+of GSL, eg. on RedHat/Fedora, this may require to
 additionally install a rpm-package named \verb+gsl-devel+. If GSL has been
-installed from the tar-ball as provided by gnu.org no problems should
+installed from the tar-ball as provided by gnu.org, no problems should
 occur. Furthermore you need a working \verb+make+ to build and install
@@
 \GLOBES\ you will need an ANSI C-compiler.  After unpacking the
-distribution the Makefiles can be prepared using the configure
+distribution, the Makefiles can be prepared using the configure
 command,\\
@@
 Under certain circumstances it may be useful to use a static version of
-libglobes or any of the binaries, e.g. when running on a cluster.
+libglobes or any of the binaries, e.g.\ when running on a cluster.
 
 The \verb+configure+ script accepts the option \verb+--disable-shared+, in
-which case only static objects are built, i.e. only a static
+which case only static objects are built, i.e.\ only a static
 version of libglobes. In case your system does not support shared
@@
 linking. If you want to build static binaries, use
-LDFLAGS=+-all-static+ for building them.
+\verb+LDFLAGS='-all-static'+ for building them.
 
 Sometimes it is convenient, eg. for debugging purposes, to have a
-statically linked version of a program using \GLOBES\, which is easiest
+statically linked version of a program using \GLOBES, which is easiest
 achieved by just linking with \verb+libglobes.a+. If you need a completely
@@
 All these options rely on a working \verb+gcc+ installation. It seems that
-\verb+gcc 3.x+ is broken in subtle way which makes it necessary to add a
+\verb+gcc 3.x+ is broken in a subtle way which makes it necessary to add a
 symbolic link in the gcc library directory. The diagnostic for this
@@
 It may be necessary to prefix the call of \verb+globes-config+ with
-the path to it, in case that location is not in \verb+$PATH+.
+the path to it, in case that this location is not in \verb+$PATH+.
 \index{norm}{Condor|)}
@@
 
-Sometimes the GNU scientific library is not available or is installed
+Sometimes, the GNU scientific library is not available or is installed
 in a non-standard location. This situation can arise in an
-installation without root privileges. In this case one can specify
+installation without root privileges. In this case, one can specify
 \verb+--with-gsl-prefix=path_to_gsl+ as option to the \verb+configure+ script.
 If one wants to use a shared version of \verb+libgsl+ then one has to make
-sure that the linker find the library at run-time. This can be
+sure that the linker can find the library at run-time. This can be
 achieved by setting the environment variable \verb+LD_LIBRARY_PATH+
@@
 The standard rpm-based installation of GSL does not provide any header
-files for GSL, which are however needed to compile \GLOBES\. You have to
+files for GSL, which are however needed to compile \GLOBES. You have to
 install an additional rpm-package called \verb+gsl-devel+. Alternatively
@@
 
-   By default, \verb+make install+ will install the package+s files in
+   By default, \verb+make install+ will install the package's files in
 \verb+/usr/local/bin+, \verb+/usr/local/man+, etc.  You can specify an
@@
 give \verb+configure+ the option \verb+--exec-prefix=PATH+, the package will use
-PATH as the prefix for installing programs and libraries.
+\verb+PATH+ as the prefix for installing programs and libraries.
 Documentation and other data files will still use the regular prefix.
@@
    Some packages pay attention to \verb+--enable-FEATURE+ options to
-\verb+configure+, where FEATURE indicates an optional part of the package.
-They may also pay attention to \verb+--with-PACKAGE+ options, where PACKAGE
+\verb+configure+, where \verb+FEATURE+ indicates an optional part of the package.
+They may also pay attention to \verb+--with-PACKAGE+ options, where \verb+PACKAGE+
 is something like \verb+gnu-as+ or \verb+x+ (for the X Window System).  The
@@
 
-This feature is experimental and your mileage my vary!
+This feature is experimental and your mileage may vary!
 
@@
 
-The trick here, is that we use SWIG (\verb+www.swig.org+) to generate
+The trick here is, that we use SWIG (\verb+www.swig.org+) to generate
 a wrapper file for \GLOBES. The wrapper file is part of the \GLOBES\
@@
 feature \verb+--enable-rpm-rules+ which should produce all the necessary
-makefile rules for RPM building. To actually build RPMs requires that
-your system is properly setup for that. How you can do that you can
-learn at \verb+http://www.rpm.org+. You then can use \verb+make rpm+, most likely
+Makefile rules for RPM building. To actually build RPMs requires that
+your system is properly setup for that. You can learn how to do that
+at \verb+http://www.rpm.org+. You then can use \verb+make rpm+, most likely
 you will need to be root to do that (\verb+sudo+ won't work!).
@@
 a message saying it can not guess the host type, give it the
-\verb+--host=TYPE+ option.  TYPE can either be a short name for the system
+\verb+--host=TYPE+ option. \verb+TYPE+ can either be a short name for the system
 type, such as \verb+sun4+, or a canonical name with three fields\\
@@
 \multicolumn{3}{|l|}{Disappearance -- $\mu^+$-stored} & $\sigma_\mathrm{norm}$ & $\sigma_\mathrm{cal}$ \\ \hline
-Signal & $0.35 \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_\mu)_\mathrm{CC}$ & & 0.001 & $10^{-4}$ \\
+Signal & $0.35 \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_\mu)_\mathrm{CC}$ & & 0.025 & $10^{-4}$ \\
  & & & & \\
@@
 \multicolumn{3}{|l|}{Appearance -- $\mu^+$-stored} & & \\ \hline
-Signal &  $0.45 \, \otimes \, (\nu_e \rightarrow \nu_\mu)_\mathrm{CC}$ & & 0.001 & $10^{-4}$ \\
+Signal &  $0.45 \, \otimes \, (\nu_e \rightarrow \nu_\mu)_\mathrm{CC}$ & & 0.025 & $10^{-4}$ \\
  & & & & \\
@@
 \multicolumn{3}{|l|}{Disappearance -- $\mu^-$-stored} & & \\ \hline
-Signal &  $0.45 \, \otimes \, (\nu_\mu \rightarrow \nu_\mu)_\mathrm{CC}$ & & 0.001& $10^{-4}$\\
+Signal &  $0.45 \, \otimes \, (\nu_\mu \rightarrow \nu_\mu)_\mathrm{CC}$ & & 0.025& $10^{-4}$\\
  & & & & \\
@@
 \multicolumn{3}{|l|}{Appearance -- $\mu^-$-stored} & & \\ \hline
-Signal & $0.35 \, \otimes \, (\bar{\nu}_e \rightarrow \bar{\nu}_\mu)_\mathrm{CC}$  & & 0.001& $10^{-4}$\\
+Signal & $0.35 \, \otimes \, (\bar{\nu}_e \rightarrow \bar{\nu}_\mu)_\mathrm{CC}$  & & 0.025& $10^{-4}$\\
  & & & & \\
@@
 the input flux files for use in \AEDL\ experiment descriptions. Source of
-the confusion is an undocumented factor 5.2 in \GLOBES\ versions older
-than 3.0 the fluxes are multiplied with (see below). In Version 3.0 and 
+the confusion is an undocumented factor 5.2 with which the fluxes
+are multiplied in \GLOBES\ versions older than 3.0 (see below). In Version 3.0 and 
 higher, the alternative flux environment {\tt nuflux} is provided, 
@@
 This replacement is not necessary for neutrino factory built-in fluxes,
-and built-in beta beams fluxes where not supported by earlier versions
+and built-in beta beam fluxes were not supported by earlier versions
 of \GLOBES .

### 2007-01-16 Documentation/appendix.tex
JK - Minor corrections in appendix B

@@
 
-\begin{appendix}
-
+%\begin{appendix}
+\appendix
 \chapter{\GLOBES\ installation}
@@
 
-\section*{Prerequisites for installation of \GLOBES}
+\section{Prerequisites for installation of \GLOBES}
 
@@
 
-\section*{Installation Instructions}
+\section{Installation Instructions}
 
@@
 neutrino factory setups. A complete list
-of all pre-defined experiment files an be found in \tabl{experiments}. More detailed descriptions of the
+of all pre-defined experiment files can be found in \tabl{experiments}. More detailed descriptions of the
 corresponding files, the assumptions, requirements, and references are given in the following.
@@
 The \TtoK\ experiment can be simulated with the file {\tt T2K.glb}. This file tries to approximate as closely as
-possible the LOI \cite{Itow:2001ee} and the basic version was used within \cite{Huber:2002mx}. These references
+possible the LOI \cite{Itow:2001ee}, and the basic version was used within \cite{Huber:2002mx}. These references
 should be cited if the file {\tt T2K.glb} is used for a scientific publication or a talk. For calculations that
@@
 $\mathrm{\mathcal{L} = 400 \,t \, GW\, yr}$, \eg\ a 20~t detector, a reactor with a thermal power of 4~GW
-and a running period of 5 years. As detector technology a liquid scintillator 
+and a running period of 5 years. As detector technology, a liquid scintillator 
 detector is assumed, a far detector at a baseline of L~=~1.7~km and a near detector which is
@@
 The files {\tt D-Chooz\_near.glb} and {\tt D-Chooz\_far.glb} allow to simulate the \DC\ reactor experiment in France. 
-The basic version of these files were used within \cite{Huber:2006vr} which should be cited if the files 
+The basic versions of these files were used within \cite{Huber:2006vr}, which should be cited if the files 
 {\tt D-Chooz\_near.glb} and {\tt D-Chooz\_far.glb} are
@@
 \ei
-The \DC\ experiment is located at the Chooz reactor complex and the two reactor cores serve as $\bar{\nu}_e$
+The \DC\ experiment is located at the Chooz reactor complex, and the two reactor cores serve as $\bar{\nu}_e$
 neutrino source, so the thermal power is $\mathrm{2\cdot4.2\,GW}$. Two identical liquid scintillator detectors
@@
 NFstandard.glb}. Only the treatment of the disappearance channels is different. Whereas the CID threshold is
-also applied in the disappearance channels of {\tt NFstandard.glb} this is not the case for the disapperance
-channel definition within {\tt NFvar.glb} where no CID is applied and a threshold similar to the threshold of
+also applied in the disappearance channels of {\tt NFstandard.glb}, this is not the case for the disapperance
+channel definition within {\tt NFvar.glb}, where no CID is applied, and a threshold similar to the threshold of
 the MINOS experiment \cite{Ables:1995wq} is applied. The following rules are defined within
@@
 in the range $\mathrm{10\,GeV\lesssim E_\mu\lesssim 80\, GeV}$. The baseline is set by the {\sf
-AEDL}-Variable {\tt BASELINE}. For the silver channel an additional ECC detector with a fiducial mass
+AEDL}-Variable {\tt BASELINE}. For the silver channel, an additional ECC detector with a fiducial mass
 $\mathrm{m_{ECC} = 5\,kt}$ is assumed to be located at the same baseline as the MID detector. The energy
 resolution of the silver channel is set to $\mathrm{\sigma_e=20\%\cdot E}$. The following additional rule
-compared to {\tt NFvar.glb} is introduced to {\tt NF\_GoldSilver.glb}:
+compared to {\tt NFvar.glb} is introduced in {\tt NF\_GoldSilver.glb}:
 \begin{center}
@@
 implements a lower threshold ($\mathrm{\sim\,1\,GeV}$) at a higher energy resolution
-$\mathrm{\sigma_e=15\%\cdot E + 0.085\, MeV}$ where the constant term represents 
-the effects from Fermi Motion. The background rejection is introduced energy dependent as 
-$10^{-3}/\mathrm{E^2}$ and matches {\tt NFvar.glb} at higher energies. The following rules are defined within {\tt NF\_hR\_lT.glb}: 
+$\mathrm{\sigma_e=15\%\cdot E + 0.085\, MeV}$, where the constant term represents 
+the effects from Fermi Motion. The background rejection is energy dependent according to 
+$10^{-3}/\mathrm{E^2}$, and matches {\tt NFvar.glb} at higher energies. The following rules are defined within {\tt NF\_hR\_lT.glb}: 
 \begin{center}
@@
 
-\end{appendix}
+%\end{appendix}
 

### 2007-01-16 Documentation/Paper2/paper2.tex
JK - Minor typos in paper2 corrected

@@
 \eg , \Ref~\cite{Barger:2003qi}. Future facilities, using accelerator-based
- neutrino beams or a nuclear reactors as neutrino sources, are proposed
+ neutrino beams or nuclear reactors as neutrino sources, are proposed
 for precision measurements of the neutrino oscillation parameters. In the
@@
 affect the performance.
-Thus optimization strategies are required which maximally exploit 
+Thus, optimization strategies are required which maximally exploit 
 complementarity between experiments. The \GLOBES\ software package~\cite{Huber:2004ka}
@@
 which can be used for systematical parameter space scans. In addition, it 
-simplified the direct comparison of experiments.
+simplifies the direct comparison of experiments.
 
@@
 atmosphere, can not be described. In addition, sources with a physically 
-significant time dependencies can not be studied, such as  supernov\ae. 
+significant time dependence can not be studied, such as  supernov\ae. 
 However, in \GLOBESN\ and higher, new flexibility is introduced by the
@@
 Hamiltonian in layers of constant matter density using the standard three flavor
-scenario of neutrino oscillations. In flavor base, we have
+scenario of neutrino oscillations. In the flavor basis, we have
 \begin{equation}
@@
 \end{equation}
-with $U$ the mixing matrix (described by four parameters $\theta_{12}$, $\theta_{13}$, $\theta_{23}$, and $\deltacp$), and $n_e$ the constant electron density
-in that matter density layer. The electron density is related to the matter density $\rho$ by 
-$n_e \simeq 0.5 \, \rho/m_N$ with $m_N$ being the nucleon mass. The sign in the second term depends on if one
+with $U$ being the mixing matrix (described by four parameters $\theta_{12}$, $\theta_{13}$, $\theta_{23}$, and $\deltacp$), and $n_e$ the constant electron density
+in the respective matter density layer. The electron density is related to the matter density $\rho$ by 
+$n_e \simeq 0.5 \, \rho/m_N$ with $m_N$ being the nucleon mass. The sign in the second term depends on whether one
 uses neutrinos or antineutrinos. For $k$ matter density layers with densities
@@
 re-compilation of the \GLOBES\ software. This is implemented
-by using pointers to the functions of the standard probability engine, which can be changed if
-a different probability engine is registered. For example, a user may just copy the standard probability
+by using pointers to the functions of the standard probability engine, which can be changed by
+registering a different probability engine. For example, a user may just copy the standard probability
 engine of \GLOBES , modify it, and register it. An important prerequisite for the ability to handle
@@
 
-\GLOBES\ supports four types of systematical errors as a standard: Signal and background
+\GLOBES\ supports four types of systematical errors by default: Signal and background
 normalization errors, as well as signal and background tilt or energy calibration errors.
@@
 $\sigma_{\zeta_j}$ corresponds to the actual systematical error.
-Core part of \equ{chipull} is the definition of $\chi^2$. For example, for a background-free measurement and
+The core part of \equ{chipull} is the definition of $\chi^2$. For example, for a background-free measurement and
 a signal normalization error only, it reads
@@
 $\boldsymbol{\lambda}$ is projected onto the subspace $\boldsymbol{\lambda_1}$.
-If one wants to include external input which only depends on the oscillation
-parameters and not on systematics, one can do this at this level. \GLOBES\ provides
+If one wants to include external input, which only depends on the oscillation
+parameters and not on systematics, one can do so at this level. \GLOBES\ provides
 the pre-defined possibility to include external Gaussian constraints on the 
@@
 Pre-defined experiment prototypes coming with \GLOBESN : The columns represent the \AEDL\ filename, 
-a short description, and the references in which they are originally used and 
-discussed (except from minor modifications, such as a different implementation of 
+a short description, and the references in which the files are originally used and 
+discussed (except for minor modifications, such as a different implementation of 
 the energy threshold function). 
@@
 \AEDL\ (Abstract Experiment Definition Language) is a special syntax developed for the description of
-experiments in \GLOBES\ represented by simple text files. Core part of \AEDL\ are the following constructions
+experiments in \GLOBES, represented by simple text files. Core part of \AEDL\ are the following constructions
 (for details, see \Ref~\cite{Manual}):
@@
 appearance and disappearance channels, neutrino and antineutrino operations modes, \etc.
-An experiment has a common detector and baseline, but may use more different fluxes (such as for neutrinos and antineutrinos).
+An experiment has a common detector and baseline, but may use several different fluxes (such as for neutrinos and antineutrinos).
 Therefore,
@@
 the user to break up the strict coupling between systematics and rules. For example, systematics can be
-correlated between different rules or experiments, such as for an reactor experiment with identical
+correlated between different rules or experiments, such as for a reactor experiment with identical
 near and far detectors. In addition, one can introduce user-defined systematical errors.
@@
 %
-Further changes in \AEDL\ are the requirement to specify the minimum \GLOBES\ version the \AEDL\ is running with,
+Further changes in \AEDL\ are the requirement to specify the minimum \GLOBES\ version the \AEDL\ file can be used with,
 and the extensions by lists as \AEDL\ variables and by an interpolation feature.
@@
 Firstly, the \GLOBES\ minimizers distinguish between minimization over
-oscillation parameters and minimization over systematics parameters.
+oscillation parameters and minimization over systematics (nuisance) parameters.
 This is necessary because each step in the oscillation parameter space
@@
 The numerical method that does the minimization is the Powell
-algorithm HERE ORIGINAL POWELL REFERENCE [WW].
+algorithm~\cite{Fletcher:1963:RCD}.
 
@@
 algorithm discussed in \Ref~\cite{Kopp:2006wp}, but for the rare cases where this may become
-inaccurate, it falls back to the QL (WRITE OUT!)
-method from~\cite{Press:NumRecip} REPLACE BY ORIGINAL REFERENCE!.
+inaccurate, it falls back to the QL method~\cite{Fra61,Stoer:NumAnalysis}.
 
@@
 of \GLOBESN\ to carry more than six oscillation parameters. New experiment
-prototypes make extensive use these new features, as well as they are illustrated by
+prototypes make extensive use of these new features, as well as they are illustrated by
 examples in the manual. One prominent example is the Double Chooz experiment,
@@
 will make an easy adjustment of the software possible for both experimentalists and theorists.
-Finally, note that the past has shown that it is often more than one possibility 
+Finally, note that the past has shown that there is often more than one possibility 
 how a specific problem can be implemented in \GLOBES .
@@
 %
-The development of \GLOBES\ is currently being supported by the Studienstiftung des Deutschen Volkes [JK] and the
+The development of \GLOBES\ is currently being supported by the
 Emmy Noether-Program of Deutsche Forschungsgemeinschaft [WW].

### 2007-01-17 Documentation/appendix.tex
*** empty log message ***

@@
 
-A $\gamma=350$ medium gamma $\beta$-beam green-field scenario involving a megaton Water Cerenkov detector can be simulated with the file {\tt
+A $\gamma=350$ medium gamma $\beta$-beam scenario involving a megaton Water Cerenkov detector can be simulated with the file {\tt
 BB\_350.glb}. This file tries to approximate as closely as possible the scenario {\sf Setup III} from
@@
 The acceleration factor is $\gamma=350$ for both types of isotopes and $2.2\cdot10^{18}$ $^{18}$Ne decays per
-year and $5.8\cdot10^{18}$ $^{6}$He decays per year are assumed. The baseline is L~=~730~km, the
+year and $5.8\cdot10^{18}$ $^{6}$He decays per year are assumed. The ion acceleration would require either a refurbished SPS (with
+superconducting magnets) or a more powerful accelerator, such as the
+Tevatron or LHC. The baseline is L~=~730~km, the
 fiducial mass of the detector is $\mathrm{m_{det} = 500 \,kt}$ and 4~years $\nu$-running and 4~years
@@
 \bi
-\item {\tt ThresholdEnergies.dat} (list of energies)
-\item {\tt CIDthreshold.dat} (list for CID threshold)
-\item {\tt MINOSthreshold.dat} (list for MINOS threshold)
 \item {\tt XCC.dat} (charged current cross sections)
@@
 \bi
-\item {\tt ThresholdEnergies.dat} (list of energies)
-\item {\tt CIDthreshold.dat} (list for CID threshold)
-\item {\tt MINOSthreshold.dat} (list for MINOS threshold)
-\item {\tt ECCthreshold.dat} (list for ECC threshold)
 \item {\tt XCC.dat} (charged current cross sections)

### 2007-01-17 Documentation/part1.tex
*** empty log message ***

@@
 
-medium $\gamma$ & {\tt BB\_350.glb} & $\gamma=350$ frogreen field scenario, 4~yr $\nu$- and
+medium $\gamma$ & {\tt BB\_350.glb} & $\gamma=350$, refurbished SPS scenario, 4~yr $\nu$- and
 4~yr $\bar{\nu}$-running & \cite{Burguet-Castell:2005pa} \\[0.2cm]
@@
   \hspace*{1.5cm} if(i==GLB\_THETA\_12) /* Prior on $\sin^2 \theta_{12}$ */ \\ 
-\hspace*{2.0cm} pv+=square((square(sin(startvalue))-fitvalue)/inputerror);\\ 
+\hspace*{2.0cm} pv+=square((startvalue-square(sin(fitvalue)))/inputerror);\\ 
   \hspace*{1.5cm} else \\ 

### 2007-01-17 Documentation/part1.tex
*** empty log message ***

@@
 \index{norm}{GLOBES tour@\GLOBES~tour}
-
-FOR WW MYSELF: TOUR ITSELF (+EXAMPLE FILE) TO BE UPDATED (VALUES MAY HAVE CHANGED DUE TO AEDL FILE DEFINITION
-CHANGES ETC. ALSO EXPERIMENT (OLD) NUFACT DOES NOT EXIST ANYMORE. AND: SOME ERROR IN THIS OLD
-NUFACT-FILE [WW].
-
-CHECK: GLBSETPROJECTION REALLY USED WITH GLB\_FIXED FOR DENSITY PARAMS? IS NFSTANDARD USED?
-(QUESTIONS FOR WW MYSELF!)
 
@@
 The output is written to {\tt stream}, which can be either {\tt stdout},
-or a file. Details about how to use \GLOBES\ with C can found in \Chapt~\ref{chapt:gettingstarted} and the following chapters.
+or a file.\footnote{Note that the output in this section can be slightly
+different from yours depending on the current version of the 
+probability engine, systematics implementation, and \AEDL\ file used.}
+Details about how to use \GLOBES\ with C can found in \Chapt~\ref{chapt:gettingstarted} and the following chapters.
 You can also find a summary of the most important \GLOBES\ $\chi^2$-functions in \tabl{stdfunctions}. Note that this chapter
@@
 glb\_params fit\_values = glbAllocParams();\\
-glb\_params starting\_values = glbAllocParams();\\
+glb\_params central\_values = glbAllocParams();\\
 glb\_params input\_errors = glbAllocParams();\\
@@
 \go{
-Oscillation probabilities in vacuum: 1->1: 0.999953 1->2: 2.69441e-05 1->3: 1.98019e-05 \\
-Oscillation probabilities in matter: 1->1: 0.999965 1->2: 2.02573e-05 1->3: 1.49021e-05 
+Oscillation probabilities in vacuum: 1->1: 0.999955 1->2: 2.58628e-05 1->3: 1.92142e-05 \\
+Oscillation probabilities in matter: 1->1: 0.999965 1->2: 2.01364e-05 1->3: 1.49644e-05 
 }
@@
 \go{
-chi2 with systematics only: 22.3984
+chi2 with systematics only: 22.433
 }
@@
 \go{
-This we would have from the CP-even appearance channel only: 21.6223
+This we would have from the CP-even appearance channel only: 21.1569
 }
@@
 \go{
-The sum over all rules gives again: 22.3984
+The sum over all rules gives again: 22.433
 }
@@
 \go{
-chi2 with correlations: 2.1038 \\
+chi2 with correlations: 1.99794 \\
 Position of minimum: theta12,theta13,theta23,delta,sdm,ldm,rho \\
-0.542002 0.0193698 0.747915 1.77688 6.66156e-05 0.00200817 \\
-1.00434 \\
-Iterations: 1693 \\
+0.541226 0.0193698 0.746156 1.74968 6.64399e-05 0.00200514 \\
+1.00341 \\
+Iterations: 1988 \\
 Note that s22theta13 is unchanged/kept fixed: 0.0015!  
@@
 \go{
-chi2 with correlations other than with deltacp: 4.32831 
+chi2 with correlations other than with deltacp: 4.02974  
 }
@@
 \go{
-chi2 with correlation only with deltacp: 2.80651 
+chi2 with correlation only with deltacp: 2.72943
 }
@@
 \go{
-chi2 with statistics only: 39.143
+chi2 with statistics only: 37.9736
 }
@@
   glbDefineParams(input\_errors,theta12*0.1,0,0,0,sdm*0.1,ldm/3); \\
-  glbDefineParams(starting\_values,theta12,theta13,theta23, \\
+  glbDefineParams(central\_values,theta12,theta13,theta23, \\
   \hspace*{0.5cm} deltacp,sdm,-ldm); \\
@@
   glbSetDensityParams(starting\_errors,1.0,GLB\_ALL); \\
-  glbSetCentralValues(starting\_values);\\
+  glbSetCentralValues(central\_values);\\
   glbSetInputErrors(input\_errors);\\
-  chi2=glbChiAll(starting\_values,minimum,GLB\_ALL);\\ 
+  chi2=glbChiAll(central\_values,minimum,GLB\_ALL);\\ 
   fprintf(stream,"chi2 at minimum: \%g $\backslash$n",chi2);\\
@@
 \go{
-chi2 at minimum: 6.20025 \\
+chi2 at minimum: 4.2137 \\
 Position of minimum: theta12,theta13,theta23,delta,sdm,ldm,rho \\
-0.591812 0.0264717 0.72763 1.08709 8.0004e-05 -0.00206094 \\
-0.970685  \\
-Iterations: 1946 \\
+0.585006 0.0275073 0.725679 1.05251 7.80256e-05 -0.00191273  \\
+0.979049  \\
+Iterations: 3184
 }
@@
  glbFreeParams(fit\_values); \\ 
- glbFreeParams(starting\_values);\\
+ glbFreeParams(central\_values);\\
  glbFreeParams(input\_errors);\\
@@
  fit\_values = glbAllocParams(); \\
- starting\_values = glbAllocParams(); \\
+ central\_values = glbAllocParams(); \\
  input\_errors = glbAllocParams(); \\
  minimum = glbAllocParams(); \\
- glb\_params minimum2 = glbAllocParams(); 
 }
@@
 \go{
- chi2 with systematics for all exps: 31.0797
+ chi2 with systematics for all exps: 30.5678
 }
@@
 \go{
-chi2 with systematics for 3000km: 22.3984 \\
-chi2 with systematics for 7500km: 8.68131 \\
+chi2 with systematics for 3000km: 22.433 \\
+chi2 with systematics for 7500km: 8.1348 \\
 The two add again to: 31.0797
@@
 \go{
-chi2 with correlations for 3000km: 2.1038 \\
-0.542002 0.0193698 0.747915 1.77688 6.66156e-05 0.00200817 \\ 
-1.00434 1 \\
-Iterations: 1693 \\
-\\
-chi2 with correlations for 7500km: 1.08421 \\
-0.557356 0.0193698 0.771359 4.77751 7.00762e-05 0.00200105 \\
-1 1.01517 \\
-Iterations: 661 \\
-\\
-chi2 with correlations for combination: 3.90835  \\
-0.544432 0.0193698 0.770175 1.78502 6.61621e-05 0.00200303 \\ 
-1.00431 1.03679 \\
-Iterations: 1636 \\
-\\
-The sum of the two chi2s is 3.18801, whereas the total chi2 is 3.90835!
-}
-Now find the $\mathrm{sgn}(\ldm)$-degeneracies for both individual
-experiments and test if they are still there in the combination of the
+chi2 with correlations for 3000km: 1.99794 \\
+0.541226 0.0193698 0.746156 1.74968 6.64399e-05 0.00200514 \\ 
+1.00341 1  \\
+Iterations: 1988 \\
+\\
+chi2 with correlations for 7500km: 0.803787 \\
+0.556503 0.0193698 0.771812 4.75971 7.00813e-05 0.0020012 \\ 
+1 1.01201 \\
+Iterations: 798 \\
+\\
+chi2 with correlations for combination: 3.68133 \\
+0.543335 0.0193698 0.77004 1.75576 6.59505e-05 0.00199913 \\ 
+1.00333 1.03269 \\
+Iterations: 2187 \\
+\\
+The sum of the two chi2s is 2.80173, whereas the total chi2 is 3.68133!
+}
+Now find the $\mathrm{sgn}(\ldm)$-degeneracy for the neutrino factory
+at the short baseline and test if it is still present in the combination of the
 experiments.
@@
  \hspace*{0.5cm} deltacp,sdm*0.1,ldm/3); \\
-glbDefineParams(starting\_values,theta12,theta13,theta23, \\
+glbDefineParams(central\_values,theta12,theta13,theta23, \\
 \hspace*{0.5cm} deltacp,sdm,-ldm);\\
   glbSetDensityParams(input\_errors,0.05,GLB\_ALL);\\
-  glbSetDensityParams(starting\_values,1.0,GLB\_ALL); \\
-  glbSetCentralValues(starting\_values);\\
+  glbSetDensityParams(central\_values,1.0,GLB\_ALL); \\
+  glbSetCentralValues(central\_values);\\
   glbSetInputErrors(input\_errors);\\
   \\
-  chi2=glbChiAll(starting\_values,minimum,0); \\ 
+  chi2=glbChiAll(central\_values,minimum,0); \\ 
   fprintf(stream,"chi2 at minimum, L=3000km: \%g $\backslash$n",chi2); \\
   glbPrintParams(stream,minimum);   \\
-  \\
-  chi2b=glbChiAll(starting\_values,minimum2,1);  \\
-  \mbox{fprintf(stream,"$\backslash$nchi2 at minimum, L=7500km: \%g \\ $\backslash$n",chi2b);} \\
-  glbPrintParams(stream,minimum2);  \\
   \\
@@
   glbPrintParams(stream,minimum);  \\
-  \\
-  chi2b=glbChiAll(minimum2,minimum2,GLB\_ALL); \\
-  fprintf(stream,"$\backslash$nchi2 for combination at minimum of Exp. 2: \\
-  \hspace*{0.5cm} \%g $\backslash$n",chi2b); \\
-  glbPrintParams(stream,minimum2);  
 }
 \go{
-chi2 at minimum, L=3000km: 6.71794 \\
-0.591497 0.0257396 0.729058 1.11537 7.98867e-05 -0.00206005 \\
-0.970499 1 \\
-Iterations: 2104\\
-\\
-chi2 at minimum, L=7500km: 47.1013 \\
-0.590347 0.0018489 0.768372 0.984827 8.23415e-05 -0.00204588 \\
-1 0.780995 \\
-Iterations: 1270\\
-\\
-chi2 for combination at minimum of Exp. 1: 70.6353 \\
-0.607988 0.0165985 0.767682 1.41422 8.44573e-05 -0.00204853 \\
-0.96147 1.1831 \\
-Iterations: 1549 \\
-\\
-chi2 for combination at minimum of Exp. 2: 70.6357 \\
-0.608454 0.0165823 0.767757 1.41481 8.43864e-05 -0.00204853 \\
-0.961129 1.18304 \\
-Iterations: 1447 
+chi2 at minimum, L=3000km: 4.82021  \\
+0.584568 0.026488 0.727714 1.09015 7.78747e-05 -0.00191163 \\ 
+0.978867 1 \\
+Iterations: 3298 \\
+\\
+chi2 for combination at minimum of Exp. 1: 59.4879 \\
+0.599908 0.0172467 0.768342 1.39186 8.198e-05 -0.00189508 \\ 
+0.970733 1.17125 \\
+Iterations: 2327 \\
 }
@@
  glbFreeParams(fit\_values); \\
- glbFreeParams(starting\_values); \\
+ glbFreeParams(central\_values); \\
  glbFreeParams(input\_errors); \\
  glbFreeParams(minimum); \\
- glbFreeParams(minimum2); \\
 }
@@
   /* Set starting vales to suspected position at opposite sign of ldm */ \\
-  \mbox{glbDefineParams(starting\_values,theta12,theta13,theta23,deltacp,sdm,-ldm);} \\
-  glbSetDensityParams(starting\_values,1.0,GLB\_ALL); \\
+  \mbox{glbDefineParams(central\_values,theta12,theta13,theta23,deltacp,sdm,-ldm);} \\
+  glbSetDensityParams(central\_values,1.0,GLB\_ALL); \\
   \\
@@
   \\
-  glbSetCentralValues(starting\_values); \\
+  glbSetCentralValues(central\_values); \\
   glbSetInputErrors(input\_errors); \\
@@
   /* Localize degenerate solution by minimization over all parameters */ \\
-  double CL=glbChiAll(starting\_values,deg\_pos,GLB\_ALL); \\
+  double CL=glbChiAll(central\_values,deg\_pos,GLB\_ALL); \\
    \\

### 2007-01-17 Distribution/examples/example-tour.c
*** empty log message ***

@@ -236,7 +236,7 @@ int main(int argc, char *argv[])
   fprintf(stream,"chi2 at minimum, L=3000km: %g \n",chi2);
   glbPrintParams(stream,minimum);  
   chi2=glbChiAll(minimum,minimum,GLB_ALL); 
-  fprintf(stream,"\nchi2 for combination at minimum of Exp. 1: %g \n",chi2);
+  fprintf(stream,"\nchi2 for combination at minimum of L=3000km: %g \n",chi2);
   glbPrintParams(stream,minimum);  
  
   /* Destroy parameter vector(s) */

### 2007-01-17 Distribution/examples/example2.c

@@ -79,7 +79,7 @@ int main(int argc, char *argv[])
   /* Set two-parameter projection onto s22th13-axis: only deltacp free! */
   glbDefineProjection(theta13_projection,GLB_FIXED,GLB_FIXED,GLB_FIXED,
     GLB_FREE,GLB_FIXED,GLB_FIXED);
-  glbSetDensityProjectionFlag(theta13_projection, GLB_FREE, GLB_ALL);
+  glbSetDensityProjectionFlag(theta13_projection, GLB_FIXED, GLB_ALL);
   glbSetProjection(theta13_projection); 
 
   /* Iteration over all values to be computed */

### 2007-01-17 Distribution/examples/example4.c

@@ -93,18 +93,18 @@ double CalcProjection(double theldm,double thex,glb_params start_vector)
   double thesign=Sign(glbGetOscParams(start_vector,GLB_DM_ATM));
 
   glb_params input_errors = glbAllocParams();
-  glb_params starting_values = glbAllocParams();
+  glb_params central_values = glbAllocParams();
   glb_params minimum = glbAllocParams();
 
   glbDefineParams(input_errors,theta12*0.1,0,0,0,sdm*0.1,theldm/3);  
   glbSetDensityParams(input_errors,0.05,GLB_ALL);
-  /* Set starting values to respective +-dm31^2 to avoid falling into unwanted solution;
+  /* Set central values to respective +-dm31^2 to avoid falling into unwanted solution;
      Note that the error on dm31^2 should not be too small in order to avoid large
      contributions from the priors */  
-  glbDefineParams(starting_values,theta12,theta13,
+  glbDefineParams(central_values,theta12,theta13,
                                    theta23,deltacp,sdm,theldm*thesign);
-  glbSetDensityParams(starting_values,1.0,GLB_ALL);
-  glbSetCentralValues(starting_values);
+  glbSetDensityParams(central_values,1.0,GLB_ALL);
+  glbSetCentralValues(central_values);
   glbSetInputErrors(input_errors);
 
   start_vector=glbSetOscParams(start_vector,thetheta13,GLB_THETA_13);
@@ -115,7 +115,7 @@ double CalcProjection(double theldm,double thex,glb_params start_vector)
   glbSetOscParams(start_vector,glbGetOscParams(minimum,GLB_DELTA_CP),GLB_DELTA_CP);
  
   glbFreeParams(input_errors);
-  glbFreeParams(starting_values);
+  glbFreeParams(central_values);
   glbFreeParams(minimum);
   
   return res; 
@@ -127,18 +127,18 @@ double FindDeg(glb_params deg_pos,double theldm)
   SetOscParams(theldm);
 
   glb_params input_errors = glbAllocParams();
-  glb_params starting_values = glbAllocParams();
+  glb_params central_values = glbAllocParams();
 
-  glbDefineParams(starting_values,theta12,theta13,theta23,deltacp,sdm,-theldm);  
-  glbSetDensityParams(starting_values,1.0,GLB_ALL);
[…]

### 2007-01-17 Documentation/appendix.tex
*** empty log message ***

@@
 
-\section*{Superbeam experiments}
+\section{Superbeam experiments}
 \subsection*{T2K -- {\tt T2K.glb}}
@@
 from the QE sample with a free normalization at an energy resolution of $\mathrm{\sigma_e=0.085\, GeV}$ due to the
-Fermi Motion as is described in more detail in~\cite{Huber:2002mx}. The treatment of systematics is similar to
-\cite{Ishitsuka:2005qi}. The following rules are defined within {\tt T2K.glb}:  
+Fermi Motion. The normalization of the QE samples is kept free in order to avoid double counting of events since all QE
+events are also contained in the CC samples. The free normalization is introduced within a rule with the line 
+{\tt \@signalerror = 10.:0.0001}. In case of systematics switched on the systematics functions {\tt chiSpectrumTilt}
+for the QE sample and {\tt chiTotalRatesTilt} for the CC sample are used. However, for systematics switched off, only the
+systematics function for the CC sample is changed to {\tt chiNoSysTotalRates}, but the systematics function for the QE
+sample stays {\tt chiSpectrumTilt}. Note, that otherwise the free normalization would be swiched off and all events from
+the QE sample would be counted twice. A detailed discussion of this QE/CC sample splitting can be found in~\cite{Huber:2002mx}. 
+For the disappearance channels only the QE sample is used since statistics is already quite large and a treatment like for the
+appearance channels would only shlightly modify the results. The quantitative treatment of systematics is similar to \cite{Ishitsuka:2005qi}. The following rules are 
+defined within {\tt T2K.glb}:  
 \begin{center}
@@
 
-\section*{Reactor experiments}
+\section{Reactor experiments}
 \subsection*{Small reactor experiment -- {\tt Reactor1.glb}}
@@
 
-\section*{Beta beam experiments}
+\section{Beta beam experiments}
 \subsection*{CERN-Fr\'{e}jus baseline scenario -- {\tt BB\_100.glb}}
@@
 from the QE sample with a free normalization at an energy resolution of $\mathrm{\sigma_e=0.085GeV}$ due to the
-Fermi Motion identical to the treatment of systematics within the \TtoK\ and \TtoHK\ files. The following rules are defined within {\tt BBvar\_WC.glb}:
+Fermi Motion identical to the treatment of systematics within the \TtoK\ and \TtoHK\ files. Note, that here also the systematics
+function for the appearance sample is {\tt chiSpectrumTilt} in case of systematics on {\em and} off to avoid double counting of the
+QE events for systematics switched off. The following rules are defined within {\tt BBvar\_WC.glb}:
 \begin{center}
@@
 
-\section*{Neutrino factory experiments}
+\section{Neutrino factory experiments}
 \subsection*{Standard neutrino factory -- {\tt NFstandard.glb}}

### 2007-01-18 Documentation/appendix.tex
*** empty log message ***

@@
 
-\section{Superbeam experiments}
+\section{Superbeam Experiments}
+
+\begin{center}
+\begin{tabular}{lllllll} 
+Experiment & File & Runtime & Power & Baseline & Detector & Mass \\ \hline \hline
+T2K & {\tt T2K.glb} & 2~yr $\nu$/6~yr $\bar{\nu}$ & 0.77~MW & 295~km & WC & 22.5~kt \\ \hline
+T2HK & {\tt T2HK.glb} & 4~yr $\nu$/4~yr $\bar{\nu}$ & 4~MW & 295~km & WC & 500~kt \\ \hline
+NO$\nu$A & {\tt NOvA.glb} & 3~yr $\nu$/3~yr $\bar{\nu}$ & 1.12~MW & 812~km & TASD & 25~kt \\ \hline
+SPL & {\tt SPL.glb} & 2~yr $\nu$/8~yr $\bar{\nu}$ & 4~MW & 130~km & WC & 500~kt \\ 
+\end{tabular}
+\end{center}
+
 \subsection*{T2K -- {\tt T2K.glb}}
@@
 Background & $0.0056 \, \otimes \, (\nu_\mu \rightarrow \nu_x)_\mathrm{NC}$ & & 0.2 & $10^{-4}$ \\ \hline \hline 
-\multicolumn{3}{|l|}{Appearance (--) -- Spectrum}  & & \\ \hline
+\end{tabular}
+\end{center}
+\begin{center}
+\begin{tabular}{|l|ll|c|c|}
+\hline \hline
+\multicolumn{3}{|l|}{Appearance (--) -- Spectrum}  & $\sigma_\mathrm{norm}$ & $\sigma_\mathrm{cal}$ \\ \hline
 Signal & $0.505 \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_e)_\mathrm{QE}$ & & 10.0 & $10^{-4}$\\
@@
 Beam Background & $0.677 \, \otimes \, (\bar{\nu}_e \rightarrow \bar{\nu}_e )_\mathrm{CC}$ & $0.707 \, \otimes \, (\nu_e \rightarrow \nu_e )_\mathrm{CC}$ & 0.02 & $10^{-4}$ \\ \hline \hline
-\end{tabular}
-\end{center}
-\begin{center}
-\begin{tabular}{|l|ll|c|c|}
-\hline \hline
 \multicolumn{3}{|l|}{Disappearance (--)} & & \\ \hline 
@@
 
-\section{Reactor experiments}
-\subsection*{Small reactor experiment -- {\tt Reactor1.glb}}
+\section{Reactor Experiments}
+
+\begin{center}
+\begin{tabular}{lllllll} 
+Experiment & File & Runtime & th. Power & Baseline & Det. & Mass \\ \hline \hline
+Small & {\tt Reactor1.glb} & 5~yr $\bar{\nu}$ & 4~GW & 1.7~km & LS & 20~t \\ \hline
+Large & {\tt Reactor.glb} & 8~yr $\bar{\nu}$ & 10~GW & 1.7~km & LS & 100~t \\ \hline
+DoubleCHOOZ & {\tt D-Chooz\_near.glb} & 5~yr $\bar{\nu}$ & 2$\times$4.2~GW & 0.1~km & LS & 10.16~t \\ 
+ & {\tt D-Chooz\_far.glb} & 5~yr $\bar{\nu}$ & 2$\times$4.2~GW & 1.05~km & LS & 10.16~t \\ 
+\end{tabular}
+\end{center}
+
+\subsection*{Small Reactor Experiment -- {\tt Reactor1.glb}}
 
@@
 
-\subsection*{Large reactor experiment -- {\tt Reactor2.glb}}
+\subsection*{Large Reactor Experiment -- {\tt Reactor2.glb}}
 
@@
 
-\section{Beta beam experiments}
-\subsection*{CERN-Fr\'{e}jus baseline scenario -- {\tt BB\_100.glb}}
+\section{Beta Beam Experiments}
+
+\begin{center}
+\begin{tabular}{lllllll} 
+Experiment & File & Runtime & $\gamma$ & Baseline & Det. & Mass \\ \hline \hline
+Low $\gamma$ & {\tt BB\_100.glb} & 4~yr $\nu$/4~yr $\bar{\nu}$ & 100 & 130~km & WC & 500~kt \\ \hline
+Medium $\gamma$ & {\tt BB\_350.glb} & 4~yr $\nu$/4~yr $\bar{\nu}$ & 350 & 730~km & WC & 500kt \\  \hline
+Variable & {\tt BBvar\_WC.glb} & 4~yr $\nu$/4~yr $\bar{\nu}$ & variable & variable & WC & 500~kt \\  \hline
+Variable & {\tt BBvar\_TASD.glb} & 4~yr $\nu$/4~yr $\bar{\nu}$ & variable & variable & TASD & 50~kt \\ 
+\end{tabular}
+\end{center}
+
+\subsection*{CERN-Fr\'{e}jus Baseline Scenario -- {\tt BB\_100.glb}}
 
@@
 Atm. background & & (from external file) & 0.02 & $10^{-4}$ \\ \hline \hline
-\end{tabular}
-\end{center}
-\begin{center}
-\begin{tabular}{|l|ll|c|c|}
-\hline \hline
 \multicolumn{3}{|l|}{Disappearance -- $^6$He stored} & &  \\ \hline
@@
 
-\subsection*{Higher gamma scenario -- {\tt BB\_350.glb}}
+\subsection*{Higher Gamma Scenario -- {\tt BB\_350.glb}}
 
@@
 
-\subsection*{Variable beta beam (Water Cerenkov) -- {\tt BBvar\_WC.glb}}
+\subsection*{Variable Beta Beam (Water Cerenkov) -- {\tt BBvar\_WC.glb}}
 
@@
 EXP\_FACTOR} is a parameter that describes ion decay scaling. As default value {\tt EXP\_FACTOR=0} should be chosen.
-See \cite{Huber:2005jk} for a detailed discussion of this parameter. The baseline is
+Technically {\tt EXP\_FACTOR} can be chosen completely free, but the value should not deviate far from zero to stay
+meaningful. See \cite{Huber:2005jk} for a detailed discussion of this parameter. The $\gamma$ value has to be chosen above 50. {\tt BBvar\_WC.glb} is optimized for $\gamma\lesssim350$ but
+should be satisfactory up to $\gamma=500$. 
+ The baseline is
 L~=~{\tt baselinefactor}$\cdot\gamma$~km, the
 fiducial mass of the detector is $\mathrm{m_{det} = 500 \,kt}$ and 4~years $\nu$-running and 4~years
-$\bar{\nu}$-running are assumed. The appearance measurement involves the total rates data from all CC events and the spectral data
+$\bar{\nu}$-running are assumed. The {\sf AEDL}-Variable {\tt baselinefactor} must be chosen such that the baseline lies
+in the interval $\mathrm{1\,km\lesssim\, L \, \lesssim\, 2\cdot R_{EARTH}}$. The appearance measurement involves the total rates data from all CC events and the spectral data
 from the QE sample with a free normalization at an energy resolution of $\mathrm{\sigma_e=0.085GeV}$ due to the
@@
 Background & $0.003 \, \otimes \, (\nu_e \rightarrow \nu_x)_\mathrm{NC}$ & & 0.05 & $10^{-4}$ \\ \hline \hline 
-\multicolumn{3}{|l|}{Appearance -- $^{18}$Ne stored -- Spectrum} & & \\ \hline
-Signal & $0.55 \, \otimes \, (\nu_e \rightarrow \nu_\mu)_\mathrm{QE}$ & & 10.0 & $10^{-4}$ \\
+\end{tabular}
+\end{center}
+\begin{center}
+\begin{tabular}{|l|ll|c|c|}
+\hline \hline
+\multicolumn{3}{|l|}{Appearance -- $^{18}$Ne stored -- Spectrum} & $\sigma_\mathrm{norm}$ & $\sigma_\mathrm{cal}$\\ \hline
+Signal & $0.55 \, \otimes \, (\nu_e \rightarrow \nu_\mu)_\mathrm{QE}$ & \hspace{6cm} & 10.0 & $10^{-4}$ \\
  & & & & \\
@@
 
-\subsection*{Variable beta beam (TASD) -- {\tt BBvar\_TASD.glb}}
+\subsection*{Variable Beta Beam (TASD) -- {\tt BBvar\_TASD.glb}}
 
@@
 year and $(60/\gamma)^\alpha\cdot5.8\cdot10^{18}$ $^{6}$He decays per year are assumed where $\alpha$~=~{\tt
-EXP\_FACTOR} is a parameter that describes ion decay scaling. As default value {\tt EXP\_FACTOR=0} should be chosen.
-See \cite{Huber:2005jk} for a detailed discussion of this parameter. The baseline is L~=~{\tt baselinefactor}$\cdot\gamma$~km, the
+EXP\_FACTOR} is a parameter that describes ion decay scaling.
+As default value {\tt EXP\_FACTOR=0} should be chosen.
+Technically {\tt EXP\_FACTOR} can be chosen completely free, but the value should not deviate far from zero to stay
+meaningful. See \cite{Huber:2005jk} for a detailed discussion of this parameter. The $\gamma$ value has to be 
+chosen above 80. The baseline is L~=~{\tt baselinefactor}$\cdot\gamma$~km, the
 fiducial mass of the detector is $\mathrm{m_{det} = 50 \,kt}$ and 4~years $\nu$-running and 4~years
-$\bar{\nu}$-running are assumed. The energy resolution is $\mathrm{\sigma_e=6 \% \cdot \sqrt{E}}$ for electrons and  $\mathrm{\sigma_e=3 \% \cdot \sqrt{E}}$
+$\bar{\nu}$-running are assumed. 
+The {\sf AEDL}-Variable {\tt baselinefactor} must be chosen such that the baseline lies
+in the interval $\mathrm{1\,km\lesssim\, L \, \lesssim\, 2\cdot R_{EARTH}}$.
+The energy resolution is $\mathrm{\sigma_e=6 \% \cdot \sqrt{E}}$ for electrons and  $\mathrm{\sigma_e=3 \% \cdot \sqrt{E}}$
 for muons.  The following rules are defined within {\tt BBvar\_TASD.glb}:
@@
 
-\section{Neutrino factory experiments}
-\subsection*{Standard neutrino factory -- {\tt NFstandard.glb}}
+\section{Neutrino Factory Experiments}
+
+\begin{center}
+\begin{tabular}{lllllll} 
+Experiment & File & Runtime & $\mathrm{E_\mu}$ & Baseline & Det. & Mass \\ \hline \hline
+Standard & {\tt NFstandard.glb} & \begin{tabular}{l}4~yr $\nu$ \\ 4~yr $\bar{\nu}$\end{tabular} & 50~GeV & 3000~km & \begin{tabular}{l}MID\end{tabular} & \begin{tabular}{l}50~kt\end{tabular} \\ \hline
+Variable & {\tt NFvar.glb} & \begin{tabular}{l}4~yr $\nu$ \\ 4~yr $\bar{\nu}$\end{tabular} & variable & variable & \begin{tabular}{l}MID\end{tabular} & \begin{tabular}{l}50~kt\end{tabular} \\ \hline
+Silver & {\tt NF\_GoldSilver.glb} & \begin{tabular}{l}4~yr $\nu$ \\ 4~yr $\bar{\nu}$\end{tabular} & variable & variable
+&\begin{tabular}{l}MID\\ECC\end{tabular}&\begin{tabular}{l}50~kt\\5~kt\end{tabular}\\ \hline
+Hybrid & {\tt NF\_hR\_lT.glb} & \begin{tabular}{l}4~yr $\nu$ \\ 4~yr $\bar{\nu}$\end{tabular} & variable & variable & \begin{tabular}{l}Hybrid\end{tabular} & \begin{tabular}{l}50~kt\end{tabular} \\ 
+\end{tabular}
+\end{center}
+
+\subsection*{Standard Neutrino Factory -- {\tt NFstandard.glb}}
 
@@
 
-\subsection*{Variable neutrino factory -- {\tt NFvar.glb}}
+\subsection*{Variable Feutrino Factory -- {\tt NFvar.glb}}
 
@@
 The neutrino beam is produced by the decay of muons stored in a storage ring at a parent energy of
-$\mathrm{E_\mu =}$~{\tt emax}. The parent energy of the muons can be set in the range
-$\mathrm{10\,GeV\lesssim E_\mu\lesssim 80\, GeV}$. The baseline of the scenario is L~=~{\tt BASELINE}. 
+$\mathrm{E_\mu =}$~{\tt emax}. The parent energy of the muons can be appropriately set in the range
+$\mathrm{10\,GeV\lesssim E_\mu\lesssim 80\, GeV}$ (technically it is possible to arrange {\tt emax} for $\mathrm{E_\mu=5\,
+GeV}$). The baseline of the scenario is L~=~{\tt BASELINE} and has to be chosen within the interval $\mathrm{1\,km\lesssim\, L \, \lesssim\, 2\cdot R_{EARTH}}$. 
 Besides these settings the other attributes of {\tt NFvar.glb} are similar to the ones from {\tt
@@
 
-\subsection*{Variable neutrino factory with Silver Channel -- \\{\tt NF\_GoldSilver.glb}}
+\subsection*{Variable Feutrino Factory with Silver Channel -- \\{\tt NF\_GoldSilver.glb}}
 
@@
 \ei
-The beam and golden channel attributes are similar to {\tt NFvar.glb}. The parent energy of the muons can be set
-in the range $\mathrm{10\,GeV\lesssim E_\mu\lesssim 80\, GeV}$. The baseline is set by the {\sf
-AEDL}-Variable {\tt BASELINE}. For the silver channel, an additional ECC detector with a fiducial mass
+The beam and golden channel attributes are similar to {\tt NFvar.glb}. The parent energy of the muons can be appropriately
+set in the range $\mathrm{10\,GeV\lesssim E_\mu\lesssim 80\, GeV}$ (technically it is possible to arrange {\tt emax} for $\mathrm{E_\mu=5\,
+GeV}$).
+The baseline is set by the {\sf AEDL}-Variable {\tt BASELINE} and has to be chosen within the interval $\mathrm{1\,km\lesssim\, L \, \lesssim\, 2\cdot R_{EARTH}}$.
+For the silver channel, an additional ECC detector with a fiducial mass
 $\mathrm{m_{ECC} = 5\,kt}$ is assumed to be located at the same baseline as the MID detector. The energy
@@
 
-\subsection*{High Resolution/Low Threshold neutrino factory -- \\{\tt NF\_hR\_lT.glb}}
+\subsection*{High Resolution/Low Threshold Neutrino Factory -- \\{\tt NF\_hR\_lT.glb}}
 
@@
 For calculations that involve {\tt NF\_hR\_lT.glb}, the same additional files as for {\tt NFvar.glb} are
-required and the {\sf AEDL}-Variables {\tt emax} and {\tt BASELINE} have to be set. {\tt NF\_hR\_lT.glb} 
+required and the {\sf AEDL}-Variables {\tt emax} and {\tt BASELINE} have to be set within the same constraints as for {\tt
+NFvar.glb}. {\tt NF\_hR\_lT.glb} 
 implements a lower threshold ($\mathrm{\sim\,1\,GeV}$) at a higher energy resolution
@@
 \multicolumn{3}{|l|}{Disappearance -- $\mu^+$-stored} & $\sigma_\mathrm{norm}$ & $\sigma_\mathrm{cal}$ \\ \hline
-Signal & $0.9 \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_\mu)_\mathrm{CC}$ & $0.9 \, \otimes \, (\nu_e \rightarrow \nu_\mu)_\mathrm{CC}$& 0.025 & $10^{-4}$ \\
+Signal & $0.9 \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_\mu)_\mathrm{CC}$ & $0.9 \, \otimes \,
+(\nu_e \rightarrow \nu_\mu)_\mathrm{CC}$ \hspace{1.6cm} & 0.025 & $10^{-4}$ \\
  & & & & \\
 Background & $1.0 \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_x)_\mathrm{NC}$ & & 0.2 & $10^{-4}$ \\ \hline \hline
+\end{tabular}
+\end{center}
+\begin{center}
+\begin{tabular}{|l|ll|c|c|}
+\hline \hline
 \multicolumn{3}{|l|}{Appearance -- $\mu^+$-stored} & & \\ \hline
@@
 Signal &  $0.9 \, \otimes \, (\nu_\mu \rightarrow \nu_\mu)_\mathrm{CC}$ & $0.9 \, \otimes \, (\bar{\nu}_e
-\rightarrow \bar{\nu}_\mu)_\mathrm{CC}$& 0.025& $10^{-4}$\\
+\rightarrow \bar{\nu}_\mu)_\mathrm{CC}$ & 0.025& $10^{-4}$\\
  & & & & \\

### 2007-01-18 Documentation/part2.tex
JK - Minor modifications in the appendices (except appendix B)
*** empty log message ***

@@
 For each of these layers, the Hamiltonian in matter is diagonalized in order to propagate the neutrino transition amplitudes.
+Since this step is computationally expensive, a specialized algorithm is used~\cite{Kopp:2006wp}.
 Finally, the transition probability is obtained as the absolute square of the total neutrino transition amplitudes.
@@
 \mycaption{\label{tab:binfluxes} Built-in fluxes currently supported by \GLOBES . For details on
-the beta beams, see \Ref~\cite{Huber:2005jk}.
-MR ABOUT BETA BEAMS: DOES THIS REALLY WORK FOR ALL ISOTOPES?
-IS THE DIFFERENCE BETWEEN 3 AND 4 CORRECTLY DESCRIBED?
-WHY IS THERE A NORMALIZATION FACTOR 50.9 IN THE FILES?}
+the beta beams, see \Ref~\cite{Huber:2005jk}. }
 \end{table}
@@
 \end{quote}
-MR: SOME MORE ELABORATION AND AN EXAMPLE FOR THE BETA BEAMS MAY BE APPROPRIATE HERE [WW].
+%
+Furthermore, two beta beam neutrino fluxes are available: inverse beta decay ({\tt @builtin = 3}, \ie\ from the decay of
+$^{18}$Ne-isotopes) and beta decay ({\tt @builtin = 4}, \ie\ from the decay of $^{6}$He-isotopes). For a beta beam file, using
+the beta beam built-in fluxes, the end point energy of the decay ({\tt @end\_point}), the number of decays per year 
+({\tt @stored\_muons}), and the acceleration factor of the ions $\gamma$ ({\tt @gamma}) have to be specified. An example for
+setting up a beta beam neutrino flux from the decay of $^{18}$Ne isotopes is
+\begin{quote}
+{\tt nuflux(\#nu\_e\_flux)<\\
+\tb  @builtin = 3\\
+\tb  @gamma = 130.0\\
+\tb  @end\_point = 0.0034\\
+\tb  @stored\_muons = 2.2e+18\\
+\tb  @time = 4.0\\
+>}
+\end{quote}
 %
@@
 \index{aedl}{nuflux@{\tt nuflux}!flux file@{\tt "@flux\_file}}
+%
 For a user-defined flux, one has to specify the file name:

### 2007-01-18 Documentation/appendix.tex
*** empty log message ***

@@
 events are also contained in the CC samples. The free normalization is introduced within a rule with the line 
-{\tt \@signalerror = 10.:0.0001}. In case of systematics switched on the systematics functions {\tt chiSpectrumTilt}
+{\tt @signalerror = 10.:0.0001}. In case of systematics switched on the systematics functions {\tt chiSpectrumTilt}
 for the QE sample and {\tt chiTotalRatesTilt} for the CC sample are used. However, for systematics switched off, only the
@@
 the QE sample would be counted twice. A detailed discussion of this QE/CC sample splitting can be found in~\cite{Huber:2002mx}. 
-For the disappearance channels only the QE sample is used since statistics is already quite large and a treatment like for the
-appearance channels would only shlightly modify the results. The quantitative treatment of systematics is similar to \cite{Ishitsuka:2005qi}. The following rules are 
+For the disappearance channels only the QE sample is used since statistics is already quite large and a treatment as for the
+appearance channels would only slightly modify the results. The quantitative treatment of systematics is similar to \cite{Ishitsuka:2005qi}. The following rules are 
 defined within {\tt T2K.glb}:  
@@
 $\mathrm{m_{det} = 500 \,kt}$ at a baseline of L~=~130~km. The energy resolution is introduced manually by
-four migration matrices (for $\nu_e , \bar{\nu}_e ,\nu_{\mu} ,\bar{\nu}_\mu$) that describe energy smearing mainly
-from Fermi Motion. The following rules are defined within {\tt SPL.glb}: 
+four migration matrices (for $\nu_e , \bar{\nu}_e ,\nu_{\mu} ,\bar{\nu}_\mu$) that describe energy smearing 
+due to Fermi Motion. The following rules are defined within {\tt SPL.glb}: 
 \begin{center}
@@
 individual uncertainties (see~\cite{Huber:2003pm}). The energy resolution is
-$\mathrm{\sigma_e=5\%\cdot\sqrt{E_{vis}}}$ and the choice for {\tt sigma\_function} is {\tt \#inverse\_beta}. The following rules are defined within {\tt Reactor1.glb}: 
+$\mathrm{\sigma_e=5\%\cdot\sqrt{E_{vis}}}$ and the choice for {\tt @sigma\_function} is {\tt \#inverse\_beta}. The following rules are defined within {\tt Reactor1.glb}: 
 \begin{center}
@@
 The energy resolution is
-$\mathrm{\sigma_e=5\%\cdot\sqrt{E_{vis}}}$ and the choice for {\tt sigma\_function} is {\tt \#inverse\_beta}.
+$\mathrm{\sigma_e=5\%\cdot\sqrt{E_{vis}}}$ and the choice for {\tt @sigma\_function} is {\tt \#inverse\_beta}.
 The following rules are defined within {\tt D-Chooz\_near.glb} and {\tt D-Chooz\_far.glb}:
@@
 $\bar{\nu}$-running are assumed. The energy resolution is introduced manually by
-four migration matrices (for $\nu_e , \bar{\nu}_e ,\nu_{\mu} ,\bar{\nu}_\mu$) that describes energy smearing mainly
-from Fermi Motion. The following rules are defined within {\tt BB\_100.glb}:
+four migration matrices (for $\nu_e , \bar{\nu}_e ,\nu_{\mu} ,\bar{\nu}_\mu$) that describe the energy smearing 
+due to Fermi Motion. The following rules are defined within {\tt BB\_100.glb}:
 \begin{center}
@@
 The neutrino beam is produced by the decay of muons stored in a storage ring at a parent energy of
-$\mathrm{E_\mu = 50\,GeV}$. A number of $1.06\cdot10^{21}$ useful muon decays per year are assumed in each
+$\mathrm{E_\mu = 50\,GeV}$. A number of $1.06\cdot10^{21}$ useful muon decays per year is assumed in each
 polarity (corresponding to $5.3\cdot10^{20}$ useful muon decays per year and polarity for simultaneous operation
@@
 \bi
-\item {\tt emax} (parent energy of the stored muons $\left[\mathrm{km}\right]$)
-\item {\tt BASELINE} (experiment baseline $\left[\mathrm{GeV}\right]$)
+\item {\tt emax} (parent energy of the stored muons $\left[\mathrm{GeV}\right]$)
+\item {\tt BASELINE} (experiment baseline $\left[\mathrm{km}\right]$)
 \ei

### 2007-01-19 Documentation/tutorials/Features-tutorial/tex/Features.tex
PH -- renamed a few files and added my tex-sources

\documentclass[12pt,a4paper]{article}

%\usepackage{german}
\usepackage{a4wide}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
%\usepackage{ae,aecompl}

\pagestyle{empty}
\setlength{\parindent}{0pt}
\renewcommand{\labelenumi}{\alph{enumi})}
\renewcommand{\labelenumii}{\roman{enumii})}

\newcommand{\aufg}[1]{\vspace{4mm}{\bf\underline{Problem #1:}}\vspace{3mm}}

\begin{document}
\vspace*{-3cm}
\begin{center}
{\large\bf GLoBES Tutorial: Using user-defined systematics and
  oscillation probabilities}\\[0.3cm]
GLoBES workshop in Heidelberg, Germany, January 24-26, 2007
\end{center}
\vspace*{2mm}

Patrick\ Huber \hfill \today

\bigskip
\hrule
\vspace*{4mm}

{\small In this advanced-level tutorial, we show how to use
  user-defined $\chi^2$-functions to treat a two detector setup with
  various systematical errors. This configuration is relevant for
  reactor neutrino experiments like Double Chooz.

  In a second part we will provide an example of how define an
  oscillation probability engine which includes non-standard physics.
  In the case at hand we deal with some form of 'decoherence' which
  leads to an exponential damping of the oscillation signature.  }

\vspace*{0.5cm}

\subsection*{Part 1: Reactor neutrino experiments}

In a reactor neutrino experiment, one tries to determine
$\sin^22\theta_{13}$ by measuring the disappearance of $\bar\nu_e$. In
order to be able to measure this small effect it is crucial to know
how many neutrinos are emitted from the source. To this end a near
detector is used. The comparison of event rates between near and far
detector allows to cancel most of the systematics associated with
source and only the relative normalization of the two detectors remains
(at first order). This setup cannot be described within GLoBES without
using a special user-defined $\chi^2$-function. The example shown here
is taken from~\cite{Huber:2006vr} and quite accurately describes
Double Chooz.

The actual C-code is in {\tt systematic.c}, typing {\tt make} should
compile the software and {\tt ./systenatic} executes the program,
which takes about 1 minute to finish. It will produce four data files
{\tt sys-data0} till {\tt sys-data3}. You can use {\tt gnuplot <
  systematic.gnuplot} to produce a plot, {\tt systematic.eps}. We use
{\tt gnuplot} in this example because it is GPL'd and widely
available, but of course, you can use whatever software you prefer for
plotting. In some cases you might want to adjust the output format to
fit your preference.

\aufg{1} {\bf The AEDL files}

First, we need to have a description of the two detectors. Have a look
a {\tt D-Chooz\_far.glb} and {\tt D-Chooz\_near.glb}. What are the
differences beyond the obvious baseline?

\begin{enumerate}
\item {\tt @norm} -- the near detector is less efficient, basically
  because it has less overburden, hence higher backgrounds and thus
  less life time.
\item {\tt @sys\_on\_function} -- {\tt chiZero} is used for the near
  detector since we want to compute the $\chi^2$ only once.
  Otherwise GLoBES would compute it for each detector since they are
  regarded as two experiments. {\tt
    chiDCNorm} is used for the far detector and describes a $\chi^2$
  definition which includes an error on the flux, the fiducial mass of
  the near detector, the fiducial mass of the far detector, the energy
  scale of the near detector and the energy scale of the far detector.
The definition of the $\chi^2$ is
\begin{align}
  \chi^2 &= \chi^2_{F} + \chi^2_{N} + \chi^2_\mathrm{pull}
\, ,
  \label{eq:chi2total}
\end{align}

where
\begin{align}
  \hspace{-1 cm} \chi^2_{F} &= \sum_i
     \frac{\left[ (1 + a_{F,\rm fid} + a_{\rm norm} ) T_{
F,i}\, - O_{F,i} \right]^2}{O_{F,i}} \, , \label{eq:chi2F1} \\
  \hspace{-1 cm} \chi^2_{N} &= \sum_i
    \frac{\left[ (1 + a_{N,\rm fid} + a_{\rm norm}) T_{N,i
}
        - O_{N,i} \right]^2}{O_{N,i}}     \, ,    \label{eq:chi2N2} \\
  \hspace{-1 cm} \chi^2_\mathrm{pull} &= \frac{a_{F,\rm fid}^2}{\sigma_{F,\rm fi
d}^2}
       \, +\, \frac{a_{N,\rm fid}^2}{\sigma_{N,\rm fid}^2}
       \, +\, \frac{a_{\rm norm}^2}{\sigma_{\rm norm}^2} \, .
       \label{eq:chi2pull}
\end{align}

Finally, we introduce an energy calibration error which is implemented
as a re-binning of $T_{F,i}$, and $T_{N,i}$ before
the $\chi^2$ analysis (see App.~A of~\cite{Huber:2003pm}). It is uncorrelated
between the two detectors.

\item {\tt @sys\_on\_errors} -- of course contains only values if {\tt
    chiDCNorm} is used.
\end{enumerate}

\aufg{2} {\bf The $\mathbf{\chi}^2$-function}

Now that we have the AEDL files, we need to define the $\chi^2$-function
and have to make sure that GLoBES knows about it. It is quite
straightforward to write code which realizes above definition.

There are two things in the definition of {\tt chiDCNorm}, which need
to be mentioned: For implementing the energy scale error we use the
GLoBES function {\tt  glbShiftEnergyScale}. Second, in order to ensure
good performance we just get the memory address from GLoBES where is
stores all its event rates (pass by reference). This address does not
change during the computation and therefore our own $\chi^2$-function
is as effective as the GLoBES built-ins. The functions used for that
is {\tt  glbGetRuleRatePtr}.

The new $\chi^2$-function has to be registered. This serves a
two-fold purpose: GLoBES needs to know what function to actually use
and secondly, in that process you give  your new
function a name. This name will be used in the AEDL file to identify your
function and thus prevents errors from using the wrong function.

\aufg{3} {\bf Adding a shape error}

So far, we have no way to account for the fact the neutrino spectrum
of  a nuclear reactor is only known up to some accuracy. There are two
reasons for that: the neutrinos are produced by a mixture of isotopes
and this mixture initially is only known with some error and moreover
changes during reactor operation (burn-up). The other reason is that
the neutrino spectrum for each isotope is measured only with some
accuracy. We lump all these effects together in what is called a shape
error. Basically, this is an error which changes the number of events
in a bin uncorrelated with any other bin, but fully correlated between
detectors. This error largely cancels in a two detector setup.
\begin{align}
  \hspace{-1cm} \chi^2_{F} &= \sum_i
     \frac{\left[ (1 + a_{F,\rm fid} + a_{\rm norm} + a_{\mathrm{shape}, i}) T_{F,i}
        - O_{F,i} \right]^2}{O_{F,i}}      \, ,    \\
  \hspace{-1 cm} \chi^2_{N} &= \sum_i
    \frac{\left[ (1 + a_{N,\rm fid} + a_{\rm norm}+ a_{\mathrm{shape},i}) T_{N,i}
       - O_{N,i} \right]^2}{O_{N,i}}     \, ,    \ \\
  \hspace{-1 cm} \chi^2_\mathrm{pull} &= \frac{a_{F,\rm fid}^2}{\sigma_{F,\rm fid}^2}
       \, +\, \frac{a_{N,\rm fid}^2}{\sigma_{N,\rm fid}^2}
       \, +\, \frac{a_{\rm norm}^2}{\sigma_{\rm norm}^2}
       \, +\, \sum_i \frac{a_{\mathrm{shape},i}^2}{\sigma_{\mathrm{shape},i}^2}\, .      
\end{align}
This definition increases the number of nuisance parameters quite a
lot and that creates some issues with the convergence of the numerical
minimization. This can be alleviated by a clever way to chose the
starting values for this minimization. We always will start with a
very small $\sin^22\theta_{13}$ and save the location of the minimum.
We use that as starting value for the next larger value of
$\sin^22\theta_{13}$. If the step size in $\sin^22\theta_{13}$ is
small enough we get good convergence. Just try setting {\tt
  step\_size} to $0.1$ (instead of $0.01$) and you will see that
introducing the shape error makes the $\chi^2$ larger! This is a clear
sign of failure to converge.

\aufg{4} {\bf Adding a bin-to-bin error}

A bin-to-bin error is a theorist's way to account for all those little
real world things which you only know once you have done the
experiment. It is like a shape error but now also uncorrelated between
near and far detector. That is the $a_{\mathrm{shape},i}$ acquire one more
index and there are twice as many.

\begin{center}
\includegraphics[width=0.8\textwidth]{systematic}
\end{center}

\subsection*{Part 2: Non-standard physics}

In principle one can imagine an arbitrary number of non-standard
scenarios, like 4 neutrinos, modified matter effect,
Lorentz-invariance violation asf. GLoBES now allows the user to supply
his own oscillation probability routines. This feature is not only
useful to deal with non-standard effects, but it can be used to write
debugging  tools (like watching the minimizers step-by-step) or to
provide routines which are optimized for a certain application.

Here we will discuss a so called damping effect in a reactor neutrino
experiment. The reactor experiment will be much simpler than in the
previous section and consists only of one detector. The oscillation
physics is taken from~\cite{Blennow:2005yk}.

The actual C-code is in {\tt oscillation.c}, typing {\tt make} should
compile the software and {\tt ./oscillation} executes the program,
which takes about 1 minute to finish. It will produce two data files
{\tt osc-data1} and {\tt osc-data2}. You can use {\tt gnuplot <
  oscillation.gnuplot} to produce a plot, {\tt oscillation.eps}.

\aufg{1} {\bf Implementing the oscillation engine}

The oscillation probability for our example is given by

\begin{eqnarray}
P_{\bar e\bar e} &=&
c_{13}^4 \left\{ 1-2 s_{12}^2 c_{12}^2 ( 1-D_{21}\cos(2\Delta_{21}))
  \right\}\nonumber\\
&&+2 s_{13}^2 c_{13}^2 [D_{21} c_{12}^2 \cos(2\Delta_{31})\nonumber\\
&&+D_{32}s_{12}^2\cos(2\Delta_{32})] + s_{13}^2 
\,,
\end{eqnarray}

where

\begin{equation}
D_{ij}=\exp \left[ -4 \left(\Delta_{ij} \frac{\sigma_E}{E}\right)^2\right]\,.
\end{equation}
Thus we have one additional parameter $\sigma_E$ which has unit GeV.

\aufg{2} {\bf Using the new parameter}

Once we have an oscillation routine which can handle non-standard
physics, we have to maintain this new parameter and understand how the
minimizer deals with additional parameters.

\begin{center}
\includegraphics[width=0.8\textwidth]{oscillation}
\end{center}

\bibliographystyle{./apsrev}
\bibliography{references}

\end{document}

### 2007-01-19 Documentation/tutorials/Deg-tutorial/tex/DegTut.tex
*** empty log message ***

\documentclass[12pt,a4paper]{article}

%\usepackage{german}
\usepackage{a4wide}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{ae,aecompl}

\pagestyle{empty}
\setlength{\parindent}{0pt}
\renewcommand{\labelenumi}{\alph{enumi})}
\renewcommand{\labelenumii}{\roman{enumii})}

\newcommand{\aufg}[1]{\vspace{4mm}{\bf\underline{Problem #1:}}\vspace{3mm}}

\begin{document}
\vspace*{-3cm}
\begin{center}
{\large\bf GLoBES Tutorial: Finding Degeneracies}\\[0.3cm]
GLoBES workshop in Heidelberg, Germany, January 24-26, 2007
\end{center}
\vspace*{2mm}

Walter\ Winter \hfill January 25, 2007

\bigskip
\hrule
\vspace*{4mm}

{\small
In this advanced-level tutorial, we illustrate several techniques for discrete degeneracy localization in GLoBES (General Long Baseline Experiment Simulator). In order to include correlations as well, the unused parameters are usually marginalized/minimized over for a given set of simulated parameters. In principle, one can include all degenerate solutions in GLoBES by a complete scan of the parameter space with {\tt glbChiSys}. However, such a ``grid-based'' approach takes $g^d$ steps, where $g$ is the number of steps in one dimension (such as $\theta_{13}$), and $d$ is the dimension of the parameter space (typically $d=6$ for neutrino oscillations). For example, for 
$g=20$ steps in one direction, $20^6 = 6.4 \cdot 10^7$ evaluations of the systematics $\chi^2$
would be necessary, costing many hours of CPU time on a modern computer -- and this is only
for one set of input parameters. Therefore, GLoBES provides the concept of local minimization: Instead of scanning the complete parameter space, most of which is uninteresting because of a too large $\chi^2$,
a minimizer is started at the position of an ``educated guess'' and runs into the local minimum to include the correlations with the other parameters.
This reduces the computational effort to about $n \times 1 \, 000$ evaluations of the systematics $\chi^2$, where $n$ is the number of discrete degeneracies. For example, for $n=8$, the computational effort can be reduced from several hours to a few seconds. Since the local minimizer may end up in the wrong minimum if started too far away from the actual solution, finding a proper educated guess for the discrete degeneracies is crucial for this approach. Therefore, most of this tutorial deals with {\em methods to find an educated guess}. Note that the chosen examples do not represent the standard level of sophistication needed to use GLoBES, such as to simulate superbeams. They are chosen from a large set of calculations to illustrate cases where the straightforward methods fail.
}

\vspace*{0.5cm}

\subsubsection*{Finding the $\boldsymbol{\mathrm{sgn}(\Delta m_{31}^2)}$-degeneracy at T2HK}

\aufg{1} {\bf Warm-up}

In this tutorial, all programs are labeled {\tt deg\_tut\_n.c}, where {\tt n} is the problem number.
In every problem, the program, which we have provided for you, illustrates a straightforward solution. As you will see, we have chosen examples where the straightforward approach will fail. Familiarize yourself with the source code, compile the program with {\tt make deg\_tut\_n}, and run it with {\tt ./deg\_tut\_n}. You will see some additional information on screen while the program is executed. The output is, as a standard, written into a file named {\tt tutn.dat}, {\tt n} being the problem number again. Take a look at the output, which is a simple two-column file, with the editor or plot program of your choice. For example, start {\tt gnuplot} and enter {\tt plot "tutn.dat" with lines}.
%
As the next step, we will describe a possible solution to the problem, and is up to you to implement it in the given program file. We will demonstrate how the output should look like, but we will not provide a specific program code for the solution.

\vspace*{3mm}

Let's try that. In this example, we want to find the minimum $\chi^2$ at the $\mathrm{sgn}(\Delta m_{31}^2)$-degeneracy for T2HK. This values describes the sensitivity to the normal hierarchy as function of $\mathrm{log}( \sin^2 2 \theta_{13})$. We compute it as function of the simulated (true) $\mathrm{log}( \sin^2 2 \theta_{13})$ for the simulated $\delta_\mathrm{CP}=3 \pi/2$ in the program {\tt deg\_tut\_1.c}.
Try to follow the procedure above. This is what your result should look like:

\begin{center}
\includegraphics[height=4.5cm]{problem1}
\end{center}

The result is a smooth $\chi^2$, and the straightforward algorithm seems to work. Nothing to do here - but just wait, until we change $\delta_\mathrm{CP}$ in the next problem.

\aufg{2} {\bf Using analytical knowledge}

In this problem, we simply change $\delta_{\mathrm{CP}}$, {\it i.e.}, we compute the $\chi^2$ for the normal mass hierarchy sensitivity as function $\mathrm{log}( \sin^2 2 \theta_{13})$ for the simulated $\delta_{\mathrm{CP}}=\pi/8$. The result is the solid curve in the following plot:

\begin{center}
\includegraphics[height=5cm]{problem2}
\end{center}

Obviously, the calculated $\chi^2$ is much to large for the experiment considered: T2HK cannot measure the mass hierarchy very well because of the rather short baseline. The reason for the poor performance of our method is the contribution from both appearance and disappearance channels used in this experiment. In certain regions of the parameter space, the disappearance information is at least comparable to the appearance $\chi^2$, which means that one cannot rely on one specific topology (such as of the appearance channel). The educated guess for the minimizer is obviously not good enough.

\vspace*{3mm}

\underline{Possible improvement:}
 For very small $\theta_{13}$, the appearance rate in T2HK will be almost zero. Therefore, the topology is determined by the disappearance channel.

Fortunately, one can in this case obtain a very simple guess for the location of the degeneracy in $\Delta m_{31}^2$ (see, {\it e.g.}, Ref.~\cite{deGouvea:2005mi}):
\begin{equation}
(\Delta m_{31}^2)^- = - (\Delta m_{31}^2)^+ + 2 \, \Delta m_{21}^2 \, \cos^2 \theta_{12} \, .
\end{equation}
Use this value to start the minimizer there. As you see in the plot (dashed curve), this works very well for small $\theta_{13}$, but it fails for larger $\theta_{13}$ where the appearance channel adds information. In the next
problem, we will use a different method, which is more successful in this particular example. However, note that analytical knowledge can also be used for the appearance channel and for other problems. We have just restricted ourselves to the simplest example we could think of.

\aufg{3} {\bf Tracking algorithm}

Let us now improve the solution from last example by using a different method.

\vspace*{3mm}

\underline{Possible solution:} Assume that the straightforward approach works in a part of the parameter space, and we find a solution there which is good enough. A {\em tracking algorithm} can then use the output from {\tt glbChiAll} (local minimum, returned as the second parameter) as the input for the next minimization if the points are not too far apart from each other. In this case, the parameter space is ``adiabatically'' changed. Use, for example, $\theta_{13}=0$ as a starting guess and, from there, move gradually upward in $\theta_{13}$. 
The result should then resemble the dashed curve in the following figure:
\begin{center}
\includegraphics[height=5cm]{problem3}
\end{center}
In this particular case, it is a very good solution. If you scan the parameter space in the (true) $\delta_\mathrm{CP}$-direction (and other true parameter directions)
continously, you will find cases where the method fails for small $\theta_{13}$. Combine then the solutions
of problem~2 and problem~3, and you will have a working algorithm.

\subsubsection*{Locating the intrinsic $\boldsymbol{(\delta_\mathrm{CP},\theta_{13})}$-degeneracy at a neutrino factory}

\aufg{4} {\bf Pre-scanning a parameter sub-space}

We now consider a different experiment. A neutrino factory typically faces the intrinsic degeneracy
problem in a part of the parameter space. In this example, we compute the $\sin^2 \theta_{13}$ precision, {\it i.e.}, the $\chi^2$ as function of the fit $\sin^2 \theta_{13}$ for a given set of simulated values. We project onto the $\sin^2 \theta_{13}$-axis, because we want to read off the precision of $\sin^2 \theta_{13}$ including all correlations. In order to start the minimizer, we use the set of true/simulated values as an educated guess for the minimizer. We obtain the solid curve in the following figure:

\begin{center}
\includegraphics[height=5cm]{problem4}
\end{center}

Obviously, the $(\delta_\mathrm{CP},\theta_{13})$-degenerate solution on the right is not located properly. The reason is that the minimizer ends up in the local minimum corresponding to the best-fit point, whereas the degeneracy is located at a different value of $\delta_\mathrm{CP}$.

\vspace*{3mm}

\underline{Possible solution:}
For each $\sin^2 2 \theta_{13}$, use {\tt glbChiSys} to pre-scan the parameter space in the $\delta_{\mathrm{CP}}$ direction only. Since {\tt glbChiSys} is very fast, one can just scan the $\delta_{\mathrm{CP}}$-direction for the local minimum in the $\delta_{\mathrm{CP}}$ direction, and then use this
as a guess for the minimizer (such as using 50 steps from $0$ to $2 \pi$). This guess does not include the full correlation, but it exploits the dominant effect in an efficient manner. The full correlation is then obtained by the following run of {\tt glbChiTheta13}. The result is the dashed curve in the above figure.

\subsubsection*{Octant $\boldsymbol{(\theta_{23},\pi/2-\theta_{23})}$-degeneracy }

\aufg{5} {\bf Manual scan of a sub-space}

Locating the octant degeneracy is a computationally equivalent to the ability of an experiment to exclude the
other (``wrong'') octant. Naturally, one would use 
$\pi/2 - \theta_{23}$ as a guess for the starting point of the minimizer.
The octant degeneracy is especially tricky from the computational point of view 
for $\theta_{23}$ very close to maximal mixing: In this case, the local minimizer tends to end up
in the best-fit octant -- there is no barrier for preventing it from doing so. We illustrate this in this problem using T2HK, which has, as a standalone experiment without atmospheric data, a very weak octant resolution potential. 
For example, for $\sin^2 2 \theta_{13}=0.1$, we obtain the solid curve in the following figure as function of the simulated $\sin^2 \theta_{23}$:
\begin{center}
\includegraphics[height=5cm]{problem5}
\end{center}
As the algorithm tells you when you run it (see source code), the minimizer ends up in the wrong octant
for four points left and two points right adjacent to the vertical line (note that at the best-fit solution, $\chi^2 = 0$). So how can we prevent it from doing so? One possibility is to add an external constraint on $\theta_{23}$. However, this does not guarantee ending up in the targeted octant. In addition, a too high external precision in this constraint will make the $\chi^2$ artificially higher. In this and the next problem, we therefore discuss two possibilities which end up in the right octant for sure.

\vspace*{3mm}

\underline{Possible solution:}
Instead of minimizing over all parameters simultaneously, scan the $\theta_{23}$-direction separately and
use {\tt glbChiNP} with a user-defined projection to marginalize over the other parameters (keeping $\theta_{23}$ fixed). Therefore, we split the marginalization procedure in a grid-based and local minimization process. For example, for the simulated $\sin^2 \theta_{23} = 0.45$, scan the fit $\sin^2 \theta_{23}$ from $0.5$ to $0.6$ in steps of $0.005$ (or larger steps, if too slow), and minimize over the other parameters for each $\sin^2 \theta_{23}$. Choose the minimum $\chi^2$ from the $\theta_{23}$ scan. Make sure to cover a large enough range for the $\theta_{23}$ scan in the wrong octant. Using this procedure, it is obvious that one cannot end up in the unwanted octant. As a result, you should find the dashed curve in the above figure (for a stepsize of $0.005$). However, as you can check, the result somewhat depends on the grid width/stepsize (which explains the discrepancy for the points ending up in the correct octant). 
For a precision calculation, many steps in $\theta_{23}$ are required (or interpolation techniques in the $\theta_{23}$ direction). In addition, the procedure is extremely slow.

\aufg{6} {\bf Using Schwetz-priors} (requires GLoBES 3.0 or higher!)

Here we will use a more advanced and very fast approach to the last problem, which is named
after its inventor Thomas Schwetz. It requires GLoBES 3.0 or higher and uses user-defined
priors. However, you will not need any particular knowledge about the implementation: The
example {\tt deg\_tut\_6.c} already defines a standard prior for you. You will only need
to modify the prior where indicated to include the solution.

\vspace*{3mm}

\underline{Possible solution:}
Add a very large value to the $\chi^2$ if the minimizer ends up in the wrong octant. This can be
done with a user-defined prior, which is a function of the oscillation parameters only to be added to the $\chi^2$ {\em before} the oscillation parameter marginalization.  Modify the already implemented user-defined prior function where marked in the code.  Use the {\tt central\_values} to find out
which the right and wrong octant is. Use {\tt in} to extract the tested $\theta_{23}$ and compare it
to the one from {\tt central\_values}. If in the unwanted octant, add a very large number to the $\chi^2$.

As a result, you should find the following dashed-dotted curve with triangles:
\begin{center}
\includegraphics[height=5cm]{problem6}
\end{center}
As you can easily see, the procedure reproduces the correct solutions from the ``straightforward'' curve, but it does not fall in the unwanted octant in the few points close to the vertical line.

\subsubsection*{Advanced topic: Eight-fold degeneracy}

%\aufg{8} {\bf Pitfalls}

\aufg{7} {\bf Using iso-rate curves} (homework: requires extensive programming!)

In practice, one wants to have a stable and reliable tool for discrete degeneracy localization.
Unfortunately, such a tool depends on the topology and therefore on the experiment.
A typical very sophisticated experiment with all degeneracies being relevant, is a neutrino factory at a baseline of about $2 \, 000 \, \mathrm{km}$ to $4 \, 000 \, \mathrm{km}$. In the literature, the degeneracies
are defined by equal probabilities for both neutrinos and antineutrinos
\begin{equation}
P_{e \mu} (\boldsymbol{\eta_\mathrm{best-fit}}) = P_{e \mu} (\boldsymbol{\eta'}) \, , \quad  P_{\bar{e} \bar{\mu}}(\boldsymbol{\eta_\mathrm{best-fit}}) = P_{\bar{e} \bar{\mu}}(\boldsymbol{\eta'})\, 
\end{equation}
at an oscillation parameter point $\boldsymbol{\eta'}$ and the best-fit value $\boldsymbol{\eta_{\mathrm{best-fit}}}$ (see, {\em e.g.}, Ref.~\cite{Barger:2001yr}).
In principle, one can use analytical knowledge to find all such points $\boldsymbol{\eta'}$ corresponding to the different discrete degeneracies. However, for an experiment using spectral information, the dominating effect will we a weighted average over different energies including the convolution with cross sections, flux, {\em etc.} Therefore, in practice, two different methods turn out to be very efficient:  
\begin{enumerate}
\item
 Winter's method: Use {\tt glbTotalRuleRate} to obtain the total event rates for the neutrino and antineutrino appearance
channels. Plot the curves with equal total rate in the $\sin^2 2 \theta_{13}$-$\delta_{\mathrm{CP}}$-plane for both neutrinos and antineutrinos using the same rates as in the best-fit point. The curves will intersect at the intrinsic degeneracy (if it all). Plot the same curves for the same rates with $\mathrm{sgn}(\Delta m_{31}^2)$ flipped. Again you will find a maximum of two intersection points. Now do the same for $\pi/2 - \theta_{23}$ flipped, and for  $\mathrm{sgn}(\Delta m_{31}^2)$ combined with $\pi/2 - \theta_{23}$ flipped (mixed degeneracy). You will find at most two more intersection points in each case.
The results should look somewhat like this, where the best-fit point is not marked (the marks correspond to the points where the discrete degeneracies are located according to this specific algorithm):
\begin{center}
\includegraphics[width=0.95\textwidth]{ircurves}
\end{center}
 Altogether, there is a maximum of eight intersection points in the $\sin^2 2 \theta_{13}$-$\delta_{\mathrm{CP}}$-plane, one of which is the best-fit point. These points can be using as starting points of the minimizer to locate the eight-fold degeneracy.
\item
 Huber's method:
 Pre-scan the $\chi^2$ in the $\sin^2 2 \theta_{13}$-$\delta_{\mathrm{CP}}$-plane for the cases described in~a). Find the local minimum which is separated from the best-fit solution. This method is similar to a), but it uses the $\chi^2$ instead of the total event rate. 
\end{enumerate}

\bibliographystyle{apsrev}
\bibliography{references}

\end{document}

### 2007-01-19 Documentation/Manual.tex
*** empty log message ***

@@
 one hand, it contains a comprehensive abstract experiment definition
-language (AEDL), which allows to describe most classes 
+language (\AEDL), which allows to describe most classes 
 of long baseline experiments
@@
 time dependence of the neutrino source is physically only important to suppress backgrounds. 
+Furthermore, experiments with discrete numbers of sources and detectors can
+be implemented by user-defined systematics in \GLOBES\ 3.0 and higher.
 
@@
 degenerate solutions.
+
+In the newest version \GLOBES\ 3.0 flexibility is introduced at all levels. At the probability level,
+the transition probabilities can be modified to introduce new physics. At the systematics level,
+user-defined systematical errors and correlations between sources or detectors can be simulated,
+and at the analysis level, arbitrary  input from external measurements can be added.
+Therefore, \GLOBES\ now provides solutions for new classes of problems.
 
@@
 When using the built-in Earth matter density profile, the 
-original source is \Ref~\cite{Stacey}.
+original source is \Refs~\cite{Dziewonski:1981xy,Stacey}.
 
 All files ending with \verb+.dat+ or \verb+.glb+ 
-in the \verb+data+ subdirectory of the \GLOBES\ tar-ball have on top a comment field which clearly indicates which works should be
+in the \verb+data+ subdirectory of the \GLOBES\ tar-ball have on top a comment field which clearly indicates which studies should be
 cited when using a certain file. Make sure that dependencies are correctly
@@
 checked, too (for example, cross section or flux files). One can use 
-the \verb+-v3+ option to \verb+globes+ to see which files
-are included.
-
+the \verb+-v3+ option to the \verb+globes+ command to see which files
+are included (\cf, \Chapt~\ref{chap:exp_def}).
 It is recommended that you use the same style for your own input files, since then, in case they are distributed, everybody will know how to correctly reference your work.
@@
  \item
-  User-defined systematics, which can be used to simulate reactor experiments \etc ; see \Sec~\ref{sec:userchi}
+  User-defined systematics, which can be used to simulate reactor experiments \etc ; see \Secs~\ref{sec:userchi}
 and~\ref{sec:rules}
@@
 \item
-  Updated AEDL files; see \Tab~\ref{tab:experiments}
+  Updated \AEDL\ files; see \Tab~\ref{tab:experiments}
 \item
@@
 Most of the modifications should not require that old pieces of software be changed. 
-However, the following changes are relevant:
+However, the following changes could be relevant:
 \begin{itemize}
 \item
- AEDL requires now that {\tt \$version} be used to define the minimum version number this AEDL
-file is to be used with. With this requirement one can easily avoid that new AEDL
+ \GLB{SetDensityParams} has to be used with \GLB{DefineParams} and \GLB{SetDensityProjectionFlag} together
+with \GLB{DefineProjection}, because unexpected pre-defined behavior should be avoided.
+\item
+ Functions \GLB{SetFilter}, \GLB{GetFilter}, \GLB{SetFilterState}, and \GLB{GetFilterState} replaced
+by functions {\tt ...InExperiment}.
+\item
+ \AEDL\ requires now that {\tt \$version} be used to define the minimum version number this \AEDL\
+file is to be used with. With this requirement one can easily avoid that new \AEDL\
 files with new features be used with old versions of \GLOBES\ which may not recognize these features.
 \item
- Some of the earlier AEDL files have been updated, changed names, or have been removed.
+ Some of the earlier \AEDL\ files have been updated, changed names, or have been removed.
 In addition, new files have been added.
-Although old AEDL files will run as usual for compatibility, they will not be
-supported by the \GLOBES\ team anymore. You should make sure to keep these files when updating \GLOBES . 
+Although old \AEDL\ files will run as usual for compatibility, they will not be
+supported by the \GLOBES\ Team anymore. You should make sure to keep these files when updating \GLOBES . 
+\item 
+ The implementation of the tilt (systematics) has slightly changed. The tilt also works for
+variable bin widths. Therefore, you will obtain slightly different results when you run
+the same \AEDL\ between older and newer versions of \GLOBES .
 \end{itemize}
@@
  from the respective directory in the \GLOBES\ software package.
+The corresponding figures are produced with the
+Mathematica Notebook {\tt DocPlots.nb}, which can be found
+in the example directory as well.
 
@@
 in the last chapter of \Part~\ref{part:2}.
+
+In this version of the manual, introductory topics and advanced topics are
+mixed if they belong to the same subject. Therefore, we have marked more
+advanced material by a star~($^*$). This material can be skipped in a first
+reading of the manual. In some cases, it may be even recommendable to do
+so because knowledge of \AEDL\ is required (which is introduced in
+the second part).
 
@@
 
-This work and the development of \GLOBES\ have over the years been supported by:
+This work and the development of \GLOBES\ have over the years been supported by
+(in chronological order):
 \begin{itemize}
@@
 
-\printindex{aedl}{AEDL reference}
+\printindex{aedl}{\AEDL\ reference}
 

### 2007-01-19 Documentation/appendix.tex
*** empty log message ***

@@
 \label{app:aedlfiles}
+%\index{norm}{glb files@{\tt glb}-files}
+
 Along with the \GLOBES\ package comes a catalogue of pre-defined experiment \AEDL\ files for different
@@
 \begin{tabular}{lllllll}
+\hline 
 Experiment & File & Runtime & Power & Baseline & Detector & Mass \\ \hline \hline
-T2K & {\tt T2K.glb} & 2~yr $\nu$/6~yr $\bar{\nu}$ & 0.77~MW & 295~km & WC & 22.5~kt \\ \hline
-T2HK & {\tt T2HK.glb} & 4~yr $\nu$/4~yr $\bar{\nu}$ & 4~MW & 295~km & WC & 500~kt \\ \hline
-NO$\nu$A & {\tt NOvA.glb} & 3~yr $\nu$/3~yr $\bar{\nu}$ & 1.12~MW & 812~km & TASD & 25~kt \\ \hline
+T2K & {\tt T2K.glb} & 2~yr $\nu$/6~yr $\bar{\nu}$ & 0.77~MW & 295~km & WC & 22.5~kt \\ 
+T2HK & {\tt T2HK.glb} & 4~yr $\nu$/4~yr $\bar{\nu}$ & 4~MW & 295~km & WC & 500~kt \\ 
+NO$\nu$A & {\tt NOvA.glb} & 3~yr $\nu$/3~yr $\bar{\nu}$ & 1.12~MW & 812~km & TASD & 25~kt \\
 SPL & {\tt SPL.glb} & 2~yr $\nu$/8~yr $\bar{\nu}$ & 4~MW & 130~km & WC & 500~kt \\ 
+\hline
 \end{tabular}
@@
 \subsection*{T2K -- {\tt T2K.glb}}
+\index{aedl}{T2K.glb@{\tt T2K.glb}}
 
@@
 systematics function for the CC sample is changed to {\tt chiNoSysTotalRates}, but the systematics function for the QE
-sample stays {\tt chiSpectrumTilt}. Note, that otherwise the free normalization would be swiched off and all events from
+sample stays {\tt chiSpectrumTilt}. Note, that otherwise the free normalization would be switched off and all events from
 the QE sample would be counted twice. A detailed discussion of this QE/CC sample splitting can be found in~\cite{Huber:2002mx}. 
@@
 \subsection*{T2HK -- {\tt T2HK.glb}}
+\index{aedl}{T2HK.glb@{\tt T2HK.glb}}
 
@@
 \subsection*{NO$\nu$A -- {\tt NOvA.glb}}
-
-The \NOVA\ experiment can be simulated with the file {\tt NOvA.glb}. The description of the disapperance 
+\index{aedl}{NOvA.glb@{\tt NOvA.glb}}
+
+The \NOVA\ experiment can be simulated with the file {\tt NOvA.glb}. The description of the disappearance 
 channels is taken from \cite{Yang_2004} and the description of the appearance channels follows 
@@
 \ei
-The \NOVA\ experiment uses a neutrino beam from the Fermilab NuMI beamline. The target power is 
-1.12~MV which results in $10^{21}$~pot~$\mathrm{yr^{-1}}$ and 3~years $\nu$-running and 3~years $\bar{\nu}$-running 
-is assumed. The fiducial mass of the Totally Liquid Scintillator Detector (TASD) is taken to be 
+The \NOVA\ experiment uses a neutrino beam from the Fermilab NuMI beamline. The source power is $10^{21}$~pot~$\mathrm{yr^{-1}}$, corresponding to a (naively computed) target power of
+1.12~MW. For the running times, 3~years $\nu$-running and 3~years $\bar{\nu}$-running 
+are assumed. The fiducial mass of the Totally Liquid Scintillator Detector (TASD) is taken to be 
 $\mathrm{m_{det} = 25 \,kt}$ at a baseline of L~=~812~km approximately 12~km off-axis to the beamline. The
@@
 \subsection*{SPL - {\tt SPL.glb}}
+\index{aedl}{SPL.glb@{\tt SPL.glb}}
 
@@
 The \SPL\ experiment uses a neutrino beam from the CERN to Fr\'{e}jus. The target power is 
-4~MV and 2~years $\nu$-running and 8~years $\bar{\nu}$-running 
+4~MW and 2~years $\nu$-running and 8~years $\bar{\nu}$-running 
 is assumed. The fiducial mass of the Water Cerenkov detector is taken to be 
@@
 \begin{tabular}{lllllll} 
+\hline
 Experiment & File & Runtime & th. Power & Baseline & Det. & Mass \\ \hline \hline
-Small & {\tt Reactor1.glb} & 5~yr $\bar{\nu}$ & 4~GW & 1.7~km & LS & 20~t \\ \hline
-Large & {\tt Reactor.glb} & 8~yr $\bar{\nu}$ & 10~GW & 1.7~km & LS & 100~t \\ \hline
-DoubleCHOOZ & {\tt D-Chooz\_near.glb} & 5~yr $\bar{\nu}$ & 2$\times$4.2~GW & 0.1~km & LS & 10.16~t \\ 
+Small & {\tt Reactor1.glb} & 5~yr $\bar{\nu}$ & 4~GW & 1.7~km & LS & 20~t \\ 
+Large & {\tt Reactor.glb} & 8~yr $\bar{\nu}$ & 10~GW & 1.7~km & LS & 100~t \\
+Double Chooz & {\tt D-Chooz\_near.glb} & 5~yr $\bar{\nu}$ & 2$\times$4.2~GW & 0.1~km & LS & 10.16~t \\ 
  & {\tt D-Chooz\_far.glb} & 5~yr $\bar{\nu}$ & 2$\times$4.2~GW & 1.05~km & LS & 10.16~t \\ 
+\hline 
 \end{tabular}
@@
 \subsection*{Small Reactor Experiment -- {\tt Reactor1.glb}}
+\index{aedl}{Reactor1.glb@{\tt Reactor1.glb}}
 
@@
 The neutrino source is the core of a nuclear power reactor. The integrated luminosity is assumed to be
-$\mathrm{\mathcal{L} = 400 \,t \, GW\, yr}$, \eg\ a 20~t detector, a reactor with a thermal power of 4~GW
+$\mathrm{\mathcal{L} = 400 \,t \, GW\, yr}$, such as for a 20~t detector, a reactor with a thermal power of 4~GW,
 and a running period of 5 years. As detector technology, a liquid scintillator 
 detector is assumed, a far detector at a baseline of L~=~1.7~km and a near detector which is
-assumed to be identical to the far detector (maybe apart from the size) in order to minimize the impact of systematical uncertainties. The normalization error
+assumed to be identical to the far detector (maybe apart from the size) in order to minimize the impact of systematical uncertainties.
+The near detector is simulated implicitly by lower effective systematical errors,
+and there are no external backgrounds assumed.
+The normalization error
 used in the file {\tt Reactor1.glb} has to be considered as an effective error, receiving contributions from
-individual uncertainties (see~\cite{Huber:2003pm}). The energy resolution is
+individual uncertainties (see \Ref~\cite{Huber:2003pm}). The energy resolution is
 $\mathrm{\sigma_e=5\%\cdot\sqrt{E_{vis}}}$ and the choice for {\tt @sigma\_function} is {\tt \#inverse\_beta}. The following rules are defined within {\tt Reactor1.glb}: 
@@
 \subsection*{Large Reactor Experiment -- {\tt Reactor2.glb}}
+\index{aedl}{Reactor2.glb@{\tt Reactor2.glb}}
 
@@
 
-\subsection*{DoubleCHOOZ -- {\tt D-Chooz\_near.glb} and {\tt D-Chooz\_far.glb}}
+\subsection*{Double Chooz -- {\tt D-Chooz\_near.glb} and {\tt D-Chooz\_far.glb}}
+\index{aedl}{D-Chooznear.glb@{\tt D-Chooz\_near.glb}}
+\index{aedl}{D-Choozfar.glb@{\tt D-Chooz\_far.glb}}
 
 The files {\tt D-Chooz\_near.glb} and {\tt D-Chooz\_far.glb} allow to simulate the \DC\ reactor experiment in France.
+They require user-defined systematics and \GLOBES\ 3.0 or higher, where the user-defined
+systematics function can be found in the header of  {\tt D-Chooz\_near.glb} or {\tt example5.c}.
 The basic versions of these files were used within \cite{Huber:2006vr}, which should be cited if the files 
@@
 \begin{tabular}{lllllll} 
+\hline 
 Experiment & File & Runtime & $\gamma$ & Baseline & Det. & Mass \\ \hline \hline
-Low $\gamma$ & {\tt BB\_100.glb} & 4~yr $\nu$/4~yr $\bar{\nu}$ & 100 & 130~km & WC & 500~kt \\ \hline
-Medium $\gamma$ & {\tt BB\_350.glb} & 4~yr $\nu$/4~yr $\bar{\nu}$ & 350 & 730~km & WC & 500kt \\  \hline
-Variable & {\tt BBvar\_WC.glb} & 4~yr $\nu$/4~yr $\bar{\nu}$ & variable & variable & WC & 500~kt \\  \hline
-Variable & {\tt BBvar\_TASD.glb} & 4~yr $\nu$/4~yr $\bar{\nu}$ & variable & variable & TASD & 50~kt \\ 
+Low $\gamma$ & {\tt BB\_100.glb} & 4~yr $\nu$/4~yr $\bar{\nu}$ & 100 & 130~km & WC & 500~kt \\ 
+Medium $\gamma$ & {\tt BB\_350.glb} & 4~yr $\nu$/4~yr $\bar{\nu}$ & 350 & 730~km & WC & 500kt \\
+Variable $\gamma$ & {\tt BBvar\_WC.glb} & 4~yr $\nu$/4~yr $\bar{\nu}$ & variable & variable & WC & 500~kt \\  
+Variable $\gamma$ & {\tt BBvar\_TASD.glb} & 4~yr $\nu$/4~yr $\bar{\nu}$ & variable & variable & TASD & 50~kt \\ 
+\hline 
 \end{tabular}
@@
 \subsection*{CERN-Fr\'{e}jus Baseline Scenario -- {\tt BB\_100.glb}}
+\index{aedl}{BB100.glb@{\tt BB\_100.glb}}
 
@@
 \subsection*{Higher Gamma Scenario -- {\tt BB\_350.glb}}
+\index{aedl}{BB350.glb@{\tt BB\_350.glb}}
 
@@
 for $^{18}$Ne and $^6$He) that describes energy smearing. These migration matrices also already include energy
-dependend efficiencies and background rejection factors. They are taken from the appendix of
+dependent efficiencies and background rejection factors. They are taken from the appendix of
 \cite{Burguet-Castell:2005pa}. The following rules are defined within {\tt BB\_350.glb}:
@@
 \subsection*{Variable Beta Beam (Water Cerenkov) -- {\tt BBvar\_WC.glb}}
+\index{aedl}{BBvarWC.glb@{\tt BBvar\_WC.glb}}
 
@@
 \subsection*{Variable Beta Beam (TASD) -- {\tt BBvar\_TASD.glb}}
+\index{aedl}{BBvarTASD.glb@{\tt BBvar\_TASD.glb}}
 
@@
 
+{\small
 \begin{center}
 \begin{tabular}{lllllll}
+\hline 
 Experiment & File & Runtime & $\mathrm{E_\mu}$ & Baseline & Det. & Mass \\ \hline \hline
-Standard & {\tt NFstandard.glb} & \begin{tabular}{l}4~yr $\nu$ \\ 4~yr $\bar{\nu}$\end{tabular} & 50~GeV & 3000~km & \begin{tabular}{l}MID\end{tabular} & \begin{tabular}{l}50~kt\end{tabular} \\ \hline
-Variable & {\tt NFvar.glb} & \begin{tabular}{l}4~yr $\nu$ \\ 4~yr $\bar{\nu}$\end{tabular} & variable & variable & \begin{tabular}{l}MID\end{tabular} & \begin{tabular}{l}50~kt\end{tabular} \\ \hline
-Silver & {\tt NF\_GoldSilver.glb} & \begin{tabular}{l}4~yr $\nu$ \\ 4~yr $\bar{\nu}$\end{tabular} & variable & variable
-&\begin{tabular}{l}MID\\ECC\end{tabular}&\begin{tabular}{l}50~kt\\5~kt\end{tabular}\\ \hline
-Hybrid & {\tt NF\_hR\_lT.glb} & \begin{tabular}{l}4~yr $\nu$ \\ 4~yr $\bar{\nu}$\end{tabular} & variable & variable & \begin{tabular}{l}Hybrid\end{tabular} & \begin{tabular}{l}50~kt\end{tabular} \\ 
-\end{tabular}
-\end{center}
+Standard & {\tt NFstandard.glb} & \begin{tabular}{l}4~yr $\nu$ \\ 4~yr $\bar{\nu}$\end{tabular} & 50~GeV & 3000~km & \begin{tabular}{l}MID\end{tabular} & \begin{tabular}{l}50~kt\end{tabular} \\ 
+Variable $E_\mu$ & {\tt NFvar.glb} & \begin{tabular}{l}4~yr $\nu$ \\ 4~yr $\bar{\nu}$\end{tabular} & variable & variable & \begin{tabular}{l}MID\end{tabular} & \begin{tabular}{l}50~kt\end{tabular} \\ 
+Gold+Silver & {\tt NF\_GoldSilver.glb} & \begin{tabular}{l}4~yr $\nu$ \\ 4~yr $\bar{\nu}$\end{tabular} & variable & variable
+&\begin{tabular}{l}MID\\ECC\end{tabular}&\begin{tabular}{l}50~kt\\5~kt\end{tabular}\\
+Hybrid det.
+& {\tt NF\_hR\_lT.glb} & \begin{tabular}{l}4~yr $\nu$ \\ 4~yr $\bar{\nu}$\end{tabular} & variable & variable & \begin{tabular}{l}Hybrid\end{tabular} & \begin{tabular}{l}50~kt\end{tabular} \\
+\hline
+\end{tabular}
+\end{center}
+}
 
 \subsection*{Standard Neutrino Factory -- {\tt NFstandard.glb}}
+\index{aedl}{NFstandard.glb@{\tt NFstandard.glb}}
 
 A standard neutrino factory scenario can be simulated with the file {\tt NFstandard.glb}. The basic version was
-used within \cite{Huber:2002mx} (NuFact-II scenario). This reference should be cited if the file {\tt
+used within \cite{Huber:2002mx} (NuFact-II scenario, but some changes in normalization errors). 
+This reference should be cited if the file {\tt
 NFstandard.glb} is used for a scientific publication or a talk. For calculations that involve {\tt
@@
 
-\subsection*{Variable Feutrino Factory -- {\tt NFvar.glb}}
+\subsection*{Variable $\boldsymbol{E_\mu}$ Neutrino Factory -- {\tt NFvar.glb}}
+\index{aedl}{NFvar.glb@{\tt NFvar.glb}}
 
@@
 \bi
-\item {\tt emax} (parent energy of the stored muons $\left[\mathrm{GeV}\right]$)
-\item {\tt BASELINE} (experiment baseline $\left[\mathrm{km}\right]$)
+\item {\tt emax} (parent energy of the stored muons $\left[\mathrm{km}\right]$)
+\item {\tt BASELINE} (experiment baseline $\left[\mathrm{GeV}\right]$)
 \ei
@@
 $\mathrm{E_\mu =}$~{\tt emax}. The parent energy of the muons can be appropriately set in the range
-$\mathrm{10\,GeV\lesssim E_\mu\lesssim 80\, GeV}$ (technically it is possible to arrange {\tt emax} for $\mathrm{E_\mu=5\,
-GeV}$). The baseline of the scenario is L~=~{\tt BASELINE} and has to be chosen within the interval $\mathrm{1\,km\lesssim\, L \, \lesssim\, 2\cdot R_{EARTH}}$. 
+$\mathrm{10\,GeV\lesssim E_\mu\lesssim 80\, GeV}$.
+The baseline of the scenario is L~=~{\tt BASELINE} and has to be chosen within the interval $\mathrm{1\,km\lesssim\, L \, \lesssim\, 2\cdot R_{EARTH}}$.
 Besides these settings the other attributes of {\tt NFvar.glb} are similar to the ones from {\tt
-NFstandard.glb}. Only the treatment of the disappearance channels is different. Whereas the CID threshold is
-also applied in the disappearance channels of {\tt NFstandard.glb}, this is not the case for the disapperance
-channel definition within {\tt NFvar.glb}, where no CID is applied, and a threshold similar to the threshold of
-the MINOS experiment \cite{Ables:1995wq} is applied. The following rules are defined within
+NFstandard.glb}.
+Only the treatment of the disappearance channels is different.
+In the disappearance channels of {\tt NFstandard.glb}, the charge identification is used to reduce the background from the appearance neutrinos.
+In {\tt NFvar.glb}, however, a threshold similar to the MINOS experiment~\cite{Ables:1995wq} is applied and the appearance and disappearance rates are assumed to be indistinguishable for the disappearance channel.
+The following rules are defined within
 {\tt NFvar.glb}:
@@
 \subsection*{Variable Feutrino Factory with Silver Channel -- \\{\tt NF\_GoldSilver.glb}}
+\index{aedl}{NFGoldSilver.glb@{\tt NF\_GoldSilver.glb}}
 
@@
 \subsection*{High Resolution/Low Threshold Neutrino Factory -- \\{\tt NF\_hR\_lT.glb}}
+\index{aedl}{NFhRlT.glb@{\tt NF\_hR\_lT.glb}}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Flux normalization in \GLOBES }
+\chapter{Flux normalization in \GLOBES $^*$ }
 \index{norm}{Normalization of fluxes}
@@
 correspondence with the number of neutrinos produced. Another sensible
-unit could be {\it megawatt on target (MW)}, again this number is directly
+unit could be {\it megawatts on target (MW)}, again this number is directly
 correlated with the number of neutrinos and moreover there is a unique

### 2007-01-19 Documentation/part1.tex
*** empty log message ***

@@
  fprintf(stream,"$\backslash$nOscillation probabilities in matter: ");\\
- for(i=1;i<4;i++) fprintf(stream,"1->\%i: \%g ",i,\\ \hspace*{4cm} glbProfileProbability(0,1,i,+1,50)); \\
+ for(i=1;i<4;i++) fprintf(stream,"1->\%i: \%g ",i,\\ \hspace*{4cm} glbProfileProbability(0,1,i,+1,50)); 
 }
@@
  input\_errors = glbAllocParams(); \\
- minimum = glbAllocParams(); \\
+ minimum = glbAllocParams(); 
 }
@@
 at the short baseline and test if it is still present in the combination of the
-experiments.
+two baselines:
 \gq{
@@
   chi2=glbChiAll(minimum,minimum,GLB\_ALL); \\
-  fprintf(stream,"$\backslash$nchi2 for combination at minimum of Exp. 1:\\ \hspace*{0.5cm} \%g $\backslash$n",chi2); \\
+  fprintf(stream,"$\backslash$nchi2 for combination at minimum of L=3000km:\\ \hspace*{0.5cm} \%g $\backslash$n",chi2); \\
   glbPrintParams(stream,minimum);  \\
@@
 \\
-chi2 for combination at minimum of Exp. 1: 59.4879 \\
+chi2 for combination at minimum of L=3000km: 59.4879 \\
 0.599908 0.0172467 0.768342 1.39186 8.198e-05 -0.00189508 \\ 
@@
 definition. The pre-defined experiment prototypes in the {\tt data}
-subdirectory are summarized in \tabl{experiments}. They correspond
+subdirectory are summarized in \tabl{experiments} and described in
+\App~\ref{app:aedlfiles}. They correspond
 (except from minor modifications) to the experiments in the
@@
 \begin{center}
-\begin{tabular}{llp{7.5cm}c}
+{\small
+\begin{tabular}{llp{7.1cm}c}
 \hline
-Experiment & Filename & Short description &  Ref. \\[0.2cm]
+Experiment & Filename & Short description &  \Ref \\
 \hline 
@@
 
-Double{\sc Chooz} & {\tt D-Chooz\_near.glb} & DoubleCHOOZ near 
+Double{\sc Chooz} & {\tt D-Chooz\_near.glb} & Double Chooz near 
 detector, 5 years data taking & \cite{Huber:2006vr} \\ 
- & {\tt D-Chooz\_far.glb} & DoubleCHOOZ near detector, 5 years data taking & \\[0.2cm]
+ & {\tt D-Chooz\_far.glb} & Double Chooz far detector, 5 years data taking & \\[0.2cm]
 
@@
 
-low $\gamma$ & {\tt BB\_100.glb} & $\gamma=100$ CERN to Fr\'{e}jus baseline 
+Low $\gamma$ & {\tt BB\_100.glb} & $\gamma=100$ CERN to Fr\'{e}jus baseline 
 scenario, 4~yr $\nu$- and 4~yr $\bar{\nu}$-running & \cite{Campagne:2006yx} \\[0.2cm]
 
-medium $\gamma$ & {\tt BB\_350.glb} & $\gamma=350$, refurbished SPS scenario, 4~yr $\nu$- and
+Medium $\gamma$ & {\tt BB\_350.glb} & $\gamma=350$, refurbished SPS scenario, 4~yr $\nu$- and
 4~yr $\bar{\nu}$-running & \cite{Burguet-Castell:2005pa} \\[0.2cm]
 
-Variable & {\tt BBvar\_WC.glb} & Variable $\beta$-beam, 4~yr $\nu$- and 
+Variable $\gamma$ & {\tt BBvar\_WC.glb} & Variable $\beta$-beam, 4~yr $\nu$- and 
 4~yr $\bar{\nu}$-running, {\sf AEDL}-Variables {\tt gammafactor}, 
@@
 
-	& {\tt BBvar\_TASD.glb} & like {\tt BBvar\_WC.glb} but with a 50~kt TASD detector  
+	& {\tt BBvar\_TASD.glb} & As {\tt BBvar\_WC.glb}, but with a 50~kt TASD
 & \cite{Huber:2005jk} \\[0.2cm]
@@
 
-Variable  & {\tt NFvar.glb} & Variable neutrino factory, 4~yr 
-$\nu$- and 4~yr $\bar{\nu}$-running, {\sf AEDL}-Variables {\tt emax} and {\tt BASELINE}, disappearance channels
+Variable $E_\mu$ & {\tt NFvar.glb} & Variable $E_\mu$ neutrino factory, 4~yr 
+$\nu$- and 4~yr $\bar{\nu}$-running, {\sf AEDL}-Variables {\tt emax} and {\tt BASELINE}, disapp. channels
 {\em without} CID & \cite{Huber:2002mx,Huber:2006wb}  \\[0.2cm]
 
-Gold + Silver & {\tt NF\_GoldSilver.glb} & as {\tt NFvar.glb} 
+Gold + Silver & {\tt NF\_GoldSilver.glb} & As {\tt NFvar.glb}, 
 plus 5~kt ECC detector for Silver Channel
@@
 
-Hybrid detector & {\tt NF\_hR\_lT.glb} & like {\tt NFvar.glb} 
-with lower threshold and higher energy resolution & \cite{Huber:2002mx,Huber:2006wb} \\[0.2cm]
+Hybrid det. & {\tt NF\_hR\_lT.glb} & As {\tt NFvar.glb}, 
+but lower threshold and higher energy resolution & \cite{Huber:2002mx,Huber:2006wb} \\[0.2cm]
 
@@
 \end{tabular}
+} % small
 \end{center}
-\mycaption{\label{tab:experiments} \index{norm}{Experiment files (table)}
+\mycaption{\label{tab:experiments} 
+%
+\index{norm}{Experiment files (table)}
 \index{norm}{glb files@{\tt glb}-files} 
+% DOES NOT WORK. FOR SOME REASON, THERE ARE DOUBLE ENTRIES IN THE INDEX, IF THE FOLLOWING CODE IS USED [WW]:
+% \index{aedl}{T2K.glb@{\tt T2K.glb}}
+% \index{aedl}{T2HK.glb@{\tt T2HK.glb}}
+% \index{aedl}{NOvA.glb@{\tt NOvA.glb}}
+% \index{aedl}{SPL.glb@{\tt SPL.glb}}
+% \index{aedl}{Reactor1.glb@{\tt Reactor1.glb}}
+% \index{aedl}{Reactor2.glb@{\tt Reactor2.glb}}
+% \index{aedl}{D-Chooznear.glb@{\tt D-Chooz\_near.glb}}
+% \index{aedl}{D-Choozfar.glb@{\tt D-Chooz\_far.glb}}
+% \index{aedl}{BB100.glb@{\tt BB\_100.glb}}
+% \index{aedl}{BB350.glb@{\tt BB\_350.glb}}
+% \index{aedl}{BBvarWC.glb@{\tt BBvar\_WC.glb}}
+% \index{aedl}{BBvarTASD.glb@{\tt BBvar\_TASD.glb}}
+% \index{aedl}{NFstandard.glb@{\tt NFstandard.glb}}
+% \index{aedl}{NFvar.glb@{\tt NFvar.glb}}
+% \index{aedl}{NFGoldSilver.glb@{\tt NF\_GoldSilver.glb}}
+% \index{aedl}{NFhRlT.glb@{\tt NF\_hR\_lT.glb}}
+%
 Pre-defined experiment prototypes, their filenames (to be used in \GLB{InitExperiment}), 
@@
 discussed (except from minor modifications, such as a different implementation of 
-the energy threshold function). Note that some of these experiments are outdated 
-in terms of integrated luminosities, baseline, fluxes, efficiencies, or other factors. 
-In any case these file are installed along with \GLOBES.}
+the energy threshold function). Note that some of these experiments may want to be
+adjusted in terms of integrated luminosity, baseline, flux, efficiencies, or other factors. 
+In any case these file are installed along with \GLOBES . For more details, see \App~\ref{app:aedlfiles}.}
 \end{table}
@@
 In principle, the \GLOBES\ user interface can currently handle up to 32 of different long-baseline experiments simultaneously, where the number
-of existing experiment definition files can, of course, be unlimited.
-This means that their $\Delta \chi^2$-values are added {\em after} the minimization over the systematics parameters, and {\em before} any minimization over the oscillation parameters.
-Note that each experiment
+of existing experiment definition files can, of course, be unlimited.  Note that each experiment
 assumes a specific matter density profile, which means
@@
 \index{norm}{Path resolution}
-A list of pre-defined experiment prototypes, their filenames, their short descriptions, and the references of their definitions can be found in \tabl{experiments}.
+A list of pre-defined experiment prototypes, their filenames, their short descriptions, and the references of their definitions can be found in \tabl{experiments}; for more details, see \App~\ref{app:aedlfiles}.
 If the program cannot find these files, or some of them are syntactically not correct, it will break at this place.
@@
 Compared to existing experiments, which use real data, future experiments
-use simulated data. Thus, the {\em true parameter values} and their results in 
-form of the reference event rate vectors are simulated. After setting the 
+use simulated data. Thus, the {\em true parameter values} (or simulated
+parameter values) are used to calculate the reference event rate vectors
+corresponding to the simulated experiment result. After setting the 
 true parameter values, the {\em fit parameter values} can be varied in order to 
@@
 \item
- The program and AEDL file versions are compatible.
+ The program and \AEDL\ file versions are compatible.
 \end{itemize}
-However, note that GLOBES~3.0 and higher requires that the \GLOBES\ version be at least as new as the
-version the AEDL file was written for.
+However, note that \GLOBES ~3.0 and higher requires that the \GLOBES\ version be at least as new as the
+version the \AEDL\ file was written for.
 %
@@
 the version string of the experiment number {\tt experiment} (set by
-{\tt \$version} in AEDL).
+{\tt \$version} in \AEDL ).
 The version string is allocated within the experiment structure, which means
@@
 
-Calculating a $\chi^2$-value with or without systematics, but no correlations and degeneracies, is the simplest and fastest possibility to obtain high-level information on an experiment.
-In general, \GLOBES\ uses the six independent oscillation parameters $\theta_{12}$, $\theta_{13}$, $\theta_{23}$, $\deltacp$, $\sdm$, $\ldm$, as well as the matter density
-scaling factor $\hat{\rho}$ of each experiment.\index{norm}{Matter density!
-scaling factor} Thus, there are six plus the number of experiments parameters determining the rate vectors.
-Using the matter density scaling factors  in addition to the oscillation parameters will allow the simulation of the correlations with matter density uncertainties: In this
-approach, the matter density profile normalization $\hat{\rho}$ can 
-be treated
-as parameter to be measured by the experiment, where an external precision
-given by observations is imposed (typically up to $5\%$). 
-\index{norm}{Matter density!uncertainty}
- For this section, it is important to keep in mind
-that there are more parameters than just the oscillation parameters
-determining the simple $\chi^2$. However, as we have described in the
-last section, the mechanism for the matter density scaling factors
-is hidden in the definition of \GLB{\_params}: Each of the scaling
-factors is initially set to $1.0$. Therefore, for the calculation of
-$\chi^2$ with systematics only, we do not have to care about the
-matter density scaling factors.
+Calculating a $\chi^2$-value including systematics, but without correlations and degeneracies, is the simplest and fastest possibility to obtain high-level information on an experiment.
+Here we describe the use of built-in
+systematics, and the rather advanced topic of user-defined systematics. Note that the matter density 
+is treated as an oscillation parameter in \GLOBES , which means that it is not dealt with at the
+systematics level.
 
@@
 \begin{center}
-\colorbox{white}{\includegraphics[width=8cm]{correx}}
+\colorbox{white}{\includegraphics[width=7.7cm]{correx}}
 \end{center}
@@
 corresponds to the sum of all of the individual {\tt glbChiSys} calls. 
-This equality will not hold for the minimizers in the next sections anymore. 
+This equality will not hold for the minimizers in the next chapters anymore. 
  An example how to use  {\tt glbChiSys} can be found on page~\pageref{ex:corrth13dcp}.
@@
 The treatment of systematics in \GLOBES\ is performed by the so-called
-{\em pull method} with the help of auxiliary systematics parameters.
+{\em pull method} with the help of nuisance systematics parameters.
 They are taken to be completely uncorrelated among different rules,
@@
  One example for a systematics parameter is the signal normalization error, \ie, an error on the overall normalization of the signal.
-For illustration, we assume that the signal event rate in the $i$th bin $s_i^0$ of one oscillation channel is altered by the overall normalization auxiliary parameter\index{norm}{Auxiliary parameter} of this channel, \ie ,
+For illustration, we assume that the signal event rate in the $i$th bin $s_i^0$ of one oscillation channel is altered by the overall normalization nuisance parameter\index{norm}{Nuisance parameter} of this channel, \ie ,
 \be
@@
 The total number of events in the $i$th bin $x_i$ also includes the background event rates $b_i$, \ie, $x_i = s_i + b_i$, which may have their own systematics parameters.
-In order to implement an overall signal normalization error $\sigma_{n_s}$,  the $\chi^2$, which includes all event rates $x_i$ of all bins, is minimized over the auxiliary parameter $n_s$:
+In order to implement an overall signal normalization error $\sigma_{n_s}$,  the $\chi^2$, which includes all event rates $x_i$ of all bins, is minimized over the nuisance parameter $n_s$:
 \be
@@
 \ee 
-This minimization is done independently for all auxiliary parameters of the rule.
+This minimization is done independently for all nuisance parameters of the rule.
 The total $\chi^2$ for the considered experiment is finally obtained by repeating this procedure for all rules and adding their $\chi^2$-values.
@@
 
-\section{User-defined systematics calculation}
+\section{User-defined systematics calculation$^*$}
 \index{norm}{Systematics!user-defined}
@@
 to override the default $\chi^2$ calculation. The reader who wants to use this
-feature should be familiar with \Secs~\ref{sec:event_rates},~\ref{sec:channel}, and~\ref{sec:rules}
+feature should be familiar with \Secs~\ref{sec:event_rates},~\ref{sec:systematics},~\ref{sec:channel}, and~\ref{sec:rules}
 of the manual. Note that this is an advanced topic, which requires \GLOBES\ 3.0 or
@@
 User-defined systematics is, compared to the built-in systematics, a compound between
-the AEDL definition and some source code defining the behavior. In the simplest case,
-the AEDL file defines user-defined systematics for a specific rule, gives it a 
-name/identifier, and defines the systematical errors. The source code has then to match this AEDL
+the \AEDL\ definition and some source code defining the behavior. In the simplest case,
+the \AEDL\ file defines user-defined systematics for a specific rule, gives it a 
+name/identifier, and defines the systematical errors. The source code has then to match this \AEDL\
 definition -- therefore, make sure to use only unique identifiers. 
@@
 {\tt double (*glb\_chi\_function)(int exp, int rule, int dim, double *params, double *errors, void* user\_data)}.
-In this function type, {\tt exp} is the experiment number, {\tt rule} is the rule number, {\tt dim} is the number of systematics parameters, {\tt params} is an array of the systematics auxiliary parameters themselves, and {\tt errors} is an array with the systematical errors.
+In this function type, {\tt exp} is the experiment number, {\tt rule} is the rule number, {\tt dim} is the number of systematics parameters, {\tt params} is an array of the systematics nuisance parameters themselves, and {\tt errors} is an array with the systematical errors.
 The parameter {\tt user\_data} is set as defined with
@@
 A typical implementation may look like the following code, where ``{\tt chiDCNorm}'' is the identifier
-assigned in the AEDL definition:
+assigned in the \AEDL\ definition:
 \begin{quote}
@@
 \end{quote}
-During running time, it is also possible to change the systematics of an experiment or
-rule (as compared to the systematics assigned in the AEDL file or earlier) with the following function:
-\begin{function}
-\GLBNS{SetChiFunction}
-{\tt int glbSetChiFunction(int exp, int rule, int on\_off, const char *name, double *errors)}
- tells \GLOBES\ to use the registered user-defined systematics identified by the string {\tt name}
-instead of the built-in systematics
-to calculate $\chi^2$ for the experiment {\tt exp} and the rule {\tt rule}. 
-Both of the parameters {\tt exp} and {\tt rule} can take the value \GLBC{GLB\_ALL} to specify that the
-given systematics function should be used for each experiment or each rule. The parameter 
-{\tt on\_off} determines if the systematics function should be used when systematical errors
-are switched on ({\tt GLB\_ON}), or when they are switched off ({\tt GLB\_OFF}).
-The array {\tt errors} sets the systematical errors in the order in which they
-are expected by the systematics function (indices run from $0$ to
-the number of systematics parameters-1). The function returns zero, if successful.
-\end{function}
-Note that new user-defined systematics functions have to be registered with {\tt glbDefineChiFunction} first. 
-One can also request the systematics function name by 
-\begin{function}
-\GLBNS{GetChiFunction}
-{\tt int glbGetChiFunction(int exp, int rule, int on\_off, char *sys\_id, size\_t max\_len)}
-returns the name of the systematics $\chi^2$ function of a given experiment {\tt exp} and rule {\tt rule} for
-systematics on or off as given by {\tt on\_off}. The name is copied to the string
-{\tt sys\_id}, the maximum length of which is specified by {\tt max\_len}. If
-{\tt max\_len} is too small, or if any other error occurs, the return value is $< 0$.
-\end{function}
-Similarly, a pointer to the systematics function can be returned either by name, or by experiment and rule selection:
+
+During running time, the systematics can be changed with 
+\GLB{SetChiFunction} and \GLB{GetChiFunction} as described in \Sec~\ref{sec:systematics}.
+In addition, especially useful in the context of user-defined systematics, a pointer to the systematics function can be returned either by name, or by experiment and rule selection:
 \begin{function}
@@
 
-The $\chi^2$ function {\tt glb\_chi\_function} may use a number of helper functions for
+The user-defined $\chi^2$ function of type {\tt glb\_chi\_function} may use a number of helper functions for
 for the $\chi^2$ calculation. 
@@
 For a setup with near and far detectors,
-the formula for $\chi^2$ is
+the Gaussian formula for $\chi^2$ is
 \begin{align}
-  \chi^2 =& \sum_{i=1}^{\textrm{\# of bins}} \sum_{d = N,F}2
-     \left( O_{d,i} - (1 + a_R + a_d) T_{d,i} +
-       (1 + a_R + a_d) T_{d,i} \log \frac{(1 + a_R + a_d) T_{d,i}}{O_{d,i}}  \right) \nonumber\\
-         & \ \ + \frac{a_R^2}{\sigma_R^2} + \frac{a_N^2}{\sigma_N^2} + \frac{a_F^2}{\sigma_F^2}
+  \chi^2 =& \sum_{i=1}^{\textrm{\# of bins}} \sum_{d = N,F} 
+     \frac{\left( O_{d,i} - (1 + a_R + a_d) T_{d,i} \right)^2 }{O_{d,i}} + \frac{a_R^2}{\sigma_R^2} + \frac{a_N^2}{\sigma_N^2} + \frac{a_F^2}{\sigma_F^2}
 \label{equ:chi2reactor}
 \end{align}
+% WAS POSSONIAN
+% \begin{align}
+%   \chi^2 =& \sum_{i=1}^{\textrm{\# of bins}} \sum_{d = N,F} 2
+%      \left( O_{d,i} - (1 + a_R + a_d) T_{d,i} +
+%        (1 + a_R + a_d) T_{d,i} \log \frac{(1 + a_R + a_d) T_{d,i}}{O_{d,i}}  \right) \nonumber\\
+%          & \ \ + \frac{a_R^2}{\sigma_R^2} + \frac{a_N^2}{\sigma_N^2} + \frac{a_F^2}{\sigma_F^2}
+% \label{equ:chi2reactor}
+% \end{align}
 where $O_{N,i}$ and $O_{F,i}$ are the event rates for the $i$-th bin in the near and
@@
 $T_{d,i}$ are the event rates for the parameter values that are currently being tested.
-$a_R$, $a_N$ and $a_F$ parametrize the small uncertainties in the reactor flux and the fiducial
+$a_R$, $a_N$ and $a_F$ parameterize the small uncertainties in the reactor flux and the fiducial
 mass of the two detectors. In this example, their central values are assumed to be zero,
 while their standard deviations are $\sigma_R$, $\sigma_N$ and $\sigma_F$.
+%
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=10cm]{reactor}
+\end{center}
+\mycaption{\label{fig:reactor} The result from {\tt example5.c} (figure similar to the one
+in \Ref~\cite{Huber:2006vr}). The thick curve corresponds to the systematics on page~\pageref{ex:syst}.}
+\end{figure}
 %
@@
 %
-We use two AEDL files for this experiment, one for the far detector and one for the
+We use two \AEDL\ files for this experiment, one for the far detector and one for the
 near detector.
-The corresponding rule definition ($\nu_e$ disappearance) in the far detector AEDL-file is
+The corresponding rule definition ($\nu_e$ disappearance) in the far detector \AEDL -file is
 \begin{quote}
@@
 off is computed with spectral information, but without systematical errors.
-For the near detector AEDL-file, we have instead
+For the near detector \AEDL -file, we have instead
 \begin{quote}
@@
 %
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=10cm]{reactor}
-\end{center}
-\mycaption{\label{fig:reactor} The result from {\tt example5.c} (figure similar to the one
-in \Ref~\cite{Huber:2006vr}). The thick curve corresponds to the systematics on page~\pageref{ex:syst}.}
-\end{figure}
-
-Further functions useful in the context of user-defined systematics are:
-\begin{function}
-\GLBNS{SetSysErrorsList} 
-{\tt int glbSetSysErrorsList(int exp, int rule, int on\_off, const double *sys\_list)}
-changes the systematical errors defined in the AEDL file in experiment {\tt exp} and 
-rule {\tt rule} to the values in {\tt sys\_list}. This change can be perfomed for
-systematics on or systematics off by using {\tt GLB\_ON} or {\tt GLB\_OFF} for {\tt on\_off}. 
-\end{function}
-\begin{function}
-\GLBNS{GetSysErrorsListPtr} 
-{\tt double* glbGetSysErrorsListPtr(int exp, int rule, int on\_off)}
-returns a pointer to the list of systematical errors as defined in the AEDL file in experiment {\tt exp} and 
-rule {\tt rule}. Choose
-systematics on or systematics off by using {\tt GLB\_ON} or {\tt GLB\_OFF} for {\tt on\_off}. 
-\end{function}
+
 In cases where the systematics minimization does not converge fast enough or ends
@@
 sets the starting values of the local systematics minimizer in experiment {\tt exp} and 
-rule {\tt rule} to the values in {\tt sys\_list}. This change can be perfomed for
+rule {\tt rule} to the values in {\tt sys\_list}. This change can be performed for
 systematics on or systematics off by using {\tt GLB\_ON} or {\tt GLB\_OFF} for {\tt on\_off}. 
@@
 A useful trick is often to use the minimum from the last run as starting values for the next one
-if the input parameters are only slighltly changed. One can obtain the minimum from the last run 
+if the input parameters are only slightly changed. One can obtain the minimum from the last run 
 from the last call of the $\chi^2$ function (\cf, {\tt example5.c}).
@@
 
-\index{norm}{Correlation!multi-parameter}
 This chapter deals with the rather complicated issue of $n$-parameter correlations. It is one of the greatest strengths of this software 
@@
 Therefore, we use a simple step by step introduction to the problem.
+
+\index{norm}{Correlation!multi-parameter}
 
@@
 
-Thus, for each point on the projection axis/plane, one can obtain a result within about $10$ to $30$ seconds on a modern computer, which means that the complete measurement precision for one fixed true parameter set can be obtained in as much as $10$ to $15$ minutes.
+Thus, for each point on the projection axis/plane, one can obtain a result within several seconds on a modern computer, which means that the complete measurement precision for one fixed true parameter set can be obtained in a few minutes.
 One can easily imagine that such a minimizer makes more sophisticated applications possible with the help of overnight calculations, such as showing the dependencies on the true parameter values.
@@
 For example, if the disappearance channels of the experiment determine the leading oscillation parameters with unprecedented
-precisions, one can omit the respective input errors. In \GLOBES , a
+precisions, one can omit the respective input errors. In \GLOBES , an input error
 value of $0$ corresponds\footnote{To be precise, a value for the error
@@
 \item
-Currently, \GLOBES\ only supports Gau\ss ian priors for the individual
+Currently, \GLOBES\ only supports Gau\ss ian pre-defined priors for the individual
 oscillation parameters. Especially for the solar parameters, this is only
  an approximation, since they are imposed on $\theta_{12}$ and not
- on $\sin 2 \theta_{12}$, $\sin 2 \theta_{12} \cdot \sdm$, or $\sin  
- \theta_{12}$. \GLOBES\ 3.0 and higher provide an alternative
-to that by the concept of user-defined priors.
+ on $\ldm$ or $\sin^2  \theta_{12}$. \GLOBES\ 3.0 and higher provide an alternative
+to that by the concept of user-defined priors (\cf, \Sec~\ref{sec:userdefined}).
  \end{enumerate}
@@
   glbDefineProjection(th13\_projection,GLB\_FIXED,GLB\_FIXED,GLB\_FIXED,GLB\_FREE,GLB\_FIXED,GLB\_FIXED); \\
-  glbSetDensityProjectionFlag(th13\_projection,GLB\_FREE,GLB\_ALL); \\
+ % HAS TO BE GLB_FIXED HERE; OTHEREWISE FIG. 4.1 DOES NOT MAKE SENSE!
+  glbSetDensityProjectionFlag(th13\_projection,GLB\_FIXED,GLB\_ALL); \\
   glbSetProjection(th13\_projection); \\ 
@@
 \hspace*{0.5cm} AddToOutput(x,res1,res2);\\
-  \} \\
+  \}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section[Projection onto any hyperplane]{Projection onto any hyperplane}
+\section[Projection onto any hyperplane$^*$]{Projection onto any hyperplane$^*$}
 \index{norm}{Projection!hyperplane}
@@
 \end{function}
-This function works analogously to the ones in the last section. It can, for example, be used to obtain a figure similar to \figu{projex}, left.
-The example on page~\pageref{ex:corrproj} illustrates then the difference
-between the projections of the ``eggs'' within the 
+This function works analogously to the ones in the last section. It can, for example, be used to obtain a figure similar to \figu{projex}, left, but with all other parameters marginalized over.
+The example on page~\pageref{ex:corrproj} illustrates then the result of the
+projection of the ``eggs'' within the 
 $\stheta$-$\deltacp$-plane onto the $\theta_{13}$-axis. 
@@
   \hspace*{2cm} GLB\_FREE,GLB\_FIXED,GLB\_FIXED); \\
-  glbSetDensityProjectionFlag(t13\_projection,GLB\_FREE,GLB\_ALL); \\
+  glbSetDensityProjectionFlag(t13\_projection,GLB\_FIXED,GLB\_ALL); \\
   glbSetProjection(th13\_projection); \\ 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{User-defined priors}
+\section{User-defined priors$^*$}
 \index{norm}{User-defined priors}
@@
 this section can be skipped in a first reading of the manual.
-They allow for arbitrary priors being added to the $\chi^2$ after the systematics
-minimization and before the marginalization over the oscillation parameters (\cf, \equ{priors}).
+They allow for arbitrary priors which depend on the oscillation parameters only (as opposed
+to systematics; \cf, \equ{priors}).
 Therefore, compared to user-defined systematics, they depend on the oscillation
@@
 In addition, it is possible to register
-two functions {\tt central} and {\tt error} being called everytime \GLB{SetCentralValues} or \GLB{SetInputErrors} are called with the same argument.
+two functions {\tt central} and {\tt error} being called every time \GLB{SetCentralValues} or \GLB{SetInputErrors} are called with the same argument.
 For example, the central values and input errors can be stored in
@@
 
-The prior function should expect the parameter structre {\tt glb\_params} containing the
+The prior function should expect the parameter structure {\tt glb\_params} containing the
 current fit values and the {\tt void}-pointer {\tt user\_data} as arguments.
@@
 one does not need to know if the prior was called for one experiment only or all experiments,
-since this information is implicitely given by the different density projection flags being set
+since this information is implicitly given by the different density projection flags being set
 accordingly.
@@
 
-In the last chapter, we introduced the projection of any set of $k$ parameters onto any $n-k$ dimensional hyperplane, which was performed by the minimization over the $k$ free fit parameters.
+Here we describe how one can locate degenerate solutions in \GLOBES , and we discuss
+several techniques for the application software.
+
+\section{Minimization over all oscillation parameters}
+
+In the last chapter, we introduced marginalizations over different parameters to obtain measurement precisions.
 Similarly, one can minimize over {\em all} $n$ parameters to find the local minimum close to any starting point.
@@
 m^2_{31}\rightarrow - \Delta m^2_{31} + \Delta m^2_{21}$, because
-with this defintion the absolute value of  $\Delta m^2_{32}$ is unchanged
+with this definition the absolute value of  $\Delta m^2_{32}$ is unchanged
 and no new frequency is introduced.\\
 
+\section{Advanced tricks for degeneracy localization$^*$}
 \index{norm}{Advanced tricks}
-For the more advanced reader, a number of tricks can be useful for the numerical localization of degenerate solutions:
+For the more advanced reader, a number of tricks can be useful for the numerical localization of degenerate solutions. Here we give a qualitative, incomplete list:
 \begin{description}
@@
 In this case, one can scan one parameter (such as $\theta_{23}$) on a grid by choosing only valid values, and use {\tt glbChiNP} to marginalize over the other ones (while keeping \eg\ $\theta_{23}$ fixed).
-Therefore, the 6-parameter minimization is split into a 1-parameter manual minimization and a 5-parameter simultaneous minimization.
+Therefore, the 6-parameter minimization is split into a 1-parameter grid-based minimization and a 5-parameter simultaneous minimization.
 Advantage: Runs with all \GLOBES\ versions.
@@
 You will find at most two more intersection points in each case.
-The results should look somewhat like this, where the best-fit point is not marked (the marks correspond to the points where the discrete degeneracies are located according to this specific algorithm):
+The results should look somewhat like this, where the best-fit point is not marked\footnote{The marks correspond to the points where the discrete degeneracies are located according to this specific algorithm.
+Note that
+one also wants to find the positions of close-to-intersections, because statistical errors may be as large
+as that the corresponding degenerate solution may still be present at the chosen confidence level.
+Therefore, in the third plot these close-to-intersections are marked as well.}:
 \begin{center}
@@
 Note that similar methods using the $\chi^2$ instead of the total rates have also been successfully used in the past.
+In this case, one would scan for local minimas
+disconnected from the best-fit solution.
 \end{description}
@@
 \end{function}
+This function ignores the filter state, and it does not use the filter if switched on.
 For a constant matter density profile, it is sufficient to specify the oscillation channel with
@@
 \end{function}
+This function uses the filter depending on the filter state, \ie, if it is switched off, it will not be used.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Information from \AEDL\ files}
+\section{Information from \AEDL\ files$^*$}
 \label{sec:aedl_names}
 
-In some cases, it is necessary to obtain information from the \AEDL\
-files. Very basic information is supplied with
+In some cases, it is necessary to obtain information from the loaded \AEDL\
+files. This and the next sections are marked as advanced because knowledge of \AEDL\ is
+required, \ie,  the reader should be familiar with Part~\ref{part:2} of the manual.
+%
+Very basic information is supplied with
 \begin{function}
@@
 {\tt int glbGetEminEmax(int experiment, double *emin, double *emax)} returns the
-energy range [{\tt emin}, {\tt emax}] covered by the binning as defined in the AEDL file for the experiment
+energy range [{\tt emin}, {\tt emax}] covered by the binning as defined in the \AEDL\ file for the experiment
 {\tt experiment}.
@@
 {\tt int glbGetEnergyWindow(int exp, int rule, double *low, double *high)} returns the
-energy window [{\tt low}, {\tt high}] defined in the AEDL file for the experiment
+energy window [{\tt low}, {\tt high}] defined in the \AEDL\ file for the experiment
 {\tt exp} and rule {\tt rule}.
@@
 {\tt int glbGetEnergyWindowBins(int exp, int rule, int *lowbin, int *highbin)} returns the
-energy window in terms of bin numbers [{\tt lowbin}, {\tt highbin}] corresponding to the energy window defined in the AEDL file for the experiment {\tt exp} and rule {\tt rule}.
+energy window in terms of bin numbers [{\tt lowbin}, {\tt highbin}] corresponding to the energy window defined in the \AEDL\ file for the experiment {\tt exp} and rule {\tt rule}.
 \end{function}
@@
 \end{function}
-\begin{function}
-\GLBNS{GetNormalizationInRule}
-{\tt double glbGetNormalizationInRule(int exp, int rule, int signal)}
-returns the normalization (normalization or background center)
-in rule {\tt rule} of the experiment {\tt exp}.
-The parameter {\tt signal} refers to signal (\GLBC{GLB\_SIG}) or background
-(\GLBC{GLB\_BG}).
-\end{function}
+% NOT DOCUMENTED, BECAUSE THESE NORMALIZATIONS ARE NOW PRE-DEFINED:
+%
+% \begin{function}
+% \GLBNS{GetNormalizationInRule}
+% {\tt double glbGetNormalizationInRule(int exp, int rule, int signal)}
+% returns the normalization (normalization or background center)
+% in rule {\tt rule} of the experiment {\tt exp}.
+% The parameter {\tt signal} refers to signal (\GLBC{GLB\_SIG}) or background
+% (\GLBC{GLB\_BG}).
+% \end{function}
 \begin{function}
@@
 \end{function} 
+For each channel, the efficiencies and backgrounds can be returned with the following functions:
+\begin{function}
+\GLBNS{GetEfficiencyPtr}
+{\tt double *glbGetEfficiencyPtr(int exp, int ch, int pre\_post)} returns a pointer to
+the efficiency list for experiment {\tt exp} and channel {\tt ch}. The pre- or post-smearing
+efficiencies are returned with {\tt pre\_post} set to {\tt GLB\_PRE} and {\tt GLB\_POST}, respectively.
+\end{function}
+\begin{function}
+\GLBNS{GetBackgroundPtr}
+{\tt double *glbGetBackgroundPtr(int exp, int ch, int pre\_post)} returns a pointer to
+the background list for experiment {\tt exp} and channel {\tt ch}. The pre- or post-smearing
+backgrounds are returned with {\tt pre\_post} set to {\tt GLB\_PRE} and {\tt GLB\_POST}, respectively.
+\end{function}
 
@@
 Since in \AEDL\, rules, cross section, fluxes etc. carry a `name' by
-which they can be refered to, while in C they carry only an integer
+which they can be referred to, while in C they carry only an integer
 index, it is sometimes difficult to figure out the correct correspondence.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Event rates}
+\section{Event rates$^*$}
 \index{norm}{Event rates}
@@
 
-\section{Fluxes and cross sections}
+\section{Fluxes and cross sections$^*$}
 \index{norm}{Flux} \index{norm}{Cross section}
@@
 
-\section{Baseline and matter density profile}
+\section{Systematics}
+\label{sec:systematics}
+\index{norm}{Systematics}
+\index{norm}{Systematics!on/off}
+
+Changing the systematics at running time can be useful to investigate
+the impact factors affecting the measurement.
+In \GLOBES , the systematics is defined rule-based, \ie, each rule
+has its own systematics. 
+In addition, \GLOBES\ supports dual systematics, \ie, \AEDL\ requires that it has to be
+defined in each rule what ``Systematics on'' and ``Systematics off'' means.
+In principle, these two sets correspond to two completely different systematics
+implementations, and it is up to the \AEDL\ authors to define what that means.
+From the point of view of the API, it is very simple to switch the systematics on and off, \ie,
+to switch between the two systematics modes:
+\begin{function}
+\GLBNS{SwitchSystematics}
+{\tt int glbSwitchSystematics(int exp, int rule, int on\_off)}
+switches the systematics in experiment {\tt exp} and rule {\tt rule}
+on ({\tt on\_off} is \GLBC{GLB\_ON}) or off ({\tt on\_off} is \GLBC{GLB\_OFF}). For the experiment or
+rule index, one can also use \GLBC{GLB\_ALL}. 
+\end{function}
+One can also return the systematics state with
+\begin{function}
+\GLBNS{GetSysOnOffState}
+{\tt int glbGetSysOnOffState(int exp, int rule)}
+returns the systematics state in experiment {\tt exp} and rule {\tt rule}. 
+\end{function}
+In the example on page~\pageref{ex:barcharts}, the application of
+{\tt glbSwitchSystematics} is demonstrated to show the impact of
+systematics, correlations, and degeneracies.
+
+\example{The impact of systematics, correlations, and degeneracies}{
+\label{ex:barcharts}
+\index{norm}{Bar plots}
+\index{norm}{Systematics!on/off}
+Here, we demonstrate how systematics, correlations, and degeneracies
+can be successively included in the calculation of the $\stheta$-sensitivity limit.
+The following code fragment shows how systematics can be switched off
+in order to compute the sensitivity limit from statistics only:
+\begin{quote}
+{\tt 
+/* Calculate chi2 with statistics only */ \\
+double CalcNoSystematics(double theldm,double thex) \\
+\{ \\
+\hspace*{0.5cm} /* Switch systematics off for all exps and all rules */ \\
+\hspace*{0.5cm} glbSwitchSystematics(GLB\_ALL,GLB\_ALL,GLB\_OFF); \\
+\\
+\hspace*{0.5cm} /* Calculate Chi2-list as if systematics were on */ \\
+\hspace*{0.5cm} double res=CalcSystematics(theldm,thex); \\
+\\
+\hspace*{0.5cm} /* Switch systematics on for all exps and all rules */ \\
+\hspace*{0.5cm} glbSwitchSystematics(GLB\_ALL,GLB\_ALL,GLB\_ON); \\
+\\
+\hspace*{0.5cm} return res; \\
+\} \\
+}
+\end{quote}
+
+\vspace*{-0.5cm}
+
+The complete code is very advanced and can be found in {\tt example4.c}.
+It includes many concepts from earlier examples, and, in addition,
+it uses a little trick: It avoids falling into the wrong solution with
+{\tt glbChiTheta13} by using the fit value of $\deltacp$ from the step
+before as prediction for the position of the minimum in the current calculation.
+
+The returned lists of data from the example represent $\chi^2$ 
+as function of the fit value of $\stheta$. The intersections of these
+curves with the line $\chi^2 = 9$ give the $\stheta$ sensitivity
+limits at the $3 \sigma$ confidence level. Note that, in the following
+plot, the $\mathrm{sgn}(\ldm)$- and $(\deltacp,\theta_{13})$-degeneracies
+are not included in the sensitivity limit with correlations only (green bar),
+but only in the limit with degeneracies (yellow bar):
+\begin{center}
+\colorbox{white}{\includegraphics[width=6.5cm]{barsex}}
+\end{center}
+}
+
+The following material requires knowledge of \AEDL , which means that it can be skipped
+at a first reading. 
+%
+During running time, it is possible to change the systematics of an experiment or
+rule (as compared to the systematics assigned in the \AEDL\ file) with the following function:
+\begin{function}
+\GLBNS{SetChiFunction}
+{\tt int glbSetChiFunction(int exp, int rule, int on\_off, const char *name, double *errors)}
+ tells \GLOBES\ to use the registered user-defined systematics identified by the string {\tt name}
+or any built-in systematics
+to calculate $\chi^2$ for the experiment {\tt exp} and the rule {\tt rule}. 
+Both of the parameters {\tt exp} and {\tt rule} can take the value \GLBC{GLB\_ALL} to specify that the
+given systematics function should be used for each experiment or each rule. The parameter 
+{\tt on\_off} determines if the systematics function should be used when systematical errors
+are switched on ({\tt GLB\_ON}), or when they are switched off ({\tt GLB\_OFF}).
+The array {\tt errors} sets the systematical errors in the order in which they
+are expected by the systematics function (indices run from $0$ to
+the number of systematics parameters-1). The function returns zero, if successful.
+\end{function}
+Note that user-defined systematics functions have to be registered with \GLB{DefineChiFunction} first. 
+One can also request the systematics function by 
+\begin{function}
+\GLBNS{GetChiFunction}
+{\tt int glbGetChiFunction(int exp, int rule, int on\_off, char *sys\_id, size\_t max\_len)}
+returns the name of the systematics $\chi^2$ function of a given experiment {\tt exp} and rule {\tt rule} for
+systematics on or off as given by {\tt on\_off}. The name is copied to the string
+{\tt sys\_id}, the maximum length of which is specified by {\tt max\_len}. If
+{\tt max\_len} is too small, or if any other error occurs, the return value is $< 0$.
+\end{function}
+
+% \index{norm}{Error dimension}
+% The error dimension (for the definition, see \Sec~\ref{sec:rules}) can also be accessed directly with 
+% \begin{function}
+% \GLBNS{SetErrorDim}
+% {\tt int glbSetErrorDim(int exp, int rule, int on\_off, int value)}
+% sets the error dimension for systematics on ({\tt on\_off} is \GLBC{GLB\_ON}) or off ({\tt on\_off} is \GLBC{GLB\_OFF}) of experiment {\tt exp} and rule {\tt rule} to the value {\tt value}.
+The function returns
+% $-1$ if not successful.
+% \end{function}
+% \begin{function}
+% \GLBNS{GetErrorDim}
+% {\tt int glbGetErrorDim(int exp, int rule, int on\_off)}
+% returns the error dimension for systematics on ({\tt on\_off} is \GLBC{GLB\_ON}) or off ({\tt on\_off} is \GLBC{GLB\_OFF}) of experiment {\tt exp} and rule {\tt rule}.
+% \end{function}
+
+\index{norm}{Signal!errors}
+\index{norm}{Background!errors}
+\index{norm}{Background!centers}
+Except from the general treatment of systematics, one can read out
+and change the signal and background errors for standard pre-defined systematics
+during running time:
+\begin{function}
+\GLBNS{SetSignalErrors}
+{\tt int glbSetSignalErrors(int exp, int rule, double norm, double tilt)}
+sets the signal errors of experiment {\tt exp} and rule {\tt rule}
+to {\tt norm} (normalization error) and {\tt tilt} (tilt/calibration error).
+\end{function}
+\begin{function}
+\GLBNS{GetSignalErrors}
+{\tt int glbGetSignalErrors(int exp, int rule, double* norm, double* tilt)}
+writes the signal errors of experiment {\tt exp} and rule {\tt rule}
+to {\tt norm} (normalization error) and {\tt tilt} (tilt/calibration error).
+\end{function}
+\begin{function}
+\GLBNS{SetBGErrors}
+{\tt int glbSetBGErrors(int exp, int rule, double norm, double tilt)}
+sets the background errors of experiment {\tt exp} and rule {\tt rule}
+to {\tt norm} (normalization error) and {\tt tilt} (tilt/calibration error).
+\end{function}
+\begin{function}
+\GLBNS{GetBGErrors}
+{\tt int glbGetBGErrors(int exp, int rule, double* norm, double* tilt)}
+writes the background errors of experiment {\tt exp} and rule {\tt rule}
+to {\tt norm} (normalization error) and {\tt tilt} (tilt/calibration error).
+\end{function}
+% ELIMINATED FROM DOC, BECAUSE WE WANT TO AVOID THE BG CENTERS IN THE FUTURE (CONFUSION) [WW]
+%
+% \begin{function}
+% \GLBNS{SetBGCenters}
+% {\tt int glbSetBGCenters(int exp, int rule, double norm, double tilt)}
+% sets the background centers of experiment {\tt exp} and rule {\tt rule}
+% to {\tt norm} (normalization center) and {\tt tilt} (tilt/calibration center).
+% \end{function}
+% \begin{function}
+% \GLBNS{GetBGCenters}
+% {\tt int glbGetBGCenters(int exp, int rule, double* norm, double* tilt)}
+% writes the background centers of experiment {\tt exp} and rule {\tt rule}
+% to {\tt norm} (normalization center) and {\tt tilt} (tilt/calibration center).
+% \end{function}
+As a more flexible concept, the systematical errors may be given as lists for
+both user-defined and built-in systematics. For example, the standard signal and background
+pre-defined errors can be accessed as lists with four elements (signal
+normalization, signal tilt, background normalization, background tilt) if built-in
+systematics is used. The corresponding 
+functions are:
+\begin{function}
+\GLBNS{SetSysErrorsList} 
+{\tt int glbSetSysErrorsList(int exp, int rule, int on\_off, const double *sys\_list)}
+changes the systematical errors defined in the \AEDL\ file in experiment {\tt exp} and 
+rule {\tt rule} to the values in {\tt sys\_list}. This change can be perfomed for
+systematics on or systematics off by using {\tt GLB\_ON} or {\tt GLB\_OFF} for {\tt on\_off}. 
+\end{function}
+\begin{function}
+\GLBNS{GetSysErrorsListPtr} 
+{\tt double* glbGetSysErrorsListPtr(int exp, int rule, int on\_off)}
+returns a pointer to the list of systematical errors as defined in the \AEDL\ file in experiment {\tt exp} and 
+rule {\tt rule}. Choose
+systematics on or systematics off by using {\tt GLB\_ON} or {\tt GLB\_OFF} for {\tt on\_off}. 
+\end{function}
+As usual, all these functions return $-1$ if they were not successful.
+%.
+For user-defined systematics, see \Sec~\ref{sec:userchi}, and 
+for the definitions of these quantities, see \Sec~\ref{sec:rules}.
+
+\section{Baseline and matter density profile$^*$}
 \label{sec:baselinemd}
@@
 in the \AEDL\ file (average density, PREM profile with a given
-number of steps, or arbitrary profile). One can check the currently
+number of steps, or arbitrary profile; \cf, \Tab~\ref{tab:profiletypes}). One can check the currently
 used profile type with
@@
 where the average density or the PREM profile are re-computed, or the
-steps in the arbitrary profile are re-scaled. If this behaviour is 
+steps in the arbitrary profile are re-scaled. If this behavior is 
 not the desired one, one has to use {\tt glbSetProfilDataInExperiment}
@@
 during running time. However, one can change the matter density
-profile, where the profile type is automatically changed to 3, \ie\ 
+profile, where the profile type is automatically switched to 3, \ie\ 
 arbitrary matter density profile. In addition, a number of functions 
@@
 In addition, they normally provide the length of the lists $N$
-by means of an additinal argument which is a pointer to {\tt size\_t}. 
+by means of an additional argument which is a pointer to {\tt size\_t}. 
 Normally, it is enough to declare
@@
 \end{function}
+Note that this function does not interpolate or average within individual layers.
 \begin{function}
@@
 \end{function}
-Note that this function requires that the memory space for the lists
-be reserved already.
 
@@
 
-\example{The impact of systematics, correlations, and degeneracies}{
-\label{ex:barcharts}
-\index{norm}{Bar plots}
-\index{norm}{Systematics!on/off}
-Here, we demonstrate how systematics, correlations, and degeneracies
-can be succesively included in the calculation of the $\stheta$-sensitivity limit.
-The following code fragment shows how systematics can be switched off
-in order to compute the sensitivity limit from statistics only:
-\begin{quote}
-{\tt 
-/* Calculate chi2 with statistics only */ \\
-double CalcNoSystematics(double theldm,double thex) \\
-\{ \\
-\hspace*{0.5cm} /* Switch systematics off for all exps and all rules */ \\
-\hspace*{0.5cm} glbSwitchSystematics(GLB\_ALL,GLB\_ALL,GLB\_OFF); \\
-\\
-\hspace*{0.5cm} /* Calculate Chi2-list as if systematics were on */ \\
-\hspace*{0.5cm} double res=CalcSystematics(theldm,thex); \\
-\\
-\hspace*{0.5cm} /* Switch systematics on for all exps and all rules */ \\
-\hspace*{0.5cm} glbSwitchSystematics(GLB\_ALL,GLB\_ALL,GLB\_ON); \\
-\\
-\hspace*{0.5cm} return res; \\
-\} \\
-}
-\end{quote}
-
-\vspace*{-0.5cm}
-
-The complete code is very advanced and can be found in {\tt example4.c}.
-It includes many concepts from earlier examples, and, in addition,
-it uses a little trick: It avoids falling into the wrong solution with
-{\tt glbChiTheta13} by using the fit value of $\deltacp$ from the step
-before as prediction for the position of the minimum in the current calculation.
-
-The returned lists of data from the example represent $\chi^2$ 
-as function of the fit value of $\stheta$. The intersections of these
-curves with the line $\chi^2 = 9$ give the $\stheta$ sensitivity
-limits at the $3 \sigma$ confidence level. Note that, in the following
-plot, the $\mathrm{sgn}(\ldm)$- and $(\deltacp,\theta_{13})$-degeneracies
-are not included in the sensitivity limit with correlations only (green bar),
-but only in the limit with degeneracies (yellow bar):
-\begin{center}
-\colorbox{white}{\includegraphics[width=6.5cm]{barsex}}
-\end{center}
-}
-
-\section{Systematics}
-\label{sec:systematics}
-\index{norm}{Systematics}
-\index{norm}{Systematics!on/off}
-
-Changing the systematics at running time can be useful to investigate
-the impact factors affecting the measurement.
-In \GLOBES , the systematics is defined rule-based, \ie, each rule
-has its own systematics. 
-In addition, \GLOBES\ supports dual systematics, \ie, \AEDL\ requires that it has to be
-defined in each rule what ``Systematics on'' and ``Systematics off'' means.
-In principle, these two sets correspond to two completely different systematics
-implementations, and it is up to the \AEDL\ authors to define what that means.
-From the point of view of the API, it is very simple to switch the systematics on and off, \ie,
-to switch between the two systematics modes:
-\begin{function}
-\GLBNS{SwitchSystematics}
-{\tt int glbSwitchSystematics(int exp, int rule, int on\_off)}
-switches the systematics in experiment {\tt exp} and rule {\tt rule}
-on ({\tt on\_off} is \GLBC{GLB\_ON}) or off ({\tt on\_off} is \GLBC{GLB\_OFF}). For the experiment or
-rule index, one can also use \GLBC{GLB\_ALL}. 
-\end{function}
-One can also return the systematics state with
-\begin{function}
-\GLBNS{GetSysOnOffState}
-{\tt int glbGetSysOnOffState(int exp, int rule)}
-returns the systematics state in experiment {\tt exp} and rule {\tt rule}. 
-\end{function}
-In the example on page~\pageref{ex:barcharts}, the application of
-{\tt glbSwitchSystematics} is demonstrated to show the impact of
-systematics, correlations, and degeneracies.
-Note that also the systematics function itself may be changed during running time
-using \GLB{SetChiFunction}, as it is described in \Sec~\ref{sec:userchi}.
-
-% \index{norm}{Error dimension}
-% The error dimension (for the definition, see \Sec~\ref{sec:rules}) can also be accessed directly with 
-% \begin{function}
-% \GLBNS{SetErrorDim}
-% {\tt int glbSetErrorDim(int exp, int rule, int on\_off, int value)}
-% sets the error dimension for systematics on ({\tt on\_off} is \GLBC{GLB\_ON}) or off ({\tt on\_off} is \GLBC{GLB\_OFF}) of experiment {\tt exp} and rule {\tt rule} to the value {\tt value}.
-The function returns
-% $-1$ if not successful.
-% \end{function}
-% \begin{function}
-% \GLBNS{GetErrorDim}
-% {\tt int glbGetErrorDim(int exp, int rule, int on\_off)}
-% returns the error dimension for systematics on ({\tt on\_off} is \GLBC{GLB\_ON}) or off ({\tt on\_off} is \GLBC{GLB\_OFF}) of experiment {\tt exp} and rule {\tt rule}.
-% \end{function}
-
-\index{norm}{Signal!errors}
-\index{norm}{Background!errors}
-\index{norm}{Background!centers}
-Except from the general treatment of systematics, one can read out
-and write the signal and background errors for standard pre-defined systematics.
-For user-defined systematics, see \Sec~\ref{sec:userchi}, and 
-for the definitions of these quantities, see \Sec~\ref{sec:rules}.
-\begin{function}
-\GLBNS{SetSignalErrors}
-{\tt int glbSetSignalErrors(int exp, int rule, double norm, double tilt)}
-sets the signal errors of experiment {\tt exp} and rule {\tt rule}
-to {\tt norm} (normalization error) and {\tt tilt} (tilt/calibration error).
-\end{function}
-\begin{function}
-\GLBNS{GetSignalErrors}
-{\tt int glbGetSignalErrors(int exp, int rule, double* norm, double* tilt)}
-writes the signal errors of experiment {\tt exp} and rule {\tt rule}
-to {\tt norm} (normalization error) and {\tt tilt} (tilt/calibration error).
-\end{function}
-\begin{function}
-\GLBNS{SetBGErrors}
-{\tt int glbSetBGErrors(int exp, int rule, double norm, double tilt)}
-sets the background errors of experiment {\tt exp} and rule {\tt rule}
-to {\tt norm} (normalization error) and {\tt tilt} (tilt/calibration error).
-\end{function}
-\begin{function}
-\GLBNS{GetBGErrors}
-{\tt int glbGetBGErrors(int exp, int rule, double* norm, double* tilt)}
-writes the background errors of experiment {\tt exp} and rule {\tt rule}
-to {\tt norm} (normalization error) and {\tt tilt} (tilt/calibration error).
-\end{function}
-% ELIMINATED FROM DOC, BECAUSE WE WANT TO AVOID THE BG CENTERS IN THE FUTURE (CONFUSION) [WW]
-%
-% \begin{function}
-% \GLBNS{SetBGCenters}
-% {\tt int glbSetBGCenters(int exp, int rule, double norm, double tilt)}
-% sets the background centers of experiment {\tt exp} and rule {\tt rule}
-% to {\tt norm} (normalization center) and {\tt tilt} (tilt/calibration center).
-% \end{function}
-% \begin{function}
-% \GLBNS{GetBGCenters}
-% {\tt int glbGetBGCenters(int exp, int rule, double* norm, double* tilt)}
-% writes the background centers of experiment {\tt exp} and rule {\tt rule}
-% to {\tt norm} (normalization center) and {\tt tilt} (tilt/calibration center).
-% \end{function}
-As usual, all these functions return $-1$ if they were not successful.
-
-\section{External parameters in \AEDL\ files}
+\section{External parameters in \AEDL\ files$^*$}
 \label{sec:aedlparams}
@@
 
-\section{Algorithm parameters: Filter functions}
+\section{Algorithm parameters: Filter functions$^*$}
 \index{norm}{Filter!functions}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Simulating non-standard physics}
+\chapter{Simulating non-standard physics$^*$}
 \label{chapt:nsphysics}
@@
 In this chapter, we discuss how to simulate non-standard physics with \GLOBES , \ie, physics
-beyond three-flavor neutrino oscillations. For the first time, this feature was used as experimental 
+beyond the standard three-flavor neutrino oscillation scenario. For the first time, this feature was used as experimental 
 feature in \Ref~\cite{Blennow:2005yk},
@@
 computation of oscillation probabilities, and therefore the flavor transition probabilities, is the core basic
-element of \GLOBES , the introduction of non-standard physics requires fimiliarity with the
+element of \GLOBES , the introduction of non-standard physics requires familiarity with the
 probability calculation in \GLOBES . It is therefore an advanced topic.
@@
 The function should store them into variables accessible to {\tt glb\_probability\_matrix\_function}.
-In addition, it can be used to pre-compute quantities which are dependent on the oscillation paramaters for efficiency. 
+In addition, it can be used to pre-compute quantities which are dependent on the oscillation parameters only, such
+as the mixing matrix. 
 In order to circumvent global variables, arbitrary additional parameters can be
@@
 \\
-  return 0; \\
+ \hspace*{0.5cm}  return 0; \\
 \} \\
@@
 For example, there are no standard routines in \GLOBES\ which define projections including more than six parameters, 
-which means that you should only use {\tt glbChiNP} further on. In addition, you can still use {\tt glbDefineParams}, but this 
+which means that you should only use {\tt glbChiNP} further on in order to have a defined behavior
+of the projection. In addition, you can still use {\tt glbDefineParams}, but this 
 will only access the six standard parameters. You will need to set the additional ones manually using {\tt glbSetOscParams}. 
@@
 can be used to copy all parameters inluding the non-standard ones. For projections,
-do not use any of the {\tt glbChi...} functions anymore except from {\tt glbChiNP} in order to have a
-predictive behavior of the extra dimenstions. Similar to the oscillation parameters, you can define
+it is highly recommended not to use any of the {\tt glbChi...} functions anymore except from {\tt glbChiNP},
+ in order to have a
+predictive behavior of the ``extra'' dimensions.\footnote{It is easy to write headers for new functions
+with a self-defined behavior with respect to the new dimensions using {\tt glbChiNP}.} Similar to the oscillation parameters, you can define
 the marginalization over the extra parameter(s) by {\tt glbSetProjectionFlag}. You can find a simple
@@
 
-\chapter{Experimental features}
+\chapter{Experimental features$^*$}
 \label{chapt:experimental}
@@
 \GLBNS{SelectMinimizer}
-{\tt int glbSelectMinimizer(int minimizer\_ID)} selects the minimizer {\tt GLB\_MIN\_NESTED\_POWELL}, {\tt GLB\_MIN\_DEFAULT}, or the hybrid minimzer {\tt GLB\_MIN\_POWELL} by the parameter {\tt minimizer\_ID}.
+{\tt int glbSelectMinimizer(int minimizer\_ID)} selects the minimizer \GLBC{GLB\_MIN\_NESTED\_POWELL}, \GLBC{GLB\_MIN\_DEFAULT}, or the hybrid minimizer \GLBC{GLB\_MIN\_POWELL} by the parameter {\tt minimizer\_ID}.
 {\tt GLB\_MIN\_NESTED\_POWELL} is the currently implemented standard algorithm, {\tt GLB\_MIN\_DEFAULT} chooses the standard algorithm at any given time, and {\tt GLB\_MIN\_POWELL} is a faster, yet rarely tested hybrid
@@
 This method is much faster,
-but correlations between systematics parameters and oscillation parameters might lead to a different convergence behaviour in some situations.
+but correlations between systematics parameters and oscillation parameters might lead to a different convergence behavior in some situations.
 Thus, when switching to the hybrid minimizer in an application program, one has to reconsider the question whether all degeneracies are found.
 Since one can change the minimizer at any time, it is recommended that one cross check the minimization for a particular experiment.
-The danger of modifying the convergence behaviour in existing application programs is another reason why the new, faster minimizer is still declared as experimental and not used by default
+The danger of modifying the convergence behavior in existing application programs is another reason why the new, faster minimizer is still declared as experimental and not used by default.
+Note that the meaning of the number of iterations in {\tt glb\_params} changes for the hybrid minimizer.
+Since the systematics and oscillation parameter minimizations are not strictly separated anymore,  the minimizer does not count the iterations separately, and the total number of iterations is returned.
+Therefore, it appears that the hybrid minimizer needs more iterations, but, in fact, the default minimizer only counts the oscillation parameter level.
 

### 2007-01-19 Documentation/part2.tex
*** empty log message ***

@@
 \index{norm}{AEDL@\AEDL|(}
-\part{The Abstract Experiment Definition Language -- AEDL}
+\part{The Abstract Experiment Definition Language -- \AEDL }
 \label{part:2}
@@
 
-Here, the general concept of the AEDL is described and illustrated by an example. In addition, a short introduction to the syntax of the AEDL is given.
+Here, the general concept of AEDL\ is described and illustrated by an example. In addition, a short introduction to the syntax of \AEDL\ is given.
 
@@
 
-The goal of AEDL is to describe a large number of complex and very 
+The goal of \AEDL\ is to describe a large number of complex and very 
 different experiments by a limited number of parameters. It allows a
@@
  which is the link between 
-the oscillation physics and the detection properties for a specfific oscillation pattern (\cf, \figu{channel}).
+the oscillation physics and the detection properties for a specific oscillation pattern (\cf, \figu{channel}).
 A channel specifies the mapping of a specific neutrino flavor produced by the source onto a reconstructed neutrino flavor.
@@
 Eventually, each rule gives a $\Delta \chi^2$-value, and the total $\Delta \chi^2$ of one experiment is obtained by adding the $\Delta \chi^2$'s of all rules (\cf, \figu{experiment}).
+Note that each experiment shares a common matter density profile.
  An example for a rule could look like this: We want to detect electron
@@
 \item
- Within each rule, the event numbers are added.
+ Within each rule, the event numbers from different channels are added.
 \item
@@
 \item
- For each rule, the $\Delta \chi^2$ is computed; the $\Delta \chi^2$'s from all rules are added.
+ For each rule, the $\Delta \chi^2$ is computed; the $\Delta \chi^2$'s from all rules are added in an
+experiment.
 \end{itemize}
@@
 
-\section{A simple example for AEDL}
-
-Experiments in \GLOBES\ are defined by the Abstract Experiment Definition Language (AEDL).
-The experiment definition is written into a text file using the AEDL syntax.
+\section{A simple example for \AEDL }
+
+Experiments in \GLOBES\ are defined by the Abstract Experiment Definition Language (\AEDL ).
+The experiment definition is written into a text file using the \AEDL\ syntax.
 Currently, a number of pre-defined experiment definition files are provided with \GLOBES , which have to be modified manually in order to define new experiments.
 The application software then uses this text file to initialize the experiment, where other secondary files might be read for source fluxes, cross sections \etc .
-In this section, we show the definition of a very simple neutrino factory in AEDL, where we do not go into details.
+In this section, we show the definition of a very simple neutrino factory in \AEDL , where we do not go into details.
 In the next chapter, we will discuss each of the individual steps in detail.
@@
 The experiment defined here represents a first simplified version of a neutrino factory experiment.
-It still lacks the correct energy dependence of the efficiencies, the antineutrino disappearance channel, and the channels and rules for the symmetric operation with $\mu^-$ stored.
+It still lacks the correct energy dependence of the efficiencies, the antineutrino disappearance rule, and the channels and rules for the symmetric operation with $\mu^-$ stored.
 However, it may serve as a simple, introductory example.
-In the next chapter, we will demonstrate that AEDL is much more powerful than illustrated here.
+In the next chapter, we will demonstrate that \AEDL\ is much more powerful than illustrated here.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction to the syntax of AEDL}
+\section{Introduction to the syntax of \AEDL }
 \label{sec:syntax}
 
-We now give a short introduction to the syntax of AEDL.
+We now give a short introduction to the syntax of \AEDL .
  The first eight characters have to be {\tt \%!GLoBES}
@@
 The first part is the type of environment, which is {\tt channel} here. 
-There are the following types of environments in AEDL:
+There are the following types of environments in \AEDL :
 \begin{quote}
@@
 channel defined in {\tt <$\ldots$>}. Those names are so-called 
-``automatic variables'' and have to start with {\tt \#}. Note that these names have to be unique and can only be refered to after their definition.
+``automatic variables'' and have to start with {\tt \#}. Note that these names have to be unique and can only be referred to after their definition.
 However, similar to C, one can give a declaration without definition before:
@@
 variables are numbered in the order of their definition, and the number
-can later be used to refer to them in the C code (from $0$ to {\tt numof...}$-1$). In order to facilitate the the mapping from names in AEDL to indices
+can later be used to refer to them in the C code (from $0$ to {\tt numof...}$-1$). In order to facilitate the the mapping from names in \AEDL\ to indices
 in C there are two functions \GLB{NameToValue} and \GLB{ValueToName} which
@@
 \index{aedl}{include@{\tt include}}
-As another feature of AEDL one, can use include files with the {\tt include} command. Includes can be nested up to {\tt MAX\_INCLUSION\_DEPTH}, which is currently set to $10$. Error reporting works 
+As another feature of \AEDL , can use include files with the {\tt include} command. Includes can be nested up to {\tt MAX\_INCLUSION\_DEPTH}, which is currently set to $10$. Error reporting works 
  for nested includes, too. The included file is not required to begin 
@@
 These functions can be used everywhere, where
-otherwise only a scalar number would appear. However, they can not be
-applied to lists, \ie, expressions like {\tt sin(\{1,2,3\})} will not work.
+otherwise only a scalar number would appear.
 
@@
 
-\section{More advanced AEDL features}
+\section{More advanced \AEDL\ features$^*$}
 \label{sec:advaedl}
@@
 In \GLOBES\ 3.0 and higher, a number of new features can be used.  The
-features in this section are slighlty more advanced and can be skipped in a first reading of
-the manual. The
 most important one are lists as variables in \AEDL . They start with {\tt \%}, such as
@@
 use the original and the threading result. For that purpose, the {\tt copy}
-funcion is provided:
+function is provided:
 \begin{quote}
@@
  listb = copy(lista) \\
- listb := lista // Alternative method 
+ listb := lista // Alternative method in environments
 }
@@
 Two helper functions {\tt bincenter()} 
-\index{aedl}{bincenter@{\tt bincenter}} and {\tt samplingbincenter} \index{aedl}{samplingbincenter@{\tt samplingbincenter}} return lists with the central energies of the bins or sampling points, respectively.
+\index{aedl}{bincenter@{\tt bincenter}} and {\tt samplingbincenter()} \index{aedl}{samplingbincenter@{\tt samplingbincenter}} return lists with the central energies of the bins or sampling points, respectively.
 For example,
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Experiment definition with AEDL}
+\chapter{Experiment definition with \AEDL }
 
@@
 
-As we have disussed before, \GLOBES\ can only deal with point sources.
+As we have discussed before, \GLOBES\ can only deal with point sources.
 Thus,  it is not possible to study effects of the finite size of the neutrino production region, such as in the sun or in reactor experiments with many
@@
 \begin{equation}
-\mathrm{Fid.~detector~mass}\,\left[\mathrm{kt/t}\right]\times 
+N \times \mathrm{Fid.~det.~mass}\,\left[\mathrm{kt/t}\right]\times 
 \mathrm{Running~time} \,\left[\mathrm{yr}\right]\times\left\{ \begin{array}{c}
@@
 \mathrm{Useful~parent~decays}\,\left[\mathrm{yr}^{-1}\right]
-\end{array}\right.\,.
-\end{equation}
+\end{array}\right.
+\end{equation}
+with $N$ being a normalization constant.
 Thus, the source power corresponds to either the amount of energy produced per time frame in the target (such as for nuclear reactors or sources based on pion decay), or the useful parent particle decays per time frame (neutrino factories, beta beams).
@@
 idea to document the choices made by the user by corresponding comments
-in \AEDL. For more details on the luminosity implementation, see appendix on page~\pageref{app:flux}.
+in \AEDL. For more details on the luminosity implementation, see \App~\ref{app:flux}.
 
@@
 Note that \GLOBES\ versions older than 3.0 use the still supported {\tt flux} environment, which is different from {\tt nuflux} by an undocumented normalization factor 5.2 for user-defined fluxes.
-This difference is explained in the appendix on page~\pageref{app:flux}.
+This difference is explained in \App~\ref{app:flux}.
 
@@
 For example, two built-in neutrino factory fluxes are available: $\mu^+$-decay ({\tt @builtin = 1}) and $\mu^-$-decay ({\tt @builtin = 2}).
-In these cases, the muon energy (enery of the parent particle) has to be specified together with the number of useful decays  per year.
+In these cases, the muon energy (energy of the parent particle) has to be specified together with the number of useful decays  per year.
 Thus, an example to set up a neutrino factory flux is
@@
 neutrino source, such as neutrinos per proton on target per area per time frame. Right now, each flux has its own normalization factor, which is
-not always straightfoward to calculate. Often, one has to take into account
+not always straightforward to calculate. Often, one has to take into account
 many things, such as the number of target particles per unit mass. 
@@
 Usually this adjustment of {\tt @norm} is performed by comparison with known energy spectra for a specific experiment.
-For more details on the flux definition, see page~\pageref{app:flux} in the appendix.
+For more details on the flux definition, see \App~\ref{app:flux}.
 
@@
 if part of this interval is omitted in the flux file, zero will be
-used there. If some neutrino flavours are not used in the simulation,
+used there. If some neutrino flavors are not used in the simulation,
 the corresponding columns in the flux file have to be filled nevertheless,
@@
 has to be specified.
-The simplest profile is a constant matter density equal to the average matter density from the PREM~\cite{Stacey} onion shell model of the earth:\index{norm}{PREM| \see{Matter density}}
+The simplest profile is a constant matter density equal to the average matter density from the PREM~\cite{Dziewonski:1981xy,Stacey} onion shell model of the earth:\index{norm}{PREM| \see{Matter density}}
 \index{norm}{Matter density!of the earth}
@@
 %
-If you are using this option please cite reference~\cite{Stacey}\index{norm}{Referencing!matter profile data}.
+If you are using this option please cite \Refs~\cite{Dziewonski:1981xy,Stacey}\index{norm}{Referencing!matter profile data}.
 
@@
 \end{quote}  
-This cross section can later be refered to by {\tt \#name}.
+This cross section can later be referred to by {\tt \#name}.
 
@@
 In general, it is advisable to provide the cross sections in the range between {\tt \$sampling\_min} and {\tt \$sampling\_max} (\cf, \Sec~\ref{sec:energy}).
-Cross sections for unused neutrino flavours have to be filled with zeros, and
+Cross sections for unused neutrino flavors have to be filled with zeros, and
 can not just be omitted.
@@
 initial neutrino flavor, its CP eigenvalue 
-(neutrino or antineutrino)\footnote{Currently, \GLOBES\ does not support lepton
-number violating transitions, \ie\ no transitions from neutrino to 
-antineutrino (or vice versa) are considered.}, 
+(neutrino or antineutrino), 
 the detected neutrino flavor, the interaction cross sections for the chosen interaction type, and the energy resolution function of the detector.
 
-Before we come to the definition of channels in AEDL, we introduce the general concept for the calculation of event rates. The first step is to
+Before we come to the definition of channels in \AEDL , we introduce the general concept for the calculation of event rates. The first step is to
 compute the number of events for each IT in the detector for each 
@@
 in the software is a smearing matrix\index{norm}{Energy!resolution}\index{norm}{Smear matrix}.
-The function $\epsilon_\beta^{\text{IT}}(E')$ will later be refered to as ``post-smearing efficiencies'', since it will allow us to define cuts and threshold functions {\em after} the smearing is performed, \ie, as function of $E'$.
+The function $\epsilon_\beta^{\text{IT}}(E')$ will later be referred to as ``post-smearing efficiencies'', since it will allow us to define cuts and threshold functions {\em after} the smearing is performed, \ie, as function of $E'$.
 The detailed definition and initialization of the energy resolution function is described in \Sec~\ref{sec:energy}.
@@
 prefix {\tt NOSC\_}\index{aedl}{channel@{\tt channel}!{\tt NOSC\_}} 
-in either the initial  flavour or 
-the final flavour, like this
+in either the initial  flavor or 
+the final flavor, like this
 \begin{quote}
@@
 %
-In this case all diagonal probabilities $P_{\alpha\alpha}$ are unity, and all off-diagonal probablities  $P_{\alpha\beta}$ are zero. This is,
+In this case all diagonal probabilities $P_{\alpha\alpha}$ are unity, and all off-diagonal probabilities  $P_{\alpha\beta}$ are zero. This is,
 for instance, useful for
@@
 In the following subsections we will explain the energy resolution function.
-All energy resolution functions are defined within an {\tt energy} environment and can be refered to by {\tt \#name}:
+All energy resolution functions are defined within an {\tt energy} environment and can be referred to by {\tt \#name}:
 \begin{quote}
@@
 \item
- The neutrino oscillations are slow on a scale of the smapling point distance.
+ The neutrino oscillations are slow on a scale of the sampling point distance.
 \end{itemize}
@@
 
-\subsection{Low-pass filter}
+\subsection{Low-pass filter$^*$}
 
@@
 
-\subsection{Manual energy smearing}
+\subsection{Manual energy smearing$^*$}
 
@@
 \end{equation}
-where the unshown entries are zero. Thus, the values of $K_{ij}$ have to be specified between $k_l^i$ and $k_u^i$ in the form $\{ k_l^i,k_u^i, K_{i \, k_l^i}, K_{i k_l^i+1} , \hdots , K_{i k_u^i} \}$:
+where the un-shown entries are zero.
+Thus, the values of $K_{ij}$ have to be specified between $k_l^i$ and $k_u^i$ in the form $\{ k_l^i,k_u^i, K_{i \, k_l^i}, K_{i k_l^i+1} , \hdots , K_{i k_u^i} \}$:
 \index{aedl}{energy@{\tt energy}!energy@{\tt "@energy}}
@@
 lead to sums smaller than unity, such as in the case of truncated Gau\ss ian
-distributions. The sum of entries in each {\em row} is not defined, since the events might be unevently distributed into the energy bins
+distributions. The sum of entries in each {\em row} is not defined, since the events might be unevenly distributed into the energy bins
 according to the energy resolution function.
@@
 For a rule, the splitting
-into signal and background is useful for the treatment of systematics, as we will se later. Each rule will lead to a $\Delta \chi^2$-value,
+into signal and background is useful for the treatment of systematics, as we will see later. Each rule will lead to a $\Delta \chi^2$-value,
 which means that all $\Delta \chi^2$'s of the different rules will be added
@@
 is more accurate, but it requires a careful choice of parameters. 
-The first option (method ``T'') is\footnote{Note that this behaviour
+The first option (method ``T'') is\footnote{Note that this behavior
 has slightly changed compared to previous \GLOBES\ releases.}
@@
 energy of the $i$th bin.
-This method is often refered to as a ``tilt'' of the
+This method is often referred to as a ``tilt'' of the
 spectrum, since it describes a linear distortion 
-of the event rate spectrum.
+of the event rate spectrum. It also works for a variable bin width.
 %Note that this definition of a tilt makes
@@
 {\tt chiTotalRatesTilt} &+&+&+&+&T&4&Total rates\\
-{\tt chiSpectrumOnly} & $\infty$ &-& $\infty$ &-&-&7&Spectrum only\\
+{\tt chiSpectrumOnly} & $\infty$ &-&- &-&-&7&Spectrum only\\
 {\tt chiNoSysTotalRates} &-&-&-&-&-&8&Total rates, no syst.\\
@@
 {\tt @backgrounderror} are possible. In this case, {\tt @sys\_on\_errors} and {\tt @sys\_off\_errors}
-have priority. If one or both of these are not present, {\tt @signalerror} or
+have priority.
+In addition, {\tt @sys\_on\_errors} and {\tt @sys\_off\_errors} can be used for built-in systematics in the order: signal normalization, signal tilt (energy calibration), background normalization, background tilt (energy calibration).
+If one or both of these are not present, {\tt @signalerror} or
 {\tt @backgrounderror} will be used.
@@
 \end{quote}
-Note that in the case of additional non-standard parameters, these will have to 
-be included at the end of these lists.
+Note that in the case of additional non-standard parameters, these can not be included in {\tt globes}
+(the command behaves as for the standard three-flavor case, since otherwise a re-compilation of the software
+were necessary).
 Furthermore it is possible to switch off oscillations with the {\tt -N} option and to switch them on again with {\tt -O} (the default).
@@
 
-\subsection*{\AEDL\ external variable substitution}
+\subsection*{\AEDL\ external variable substitution$^*$}
 \index{norm}{globes@{\tt globes}!variable substitution}

### 2007-01-22 Documentation/Paper2/paper2.tex
*** empty log message ***

@@
 \begin{flushright}
-TUM-HEP-YYY/0X\\
+TUM-HEP-656/07\\
 %hep-ph/
@@
 {\Large\bf New features in the simulation of neutrino oscillation experiments with GLoBES 3.0}
-
-(Alternative: ``GLoBES 3.0: From the simulation of non-standard physics to user-defined systematics'')
-
-(Alternative: ``GLoBES 3.0: Simulation of non-standard physics and multi-source multi-detector
-setups'')
 
@@
 We present Version 3.0 of the GLoBES (``General Long Baseline Experiment Simulator'') software, which is
-a simulation tool for short- and long-baseline neutrino oscillation experiments. As new features, \GLOBESN\ allows the simulation of experiments with multiple discrete sources and detectors by the 
-concept of user-defined systematics.
+a simulation tool for short- and long-baseline neutrino oscillation experiments.
+As new features, \GLOBESN\ allows for user-defined systematical errors, which can also be used to simulate experiments with
+multiple discrete sources and detectors.
 In addition, the combination with external information, such as from different experiment classes, is 
-simplified by the concept of user-defined priors. As far as the probability calculation is concerned, \GLOBES\ 
+simplified. As far as the probability calculation is concerned, \GLOBES\ 
 now provides an interface for the inclusion of non-standard physics without re-compilation of the software. 
-The set of experiment descriptions coming with \GLOBES\ has been updated. For example, built-in fluxes are now
+The set of experiment prototypes coming with \GLOBES\ has been updated. For example, built-in fluxes are now
 provided for the simulation of beta beams.
@@
 transition mechanism for neutrinos in a long history of many experiments, see
-\eg , \Ref~\cite{Barger:2003qi}. Future facilities, using accelerator-based
+\eg\ \Ref~\cite{Barger:2003qi} and references therein. Future facilities, using accelerator-based
  neutrino beams or nuclear reactors as neutrino sources, are proposed
@@
 the $\stheta$-axis and the $\stheta$-$\deltacp$-plane. In addition, 
-all oscillation parameters can be kept free to precisely localize 
+all oscillation parameters can be kept free to numerically localize 
 degenerate solutions.
@@
 Hamiltonian in layers of constant matter density using the standard three flavor
-scenario of neutrino oscillations. In the flavor basis, we have
+scenario of neutrino oscillations. In the flavor base, we have
 \begin{equation}
@@
 \end{equation}
-with the evolution operators (CORRECT REFERENCE?)
+with the evolution operators (see, \eg, \Ref~\cite{Ohlsson:1999um})
 \begin{equation}
@@
 event rate computation, which is described in greater detail in \Ref~\cite{Huber:2004ka}. In the next section,
-we will demonstrate how systematics is implemented thereafter.
+we will demonstrate how systematics is implemented.
 
@@
 Here $\sigma_E$ is an additional non-standard parameter describing the wave packet width of the neutrino oscillation wave packet.
+If this parameter is
+kept free (vertical axis), the $\stheta$ sensitivity (horizontal axis) will be affected. Note that
+the standard $\stheta$ sensitivity is given by the curves for $\sigma_E \equiv 0$.
 The figure is computed with
@@
 registering a different probability engine. For example, a user may just copy the standard probability
-engine of \GLOBES , modify it, and register it. An important prerequisite for the ability to handle
+engine of \GLOBES , modify it, and register it. An important prerequisite is the ability to handle
 more than the standard six oscillation parameters (plus matter density) $\boldsymbol{\lambda}=(\theta_{12},\theta_{13}, \theta_{23}, \deltacp, \sdm, \ldm, \rho)$, \ie,
@@
 In the literature, the non-standard physics feature in \GLOBES\ as experimental feature has been used for damping effects (such as neutrino decoherence, decay, \etc) in \Ref~\cite{Blennow:2005yk}, Hamiltonian-level effects (such as non-standard matter effects) in \Ref~\cite{Blennow:2005qj}, and mass-varying neutrinos in \Ref~\cite{Schwetz:2005fy} by implementing an environment dependence of neutrino mass.
-ANY OTHERS (PH!)?
 We show in \figu{reactorNSI} a result from \Ref~\cite{Blennow:2005yk} provided as {\tt example6.c} in the \GLOBES\
-distribution. Compared to \equ{nsi}, the underlying physics takes place at the probability level, \ie, 
-\equ{probs} (for details, see \Ref~\cite{Blennow:2005yk}). In {\tt example6.c}, an analytical implementation
+distribution. In this case, the non-standard physics is a loss of coherence in neutrino oscillations
+described by an intrinsic width of the mass eigenstate wave packets $\sigma_E$ (in energy space).
+Compared to \equ{nsi}, the underlying physics takes place at the probability level, \ie, 
+\equ{probs}, and it can be easily implemented analytically (for details, see \Ref~\cite{Blennow:2005yk}). 
+In {\tt example6.c}, an analytical implementation
 of this physics is used for simplicity.
@@
 \end{equation}
-(JK/PH: CHECK THIS EQUATION IF THIS IS THE ACTUAL IMPLEMENTATION. HERE THE SYSTEMATICS IS\\
-IMPLEMENTED IN THE SIMULATED RATES. IN ADDITION, CHECK IF THEORETICAL AND EXP. RATES
-ARE CORRECTLY GIVEN. [WW])
 
@@
 The thick curve corresponds to \equ{chi2reactor} for $\sigma_R=\sigma_{\mathrm{corr}}=2.8\%$, $\sigma_N=\sigma_F=\sigma_{\mathrm{uncorr}}=0.6\%$ and an additional calibration error.
-The figure is computed with {\tt example5.c} provided with the \GLOBES\ software.
- MISLEADING: WHY IS THE LUMINOSITY
-OF THE ``FAR DETECTOR'' ON THE X-AXIS? SHOULD THIS NOT BE THE SAME AS THE ONE FOR THE NEAR DETECTOR? 
-AT LEAST IF I INTERPRET THE CODE CORRECTLY [WW]}
+The figure is computed with {\tt example5.c} provided with the \GLOBES\ software.}
 \end{figure}
@@
   \chi^2(\boldsymbol{\lambda},a_N,a_F,a_R) = & \sum_{i=1}^{\textrm{\# of bins}} \sum_{d = N,F} 2
-     \left( O_{d,i} - (1 + a_R + a_d) T_{d,i} + \right. \nonumber \\
+     \Big( O_{d,i} - (1 + a_R + a_d) T_{d,i} +  \nonumber \\
 & \left. + (1 + a_R + a_d) T_{d,i} \log \frac{(1 + a_R + a_d) T_{d,i}}{O_{d,i}}  \right) 
@@
 \end{align}
-HOW CAN ONE MAKE THE LEFT BRACKET LARGER? [WW]
 This function can be defined with user-defined systematics. Similarly, calibration errors, shape
@@
 such a calculation, where the thick curve corresponds to \equ{chi2reactor} plus an additional 
-energy calibration error.
+energy calibration error.\footnote{In fact, \figu{reactor} was computed with Gaussian statistics for
+efficiency reasons, whereas \GLOBES\ internally uses Poissonian statistics as a standard; see \Ref~\cite{Manual}.}
 Furthermore, more complicated setups using multiple discrete sources and detectors can be simulated. For details, we refer to the manual~\cite{Manual}.
 In the past, user-defined systematics has (as experimental feature) been used in \Ref~\cite{Huber:2006vr} to
-simulate Double Chooz and in \Ref~\cite{Barger:2006kp} to simulate T2KK. ANYTHING ELSE? [WW]
+simulate Double Chooz, in \Ref~\cite{Barger:2006kp} to simulate T2KK, and in
+\Ref~\cite{Kopp:2006mw} to simulate various reactor experiment setups and complicated
+ backgrounds from geoneutrinos.
 
@@
 \item[{\tt chiSpectrumOnly}]
- Spectrum only, no other systematics 
+ Spectrum only (free signal normalization), no other systematics 
 \item[{\tt chiNoSysTotalRates}]
@@
 after the systematics $\chi^2_{\mathrm{pull}}$ in \equ{chipull} (or similar) 
-has been determined:
+has been determined:\footnote{The marginalization order systematics - correlations
+is, strictly speaking, not valid anymore for the new hybrid minimizer provided
+with \GLOBESN\ as an experimental feature. However, we keep this description
+for pedagogical reasons.}
 \begin{equation}
@@
 this feature has been used as an experimental feature are the combination of terrestrial neutrino
-data with the information from neutrino telescopes~\cite{Winter:2006ce} and the combination of long-baseline data with atmospheric neutrino data~\cite{Huber:2005ep,Campagne:2006yx} (CORRECT?).
+data with the information from neutrino telescopes~\cite{Winter:2006ce} and the combination of long-baseline data with atmospheric neutrino data~\cite{Huber:2005ep,Campagne:2006yx}.
 Another very interesting option is the use of penalties in degeneracy localization, as suggested by Thomas
@@
 scenario & \cite{Campagne:2006yx} \\
-Medium $\gamma$ & {\tt BB\_350.glb} & $\gamma=350$ green field scenario 
+Medium $\gamma$ & {\tt BB\_350.glb} & $\gamma=350$ ``refurbished SPS'' scenario (or other accelerator)
 & \cite{Burguet-Castell:2005pa} \\
@@
 the energy threshold function). 
-%
-IS THE GAMMA=350 OPTION REALLY GREEN-FIELD? I THOUGHT THAT THIS MAY BE THE MAXIMUM
-ONE COULD GET OUT OF THE SPS WITH ACCELERATOR MODIFICATIONS? [WW]
 }
@@
 
-PH, JK: IS THERE ANYTHING ELSE YOU WANT TO SAY? ABOUT NEW PLATFORMS,
-PERL, WINDOWS, PERFORMANCE, 64BIT, INSTALLATION?
-
 \section{Summary and conclusions}
@@
 allow for more space in problem realizations. In this version of \GLOBES ,
-three new such flexible concepts have been introduced: User-defined systematics,
-user-defined priors, and user-defined oscillation probabilities. These approaches
-will make an easy adjustment of the software possible for both experimentalists and theorists.
+three new such flexible concepts have been introduced: user-defined oscillation 
+probabilities, user-defined systematics, and user-defined priors. These approaches
+will make an easy adjustment of the software possible for both experimentalists and theorists,
+and will allow for flexibility at all levels.
 Finally, note that the past has shown that there is often more than one possibility 

### 2007-01-22 Documentation/Paper2/paper2.tex
*** empty log message ***

@@
 a simulation tool for short- and long-baseline neutrino oscillation experiments.
-As new features, \GLOBESN\ allows for user-defined systematical errors, which can also be used to simulate experiments with
+As a new feature, \GLOBESN\ allows for user-defined systematical errors, which can also be used to simulate experiments with
 multiple discrete sources and detectors.

### 2007-01-23 Documentation/tutorials/T2K-tutorial/tex/T2K-tutorial.tex
JK - Added TeX-Files for T2K-tutorial

\documentclass[12pt,a4paper]{article}

\usepackage{a4wide}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{ae,aecompl}
\usepackage{amsthm}

%\pagestyle{empty}
\setlength{\parindent}{0pt}
\renewcommand{\labelenumi}{\alph{enumi})}
\renewcommand{\labelenumii}{\roman{enumii})}

% Neutrino oscillation channels
\newcommand{\ee}{\ensuremath{\nu_e \rightarrow \nu_e}}
\newcommand{\emu}{\ensuremath{\nu_e \rightarrow \nu_\mu}}
\newcommand{\etau}{\ensuremath{\nu_e \rightarrow \nu_\tau}}
\newcommand{\mue}{\ensuremath{\nu_\mu \rightarrow \nu_e}}
\newcommand{\mumu}{\ensuremath{\nu_\mu \rightarrow \nu_\mu}}
\newcommand{\mutau}{\ensuremath{\nu_\mu \rightarrow \nu_\tau}}
\newcommand{\taue}{\ensuremath{\nu_\tau \rightarrow \nu_e}}
\newcommand{\taumu}{\ensuremath{\nu_\tau \rightarrow \nu_\mu}}
\newcommand{\tautau}{\ensuremath{\nu_\tau \rightarrow \nu_\tau}}

% Neutrino oscillation parameters
\newcommand{\sthsol}{\ensuremath{\sin^2 2\theta_{12}}}     % Mixing angles
\newcommand{\sthchooz}{\ensuremath{\sin^2 2\theta_{13}}}
\newcommand{\sthatm}{\ensuremath{\sin^2 2\theta_{23}}}
\newcommand{\sthtf}{\ensuremath{\sin^2 2\theta_{2f}}}
\newcommand{\ldm}{\ensuremath{{\Delta m_{31}^2}}}          % Mass squared differences
\newcommand{\sdm}{\ensuremath{{\Delta m_{21}^2}}}

% Important neutrino experiments
\newcommand{\TtoK}{{\sffamily T2K}}

% Miscellaneous commands
\newcommand{\aufg}[2]{\vspace{4mm}{\bf\underline{Problem #1:} {#2}} \vspace{3mm}}

% Environment for API reference entries
\newtheoremstyle{dotless}{}{}{\itshape}{}{\bfseries}{}{0cm}{}
\theoremstyle{dotless}
\newtheorem*{function}{}

\begin{document}
\vspace*{-3cm}
\begin{center}
{\large\bf GLoBES Tutorial: Simulating accelerator neutrino experiments}\\[0.3cm]
GLoBES workshop in Heidelberg, Germany, January 24-26, 2007
\end{center}
\vspace*{2mm}

Joachim Kopp \hfill 25.01.07

\bigskip
\hrule
\vspace*{4mm}

{\small
This tutorial will introduce some of the basic concepts required to simulate
accelerator neutrino experiments with GLoBES (General Long Baseline Experiment
Simulator). We will start from a simple and incomplete implementation of the
\TtoK\ superbeam experiment, which we will gradually improve. We will first consider
the precision measurement of the leading atmospheric oscillation parameters
$\theta_{23}$ and $\ldm$, and later discuss the possible detection of
non-zero $\theta_{13}$ and $\delta_{CP}$.
}

% ----------------------------------------------------------------------------
\section*{Part 1: Precision measurement of $\theta_{23}$ and $\ldm$}
% ----------------------------------------------------------------------------

\aufg{1}{Warm-up}

Consider first the program in the directory {\tt globes-tutorials/th23dm31/},
which computes $\chi^2$ as a function of the fit values of $\theta_{23}$
and $\ldm$, but is still lacking several important features. Compile
the program by typing {\tt make} and run it with the command {\tt ./th23dm31}.
Each line of the output file {\tt th23dm31.dat} will contain three numbers:
the fit value for $\theta_{23}$ in degrees, the fit value for $\ldm$ in eV$^2$,
and the corresponding $\chi^2$. Use the script {\tt th23dm31.gnuplot}
to view the results as an EPS plot, which should look like
Fig.~\ref{fig:th23dm31-orig}.

Now, familiarize yourself with the C code ({\tt th23dm31.c}) and the AEDL
experiment definition ({\tt T2K-tutorial.glb}). In doing so, you
can already watch out for the aforementioned shortcomings of the code.
There are essentially three of them, which we will one by one discuss
below in problems 2, 4, and 5.

\begin{figure}
  \begin{center}
    \includegraphics{T2K-1}
  \end{center}
  \vspace{-0.7 cm}
  \caption{Output of {\tt th23dm31}: Confidence regions in the $\theta_{23}$--$\ldm$
    plane for a (too) simple implementation of the \TtoK\ experiment.}
  \label{fig:th23dm31-orig}
\end{figure}

\aufg{2}{Spectral analysis vs.\ total rates}

Let us examine Fig.~\ref{fig:th23dm31-orig} more closely: There is
a very strong correlation between $\theta_{23}$ and $\ldm$: A mixing angle
far from its assumed ``true'' value of $45^\circ$ is still compatible with the
data, if, at the same time, a larger $\ldm$ is assumed. The reason for this
is, that increasing $\ldm$ means that larger parts of the neutrino energy
spectrum are affected by oscillations, which compensates the smaller
oscillation amplitude. This woul be a severe limitation to the sensitivity
of the experiments, but luckily, it can be remedied by an improved data
analysis: Instead of performing a total rates analysis, we should incorporate
spectral information. Then, we can determine the energy at which the first
oscillation maximum occurs, and thus pin down $\ldm$.

Accordingly, we have to increase the number of analysis bins ({\tt \$bins}) from {\tt 1}
to {\tt 20} and select a different $\chi^2$ function in {\tt T2K-tutorial.glb}.
Change the command {\tt @sys\_on\_function = "chiTotalRatesTilt"} to
{\tt @sys\_on\_function = "chiSpectrumTilt"} in the first two rules
({\tt \#NU\_E\_Appearance\_QE} and {\tt \#NU\_MU\_Disappearance\_QE}).
Leave the charged current $\nu_e$ appearance rule unchanged to reflect the
fact that the Super-Kamiokande detector cannot reconstruct the energy of
CC $\nu_e$ events very well, so that a spectral analysis is not possible
for them. The plot you will obtain after re-running the program,
Fig.~\ref{fig:th23dm31-spectral}, looks much more like what we expect
from \TtoK. The correlation between $\theta_{23}$ and $\ldm$ is still visible,
but much less pronounced than before.

\begin{figure}
  \begin{center}
    \includegraphics{T2K-2}
  \end{center}
  \vspace{-0.7 cm}
  \caption{Solution of problem 2. Confidence regions in the $\theta_{23}$--$\ldm$
    plane with the inclusion of spectral information.}
  \label{fig:th23dm31-spectral}
\end{figure}

\aufg{3}{The octant degeneracy}

Up to now, we have considered $\theta_{23}$ to be maximal. However, deviations
Of about $\pm 5^\circ$ from this best fit value are still compatible with
current data. Therefore, let us now consider the effect of $\theta_{23} = 40.0^\circ$.
Modify {\tt th23dm31.c} accordingly and compute the resulting confidence regions,
which should resemble Fig.~\ref{fig:th23dm31-nonmax}a. There are now two
distinct regions in the parameter space, which are both compatible with the
data at less than $1\sigma$: The correct one around $\theta_{23} = 40.0^\circ$,
and the degenerate one at $\theta_{23} = 90^\circ - 40.0^\circ$ (the so-called
octant degeneracy).

\begin{figure}
  \hspace{-0.7 cm}
  \begin{tabular}{cc}
    \includegraphics[width=8cm]{T2K-3}  & 
    \includegraphics[width=8cm]{T2K-4} \\
           (a)                   &          (b)
  \end{tabular}
  \caption{Solution of problem 3: Confidence regions in the $\theta_{23}$--$\ldm$
  plane for non-maximal $\theta_{23}$ and (a) zero $\theta_{13}$,
  (b) large $\theta_{13}$.}
  \label{fig:th23dm31-nonmax}
\end{figure}

This looks correct, but let us now increase $\theta_{13}$ to a value close
to the current upper bound, say, $\sthchooz = 0.1$. You will find that the
degenerate solution is now excluded at almost $2\sigma$ (Fig.~\ref{fig:th23dm31-nonmax}b).
Try to understand this feature from the (approximate) analytical expressions
for the neutrino oscillation probabilities:
\begin{align*}
  P_{\mu \mu} &= 1 - \sin^2 2\theta_{23} \, \sin^2 \Delta +
                    \alpha \, c_{12}^2 \, \sin^2 2\theta_{23} \, \Delta \, \sin 2\Delta \\
              & \ \ \ \ - \alpha^2 \, \Delta^2 \left[ \sin^2 2\theta_{12} \, c_{23}^2
                          + c_{12}^2 \, \sin^2 2\theta_{23} \left(\cos 2\Delta 
                          - s^2_{12} \right) \right]
                          + 4 \, s_{13}^2 \, s_{23}^2 \, \cos 2\theta_{23} \, \sin^2 \Delta \\
              & \ \ \ \ - 2 \, \alpha \, s_{13} \, \sin 2\theta_{12} \, s_{23}^2 \,
                                \sin 2\theta_{23} \, \cos \delta_{\rm CP} \, \Delta \, \sin 2\Delta\,,
    \\[3mm]
%
  P_{\mu e}^\mathrm{vac} &= \alpha^2 \, \sin^2 2\theta_{12} \, c_{23}^2 \, \Delta^2
                          + 4 \, s_{13}^2 \, s_{23}^2 \, \sin^2\Delta 
                          + 2 \, \alpha \, s_{13} \, \sin2\theta_{12} \sin2\theta_{23} \,
                              \cos(\Delta+\delta_{\rm CP}) \, \Delta \, \sin\Delta \,,
\end{align*}
where $\alpha = \sdm / \ldm$, $\Delta = \ldm L / 4E$, 
$s_{ij} = \sin\theta_{ij}$, and $c_{ij} = \cos\theta_{ij}$.

\aufg{4}{Incorporation of correlations with $\theta_{13}$ and $\delta_{CP}$}

Fig.~\ref{fig:th23dm31-nonmax}b should make us suspicious, since we do not
expect \TtoK\ to have the capability to resolve the octant degeneracy. Indeed,
we have so far assumed full knowledge about all oscillation parameters
except $\theta_{23}$ and $\delta_{CP}$, i.e.\ we have kept them fixed
at their ``true'' values in our fits. Unless there were some theoretical
argument (e.g.\ a flavor symmetry) pinning down these parameters, we
should allow them to vary within their presently allowed ranges in the fit.
This will yield smaller $\chi^2$ values, i.e.\ the sensitivity will decrease.

The marginalization over multi-dimensional subspaces of the oscillation
parameter space is one of the most powerful features of GLoBES.
To enable it in our code, we use the API functions {\tt glbAllocProjection},
{\tt glbDefineProjection}, {\tt glbSetDensityProjectionFlag}, and
{\tt glbSetProjection} to specify that $\theta_{12}$, $\theta_{13}$,
$\sdm$, and $\delta_{CP}$ should be marginalized over, while $\theta_{23}$,
$\ldm$, and the matter density should be kept at their initial values.
A short documentation of the required API functions is given in the
appendix.

Furthermore, we have to specify the parameters of the prior terms
\begin{align*}
  \chi^2_{\rm prior} = \frac{(x - x_0)^2}{\sigma_x^2},
\end{align*}
which can be added to $\chi^2$ for all oscillation parameters $x$ to
include external information on these parameters. Set all $x_0$ to the
{\tt true\_values} with the function {\tt glbSetCentralValues}, and use
{\tt glbAllocParams}, {\tt glbDefineParams}, {\tt glbSetDensityParams},
and {\tt glbSetInputErrors} to set
\begin{align*}
  \sigma_{\theta_{12}} &= 0.1 \cdot \theta_{12, \rm true} \\
  \sigma_\sdm          &= 0.1 \cdot \sdm_{\rm, true}.
\end{align*}
For all other parameters (including the matter density), we choose to omit
the prior terms by setting the respective $\sigma_x$ to 0.

Finally, we have to replace the call to {\tt glbChiSys} by one to {\tt glbChiNP}.
Due to the multi-dimensional minimization in the oscillation parameter space,
the calculation will now take a bit longer.\footnote{Note that our example
programs already use the faster, alternative minimization method GLB\_MIN\_POWELL
instead of the default GLB\_MIN\_NESTED\_POWELL. We emphasize, however, that
this is still an experimental feature in GLoBES 3.0.}
%For debugging, you
%can speed things up by switching off systematical errors with the command
%\begin{quote}
%  \ttfamily
%  glbSwitchSystematics(GLB\_ALL, GLB\_ALL, GLB\_OFF);
%\end{quote}
%This tells GLoBES to calculate $\chi^2$ according to the
%{\tt @sys\_off\_function} directives instead of the default
%{\tt @sys\_on\_function}. For \TtoK, this means using the $\chi^2$
%functions {\tt chiNoSysSpectrum} and {\tt chiNoSysTotalRates},
%which do not include any free systematical biases. In contrast to this,
%the default functions {\tt chiSpectrumTilt} and {\tt chiTotalRatesTilt}
%leave the normalization and tilt of the neutrino spectrum variable
%within the ranges defined by {\tt @signalerror} and {\tt @backgrounderror}.
%Thus, GLoBES has to fit these systematics parameters in addition to the
%oscillation parameters, which greatly increases the computational
%effort.

\begin{figure}
  \begin{center}
    \includegraphics{T2K-5}
  \end{center}
  \vspace{-0.7 cm}
  \caption{Solution of problem 4: Confidence regions in the $\theta_{23}$--$\ldm$
    plane for non-maximal $\theta_{23}$ and large $\theta_{13}$, including
    parameter correlations.}
  \label{fig:th23dm31-corr}
\end{figure}

\aufg{5}{The $\mathrm{sgn}(\ldm)$ degeneracy}

Besides the octant degeneracy, there is another degeneracy in the three-flavor
neutrino oscillation probabilities, which we have not considered ao far:
The ambiguity in the sign of $\ldm$. Indeed, if we considered the
half-plane of negative $\ldm$, we would find a mirror image of
Fig.~\ref{fig:th23dm31-corr}.

For the moment, however, we are only interested in the positions of the
two solutions with $\ldm < 0$. To find them, use the function {\tt glbChiAll},
and choose the starting values such that the minimizer will converge
to the appropriate local minima of $\chi^2$. Create a new {\tt glb\_params}
data structure and pass it to {\tt glbChiAll} as the second argument to
retrieve the position of the minimum.

Why is $|\ldm|$ slightly smaller at the degenerate solutions than at
the true one?

% ----------------------------------------------------------------------------
\section*{Part 2: Generic three-flavor effects: $\theta_{13}$ and $\delta_{CP}$}
% ----------------------------------------------------------------------------

\aufg{6}{Confidence regions in the $\theta_{13}$--$\delta_{CP}$ plane}

One of the main aims of superbeam experiments is the detection of
generic three-flavor effects, in particular of non-zero $\theta_{13}$
or $\delta_{CP}$. To examine the capability of \TtoK\ to measure these
parameters, we provide the program {\tt th13delta} and the accompanying
script {\tt th13delta.gnuplot}, which produces a confidence plot 
in the $\theta_{13}$--$\delta_{CP}$ plane. For simplicity, the program
only considers the normal mass hierarchy in the fit, i.e.\ we assume
that the $\mathrm{sgn}(\ldm)$ degeneracy has been resolved independently,
e.g.\ with supernova neutrinos. Furthermore, we have set the ``true''
$\theta_{23}$ back to $45^\circ$, so that we can also neglect the octant
degeneracy.

Fig.~\ref{fig:th13delta-orig} shows the results of the simulation.
As we can see from the plot, there is a strong correlation between
$\theta_{13}$ and $\delta_{CP}$, which makes it impossible to
efficiently constrain the CP phase. In the remainder of this
tutorial, we will consider two strategies for breaking this
correlation.

\begin{figure}
  \begin{center}
    \includegraphics{T2K-6}
  \end{center}
  \vspace{-0.7 cm}
  \caption{Output of {\tt th13delta}: Confidence regions in the
    $\theta_{13}$--$\delta_{CP}$ plane for a fit with the normal
    mass hierarchy.
    The ``true'' parameter values are $\theta_{23} = 45^\circ$,
    $\ldm = 2.6 \cdot 10^{-3}$~eV$^2$, $\sthchooz = 0.1$, and
    $\delta_{CP} = 90^\circ$.}
  \label{fig:th13delta-orig}
\end{figure}

\aufg{7}{Improving the sensitivity by anti-neutrino running}

One idea for improving the sensitivity of superbeam experiments is to operate
them in the anti-neutrino mode for several years after the neutrino running. 
Although the cross sections, and thus the event numbers, in the anti-neutrino
mode are smaller by a factor of three than those for neutrinos, one can
benefit from the fact that the dependence of the anti-neutrino oscillation
probabilities on $\delta_{CP}$ is different. To see the effect of this,
we need to incorporate a new flux definition, several new channels, and the
appropriate rules in th AEDL file. The parameters of these new AEDL
environments are mostly the same as for the existing ones, so you can start by
simply duplicating these. Give them new names, change the flux file to
{\tt JHFminus.dat}, and reverse the CP signs for the new channels.
Finnaly, increase the running time for the anti-neutrino mode to 6~years.

\begin{figure}
  \begin{center}
    \includegraphics{T2K-7}
  \end{center}
  \vspace{-0.7 cm}
  \caption{Solution of problem 7: Confidence regions in the
    $\theta_{13}$--$\delta_{CP}$ plane for 3~years of neutrino
    running, followed by 3~years of anti-neutrino running, and
    assuming the normal mass hierarchy.}
  \label{fig:th13delta-anti}
\end{figure}

As you can see from Fig.~\ref{fig:th13delta-anti}, the anti-neutrino running
will exclude almost half of the $\delta_{CP}$ values at the 90\% C.L., but
at $3\sigma$, the excluded region is much smaller.

\aufg{8}{Improving the sensitivity by incorporating reactor results}

Another option for breaking some of the correlations in \TtoK\ is the combination
with an advanced reactor neutrino experiment. To study this, incorporate
{\tt Reactor2.glb} into your simulation by adding an appropriate call to
{\tt glbInitExperiment}. Re-run the program to obtain
Fig.~\ref{fig:th13delta-reactor}, which shows that an advanced reactor
experiment can measure $\theta_{13}$ with a great precision. Moreover,
since the $\nu_e$ disappearance channel is independent of $\delta_{CP}$,
it does not suffer from parameter correlations. Therefore, the parameter
constraints that can be set by a combination of \TtoK\ and reactor data, are
competitive to those obtainable with the much more expensive anti-neutrino running.
Note, however, that this is only true as long is $\theta_{13}$ is
large. For smaller values, one cannot so easily benefit from reactor data.

\begin{figure}
  \begin{center}
    \includegraphics{T2K-8}
  \end{center}
  \vspace{-0.7 cm}
  \caption{Solution of problem 8: Confidence regions in the
    $\theta_{13}$--$\delta_{CP}$ plane for 3~years of \TtoK\ neutrino
    running, combined with {\tt Reactor2} data (exposure of
    8000~GW~kt~yrs), and assuming the normal mass hierarchy.}
  \label{fig:th13delta-reactor}
\end{figure}

% ----------------------------------------------------------------------------
\appendix
\section*{Appendix: Documentation of required API functions}
% ----------------------------------------------------------------------------

\begin{function}
{\tt glb\_params glbAllocParams()} allocates the memory space
needed for a parameter vector and returns a pointer to it.
\end{function}

\begin{function}
{\tt glb\_projection glbAllocProjection()} allocates the memory space
needed for a projection definition, and returns a pointer to it.
\end{function}

\begin{function}
{\tt double glbChiAll(const glb\_params in, glb\_params out, int exp)}
returns the minimized $\chi^2$ over all parameters for the experiment
{\tt exp}. For the simulation of all initialized experiments, use
{\tt GLB\_ALL} for {\tt exp}. The values in {\tt in} are the guessed
fit values for the minimizer. The actually determined parameters at
the minimum are returned in {\tt out}. If {\tt out} is set to {\tt NULL},
this information will not be returned.
\end{function}

\begin{function}
{\tt double glbChiNP(const glb\_params in, glb\_params out, int exp)}
returns the projected $\chi^2$ onto the hyperplane specified by
{\tt glbSetProjection} for the  experiment {\tt exp}.
For the simulation of all initialized experiments,
use {\tt GLB\_ALL} for {\tt exp}. The values in {\tt in} are the guessed
fit values for the minimizer (all free parameters) and the fit values
on the hyperplane (all fixed parameters). The actually determined parameters
at the minimum are returned in {\tt out}, where the fixed parameters are
still at their input values. If {\tt out} is set to {\tt NULL}, this
information will not be returned.
\end{function}

\begin{function}
{\tt glb\_params glbDefineParams(glb\_params in, double theta12, double
theta13, \\ double theta23, double delta, double dms, double dma)}
assigns a set of oscillation parameters to the vector {\tt in}, which
has to be allocated before. The return value is the pointer to {\tt in}
if the assignment was successful, and {\tt NULL} otherwise.
\end{function}

\begin{function}
{\tt glb\_projection glbDefineProjection(glb\_projection in, int theta12,
int \\ theta13,int theta23, int delta, int dm21, int dm31)} defines a
projection in the oscillation parameter space. For each parameter,
the function expects one of the flags {\tt GLB\_FREE} or {\tt GLB\_FIXED},
indicating whether the respective parameter is to me marginalized over
or not. The return value is the pointer to {\tt in} if
the assignment was successful, and {\tt NULL} otherwise.
\end{function}

\begin{function}
{\tt void glbFreeParams(glb\_params stale)} frees the memory
needed for a parameter vector {\tt stale}.
\end{function}

\begin{function}
{\tt void glbFreeProjection(glb\_projection stale)} frees the memory
needed for the projection definition vector {\tt stale}.
\end{function}

\begin{function}
{\tt int glbInitExperiment(char *inf, glb\_exp *in, int *counter)}
adds a single experiment with the filename {\tt inf} to the list of currently
loaded experiments. The {\tt counter} is a pointer to the variable containing
the number of experiments (normally {\tt \&glb\_num\_of\_exps}), and
{\tt in} points to the beginning of the experiment list (normally
{\tt \&glb\_experiment\_list[0]}). The function returns zero if it was
successful.
\end{function}

\begin{function}
{\tt int glbSetCentralValues(const glb\_params in)} sets the central
values for the prior terms to {\tt in}.
\end{function}

\begin{function}
{\tt glb\_params glbSetDensityParams(glb\_params in, double dens, int which)}
sets the density parameter for experiment number {\tt which} in the structure
{\tt in} to the value {\tt dens}. If the assignment was unsuccessful, the
function returns {\tt NULL}. If {\tt GLB\_ALL} is used for {\tt which},
the density parameters of all experiments will be set.
\end{function}

\begin{function}
{\tt glb\_projection glbSetDensityProjectionFlag(glb\_projection in,
int flag, int which)} specifies whether the matter density in experiment
number {\tt which} should be marginalized over ({\tt flag} = {\tt GLB\_FREE})
or not ({\tt flag} = {\tt GLB\_FIXED}). The return value is the pointer to {\tt in}
if the assignment was successful, and {\tt NULL} otherwise.
\end{function}

\begin{function}
{\tt int glbSetInputErrors(const glb\_params in)}
sets the input errors for all of the following minimizer calls to {\tt in}.
An input error of $0$ corresponds to not taking into account the
respective prior.
\end{function}

\begin{function}
{\tt int glbSetProjection(const glb\_projection in)} sets the projection
to {\tt in}. The return value is $0$ if successful, and $-1$ if
unsuccessful.
\end{function}

\end{document}

### 2007-01-23 Documentation/tutorials/AEDL-tutorial/source/AEDL-tutorial.tex
*** empty log message ***

\documentclass[12pt,a4paper]{article}

\usepackage{a4wide}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{ae,aecompl}
\usepackage{amsthm}

%\pagestyle{empty}
\setlength{\parindent}{0pt}
\renewcommand{\labelenumi}{\alph{enumi})}
\renewcommand{\labelenumii}{\roman{enumii})}

% Neutrino oscillation channels
\newcommand{\ee}{\ensuremath{\nu_e \rightarrow \nu_e}}
\newcommand{\emu}{\ensuremath{\nu_e \rightarrow \nu_\mu}}
\newcommand{\etau}{\ensuremath{\nu_e \rightarrow \nu_\tau}}
\newcommand{\mue}{\ensuremath{\nu_\mu \rightarrow \nu_e}}
\newcommand{\mumu}{\ensuremath{\nu_\mu \rightarrow \nu_\mu}}
\newcommand{\mutau}{\ensuremath{\nu_\mu \rightarrow \nu_\tau}}
\newcommand{\taue}{\ensuremath{\nu_\tau \rightarrow \nu_e}}
\newcommand{\taumu}{\ensuremath{\nu_\tau \rightarrow \nu_\mu}}
\newcommand{\tautau}{\ensuremath{\nu_\tau \rightarrow \nu_\tau}}

% Neutrino oscillation parameters
\newcommand{\sthsol}{\ensuremath{\sin^2 2\theta_{12}}}     % Mixing angles
\newcommand{\sthchooz}{\ensuremath{\sin^2 2\theta_{13}}}
\newcommand{\sthatm}{\ensuremath{\sin^2 2\theta_{23}}}
\newcommand{\sthtf}{\ensuremath{\sin^2 2\theta_{2f}}}
\newcommand{\ldm}{\ensuremath{{\Delta m_{31}^2}}}          % Mass squared differences
\newcommand{\sdm}{\ensuremath{{\Delta m_{21}^2}}}

% Important neutrino experiments
\newcommand{\TtoK}{{\sffamily T2K}}

% Miscellaneous commands
\newcommand{\aufg}[2]{\vspace{4mm}{\bf\underline{Problem #1:} {#2}} \vspace{3mm}}

% Environment for API reference entries
\newtheoremstyle{dotless}{}{}{\itshape}{}{\bfseries}{}{0cm}{}
\theoremstyle{dotless}
\newtheorem*{function}{}

\begin{document}
\vspace*{-3cm}
\begin{center}
{\large\bf GLoBES Tutorial: Advanced Usage of AEDL}\\ [0.3cm]
GLoBES workshop in Heidelberg, Germany, January 24-26, 2007
\end{center}
\vspace*{2mm}

Mark Rolinec \hfill 25.01.07

\bigskip
\hrule
\vspace*{4mm}

{\small
This tutorial will adress basic and advanced features of the {\bf A}bstract 
{\bf E}xperiment {\bf D}efinition {\bf L}anguage AEDL. The syntax of AEDL is
comprehensive and thus allows to define all different kinds of neutrino 
oscillation experiments with different neutrino sources and different detector
technologies. Within this tutorial we will define a toy experiment by using 
the most important features of AEDL and get familiar with the AEDL syntax.
Then we will modifiy the experiment description going from basic usage of AEDL
to more advanced features. After that we will explore some of the pre-defined
AEDL files that come with GLoBES 3.0 and analyze advanced features in their AEDL
description. \\

The toy experiment for this tutorial is absolutely fictious, although the attributes 
of the detector follow roughly the response of a huge Water Cerenkov detector.
However, all efficiencies and migration matrices are coming from a fake 
Monte Carlo simulation and do {\bf not} exactly correspond to the true 
attributes of a Water Cerenkov detector. So, all files prepared for this tutorial 
do {\bf not} correspond to a real detector response and all files are just to 
be used for this tutorial.  
}

% ----------------------------------------------------------------------------
\section*{Part 1: AEDL implementation of a toy experiment}
% ----------------------------------------------------------------------------

Suppose you work for a collaboration that plans to build a long baseline 
neutrino oscillation experiment with a flavor-pure neutrino beam, that consists 
either of $\nu_e$ or $\bar{\nu}_e$. It is planned to build a huge {\em modified} 
Water Cerenkov detector. Now, you want to analyze the future potential of the 
experiment and maybe try to optimize some of the attributes. Your decision is to
do the analysis with GLoBES. So, first you have to build the experiment description
in AEDL,~{\it i.e.} write the corresponding {\tt glb}-file. One of your colleagues
sends the results from a Monte Carlo analysis. 
\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=0.36\textwidth]{CCevents.eps} \hspace{0.5cm}
    \includegraphics[width=0.36\textwidth]{QEevents.eps}
  \end{center}
  \vspace{-0.7 cm}
  \caption{\small The sample of 20000 charged current $\nu_\mu$ events on the left hand side. The
  quasi-elastic sample is shown in dark grey while the inelastic part of the sample is
  shown in bright grey. The quasi-elastic part of the sample alone is plotted at the right-hand
  side. It can be seen that this sample allows to recontruct the true neutrino energy at a
  certain level, while this is not possible for the inelastic part of the sample.}
  \label{fig:mccc}
\end{figure}

It is also important to implement the background from single pion production in 
neutral current events. Therefore, the colleague sends the Monte Carlo sample of NC events.

\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=0.45\textwidth]{NCevents.eps} 
  \end{center}
  \vspace{-0.7 cm}
  \caption{\small The sample of 5000 neutral current events. It can be seen that the
  reconstructed energy is systematically smaller than the true neutrino energy since
  there is missing energy due to the undetected neutrino in the NC event.}
  \label{fig:mcnc}
\end{figure}

The colleague also sends the energy dependent signal efficiencies of the charged
current $\nu_\mu$/$\bar{\nu}_\mu$ events from the Monte Carlo simulations. These 
include the Cerenkov threshold for muons. They are given for 18 bins of 
$\mathrm{\Delta E=0.1\, GeV}$ between 0.2~GeV and 2.0~GeV (The list of these energy dependent
efficiecies is given in the files {\tt NUMUeff.dat} and {\tt NUMUBAReff.dat}).
Furthermore, he tells you that the rejection factor for the NC events should be $10^-3$.
This is all the information needed to reconstruct smearing matrices out of the data
that can be used to describe the experiment in an AEDL file.

\aufg{1}{Warm-Up}

The smearing matrices are provided in the files {\tt CCmigration.dat}, {\tt CCBARmigration.dat},
and {\tt NCmigration.dat}. They are already given in the right syntax, so that they can be
implemented into an AEDL file. However, try to qualitatively understand, how these matrices 
can be reconstructed from the data given above. If you are not familiar with the usage of
manually defined smearing matrices, please consult the GLoBES manual on pages 101-102.

\aufg{2}{Finding the Normalization}

Another colleague of yours provides you with the data of the neutrino beam flux. This is given
in the files {\tt TOYplus.dat} ($\nu_e$) and {\tt TOYminus.dat} ($\bar{\nu}_e$), already provided 
in the right syntax for GLoBES. The implementation of neutrino fluxes from a data file is described in the 
GLoBES manual on pages 89-90.

\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=0.45\textwidth]{FLUX.eps} 
  \end{center}
  \vspace{-0.7 cm}
  \caption{\small The neutrino flux of the toy experiment as a function of the neutrino energy.
  The flux is given in arbitrary units so the right normalization has to be found.}
  \label{fig:flux}
\end{figure}

Unfortunately, your colleague does not know the units of the flux data anymore, but he 
tells you that his Monte Carlo simulation showed that in case of no oscillations 
he would get at total number of 13503.5 $\nu_e$ charged current events and 
4612.18 $\bar{\nu}_e$ charged current events per kt and year at the planned baseline 
of L~=~350~km. He tells you that they estimated a new average matter density along the
baseline of $\rho=2.95\,g/cm^3$. \\

Start from the draft file {\tt TOYdraft.glb} and build an AEDL file that
uses the flux files and allows to calculate the unoscillated $\nu_e/\bar{\nu}_e$ CC events per kt and
year. Use 18 bins from 0.2~GeV to 2~GeV (necessary, since the migration matrices are given for this binning). 
Calculate the event rates with the
command {\tt globes} from the {\tt /source} directory. Within {\tt TOYdraft.glb}  
you can find some hints where the corresponding AEDL features are documented in the GLoBES
manual.

\aufg{3}{Describing the Experiment}

Now, that we have found the right normalization, we can describe the toy experiment. The detector will
be located at the baseline of L~=~350~km and have a fiducial mass of 500~kt. The experiment
will measure $\nu_\mu$-appearance and $\bar{\nu}_\mu$-appearance. For collecting comparable
amount of data in both channels, it is planned to have split mode of operation,~{\it i.e.}
2 years in $\nu_e$ mode and 6 years in $\bar{\nu}_e$ mode. It is assumed that the overall
normalization uncertainty of the signal events will be 2.5\% and 20\% on the background events. \\

Build up an AEDL file that decribes the experiment and makes use of the migration matrices.
Note, that the migration matrices are for CC and NC events. \\

Check your AEDL file with the {\tt globes} command for the following oscillation parameters:
\begin{eqnarray}
\sin^2\theta_{12} & = & 0.3 \nonumber \\
\sin^22\theta_{13} & = & 0.1  \nonumber \\
\sin^22\theta_{23} & = & 1  \nonumber \\
\delta_{CP} & = & \pi/2  \nonumber \\
\Delta^2_{21} & = & 7.9\cdot 10^{-5}\, eV^2  \nonumber \\
\Delta^2_{31} & = & 2.6\cdot 10^{-3}\, eV^2  
\end{eqnarray}
This can be done with the command 
\begin{center}
{\tt ./globes -p'0.5796,0.1609,0.7854,1.5708,7.9e-5,2.6e-3' NAME.glb}
\end{center}

This is what you should be obtained (signal and background event numbers):\\

{\tt 
\hspace*{2cm} --------- \#NU\_MU\_Appearance ---------\\
\hspace*{2cm}               233787 ||       233787\\
\hspace*{2cm}              4640.85 ||      4640.85\\
\hspace*{2cm} \\
\hspace*{2cm} \\
\hspace*{2cm} --------- \#NU\_MU\_BAR\_Appearance ---------\\
\hspace*{2cm}               173047 ||       173047\\
\hspace*{2cm}              4911.57 ||      4911.57\\
}

The binned event rates can be calculated with the command {\tt ./globes -s NAME.glb}. The screen output that should be
derived is given in the appendix.  \\

Furthermore, the program {\tt th13delta.c} and the script {\tt th13delta.gnuplot} can be used
to produce a contour plot of the best fit solution in the $\sin^22\theta_{13}$-$\delta_{CP}$-plane (there are still degenerate solutions left 
in different locations of the parameter space,~{\it i.e.} the intrinsic and the
sign-degeneracy). Do not forget to enter the Name of your AEDL file into {\tt th13delta.c}.
Note, that another AEDL file is loaded within {\tt th13delta.c}, {\tt T2K\_disappearance\_only.glb}. This is due to
the fact, that the leading atmospheric parameters cannot be measured at the toy experiment. This can spoil the
results through parameter correlations. Thus, the simulated disappearance data from T2K is included. The T2K appearance
data is missing, so that information on $\sin^22\theta_{13}$ and $\delta_{CP}$ can only be derived from the toy
experiment data. The effect can be seen in Fig.~\ref{fig:withoutT2K} in the appendix.

\aufg{4}{Experiment Description without Smearing Matrices}

One of your colleagues tells you, that the energy reconstruction of the quasi-elastic
events follows a gaussian energy resolution function with a width of $\sigma=0.085\, GeV$.
The smearing is due to Fermi Motion and so in his analysis he found that this width is
independent of the neutrino energy over the whole energy window up to 2 GeV. \\

Modify the AEDL file in such a way that no migration matrices are used anymore. You will need
the efficiency and background rejection data from the files {\tt NUMUeff.dat}, {\tt
NUMUBAReff.dat}, {\tt NC\_bckg\_rej.dat}, and {\tt NCBAR\_bckg\_rej.dat}. Make use of the
GLoBES built-in energy resolution function. Then, compare the event rates and the best-fit
solution you get from this file with the file from the former problem. You should approximately
end up with this output:\\

{\tt  
\hspace*{2cm} --------- \#NU\_MU\_Appearance --------- \\
\hspace*{2cm}               222833 ||       222833 \\
\hspace*{2cm}               4640.5 ||       4640.5 \\
\hspace*{2cm}  \\
\hspace*{2cm}  \\
\hspace*{2cm} --------- \#NU\_MU\_BAR\_Appearance --------- \\
\hspace*{2cm}               171644 ||       171644 \\
\hspace*{2cm}              4911.57 ||      4911.57 \\
}

The binned event rates again can be derived with the command {\tt ./globes -s NAME.glb}. The screen output 
that should be obtained is given in the appendix. \\

Now we can compare the best-fit solution from the file with migration matrices and the 
file without migration matrices:

\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=0.48\textwidth]{WithMatrices.eps}   
    \includegraphics[width=0.48\textwidth]{WithoutMatrices.eps} 
  \end{center}
  \vspace{-0.7 cm}
  \caption{\small Left-Hand Side: The allowed best-fit solution obtained with {\tt th13delta.c}
  and the toy experiment description using manually defined migration matrices. Right-Hand Side: 
  The allowed best-fit solution obtained with {\tt th13delta.c}
  and the toy experiment description using the GLoBES internal energy resolution function.}
  \label{fig:Comp}
\end{figure}

\aufg{5}{List Interpolation in AEDL}

The description of the experiment without migration matrices has one big advantage
in comparison to the description with migration matrices. Suppose, one of your colleagues
wants to compare your analyisis with your analysis. Unfortunately, he uses 16 bins instead of
18 and asks you to do the same. One could either do the same procedure again and prepare
completely new migration matrices, this time $16\times16$. But, this would of course be very
time consuming. Or one can use the former description and introduce the efficiencies 
by applying the AEDL feature of list interpolation as described in the GLoBES manual 
on pages 84-85. The result should of course look extremely similar.

\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=0.5\textwidth]{Bins.eps}   
      \end{center}
  \vspace{-0.7 cm}
  \caption{\small The allowed best-fit solution obtained with {\tt th13delta.c}
  and the toy experiment description using the GLoBES internal energy resolution function.
  and 16 bins with efficiencies obtained with the AEDL list interpolation feature.}
  \label{fig:Bins}
\end{figure}

\aufg{6}{Usage of Total Rates}

The former files only make use of the quasi-elastic event sample, since there energy
reconstruction is possible to a certain level. However, this reduces the statistics of the
experiment, because all inelastic events are omitted.\\

The CC current events can be included to the file and the systematics function
{\tt chiTotalRatesTilt} within the corresponding rule can be used to only count 
total rates. The spectral information still comes from the QE events. However, now double
counting of events has to be avoided. Therefore the QE events have to be taken into account
with a free normalization. The explanations about the needed systematics functions can be found
in the GLoBES manual on pages 105-106. You can also check the {\tt T2K.glb} file. There, this
description is already implemented. Please discuss the systematics function, that has to be
used for the QE events in case of no treatment of systematics.\\

In the end you should end up at:\\

{\tt
\hspace*{2cm} --------- \#NU\_MU\_Appearance\_QE ---------\\
\hspace*{2cm}               222706 ||       222706\\
\hspace*{2cm}              4840.44 ||      4840.44\\
\hspace*{2cm} \\
\hspace*{2cm} \\
\hspace*{2cm} --------- \#NU\_MU\_BAR\_Appearance\_QE ---------\\
\hspace*{2cm}               171630 ||       171630\\
\hspace*{2cm}              5169.42 ||      5169.42\\
\hspace*{2cm} \\
\hspace*{2cm} \\
\hspace*{2cm} --------- \#NU\_MU\_Appearance\_CC ---------\\
\hspace*{2cm}               379842 ||       379842\\
\hspace*{2cm}              4840.44 ||      4840.44\\
\hspace*{2cm} \\
\hspace*{2cm} \\
\hspace*{2cm} --------- \#NU\_MU\_BAR\_Appearance\_CC ---------\\
\hspace*{2cm}               265792 ||       265792\\
\hspace*{2cm}              5169.42 ||      5169.42\\
}

\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=0.5\textwidth]{CCtotal.eps}   
      \end{center}
  \vspace{-0.7 cm}
  \caption{\small The allowed best-fit solution obtained with {\tt th13delta.c}
  and the toy experiment description using the GLoBES internal energy resolution function.
  Here, the total rates data from all CC events and the spectral data from the QE sample
  with a free normalization is used within the analysis.}
  \label{fig:total}
\end{figure}

The difference to the former simulations is not very large. This is due to the fact that
the true value of $\sin^22\theta_{13}$ is chosen quite large. For smaller $\sin^22\theta_{13}$
the effect becomes more important, since overall statistics decrease and the additional statistics
improves the results in a more dramatic way.

% ----------------------------------------------------------------------------
\section*{Part 2: Features of the Pre-Defined AEDL Experiments}
% ----------------------------------------------------------------------------

Now, we can have a look at the pre-defined AEDL experiments. The problems of
Part 1 have made us familiar with the syntax of AEDL. One of the most important
AEDL features used in the variable AEDL files is the usage of AEDL variables.
For instance the the file {\tt data/NuFact/Variable/NFvar.glb} is defined with free
parameters. First, we can have a look at the implementation of the neutrino factory
flux. In contary to the toy experiment, the fluxes are not loaded from a data file, but
derived from the built-in flux feature of GLoBES. The parent energy and the baseline of the
experiment is kept free and can be specified by usage of the AEDL variables {\tt emax} and
{\tt BASELINE}.\\

Copy the file {\tt source/globes} to the directory with the {\tt NFvar.glb} file and try the
command:
\begin{center}
{\tt ./globes -DBASELINE=3000 -Demax=50 NFvar.glb}
\end{center}

Furthermore, the file {\tt NFvar.glb} makes use of the AEDL feature to allow variable
binwidths. The binwiths are chosen such that the bins at smaller energies are more narrow than
for larger energies. This is done, because the energy resolution is proportional to the
neutrino energy and a better resolution is obtained at smaller energies, which requires
smaller binning. However, the absolute position of the bins depends on {\tt emax} so that
it is not possible to give fixed efficiency lists. Here, the new interpolation feature
of GLoBES is necessary.\\

The implementation of a built-in beta beam flux can be explored
in the variable beta beam file {\tt data/BetaBeam/BB\_WC/BBvar\_WC.glb}. Here, also AEDL variables have been
introduced to allow an analysis of beta beams by modifying the $\gamma$ factor and the
baseline. Since for this experiment migration matrices cannot be used to maintain the
variability, the same technique as in Problem 6 is used.

\newpage
\begin{appendix}
\section{Effect of T2K Disappearance Data}

\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=0.5\textwidth]{WITHOUT_T2K.eps}   
      \end{center}
  \vspace{-0.7 cm}
  \caption{\small The allowed best-fit solution obtained with {\tt th13delta.c}
  and the toy experiment description using migration matrices. The disappearance data from
  T2K is {\bf not} used here.}
  \label{fig:withoutT2K}
\end{figure}

\section{Spectral Event Rates}

\footnotesize \tt

\begin {center}
\begin{tabular}{lcl}

{\large Migration Matrices} && {\large Energy Resolution Function} \\ \hline && \\ 
--------- \#NU\_MU\_Appearance ---------&&--------- \#NU\_MU\_Appearance ---------\\
&&\\
  0.25       3235.32&&  0.25       872.803\\
  0.35        8516.4&&  0.35       4161.83\\
  0.45       18670.5&&  0.45       13747.8\\
  0.55       30845.9&&  0.55       27084.3\\
  0.65         37326&&  0.65       35648.9\\
  0.75       36255.2&&  0.75       36404.7\\
  0.85       30234.2&&  0.85       32152.6\\
  0.95       23895.9&&  0.95       25359.3\\
  1.05         17472&&  1.05       18206.2\\
  1.15       11164.9&&  1.15       12112.7\\
  1.25       7213.56&&  1.25       7574.92\\
  1.35       4387.87&&  1.35       4499.13\\
  1.45        2290.2&&  1.45       2501.03\\
  1.55       1175.12&&  1.55        1317.7\\
  1.65       573.792&&  1.65       668.669\\
  1.75       287.307&&  1.75       317.866\\
  1.85       158.537&&  1.85       143.546\\
  1.95        83.849&&  1.95       58.5287\\
----------------------&&----------------------\\
Total:        233787&&Total:        222833\\
&&\\
  0.25       1899.59&&  0.25       1899.63\\
  0.35       1379.43&&  0.35       1379.05\\
  0.45       707.324&&  0.45       707.319\\
  0.55       347.565&&  0.55       347.565\\
  0.65       161.629&&  0.65       161.629\\
  0.75       75.6928&&  0.75       75.6928\\
  0.85       41.3034&&  0.85       41.3034\\
  0.95       17.3519&&  0.95       17.3519\\
  1.05        6.3722&&  1.05        6.3722\\
  1.15       3.16321&&  1.15       3.16321\\
  1.25       0.53009&&  1.25       0.53009\\
  1.35      0.598437&&  1.35      0.598437\\
  1.45             0&&  1.45   0.000143515\\
  1.55             0&&  1.55   8.79703e-05\\
  1.65             0&&  1.65   5.08887e-05\\
  1.75             0&&  1.75   2.78955e-05\\
  1.85      0.298004&&  1.85      0.298002\\
  1.95             0&&  1.95    7.1521e-06\\
----------------------&&----------------------\\
Total:       4640.85&&Total:        4640.5\\
\end{tabular}
\end{center}

\newpage

\begin {center}
\begin{tabular}{lcl}

{\large Migration Matrices} && {\large Energy Resolution Function} \\ \hline && \\ 
--------- \#NU\_MU\_BAR\_Appearance ---------&&--------- \#NU\_MU\_BAR\_Appearance ---------\\
&&\\
  0.25       1110.29&&  0.25       430.704\\
  0.35        3257.3&&  0.35       1595.73\\
  0.45       9112.41&&  0.45       5209.66\\
  0.55       15748.8&&  0.55         12256\\
  0.65         19876&&  0.65       20000.1\\
  0.75       24939.8&&  0.75       25053.5\\
  0.85       24582.1&&  0.85         25888\\
  0.95       20723.2&&  0.95       23282.3\\
  1.05       17965.5&&  1.05       19168.5\\
  1.15       13890.8&&  1.15       14457.6\\
  1.25       8845.95&&  1.25       10004.5\\
  1.35       5449.44&&  1.35       6349.17\\
  1.45       3585.94&&  1.45       3729.31\\
  1.55       1870.27&&  1.55       2099.19\\
  1.65       1027.29&&  1.65       1137.16\\
  1.75       559.153&&  1.75       582.734\\
  1.85       330.635&&  1.85        280.22\\
  1.95       172.192&&  1.95       119.721\\
----------------------&&----------------------\\
Total:        173047&&Total:        171644\\
&&\\
  0.25       1882.64&&  0.25       1882.64\\
  0.35        1436.5&&  0.35       1436.51\\
  0.45       790.672&&  0.45       790.666\\
  0.55       412.492&&  0.55       412.492\\
  0.65        201.96&&  0.65        201.96\\
  0.75       95.6634&&  0.75       95.6634\\
  0.85       53.6873&&  0.85       53.6873\\
  0.95       22.8585&&  0.95       22.8585\\
  1.05       8.63062&&  1.05       8.63062\\
  1.15       4.40415&&  1.15       4.40415\\
  1.25      0.765549&&  1.25      0.765549\\
  1.35      0.843628&&  1.35      0.843628\\
  1.45             0&&  1.45    0.00019704\\
  1.55             0&&  1.55   0.000124013\\
  1.65             0&&  1.65   7.34927e-05\\
  1.75             0&&  1.75   4.11518e-05\\
  1.85      0.455293&&  1.85      0.455293\\
  1.95             0&&  1.95    1.0927e-05\\
----------------------&&----------------------\\
Total:       4911.57&&Total:       4911.57

\end{tabular}
\end{center}

\end{appendix}

\end{document}

### 2007-01-24 Documentation/tutorials/AEDL-tutorial/source/AEDL-tutorial.tex
*** empty log message ***

@@
 The toy experiment for this tutorial is absolutely fictious, although the attributes 
-of the detector follow roughly the response of a huge Water Cerenkov detector.
+of the detector roughly follow the response of a huge Water Cerenkov detector.
 However, all efficiencies and migration matrices are coming from a fake 
@@
 attributes of a Water Cerenkov detector. So, all files prepared for this tutorial 
-do {\bf not} correspond to a real detector response and all files are just to 
+do {\bf not} correspond to a real detector response and are just to 
 be used for this tutorial.  
@@
 in AEDL,~{\it i.e.} write the corresponding {\tt glb}-file. One of your colleagues
-sends the results from a Monte Carlo analysis. 
+sends the results from a Monte Carlo analysis (see Fig.~\ref{fig:mccc}). 
+%
 \begin{figure}[h!]
@@
   \vspace{-0.7 cm}
-  \caption{\small The sample of 20000 charged current $\nu_\mu$ events on the left hand side. The
+  \caption{\small The sample of 20000 charged current $\nu_\mu$ events on the left-hand side. The
   quasi-elastic sample is shown in dark grey while the inelastic part of the sample is
@@
 \end{figure}
-
+%
 It is also important to implement the background from single pion production in 
-neutral current events. Therefore, the colleague sends the Monte Carlo sample of NC events.
-
+neutral current events. Therefore, the colleague sends the Monte Carlo sample of NC events
+(see Fig.~\ref{fig:mcnc}). 
+%
 \begin{figure}[h!]
   \begin{center}
-    \includegraphics[width=0.45\textwidth]{NCevents.eps} 
+    \includegraphics[width=0.36\textwidth]{NCevents.eps} 
   \end{center}
@@
 \end{figure}
-
-The colleague also sends the energy dependent signal efficiencies of the charged
+%
+Additionally you receive the energy dependent signal efficiencies of the charged
 current $\nu_\mu$/$\bar{\nu}_\mu$ events from the Monte Carlo simulations. These 
-include the Cerenkov threshold for muons. They are given for 18 bins of 
-$\mathrm{\Delta E=0.1\, GeV}$ between 0.2~GeV and 2.0~GeV (The list of these energy dependent
+include the Cerenkov threshold for muons and are given for 18 bins of 
+$\mathrm{\Delta E=0.1\, GeV}$ between 0.2~GeV and 2.0~GeV (the list of these energy dependent
 efficiecies is given in the files {\tt NUMUeff.dat} and {\tt NUMUBAReff.dat}).
-Furthermore, he tells you that the rejection factor for the NC events should be $10^-3$.
+Furthermore, he tells you that the rejection factor for the NC events should be $10^{-3}$.
 This is all the information needed to reconstruct smearing matrices out of the data
@@
 
-Another colleague of yours provides you with the data of the neutrino beam flux. This is given
+Another colleague of yours provides you with the data of the neutrino beam flux. It is given
 in the files {\tt TOYplus.dat} ($\nu_e$) and {\tt TOYminus.dat} ($\bar{\nu}_e$), already provided 
@@
 GLoBES manual on pages 89-90.
-
+%
 \begin{figure}[h!]
@@
 \end{figure}
-
+%
 Unfortunately, your colleague does not know the units of the flux data anymore, but he 
@@
 Calculate the event rates with the
-command {\tt globes} from the {\tt /source} directory. Within {\tt TOYdraft.glb}  
+command {\tt globes} from the {\tt /source} directory of GLoBES. Within {\tt TOYdraft.glb}  
 you can find some hints where the corresponding AEDL features are documented in the GLoBES
@@
 \end{eqnarray}
-This can be done with the command 
+This can either be done with the command 
 \begin{center}
@@
 \end{center}
+or just with {\tt ./globes NAME.glb} within a shell session if the values are defined as the default parameters:
+ \begin{center}
+{\tt export GLB\_CENTRAL\_VALUES='0.5796,0.1609,0.7854,1.5708,7.9e-5,2.6e-3'}
+\end{center}
 
-This is what you should be obtained (signal and background event numbers):\\
+This is what approximately should be obtained (signal and background event numbers):\\
 
@@
 
+\newpage
 \aufg{4}{Experiment Description without Smearing Matrices}
@@
 
-The binned event rates again can be derived with the command {\tt ./globes -s NAME.glb}. The screen output 
-that should be obtained is given in the appendix. \\
+The 
+binned event rates again can be derived with the command {\tt ./globes -s NAME.glb}. The screen output 
+that should be obtained is given in the appendix.
 
@@
 file without migration matrices:
-
+%
 \begin{figure}[h!]
@@
 on pages 84-85. The result should of course look extremely similar.
-
+%
 \begin{figure}[h!]
@@
 
-The former files only make use of the quasi-elastic event sample, since there energy
+The former files only make use of the quasi-elastic event sample, there energy
 reconstruction is possible to a certain level. However, this reduces the statistics of the
@@
 
-The CC current events can be included to the file and the systematics function
-{\tt chiTotalRatesTilt} within the corresponding rule can be used to only count 
+The CC events can be included to the file and the systematics function
+{\tt chiTotalRatesTilt} within the corresponding rule can be used to only account for 
 total rates. The spectral information still comes from the QE events. However, now double
 counting of events has to be avoided. Therefore the QE events have to be taken into account
-with a free normalization. The explanations about the needed systematics functions can be found
-in the GLoBES manual on pages 105-106. You can also check the {\tt T2K.glb} file. There, this
-description is already implemented. Please discuss the systematics function, that has to be
-used for the QE events in case of no treatment of systematics.\\
+with a free normalization. The explanations about the necessary systematics functions can be found
+in the GLoBES manual on pages 105-106. You can also explore the {\tt T2K.glb} file. There, this
+technique is already implemented. Please discuss the systematics function, that has to be
+used for the QE events in case of no treatment of systematics.
 
+\newpage
 In the end you should end up at:\\
@@
 AEDL features used in the variable AEDL files is the usage of AEDL variables.
-For instance the the file {\tt data/NuFact/Variable/NFvar.glb} is defined with free
-parameters. First, we can have a look at the implementation of the neutrino factory
+For instance the file {\tt data/NuFact/Variable/NFvar.glb} is defined with free
+variables. First, we can have a look at the implementation of the neutrino factory
 flux. In contary to the toy experiment, the fluxes are not loaded from a data file, but
 derived from the built-in flux feature of GLoBES. The parent energy and the baseline of the
-experiment is kept free and can be specified by usage of the AEDL variables {\tt emax} and
+experiment are kept free and can be specified by the usage of the AEDL variables {\tt emax} and
 {\tt BASELINE}.\\

### 2007-02-09 Documentation/part1.tex
*** empty log message ***

@@
 
+\vspace*{-0.3cm}
+
 \begin{quote}{\tt \footnotesize
-
 double likelihood(double true\_rate, double fit\_rate, double sqr\_sigma) \\
@@
 
-double chiDCNorm(int exp, int rule, int np, double *x, double *errors) \\
+double chiDCNorm(int exp, int rule, int np, double *x, double *errors, \\
+  \hspace*{0.3cm}  void* user\_data) \\
 \{ \\
@@
   double chiDCNorm(int exp, int rule, int dim, double *x, \\
-  \hspace*{0.5cm} double *errors) \\
+  \hspace*{0.5cm} double *errors, void* user\_data) \\
   \{ \\
@@
   \hspace*{0.5cm} glbInit(agrv[0]); \\
-  \hspace*{0.5cm} glbDefineChiFunction(\&chiDCNorm,5,"chiDCNorm"); \\
+  \hspace*{0.5cm} glbDefineChiFunction(\&chiDCNorm,5,"chiDCNorm",NULL); \\
   \hspace*{0.5cm} ... \\
@@
 \hspace*{0.5cm}        @sys\_on\_function  = "chiDCNorm" \\
-\hspace*{0.5cm}        @sys\_on\_errors    = \{ 0.02,    0.006,        0.006,       0.005,     0.005   \} \\
+\hspace*{0.5cm}        @sys\_on\_errors    = \{ 0.028,    0.006,        0.006,       0.005,     0.005   \} \\
 \hspace*{0.5cm}        \mbox{ // \{ Flux, Fid. mass FD, Fid. mass ND, Energy FD, Energy ND \} } \\

### 2007-03-29 Distribution/source/glb_parser.c
JK - Fixed bug in glb_likelihood (true and fitted rates exchanged)

@@ -1,7 +1,7 @@
-/* A Bison parser, made by GNU Bison 1.875.  */
+/* A Bison parser, made by GNU Bison 2.1.  */
 
 /* Skeleton parser for Yacc-like parsing with Bison,
-   Copyright (C) 1984, 1989, 1990, 2000, 2001, 2002 Free Software Foundation, Inc.
+   Copyright (C) 1984, 1989, 1990, 2000, 2001, 2002, 2003, 2004, 2005 Free Software Foundation, Inc.
 
    This program is free software; you can redistribute it and/or modify
    it under the terms of the GNU General Public License as published by
@@ -15,8 +15,8 @@
 
    You should have received a copy of the GNU General Public License
    along with this program; if not, write to the Free Software
-   Foundation, Inc., 59 Temple Place - Suite 330,
-   Boston, MA 02111-1307, USA.  */
+   Foundation, Inc., 51 Franklin Street, Fifth Floor,
+   Boston, MA 02110-1301, USA.  */
 
 /* As a special exception, when this file is copied by Bison into a
    Bison output file, you may use that output file without restriction.
@@ -35,6 +35,9 @@
 
 /* Identify Bison output.  */
 #define YYBISON 1
+
+/* Bison version.  */
+#define YYBISON_VERSION "2.1"
 
 /* Skeleton name.  */
 #define YYSKELETON_NAME "yacc.c"
@@ -87,6 +90,7 @@
      NEG = 289
    };
 #endif
+/* Tokens.  */
 #define NUM 258
 #define SFNCT 259
 #define BOGUS 260
@@ -1135,6 +1139,11 @@ static int set_exp_energy(char *name, glb_List **value)
 # define YYERROR_VERBOSE 0
 #endif
 
+/* Enabling the token table.  */
+#ifndef YYTOKEN_TABLE
+# define YYTOKEN_TABLE 0
+#endif
+
 #if ! defined (YYSTYPE) && ! defined (YYSTYPE_IS_DECLARED)
 #line 1022 "glb_parser.y"
 typedef union YYSTYPE {
@@ -1148,8 +1157,8 @@ typedef union YYSTYPE {
   int in;
   glb_namerec *nameptr;
 } YYSTYPE;
-/* Line 191 of yacc.c.  */
[…]

### 2007-03-29 Distribution/source/glb_sys.c

@@ -724,10 +724,23 @@ int glbSetSysStartingValuesListInRule(int exp, int rule, int on_off, const doubl
  ***************************************************************************/
 inline double glb_likelihood(double true_rate, double fit_rate)
 {
+  double res; 
+  res = fit_rate - true_rate;
+  if (true_rate > 0)
+  {
+    if (fit_rate <= 0.0) 
+      res = 1e100;
+    else
+      res += true_rate * log(true_rate/fit_rate);
+  }
+  return 2.0 * res;
+  
+/* JK - 29.03.2007
   if(fit_rate > 0)
     return 2*(true_rate - fit_rate) + 2*fit_rate * log(fit_rate/true_rate);
   else
     return 2*(true_rate - fit_rate);
+*/
 }
 
 /* Alternative version of likelihood, using Gauss distribution */

### 2007-05-11 Documentation/Paper2/GLoBESCPC.tex
*** empty log message ***

%% This `backbone' file can be used to write a paper for
%% Computer Physics Communications using LaTeX.
%% For authors who want to write a computer program description,
%% an example Program Summary is included that only has to be
%% completed and which will give the correct layout in the
%% preprint and the journal.
%% The `elsart' style that is used and more information on this style
%% can be found at the Author Gateway (http://authors.elsevier.com) and follow %% the link to "Guide to publishing with Elsevier".
\documentclass{elsart}

%% This list environment is used for the references in the
%% Program Summary
%%
\newcounter{bla}
\newenvironment{refnummer}{%
\list{[\arabic{bla}]}%
{\usecounter{bla}%
 \setlength{\itemindent}{0pt}%
 \setlength{\topsep}{0pt}%
 \setlength{\itemsep}{0pt}%
 \setlength{\labelsep}{2pt}%
 \setlength{\listparindent}{0pt}%
 \settowidth{\labelwidth}{[9]}%
 \setlength{\leftmargin}{\labelwidth}%
 \addtolength{\leftmargin}{\labelsep}%
 \setlength{\rightmargin}{0pt}}}
 {\endlist}

\begin{document}
\begin{frontmatter}

\title{GLoBES: General Long Baseline Experiment Simulator}

\author[a]{Patrick Huber},
\author[b]{Joachim Kopp},
\author[b]{Manfred Lindner},
\author[c]{Mark Rolinec},
\author[d]{Walter Winter}

\address[a]{University of Wisconsin, Physics Department, 1150 University Av. Madison, WI 53706, USA}
\address[b]{Max-Planck-Institut f\"{u}r Kernphysik, Postfach 10 39 80, D-69029 Heidelberg, Germany}
\address[c]{Technische Universit\"{a}t M\"{u}nchen, Institut f\"{u}r
Theoretische Physik, Physik-Department, James-Franck-Strasse,
D-85748 Garching, Germany}
\address[d]{Universit\"{a}t W\"{u}rzburg, Lehrstuhl f\"{u}r theoretische Physik II, Institut f\"{u}r theoretische Physik und Astrophysik, Am Hubland, D-97074 W\"{u}rzburg, Germany}

\begin{abstract}
GLoBES (General Long Baseline Experiment Simulator) is a flexible software package to simulate
neutrino oscillation long baseline and reactor experiments. On the one hand, it contains a
comprehensive abstract experiment definition language (AEDL), which allows to describe most classes of
long baseline experiments at an abstract level. On the other hand, it provides a C-library to process
the experiment information in order to obtain oscillation probabilities, rate vectors, and
$\Delta_{\chi^{2}}$-values. Currently, GLoBES is available for GNU/Linux. Since the source code is
included, the port to other operating systems is in principle possible.

GLoBES is an open source code that has previously been described in Computer Physics Communications
167~(2005)~195 and in this issue, xx (2007) yyy). The source code and a comprehensive User Manual for
GLoBES v3.0.8 is now available from the CPC Program Library as described in the Program Summary below.
The home of GLobES is http://www.mpi-hd.mpg.de/$\sim$globes/.

\begin{flushleft}

PACS: 14.60.Pq

\end{flushleft}

\begin{keyword}
Neutrino oscillations; Long-baseline experiment; GLoBES;
\end{keyword}

\end{abstract}

\end{frontmatter}

% Computer program descriptions should contain the following
% PROGRAM SUMMARY.

{\bf PROGRAM SUMMARY}
  %Delete as appropriate.

\begin{small}
\noindent
{\em Manuscript Title:} GLoBES: General Long Baseline Experiment Simulator\\
{\em Authors:} Patrick Huber, Joachim Kopp, Manfred Lindner, Mark Rolinec, Walter Winter\\
{\em Program Title:} GLoBES version 3.0.8\\
{\em Journal Reference:}                                      \\
  %Leave blank, supplied by Elsevier.
{\em Catalogue identifier:}                                   \\
  %Leave blank, supplied by Elsevier.
{\em Licensing provisions:} none                                  \\
  %enter "none" if CPC non-profit use license is sufficient.
{\em Programming language:} C                                  \\
{\em Computer:} GLoBES builds and installs on 32bit and 64bit Linux systems.\\
  %Computer(s) for which program has been designed.
{\em Operating system:} 32bit or 64bit Linux                                      \\
  %Operating system(s) for which program has been designed.
{\em RAM:} Typically a few MBs                                         \\
  %RAM in bytes required to execute program with typical data.
{\em Keywords:} Neutrino oscillations, Long-baseline experiment,
GLoBES \\
{\em PACS:} 14.60Pq    \\
{\em Classification:} 11.1  General, High Energy Physics and Computing, 
  11.7 Detector Design and Simulation, 11.10 Accelerators and Particle Beams    \\
  %Classify using CPC Program Library Subject Index, see (
  % http://cpc.cs.qub.ac.uk/subjectIndex/SUBJECT_index.html)
  %e.g. 4.4 Feynman diagrams, 5 Computer Algebra.
{\em External routines/libraries:} GSL - The GNU Scientific Library,
www.gnu.org/software/gsl/\\

{\em Nature of problem:}\\
  Neutrino oscillations are now established as the leading flavor
transition mechanism for neutrinos. In a long history of many experiments, see, e.g., [1], two
oscillation frequencies have been identified: The fast atmospheric and the slow solar
oscillations, which are driven by the respective mass squared differences. In addition, there could be
interference effects between these two oscillations, provided that the coupling given by the small
mixing angle $\theta_{13}$ is large enough. Such interference effects include, for example, leptonic
CP violation. In order to test the unknown oscillation parameters, i.e. the mixing angle
$\theta_{13}$, the leptonic CP phase, and the neutrino mass hierarchy, new long-baseline and reactor
experiments are proposed. These experiments send an artificial neutrino beam to a detector, or detect
the neutrinos produced by a nuclear fission reactor. However, the presence of multiple solutions which
are intrinsic to neutrino oscillation probabilities [2-5] affect these measurements. Thus optimization
strategies are required which maximally exploit complementarity between experiments. Therefore, a
modern, complete experiment simulation and analysis tool does not only need to have a highly accurate
beam and detector simulation, but also powerful means to analyze correlations and degeneracies,
especially for the combination of several experiments. The GLoBES software package is such a
tool~[6-7].
   \\
{\em Solution method:}\\
   GLoBES is a flexible software tool to simulate and analyze
neutrino oscillation long-baseline and reactor experiments using a
complete three-flavor description. On the one hand, it contains a
comprehensive abstract experiment definition language (AEDL), which
makes it possible to describe most classes of long baseline and
reactor experiments at an abstract level. On the other hand, it
provides a C-library to process the experiment information in order
to obtain oscillation probabilities, rate vectors, and
$\Delta {\chi^{2}}$-values. In addition, it provides a binary
program to test experiment definitions very quickly, before they are
used by the application software.
   \\
{\em Restrictions:}\\
  Currently restriced to discrete sets of sources and detectors. For example, the
simulation of an atmospheric neutrino flux is not supported. 
  %Describe any restrictions on the complexity of the problem here.
   \\
{\em Unusual features:}\\
   Clear separation between experiment description and the simulation software.
  %Describe any unusual features of the program/problem here.
   \\
{\em Additional comments:}\\
   To find information on the latest version of the software and user manual, please check the author's
  web site,  http://www.mpi-hd.mpg.de/$\sim$globes
   \\
{\em Running time:}\\
The examples included in the distribution take only a few minutes to
complete. More sophisticated problems can take up to several days.
  %Give an indication of the typical running time here.
   \\
{\em References:}
\begin{refnummer}
\item V. Barger, D. Marfatia, K. Whisnant, Int. J. Mod. Phys. E 12(2003)569, and references therein,
hep-ph/0308123.
\item G.L. Fogli, E. Lisi, Phys. Rev. D 54 (1996) 3667, hepph/
9604415.
\item J. Burguet-Castell, M.B. Gavela, J.J. Gomez-Cadenas, P. Hernandez,
O. Mena, Nucl. Phys. B 608 (2001) 301, hep-ph/ 0103258.
\item H. Minakata, H. Nunokawa, JHEP 0110 (2001) 001, hep-ph/
0108085.
\item V. Barger, D. Marfatia, K. Whisnant, Phys. Rev. D 65 (2002)
073023, hep-ph/0112119.
\item P. Huber, M. Lindner and W. Winter, Comput. Phys. Commun. 167 (2005) 195.
\item Patrick Huber, Joachim Kopp, Manfred Lindner, Mark Rolinec, Walter Winter, Comput. Phys. Commun. xxx (2007) yyy.
\end{refnummer}

\end{small}

\newpage

\end{document}

### 2007-11-08 Documentation/Manual.tex
Another test

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\usepackage{ae,aecompl} % only for Type 1 fonts for readable pdf convert
+\usepackage{ae,aecompl} % only for Type 1 fonts for readable pdf convert; may not be necessary anymore!
 
 \newcommand{\glbxxx}{hep-ph/0407333}
-\newcommand{\glbxxxtwo}{hep-ph/XXYYZZZ}
+\newcommand{\glbxxxtwo}{hep-ph/0701187}
 \newcommand{\capdef}{}
@@
 P. Huber, M. Lindner and W. Winter\\
-Simulation of long baseline neutrino oscillation experiments with \GLOBES\\
+Simulation of long baseline neutrino oscillation experiments with \GLOBES ,\\
 Comput. Phys. Commun. 167 (2005) 195, arXiv:\glbxxx,
@@
 P. Huber, J. Kopp, M. Lindner, M. Rolinec, and W. Winter\\
-TO BE DEFINED, NEEDS TO BE UPDATED [WW]\\
+ New features in the simulation of neutrino oscillation experiments with \GLOBES~3.0,
 arXiv:\glbxxxtwo,

### 2010-05-07 Documentation/tutorials/LBNE-tutorial/tex/LBNE-tutorial.tex
- Added LBNE tutorial to Documentation/tutorials folder

\documentclass[12pt,a4paper]{article}

\usepackage{a4wide}
\usepackage{graphicx}
\usepackage{amsmath,amssymb}
\usepackage{ae,aecompl}
\usepackage{amsthm}

%\pagestyle{empty}
\setlength{\parindent}{0pt}
\renewcommand{\labelenumi}{\alph{enumi})}
\renewcommand{\labelenumii}{\roman{enumii})}

% Neutrino oscillation channels
\newcommand{\ee}{\ensuremath{\nu_e \rightarrow \nu_e}}
\newcommand{\emu}{\ensuremath{\nu_e \rightarrow \nu_\mu}}
\newcommand{\etau}{\ensuremath{\nu_e \rightarrow \nu_\tau}}
\newcommand{\mue}{\ensuremath{\nu_\mu \rightarrow \nu_e}}
\newcommand{\mumu}{\ensuremath{\nu_\mu \rightarrow \nu_\mu}}
\newcommand{\mutau}{\ensuremath{\nu_\mu \rightarrow \nu_\tau}}
\newcommand{\taue}{\ensuremath{\nu_\tau \rightarrow \nu_e}}
\newcommand{\taumu}{\ensuremath{\nu_\tau \rightarrow \nu_\mu}}
\newcommand{\tautau}{\ensuremath{\nu_\tau \rightarrow \nu_\tau}}

% Neutrino oscillation parameters
\newcommand{\sthsol}{\ensuremath{\sin^2 2\theta_{12}}}     % Mixing angles
\newcommand{\sthchooz}{\ensuremath{\sin^2 2\theta_{13}}}
\newcommand{\sthatm}{\ensuremath{\sin^2 2\theta_{23}}}
\newcommand{\sthtf}{\ensuremath{\sin^2 2\theta_{2f}}}
\newcommand{\ldm}{\ensuremath{{\Delta m_{31}^2}}}          % Mass squared differences
\newcommand{\sdm}{\ensuremath{{\Delta m_{21}^2}}}

% Important neutrino experiments
\newcommand{\TtoK}{{\sffamily T2K}}

% Miscellaneous commands
\newcommand{\aufg}[2]{\vspace{4mm}{\bf\underline{Task #1:} {#2}} \vspace{3mm}}

% Environment for API reference entries
\newtheoremstyle{dotless}{}{}{\itshape}{}{\bfseries}{}{0cm}{}
\theoremstyle{dotless}
\newtheorem*{function}{}

\begin{document}
\vspace*{-3cm}
\begin{center}
{\large\bf GLoBES Tutorial: Implementing LBNE}\\[0.3cm]
GLoBES mini-workshop at Fermilab
\end{center}
\vspace*{2mm}

Joachim Kopp \hfill May 13, 2010

\bigskip
\hrule
\vspace*{4mm}

{\small
In this tutorial, we will implement LBNE (with a Water Cherenkov detector) in
GLoBES, starting from the existing AEDL (Abstract Experiment Definition
Language) file for NOvA. We will see how fluxes, efficiencies, smearing
matrices, etc.\ taken from Monte Carlo simulation can be used as input for
GLoBES. In the second part, we will discuss strategies for reliably finding
degenerate solutions in parameter space.
}

% ----------------------------------------------------------------------------
\section*{Part 1: An AEDL file for LBNE}
% ----------------------------------------------------------------------------

\aufg{1}{Warm-up: Familiarize yourself with the NOvA simulation}

Look at the file NOvA.glb: The first part defines the properties of the beam;
the spectrum is read from two external files {\tt NOvAplus.dat} and {\tt
NOvAminus.dat}. Below, we define the target mass, the binning used in the
analysis, and the baseline. The energy resolution function has the simple form
$\sigma_E = 0.1 \sqrt{E/\text{GeV}}$~GeV for $\nu_e$ and $\sigma_E = 0.05
\sqrt{E/\text{GeV}}$~GeV for $\nu_\mu$. Next comes the definition of the
various CC and NC oscillation channels, which are finally combined in
\emph{rules}, i.e.\ realistic event samples consisting of signal and background
contributions and having systematic uncertainties (normalization and spectral
tilt) associated with them.

Compile ({\tt make}) and run the program {\tt th13delta} to compute the
$\theta_{13}$ and $\delta_{CP}$ sensitivity of NOvA.%
\footnote{Note that, to speed things up, this program does not take into
account all parameter correlations and should therefore not be used for real
production runs unless appropriate modifications are made.} You can plot the
output of {\tt th13delta} using gnuplot (call the script {\tt
./th13delta.gnuplot}) or your favorite plotting program.

Familiarize yourself with the source code {\tt th13delta.c}, which
illustrates the basic structure of all GLoBES application programs:
After initialization of the library and definition of the true parameter
values and external priors, the code computes the ``true'' event rates
and then performs a $\chi^2$ fit. For each test value in the $\theta_{13}$-$\delta_{CP}$
plane, it computes the corresponding $\chi^2$, marginalizing over (some of) the other
oscillation parameters, as specified by the ``projection'' object
{\tt th13delta\_projection}.

\aufg{2}{Implenting the basic parameters of LBNE}

Let us now begin to implement LBNE. While modifying the AEDL files, you can use
the command {\tt globes LBNE.glb} to see the predicted total event rates for
each rule.  (Use {\tt globes -?} for a full list of command line options.)

For simplicity, we will only implement the case where the target is operated in
neutrino mode.  Therefore, copy {\tt NOvA.glb} to a new file {\tt LBNE.glb},
and remove everything that is related to anti-neutrino running (flux, channel,
and rule definitions).  Then, change the beam spectrum for the neutrino mode to
the simluted spectrum of LBNE:
\begin{verbatim}
  dusel120e250i002dr280dz1300km\_flux.txt
\end{verbatim}
(thanks, Mary, for providing the flux files!).  Next, we have to properly
normalize the flux by choosing appropriate values for {\tt @time}, {\tt
@power}, and {\tt @norm}. (Note that it is only important that the product of
these numbers is correct.) Choose {\tt @time = 3}~[yrs], {\tt @power =
7.3}~[$\times 10^{20}$~pot/yr], and {\tt @norm = $1.0544 \times 10^{17}$}.%
\footnote{In principle, it is possible to compute {\tt @norm}
from first principles (see GLoBES manual, appendix C), but often it is easier
to choose a value that yields a predetermined number of events.}

Next, adjust the target mass to 100~kt, the baseline to 1\,300~km and the
energy range to 0.5--60~GeV (use 476 bins, i.e.\ a bin width of 0.125~GeV).
Don't forget to also adjust {\tt \$sampling\_min}, {\tt \$sampling\_max}, and
{\tt \$sampling\_points}, which give the binning before smearing. Use again the
energy range 0.05--60~GeV, divided into 476 bins (boundary effects are
negligible). Remove the {\tt @energy\_window} directives which would limit the
analysis to a smaller energy range.

For the energy resolution in a WC detector, it is appropriate to use a constant
value of 0.085 GeV ({\tt @sigma\_e = \{0.0,0.0,0.085\}}) (we will proceed to
more realistic energy smearing later).

Reliable reconstruction of the neutrino energy is easiest for quasi-elastic
events, therefore we will only include quasi-elastic events in the
disappearance channel, which is not statistics-limited. Consequently, include a
new set of cross sections, {\tt XQE.dat}, in your glb file, define a new
channel {\tt \#nu\_mu\_QE}, and use this channel instead of {\tt \#nu\_mu\_CC}
in rule {\tt \#Nu\_Mu\_Disappearance}.

Finally, change the efficiencies in the rules to the values given in
Table~\ref{tab:eff} (these are effective constant efficiencies, chosen such
that the number of events in all signal and background contributions matches
the results of the full LBNE simulation), and the systematic uncertainties to
the values given in Table~\ref{tab:sys}.  Since spectral errors are negligible
in this simulation, you can choose the faster $\chi^2$ function {\tt
chiSpectrumTilt} instead of the slower {\tt chiSpectrumCalib}.

Use the {\tt th13delta} program to see the effect of your modifications.

\begin{table}
  \centering
  \begin{tabular}{llc}
    \hline
    Rule                               & Channel                    & Efficiency \\
    \hline
    {\tt \#Nu\_E\_Appearance}          & {\tt \#nu\_e\_signal}      & 0.137      \\
                                       & {\tt \#nu\_mu\_CC}         & $2.3 \times 10^{-4}$ \\
                                       & {\tt \#nu\_mu\_NC}         & 0.0034     \\
                                       & {\tt \#nu\_e\_beam}        & 0.096      \\\hline
    {\tt \#Nu\_Mu\_Disappearance}      & {\tt \#nu\_mu\_CC}         & 0.973      \\
                                       & {\tt \#nu\_mu\_NC}         & 0.034      \\\hline
    \hline
  \end{tabular}
  \caption{Efficiencies for LBNE (taken from the T2K simulation).}
  \label{tab:eff}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{lcccc}
    \hline
    Rule                               & Signal norm. & Signal tilt & BG norm & BG tilt \\
    \hline
    {\tt \#Nu\_E\_Appearance}          & 0.01         & 0.0001      & 0.05    & 0.0001  \\
    {\tt \#Nu\_Mu\_Disappearance}      & 0.01         & 0.0001      & 0.10    & 0.0001  \\
    \hline
  \end{tabular}
  \caption{Systematic uncertainties in LBNE.}
  \label{tab:sys}
\end{table}

\aufg{3}{Advanced AEDL: Variable bin widths, energy-dependent efficiencies,
         and user-defined smearing matrices}

To keep event numbers in each bin reasonably large (and to improve computational
efficiency), it can be helpful to use a coarser binning in energy intervals where
only few events are expected. For LBNE, we use 0.125~GeV wide bins between
$E_\nu = 0.5$~GeV and $E_\nu = 8$~GeV, and 2~GeV wide bins above. Replace the
{\tt \$bins} and {\tt \$sampling\_points} statements by
\begin{verbatim}
$binsize = {0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125,
 0.125, 0.125,0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125,
 0.125, 0.125,0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125,
 0.125, 0.125,0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125,
 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125,
 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 2, 2, 2, 2, 2, 2, 2, 2, 2,
 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 4}
\end{verbatim}
and
\begin{verbatim}
$sampling_stepsize= {0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125,
 0.125, 0.125, 0.125,0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125,
 0.125, 0.125, 0.125,0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125,
 0.125, 0.125, 0.125,0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125,
 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125,
 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 2, 2, 2, 2, 2, 2, 2,
 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2}
\end{verbatim}
respectively. (To avoid the typing, copy these statements from {\tt
LBNE-2.glb}.) Use the {\tt globes} command to check that the total event rates
(especially those in the appearance channels) did not change significantly.

The next important modification is to incorporate the energy dependence of the
detection efficiencies. GLoBES uses two different types of efficiencies:
pre-smearing efficiencies $\epsilon_{\rm pre}(E')$ and post-smearing efficiencies
$\epsilon_{\rm post}(E)$, and we will use both of them. The event rate is
proportional to
\begin{align}
  N(E) = \int\! dE' \, \epsilon_{\rm post}(E) \times R(E, E') \times
         \epsilon_{\rm pre}(E') \,,
\end{align}
where $R(E, E')$ is the smearing matrix that maps the neutrino energy $E'$ to
the reconstructed energy $E$. The AEDL commands to define pre- and
post-smearing efficiencies are {\tt @pre\_smearing\_efficiencies} and {\tt
@post\_smearing\_efficiencies}, respectively. Additionally the coefficients in
the {\tt @signal} and {\tt @background} directives can be used to implement
energy-independent contributions to the post-smearing efficiencies, as we have
done above.

Copy the efficiency vectors from LBNE-2.glb to your own file, and change the
coefficients of the corresponding channels in the {\tt @signal} and
{\tt @background} directives to 1.0 (all efficiency effects are included in
the energy-dependent efficiency vectors, so these constant factors are
no longer necessary).

The final step is to introduce smearing matrices obtained from detector
Monte Carlo simulations instead of the simple Gaussians provided by
GLoBES. A typical smearing matrix has the form
\begin{align}
  \underbrace{
  \begin{pmatrix}
    a_{00} & a_{01} & a_{02} & a_{03} &  &  &  &  &  & \\
    a_{10} & a_{11} & a_{12} & a_{13} & a_{14} &  &  &  &  &  \\
           & a_{21} & a_{22} & a_{23} & a_{24} & a_{25} &  &  &  &  \\
           &  & a_{32} & a_{33} & a_{34} & a_{35} & a_{36} &  &  &  \\
           &  &  & a_{43} & a_{44} & a_{45} & a_{46} & a_{47} &  &  \\
           & & & \uparrow  & & \ddots & & \uparrow & & \\
           & & & k_l^i & & & & k_u^i & &  
  \end{pmatrix}
   }_{\mathtt{\$sampling\_points}  \, \, \text{columns}}
   \quad \leftarrow \quad \text{\footnotesize{{\tt \$bins} rows}} \,,
\end{align}
where we use $k_l^i$ and $k_u^i$ denote the column indices of the
first and last nonzero entries in the $i$th row.
Due to the many zeros in each row, GLoBES uses the following syntax
for specifying smearing matrices:
\begin{tabbing}
  \tt energy(\#name\_of\_smearing\_matrix)<   \\
  \ \ \tt @energy = \=\tt\{ $k^0_l$, $k^0_u$, $a_{0 k^0_l}$, \dots,  $a_{0 k^0_u}$ \}: \\
  \>  \tt                \{ $k^1_l$, $k^1_u$, $a_{1 k^1_l}$, \dots,  $a_{1 k^1_u}$ \}: \\
  \>  \tt             \vdots \\
  \tt >
\end{tabbing}
For this tutorial, we provide smearing matrices in this format
in the files listed in table~\ref{tab:smearing}. Use the {\tt include}
directive (synatx {\tt include "filename"}) to incorporate them
into your AEDL file, and change the references to the smearing matrices
in the {\tt channel} definitions according to table~\ref{tab:smearing}.
Note that the neutral current backgrounds to the $\nu_e$ and $\nu_\mu$
channels have different smearing matrices, so you will have to
create a new channel {\tt \#nu\_mu\_bg\_QE} for the background in the
disappearance channel. Use a constant efficiency 0.204 for this channel.

\begin{table}
  \centering
  \begin{tabular}{llp{6.5cm}}
    \hline
    File                  & Name of smearing matrix & used in channels \\ \hline
    \tt smear\_signal.dat & \tt \#signal            & \tt \#nu\_e\_signal,
                                                          \#nu\_e\_beam \\
    \tt smear\_disapp.dat & \tt \#dis               & \tt \#nu\_mu\_CC,
                                                          \#nu\_mu\_QE \\
    \tt smear\_nc.dat     & \tt \#nc                & \tt \#nu\_mu\_NC \\
    \tt smear\_mpip.dat   & \tt \#mpip              & \tt \#nu\_mu\_bg\_QE \\
    \hline
  \end{tabular}
  \caption{Smearing matrices for the LBNE simulation}
  \label{tab:smearing}
\end{table}

% ----------------------------------------------------------------------------
\section*{Part 2: Finding degenerate solutions in parameter space}
% ----------------------------------------------------------------------------

To determine the allowed ranges for a given oscillation parameter $x$, GLoBES
minimizes $\chi^2$ over all other parameters, while keeping $x$ fixed. This is
done for many values of $x$, so that $\chi^2$ is obtained as a function of the
test value for $x$.  The minimization over oscillation parameters is done using
a \emph{local} minimization algorithm. If the $\chi^2$ manifold has several
local minima (``degenerate solutions''), this minimizer may not find the
\emph{global} minimum $\chi^2$, resulting in overly optimistic sensitivity
estimates.

Use the program {\tt th13chi2}, and the script {\tt th13chi2.gnuplot} to see
an example of this effect. {\tt th13chi2} computes the $\chi^2$ for the wrong
hierarchy solution as a function of $\sin^2 2\theta_{13}$ for a simple neutrino
factory setup. (We use a neutrino factory as an example here, because degenerate
solutions are more of a problem there --- but you should worry about them also
in a superbeam experiment!) It is obvious from the resulting plot,
fig.~\ref{fig:th13chi2-1} (left panel), that something is going wrong at $\sthchooz \sim
0.002$. The right panel of fig.~\ref{fig:th13chi2-1} shows why this happens.
(This plot has been generated with a program similar to {\tt th13delta}.) For
$\sthchooz \gtrsim 0.001$, there are several local dips in $\chi^2$ at
$\delta_{CP} \sim 200-350^\circ$, and for the starting value of the minimization
chosen in our code, $\delta_{CP} = 45^\circ$, the minimizer happens to converge
into the wrong minimum.

\begin{figure}
  \begin{center}
    \includegraphics[height=5cm]{th13chi2-1.eps}
    \raisebox{-.5cm}{\includegraphics[height=6cm]{th13delta-NF.eps}}
  \end{center}
  \caption{Left: $\chi^2$ as a function of $\sthchooz$ for the wrong
    (inverted) hierarchy solution in a neutrino factory. Right: $\chi^2$
    contours in the $\sthchooz$--$\delta_{CP}$ plane.}
  \label{fig:th13chi2-1}
\end{figure}

\aufg{4}{Choosing better starting values}

The above discussion already suggests a solution to the degeneracy problem in
this particular case: Choose more appropriate starting values for the
minimization.  Use the function {\tt glbSetOscParams} to set $\delta_{CP}$ to
$\pi/2$ in {\tt test\_values} before calling {\tt glbChiTheta13}. The result,
shown in fig.~\ref{fig:th13chi2-2}, shows that this solves the problem in this
particular case.

\begin{figure}
  \begin{center}
    \includegraphics[height=5cm]{th13chi2-2.eps}
  \end{center}
  \caption{Solution to problems 4 and 5: The correct $\chi^2$ function for the
    wrong hierarchy solution at the neutrino factory.}
  \label{fig:th13chi2-2}
\end{figure}

\aufg{5}{A tracking algorithm}

Choosing appropriate starting values for the minimization is not a very
robust solution to the degeneracy problem. In particular, it will fail once we
choose different ``true'' parameter values, but do not adjust our starting
values. (Try choosing $\delta_{CP}^{\rm true} = \tfrac{5}{4} \pi$ to see this.)
A more robust approach is the so-called tracking algorithm: When scanning over
one parameter ($\sthchooz$ in our case), use the location of the minimum
found for one value of $\sthchooz$ as the starting point for the minimization
for the next $\sthchooz$.

To implement this algorithm, modify {\tt th13chi2.c} by allocationg an
additional parameter vector, called e.g.\ {\tt loc\_min}, and pass it to
{\tt glbChiTheta13} as its second argument. It will be filled with the location
of the minimum. In the next iteration, use {\tt loc\_min} to initialize
{\tt test\_values}. Your code should now produce a result like
fig.~\ref{fig:th13chi2-2}, but will still work when you change the ``true''
parameter values, e.g.\ to $\delta_{CP}^{\rm true} = \tfrac{5}{4} \pi$.

\aufg{6}{{\tt degfinder}, a flexible degeneracy finder using pre-scanning}

One of the most elegant solution to the problem of degeneracy finding is to
perform a quick and dirty preliminary scan of the relevant parts of parameter
space to get the rough locations of the degenerate solutions, and then use
these as starting points for local minimization. Here, quick and dirty means
that no minimization is performed over less relevant parameters (like
$\theta_{12}$ and $\sdm$), systematical errors are not taken into account, etc.
While it is straightforward to implement this algorithm, we will here use an
existing routine called {\tt degfinder}.

To use {\tt degfinder}, the following steps are necessary:
\begin{enumerate}
  \item Define the parameters that should be included in the preliminary
    scan of the parameter space, and the ranges over which you want to
    scan. In our case, we are only going to scane over $\delta_{CP}$, and
    we are going to use 8 steps (i.e.\ a step size of $\pi/4$.
    \begin{verbatim}
int n_prescan_params = 1;              /*# of params for prescan    */
int prescan_params[] = {GLB_DELTA_CP}; /*List of params to scan over*/
double prescan_min[] = {0.0         }; /*Lower edge of scan range   */
double prescan_max[] = {2*M_PI      }; /*Upper edge of scan range   */
int prescan_steps[]  = {8           }; /*Number of steps            */
    \end{verbatim}

  \item Define the {\tt glb\_projection} for the preliminary scan and
    for the finial fit. In our case, we will keep all parameters
    fixed in the prescan, and marginalize over all of them (except,
    of course, $\theta_{13}$), in the final fit.
    \begin{verbatim}
glb_projection prescan_proj = glbAllocProjection();
glbDefineProjection(prescan_proj, GLB_FIXED, GLB_FIXED, GLB_FIXED,
                                  GLB_FIXED, GLB_FIXED, GLB_FIXED);
glbSetDensityProjectionFlag(prescan_proj, GLB_FIXED, GLB_ALL);

glb_projection proj = glbAllocProjection();
glbDefineProjection(proj, GLB_FREE, GLB_FIXED, GLB_FREE,
                          GLB_FREE, GLB_FREE, GLB_FREE);
glbSetDensityProjectionFlag(proj, GLB_FREE, GLB_ALL);
    \end{verbatim}

  \item Allocate data structures where {\tt degfinder} can store the
    locations of the minima and the corresponding $\chi^2$ values.
    \begin{verbatim}
const int max_deg = 32;          /* Max # of degeneracies expected */
int n_deg;                       /* # of degeneracies found        */
double deg_chi2[max_deg];        /* chi^2 values at degeneracies   */
glb_params deg_pos[max_deg];     /* Locations of degeneracies      */
for (i=0; i < max_deg; i++)
  deg_pos[i] = glbAllocParams();
    \end{verbatim}

  \item Call {\tt degfinder} where you would normally call {\tt glbChiNP}, {\tt
    glbChiTheta13}, etc. In particular, make sure that {\tt
    glbSetCentralValues}, {\tt glbSetInputErrors} (defining external priors),
    {\tt glbSetOscillationParameters}, and {\tt glbSetRates} (computing
    ``true'' event rates) are called before any invocation of {\tt degfinder}.
    \begin{verbatim}
degfinder(test_values, n_prescan_params, prescan_params,
          prescan_min, prescan_max, prescan_steps,
          prescan_proj, proj, &n_deg, deg_pos, deg_chi2,
          DEG_NO_NH);    
    \end{verbatim}
    Besides the arguments discussed above, {\tt degfinder} accepts the
    following flags (which can be combined using binary OR):

    \begin{tabular}{lp{12cm}}
      \tt DEG\_NO\_NH    & Ignore normal hierarchy solutions in prescan. Note
                           that there is a small danger that the final minimization
                           might still converge into a normal hierarchy solution.    \\
      \tt DEG\_NO\_IH    & Ignore normal hierarchy solutions in prescan. Note
                           that there is a small danger that the final minimization
                           might still converge into an inverted hierarchy solution. \\
      \tt DEG\_NO\_SYS   & Switch off systematics in final fit. (Systematics are alway
                           off during prescan.)                                      \\
      \tt DEG\_NO\_CORR  & Switch off all parameter correlations, i.e.\ do not
                           minimize over oscillation parameters and matter density.
                           This overrides any contrary setting in {\tt prescan\_proj}
                           or {\tt proj}.
    \end{tabular}

  \item Evaluate the results from {\tt degfinder}
    \begin{verbatim}
double chi2 = 1.0e10;
for (i=0; i < n_deg; i++)
{     
  if (glbGetOscParams(deg_pos[i], GLB_DM_31) < 0.0)
  { 
    if (deg_chi2[i] < chi2)
      chi2 = deg_chi2[i];
  } 
  else
    fprintf(stderr, "Error: Minimizer converged into NH minimum."
                    "This should not have happened.\n");
}     
    \end{verbatim}

  \item Clean up
    \begin{verbatim}
for (i=0; i < max_deg; i++)
  if (deg_pos[i] != NULL)  { glbFreeParams(deg_pos[i]); deg_pos[i] = NULL; }
glbFreeProjection(proj);
glbFreeProjection(prescan_proj);    
    \end{verbatim}
\end{enumerate}

If you run the resulting program, you should again obtain a result similar to
fig.~\ref{fig:th13chi2-2}. By adding a few lines of debug output to your code,
you can easily verify that, for $\sthchooz \gtrsim 0.002$, {\tt degfinder}
indeed finds two local minima, but picks the correct one in the end.

\end{document}

### 2010-05-07 Documentation/tutorials/LBNE-tutorial/degfinder.c

@@ -1,0 +1,427 @@
+/* ---------------------------------------------------------------------------- */
+/* Finding degeneracies in neutrino oscillations                                */
+/* ---------------------------------------------------------------------------- */
+/* Author: Joachim Kopp (partly based on ideas by P. Huber and W. Winter)       */
+/* ---------------------------------------------------------------------------- */
+#include <stdlib.h>
+#include <stdio.h>
+#include <string.h>
+#include <math.h>
+
+#include <globes/globes.h>
+#include "degfinder.h"
+
+/* Global variables */
+const int debug_level = 0;
+
+#define MIN(X,Y)    ( ((X) < (Y)) ? (X) : (Y) )
+#define MAX(X,Y)    ( ((X) > (Y)) ? (X) : (Y) )
+
+
+/* ---------------------------------------------------------------------------- */
+double ChiNPWrapper(glb_params base_values, double th12, double th13, double th23,
+                    double delta, double dm21, double dm31, int hierarchy,
+                    glb_params fit_values)
+/* ---------------------------------------------------------------------------- */
+/* Perform minimization for a certain set of test values                        */
+/* ---------------------------------------------------------------------------- */
+{
+  double result;
+  glb_params tv     = glbAllocParams();      /* Test values             */
+  glb_params cv     = glbAllocParams();      /* Central values          */
+  glb_params old_cv = glbAllocParams();      /* Previous central values */
+  glbCopyParams(base_values, tv);
+  glbGetCentralValues(cv);
+  glbGetCentralValues(old_cv);
+
+  if (!GLB_ISNAN(th12))
+    glbSetOscParams(tv, th12, GLB_THETA_12);
+  if (!GLB_ISNAN(th13))
+    glbSetOscParams(tv, th13, GLB_THETA_13);
+  if (!GLB_ISNAN(th23))
+    glbSetOscParams(tv, th23, GLB_THETA_23);
+  if (!GLB_ISNAN(delta))
+    glbSetOscParams(tv, delta, GLB_DELTA_CP);
+  if (!GLB_ISNAN(dm21))
+    glbSetOscParams(tv, dm21, GLB_DM_21);
+  if (!GLB_ISNAN(dm31))
[…]

### 2010-09-18 Documentation/tutorials/norm.tex
meeting-URL corrected. -ML

@@
 the number of target nuclei $\tau$ (or protons or whatever applies to the
-given cross section) per unit target mass $m_u$ (which usually is {\it
+given cross section\footnote{The cross section which come with
+  \GLOBES\ always are per nucleon.}) per unit target mass $m_u$ (which usually is {\it
   kt}), is dealt with in the
@@
 \begin{eqnarray}
-(\mathtt{@norm})^{-1}=5.20034\left(\frac{\Delta E}{\mathrm{GeV}}\right)\left(\frac{A}{\mathrm{cm}^2}\right)\left(\frac{\mathrm{km}}{L}\right)^2\left(\frac{\tau}{m_u}\right)\times10^{-38}\times\left(\frac{\mathcal{L}}{\mathcal{L}_u}\right)
+\mathtt{@norm}=\frac{1}{5.20034}\left(\frac{\mathrm{GeV}}{\Delta
+      E}\right)
+\left(\frac{\mathrm{cm}^2}{A}\right)\left(\frac{L}{\mathrm{km}}\right)^2\left(\frac{\tau}{m_u}\right)\times10^{-38}\times\left(\frac{\mathcal{L}_u}{\mathcal{L}}\right)
 \end{eqnarray}

### 2010-09-18 Distribution/config.h

@@ -1,920 +1,0 @@
-/* config.h.  Generated from config-h.in by configure.  */
-/* config-h.in.  Generated from configure.in by autoheader.  */
-
-/* Define if the compiler is building for multiple architectures of Apple
-   platforms at once. */
-/* #undef AA_APPLE_UNIVERSAL_BUILD */
-
-/* Define to the number of bits in type 'ptrdiff_t'. */
-/* #undef BITSIZEOF_PTRDIFF_T */
-
-/* Define to the number of bits in type 'sig_atomic_t'. */
-/* #undef BITSIZEOF_SIG_ATOMIC_T */
-
-/* Define to the number of bits in type 'size_t'. */
-/* #undef BITSIZEOF_SIZE_T */
-
-/* Define to the number of bits in type 'wchar_t'. */
-/* #undef BITSIZEOF_WCHAR_T */
-
-/* Define to the number of bits in type 'wint_t'. */
-/* #undef BITSIZEOF_WINT_T */
-
-/* Define to one of `_getb67', `GETB67', `getb67' for Cray-2 and Cray-YMP
-   systems. This function is required for `alloca.c' support on those systems.
-   */
-/* #undef CRAY_STACKSEG_END */
-
-/* Define to 1 if using `alloca.c'. */
-/* #undef C_ALLOCA */
-
-/* Define to 1 if // is a file system root distinct from /. */
-/* #undef DOUBLE_SLASH_IS_DISTINCT_ROOT */
-
-/* Define on systems for which file names may have a so-called `drive letter'
-   prefix, define this to compute the length of that prefix, including the
-   colon. */
-#define FILE_SYSTEM_ACCEPTS_DRIVE_LETTER_PREFIX 0
-
-/* Define if the backslash character may also serve as a file name component
-   separator. */
-#define FILE_SYSTEM_BACKSLASH_IS_FILE_NAME_SEPARATOR 0
-
-/* Define if a drive letter prefix denotes a relative path if it is not
-   followed by a file name component separator. */
-#define FILE_SYSTEM_DRIVE_PREFIX_CAN_BE_RELATIVE 0
-
-/* The version of libglb in libtool format c:r:a */
-#define GLB_LIBVERSION "9:0:1"
-
-/* Define to indicate the 'malloc' module. */
-#define GNULIB_MALLOC_GNU 1
-
[…]

### 2010-09-18 Distribution/globes/globes.h

@@ -59,7 +59,7 @@ enum glb_enum_efficiency_types
 
 /* Available minimization algorithms */
 enum glb_enum_minimizers
-       { GLB_MIN_NESTED_POWELL, GLB_MIN_POWELL, GLB_MIN_SIMAN };
+       { GLB_MIN_NESTED_POWELL, GLB_MIN_POWELL };
 #define GLB_MIN_DEFAULT  GLB_MIN_NESTED_POWELL
 
 
@@ -76,12 +76,6 @@ enum glb_enum_minimizers
 #define GLB_EV_TO_KM_FACTOR  1.9747235e-10
 #define GLB_KM_TO_EV(x)      ((x) / GLB_EV_TO_KM_FACTOR)
 #define GLB_EV_TO_KM(x)      ((x) * GLB_EV_TO_KM_FACTOR)
-
-
-/* maximum number of experiments */
-
-#define GLB_MAX_EXP 300
-
 
 
 
@@ -119,9 +113,9 @@ extern int glb_single_experiment_number;
  * declare this one to SWIG and the other global 
  * variables not
  */
-glb_exp glb_experiment_list[GLB_MAX_EXP];
+glb_exp glb_experiment_list[32];
 #else
-extern glb_exp glb_experiment_list[GLB_MAX_EXP];
+extern glb_exp glb_experiment_list[32];
 #endif
 extern int glb_rule_number;
 
@@ -139,7 +133,6 @@ void glbInit(char *name);
 void glbClearExperimentList();
 void glbDefineAEDLVariable(const char *name, double value);
 void glbDefineAEDLList(const char *name, double *list, size_t length);
-double glbGetAEDLVariable(const char *name);
 void glbClearAEDLVariables();
 int glbInitExperiment(char *inf, glb_exp *in, int *counter);
 
@@ -243,8 +236,6 @@ int glbSetSourcePower(int experiment, int flux_ident, double power);
 int glbSetRunningTime(int experiment, int flux_ident, double time);
 int glbSetFilterStateInExperiment(int experiment,int on_off);
 int glbSetFilterInExperiment(int experiment,double filter);
-int glbCompensateFilterInExperiment(int experiment);
-int glbOptimizeSmearingMatrixInExperiment(int experiment);
 
 double glbGetTargetMass(int experiment);
 double glbGetSourcePower(int experiment, int flux_ident);
@@ -254,7 +245,6 @@ double glbGetFilterInExperiment(int experiment);
 int glbGetEminEmax(int experiment, double *emin, double *emax);
[…]

### 2010-09-18 Distribution/globes/globes_perl.c

@@ -1,0 +1,5746 @@
+/* ----------------------------------------------------------------------------
+ * This file was automatically generated by SWIG (http://www.swig.org).
+ * Version 1.3.21
+ * 
+ * This file is not intended to be easily readable and contains a number of 
+ * coding conventions designed to improve portability and efficiency. Do not make
+ * changes to this file unless you know what you are doing--modify the SWIG 
+ * interface file instead. 
+ * ----------------------------------------------------------------------------- */
+
+/*************************************************************** -*- c -*-
+ * perl5/precommon.swg
+ *
+ * Rename all exported symbols from common.swg, to avoid symbol
+ * clashes if multiple interpreters are included
+ *
+ ************************************************************************/
+
+#define SWIG_TypeRegister    SWIG_Perl_TypeRegister
+#define SWIG_TypeCheck       SWIG_Perl_TypeCheck
+#define SWIG_TypeCast        SWIG_Perl_TypeCast
+#define SWIG_TypeDynamicCast SWIG_Perl_TypeDynamicCast
+#define SWIG_TypeName        SWIG_Perl_TypeName
+#define SWIG_TypeQuery       SWIG_Perl_TypeQuery
+#define SWIG_TypeClientData  SWIG_Perl_TypeClientData
+#define SWIG_PackData        SWIG_Perl_PackData 
+#define SWIG_UnpackData      SWIG_Perl_UnpackData 
+
+
+/***********************************************************************
+ * common.swg
+ *
+ *     This file contains generic SWIG runtime support for pointer
+ *     type checking as well as a few commonly used macros to control
+ *     external linkage.
+ *
+ * Author : David Beazley (beazley@cs.uchicago.edu)
+ *
+ * Copyright (c) 1999-2000, The University of Chicago
+ * 
+ * This file may be freely redistributed without license or fee provided
+ * this copyright message remains intact.
[…]

### 2010-09-18 Distribution/lib/alloca.c

@@ -21,7 +21,9 @@
    allocating any.  It is a good idea to use alloca(0) in
    your main control loop, etc. to force garbage collection.  */
 
-#include <config.h>
+#ifdef HAVE_CONFIG_H
+# include <config.h>
+#endif
 
 #include <alloca.h>
 
@@ -81,37 +83,37 @@ long i00afunc ();
    STACK_DIRECTION = 0 => direction of growth unknown  */
 
 #  ifndef STACK_DIRECTION
-#   define STACK_DIRECTION      0       /* Direction unknown.  */
+#   define STACK_DIRECTION	0	/* Direction unknown.  */
 #  endif
 
 #  if STACK_DIRECTION != 0
 
-#   define STACK_DIR    STACK_DIRECTION /* Known at compile-time.  */
+#   define STACK_DIR	STACK_DIRECTION	/* Known at compile-time.  */
 
 #  else /* STACK_DIRECTION == 0; need run-time code.  */
 
-static int stack_dir;           /* 1 or -1 once known.  */
-#   define STACK_DIR    stack_dir
+static int stack_dir;		/* 1 or -1 once known.  */
+#   define STACK_DIR	stack_dir
 
 static void
 find_stack_direction (void)
 {
-  static char *addr = NULL;     /* Address of first `dummy', once known.  */
-  auto char dummy;              /* To get stack address.  */
+  static char *addr = NULL;	/* Address of first `dummy', once known.  */
+  auto char dummy;		/* To get stack address.  */
 
   if (addr == NULL)
-    {                           /* Initial entry.  */
+    {				/* Initial entry.  */
       addr = ADDRESS_FUNCTION (dummy);
 
-      find_stack_direction ();  /* Recurse once.  */
+      find_stack_direction ();	/* Recurse once.  */
     }
   else
     {
       /* Second entry.  */
       if (ADDRESS_FUNCTION (dummy) > addr)
-        stack_dir = 1;          /* Stack grew upward.  */
+	stack_dir = 1;		/* Stack grew upward.  */
       else
-        stack_dir = -1;         /* Stack grew downward.  */
+	stack_dir = -1;		/* Stack grew downward.  */
     }
 }
 
@@ -124,21 +126,21 @@ find_stack_direction (void)
    It is very important that sizeof(header) agree with malloc
    alignment chunk size.  The following default should work okay.  */
 
[…]

### 2010-09-18 Distribution/lib/argp-fs-xinl.c

@@ -1,28 +1,29 @@
 /* Real definitions for extern inline functions in argp-fmtstream.h
-   Copyright (C) 1997, 2003, 2004, 2009, 2010 Free Software Foundation, Inc.
+   Copyright (C) 1997, 2003 Free Software Foundation, Inc.
    This file is part of the GNU C Library.
    Written by Miles Bader <miles@gnu.ai.mit.edu>.
 
-   This program is free software: you can redistribute it and/or modify
+   This program is free software; you can redistribute it and/or modify
    it under the terms of the GNU General Public License as published by
-   the Free Software Foundation; either version 3 of the License, or
-   (at your option) any later version.
+   the Free Software Foundation; either version 2, or (at your option)
+   any later version.
 
    This program is distributed in the hope that it will be useful,
    but WITHOUT ANY WARRANTY; without even the implied warranty of
    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
    GNU General Public License for more details.
 
-   You should have received a copy of the GNU General Public License
-   along with this program.  If not, see <http://www.gnu.org/licenses/>.  */
+   You should have received a copy of the GNU General Public License along
+   with this program; if not, write to the Free Software Foundation,
+   Inc., 59 Temple Place - Suite 330, Boston, MA 02111-1307, USA.  */
 
 #ifdef HAVE_CONFIG_H
-# include <config.h>
+#include <config.h>
 #endif
 
 #define ARGP_FS_EI
 #undef __OPTIMIZE__
-#define __OPTIMIZE__ 1
+#define __OPTIMIZE__
 #include "argp-fmtstream.h"
 
 #if 0

### 2010-09-18 Distribution/lib/argp-namefrob.h

@@ -1,20 +1,21 @@
 /* Name frobnication for compiling argp outside of glibc
-   Copyright (C) 1997, 2003, 2007, 2009, 2010 Free Software Foundation, Inc.
+   Copyright (C) 1997, 2003 Free Software Foundation, Inc.
    This file is part of the GNU C Library.
    Written by Miles Bader <miles@gnu.ai.mit.edu>.
 
-   This program is free software: you can redistribute it and/or modify
+   This program is free software; you can redistribute it and/or modify
    it under the terms of the GNU General Public License as published by
-   the Free Software Foundation; either version 3 of the License, or
-   (at your option) any later version.
+   the Free Software Foundation; either version 2, or (at your option)
+   any later version.
 
    This program is distributed in the hope that it will be useful,
    but WITHOUT ANY WARRANTY; without even the implied warranty of
    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
    GNU General Public License for more details.
 
-   You should have received a copy of the GNU General Public License
-   along with this program.  If not, see <http://www.gnu.org/licenses/>.  */
+   You should have received a copy of the GNU General Public License along
+   with this program; if not, write to the Free Software Foundation,
+   Inc., 59 Temple Place - Suite 330, Boston, MA 02111-1307, USA.  */
 
 #if !_LIBC
 /* This code is written for inclusion in gnu-libc, and uses names in the
@@ -74,6 +75,11 @@
 #define __argp_fmtstream_rmargin argp_fmtstream_rmargin
 #undef __argp_fmtstream_wmargin
 #define __argp_fmtstream_wmargin argp_fmtstream_wmargin
+
+#include "mempcpy.h"
+#include "strcase.h"
+#include "strchrnul.h"
+#include "strndup.h"
 
 /* normal libc functions we call */
 #undef __flockfile
@@ -135,23 +141,16 @@
 # define putchar_unlocked(x) putchar (x)
 # endif
 
+extern char *__argp_basename (char *name);
+
 #endif /* !_LIBC */
 
 #ifndef __set_errno
 #define __set_errno(e) (errno = (e))
 #endif
 
-#if defined GNULIB_ARGP_DISABLE_DIRNAME
-# define __argp_base_name(arg) arg
-#elif defined GNULIB_ARGP_EXTERN_BASENAME
-extern char *__argp_base_name (const char *arg);
-#else
-# include "dirname.h"
[…]

### 2010-09-18 Distribution/lib/argp-parse.c

@@ -1,27 +1,27 @@
 /* Hierarchial argument parsing, layered over getopt
-   Copyright (C) 1995-2000, 2002-2004, 2009-2010 Free Software Foundation, Inc.
+   Copyright (C) 1995-2000, 2002, 2003, 2004 Free Software Foundation, Inc.
    This file is part of the GNU C Library.
    Written by Miles Bader <miles@gnu.ai.mit.edu>.
 
-   This program is free software: you can redistribute it and/or modify
+   This program is free software; you can redistribute it and/or modify
    it under the terms of the GNU General Public License as published by
-   the Free Software Foundation; either version 3 of the License, or
-   (at your option) any later version.
+   the Free Software Foundation; either version 2, or (at your option)
+   any later version.
 
    This program is distributed in the hope that it will be useful,
    but WITHOUT ANY WARRANTY; without even the implied warranty of
    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
    GNU General Public License for more details.
 
-   You should have received a copy of the GNU General Public License
-   along with this program.  If not, see <http://www.gnu.org/licenses/>.  */
+   You should have received a copy of the GNU General Public License along
+   with this program; if not, write to the Free Software Foundation,
+   Inc., 59 Temple Place - Suite 330, Boston, MA 02111-1307, USA.  */
 
 #ifdef HAVE_CONFIG_H
-# include <config.h>
+#include <config.h>
 #endif
 
 #include <alloca.h>
-#include <stddef.h>
 #include <stdlib.h>
 #include <string.h>
 #include <unistd.h>
@@ -37,18 +37,15 @@
 #else
 # include "gettext.h"
 #endif
-#define N_(msgid) msgid
+#define N_(msgid) (msgid)
 
 #include "argp.h"
 #include "argp-namefrob.h"
 
-#define alignof(type) offsetof (struct { char c; type x; }, x)
-#define alignto(n, d) ((((n) + (d) - 1) / (d)) * (d))
-
 /* Getopt return values.  */
-#define KEY_END (-1)            /* The end of the options.  */
-#define KEY_ARG 1               /* A non-option argument.  */
-#define KEY_ERR '?'             /* An error parsing the options.  */
+#define KEY_END (-1)		/* The end of the options.  */
+#define KEY_ARG 1		/* A non-option argument.  */
+#define KEY_ERR '?'		/* An error parsing the options.  */
 
[…]

### 2010-09-18 Distribution/lib/argp-xinl.c

@@ -1,23 +1,24 @@
 /* Real definitions for extern inline functions in argp.h
-   Copyright (C) 1997, 1998, 2004, 2009, 2010 Free Software Foundation, Inc.
+   Copyright (C) 1997, 1998 Free Software Foundation, Inc.
    This file is part of the GNU C Library.
    Written by Miles Bader <miles@gnu.ai.mit.edu>.
 
-   This program is free software: you can redistribute it and/or modify
+   This program is free software; you can redistribute it and/or modify
    it under the terms of the GNU General Public License as published by
-   the Free Software Foundation; either version 3 of the License, or
-   (at your option) any later version.
+   the Free Software Foundation; either version 2, or (at your option)
+   any later version.
 
    This program is distributed in the hope that it will be useful,
    but WITHOUT ANY WARRANTY; without even the implied warranty of
    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
    GNU General Public License for more details.
 
-   You should have received a copy of the GNU General Public License
-   along with this program.  If not, see <http://www.gnu.org/licenses/>.  */
+   You should have received a copy of the GNU General Public License along
+   with this program; if not, write to the Free Software Foundation,
+   Inc., 59 Temple Place - Suite 330, Boston, MA 02111-1307, USA.  */
 
 #ifdef HAVE_CONFIG_H
-# include <config.h>
+#include <config.h>
 #endif
 
 #if defined _LIBC || defined HAVE_FEATURES_H
@@ -25,11 +26,11 @@
 #endif
 
 #ifndef __USE_EXTERN_INLINES
-# define __USE_EXTERN_INLINES   1
+# define __USE_EXTERN_INLINES	1
 #endif
 #define ARGP_EI
 #undef __OPTIMIZE__
-#define __OPTIMIZE__ 1
+#define __OPTIMIZE__
 #include "argp.h"
 
 /* Add weak aliases.  */

### 2010-09-18 Distribution/lib/argp.h

@@ -1,20 +1,21 @@
 /* Hierarchial argument parsing, layered over getopt.
-   Copyright (C) 1995-1999, 2003-2010 Free Software Foundation, Inc.
+   Copyright (C) 1995-1999,2003,2004 Free Software Foundation, Inc.
    This file is part of the GNU C Library.
    Written by Miles Bader <miles@gnu.ai.mit.edu>.
 
-   This program is free software: you can redistribute it and/or modify
+   This program is free software; you can redistribute it and/or modify
    it under the terms of the GNU General Public License as published by
-   the Free Software Foundation; either version 3 of the License, or
-   (at your option) any later version.
+   the Free Software Foundation; either version 2, or (at your option)
+   any later version.
 
    This program is distributed in the hope that it will be useful,
    but WITHOUT ANY WARRANTY; without even the implied warranty of
    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
    GNU General Public License for more details.
 
-   You should have received a copy of the GNU General Public License
-   along with this program.  If not, see <http://www.gnu.org/licenses/>.  */
+   You should have received a copy of the GNU General Public License along
+   with this program; if not, write to the Free Software Foundation,
+   Inc., 59 Temple Place - Suite 330, Boston, MA 02111-1307, USA.  */
 
 #ifndef _ARGP_H
 #define _ARGP_H
@@ -22,10 +23,13 @@
 #include <stdio.h>
 #include <ctype.h>
 #include <getopt.h>
-#include <limits.h>
 
 #define __need_error_t
 #include <errno.h>
+
+#ifndef __const
+# define __const const
+#endif
 
 #ifndef __THROW
 # define __THROW
@@ -35,24 +39,23 @@
 #endif
 
 #ifndef __attribute__
-/* The __attribute__ feature is available in gcc versions 2.5 and later.
-   The __-protected variants of the attributes 'format' and 'printf' are
-   accepted by gcc versions 2.6.4 (effectively 2.7) and later.
-   We enable __attribute__ only if these are supported too, because
-   gnulib and libintl do '#define printf __printf__' when they override
-   the 'printf' function.  */
-# if __GNUC__ < 2 || (__GNUC__ == 2 && __GNUC_MINOR__ < 7)
-#  define __attribute__(Spec)   /* empty */
+/* This feature is available in gcc versions 2.5 and later.  */
[…]

### 2010-09-18 Distribution/lib/exitfail.c

@@ -1,12 +1,11 @@
 /* Failure exit status
 
-   Copyright (C) 2002, 2003, 2005, 2006, 2007, 2009, 2010 Free Software
-   Foundation, Inc.
+   Copyright (C) 2002, 2003 Free Software Foundation, Inc.
 
-   This program is free software: you can redistribute it and/or modify
+   This program is free software; you can redistribute it and/or modify
    it under the terms of the GNU General Public License as published by
-   the Free Software Foundation; either version 3 of the License, or
-   (at your option) any later version.
+   the Free Software Foundation; either version 2, or (at your option)
+   any later version.
 
    This program is distributed in the hope that it will be useful,
    but WITHOUT ANY WARRANTY; without even the implied warranty of
@@ -14,12 +13,15 @@
    GNU General Public License for more details.
 
    You should have received a copy of the GNU General Public License
-   along with this program.  If not, see <http://www.gnu.org/licenses/>.  */
+   along with this program; see the file COPYING.
+   If not, write to the Free Software Foundation,
+   59 Temple Place - Suite 330, Boston, MA 02111-1307, USA.  */
 
-#include <config.h>
+#if HAVE_CONFIG_H
+# include <config.h>
+#endif
 
 #include "exitfail.h"
-
-#include <stdlib.h>
+#include "exit.h"
 
 int volatile exit_failure = EXIT_FAILURE;

### 2010-09-18 Distribution/lib/exitfail.h

@@ -1,11 +1,11 @@
 /* Failure exit status
 
-   Copyright (C) 2002, 2009, 2010 Free Software Foundation, Inc.
+   Copyright (C) 2002 Free Software Foundation, Inc.
 
-   This program is free software: you can redistribute it and/or modify
+   This program is free software; you can redistribute it and/or modify
    it under the terms of the GNU General Public License as published by
-   the Free Software Foundation; either version 3 of the License, or
-   (at your option) any later version.
+   the Free Software Foundation; either version 2, or (at your option)
+   any later version.
 
    This program is distributed in the hope that it will be useful,
    but WITHOUT ANY WARRANTY; without even the implied warranty of
@@ -13,6 +13,8 @@
    GNU General Public License for more details.
 
    You should have received a copy of the GNU General Public License
-   along with this program.  If not, see <http://www.gnu.org/licenses/>.  */
+   along with this program; see the file COPYING.
+   If not, write to the Free Software Foundation,
+   59 Temple Place - Suite 330, Boston, MA 02111-1307, USA.  */
 
 extern int volatile exit_failure;

### 2010-09-18 Distribution/lib/getopt1.c

@@ -1,25 +1,29 @@
 /* getopt_long and getopt_long_only entry points for GNU getopt.
-   Copyright (C) 1987, 1988, 1989, 1990, 1991, 1992, 1993, 1994, 1996, 1997,
-   1998, 2004, 2006, 2009, 2010 Free Software Foundation, Inc.
+   Copyright (C) 1987,88,89,90,91,92,93,94,96,97,98,2004
+     Free Software Foundation, Inc.
    This file is part of the GNU C Library.
 
-   This program is free software: you can redistribute it and/or modify
+   This program is free software; you can redistribute it and/or modify
    it under the terms of the GNU General Public License as published by
-   the Free Software Foundation; either version 3 of the License, or
-   (at your option) any later version.
+   the Free Software Foundation; either version 2, or (at your option)
+   any later version.
 
    This program is distributed in the hope that it will be useful,
    but WITHOUT ANY WARRANTY; without even the implied warranty of
    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
    GNU General Public License for more details.
 
-   You should have received a copy of the GNU General Public License
-   along with this program.  If not, see <http://www.gnu.org/licenses/>.  */
+   You should have received a copy of the GNU General Public License along
+   with this program; if not, write to the Free Software Foundation,
+   Inc., 59 Temple Place - Suite 330, Boston, MA 02111-1307, USA.  */
 
+#ifdef HAVE_CONFIG_H
+#include <config.h>
+#endif
+
 #ifdef _LIBC
 # include <getopt.h>
 #else
-# include <config.h>
 # include "getopt.h"
 #endif
 #include "getopt_int.h"
@@ -32,25 +36,24 @@
 #include <stdlib.h>
 #endif
 
-#ifndef NULL
+#ifndef	NULL
 #define NULL 0
 #endif
 
 int
-getopt_long (int argc, char *__getopt_argv_const *argv, const char *options,
-             const struct option *long_options, int *opt_index)
+getopt_long (int argc, char *const *argv, const char *options,
+	     const struct option *long_options, int *opt_index)
 {
-  return _getopt_internal (argc, (char **) argv, options, long_options,
-                           opt_index, 0, 0);
+  return _getopt_internal (argc, argv, options, long_options, opt_index, 0);
 }
 
 int
-_getopt_long_r (int argc, char **argv, const char *options,
[…]

### 2010-09-18 Distribution/lib/getopt_int.h

@@ -1,68 +1,33 @@
 /* Internal declarations for getopt.
-   Copyright (C) 1989-1994, 1996-1999, 2001, 2003-2004, 2009-2010 Free Software
-   Foundation, Inc.
+   Copyright (C) 1989-1994,1996-1999,2001,2003,2004
+   Free Software Foundation, Inc.
    This file is part of the GNU C Library.
 
-   This program is free software: you can redistribute it and/or modify
+   This program is free software; you can redistribute it and/or modify
    it under the terms of the GNU General Public License as published by
-   the Free Software Foundation; either version 3 of the License, or
-   (at your option) any later version.
+   the Free Software Foundation; either version 2, or (at your option)
+   any later version.
 
    This program is distributed in the hope that it will be useful,
    but WITHOUT ANY WARRANTY; without even the implied warranty of
    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
    GNU General Public License for more details.
 
-   You should have received a copy of the GNU General Public License
-   along with this program.  If not, see <http://www.gnu.org/licenses/>.  */
+   You should have received a copy of the GNU General Public License along
+   with this program; if not, write to the Free Software Foundation,
+   Inc., 59 Temple Place - Suite 330, Boston, MA 02111-1307, USA.  */
 
 #ifndef _GETOPT_INT_H
-#define _GETOPT_INT_H   1
+#define _GETOPT_INT_H	1
 
-#include <getopt.h>
-
-extern int _getopt_internal (int ___argc, char **___argv,
-                             const char *__shortopts,
-                             const struct option *__longopts, int *__longind,
-                             int __long_only, int __posixly_correct);
+extern int _getopt_internal (int ___argc, char *const *___argv,
+			     const char *__shortopts,
+		             const struct option *__longopts, int *__longind,
+			     int __long_only);
 
 
 /* Reentrant versions which can handle parsing multiple argument
    vectors at the same time.  */
-
-/* Describe how to deal with options that follow non-option ARGV-elements.
-
-   If the caller did not specify anything,
-   the default is REQUIRE_ORDER if the environment variable
-   POSIXLY_CORRECT is defined, PERMUTE otherwise.
-
-   REQUIRE_ORDER means don't recognize them as options;
-   stop option processing when the first non-option is seen.
-   This is what Unix does.
-   This mode of operation is selected by either setting the environment
-   variable POSIXLY_CORRECT, or using `+' as the first character
[…]

### 2010-09-18 Distribution/lib/gettext.h

@@ -1,10 +1,9 @@
 /* Convenience header for conditional use of GNU <libintl.h>.
-   Copyright (C) 1995-1998, 2000-2002, 2004-2006, 2009-2010 Free Software
-   Foundation, Inc.
+   Copyright (C) 1995-1998, 2000-2002 Free Software Foundation, Inc.
 
    This program is free software; you can redistribute it and/or modify
    it under the terms of the GNU General Public License as published by
-   the Free Software Foundation; either version 3, or (at your option)
+   the Free Software Foundation; either version 2, or (at your option)
    any later version.
 
    This program is distributed in the hope that it will be useful,
@@ -14,7 +13,7 @@
 
    You should have received a copy of the GNU General Public License along
    with this program; if not, write to the Free Software Foundation,
-   Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.  */
+   Inc., 59 Temple Place - Suite 330, Boston, MA 02111-1307, USA.  */
 
 #ifndef _LIBGETTEXT_H
 #define _LIBGETTEXT_H 1
@@ -24,18 +23,6 @@
 
 /* Get declarations of GNU message catalog functions.  */
 # include <libintl.h>
-
-/* You can set the DEFAULT_TEXT_DOMAIN macro to specify the domain used by
-   the gettext() and ngettext() macros.  This is an alternative to calling
-   textdomain(), and is useful for libraries.  */
-# ifdef DEFAULT_TEXT_DOMAIN
-#  undef gettext
-#  define gettext(Msgid) \
-     dgettext (DEFAULT_TEXT_DOMAIN, Msgid)
-#  undef ngettext
-#  define ngettext(Msgid1, Msgid2, N) \
-     dngettext (DEFAULT_TEXT_DOMAIN, Msgid1, Msgid2, N)
-# endif
 
 #else
 
@@ -49,47 +36,23 @@
 # include <locale.h>
 #endif
 
-/* Many header files from the libstdc++ coming with g++ 3.3 or newer include
-   <libintl.h>, which chokes if dcgettext is defined as a macro.  So include
-   it now, to make later inclusions of <libintl.h> a NOP.  */
[…]

### 2010-09-18 Distribution/lib/mempcpy.c

@@ -1,9 +1,9 @@
 /* Copy memory area and return pointer after last written byte.
-   Copyright (C) 2003, 2007, 2009, 2010 Free Software Foundation, Inc.
+   Copyright (C) 2003 Free Software Foundation, Inc.
 
    This program is free software; you can redistribute it and/or modify
    it under the terms of the GNU General Public License as published by
-   the Free Software Foundation; either version 3, or (at your option)
+   the Free Software Foundation; either version 2, or (at your option)
    any later version.
 
    This program is distributed in the hope that it will be useful,
@@ -13,11 +13,11 @@
 
    You should have received a copy of the GNU General Public License
    along with this program; if not, write to the Free Software Foundation,
-   Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.  */
-
-#include <config.h>
+   Inc., 59 Temple Place - Suite 330, Boston, MA 02111-1307, USA.  */
 
 /* Specification.  */
+#include "mempcpy.h"
+
 #include <string.h>
 
 /* Copy N bytes of SRC to DEST, return pointer to bytes after the

### 2010-09-18 Distribution/lib/obstack.c

@@ -1,27 +1,31 @@
 /* obstack.c - subroutines used implicitly by object stack macros
 
-   Copyright (C) 1988, 1989, 1990, 1991, 1992, 1993, 1994, 1996, 1997, 1998,
-   1999, 2000, 2001, 2002, 2003, 2004, 2005, 2006, 2009, 2010 Free Software
-   Foundation, Inc.
+   Copyright (C) 1988, 1989, 1990, 1991, 1992, 1993, 1994, 1996, 1997,
+   1998, 1999, 2000, 2001, 2002, 2003, 2004 Free Software Foundation,
+   Inc.
 
-   This program is free software: you can redistribute it and/or modify
+   This program is free software; you can redistribute it and/or modify
    it under the terms of the GNU General Public License as published by
-   the Free Software Foundation; either version 3 of the License, or
-   (at your option) any later version.
+   the Free Software Foundation; either version 2, or (at your option)
+   any later version.
 
    This program is distributed in the hope that it will be useful,
    but WITHOUT ANY WARRANTY; without even the implied warranty of
    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
    GNU General Public License for more details.
 
-   You should have received a copy of the GNU General Public License
-   along with this program.  If not, see <http://www.gnu.org/licenses/>.  */
+   You should have received a copy of the GNU General Public License along
+   with this program; if not, write to the Free Software Foundation,
+   Inc., 59 Temple Place - Suite 330, Boston, MA 02111-1307, USA.  */
+
+#ifdef HAVE_CONFIG_H
+# include <config.h>
+#endif
 
 #ifdef _LIBC
 # include <obstack.h>
 # include <shlib-compat.h>
 #else
-# include <config.h>
 # include "obstack.h"
 #endif
 
@@ -39,7 +43,7 @@
    program understand `configure --with-gnu-libc' and omit the object
    files, it is simpler to just do this in the source for each such file.  */
 
-#include <stdio.h>              /* Random thing to get __GNU_LIBRARY__.  */
+#include <stdio.h>		/* Random thing to get __GNU_LIBRARY__.  */
 #if !defined _LIBC && defined __GNU_LIBRARY__ && __GNU_LIBRARY__ > 1
 # include <gnu-versions.h>
 # if _GNU_OBSTACK_INTERFACE_VERSION == OBSTACK_INTERFACE_VERSION
[…]

### 2010-09-18 Distribution/lib/stdio.h

@@ -1,1349 +1,0 @@
-/* DO NOT EDIT! GENERATED AUTOMATICALLY! */
-/* A GNU-like <stdio.h>.
-
-   Copyright (C) 2004, 2007-2010 Free Software Foundation, Inc.
-
-   This program is free software; you can redistribute it and/or modify
-   it under the terms of the GNU General Public License as published by
-   the Free Software Foundation; either version 3, or (at your option)
-   any later version.
-
-   This program is distributed in the hope that it will be useful,
-   but WITHOUT ANY WARRANTY; without even the implied warranty of
-   MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-   GNU General Public License for more details.
-
-   You should have received a copy of the GNU General Public License
-   along with this program; if not, write to the Free Software Foundation,
-   Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.  */
-
-#if __GNUC__ >= 3
-#pragma GCC system_header
-#endif
-
-#if defined __need_FILE || defined __need___FILE
-/* Special invocation convention inside glibc header files.  */
-
-#include_next <stdio.h>
-
-#else
-/* Normal invocation convention.  */
-
-#ifndef _GL_STDIO_H
-
-/* The include_next requires a split double-inclusion guard.  */
-#include_next <stdio.h>
-
-#ifndef _GL_STDIO_H
-#define _GL_STDIO_H
-
-/* Get va_list.  Needed on many systems, including glibc 2.8.  */
-#include <stdarg.h>
-
-#include <stddef.h>
-
-/* Get off_t and ssize_t.  Needed on many systems, including glibc 2.8.  */
-#include <sys/types.h>
-
-#ifndef __attribute__
-/* The __attribute__ feature is available in gcc versions 2.5 and later.
-   The __-protected variants of the attributes 'format' and 'printf' are
-   accepted by gcc versions 2.6.4 (effectively 2.7) and later.
-   We enable __attribute__ only if these are supported too, because
-   gnulib and libintl do '#define printf __printf__' when they override
-   the 'printf' function.  */
-# if __GNUC__ < 2 || (__GNUC__ == 2 && __GNUC_MINOR__ < 7)
-#  define __attribute__(Spec)   /* empty */
-# endif
-#endif
-
-
-/* The definitions of _GL_FUNCDECL_RPL etc. are copied here.  */
-#ifndef _GL_CXXDEFS_H
[…]

### 2010-09-18 Distribution/lib/strcasecmp.c

@@ -1,9 +1,9 @@
-/* Case-insensitive string comparison function.
-   Copyright (C) 1998-1999, 2005-2007, 2009-2010 Free Software Foundation, Inc.
+/* strcasecmp.c -- case insensitive string comparator
+   Copyright (C) 1998, 1999 Free Software Foundation, Inc.
 
    This program is free software; you can redistribute it and/or modify
    it under the terms of the GNU General Public License as published by
-   the Free Software Foundation; either version 3, or (at your option)
+   the Free Software Foundation; either version 2, or (at your option)
    any later version.
 
    This program is distributed in the hope that it will be useful,
@@ -13,31 +13,40 @@
 
    You should have received a copy of the GNU General Public License
    along with this program; if not, write to the Free Software Foundation,
-   Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.  */
+   Inc., 59 Temple Place - Suite 330, Boston, MA 02111-1307, USA.  */
 
-#include <config.h>
+#if HAVE_CONFIG_H
+# include <config.h>
+#endif
 
-/* Specification.  */
-#include <string.h>
+#ifdef LENGTH_LIMIT
+# define STRXCASECMP_FUNCTION strncasecmp
+# define STRXCASECMP_DECLARE_N , size_t n
+# define LENGTH_LIMIT_EXPR(Expr) Expr
+#else
+# define STRXCASECMP_FUNCTION strcasecmp
+# define STRXCASECMP_DECLARE_N /* empty */
+# define LENGTH_LIMIT_EXPR(Expr) 0
+#endif
 
+#include <stddef.h>
 #include <ctype.h>
-#include <limits.h>
 
 #define TOLOWER(Ch) (isupper (Ch) ? tolower (Ch) : (Ch))
 
-/* Compare strings S1 and S2, ignoring case, returning less than, equal to or
-   greater than zero if S1 is lexicographically less than, equal to or greater
-   than S2.
-   Note: This function does not work with multibyte strings!  */
+/* Compare {{no more than N characters of }}strings S1 and S2,
+   ignoring case, returning less than, equal to or
+   greater than zero if S1 is lexicographically less
+   than, equal to or greater than S2.  */
 
 int
-strcasecmp (const char *s1, const char *s2)
[…]

### 2010-09-18 Distribution/lib/strchrnul.c

@@ -1,10 +1,10 @@
 /* Searching in a string.
-   Copyright (C) 2003, 2007, 2008, 2009, 2010 Free Software Foundation, Inc.
+   Copyright (C) 2003 Free Software Foundation, Inc.
 
-   This program is free software: you can redistribute it and/or modify
+   This program is free software; you can redistribute it and/or modify
    it under the terms of the GNU General Public License as published by
-   the Free Software Foundation; either version 3 of the License, or
-   (at your option) any later version.
+   the Free Software Foundation; either version 2, or (at your option)
+   any later version.
 
    This program is distributed in the hope that it will be useful,
    but WITHOUT ANY WARRANTY; without even the implied warranty of
@@ -12,131 +12,19 @@
    GNU General Public License for more details.
 
    You should have received a copy of the GNU General Public License
-   along with this program.  If not, see <http://www.gnu.org/licenses/>.  */
-
-#include <config.h>
+   along with this program; if not, write to the Free Software Foundation,
+   Inc., 59 Temple Place - Suite 330, Boston, MA 02111-1307, USA.  */
 
 /* Specification.  */
-#include <string.h>
+#include "strchrnul.h"
 
 /* Find the first occurrence of C in S or the final NUL byte.  */
 char *
 strchrnul (const char *s, int c_in)
 {
-  /* On 32-bit hardware, choosing longword to be a 32-bit unsigned
-     long instead of a 64-bit uintmax_t tends to give better
-     performance.  On 64-bit hardware, unsigned long is generally 64
-     bits already.  Change this typedef to experiment with
-     performance.  */
-  typedef unsigned long int longword;
+  char c = c_in;
+  while (*s && (*s != c))
+    s++;
 
-  const unsigned char *char_ptr;
-  const longword *longword_ptr;
-  longword repeated_one;
-  longword repeated_c;
-  unsigned char c;
-
-  c = (unsigned char) c_in;
-  if (!c)
-    return rawmemchr (s, 0);
-
-  /* Handle the first few bytes by reading one byte at a time.
-     Do this until CHAR_PTR is aligned on a longword boundary.  */
-  for (char_ptr = (const unsigned char *) s;
-       (size_t) char_ptr % sizeof (longword) != 0;
-       ++char_ptr)
-    if (!*char_ptr || *char_ptr == c)
-      return (char *) char_ptr;
-
-  longword_ptr = (const longword *) char_ptr;
-
-  /* All these elucidatory comments refer to 4-byte longwords,
[…]

### 2010-09-18 Distribution/lib/strndup.c

@@ -1,11 +1,11 @@
-/* A replacement function, for systems that lack strndup.
+/* Copyright (C) 1996, 1997, 1998, 2000, 2003 Free Software Foundation, Inc.
 
-   Copyright (C) 1996, 1997, 1998, 2001, 2002, 2003, 2005, 2006, 2007, 2009,
-   2010 Free Software Foundation, Inc.
+   NOTE: The canonical source of this file is maintained with the GNU C Library.
+   Bugs can be reported to bug-glibc@prep.ai.mit.edu.
 
    This program is free software; you can redistribute it and/or modify it
    under the terms of the GNU General Public License as published by the
-   Free Software Foundation; either version 3, or (at your option) any
+   Free Software Foundation; either version 2, or (at your option) any
    later version.
 
    This program is distributed in the hope that it will be useful,
@@ -15,16 +15,31 @@
 
    You should have received a copy of the GNU General Public License
    along with this program; if not, write to the Free Software Foundation,
-   Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.  */
+   Inc., 59 Temple Place - Suite 330, Boston, MA 02111-1307, USA.  */
 
-#include <config.h>
-
-#include <string.h>
+#ifdef HAVE_CONFIG_H
+# include "config.h"
+#endif
 
 #include <stdlib.h>
+#include <string.h>
+
+#ifndef HAVE_DECL_STRNLEN
+"this configure-time declaration test was not run"
+#endif
+#if !HAVE_DECL_STRNLEN
+size_t strnlen ();
+#endif
+
+#undef __strndup
+#undef strndup
+
+#ifndef weak_alias
+# define __strndup strndup
+#endif
 
 char *
-strndup (char const *s, size_t n)
+__strndup (const char *s, size_t n)
 {
   size_t len = strnlen (s, n);
   char *new = malloc (len + 1);
@@ -35,3 +50,6 @@ strndup (char const *s, size_t n)
   new[len] = '\0';
   return memcpy (new, s, len);
 }
+#ifdef weak_alias
+weak_alias (__strndup, strndup)
+#endif

### 2010-09-18 Distribution/lib/strnlen.c

@@ -1,10 +1,10 @@
 /* Find the length of STRING, but scan at most MAXLEN characters.
-   Copyright (C) 2005, 2006, 2007, 2009, 2010 Free Software Foundation, Inc.
-   Written by Simon Josefsson.
+   Copyright (C) 1996, 1997, 1998, 2000-2003 Free Software Foundation, Inc.
+   This file is part of the GNU C Library.
 
    This program is free software; you can redistribute it and/or modify
    it under the terms of the GNU General Public License as published by
-   the Free Software Foundation; either version 3, or (at your option)
+   the Free Software Foundation; either version 2, or (at your option)
    any later version.
 
    This program is distributed in the hope that it will be useful,
@@ -12,20 +12,37 @@
    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
    GNU General Public License for more details.
 
-   You should have received a copy of the GNU General Public License
-   along with this program; if not, write to the Free Software Foundation,
-   Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.  */
+   You should have received a copy of the GNU General Public License along
+   with this program; if not, write to the Free Software Foundation,
+   Inc., 59 Temple Place - Suite 330, Boston, MA 02111-1307, USA.  */
 
-#include <config.h>
+#if HAVE_CONFIG_H
+# include <config.h>
+#endif
+#undef strnlen
 
 #include <string.h>
+
+#undef __strnlen
+#undef strnlen
+
+#ifndef _LIBC
+# define strnlen rpl_strnlen
+#endif
+
+#ifndef weak_alias
+# define __strnlen strnlen
+#endif
 
 /* Find the length of STRING, but scan at most MAXLEN characters.
    If no '\0' terminator is found in that many characters, return MAXLEN.  */
 
 size_t
-strnlen (const char *string, size_t maxlen)
+__strnlen (const char *string, size_t maxlen)
 {
   const char *end = memchr (string, '\0', maxlen);
   return end ? (size_t) (end - string) : maxlen;
 }
+#ifdef weak_alias
+weak_alias (__strnlen, strnlen)
+#endif

### 2010-09-18 Distribution/source/glb_error.c

@@ -27,9 +27,6 @@
 //Gary V. Vaughan.
 
 
-#if HAVE_CONFIG_H   /* config.h should come before any other includes */
-#  include "config.h"
-#endif
 
 #include <stdlib.h>
 #include <string.h>
@@ -67,8 +64,8 @@ int glbSetVerbosityLevel(int level)
 static void
 error (int exit_status, const char *mode, const char *message,int verb_level)
 {
-
-  if((verb_level==1)||(exit_status)>=0)
+  
+  if((verb_level==1)||(exit_status)>=0) 
     fprintf (stderr, "%s: %s: %s.\n", glb_prog_name, mode, message);
 
   if (exit_status >= 0)
@@ -91,7 +88,7 @@ glb_error (const char *message)
   error (-1, "ERROR", message,v);
 }
 
-void
+void  
 glb_fatal (const char *message)
 {
   /* A FATAL message always is displayed */
@@ -161,81 +158,53 @@ FILE *glb_fopen(const char *filename, const char *mode)
   size_t i,a,b;
   char *test_name=NULL;
   char *new_name=NULL;
-
+ 
   FILE *t;
-  register FILE *value;
-
- /* handling an absolute path */
-  if(strlen(filename)>0 && filename[0] == '/')
+  register FILE *value; 
+  for(i=0;i<glb_path_vector_length;i++)
     {
-      if(verbosity_level >= 4)
-	fprintf(stderr,"Searched path: %s\n",filename);
-      t=fopen(filename,"r");
-      if(t!=NULL)
+      a=strlen(filename);
+      b=strlen(glb_path_vector[i]);
+      test_name=glb_malloc((a+b+2)*sizeof(char));
+      test_name=strcpy(test_name,glb_path_vector[i]);
+      test_name=strcat(test_name,filename);
+      if(verbosity_level >= 4) fprintf(stderr,"Searched path: %s\n",test_name);
+      t=fopen(test_name,"r");
+      if(t!=NULL) 
 	{
-	  new_name=strdup(filename);
+	  new_name=strdup(test_name);
 	  fclose(t);
+	  glb_free(test_name);
+	  break;
 	}
+      glb_free(test_name);
+      /* Repeating the exercise with an additional '/' between
+       * path and filename.
+       */
+      a=strlen(filename);
+      b=strlen(glb_path_vector[i]);
+      test_name=glb_malloc((a+b+2)*sizeof(char));
[…]

### 2010-09-18 Distribution/source/glb_error.h

@@ -19,12 +19,12 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
+
+
+
+
 #ifndef GLB_ERROR_H
 #define GLB_ERROR_H 1
-
-#if HAVE_CONFIG_H   /* config.h should come before any other includes */
-#  include "config.h"
-#endif
 
 #include <stdlib.h>
 #include <stdio.h>

### 2010-09-18 Distribution/source/glb_fluxes.c

@@ -19,9 +19,7 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
-#if HAVE_CONFIG_H   /* config.h should come before any other includes */
-#  include "config.h"
-#endif
+
 
 
 /* Units and Normalization */
@@ -37,123 +35,37 @@
 
 #include <stdio.h>
 #include <stdlib.h>
-#include <string.h>
 #include <math.h>
 
 #include "glb_types.h"
 #include "glb_error.h"
-#include "glb_fluxes.h"
 #include "glb_multiex.h"
 
 #define PI 3.1415
 #define NORM 6.02204*1E-12 //10^-38/u (atomic mass unit)
-#define GLB_ELECTRON_MASS 0.511E-3
+#define GLB_ELECTRON_MASS 0.511E-3 
 
 static double mmu = 0.1056;           // muon mass in GeV
 static double beam_pol = 0;           // beam polaristaion
 static double beam_en = 20;           // beam energy in GeV
 static double beta,rgamma;             // relativistic quantities
-static double stored_muons = 2.0E20;  // stored muons per year
+static double stored_muons = 2.0E20;  // stored muons per year 
 // or protons on target
 // care must be taken with this, it is a different number for each experiment
 static double power;
 static int    channel = 1;            // channel
-static int FLUX_STEPS = 500;          // Number of sampling points for builtin fluxes
+static int NFLUX = 500;
 
 
-
-/***************************************************************************
- *                    H E L P E R   F U N C T I O N S                      *
- ***************************************************************************/
-
-/***************************************************************************
- * Function glb_load_n_columns                                             *
- ***************************************************************************
- * Load data from file. The expected format is n_columns columns and an    *
- * arbitray number of rows. The resulting number of rows will be stored in *
- * n_lines, while for the actual data, arrays will be created (one for     *
- * each column), i.e. data[] should have length n_columns                  *
[…]

### 2010-09-18 Distribution/source/glb_glb.c

@@ -19,6 +19,7 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
+
 #ifdef HAVE_CONFIG_H
 #include "config.h"
 #endif
@@ -35,18 +36,18 @@
 
 /* Moved from glb_wrapper.c, in order to have the module support stuff
  * only used in this file.
- */
-static void final_clean()
+ */ 
+static void final_clean() 
 {
   glb_builtin_prior_clean();
-  glb_clean_up();
-  return;
+  glb_clean_up(); 
+  return; 
 }
 
 static int test=1234;
 void glbInit(char *name)
 {
-
+  
 
   atexit(final_clean);
   glb_init(name);
@@ -65,12 +66,12 @@ void glbInit(char *name)
   /* Select default minimizer */
   glbSelectMinimizer(GLB_MIN_DEFAULT);
 
-  /* Setup default priors */
+  /* Setup default priors */  
   glbRegisterPriorFunction(glb_builtin_prior_prior,
 			   glb_builtin_prior_starting_values,
 			   glb_builtin_prior_input_errors,
                            &test);
-
+  
   /* Register built-in chi^2 functions */
   /* When making any changes here, don't forget to update the array
    * glb_2011_compatible_chi_functions in glb_sys.c and the functions

### 2010-09-18 Distribution/source/glb_lexer.c

@@ -8,7 +8,7 @@
 #define FLEX_SCANNER
 #define YY_FLEX_MAJOR_VERSION 2
 #define YY_FLEX_MINOR_VERSION 5
-#define YY_FLEX_SUBMINOR_VERSION 35
+#define YY_FLEX_SUBMINOR_VERSION 31
 #if YY_FLEX_SUBMINOR_VERSION > 0
 #define FLEX_BETA
 #endif
@@ -30,15 +30,7 @@
 
 /* C99 systems have <inttypes.h>. Non-C99 systems may or may not. */
 
-#if defined (__STDC_VERSION__) && __STDC_VERSION__ >= 199901L
-
-/* C99 says to define __STDC_LIMIT_MACROS before including stdint.h,
- * if you want the limit (max/min) macros for int types. 
- */
-#ifndef __STDC_LIMIT_MACROS
-#define __STDC_LIMIT_MACROS 1
-#endif
-
+#if defined __STDC_VERSION__ && __STDC_VERSION__ >= 199901L
 #include <inttypes.h>
 typedef int8_t flex_int8_t;
 typedef uint8_t flex_uint8_t;
@@ -93,12 +85,11 @@ typedef unsigned int flex_uint32_t;
 
 #else	/* ! __cplusplus */
 
-/* C99 requires __STDC__ to be defined as 1. */
-#if defined (__STDC__)
+#if __STDC__
 
 #define YY_USE_CONST
 
-#endif	/* defined (__STDC__) */
+#endif	/* __STDC__ */
 #endif	/* ! __cplusplus */
 
 #ifdef YY_USE_CONST
@@ -178,9 +169,14 @@ extern FILE *yyin, *yyout;
 
 #define unput(c) yyunput( c, (yytext_ptr)  )
 
+/* The following is because we cannot portably get our hands on size_t
+ * (without autoconf's help, which isn't available because we want
+ * flex-generated scanners to compile on their own).
+ */
+
 #ifndef YY_TYPEDEF_YY_SIZE_T
 #define YY_TYPEDEF_YY_SIZE_T
-typedef size_t yy_size_t;
+typedef unsigned int yy_size_t;
 #endif
 
 #ifndef YY_STRUCT_YY_BUFFER_STATE
@@ -275,7 +271,7 @@ int yyleng;
 
 /* Points to current character in buffer. */
 static char *yy_c_buf_p = (char *) 0;
-static int yy_init = 0;		/* whether we need to initialize */
[…]

### 2010-09-18 Distribution/source/glb_minimize.c

@@ -19,9 +19,8 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
-#if HAVE_CONFIG_H   /* config.h should come before any other includes */
-#  include "config.h"
-#endif
+
+
 
 
 /* --------------------------------------------------------
@@ -47,7 +46,7 @@
  * and in which order the truncations are performed. This may
  * also change under compiler optimization. In any case the errors
  * should be very small (i.e. <<1E-6) and should not have any impact
- * on the physics results.
+ * on the physics results. 
  */
 
 
@@ -57,7 +56,6 @@
 #include <math.h>
 #include <setjmp.h>
 #include <globes/globes.h>
-#include <gsl/gsl_siman.h>
 
 #include "glb_probability.h"
 #include "glb_fluxes.h"
@@ -79,7 +77,7 @@ double glb_last_priors=0.0; /* Used to store prior terms between calls to glb_hy
 /* Warning-- fiddling around with these in general may
  * influence the stability of the minimization process !
  *
- * The latest Recator (D-CHOOZ as in hep-ph/0403068) setups
+ * The latest Recator (D-CHOOZ as in hep-ph/0403068) setups 
  * have TOLSYS 1.0E-12, whereas all the other setups have TOLSYS 1.0E-5
  * and TOLSYS 1.0E-8 may be faster by 20% for Nufact (improves convergence
  * of the second layer of minimization)
@@ -89,7 +87,7 @@ double glb_last_priors=0.0; /* Used to store prior terms between calls to glb_hy
 #define TOLOSC 1.0E-5
 
 
-/* defines wether the program will run
+/* defines wether the program will run 
  * with (1) or w/o mathematica (0)
  */
 
@@ -106,7 +104,7 @@ double glb_last_priors=0.0; /* Used to store prior terms between calls to glb_hy
 /* this is for making the Chi... functions abortable */
 static int okay_flag = 0;
 /* this needed by the wrappers in order not to return garbage */
-static jmp_buf env;
+static jmp_buf env; 
 /* this is need by setjmp() longjmp() */
 //--------------------------------------------------------
 
[…]

### 2010-09-18 Distribution/source/glb_minimize.h

@@ -22,10 +22,6 @@
 #ifndef GLB_MMINIMIZE_H
 #define GLB_MMINIMIZE_H 1
 
-#if HAVE_CONFIG_H   /* config.h should come before any other includes */
-#  include "config.h"
-#endif
-
 #include <glb_min_sup.h>
 
 
@@ -61,7 +57,6 @@ double* glb_return_input_values();
 
 
 int glb_invoke_hybrid_minimizer(int exp, int rule, double *x, double *chi2);
-int glb_invoke_siman_minimizer(int exp, int rule, double *x, double *chi2);
 
 
 // various projections with th12 fixed for a single experiment

### 2010-09-18 Distribution/source/glb_modules.c

@@ -19,9 +19,6 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
-#if HAVE_CONFIG_H   /* config.h should come before any other includes */
-#  include "config.h"
-#endif
 
 #include <ltdl.h>
 #include <stdio.h>
@@ -29,6 +26,7 @@
 #include <unistd.h>
 #include <globes/globes.h>
 #include <globes/glb-modules.h>
+
 #include "glb_path.h"
 #include "glb_error.h"
 
@@ -64,7 +62,7 @@ void glb_close_module_support()
 /* Set up path to be searched for modules */
 int glb_setup_module_search_path()
 {
-
+ 
   int i,s=1;
   char *in_path;
   char *path;
@@ -79,11 +77,11 @@ int glb_setup_module_search_path()
   /* Parse the path -- try this aloud twenty times as fast as you can ;-) */
   s=glb_break_up_path(path,&path_vector,&path_vector_length);
   glb_free(path);
-
+  
   /* Setup the correct directory to find the builtin modules */
   s += lt_dladdsearchdir(GLB_MODULE_DIR);
 
-  /* Adding the elements of path vector to the module search path
+  /* Adding the elements of path vector to the module search path 
    * ommiting the first (which would be the local directory) */
   for(i=1;i<path_vector_length;i++) s += lt_dladdsearchdir(path_vector[i]);
 
@@ -91,14 +89,14 @@ int glb_setup_module_search_path()
   for(i=0;i<path_vector_length;i++)
     glb_free((char *) path_vector[i]);
   glb_free((char **) path_vector);
-
+  
   return s;
 }
 
 #define GLB_MODULE_NOT_FOUND -2
 #define GLB_MODULE_NO_INIT -1
 
-/* Checking wether a module exists and is a GLoBES module
+/* Checking wether a module exists and is a GLoBES module 
  * The idea is to adjust verbosity level to whatever the
  * user needs, if he wants to avoid later error messages
  * he has to explicitely call glb_probe_module before any
@@ -137,13 +135,13 @@ int glbProbeModule(const char *module_name, int verbosity)
       fprintf(stderr,"glb_probe_module: ERROR: in module %s: %s\n",
 	      module_name,error);
[…]

### 2010-09-18 Distribution/source/glb_parser_addons.h

@@ -19,12 +19,13 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
+
+
+
+
+
 #ifndef GLB_MY_PARSER_H
 #define GLB_MY_PARSER_H 1
-
-#if HAVE_CONFIG_H   /* config.h should come before any other includes */
-#  include "config.h"
-#endif
 
 #include <globes/globes.h>
 

### 2010-09-18 Distribution/source/glb_parser_type.h

@@ -19,12 +19,12 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
+
+
+
+
 #ifndef GLB_PARSER_DATA_H
 #define GLB_PARSER_DATA_H 1
-
-#if HAVE_CONFIG_H   /* config.h should come before any other includes */
-#  include "config.h"
-#endif
 
 #include "glb_types.h"
 

### 2010-09-18 Distribution/source/glb_path.c

@@ -19,9 +19,7 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
-#if HAVE_CONFIG_H   /* config.h should come before any other includes */
-#  include "config.h"
-#endif
+
 
 
 /* Resolution of a path given by an environment variable.
@@ -80,7 +78,7 @@ int glb_break_up_path(char *in_path, char ***pathv, size_t *len)
       return 0;
     }
   token=strtok(in_path,delim);
-  if(token!=NULL)
+  if(token!=NULL) 
     {
       vec=(char **) glb_realloc(vec,sizeof(char*)*(length+1));
       vec[length]=strdup(token);
@@ -96,7 +94,7 @@ int glb_break_up_path(char *in_path, char ***pathv, size_t *len)
 	  length++;
 	}
     }
-
+  
   *len=length;
   *pathv=vec;
   return 0;

### 2010-09-18 Distribution/source/glb_prior.c

@@ -18,16 +18,19 @@
  * You should have received a copy of the GNU General Public License
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
+
  */
-#if HAVE_CONFIG_H   /* config.h should come before any other includes */
-#  include "config.h"
-#endif
+
 
 /* Headers needed by this module */
 #include <stdio.h>
 #include <stdlib.h>
 #include <math.h>
 #include <string.h>
+
+#ifdef HAVE_CONFIG_H
+#include "config.h"
+#endif
 
 #include "globes/globes.h"
 #include "glb_prior.h"
@@ -36,17 +39,17 @@
 static glb_params sv;
 static glb_params er;
 
-/* glb_module_init is mandatory. Any module must have glb_module_init,
- * which returns 0 on succes or -1 on failure. This function is called
+/* glb_module_init is mandatory. Any module must have glb_module_init, 
+ * which returns 0 on succes or -1 on failure. This function is called 
  * when the module is loaded.
  *
- * In our case it allocates memory for sv and er and sets sv/er to any
- * input errors starting values which have already been defined
- */
+ * In our case it allocates memory for sv and er and sets sv/er to any 
+ * input errors starting values which have already been defined 
+ */   
 int glb_builtin_prior_init()
 {
   sv=glbAllocParams();
-  er=glbAllocParams();
+  er=glbAllocParams(); 
   glbGetStartingValues(sv);
   glbGetInputErrors(er);
 
@@ -66,20 +69,20 @@ void glb_builtin_prior_clean()
 }
 
 /* This is a local function which implements the prior calculation
- * and ensure that the builtin prior behaves as described in the
+ * and ensure that the builtin prior behaves as described in the 
  * documentation, i.e. it returns 0 when the error is below 1E-12
  */
 static double sprior(double x, double center, double sigma)
 {
   if(fabs(sigma-0)<1E-12) return 0;
   return (x-center)*(x-center)/sigma/sigma;
-}
+}  
 
 /* glb_builtin_prior_prior is mandatory (for prior modules).
  *
  * This is the default prior, i.e. priors are only added for free
  * parameters, and the prior accepts starting values as stored in sv
[…]

### 2010-09-18 Distribution/source/glb_probability.c

@@ -19,10 +19,6 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
-#if HAVE_CONFIG_H   /* config.h should come before any other includes */
-#  include "config.h"
-#endif
-
 #include <stdio.h>
 #include <stdlib.h>
 #include <float.h>
@@ -68,11 +64,8 @@ gsl_matrix_complex *S1=NULL, *T0=NULL; /* Temporary matrix storage              
  ***************************************************************************/
 
 // ----------------------------------------------------------------------------
-//JK 2008-08-09 Temporary workaround for bug #22
-//inline void zhetrd3(double complex A[3][3], double complex Q[3][3],
-//                    double d[3], double e[2])
-void zhetrd3(double complex A[3][3], double complex Q[3][3],
-             double d[3], double e[2])
+inline void zhetrd3(double complex A[3][3], double complex Q[3][3],
+                    double d[3], double e[2])
 // ----------------------------------------------------------------------------
 // Reduces a hermitian 3x3 matrix to real tridiagonal form by applying
 // (unitary) Householder transformations:
@@ -87,7 +80,7 @@ void zhetrd3(double complex A[3][3], double complex Q[3][3],
   double complex u[n], q[n];
   double complex omega, f;
   double K, h, g;
-
+  
   // Initialize Q to the identitity matrix
 #ifndef EVALS_ONLY
   for (int i=0; i < n; i++)
@@ -98,7 +91,7 @@ void zhetrd3(double complex A[3][3], double complex Q[3][3],
   }
 #endif
 
-  // Bring first row and column to the desired form
+  // Bring first row and column to the desired form 
   h = SQR_ABS(A[0][1]) + SQR_ABS(A[0][2]);
   if (creal(A[0][1]) > 0)
     g = -sqrt(h);
[…]

### 2010-09-18 Distribution/source/glb_rate_engine.c

@@ -19,9 +19,10 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
-#if HAVE_CONFIG_H   /* config.h should come before any other includes */
-#  include "config.h"
-#endif
+
+
+
+
 
 #include <stdio.h>
 #include <stdlib.h>
@@ -36,7 +37,6 @@
 #include "glb_multiex.h"
 #include "glb_fluxes.h"
 #include "glb_error.h"
-#include "glb_rate_engine.h"
 
 
 #define PI 3.1415
@@ -61,12 +61,12 @@ int* glb_calc_uprange[32];
 int* glb_calc_lowrange[32];
 static int* rules[32];
 static int rule_length[32];
-static double* rule_coeff[32];
+static double* rule_coeff[32]; 
 static int* BGrules[32];
 static int BGrule_length[32];
 static double* BGrule_coeff[32];
  double* glb_calc_signal_rates[32];
- double* glb_calc_bg_rates[32];
+ double* glb_calc_bg_rates[32]; 
 // ---------------------------------
 double* glb_calc_rates_0[32];
 double* glb_calc_rates_1[32];
@@ -95,7 +95,6 @@ double *glb_calc_chrb_1[32];
 double *glb_calc_chra_0[32];
 double *glb_calc_chra_1[32];
 
-double *glb_calc_chr_template[32];
 
 //binwise glb_calc_efficiencies
 double* glb_calc_efficiencies[32];
@@ -143,7 +142,7 @@ void glb_set_baseline(double l)
   ll = e->lengthtab;
   sum=0;
   if(s<=0)
-    {
+    { 
       glb_error("No baseline data to change");
       return;
     }
@@ -153,7 +152,7 @@ void glb_set_baseline(double l)
       glb_error("Sum of lengths of layers does not match baseline");
       return;
     }
-  baseline = sum;
+  baseline = sum;  
 }
 
 double glb_check_baseline()
@@ -212,19 +211,19 @@ int glb_get_number_of_bins()
   return bins;
 }
 
-/* Initialize the controlstructure for signal and backgrounds
- * Here I introduce the concept of channels. In general there are
- * three flavours in a beam, with to CP-signs, thus there six initial
- * states, which can oscillate to six final states. The reaction in
- * the detector can be anything, e.g. CC. This are altogether 72
- * possibilities, of which you usually need only a few. Therefore we
[…]

### 2010-09-18 Distribution/source/glb_rate_engine.h

@@ -19,17 +19,14 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
+
+
+
+
 #ifndef GLB_CALC_H
 #define GLB_CALC_H 1
 
-#if HAVE_CONFIG_H   /* config.h should come before any other includes */
-#  include "config.h"
-#endif
-
 #include "glb_types.h"
-
-#define GLB_FAST_RATES 1
-#define GLB_SLOW_RATES 0
 
 extern int glb_num_of_rules;
 extern double glb_glb_sig_norm_error[32];
@@ -71,7 +68,6 @@ void glb_set_bg_rule(int i, int cn, int *rule, double *coeff);
 
 void glb_set_rates();
 void glb_set_new_rates();
-void glb_rate_template();
 void glb_set_signal_errors(int i,double norm, double tilt);
 void glb_set_bg_errors(int i,double norm, double tilt);
 void glb_set_bg_center(int i,double norm, double tilt);
@@ -142,7 +138,6 @@ extern double *glb_calc_chrb_0[32];
 extern double *glb_calc_chrb_1[32];
 extern double *glb_calc_chra_0[32];
 extern double *glb_calc_chra_1[32];
-extern double *glb_calc_chr_template[32];
 
 extern double *glb_calc_user_pre_sm_channel[32];
 extern double *glb_calc_user_post_sm_channel[32];

### 2010-09-18 Distribution/source/glb_smear.c

@@ -19,9 +19,8 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
-#if HAVE_CONFIG_H   /* config.h should come before any other includes */
-#  include "config.h"
-#endif
+
+
 
 /* These routines compute the smear matrix in the same way as CalcSmearA
  * in GlobesMESbeta.m. CalcSmearB is also implemented, but it seems
@@ -40,11 +39,6 @@
 #include <gsl/gsl_math.h>
 #include <gsl/gsl_sf_erf.h>
 #include <gsl/gsl_errno.h>
-#include <gsl/gsl_matrix.h>
-#include <gsl/gsl_permutation.h>
-#include <gsl/gsl_blas.h>
-#include <gsl/gsl_linalg.h>
-#include <gsl/gsl_sf_gamma.h>
 
 
 
@@ -90,7 +84,7 @@ void glb_init_bin_data(glb_smear *in);
 glb_option_type *glb_option_type_alloc()
 {
   glb_option_type *in;
-  in=(glb_option_type *) glb_malloc(sizeof(glb_option_type));
+  in=(glb_option_type *) glb_malloc(sizeof(glb_option_type)); 
   in->corr_fac=-1;
   in->confidence_level=-1;
   in->offset=-1;
@@ -100,7 +94,7 @@ glb_option_type *glb_option_type_alloc()
 }
 
 glb_option_type *glb_option_type_reset(glb_option_type *in)
-{
+{ 
   in->corr_fac=-1;
   in->confidence_level=-1;
   in->offset=-1;
@@ -110,7 +104,7 @@ glb_option_type *glb_option_type_reset(glb_option_type *in)
 }
 
 void glb_option_type_free(glb_option_type *in)
-{
+{    
     glb_free(in);
 }
 
@@ -127,8 +121,8 @@ const struct glb_experiment *headex)
     {glb_error("The confidence level has to be inbetween 0 and 1!");s=-1;}
   if(in->offset==-1)  {in->offset=0;d=2;}
   if(in->offset<0) {glb_error("Offset must not be smaller than 0!");s=-1;}
-
-  if(head==NULL)
+  
+  if(head==NULL)  
     {
       if(in->low_bound==-1) {glb_error("No low_bound given!");s=-1;}
       if(in->up_bound==-1) {glb_error("No up_bound given!");s=-1;}
@@ -140,7 +134,7 @@ const struct glb_experiment *headex)
     }
[…]

### 2010-09-18 Distribution/source/glb_sys.h

@@ -22,10 +22,6 @@
 #ifndef GLB_SYS_H
 #define GLB_SYS_H 1
 
-#if HAVE_CONFIG_H   /* config.h should come before any other includes */
-#  include "config.h"
-#endif
-
 #include "glb_types.h"
 
 extern glb_systematic *glb_sys_list;

### 2010-09-18 Distribution/source/glb_version.c

@@ -19,9 +19,7 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
-#if HAVE_CONFIG_H   /* config.h should come before any other includes */
-#  include "config.h"
-#endif
+
 
 #include <stdlib.h>
 #include <string.h>
@@ -36,16 +34,16 @@
 #include "glb_error.h"
 
 const char *glb_release_version = VERSION;
-static const char *library_version = GLB_LIBVERSION;
+static const char *library_version = GLB_LIBVERSION; 
 
-static int
+static int 
 break_up_version(char const *version,int *first,int *second, int *third)
 {
-
+  
   char c;
   size_t length,i,k,ct=0,ctb=0;
   char *vec=NULL;
-  int res[4]={-1,-1,-1,-1};
+  int res[3]={-1,-1,-1};
   if(version==NULL)
     {
       *third=-1;
@@ -63,11 +61,11 @@ break_up_version(char const *version,int *first,int *second, int *third)
 	  vec=(char *) glb_realloc(vec,sizeof(char)*(ct+1));
 	  vec[ct-1]=c;
 	  vec[ct]='\0';
-	}
+	} 
       else
 	{
 	  ctb++;
-	  if((ctb>4)||(vec==NULL))
+	  if((ctb>3)||(vec==NULL)) 
 	    {glb_fatal("Invalid version string");exit(1);}
 	  sscanf(vec,"%d",&res[ctb-1]);
 	  ct=0;
@@ -88,12 +86,12 @@ glbTestReleaseVersion(const char *version)
   int major,minor,bug,tmajor,tminor,tbug;
   break_up_version(version,&major,&minor,&bug);
   break_up_version(glb_release_version,&tmajor,&tminor,&tbug);
-
-  if(major==tmajor)
+  
+  if(major==tmajor) 
     {
-      if(minor==tminor)
+      if(minor==tminor) 
 	{
-	  if(bug==tbug)
+	  if(bug==tbug) 
 	    {
 	      s=0;
 	    }
@@ -115,11 +113,11 @@ glbTestLibraryVersion(const char *version)
   int major,minor,bug,tmajor,tminor,tbug;
   break_up_version(version,&major,&bug,&minor);
   break_up_version(library_version,&tmajor,&tbug,&tminor);
-  if(major==tmajor)
[…]

### 2010-09-18 Distribution/source/glb_version.h

@@ -19,12 +19,10 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
+
+
 #ifndef GLB_VERSION_H
 #define GLB_VERSION_H 1
-
-#if HAVE_CONFIG_H   /* config.h should come before any other includes */
-#  include "config.h"
-#endif
 
 extern const char *glb_release_version;
 

### 2010-09-18 Distribution/source/glb_wrapper.c

@@ -19,9 +19,9 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
-#if HAVE_CONFIG_H   /* config.h should come before any other includes */
-#  include "config.h"
-#endif
+
+
+
 
 #include <stdio.h>
 #include <string.h>
@@ -48,7 +48,7 @@
 
 /* The global variables */
 int glb_num_of_exps;
-glb_exp glb_experiment_list[GLB_MAX_EXP];
+glb_exp glb_experiment_list[32];
 int glb_rule_number;
 
 
@@ -64,11 +64,11 @@ glb_osc_type *glb_alloc_osc_type()
 {
   glb_osc_type *temp;
   int i;
-
+  
   temp = (glb_osc_type *) glb_malloc(sizeof(glb_osc_type));
   if (temp != NULL)
   {
-    temp->osc_params = (double *) glb_malloc(sizeof(double) * glbGetNumOfOscParams());
+    temp->osc_params = (double *) glb_malloc(sizeof(double) * glbGetNumOfOscParams()); 
     if (temp->osc_params != NULL)
     {
       for (i=0; i < glbGetNumOfOscParams(); i++)
@@ -89,13 +89,11 @@ glb_density_type *glb_alloc_density_type()
 {
   glb_density_type *temp;
   int i;
-
+  
   temp = (glb_density_type *) glb_malloc(sizeof(glb_density_type));
   if (temp != NULL)
   {
     temp->density_params = (double *) glb_malloc(sizeof(double)*glb_num_of_exps);
-
-
     if (temp->density_params != NULL)
     {
       for(i=0; i < glb_num_of_exps; i++)
@@ -147,7 +145,7 @@ int glbGetIteration(const glb_params in)
   return in->iterations;
 }
 
-glb_params glbSetDensityParams(glb_params in,
+glb_params glbSetDensityParams(glb_params in, 
 				 double dens, int which)
 {
   int i;
@@ -159,7 +157,7 @@ glb_params glbSetDensityParams(glb_params in,
     {
       (in->density)->density_params[which]=dens;
     }
-  else
+  else 
     {
       glb_error("glbSetDensityParams: Density list length mismatch");
       return NULL;
@@ -174,7 +172,7 @@ double glbGetDensityParams(const glb_params in, int which)
     {
       out=(in->density)->density_params[which];
     }
-  else
[…]

### 2010-09-18 Distribution/source/glb_wrapper.h

@@ -19,12 +19,12 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
+
+
+
+
 #ifndef glb_WRAPPER_H
 #define  glb_WRAPPER_H 1
-
-#if HAVE_CONFIG_H   /* config.h should come before any other includes */
-#  include "config.h"
-#endif
 
 #include <globes/globes.h>
 

### 2010-09-18 Distribution/source/globes.c

@@ -19,9 +19,10 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
-#if HAVE_CONFIG_H   /* config.h should come before any other includes */
-#  include "config.h"
-#endif
+
+
+
+
 
 #include <stdio.h>
 #include <stdlib.h>
@@ -31,14 +32,21 @@
 #include <argp.h>
 #include <ctype.h>
 #include "glb_error.h"
+
+
 #include "globes/globes.h"
+
+#if HAVE_CONFIG_H
+#  include "config.h"
+#endif
+
 
 #ifdef TEST
 #include "glb_modules.h"
 #endif /* TEST */
 
 const char *argp_program_version =
-"globes "VERSION"\n(C) 2002 - 2010 The GLoBES Team\n"
+"globes "VERSION"\n(C) 2002 - 2007 The GLoBES Team\n"
 "This is free software see the source for copying conditions. There is NO\n"
 "warranty; not even for MERCHANTABILITY or"
 " FITNESS FOR A PARTICULAR PURPOSE.";
@@ -74,13 +82,13 @@ static int parse_definition(const char *str)
   inc=strdup(str);
   if(inc==NULL) return 0;
   token=strtok(inc,delim);
-  if(token!=NULL)
+  if(token!=NULL) 
     {
      lhs=strdup(token);
      length++;
     }
-  token=strtok(NULL,delim);
-  if(token!=NULL)
+  token=strtok(NULL,delim); 
+  if(token!=NULL) 
     {
      rhs=strdup(token);
      length++;
@@ -95,21 +103,21 @@ static int parse_definition(const char *str)
   wrk=dummy;
   if(rhs[0]=='{' && rhs[s-1]=='}') {strncpy(wrk,&rhs[1],s);vec=1;wrk[s-2]='\0';}
   else strncpy(wrk,&rhs[0],s+1);
-
+  
   while(wrk)
     {
         errno=0;
       res=strtod(wrk,&endp);
-
+     
       if(errno) {fprintf(stderr,"globes: ERROR: While parsing input the following"
 			 " error occured\nFATAL: '%s'\n",
                          strerror(errno));exit(1);}
       c++;
       result = (double* ) glb_realloc(result, c * sizeof(double));
       result[c-1]=res;
-
+  
[…]

### 2010-09-18 Documentation/tutorials/LBNE-tutorial/degfinder.h

@@ -1,37 +1,0 @@
-#ifndef __DEGFINDER_H
-#define __DEGFINDER_H
-
-#include <math.h>
-#include <float.h>
-#include <globes/globes.h>
-
-/* Mass hierarchies */
-#define HIERARCHY_NORMAL     1
-#define HIERARCHY_INVERTED  -1
-
-/* Options for degfinder */
-#define DEG_NO_NH      0x01   /* Omit normal hierarchy fits                    */
-#define DEG_NO_IH      0x02   /* Omit inverted hierarchy fits                  */
-#define DEG_NO_SYS     0x04   /* Switch off systematics                        */
-#define DEG_NO_CORR    0x10   /* Switch off correlations                       */
-#define DEG_NO_DEG     0x20   /* Switch off all degeneracies                   */
-
-/* Define NaN (not a number) */
-#ifdef NAN
-  #define GLB_NAN      NAN
-  #define GLB_ISNAN(x) isnan(x)
-#else
-  #define GLB_NAN      DBL_MAX
-  #define GLB_ISNAN(x) ((x) == DBL_MAX)
-#endif
-
-
-/* degfinder.c */
-int degfinder(const glb_params base_values, const int n_prescan_params,
-      const int *prescan_params, const double *prescan_min,
-      const double *prescan_max, const int *prescan_steps,
-      const glb_projection prescan_proj, const glb_projection fit_proj,
-      int *n_deg, glb_params *deg_pos, double *deg_chi2, const long flags);
-
-#endif
-

### 2010-09-18 Documentation/tutorials/LBNE-tutorial/th13chi2-p4.c

@@ -1,124 +1,0 @@
-/* GLoBES -- General LOng Baseline Experiment Simulator
- * (C) 2002 - 2007,  The GLoBES Team
- *
- * GLoBES is mainly intended for academic purposes. Proper
- * credit must be given if you use GLoBES or parts of it. Please
- * read the section 'Credit' in the README file.
- *
- * This program is free software; you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation; either version 2 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program; if not, write to the Free Software
- * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
- */
-
-/***************************************************************************
- *            G L o B E S   -   M I N I - W O R K S H O P                  *
- *                                                                         *
- *                     May 13, 2010, Fermilab                              *
- *                                                                         *
- * Hands-on session on LBNE: Computing chi^2 for the wrong hierarchy       *
- * as a function of th13                                                   *
- *                                                                         *
- * Solution to problem 4: Code is similar to th13chi2.c, but with a        *
- * different starting value for the minimization                           *
- ***************************************************************************/
-
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <math.h>
-#include <string.h>
-
-#include <globes/globes.h>            /* GLoBES library */
-
-/* Output file, empty string = stdout */
-char MYFILE[]="th13chi2.dat";
-FILE *outfile = NULL;
-
-int main(int argc, char *argv[])
-{
-  /* Initialize libglobes */
-  glbInit(argv[0]);
-
[…]

### 2010-09-18 Documentation/tutorials/LBNE-tutorial/th13chi2-p5.c

@@ -1,125 +1,0 @@
-/* GLoBES -- General LOng Baseline Experiment Simulator
- * (C) 2002 - 2007,  The GLoBES Team
- *
- * GLoBES is mainly intended for academic purposes. Proper
- * credit must be given if you use GLoBES or parts of it. Please
- * read the section 'Credit' in the README file.
- *
- * This program is free software; you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation; either version 2 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program; if not, write to the Free Software
- * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
- */
-
-/***************************************************************************
- *            G L o B E S   -   M I N I - W O R K S H O P                  *
- *                                                                         *
- *                     May 13, 2010, Fermilab                              *
- *                                                                         *
- * Hands-on session on LBNE: Computing chi^2 for the wrong hierarchy       *
- * as a function of th13                                                   *
- *                                                                         *
- * Solution to problem 5: Implementation of the tracking algorithm         *
- ***************************************************************************/
-
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <math.h>
-#include <string.h>
-
-#include <globes/globes.h>            /* GLoBES library */
-
-/* Output file, empty string = stdout */
-char MYFILE[]="th13chi2.dat";
-FILE *outfile = NULL;
-
-int main(int argc, char *argv[])
-{
-  /* Initialize libglobes */
-  glbInit(argv[0]);
-
-  /* New minimization algorithm -> significant speed-up */
-  glbSelectMinimizer(GLB_MIN_POWELL);
-
[…]

### 2010-09-18 Documentation/tutorials/LBNE-tutorial/th13delta-NF.c

@@ -1,142 +1,0 @@
-/* GLoBES -- General LOng Baseline Experiment Simulator
- * (C) 2002 - 2004,  The GLoBES Team
- *
- * GLoBES is mainly intended for academic purposes. Proper
- * credit must be given if you use GLoBES or parts of it. Please
- * read the section 'Credit' in the README file.
- *
- * This program is free software; you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation; either version 2 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program; if not, write to the Free Software
- * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
- */
-
-/***************************************************************************
- *            G L o B E S   -   M I N I - W O R K S H O P                  *
- *                                                                         *
- *                     May 13, 2010, Fermilab                              *
- *                                                                         *
- * Hands-on session on LBNE: Computing confidence regions in the           *
- * th13-delta_CP plane for a neutrino factory                              *
- ***************************************************************************/
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <math.h>
-#include <string.h>
-
-#include <globes/globes.h>   /* GLoBES library */
-
-/* Output file, empty string = stdout */
-char MYFILE[]="th13delta-NF.dat";
-FILE *outfile = NULL;
-
-int main(int argc, char *argv[])
-{ 
-  /* Initialize libglobes */
-  glbInit(argv[0]); 
-
-  /* New minimization algorithm -> significant speed-up */
-  glbSelectMinimizer(GLB_MIN_POWELL);
-
-  /* Initialize experiment */
[…]

### 2010-09-21 Documentation/part2.tex
*** empty log message ***
Improved documenation of #inverse_beta
Changes in manual for version 3.1.7

@@
 & & {\tt @stored\_muons} &  \\
-3 & Beta beam, inv. beta decay & {\tt @end\_point} [GeV], & 3.0 \\
+3 & Beta beam, $\beta^+$ decay & {\tt @end\_point} [GeV], & 3.0 \\
 & & {\tt @stored\_ions}, &  \\
 & & {\tt @gamma} & \\
-4 & Beta beam, beta decay & {\tt @end\_point} [GeV], & 3.0 \\
+4 & Beta beam, $\beta^-$ decay & {\tt @end\_point} [GeV], & 3.0 \\
 & & {\tt @stored\_ions}, &  \\
@@
 The software assumes that the given flux file has seven columns and
-501 lines with equidistant energies. The format is:
+501 lines\footnote{The requirement of exactly $501$ lines 
+has been removed in versions 3.1.7 and higher.} with equidistant energies. The format is:
 \begin{quotation}
@@
 The software assumes that the cross section files are text files with 
-seven columns and $1001$ lines of the form
+seven columns and $1001$ lines\footnote{The requirement of exactly $1001$ lines 
+has been removed in versions 3.1.7 and higher.}
+of the form
 \index{norm}{Cross section!file}
@@
 For the case of large values of the integrand in \eq~(\ref{eq:simple_int}) at the energy range limits, it is recommended to exceed the analysis energy window by about three times the energy calibration error in order
-to avoid cutoff effects.
+to avoid cutoff effects. In addition, note that an energy resolution larger than about the
+bin range will distribute events out of this range. Therefore, the normalization
+will be affected.
 
@@
 R^c(E,E')=\frac{1}{\sigma(E)\,\sqrt{2\pi}}\,e^{-\frac{(E-E')^2}{2\sigma^2(E)}} \, .
+\label{eq:respfun}
 \end{equation} 
@@
 it is possible to impose a low-pass filter already during the calculation
-of the probabilities itself. This highly
-experimental feature will be called ``filter''\index{norm}{Filter} in the following. 
+of the probabilities itself. This in the meanwhile frequently used feature will be called ``filter''\index{norm}{Filter} in the following. 
 The calculation of oscillation
@@
 e^{-i\Phi_{ij}}\times 
-e^{ -\Phi_{ij}^2/\sigma_f(E)^2 }\,,
+e^{ - \sigma_e^2 \, \Phi_{ij}^2/(2 E^2) }\,,
 \end{eqnarray}
 where $\Phi_{ij}:=\Delta m_{ij}^2 L/2E$ is the usual phase difference and
-the last term is a Gau\ss ian filter with width $\sigma_f(E)$. Choosing
-$\sigma_f(E):=\sigma_f^0 \cdot E$ ensures that this filter behaves 
-approximately as an energy resolution function with constant width 
-$\sigma_e=\sqrt{2}/\sigma_f^0$, \ie\
-\begin{equation}
-\label{eq:filter_b}
-\int d\tilde E\quad P(\tilde E) \frac{1}{\sigma_e\,\sqrt{2\pi}}\,
-e^{-\frac{(E-\tilde E)^2}{2\sigma^2_e}}\,.
-\end{equation}
-The relationship between \eqs~\eqref{eq:filter_a} and~\eqref{eq:filter_b}
-is not obvious and connected to the properties of $P_{\alpha\beta}$: 
-see \Refs~\cite{Kiers:1996zj,Giunti:2003ax}. This feature works \emph{only} 
-for vacuum and constant densities and is controlled
+the last term is a Gau\ss ian filter with width $\sigma_e$ [GeV]; for details,
+see  theory of neutrino oscillations with wave packets, \eg, 
+\Refs~\cite{Kiers:1996zj,Giunti:2003ax}. In fact, it can be shown that this filter
+corresponds to \eq~(\ref{eq:respfun}) with a constant function $\sigma(E)=\sigma_e$.
+ This feature works 
+for vacuum and constant densities \emph{only}, and it is controlled
 by the filer state variable. In addition, $\sigma_e$ is set by the filter value variable:

### 2010-09-21 Documentation/part1.tex
JK - Fixed bug in glb_likelihood (true and fitted rates exchanged)
Revert "meeting-URL corrected. -ML"
This reverts commit 7f96a80099a797c54ea125beeff5e7fdc61f7332.

@@
 The first line of \equ{chi2reactor} is the standard $\chi^2$ expression for
-the Poisson distribution, while the terms in the second line are penalties
+the Gauss distribution, while the terms in the second line are penalties
 for deviations of the systematics parameters from their central values. 
@@
   \hspace*{1.5cm} if(i==GLB\_THETA\_12) /* Prior on $\sin^2 \theta_{12}$ */ \\ 
-\hspace*{2.0cm} pv+=square((startvalue-square(sin(fitvalue)))/inputerror);\\ 
+  \hspace*{2.0cm} pv+=square((centralvalue-square(sin(fitvalue)))/inputerror);\\
   \hspace*{1.5cm} else \\ 
-  \hspace*{2.0cm}  pv+=square((startvalue-fitvalue)/inputerror); \\
+  \hspace*{2.0cm}  pv+=square((centralvalue-fitvalue)/inputerror); \\
   \hspace*{1cm} \} \\

### 2010-09-21 Documentation/tutorials/norm.tex
Revert "meeting-URL corrected. -ML"
This reverts commit 7f96a80099a797c54ea125beeff5e7fdc61f7332.

@@
 A common issue with \GLOBES\ is confusion about the proper units for
-the input flux files for use in \AEDL\ experiment descriptions. 
+the input flux files for use in \AEDL\ experiment descriptions. This
+version describes globes versions 3.0 and later.
 \end{abstract}
@@
 \begin{eqnarray}
-n &=& 5.20034\times x\times E\times f\times\\
+n &=&  x\times E\times f\times\\
 &&\mathtt{@norm}\times\mathtt{@power}\times\mathtt{@stored\_muons}\times\mathtt{@time}\times\mathtt{\$target\_mass}\times(\mathtt{\$baseline})^{-2}\nonumber
 \end{eqnarray}
- Beware: $5.2\ldots$ is a undocumented fudge factor!
 
@@
 \begin{eqnarray}
-\mathtt{@norm}=\frac{1}{5.20034}\left(\frac{\mathrm{GeV}}{\Delta
+\mathtt{@norm}=\left(\frac{\mathrm{GeV}}{\Delta
       E}\right)

### 2010-09-21 Distribution/globes/globes.h

@@ -59,7 +59,7 @@ enum glb_enum_efficiency_types
 
 /* Available minimization algorithms */
 enum glb_enum_minimizers
-       { GLB_MIN_NESTED_POWELL, GLB_MIN_POWELL };
+       { GLB_MIN_NESTED_POWELL, GLB_MIN_POWELL, GLB_MIN_SIMAN };
 #define GLB_MIN_DEFAULT  GLB_MIN_NESTED_POWELL
 
 
@@ -76,6 +76,12 @@ enum glb_enum_minimizers
 #define GLB_EV_TO_KM_FACTOR  1.9747235e-10
 #define GLB_KM_TO_EV(x)      ((x) / GLB_EV_TO_KM_FACTOR)
 #define GLB_EV_TO_KM(x)      ((x) * GLB_EV_TO_KM_FACTOR)
+
+
+/* maximum number of experiments */
+
+#define GLB_MAX_EXP 300
+
 
 
 
@@ -113,9 +119,9 @@ extern int glb_single_experiment_number;
  * declare this one to SWIG and the other global 
  * variables not
  */
-glb_exp glb_experiment_list[32];
+glb_exp glb_experiment_list[GLB_MAX_EXP];
 #else
-extern glb_exp glb_experiment_list[32];
+extern glb_exp glb_experiment_list[GLB_MAX_EXP];
 #endif
 extern int glb_rule_number;
 
@@ -133,6 +139,7 @@ void glbInit(char *name);
 void glbClearExperimentList();
 void glbDefineAEDLVariable(const char *name, double value);
 void glbDefineAEDLList(const char *name, double *list, size_t length);
+double glbGetAEDLVariable(const char *name);
 void glbClearAEDLVariables();
 int glbInitExperiment(char *inf, glb_exp *in, int *counter);
 
@@ -236,6 +243,8 @@ int glbSetSourcePower(int experiment, int flux_ident, double power);
 int glbSetRunningTime(int experiment, int flux_ident, double time);
 int glbSetFilterStateInExperiment(int experiment,int on_off);
 int glbSetFilterInExperiment(int experiment,double filter);
+int glbCompensateFilterInExperiment(int experiment);
+int glbOptimizeSmearingMatrixInExperiment(int experiment);
 
 double glbGetTargetMass(int experiment);
 double glbGetSourcePower(int experiment, int flux_ident);
@@ -245,6 +254,7 @@ double glbGetFilterInExperiment(int experiment);
 int glbGetEminEmax(int experiment, double *emin, double *emax);
[…]

### 2010-09-21 Distribution/lib/argp-fmtstream.c

@@ -1,27 +1,27 @@
 /* Word-wrapping and line-truncating streams
-   Copyright (C) 1997,1998,1999,2001,2002,2003 Free Software Foundation, Inc.
+   Copyright (C) 1997-1999, 2001-2003, 2005, 2009-2010 Free Software
+   Foundation, Inc.
    This file is part of the GNU C Library.
    Written by Miles Bader <miles@gnu.ai.mit.edu>.
 
-   This program is free software; you can redistribute it and/or modify
+   This program is free software: you can redistribute it and/or modify
    it under the terms of the GNU General Public License as published by
-   the Free Software Foundation; either version 2, or (at your option)
-   any later version.
+   the Free Software Foundation; either version 3 of the License, or
+   (at your option) any later version.
 
    This program is distributed in the hope that it will be useful,
    but WITHOUT ANY WARRANTY; without even the implied warranty of
    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
    GNU General Public License for more details.
 
-   You should have received a copy of the GNU General Public License along
-   with this program; if not, write to the Free Software Foundation,
-   Inc., 59 Temple Place - Suite 330, Boston, MA 02111-1307, USA.  */
+   You should have received a copy of the GNU General Public License
+   along with this program.  If not, see <http://www.gnu.org/licenses/>.  */
 
 /* This package emulates glibc `line_wrap_stream' semantics for systems that
    don't have that.  */
 
 #ifdef HAVE_CONFIG_H
-#include <config.h>
+# include <config.h>
 #endif
 
 #include <stdlib.h>
@@ -56,7 +56,7 @@
    Returns NULL if there was an error.  */
 argp_fmtstream_t
 __argp_make_fmtstream (FILE *stream,
-		       size_t lmargin, size_t rmargin, ssize_t wmargin)
+                       size_t lmargin, size_t rmargin, ssize_t wmargin)
 {
   argp_fmtstream_t fs;
 
@@ -73,15 +73,15 @@ __argp_make_fmtstream (FILE *stream,
 
       fs->buf = (char *) malloc (INIT_BUF_SIZE);
       if (! fs->buf)
-	{
-	  free (fs);
-	  fs = 0;
-	}
+        {
+          free (fs);
+          fs = 0;
+        }
       else
-	{
-	  fs->p = fs->buf;
-	  fs->end = fs->buf + INIT_BUF_SIZE;
[…]

### 2010-09-21 Distribution/lib/argp-help.c

@@ -1,28 +1,27 @@
 /* Hierarchial argument parsing help output
-   Copyright (C) 1995-2003, 2004 Free Software Foundation, Inc.
+   Copyright (C) 1995-2005, 2007, 2009-2010 Free Software Foundation, Inc.
    This file is part of the GNU C Library.
    Written by Miles Bader <miles@gnu.ai.mit.edu>.
 
-   This program is free software; you can redistribute it and/or modify
+   This program is free software: you can redistribute it and/or modify
    it under the terms of the GNU General Public License as published by
-   the Free Software Foundation; either version 2, or (at your option)
-   any later version.
+   the Free Software Foundation; either version 3 of the License, or
+   (at your option) any later version.
 
    This program is distributed in the hope that it will be useful,
    but WITHOUT ANY WARRANTY; without even the implied warranty of
    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
    GNU General Public License for more details.
 
-   You should have received a copy of the GNU General Public License along
-   with this program; if not, write to the Free Software Foundation,
-   Inc., 59 Temple Place - Suite 330, Boston, MA 02111-1307, USA.  */
+   You should have received a copy of the GNU General Public License
+   along with this program.  If not, see <http://www.gnu.org/licenses/>.  */
 
 #ifndef _GNU_SOURCE
-# define _GNU_SOURCE	1
+# define _GNU_SOURCE    1
 #endif
 
 #ifdef HAVE_CONFIG_H
-#include <config.h>
+# include <config.h>
 #endif
 
 #include <alloca.h>
@@ -63,15 +62,15 @@
    UPARAM_NAMES vector, below.  */
 
 /* Default parameters.  */
-#define DUP_ARGS      0		/* True if option argument can be duplicated. */
-#define DUP_ARGS_NOTE 1		/* True to print a note about duplicate args. */
-#define SHORT_OPT_COL 2		/* column in which short options start */
-#define LONG_OPT_COL  6		/* column in which long options start */
-#define DOC_OPT_COL   2		/* column in which doc options start */
-#define OPT_DOC_COL  29		/* column in which option text starts */
-#define HEADER_COL    1		/* column in which group headers are printed */
-#define USAGE_INDENT 12		/* indentation of wrapped usage lines */
-#define RMARGIN      79		/* right margin used for wrapping */
+#define DUP_ARGS      0         /* True if option argument can be duplicated. */
+#define DUP_ARGS_NOTE 1         /* True to print a note about duplicate args. */
[…]

### 2010-09-21 Distribution/lib/argp-pvh.c

@@ -1,24 +1,24 @@
 /* Default definition for ARGP_PROGRAM_VERSION_HOOK.
-   Copyright (C) 1996, 1997, 1999, 2004 Free Software Foundation, Inc.
+   Copyright (C) 1996, 1997, 1999, 2004, 2009, 2010 Free Software Foundation,
+   Inc.
    This file is part of the GNU C Library.
    Written by Miles Bader <miles@gnu.ai.mit.edu>.
 
-   This program is free software; you can redistribute it and/or modify
+   This program is free software: you can redistribute it and/or modify
    it under the terms of the GNU General Public License as published by
-   the Free Software Foundation; either version 2, or (at your option)
-   any later version.
+   the Free Software Foundation; either version 3 of the License, or
+   (at your option) any later version.
 
    This program is distributed in the hope that it will be useful,
    but WITHOUT ANY WARRANTY; without even the implied warranty of
    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
    GNU General Public License for more details.
 
-   You should have received a copy of the GNU General Public License along
-   with this program; if not, write to the Free Software Foundation,
-   Inc., 59 Temple Place - Suite 330, Boston, MA 02111-1307, USA.  */
+   You should have received a copy of the GNU General Public License
+   along with this program.  If not, see <http://www.gnu.org/licenses/>.  */
 
 #ifdef HAVE_CONFIG_H
-#include <config.h>
+# include <config.h>
 #endif
 
 #include "argp.h"

### 2010-09-21 Distribution/lib/mempcpy.c

@@ -1,9 +1,9 @@
 /* Copy memory area and return pointer after last written byte.
-   Copyright (C) 2003 Free Software Foundation, Inc.
+   Copyright (C) 2003, 2007, 2009, 2010 Free Software Foundation, Inc.
 
    This program is free software; you can redistribute it and/or modify
    it under the terms of the GNU General Public License as published by
-   the Free Software Foundation; either version 2, or (at your option)
+   the Free Software Foundation; either version 3, or (at your option)
    any later version.
 
    This program is distributed in the hope that it will be useful,
@@ -13,11 +13,11 @@
 
    You should have received a copy of the GNU General Public License
    along with this program; if not, write to the Free Software Foundation,
-   Inc., 59 Temple Place - Suite 330, Boston, MA 02111-1307, USA.  */
+   Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.  */
+
+#include <config.h>
 
 /* Specification.  */
-#include "mempcpy.h"
-
 #include <string.h>
 
 /* Copy N bytes of SRC to DEST, return pointer to bytes after the

### 2010-09-21 Distribution/lib/obstack.h

@@ -1,25 +1,20 @@
 /* obstack.h - object stack macros
-   Copyright (C) 1988-1994,1996-1999,2003,2004 Free Software Foundation, Inc.
+   Copyright (C) 1988-1994, 1996-1999, 2003-2006, 2009-2010 Free Software
+   Foundation, Inc.
+   This file is part of the GNU C Library.
 
-   This file is part of the GNU C Library.  Its master source is NOT part of
-   the C library, however.  The master source lives in /gd/gnu/lib.
-
-   NOTE: The canonical source of this file is maintained with the GNU C Library.
-   Bugs can be reported to bug-glibc@gnu.org.
-
-   This program is free software; you can redistribute it and/or modify
+   This program is free software: you can redistribute it and/or modify
    it under the terms of the GNU General Public License as published by
-   the Free Software Foundation; either version 2, or (at your option)
-   any later version.
+   the Free Software Foundation; either version 3 of the License, or
+   (at your option) any later version.
 
    This program is distributed in the hope that it will be useful,
    but WITHOUT ANY WARRANTY; without even the implied warranty of
    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
    GNU General Public License for more details.
 
-   You should have received a copy of the GNU General Public License along
-   with this program; if not, write to the Free Software Foundation,
-   Inc., 59 Temple Place - Suite 330, Boston, MA 02111-1307, USA.  */
+   You should have received a copy of the GNU General Public License
+   along with this program.  If not, see <http://www.gnu.org/licenses/>.  */
 
 /* Summary:
 
@@ -90,17 +85,17 @@ growing object.  This allows the use of register variables, which
 break the ordinary 'growth' macro.
 
 Summary:
-	We allocate large chunks.
-	We carve out one object at a time from the current chunk.
-	Once carved, an object never moves.
-	We are free to append data of any size to the currently
-	  growing object.
-	Exactly one object is growing in an obstack at any one time.
-	You can run one obstack per control block.
-	You may have as many control blocks as you dare.
-	Because of the way we do it, you can `unwind' an obstack
-	  back to a previous state. (You may remove objects much
-	  as you would with a stack.)
+        We allocate large chunks.
+        We carve out one object at a time from the current chunk.
+        Once carved, an object never moves.
[…]

### 2010-09-21 Distribution/lib/strndup.c

@@ -1,11 +1,11 @@
-/* Copyright (C) 1996, 1997, 1998, 2000, 2003 Free Software Foundation, Inc.
+/* A replacement function, for systems that lack strndup.
 
-   NOTE: The canonical source of this file is maintained with the GNU C Library.
-   Bugs can be reported to bug-glibc@prep.ai.mit.edu.
+   Copyright (C) 1996, 1997, 1998, 2001, 2002, 2003, 2005, 2006, 2007, 2009,
+   2010 Free Software Foundation, Inc.
 
    This program is free software; you can redistribute it and/or modify it
    under the terms of the GNU General Public License as published by the
-   Free Software Foundation; either version 2, or (at your option) any
+   Free Software Foundation; either version 3, or (at your option) any
    later version.
 
    This program is distributed in the hope that it will be useful,
@@ -15,31 +15,16 @@
 
    You should have received a copy of the GNU General Public License
    along with this program; if not, write to the Free Software Foundation,
-   Inc., 59 Temple Place - Suite 330, Boston, MA 02111-1307, USA.  */
+   Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.  */
 
-#ifdef HAVE_CONFIG_H
-# include "config.h"
-#endif
+#include <config.h>
 
-#include <stdlib.h>
 #include <string.h>
 
-#ifndef HAVE_DECL_STRNLEN
-"this configure-time declaration test was not run"
-#endif
-#if !HAVE_DECL_STRNLEN
-size_t strnlen ();
-#endif
-
-#undef __strndup
-#undef strndup
-
-#ifndef weak_alias
-# define __strndup strndup
-#endif
+#include <stdlib.h>
 
 char *
-__strndup (const char *s, size_t n)
+strndup (char const *s, size_t n)
 {
   size_t len = strnlen (s, n);
   char *new = malloc (len + 1);
@@ -50,6 +35,3 @@ __strndup (const char *s, size_t n)
   new[len] = '\0';
   return memcpy (new, s, len);
 }
-#ifdef weak_alias
-weak_alias (__strndup, strndup)
-#endif

### 2010-09-21 Distribution/lib/strnlen.c

@@ -1,10 +1,10 @@
 /* Find the length of STRING, but scan at most MAXLEN characters.
-   Copyright (C) 1996, 1997, 1998, 2000-2003 Free Software Foundation, Inc.
-   This file is part of the GNU C Library.
+   Copyright (C) 2005, 2006, 2007, 2009, 2010 Free Software Foundation, Inc.
+   Written by Simon Josefsson.
 
    This program is free software; you can redistribute it and/or modify
    it under the terms of the GNU General Public License as published by
-   the Free Software Foundation; either version 2, or (at your option)
+   the Free Software Foundation; either version 3, or (at your option)
    any later version.
 
    This program is distributed in the hope that it will be useful,
@@ -12,37 +12,20 @@
    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
    GNU General Public License for more details.
 
-   You should have received a copy of the GNU General Public License along
-   with this program; if not, write to the Free Software Foundation,
-   Inc., 59 Temple Place - Suite 330, Boston, MA 02111-1307, USA.  */
+   You should have received a copy of the GNU General Public License
+   along with this program; if not, write to the Free Software Foundation,
+   Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.  */
 
-#if HAVE_CONFIG_H
-# include <config.h>
-#endif
-#undef strnlen
+#include <config.h>
 
 #include <string.h>
-
-#undef __strnlen
-#undef strnlen
-
-#ifndef _LIBC
-# define strnlen rpl_strnlen
-#endif
-
-#ifndef weak_alias
-# define __strnlen strnlen
-#endif
 
 /* Find the length of STRING, but scan at most MAXLEN characters.
    If no '\0' terminator is found in that many characters, return MAXLEN.  */
 
 size_t
-__strnlen (const char *string, size_t maxlen)
+strnlen (const char *string, size_t maxlen)
 {
   const char *end = memchr (string, '\0', maxlen);
   return end ? (size_t) (end - string) : maxlen;
 }
-#ifdef weak_alias
-weak_alias (__strnlen, strnlen)
-#endif

### 2010-09-21 Distribution/source/glb_compat.c

@@ -19,6 +19,10 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
+#if HAVE_CONFIG_H   /* config.h should come before any other includes */
+#  include "config.h"
+#endif
+
 #include <globes/globes.h>
 
 /* This file contains some deprecated functions, which are provided for
@@ -67,4 +71,4 @@ int glbGetNumberOfSimBins(int exp)
 {
   return glbGetNumberOfSamplingPoints(exp);
 }
- 
+

### 2010-09-21 Distribution/source/glb_error.c

@@ -27,6 +27,9 @@
 //Gary V. Vaughan.
 
 
+#if HAVE_CONFIG_H   /* config.h should come before any other includes */
+#  include "config.h"
+#endif
 
 #include <stdlib.h>
 #include <string.h>
@@ -64,8 +67,8 @@ int glbSetVerbosityLevel(int level)
 static void
 error (int exit_status, const char *mode, const char *message,int verb_level)
 {
-  
-  if((verb_level==1)||(exit_status)>=0) 
+
+  if((verb_level==1)||(exit_status)>=0)
     fprintf (stderr, "%s: %s: %s.\n", glb_prog_name, mode, message);
 
   if (exit_status >= 0)
@@ -88,7 +91,7 @@ glb_error (const char *message)
   error (-1, "ERROR", message,v);
 }
 
-void  
+void
 glb_fatal (const char *message)
 {
   /* A FATAL message always is displayed */
@@ -158,53 +161,81 @@ FILE *glb_fopen(const char *filename, const char *mode)
   size_t i,a,b;
   char *test_name=NULL;
   char *new_name=NULL;
- 
+
   FILE *t;
-  register FILE *value; 
-  for(i=0;i<glb_path_vector_length;i++)
+  register FILE *value;
+
+ /* handling an absolute path */
+  if(strlen(filename)>0 && filename[0] == '/')
     {
-      a=strlen(filename);
-      b=strlen(glb_path_vector[i]);
-      test_name=glb_malloc((a+b+2)*sizeof(char));
-      test_name=strcpy(test_name,glb_path_vector[i]);
-      test_name=strcat(test_name,filename);
-      if(verbosity_level >= 4) fprintf(stderr,"Searched path: %s\n",test_name);
-      t=fopen(test_name,"r");
-      if(t!=NULL) 
+      if(verbosity_level >= 4)
+	fprintf(stderr,"Searched path: %s\n",filename);
+      t=fopen(filename,"r");
+      if(t!=NULL)
 	{
-	  new_name=strdup(test_name);
+	  new_name=strdup(filename);
 	  fclose(t);
-	  glb_free(test_name);
-	  break;
 	}
-      glb_free(test_name);
-      /* Repeating the exercise with an additional '/' between
-       * path and filename.
-       */
-      a=strlen(filename);
-      b=strlen(glb_path_vector[i]);
-      test_name=glb_malloc((a+b+2)*sizeof(char));
[…]

### 2010-09-21 Distribution/source/glb_fluxes.c

@@ -19,7 +19,9 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
-
+#if HAVE_CONFIG_H   /* config.h should come before any other includes */
+#  include "config.h"
+#endif
 
 
 /* Units and Normalization */
@@ -35,37 +37,123 @@
 
 #include <stdio.h>
 #include <stdlib.h>
+#include <string.h>
 #include <math.h>
 
 #include "glb_types.h"
 #include "glb_error.h"
+#include "glb_fluxes.h"
 #include "glb_multiex.h"
 
 #define PI 3.1415
 #define NORM 6.02204*1E-12 //10^-38/u (atomic mass unit)
-#define GLB_ELECTRON_MASS 0.511E-3 
+#define GLB_ELECTRON_MASS 0.511E-3
 
 static double mmu = 0.1056;           // muon mass in GeV
 static double beam_pol = 0;           // beam polaristaion
 static double beam_en = 20;           // beam energy in GeV
 static double beta,rgamma;             // relativistic quantities
-static double stored_muons = 2.0E20;  // stored muons per year 
+static double stored_muons = 2.0E20;  // stored muons per year
 // or protons on target
 // care must be taken with this, it is a different number for each experiment
 static double power;
 static int    channel = 1;            // channel
-static int NFLUX = 500;
+static int FLUX_STEPS = 500;          // Number of sampling points for builtin fluxes
 
 
-void glb_flux_loader(glb_flux *data, int number, int polarity);
+
+/***************************************************************************
+ *                    H E L P E R   F U N C T I O N S                      *
+ ***************************************************************************/
+
+/***************************************************************************
+ * Function glb_load_n_columns                                             *
+ ***************************************************************************
+ * Load data from file. The expected format is n_columns columns and an    *
+ * arbitray number of rows. The resulting number of rows will be stored in *
+ * n_lines, while for the actual data, arrays will be created (one for     *
[…]

### 2010-09-21 Distribution/source/glb_glb.c

@@ -19,7 +19,6 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
-
 #ifdef HAVE_CONFIG_H
 #include "config.h"
 #endif
@@ -36,18 +35,18 @@
 
 /* Moved from glb_wrapper.c, in order to have the module support stuff
  * only used in this file.
- */ 
-static void final_clean() 
+ */
+static void final_clean()
 {
   glb_builtin_prior_clean();
-  glb_clean_up(); 
-  return; 
+  glb_clean_up();
+  return;
 }
 
 static int test=1234;
 void glbInit(char *name)
 {
-  
+
 
   atexit(final_clean);
   glb_init(name);
@@ -66,12 +65,12 @@ void glbInit(char *name)
   /* Select default minimizer */
   glbSelectMinimizer(GLB_MIN_DEFAULT);
 
-  /* Setup default priors */  
+  /* Setup default priors */
   glbRegisterPriorFunction(glb_builtin_prior_prior,
 			   glb_builtin_prior_starting_values,
 			   glb_builtin_prior_input_errors,
                            &test);
-  
+
   /* Register built-in chi^2 functions */
   /* When making any changes here, don't forget to update the array
    * glb_2011_compatible_chi_functions in glb_sys.c and the functions

### 2010-09-21 Distribution/source/glb_min_sup.c

@@ -19,435 +19,26 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
-
-
+#if HAVE_CONFIG_H   /* config.h should come before any other includes */
+#  include "config.h"
+#endif
 
 #include <math.h>
-
 #include <stdio.h>
-
 #include "glb_error.h"
 #include "glb_min_sup.h"
-
-static void glb_minimizer_error(char error_text[])
-{
-  glb_warning(error_text);  
-}
-
-/* This here is code which works fine, but somehow valgrind complains
- * about leaking memory. Therefore we use the version just below this
- *  comment, which works fine and makes valgrind happy
-
- double *glb_alloc_vec(int nl,int nh)
- {
- double *v;
- v=(double *)glb_malloc((unsigned) (nh-nl+1)*sizeof(double));
- return v-nl;
- }
- 
- double **glb_alloc_mat(int nrl,int nrh, int ncl,int nch)
- {
- int i;
- double **m;
- 
- m=(double **) glb_malloc((unsigned) (nrh-nrl+1)*sizeof(double*));
- m -= nrl;
- for(i=nrl;i<=nrh;i++) 
- {
- m[i]=(double *) glb_malloc((unsigned) 
- (nch-ncl+1)*sizeof(double));
- m[i] -= ncl;
- }
- return m;
- }
- 
- void glb_free_vec(double *v,int nl,int nh)
- {
- glb_free((char*) (v+nl));
- }
- 
- void glb_free_mat(double **m,int nrl,int nrh,int ncl,int nch)
- {
- int i;
- for(i=nrh;i>=nrl;i--) glb_free((char*) (m[i]+ncl));
- glb_free((char*) (m+nrl));
- }
- */
-
-double *glb_alloc_vec(int nl,int nh)
-{
-	double *v;
-	v=(double *)glb_malloc((unsigned) (nh+1)*sizeof(double));
-	return v;
-}
-
-double **glb_alloc_mat(int nrl,int nrh, int ncl,int nch)
-{
-	int i;
-	double **m;
[…]

### 2010-09-21 Distribution/source/glb_min_sup.h

@@ -19,10 +19,17 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
-
 #ifndef GLB_MIN_SUP
 #define GLB_MIN_SUP 1
 
+#if HAVE_CONFIG_H   /* config.h should come before any other includes */
+#  include "config.h"
+#endif
+
+#define SQR(x)    ((x)*(x))
+#define MIN(a,b)  ((a) < (b) ? (a) : (b))
+#define MAX(a,b)  ((a) > (b) ? (a) : (b))
+#define SIGN(a,b) ((b) > 0.0 ? fabs(a) : -fabs(a))
 double *glb_alloc_vec(int nl,int nh);
 double **glb_alloc_mat(int nrl,int nrh, int ncl,int nch);
 void glb_free_vec(double *v,int nl,int nh);

### 2010-09-21 Distribution/source/glb_modules.c

@@ -19,6 +19,9 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
+#if HAVE_CONFIG_H   /* config.h should come before any other includes */
+#  include "config.h"
+#endif
 
 #include <ltdl.h>
 #include <stdio.h>
@@ -26,7 +29,6 @@
 #include <unistd.h>
 #include <globes/globes.h>
 #include <globes/glb-modules.h>
-
 #include "glb_path.h"
 #include "glb_error.h"
 
@@ -62,7 +64,7 @@ void glb_close_module_support()
 /* Set up path to be searched for modules */
 int glb_setup_module_search_path()
 {
- 
+
   int i,s=1;
   char *in_path;
   char *path;
@@ -77,11 +79,11 @@ int glb_setup_module_search_path()
   /* Parse the path -- try this aloud twenty times as fast as you can ;-) */
   s=glb_break_up_path(path,&path_vector,&path_vector_length);
   glb_free(path);
-  
+
   /* Setup the correct directory to find the builtin modules */
   s += lt_dladdsearchdir(GLB_MODULE_DIR);
 
-  /* Adding the elements of path vector to the module search path 
+  /* Adding the elements of path vector to the module search path
    * ommiting the first (which would be the local directory) */
   for(i=1;i<path_vector_length;i++) s += lt_dladdsearchdir(path_vector[i]);
 
@@ -89,14 +91,14 @@ int glb_setup_module_search_path()
   for(i=0;i<path_vector_length;i++)
     glb_free((char *) path_vector[i]);
   glb_free((char **) path_vector);
-  
+
   return s;
 }
 
 #define GLB_MODULE_NOT_FOUND -2
 #define GLB_MODULE_NO_INIT -1
 
-/* Checking wether a module exists and is a GLoBES module 
+/* Checking wether a module exists and is a GLoBES module
  * The idea is to adjust verbosity level to whatever the
  * user needs, if he wants to avoid later error messages
  * he has to explicitely call glb_probe_module before any
@@ -135,13 +137,13 @@ int glbProbeModule(const char *module_name, int verbosity)
       fprintf(stderr,"glb_probe_module: ERROR: in module %s: %s\n",
 	      module_name,error);
[…]

### 2010-09-21 Distribution/source/glb_modules.h

@@ -19,9 +19,12 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
-
 #ifndef GLB_MODULES_H
 #define GLB_MODULES_H 1
+
+#if HAVE_CONFIG_H   /* config.h should come before any other includes */
+#  include "config.h"
+#endif
 
 #include <globes/globes.h>
 

### 2010-09-21 Distribution/source/glb_parser.c

@@ -1,7 +1,9 @@
-/* A Bison parser, made by GNU Bison 1.875.  */
+/* A Bison parser, made by GNU Bison 2.3.  */
 
-/* Skeleton parser for Yacc-like parsing with Bison,
-   Copyright (C) 1984, 1989, 1990, 2000, 2001, 2002 Free Software Foundation, Inc.
+/* Skeleton implementation for Bison's Yacc-like parsers in C
+
+   Copyright (C) 1984, 1989, 1990, 2000, 2001, 2002, 2003, 2004, 2005, 2006
+   Free Software Foundation, Inc.
 
    This program is free software; you can redistribute it and/or modify
    it under the terms of the GNU General Public License as published by
@@ -15,16 +17,24 @@
 
    You should have received a copy of the GNU General Public License
    along with this program; if not, write to the Free Software
-   Foundation, Inc., 59 Temple Place - Suite 330,
-   Boston, MA 02111-1307, USA.  */
+   Foundation, Inc., 51 Franklin Street, Fifth Floor,
+   Boston, MA 02110-1301, USA.  */
 
-/* As a special exception, when this file is copied by Bison into a
-   Bison output file, you may use that output file without restriction.
-   This special exception was added by the Free Software Foundation
-   in version 1.24 of Bison.  */
+/* As a special exception, you may create a larger work that contains
+   part or all of the Bison parser skeleton and distribute that work
+   under terms of your choice, so long as that work isn't itself a
+   parser generator using the skeleton or a modified version thereof
+   as a parser skeleton.  Alternatively, if you modify or redistribute
+   the parser skeleton itself, you may (at your option) remove this
+   special exception, which will cause the skeleton and the resulting
+   Bison output files to be licensed under the GNU General Public
+   License without this special exception.
 
-/* Written by Richard Stallman by simplifying the original so called
[…]

### 2010-09-21 Distribution/source/glb_parser.h

@@ -1,7 +1,9 @@
-/* A Bison parser, made by GNU Bison 1.875.  */
+/* A Bison parser, made by GNU Bison 2.3.  */
 
-/* Skeleton parser for Yacc-like parsing with Bison,
-   Copyright (C) 1984, 1989, 1990, 2000, 2001, 2002 Free Software Foundation, Inc.
+/* Skeleton interface for Bison's Yacc-like parsers in C
+
+   Copyright (C) 1984, 1989, 1990, 2000, 2001, 2002, 2003, 2004, 2005, 2006
+   Free Software Foundation, Inc.
 
    This program is free software; you can redistribute it and/or modify
    it under the terms of the GNU General Public License as published by
@@ -15,13 +17,21 @@
 
    You should have received a copy of the GNU General Public License
    along with this program; if not, write to the Free Software
-   Foundation, Inc., 59 Temple Place - Suite 330,
-   Boston, MA 02111-1307, USA.  */
+   Foundation, Inc., 51 Franklin Street, Fifth Floor,
+   Boston, MA 02110-1301, USA.  */
 
-/* As a special exception, when this file is copied by Bison into a
-   Bison output file, you may use that output file without restriction.
-   This special exception was added by the Free Software Foundation
-   in version 1.24 of Bison.  */
+/* As a special exception, you may create a larger work that contains
+   part or all of the Bison parser skeleton and distribute that work
+   under terms of your choice, so long as that work isn't itself a
+   parser generator using the skeleton or a modified version thereof
+   as a parser skeleton.  Alternatively, if you modify or redistribute
+   the parser skeleton itself, you may (at your option) remove this
+   special exception, which will cause the skeleton and the resulting
+   Bison output files to be licensed under the GNU General Public
+   License without this special exception.
+
+   This special exception was added by the Free Software Foundation in
+   version 2.2 of Bison.  */
 
 /* Tokens.  */
 #ifndef YYTOKENTYPE
@@ -48,21 +58,22 @@
      VERS = 274,
[…]

### 2010-09-21 Distribution/source/glb_parser_addons.h

@@ -19,13 +19,12 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
-
-
-
-
-
 #ifndef GLB_MY_PARSER_H
 #define GLB_MY_PARSER_H 1
+
+#if HAVE_CONFIG_H   /* config.h should come before any other includes */
+#  include "config.h"
+#endif
 
 #include <globes/globes.h>
 

### 2010-09-21 Distribution/source/glb_path.h

@@ -19,12 +19,12 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
-
-
-
-
 #ifndef GLB_PATH_H
 #define GLB_PATH_H 1
+
+#if HAVE_CONFIG_H   /* config.h should come before any other includes */
+#  include "config.h"
+#endif
 
 int glb_break_up_path(char *in_path, char ***pathv, size_t *len);
 int glb_setup_path();

### 2010-09-21 Distribution/source/glb_prior.c

@@ -18,19 +18,16 @@
  * You should have received a copy of the GNU General Public License
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
-
  */
-
+#if HAVE_CONFIG_H   /* config.h should come before any other includes */
+#  include "config.h"
+#endif
 
 /* Headers needed by this module */
 #include <stdio.h>
 #include <stdlib.h>
 #include <math.h>
 #include <string.h>
-
-#ifdef HAVE_CONFIG_H
-#include "config.h"
-#endif
 
 #include "globes/globes.h"
 #include "glb_prior.h"
@@ -39,17 +36,17 @@
 static glb_params sv;
 static glb_params er;
 
-/* glb_module_init is mandatory. Any module must have glb_module_init, 
- * which returns 0 on succes or -1 on failure. This function is called 
+/* glb_module_init is mandatory. Any module must have glb_module_init,
+ * which returns 0 on succes or -1 on failure. This function is called
  * when the module is loaded.
  *
- * In our case it allocates memory for sv and er and sets sv/er to any 
- * input errors starting values which have already been defined 
- */   
+ * In our case it allocates memory for sv and er and sets sv/er to any
+ * input errors starting values which have already been defined
+ */
 int glb_builtin_prior_init()
 {
   sv=glbAllocParams();
-  er=glbAllocParams(); 
+  er=glbAllocParams();
   glbGetStartingValues(sv);
   glbGetInputErrors(er);
 
@@ -69,20 +66,20 @@ void glb_builtin_prior_clean()
 }
 
 /* This is a local function which implements the prior calculation
- * and ensure that the builtin prior behaves as described in the 
+ * and ensure that the builtin prior behaves as described in the
  * documentation, i.e. it returns 0 when the error is below 1E-12
  */
 static double sprior(double x, double center, double sigma)
 {
   if(fabs(sigma-0)<1E-12) return 0;
   return (x-center)*(x-center)/sigma/sigma;
-}  
+}
 
 /* glb_builtin_prior_prior is mandatory (for prior modules).
  *
  * This is the default prior, i.e. priors are only added for free
  * parameters, and the prior accepts starting values as stored in sv
[…]

### 2010-09-21 Distribution/source/glb_probability.h

@@ -19,13 +19,12 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
-
-
-
-
-
 #ifndef GLB_OSZPROB_H
 #define GLB_OSZPROB_H 1
+
+#if HAVE_CONFIG_H   /* config.h should come before any other includes */
+#  include "config.h"
+#endif
 
 #include <complex.h>
 #include <globes/globes.h>
@@ -55,8 +54,8 @@ extern void *glb_probability_user_data;
 
 
 /* Functions */
-inline void zhetrd3(double complex A[3][3], double complex Q[3][3],
-                    double d[3], double e[2]);
+//JK 2008-07-22 inline void zhetrd3(double complex A[3][3], double complex Q[3][3],
+//                    double d[3], double e[2]);
 int zheevc3(double complex A[3][3], double w[3]);
 int zheevq3(double complex A[3][3], double complex Q[3][3], double w[3]);
 int zheevh3(double complex A[3][3], double complex Q[3][3], double w[3]);

### 2010-09-21 Distribution/source/glb_profile.c

@@ -19,9 +19,9 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
-
-
-
+#if HAVE_CONFIG_H   /* config.h should come before any other includes */
+#  include "config.h"
+#endif
 
 #include <stdio.h>
 #include <stdlib.h>
@@ -41,44 +41,44 @@
  * -----------------------------------------------------------
  */
 static const double stacey[501] = {
- 2.8, 2.92, 3.05, 3.18, 3.30, 3.30, 3.31, 3.32, 3.32, 3.33, 3.34, 3.35, 3.35, 
-3.36, 3.37, 3.38, 3.38, 3.39, 3.40, 3.40, 3.41, 3.42, 3.43, 3.43, 3.44, 3.45, 
-3.46, 3.46, 3.47, 3.48, 3.48, 3.49, 3.73, 3.81, 3.82, 3.84, 3.86, 3.87, 3.89, 
-3.90, 3.92, 3.94, 3.95, 3.97, 3.98, 4.00, 4.02, 4.03, 4.05, 4.06, 4.08, 4.09, 
-4.40, 4.40, 4.41, 4.42, 4.42, 4.43, 4.44, 4.45, 4.45, 4.46, 4.47, 4.47, 4.48, 
-4.49, 4.49, 4.50, 4.51, 4.52, 4.52, 4.53, 4.54, 4.54, 4.55, 4.56, 4.56, 4.57, 
-4.58, 4.58, 4.59, 4.60, 4.61, 4.61, 4.62, 4.63, 4.63, 4.64, 4.65, 4.65, 4.66, 
-4.67, 4.68, 4.68, 4.69, 4.70, 4.70, 4.71, 4.72, 4.72, 4.73, 4.74, 4.75, 4.75, 
-4.76, 4.77, 4.77, 4.78, 4.79, 4.79, 4.80, 4.81, 4.82, 4.82, 4.83, 4.84, 4.84, 
-4.85, 4.86, 4.86, 4.87, 4.88, 4.89, 4.89, 4.90, 4.91, 4.91, 4.92, 4.93, 4.93, 
-4.94, 4.95, 4.95, 4.96, 4.97, 4.98, 4.98, 4.99, 5.00, 5.00, 5.01, 5.02, 5.02, 
[…]

### 2010-09-21 Distribution/source/glb_rate_engine.h

@@ -19,14 +19,17 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
-
-
-
-
 #ifndef GLB_CALC_H
 #define GLB_CALC_H 1
 
+#if HAVE_CONFIG_H   /* config.h should come before any other includes */
+#  include "config.h"
+#endif
+
 #include "glb_types.h"
+
+#define GLB_FAST_RATES 1
+#define GLB_SLOW_RATES 0
 
 extern int glb_num_of_rules;
 extern double glb_glb_sig_norm_error[32];
@@ -68,6 +71,7 @@ void glb_set_bg_rule(int i, int cn, int *rule, double *coeff);
 
 void glb_set_rates();
 void glb_set_new_rates();
+void glb_rate_template();
 void glb_set_signal_errors(int i,double norm, double tilt);
 void glb_set_bg_errors(int i,double norm, double tilt);
 void glb_set_bg_center(int i,double norm, double tilt);
@@ -138,6 +142,7 @@ extern double *glb_calc_chrb_0[32];
 extern double *glb_calc_chrb_1[32];
 extern double *glb_calc_chra_0[32];
 extern double *glb_calc_chra_1[32];
+extern double *glb_calc_chr_template[32];
 
 extern double *glb_calc_user_pre_sm_channel[32];
 extern double *glb_calc_user_post_sm_channel[32];

### 2010-09-21 Distribution/source/glb_smear.h

@@ -20,8 +20,6 @@
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
 
-
-
 /* This routines computes the smear matrix in the same way as CalcSmearA
    in GlobesMESbeta.m.
 
@@ -31,6 +29,10 @@
 
 #ifndef GLB_SMEAR_H
 #define GLB_SMEAR_H 1
+
+#if HAVE_CONFIG_H   /* config.h should come before any other includes */
+#  include "config.h"
+#endif
 
 #include "glb_types.h"
 
@@ -57,7 +59,8 @@ glb_smear  *glb_smear_reset(glb_smear *in);
 void glb_compute_smearing_matrix(double ***matrix, 
 			   int **low, int **up, glb_smear *in
 			   ,const struct glb_experiment *head);
-
+void glb_optimize_smearing_matrix(glb_smear *s, double **matrix, int *lower, int *upper);
+void glb_filter_compensate(glb_smear *s, double **matrix, int *lower, int *upper);
 
 
 void glb_set_up_smear_data(glb_smear *test,const struct glb_experiment *head);

### 2010-09-21 Distribution/source/glb_sys.c

@@ -19,6 +19,10 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
+#if HAVE_CONFIG_H   /* config.h should come before any other includes */
+#  include "config.h"
+#endif
+
 #include <math.h>
 #include <string.h>
 #include <globes/globes.h>
@@ -31,7 +35,7 @@
 /* Global variables */
 glb_systematic *glb_sys_list;   /* Connected list of systematics definitions */
 
-static const char *glb_2011_compatible_chi_functions[] = 
+static const char *glb_2011_compatible_chi_functions[] =
   {
     "chiSpectrumTilt",
     "chiNoSysSpectrum",
@@ -77,7 +81,7 @@ int glbIsChiFunction2011Compatible(const char *name)
 glb_systematic *glbFindChiFunctionByName(const char *name)
 {
   glb_systematic *sys;
-  
+
   for (sys = glb_sys_list; sys != NULL; sys = sys->next)
     if (strcmp(name, sys->name) == 0)
       break;
@@ -103,7 +107,7 @@ glb_systematic *glbFindChiFunctionByName(const char *name)
 char *glbConvertErrorDim(int errordim)
 {
   char name[50];
-  
+
   switch (errordim)
   {
     case 0:
@@ -166,7 +170,7 @@ int glbSetChiFunctionInRule(struct glb_experiment *exp, int rule, int on_off,
   glb_systematic *sys;
   double tmp_errorlist[5];
   int k;
-  
+
   /* Something tells me that people will come up with the idea to pass
    * exp->sys_on_errors; this has to be blockes */
   if (errors != NULL  &&
@@ -175,7 +179,7 @@ int glbSetChiFunctionInRule(struct glb_experiment *exp, int rule, int on_off,
     glb_error("glbSetChiFunctionInRule: Invalid error array");
     return -1;
   }
-  
+
   /* Find requested glb_sys_list entry */
   if ((sys=glbFindChiFunctionByName(sys_id)) == NULL)
   {
@@ -211,7 +215,7 @@ int glbSetChiFunctionInRule(struct glb_experiment *exp, int rule, int on_off,
     }
   }
 
- 
+
   if (on_off == GLB_ON)
   {
     if (rule >= 0 && rule < exp->numofrules)
[…]

### 2010-09-21 Distribution/source/glb_version.c

@@ -19,7 +19,9 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
-
+#if HAVE_CONFIG_H   /* config.h should come before any other includes */
+#  include "config.h"
+#endif
 
 #include <stdlib.h>
 #include <string.h>
@@ -34,16 +36,16 @@
 #include "glb_error.h"
 
 const char *glb_release_version = VERSION;
-static const char *library_version = GLB_LIBVERSION; 
+static const char *library_version = GLB_LIBVERSION;
 
-static int 
+static int
 break_up_version(char const *version,int *first,int *second, int *third)
 {
-  
+
   char c;
   size_t length,i,k,ct=0,ctb=0;
   char *vec=NULL;
-  int res[3]={-1,-1,-1};
+  int res[4]={-1,-1,-1,-1};
   if(version==NULL)
     {
       *third=-1;
@@ -61,11 +63,11 @@ break_up_version(char const *version,int *first,int *second, int *third)
 	  vec=(char *) glb_realloc(vec,sizeof(char)*(ct+1));
 	  vec[ct-1]=c;
 	  vec[ct]='\0';
-	} 
+	}
       else
 	{
 	  ctb++;
-	  if((ctb>3)||(vec==NULL)) 
+	  if((ctb>4)||(vec==NULL))
 	    {glb_fatal("Invalid version string");exit(1);}
 	  sscanf(vec,"%d",&res[ctb-1]);
 	  ct=0;
@@ -86,12 +88,12 @@ glbTestReleaseVersion(const char *version)
   int major,minor,bug,tmajor,tminor,tbug;
   break_up_version(version,&major,&minor,&bug);
   break_up_version(glb_release_version,&tmajor,&tminor,&tbug);
-  
-  if(major==tmajor) 
+
+  if(major==tmajor)
     {
-      if(minor==tminor) 
+      if(minor==tminor)
 	{
-	  if(bug==tbug) 
+	  if(bug==tbug)
 	    {
 	      s=0;
 	    }
@@ -113,11 +115,11 @@ glbTestLibraryVersion(const char *version)
   int major,minor,bug,tmajor,tminor,tbug;
   break_up_version(version,&major,&bug,&minor);
   break_up_version(library_version,&tmajor,&tbug,&tminor);
-  if(major==tmajor) 
[…]

### 2010-09-21 Distribution/source/glb_wrapper.h

@@ -19,12 +19,12 @@
  * along with this program; if not, write to the Free Software
  * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
  */
-
-
-
-
 #ifndef glb_WRAPPER_H
 #define  glb_WRAPPER_H 1
+
+#if HAVE_CONFIG_H   /* config.h should come before any other includes */
+#  include "config.h"
+#endif
 
 #include <globes/globes.h>
 

### 2010-09-21 Documentation/tutorials/LBNE-tutorial/th13chi2-p5.c

@@ -1,0 +1,125 @@
+/* GLoBES -- General LOng Baseline Experiment Simulator
+ * (C) 2002 - 2007,  The GLoBES Team
+ *
+ * GLoBES is mainly intended for academic purposes. Proper
+ * credit must be given if you use GLoBES or parts of it. Please
+ * read the section 'Credit' in the README file.
+ *
+ * This program is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU General Public License as published by
+ * the Free Software Foundation; either version 2 of the License, or
+ * (at your option) any later version.
+ *
+ * This program is distributed in the hope that it will be useful,
+ * but WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ * GNU General Public License for more details.
+ *
+ * You should have received a copy of the GNU General Public License
+ * along with this program; if not, write to the Free Software
+ * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
+ */
+
+/***************************************************************************
+ *            G L o B E S   -   M I N I - W O R K S H O P                  *
+ *                                                                         *
+ *                     May 13, 2010, Fermilab                              *
+ *                                                                         *
+ * Hands-on session on LBNE: Computing chi^2 for the wrong hierarchy       *
+ * as a function of th13                                                   *
+ *                                                                         *
+ * Solution to problem 5: Implementation of the tracking algorithm         *
+ ***************************************************************************/
+
+
+#include <stdio.h>
+#include <stdlib.h>
+#include <math.h>
+#include <string.h>
+
+#include <globes/globes.h>            /* GLoBES library */
+
+/* Output file, empty string = stdout */
+char MYFILE[]="th13chi2.dat";
+FILE *outfile = NULL;
+
+int main(int argc, char *argv[])
+{
+  /* Initialize libglobes */
+  glbInit(argv[0]);
+
+  /* New minimization algorithm -> significant speed-up */
[…]

### 2010-09-21 Documentation/tutorials/LBNE-tutorial/th13chi2-p6.c

@@ -1,0 +1,184 @@
+/* GLoBES -- General LOng Baseline Experiment Simulator
+ * (C) 2002 - 2007,  The GLoBES Team
+ *
+ * GLoBES is mainly intended for academic purposes. Proper
+ * credit must be given if you use GLoBES or parts of it. Please
+ * read the section 'Credit' in the README file.
+ *
+ * This program is free software; you can redistribute it and/or modify
+ * it under the terms of the GNU General Public License as published by
+ * the Free Software Foundation; either version 2 of the License, or
+ * (at your option) any later version.
+ *
+ * This program is distributed in the hope that it will be useful,
+ * but WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ * GNU General Public License for more details.
+ *
+ * You should have received a copy of the GNU General Public License
+ * along with this program; if not, write to the Free Software
+ * Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
+ */
+
+/***************************************************************************
+ *            G L o B E S   -   M I N I - W O R K S H O P                  *
+ *                                                                         *
+ *                     May 13, 2010, Fermilab                              *
+ *                                                                         *
+ * Hands-on session on LBNE: Computing chi^2 for the wrong hierarchy       *
+ * as a function of th13                                                   *
+ *                                                                         *
+ * Solution to problem 6: Usage of degfinder                               *
+ ***************************************************************************/
+
+
+#include <stdio.h>
+#include <stdlib.h>
+#include <math.h>
+#include <string.h>
+
+#include <globes/globes.h>            /* GLoBES library */
+#include "degfinder.h"
+
+/* Output file, empty string = stdout */
+char MYFILE[]="th13chi2.dat";
+FILE *outfile = NULL;
+
+int main(int argc, char *argv[])
+{
+  int i;
+
+  /* Initialize libglobes */
+  glbInit(argv[0]);
+
[…]

### 2010-09-27 Documentation/Manual.tex
meeting-URL corrected. -ML
Revert "meeting-URL corrected. -ML"
This reverts commit 7f96a80099a797c54ea125beeff5e7fdc61f7332.
Manual changes performed, again
Small addl. manual changes

@@
 
+\usepackage{hyperref}
+\hypersetup{colorlinks,citecolor=red,filecolor=red,linkcolor=blue,urlcolor=green}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\usepackage{ae,aecompl} % only for Type 1 fonts for readable pdf convert; may not be necessary anymore!
+\usepackage{ae,aecompl} % only for Type 1 fonts for readable pdf convert
 
@@
 \begin{center}
-Version from \today\ for \GLOBES\ 3.0
+Version from \today\ for \GLOBES\ 3.1
 \end{center}
@@
 \begin{quote}
-    Copyright \copyright  2004--2007  The GLoBES Team.
+    Copyright \copyright  2004--2010  The GLoBES Team.
     Permission is granted to copy, distribute and/or modify this document
@@
  New features in the simulation of neutrino oscillation experiments with \GLOBES~3.0,
-arXiv:\glbxxxtwo,
+Comput. Phys. Commun. 177 (2007) 432, arXiv:\glbxxxtwo,
 \end{quote}

### 2010-09-27 Documentation/part1.tex

--- 2007-02-09 Documentation/part1.tex
@@
 \begin{function}
-\GLBNS{GetEnergyWindow}
+\GLBNS{GetEnergyWindow} \label{glbGetEnergyWindow}
 {\tt int glbGetEnergyWindow(int exp, int rule, double *low, double *high)} returns the
 energy window [{\tt low}, {\tt high}] defined in the \AEDL\ file for the experiment
-{\tt exp} and rule {\tt rule}.
+{\tt exp} and rule {\tt rule}. See also {\tt glbSetEnergyWindow} on page~\pageref{glbSetEnergyWindow}.
 \end{function}
@@
 \begin{function}
-\GLBNS{GetCoefficientInRule}
+\GLBNS{GetCoefficientInRule} \label{glbGetCoefficientInRule}
 {\tt double glbGetCoefficientInRule(int exp, int rule, int pos, int signal)}
@@
 The parameter {\tt signal} refers to signal (\GLBC{GLB\_SIG}) or background
-(\GLBC{GLB\_BG}).
+(\GLBC{GLB\_BG}). See also {\tt glbSetCoefficientInRule} on page~\pageref{glbSetCoefficientInRule}.
 \end{function}
@@
 \end{function}
+In addition, the coefficients in the rules can be manipulated by\footnote{Available in GLoBES 3.1.7 and higher.}
+\begin{function}
+\GLBNS{SetCoefficientInRule} \label{glbSetCoefficientInRule}
+{\tt int glbSetCoefficientInRule(int exp, int rule, int pos, int signal, double coeff)}
+sets the coefficient of the component {\tt pos} in rule {\tt rule} 
+of the experiment {\tt exp} to value {\tt coeff}.
+The parameter {\tt signal} refers to signal (\GLBC{GLB\_SIG}) or background
+(\GLBC{GLB\_BG}). See also {\tt glbGetCoefficientInRule} on page~\pageref{glbGetCoefficientInRule}.
+\end{function}
+
 % ELIMINATED FROM DOC, BECAUSE WE WANT TO AVOID THE BG CENTERS IN THE FUTURE (CONFUSION) [WW]
@@
 \end{function}
-As usual, all these functions return $-1$ if they were not successful.
+Furthermore, the energy window can be manipulated by\footnote{Available in GLoBES 3.1.7 and higher.}
+\begin{function}
+\GLBNS{SetEnergyWindow} \label{glbSetEnergyWindow}
+{\tt int glbSetEnergyWindow(int experiment, int rule, double low, double high)}
+sets the energy window in {\tt experiment} and {\tt rule}, which can be also {\tt GLB\_ALL}.
+The energy window is limited by the energies {\tt low} and {\tt high}.
+See also {\tt glbGetEnergyWindow} on page~\pageref{glbGetEnergyWindow}.
+\end{function}
+As usual, all the set-functions return $-1$ if they were not successful.
 %.
@@
 parameter vectors every time within the loop as long as the
-number of experiments does not change.
+number of experiments does not change. The counterpart of that
+function is\footnote{Supported by version 3.1.7 and higher.}
+\begin{function}
+\GLBNS{GetAEDLVariable}
+{\tt double glbGetAEDLVariable(const char* name)}
+returns the value of the \AEDL\ variable {\tt name}.
+\end{function}
 

### 2010-09-28 Documentation/part1.tex
Further functions added to manual

@@
 %
+The file name of an \AEDL\ file can be retrived with\footnote{The following two functions are available in \GLOBES\ 3.1.7 and higher.}
+\begin{function}
+\GLBNS{GetFilenameOfExperiment}
+{\tt const char *glbGetFilenameOfExperiment(int experiment)} returns the file name of the \AEDL\ file for the experiment number {\tt experiment}. Note that the returned pointer cannot be modified.
+\end{function}
+and the references to be cited with
+\begin{function}
+\GLBNS{GetCitationForExperiment} \label{glbGetCitationForExperiment}
+{\tt const char *glbGetCitationForExperiment(int experiment)} returns the citation defined in the \AEDL\ file for the experiment number {\tt experiment}.
+Note that the returned pointer cannot be modified.
+See also {\tt \$citation} in \AEDL\ on  page~\pageref{citation}.
+\end{function}
 Very basic information is supplied with
@@
 \end{function}
-In addition, the coefficients in the rules can be manipulated by\footnote{Available in GLoBES 3.1.7 and higher.}
+In addition, the coefficients in the rules can be manipulated by\footnote{Available in \GLOBES\ 3.1.7 and higher.}
 \begin{function}
@@
 \end{function}
-Furthermore, the energy window can be manipulated by\footnote{Available in GLoBES 3.1.7 and higher.}
+Furthermore, the energy window can be manipulated by\footnote{Available in \GLOBES\ 3.1.7 and higher.}
 \begin{function}
@@
 
+\section{Defining oscillation parameter names at running times}
+
+One can either define constants for the non-standard physics parameters, as in our example
+\begin{quote}
+{\tt
+  const int GLB\_SIGMA\_E = 6;  
+}
+\end{quote}
+or assign and read out names dynamically. This feature is available in \GLOBES\ 3.1.7 and higher. The following functions are supported:
+\begin{function}
+\GLBNS{SetParamName}
+{\tt int glbSetParamName(const char *name, int i)} assigns {\tt name} to the oscillation parameter {\tt i} (starting at 0 for {\tt GLB\_THETA\_12}).  
+\end{function}
+As usual, the return value is $0$ ({\tt GLB\_SUCCESS}) if successful.
+Note that the string is copied, which means that the original pointer can be modied or destroyed thereafter.
+\begin{function}
+\GLBNS{SetParamNames}
+{\tt int glbSetParamNames(char **names)} assigns names to all of the oscillation parameters at once.
+The array of strings {\tt names} must be of length {\tt glbGetNumOfOscParams()}.
+Individual entries are allowed to be {\tt NULL}.
+\end{function}
+\begin{function}
+\GLBNS{GetParamName}
+{\tt char *glbGetParamName(int i)} returns the name of the oscillation parameter {\tt i}.  
+\end{function}
+\begin{function}
+\GLBNS{FindParamByName}
+{\tt int glbFindParamByName(const char *name)} returns the number of the oscillation parameter named {\tt name}.
+\end{function}
+\begin{function}
+\GLBNS{SetOscParamByName}
+{\tt int glbSetOscParamByName(glb\_params in, double value, const char *name)} sets an oscillation parameter in the structure {\tt in} to {\tt value}.
+Compared to the similar function {\tt glbSetOscParams}, which parameter is to be accessed is determined by the {\tt name}, not the number.
+\end{function}
+\begin{function}
+\GLBNS{GetOscParamByName}
+{\tt double glbGetOscParamByName(glb\_params in, const char *name)} returns an oscillation parameter in the structure {\tt in}.
+Compared to the similar function {\tt glbGetOscParams}, which parameter is to be accessed is determined by the {\tt name}, not the number.
+\end{function}
+\begin{function}
+\GLBNS{ClearParamNames}
+{\tt int glbClearParamNames()} clears all names assigned to oscillation parameters and returns the allocated memory.
+\end{function}
+
 \chapter{Experimental features$^*$}

### 2010-11-20 Tools/degfinder/doc/degfinder.tex
- ``Tools'' section of website now contains most recent degfinder and   sterile neutrino engine + documentation - New functions glbSetProjectionFlagByName, glbGetProjectionFlagByName

\documentclass{article}

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{ctable}
\usepackage{longtable}

% Define bold face typewriter font
\DeclareFontShape{OT1}{cmtt}{bx}{n}{
  <5><6><7><8><9><10><10.95><12><14.4><17.28><20.74><24.88>cmttb10}{}

\title{{\tt degfinder} --- Event rate based degeneracy finding in GLoBES}
\author{Joachim Kopp}
\date{v1.0 (Nov 2010)}

\begin{document}

\maketitle

\begin{center}
\framebox{
\begin{minipage}{11cm}
{\bf IMPORTANT NOTICE:} It is important to keep in mind that degeneracy finding is
not a foolproof process. The performance of {\tt degfinder} depends critically
on the way it is configured by the user. For example, it is the user's
responsibility to determine which parameters need to be scanned over in
the pre-scan, which parameter correlations need to be taken into account,
etc. If a mistake is made, results may be too optimistic, and it is often
hard to notice when this happens. If your plots show
funny features (kinks, spikes, tiny islands of good sensitivity), this
is often an indication that degeneracy finding has gone wrong.
Once again, {\bf \texttt{\textbf{degfinder}} MUST NOT BE USED AS A BLACK BOX!}
\end{minipage}}
\end{center}

\section{Introduction}

In a fit to data from a high-precision neutrinon oscillation experiment, one
often has to deal with the problem of multiple degenerate solutions in the
high-dimensional space of oscillation parameters. For example, the two cases
$\Delta m_{31}^2 > 0$ (normal mass hierarchy) and $\Delta m_{31}^2 < 0$ (inverted
mass hierarchy) often cannot be distinguished
(``mass hierarchy degneracy''). The fitting algorithm in GLoBES is based on 
\emph{local} minimization of $\chi^2$ in the space of oscillation parameters
and systematic nuisance parameters. This local minimization can only find
one degenerate solution at a time. For example, if a $\Delta m_{31}^2 > 0$
is used as a starting value (the {\tt in} parameters passed to {\tt glbChiNP}
etc.), the minimizer will usually converge into a normal hierarchy solution
and ignore the inverted hierarchy solution. This may lead to results that
are too optimistic, and hence it is mandatory to address the problem of
degenerate solutions carefully.

This is the aim of the {\tt degfinder} add-pon. The function {\tt degfinder}
first performs a rough scan (the so-called \emph{pre-scan}) of part of the
parameter space during which many features like systematic uncertainties
and parameter correlations are switched off to improve speed. Which part of
the parameter space is scanned over is determined by the user. The pre-scan
provides rough estimates for the locations of the local $\chi^2$ minima,
and each of these is then used as a starting points for a full-fledged
minimization including all features.

\section{Installation}

To use degfinder in your GLoBES application, simply {\tt \#include} the header file
{\tt degfinder.h} in your source code and add the source file {\tt degfinder.c} to
your {\tt Makefile} so that it is compiled and linked to your executable.

Note that {\tt degfinder} is written in the GNU version of C99. To compile it
with {\tt gcc}, use the option {\tt -std=gnu99}. I haven't tested {\tt
degfinder} with other compilers, so I cannot exclude that some tweaking may be
necessary to compile it with them.

\section{Usage}

The syntax of {\tt degfinder} is

\begin{verbatim}
int degfinder(const glb_params base_values,
  const int n_prescan_params, const int *prescan_params,
  const double *prescan_min, const double *prescan_max,
  const int *prescan_steps, const glb_projection prescan_proj,
  const glb_projection fit_proj, int *n_deg, glb_params *deg_pos,
  double *deg_chi2, const long flags);
\end{verbatim}

\noindent The arguments are \\
\begin{longtable}{p{3cm}p{9cm}}
  \toprule
  {\tt base\_values} & Vector of oscillation parameters.
                       Parameters that are not scanned in the pre-scan (i.e.\ parameters
                       that are not included in {\tt prescan\_params}) are treated in
                       the same way as {\tt in} parameters passed to {\tt glbChiNP}
                       and similar function. This means that if a parameter is declared as
                       {\tt GLB\_FIXED} in {\tt prescan\_proj} or {\tt fit\_proj}, it is
                       assumed to be known with a zero uncertainty in the pre-scan or
                       during the final fit, respectively. The value of a parameter that
                       is declared as {\tt GLB\_FREE} in {\tt prescan\_proj} or {\tt fit\_proj}
                       is used as starting values for local minimizations of $\chi^2$ with
                       respect to that parameter during the pre-scan or during the final
                       minimization, respectively. The {\tt base\_values} of parameters included in
                       {\tt prescan\_params} are ignored. \\\midrule
  {\tt n\_prescan\_params} & The number of parameters that should be scanned over during
                             the pre-scan. \\\midrule
  {\tt prescan\_params}    & A list of parameters to be scanned over during the pre-scan.
                             Parameters are referred to by their numerical index, e.g.\
                             {\tt GLB\_THETA\_12}, {\tt GLB\_THETA\_13}, etc. There should
                             be {\tt n\_prescan\_params} entries in {\tt prescan\_params}. \\\midrule
  {\tt prescan\_min}, {\tt prescan\_max}, {\tt prescan\_steps}
                           & Parameter ranges for the pre-scan. For each entry of
                             {\tt prescan\_params}, a minimum value, a maximum value, and
                             a number of steps is given. The number of points sampled for
                             each parameter is the number of steps plus one. Note that all
                             parameters except $\theta_{13}$ are scanned on a linear scale,
                             while $\theta_{13}$ is scanned on a logarithmic scale in
                             $\sin^2 2\theta_{13}$. For example, if {\tt prescan\_min = -3},
                             {\tt prescan\_max = -1}, $\sin^2 2\theta_{13}$ will be varied
                             between $10^{-3}$ and $10^{-1}$, using {\tt prescan\_steps+1}
                             logarithmically spaced sampling points. \\\midrule
  {\tt prescan\_proj}      & The projection to be used in the pre-scan. Typically, most or
                             all entries will be declared as {\tt GLB\_FIXED} to make
                             the pre-scan as efficient as possible. Sometimes, however,
                             it may be necessary to declare one or several parameters as
                             {\tt GLB\_FREE} to tell {\tt degfinder} to do a local minimization
                             over that parameter even during the pre-scan. \\\midrule
  {\tt fit\_proj}          & The projection to be used in the final fit. Typically, most
                             parameters will be declared as {\tt GLB\_FREE} here. \\\midrule
  {\tt n\_deg}             & Input: Pointer to an integer giving the maximum number of degenerate
                             solutions to accept (the length of the vectors {\tt deg\_pos}
                             and {\tt deg\_chi}). \newline
                             Output: The number of degenerate solutions actually found. \\\midrule
  {\tt deg\_pos}           & An array of pointers to {\tt glb\_params} structures that will be
                             filled with the locations of the degenerate solutions in the space
                             of oscillation parameters. \\\midrule
  {\tt deg\_chi}           & This array will be filled with the $\chi^2$ values of the
                             degenerate solutions. \\\midrule
  {\tt flags}              & Some flags that control the behavior of {\tt degfinder}. \newline
                             \begin{tabular}{lp{6.3cm}}
                               {\tt DEG\_NO\_NH}   & Omit normal hierarchy solutions. \\
                               {\tt DEG\_NO\_IH}   & Omit inverted hierarchy solutions. \\
                               {\tt DEG\_NO\_SYS}  & Switch off systematics even during the final fit
                                                     (systematics are always off during the pre-scan). \\
                               {\tt DEG\_NO\_CORR} & Switch off parameter correlations. Equivalent to
                                                     declaring all parameters as {\tt GLB\_FIXED} in both
                                                     {\tt prescan\_proj} and {\tt fit\_proj}. \\
                               {\tt DEG\_NO\_DEG}  & Switch off degeneracies. Equivalent to
                                                     setting {\tt n\_prescan\_params = 0}. Note that in
                                                     this case, {\tt degfinder} should behave just like
                                                     {\tt glbChiNP}. \\
                             \end{tabular} \\
  \bottomrule
\end{longtable}

\section{{\tt degfinder} and non-standard interactions}

{\tt degfinder} works together with my non-standard interaction code. If you want to
use this feature, include {\tt nsi\_probability.h} in the header of {\tt degfinder.c}. If non-standard
parameters are included in the pre-scan, {\tt degfinder} will step through the non-standard phases
on a linear scale, while the absolute values of the non-standard parameters are scanned
on a log scale. Consequently, for the absolute values of the non-standard paremeters,
{\tt prescan\_min} and {\tt prescan\_max} should be logarithmic. For example, a range
$[-3, -1]$ means that this parameter will be varied between $10^{-3}$ and $10^{-1}$,
with logarithmically spaced sampling points. Note that degeneracy finding in high
dimensional NSI parameter spaces can be extremely inefficient.

\end{document}

### 2010-11-20 Tools/snu/doc/snu.tex
- ``Tools'' section of website now contains most recent degfinder and   sterile neutrino engine + documentation - New functions glbSetProjectionFlagByName, glbGetProjectionFlagByName

\documentclass{article}

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{ctable}
\usepackage{longtable}

% Define bold face typewriter font
\DeclareFontShape{OT1}{cmtt}{bx}{n}{
  <5><6><7><8><9><10><10.95><12><14.4><17.28><20.74><24.88>cmttb10}{}

\title{Sterile neutrinos and non-standard neutrino interactions in GLoBES}
\author{Joachim Kopp}
\date{v1.0 (Nov 2010)}

\begin{document}

\maketitle

Sterile neutrinos and non-standard neutrino interactions (NSI) are two of
the most generic types of new physics possible in the neutrino sector.
For an introduction to the formalism of NSI, see e.g.\ \cite{Kopp:2007ne}.

There are several implementations of NSI in GLoBES (e.g.\ the MonteCubes plugin
\cite{Blennow:2009pk}, for instance, can handle NSI), but to the best of
my knowledge, there is no publicly available implementation of sterile neutrinos.
The code discussed here can handle both: It works with up to 9 neutrino flavors,
and can handle NSI in the production and detection processes as well as
non-standard matter effects (NSI in propagation). The non-standard operators
can affect the active neutrino species as well as the sterile ones.

To use the code, include {\tt snu.c} in your project by modifying your
{\tt Makefile} accordingly, and {\tt \#include} the header file {\tt snu.h}
in your source code.

The first step is to initialize the sterile neutrino/NSI engine. If you only
need NSI, but no sterile neutrinos, you may use the simple command
\begin{verbatim}
  snu_init_probability_engine();
\end{verbatim}
On the other hand, if you need sterile neutrinos, you need to specify how
many neutrino species you would like to include, and how the mixing matrix
is defined. The syntax for this is
\begin{verbatim}
  int snu_init_probability_engine(int n_flavors,
    int rotation_order\left[2], int phase_order[]);
\end{verbatim}
\noindent The arguments are
\begin{longtable}{p{2.5cm}p{8.5cm}}
  \toprule
  {\tt n\_flavors}      & The number of neutrino flavors. Must be $\geq 3$.
                          Note that for ${\tt n\_flavors} = 3$, the optimized
                          algorithms from \cite{Kopp:2006wp} are used to
                          diagonalize the Hamiltonian, while for more than four
                          flavors, the GNU Scientific Library is used. \\\midrule
  {\tt rotation\_order}, {\tt phase\_order} &
    Specifies how the neutrino mixing matrix $U$ is composed of individual rotation
    matrices, and which of the rotation matrices carry the complex phases. A
    definition of the form
    \begin{flushleft}
      {\tt rotation\_order} = \{\{$i_1$, $j_1$\}, \{$i_2$, $j_2$\}, \dots \} \\
      {\tt phase\_order} = $\{k_1, k_2, \dots\}$
    \end{flushleft}
    means $U = R(\theta_{i_1 j_1},
    \delta_{k_1}) R(\theta_{i_2 j_2}, \delta_{k_2}) \cdots$, where
    \begin{displaymath}
      R(\theta_{ij},\delta_k) =
        \begin{pmatrix}
          0      & \cdots &   0             & \cdots &   0                    & \cdots & 0 \\
          \vdots &        & \vdots          &        & \vdots                 &        & \vdots \\
          0      & \cdots & \cos\theta_{ij} & \cdots & e^{i\delta_k} \sin\theta_{ij} & \cdots & 0 \\
          \vdots &        & \vdots          &        & \vdots                 &        & \vdots \\
          0      & \cdots & -e^{-i\delta_k} \sin\theta_{ij} & \cdots & \cos\theta_{ij} & \cdots & 0 \\
          \vdots &        & \vdots          &        & \vdots                 &        & \vdots \\
          0      & \cdots &   0             & \cdots &   0                    & \cdots & 0 \\
        \end{pmatrix}
    \end{displaymath}
    is a rotation matrix in the $ij$-sector with complex phase $\delta_k$. The
    indices $i$ and $j$ run from $1$ to {\tt n\_flavors}. Both {\tt
    rotation\_order} and {\tt phase\_order} should have ${\tt n\_flavors}
    ({\tt n\_flavors}-1)/2$ entries. Since there are only
    $({\tt n\_flavors}-1)({\tt n\_flavors}-2)/2$ phases, some of the entries
    of {\tt phase\_order} should be $-1$, indicating that the respective
    rotation matrices will be treated as real. In the above syntax, the
    standard three-flavor mixing matrix $U_{3\times3} = R(\theta_{23},0)
    R(\theta_{13},\delta) R(\theta_{12},0)$ would be given by
    \begin{flushleft}
      {\tt rotation\_order} = \{\{2, 3\}, \{1, 3\}, \{1, 2\} \} \\
      {\tt phase\_order} = \{-1, 0, -1\}\,.
    \end{flushleft} \\
  \bottomrule
\end{longtable}
After initializing the sterile neutrino/NSI engine, you have to make it known
to GLoBES by calling
\begin{verbatim}
  glbRegisterProbabilityEngine(6*SQR(n_flavors) - n_flavors,
                               &snu_probability_matrix,
                               &snu_set_oscillation_parameters,
                               &snu_get_oscillation_parameters,
                               NULL);
\end{verbatim}
Note the expression for the number of oscillation parameters in the first line.
As with any new probability engine, the call to {\tt
glbRegisterProbabilityEngine} has to occur before any calls to {\tt
glbAllocParams} or {\tt glbAllocProjection} to make sure that all parameter and
projection vectors have the correct length.

It is recommended to use {\tt glbSetParamName} to assign human-readable
names to the oscillation parameters by which they can be referred to in
subsequent calls to {\tt glbSetOscParamByName} and {\tt
glbSetProjectionFlagByName}. A list of suitable names is provided in the global
array {\tt char snu\_param\_strings[][64]}. The syntax is {\tt TH12},
{\tt TH13}, {\tt TH14}, etc. for the mixing angles, {\tt DELTA\_0},
{\tt DELTA\_1}, \dots for the CP phases, {\tt ABS\_EPS\_$c$\_$\alpha\beta$}
for the absolute values of the non-standard parameters, and
{\tt ARG\_EPS\_$c$\_$\alpha\beta$} for their phases. Here, $c = {\tt S},
{\tt M}, {\tt D}$ to identify NSI in the source, in propagation (non-standard
matter effects), and in detection. $\alpha$ and $\beta$ are neutrino flavors,
where the standard flavors are denoted by {\tt E}, {\tt MU}, {\tt TAU},
and the sterile flavors are {\tt S1}, {\tt S2}, {\tt S3}, \dots. For
example, {\tt ABS\_EPS\_M\_ETAU} refers to $|\varepsilon^m_{e\tau}|$
in the notation of \cite{Kopp:2007ne}.

Having initialized and registered the non-standard probability engine, you can
use all GLoBES functions in the same way as you would for standard
oscillations.  Note, however that at the moment the initial and final flavors
in the definition of an oscillation channel can only involve the three active
flavors. It is for instance not possible to compute sterile neutrino
appearance.

After using the sterile neutrino/NSI engine, you may want to release the
(small) amount of memory allocated by it by calling
\begin{verbatim}
  snu_free_probability_engine();
\end{verbatim}

\bibliographystyle{plain}
\bibliography{snu}

\end{document}

### 2010-11-20 Distribution/globes/globes.h

@@ -216,6 +216,8 @@ glb_projection glbSetDensityProjectionFlag(glb_projection in, int flag, int whic
 int glbGetProjectionFlag(const glb_projection in, int which);
 int glbGetDensityProjectionFlag(const glb_projection in, int which);
 void glbPrintProjection(FILE *stream, const glb_projection in);
+int glbSetProjectionFlagByName(glb_projection in, int flag, const char *name);
+int glbGetProjectionFlagByName(const glb_projection in, const char *name);
 
 int glbSetProjection(const glb_projection in);
 int glbGetProjection(glb_projection in);

### 2010-11-20 Distribution/globes/globes_perl.c

@@ -4022,6 +4022,94 @@ XS(_wrap_glbPrintProjection) {
 }
 
 
+XS(_wrap_glbSetProjectionFlagByName) {
+  {
+    glb_projection arg1 = (glb_projection) 0 ;
+    int arg2 ;
+    char *arg3 = (char *) 0 ;
+    int result;
+    void *argp1 = 0 ;
+    int res1 = 0 ;
+    int val2 ;
+    int ecode2 = 0 ;
+    int res3 ;
+    char *buf3 = 0 ;
+    int alloc3 = 0 ;
+    int argvi = 0;
+    dXSARGS;
+    
+    if ((items < 3) || (items > 3)) {
+      SWIG_croak("Usage: glbSetProjectionFlagByName(in,flag,name);");
+    }
+    res1 = SWIG_ConvertPtr(ST(0), &argp1,SWIGTYPE_p_glb_projection_type, 0 |  0 );
+    if (!SWIG_IsOK(res1)) {
+      SWIG_exception_fail(SWIG_ArgError(res1), "in method '" "glbSetProjectionFlagByName" "', argument " "1"" of type '" "glb_projection""'"); 
+    }
+    arg1 = (glb_projection)(argp1);
+    ecode2 = SWIG_AsVal_int SWIG_PERL_CALL_ARGS_2(ST(1), &val2);
+    if (!SWIG_IsOK(ecode2)) {
+      SWIG_exception_fail(SWIG_ArgError(ecode2), "in method '" "glbSetProjectionFlagByName" "', argument " "2"" of type '" "int""'");
+    } 
+    arg2 = (int)(val2);
+    res3 = SWIG_AsCharPtrAndSize(ST(2), &buf3, NULL, &alloc3);
+    if (!SWIG_IsOK(res3)) {
+      SWIG_exception_fail(SWIG_ArgError(res3), "in method '" "glbSetProjectionFlagByName" "', argument " "3"" of type '" "char const *""'");
+    }
+    arg3 = buf3;
+    result = (int)glbSetProjectionFlagByName(arg1,arg2,(char const *)arg3);
+    ST(argvi) = SWIG_From_int  SWIG_PERL_CALL_ARGS_1((int)(result)); argvi++ ;
+    
+    
+    if (alloc3 == SWIG_NEWOBJ) free((char*)buf3);
+    XSRETURN(argvi);
+  fail:
+    
+    
+    if (alloc3 == SWIG_NEWOBJ) free((char*)buf3);
[…]

### 2010-11-20 Distribution/source/glb_wrapper.c

@@ -706,6 +706,65 @@ void glbPrintProjection(FILE *stream, const glb_projection in)
 }
 
 
+/***************************************************************************
+ * Function glbSetProjectionFlagByName                                     *
+ ***************************************************************************
+ * Same as glbSetProjectionFlag, but identifies the parameter by name.     *
+ * Return value is GLB_SUCCESS or one of the (negative) GLBERR_XXX error   *
+ * codes                                                                   *
+ ***************************************************************************/
+int glbSetProjectionFlagByName(glb_projection in, int flag, const char *name)
+{
+  int index;
+
+  if((flag!=GLB_FREE)&&(flag!=GLB_FIXED))
+  {
+    glb_error("glbSetProjectionFlagByName: Projection flag must be GLB_FREE or GLB_FIXED");
+    return NULL;
+  }
+
+  if (!name)
+  {
+    glb_error("glbSetProjectionFlagByName: Invalid parameter name (NULL)");
+    return GLBERR_INVALID_ARGS;
+  }
+
+  if ((index=glbFindParamByName(name)) < 0)
+    return index;
+  else
+  {
+    if (!glbSetProjectionFlag(in, flag, index))
+      return GLBERR_GENERIC;
+    else
+      return GLB_SUCCESS;
+  }
+}
+
+
+/***************************************************************************
+ * Function glbGetProjectionFlagByName                                     *
+ ***************************************************************************
+ * Same as glbGetProjectionFlag, but identifies the parameter by name.     *
+ * In case of an error, the return value is negative                       *
+ ***************************************************************************/
+int glbGetProjectionFlagByName(const glb_projection in, const char *name)
+{
+  int index;
+  
+  if (!name)
+  {
+    glb_error("glbSetProjectionFlagByName: Invalid parameter name (NULL)");
+    return GLBERR_INVALID_ARGS;
+  }
+
+  if ((index=glbFindParamByName(name)) < 0)
+    return index;
+  else
+    return glbGetProjectionFlag(in, index);
+}
+
+
+
 /* Wrappers for the old functions dealing with setting the oscillation
  * parameters
  */

### 2010-11-20 Distribution/source/globes.c

@@ -154,7 +154,7 @@ static struct argp_option options[] ={
   {"parameters",'p',"PARAMETERS",0,
    "Set of oscillation parameters\n   for which the rates are computed\n"
    "   Format 'th12,th13,th23,delta,dm21,dm31'"},
-  {"before",'b',0,0,"Channel rates before smearing (implies -c)"},
+  {"before",'b',0,0,"Channel rates before smearing (ignored if -c is not used)"},
   {"after",'a',0,0,"Rates after smearing"},
   {"efficiencies",'f',0,0,"Rates without efficiencies"},
   {"backgrounds",'g',0,0,"Rates without backgrounds"},

### 2010-11-20 Tools/degfinder/degfinder.c

@@ -30,7 +30,20 @@
 
 #include <globes/globes.h>
 #include "degfinder.h"
-#include "nsi_probability.h"
+/*#include "nsi_probability.h"*/ /* Uncomment if you want to use degfinder with JK's NSI engine */
+
+#ifndef FIRST_NSI_INDEX
+#define FIRST_NSI_INDEX -100
+#define LAST_NSI_INDEX  -101
+#endif
+
+#define SQR(x)      ((x)*(x))                              // x^2 
+#define POW10(x)    (exp(M_LN10*(x)))                      // 10^x
+#define MIN(X,Y)    ( ((X) < (Y)) ? (X) : (Y) )
+#define MAX(X,Y)    ( ((X) > (Y)) ? (X) : (Y) )
+#define SIGN(a,b)   ( ((b) > 0.0) ? (fabs(a)) : (-fabs(a)) )
+#define SGN(a)      ( ((a) >= 0.0) ? (1) : (-1) )
+#define ROUND(x)    ( (int)((x) + 0.5) )
 
 /* Global variables */
 extern const char *param_strings[];

### 2010-11-20 Tools/snu/snu.c

@@ -1,0 +1,1336 @@
+// ----------------------------------------------------------------------------
+// Probability engine for simulating sterile neutrinos + non-standard
+// interactions
+// ----------------------------------------------------------------------------
+// Author: Joachim Kopp (jkopp@fnal.gov)
+// ----------------------------------------------------------------------------
+// GLoBES -- General LOng Baseline Experiment Simulator
+// (C) 2002 - 2010,  The GLoBES Team
+//
+// GLoBES as well as this add-on are mainly intended for academic purposes.
+// Proper credit must be given if you use GLoBES or parts of it. Please
+// read the section 'Credit' in the README file.
+//
+// This program is free software; you can redistribute it and/or modify
+// it under the terms of the GNU General Public License as published by
+// the Free Software Foundation; either version 2 of the License, or
+// (at your option) any later version.
+//
+// This program is distributed in the hope that it will be useful,
+// but WITHOUT ANY WARRANTY; without even the implied warranty of
+// MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+// GNU General Public License for more details.
+//
+// You should have received a copy of the GNU General Public License
+// along with this program; if not, write to the Free Software
+// Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA  02111-1307  USA
+// ----------------------------------------------------------------------------
+// Note: This file is written in C99. To compile in gcc use -std=c99 or
+// -std=gnu99
+// ----------------------------------------------------------------------------
+// Citation information:
+//
+//      @Article{Kopp:2006wp,
+//        author    = "Kopp, Joachim",
+//        title     = "{Efficient numerical diagonalization of hermitian
+//                     $3 \times 3$ matrices}",
+//        journal   = "Int. J. Mod. Phys.",
+//        volume    = "C19",
+//        year      = "2008",
+//        pages     = "523-548",
+//        eprint    = "physics/0610206",
[…]

### 2020-05-01 Documentation/Manual.tex
Updated my affiliation and the acknowledgments (included my DOE grant)
Finally (!) updated the manual.

@@
 \newcommand{\Figs}{Figs.}
-\newcommand{\Ref}{Ref.}
+\renewcommand{\Ref}{Ref.}
 \newcommand{\Refs}{Refs.}
@@
 
-%\vspace*{1.5cm}
-%{\large \bf !!!DRAFT IN PROGRESS!!! }
+\vspace*{1.5cm}
+{\large \bf !!!DRAFT IN PROGRESS!!! }
 \end{center}
@@
 \begin{center}
-{\large Patrick Huber$^a$, Joachim Kopp$^b$, Manfred Lindner$^b$, \\ Mark Rolinec$^c$, Walter Winter$^d$}
+{\large Patrick Huber$^a$, Joachim Kopp$^b$, Manfred Lindner$^c$, \\ Walter Winter$^d$}
 \end{center}
@@
 \begin{center}
-Version from \today\ for \GLOBES\ 3.1
+Version from \today\ for \GLOBES\ 3.2.18
 \end{center}
@@
 {\em\small $^{\mathrm{a}}$%
-       University of Wisconsin,  Physics Department,
-       1150 University Av., Madison, WI 53706, USA}
+       Virginia Tech, Center for Neutrino Physics, Blacksburg, VA 24061, USA}
 
 {\em\small $^{\mathrm{b}}$%
+  Theoretical Physics Department,
+  CERN and Johannes Gutenberg University Mainz (Germany),
+  1211 Geneve 23, Switzerland
+}
+
+{\em\small $^{\mathrm{c}}$%
        Max--Planck--Institut f\"ur Kernphysik, 
@@
 
-{\em\small $^{\mathrm{c}}$%
-       Technische Universit\"at M\"unchen,
-       Institut f\"ur Theoretische Physik, Physik--Department,\\
-       James--Franck--Strasse, D--85748 Garching, Germany}
-
 {\em\small $^{\mathrm{d}}$%
-       Universit\"at W\"urzburg, 
-       Lehrstuhl f\"ur theoretische Physik II, \\
-       Institut f\"ur theoretische Physik und Astrophysik, 
-       Am Hubland,
-       D-97074 W\"urzburg, Germany}
+Deutsches Elektronen-Synchrotron (DESY)
+Platanenallee 6
+15738 Zeuthen, Germany}
 
@@
 \begin{quote}
-    Copyright \copyright  2004--2010  The GLoBES Team.
+    Copyright \copyright  2004--2020  The GLoBES Team.
     Permission is granted to copy, distribute and/or modify this document
@@
 evolve during time, which means that outdated features may not be documented anymore.
+
+\section*{Version 3.2.18}
+\index{norm}{Version 3.2.18}
+
+\begin{itemize}
+
+  \item Systematics correlated across experiments/detector; see
+    \Sec~\ref{sec:newsys}. This feature has been extensively tested
+    and used, amongst others, for the DUNE CDR~\cite{Acciarri:2015uup}.
+\item Fitting of actual data; see \Sec~\ref{sec:fitting}.
+\end{itemize}
 
@@
 \item Universit\"at W\"urzburg [WW]
+\item US Department of Energy under award number \protect{DE-SC0003915} [PH]
 \end{itemize}

### 2020-05-01 Documentation/part1.tex
Finally (!) updated the manual.

@@
 
+A related function which allows to name an oscillation engine without invoking it
+\begin{function}
+  \GLBNS{DefineOscEngine}
+{\tt int glbDefineOscillationEngine(int n\_parameters,
+                 glb\_probability\_matrix\_function prob\_func, \\
+                 glb\_set\_oscillation\_parameters\_function set\_params\_func, \\
+                 glb\_get\_oscillation\_parameters\_function get\_params\_func, \\
+                 const char[] name,\\
+                 void* user\_data)} defines a probability
+engine for the simulation of non-standard physics with {\tt n\_parameters} oscillation parameters ({\tt n\_parameters} $\ge6$).
+The three functions {\tt prob\_func}, {\tt set\_params\_func}, and {\tt get\_params\_func}
+will be used with the function types defined above.
+In order to circumvent global variables, an arbitrary pointer {\tt user\_data} can be defined, that will be passed to the probability engine in each function call.
+{\tt name} can be use to refer to the so defined oscillation engine within {\sf AEDL} using the {\tt \$oscillation\_engine} directive.
+\end{function}
+\index{aedl}{oscillation engine@{\tt "\$oscillation\_engine}}
+
 Let us now illustrate the implementation of non-standard physics with a simple example, which can

### 2020-05-01 Documentation/part2.tex
Further functions added to manual
Finally (!) updated the manual.

@@
 in order to avoid parsing megabytes of chunk
- and producing thousands of error messages.  In addition, the minimum
+ and producing thousands of error messages.  The minimum
 \GLOBES\ version that the \AEDL\ file is supposed to run with has to be
@@
 \index{aedl}{version@{\tt \$version}} 
+Furthermore, references linked to the \AEDL\ file can be defined with {\tt \$citation="..."},
+which can later be retrieved with {\tt glbGetCitationForExperiment}, see page~\pageref{glbGetCitationForExperiment}.\footnote{Available in \GLOBES\ 3.1.7 and higher.}
+\index{aedl}{citation@{\tt \$citation}} \label{citation}
 %
@@
 \end{quote}
-This command resets the counters for channels, rules, fluxes, cross section 
-and energy resolution environments. All variables have their scope limited 
-by either {\tt \%!GLoBES, \#NEXT\#} or {\tt EOF}.  This allows 
-a consistent treatment of various experiments in one file.
+This command resets the counters for channels, rules, fluxes, cross
+section and energy resolution environments. All, except named nuisance
+parameters (see. Sec.~\ref{sec:newsys}), variables have their scope limited by either {\tt
+  \%!GLoBES, \#NEXT\#} or {\tt EOF}.  This allows a consistent
+treatment of various experiments in one file.
+
+\index{aedl}{DETECTOR@{\tt \#DETECTOR\#}}
+If you want to have several detectors in the same beam, separate the different
+ detectors by 
+\begin{quote}
+{\tt    \#DETECTOR\#}
+\end{quote}
+This command resets the counters and names \emph{only} for rules. All
+other variables have global scope.  This feature is useful to define
+correlated systematics for near and far detector type setups as
+described in Sec.~\ref{sec:newsys}.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Systematics correlated accross experiments and rules*}
+\label{sec:newsys}
+
+\index{norm}{Systematics!correlated|(} 
+\index{aedl}{sys@{\tt sys}|(} In many cases the simple rule-based
+style of systematics is sufficient, but there are cases, notably DUNE
+and almost all modern reactor experiments, where a more fine-grained
+systematics implementation is necessary.  The way to accomplish is to introduce named nuisance parameters:
+\begin{quote}
+  {\tt sys(\#norm)<\\
+    \tb @systype=0\\
+\tb  @error = 0.02\\
+>}
+\end{quote}
+
+\index{aedl}{sys@{\tt sys}!error@{\tt "@error}}
+
+\index{aedl}{sys@{\tt sys}!systype@{\tt "@systype}}
+
+{\tt @systype} defaults to $0$, a common multiplicative nuisance
+parameter.  {\tt @systype} can also be set to $1$ where this becomes a
+energy calibration error as desribed in Sec.~\ref{sec:energy}.
+
+In this case we introduced a nuisance parameter {\tt \#norm} with an
+error of $0.02$ and a central value of $1$. These named nuisance
+parameters now can be used in the definition of a rule:
+
+\begin{quote}
+  {\tt rule(\#NU\_E\_Disappearance)<\\
+\tb    @signal     = 0.4@\#nu\_e\_disappearance\_CC\_U235 \\
+\tb    @sys\_on\_multiex\_errors\_sig =\{\#norm,\#flux235\} \\
+\tb    @background = 0.1@\#nu\_e\_bg\\
+\tb     @sys\_on\_multiex\_errors\_bg  = {\#bgnorm}\\
+\tb        @sys\_on\_function  = "chiMultiExp"\\
+\tb        @sys\_off\_function = "chiNoSysSpectrum"\\
+>}
+\end{quote}
+
+The key here is the line {\tt @sys\_on\_multiex\_errors\_sig
+  =\{\#norm,\#flux235\} }, where the tag {\tt
+  @sys\_on\_multiex\_errors\_sig } is invoked to signal the use of
+this feature and each {\tt \{ ...\}} contains all named nuisance
+parameters applicable to the corresponding part of {\tt @signal}
+definition; the same syntax applies to {\tt
+  @sys\_on\_multiex\_errors\_bg } for the {\tt @background} part of
+the rule.  {\tt @sys\_on\_function = "chiMultiExp"} ensures that the
+appropriate $\chi^2$-function is used to handle named nuisance
+parameters.%
+
+\index{aedl}{rule@{\tt rule}!sys on multiex errors  sig@{\tt "@sys\_on\_multiex\_errors\_sig}}
+
+\index{aedl}{rule@{\tt rule}!sys on multiex errors bg@{\tt "@sys\_on\_multiex\_errors\_bg}}
+
+Now this very same technique can be used in different rules of the
+same experiment but also in rules defined in \emph{different}
+experiments. All instance of the named parameter, in this example {\tt
+  \#norm}, will be fully correlated. In the context of
+near/far-detector experiments also the {\tt \#DETECTOR\#} directive is
+interesting since it allows to reuse flux, channel etc. definitions
+seamlessly. It is important that all named parameters have been
+defined using the {\tt sys<>} environment before they are used. It is
+good practive to put all {\tt sys<>} definitions in one file and to
+include this file into all subsequent {\sf AEDL} files.
+
+Sometimes, systematics are energy dependent and this can be implemented within the same framework as
+\begin{quote}
+  {\tt sys(\#norm)<\\
+    \tb @energy\_list =\{ 0.002,0.005,0.006,0.009\}\\
+    \tb  @error\_list = \{0.02,0.01,0.005,0\}\\
+>}
+\end{quote}
+\index{aedl}{sys@{\tt sys}!error list@{\tt "@error\_list}}
+\index{aedl}{sys@{\tt sys}!energy list@{\tt "@energy\_list}}
+
+This effectively introduces 1 nuisance parameter in the fit which
+varies the rate by 0.02 at 0.002\,GeV and by 0.01 at 0.005\,GeV
+asf. fully correlated among all energies. For bin center energies
+between the support points in {\tt @energy\_list} linear interpolation is
+used. Both lists need to have the same number of elements.
+
+\index{aedl}{sys@{\tt sys}|)}
+\index{norm}{Systematics!correlated|)}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Fitting actual experimental data*}
+\label{sec:fitting}
+
+\index{norm}{Fitting of data|(}
+The framework presented so far is general enough to achieve an
+accurate decsription of real experiments and thus, the user may want
+to fit actual data with {\sf GLoBES}. This can be conveniently achieved by using
+
+\begin{quote}
+  {\tt rule(\#foo)<\\
+    \tb all the usual definitions of a rule\\
+    \\
+    \tb @data\_flag=1\\
+    \tb @data=\{1.23,2.34,3.45,...,42\}\\
+>}
+\end{quote}
+\index{aedl}{rule@{\tt rule}!data flag@{\tt "@data\_flag}}
+\index{aedl}{rule@{\tt rule}!data@{\tt "@data}}
+
+where {\tt @data} is a list of the actual data which needs to have
+{\tt \$bins} elements and the {\tt @data\_flag} can be use to turn
+fitting of this data on (1) or off (0).
+
+In some cases a non-standard oscillation probability calculaton may be
+required, to include for instance the size of the neutrino production
+region. This requires to use the mechanism to define an oscillation
+engine and to name it, this is done with {\tt DefineOscEngine} and to
+then call the corresponding name from withing {\sf AEDL}, which is
+accomplished with the {\tt \$oscillation\_engine} directive. Note,
+that the scope of {\tt \$oscillation\_engine} is not limited by {\tt
+  \#DETECTOR\#}.
+
+\GLBNS{DefineOscEngine}
+\index{aedl}{oscillation engine@{\tt "\$oscillation\_engine}}
+
+\index{norm}{Fitting of data|)}
+
 % THIS PART HAS GAINED IMPORTANCE AND THEREFORE BEEN EXPLAINED ALREADY IN THE LAST CHAPTER!

### 2023-09-27 Distribution/examples/Dune_EFT_study2/Fluxes_smeft/AllFluxesDUNE/interpolation.Rmd
Dune analysis program

---
title: "R Notebook"
output: html_notebook
---

This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*.

```{r}
# Load required libraries

library(ggplot2)

v <- c(0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 10, 10, 10, 10, 10, 10, 10, 10, 10)

energy_values <- cumsum(v)
# List all .txt files in the current directory
file_list <- list.files(pattern = "*.txt")

# Process each file and store the results in a list
result_list <- lapply(file_list, function(file) {

  # Read the data file
  data <- read.table(file, header = FALSE)

  # Interpolate and evaluate each function
  interpolated_data <- data.frame(V1 = energy_values)
  for (i in 2:ncol(data)) {
    f <- approxfun(data$V1, data[[i]], method = "linear")
    interpolated_data[[i]] <- f(energy_values)
  }

  # Replace NaN values with 0
  interpolated_data[is.na(interpolated_data)] <- 0

  # Save the interpolated data to a new file
  write.table(interpolated_data, paste0(file),
              row.names = FALSE, col.names = FALSE, quote = FALSE)

  return(interpolated_data)
})

# Combine all interpolated data for plotting
combined_data <- do.call(rbind, result_list)

# Plot the combined interpolated functions
ggplot(combined_data, aes(x = V1)) +
  geom_line(aes(y = V2), color = "red") +
  geom_line(aes(y = V3), color = "blue") +
  geom_line(aes(y = V4), color = "green") +
  xlab("Energy") +
  ylab("Function Value")

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

### 2023-09-27 Distribution/examples/Dune_EFT_study2/Fluxes_smeft/ProductionCoefficients_DUNE_separated/extraction.Rmd
Dune analysis program

--- 2023-09-27 Distribution/examples/Dune_EFT_study2/Fluxes_smeft/AllFluxesDUNE/interpolation.Rmd
@@
 ```{r}
-# Load required libraries
 
-library(ggplot2)
+vector1 = seq(2,151,6)
+vector2 = seq(3,151,6)
+vector3 = seq(4,151,6)
+vector4 = seq(5,151,6)
+vector5 = seq(6,151,6)
+vector6 = seq(7,151,6)
 
-v <- c(0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 10, 10, 10, 10, 10, 10, 10, 10, 10)
+# Set the path to the directory where the input files are located
+input_dir <- "."
 
-energy_values <- cumsum(v)
-# List all .txt files in the current directory
-file_list <- list.files(pattern = "*.txt")
+# Get a list of all the txt files in the directory
+input_files <- list.files(input_dir, pattern = "\\.txt$", full.names = TRUE)
 
-# Process each file and store the results in a list
-result_list <- lapply(file_list, function(file) {
+# Loop through the input files and extract the desired columns
+for (file in input_files) {
+  # Read the input file
+  data <- read.table(file, header = FALSE, stringsAsFactors = FALSE)
 
-  # Read the data file
-  data <- read.table(file, header = FALSE)
+  # Extract the desired columns and save them to separate files
+  output_file1 <- paste0("plus_", basename(file))
+  output_file2 <- paste0("minus_", basename(file))
+  vec = c(vector1,vector2,vector3)
+  vec2 =  c(vector4,vector5,vector6)
+  columns1 <- c(vector1,vector2,vector3)[order(vec, decreasing = FALSE)]
+  columns2 <- c(vector4,vector5,vector6)[order(vec2, decreasing = FALSE)]
 
-  # Interpolate and evaluate each function
-  interpolated_data <- data.frame(V1 = energy_values)
-  for (i in 2:ncol(data)) {
-    f <- approxfun(data$V1, data[[i]], method = "linear")
-    interpolated_data[[i]] <- f(energy_values)
+  data1 <- data[, c(1, columns1)]
+  data2 <- data[, c(1, columns2)]
+
+  write.table(data1, file = output_file1, sep = "\t", quote = FALSE, row.names = FALSE,col.names = FALSE)
+  write.table(data2, file = output_file2, sep = "\t", quote = FALSE, row.names = FALSE,col.names = FALSE)
 }
 
-  # Replace NaN values with 0
-  interpolated_data[is.na(interpolated_data)] <- 0
+```
 
-  # Save the interpolated data to a new file
-  write.table(interpolated_data, paste0(file),
-              row.names = FALSE, col.names = FALSE, quote = FALSE)
-
-  return(interpolated_data)
-})
-
-# Combine all interpolated data for plotting
-combined_data <- do.call(rbind, result_list)
-
-# Plot the combined interpolated functions
-ggplot(combined_data, aes(x = V1)) +
-  geom_line(aes(y = V2), color = "red") +
-  geom_line(aes(y = V3), color = "blue") +
-  geom_line(aes(y = V4), color = "green") +
-  xlab("Energy") +
-  ylab("Function Value")
+```{r}
+data <- read.table("ProdALLDpDecay.txt", header = FALSE)
 

### 2023-09-27 Distribution/examples/Dune_EFT_study2/xsec/detection_coefficients/interpolation.Rmd
Dune analysis program

--- 2023-09-27 Distribution/examples/Dune_EFT_study2/Fluxes_smeft/AllFluxesDUNE/interpolation.Rmd
@@
 energy_values <- cumsum(v)
+
 # List all .txt files in the current directory
@@
   for (i in 2:ncol(data)) {
-    f <- approxfun(data$V1, data[[i]], method = "linear")
+    f <- approxfun(10^(data$V1), data[[i]], method = "linear")
     interpolated_data[[i]] <- f(energy_values)
@@
   # Save the interpolated data to a new file
-  write.table(interpolated_data, paste0(file),
+  write.table(interpolated_data, paste0("modified_",file,sep=""),
               row.names = FALSE, col.names = FALSE, quote = FALSE)

### 2023-09-27 Distribution/examples/Dune_EFT_study2/old/WEFT_Analysis.c

@@ -1,0 +1,164 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include <math.h>
+#include <string.h>
+
+
+
+
+#include <globes/globes.h>   /* GLoBES library */
+#include <gsl/gsl_complex.h>
+#include <gsl/gsl_complex_math.h>
+#include <gsl/gsl_matrix.h>
+#include <gsl/gsl_blas.h>
+#include <gsl/gsl_eigen.h>
+#include "smeft.h"
+
+#define FLAVS 4
+#define ROTS 6
+
+#define EXP_FAR  0
+#define EXP_NEAR 1
+int number_param = 600;
+
+
+int smeft_remove(glb_params p){
+    int i;
+    for(i = 6; i < number_param; i++){
+       
+            glbSetOscParams(p,0.0,i);
+            
+        
+    }
+    return 0;
+}
+
+int get_index(char smeft_param_strings[600][64], int size, char *str) {
+    for(int i = 0; i < size; i++) {
+        if(strcmp(smeft_param_strings[i], str) == 0)
+            return i;
+    }
+    return -1;
+}
+
+
+int main(int argc, char *argv[]) {
+
+
+   glbInit(argv[0]);	
+
+  /* Intitialize output */
+
+  
+  /* Define my standard oscillation parameters */
+    double theta12 = 0.5903;
+    double theta13 = 0.1503;
+    double theta23 = 0.8674;
+    double deltacp = 217*M_PI/180;
+    double sdm = 7.39e-5;
+    double ldm = 2.525e-3;
+    
+  
+	
+   /* Initialize one experiment */
+   
+    smeft_init_probability_engine_3();
+    glbRegisterProbabilityEngine(number_param, &smeft_probability_matrix, &smeft_set_oscillation_parameters, &smeft_get_oscillation_parameters, NULL);
+   
+    glbInitExperiment("DUNE_GLoBES_FD.glb",&glb_experiment_list[0],&glb_num_of_exps);
+    glbInitExperiment("DUNE_GLoBES_ND.glb",&glb_experiment_list[0],&glb_num_of_exps);
+    
+    /* Initialize a number of parameter vector(s) */
+    glb_params true_values = glbAllocParams();
+    glb_params test_values = glbAllocParams();
+    glb_params central_values = glbAllocParams();
+    glb_params input_errors = glbAllocParams();
+    
[…]

### 2023-09-27 Distribution/examples/Dune_EFT_study2/results/plot_BOTH_us.py

@@ -1,0 +1,78 @@
+import matplotlib.pyplot as plt
+from matplotlib.colors import LinearSegmentedColormap
+from scipy.interpolate import interp1d
+import numpy as np
+
+# Enabling LaTeX rendering
+plt.rc('text', usetex=True)
+plt.rc('font', family='serif')
+
+def convert_name_to_latex(name):
+    last_part = name.split('_')[-1].lower()
+    latex_mapping = {
+        'ee': '\mathrm{ee}',
+        'emu': '\mathrm{e} \mu',
+        'etau': '\mathrm{e} \mathrm{\\tau}',
+        'mue': '\mu \mathrm{e}',
+        'mumu': '\mu \mu',
+        'mutau': '\mu \mathrm{\\tau}',
+        'taue': '\\tau \mathrm{e}',
+        'taumu': '\\tau \mu',
+        'tautau': '\\tau \\tau'
+    }
+    return r'$\varepsilon_P^{u d}\left\{' + latex_mapping[last_part] + r'\right\}$'
+
+file_path = 'CC_P_us_new_resultsBOTH.txt' # Update the path if needed
+
+interpolation_results = {}
+
+with open(file_path, 'r') as file:
+    current_name = None
+    x_values = []
+    y_values = []
+    
+    for line in file:
+        if not line.strip():
+            continue
+            
+        name, param, chi2 = line.strip().split()
+        param, chi2 = float(param), float(chi2)
+
+        if name != current_name and current_name is not None:
+            interp_func = interp1d(y_values, x_values, kind='linear', bounds_error=False, fill_value='extrapolate')
+            param_value_at_chi2_2_71 = interp_func(2.71)
+            interpolation_results[current_name] = param_value_at_chi2_2_71
+            x_values, y_values = [], []
+            
+        current_name = name
+        x_values.append(param)
+        y_values.append(chi2)
+
+    interp_func = interp1d(y_values, x_values, kind='linear', bounds_error=False, fill_value='extrapolate')
+    param_value_at_chi2_2_71 = interp_func(2.71)
+    interpolation_results[current_name] = param_value_at_chi2_2_71
+
+names_latex = [convert_name_to_latex(name) for name in interpolation_results.keys()]
+param_values = [value for value in interpolation_results.values()]
+
[…]
